### 2013-05-13 Rcode/BFcompare.rmd
first commit

try comparing BF for normal vs servin and stephens bf

```{r}
lABF = function(k,T ){return(((k/2)*T^2))}
lBF = function(k,T,n){return((-n/2)*log(1-(k/n)*T^2))}

n=seq(1,100,length=100)
T=2
k=.4/T^2

plot(n,lBF(k,T,n)-lABF(k,T),type="l")
abline(h=0,col=2)
```
Note that lBF>lABF always. So lABF is "biased" towards the null?
Is that right? Seems wrong, as the normal assumption is somehow
less conservative than the t?

Of course x<log(1+x)
```{r}
x=seq(0,1,length=100)
plot(x,log(1+x))
abline(a=0,b=1)
```
which seems to imply it is true.

Now let's compare
H0: \hat{\beta} \sim S T_\nu
H1: \hat{\beta} \sim \beta + S T_\nu
with \beta \sim N(0,W)
Here S is an estimate of the standard error of betahat.

the BF is the integral over beta of 
the ratio of T densities:
```{r}
BFt=function(W,betahat,S,nu){
  beta = rnorm(1000000,0,sd=sqrt(W))
  bft=mean(dt((betahat-beta)/S,nu) / dt(betahat/S,nu))
  #bfn= mean(dnorm((betahat-beta)/S)/dnorm(betahat/S))
  bfn= dnorm(betahat,0,sd=sqrt(W+S^2))/dnorm(betahat,0,sd=S)
  Zeff = effective.Zscore(betahat/S,nu)
  Seff = betahat/Zeff
  betahat.eff = Zeff*S
  cat(pnorm(Zeff),pnorm(betahat.eff/S),Zeff,betahat,Seff,pnorm(betahat/Seff),"\n")
  bfn.Seff= dnorm(betahat,0,sd=sqrt(W+Seff^2))/dnorm(betahat,0,sd=Seff)
  bfn.beff= dnorm(betahat.eff,0,sd=sqrt(W+S^2))/dnorm(betahat.eff,0,sd=S)
  return(list(bft=bft,bfn=bfn,bfn.Seff=bfn.Seff,bfn.beff=bfn.beff))
}
BFt(1,10,1,2)

```

Now transform T to Z be
```{r}
effective.Zscore = function(T,nu){
  return(qnorm(pt(T,nu)))
}
Z = effective.Zscore(T,n)
plot(n,lBF(k,T,n)-lABF(k,Z),type="l")
abline(h=0,col=2)

```

### 2013-05-13 Rcode/intro.rmd
first commit

It is common in statistics that you measure many ``similar" things imperfectly, and wish to estimate their values. The situation arises commonly in the kinds of genomics applications I am often involved in, but also in other areas of statistics.
In genomics, for example, a very common goal
is to compare the mean
expression (activity) level of many genes in two conditions.
Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
measurements are made on only a small number of
samples in each condition - sometimes as few as one
sample in each condition. Thus the error in estimates
of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
in estimates of $\beta_j$ still greater.

A fundamental idea is
that the measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.

### 2013-05-13 Rcode/readme.rmd
first commit

Suppose that we measure, with error, a series of ``effects", $\beta_1,\dots,\beta_J$. To take just one concrete 
example, $\beta_j$ could be 
the difference in the mean (log) expression levels of gene $j$ ($j=1,\dots,J$) between 2 conditions. In this case, a measurement of
the effect might be the difference in sample means obtained in the two conditions. We will let $\hat\beta_j$ denote the measured value of $\beta_j$, and assume that
each measurement comes with an associated standard error, 
$s_j$. A key aim here will be to take proper account of the fact that some measurements may be more precise than others: that is, to take proper account of variation in $s_j$ across $j$.

A common goal, particularly in genomic studies, is to identify which $\beta_j$ differ from zero. This is commonly tackled
by first computing a $p$ value for each $j$, testing $H_j:\beta_j=0$, and then applying methods based on False Discovery Rates, such as the qvalue package. For example, we could compute $Z$ scores $Z_j = \beta_j/s_j$, and translate $Z_j$ to a $p$ value, 
$p_j$, and then use now-standard methods to select a $p$ value threshold to control the FDR.

There are two issues with this approach that I would like to address here. The first is that it really does not take proper account of the measurement errors. To see this, consider an example where half the measurements are quite precise, and the other half are really, really, poor. Intuitively, the poor measurements tell us nothing, and any sane analysis should effectively ignore them. However, in a standard FDR-type analysis, these poor measurements add ``noise" and affect estimated FDRs. This is because the $p$ values from the poor measurements will be effectively uniformly distributed, and some will be significant at any given threshold.

I will illustrate this with an example. The following R code
simulates some data where the effects are normally distributed,
and the first 500 observations have good precision, and the next 500 have poor precision.

```{r}
#install q value package
#source("http://bioconductor.org/biocLite.R")
#biocLite("qvalue")
setwd("~/Documents/BayesFDR/Rcode/")
library("qvalue")
library("lattice") #library for some of the plots

#set up some data with mixture of values of s
set.seed(100)
s.good = 0.5
s.poor = 10
J.good= 500
J.poor = 500
J=J.good+J.poor
beta = c(rnorm(J,0,1))
s = c(rep(s.good,J.good),rep(s.poor,J.poor))
betahat = beta + rnorm(J,0,s)
#compute the usual zscore and corresponding p value
zscore = betahat/s
pval = pchisq(zscore^2,df=1,lower.tail=F)
```
As expected the $p$ values from the poor observations are approximately uniform, whereas those from the good observations  are enriched for small $p$ values:

```{r}
p= data.frame(pval=pval,type=c(rep("GOOD",500),rep("POOR",500)))
histogram(~pval|type,p,layout=c(1,2),xlim=c(0,1),breaks=seq(from=0,to=1,length=20),type="count")
```

Now what happens if we apply FDR methods to the all the data is
that the uniform $p$ values from the poor observations
add noise relative to looking at the good observations only. This impacts both the estimate of $pi_0$ from qvalue, and the number of findings that are significant at a given FDR.
```{r}
qq.all = qvalue(p$pval)
qq.good = qvalue(p$pval[p$type=="GOOD"])
print(c(qq.good$pi0, qq.all$pi0))
print(c(sum(qq.good$qvalues<0.05),sum(qq.all$qvalues<0.05)))
```

The two main assumptions we will make here
are that 
* these effects are exchangeable: that is (by de Finetti's theorem) they can be
thought of as being independent and identically distributed from some unknown distribution $g(\beta)$. 
* these effects are symmetrically distributed about 0, with density monotonically decaying as we move away from 0.

Specifically we will assume a parameteric form for $g$
as a mixture of 0-centered normal distributions

A key issue I want to address here is that if these measurements are made with different precisions, then we want to
take this into account in our analysis.
We can do this by assuming that the likelihood for $\beta$ is a normal, with mean $\hat\beta_p$ and standard deviation $s_p$. (Note that this is equivalent to the likelihood we would get
if we ``observed" data $\hat\beta_p \sim N(\beta_p, s_p)$.)

```{r}
#install q value package, and load in ash code
source("../Rcode/ash.R")
beta.ash.all = ash(betahat,s)
beta.ash.good = ash(betahat[p$type=="GOOD"],s[p$type=="GOOD"])
plot(beta.ash.good$PosteriorMean,beta.ash.all$PosteriorMean[p$type=="GOOD"])
plot(beta.ash.good$PositiveProb,beta.ash.all$PositiveProb[p$type=="GOOD"])
```
So here it makes no difference whether you use all the observations or the just the good ones -
the noise is ignored (as it should be!)

### Analyse Hedenfalk et al data.
Note that there were a few genes with odd observations (eg in the thousands, rather than the usual expression measures), and 
following Storey's ODP paper I removed the genes with any observation >20.
```{r}

#note: I edited nejm_brca_release.txt by removing columns 1-3
print(getwd())
hh = read.table("../data/nejm_brca_release_edit.csv",sep=",",skip=3)
subset = apply(hh, 1, max)<20
hh = hh[subset,]

labs = read.table("../data/nejm_brca_release_edit.csv",sep=",",skip=1,nrows=1)
labs = 1*(labs=="BRCA1") + 2 * (labs=="BRCA2")

hh.betahat = apply(hh[,labs==1],1,mean) - apply(hh[,labs==2],1,mean)
n1 = sum(labs==1)
n2 = sum(labs==2)
hh.sebetahat = sqrt(apply(hh[,labs==1],1,var)/n1 + apply(hh[,labs==2],1,var)/n2)
hh.zscore = hh.betahat/hh.sebetahat
hh.pval = pchisq(hh.zscore^2,df=1,lower.tail=F)
hist(hh.pval,probability=TRUE,xlab="p value",main="Distribution of p values for Hedenfalk et al data",nclass=40,col=5)
hh.q = qvalue(hh.pval)
abline(h=hh.q$pi0,col=2,lwd=2)
sum(hh.q$q<0.05)
```

Now compare this with ASH
```{r}
hh.ash = ash(hh.betahat,hh.sebetahat,df=n1+n2-2)
sum(hh.ash$qvalue<0.05)
plot(hh.ash$localfdr, log(hh.pval),main="Illustration of how ASH reorders significance of observations",ylim=c(-10,-1),xlim=c(0,0.1),cex=0.2,xlab="ASH local fdr",ylab="Original p value")
```

Now see how it compares with ignoring the variability in
measurement error, by applying ASH to the z scores.
```{r}
hh.ash2 = ash(hh.zscore,rep(1,length(hh.zscore)),df=20)
plot(hh.ash$localfdr,hh.ash2$localfdr,cex=0.2,ylab="ignoring sd",xlab="accounting for sd",main="Effect of accounting for measurement precision on local fdr")
abline(a=0,b=1,col=2)
cat(sum(hh.ash$qvalue<0.05),sum(hh.ash2$qvalue<0.05))
cat(sum(hh.ash$localfdr<0.05),sum(hh.ash2$localfdr<0.05))
plot(hh.ash2$localfdr, log(hh.pval),main="Illustration of how ASH without precision does not reorder significance?",ylim=c(-10,-1),xlim=c(0,0.1),cex=0.2,xlab="ASH local fdr",ylab="Original p value")
```
Hmm - this was unexpected. Why are more being called
in the second analysis? Could it be that the effect sizes
are larger for larger standard errors, as using the 
z scores effectively assumes?

It would be nice to be able to assess the fit of the different models
(ie are effect sizes larger for larger  standard errors?).
In principle this should be assessable via likelihood, but
need to code it. This section is work in progress.

The following computes the log likelihood for the two fitted ash models,
first with $\beta \sim \sum_k \pi_k N(0,\sigma_k^2)$, and then $\beta_s \sim \sum_k \pi_k N(0, \sigma^2_k se^2_s)$:
```{r}
cat(mixseLoglik(hh.betahat,hh.ash$fitted.f$pi, hh.ash$fitted.f$mu, hh.ash$fitted.f$sigma, hh.sebetahat),mixseLoglik(hh.betahat,hh.ash2$fitted.f$pi, hh.ash2$fitted.f$mu, sqrt(hh.ash2$fitted.f$sigma^2+1), hh.sebetahat,FUN="*"))
```
Looks like the second model fits better.

```{r}
plot(hh.betahat, hh.sebetahat)
```

Another unexpected feature is that the ASH analysis based
on z scores is rather different from the standard q value analysis, even though they are both based on z scores.
Look:
```{r}
plot(as.numeric(hh.q$qvalue),hh.ash2$qval,cex=0.2,ylab="ASH q value",xlab="standard q value",main="ASH vs standard q value")
cat(hh.ash2$fitted.f$pi[1],hh.q$pi0)
```
At first I was really worried about this - the pi0 estimate
is much bigger using the p values. However, this is all to
do with the assumptions. The q value approach estimates
pi0 by looking at the p values near 1. Effectively it assumes
that all the p values near 1 are null. At first this seems OK.
But note that it actually corresponds to assuming that all the z scores near 0 are null - that is, you can never see a z score of 0 under the alternative. In contrast we assume that the distribution of z scores under the alternative also has its mode at 0 - under this assumption not all z scores near 0 will be null at all.... these are very different assumptions. 
(Of course, z scores near 0 won't give "significant"" results in either approach.)

Now I wanted to investigate what distribution for z values under the alternative the standard approach implicitly corresponds to.
```{r}
x= seq(-6,8,length=100)
hist(hh.zscore,prob=TRUE,nclass=100)
lines(x,dnorm(x)*hh.q$pi0,col=3,lwd=2)
z.kerneld=density(hh.zscore)
lines(z.kerneld$x,z.kerneld$y,col=2,lwd=2)
plot(z.kerneld$x,z.kerneld$y-dnorm(z.kerneld$x)*0.6,type="l")
abline(h=0)
```

#### Miscellaneous

code below here is not "tidied"

This was to test the mixseLoglik function.
```{r}
n= length(hh.sebetahat)
se.test =median(hh.sebetahat)
bhat.test = rnorm(n,0,se.test)+rnorm(n,0,hh.sebetahat)
bhat2.test = rnorm(n,0,hh.sebetahat) + rnorm(n,0,hh.sebetahat)

mixseLoglik(bhat.test,c(0,1),c(0,0),c(se.test,se.test),hh.sebetahat)
mixseLoglik(bhat.test,c(0,1),c(0,0),c(se.test,se.test),hh.sebetahat,FUN="*")

mhat = mean(bhat2.test)
sdhat = sd(bhat2.test)
sum(dnorm(bhat2.test,mhat,sdhat,log=TRUE))
mixseLoglik(bhat2.test,c(0,1),c(0,0),sqrt(c(2,2)),hh.sebetahat,FUN="*")

mixseLoglik(bhat.test,c(0,1),c(0,0),c(se.test,se.test),hh.sebetahat)
mixseLoglik(bhat.test,c(0,1),c(0,0),c(se.test,se.test),hh.sebetahat,FUN="*")

z.test = bhat.test/hh.sebetahat
z2.test = bhat2.test/hh.sebetahat

p.test = pchisq(z.test^2,df=1,lower.tail=F)
p2.test = pchisq(z2.test^2,df=1,lower.tail=F)

hist(p.test)
hist(p2.test)

test1.ash1 = ash(bhat.test, hh.sebetahat)
test1.ash2 = ash(z.test, rep(1,n))
cat(sum(test1.ash1$qval<0.05),sum(test1.ash2$qval<0.05))

test2.ash1 = ash(bhat2.test, hh.sebetahat)
test2.ash2 = ash(z2.test, rep(1,n))
cat(sum(test2.ash1$qval<0.05),sum(test2.ash2$qval<0.05))

mixseLoglik(bhat.test,test1.ash1$fitted.f$pi,test1.ash1$fitted.f$mu,test1.ash1$fitted.f$sigma,hh.sebetahat)
mixseLoglik(bhat.test,test1.ash2$fitted.f$pi,test1.ash2$fitted.f$mu,sqrt(test1.ash2$fitted.f$sigma^2+1),hh.sebetahat,FUN="*")

mixseLoglik(bhat2.test,test2.ash1$fitted.f$pi,test2.ash1$fitted.f$mu,test2.ash1$fitted.f$sigma,hh.sebetahat)
mixseLoglik(bhat2.test,test2.ash2$fitted.f$pi,test2.ash2$fitted.f$mu,sqrt(test2.ash2$fitted.f$sigma^2+1),hh.sebetahat,FUN="*")
```

Here we look at the distribution of the average of p ABFs,
under the null, for different values of p. 
```{r}
nsim = 1000
pmax = 10000
z = matrix(rnorm(nsim*pmax),ncol=pmax)
abf = ABF(z,1,1)
pvec=c(5,10,20,50,100,1000,10000)
q=rep(0,length(pvec))
for(i in 1:length(pvec)){
abf.m = rowMeans(abf[,1:pvec[i]])
q[i] = quantile(abf.m,0.95)
}
plot(q)
```

Here look at the 95th percentile of ABF for different sigma:
```{r}
nsim = 10000
z = rnorm(nsim)
Wvec = seq(0,10,length=101)
q= rep(0,length(Wvec))
for(i in 1:length(Wvec)){
  abf = ABF(z,1,Wvec[i])
  q[i] = quantile(abf,0.95)
}
plot(q)
```
so it seems that even in the worst case, the 95th quantile is never much more than 2. It seems this could be used to bound the probability of making a false discovery even for p=1. Then (I believe) under the null, the probability of making a false discovery under the null should be even smaller for large p.

And idea; this seems related to the sellke and berger result. at the 0.05 point, the ABF never exceeds about 2. Can we make arguments about false positive rates based on this?

An Idea: for a single datum
Pr(data | pi) = pi ABF + 1-pi

so if a priori pi \sim Di(1,1)
then pi | data \sim w1 Di(2,1) + w2 Di(1,2)
so worst case is pi is \Di(2,1) and posterior mean of pi is 2/3.

Or P(data and H1) = pi ABF
P(data and H0) = (1-pi)

```{r}

#set up some data with mixture of values of sigmaa
set.seed(100)
#sebetahat = sample(c(1,0.1),size=1500,replace=T) 
sebetahat = 0.01*rgamma(1500,.1,.1)
beta =  c(rnorm(500,0,1),rnorm(500,0,0.5),rnorm(500,0,0.000001))
betahat = beta + rnorm(1500,0,sebetahat)

#compute the usual zscore and corresponding p value
zscore = betahat/sebetahat
pval = pchisq(zscore^2,df=1,lower.tail=F)

#apply the ash method to do adaptive shrinkage of betahat
#and the qvalue method to get 
beta.ash = ash(betahat,sebetahat)
qq = qvalue(pval)
```

Now we want to see how the results compare
```{r}

attach(beta.ash)
conf=ifelse(PositiveProb>0.5,PositiveProb,1-PositiveProb)
sum(conf>0.95)
sum(qq$qvalues<0.05)
```
So there are more significant at FDR=0.05, than at local bayes FDR 0.05. But this is not suprising because of the difference between local FDR and FDR.

Indeed, let's look at the errors
```{r}
err = (sign(betahat) != sign(beta))
table(err[conf>0.95])
table(err[qq$qvalues<0.05])

#check whether ordering by q values does better or worse
#job than ordering by confidence, in terms of identifying
#betas with the right sign

plot(cumsum(err[order(qq$qvalues)]),type="l")
lines(cumsum(err[order(conf,decreasing=TRUE)]),col=2)
```

### 2013-05-13 talks/009-slides.Rmd
first commit

% Writing beautiful and reproducible slides quickly
% Yihui Xie
% 2012/04/30

```{r setup, include=FALSE}
# set global chunk options
opts_chunk$set(cache=TRUE)
```

# Why

- after you finished typing `\documentclass{beamer}` and `\title{}`, I have finished my first slide with markdown
- much less commands to remember, e.g. to write bullet points, just begin with a dash "`-`" instead of `\begin{itemize}` and `\item`; how things can be simpler?
- I know you want math to show you are a statistician, e.g. $f(k)={n \choose k}p^{k}(1-p)^{n-k}$
- you do not need to maintain output -- only maintain a source file
- HTML5/CSS3 is much more fun than LaTeX

# A bit R code

```{r computing}
head(cars)
cor(cars)
```

# Graphics too

```{r graphics, message=FALSE, fig.cap='A scatterplot of `cars`', fig.width=11, fig.height=5}
library(ggplot2)
qplot(speed, dist, data = cars) + geom_smooth()
```

# How

- source editor: [RStudio](http://www.rstudio.org/) (perfect integration with [**knitr**](http://yihui.name/knitr/); one-click compilation); currently you have to use the version >= 0.96.109
- HTML5 slides converter: [pandoc](http://johnmacfarlane.net/pandoc/); this document was generated by: `pandoc -s -S -i -t dzslides --mathjax knitr-slides.md -o knitr-slides.html`
- the file [`knitr-slides.md`](https://github.com/yihui/knitr-examples/blob/master/009-slides.md) is the markdown output from its [source](https://github.com/yihui/knitr-examples/blob/master/009-slides.Rmd): `library(knitr); knit('knitr-slides.Rmd')`
- or simple click the button `Knit HTML` in RStudio

# For ninjas

- you should tweak the default style; why not try some [Google web fonts](http://www.google.com/webfonts)? (think how painful it is to wrestle with fonts in LaTeX)
- pandoc provides 3 types of HTML5 slides (dzslides is one of them)
- you can tweak the default template to get better appearances
- if you have come up with a better dzslides template, please let me know or contribute to pandoc directly (e.g. `pre` blocks should have `max-width` and `max-height`)

# For beamer lovers

- pandoc supports conversion to beamer as well. period.

# For Powerpoint lovers

- ...

# Reproducible research

It is good to include the session info, e.g. this document is produced with **knitr**. Here is my session info:

```{r session-info}
print(sessionInfo(), locale=FALSE)
```

# Misc issues

- the plots are too wide? use the chunk option `out.width` which will be used in `<img width=... />`, e.g. `out.width=400px`

# Life is short

- so keep your audience awake!

### 2013-05-13 talks/ilike-slides.Rmd
first commit

% Apaptive Shrinkage and False Discovery Rates by Laplace Approximation
% Matthew Stephens
% 2013/5/13

```{r setup, include=FALSE}
# set global chunk options
opts_chunk$set(cache=TRUE,autodep=TRUE)
dep_auto()
library("qvalue")
```

# Motivation

- A common problem: you have imperfect measurements
of many ``similar" things, and wish to estimate their values.

- Particularly common in genomics. For example,
 a very common goal is to compare the mean
expression (activity) level of many genes in two conditions.

# Example: Mouse Heart Data

```{r, include=FALSE}
setwd("~/Documents/git/BayesFDR/talks/")
## load Poisson_binomial and ash functions 
source("../../stat45800/code/Rcode/PoissonBinomial.funcs.R")  
source("../Rcode/ash.R") 
x = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
```

- Data on 150 mouse hearts, dissected into left and right ventricle
(courtesy Scott Schmemo, Marcelo Nobrega)

# Example: Mouse Heart Data
```{r}
head(x)
```

```{r, echo=FALSE}
cc = x[, 2:5]
g = c(-1, -1, 1, 1)
ngene = dim(x)[1]

cc.assoc = counts.associate(cc, g, 1)
zdat.ash = cc.assoc$zdat.ash
zdat = cc.assoc$zdat

# two-sided test
ttest.pval = function(t, df) {
    pval = pt(t, df = df, lower.tail = T)
    pval = 2 * ifelse(pval < 0.5, pval, 1 - pval)
    return(pval)
}

tscore = zdat[3, ]/zdat[4, ]
pval = ttest.pval(tscore, df = 2)

```

# Borrowing Strength

- For gene $j$, denote the difference in mean between condition 0 and condition 1 by $$\beta_j:= \mu^0_j - \mu^1_j.$$

- As measurements are made on only two
samples in each condition, error in estimates
of $\beta_j$ are appreciable.

- Fundamental idea: measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.

# False Discovery Rates

-Standard practice is to estimate an effect size $\beta_j$ and standard error $s_j$ for each gene.

-Then convert this to a $p$ value for each gene, e.g. by a $t$ test on $\beta_j/s_j$.

-Then use the distribution of $p$ values to estimate the false discovery rate (FDR) at a given threshold.

# False Discovery Rates

```{r, echo=FALSE}
hist(pval,prob=TRUE)
qval = qvalue(pval)
abline(h=qval$pi0,col="red",lwd=2)
```

# False Discovery Rates

```{r, echo=FALSE}
h=hist(pval,prob=TRUE)
qval = qvalue(pval)
abline(h=qval$pi0,col="red",lwd=2)
rect(0,0,0.05,qval$pi0,col="red")
rect(0,qval$pi0,0.05,h$intensities[1],col="green")
abline(v=0.05,lwd=3)
```

#FDR methods are very widely used!

- Why?

- Partly, they make sense.

- Partly, they are generic. Once you have the $p$ values, you don't really care where they came from.

#Issues

- Different precision of different measurements 
is forgotten once the $p$ value is computed.

- Borrows information for testing, not for estimation.

#An Alternative

- Instead of summarizing each gene by a single number ($p$ value)
summarize by two numbers, $\hat\beta_j$ and its standard error $\hat\s_j$.

- Then do inference for $\beta_j$ by Laplace approximation.

# Example: Wavelets

# Reproducible research

This document is produced with **knitr**, **Rstudio** and **Pandoc**.

Here is my session info:

```{r session-info}
print(sessionInfo(), locale=FALSE)
```

# Misc issues

- the plots are too wide? use the chunk option `out.width` which will be used in `<img width=... />`, e.g. `out.width=400px`

### 2013-05-13 talks/ilike.2013.tex
first commit

% CVSId: $Id: Example.tex,v 1.1.1.1 2000/11/28 11:15:12 exupery Exp $
\documentclass[ignorenonframetext]{beamer}

\usepackage{tikz}
\usepackage{amsbsy}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{latexsym}
\usepackage{algorithm}
\usepackage{multirow}
\usepackage{algorithmic}
\usepackage{color} 
\usepackage{graphicx}
\usepackage{tabularx}
%\usepackage{hyperref}
\usepackage{eepic}
\usepackage{booktabs}
\usepackage{epic}
\def \hatd { \widehat{ d } }
\def \dhatin {\hatd^{\mbox{\tiny IN}}}
\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
\def \mvn{\text{MVN}}
\def\mn{\text{MN}}
\def \iw{\text{IW}}
\def \fp {fastPHASE\ }
\def\var{\rm Var}
\def\BF{\text{BF}}
\def\BFfull{\text{BF}_\text{all}}
\def\mvr{\text{BMVR}}
\def\one{\mbf{1}}
\def\zero{\mbf{0}}
\def\sxx{S_{xx}}
\def\syy{S_{yy}}
\def\rss{\text{\rm RSS}}
\def\I{I}
\def\D{D}
\def\U{U}
\def\p{p}
\def\g{G}
\def\bfav{BF_{\text{av}}}
\def\pall{p_\text{all}}

%\ColorFoot{2}
\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
                                % mode too
\def\BF{\rm BF}
\def\hfull{\textcolor{red}{H_\text{all}}}
\def\hnull{\textcolor{blue}{H_0}}

\mode<article>
{
  \usepackage{fullpage}
  \usepackage{pgf}
  \usepackage{hyperref}
  \setjobnamebeamerversion{sanger.beamer}
}

\mode<presentation>
{
  \usetheme{Dresden}

  \setbeamercovered{transparent}
}

\usepackage[latin1]{inputenc}
\usepackage[english]{babel}
\def \fp {fastPHASE\ }

\title[Adaptive Shrinkage by Laplace Approximation]{Adaptive Shrinkage and False Discovery Rates by Laplace Approximation}
\author{Matthew Stephens} \date{i-like, Warwick,  May 2013 }

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Motivation}  
A common problem: you have imperfect measurements
of many ``similar" things, and wish to estimate their values.

This is particularly common in genomics. For example,
 a very common goal is to compare the mean
expression (activity) level of many genes in two conditions.

Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference.

Typically expression
measurements are made on only a small number of
samples in each condition - sometimes as few as one
sample in each condition. Thus the error in estimates
of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
in estimates of $\beta_j$ still greater.

Marcelo's Heart data as an example.

Wavelets as a second application.
\end{frame}

\begin{frame}{Borrowing Strength}
A fundamental idea is
that the measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.
\end{frame}

\begin{frame}{False Discovery Rates}

What is actually done in practice?

\end{frame}

\begin{frame}{Genetic Association Studies}

So a typical GWAS analysis involves fitting millions of simple regressions, and testing effects for significance.

\bigskip

Notation: $y$ for phenotype, $g$ for genotype, $\beta$ for genetic effects:

$$y_i = \mu + \beta g_i + \epsilon_i \qquad (i=1,\dots,n)$$
\end{frame}

\begin{frame}{Genetic Association Studies and Heterogeneity}

We have been developing statistical methods for association mapping in multiple subgroups, incorporating heterogeneity of effects.
See also Lebrec et al (2010); Han and Eskin (2011,2012). 
\bigskip

Motivating examples include:
\begin{enumerate}
 \item The ``Global Lipids Consortium" Genome-wide Association Study meta-analysis (Teslovich et al, 2010).
 \item Gene expression analysis among multiple tissues (e.g.~Dimas et al, 2009).
 \end{enumerate}
\end{frame}

 \begin{frame}{Example 1: Global Lipids meta-analysis}

GWAS data from the Global Lipids consortium (Teslovich et al, 2010) on $>100,000$ individuals
from at least 25 separate studies.

\bigskip 
Four phenotypes: total cholesterol (TC), low-density lipoprotein cholesterol
(LDL-C), high-density lipoprotein cholesterol (HDL-C) and triglycerides (TG).

\bigskip
The original association analyses performed a {\it fixed-effects meta-analysis}.
That is, assume the effects in each subgroup  are all equal ($\beta_s = \beta \, \forall s$ ), and test $H_0: \beta=0$.

\bigskip
They reported a total of 95
SNPs as being associated with one or more phenotypes.

\end{frame}

\begin{frame}{Example 1: Global Lipids meta-analysis}

{\bf Question:} Given that this study involved 25 separate subgroups, many with quite different recruitment criteria
(e.g. some were recruited as cases for a particular disease; others were recruited as controls; etc),
would an analysis that allows for heterogeneity across studies identify more associations?
 \end{frame}

 \begin{frame}{Example 2: eQTL studies across multiple cell-types}

 Dimas et al (2009), measured expression data in 75 individuals, in 3 cell types: Fibroblasts, LCLs and T-cells.

 \bigskip

A key goal was to identify genetic variants associated with expression that were shared among
 cell types, or were specific to some subset of cell-types.  (Identifying eQTLs specific to individual cell types may shed insight into
 cell-type-specific regulation mechanisms.)

 \bigskip

 Original analysis performed association analysis separately in each cell type, and then looked at the overlap
 of the resulting associations.  The overlap was small (14\%), and they concluded that many eQTLs occur in only one cell type.

 \end{frame}

 \begin{frame}{Example 2: eQTL studies across multiple cell-types}

 {\bf Question:} Incomplete power may cause this analysis to underestimate sharing; does a joint analysis of all cell types
 come to the same conclusion?
\end{frame}

\begin{frame}{Methods}

Focus first on meta-analysis, where effect $\beta$ may vary across subgroups $s$:
$$y_{si} = \mu_s + \beta_s g_i + \epsilon_{si} \; \text{ with }\epsilon_{si} \sim \mathcal{N}(0,\sigma_s^2).$$

\bigskip

Primary goal of the meta-analysis is to identify SNPs for which there is strong evidence
against $H_0: \beta_s \equiv 0 \, \forall s$. 
\end{frame}

\begin{frame}{Alternatives to $H_0$}

To assess evidence against $H_0$ we introduce a set of alternative models, indexed
by parameters $\phi, \omega$, to be compared with $H_0$.

\bigskip

$H_1(\phi, \omega): \beta_s \text{ normally distributed about common mean } \bar \beta$.
$$\beta_s |\bar \beta \sim N(\bar\beta,\phi^2); \qquad \bar\beta \sim N(0, \omega^2).$$

\bigskip

{\bf Note 1:} $\phi=0$ corresponds to the usual ``fixed effects" alternative, $\beta_s = \bar \beta \, \forall s$.

\bigskip
{\bf Note 2:} Can alternatively work with the ``standardized" effect sizes, $b_s = \beta_s/\sigma_s$, which 
generally yields similar (but not identical) results.
\end{frame}

\begin{frame}{Bayes Factors}

The Bayes Factor
$$\BF(\phi,\omega) = p(y | g, H_1(\phi, \omega)) / p(y | H_0)$$
measures the support for $H_1(\phi,\omega)$ vs $H_0$, with large
values indicating strong evidence against $H_0$.

\bigskip
Although $\BF(\phi,\omega)$ depends on priors for nuisance parameters $(\mu,\sigma^2_s)$, it is not very sensitive,
and sensible default choices exist.

\bigskip 
Hyperparameters $\phi$ and $\omega$ must be chosen to reflect expected effect sizes and levels of heterogeneity
(but can average over several values to reflect uncertainty in choice of appropriate values).
\end{frame}

\begin{frame}{Computation}

$\BF(\phi,\omega)$ can be quickly and accurately approximated by Laplace approximation.

\bigskip

In the simplest cases these approximations depend only on the summary statistics
in each study, $\hat\beta_s$ and $\rm{se}(\hat\beta_s)$.
(Details: Wen and Stephens, 2011).
\end{frame}

\begin{frame}{Bayes Factors and standard test statistics}

This framework includes some commonly-used frequentist test statistics as special cases.
\bigskip

For example, if we allow $\omega$ to vary across SNPs according to the inverse
of the standard error of $\bar \beta$ then $BF(\phi=0, \omega)$ is monotonic with the weighted $Z$ score
\begin{equation} \label{weighted.Zscores.eqn}
\mathcal{Z} = \frac{\sum_s w_s Z_s}{\sqrt{\sum_{s'} w_{s'}^2}}
\end{equation}
where $Z_s=\hat\beta_s/ {\rm se}(\hat{\beta}_s)$ and $w_s =  {\rm se}(\hat{\beta}_s)^{-1}$. (Details: Wen and Stephens, 2011.)

\bigskip
In other words, we can see what implicit models for $\beta$ are assumed by standard methods.
\end{frame}

\begin{frame}{Example 1: Global Lipids meta-analysis}

{\bf Question:} Given that this study involved 25 separate subgroups, many with quite different recruitment criteria
(e.g. some were recruited as cases for a particular disease; others were recruited as controls; etc),
would an analysis that allows for heterogeneity across studies identify more associations?

 \bigskip

 \pause{{\bf Answer:} Not much!}
 \end{frame}

 \begin{frame}{Results: Global Lipids GWAS}

\begin{itemize}
\item Searched genome-wide for SNPs with strong signal when allowing
for heterogeneity ($\BF_{\rm het}>10^6$) but not when assuming no heterogeneity ($\BF_{\rm no-het}<10^6$).
\item 42 SNPs satisfied these criteria.
\item But 36 of these were driven by apparently strong associations in a single study (Framingham heart Study), and seemed likely to be due to data processing errors.
\item Two more showed similarly suspicious patterns (association in just one study, a subset of the WTCCC). 
\end{itemize}
\end{frame}

\begin{frame}{Results: Global Lipids GWAS}

\begin{table}[h!t]
\begin{center}
\begin{tabular}{ l c c c c }
\toprule
Phenotype & SNP & Gene & $\log_{10} ({\rm BF}_{\rm no-het})$ &  $\log_{10}({\rm BF}_{\rm het})$ \\
\midrule
LDL & rs1800978 & {\footnotesize  ABCA1} & 5.2  & 6.0  \\
TG  & rs1562398 & {\footnotesize nr KLF14} & 5.3 &  6.5 \\
%TG &  rs765547  & {\footnotesize Flanking LPL}   & 1.7 & 7.9 & 5.8  \\ 
HDL & rs11229165 & {\footnotesize nr OR4A16} & 4.6 & 6.4 \\
HDL & rs7108164 & {\footnotesize nr OR4A42P} & 4.2 & 6.3 \\ 
%HDL & rs11984900 & {\footnotesize N.A. } &-1.1 & 16.6 & 6.2 \\
%HDL & rs6995137 & {\footnotesize Flanking SFRP1} & -0.4 & 6.9 & 4.8 \\
\bottomrule
\end{tabular}
%\caption{\label{lipids.miss.tbl} Association signals that show strong association under the models allowing for heterogenetiy (${\rm ABF^{EE}_{cefn}} \text{ or } {\rm ABF_{maxH}^{EE}} \ge 10^6$) but less strong under a model with no heterogeneity (${\rm ABF^{EE}_{fix}}< 10^6$). It seems likely that the last two of these represent false positive associations, but we include them in the table for completeness (see text for discussion).}
\end{center}
\end{table}
\end{frame}

\begin{frame}{Example 2: eQTL sharing across cell-types}

 In meta-analysis application the primary goal was to reject the global null, $H_0: \beta_s =0 \, \forall s$.

 \bigskip

 Mapping eQTLs among multiple cell-types (subgroups) differs 
 in that we care more about {\it which} $\beta_s$ are non-zero, and patterns
 of sharing among subgroups.

 \bigskip

E.g. Dimas et al (2009), identified eQTLs separately in 3 cell types, and asked which eQTLs are shared among cell types (subgroups).

 \end{frame}

 \begin{frame}{Example 2: eQTL sharing across cell-types}

 To address this we expand our alternative models  $H_1(\phi,\omega)$ to
 allow that effects may be zero in some subgroups.

\bigskip

Introduce a \alert{configuration} $\gamma$ indicating which subgroups have non-zero effect.
\begin{itemize}
\item E.g. $\gamma=[110]$ corresponds to non-zero effect in the first two subgroups.
\end{itemize}

\bigskip

See also Han \& Eskin (PloS Genetics, 2012).

%\pause
%\bigskip
%An eQTL can have different effects in different tissues ($\beta_1 \neq \beta_2$):
%\begin{itemize}
% \item $b_s \sim \mathcal{N}(\bar{b}, \alert{\phi^2})$ and $\bar{b} \sim \mathcal{N}(0,\alert{\omega^2})$
%\item $\omega$: magnitude of the average effect
%\item $\phi$: magnitude of heterogeneity
% \item both are specified by a grid of values
%\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Bayesian Model Averaging and hierarchical modeling}

The support in the data for configuration $\gamma$ can be measured by the Bayes Factor
$$\BF_\gamma(\phi,\omega) = \frac{p(y | H_1(\gamma,\phi,\omega))}{p(y | H_0)}$$.

\bigskip

Overall evidence against $H_0$ can be measured by averaging over $\gamma, \phi, \omega$:
$\text{BMA} = \sum_{\gamma,\phi,\omega} \eta_{\gamma,\phi,\omega} \text{BF}_\gamma(\phi,\omega)$

\bigskip

Estimate proportions $\eta_{\gamma,\phi,\omega}$ using a hierarchical model to combine information across genes.

\end{frame}

%\begin{frame}
%\frametitle{Bayes Factor (BF) and hierarchical model (HM)}
%\begin{itemize}
%\item Quantify evidence for an eQTL in configuration $\gamma$:
%\bigskip
%$\text{BF}_\gamma = \frac{\mathsf{P}(\text{data } | \text{ eQTL in configuration }\gamma)}{\mathsf{P}(\text{data } | \text{ no eQTL in any tissue})}$
%\pause
%\bigskip
%\bigskip
%\item Estimate configuration proportions $\eta_\gamma$ by borrowing information across genes (pooling).
%\end{itemize}
%\end{frame}

%\begin{frame}
%\frametitle{Tools to fulfill the two goals}
%\begin{itemize}
%\item Detect eQTLs by Bayesian Model Averaging (BMA):
%\bigskip
%$\text{BMA} = \sum_\gamma \eta_\gamma \text{BF}_\gamma$
%\pause
%\bigskip
%\bigskip
%\item Identify in which tissue(s) they are active via their posterior:
%\bigskip
%\small{$\mathsf{P}(\text{SNP is in configuration }\gamma \; | \text{ data, SNP is eQTL}) = \frac{\eta_\gamma \text{BF}_\gamma}{\sum_\gamma \eta_\gamma \text{BF}_\gamma}$}
%\end{itemize}
%\end{frame}

%-----------------------------------------------------------------------------

%\section{Simulations}

%-----------------------------------------------------------------------------

%\section{Analysis of three cell types from Dimas \textit{et al.}}

\begin{frame}
\frametitle{Example 2: eQTL studies across multiple cell-types}

Dimas et al (2009), measured expression data in 75 individuals, in 3 cell types: Fibroblasts, LCLs and T-cells.

\bigskip

They identified eQTLs separately in each cell-type, and found small overlap of results (14\%).

\bigskip

{\bf Question:} Incomplete power may cause this analysis to underestimate sharing; does a joint analysis of all cell types
 come to the same conclusion?

\end{frame}

\begin{frame}
\frametitle{Joint Analysis Increases Power}
\begin{center}
  \begin{tikzpicture}
    \node[anchor=south west,inner sep=0] (image) at (0,0) {\includegraphics[width=\textwidth,height=\textheight,keepaspectratio=true]{figures/box-forest_weak-consistent_ENSG00000090924-rs755690}};
  \end{tikzpicture}
\end{center}
%\begin{center}
%  \small{Example of gene ENSG00000090924 and SNP rs755690.}
%\end{center}
\end{frame}

\begin{frame}
\frametitle{Gain in power from the joint analysis}
%\only<1>{
%\begin{center}
%%\includegraphics[width=0.95\textwidth,height=0.80\textheight,keepaspectratio=true,clip,trim=0cm 0cm 5.6cm 0cm]{dimas}%
%  \begin{tikzpicture}
%    \node[anchor=south west,inner sep=0] (image) at (0,0) {\includegraphics[width=0.95\textwidth,height=0.80\textheight,keepaspectratio=true,clip,trim=0cm 0cm 5.6cm 0cm]{figures/dimas}};
%    \begin{scope}[x={(image.south east)},y={(image.north west)}]
%      % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
%      % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
%      % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y}; }
%      \draw[red,ultra thick,rounded corners] (0.19,0.64) rectangle (0.44,0.81);
%      \draw[red,ultra thick,rounded corners] (0.68,0.64) rectangle (0.94,0.81);
%    \end{scope}
%  \end{tikzpicture}
%\end{center}
%}
%\only<2>{
\begin{center}
\includegraphics[width=0.95\textwidth,height=0.80\textheight,keepaspectratio=true,clip,trim=11.6cm 0cm 0cm 0cm]{figures/dimas}%
\end{center}
%}
\end{frame}

\begin{frame}
\frametitle{Joint analysis suggests much more sharing of eQTLs}
%\begin{table}[!ht]
%\begin{tabular}{|c|c|c|}
%\hline
%Configuration & Hierarchical model & Tissue-by-tissue \\
%\hline
%F-L-T & \alert<2-3>{0.865} [0.807, 0.983] & \alert<2-3>{0.142} \\
%L-T & \alert<4>{0.046} [0.003, 0.104] & 0.056 \\
%F-L & 0.005 [0.000, 0.028] & 0.099 \\
%F-T & 0.001 [0.000, 0.016] & 0.056 \\
%F & \alert<3>{0.025} [0.000, 0.070] & \alert<3>{0.253} \\
%L & \alert<3>{0.051} [0.000, 0.115] & \alert<3>{0.265} \\
%T & \alert<3>{0.007} [0.000, 0.033] & \alert<3>{0.130} \\
%\hline
%\end{tabular}
%\begin{flushleft}\small{The results for the hierarchical model were obtained with the correlated Bayes Factors and the EM algorithm. The results for the tissue-by-tissue analysis were obtained by fixing an FDR of 0.05 in each tissue.}\end{flushleft}
%\end{table}
%\only<5>{
%\alert{Tissue-by-tissue analysis} doesn't account for different power between tissues $\Rightarrow$ likely to \alert{over-estimate tissue specificity}.
%}
\begin{center}
%\includegraphics[width=0.95\textwidth,height=0.80\textheight,keepaspectratio=true]{venn-diag_config-props_sep-vs-join}%
\begin{tikzpicture}
  \node[anchor=south west,inner sep=0] (image) at (0,0) {\includegraphics[width=0.95\textwidth,height=0.80\textheight,keepaspectratio=true]{figures/venn-diag_config-props_sep-vs-join}};
  \begin{scope}[x={(image.south east)},y={(image.north west)}]
    % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
    % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
    % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y}; }
    \only<1>{
      \draw[red,ultra thick,rounded corners] (0.43,0.49) rectangle (0.57,0.6);
    }
%    \only<2>{
%      \draw[red,ultra thick,rounded corners] (0.14,0.67) rectangle (0.27,0.78);
%      \draw[red,ultra thick,rounded corners] (0.7,0.67) rectangle (0.84,0.78);
%      \draw[red,ultra thick,rounded corners] (0.43,0.12) rectangle (0.57,0.23);
%    }
    % \only<3>{
    %   \draw[red,ultra thick,rounded corners] (0.6,0.40) rectangle (0.7,0.51);
    % }
  \end{scope}
\end{tikzpicture}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Wrong tissue-specific call by the tissue-by-tissue analysis}
\only<1>{
  \begin{center}
    \begin{tikzpicture}
      \node[anchor=south west,inner sep=0] (image) at (0,0) {\includegraphics[width=\textwidth,height=\textheight,keepaspectratio=true]{figures/box-forest_wrong-specific_ENSG00000106153-rs4948093}};
      \begin{scope}[x={(image.south east)},y={(image.north west)}]
        % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
        % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
        % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y}; }

        %\only<2>{
          \draw[red,ultra thick,rounded corners] (0.77,0.65) rectangle (0.85,0.72);
                  \draw[red,ultra thick,rounded corners] (0.77,0.47) rectangle (0.85,0.54);
            \draw[red,ultra thick,rounded corners] (0.77,0.28) rectangle (0.85,0.35);
%              \draw[red,ultra thick,rounded corners] (0.77,0.65) rectangle (0.85,0.52);
          %\draw[red,ultra thick,rounded corners] (0.71,0.91) rectangle (0.92,0.99);
        %}
      \end{scope}
    \end{tikzpicture}
  \end{center}

  \begin{center}
    \small{Example of gene ENSG00000106153 and SNP rs4948093 (MAF=0.23).}

    \small{See also Ding \textit{et al.} (2010, AJHG).}
  \end{center}
}
\end{frame}

%-----------------------------------------------------------------------------

%\section{Conclusions and perspectives}

\begin{frame}[fragile]
\frametitle{The next challenge - more subgroups!}
\begin{itemize}
\item This ``configuration-based" framework can deal satisfactorily with, perhaps, 6-10 subgroups.
\item The NIH GTEX project is currently collecting data on upwards of 20 tissues.
\item More generally, in genomics, one might have hundreds of  ``observations" on each unit...
\item ... and, potentially, relevant covariates.
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{The next challenge - a general framework for data integration?}

\begin{itemize}
\item Summarize the ``data" on each SNP in each subgroup (or experimental condition) as $(\hat\beta, se(\hat\beta))$.
\item Arrange these in a big $p$ by $S$ matrix.
\item Goal: identify the elements that correspond to non-zero (or ``large") $\beta$, exploiting combined structure
across rows of the matrix.
\end{itemize}

Challenge: exploit the many available tools -- clustering, PCA, factor analysis,etc -- to do this in a flexible and powerful way.

\end{frame}

\begin{frame}
\frametitle{Acknowledgments}

\begin{itemize}
\item  Xiaoquan Wen,  Timoth\'ee Flutre, Jonathan Pritchard. 
\item Global Lipids Consortium for making data available
\item Manoulis Dermitzakis for expression data.
\item Funding: NHGRI and NIH GTEX consortium.
\end{itemize}

Selected References:
\begin{itemize}
\item Wen \& Stephens (2011, arXiv), Wen (2012, arXiv), Flutre et al (2013, PloS Genetics).
\item Han \& Eskin (2011, AJHG; 2012, PloS Genetics).
\item Ding \textit{et al.} (2010, AJHG).
\item Lebrec \textit{et al.} (2010, SAGMB).
\end{itemize}
\url{http://stephenslab.uchicago.edu/publications.html}
\end{frame}

\begin{frame}
\frametitle{Gain in power from the joint analysis}
\only<1>{
\begin{center}
%\includegraphics[width=0.95\textwidth,height=0.80\textheight,keepaspectratio=true,clip,trim=0cm 0cm 5.6cm 0cm]{dimas}%
  \begin{tikzpicture}
    \node[anchor=south west,inner sep=0] (image) at (0,0) {\includegraphics[width=0.95\textwidth,height=0.80\textheight,keepaspectratio=true,clip,trim=0cm 0cm 5.6cm 0cm]{figures/dimas}};
    \begin{scope}[x={(image.south east)},y={(image.north west)}]
      % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
      % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
      % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y}; }
      \draw[red,ultra thick,rounded corners] (0.19,0.64) rectangle (0.44,0.81);
      \draw[red,ultra thick,rounded corners] (0.68,0.64) rectangle (0.94,0.81);
    \end{scope}
  \end{tikzpicture}
\end{center}
}
\only<2>{
\begin{center}
\includegraphics[width=0.95\textwidth,height=0.80\textheight,keepaspectratio=true,clip,trim=11.6cm 0cm 0cm 0cm]{figures/dimas}%
\end{center}
}
\end{frame}

%  \begin{frame}{Why Jointly Analyze?}
%\begin{enumerate}
%\item Power gains if eQTLs are shared among tissues.
%\item Without a joint analysis, very hard to assess tissue specificity due to lack of power.
% \end{enumerate}
% \end{frame}
% 
%

% Similar, but different, issues, arise when analyzing eQTLs in multiple tissues

%\begin{frame}{LDL Subfractions}  
%\begin{center}
%\includegraphics[width=2in,height=2in]{subfrac_top3_av.pdf}
%\end{center}
%\end{frame}

\begin{frame}{Where next?}
\begin{itemize}
\item Larger-scale problems (e.g.~GTEx collecting data on 30 tissues)
%\item More complex outcome structures (e.g. small time series).
\item Multi-SNP multi-phenotype? (e.g.~Verzilli et al (2005); Banerjee et al (2008)).
%\item Clustering problems generally (Genotype unobserved).
\item Dealing with non-normality; Outliers; Binary outcome with intermediate quantitative phenotypes.
\item ``Response" phenotypes. (Maranville et al, PloS Genetics, 2011).
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Some eQTLs may be shared; others tissue-specific}
\only<1>{
\begin{center}
\includegraphics[width=0.95\textwidth,height=0.8\textheight,keepaspectratio=true]{figures/ex_config11_var0-5_boxplot}%
\end{center}
}
\only<2>{
\begin{center}
\includegraphics[width=0.95\textwidth,height=0.8\textheight,keepaspectratio=true]{figures/ex_config10_var0-5_boxplot}%
\end{center}
}
\only<3>{
\begin{center}
\includegraphics[width=0.95\textwidth,height=0.8\textheight,keepaspectratio=true]{figures/ex_config00_var0-5_boxplot}%
\end{center}
}
\end{frame}

\end{document}

### 2013-05-13 Rcode/ash.R

@@ -1,0 +1,391 @@
+#return ABF for vector of betahat and standard errors
+ABF = function(betahat, sebetahat,sigmaa){
+  T = betahat/sebetahat
+  lambda = sebetahat^2/(sebetahat^2+sigmaa^2)
+  return((sqrt(lambda) * exp(0.5*T^2 *(1-lambda))))
+}
+
+logABF = function(betahat,sebetahat,sigmaa){
+  T = betahat/sebetahat
+  lambda = sebetahat^2/(sebetahat^2+sigmaa^2)
+  return(0.5*log(lambda) + 0.5*T^2 *(1-lambda))
+# the following line is same up to a constant, and probably faster:
+# return(dnorm(betahat,0,sqrt(sebetahat^2+sigmaa^2),log=TRUE))
+}
+
+#compute normal density for n-vector x
+#at each of k values of mu and sigma
+#OUTPUT: k by n matrix of normal densities
+matdnorm = function (x, mu, sigma) 
+{
+  k=length(mu)
+  n=length(x)
+  d = matrix(rep(x,rep(k,n)),nrow=k)
+  return(matrix(dnorm(d, mu, sigma),nrow=k))
+}
+
+#compute mixture log likelihood for data x, for mixture of k normals
+#INPUT: x an n vector of data
+#mixture component parameters pi, mu, sigma, each being k vectors
+#
+mixLoglik = function(x,pi,mu,sigma){
+  dd=matdnorm(x,mu,sigma) 
+  return(sum(log(pi %*% dd)))
+}
+
+#compute mixture log likelihood for data x, for mixture of k normals
+#plus observation-specific errors with observation-specific variances
+#ie x_s is sum_k pi_k N(x_s; mu_k, sigma^2_k + sigma^2_s)
+# [optionally, use sigma^2_k * sigma^2_s if FUN="*"]
+#INPUT: x an n vector of data
+#mixture component parameters pi, mu, sigma, each being k vectors
+#se, the observation-specific standard errrors
+mixseLoglik = function(x,pi,mu,sigma,se,FUN="+"){
+  k=length(mu)
+  n=length(x)
+  d = matrix(rep(x,rep(k,n)),nrow=k)
+  s = sqrt(outer(sigma^2,se^2, FUN)) # k by n matrix of standard errors
+  dd = matrix(dnorm(d, mu, s),nrow=k) 
+  return(sum(log(pi %*% dd)))
[…]

### 2013-05-13 Rcode/cash (Scott's stuff)/cash.R

@@ -1,0 +1,108 @@
+library(ordinal)
+library(VGAM)
+library(catdata)
+source('plum.R')
+source('../ash.R')
+
+# EXPECTATION STEP
+expectation=function(X,likely,theta,alpha){
+	n=length(beta)
+	linear=matrix(rep(theta,n),n,byrow=T)-colSums(t(X)*alpha)
+	priorp=t(apply(cbind(rep(0,n),exp(linear)/(1+exp(linear)),rep(1,n)),1,diff))
+	margin=sum(log(colSums(likely*priorp)))
+	print(margin)
+	poster=likely*priorp/rowSums(likely*priorp)
+	return(poster)}
+
+# MAXIMIZATION STEP (CLM)
+maximization.clm=function(poster,X){
+	K=dim(poster)[2]
+	n=dim(X)[1]
+	p=dim(X)[2]
+	output=ordered(rep(1:K,each=n))
+	weight=array(poster)
+	design=kronecker(array(1,K),X)
+	fitted=clm(output~design,weights=weight)$coe
+	theta=fitted[1:(K-1)]
+	alpha=fitted[K:(K+p-1)]
+	return(list(theta=theta,alpha=alpha))}
+
+# MAXIMIZATION STEP (PLUM)
+maximization.plum=function(poster,X){
+	K=dim(poster)[2]
+	p=dim(X)[2]
+	fitted=plum(poster,X)$coef
+	theta=fitted[1:(K-1)]
+	alpha=-fitted[K:(K+p-1)]
+	return(list(theta=theta,alpha=alpha))}
+	
+# MAXIMIZATION STEP (POLR) ?
+	
+# MAXIMIZATION STEP (VGLM)
+maximization.vglm=function(poster,X){
+	K=dim(poster)[2]
+	n=dim(X)[1]
+	p=dim(X)[2]
+	fitted=Coef.vlm(vglm(round(1e16*poster)~X,family=propodds(F)))
+	theta=fitted[1:(K-1)]
+	alpha=-fitted[K:(K+p-1)]
+	return(list(theta=theta,alpha=alpha))}
+	
+# CASH
+# GIVEN: ESTIMATED EFFECTS (BETA), STANDARD ERRORS (S), COVARIATE MATRIX (X) AND METHOD
+# beta: n-vector
+# s: n-vector
+# X: n-by-p matrix
[…]

### 2013-05-13 Rcode/cash (Scott's stuff)/plum.R

@@ -1,0 +1,92 @@
+### glm p.179
+#data <- c(98, 0, 0, 51, 2, 1, 34, 6, 3, 35, 5, 8,
+#	32, 10, 9, 23, 7, 8, 12, 6, 10, 4, 2,5)
+#yrs <- c(5.8, 15.0, 21.5, 27.5, 33.5, 39.5, 46.0, 51.5)
+#n <- 8
+#k <- 3
+#data <- matrix(data, n, k, byrow=T)
+#fit1 <- plum(data, log(yrs))
+#fit2 <- plum(data, cbind(yrs,log(yrs)))
+#fit3 <- plum(data, log(yrs), link="c-log log")
+###%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+#nrows <- 4
+#ncols <- 9
+#freq <- scan("http://www.stat.uchicago.edu/~pmcc/glm/GLM175.txt")
+#data <- matrix(freq, ncol=ncols, byrow=T)
+#cheese <- as.factor(1:4)
+#fit0 <- plum(data)
+#fit1 <- plum(data, cheese)
+##%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+plum <- function(data, X, link="logit", maxcyc=15, tol=1.0e-4){
+#
+# fits the proportional odds model or related model
+# the matrix X should be of full rank, as should the augmented matrix 1X
+#
+if(is.factor(data)) data <- model.matrix(~data-1)
+n <- dim(data)[1]
+k <- dim(data)[2]
+nk <- n*k
+nk1 <- nk - n
+p <- 0
+if(missing(X)) X <- matrix(rep(1, n), n, 1)
+if(!missing(X)){
+	if(is.factor(X)) X <- model.matrix(~X)[, 2:length(levels(X))]
+	if(is.vector(X)) X <- matrix(X, n, 1)
+	}
+qr <- qr(cbind(rep(1, n), X))
+X <- X[, qr$pivot[2:qr$rank] - 1]
+p <- qr$rank - 1 
+m <- data %*% rep(1, k)
+tot <- t(rep(1, n)) %*% data
+ctot <- cumsum(tot)[1:(k-1)]/sum(m)
+# initial values
+beta <- c(log(ctot/(1-ctot)), rep(0, p))
[…]

### 2013-05-14 talks/ilike-slides.Rmd
final draft of ilike slides; added papers

@@
-% Apaptive Shrinkage and False Discovery Rates by Laplace Approximation
+% Adaptive Shrinkage and False Discovery Rates by Laplace Approximation
 % Matthew Stephens
@@
 
-# Motivation
+# Outline
 
-- A common problem: you have imperfect measurements
+* Prelude
+* Allegro (ma non troppo)
+
+# Prelude
+
+- Consider testing the null hypothesis $H_0: \beta=0$, vs the alternative $H_1: \beta \neq 0$ in the
+logistic regression model:
+$$\log \frac{p(Y_i=1|X_i=x)}{p(Y_i=0|X_i=x)}  = \mu + x \beta$$
+
+- Specifically, consider computing the Bayes Factor
+$$BF:=\frac{p(Y | X, H_1)}{p(Y| X ,H_0)}.$$
+
+- In genome-wide association studies, we may wish to do this 
+for millions of different genetic variants ($X$).
+
+# Prelude
+
+$$BF= \frac{\int p(Y | \mu, \beta, X) p_1(\mu,\beta | X) \, d\mu d\beta}{\int p(Y | \mu, \beta=0, X) p_0(\mu | X) \, d\mu},$$
+where $p_0$ and $p_1$ denote priors under $H_0$ and $H_1$.
+
+- These integrals generally don't have closed forms, but being 
+low-dimensional they are simple to approximate.
+
+- For $p_1: \beta \sim N(0,\phi^2)$, 
+Wakefield, 2009 (see also Johnson, 2008) suggested a particularly simple \emph{Approximate Bayes Factor} (ABF)
+based on the maximum likelihood estimate, $\hat\beta$, and 
+its (estimated) standard error $s$.
+
+# Prelude
+
+$$ABF = \sqrt{1-k} \exp(0.5 k T^2)$$
+where $k:= \phi^2/(s^2 + \phi^2)$ and $T:= \hat\beta/s$.
+
+- ABF arises if we assume $\hat{\beta} | s, \beta \sim N(\beta, s^2)$ and treat $\hat{\beta}$ as the observed ``data".
+
+- Equivalently ABF can be derived as a ``Laplace approximation",
+approximating the likelihood
+$L(\beta)$ as Normal, centered on $\hat\beta$,
+with variance $s^2$: $$L(\beta) \propto \exp[- 0.5(\beta-\hat\beta)^2/s^2].$$
+
+# Comments on ABF
+
+* This is not the moxt accurate Laplace approximation one might consider.
+
+* However, it has some nice features.  
+    * The approximation is independent of prior.
+    * Applicable to any regression where $\hat\beta$ and $s$ are available.
+    * Easily computed using results of standard software or published analyses (e.g. CI).
+
+* A simple transformation of $T$ can improve accuracy for small samples (analagous to $t$ test vs $Z$ test); Wen and Stephens, Arxiv.
+
+# Extensions of ABF
+
+- Similar ideas can be used to compute 
+ABFs in slightly more complex settings.
+
+- Eg In Wen and Stephens, we consider
+$S$ subgroups, and approximate the BF for
+$H_0: \beta_s =0$ for all $s$, vs a general alternative
+$H_0: \beta_s \neq 0$.
+
+# Allegro (ma non troppo)
+
+- The problem: you have imperfect measurements
 of many ``similar" things, and wish to estimate their values.
@@
 
-```{r, include=FALSE}
-setwd("~/Documents/git/BayesFDR/talks/")
+```{r, echo=FALSE}
+setwd("~/Documents/git/ash/talks/")
 ## load Poisson_binomial and ash functions 
@@
 source("../Rcode/ash.R")
+
 x = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
+xx = rowSums(x[,2:5])
+x = x[xx>0,]
+xx = xx[xx>0]
 ```
@@
 
-# Example: Mouse Heart Data
-```{r}
+```{r echo=FALSE}
 head(x)
@@
 ```{r, echo=FALSE}
+
 cc = x[, 2:5]
+
 g = c(-1, -1, 1, 1)
@@
 pval = ttest.pval(tscore, df = 2)
+qval = qvalue(pval)
+
+highxp = xx>1000 # select high expressed genes
+pval.high = pval[highxp]
+qval.high = qvalue(pval.high)
+cc.assoc.high = counts.associate(cc[highxp,],g,1)
+zdat.ash.high = cc.assoc.high$zdat.ash
+zdat.high = cc.assoc.high$zdat
 
@@
 
-# Borrowing Strength
+# False Discovery Rate analyses
 
-- For gene $j$, denote the difference in mean between condition 0 and condition 1 by $$\beta_j:= \mu^0_j - \mu^1_j.$$
+- Standard practice: analyses use False Discovery Rates
 
-- As measurements are made on only two
-samples in each condition, error in estimates
-of $\beta_j$ are appreciable.
+    - e.g. Benjamini and Hochberg, 1995; Storey and Tibshirani, 2003, which have roughly 18k and 4k citations respectively!
 
-- Fundamental idea: measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.
+- Typical analysis proceeds roughly as follows:
+
+    - Estimate an effect size $\beta_j$ and standard error $s_j$ for each gene.
+
+    - Convert this to a $p$ value for each gene, e.g. by a $t$ test on $\beta_j/s_j$.
+
+    - Use the distribution of $p$ values to estimate the false discovery rate (FDR) at a given threshold.
 
@@
 
--Standard practice is to estimate an effect size $\beta_j$ and standard error $s_j$ for each gene.
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+hist(pval,prob=TRUE,main="p value distribution, all genes",ylim=c(0,4),xlab="p value")
+abline(h=qval$pi0,col="red",lwd=2)
 
--Then convert this to a $p$ value for each gene, e.g. by a $t$ test on $\beta_j/s_j$.
-
--Then use the distribution of $p$ values to estimate the false discovery rate (FDR) at a given threshold.
-
-# False Discovery Rates
-
-```{r, echo=FALSE}
-hist(pval,prob=TRUE)
-qval = qvalue(pval)
-abline(h=qval$pi0,col="red",lwd=2)
 ```
@@
 
-```{r, echo=FALSE}
-h=hist(pval,prob=TRUE)
+```{r, echo=FALSE,fig.height=4,fig.cap="FDR=0.21"}
+h=hist(pval,prob=TRUE,main="p value distribution, all genes",ylim=c(0,4),xlab="p value")
 qval = qvalue(pval)
@@
 
-#FDR methods are very widely used!
+# FDR problem: different genes have different precision/power
 
-- Why?
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+hist(log10(xx),main="Counts vary considerably across genes", xlab="log10(counts)")
+```
 
-- Partly, they make sense.
+# FDR problem: lower count genes, less power, add noise
 
-- Partly, they are generic. Once you have the $p$ values, you don't really care where they came from.
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+h=hist(pval[xx<1000],prob=TRUE,ylim=c(0,4),main="p values, low count genes",xlab="p value",breaks=seq(0,1,length=21))
+qval.low=qvalue(pval[xx<1000])
+abline(h=qval.low$pi0,col="red",lwd=2)
+rect(0,0,0.05,qval.low$pi0,col="red")
+abline(v=0.05,lwd=3)
+```
 
-#Issues
+# FDR problem: higher count genes, more power
+```{r, echo=FALSE, fig.height=4,fig.cap="FDR=0.10"}
 
-- Different precision of different measurements 
-is forgotten once the $p$ value is computed.
+h=hist(pval.high,prob=TRUE,ylim=c(0,4),main="p values, high count genes",xlab="p value")
 
-- Borrows information for testing, not for estimation.
+abline(h=qval.high$pi0,col="red",lwd=2)
+rect(0,0,0.05,qval.high$pi0,col="red")
+rect(0,qval.high$pi0,0.05,h$intensities[1],col="green")
+abline(v=0.05,lwd=3)
+#qval.high$q[pval.high<0.05] gives FDR for p<0.05
+```
 
-#An Alternative
+# Adaptive Shrinkage
 
-- Instead of summarizing each gene by a single number ($p$ value)
-summarize by two numbers, $\hat\beta_j$ and its standard error $\hat\s_j$.
+- Fundamental idea: use hierarchical modelling so measurements of $\beta_j$ for each gene improve inference for $\beta$ at other genes.
 
-- Then do inference for $\beta_j$ by Laplace approximation.
+- Despite a long-standing literature on these types of methods - e.g. Greenland and Robins 1991, Efron and Tibshirani 2002, Gelman et al 2012 -  they are much less widely used (in genomics at least).
 
-# Example: Wavelets
+- Possibly this is due, in part, to the lack of a simple, flexible, and generic implementation?
+
+# Generic adaptive shrinkage via Laplace approximation
+
+- Summarize data on each gene by two numbers, $\hat\beta_j$ and its standard error $s_j$. (a la Wakefield; Greenland and Robins 1991)
+
+- Approximate likelihood for $\beta_j$ by 
+$$L(\beta_j) \propto \exp(-0.5 (\beta_j - \hat\beta_j)^2/s_j^2).$$
+(``Laplace Approximation")
+
+- Borrow information by assuming $\beta_j$ are iid $\sim g(\cdot; \pi)$, where $\pi$ are hyperparameters to be estimated.
+
+- Letting $g(\cdot; \pi)$ be a mixture of normal distributions provides
+both flexibility, and analytic calculations.
+
+    - very small variances can capture effects that are ``effectively" zero.
+
+# An important special case
+
+- Focus on the special case where $g(\cdot; \pi)$ can be
+assumed unimodal and symmetric about zero.
+
+- Then the posterior mean, $E(\beta_j | \hat\beta, s, \hat\pi)$ is a ``shrinkage" estimate of $\beta_j$.
+
+- And $p(\beta_j > 0 | \hat\beta, s, \hat\pi)$ can be used
+to identify $j$ for which the sign of $\beta_j$ can be confidently determined (analogous to test of $\beta_j =0$; Gelman et al, 2012).
+
+- Because $\pi$ is estimated from the data, the amount
+of shrinkage is adaptive to the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each gene.
+
+# Example: ASH applied to mouse data
+
+```{r fig.height=4,echo=FALSE,fig.cap=""}
+hist(zdat[3,],main="Raw effect size estimates",xlab="betahat",xlim=c(-2,2),prob=T,ylim=c(0,3))
+```
+
+# Example: ASH applied to mouse data
+
+```{r fig.height=4,echo=FALSE,fig.cap=""}
+hist(zdat[3,],main="Raw effect size estimates",xlab="betahat",xlim=c(-2,2),prob=T,ylim=c(0,3))
+t=seq(-2,2,length=201)
+lines(t,dnorm(t,0,sd=0.032), col=2, lwd=2)
+```
+
+# Example: ASH applied to mouse data
+
+```{r fig.height=4,echo=FALSE,fig.cap=""}
+hist(zdat.ash$PosteriorMean,main="Shrunken estimates",xlab="shrunk betahat",xlim=c(-0.1,0.1),prob=T)
+lines(t,dnorm(t,0,sd=0.032), col=2, lwd=2)
+```
+
+# Shrinkage is adaptive to information
+
+```{r, echo=FALSE,fig.height=4}
+#hist(zdat[3,])
+#plot(zdat[3,],zdat.ash$PosteriorMean,xlim=c(-0.2,0.2))
+#points(zdat[3,16677],zdat.ash$PosteriorMean[16677],col=2,pch=16)
+#points(zdat[3,16079],zdat.ash$PosteriorMean[16079],col=2,pch=16)
+#x[16677,]
+#x[16079,]
+```
+
+```{r, echo=FALSE}
+#temp = counts.associate(cc,c(-1,1,-1,1),1)
+#temp.tscore = temp$zdat[3,]/temp$zdat[4,]
+#temp.pval = ttest.pval(temp.tscore,df=2)
+#plot(temp$zdat.ash$localfdr,temp.pval)
+#identify(temp$zdat.ash$localfdr,temp.pval)
+```
+
+```{r, echo=FALSE,fig.height=3,fig.width=3,fig.cap=""}
+plot(zdat.ash$localfdr,pval,ylab="p value", xlab="ASH local fdr")
+```
+
+# Shrinkage is adaptive to information
+
+```{r, echo=FALSE,fig.height=3,fig.width=3,fig.cap=""}
+plot(zdat.ash$localfdr,pval,ylab="p value", xlab="ASH local fdr")
+points(zdat.ash$localfdr[15325],pval[15325],col=2,pch=16)
+points(zdat.ash$localfdr[16123],pval[16123],col=2,pch=16)
+```
+
+# Shrinkage is adaptive to information
+
+```{r, echo=FALSE}
+cbind(x[,1:5],pval,zdat.ash$localfdr)[c(15325,16123),]
+```
+
+# Summary: FDR vs ASH
+
+- Both provide a rational approach to identifying ``significant" findings.
+
+- Both are generic and modular: once you have the summary data, you can forget where they came from.
+
+- But by using two numbers ($\hat\beta,s$) instead of one ($p$ values) precision of different measurements 
+can be better accounted for.
+
+- ASH borrows information for estimation, as well as testing.
+
+# Next steps?
+
+- Extend to allow $g(\cdot;\pi)$ to depend on covariates $X$.
+
+- Extend to allow for correlations in the measured $\hat\beta_j$.
+
+# Thanks
+
+- to the several postdoctoral researchers and students
+who have worked with me on related topics.
+
+- Especially: William Wen, Timothee Flutre, Scott Powers, Heejung Shim, Zhengrong Xing, and Ester Pantaleo.
+
+- And to the NIH for funding, and i-like for inviting me.
 
@@
 
-This document is produced with **knitr**, **Rstudio** and **Pandoc**.
+- This document is produced with **knitr**, **Rstudio** and **Pandoc**.
+
+- For more details see my \tt{stephens999/ash} repository at http://www.github.com/stephens999/ash
+
+- Website: http://stephenslab.uchicago.edu
+
+# Pandoc Command used
+
+pandoc -s -S -i --template=my.beamer -t beamer -V theme:CambridgeUS -V colortheme:beaver  ilike-slides.md -o ilike-slides.pdf
 
@@
 
-# Misc issues
+# FDRs for higher count genes affected by lower count genes
 
-- the plots are too wide? use the chunk option `out.width` which will be used in `<img width=... />`, e.g. `out.width=400px`
+```{r,echo=FALSE,fig.height=4,fig.cap=""}
+plot(qval.high$q, qval$q[xx>1000],main="",xlab="FDR, high count genes", ylab="FDR, all genes")
+```
+
+# Some odd things in the data
+
+```{r, echo=FALSE,fig.height=3,fig.width=3}
+  plot(zdat[3,],zdat[4,],ylab="standard error",xlab="beta-hat")
+  x[tail(order(zdat[3,])),]
+```

### 2013-05-14 papers/PoissssonBinomial_etc02.tex

\documentclass[12pt,letterpaper]{article}

% Page Layout
\usepackage[hmargin={1.5in, 1.5in}, vmargin={1.5in, 1.5in}]{geometry}

\usepackage{rotating}

% Spacing
\usepackage{setspace}
% Use \singlespacing, \onehalfspacing, or \doublespacing,
% or alternatively \setstretch{3} for triple spacing (or any other number).

% Mathematical Notation
\usepackage{amsmath,amstext,amssymb,amsfonts}

% Colors				
\usepackage{xcolor,colortbl}

\renewcommand{\Pr}{\mathsf{P}}
\newcommand{\prob}[1]{\Pr\left(#1\right)}
\newcommand{\given}{\mid}
\newcommand{\me}{\mathrm{e}} % use for base of the natural logarithm
\newcommand{\md}{\mathrm{d}} % use for base of the natural logarithm
\newcommand{\var}{\mathrm{Var}}
\newcommand{\mean}{\mathrm{E}}
\newcommand{\Normal}{\mathcal{N}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\like}{\mathcal{L}}
\newcommand{\loglike}{\mathcal{L}\mathcal{L}}
\newcommand{\eps}{\epsilon}
\newcommand{\nin}{\noindent}

\def\logten{\log_{10}}
\def\BFav{{\rm BF}_\text{av}}
\def\BFuni{{\rm BF}_\text{uni}}
\def\BFldl{{\rm BF}_\text{ldl}}
\def\BF{\rm BF}
\def\ABF{\rm ABF}
\def\BFall{{\rm BF}_\text{all}}

\newcommand{\zuk}[1]{\textcolor{red}{[#1 --Or]}}
\newcommand{\heejung}[1]{\textcolor{blue}{[#1 --Heejung]}}

%\usepackage{alltt}
% for mathematical notation in a verbatim-like environment
% \begin{alltt} ... \end{alltt}

% Graphics
\usepackage{graphicx}
%\usepackage[small]{subfigure}
% for subfigures in a single figure

%\usepackage{epsfig,rotating,epsf,psfrag,lscape}

% Citations
\usepackage{natbib}

% Style for sections and such.
%\usepackage{mbe}
%\usepackage{genetics}

\begin{document}

%\begin{abstract}
%Template for papers.
%In particular, check out the way to automatically create a bibliography.
%\end{abstract}

\section{Model in Ash}
\label{sec:Model_Ash}
The model considered in Ash is for each $i$,
\begin{equation}
x_i \sim \Normal(\alpha_i, y_i^2).
\end{equation}
Note that in a PoissonBinomial model, $x_i = \hat{\alpha}_i$ and $y_i = se(\hat{\alpha}_i)$ are estimate (MLE) for $\alpha_i = logit(p_i)$ and its standard error which are estimated by using a glm function in R. 
Ash considers a mixture of normal distributions as a prior on $\alpha_i$. Specifically, for each $i$,
\begin{equation}
\alpha_i \given \pi, \sigma^2 \sim \sum_{m=1}^M \pi_m \Normal(0, \sigma^2_m),\label{eqn:ash_prior}
\end{equation}
where $\pi = (\pi_1, \ldots, \pi_M)$ are the mixture proportions which are constrained to be non-negative and sum to one, and $\sigma^2 = (\sigma^2_1, \ldots, \sigma^2_M)$ are the variances for each normal distribution. For now, Ash assumes that the Gaussian
 means are zero, that the variances vector $\sigma^2$ is known, and estimates $\pi$ by using an empirical Bayes procedure - that is
 we find the maximum-likelihood estimator $\hat{\pi}_{MLE}$.

\section{EM algorithm in Ash}
\label{sec:EM_Ash}
The MLE for $\pi$ can be obtained by using the following EM algorithm. Let $D_i = (x_i, y_i)$ and $D = (D_1, \ldots, D_n)$. Consider unobserved latent variables $Z = (Z_1, \ldots, Z_n)$, where $Z_i \in \{1, \ldots, M\}$ and $\Pr(Z_i = m) = \pi_m$. Then, a complete data likelihood can be written
\begin{eqnarray}
\Pr(D, Z \given \pi) &=& \prod_{i=1}^{n}\Pr(D_i, Z_i \given \pi)\\
&=& \prod_{i=1}^{n} \prod_{m=1}^M\Pr(D_i, Z_i = m \given \pi)^{I(Z_i = m)}\\
&=& \prod_{i=1}^{n} \prod_{m=1}^M[\Pr(D_i \given Z_i = m, \pi)\pi_m] ^{I(Z_i = m)},
\end{eqnarray}
yielding a log likelihood
\begin{eqnarray}
\log\Pr(D, Z \given \pi) &=& \sum_{i=1}^{n} \sum_{m=1}^M I(Z_i = m)\Big[\log{\Pr(D_i \given Z_i = m)} + \log{\pi_m}\Big].
\end{eqnarray}

We denote by $\pi^{l}$ the vector of probabilities at step $l$ of the EM algorithm. In each step, we update the vector, i.e.
compute in the $l$-th step $\pi^{l+1}$ from $\pi^l$ and the data.

{\bf E-step:} For each $i$ and $m$,
\begin{eqnarray}
A_{im}^l \equiv \Pr(Z_i = m \given D_i, \pi^l) &=& \frac{\Pr(Z_i = m, D_i \given \pi^l)}{\sum_{n=1}^M\Pr(Z_i = n, D_i \given \pi^l)} \\
&=& \frac{\pi_m^l\Pr(D_i \given Z_i = m)}{\sum_{n=1}^M\pi_n^l\Pr(D_i \given Z_i = n)} \\
&=& \frac{\pi_m^l\BF_i(\sigma^2_m)}{\sum_{n=1}^M\pi_n^l\BF_i(\sigma^2_n)},
\end{eqnarray}
where
\begin{eqnarray}
\label{eq:base_factor}
\BF_i(\sigma^2_m) =  \frac{\Pr(D_i \given Z_i = m)}{\Pr(D_i \given \alpha_i = 0)}.
\end{eqnarray}
{\bf M-step:} Find the parameters $\pi$ which maximizes $\mean_{Z \given D, \pi^l}[\log\Pr(D, Z \given \pi)]$.
\begin{eqnarray}
\pi^{l+1} &=& \argmax_{\pi}\mean_{Z \given D, \pi^l}[\log\Pr(D, Z \given \pi)]\\
	       &=& \argmax_{\pi} \sum_{i=1}^{n} \sum_{m=1}^M A_{im}^l[\log{\BF_{im}} + \log{\pi_m}]\\
	       &\equiv& \argmax_{\pi} Q(\pi \given \pi^l),
\end{eqnarray}
\nin where we used the fact that $\log \Pr(D_i \given \alpha_i=0)$ is constant.
%where $A_{im} = \Pr(Z_i = m \given D_i, \pi^l)$ and $\BF_{im} = \BF_i(\sigma^2_m)$.
For each $m = 1, \ldots, M-1$,
\begin{eqnarray}
\frac{\partial Q(\pi \given \pi^l)}{\partial \pi_m} &=& \sum_{i=1}^n[\frac{A_{im}^l}{\pi_m} + \frac{-A_{iM}^l}{\pi_M}]\\
			&=& \sum_{i=1}^n[\frac{A_{im}^l\pi_M -A_{iM}^l\pi_m}{\pi_m\pi_M}]\\
			&=& \frac{\pi_M \sum_{i=1}^nA_{im}^l -\pi_m \sum_{i=1}^nA_{iM}^l}{\pi_m\pi_M},
\end{eqnarray}
%\zuk{I don't understand how do you get the first equation here (14). If you just take the derivative of (12) you get something more complicated }
%\heejung{if you take the derivative of (12) wrt $\pi_m$, only $\pi_m$ and $\pi_M = 1 - (\pi_{1} + \ldots, + \pi_{M-1} )$ will matter. The remaining terms are just constant. You should use the relationship in (17) and (18) to make it simple. If it doesn't explain well, I will explain it when we get together.}
where
\begin{eqnarray}
	A_{iM}^l = 1 - (A_{i1}^l + \ldots, + A_{i(M-1)}^l),\\
	\pi_{M} = 1 - (\pi_{1} + \ldots, + \pi_{M-1}).
\end{eqnarray}
Then,
\begin{eqnarray}
\pi_M \sum_{i=1}^nA_{im}^l = \pi_m \sum_{i=1}^nA_{iM}^l \quad \text{for} \quad m = 1, \ldots, M-1. \label{eqn:temp}
\end{eqnarray}
Summing $M-1$ equations in (\ref{eqn:temp}) leads to
\begin{eqnarray}
	\pi_M [\sum_{i=1}^n\sum_{m=1}^{M-1}A_{im}^l] = (1-\pi_M) \sum_{i=1}^nA_{iM}^l.
\end{eqnarray}
Then,
\begin{eqnarray}
	\pi_M  &=& \frac{\sum_{i=1}^nA_{iM}^l} {\sum_{i=1}^n\sum_{m=1}^{M}A_{im}^l}\\
		  &=& \frac{\sum_{i=1}^nA_{iM}^l}{n},
\end{eqnarray}
and for each $m= 1, \ldots, M-1$,
\begin{eqnarray}
	\pi_m  &=& \frac{\sum_{i=1}^nA_{im}^l}{n}.
\end{eqnarray}

To implement the EM algorithm we need an explicit expression for the Bayes Factor in eq. (\ref{eq:base_factor}).
We use an approximate likelihood approach to approximate this factor, shown in the next two paragraphs.

\subsection{Likelihood approximation in a PoissonBinomial Model}
\label{sec:app_PB_like}
Under a PoissonBinomial model, a log likelihood function for $\alpha_i = logit(p_i)$ can be written as
\begin{eqnarray}
\loglike(\alpha_i) &=& \log{P(D_i \given \alpha_i)}\\
		&=& \log{\prod_j{{n_{ij} \choose x_{ij}} \left(\frac{1}{1+e^{-\alpha_i}}\right)^{x_{ij}} \left(\frac{e^{-\alpha_i}}{1+e^{-\alpha_i}}\right)^{n_{ij} - x_{ij}}}}.
\end{eqnarray}
Taking three elements in Taylor series of $\loglike(\alpha_i)$ about a MLE $\hat{\alpha}_i$, $\loglike(\alpha)$ can be approximated by
\begin{eqnarray}
\loglike(\alpha_i) &\approx& \loglike(\hat{\alpha}_i) + \frac{\loglike''(\hat{\alpha}_i)(\alpha_i - \hat{\alpha}_i)^2}{2}\\
		&\approx& \loglike(\hat{\alpha}_i) - \frac{(\alpha_i - \hat{\alpha}_i)^2}{2se(\hat{\alpha}_i)^2}.
\end{eqnarray}
Then, a likelihood function for $\alpha_i$, $\like(\alpha_i)$ can be approximated by
\begin{eqnarray}
\like(\alpha_i) = exp[\loglike(\alpha_i)] \propto \phi(\alpha_i ; \hat{\alpha}_i, se(\hat{\alpha}_i)^2) %\Normal(\hat{\alpha}_i, se(\hat{\alpha}_i)^2).
\end{eqnarray}

\nin where $\phi$ is the Gaussian density function, $\phi(x ; \mu, \sigma^2) = \frac{1}{\sqrt{2\pi} \sigma} exp [\frac{-(x-\mu)^2}{2\sigma^2}]$.

\subsection{BF and posterior prob approximation in Ash}
\label{sec:BF_post_Ash}
This section describes derivations for the approximate Bayes Factor ($\ABF$) and posterior on $\alpha_i$ when $Z_i = m$ in the prior from equation (\ref{eqn:ash_prior}).
\begin{eqnarray}
\Pr(D_i \given Z_i = m) &=& \int \frac{C(\hat{\alpha}_i, se(\hat{\alpha}_i)^2)}{\sqrt {2\pi se(\hat{\alpha}_i)^2 } }\exp{[-\frac{\left(\alpha_i -  \hat{\alpha}_i\right)^2}{2se(\hat{\alpha}_i)^2}]} \frac{1}{\sqrt {2\pi \sigma^2_m } }\exp{[-\frac{\alpha_i^2}{2\sigma^2_m}]}\mathrm{d}\alpha_i\\
	&=&  \frac{C(\hat{\alpha}_i, se(\hat{\alpha}_i)^2)}{2\pi \sqrt {se(\hat{\alpha}_i)^2  \sigma^2_m }} \int \exp{[-\frac{\sigma^2_m\left(\alpha_i -  \hat{\alpha}_i\right)^2 + se(\hat{\alpha}_i)^2\alpha_i^2}{2se(\hat{\alpha}_i)^2\sigma^2_m}]}\mathrm{d}\alpha_i\\
	&=&  \frac{C(\hat{\alpha}_i, se(\hat{\alpha}_i)^2)}{2\pi \sqrt {se(\hat{\alpha}_i)^2  \sigma^2_m }} \int \exp{[-\frac{ \left(\sigma^2_m + se(\hat{\alpha}_i)^2\right)\alpha_i^2 - 2\sigma^2_m \hat{\alpha}_i\alpha_i + \sigma^2_m\hat{\alpha}^2_i}{2se(\hat{\alpha}_i)^2\sigma^2_m}]}\mathrm{d}\alpha_i\\
	&=&  \frac{C(\hat{\alpha}_i, se(\hat{\alpha}_i)^2)}{2\pi \sqrt {se(\hat{\alpha}_i)^2  \sigma^2_m }} \int \exp{[-\frac{\left(\sigma^2_m + se(\hat{\alpha}_i)^2\right)[\alpha_i- \frac{\sigma^2_m \hat{\alpha}_i}{\sigma^2_m + se(\hat{\alpha}_i)^2}]^2 + \frac{\sigma^2_m \hat{\alpha}^2_i se(\hat{\alpha}_i)^2}{\sigma^2_m + se(\hat{\alpha}_i)^2}}{2se(\hat{\alpha}_i)^2\sigma^2_m}]}\mathrm{d}\alpha_i\\
	&=&  \frac{C(\hat{\alpha}_i, se(\hat{\alpha}_i)^2)\sqrt{2\pi \frac{\sigma^2_m se(\hat{\alpha}_i)^2}{\sigma^2_m + se(\hat{\alpha}_i)^2}}}{2\pi \sqrt {se(\hat{\alpha}_i)^2  \sigma^2_m }} \exp{[-\frac{\hat{\alpha}^2_i}{2(se(\hat{\alpha}_i)^2+\sigma^2_m)}]}\\
	&=&  \frac{C(\hat{\alpha}_i, se(\hat{\alpha}_i)^2)}{\sqrt{2\pi (\sigma^2_m + se(\hat{\alpha}_i)^2)}} \exp{[-\frac{\hat{\alpha}^2_i}{2(se(\hat{\alpha}_i)^2+\sigma^2_m)}]},
\end{eqnarray}
where $C(\hat{\alpha}_i, se(\hat{\alpha}_i)^2)$ is a constant depending on $\hat{\alpha}_i, se(\hat{\alpha}_i)^2$ but does not 
depend on $\alpha_i$. In fact $C(\hat{\alpha}_i, se(\hat{\alpha}_i)^2) = e^{\loglike(\hat{\alpha}_i)} \sqrt{2\pi se(\hat{\alpha}_i)^2}$.

%\zuk{Why do you need the constant C? isn't it simply 1?}
%\heejung{Actually $\like(\alpha_i) \propto  \phi(\alpha_i ; \hat{\alpha}_i, se(\hat{\alpha}_i)^2)$. $C$ contains all constant ($C \propto \exp({f(\hat{\alpha}_i)})$) and will be canceled out when you take a ratio for BF calculation.}
and
\begin{eqnarray}
\Pr(D_i \given \alpha_i = 0) &=& \frac{C(\hat{\alpha}_i, se(\hat{\alpha}_i)^2)}{\sqrt {2\pi se(\hat{\alpha}_i)^2 } }\exp{[-\frac{\hat{\alpha}_i^2}{2se(\hat{\alpha}_i)^2}]}.
\end{eqnarray}
Then, $\ABF$ can be written as
\begin{eqnarray}
\BF_i(\sigma^2_m) &=&  \frac{\Pr(D_i \given Z_i = m)}{\Pr(D_i \given \alpha_i = 0)}\\
			       &=&  \sqrt{\frac{se(\hat{\alpha}_i)^2}{\sigma^2_m + se(\hat{\alpha}_i)^2}}  \exp{[\frac{\hat{\alpha}_i^2}{2se(\hat{\alpha}_i)^2} \frac{\sigma^2_m}{\sigma^2_m + se(\hat{\alpha}_i)^2}]}\\
			       &=& \sqrt{\lambda} \exp{[T^2 (1-\lambda)/2]},
\end{eqnarray}

where
\begin{eqnarray}
\lambda &=& \frac{se(\hat{\alpha}_i)^2}{se(\hat{\alpha}_i)^2 + \sigma^2_m} ,\\
T &=&  \frac{\hat{\alpha}_i}{se(\hat{\alpha}_i)}.
\end{eqnarray}
And a posterior on $\alpha_i$ is
\begin{eqnarray}
\Pr(\alpha_i \given D_i, Z_i = m) &\propto& \like(\alpha_i)\Pr(\alpha_i \given Z_i)\\
					   &\propto&  \exp{[-\frac{\left(\alpha_i -  \hat{\alpha}_i\right)^2}{2se(\hat{\alpha}_i)^2}]} \exp{[-\frac{\alpha_i^2}{2\sigma^2_m}]}\\
					   &\propto&  \exp{[-\frac{\left(\sigma^2_m + se(\hat{\alpha}_i)^2\right)[\alpha_i- \frac{\sigma^2_m \hat{\alpha}_i}{\sigma^2_m + se(\hat{\alpha}_i)^2}]^2}{2se(\hat{\alpha}_i)^2\sigma^2_m}]},
\end{eqnarray}
leading to
\begin{eqnarray}
\alpha_i \given D_i, Z_i = m &\sim& \Normal(\frac{\sigma^2_m \hat{\alpha}_i}{\sigma^2_m + se(\hat{\alpha}_i)^2}, \frac{\sigma^2_mse(\hat{\alpha}_i)^2}{\sigma^2_m + se(\hat{\alpha}_i)^2}).
\end{eqnarray}

%\zuk{From this we can compute the posterior probability distribution on $\alpha_i$ - right? it's not just plugging in the fitted $\pi$ vector, since
%conditional on the data we get different probabilities for different Gaussians?}
%\heejung{Yes. You're right}
\begin{eqnarray}
Pr(\alpha_i \given D_i) &=& \sum_{m=1}^M Pr(Z_i = m | D_i) Pr(\alpha_i \given D_i, Z_i = m) \\
&=& \sum_{m=1}^M \frac{\pi_m Pr(D_i | Z_i=m)}{\sum_n \pi_n Pr(D_i | Z_i=n)} Pr(\alpha_i \given D_i, Z_i = m) \\
&=& \sum_{m=1}^M A_{im}^l \Normal(\frac{\sigma^2_m \hat{\alpha}_i}{\sigma^2_m + se(\hat{\alpha}_i)^2}, \frac{\sigma^2_mse(\hat{\alpha}_i)^2}{\sigma^2_m + se(\hat{\alpha}_i)^2})
\end{eqnarray}

\subsection{EM-Algorithm Description}
We can now write explicitly the EM-algorithm:

\nin {\bf Algorithm: EM-Ash} \\
{\bf Input: Data $D = (\hat{\alpha}_i, se(\hat{\alpha}_i)^2), i=1, .., n$} \\
{\bf Output: $\hat{\pi}_{MLE}$} \\
\begin{enumerate}
\item
Initialize $\pi_m^1 = \frac{1}{M}$ for $m=1,..,M$. Set $l=1$.

\item
While $|| \pi^{l+1} - \pi^l || > \eps$  (i.e. repeat until $\pi^l$ value converges)

\begin{enumerate}
\item E-Step: For each $i=1,..,n$ set
\begin{eqnarray}
\BF_i(\sigma^2_m) &=& \sqrt{\frac{se(\hat{\alpha}_i)^2}{\sigma^2_m + se(\hat{\alpha}_i)^2}}  \exp{[\frac{\hat{\alpha}_i^2}{2se(\hat{\alpha}_i)^2} \frac{\sigma^2_m}{\sigma^2_m + se(\hat{\alpha}_i)^2}]}
\end{eqnarray}

For each $i=1,..,n$ and $m=1,..,M$ set
\begin{eqnarray}
A_{im}^l &=& \frac{\pi_m^l\BF_i(\sigma^2_m)}{\sum_{n=1}^M\pi_n^l\BF_i(\sigma^2_n)}
\end{eqnarray}

\item M-Step: For each $m=1,..,M$, set
\begin{eqnarray}
	\pi_m^{l+1}  &=& \frac{\sum_{i=1}^nA_{im}^l}{n}.
\end{eqnarray}
set $l=l+1$.

\end{enumerate}

\item Output $\hat{\pi}_{MLE} = \pi^{l}$.

\end{enumerate}

\section{Mean and Variance for a mixture of distributions}
\label{sec:mean_var_mixture}
We will describe how to get mean and variance for a mixture of distributions in a general context.
Consider a mixture of distributions whose mean and variance are given. Specifically,
\begin{equation}
Y \given \pi, \mu, \sigma^2 \sim \sum_{m=1}^M \pi_mD(\mu_m, \sigma^2_m),
\end{equation}
where $\pi = (\pi_1, \ldots, \pi_M)$ are the mixture proportions which are constrained to be non-negative and sum to one and $\mu = (\mu_1, \ldots, \mu_M)$ and $\sigma^2 = (\sigma^2_1, \ldots, \sigma^2_M)$ are the means and variances for each distribution.
This distribution can be written as
\begin{eqnarray}
Y \given Z=m &\sim& D(\mu_m, \sigma^2_m)
\end{eqnarray}
where a latent variable $Z \in \{1, \ldots, M\}$ has a distribution $\Pr(Z = m) = \pi_m$.
Then,
\begin{eqnarray}
\mean(Y) &=& \mean(\mean(Y \given Z)) \\
	        &=& \sum_{m=1}^M \pi_m \mu_m,\\
\mean(Y^2) &=&  \mean(\mean(Y^2 \given Z)) \\
	        &=& \sum_{m=1}^M \pi_m (\mu_m^2 + \sigma_m^2), \\
\var(Y)  &=& \mean(Y^2) - \mean(Y)^2  \\
        &=& \sum_{m=1}^M \pi_m (\mu_m^2 + \sigma_m^2) - \big(\sum_{m=1}^M \pi_m \mu_m\big)^2
\end{eqnarray}

\bibliographystyle{rss}
\bibliography{sum}

\end{document}

### 2013-05-15 talks/ilike-slides.Rmd
final slides

@@
 * Allegro (ma non troppo)
+* Coda
 
@@
 
+# Other Applications
+
+- Widely applicable: perhaps anywhere (?) 
+where shrinkage is appropriate, requiring only an estimated
+effect size and standard error for each object.
+
+- Could also use effect size estimate and $p$ value for each variable, by converting to effect size estimate and (pseudo-) standard error.
+
+- Currently applying it to wavelet shrinkage applications.
+
+# Guarantees?
+
+- ``I think you have some nice ideas. How will you convince
+people to use them?" (C Morris)
+
 # Next steps?
@@
 - Extend to allow for correlations in the measured $\hat\beta_j$.
+
+# Coda: Other i-like related work
+
+- Bayesian variable selection for large-scale linear regression.
+
+- BSLMM: $$Y = X \beta + \epsilon,$$ with $\beta_j \sim 
+\pi N(0,\sigma_b^2) + (1-\pi) N(0,\sigma_a^2+\sigma_b^2)$.
+
+- Particular focus on prior specification (reparameterize in 
+terms of regression $R^2$).
+
+- BSLMM software, runs with thousands of individuals, hundreds
+of thousands of variables. (Zhou et al, PloS Genetics 2013)
+
+- Also variational approximations (Carbonetto and Stephens, Bayesian Analysis, 2012)
 

### 2013-05-20 Rcode/intro.rmd
extended intro.rmd

@@
+### Project Goals
+
+The initial goal of the ASH (Adaptive SHrinkage) project is to provide simple, generic, and flexible methods to derive ``shrinkage-based" estimates and credible intervals for unknown quantities $\beta=(\beta_1,\dots,\beta_J)$, given only estimates of those quantities ($\hat\beta=(\hat\beta_1,\dots, \hat\beta_J)$) and their corresponding estimated standard errors ($s=(s_1,\dots,s_J)$).
+
+Although shrinkage-based estimation can be motivated in various ways, our key goal here is to combine information across the multiple measurements $j=1,\dots,J$ to improve inference for each individual $\beta_j$.
+By improved inference, we mean both
+improved average accuracy of point estimates, which is
+the traditional focus of shrinkage-based methods, \emph{and} improved assessments of uncertainty.
+
+By ``adaptive" shrinkage we 
+have two key properties in mind.
+First, the appropriate amount of shrinkage is determined from the data, rather than being pre-specified.
+Second, the amount of shrinkage undergone by each $\hat\beta_j$ will depend on the standard error $s_j$: measurements with high standard error will undergo more shrinkage than measurements with low standard error.
+
+As an important special case, we intend these methods to address the "multiple comparisons" problem, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
+Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\hat\beta_j/s_j$,
+and then applying a generic procedure, such as that of Benjamini 
+and Hochberg (1995?) or Storey (2001?), designed to control or
+estimate the false discovery rate (FDR) or the positive FDR (Storey, 2001?). In essence we aim to provide analagous
+generic methods that work directly with two numbers for each 
+measurement $(\hat\beta_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
+Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
+uncertainty in each measurement $\hat\beta_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
+The potential for shrinkage-based estimation to
+help address multiple comparisons problems has been highlighted
+previously, including Greenland and Robins (1991),
+Efron (2008) and Gelman et al (2012). [Note, check also Louis, JASA, 1984]
+
+### Methods Outline
+
+The methods are based on treating the vectors $\hat\beta$
+and $s$ as ``observed data", and then performing inference for $\beta$ from these observed data, using a standard hierarchical modelling framework
+to combine information across $j=1,\dots,J$.
+
+Specifically, 
+we assume that the true 
+$\beta_j$ values are independent
+and identically distributed from some distribution $g(\cdot ; \pi)$, where $\pi$ is a hyper-parameter to be estimated. 
+Then, given $\beta$, we assume that $(\hat\beta_j,s_j)$ are independent across $j$, and depend only on $\beta_j$.
+Putting these together, the joint model for the unobserved $\beta$ and the observed $\hat\beta, s$ is:
+\begin{align}
+p(\hat\beta, s, \beta | \pi) & = \prod_j g(\beta_j ; \pi) p(\hat\beta_j, s_j | \beta_j) \\
+& = \prod_j g(\beta_j ; \pi) L(\beta_j; \hat\beta_j,s_j).
+\end{align}
+The specific choices of $g$ and $L$ are described below.
+
+We fit this hierarchical model using the following "Empirical Bayes" approach. First we estimate the hyper-parameters $\pi$ by maximizing the likelihood
+$$L(\pi; \hat\beta, s) := p(\hat\beta, s | \pi) = \int p(\hat\beta, s, \beta | \pi) d\beta.$$
+Then, given this estimate $\hat\pi$, we compute the conditional distributions $$p(\beta_j | \hat\pi, \hat\beta, s) \propto g(\beta_j; \pi) L(\beta_j; \hat\beta_j, s_j).$$ 
+In principle we would
+prefer to take a full Bayes approach that accounts for uncertainty in $\pi$, but, at least for now, we compromise this principle for the simplicity of the EB approach.
+
+[put picture of hierarchical model here]
+
+The conditional distributions $p(\beta_j | \hat\pi, \hat\beta, s)$ 
+encapsulate uncertainty in the values for $\beta_j$, combining information across
+$j=1,\dots,J$. The combining of the information occurs through estimation of
+$\pi$, which involves all of the data, and it is
+
+These conditional distributions can be conveniently summarized
+in various ways, including point estimates (e.g. the posterior means or medians),
+and credible intervals/regions.
+
+The key components of this hierarchical model
+are the distribution $g$ and the likelihood $L(\beta_j; \hat\beta_j, s_j)$. We make the following choices for these.
+
+1. The likelihood for $\beta_j$ is normal, centered on $\hat\beta_j$, with standard deviation $s_j$.
+That is, 
+$$L(\beta_j; \hat\beta_j, s_j) \propto \exp[-0.5(\beta_j-\hat\beta_j)^2/s_j^2]. \quad (**)$$
+
+2. The distribution $g(\cdot; \pi)$ is a mixture of zero-centered normal distributions, 
+$$g(\cdot; \pi) = \sum_{k=1}^K pi_k N(\cdot; 0, \sigma^2_k).$$
+In practice, we currently fix the number of components $K$ to be large, and take the variances $\sigma_1<\sigma_2<\dots<\sigma_K$ to be fixed, and vary from very small (possibly 0), to very large --  sufficiently large that typically $\hat\pi_K=0$.
+
+The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
+For example, we can write $p(\hat\beta_j, s_j | \beta_j) = p(\hat\beta_j | s_j, \beta_j)p(s_j | \beta_j)$.
+Now, if we are willing to assume that
+$s_j$ alone contains no information about $\beta_j$, or equivalently that $p(s_j | \beta_j)$ does not depend on $\beta_j$, then
+$L(\beta_j) \propto p(\hat\beta_j | s_j, \beta_j)$,
+and if $\hat\beta_j | s_j, \beta_j \sim N(\beta_j, s_j^2)$,
+as is often asymptotically the case, then we obtain the likelihood (**) above.
+
+An alternative motivation is to think of this as a normal
+approximation to the likelihood from the 
+raw data $D_j$ that were used to compute
+$\hat\beta_j$ and $s_j$. Then if we observed these data
+the likelihood for $\beta$ would be 
+$p(D_j | \beta_j)$, and a Taylor series expansion of the log likelihood around the maximum likelihood estimate $\hat\beta_j$  yields $$l(\beta_j) \approx l(\hat\beta_j) + 0.5* (\beta_j - \hat\beta_j)^2 l''(\hat\beta_j).$$ [Fill in details?]
+
+The idea of using a mixture of normal distributions for $g$ 
+also seems very natural: mixtures of normals provide a flexible family of distributions able to provide a good approximation to any true underlying $g$; and 
+when combined with the normal likelihood they give
+an analytic form for the conditional distribution $p(\beta_j | \hat\pi, \hat\beta_j, s_j)$ (also a mixture of normals).
+The constraint that these normals be centered at zero may seem initially less natural. 
+Certainly this constraint could be relaxed in principle.
+However, we view it as a convenient way to impose an assumption 
+that $g$ is unimodal with its mode at 0,
+which we view as a plausible assumption in many 
+settings, and one which may be helpful to avoid
+"overfitting" of $g$. (Using normal distributions centered at 0
+also imposes an assumption
+that $g$ is symmetric about zero, which we view as less
+plausible, but represents a compromise between simplicity
+and flexibility.
+In cases where this assumption seems wildly inappropriate one could perhaps
+improve results by applying the
+model separately to positive and negative values of $\betahat$.)
+
+Finally, using a large number of normal components with a wide range of variances,
+rather than, say, a smaller number of components with the variances
+to be estimated, is simply for computational convenience. With fixed
+variances there exists a very simple EM algorithm to maximise the likelihood
+in $\pi$. Obtaining maximum likelihood estimates for the variances could certainly
+be implemented with a little
+more work, but it is unclear whether this would result in practically-important gains in many situations of interest.
+
+### Computation Outline
+
+As outlined above, we fit the model using the following Empirical Bayes procedure:
+1. Estimate $\pi$ by maximizing the likelihood $L(\pi)$.
+2. Compute the conditional distributions $p(\beta_j | \hat\beta_j, s_j, \hat\pi)$.
+
+Using a normal likelihood $L(\beta_j)$, and assuming
+$g$ to be a mixture of normal distributions with fixed variances, 
+yields a simple EM algorithm
+for estimating $\pi$ in Step 1, and simple analytic forms for the conditional
+distributions in Step 2.
+
+### A simple example
+
+Load in some functions.
+```{r}
+setwd("~/Documents/git/ash/Rcode/")
+set.seed(32327)
+## load Poisson_binomial and ash functions 
+source("../Rcode/ash.R") 
+library("qvalue")
+```
+
+Here we simulate data, effectively as in Efron, 2008, Section 7.
+```{r}
+truenull = c(rep(0,1000),rep(1,9000))
+beta = c(rnorm(1000,-3,1),rep(0,9000))
+s= rep(1,10000)
+betahat = beta+rnorm(10000,beta,s)
+
+beta.ash=ash(betahat,s)
+#compute the usual zscore and corresponding p value
+zscore = betahat/s
+pval = pchisq(zscore^2,df=1,lower.tail=F)
+qval = qvalue(pval)
+
+plot(betahat,beta.ash$PosteriorMean,xlab="Observed betahat", ylab="Estimated beta (posterior mean)",ylim=c(-7,4),xlim=c(-7,4))
+abline(h=0)
+abline(a=0,b=1,col=2)
+
+plot(qval$q,2*beta.ash$qval,main="comparison of ash and q value qvalues")
+abline(a=0,b=1)
+
+o = order(beta.ash$qval)
+
+plot(cumsum(truenull[o])/(1:10000),qval$qval[o],col=2,type="l")
+lines(cumsum(truenull[o])/(1:10000),2*beta.ash$qval[o])
+abline(a=0,b=1)
+```
+
+It seems that in this case the ash q values underestimate the
+FDR slightly. Possibly this is the assymetry.
+Try shrinking positive and negatives separately:
+```{r}
+pos= betahat>0
+betapos.ash=ash(betahat[pos],s[pos])
+betaneg.ash = ash(betahat[!pos],s[!pos])
+lfdr = rep(0,length(betahat))
+lfdr[pos] = betapos.ash$localfdr
+lfdr[!pos] = betaneg.ash$localfdr
+qv = qval.from.localfdr(lfdr)
+o = order(qv)
+plot(cumsum(truenull[o])/(1:10000),2*qv[o],type="l")
+abline(a=0,b=1)
+```
+No, that's not it.
+
+So maybe it is the non-monotonic behaviour of the true beta values in this case.
+Let's try putting in a component at -1.5 to make the data fit this assumption better:
+
+```{r}
+truenull = c(rep(0,2000),rep(1,8000))
+beta = c(rnorm(1000,-3,1),rnorm(1000,-1.5,1),rep(0,8000))
+s= rep(1,10000)
+betahat = beta+rnorm(10000,beta,s)
+
+beta.ash=ash(betahat,s)
+#compute the usual zscore and corresponding p value
+zscore = betahat/s
+pval = pchisq(zscore^2,df=1,lower.tail=F)
+qval = qvalue(pval)
+plot(qval$q,2*beta.ash$qval,main="comparison of ash and q value qvalues")
+abline(a=0,b=1)
+
+o = order(beta.ash$qval)
+
+plot(cumsum(truenull[o])/(1:10000),qval$qval[o],col=2,type="l")
+lines(cumsum(truenull[o])/(1:10000),2*beta.ash$qval[o])
+abline(a=0,b=1)
+
+```
+
+### Motivation
+
 It is common in statistics that you measure many ``similar" things imperfectly, and wish to estimate their values.
@@
 that the measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.
+
+Note on multiple comparisons: it isn't really a "problem" but an "opportunity". This viewpoint also espoused by Greenland and Robins. It isn't the number of tests that is relevant (the false
+discovery rate at a given threshold does not depend on the number of tests). It is the *results* of the tests that are relevant.

### 2013-05-20 Rcode/readme.rmd
extended intro.rmd

@@
 #biocLite("qvalue")
-setwd("~/Documents/BayesFDR/Rcode/")
+setwd("~/Documents/git/ash/Rcode/")
 library("qvalue")
@@
 
+### Errors in sign, and true 0s
+
+If the true value is 0 and you say it is positive,
+is this an error in sign? Well, yes, it is. But it is not
+negative, and it turns out you need to count errors of sign
+when truth is 0 as 0.5 of a false discovery to make things
+well calibrated. This is because the method gives the same
+results if the true value of mu is exactly 0, or if it has
+epsilon variance around 0. But of course the errors in sign
+in the latter occur at rate 0.5 because of random guessing.
+
+The following simulation illustrates this. It also
+illustrates that the method seems to be relatively robust to
+the situation where the true $\beta$ are not symmetric around 0.
+```{r}
+musim2 = c(rnorm(1000,-3,1),rnorm(1000,-1.5,1),rnorm(1000,0,1),rep(0,7000))
+zsim2 = musim2+rnorm(10000)
+zsim2.ash=ash(zsim2,rep(1,10000))
+hist(musim2[1:3000])
+errorinsign = ifelse(musim2==0,0.5,(zsim2.ash$PosteriorMean/musim2)<0)
+sum(errorinsign[zsim2.ash$qvalue<0.05])/sum(zsim2.ash$qvalue<0.05)
+```
+
 #### Miscellaneous
+
+Simulate data as in Efron, 2008, Section 7. 
+```{r}
+musim = c(rnorm(1000,-3,1),rep(0,9000))
+zsim = musim+rnorm(10000)
+zsim.ash=ash(zsim,rep(1,10000))
+```
+Need to implement a method to find CIs to look at this.
 

### 2013-05-28 Rcode/intro.rmd
added simple example to intro.rmd

@@
 
+```{r}
+#simulate n beta-hat values, nnull under the null
+#with altmean and altsd being the mean and sd of beta under the alternative
+simdata=function(n, nnull, altmean, altsd, betasd){
+  null = c(rep(1,nnull),rep(0,n-nnull))
+  beta = c(rep(0,nnull),rnorm(n-nnull,altmean,altsd))
+  betahat = rnorm(n,beta,betasd)
+  return(list(null=null,beta=beta,betahat=betahat,betasd=betasd))
+}
+```
+
+Simulate 10000 tests, with 2000 alternatives, $\beta \sim N(0,sd=2)$, 
+and 8000 null values ($\beta=0$).
+```{r}
+ss = simdata(10000,8000,0,2,1)
+
+beta.ash=ash(ss$betahat,ss$betasd)
+#compute the usual zscore and corresponding p value
+zscore = ss$betahat/ss$betasd
+pval = pchisq(zscore^2,df=1,lower.tail=F)
+qval = qvalue(pval)
+```
+
+Now, we find the fitted values for $\pi$ are mostly near 0, or at 2 (the alternative value).
+```{r}
+beta.ash$fitted.f
+```
+
+Also, we can see the effects of shrinkage: small estimates of $\hat\beta$ are
+shrunk to close to 0. Large estimates of $\hat\beta$ are shrunk less strongly because ash recognizes that these larger $\hat\beta$ are likely
+from the alternative, rather than the null.
+```{r}
+plot(ss$betahat,beta.ash$PosteriorMean,xlab="Observed betahat", ylab="Estimated beta (posterior mean)",ylim=c(-7,4),xlim=c(-7,4))
+abline(h=0)
+abline(a=0,b=1,col=2)
+```
+
+Note that to compare ash $q$ values with fdr-type $q$ values we
+multiply them by 2 (this is because the ash $q$ values indicate the
+probability of an error of sign, which is half as large as the
+probability of a false discovery: half the false discoveries would
+have the right sign by random guessing)
+```{r}
+plot(qval$q,2*beta.ash$qval,main="comparison of ash and q value qvalues",xlab="qvalue", ylab="twice ash q values")
+abline(a=0,b=1)
+```
+
+In this example we see that qval overestimates the actual FDR. (This
+is because it assumes all the $p$ values near 1 are null, when they are not.)
+```{r}
+o = order(beta.ash$qval)
+plot(cumsum(ss$null[o])/(1:10000),qval$qval[o],col=2,type="l", xlab="actual FDR", ylab="q value", main="Comparison of actual FDR with q value")
+lines(cumsum(ss$null[o])/(1:10000),2*beta.ash$qval[o])
+abline(a=0,b=1)
+```
+
+### Miscellaneous
+
+code and text below here is work in progress and untidied.
+
 Here we simulate data, effectively as in Efron, 2008, Section 7.
@@
 s= rep(1,10000)
-betahat = beta+rnorm(10000,beta,s)
+betahat = rnorm(10000,beta,s)
 

### 2013-05-28 Rcode/ash.R

@@ -294,6 +294,10 @@ ash = function(betahat,sebetahat,nullcheck=TRUE,df=NULL,randomstart=FALSE, usePo
   if(!is.null(df)){
     betahat = effective.effect(betahat,sebetahat,df)
   }	
+  
+  if(length(sebetahat)==1){
+    sebetahat = rep(sebetahat,length(betahat))
+  }
   if(usePointMass){
         sigmaavec = c(0,0.0005,0.001,0.002,0.004,0.008,0.016,0.032,0.064,0.128,0.256,0.512,1.024,2.048,4.096,8.192)
   }else{		

### 2013-06-07 Rcode/intro.rmd
added ability to specify sigmaavec

@@
 zscore = ss$betahat/ss$betasd
+
 pval = pchisq(zscore^2,df=1,lower.tail=F)
 qval = qvalue(pval)
+hist(zscore)
 ```

### 2013-06-07 Rcode/ash.R

@@ -243,6 +243,7 @@ qval.from.localfdr = function(localfdr){
   return(qvalue)
 }
 
+
 #main adaptive shrinkage function
 #takes a vector of betahats and ses;
 #fits a mixture of normals to it
@@ -258,7 +259,7 @@ qval.from.localfdr = function(localfdr){
 #Things to do: automate choice of sigmavec
 # check sampling routin
 # check number of iterations
-ash = function(betahat,sebetahat,nullcheck=TRUE,df=NULL,randomstart=FALSE, usePointMass = FALSE, onlylogLR = FALSE, localfdr = TRUE, prior=NULL){
+ash = function(betahat,sebetahat,nullcheck=TRUE,df=NULL,randomstart=FALSE, usePointMass = FALSE, onlylogLR = FALSE, localfdr = TRUE, prior=NULL, sigmaavec=NULL){
   #if df specified, convert betahat so that bethata/sebetahat gives the same p value
   #from a z test as the original effects would give under a t test with df=df
   if(!is.null(df)){
@@ -268,11 +269,13 @@ ash = function(betahat,sebetahat,nullcheck=TRUE,df=NULL,randomstart=FALSE, usePo
   if(length(sebetahat)==1){
     sebetahat = rep(sebetahat,length(betahat))
   }
-  if(usePointMass){
-        sigmaavec = c(0,0.0005,0.001,0.002,0.004,0.008,0.016,0.032,0.064,0.128,0.256,0.512,1.024,2.048,4.096,8.192)
-  }else{		
-  	sigmaavec = c(0.00025,0.0005,0.001,0.002,0.004,0.008,0.016,0.032,0.064,0.128,0.256,0.512,1.024,2.048,4.096,8.192)
+  if(is.null(sigmaavec)){
+    sigmaavec = c(0.00025,0.0005,0.001,0.002,0.004,0.008,0.016,0.032,0.064,0.128,0.256,0.512,1.024,2.048,4.096,8.192)
   }
+  if(usePointMass){
+        sigmaavec = c(0,sigmaavec)
+  }
+    
   pi = rep(1, length(sigmaavec))
   pi[1]=length(sigmaavec)
   pi=normalize(pi)
@@ -306,85 +309,6 @@ ash = function(betahat,sebetahat,nullcheck=TRUE,df=NULL,randomstart=FALSE, usePo
 
 
 
-
-
-
-
-
[…]

### 2013-06-07 Rcode/intro.rmd
added automatic say to select sigmaavec

@@
 beta.ash=ash(ss$betahat,ss$betasd)
+beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)
 #compute the usual zscore and corresponding p value
@@
 beta.ash$fitted.f
+```
+
+Plot the fitted underlying distribution on top of true values for beta
+```{r}
+hist(ss$beta,prob=TRUE,breaks=seq(-7,7,length=20))
+x= seq(-4,4,length=1000)
+lines(x,density(beta.ash,x),col=2)
+```
+
+[for testing: compare with the results from the automatic way for selecting sigma]
+```{r}
+hist(ss$beta,prob=TRUE,breaks=seq(-7,7,length=20))
+x= seq(-4,4,length=1000)
+lines(x,density(beta.ash.auto,x),col=2)
 ```

### 2013-06-10 Rcode/intro.rmd
added comparison of localfdr from auto and non-auto methods

@@
 
+[for testing: note that the PosteriorMean and PositiveProb don't depend much on sigmaa vec used ]
+```{r}
+plot(beta.ash.auto$PosteriorMean, beta.ash$PosteriorMean, xlab="Shrunk estimate from auto method", ylab="Shrunk estimate from fixed method")
+abline(a=0,b=1)
+plot(beta.ash.auto$localfdr, beta.ash$localfdr, xlab="lfdr from auto method", ylab="ldfr from fixed method")
+abline(a=0,b=1)
+```
+
 Also, we can see the effects of shrinkage: small estimates of $\hat\beta$ are

### 2013-06-28 Rcode/intro.rmd
defined n in VB EM; change log(gamma) to lgamma in KL

@@
 beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)
+beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior="uniform")
+beta.ash.vb.null = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior=NULL)
+
 #compute the usual zscore and corresponding p value
@@
 beta.ash$fitted.f
+beta.ash.auto$fitted.f
+beta.ash.vb.uniform$fitted.f
+beta.ash.vb.null$fitted.f
 ```

### 2013-06-30 Rcode/intro.rmd
corrected error introduced in diriKL
added text to describe the issue with the factor of 2 in local false sign rates; corrected bug in computation of localfdr with point mass

--- 2013-06-10 Rcode/intro.rmd
@@
 beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)
+#these to test the VB version 
+#beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior="uniform")
+#beta.ash.vb.null = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior=NULL)
+
 #compute the usual zscore and corresponding p value
@@
 beta.ash$fitted.f
+beta.ash.auto$fitted.f
+#beta.ash.vb.uniform$fitted.f
+#beta.ash.vb.null$fitted.f
 ```
@@
 ```
+
+### Some additional notes
+
+#### Do we need a point mass at zero?
+
+In some settings it is the convention to focus on testing whether $\beta_j=0$.
+However some dislike this focus, objecting that it is unlikely to be the case that $\beta_j=0$ exactly.
+For example, when comparing the average expression of a gene in human samples vs chimp samples, it might be considered unlikely that the expression
+is *exactly* the same in both. Whether or not $\beta_j=0$
+is considered unlikely may depend on the context.
+However, in most contexts, finite data cannot
+distinguish between $\beta_j=0$ and $\beta_j$ being very close to zero.
+Thus finite data cannot usually convince a skeptic that $\beta_j$ is exactly zero, rather than just very small.
+In contrast it is easy to imagine data that would convince a doubter that $\beta_j$ is truly non-zero.
+In this sense there is an assymetry between the inferences "$\beta_j$ is zero" and "$\beta_j$ is non-zero", an assymetry that is reflected in the admonition "failure to reject the null does not imply it to be true".
+
+Thus any analysis that purports to distinguish between these cases must be making an assumption.
+
+Consider two analyses of the same data, using two different "priors" $g$ for $\beta_j$, that effectively differ only in their assumptions about whether or not $\beta_j$ can be exactly zero.
+For concreteness, consider
+\[ g_1(\cdot) = \pi \delta_0(\cdot) + (1-\pi) N(\cdot; 0,\sigma^2) \]
+and
+\[g_2(\cdot) = \pi N(\cdot; 0, \epsilon^2) + (1-\pi) N(\cdot; 0, \sigma^2).\]
+If $\epsilon^2$ is sufficiently small, then these 
+priors are "approximately the same", and will lead to "approximately the same" posteriors and inferences in many senses.
+To discuss these, let $p_j$ denote the posterior under prior $g_j$.
+Then, for any given (small) $\delta$, we will have $p_1(|\beta_j|<\delta) \approx p_2(|\beta_j|< \delta)$.
+However, we will not have $p_1(\beta_j=0) \approx p_2(\beta_j=0)$: the latter will always be zero, while the former could be appreciable.
+
+ What if, instead, we examine $p_1(\beta_j >0)$ and $p_2(\beta_j >0)$?
+Again, these will differ.
+If this probability is big in the first analysis, say $1-\alpha$ with $\alpha$ small, then it could be as big as $1-\alpha/2$ in the second analysis.
+This is because if $p_1(\beta_j>0)=1-\alpha$, then $p_1(\beta_j=0)$ will often be close to $\alpha$, so for small $\epsilon$ $p_2(\beta_j)$ will have mass $\alpha$ near 0, of which half will be positive and half will be negative.
+Thus if we do an analysis without a point mass, but allow
+for mass near 0, then we may predict what the results would have been if we had used a point mass.
+
+Let's try: 
+```{r}
+beta.ash.pm = ash(ss$betahat, ss$betasd, auto=TRUE, usePointMass=TRUE)
+beta.ash.pm$fitted.f
+beta.ash.auto$fitted.f
+plot(beta.ash.auto$localfdr,beta.ash.pm$localfdr,main="comparison of ash localfdr, with and without point mass",xlab="no point mass", ylab="with point mass",xlim=c(0,1),ylim=c(0,1))
+abline(a=0,b=2)
+```
+
+Our conclusion: if we simulate data with a point mass,
+and we analyze it without a point mass, we may underestimate
+the false sign rate by a factor of 2.
+Therefore, to be conservative, we might prefer to analyze the data allowing for the point mass, or, if analyzed without a point mass, multiply estimated false sign rates by 2.
 

### 2013-06-30 Rcode/intro.rmd
minor update

@@
 
-Note that to compare ash $q$ values with fdr-type $q$ values we
-multiply them by 2 (this is because the ash $q$ values indicate the
-probability of an error of sign, which is half as large as the
-probability of a false discovery: half the false discoveries would
-have the right sign by random guessing)
+#### Comparison with qvalue
+
+Here we compare ash $q$ values with those from the qvalue package. Because of the issue above, we multiply the ash  $q$ values by 2.
 ```{r}

### 2013-06-30 Rcode/intro.rmd
added text on false sign rate

@@
 prefer to take a full Bayes approach that accounts for uncertainty in $\pi$, but, at least for now, we compromise this principle for the simplicity of the EB approach.
+[Note: a Variational Bayes version of this is
+also implemented, and may become our preferred approach
+after testing]
 
@@
 
-The idea of using a mixture of normal distributions for $g$ 
+Using a mixture of normal distributions for $g$ 
 also seems very natural: mixtures of normals provide a flexible family of distributions able to provide a good approximation to any true underlying $g$; and 
@@
 improve results by applying the
-model separately to positive and negative values of $\betahat$.)
+model separately to positive and negative values of $\hat\beta$.)
 
@@
 more work, but it is unclear whether this would result in practically-important gains in many situations of interest.
+
+### The False Sign Rate
+
+The usual definition of the local False Discovery Rate (lfdr) for observation $j$ is
+$\text{lfdr}_j = p(\beta_j = 0 | \hat\beta, s)$.
+The lfdr terminology comes from using "discovery" to refer to rejecting the null ($H_j:\beta_j=0$), so lfdr 
+gives the probability that we make a mistake if we reject
+$H_j$, that is the probability that, if we reject $H_j$,
+it is a  "false discovery".
+
+As pointed out by Gelman et al, there are settings where
+$\beta_j=0$ is implausible, in which case the lfdr is not
+a useful concept: if every $\beta_j$ is non-zero then there
+is no such thing as a false discovery and the lfdr will be 0. Gelman et al suggest that in such settings we
+might replace the concept of a false discovery with the
+concept of an "error in sign". The idea is that, in settings where $\beta_j=0$ is implausible, the most fundamental
+inference objective is to ask which $\beta_j$ are positive and which are negative. Indeed, even in settings where some
+$\beta_j$ are exactly zero, it could be argued that identifying which are positive and which negative is
+fundamentally more interesting and useful than identifying which are non-zero.
+However, in such settings this usually comes down to something similar in practice: the observations that one can tell are non-zero are also the ones whose
+sign can be reliably inferred. In this sense the
+local false sign rate is a natural generalization of
+the local false discovery rate.
+
+Following these ideas of Gelman et al, we define the local False Sign Rate (lfsr) as 
+$$\text{lfsr}_j = \min[ p(\beta_j \geq 0| \hat\beta, s), p(\beta_j \leq 0| \hat\beta, s) ].$$
+To give the intuition behind this, suppose for concreteness
+that the minimum is achieved by the first term, $p(\beta_j \geq 0| \hat\beta, s)=0.05$ say. Then, if we were to declare that $\beta_j$ is negative, the probability that we have
+made an error in sign would be 0.05.
+
+Note that $\text{lfsr}_j \geq \text{lfdr}_j$ 
+because both the events $\beta_j \geq 0$
+and $\beta_j \leq 0$ include the event $\beta_j=0$.
+Thus, lfsr gives an upper bound for the lfdr,
+and so can be used as a conservative estimate of the lfdr
+if an lfdr is desired. (This may be helpful when comparing
+with methods that compute an lfdr.)
 

### 2013-07-09 Rcode/intro.rmd

@@
 
-As an important special case, we intend these methods to address the "multiple comparisons" problem, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
+As an important special case, these methods address the "multiple comparisons" setting, where interest focuses on which $\beta_j$ can be confidently inferred to be non-zero.
 Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\hat\beta_j/s_j$,
@@
 uncertainty in each measurement $\hat\beta_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
+
 The potential for shrinkage-based estimation to
-help address multiple comparisons problems has been highlighted
+address the multiple comparisons setting 
+has been highlighted
 previously, including Greenland and Robins (1991),
 Efron (2008) and Gelman et al (2012). [Note, check also Louis, JASA, 1984]
+
+Performing multiple comparisons, or multiple tests, is often
+regarded as a "problem". However, here we regard it instead as an opportunity - an opportunity to combine (or "pool") information across tests or comparisons.
+
+Imagine that you are preparing to perform 1 million tests,
+each based on a $Z$ score that is assumed to be $N(0,1)$ under the null.
+You first order these tests randomly, and begin by 
+performing the first test, which returns a $Z$ score of 4.
+At this point you are interrupted by a friend, who asks how the analysis is going.
+"It's early days, but looking promising" you reply.
+Well, who wouldn't?
+If the aim is to find lots of significant differences, a strong first result is surely a good outcome.
+
+At this point you have reason to expect that many of the
+subsequent tests also output strong results.
+
+Now consider two alternative scenarios for the remaining 999,999 tests. In the first scenario, the remaining tests
+produce $Z$ values that fit very well with the null, closely following a standard normal distribution; in the second
+scenario a large proportion of the remaining tests, say 50 percent, show outcomes that lie outside of $[-4,4]$.
+
+If your friend enquired after your analysis again, your
+response would surely differ in the first scenario ("Oh, it didn't pan out so well after all") vs the second ("It went great"). Further,
+in the first scenario, if your friend pressed you further about the results of the first test,
+you would likely, I think, be inclined to put them down to chance.
+In contrast, in the second scenario, the first test turned out to be, as you hoped, a harbinger of good things to come, and in this scenario you would likely regard that test as likely corresponding to a true discovery.
+
+The key point is that it is the *outcomes* of the tests, not the *number* of tests, that impacts interpretation of that first test.
+
+(Some may by pondering whether the fact that you are about to perform another
+999,999 tests should be considered pertinent in responding to your friend. Our view is that it is  irrelevant.
+The standard frequentist framework would disagree, because
+it requires the analyst to consider hypothetical repetitions
+of the "experiment", and so the fact that the experiment
+consists of a million tests is pertinent. However, this
+argument is a distraction from the main point.)
+
+Indeed, we believe that the practice of focussing on the *number* of tests performed is
+
+Focussing on the number of tests performed can be
+seen as an approximation.
+
+The standard argument is that,
+when performing multiple tests, some will be significant
+just by chance.
 

### 2013-07-15 Rcode/intro.rmd
updated test of vb method; updated intro

@@
 Efron (2008) and Gelman et al (2012). [Note, check also Louis, JASA, 1984]
-
-Performing multiple comparisons, or multiple tests, is often
-regarded as a "problem". However, here we regard it instead as an opportunity - an opportunity to combine (or "pool") information across tests or comparisons.
-
-Imagine that you are preparing to perform 1 million tests,
-each based on a $Z$ score that is assumed to be $N(0,1)$ under the null.
-You first order these tests randomly, and begin by 
-performing the first test, which returns a $Z$ score of 4.
-At this point you are interrupted by a friend, who asks how the analysis is going.
-"It's early days, but looking promising" you reply.
-Well, who wouldn't?
-If the aim is to find lots of significant differences, a strong first result is surely a good outcome.
-
-At this point you have reason to expect that many of the
-subsequent tests also output strong results.
-
-Now consider two alternative scenarios for the remaining 999,999 tests. In the first scenario, the remaining tests
-produce $Z$ values that fit very well with the null, closely following a standard normal distribution; in the second
-scenario a large proportion of the remaining tests, say 50 percent, show outcomes that lie outside of $[-4,4]$.
-
-If your friend enquired after your analysis again, your
-response would surely differ in the first scenario ("Oh, it didn't pan out so well after all") vs the second ("It went great"). Further,
-in the first scenario, if your friend pressed you further about the results of the first test,
-you would likely, I think, be inclined to put them down to chance.
-In contrast, in the second scenario, the first test turned out to be, as you hoped, a harbinger of good things to come, and in this scenario you would likely regard that test as likely corresponding to a true discovery.
-
-The key point is that it is the *outcomes* of the tests, not the *number* of tests, that impacts interpretation of that first test.
-
-(Some may by pondering whether the fact that you are about to perform another
-999,999 tests should be considered pertinent in responding to your friend. Our view is that it is  irrelevant.
-The standard frequentist framework would disagree, because
-it requires the analyst to consider hypothetical repetitions
-of the "experiment", and so the fact that the experiment
-consists of a million tests is pertinent. However, this
-argument is a distraction from the main point.)
-
-Indeed, we believe that the practice of focussing on the *number* of tests performed is
-
-Focussing on the number of tests performed can be
-seen as an approximation.
-
-The standard argument is that,
-when performing multiple tests, some will be significant
-just by chance.
 
@@
 with methods that compute an lfdr.)
+In addition, we intuitively expect the difference $\text{lfsr}_j - \text{lfdr}_j$ to be small, at least when $\text{lfdr}_j$ is small, because when we are confident that a particular effect is non-zero we would expect to be similarly confident about the sign of the effect.
 
@@
 #these to test the VB version 
-#beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior="uniform")
-#beta.ash.vb.null = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior=NULL)
+beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior="uniform")
+beta.ash.vb.null = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior=NULL)
 
@@
 beta.ash.auto$fitted.f
-#beta.ash.vb.uniform$fitted.f
-#beta.ash.vb.null$fitted.f
+beta.ash.vb.uniform$fitted.f
+beta.ash.vb.null$fitted.f
 ```
@@
 [for testing: note that the PosteriorMean and PositiveProb don't depend much on sigmaa vec used ]
+
 ```{r}
@@
 abline(a=0,b=1)
+
+```
+
+[And VB method produces similar results to EM method]
+```{r}
+plot(beta.ash.auto$PosteriorMean, beta.ash.vb.uniform$PosteriorMean, xlab="Shrunk estimate from auto method", ylab="Shrunk estimate from vb method")
+points(beta.ash.auto$PosteriorMean, beta.ash.vb.null$PosteriorMean, col=2)
+abline(a=0,b=1)
+plot(beta.ash.auto$localfdr, beta.ash.vb.uniform$localfdr, xlab="lfdr from auto method", ylab="ldfr from vb method")
+points(beta.ash.auto$localfdr, beta.ash.vb.null$localfdr, col=2)
+abline(a=0,b=1)
+
 ```
@@
 ```{r}
-plot(ss$betahat,beta.ash$PosteriorMean,xlab="Observed betahat", ylab="Estimated beta (posterior mean)",ylim=c(-7,4),xlim=c(-7,4))
+plot(ss$betahat,beta.ash$PosteriorMean,xlab="Observed betahat", ylab="Estimated beta (posterior mean)",ylim=c(-7,7),xlim=c(-7,7))
 abline(h=0)
@@
 beta.ash.auto$fitted.f
-plot(beta.ash.auto$localfdr,beta.ash.pm$localfdr,main="comparison of ash localfdr, with and without point mass",xlab="no point mass", ylab="with point mass",xlim=c(0,1),ylim=c(0,1))
+plot(beta.ash.auto$localfsr,beta.ash.pm$localfsr,main="comparison of ash localfsr, with and without point mass",xlab="no point mass", ylab="with point mass",xlim=c(0,1),ylim=c(0,1))
+abline(a=0,b=1)
 abline(a=0,b=2)
@@
 and we analyze it without a point mass, we may underestimate
-the false sign rate by a factor of 2.
+the lfsr by a factor of 2.
 Therefore, to be conservative, we might prefer to analyze the data allowing for the point mass, or, if analyzed without a point mass, multiply estimated false sign rates by 2.
+In fact the latter might be preferable: even if we analyze the data with a point mass, there is going to be some unidentifiability
+that means estimating the pi value on the point mass will be somewhat unreliable, and we still might underestimate the false sign rate if we rely on that estimate.
 
@@
 
-Here we compare ash $q$ values with those from the qvalue package. Because of the issue above, we multiply the ash  $q$ values by 2.
+Here we compare ash $q$ values with those from the qvalue package. 
 ```{r}
-plot(qval$q,2*beta.ash$qval,main="comparison of ash and q value qvalues",xlab="qvalue", ylab="twice ash q values")
+plot(qval$q,beta.ash$qval,main="comparison of ash and q value qvalues",xlab="qvalue", ylab="ash q values")
 abline(a=0,b=1)
@@
 plot(cumsum(ss$null[o])/(1:10000),qval$qval[o],col=2,type="l", xlab="actual FDR", ylab="q value", main="Comparison of actual FDR with q value")
-lines(cumsum(ss$null[o])/(1:10000),2*beta.ash$qval[o])
+lines(cumsum(ss$null[o])/(1:10000),beta.ash$qval[o])
 abline(a=0,b=1)
@@
 
-plot(betahat,beta.ash$PosteriorMean,xlab="Observed betahat", ylab="Estimated beta (posterior mean)",ylim=c(-7,4),xlim=c(-7,4))
+plot(betahat,beta.ash$PosteriorMean,xlab="Observed betahat", ylab="Estimated beta (posterior mean)",ylim=c(-7,7),xlim=c(-7,7))
 abline(h=0)
@@
 
-plot(qval$q,2*beta.ash$qval,main="comparison of ash and q value qvalues")
+plot(qval$q,beta.ash$qval,main="comparison of ash and q value qvalues")
 abline(a=0,b=1)
@@
 plot(cumsum(truenull[o])/(1:10000),qval$qval[o],col=2,type="l")
-lines(cumsum(truenull[o])/(1:10000),2*beta.ash$qval[o])
+lines(cumsum(truenull[o])/(1:10000),beta.ash$qval[o])
 abline(a=0,b=1)
@@
 o = order(qv)
-plot(cumsum(truenull[o])/(1:10000),2*qv[o],type="l")
+plot(cumsum(truenull[o])/(1:10000),qv[o],type="l")
 abline(a=0,b=1)
@@
 qval = qvalue(pval)
-plot(qval$q,2*beta.ash$qval,main="comparison of ash and q value qvalues")
+plot(qval$q,beta.ash$qval,main="comparison of ash and q value qvalues")
 abline(a=0,b=1)
@@
 discovery rate at a given threshold does not depend on the number of tests). It is the *results* of the tests that are relevant.
+
+Performing multiple comparisons, or multiple tests, is often
+regarded as a "problem". However, here we regard it instead as an opportunity - an opportunity to combine (or "pool") information across tests or comparisons.
+
+Imagine that you are preparing to perform 1 million tests,
+each based on a $Z$ score that is assumed to be $N(0,1)$ under the null.
+You first order these tests randomly, and begin by 
+performing the first test, which returns a $Z$ score of 4.
+At this point you are interrupted by a friend, who asks how the analysis is going.
+"It's early days, but looking promising" you reply.
+Well, who wouldn't?
+If the aim is to find lots of significant differences, a strong first result is surely a good outcome.
+
+At this point you have reason to expect that many of the
+subsequent tests also output strong results.
+
+Now consider two alternative scenarios for the remaining 999,999 tests. In the first scenario, the remaining tests
+produce $Z$ values that fit very well with the null, closely following a standard normal distribution; in the second
+scenario a large proportion of the remaining tests, say 50 percent, show outcomes that lie outside of $[-4,4]$.
+
+If your friend enquired after your analysis again, your
+response would surely differ in the first scenario ("Oh, it didn't pan out so well after all") vs the second ("It went great"). Further,
+in the first scenario, if your friend pressed you further about the results of the first test,
+you would likely, I think, be inclined to put them down to chance.
+In contrast, in the second scenario, the first test turned out to be, as you hoped, a harbinger of good things to come, and in this scenario you would likely regard that test as likely corresponding to a true discovery.
+
+The key point is that it is the *outcomes* of the tests, not the *number* of tests, that impacts interpretation of that first test.
+
+(Some may by pondering whether the fact that you are about to perform another
+999,999 tests should be considered pertinent in responding to your friend. Our view is that it is  irrelevant.
+The standard frequentist framework would disagree, because
+it requires the analyst to consider hypothetical repetitions
+of the "experiment", and so the fact that the experiment
+consists of a million tests is pertinent. However, this
+argument is a distraction from the main point.)
+
+Indeed, we believe that the practice of focussing on the *number* of tests performed is
+
+Focussing on the number of tests performed can be
+seen as an approximation.
+
+The standard argument is that,
+when performing multiple tests, some will be significant
+just by chance.

### 2013-07-15 Rcode/ash.test.R

@@ -6,3 +6,51 @@ eps = 1e-10
 abf[abf==0] = eps #replace 0 with small number
 print(all.equal(VBEM(abf,c(1,1,1,1))$post,c(2,2,4,1)))
 print(all.equal(VBEM(abf,c(1,2,1,1))$post,c(2,3,4,1)))
+
+#simulate n beta-hat values, nnull under the null
+#with altmean and altsd being the mean and sd of beta under the alternative
+simdata=function(n, nnull, altmean, altsd, betasd){
+  null = c(rep(1,nnull),rep(0,n-nnull))
+  beta = c(rep(0,nnull),rnorm(n-nnull,altmean,altsd))
+  betahat = rnorm(n,beta,betasd)
+  return(list(null=null,beta=beta,betahat=betahat,betasd=betasd))
+}
+
+ss = simdata(10000,8000,0,2,1)
+
+system.time((beta.ash=ash(ss$betahat,ss$betasd)))
+system.time((beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)))
+system.time((beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior="uniform")))
+system.time((beta.ash.vb.null = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior=NULL)))
+
+
+
+hist(ss$beta,prob=TRUE,breaks=seq(-7,7,length=20))
+x= seq(-4,4,length=10000)
+lines(x,density(beta.ash,x),col=2)
+lines(x,density(beta.ash.auto,x),col=3)
+lines(x,density(beta.ash.vb.uniform,x),col=4)
+lines(x,density(beta.ash.vb.null,x),col=5)
+beta.ash$fitted.f
+beta.ash.auto$fitted.f
+beta.ash.vb.uniform$fitted.f
+
+ss = simdata(10000,10000,0,2,1)
+system.time((beta.ash=ash(ss$betahat,ss$betasd)))
+system.time((beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)))
+system.time((beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior="uniform")))
+system.time((beta.ash.vb.null = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior=NULL)))
+
[…]

### 2013-07-15 Rcode/intro.rmd
added cdf computation for ash object

@@
 hist(ss$beta,prob=TRUE,breaks=seq(-7,7,length=20))
-x= seq(-4,4,length=1000)
+x= seq(-7,7,length=1000)
 lines(x,density(beta.ash,x),col=2)
+
+plot(sort(ss$beta),(1:length(ss$beta))/length(ss$beta),main="cdf of ss$beta, with fitted f overlaid",xlab="beta",ylab="cdf")
+lines(x,cdf.ash(beta.ash,x),col=2,lwd=2)
+lines(x,cdf.ash(beta.ash.auto,x),col=3,lwd=2)
+lines(x,cdf.ash(beta.ash.vb.uniform,x),col=4,lwd=2)
+lines(x,cdf.ash(beta.ash.vb.null,x),col=5,lwd=2)
 ```
@@
 that means estimating the pi value on the point mass will be somewhat unreliable, and we still might underestimate the false sign rate if we rely on that estimate.  
+TO THINK ABOUT: does multiplying the smaller of Pr(<0) and Pr(>0) by 2, and adding to Pr(=0) solve the problem in either case?
 

### 2013-07-15 Rcode/ash.R

@@ -324,6 +324,14 @@ pnormmix = function(T,pi1,mu1,sigma1,lower.tail=TRUE){
   return(apply(pi1 * pnorm(T,mu1,sigma1,lower.tail),2,sum))
 }
   
+#return posterior of being <T (or >T) for a mixture of Gaussians
+# each of pi1, mu1, sigma1 is a k-vector
+# jth element provides parameters for jth mixture of gauusians 
+# return a probabilities
+pnormmix.vec = function(T,pi1,mu1,sigma1,lower.tail=TRUE){
+  return(sum(pi1 * pnorm(T,mu1,sigma1,lower.tail)))
+}
+
 #return the "effective" estimate
 #that is the effect size betanew whose z score betanew/se
 #would give the same p value as betahat/se compared to a t with df
@@ -458,3 +466,10 @@ predictive=function(a,se){
 #a is an ash object
 #x is a vector at which the density should be computed
 density.ash=function(a,x){mixdnorm(x,a$fitted.f$pi,a$fitted.f$mu,a$fitted.f$sigma)}
+
+#return the cdf of the fitted underlying hierarchical f
+#a is an ash object
+#x is a vector at which the density should be computed
+cdf.ash=function(a,x,lower.tail=TRUE){
+ return(vapply(x,pnormmix.vec, 0,pi1=a$fitted.f$pi,mu1=a$fitted.f$mu,sigma1=a$fitted.f$sigma,lower.tail=lower.tail))    
+}

### 2013-08-28 Rcode/jash_test.rmd
Add things for joint ash

Model
========================================================
The observed values is in a $N\times n$ matrix $Y$ ($N$ tests and $n$ replicate observations for each tests), where \[y_{jk}|\beta_j,\tau_j\sim N(\beta_j,1/\tau_j),\ k=1,2,...,n,\ j=1,2,...,N\]
We assume that $\beta_j$ (mean), $\tau_j$ (precision) come from the Normal-Gamma mixture prior with $M\times L$ components:
\[\beta_j|\tau_j\sim N(\mu_j,1/(c_l\tau_j)),\quad \tau_j\sim Gamma(a_m,b_m)\quad w.p. \pi_{ml}\]
where $\sum_{m,l}\pi_{ml}=1,\ m=1,...,M,\ l=1,...,L$. Here $a_m>0,b_m>0,c_l>0,\mu_j$ are known constants (default setting $\mu_j=0$). Note that $c_l$ can be chosen as $\infty$, and thus result in a point mass at zero in $\beta$'s prior.

Posterior distribution
=======================
Hence, the conjugate postrior dis
tribution of $(\beta_j,\tau_j)$ given $Y$ ias also a Normal-Gamma mixture distribution, i.e. $\tau_j|Y\sim Gamma(\tilde{a},\tilde{b})$, $\beta_j|\tau_j,Y\sim N(\tilde{mu},1/\tilde{\tau})$. The full posterior distribution for $\beta_j$ is $P(\beta_j|Y)=\int P(\beta_j|\tau_j,Y)P(\tau_j|Y)d\tau_j$, which is hard to compute. For convenience, we can try using the distribution of $\beta_j|\hat{\tau_j},Y$ for further inference, where $\hat{\tau_j}=E(\tau_j|Y)$. Later we will compare the sampled $P(\beta_j|Y)$ to $P(\beta_j|\hat{\tau_j},Y)$.

Note: $P(\beta_j|Y)$ is feasible to compute, though very messy. I am working on it...

Estimate hyper-parameters
=======================
We estimate the hyper-parameters $\pi$ by maximizing the likelihood $L(\pi;Y)=\int P(Y,\beta,\tau|Y)d\beta d\tau$.

Note: I find that sometimes if $a_m,b_m,c_l$ are not well chosen (e.g. the resulting $Var(Y|\beta,\tau)$ for each components overlap), then it is hard for the EM algorithm to converge.

FDR and q-values
================
Same as ash, we use false sign rate for inference.

Some examples
=============

```{r}
setwd("/Volumes/PERSONAL/MS")
set.seed(22222)
source("../Rcode/ash.R") 
source("../Rcode/jash.R")

#simulate N observations with n replicates for each, Nnull under the null
#with altmean and altsd being the mean and sd of beta under the alternative
simdata = function(N, Nnull, altmean, altsd, betasd, n) {
  null = c(rep(1, Nnull), rep(0, N - Nnull))
  beta = c(rep(0, Nnull), rnorm(N - Nnull, altmean, altsd))
  Y = matrix(rnorm(N*n, rep(beta,n), rep(betasd,n)), ncol=n)
  return(list(null = null, beta = beta, Y = Y))
}
```

Simulate 10000 tests, with 2000 alternatives, β∼N(0,sd=3), and 8000 null values (β=0).
```{r}
ss = simdata(10000, 8000, 0, 3, 1, 3)
beta.jash = jash(ss$Y)
beta.jash.auto = jash(ss$Y,auto=TRUE)
```

We use ash for comparison. Here we use the test-wise sample-mean and sqrt(sample-variance/n) as the inputs of ash. (So the sebetahat here is not the true standard error of betahat!)
```{r}
betahat = apply(ss$Y,1,mean)
sebetahat = apply(ss$Y,1,sd)/sqrt(dim(ss$Y)[2])
beta.ash.auto = ash(betahat,sebetahat,auto=TRUE)
```

The histogram of true $\beta$:
```{r}
hist(ss$beta, prob = TRUE, breaks = seq(-8, 8, length = 20))
```

Both jash and ash give the Posterior means $E(\beta|Y)$, compare the shrinked means using two methods:
```{r}
plot(betahat,beta.jash$PosteriorMean,xlab='Observed betahat',ylab='Posterior mean',cex=0.5,pch=20)
points(betahat,beta.ash.auto$PosteriorMean, col=2, cex=0.5,pch=20)
title('Black: jash; Red: ash.auto')
```

Compare the sampled full posterior distribution $P(\beta|Y)$ to the conditional posterior distribution $P(\beta|\hat{tau},Y)$, where $\hat{tau}=E(\tau|Y)$. Take the first two tests for illustration. It seems that in this case, the conditional distribution fits the full distribution well. (However, I am worried that it will make difference when $\tau$ has large scale or wide range. NEED TESTING!)

```{r}
# Compare sampled distribution P(beta|Y) to the normal mixture distribution P(beta|Y,hat(tau))
# Red lines are the conditional posterior distribution P(beta|Y,hat(tau)) 
par(mfrow=c(1,2))
sample.jash = posterior_sample_jash(beta.jash$post,1000,1:2)
hist(sample.jash$beta[1,],prob = TRUE, breaks = seq(-2, 2, length = 20),ylim=c(0,2.5),main='First test')
x = seq(-2, 2, length = 1000)
normdens = dnorm(outer(x,rep(1,length(beta.jash$a.vec))),mean=beta.jash$condpost$normmean[rep(1,1000),],sd=beta.jash$condpost$normsd[rep(1,1000),])
mixnormdens = rowSums(beta.jash$condpost$condpi[rep(1,1000),]*normdens)
lines(x,mixnormdens,type='l',col=2)

hist(sample.jash$beta[2,],prob = TRUE, breaks = seq(-2.5, 2.5, length = 20),ylim=c(0,2.5),main='Second test')
x = seq(-2.5, 2.5, length = 1000)
normdens = dnorm(outer(x,rep(1,length(beta.jash$a.vec))),mean=beta.jash$condpost$normmean[rep(2,1000),],sd=beta.jash$condpost$normsd[rep(2,1000),])
mixnormdens = rowSums(beta.jash$condpost$condpi[rep(2,1000),]*normdens)
lines(x,mixnormdens,type='l',col=2) 
```

Compare the actual FDR and estimated FDR for ash, jash, jash.auto (a way of automatically selecting $a_m$,$b_m$,$c_l$). It seems that jash performs a bit better than ash (which makes sense, since the estimated sebetahat for ash is not accurate).
```{r}
# Compare ash.auto, jash, jash.auto
o.ash = order(beta.ash.auto$qval)
o.jash = order(beta.jash$qval)
o.jash.auto = order(beta.jash.auto$qval)
par(mfrow=c(1,1))
plot(cumsum(ss$null[o.jash])/(1:10000),type='l',ylim=c(0,1),ylab='FDR' )
lines(cumsum(ss$null[o.ash])/(1:10000), col=2)
lines(cumsum(ss$null[o.jash.auto])/(1:10000), col='blue')
lines(sort(beta.jash$qvalue),lty=2)
lines(sort(beta.ash.auto$qvalue),lty=2,col=2)
lines(sort(beta.jash.auto$qvalue),lty=2,col='blue')
title('Dotted line: estimated FDR; Solid line: actual FDR')
legend('topleft',col=c('black','blue','red'),lty=c(1,1,1),legend=c('jash','jash.auto','ash.auto'))
```

The fitted $\pi$ for jash: (sdvec gives the theoretical $SD(y_{jk}|a_m,b_m,c_l)$ for each component)
```{r}
beta.jash$fitted
bata.jash.auto$fitted
```

We can also have a point mass at zero for $\beta$'s prior, by adding components with $c_l=\infty$. Compare the jash results with or without point mass:
```{r}
# Use point mass
beta.jash.pm = jash(ss$Y,usePointMass=TRUE)
#beta.ash.pm = ash(betahat,sebetahat,auto=TRUE,usePointMass=TRUE)
beta.jash$fitted
beta.jash.pm$fitted
o.jash.pm = order(beta.ash.pm$qvalue)
plot(cumsum(ss$null[o.jash])/(1:10000),type='l',ylim=c(0,1),ylab='FDR' )
lines(cumsum(ss$null[o.jash.pm])/(1:10000), col=2)
lines(sort(beta.jash$qvalue),lty=2)
lines(sort(beta.jash.pm$qvalue),lty=2,col=2)
title('Dotted line: estimated FDR; Solid line: actual FDR')
legend('topright',col=c('black','red'),lty=c(1,1),legend=c('jash','jash.pm'))
```

For ash, maybe the ``df'' option can resolve the understimating FDR issue? It works to some extent, but still not very satisfying.
```{r}
# Use df modification for ash
beta.ash.t = ash(betahat,sebetahat,auto=TRUE,df=2)
o.ash.t = order(beta.ash.t$qval)
plot(cumsum(ss$null[o.jash])/(1:10000),type='l',ylim=c(0,1),ylab='FDR' )
lines(cumsum(ss$null[o.ash.t])/(1:10000), col=2)
lines(sort(beta.jash$qvalue),lty=2)
lines(sort(beta.ash.t$qvalue),lty=2,col=2)
title('Dotted line: estimated FDR; Solid line: actual FDR')
legend('topright',col=c('black','red'),lty=c(1,1),legend=c('jash','ash.t'))
```

If the there are many replicates for each test, ash should yield pretty good results? Yes, it is. (But I am trying to figure out why jash performs this bad, maybe it has something to do with the settings of $a_m,b_m,c_l$)
```{r}
ss.largen = simdata(10000, 8000, 0, 3, 1, 20)
beta.largen.jash = jash(ss.largen$Y)
beta.largen.jash.auto = jash(ss.largen$Y,auto=TRUE)

betahat.largen = apply(ss.largen$Y,1,mean)
sebetahat.largen = apply(ss.largen$Y,1,sd)/sqrt(dim(ss.largen$Y)[2])
beta.largen.ash = ash(betahat.largen,sebetahat.largen,auto=TRUE)

o.largen.ash = order(beta.largen.ash$qval)
o.largen.jash = order(beta.largen.jash$qvalue)
o.largen.jash.auto = order(beta.largen.jash.auto$qvalue)
plot(cumsum(ss$null[o.largen.jash])/(1:10000),type='l',ylim=c(0,1),ylab='FDR' )
lines(cumsum(ss$null[o.largen.ash])/(1:10000), col=2)
lines(cumsum(ss$null[o.largen.jash.auto])/(1:10000), col='blue')
lines(sort(beta.largen.jash$qvalue),lty=2)
lines(sort(beta.largen.ash$qvalue),lty=2,col=2)
lines(sort(beta.largen.jash.auto$qvalue),lty=2,col='blue')
title('Large num of replicates case')
legend('topleft',col=c('black','blue','red'),lty=c(1,1,1),legend=c('jash','jash.auto','ash.auto'))
```

Try some different data: the issue of underestimating FDR is still noticable.
```{r}
truenull = c(rep(0, 1000), rep(1, 9000))
beta = c(rnorm(1000, -3, 1), rep(0, 9000))
s = rep(1, 10000)
Y = matrix(rnorm(30000, rep(beta,3), rep(s,3)),ncol=3)
Y.betahat = apply(Y,1,mean)
Y.sebetahat = apply(Y,1,sd)/sqrt(dim(Y)[2])
Y.ash.auto = ash(Y.betahat,Y.sebetahat,auto=TRUE)
Y.jash = jash(Y)
Y.jash.auto = jash(Y,auto=TRUE)

o = order(Y.jash.auto$qvalue)
plot(cumsum(truenull[o])/(1:10000), Y.jash.auto$qvalue[o], col = 2, type = "l",ylim=c(0,0.9))
abline(a = 0, b = 1)
```

Works undone
=============
* Speed up the function!
* Derive full posterior distribution $P(\beta|Y)$
* Better ways to (automatically) select $a_m,b_m,c_l$?
* Issue of underestimating false sign rate
* Use other approaches to estimate $\pi$ (e.g. VB?)
* Try more data (with different scales, underlying distributions, etc)

### 2013-08-28 Rcode/jash.r
Add things for joint ash

@@ -1,0 +1,206 @@
+source('ash.R')
+
+# Y: N by n observation matrix (N tests, n replicate obs for each test)
+# Model: Y_{jk}|beta_j,tau_j ~ N(beta_j,1/tau_j)
+# Prior: beta_j|tau_j~N(mu_j,1/(c_l*tau_j)), tau_j~Gamma(a_m,b_m), w.p. pi_{ml}
+# Mixture constraint: sum(pi_{ml})=1
+# a_m, b_m, c_l, mu_j are known
+######################################################
+
+# post_pi computes the posterior probability 
+# P(beta_j,tau_j belongs to the i'th component | Y)
+post_pi = function(Y, precPrior, compprecPrior, mu, pi){
+  N = dim(Y)[1]
+  n = dim(Y)[2]
+  M = dim(precPrior)[1]
+  L = length(compprecPrior)
+  
+  a.vec = rep(precPrior[,1],L)
+  b.vec = rep(precPrior[,2],L)
+  c.vec = rep(compprecPrior,each=M)
+  Y.mean = apply(Y,1,mean)
+  Y.sst = apply(Y,1,var)*(n-1)
+  
+  post.pi.mat = outer(rep(1,N),pi*b.vec^a.vec/gamma(a.vec)*sqrt(1/(n/c.vec+1))*gamma(a.vec+n/2))/
+    (0.5*(outer(Y.sst,rep(1,M*L))+outer((Y.mean-mu)^2,n/(n/c.vec+1))+outer(rep(1,N),2*b.vec)))^
+    outer(rep(1,N),a.vec+n/2)
+  return(post.pi.mat)
+}
+
+# Normalize pi to make sum(pi)=1
+normalized_pi = function(pi.mat){
+  n = dim(pi.mat)[2]
+  pi.normalized = pi.mat/outer(rowSums(pi.mat),rep(1,n))  
+  return(pi.normalized)
+}
+
+# Compute posterior distribution P(tau|Y), P(beta|Y,tau)
+# precPrior: M by 2 matrix, all pairs of (a_m, b_m) (a_m>0, b_m>0)
+# compprecPrior: L vector, all c_l (c_l>=0)
+# pi: (M*L) vector
+# mu: N vector
+post_distn = function(Y, precPrior, compprecPrior, mu, pi){
+  N = dim(Y)[1]
+  n = dim(Y)[2]
+  M = dim(precPrior)[1]
+  L = length(compprecPrior)
+  
+  a.vec = rep(precPrior[,1],L)
[…]

### 2013-08-28 Rcode/jash_test.R

@@ -1,0 +1,132 @@
+set.seed(22222)
+source('ash.R')
+source('jash.R')
+
+# N: num of obs
+# n: num of replicates
+simdata = function(N, Nnull, altmean, altsd, betasd, n) {
+  null = c(rep(1, Nnull), rep(0, N - Nnull))
+  beta = c(rep(0, Nnull), rnorm(N - Nnull, altmean, altsd))
+  Y = matrix(rnorm(N*n, rep(beta,n), rep(betasd,n)), ncol=n)
+  return(list(null = null, beta = beta, Y = Y))
+}
+
+ss = simdata(10000, 8000, 0, 3, 1, 3)
+beta.jash = jash(ss$Y)
+beta.jash.auto = jash(ss$Y,auto=TRUE)
+
+betahat = apply(ss$Y,1,mean)
+sebetahat = apply(ss$Y,1,sd)/sqrt(dim(ss$Y)[2])
+beta.ash.auto = ash(betahat,sebetahat,auto=TRUE)
+
+# Histogram of true beta
+hist(ss$beta, prob = TRUE, breaks = seq(-8, 8, length = 20))
+
+plot(betahat,beta.jash$PosteriorMean,xlab='Observed betahat',ylab='Posterior mean',cex=0.5,pch=20)
+points(betahat,beta.ash.auto$PosteriorMean, col=2, cex=0.5,pch=20)
+title('Black: jash; Red: ash.auto')
+
+# Compare sampled distribution P(beta|Y) to the normal mixture distribution P(beta|Y,hat(tau))
+# Red lines are the conditional posterior distribution P(beta|Y,hat(tau)) 
+par(mfrow=c(1,2))
+sample.jash = posterior_sample_jash(beta.jash$post,1000,1:2)
+hist(sample.jash$beta[1,],prob = TRUE, breaks = seq(-2, 2, length = 20),ylim=c(0,2.5),main='First unit')
+x = seq(-2, 2, length = 1000)
+normdens = dnorm(outer(x,rep(1,length(beta.jash$a.vec))),mean=beta.jash$condpost$normmean[rep(1,1000),],sd=beta.jash$condpost$normsd[rep(1,1000),])
+mixnormdens = rowSums(beta.jash$condpost$condpi[rep(1,1000),]*normdens)
+lines(x,mixnormdens,type='l',col=2) 
+
[…]

### 2013-09-29 Rcode/intro.rmd
added first attempt at functions to compute convolution densities and cdfs for unimix to mix.R

@@
 ```{r}
-beta.ash$fitted.f
-beta.ash.auto$fitted.f
-beta.ash.vb.uniform$fitted.f
-beta.ash.vb.null$fitted.f
+print(beta.ash)
+print(beta.ash.auto)
+print(beta.ash.vb.uniform)
+print(beta.ash.vb.null)
 ```
@@
 beta.ash.pm = ash(ss$betahat, ss$betasd, auto=TRUE, usePointMass=TRUE)
-beta.ash.pm$fitted.f
-beta.ash.auto$fitted.f
+print(beta.ash.pm)
+print(beta.ash.auto)
 plot(beta.ash.auto$localfsr,beta.ash.pm$localfsr,main="comparison of ash localfsr, with and without point mass",xlab="no point mass", ylab="with point mass",xlim=c(0,1),ylim=c(0,1))

### 2013-09-29 Rcode/mix.R
added first attempt at functions to compute convolution densities and cdfs for unimix to mix.R

@@ -22,6 +22,14 @@ matdens.default = function(x,y,log=FALSE){
   stop("No such class")
 }
 
+#find cdf at y, a generic function
+mixcdf = function(x,y,lower.tail=TRUE){
+  UseMethod("mixcdf")
+}
+mixcdf.default = function(x,y,lower.tail=TRUE){
+  stop("No such class")
+}
+
 #find density at y, a generic function
 dens = function(x,y){
   UseMethod("dens")
@@ -37,6 +45,67 @@ LogLik = function(x,y){
 LogLik.default = function(x,y){
   sum(dens(x,y))
 }
+
+#compute the density of the components of the mixture m
+#when convoluted with a normal with standard deviation s
+#the density is evaluated at x
+#x and s are n-vectors
+#m is a mixture with k components
+#output is a k by n matrix of densities
+compdens_conv = function(m, x, s){
+  UseMethod("compdens_conv")
+}
+compdens_conv.default = function(m,x, s){
+  stop("No such class")
+}
+
+#compute density of mixture m convoluted with normal of sd (s)
+#at locations x
+#m is a mixture
+#x is an n vector
+#s is an n vector or integer
+dens_conv = function(m,x,s){
+  UseMethod("dens_conv")
+}
+dens_conv.default = function(m,x,s){
+  colSums(m$pi * compdens_conv(m,x,s))
+}
+
+#compute the posterior prob that each observation
+#came from each component of the mixture m
+#output a k by n vector of probabilities
+#computed by weighting the component densities by pi
+#and then normalizing
+comppostprob=function(m,x,s){
+ UseMethod("comppostprob") 
+}
+comppostprob.default = function(m,x,s){
+  t(t(m$pi * compdens_conv(m,x,s))/dens_conv(m,x,s))
+}
+# evaluate cdf of posterior distribution of beta at c
+# m is the prior on beta, a mixture
+# c is location of evaluation
+# assumption is betahat | beta \sim N(beta,sebetahat)
+# m is a mixture with k components
+# c a scalar
+# betahat, sebetahat are n vectors 
+# output is a k by n matrix
+compcdf_post=function(m,c,betahat,sebetahat){
+  UseMethod("compcdf_post")
[…]

### 2013-10-01 Rcode/test.mix.Rmd
added test.mix.Rmd; renamed ash.test.R as test.ash.Rmd; updated VBEM back to output post as well as pihat

################################# TESTING #####################################
```{r}
source("mix.R")
temp = normalmix(c(0.5,0.5),c(-3,2),c(1,1))
x = seq(-5,5,length=100)
plot(x,dens(temp,x),type="l")
all.equal(compdens(temp,x),matdnorm(x, temp$mean,temp$sd))

plot(x,mixcdf(temp,x),type="l")

plot(x,compdens_conv(temp,x,0.01)[1,],type="l")
plot(x,compdens_conv(temp,x,0.1)[2,],type="l")
plot(x,dens_conv(temp,x,0.2),type="l")
plot(x,dens_conv(temp,x,0.001),type="l")
plot(x,dens_conv(temp,x,10),type="l")

```

```{r}

temp2 = unimix(c(0.5,0.5),c(-3,3),c(-1,4))
plot_dens(temp2)

plot(x,compdens_conv(temp2,x,0.01)[1,],type="l")
plot(x,compdens_conv(temp2,x,0.1)[2,],type="l")
plot(x,dens_conv(temp2,x,0.2),type="l")

compcdf_post(temp2,0,c(1,2,3),c(10,10,10))
compcdf_post(temp2,-2,c(-2,3),c(0.1,1))

plot_post_cdf(temp2,betahat=c(-2),sebetahat=10)
plot_post_cdf(temp2,betahat=c(-2),sebetahat=1,col=2)
```

### 2013-10-01 Rcode/test.ash.Rmd
added test.mix.Rmd; renamed ash.test.R as test.ash.Rmd; updated VBEM back to output post as well as pihat

```{r}

source("ash.R")
set.seed(100)

#Test VBEM
abf = rbind(c(1,0,0,0),c(0,1,0,0),c(0,0,1,0),c(0,0,1,0),c(0,0,1,0))
eps = 1e-10
abf[abf==0] = eps #replace 0 with small number
print(all.equal(VBEM(abf,c(1,1,1,1))$post,c(2,2,4,1)))
print(all.equal(VBEM(abf,c(1,2,1,1))$post,c(2,3,4,1)))

#simulate n beta-hat values, nnull under the null
#with altmean and altsd being the mean and sd of beta under the alternative
simdata=function(n, nnull, altmean, altsd, betasd){
  null = c(rep(1,nnull),rep(0,n-nnull))
  beta = c(rep(0,nnull),rnorm(n-nnull,altmean,altsd))
  betahat = rnorm(n,beta,betasd)
  return(list(null=null,beta=beta,betahat=betahat,betasd=betasd))
}

ss = simdata(10000,8000,0,2,1)

system.time((beta.ash=ash(ss$betahat,ss$betasd)))
system.time((beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)))
system.time((beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior="uniform")))
system.time((beta.ash.vb.null = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior=NULL)))

hist(ss$beta,prob=TRUE,breaks=seq(-7,7,length=20))
x= seq(-4,4,length=10000)
lines(x,density(beta.ash,x),col=2)
lines(x,density(beta.ash.auto,x),col=3)
lines(x,density(beta.ash.vb.uniform,x),col=4)
lines(x,density(beta.ash.vb.null,x),col=5)
beta.ash$fitted.f
beta.ash.auto$fitted.f
beta.ash.vb.uniform$fitted.f

ss = simdata(10000,10000,0,2,1)
system.time((beta.ash=ash(ss$betahat,ss$betasd)))
system.time((beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)))
system.time((beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior="uniform")))
system.time((beta.ash.vb.null = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior=NULL)))

hist(ss$beta,prob=TRUE,breaks=seq(-7,7,length=20))
x= seq(-4,4,length=10000)
lines(x,density(beta.ash,x),col=2)
lines(x,density(beta.ash.auto,x),col=3)
lines(x,density(beta.ash.vb.uniform,x),col=4)
lines(x,density(beta.ash.vb.null,x),col=5)
beta.ash$fitted.f
beta.ash.auto$fitted.f
beta.ash.vb.uniform$fitted.f
beta.ash.vb.null$fitted.f

beta.ash.vb.uniform$fit$loglik
beta.ash.vb.uniform$fit$converged

```

### 2013-10-01 Rcode/ash.R

@@ -123,7 +123,7 @@ VBEM = function(matrix_lik, prior, tol=0.0001, maxiter=5000){
   
   if(i>maxiter){i=maxiter}
    
-  return(list(pihat = pipost/sum(pipost), B=B[1:i], niter = i, converged=(i<maxiter)))
+  return(list(pihat = pipost/sum(pipost), B=B[1:i], niter = i, converged=(i<maxiter),post=pipost))
 }
   
 

### 2013-10-01 Rcode/test.ash.Rmd
updated ash to use normalmix methods rather in mix.R to compute posterior mean, PositiveProb, NegativeProb etc

@@
 ```{r}
-
 source("ash.R")
 set.seed(100)
-
 #Test VBEM
@@
 print(all.equal(VBEM(abf,c(1,2,1,1))$post,c(2,3,4,1)))
+
+```
+
+```{r}
+
+source("ash.R")
+set.seed(100)
 
@@
 
+```
+
+```{r}
 ss = simdata(10000,8000,0,2,1)
+beta.ash=ash(ss$betahat,ss$betasd)
 
+system.time((post = posterior_dist(beta.ash$fitted.g,ss$betahat,rep(ss$betasd,length(ss$betahat)))))
+system.time((PositiveProb = pnormmix(0,post$pi,post$mu,post$sigma,lower.tail=FALSE)))
+
+system.time((PP2 = 1-cdf_post(beta.ash$fitted.g, 0, ss$betahat,rep(ss$betasd,length(ss$betahat)))))    
+all.equal(PP2,PositiveProb)
+```
+
+```{r}
 system.time((beta.ash=ash(ss$betahat,ss$betasd)))
@@
 lines(x,density(beta.ash.vb.null,x),col=5)
-beta.ash$fitted.f
-beta.ash.auto$fitted.f
-beta.ash.vb.uniform$fitted.f
-beta.ash.vb.null$fitted.f
-
-beta.ash.vb.uniform$fit$loglik
-beta.ash.vb.uniform$fit$converged
-
+summary(beta.ash)
+summary(beta.ash.auto)
+summary(beta.ash.vb.uniform)
+summary(beta.ash.vb.null)
 ```

### 2013-10-01 Rcode/mix.R

@@ -1,16 +1,3 @@
-############## HELPER FUNCTIONS #################################################
-
-#OUTPUT: k by n matrix of normal densities
-matdnorm = function (x, mu, sigma, log=FALSE) 
-{
-  k=length(mu)
-  n=length(x)
-  d = matrix(rep(x,rep(k,n)),nrow=k)
-  return(matrix(dnorm(d, mu, sigma, log),nrow=k))
-}
-
-
-
 ################################## GENERIC FUNCTIONS ############################
 # find matrix of densities at y, for each component of the mixture
 # INPUT y is an n-vector
@@ -105,6 +92,27 @@ cdf_post.default=function(m,c,betahat,sebetahat){
   colSums(comppostprob(m,betahat,sebetahat)*compcdf_post(m,c,betahat,sebetahat))
 }
 
+#output posterior mean for beta for prior mixture m,
+#given observations betahat, sebetahat
+postmean = function(m, betahat,sebetahat){
+  UseMethod("postmean")
+}
+postmean.default = function(m,betahat,sebetahat){
+   colSums(comppostprob(m,betahat,sebetahat) * comp_postmean(m,betahat,sebetahat))
+}
+
+
+
+#output posterior mean for beta for each component of prior mixture m,
+#given observations betahat, sebetahat
+comp_postmean = function(m, betahat,sebetahat){
+  UseMethod("comp_postmean")
+}
+comp_postmean.default = function(m,betahat,sebetahat){
+  stop("method comp_postmean not written for this class")
+}
+
+
 #find nice limits of mixture m for plotting
 min_lim = function(m){
   UseMethod("min_lim")
@@ -171,6 +179,29 @@ compdens_conv.normalmix = function(m, x, s){
 mixcdf.normalmix = function(x,y,lower.tail=TRUE){
   x$pi %*%  vapply(y,pnorm,x$mean,x$mean,x$sd,lower.tail)
 }
+
+#c is a scalar
+#m a mixture with k components
+#betahat a vector of n observations
+#sebetahat an n vector of standard errors
+#return a k by n matrix of the posterior cdf
+compcdf_post.normalmix=function(m,c,betahat,sebetahat){
+  k = length(m$pi)
+  n=length(betahat)
+  #compute posterior standard deviation (s1) and posterior mean (m1)
+  s1 = sqrt(outer(sebetahat^2,m$sd^2,FUN="*")/outer(sebetahat^2,m$sd^2,FUN="+"))
[…]

### 2013-10-01 Rcode/test.mix.Rmd
updated ash to use normalmix methods rather in mix.R to compute posterior mean, PositiveProb, NegativeProb etc
added computing of posterior mean for uniform prior; however it is not yet numerically stable when the prior and the likelihood disagree

@@
+```{r}
+############## HELPER FUNCTIONS #################################################
+#OUTPUT: k by n matrix of normal densities
+matdnorm = function (x, mu, sigma, log=FALSE) 
+{
+  k=length(mu)
+  n=length(x)
+  d = matrix(rep(x,rep(k,n)),nrow=k)
+  return(matrix(dnorm(d, mu, sigma, log),nrow=k))
+}
+```
+
 ################################# TESTING #####################################
@@
 ```{r}
-
 temp2 = unimix(c(0.5,0.5),c(-3,3),c(-1,4))
@@
 ```
+
+```{r}
+  comp_postmean(temp2, c(1,2,3), c(10,10,10))
+  comp_postmean(temp2, c(1,2,3), c(0.1,0.1,0.1))
+```

### 2013-10-01 Rcode/mix.R
added computing of posterior mean for uniform prior; however it is not yet numerically stable when the prior and the likelihood disagree

@@ -1,3 +1,5 @@
+library(truncnorm)
+
 ################################## GENERIC FUNCTIONS ############################
 # find matrix of densities at y, for each component of the mixture
 # INPUT y is an n-vector
@@ -7,6 +9,14 @@ compdens = function(x,y,log=FALSE){
 }
 compdens.default = function(x,y,log=FALSE){
   stop("No such class")
+}
+
+#number of components
+ncomp = function(m){
+  UseMethod("ncomp")
+}
+ncomp.default = function(m){
+  return(length(m$pi))
 }
 
 #find cdf at y, a generic function
@@ -159,7 +169,7 @@ normalmix = function(pi,mean,sd){
 
 
 compdens.normalmix = function(x,y,log=FALSE){
-  k=length(x$mean)
+  k=ncomp(x)
   n=length(y)
   d = matrix(rep(y,rep(k,n)),nrow=k)
   return(matrix(dnorm(d, x$mean, x$sd, log),nrow=k))  
@@ -214,7 +224,7 @@ unimix = function(pi,a,b){
 }
 
 compdens.unimix = function(x,y,log=FALSE){
-  k=length(x$a)
+  k=ncomp(x)
   n=length(y)
   d = matrix(rep(y,rep(k,n)),nrow=k)
   return(matrix(dunif(d, x$a, x$b, log),nrow=k))  
@@ -247,6 +257,34 @@ compcdf_post.unimix=function(m,c,betahat,sebetahat){
     tmp[subset,] = t((pnc-pna)/(pnb-pna))
   }
   tmp
+}
+
+#return posterior mean for each component of prior m, given observations betahat and sebetahat
+#input, m is a mixture with k components
+#betahat, sebetahat are n vectors
+#output is a k by n matrix
+#note that with uniform prior, posterior is truncated normal, so
+#this is computed using formula for mean of truncated normal 
+comp_postmean.unimix = function(m,betahat,sebetahat){
+#   k= ncomp(m)
+#   n=length(betahat)
+#   a = matrix(m$a,nrow=n,ncol=k,byrow=TRUE)
+#   b = matrix(m$b,nrow=n,ncol=k,byrow=TRUE)
+#   matrix(etruncnorm(a,b,betahat,sebetahat),nrow=k,byrow=TRUE)
+  #note: etruncnorm is more stable for a and b negative than positive
[…]

### 2013-10-01 Rcode/mix.R
added my_etruncnorm to improve numerical stability when [a,b] is far from the normal

@@ -259,6 +259,25 @@ compcdf_post.unimix=function(m,c,betahat,sebetahat){
   tmp
 }
 
+my_etruncnorm= function(a,b,mean=0,sd=1){
+  alpha = (a-mean)/sd
+  beta =  (b-mean)/sd
+ #Flip the onese where both are positive, as the computations are more stable
+  #when both negative
+  flip = (alpha>0 & beta>0)
+  alpha[flip]= -alpha[flip]
+  beta[flip]=-beta[flip]
+  
+  tmp= (-1)^flip * (mean+sd*etruncnorm(alpha,beta,0,1))
+  
+  max_alphabeta = ifelse(alpha<beta, beta,alpha)
+  max_ab = ifelse(alpha<beta,b,a)
+  toobig = max_alphabeta<(-30)
+  tmp[toobig] = max_ab[toobig]
+  tmp
+}
+  
+  
 #return posterior mean for each component of prior m, given observations betahat and sebetahat
 #input, m is a mixture with k components
 #betahat, sebetahat are n vectors
@@ -276,15 +295,15 @@ comp_postmean.unimix = function(m,betahat,sebetahat){
   
   alpha = outer(-betahat, m$a,FUN="+")/sebetahat
   beta = outer(-betahat, m$b, FUN="+")/sebetahat
-
-  t(
-    betahat + sebetahat* 
-      exp(dnorm(alpha,log=TRUE)- pnorm(alpha,log=TRUE))
-   * 
-      (-expm1(dnorm(beta,log=TRUE)-dnorm(alpha,log=TRUE)))
-    /
-      (expm1(pnorm(beta,log=TRUE)-pnorm(alpha,log=TRUE)))
-  )
+  t(betahat + sebetahat*my_etruncnorm(alpha,beta,0,1))
+#   t(
+#     betahat + sebetahat* 
+#       exp(dnorm(alpha,log=TRUE)- pnorm(alpha,log=TRUE))
+#    * 
+#       (-expm1(dnorm(beta,log=TRUE)-dnorm(alpha,log=TRUE)))
+#     /
+#       (expm1(pnorm(beta,log=TRUE)-pnorm(alpha,log=TRUE)))
+#   )
 }
 
 

### 2013-10-02 Rcode/intro.rmd
fixed bug in NegativeProb computation to exclude 0

@@
 
+beta.ash.pm = ash(ss$betahat,ss$betasd,usePointMass=TRUE)
+
 #compute the usual zscore and corresponding p value
@@
 print(beta.ash.vb.null)
+print(beta.ash.pm)
 ```
@@
 lines(x,cdf.ash(beta.ash.vb.null,x),col=5,lwd=2)
+lines(x,cdf.ash(beta.ash.pm,x),col=6,lwd=2)
+```
+
+[for testing: compare results with point mass and without]
+```{r}
+  plot(beta.ash$PositiveProb)
+  points(beta.ash.pm$PositiveProb,col=2)
+```
+
+```{r}
+  plot(beta.ash$ZeroProb)
+  points(beta.ash.pm$ZeroProb,col=2)
 ```

### 2013-10-02 Rcode/mix.R
fixed bug in NegativeProb computation to exclude 0

@@ -11,6 +11,14 @@ compdens.default = function(x,y,log=FALSE){
   stop("No such class")
 }
 
+#standard deviations
+comp_sd = function(m){
+  UseMethod("comp_sd")
+}
+comp_sd.default = function(m){
+  stop("method comp_sd not written for this class")
+}
+
 #number of components
 ncomp = function(m){
   UseMethod("ncomp")
@@ -167,6 +175,9 @@ normalmix = function(pi,mean,sd){
   structure(data.frame(pi,mean,sd),class="normalmix")
 }
 
+comp_sd.normalmix = function(m){
+  m$sd
+}
 
 compdens.normalmix = function(x,y,log=FALSE){
   k=ncomp(x)
@@ -223,6 +234,11 @@ unimix = function(pi,a,b){
   structure(data.frame(pi,a,b),class="unimix")
 }
 
+
+comp_sd.unimix = function(m){
+  (m$b-m$a)/sqrt(12)
+}
+
 compdens.unimix = function(x,y,log=FALSE){
   k=ncomp(x)
   n=length(y)
@@ -252,7 +268,7 @@ compcdf_post.unimix=function(m,c,betahat,sebetahat){
   subset = m$a<c & m$b>c # subset of components (1..k) with nontrivial cdf
   if(sum(subset)>0){
     pna = pnorm(outer(betahat,m$a[subset],FUN="-")/sebetahat)
-    pnc = pnorm(outer(betahat,c,FUN="-")/sebetahat)
+    pnc = pnorm(outer(betahat,rep(c,sum(subset)),FUN="-")/sebetahat)
     pnb = pnorm(outer(betahat,m$b[subset],FUN="-")/sebetahat)
     tmp[subset,] = t((pnc-pna)/(pnb-pna))
   }

### 2013-10-02 Rcode/test.ash.Rmd
dealt with missing data for betahat and betahatsd in computation of ZeroProb, PosteriorMean, NegativeProb

@@
 ```{r}
+beta.ash=ash(ss$betahat,ss$betasd,auto=TRUE)
+beta.ash.u=ash(ss$betahat,ss$betasd,mixcompdist="uniform",auto=TRUE)
+beta.ash.halfu=ash(ss$betahat,ss$betasd,mixcompdist="halfuniform",auto=TRUE)
+zscore = ss$betahat/ss$betasd
+
+pval = pchisq(zscore^2,df=1,lower.tail=F)
+qval = qvalue(pval)
+hist(zscore)
+x=seq(min(ss$beta),max(ss$beta),length=1000)
+hist(ss$beta,prob=TRUE)
+lines(x,density(beta.ash,x),col=2)
+lines(x,density(beta.ash.u,x),col=3)
+lines(x,density(beta.ash.halfu,x),col=4)
+```
+
+```{r}
+#Test missing data
+ss = simdata(10000,8000,0,2,1)
+
+beta.ash=ash(c(ss$betahat,rep(NA,100)),1)
+beta.ash.u=ash(c(ss$betahat,rep(NA,100)),1,mixcompdist="uniform")
+beta.ash.halfu=ash(c(ss$betahat,rep(NA,100)),1,mixcompdist="halfuniform")
+
+```
+
+```{r}
 system.time((beta.ash=ash(ss$betahat,ss$betasd)))

### 2013-10-02 Rcode/ash.R

@@ -381,6 +381,7 @@ autoselect.sigmaavec = function(betahat,sebetahat){
 #fits a mixture of normals to it
 # and returns posteriors
 #INPUT: betahat (p vector); sebetahat (p vector of standard errors)
+#mixcompdist: distribution of components in mixture ("normal", "uniform" or "halfuniform")
 #df: degrees of freedome used to compute sebetahat
 #randomstart: bool, indicating whether to initialize EM randomly
 #usePointMass: bool, indicating whether to use a point mass at zero as one of components for a mixture distribution
@@ -395,16 +396,19 @@ autoselect.sigmaavec = function(betahat,sebetahat){
 #Things to do:
 # check sampling routine
 # check number of iterations
-ash = function(betahat,sebetahat,nullcheck=TRUE,df=NULL,randomstart=FALSE, usePointMass = FALSE, onlylogLR = FALSE, localfdr = TRUE, localfsr = TRUE, prior=NULL, sigmaavec=NULL, auto=FALSE, sigma.est=FALSE, nc=NULL, VB=FALSE){
+ash = function(betahat,sebetahat,mixcompdist = "normal",nullcheck=TRUE,df=NULL,randomstart=FALSE, usePointMass = FALSE, onlylogLR = FALSE, localfdr = TRUE, localfsr = TRUE, prior=NULL, sigmaavec=NULL, auto=FALSE, sigma.est=FALSE, nc=NULL, VB=FALSE){
 
   #if df specified, convert betahat so that bethata/sebetahat gives the same p value
   #from a z test as the original effects would give under a t test with df=df
   if(!is.null(df)){
     betahat = effective.effect(betahat,sebetahat,df)
   }	
-  
+    
   if(length(sebetahat)==1){
     sebetahat = rep(sebetahat,length(betahat))
+  }
+  if(length(sebetahat) != length(betahat)){
+    stop("Error: sebetahat must have length 1, or same length as betahat")
   }
   if(is.null(sigmaavec)){
     sigmaavec = c(0.00025,0.0005,0.001,0.002,0.004,0.008,0.016,0.032,0.064,0.128,0.256,0.512,1.024,2.048,4.096,8.192) 
[…]

### 2013-10-02 Rcode/intro.rmd
added small example of halfuniform in intro.rmd

@@
 So maybe it is the non-monotonic behaviour of the true beta values in this case.
-Let's try putting in a component at -1.5 to make the data fit this assumption better:
+Let's try putting in a component at -1.5 to make the data fit this assumption better.
+
+### Illustration of halfuniform method
+
+I will also use this to illustrate the halfuniform method, which
+allows for asymmetry.
 
@@
 beta.ash=ash(betahat,s)
+beta.ash.halfu = ash(betahat,s,"halfuniform")
+
+hist(beta,prob=TRUE)
+x = seq(-8,4,length=1000)
+lines(x,density(beta.ash,x),col=2)
+lines(x,density(beta.ash.halfu,x),col=3)
+
+```
+
+See how the use of the asymmtric uniform component does a better job capturing
+the underlying density. Probably clearer in the cdf plot:
+```{r}
+plot(sort(beta),(1:length(beta))/length(beta),main="cdf of beta, with fitted f overlaid",xlab="beta",ylab="cdf")
+lines(x,cdf.ash(beta.ash,x),col=2,lwd=2)
+lines(x,cdf.ash(beta.ash.halfu,x),col=3,lwd=2)
+
 #compute the usual zscore and corresponding p value
@@
 
+o = order(qval$qval)
+plot(cumsum(truenull[o])/(1:10000),qval$qval[o],col=1,type="l")
 o = order(beta.ash$qval)
+lines(cumsum(truenull[o])/(1:10000),beta.ash$qval[o],col=2)
+o2 = order(beta.ash.halfu$qval)
+lines(cumsum(truenull[o2])/(1:10000),beta.ash.halfu$qval[o2],col=3)
 
-plot(cumsum(truenull[o])/(1:10000),qval$qval[o],col=2,type="l")
-lines(cumsum(truenull[o])/(1:10000),beta.ash$qval[o])
 abline(a=0,b=1)

### 2013-10-04 Rcode/readme.rmd
added LogLik_conv to mix.R; fixed bugs related to LogLik not taking logs

@@
 source("../Rcode/ash.R")
+source("../Rcode/ash.oldfuncs.R")
 beta.ash.all = ash(betahat,s)
@@
 ```{r}
-cat(mixseLoglik(hh.betahat,hh.ash$fitted.f$pi, hh.ash$fitted.f$mu, hh.ash$fitted.f$sigma, hh.sebetahat),mixseLoglik(hh.betahat,hh.ash2$fitted.f$pi, hh.ash2$fitted.f$mu, sqrt(hh.ash2$fitted.f$sigma^2+1), hh.sebetahat,FUN="*"))
+cat(LogLik_conv(hh.ash$fitted.g, hh.betahat, hh.sebetahat),mixseLoglik(hh.betahat,hh.ash$fitted.g$pi, hh.ash$fitted.g$mean, hh.ash$fitted.g$sd, hh.sebetahat),mixseLoglik(hh.betahat,hh.ash2$fitted.g$pi, hh.ash2$fitted.g$mean, sqrt(hh.ash2$fitted.g$sd^2+1), hh.sebetahat,FUN="*"))
 ```
@@
 plot(as.numeric(hh.q$qvalue),hh.ash2$qval,cex=0.2,ylab="ASH q value",xlab="standard q value",main="ASH vs standard q value")
-cat(hh.ash2$fitted.f$pi[1],hh.q$pi0)
+cat(hh.ash2$fitted.g$pi[1],hh.q$pi0)
 ```
@@
 
-mixseLoglik(bhat.test,test1.ash1$fitted.f$pi,test1.ash1$fitted.f$mu,test1.ash1$fitted.f$sigma,hh.sebetahat)
-mixseLoglik(bhat.test,test1.ash2$fitted.f$pi,test1.ash2$fitted.f$mu,sqrt(test1.ash2$fitted.f$sigma^2+1),hh.sebetahat,FUN="*")
+mixseLoglik(bhat.test,test1.ash1$fitted.g$pi,test1.ash1$fitted.g$mean,test1.ash1$fitted.g$sd,hh.sebetahat)
+mixseLoglik(bhat.test,test1.ash2$fitted.g$pi,test1.ash2$fitted.g$mean,sqrt(test1.ash2$fitted.g$sd^2+1),hh.sebetahat,FUN="*")
 
-mixseLoglik(bhat2.test,test2.ash1$fitted.f$pi,test2.ash1$fitted.f$mu,test2.ash1$fitted.f$sigma,hh.sebetahat)
-mixseLoglik(bhat2.test,test2.ash2$fitted.f$pi,test2.ash2$fitted.f$mu,sqrt(test2.ash2$fitted.f$sigma^2+1),hh.sebetahat,FUN="*")
+mixseLoglik(bhat2.test,test2.ash1$fitted.g$pi,test2.ash1$fitted.g$mean,test2.ash1$fitted.g$sd,hh.sebetahat)
+mixseLoglik(bhat2.test,test2.ash2$fitted.g$pi,test2.ash2$fitted.g$mean,sqrt(test2.ash2$fitted.g$sd^2+1),hh.sebetahat,FUN="*")
 ```

### 2013-10-04 Rcode/mix.R

@@ -76,12 +76,23 @@ dens.default = function(x,y){
   return (x$pi %*% compdens(x, y))
 }
 
-#find log likelihood of data in y
-LogLik = function(x,y){
+#find log likelihood of data in x for mixture in m
+LogLik = function(m,x){
   UseMethod("LogLik")
 }
-LogLik.default = function(x,y){
-  sum(dens(x,y))
+LogLik.default = function(m,x){
+  sum(log(dens(m,x)))
+}
+
+#find log likelihood of data in x, when 
+#the mixture m is convolved with a normal with sd s
+#s is an n vector
+#x is an n vector
+LogLik_conv = function(m,x,s){
+  UseMethod("LogLik_conv")
+}
+LogLik_conv.default = function(m,x,s){
+  sum(log(dens_conv(m,x,s)))
 }
 
 #compute the density of the components of the mixture m

### 2013-10-04 Rcode/mix.R
added posterior mean and sd computations to mix.R

@@ -19,6 +19,15 @@ comp_sd.default = function(m){
   stop("method comp_sd not written for this class")
 }
 
+#second moments
+comp_mean2 = function(m){
+  UseMethod("comp_mean2")
+}
+comp_mean2.default = function(m){
+  comp_sd(m)^2 + comp_mean(m)^2
+}
+
+
 #return the overall mean of the mixture
 mixmean = function(m){
   UseMethod("mixmean")
@@ -27,7 +36,23 @@ mixmean.default = function(m){
   sum(m$pi * comp_mean(m))
 }
 
-#standard deviations
+#return the overall second moment of the mixture
+mixmean2 = function(m){
+  UseMethod("mixmean2")
+}
+mixmean2.default = function(m){
+  sum(m$pi * comp_mean2(m))
+}
+
+#return the overall sd of the mixture
+mixsd = function(m){
+  UseMethod("mixsd")
+}
+mixsd.default = function(m){
+  sqrt(mixmean2(m)-mixmean(m)^2)
+}
+
+#means
 comp_mean = function(m){
   UseMethod("comp_mean")
 }
@@ -166,6 +191,32 @@ postmean.default = function(m,betahat,sebetahat){
   colSums(comppostprob(m,betahat,sebetahat) * comp_postmean(m,betahat,sebetahat))
 }
 
+#output posterior mean-squared value for beta for prior mixture m,
+#given observations betahat, sebetahat
+postmean2 = function(m, betahat,sebetahat){
+  UseMethod("postmean2")
+}
+postmean2.default = function(m,betahat,sebetahat){
+  colSums(comppostprob(m,betahat,sebetahat) * comp_postmean2(m,betahat,sebetahat))
+}
+
+#output posterior sd for beta for prior mixture m,
+#given observations betahat, sebetahat
+postsd = function(m, betahat,sebetahat){
+  UseMethod("postsd")
+}
+postsd.default = function(m,betahat,sebetahat){
+  sqrt(postmean2(m,betahat,sebetahat)-postmean(m,betahat,sebetahat)^2)
+}
+
+#output posterior mean-squared value for beta for prior mixture m,
+#given observations betahat, sebetahat
+comp_postmean2 = function(m, betahat,sebetahat){
+  UseMethod("comp_postmean2")
+}
+comp_postmean2.default = function(m,betahat,sebetahat){
+  comp_postsd(m,betahat,sebetahat)^2 + comp_postmean(m,betahat,sebetahat)^2
+}
 
 
[…]

### 2013-10-04 Rcode/test.mix.Rmd
added my_etruncnorm to improve numerical stability when [a,b] is far from the normal
added posterior mean and sd computations to mix.R

@@
 source("mix.R")
+library("testthat")
 temp = normalmix(c(0.5,0.5),c(-3,2),c(1,1))
+
+expect_that(comp_mean(temp),equals(c(-3,2)))
+expect_that(mixmean(temp),equals(-0.5))
+expect_that(mixsd(temp),equals(2.692582))
+
 x = seq(-5,5,length=100)
@@
 ```{r}
+etruncnorm(10,20,0,1)
+my_etruncnorm(10,20,0,1)
+
+etruncnorm(-20,-10,0,1)
+my_etruncnorm(-20,-10,0,1)
+
+etruncnorm(-20,-10,5,1)
+my_etruncnorm(-20,-10,5,1)
+
+etruncnorm(-30,-20,5,2)
+my_etruncnorm(-30,-20,5,2)
+
+etruncnorm(-Inf,-38,0,1)
+my_etruncnorm(-Inf,-38,0,1)
+
+```
+
+```{r}
   comp_postmean(temp2, c(1,2,3), c(10,10,10))

### 2013-10-04 Rcode/ash.R

@@ -345,7 +345,7 @@ if(auto==TRUE){
     ZeroProb[!completeobs] = sum(mixprop(pi.fit$g)[comp_sd(pi.fit$g)==0])
     NegativeProb[!completeobs] = mixcdf(pi.fit$g,0) 
     PosteriorMean[!completeobs] = mixmean(pi.fit$g)
-      
+    #TO implement PosteriorSD[!completeobs] =mixsd()  
     PositiveProb =  1- NegativeProb-ZeroProb    
      
     

### 2013-10-04 Rcode/readme.rmd
improved package installation

@@
 source("../Rcode/ash.oldfuncs.R")
+source("../Rcode/mix.R")
 beta.ash.all = ash(betahat,s)

### 2013-10-04 Rcode/ash.R

@@ -1,9 +1,9 @@
 #TODO: Add nullcheck for VB?
 #Separate out the optimization over sigma from the EM algorithm
 
-ash.repodir = scan(".ash.repodir.txt",what=character()) 
-source(file.path(ash.repodir, "/Rcode/mix.R"))
-
+#ash.repodir = scan(".ash.repodir.txt",what=character()) 
+#source(file.path(ash.repodir, "/Rcode/mix.R"))
+#source("mix.R")
 
 #return the KL-divergence between 2 dirichlet distributions
 #p,q are the vectors of dirichlet parameters of same lengths
@@ -246,13 +246,11 @@ autoselect.sigmaavec = function(betahat,sebetahat){
   return(2^((-npoint):0) * sigmaamax)
 }
 
-
-
 #' @title Main Adaptive SHrinkage function
 #'
 #' @description takes a vector of betahats and ses; fits a mixture of normals to it and returns posteriors
 #'
-#' @details 
+#' @details See readme for more details
 #' 
 #' @param betahat (p vector); 
 #' @param sebetahat (p vector of standard errors)
@@ -337,16 +335,18 @@ if(auto==TRUE){
    	ZeroProb = rep(0,length=n)
     NegativeProb = rep(0,length=n)
     PosteriorMean = rep(0,length=n)
+    PosteriorSD=rep(0,length=n)
     
    	ZeroProb[completeobs] = colSums(comppostprob(pi.fit$g,betahat[completeobs],sebetahat[completeobs])[comp_sd(pi.fit$g)==0,,drop=FALSE])     
    	NegativeProb[completeobs] = cdf_post(pi.fit$g, 0, betahat[completeobs],sebetahat[completeobs]) - ZeroProb[completeobs]
     PosteriorMean[completeobs] = postmean(pi.fit$g,betahat[completeobs],sebetahat[completeobs])
+    PosteriorSD[completeobs] =postsd(pi.fit$g,betahat[completeobs],sebetahat[completeobs]) 
     
     #FOR MISSING OBSERVATIONS, USE THE PRIOR INSTEAD OF THE POSTERIOR
     ZeroProb[!completeobs] = sum(mixprop(pi.fit$g)[comp_sd(pi.fit$g)==0])
     NegativeProb[!completeobs] = mixcdf(pi.fit$g,0) 
[…]

### 2013-10-04 Rcode/package.R

@@ -1,23 +1,48 @@
-#These commands were used to set up the ashr package
+#This file documents some of the learning steps I took as I set up the ashr package
 
-setwd("~/Documents/git/ash/package")
-library(devtools)
+
+require(devtools)
+require(roxygen2)
+
+#This used to create package skeleton
 create("~/Documents/git/ash/package/ashr")
 
-Then I had to add 
-export(ash,EMest)
-to NAMESPACE file
+#THEN: I had to add export(ash) to NAMESPACE file
+#and  mv ash.R 
+# mv mix.R 
+#into the R directory
 
-#Then I put ash.R and mix.R in the R directory
-#NOte: I had to remove the source("") and library("") commands from these files
-#to get them to work
+#NOte: I had to remove the source("") and library("") commands from the
+#ash.R and mix.R files to get them to work
 
-setwd("~/Documents/git/ash/package/ashr")
-# not run build(binary=TRUE)
-install("../ashr")
+# process instructions; this als sets up NAMESPACE file
+roxygenize("~/Documents/git/ash/package/ashr")
+
+#install from source: this uses R CMD INSTALL to install
+# the ashr library in the usual directory (.Library)
+#and loads it
+install("~/Documents/git/ash/package/ashr")
+#I found I need to restart R after install to avoid errors in the help commands
+library(ashr)
 ash
-EMest
-works!
+?ash
+#works!
+
+#making a tar.gz version of ashr package
+require(devtools)
+build("~/Documents/git/ash/package/ashr",binary=FALSE)
+
+#this installs the tar.gz package and puts it in the library
+install.packages("ashr_0.1.tar.gz",repos=NULL,type="source")
+
+library(ashr)
+ash
+?ash
+#works!
+
+#Try installing from github
+require(devtools) 
+install_github('ash',user='stephens999',subdir='package/ashr')
 
 
 # had problems with these,

### 2013-10-04 Rcode/intro.rmd
improved package installation

@@
 source("../Rcode/ash.R") 
+source("../Rcode/mix.R") 
 library("qvalue")

### 2013-10-04 Rcode/test.mix.Rmd
improved package installation

@@
 ```{r}
-source("mix.R")
+setwd("~/Documents/git/ash/")
+source("Rcode/mix.R")
+source("Rcode/ash.R")
+source("Rcode/ash.oldfuncs.R")
 library("testthat")
-temp = normalmix(c(0.5,0.5),c(-3,2),c(1,1))
+temp = normalmix(c(0.5,0.5),c(0,0),c(1,2))
 
-expect_that(comp_mean(temp),equals(c(-3,2)))
-expect_that(mixmean(temp),equals(-0.5))
-expect_that(mixsd(temp),equals(2.692582))
+expect_that(comp_mean(temp),equals(c(0,0)))
+expect_that(mixmean(temp),equals(0))
+expect_that(mixsd(temp),equals(sqrt(2.5)))
+
+comp_postmean(temp,0,1)
+comp_postsd(temp,0,1)
+postsd(temp,0,1)
+
+set.seed(100)
+beta = rnorm(100)
+betahatse = abs(rnorm(100))
+betahat = rnorm(100, beta,betahatse)
+
+pd= posterior_dist(temp,betahat,betahatse)
+mv=normmix.mv(pd)
+expect_that(mv$mean,equals(postmean(temp,betahat,betahatse)))
+expect_that(sqrt(mv$var),equals(postsd(temp,betahat,betahatse)))
 

### 2013-10-04 Rcode/mix.R

@@ -369,6 +369,8 @@ comp_mean.unimix = function(m){
   (m$a+m$b)/2
 }
 
+
+
 compdens.unimix = function(x,y,log=FALSE){
   k=ncomp(x)
   n=length(y)
@@ -457,4 +459,12 @@ comp_postmean.unimix = function(m,betahat,sebetahat){
 #   )
 }
 
+#not yet implemented!
+#just returns 0s for now
+comp_postsd.unimix = function(m,betahat,sebetahat){
+  print("Warning: Posterior SDs not yet implemented for uniform components")
+  k= ncomp(m)
+  n=length(betahat)
+  return(matrix(NA,nrow=k,ncol=n)) 
+}
 

### 2013-10-06 Rcode/intro.rmd
added instructions to install ashr as a package

--- 2013-10-02 Rcode/intro.rmd
@@
+```{r, echo=FALSE}
+#Note, this needs the ashr library installed.
+#To build the library run the following (from the Rcode subdirectory):
+#require(devtools)
+#install("../package/ashr")
+#Then RESTART R before knitting the following code
+```
+
 ### Project Goals
@@
 
-As an important special case, these methods address the "multiple comparisons" setting, where interest focuses on which $\beta_j$ can be confidently inferred to be non-zero.
+As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
 Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\hat\beta_j/s_j$,
@@
 2. The distribution $g(\cdot; \pi)$ is a mixture of zero-centered normal distributions, 
-$$g(\cdot; \pi) = \sum_{k=1}^K pi_k N(\cdot; 0, \sigma^2_k).$$
+$$g(\cdot; \pi) = \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma^2_k).$$
 In practice, we currently fix the number of components $K$ to be large, and take the variances $\sigma_1<\sigma_2<\dots<\sigma_K$ to be fixed, and vary from very small (possibly 0), to very large --  sufficiently large that typically $\hat\pi_K=0$.
@@
 
-### A simple example
+### Example 1: A simple simulated example to get started
 
@@
 ```{r}
-setwd("~/Documents/git/ash/Rcode/")
+#setwd("~/Documents/git/ash/Rcode/")
 set.seed(32327)
-## load Poisson_binomial and ash functions 
-source("../Rcode/ash.R") 
-library("qvalue")
+
+## load ashr library 
+require("ashr")
+require("qvalue")
 ```
@@
 
-Simulate 10000 tests, with 2000 alternatives, $\beta \sim N(0,sd=2)$, 
-and 8000 null values ($\beta=0$).
+Simulate 10000 $\beta_j$ values, 8000 from a mixture of a point mass at 0 and
+2000 from $\beta \sim N(0,sd=2)$.
+
 ```{r}
@@
 beta.ash=ash(ss$betahat,ss$betasd)
+xmin = min(ss$beta)
+xmax = max(ss$beta)
+x=seq(xmin,xmax,length=1000)
+plot(sort(ss$beta),(1:length(ss$beta))/length(ss$beta),main="cdf of ss$beta, with fitted f overlaid",xlab="beta",ylab="cdf")
+lines(x,cdf.ash(beta.ash,x),col=2,lwd=2)
+
+```
+
+```{r}
+ss = simdata(10000,8000,0,2,1)
+
+beta.ash=ash(ss$betahat,ss$betasd,auto=FALSE)
 beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)
@@
 ```{r}
-print(beta.ash)
-print(beta.ash.auto)
-print(beta.ash.vb.uniform)
-print(beta.ash.vb.null)
-print(beta.ash.pm)
+summary(beta.ash)
+summary(beta.ash.auto)
+summary(beta.ash.vb.uniform)
+summary(beta.ash.vb.null)
+summary(beta.ash.pm)
 ```

### 2013-10-06 Rcode/package.R

@@ -1,10 +1,7 @@
 #This file documents some of the learning steps I took as I set up the ashr package
 
-
-require(devtools)
-require(roxygen2)
-
 #This used to create package skeleton
+require(devtools)
 create("~/Documents/git/ash/package/ashr")
 
 #THEN: I had to add export(ash) to NAMESPACE file
@@ -16,11 +13,14 @@ create("~/Documents/git/ash/package/ashr")
 #ash.R and mix.R files to get them to work
 
 # process instructions; this als sets up NAMESPACE file
+require(roxygen2)
 roxygenize("~/Documents/git/ash/package/ashr")
 
 #install from source: this uses R CMD INSTALL to install
 # the ashr library in the usual directory (.Library)
 #and loads it
+require(devtools)
+document("~/Documents/git/ash/package/ashr")
 install("~/Documents/git/ash/package/ashr")
 #I found I need to restart R after install to avoid errors in the help commands
 library(ashr)

### 2013-10-10 Rcode/ash.R
added ... to plot.ash; extended some examples in intro

@@ -444,10 +444,10 @@ print.ash =function(a){
 #' 
 #' @export
 #' 
-plot.ash = function(a,xmin,xmax){
+plot.ash = function(a,xmin,xmax,...){
   x = seq(xmin,xmax,length=1000)
   y = density(a,x)
-  plot(x,y,type="l")
+  plot(x,y,type="l",...)
 }
 
 #compute the predictive density of an observation

### 2013-10-10 Rcode/intro.rmd
added ... to plot.ash; extended some examples in intro

@@
 Second, the amount of shrinkage undergone by each $\hat\beta_j$ will depend on the standard error $s_j$: measurements with high standard error will undergo more shrinkage than measurements with low standard error.
+
+Given that shrinkage estimation is widely recognized as a powerful tool, there are suprisingly few software packages for performing the simplest type of shrinkage estimation considered here.
+(There are more packages for the more complex setting of covariance estimation, where shrinkage is perhaps still more important.)
+The only package we have found that provides anything similar
+to the functionality provided here is mixfdr (Muralidharan). Compared with mixfdr, the key features of ashr are that it
+i) focuses on allowing for variation in the standard deviation of each observation; ii) constrains the underlying density to be unimodal (and possibly symettric).
+NOTE: should emphasise these differences in the examples.
 
@@
 ```
+
+Compare with mixfdr. Note that I tried to do small than 1000, but mixfdr returned
+errors.
+```{r}
+ss = simdata(1000,800,0,2,1)
+
+beta.ash=ash(ss$betahat,ss$betasd)
+beta.ash.u = ash(ss$betahat,ss$betasd, mixcompdist="uniform")
+beta.ash.hu = ash(ss$betahat,ss$betasd, mixcompdist="halfuniform")
+
+require(mixfdr)
+beta.mfdr.J3 = mixFdr(ss$betahat,J=3,noiseSD=1,theonull=TRUE)
+beta.mfdr.J5 = mixFdr(ss$betahat,J=5,noiseSD=1,theonull=TRUE)
+beta.mfdr.J7 = mixFdr(ss$betahat,J=7,noiseSD=1,theonull=TRUE)
+beta.mfdr.J9 = mixFdr(ss$betahat,J=9,noiseSD=1,theonull=TRUE)
+beta.mfdr.J15 = mixFdr(ss$betahat,J=15,noiseSD=1,theonull=TRUE)
+mean((beta.mfdr.J3$effectSize-ss$beta)^2)
+mean((beta.mfdr.J5$effectSize-ss$beta)^2)
+mean((beta.mfdr.J7$effectSize-ss$beta)^2)
+mean((beta.mfdr.J9$effectSize-ss$beta)^2)
+mean((beta.mfdr.J15$effectSize-ss$beta)^2)
+
+mean((beta.ash$PosteriorMean-ss$beta)^2)
+
+#This code mostly borrowed from the examples in mixFdr in R help
+m=beta.mfdr.J15
+mTrue = list(pi=c(0.8,0.2),mu=c(0,0),sig=c(1,2))
+par(mfcol=c(1,1))
+s = seq(-5,5,by=0.01)
+trueFdr = fdrMixModel(s,mTrue,c(TRUE,FALSE))
+fdrCurve = fdrMixModel(s,m,abs(m$mu - m$mu[1])<=0.002) # this tells it which indices to consider null
+ plot(s, trueFdr, t = 'l', main = "fdr - Black is true", xlab = "z", ylab = "fdr", ylim = c(0,1))
+ lines(s, fdrCurve, col = 2)
+points(ss$betahat,beta.ash$localfdr,col=3)
+points(ss$betahat,beta.ash.u$localfdr,col=4)
+points(ss$betahat,beta.ash.hu$localfdr,col=5)
+
+```
+
+Of course, if we do something where alternatives are not centered on 0 then
+ash does worse!
+```{r}
+#This is the example from the mixfdr demo
+z = c(rnorm(1000), rnorm(100,3,1))
+ m = mixFdr(z, J = 3, noiseSD = 1)
+ s = seq(-5,5,by=0.01)
+ effCurve = effectSize(s,m)[,1]
+ fdrCurve = fdrMixModel(s,m,abs(m$mu - m$mu[1])<=0.2) # this tells it which indices to consider null
+
+# compare to the Bayes estimator for this problem
+ mTrue = list(pi = c(10,1)/11, mu = c(0,3), sig = c(1,1))
+# note that the model parameters are in terms of the MARGINAL not of the prior
+ trueEff = effectSize(s,mTrue)[,1]
+ trueFdr = fdrMixModel(s,mTrue,c(TRUE,FALSE))
+ par(mfrow = c(2,1))
+ plot(s, trueEff, t = 'l', main = "Effect Size - Black is true", xlab = "z", ylab = "E(delta|z)", ylim = c(-3,3))
+ lines(s, effCurve, col = 2)
+ plot(s, trueFdr, t = 'l', main = "fdr - Black is true", xlab = "z", ylab = "fdr", ylim = c(0,1))
+ lines(s, fdrCurve, col = 2)
+
+z.ash = ash(z,1,mixcompdist="halfuniform")
+points(z,z.ash$localfdr,col=3)
+```
+
+Things to emphasise for paper:
+- the number of components is not critical; the unimodal constraint is enough. You can increase the number of components arbitrarily and the likelihood remains bounded.
+- the ability to incorporate item-specific measurement error
+- the ability to compare a model in which effects are proportional to error
+with model in which effects
+
+Some advantages of the unimodal constraint:
+-more statsitically efficient (if true); show with small sample sizes.
+- less sensitive to number of components? Can allow number to tend to infinity?
+- may be less sensitive to local optima?
+- it allows us to easily just vary sigma on a grid, and fit pi, which
+makes allowing different noise levels really easy!
 

### 2013-10-11 Rcode/ash.R
modified default prior to uniform; changed name of sigmaavec to mixsd

@@ -1,28 +1,141 @@
-#TODO: Add nullcheck for VB?
-#Separate out the optimization over sigma from the EM algorithm
+#' @title Main Adaptive SHrinkage function
+#'
+#' @description Takes vectors of estimates (betahat) and their standard errors (sebetahat), and applies
+#' shrinkage to them, using Empirical Bayes methods, to compute shrunk estimates for beta.
+#'
+#' @details See readme for more details
+#' 
+#' @param betahat, a p vector of estimates 
+#' @param sebetahat, a p vector of corresponding standard errors
+#' @param mixcompdist: distribution of components in mixture ("normal", "uniform" or "halfuniform")
+#' @param nullcheck: whether to check that any fitted model exceeds the "null" likelihood
+#' in which all weight is on the first component
 
-#ash.repodir = scan(".ash.repodir.txt",what=character()) 
-#source(file.path(ash.repodir, "/Rcode/mix.R"))
-#source("mix.R")
-
-#return the KL-divergence between 2 dirichlet distributions
-#p,q are the vectors of dirichlet parameters of same lengths
-diriKL = function(p,q){
-  p.sum = sum(p)
-  q.sum = sum(q)
-  k = length(q)
-  KL = lgamma(q.sum)-lgamma(p.sum)+sum((q-p)*(digamma(q)-digamma(rep(q.sum,k))))+sum(lgamma(p)-lgamma(q))
-  return(KL)
+#' @param df: appropriate degrees of freedom for (t) distribution of betahat/sebetahat
+#' @param randomstart: bool, indicating whether to initialize EM randomly. If FALSE, then initializes to prior mean (for EM algorithm) or prior (for VBEM)
+#' @param usePointMass: bool, indicating whether to use a point mass at zero as one of components for a mixture distribution
+#' @param onlylogLR (= FALSE) : bool, indicating whether to use this function to get logLR. Skip posterior prob, posterior mean, localfdr...
+#' @param localfdr (=TRUE) : bool,  indicating whether to compute localfdr, localfsr, and q-value
[…]

### 2013-10-11 Rcode/intro.rmd
modified default prior to uniform; changed name of sigmaavec to mixsd

@@
 
-Compare with mixfdr. Note that I tried to do small than 1000, but mixfdr returned
-errors.
+Compare with mixfdr. Note that I tried to do smaller than 1000, but mixfdr returned
+errors. (Maybe this example is not so useful for comparing fdrs (as opposed to shrinkage), since the methods tend to agree where the fdr is small, which is what matters?)
 ```{r}
-ss = simdata(1000,800,0,2,1)
+ss = simdata(1000,500,0,2,1)
 
@@
 mean((beta.ash$PosteriorMean-ss$beta)^2)
+mean((beta.ash.u$PosteriorMean-ss$beta)^2)
+mean((beta.ash.hu$PosteriorMean-ss$beta)^2)
 
@@
 ```{r}
-#This is the example from the mixfdr demo
+#This is the example from the mixfdr help example
 z = c(rnorm(1000), rnorm(100,3,1))
@@
 
+Checking coverage
+```{r}
+
+```
+
 Things to emphasise for paper:
@@
 
+- unimodal distributions could be argued to be a primary motivation for shrinkage
+
+- bayesian coverage intervals (attempt to) give much stronger guarantess than standard confidence intervals. Eg among the 0.95 intervals excluding 0,
+less than 5% should actually be 0?
+
 Some advantages of the unimodal constraint:
@@
 - may be less sensitive to local optima?
+could demonstrate this by looking at convergence more carefully, and comparing
+results with random starting points.
+
 - it allows us to easily just vary sigma on a grid, and fit pi, which
 makes allowing different noise levels really easy!
+
+## The number of components is not critical
+
+OR - this section could be more generally about selecting between
+models (normal, uniforms etc) using log-likelihood.
+
+Don't need to worry about overfitting - only need to worry
+about whether model is "complex enough".
+
+Because of the uni-modal constraint, the number of mixture components
+is generally not critical, at least provided it is ``sufficiently large". 
+Indeed, as the number of components tends to infinity, the likelihood
+is bounded above, and the procedure will converge to a sensible limit.
+The following example illustrates this.
+We take 100 observations from 0.5N(-1,1)+0.5N(1,1)
+and add N(0,1) noise.
+```{r}
+sampsize=500
+beta = c(rnorm(sampsize,-1,1),rnorm(sampsize,1,1))
+betahat = beta+rnorm(2*sampsize)
+
+beta.ash = ash(betahat,1)
+beta.ash.u=list()
+for(i in 1:10){
+  beta.ash.u[[i]] = ash(betahat,1,mixcompdist="uniform",prior="uniform",sigmaavec = seq(0.1,5,length=10*i))
+}
+
+plot(beta.ash,-4,4,col=2)
+x=seq(-4,4,length=1000)
+lines(x,0.5*dnorm(x,-1,1)+0.5*dnorm(x,1,1),col=1)
+for(i in 1:4){
+lines(x,density(beta.ash.u[[i]],x),col=2+i)
+}
+
+plot(unlist(lapply(beta.ash.u,ashr:::get_loglik)))
+
+```
+
+```{r}
+sampsize=5000
+beta = c(rnorm(sampsize,-1,1),rnorm(sampsize,1,1))
+betahat = beta+rnorm(2*sampsize)
+
+beta.ash = ash(betahat,1)
+beta.ash.u=list()
+for(i in 1:10){
+  beta.ash.u[[i]] = ash(betahat,1,mixcompdist="uniform",prior="uniform",sigmaavec = seq(0.1,5,length=10*i))
+}
+
+plot(beta.ash,-4,4,col=2)
+x=seq(-4,4,length=1000)
+lines(x,0.5*dnorm(x,-1,1)+0.5*dnorm(x,1,1),col=1)
+for(i in 1:4){
+lines(x,density(beta.ash.u[[i]],x),col=2+i)
+}
+
+plot(unlist(lapply(beta.ash.u,ashr:::get_loglik)))
+
+```
 
@@
 #these to test the VB version 
-beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior="uniform")
-beta.ash.vb.null = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior=NULL)
+beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,VB=TRUE, prior="uniform")
+beta.ash.vb.null = ash(ss$betahat, ss$betasd,VB=TRUE, prior=NULL)
 
@@
 hist(zscore)
-```
-
-Now, we find the fitted values for $\pi$ are mostly near 0, or at 2 (the alternative value).
-```{r}
-summary(beta.ash)
-summary(beta.ash.auto)
-summary(beta.ash.vb.uniform)
-summary(beta.ash.vb.null)
-summary(beta.ash.pm)
 ```
@@
 ```{r}
-beta.ash.pm = ash(ss$betahat, ss$betasd, auto=TRUE, usePointMass=TRUE)
+beta.ash.pm = ash(ss$betahat, ss$betasd, usePointMass=TRUE)
 print(beta.ash.pm)

### 2013-10-12 Rcode/intro.rmd
added some more examples to intro.rmd

@@
  plot(s, trueFdr, t = 'l', main = "fdr - Black is true", xlab = "z", ylab = "fdr", ylim = c(0,1))
- lines(s, fdrCurve, col = 2)
 
@@
 
-Don't need to worry about overfitting - only need to worry
-about whether model is "complex enough".
-
 Because of the uni-modal constraint, the number of mixture components
@@
 Indeed, as the number of components tends to infinity, the likelihood
-is bounded above, and the procedure will converge to a sensible limit.
+is bounded above, and even for large numbers of components and small amounts of data the inferred underlying distributions tend not to be too crazy, showing few signs of ``overfitting".
+
+However, it is true that the normal distributions tend to yield smoother estimated underlying distributions. Similarly, using the posterior mean for $\pi$, rather
+than the maximum likelihood estimate, tends to lead to somewhat smoother fitted distributions. In addition, with small amounts of data the underlying
+distributions are inevitably not well determined by the data, and the fits may vary depending on the underlying assumptions made (e.g.
+number of components or distribution of the components); however, even then shrinkage-based estimates
+of the betas can be relatively robust.
+
 The following example illustrates this.
@@
 ```{r}
-sampsize=500
+set.seed(200)
+sampsize=50
 beta = c(rnorm(sampsize,-1,1),rnorm(sampsize,1,1))
@@
 
-beta.ash = ash(betahat,1)
-beta.ash.u=list()
-for(i in 1:10){
-  beta.ash.u[[i]] = ash(betahat,1,mixcompdist="uniform",prior="uniform",sigmaavec = seq(0.1,5,length=10*i))
-}
+beta.ash.default = ash(betahat,1) #run with default parameters for comparison
+beta.ash.vb.default = ash(betahat,1,VB=TRUE) #run with default parameters for comparison
 
-plot(beta.ash,-4,4,col=2)
+beta.ash.100 = ash(betahat,1,mixsd = seq(0.1,5,length=100))
+beta.ash.100u = ash(betahat,1,mixcompdist="uniform",mixsd = seq(0.1,5,length=100))
+beta.ash.100hu = ash(betahat,1,mixcompdist="halfuniform",mixsd = seq(0.1,5,length=100))
+
+beta.ash.vb.100 = ash(betahat,1,mixsd = seq(0.1,5,length=100),VB=TRUE)
+beta.ash.vb.100u = ash(betahat,1,mixcompdist="uniform",mixsd = seq(0.1,5,length=100),VB=TRUE)
+beta.ash.vb.100hu = ash(betahat,1,mixcompdist="halfuniform",mixsd = seq(0.1,5,length=100),VB=TRUE)
+
 x=seq(-4,4,length=1000)
-lines(x,0.5*dnorm(x,-1,1)+0.5*dnorm(x,1,1),col=1)
-for(i in 1:4){
-lines(x,density(beta.ash.u[[i]],x),col=2+i)
-}
+plot(x,0.5*dnorm(x,-1,1)+0.5*dnorm(x,1,1),col=1,ylim=c(0,0.45),main="Fitted densities vs truth",type="l",ylab="density")
+lines(x,density(beta.ash.default,x),col=2)
+lines(x,density(beta.ash.100,x),col=2)
+lines(x,density(beta.ash.100u,x),col=3)
+lines(x,density(beta.ash.100hu,x),col=4)
 
-plot(unlist(lapply(beta.ash.u,ashr:::get_loglik)))
+plot(x,0.5*dnorm(x,-1,1)+0.5*dnorm(x,1,1),col=1,ylim=c(0,0.45),main="Fitted densities (VB) vs truth",type="l",ylab="density")
+lines(x,density(beta.ash.vb.default,x),col=2)
+lines(x,density(beta.ash.vb.100,x),col=2,lty=2)
+lines(x,density(beta.ash.vb.100u,x),col=3,lty=2)
+lines(x,density(beta.ash.vb.100hu,x),col=4,lty=2)
+
+plot(x,0.5*pnorm(x,-1,1)+0.5*pnorm(x,1,1),col=1,main="Fitted cdfs vs truth",type="l",ylab="density")
+lines(x,cdf.ash(beta.ash.100,x),col=2,lty=2)
+lines(x,cdf.ash(beta.ash.100u,x),col=3,lty=2)
+lines(x,cdf.ash(beta.ash.100hu,x),col=4,lty=2)
+
+plot(x,0.5*pnorm(x,-1,1)+0.5*pnorm(x,1,1),col=1,main="Fitted cdfs (VB) vs truth",type="l",ylab="density")
+lines(x,cdf.ash(beta.ash.vb.default,x),col=2,lty=2)
+lines(x,cdf.ash(beta.ash.vb.100,x),col=2,lty=2)
+lines(x,cdf.ash(beta.ash.vb.100u,x),col=3,lty=2)
+lines(x,cdf.ash(beta.ash.vb.100hu,x),col=4,lty=2)
+
+c(loglik.ash(beta.ash.100,betahat,1),get_loglik(beta.ash.100u),get_loglik(beta.ash.100hu))
+c(loglik.ash(beta.ash.vb.100,betahat,1),loglik.ash(beta.ash.vb.100u,betahat,1),loglik.ash(beta.ash.vb.100hu,betahat,1))
+
+rmse=function(a,b){sqrt(mean((a$PosteriorMean-b)^2))}
+c(rmse(beta.ash.default,beta),rmse(beta.ash.100,beta),rmse(beta.ash.100u,beta),rmse(beta.ash.100hu,beta),rmse(beta.ash.vb.default,beta),rmse(beta.ash.vb.100,beta),rmse(beta.ash.vb.100u,beta),rmse(beta.ash.vb.100hu,beta))
+
+plot(beta.ash.default$PosteriorMean,beta.ash.100u$PosteriorMean,col=3,cex=0.4,xlim=c(-4,4),ylim=c(-4,4))
+points(beta.ash.default$PosteriorMean,beta.ash.100hu$PosteriorMean,col=4,cex=0.4)
+abline(a=0,b=1)
+
+plot(betahat,ifelse(pnorm(betahat)<0.5,pnorm(betahat),1-pnorm(betahat)))
+points(betahat,beta.ash.default$localfsr,col=2)
+points(betahat,beta.ash.100u$localfsr,col=3)
+points(betahat,beta.ash.100hu$localfsr,col=4)
 
 ```
+
+so it seems that the localfsr is almost the same as for a "flat" prior $g$.
+That is, in this case, the underlying distribution is flat relative to the
+error terms, and so there is little shrinkage effect except at the tails
+(where the fsr is basically 0 anyway).
 
@@
 for(i in 1:10){
-  beta.ash.u[[i]] = ash(betahat,1,mixcompdist="uniform",prior="uniform",sigmaavec = seq(0.1,5,length=10*i))
+  beta.ash.u[[i]] = ash(betahat,1,mixcompdist="uniform",mixsd = seq(0.1,5,length=10*i))
 }
@@
 beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,VB=TRUE, prior="uniform")
-beta.ash.vb.null = ash(ss$betahat, ss$betasd,VB=TRUE, prior=NULL)
+beta.ash.vb.nullbiased = ash(ss$betahat, ss$betasd,VB=TRUE, prior="nullbiased")
 

### 2013-10-12 Rcode/ash.R
added some more examples to intro.rmd

@@ -435,6 +435,39 @@ predictive=function(a,se){
 }
 
 
+#' @title Get fitted loglikelihood for ash object
+#'
+#' @description Return the log-likelihood of the data under the fitted distribution
+#'
+#' @param a the fitted ash object
+#'
+#' @details None
+#' 
+#' @export
+#' 
+#'
+get_loglik = function(a){
+  return(tail(a$fit$loglik,1))
+}
+
+#' @title Compute loglikelihood for data from ash fit
+#'
+#' @description Return the log-likelihood of the data betahat, with standard errors betahatsd, 
+#' under the fitted distribution in the ash object
+#'
+#' @param a the fitted ash object
+#' @param betahat the data
+#' @param betahatsd the observed standard errors
+#' 
+#' @details None
+#' 
+#' @export
+#' 
+#'
+loglik.ash = function(a,betahat,betahatsd){
+  return(loglik_conv(a$fitted.g,betahat, betahatsd))
+}
+
 #' @title Plot method for ash object
 #'
 #' @description Return the density of the underlying fitted distribution

### 2013-10-18 Rcode/ash.test.rmd
added ash.test.rmd

--- 2013-10-10 Rcode/intro.rmd
@@
-```{r, echo=FALSE}
-#Note, this needs the ashr library installed.
-#To build the library run the following (from the Rcode subdirectory):
-#require(devtools)
-#install("../package/ashr")
-#Then RESTART R before knitting the following code
-```
-
-### Project Goals
-
-The initial goal of the ASH (Adaptive SHrinkage) project is to provide simple, generic, and flexible methods to derive ``shrinkage-based" estimates and credible intervals for unknown quantities $\beta=(\beta_1,\dots,\beta_J)$, given only estimates of those quantities ($\hat\beta=(\hat\beta_1,\dots, \hat\beta_J)$) and their corresponding estimated standard errors ($s=(s_1,\dots,s_J)$).
-
-Although shrinkage-based estimation can be motivated in various ways, our key goal here is to combine information across the multiple measurements $j=1,\dots,J$ to improve inference for each individual $\beta_j$.
-By improved inference, we mean both
-improved average accuracy of point estimates, which is
-the traditional focus of shrinkage-based methods, \emph{and} improved assessments of uncertainty.
-
-By ``adaptive" shrinkage we 
-have two key properties in mind.
-First, the appropriate amount of shrinkage is determined from the data, rather than being pre-specified.
-Second, the amount of shrinkage undergone by each $\hat\beta_j$ will depend on the standard error $s_j$: measurements with high standard error will undergo more shrinkage than measurements with low standard error.
-
-Given that shrinkage estimation is widely recognized as a powerful tool, there are suprisingly few software packages for performing the simplest type of shrinkage estimation considered here.
-(There are more packages for the more complex setting of covariance estimation, where shrinkage is perhaps still more important.)
-The only package we have found that provides anything similar
-to the functionality provided here is mixfdr (Muralidharan). Compared with mixfdr, the key features of ashr are that it
-i) focuses on allowing for variation in the standard deviation of each observation; ii) constrains the underlying density to be unimodal (and possibly symettric).
-NOTE: should emphasise these differences in the examples.
-
-As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
-Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\hat\beta_j/s_j$,
-and then applying a generic procedure, such as that of Benjamini 
-and Hochberg (1995?) or Storey (2001?), designed to control or
-estimate the false discovery rate (FDR) or the positive FDR (Storey, 2001?). In essence we aim to provide analagous
-generic methods that work directly with two numbers for each 
-measurement $(\hat\beta_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
-Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
-uncertainty in each measurement $\hat\beta_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
-
-The potential for shrinkage-based estimation to
-address the multiple comparisons setting 
-has been highlighted
-previously, including Greenland and Robins (1991),
-Efron (2008) and Gelman et al (2012). [Note, check also Louis, JASA, 1984]
-
-### Methods Outline
-
-The methods are based on treating the vectors $\hat\beta$
-and $s$ as ``observed data", and then performing inference for $\beta$ from these observed data, using a standard hierarchical modelling framework
-to combine information across $j=1,\dots,J$.
-
-Specifically, 
-we assume that the true 
-$\beta_j$ values are independent
-and identically distributed from some distribution $g(\cdot ; \pi)$, where $\pi$ is a hyper-parameter to be estimated. 
-Then, given $\beta$, we assume that $(\hat\beta_j,s_j)$ are independent across $j$, and depend only on $\beta_j$.
-Putting these together, the joint model for the unobserved $\beta$ and the observed $\hat\beta, s$ is:
-\begin{align}
-p(\hat\beta, s, \beta | \pi) & = \prod_j g(\beta_j ; \pi) p(\hat\beta_j, s_j | \beta_j) \\
-& = \prod_j g(\beta_j ; \pi) L(\beta_j; \hat\beta_j,s_j).
-\end{align}
-The specific choices of $g$ and $L$ are described below.
-
-We fit this hierarchical model using the following "Empirical Bayes" approach. First we estimate the hyper-parameters $\pi$ by maximizing the likelihood
-$$L(\pi; \hat\beta, s) := p(\hat\beta, s | \pi) = \int p(\hat\beta, s, \beta | \pi) d\beta.$$
-Then, given this estimate $\hat\pi$, we compute the conditional distributions $$p(\beta_j | \hat\pi, \hat\beta, s) \propto g(\beta_j; \pi) L(\beta_j; \hat\beta_j, s_j).$$ 
-In principle we would
-prefer to take a full Bayes approach that accounts for uncertainty in $\pi$, but, at least for now, we compromise this principle for the simplicity of the EB approach.
-[Note: a Variational Bayes version of this is
-also implemented, and may become our preferred approach
-after testing]
-
-[put picture of hierarchical model here]
-
-The conditional distributions $p(\beta_j | \hat\pi, \hat\beta, s)$ 
-encapsulate uncertainty in the values for $\beta_j$, combining information across
-$j=1,\dots,J$. The combining of the information occurs through estimation of
-$\pi$, which involves all of the data, and it is
-
-These conditional distributions can be conveniently summarized
-in various ways, including point estimates (e.g. the posterior means or medians),
-and credible intervals/regions.
-
-The key components of this hierarchical model
-are the distribution $g$ and the likelihood $L(\beta_j; \hat\beta_j, s_j)$. We make the following choices for these.
-
-1. The likelihood for $\beta_j$ is normal, centered on $\hat\beta_j$, with standard deviation $s_j$.
-That is, 
-$$L(\beta_j; \hat\beta_j, s_j) \propto \exp[-0.5(\beta_j-\hat\beta_j)^2/s_j^2]. \quad (**)$$
-
-2. The distribution $g(\cdot; \pi)$ is a mixture of zero-centered normal distributions, 
-$$g(\cdot; \pi) = \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma^2_k).$$
-In practice, we currently fix the number of components $K$ to be large, and take the variances $\sigma_1<\sigma_2<\dots<\sigma_K$ to be fixed, and vary from very small (possibly 0), to very large --  sufficiently large that typically $\hat\pi_K=0$.
-
-The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
-For example, we can write $p(\hat\beta_j, s_j | \beta_j) = p(\hat\beta_j | s_j, \beta_j)p(s_j | \beta_j)$.
-Now, if we are willing to assume that
-$s_j$ alone contains no information about $\beta_j$, or equivalently that $p(s_j | \beta_j)$ does not depend on $\beta_j$, then
-$L(\beta_j) \propto p(\hat\beta_j | s_j, \beta_j)$,
-and if $\hat\beta_j | s_j, \beta_j \sim N(\beta_j, s_j^2)$,
-as is often asymptotically the case, then we obtain the likelihood (**) above.
-
-An alternative motivation is to think of this as a normal
-approximation to the likelihood from the 
-raw data $D_j$ that were used to compute
-$\hat\beta_j$ and $s_j$. Then if we observed these data
-the likelihood for $\beta$ would be 
-$p(D_j | \beta_j)$, and a Taylor series expansion of the log likelihood around the maximum likelihood estimate $\hat\beta_j$  yields $$l(\beta_j) \approx l(\hat\beta_j) + 0.5* (\beta_j - \hat\beta_j)^2 l''(\hat\beta_j).$$ [Fill in details?]
-
-Using a mixture of normal distributions for $g$ 
-also seems very natural: mixtures of normals provide a flexible family of distributions able to provide a good approximation to any true underlying $g$; and 
-when combined with the normal likelihood they give
-an analytic form for the conditional distribution $p(\beta_j | \hat\pi, \hat\beta_j, s_j)$ (also a mixture of normals).
-The constraint that these normals be centered at zero may seem initially less natural. 
-Certainly this constraint could be relaxed in principle.
-However, we view it as a convenient way to impose an assumption 
-that $g$ is unimodal with its mode at 0,
-which we view as a plausible assumption in many 
-settings, and one which may be helpful to avoid
-"overfitting" of $g$. (Using normal distributions centered at 0
-also imposes an assumption
-that $g$ is symmetric about zero, which we view as less
-plausible, but represents a compromise between simplicity
-and flexibility.
-In cases where this assumption seems wildly inappropriate one could perhaps
-improve results by applying the
-model separately to positive and negative values of $\hat\beta$.)
-
-Finally, using a large number of normal components with a wide range of variances,
-rather than, say, a smaller number of components with the variances
-to be estimated, is simply for computational convenience. With fixed
-variances there exists a very simple EM algorithm to maximise the likelihood
-in $\pi$. Obtaining maximum likelihood estimates for the variances could certainly
-be implemented with a little
-more work, but it is unclear whether this would result in practically-important gains in many situations of interest.
-
-### The False Sign Rate
-
-The usual definition of the local False Discovery Rate (lfdr) for observation $j$ is
-$\text{lfdr}_j = p(\beta_j = 0 | \hat\beta, s)$.
-The lfdr terminology comes from using "discovery" to refer to rejecting the null ($H_j:\beta_j=0$), so lfdr 
-gives the probability that we make a mistake if we reject
-$H_j$, that is the probability that, if we reject $H_j$,
-it is a  "false discovery".
-
-As pointed out by Gelman et al, there are settings where
-$\beta_j=0$ is implausible, in which case the lfdr is not
-a useful concept: if every $\beta_j$ is non-zero then there
-is no such thing as a false discovery and the lfdr will be 0. Gelman et al suggest that in such settings we
-might replace the concept of a false discovery with the
-concept of an "error in sign". The idea is that, in settings where $\beta_j=0$ is implausible, the most fundamental
-inference objective is to ask which $\beta_j$ are positive and which are negative. Indeed, even in settings where some
-$\beta_j$ are exactly zero, it could be argued that identifying which are positive and which negative is
-fundamentally more interesting and useful than identifying which are non-zero.
-However, in such settings this usually comes down to something similar in practice: the observations that one can tell are non-zero are also the ones whose
-sign can be reliably inferred. In this sense the
-local false sign rate is a natural generalization of
-the local false discovery rate.
-
-Following these ideas of Gelman et al, we define the local False Sign Rate (lfsr) as 
-$$\text{lfsr}_j = \min[ p(\beta_j \geq 0| \hat\beta, s), p(\beta_j \leq 0| \hat\beta, s) ].$$
-To give the intuition behind this, suppose for concreteness
-that the minimum is achieved by the first term, $p(\beta_j \geq 0| \hat\beta, s)=0.05$ say. Then, if we were to declare that $\beta_j$ is negative, the probability that we have
-made an error in sign would be 0.05.
-
-Note that $\text{lfsr}_j \geq \text{lfdr}_j$ 
-because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ include the event $\beta_j=0$.
-Thus, lfsr gives an upper bound for the lfdr,
-and so can be used as a conservative estimate of the lfdr
-if an lfdr is desired. (This may be helpful when comparing
-with methods that compute an lfdr.)
-In addition, we intuitively expect the difference $\text{lfsr}_j - \text{lfdr}_j$ to be small, at least when $\text{lfdr}_j$ is small, because when we are confident that a particular effect is non-zero we would expect to be similarly confident about the sign of the effect.
-
-### Computation Outline
-
-As outlined above, we fit the model using the following Empirical Bayes procedure:
-1. Estimate $\pi$ by maximizing the likelihood $L(\pi)$.
-2. Compute the conditional distributions $p(\beta_j | \hat\beta_j, s_j, \hat\pi)$.
-
-Using a normal likelihood $L(\beta_j)$, and assuming
-$g$ to be a mixture of normal distributions with fixed variances, 
-yields a simple EM algorithm
-for estimating $\pi$ in Step 1, and simple analytic forms for the conditional
-distributions in Step 2.
-
-### Example 1: A simple simulated example to get started
-
-Load in some functions.
 ```{r}
-#setwd("~/Documents/git/ash/Rcode/")
-set.seed(32327)
-
-## load ashr library 
 require("ashr")
 require("qvalue")
-```
+require("testthat")
 
-```{r}
+set.seed(100)
+
 #simulate n beta-hat values, nnull under the null
@@
 }
+
+ss = simdata(10000,8000,0,2,1)
+
+set.seed(133)
+system.time((beta.ash=ash(ss$betahat,ss$betasd,randomstart=TRUE)))
+set.seed(133)
+system.time((beta.ash2=oldash(ss$betahat,ss$betasd,auto=TRUE,randomstart=TRUE)))
+
+expect_that(beta.ash$PosteriorMean, equals(beta.ash2$PosteriorMean))
+expect_that(beta.ash$qvalue, equals(beta.ash2$qvalue))
+expect_that(beta.ash$PositiveProb, equals(beta.ash2$PositiveProb))
+
+system.time((beta.ash.vb.uniform = ash(ss$betahat, ss$betasd, VB=TRUE, prior="uniform")))
+system.time((beta.ash.vb.uniform2 = oldash(ss$betahat, ss$betasd, VB=TRUE, auto=TRUE,prior="uniform")))
+expect_that(beta.ash.vb.uniform$PosteriorMean, equals(beta.ash.vb.uniform2$PosteriorMean))
+expect_that(beta.ash.vb.uniform$qvalue, equals(beta.ash.vb.uniform2$qvalue))
+expect_that(beta.ash.vb.uniform$PositiveProb, equals(beta.ash.vb.uniform2$PositiveProb))
+
+system.time((beta.ash.vb.null = ash(ss$betahat, ss$betasd, VB=TRUE, prior=NULL)))
+system.time((beta.ash.vb.null2 = ash(ss$betahat, ss$betasd, VB=TRUE, prior=NULL)))
+expect_that(beta.ash.vb.null$PosteriorMean, equals(beta.ash.vb.null2$PosteriorMean))
+expect_that(beta.ash.vb.null$qvalue, equals(beta.ash.vb.null2$qvalue))
+expect_that(beta.ash.vb.null$PositiveProb, equals(beta.ash.vb.null2$PositiveProb))
+
 ```
 
-Simulate 10000 $\beta_j$ values, 8000 from a mixture of a point mass at 0 and
-2000 from $\beta \sim N(0,sd=2)$.
+```{r}
+require("ashr")
+require("qvalue")
+
+set.seed(100)
+
+#Test mixVBEM
+abf = rbind(c(1,0,0,0),c(0,1,0,0),c(0,0,1,0),c(0,0,1,0),c(0,0,1,0))
+eps = 1e-10
+abf[abf==0] = eps #replace 0 with small number
+print(all.equal(mixVBEM(abf,c(1,1,1,1))$post,c(2,2,4,1)))
+print(all.equal(mixVBEM(abf,c(1,2,1,1))$post,c(2,3,4,1)))
+
+#simulate n beta-hat values, nnull under the null
+#with altmean and altsd being the mean and sd of beta under the alternative
+simdata=function(n, nnull, altmean, altsd, betasd){
+  null = c(rep(1,nnull),rep(0,n-nnull))
+  beta = c(rep(0,nnull),rnorm(n-nnull,altmean,altsd))
+  betahat = rnorm(n,beta,betasd)
+  return(list(null=null,beta=beta,betahat=betahat,betasd=betasd))
+}
+
+ss = simdata(10000,8000,0,2,1)
+
+system.time((beta.ash=ash(ss$betahat,ss$betasd,sigmaavec=c(0,0.01,0.02,0.04,0.08,0.16,0.32,0.64,1.28,2.56,5.12,10.24))))
+system.time((beta.ash.auto = ash(ss$betahat, ss$betasd)))
+system.time((beta.ash.vb.uniform = ash(ss$betahat, ss$betasd, VB=TRUE, prior="uniform")))
+system.time((beta.ash.vb.null = ash(ss$betahat, ss$betasd, VB=TRUE, prior=NULL)))
+system.time((beta.ash.vb.null = ash(ss$betahat, ss$betasd, VB=TRUE, prior="strongnull")))
+
+hist(ss$beta,prob=TRUE,breaks=seq(floor(min(ss$beta)),ceiling(max(ss$beta)),length=20))
+x= seq(-4,4,length=10000)
+lines(x,density(beta.ash,x),col=2)
+lines(x,density(beta.ash.auto,x),col=3)
+lines(x,density(beta.ash.vb.uniform,x),col=4)
+lines(x,density(beta.ash.vb.null,x),col=5)
+summary(beta.ash)
+summary(beta.ash.auto)
+summary(beta.ash.vb.uniform)
+summary(beta.ash.vb.null)
+
+ss = simdata(10000,10000,0,2,1)
+system.time((beta.ash=ash(ss$betahat,ss$betasd)))
+system.time((beta.ash.auto = ash(ss$betahat, ss$betasd)))
+system.time((beta.ash.vb.uniform = ash(ss$betahat, ss$betasd, VB=TRUE, prior="uniform")))
+system.time((beta.ash.vb.null = ash(ss$betahat, ss$betasd, VB=TRUE, prior=NULL)))
+
+hist(ss$beta,prob=TRUE,breaks=seq(floor(min(ss$beta)),ceiling(max(ss$beta)),length=20))
+x= seq(-4,4,length=10000)
+lines(x,density(beta.ash,x),col=2)
+lines(x,density(beta.ash.auto,x),col=3)
+lines(x,density(beta.ash.vb.uniform,x),col=4)
+lines(x,density(beta.ash.vb.null,x),col=5)
+beta.ash$fitted.f
+beta.ash.auto$fitted.f
+beta.ash.vb.uniform$fitted.f
+beta.ash.vb.null$fitted.f
+
+beta.ash.vb.uniform$fit$loglik
+beta.ash.vb.uniform$fit$converged
+```
 
@@
 
-beta.ash=ash(ss$betahat,ss$betasd)
-xmin = min(ss$beta)
-xmax = max(ss$beta)
-x=seq(xmin,xmax,length=1000)
-plot(sort(ss$beta),(1:length(ss$beta))/length(ss$beta),main="cdf of ss$beta, with fitted f overlaid",xlab="beta",ylab="cdf")
-lines(x,cdf.ash(beta.ash,x),col=2,lwd=2)
-
-```
-
-Compare with mixfdr. Note that I tried to do small than 1000, but mixfdr returned
-errors.
-```{r}
-ss = simdata(1000,800,0,2,1)
-
-beta.ash=ash(ss$betahat,ss$betasd)
-beta.ash.u = ash(ss$betahat,ss$betasd, mixcompdist="uniform")
-beta.ash.hu = ash(ss$betahat,ss$betasd, mixcompdist="halfuniform")
-
-require(mixfdr)
-beta.mfdr.J3 = mixFdr(ss$betahat,J=3,noiseSD=1,theonull=TRUE)
-beta.mfdr.J5 = mixFdr(ss$betahat,J=5,noiseSD=1,theonull=TRUE)
-beta.mfdr.J7 = mixFdr(ss$betahat,J=7,noiseSD=1,theonull=TRUE)
-beta.mfdr.J9 = mixFdr(ss$betahat,J=9,noiseSD=1,theonull=TRUE)
-beta.mfdr.J15 = mixFdr(ss$betahat,J=15,noiseSD=1,theonull=TRUE)
-mean((beta.mfdr.J3$effectSize-ss$beta)^2)
-mean((beta.mfdr.J5$effectSize-ss$beta)^2)
-mean((beta.mfdr.J7$effectSize-ss$beta)^2)
-mean((beta.mfdr.J9$effectSize-ss$beta)^2)
-mean((beta.mfdr.J15$effectSize-ss$beta)^2)
-
-mean((beta.ash$PosteriorMean-ss$beta)^2)
-
-#This code mostly borrowed from the examples in mixFdr in R help
-m=beta.mfdr.J15
-mTrue = list(pi=c(0.8,0.2),mu=c(0,0),sig=c(1,2))
-par(mfcol=c(1,1))
-s = seq(-5,5,by=0.01)
-trueFdr = fdrMixModel(s,mTrue,c(TRUE,FALSE))
-fdrCurve = fdrMixModel(s,m,abs(m$mu - m$mu[1])<=0.002) # this tells it which indices to consider null
- plot(s, trueFdr, t = 'l', main = "fdr - Black is true", xlab = "z", ylab = "fdr", ylim = c(0,1))
- lines(s, fdrCurve, col = 2)
-points(ss$betahat,beta.ash$localfdr,col=3)
-points(ss$betahat,beta.ash.u$localfdr,col=4)
-points(ss$betahat,beta.ash.hu$localfdr,col=5)
-
-```
-
-Of course, if we do something where alternatives are not centered on 0 then
-ash does worse!
-```{r}
-#This is the example from the mixfdr demo
-z = c(rnorm(1000), rnorm(100,3,1))
- m = mixFdr(z, J = 3, noiseSD = 1)
- s = seq(-5,5,by=0.01)
- effCurve = effectSize(s,m)[,1]
- fdrCurve = fdrMixModel(s,m,abs(m$mu - m$mu[1])<=0.2) # this tells it which indices to consider null
-
-# compare to the Bayes estimator for this problem
- mTrue = list(pi = c(10,1)/11, mu = c(0,3), sig = c(1,1))
-# note that the model parameters are in terms of the MARGINAL not of the prior
- trueEff = effectSize(s,mTrue)[,1]
- trueFdr = fdrMixModel(s,mTrue,c(TRUE,FALSE))
- par(mfrow = c(2,1))
- plot(s, trueEff, t = 'l', main = "Effect Size - Black is true", xlab = "z", ylab = "E(delta|z)", ylim = c(-3,3))
- lines(s, effCurve, col = 2)
- plot(s, trueFdr, t = 'l', main = "fdr - Black is true", xlab = "z", ylab = "fdr", ylim = c(0,1))
- lines(s, fdrCurve, col = 2)
-
-z.ash = ash(z,1,mixcompdist="halfuniform")
-points(z,z.ash$localfdr,col=3)
-```
-
-Things to emphasise for paper:
-- the number of components is not critical; the unimodal constraint is enough. You can increase the number of components arbitrarily and the likelihood remains bounded.
-- the ability to incorporate item-specific measurement error
-- the ability to compare a model in which effects are proportional to error
-with model in which effects
-
-Some advantages of the unimodal constraint:
--more statsitically efficient (if true); show with small sample sizes.
-- less sensitive to number of components? Can allow number to tend to infinity?
-- may be less sensitive to local optima?
-- it allows us to easily just vary sigma on a grid, and fit pi, which
-makes allowing different noise levels really easy!
-
-```{r}
-ss = simdata(10000,8000,0,2,1)
-
-beta.ash=ash(ss$betahat,ss$betasd,auto=FALSE)
-beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)
+beta.ash=ash(ss$betahat,ss$betasd,sigmaavec=c(0,0.01,0.02,0.04,0.08,0.16,0.32,0.64,1.28,2.56,5.12,10.24))
+beta.ash.auto = ash(ss$betahat, ss$betasd)
 #these to test the VB version 
-beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior="uniform")
-beta.ash.vb.null = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior=NULL)
+beta.ash.vb.uniform = ash(ss$betahat, ss$betasd, VB=TRUE, prior="uniform")
+beta.ash.vb.null = ash(ss$betahat, ss$betasd, VB=TRUE, prior=NULL)
 
@@
 ```
-No, that's not it.
-
-So maybe it is the non-monotonic behaviour of the true beta values in this case.
-Let's try putting in a component at -1.5 to make the data fit this assumption better.
 
@@
 ```
-
-### Motivation
-
-It is common in statistics that you measure many ``similar" things imperfectly, and wish to estimate their values.
-The situation arises commonly in the kinds of genomics applications I am often involved in, but also in other areas of statistics.
-In genomics, for example, a very common goal
-is to compare the mean
-expression (activity) level of many genes in two conditions.
-Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-measurements are made on only a small number of
-samples in each condition - sometimes as few as one
-sample in each condition. Thus the error in estimates
-of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-in estimates of $\beta_j$ still greater.
-
-A fundamental idea is
-that the measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.
-
-Note on multiple comparisons: it isn't really a "problem" but an "opportunity". This viewpoint also espoused by Greenland and Robins. It isn't the number of tests that is relevant (the false
-discovery rate at a given threshold does not depend on the number of tests). It is the *results* of the tests that are relevant.
-
-Performing multiple comparisons, or multiple tests, is often
-regarded as a "problem". However, here we regard it instead as an opportunity - an opportunity to combine (or "pool") information across tests or comparisons.
-
-Imagine that you are preparing to perform 1 million tests,
-each based on a $Z$ score that is assumed to be $N(0,1)$ under the null.
-You first order these tests randomly, and begin by 
-performing the first test, which returns a $Z$ score of 4.
-At this point you are interrupted by a friend, who asks how the analysis is going.
-"It's early days, but looking promising" you reply.
-Well, who wouldn't?
-If the aim is to find lots of significant differences, a strong first result is surely a good outcome.
-
-At this point you have reason to expect that many of the
-subsequent tests also output strong results.
-
-Now consider two alternative scenarios for the remaining 999,999 tests. In the first scenario, the remaining tests
-produce $Z$ values that fit very well with the null, closely following a standard normal distribution; in the second
-scenario a large proportion of the remaining tests, say 50 percent, show outcomes that lie outside of $[-4,4]$.
-
-If your friend enquired after your analysis again, your
-response would surely differ in the first scenario ("Oh, it didn't pan out so well after all") vs the second ("It went great"). Further,
-in the first scenario, if your friend pressed you further about the results of the first test,
-you would likely, I think, be inclined to put them down to chance.
-In contrast, in the second scenario, the first test turned out to be, as you hoped, a harbinger of good things to come, and in this scenario you would likely regard that test as likely corresponding to a true discovery.
-
-The key point is that it is the *outcomes* of the tests, not the *number* of tests, that impacts interpretation of that first test.
-
-(Some may by pondering whether the fact that you are about to perform another
-999,999 tests should be considered pertinent in responding to your friend. Our view is that it is  irrelevant.
-The standard frequentist framework would disagree, because
-it requires the analyst to consider hypothetical repetitions
-of the "experiment", and so the fact that the experiment
-consists of a million tests is pertinent. However, this
-argument is a distraction from the main point.)
-
-Indeed, we believe that the practice of focussing on the *number* of tests performed is
-
-Focussing on the number of tests performed can be
-seen as an approximation.
-
-The standard argument is that,
-when performing multiple tests, some will be significant
-just by chance.

### 2013-10-18 Rcode/ash.test.rmd
export posterior_dist

@@
 set.seed(133)
-system.time((beta.ash=ash(ss$betahat,ss$betasd,randomstart=TRUE)))
+system.time((beta.ash=ash(ss$betahat,ss$betasd,prior="uniform",randomstart=TRUE)))
 set.seed(133)
-system.time((beta.ash2=oldash(ss$betahat,ss$betasd,auto=TRUE,randomstart=TRUE)))
+system.time((beta.ash2=oldash(ss$betahat,ss$betasd,auto=TRUE,prior="uniform",randomstart=TRUE)))
 

### 2013-10-18 Rcode/ash.oldfuncs.R

@@ -48,30 +48,6 @@ mixseLoglik = function(x,pi,mu,sigma,se,FUN="+"){
   return(sum(log(pi %*% dd)))
 }
 
-
-
-#return matrix of log densities of observations (betahat) 
-# assuming betahat_j \sim N(0, sebetahat_j^2 + sigmaavec_k^2)
-#normalized by maximum of each column
-#INPUT
-#betahat is n vector, 
-#sebetahat is n vector, 
-#sigmaavec is k vector
-#return is n by k matrix of the normal likelihoods, 
-# with (j,k)th element the density of N(betahat_j; mean=0, var = sebetahat_j^2 + sigmaavec_k^2)
-#normalized to have maximum 1 in each column
-matrix_dens = function(betahat, sebetahat, sigmaavec){
-  k = length(sigmaavec)
-  n = length(betahat)
-  ldens = dnorm(betahat,0,sqrt(outer(sebetahat^2,sigmaavec^2,FUN="+")),log=TRUE)
-  maxldens = apply(ldens, 1, max)
-  ldens = ldens - maxldens
-  return(exp(ldens))
-}
-
-
-
-
 #find mean and variance of a mixture of normals
 #INPUT: x is a list with elements pi mu and sigma, each k by n matrices
 #OUTPUT; the n vectors of mean and variances of mixtures correspondign to columns of pi, mu and sigma

### 2013-10-21 Rcode/ash.test.rmd
modified mix.R to remove require(truncnorm); amended README to note need for truncnorm

@@
 
-system.time((beta.ash.vb.null = ash(ss$betahat, ss$betasd, VB=TRUE, prior=NULL)))
-system.time((beta.ash.vb.null2 = ash(ss$betahat, ss$betasd, VB=TRUE, prior=NULL)))
+system.time((beta.ash.vb.null = ash(ss$betahat, ss$betasd, VB=TRUE, prior="nullbiased")))
+system.time((beta.ash.vb.null2 = ash(ss$betahat, ss$betasd, VB=TRUE, prior="nullbiased")))
 expect_that(beta.ash.vb.null$PosteriorMean, equals(beta.ash.vb.null2$PosteriorMean))

### 2013-10-24 Rcode/readme.rmd
start draft of paper

--- 2013-10-04 Rcode/readme.rmd
@@
 #install q value package, and load in ash code
-source("../Rcode/ash.R")
-source("../Rcode/ash.oldfuncs.R")
+library("ashr")
+#source("../Rcode/ash.R")
+#source("../Rcode/ash.oldfuncs.R")
+#source("../Rcode/mix.R")
 beta.ash.all = ash(betahat,s)

### 2013-10-24 paper/Rcode/NoisyIllustration.rmd
start draft of paper

--- 2013-10-04 Rcode/readme.rmd
@@
-Suppose that we measure, with error, a series of ``effects", $\beta_1,\dots,\beta_J$. To take just one concrete 
-example, $\beta_j$ could be 
-the difference in the mean (log) expression levels of gene $j$ ($j=1,\dots,J$) between 2 conditions. In this case, a measurement of
-the effect might be the difference in sample means obtained in the two conditions. We will let $\hat\beta_j$ denote the measured value of $\beta_j$, and assume that
-each measurement comes with an associated standard error, 
-$s_j$. A key aim here will be to take proper account of the fact that some measurements may be more precise than others: that is, to take proper account of variation in $s_j$ across $j$.
-
-A common goal, particularly in genomic studies, is to identify which $\beta_j$ differ from zero. This is commonly tackled
-by first computing a $p$ value for each $j$, testing $H_j:\beta_j=0$, and then applying methods based on False Discovery Rates, such as the qvalue package.
-For example, we could compute $Z$ scores $Z_j = \beta_j/s_j$, and translate $Z_j$ to a $p$ value,
-$p_j$, and then use now-standard methods to select a $p$ value threshold to control the FDR.
-
-There are two issues with this approach that I would like to address here.
-The first is that it really does not take proper account of the measurement errors.
-To see this, consider an example where half the measurements are quite precise, and the other half are really, really, poor.
-Intuitively, the poor measurements tell us nothing, and any sane analysis should effectively ignore them.
-However, in a standard FDR-type analysis, these poor measurements add ``noise" and affect estimated FDRs.
-This is because the $p$ values from the poor measurements will be effectively uniformly distributed, and some will be significant at any given threshold.
-
-I will illustrate this with an example. The following R code
-simulates some data where the effects are normally distributed,
-and the first 500 observations have good precision, and the next 500 have poor precision.
+The following R code simulates some data where the effects are normally distributed, and the first 500 observations have good precision, and the next 500 have poor precision.
 
@@
 
-The two main assumptions we will make here
-are that 
-* these effects are exchangeable: that is (by de Finetti's theorem) they can be
-thought of as being independent and identically distributed from some unknown distribution $g(\beta)$. 
-* these effects are symmetrically distributed about 0, with density monotonically decaying as we move away from 0.
-
-Specifically we will assume a parameteric form for $g$
-as a mixture of 0-centered normal distributions
-
-A key issue I want to address here is that if these measurements are made with different precisions, then we want to
-take this into account in our analysis.
-We can do this by assuming that the likelihood for $\beta$ is a normal, with mean $\hat\beta_p$ and standard deviation $s_p$. (Note that this is equivalent to the likelihood we would get
-if we ``observed" data $\hat\beta_p \sim N(\beta_p, s_p)$.)
+In contrast, if you use adaptive shrinkage, it makes no difference whether you use all the observations or the just the good ones -
+the noise is ignored (as it should be!)
 
 ```{r}
-#install q value package, and load in ash code
-source("../Rcode/ash.R")
-source("../Rcode/ash.oldfuncs.R")
+library("ashr")
 beta.ash.all = ash(betahat,s)
@@
 ```
-So here it makes no difference whether you use all the observations or the just the good ones -
-the noise is ignored (as it should be!)
-
-### Analyse Hedenfalk et al data.
-Note that there were a few genes with odd observations (eg in the thousands, rather than the usual expression measures), and 
-following Storey's ODP paper I removed the genes with any observation >20.
-```{r}
-
-#note: I edited nejm_brca_release.txt by removing columns 1-3
-print(getwd())
-hh = read.table("../data/nejm_brca_release_edit.csv",sep=",",skip=3)
-subset = apply(hh, 1, max)<20
-hh = hh[subset,]
-
-labs = read.table("../data/nejm_brca_release_edit.csv",sep=",",skip=1,nrows=1)
-labs = 1*(labs=="BRCA1") + 2 * (labs=="BRCA2")
-
-hh.betahat = apply(hh[,labs==1],1,mean) - apply(hh[,labs==2],1,mean)
-n1 = sum(labs==1)
-n2 = sum(labs==2)
-hh.sebetahat = sqrt(apply(hh[,labs==1],1,var)/n1 + apply(hh[,labs==2],1,var)/n2)
-hh.zscore = hh.betahat/hh.sebetahat
-hh.pval = pchisq(hh.zscore^2,df=1,lower.tail=F)
-hist(hh.pval,probability=TRUE,xlab="p value",main="Distribution of p values for Hedenfalk et al data",nclass=40,col=5)
-hh.q = qvalue(hh.pval)
-abline(h=hh.q$pi0,col=2,lwd=2)
-sum(hh.q$q<0.05)
-```
-
-Now compare this with ASH
-```{r}
-hh.ash = ash(hh.betahat,hh.sebetahat,df=n1+n2-2)
-sum(hh.ash$qvalue<0.05)
-plot(hh.ash$localfdr, log(hh.pval),main="Illustration of how ASH reorders significance of observations",ylim=c(-10,-1),xlim=c(0,0.1),cex=0.2,xlab="ASH local fdr",ylab="Original p value")
-```
-
-Now see how it compares with ignoring the variability in
-measurement error, by applying ASH to the z scores.
-```{r}
-hh.ash2 = ash(hh.zscore,rep(1,length(hh.zscore)),df=20)
-plot(hh.ash$localfdr,hh.ash2$localfdr,cex=0.2,ylab="ignoring sd",xlab="accounting for sd",main="Effect of accounting for measurement precision on local fdr")
-abline(a=0,b=1,col=2)
-cat(sum(hh.ash$qvalue<0.05),sum(hh.ash2$qvalue<0.05))
-cat(sum(hh.ash$localfdr<0.05),sum(hh.ash2$localfdr<0.05))
-plot(hh.ash2$localfdr, log(hh.pval),main="Illustration of how ASH without precision does not reorder significance?",ylim=c(-10,-1),xlim=c(0,0.1),cex=0.2,xlab="ASH local fdr",ylab="Original p value")
-```
-Hmm - this was unexpected. Why are more being called
-in the second analysis? Could it be that the effect sizes
-are larger for larger standard errors, as using the 
-z scores effectively assumes?
-
-It would be nice to be able to assess the fit of the different models
-(ie are effect sizes larger for larger  standard errors?).
-In principle this should be assessable via likelihood, but
-need to code it. This section is work in progress.
-
-The following computes the log likelihood for the two fitted ash models,
-first with $\beta \sim \sum_k \pi_k N(0,\sigma_k^2)$, and then $\beta_s \sim \sum_k \pi_k N(0, \sigma^2_k se^2_s)$:
-```{r}
-cat(LogLik_conv(hh.ash$fitted.g, hh.betahat, hh.sebetahat),mixseLoglik(hh.betahat,hh.ash$fitted.g$pi, hh.ash$fitted.g$mean, hh.ash$fitted.g$sd, hh.sebetahat),mixseLoglik(hh.betahat,hh.ash2$fitted.g$pi, hh.ash2$fitted.g$mean, sqrt(hh.ash2$fitted.g$sd^2+1), hh.sebetahat,FUN="*"))
-```
-Looks like the second model fits better.
-
-```{r}
-plot(hh.betahat, hh.sebetahat)
-```
-
-Another unexpected feature is that the ASH analysis based
-on z scores is rather different from the standard q value analysis, even though they are both based on z scores.
-Look:
-```{r}
-plot(as.numeric(hh.q$qvalue),hh.ash2$qval,cex=0.2,ylab="ASH q value",xlab="standard q value",main="ASH vs standard q value")
-cat(hh.ash2$fitted.g$pi[1],hh.q$pi0)
-```
-At first I was really worried about this - the pi0 estimate
-is much bigger using the p values. However, this is all to
-do with the assumptions. The q value approach estimates
-pi0 by looking at the p values near 1. Effectively it assumes
-that all the p values near 1 are null. At first this seems OK.
-But note that it actually corresponds to assuming that all the z scores near 0 are null - that is, you can never see a z score of 0 under the alternative.
-In contrast we assume that the distribution of z scores under the alternative also has its mode at 0 - under this assumption not all z scores near 0 will be null at all....
-these are very different assumptions.
-(Of course, z scores near 0 won't give "significant"" results in either approach.)
-
-Now I wanted to investigate what distribution for z values under the alternative the standard approach implicitly corresponds to.
-```{r}
-x= seq(-6,8,length=100)
-hist(hh.zscore,prob=TRUE,nclass=100)
-lines(x,dnorm(x)*hh.q$pi0,col=3,lwd=2)
-z.kerneld=density(hh.zscore)
-lines(z.kerneld$x,z.kerneld$y,col=2,lwd=2)
-plot(z.kerneld$x,z.kerneld$y-dnorm(z.kerneld$x)*0.6,type="l")
-abline(h=0)
-```
-
-### Errors in sign, and true 0s
-
-If the true value is 0 and you say it is positive,
-is this an error in sign? Well, yes, it is. But it is not
-negative, and it turns out you need to count errors of sign
-when truth is 0 as 0.5 of a false discovery to make things
-well calibrated. This is because the method gives the same
-results if the true value of mu is exactly 0, or if it has
-epsilon variance around 0. But of course the errors in sign
-in the latter occur at rate 0.5 because of random guessing.
-
-The following simulation illustrates this. It also
-illustrates that the method seems to be relatively robust to
-the situation where the true $\beta$ are not symmetric around 0.
-```{r}
-musim2 = c(rnorm(1000,-3,1),rnorm(1000,-1.5,1),rnorm(1000,0,1),rep(0,7000))
-zsim2 = musim2+rnorm(10000)
-zsim2.ash=ash(zsim2,rep(1,10000))
-hist(musim2[1:3000])
-errorinsign = ifelse(musim2==0,0.5,(zsim2.ash$PosteriorMean/musim2)<0)
-sum(errorinsign[zsim2.ash$qvalue<0.05])/sum(zsim2.ash$qvalue<0.05)
-```
-
-#### Miscellaneous
-
-Simulate data as in Efron, 2008, Section 7. 
-```{r}
-musim = c(rnorm(1000,-3,1),rep(0,9000))
-zsim = musim+rnorm(10000)
-zsim.ash=ash(zsim,rep(1,10000))
-```
-Need to implement a method to find CIs to look at this.
-
-code below here is not "tidied"
-
-This was to test the mixseLoglik function.
-```{r}
-n= length(hh.sebetahat)
-se.test =median(hh.sebetahat)
-bhat.test = rnorm(n,0,se.test)+rnorm(n,0,hh.sebetahat)
-bhat2.test = rnorm(n,0,hh.sebetahat) + rnorm(n,0,hh.sebetahat)
-
-mixseLoglik(bhat.test,c(0,1),c(0,0),c(se.test,se.test),hh.sebetahat)
-mixseLoglik(bhat.test,c(0,1),c(0,0),c(se.test,se.test),hh.sebetahat,FUN="*")
-
-mhat = mean(bhat2.test)
-sdhat = sd(bhat2.test)
-sum(dnorm(bhat2.test,mhat,sdhat,log=TRUE))
-mixseLoglik(bhat2.test,c(0,1),c(0,0),sqrt(c(2,2)),hh.sebetahat,FUN="*")
-
-mixseLoglik(bhat.test,c(0,1),c(0,0),c(se.test,se.test),hh.sebetahat)
-mixseLoglik(bhat.test,c(0,1),c(0,0),c(se.test,se.test),hh.sebetahat,FUN="*")
-
-z.test = bhat.test/hh.sebetahat
-z2.test = bhat2.test/hh.sebetahat
-
-p.test = pchisq(z.test^2,df=1,lower.tail=F)
-p2.test = pchisq(z2.test^2,df=1,lower.tail=F)
-
-hist(p.test)
-hist(p2.test)
-
-test1.ash1 = ash(bhat.test, hh.sebetahat)
-test1.ash2 = ash(z.test, rep(1,n))
-cat(sum(test1.ash1$qval<0.05),sum(test1.ash2$qval<0.05))
-
-test2.ash1 = ash(bhat2.test, hh.sebetahat)
-test2.ash2 = ash(z2.test, rep(1,n))
-cat(sum(test2.ash1$qval<0.05),sum(test2.ash2$qval<0.05))
-
-mixseLoglik(bhat.test,test1.ash1$fitted.g$pi,test1.ash1$fitted.g$mean,test1.ash1$fitted.g$sd,hh.sebetahat)
-mixseLoglik(bhat.test,test1.ash2$fitted.g$pi,test1.ash2$fitted.g$mean,sqrt(test1.ash2$fitted.g$sd^2+1),hh.sebetahat,FUN="*")
-
-mixseLoglik(bhat2.test,test2.ash1$fitted.g$pi,test2.ash1$fitted.g$mean,test2.ash1$fitted.g$sd,hh.sebetahat)
-mixseLoglik(bhat2.test,test2.ash2$fitted.g$pi,test2.ash2$fitted.g$mean,sqrt(test2.ash2$fitted.g$sd^2+1),hh.sebetahat,FUN="*")
-```
-
-Here we look at the distribution of the average of p ABFs,
-under the null, for different values of p. 
-```{r}
-nsim = 1000
-pmax = 10000
-z = matrix(rnorm(nsim*pmax),ncol=pmax)
-abf = ABF(z,1,1)
-pvec=c(5,10,20,50,100,1000,10000)
-q=rep(0,length(pvec))
-for(i in 1:length(pvec)){
-abf.m = rowMeans(abf[,1:pvec[i]])
-q[i] = quantile(abf.m,0.95)
-}
-plot(q)
-```
-
-Here look at the 95th percentile of ABF for different sigma:
-```{r}
-nsim = 10000
-z = rnorm(nsim)
-Wvec = seq(0,10,length=101)
-q= rep(0,length(Wvec))
-for(i in 1:length(Wvec)){
-  abf = ABF(z,1,Wvec[i])
-  q[i] = quantile(abf,0.95)
-}
-plot(q)
-```
-so it seems that even in the worst case, the 95th quantile is never much more than 2.
-It seems this could be used to bound the probability of making a false discovery even for p=1.
-Then (I believe) under the null, the probability of making a false discovery under the null should be even smaller for large p.
-
-And idea; this seems related to the sellke and berger result. at the 0.05 point, the ABF never exceeds about 2. Can we make arguments about false positive rates based on this?
-
-An Idea: for a single datum
-Pr(data | pi) = pi ABF + 1-pi
-
-so if a priori pi \sim Di(1,1)
-then pi | data \sim w1 Di(2,1) + w2 Di(1,2)
-so worst case is pi is \Di(2,1) and posterior mean of pi is 2/3.
-
-Or P(data and H1) = pi ABF
-P(data and H0) = (1-pi)
-
-```{r}
-
-#set up some data with mixture of values of sigmaa
-set.seed(100)
-#sebetahat = sample(c(1,0.1),size=1500,replace=T) 
-sebetahat = 0.01*rgamma(1500,.1,.1)
-beta =  c(rnorm(500,0,1),rnorm(500,0,0.5),rnorm(500,0,0.000001))
-betahat = beta + rnorm(1500,0,sebetahat)
-
-#compute the usual zscore and corresponding p value
-zscore = betahat/sebetahat
-pval = pchisq(zscore^2,df=1,lower.tail=F)
-
-#apply the ash method to do adaptive shrinkage of betahat
-#and the qvalue method to get 
-beta.ash = ash(betahat,sebetahat)
-qq = qvalue(pval)
-```
-
-Now we want to see how the results compare
-```{r}
-
-attach(beta.ash)
-conf=ifelse(PositiveProb>0.5,PositiveProb,1-PositiveProb)
-sum(conf>0.95)
-sum(qq$qvalues<0.05)
-```
-So there are more significant at FDR=0.05, than at local bayes FDR 0.05. But this is not suprising because of the difference between local FDR and FDR.
-
-Indeed, let's look at the errors
-```{r}
-err = (sign(betahat) != sign(beta))
-table(err[conf>0.95])
-table(err[qq$qvalues<0.05])
-
-#check whether ordering by q values does better or worse
-#job than ordering by confidence, in terms of identifying
-#betas with the right sign
-
-plot(cumsum(err[order(qq$qvalues)]),type="l")
-lines(cumsum(err[order(conf,decreasing=TRUE)]),col=2)
-```

### 2013-10-24 paper/main.tex
start draft of paper

--- 2013-10-11 Rcode/intro.rmd
@@
-```{r, echo=FALSE}
-#Note, this needs the ashr library installed.
-#To build the library run the following (from the Rcode subdirectory):
-#require(devtools)
-#install("../package/ashr")
-#Then RESTART R before knitting the following code
-```
+% Template for PLoS
+% Version 1.0 January 2009
+%
+% To compile to pdf, run:
+% latex plos.template
+% bibtex plos.template
+% latex plos.template
+% latex plos.template
+% dvipdf plos.template
 
-### Project Goals
+\documentclass[10pt]{article}
+
+% amsmath package, useful for mathematical formulas
+\usepackage{amsmath}
+% amssymb package, useful for mathematical symbols
+\usepackage{amssymb}
+\usepackage{booktabs}
+% graphicx package, useful for including eps and pdf graphics
+% include graphics with the command \includegraphics
+\usepackage{graphicx}
+
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
+\usepackage{color}
+
+% Use doublespacing - comment out for single spacing
+%\usepackage{setspace} 
+%\doublespacing
+
+% Text layout
+\topmargin 0.0cm
+\oddsidemargin 0.5cm
+\evensidemargin 0.5cm
+\textwidth 16cm 
+\textheight 21cm
+
+% Bold the 'Figure #' in the caption and separate it with a period
+% Captions will be left justified
+\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
+
+% Use the PLoS provided bibtex style
+\bibliographystyle{plos2009}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Leave date blank
+\date{}
+
+\pagestyle{myheadings}
+%% ** EDIT HERE **
+\usepackage{enumerate}
+\usepackage{multirow} 
+\usepackage{url}
+\usepackage{xr} %for cross-referencing
+%% ** EDIT HERE **
+%% PLEASE INCLUDE ALL MACROS BELOW
+\newtheorem{algorithm}{Algorithm}
+\newtheorem{proposition}{Proposition}
+\newtheorem{restateproposition}{Proposition}
+\newtheorem{lemma}{Lemma}
+\newtheorem{corollary}{Corollary}
+\newtheorem{result}{Result}
+\newtheorem{note}{Note}
+\newtheorem{definition}{Definition}
+
+\def\x{\mbf{x}}
+\def\y{\mbf{y}}
+\def\e{\mbf{e}}
+\def\g{\mbf{g}}
+\def\p{p}
+\def\Yb{\hat{Y}_1^{\text{BAYES}}}
+\def\Xperp{X_{1 \perp 0}}
+\def\Pperp{P_{1 \perp 0}}
+\def\Pperpb{\Pperp^{\text{B}}}
+\def \hatd { \widehat{ d } }
+\def \dhatin {\hatd^{\mbox{\tiny IN}}}
+\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
+\def \mvn{\text{MVN}}
+\def\mn{\text{MN}}
+\def \iw{\text{W}^{-1}}
+\def \fp {fastPHASE\ }
+\def\var{\rm Var}
+\def\mvr{\text{BMVR}}
+\def\bfuni{\text{BF}_\text{uni}}
+\def\one{\mbf{1}}
+\def\zero{\mbf{0}}
+\def\hall{H_{\text{all}}}
+\def\pall{\p_{\text{all}}}
+\def\bfall{\text{BF}_\text{all}}
+\def\bfav{\text{BF}_\text{av}}
+\def\gall{\gamma_\text{all}}
+\def\limbfall{\bfall^\rightarrow}
+\def\limbfav{\bfav^\rightarrow}
+\def\limbfgamma{\BF_\gamma^\rightarrow}
+\def\sxx{S_{xx}}
+\def\syy{S_{yy}}
+\def\vyy{V_{yy}}
+\def\vyx{V_{yx}}
+\def\vxx{{V_{xx}}}
+\def\rss{\text{\rm RSS}}
+\def\dss{\text{\rm $\Delta$SS}}
+\def\U{U}
+\def\D{D}
+\def\I{I}
+\def\luu{L_{UU}}
+\def\indep{\perp}
+\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
+                                % mode too
+\newcommand\height{{\it height}}
+\newcommand\weight{{\it weight}}
+\def\BF{\text{BF}}
+
+%% END MACROS SECTION
+\externaldocument{plos_one_rev1_suppinfo}
+
+\begin{document}
+
+% Title must be 150 characters or less
+\begin{flushleft}
+{\Large
+\textbf{Shrinkage, False Discovery Rate Estimation, and an alternative to the ``Zero assumption" when precision varies across units}
+}
+% Insert Author names, affiliations and corresponding author email.
+\\
+Matthew Stephens$^{1*}$, 
+\\
+\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
+\\
+$\ast$ E-mail: Corresponding mstephens@uchicago.edu
+\end{flushleft}
+
+% Please keep the abstract between 250 and 300 words
+\section*{Abstract}
+
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLoS ONE authors please skip this step. 
+% Author Summary not valid for PLoS ONE submissions.   
+%\section*{Author Summary}
+
+\section*{Introduction}
+
+Suppose that we measure, with error, a series of ``effects", $\beta_1,\dots,\beta_J$. To take just one concrete 
+example, $\beta_j$ could be 
+the difference in the mean (log) expression levels of gene $j$ ($j=1,\dots,J$) between 2 conditions. In this case, a measurement of
+the effect might be the difference in sample means obtained in the two conditions. We will let $\hat\beta_j$ denote the measured value of $\beta_j$, and assume that
+each measurement comes with an associated standard error, 
+$s_j$. A key aim here will be to take proper account of the fact that some measurements may be more precise than others: that is, to take proper account of variation in $s_j$ across $j$.
+
+A common goal, particularly in genomic studies, is to identify which $\beta_j$ differ from zero. This is commonly tackled
+by first computing a $p$ value for each $j$, testing $H_j:\beta_j=0$, and then applying methods based on False Discovery Rates, such as the qvalue package.
+For example, we could compute $Z$ scores $Z_j = \beta_j/s_j$, and translate $Z_j$ to a $p$ value,
+$p_j$, and then use now-standard methods to select a $p$ value threshold to control the FDR.
+
+There are two issues with this approach that I would like to address here.
+The first is that it really does not take proper account of the measurement errors.
+To see this, consider an example where half the measurements are quite precise, and the other half are really, really, poor.
+Intuitively, the poor measurements tell us nothing, and any sane analysis should effectively ignore them.
+However, in a standard FDR-type analysis, these poor measurements add ``noise" and affect estimated FDRs.
+This is because the $p$ values from the poor measurements will be effectively uniformly distributed, and some will be significant at any given threshold.
 
@@
 
-### Methods Outline
+It is common in statistics that you measure many ``similar" things imperfectly, and wish to estimate their values.
+The situation arises commonly in the kinds of genomics applications I am often involved in, but also in other areas of statistics.
+In genomics, for example, a very common goal
+is to compare the mean
+expression (activity) level of many genes in two conditions.
+Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
+of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
+measurements are made on only a small number of
+samples in each condition - sometimes as few as one
+sample in each condition. Thus the error in estimates
+of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
+in estimates of $\beta_j$ still greater.
+
+A fundamental idea is
+that the measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.
+
+ \section*{Methods}
+
+ The two main assumptions we will make here are that 
+* these effects are exchangeable: that is (by de Finetti's theorem) they can be
+thought of as being independent and identically distributed from some unknown distribution $g(\beta)$. 
+* these effects are symmetrically distributed about 0, with density monotonically decaying as we move away from 0.
+
+Specifically we will assume a parameteric form for $g$
+as a mixture of 0-centered normal distributions
+
+A key issue I want to address here is that if these measurements are made with different precisions, then we want to
+take this into account in our analysis.
+We can do this by assuming that the likelihood for $\beta$ is a normal, with mean $\hat\beta_p$ and standard deviation $s_p$. (Note that this is equivalent to the likelihood we would get
+if we ``observed" data $\hat\beta_p \sim N(\beta_p, s_p)$.)
 
@@
 
-### The False Sign Rate
+\subsection{The False Sign Rate}
 
@@
 
-As pointed out by Gelman et al, there are settings where
+As pointed out by \cite{gelman.etal}, there are settings where
 $\beta_j=0$ is implausible, in which case the lfdr is not
@@
 
-Following these ideas of Gelman et al, we define the local False Sign Rate (lfsr) as 
+Following these ideas of \cite{gelman.etal}, we define the local False Sign Rate (lfsr) as 
 $$\text{lfsr}_j = \min[ p(\beta_j \geq 0| \hat\beta, s), p(\beta_j \leq 0| \hat\beta, s) ].$$
@@
 
-### Computation Outline
+\subsection{Computation Outline}
 
@@
 
-### Example 1: A simple simulated example to get started
+\section*{Results}
 
-Load in some functions.
-```{r}
-#setwd("~/Documents/git/ash/Rcode/")
-set.seed(32327)
+- Illustration that 2 local false sign rate is approximately local fdr.
 
-## load ashr library 
-require("ashr")
-require("qvalue")
-```
+- Illustration of effects of contamination on FDR by low-quality data
+comparison with ash.
 
-```{r}
-#simulate n beta-hat values, nnull under the null
-#with altmean and altsd being the mean and sd of beta under the alternative
-simdata=function(n, nnull, altmean, altsd, betasd){
-  null = c(rep(1,nnull),rep(0,n-nnull))
-  beta = c(rep(0,nnull),rnorm(n-nnull,altmean,altsd))
-  betahat = rnorm(n,beta,betasd)
-  return(list(null=null,beta=beta,betahat=betahat,betasd=betasd))
-}
-```
+- Illustration of the difference between the zero assumption and no zero assumption on FDR
+in case z is N(0,2) vs N(0,1). [no difference in precision]
 
-Simulate 10000 $\beta_j$ values, 8000 from a mixture of a point mass at 0 and
-2000 from $\beta \sim N(0,sd=2)$.
+- Comparison between exchangeable effects vs Exchangeable standardized effects; Hedenberg data.
 
-```{r}
-ss = simdata(10000,8000,0,2,1)
+- Possibly difference between CIs obtained by ash and usual CIs (e.g. ``of intervals where local fsr is <0.05, what proportion are the sign correct"?).
 
-beta.ash=ash(ss$betahat,ss$betasd)
-xmin = min(ss$beta)
-xmax = max(ss$beta)
-x=seq(xmin,xmax,length=1000)
-plot(sort(ss$beta),(1:length(ss$beta))/length(ss$beta),main="cdf of ss$beta, with fitted f overlaid",xlab="beta",ylab="cdf")
-lines(x,cdf.ash(beta.ash,x),col=2,lwd=2)
-
-```
-
-Compare with mixfdr. Note that I tried to do smaller than 1000, but mixfdr returned
-errors. (Maybe this example is not so useful for comparing fdrs (as opposed to shrinkage), since the methods tend to agree where the fdr is small, which is what matters?)
-```{r}
-ss = simdata(1000,500,0,2,1)
-
-beta.ash=ash(ss$betahat,ss$betasd)
-beta.ash.u = ash(ss$betahat,ss$betasd, mixcompdist="uniform")
-beta.ash.hu = ash(ss$betahat,ss$betasd, mixcompdist="halfuniform")
-
-require(mixfdr)
-beta.mfdr.J3 = mixFdr(ss$betahat,J=3,noiseSD=1,theonull=TRUE)
-beta.mfdr.J5 = mixFdr(ss$betahat,J=5,noiseSD=1,theonull=TRUE)
-beta.mfdr.J7 = mixFdr(ss$betahat,J=7,noiseSD=1,theonull=TRUE)
-beta.mfdr.J9 = mixFdr(ss$betahat,J=9,noiseSD=1,theonull=TRUE)
-beta.mfdr.J15 = mixFdr(ss$betahat,J=15,noiseSD=1,theonull=TRUE)
-mean((beta.mfdr.J3$effectSize-ss$beta)^2)
-mean((beta.mfdr.J5$effectSize-ss$beta)^2)
-mean((beta.mfdr.J7$effectSize-ss$beta)^2)
-mean((beta.mfdr.J9$effectSize-ss$beta)^2)
-mean((beta.mfdr.J15$effectSize-ss$beta)^2)
-
-mean((beta.ash$PosteriorMean-ss$beta)^2)
-mean((beta.ash.u$PosteriorMean-ss$beta)^2)
-mean((beta.ash.hu$PosteriorMean-ss$beta)^2)
-
-#This code mostly borrowed from the examples in mixFdr in R help
-m=beta.mfdr.J15
-mTrue = list(pi=c(0.8,0.2),mu=c(0,0),sig=c(1,2))
-par(mfcol=c(1,1))
-s = seq(-5,5,by=0.01)
-trueFdr = fdrMixModel(s,mTrue,c(TRUE,FALSE))
-fdrCurve = fdrMixModel(s,m,abs(m$mu - m$mu[1])<=0.002) # this tells it which indices to consider null
- plot(s, trueFdr, t = 'l', main = "fdr - Black is true", xlab = "z", ylab = "fdr", ylim = c(0,1))
- lines(s, fdrCurve, col = 2)
-points(ss$betahat,beta.ash$localfdr,col=3)
-points(ss$betahat,beta.ash.u$localfdr,col=4)
-points(ss$betahat,beta.ash.hu$localfdr,col=5)
-
-```
-
-Of course, if we do something where alternatives are not centered on 0 then
-ash does worse!
-```{r}
-#This is the example from the mixfdr help example
-z = c(rnorm(1000), rnorm(100,3,1))
- m = mixFdr(z, J = 3, noiseSD = 1)
- s = seq(-5,5,by=0.01)
- effCurve = effectSize(s,m)[,1]
- fdrCurve = fdrMixModel(s,m,abs(m$mu - m$mu[1])<=0.2) # this tells it which indices to consider null
-
-# compare to the Bayes estimator for this problem
- mTrue = list(pi = c(10,1)/11, mu = c(0,3), sig = c(1,1))
-# note that the model parameters are in terms of the MARGINAL not of the prior
- trueEff = effectSize(s,mTrue)[,1]
- trueFdr = fdrMixModel(s,mTrue,c(TRUE,FALSE))
- par(mfrow = c(2,1))
- plot(s, trueEff, t = 'l', main = "Effect Size - Black is true", xlab = "z", ylab = "E(delta|z)", ylim = c(-3,3))
- lines(s, effCurve, col = 2)
- plot(s, trueFdr, t = 'l', main = "fdr - Black is true", xlab = "z", ylab = "fdr", ylim = c(0,1))
- lines(s, fdrCurve, col = 2)
-
-z.ash = ash(z,1,mixcompdist="halfuniform")
-points(z,z.ash$localfdr,col=3)
-```
-
-Checking coverage
-```{r}
-
-```
+- Asymmetric case (Half uniforms)
 
@@
 
-## The number of components is not critical
-
+\subsection{The number of components is not critical}
 OR - this section could be more generally about selecting between
 models (normal, uniforms etc) using log-likelihood.
-
-Don't need to worry about overfitting - only need to worry
-about whether model is "complex enough".
 
@@
 Indeed, as the number of components tends to infinity, the likelihood
-is bounded above, and the procedure will converge to a sensible limit.
-The following example illustrates this.
-We take 100 observations from 0.5N(-1,1)+0.5N(1,1)
-and add N(0,1) noise.
-```{r}
-sampsize=500
-beta = c(rnorm(sampsize,-1,1),rnorm(sampsize,1,1))
-betahat = beta+rnorm(2*sampsize)
+is bounded above, and even for large numbers of components and small amounts of data the inferred underlying distributions tend not to be too crazy, showing few signs of ``overfitting".
 
-beta.ash = ash(betahat,1)
-beta.ash.u=list()
-for(i in 1:10){
-  beta.ash.u[[i]] = ash(betahat,1,mixcompdist="uniform",prior="uniform",sigmaavec = seq(0.1,5,length=10*i))
-}
+However, it is true that the normal distributions tend to yield smoother estimated underlying distributions. Similarly, using the posterior mean for $\pi$, rather
+than the maximum likelihood estimate, tends to lead to somewhat smoother fitted distributions. In addition, with small amounts of data the underlying
+distributions are inevitably not well determined by the data, and the fits may vary depending on the underlying assumptions made (e.g.
+number of components or distribution of the components); however, even then shrinkage-based estimates
+of the betas can be relatively robust.
 
-plot(beta.ash,-4,4,col=2)
-x=seq(-4,4,length=1000)
-lines(x,0.5*dnorm(x,-1,1)+0.5*dnorm(x,1,1),col=1)
-for(i in 1:4){
-lines(x,density(beta.ash.u[[i]],x),col=2+i)
-}
-
-plot(unlist(lapply(beta.ash.u,ashr:::get_loglik)))
-
-```
-
-```{r}
-sampsize=5000
-beta = c(rnorm(sampsize,-1,1),rnorm(sampsize,1,1))
-betahat = beta+rnorm(2*sampsize)
-
-beta.ash = ash(betahat,1)
-beta.ash.u=list()
-for(i in 1:10){
-  beta.ash.u[[i]] = ash(betahat,1,mixcompdist="uniform",prior="uniform",sigmaavec = seq(0.1,5,length=10*i))
-}
-
-plot(beta.ash,-4,4,col=2)
-x=seq(-4,4,length=1000)
-lines(x,0.5*dnorm(x,-1,1)+0.5*dnorm(x,1,1),col=1)
-for(i in 1:4){
-lines(x,density(beta.ash.u[[i]],x),col=2+i)
-}
-
-plot(unlist(lapply(beta.ash.u,ashr:::get_loglik)))
-
-```
-
-```{r}
-ss = simdata(10000,8000,0,2,1)
-
-beta.ash=ash(ss$betahat,ss$betasd,auto=FALSE)
-beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)
-#these to test the VB version 
-beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,VB=TRUE, prior="uniform")
-beta.ash.vb.null = ash(ss$betahat, ss$betasd,VB=TRUE, prior=NULL)
-
-beta.ash.pm = ash(ss$betahat,ss$betasd,usePointMass=TRUE)
-
-#compute the usual zscore and corresponding p value
-zscore = ss$betahat/ss$betasd
-
-pval = pchisq(zscore^2,df=1,lower.tail=F)
-qval = qvalue(pval)
-hist(zscore)
-```
-
-Plot the fitted underlying distribution on top of true values for beta
-```{r}
-hist(ss$beta,prob=TRUE,breaks=seq(-7,7,length=20))
-x= seq(-7,7,length=1000)
-lines(x,density(beta.ash,x),col=2)
-
-plot(sort(ss$beta),(1:length(ss$beta))/length(ss$beta),main="cdf of ss$beta, with fitted f overlaid",xlab="beta",ylab="cdf")
-lines(x,cdf.ash(beta.ash,x),col=2,lwd=2)
-lines(x,cdf.ash(beta.ash.auto,x),col=3,lwd=2)
-lines(x,cdf.ash(beta.ash.vb.uniform,x),col=4,lwd=2)
-lines(x,cdf.ash(beta.ash.vb.null,x),col=5,lwd=2)
-lines(x,cdf.ash(beta.ash.pm,x),col=6,lwd=2)
-```
-
-[for testing: compare results with point mass and without]
-```{r}
-  plot(beta.ash$PositiveProb)
-  points(beta.ash.pm$PositiveProb,col=2)
-```
-
-```{r}
-  plot(beta.ash$ZeroProb)
-  points(beta.ash.pm$ZeroProb,col=2)
-```
-
-[for testing: compare with the results from the automatic way for selecting sigma]
-```{r}
-hist(ss$beta,prob=TRUE,breaks=seq(-7,7,length=20))
-x= seq(-4,4,length=1000)
-lines(x,density(beta.ash.auto,x),col=2)
-```
-
-[for testing: note that the PosteriorMean and PositiveProb don't depend much on sigmaa vec used ]
-
-```{r}
-plot(beta.ash.auto$PosteriorMean, beta.ash$PosteriorMean, xlab="Shrunk estimate from auto method", ylab="Shrunk estimate from fixed method")
-abline(a=0,b=1)
-plot(beta.ash.auto$localfdr, beta.ash$localfdr, xlab="lfdr from auto method", ylab="ldfr from fixed method")
-abline(a=0,b=1)
-
-```
-
-[And VB method produces similar results to EM method]
-```{r}
-plot(beta.ash.auto$PosteriorMean, beta.ash.vb.uniform$PosteriorMean, xlab="Shrunk estimate from auto method", ylab="Shrunk estimate from vb method")
-points(beta.ash.auto$PosteriorMean, beta.ash.vb.null$PosteriorMean, col=2)
-abline(a=0,b=1)
-plot(beta.ash.auto$localfdr, beta.ash.vb.uniform$localfdr, xlab="lfdr from auto method", ylab="ldfr from vb method")
-points(beta.ash.auto$localfdr, beta.ash.vb.null$localfdr, col=2)
-abline(a=0,b=1)
-
-```
-
-Also, we can see the effects of shrinkage: small estimates of $\hat\beta$ are
-shrunk to close to 0. Large estimates of $\hat\beta$ are shrunk less strongly because ash recognizes that these larger $\hat\beta$ are likely
-from the alternative, rather than the null.
-```{r}
-plot(ss$betahat,beta.ash$PosteriorMean,xlab="Observed betahat", ylab="Estimated beta (posterior mean)",ylim=c(-7,7),xlim=c(-7,7))
-abline(h=0)
-abline(a=0,b=1,col=2)
-```
-
-### Some additional notes
-
-#### Do we need a point mass at zero?
+\subsection{Do we need a point mass at zero?}
 
@@
 
-#### Comparison with qvalue
-
-Here we compare ash $q$ values with those from the qvalue package. 
-```{r}
-plot(qval$q,beta.ash$qval,main="comparison of ash and q value qvalues",xlab="qvalue", ylab="ash q values")
-abline(a=0,b=1)
-```
-
-In this example we see that qval overestimates the actual FDR. (This
-is because it assumes all the $p$ values near 1 are null, when they are not.)
-```{r}
-o = order(beta.ash$qval)
-plot(cumsum(ss$null[o])/(1:10000),qval$qval[o],col=2,type="l", xlab="actual FDR", ylab="q value", main="Comparison of actual FDR with q value")
-lines(cumsum(ss$null[o])/(1:10000),beta.ash$qval[o])
-abline(a=0,b=1)
-```
-
-### Miscellaneous
-
-code and text below here is work in progress and untidied.
-
-Here we simulate data, effectively as in Efron, 2008, Section 7.
-```{r}
-truenull = c(rep(0,1000),rep(1,9000))
-beta = c(rnorm(1000,-3,1),rep(0,9000))
-s= rep(1,10000)
-betahat = rnorm(10000,beta,s)
-
-beta.ash=ash(betahat,s)
-#compute the usual zscore and corresponding p value
-zscore = betahat/s
-pval = pchisq(zscore^2,df=1,lower.tail=F)
-qval = qvalue(pval)
-
-plot(betahat,beta.ash$PosteriorMean,xlab="Observed betahat", ylab="Estimated beta (posterior mean)",ylim=c(-7,7),xlim=c(-7,7))
-abline(h=0)
-abline(a=0,b=1,col=2)
-
-plot(qval$q,beta.ash$qval,main="comparison of ash and q value qvalues")
-abline(a=0,b=1)
-
-o = order(beta.ash$qval)
-
-plot(cumsum(truenull[o])/(1:10000),qval$qval[o],col=2,type="l")
-lines(cumsum(truenull[o])/(1:10000),beta.ash$qval[o])
-abline(a=0,b=1)
-```
-
-It seems that in this case the ash q values underestimate the
-FDR slightly. Possibly this is the assymetry.
-Try shrinking positive and negatives separately:
-```{r}
-pos= betahat>0
-betapos.ash=ash(betahat[pos],s[pos])
-betaneg.ash = ash(betahat[!pos],s[!pos])
-lfdr = rep(0,length(betahat))
-lfdr[pos] = betapos.ash$localfdr
-lfdr[!pos] = betaneg.ash$localfdr
-qv = qval.from.localfdr(lfdr)
-o = order(qv)
-plot(cumsum(truenull[o])/(1:10000),qv[o],type="l")
-abline(a=0,b=1)
-```
-No, that's not it.
-
-So maybe it is the non-monotonic behaviour of the true beta values in this case.
-Let's try putting in a component at -1.5 to make the data fit this assumption better.
-
-### Illustration of halfuniform method
-
-I will also use this to illustrate the halfuniform method, which
-allows for asymmetry.
-
-```{r}
-truenull = c(rep(0,2000),rep(1,8000))
-beta = c(rnorm(1000,-3,1),rnorm(1000,-1.5,1),rep(0,8000))
-s= rep(1,10000)
-betahat = rnorm(10000,beta,s)
-
-beta.ash=ash(betahat,s)
-beta.ash.halfu = ash(betahat,s,"halfuniform")
-
-hist(beta,prob=TRUE)
-x = seq(-8,4,length=1000)
-lines(x,density(beta.ash,x),col=2)
-lines(x,density(beta.ash.halfu,x),col=3)
-
-```
-
-See how the use of the asymmtric uniform component does a better job capturing
-the underlying density. Probably clearer in the cdf plot:
-```{r}
-plot(sort(beta),(1:length(beta))/length(beta),main="cdf of beta, with fitted f overlaid",xlab="beta",ylab="cdf")
-lines(x,cdf.ash(beta.ash,x),col=2,lwd=2)
-lines(x,cdf.ash(beta.ash.halfu,x),col=3,lwd=2)
-
-#compute the usual zscore and corresponding p value
-zscore = betahat/s
-pval = pchisq(zscore^2,df=1,lower.tail=F)
-qval = qvalue(pval)
-plot(qval$q,beta.ash$qval,main="comparison of ash and q value qvalues")
-abline(a=0,b=1)
-
-o = order(qval$qval)
-plot(cumsum(truenull[o])/(1:10000),qval$qval[o],col=1,type="l")
-o = order(beta.ash$qval)
-lines(cumsum(truenull[o])/(1:10000),beta.ash$qval[o],col=2)
-o2 = order(beta.ash.halfu$qval)
-lines(cumsum(truenull[o2])/(1:10000),beta.ash.halfu$qval[o2],col=3)
-
-abline(a=0,b=1)
-
-```
-
-### Motivation
-
-It is common in statistics that you measure many ``similar" things imperfectly, and wish to estimate their values.
-The situation arises commonly in the kinds of genomics applications I am often involved in, but also in other areas of statistics.
-In genomics, for example, a very common goal
-is to compare the mean
-expression (activity) level of many genes in two conditions.
-Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-measurements are made on only a small number of
-samples in each condition - sometimes as few as one
-sample in each condition. Thus the error in estimates
-of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-in estimates of $\beta_j$ still greater.
-
-A fundamental idea is
-that the measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.
+\subsection{Side notes on Multiple comparisons}
 
@@
 just by chance.
+
+% Do NOT remove this, even if you are not including acknowledgments
+\section*{Acknowledgments}
+
+Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
+prepared using \LaTeX.
+
+%\section*{References}
+% The bibtex filename
+\bibliography{/Users/stephens/Documents/mainbib}
+
+\section*{Figure Legends}
+
+\clearpage
+
+\section*{Tables}
+
+\section*{Supporting Information Legends}
+
+Supplementary material can be found in {\bf Supplementary Information S1.}
+%\begin{table}[!ht]
+%\caption{
+%\bf{Table title}}
+%\begin{tabular}{|c|c|c|}
+%table information
+%\end{tabular}
+%\begin{flushleft}Table caption
+%\end{flushleft}
+%\label{tab:label}
+% \end{table}
+
+\end{document}

### 2013-11-01 Rcode/EmpiricalNull.rmd
added work on talk for NS lab meeting

```{r}
  require(ashr)
  hh = read.table("../data/nejm_brca_release_edit.csv",sep=",",skip=3)
  subset = apply(hh, 1, max)<20
  hh = hh[subset,]

  labs = read.table("../data/nejm_brca_release_edit.csv",sep=",",skip=1,nrows=1)
  labs = 1*(labs=="BRCA1") + 2 * (labs=="BRCA2")

  hh.betahat = apply(hh[,labs==1],1,mean) - apply(hh[,labs==2],1,mean)
  n1 = sum(labs==1)
  n2 = sum(labs==2)
  hh.sebetahat = sqrt(apply(hh[,labs==1],1,var)/n1 + apply(hh[,labs==2],1,var)/n2)
  hh.zscore = hh.betahat/hh.sebetahat
  hh.pval = pchisq(hh.zscore^2,df=1,lower.tail=F)
  hist(hh.pval,probability=TRUE,xlab="p value",main="Distribution of p values for Hedenfalk et al data",nclass=40,col=5)
  hh.q = qvalue(hh.pval)
  abline(h=hh.q$pi0,col=2,lwd=2)
  sum(hh.q$q<0.05)

```

Here lambda1>1,lambda2>0 allows for an "inflation factor": betahat is N(beta,lambda1*sebetahat+lambda2) or in other words, betahat is the sum of beta
and a "confounding" beta, which is distributed as N(0,(lambda1-1) sebetahat +lambda2)

```{r}
  hh.ash1.0 = ash(hh.betahat,hh.sebetahat,lambda1=1)
  hh.ash1.1 = ash(hh.betahat,hh.sebetahat,lambda1=1.1)
  hh.ash1.2 = ash(hh.betahat,hh.sebetahat,lambda1=1.2)
  hh.ash1.5 = ash(hh.betahat,hh.sebetahat,lambda1=1.5)
  hh.ash2.0 = ash(hh.betahat,hh.sebetahat,lambda1=2.0)
  cat(get_loglik(hh.ash2.0),get_loglik(hh.ash1.5),get_loglik(hh.ash1.2),get_loglik(hh.ash1.1),get_loglik(hh.ash1.0))

  plot(hh.ash1.0,-5,5)
  plot(hh.ash1.1,-5,5)
  plot(hh.ash1.5,-5,5)
  plot(hh.betahat/hh.sebetahat,hh.ash2$PositiveProb)
```

Now try applying empirical null to z scores, expecting that lambda1 won't be
identifiable. However, there will be a maximum value it can take.
```{r}
  hh.ashz1.0 = ash(hh.betahat/hh.sebetahat,1,lambda1=1)
  hh.ashz1.1 = ash(hh.betahat/hh.sebetahat,1,lambda1=1.1)
  hh.ashz1.2 = ash(hh.betahat/hh.sebetahat,1,lambda1=1.2)
  hh.ashz1.5 = ash(hh.betahat/hh.sebetahat,1,lambda1=1.5)
  hh.ashz1.6 = ash(hh.betahat/hh.sebetahat,1,lambda1=1.6)
  hh.ashz1.7 = ash(hh.betahat/hh.sebetahat,1,lambda1=1.7)
  hh.ashz1.8 = ash(hh.betahat/hh.sebetahat,1,lambda1=1.8)
  hh.ashz1.9 = ash(hh.betahat/hh.sebetahat,1,lambda1=1.9)

  hh.ashz2.0 = ash(hh.betahat/hh.sebetahat,1,lambda1=2.0)
  cat(get_loglik(hh.ashz2.0),get_loglik(hh.ashz1.9),get_loglik(hh.ashz1.8),get_loglik(hh.ashz1.7),get_loglik(hh.ashz1.6),get_loglik(hh.ashz1.5),get_loglik(hh.ashz1.2),get_loglik(hh.ashz1.1),get_loglik(hh.ashz1.0))

  plot(hh.betahat/hh.sebetahat,hh.ashz1.6$PositiveProb)
  points(hh.betahat/hh.sebetahat,hh.ashz1.5$PositiveProb,col=2)
```

### 2013-11-01 paper/main.tex
added work on talk for NS lab meeting

@@
 
+The second issue is that directly modeling the p values, say via non-parametric methods, without taking account of their precision, 
+can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
+distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
+The solution is to model beta hat as a convolution of some distribution g and an error component.
+
 The initial goal of the ASH (Adaptive SHrinkage) project is to provide simple, generic, and flexible methods to derive ``shrinkage-based" estimates and credible intervals for unknown quantities $\beta=(\beta_1,\dots,\beta_J)$, given only estimates of those quantities ($\hat\beta=(\hat\beta_1,\dots, \hat\beta_J)$) and their corresponding estimated standard errors ($s=(s_1,\dots,s_J)$).
@@
 and credible intervals/regions.
+
+Efron (2008) states ``most of the z-values near zero come from null genes"
+His main aim in making this assumption is to estimate an empirical null though (not assume N(0,1) for the null)
+rather than to impose identifiability.
+
+By modeling the z scores directly under the alternative, rather than the z scores as being the truth plus noise,
+maybe that is what is most problematic? Because the resulting distribution of z scores is just not possible.
+
+Note that \cite{muralidharan:2012} models $z$ scores as something plus noise under both $H_0$ and $H_1$, which
+avoids this problem. (DOes the same maybe apply to modeling beta, rather than z scores, when the errors vary?)
+
+Rice and Spiegelhalter - BRCA data?
 

### 2013-11-01 Rcode/readme.rmd
added work on talk for NS lab meeting

@@
 
+Try resampling the z scores from the alternative:
+```{r}
+  require(locfdr)
+  hh.locfdr = locfdr(hh.zscore,nulltype=0)
+  weights = 1-hh.locfdr$fdr
+  altz = hh.zscore[runif(length(weights))<weights]
+  hist(altz)
+```
+
+Directly modelling the z scores, without taking account of their errors,
+leads to distributions under the
+alternative that are simply implausible.
+```{r}
+  require("mixfdr")
+  source("~/Documents/git/ash/Rcode/mix.R")
+  altz.mixfdr = mixFdr(altz,noiseSD=1,J=100)
+  g= normalmix(altz.mixfdr$pi,altz.mixfdr$mu,altz.mixfdr$sigma)
+  x= seq(-8,8,length=100)
+  par(mfcol=c(1,1))
+  hist(altz,breaks = x,prob=TRUE)
+  lines(x,dens(g,x),col=2)
+```
+
+Directly modelling the p values is maybe even worse?
+Here we fit a monotonic decreasing function to the p values.
+Then we subtract the uniform distribution (the null) away from this, to
+give the alternative. 
+```{r}
+  hist(hh.pval,probability=TRUE,xlab="p value",main="Distribution of p values for Hedenfalk et al data",nclass=40,col=5)
+  require(fdrtool)
+  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue")
+  weights2 = 1-hh.fdrtool$lfdr
+  altz2 = hh.zscore[runif(length(weights2))<weights2]
+  hist(altz2,nclass=100)
+  hist(altz,nclass=100)
+  plot(hh.zscore,hh.locfdr$fdr)
+  points(hh.zscore,hh.fdrtool$lfdr,col=2)
+```
+
 ### Errors in sign, and true 0s

### 2013-11-01 paper/Rcode/CurrentMethodsIllustration.rmd
added work on talk for NS lab meeting

```{r}
  setwd("~/Documents/git/ash/paper/Rcode")
  require(ashr)
  require(qvalue)

  altcol="cyan" #colors to use
  nullcol="blue" 
  nc=40 #number of bins in histograms
  ncz = 100 # number of bins in z score histograms

  hh = read.table("../../data/nejm_brca_release_edit.csv",sep=",",skip=3)
  subset = apply(hh, 1, max)<20
  hh = hh[subset,]

  labs = read.table("../../data/nejm_brca_release_edit.csv",sep=",",skip=1,nrows=1)
  labs = 1*(labs=="BRCA1") + 2 * (labs=="BRCA2")

  hh.betahat = apply(hh[,labs==1],1,mean) - apply(hh[,labs==2],1,mean)
  n1 = sum(labs==1)
  n2 = sum(labs==2)
  hh.sebetahat = sqrt(apply(hh[,labs==1],1,var)/n1 + apply(hh[,labs==2],1,var)/n2)
  hh.zscore = hh.betahat/hh.sebetahat
  hh.pval = pchisq(hh.zscore^2,df=1,lower.tail=F)
  pdf("figures/FDR_illus.pdf")
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values for Hedenfalk et al data",nclass=nc,col=altcol)

  hh.q = qvalue(hh.pval)
  abline(h=hh.q$pi0,col=nullcol,lwd=2)

  hh.hist$density=rep(hh.q$pi0,length(hh.hist$density))  
  #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
  plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)

  abline(v=0.1,lwd=2,col=2)

  text(0.05,1.1,labels="A",col=2,cex=1.7)  
  text(0.05,0.4,labels="B",col=2,cex=1.7)  
  text(0.4,3,labels=paste0("FDR = B/(A+B) =  ",round(hh.q$pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.7)
  dev.off()
```

Now I wanted to investigate what distribution for z values under the alternative the standard approach implicitly corresponds to.
We use three different approaches.
The first, fdrtool, we use to directly model the p values. fdrtool assumes
that the density of p values under the alternative is monotonically decreasing away from 0, and fits a non-parametric maximum likelihood distribution under this asssumption.

The other two approaches, locfdr and mixfdr, we use to directly model the z scores.

In each case, the distribution of z scores under the alternative is implied by a) the overall distribution of z scores, f(z); and b) the localfdr for each z score. Specifically $f_1(z)=(1-lfdr(z)) * f(z)$.

Note that mixfdr does not make the zero assumption, and actually models
the z score errors, so its estimated distribution for f1 is at least theoretically possible. However, we still view thie bimodal distribution as generally implausible.

Our conclusion is that making the zero assumption leads to distributions of the z scores under the
alternative that are generally implausible.
```{r}
#plot a histogram of z scores, highlighting the alternative distribution
#of z scores that is implied by localfdr values lfdr.
  nullalthist = function(z,lfdr,...){
    h=hist(z, freq=FALSE,col=nullcol,nclass=ncz,...)
    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
    h$density = (1-avlfdr) * h$density
    plot(h,add=TRUE,col=altcol,freq=FALSE)
  }

  pdf("figures/zhist.pdf")
  par(mfcol=c(3,1))
  require(fdrtool)
  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue",plot=FALSE)
  nullalthist(hh.zscore,hh.fdrtool$lfdr,main="fdrtool")

  require(locfdr)
  hh.locfdr = locfdr(hh.zscore,nulltype=0,plot=0)
  nullalthist(hh.zscore,hh.locfdr$fdr,main="locfdr")

  require(mixfdr)
  hh.mixfdr = mixFdr(hh.zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
  nullalthist(hh.zscore,hh.mixfdr$fdr,main="mixfdr")
  dev.off()

  par(mfcol=c(1,1))
  pdf("figures/zhist_fdrtool.pdf")
  require(fdrtool)
  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue",plot=FALSE)
  nullalthist(hh.zscore,hh.fdrtool$lfdr,main="fdrtool")
  dev.off()

  pdf("figures/zhist_locfdr.pdf")
  require(locfdr)
  hh.locfdr = locfdr(hh.zscore,nulltype=0,plot=0)
  nullalthist(hh.zscore,hh.locfdr$fdr,main="locfdr")
  dev.off()

  pdf("figures/zhist_mixfdr.pdf")
  require(mixfdr)
  hh.mixfdr = mixFdr(hh.zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
  nullalthist(hh.zscore,hh.mixfdr$fdr,main="mixfdr")
  dev.off()

```

```{r}
  hh.ashz = ash(hh.zscore,1,usePointMass = TRUE, VB=TRUE, prior="nullbiased")

  #now try a different prior to check robustness
  mixsd = c(0,0.01,0.02,0.04,0.08,0.16,0.32,0.64,1.28,2.56,5.12,10.24)
  prior = rep(1,length(mixsd))
  prior[1] = 100
  hh.ashz2 = ash(hh.zscore,1,mixsd=mixsd,prior=prior,VB=TRUE)

  pdf("figures/zhist_ash.pdf")
  nullalthist(hh.zscore,hh.ashz$ZeroProb,main="ash")
  dev.off()

```

Compare pi0 estimates
```{r}
  c(hh.fdrtool$param[3],
  hh.locfdr$fp0[1,3],
  hh.mixfdr$pi[1],
  hh.ashz$fitted.g$pi[1])
```

Number signficant at 0.05 localfdr
```{r}
  c(sum(hh.fdrtool$lfdr<0.05),
  sum(hh.locfdr$fdr<0.05),
  sum(hh.mixfdr$fdr<0.05),
  sum(hh.ashz$ZeroProb<0.05))
```

### 2013-11-01 talks/NSmeet2013.Rmd
added work on talk for NS lab meeting

% Shrinkage, False Discovery Rates, and an Alternative to the Zero Assumption
% Matthew Stephens
% 2013/11/1

```{r setup, include=FALSE}
# set global chunk options
opts_chunk$set(cache=TRUE,autodep=TRUE)
dep_auto()
library("qvalue")
library("ashr")
```

# The Canonical Genomics Experiment

- Measure lots of things, with error

- Get estimates of effects $\beta_j$ ($\hat\beta_j$) and their standard errors $s_j$

- Turn these into Z-scores, $z_j = \hat\beta_j/s_j$

- Turn these into $p$ values, $p_j$

- Apply `qvalue` 
to identify findings ``significant" at a given False Discovery Rate.

- ...?

# Example: BRCA1 vs BRCA2 expression

```{r, echo=FALSE}
  #Set up Analysis of Hedenfalk BRCA data
  setwd("~/Documents/git/ash/paper/Rcode")

  altcol="cyan" #colors to use
  nullcol="blue" 
  nc=40 #number of bins in histograms
  ncz = 100 # number of bins in z score histograms

  hh = read.table("../../data/nejm_brca_release_edit.csv",sep=",",skip=3)
  subset = apply(hh, 1, max)<20
  hh = hh[subset,]

  labs = read.table("../../data/nejm_brca_release_edit.csv",sep=",",skip=1,nrows=1)
  labs = 1*(labs=="BRCA1") + 2 * (labs=="BRCA2")

  hh.betahat = apply(hh[,labs==1],1,mean) - apply(hh[,labs==2],1,mean)
  n1 = sum(labs==1)
  n2 = sum(labs==2)
  hh.sebetahat = sqrt(apply(hh[,labs==1],1,var)/n1 + apply(hh[,labs==2],1,var)/n2)
  hh.zscore = hh.betahat/hh.sebetahat
  hh.pval = pchisq(hh.zscore^2,df=1,lower.tail=F)

```

```{r, echo=FALSE,fig.height=3.5,fig.width=5,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

```

# Example: BRCA1 vs BRCA2 expression

```{r, echo=FALSE,fig.height=3.5,fig.width=5,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

  hh.q = qvalue(hh.pval)
  abline(h=hh.q$pi0,col=nullcol,lwd=2)

  hh.hist$density=rep(hh.q$pi0,length(hh.hist$density))  
  #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
  plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)

```

# Example: BRCA1 vs BRCA2 expression

```{r, echo=FALSE,fig.height=3.5,fig.width=5,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

  hh.q = qvalue(hh.pval)
  abline(h=hh.q$pi0,col=nullcol,lwd=2)

  hh.hist$density=rep(hh.q$pi0,length(hh.hist$density))  
  #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
  plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)

  abline(v=0.1,lwd=2,col=2)

  text(0.05,1.2,labels="A",col=2,cex=1.2)  
  text(0.05,0.4,labels="B",col=2,cex=1.2)  
  text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(hh.q$pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
```

# Example 2: Mouse Heart Data

```{r, echo=FALSE}
setwd("~/Documents/git/ash/talks/")
## load Poisson_binomial and ash functions 
source("../../stat45800/code/Rcode/PoissonBinomial.funcs.R")  
#source("../Rcode/ash.R")

x = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
xx = rowSums(x[,2:5])
x = x[xx>0,]
xx = xx[xx>0]
```

```{r echo=FALSE}
head(x)
```

- Data on 150 mouse hearts, dissected into left and right ventricle
(courtesy Scott Schmemo, Marcelo Nobrega)

```{r, echo=FALSE}

cc = x[, 2:5]

g = c(-1, -1, 1, 1)
ngene = dim(x)[1]

cc.assoc = counts.associate(cc, g, 1)
zdat.ash = cc.assoc$zdat.ash
zdat = cc.assoc$zdat

# two-sided test
ttest.pval = function(t, df) {
    pval = pt(t, df = df, lower.tail = T)
    pval = 2 * ifelse(pval < 0.5, pval, 1 - pval)
    return(pval)
}

tscore = zdat[3, ]/zdat[4, ]
pval = ttest.pval(tscore, df = 2)
qval = qvalue(pval)

highxp = xx>1000 # select high expressed genes
pval.high = pval[highxp]
qval.high = qvalue(pval.high)
cc.assoc.high = counts.associate(cc[highxp,],g,1)
zdat.ash.high = cc.assoc.high$zdat.ash
zdat.high = cc.assoc.high$zdat

```

# Example 2: Mouse Heart Data

```{r, echo=FALSE,fig.height=4,fig.cap=""}
h=hist(pval,prob=TRUE,ylim=c(0,4),main="p values, all genes",xlab="p value",breaks=seq(0,1,length=21),col=altcol)
abline(h=qval$pi0,col=nullcol,lwd=2)
h$density=rep(qval$pi0,length(h$density))  
plot(h,col=nullcol,add=TRUE,freq=FALSE)
abline(v=0.05,lwd=3)
text(0.6,3,labels=paste0("FDR = ",round(max(qval$qval[pval<0.05]),2)),cex=1.2)
```

# FDR problem 1: different genes have different precision/power

```{r, echo=FALSE,fig.height=4,fig.cap=""}
hist(log10(xx),main="Counts vary considerably across genes", xlab="log10(counts)")
```

# FDR problem 1: lower count genes, less power, add noise

```{r, echo=FALSE,fig.height=4,fig.cap=""}
h=hist(pval[xx<1000],prob=TRUE,ylim=c(0,4),main="p values, low count genes",xlab="p value",breaks=seq(0,1,length=21),col=altcol)
qval.low=qvalue(pval[xx<1000])
pval.low=pval[xx<1000]
abline(h=qval.low$pi0,col=nullcol,lwd=2)
h$density=rep(qval.low$pi0,length(h$density))  
plot(h,col=nullcol,add=TRUE,freq=FALSE)
abline(v=0.05,lwd=3)
#text(0.6,3,labels=paste0("FDR = ",round(max(qval.low$qval[pval.low<0.05]),2)),cex=1.2)
```

# FDR problem 1: higher count genes, more power
```{r, echo=FALSE, fig.height=4,fig.cap=""}

h=hist(pval.high,prob=TRUE,ylim=c(0,4),main="p values, high count genes",xlab="p value",col=altcol)

abline(h=qval.high$pi0,col=nullcol,lwd=2)
h$density=rep(qval.high$pi0,length(h$density))  
plot(h,col=nullcol,add=TRUE,freq=FALSE)
abline(v=0.05,lwd=3)
text(0.6,3,labels=paste0("FDR = ",round(max(qval.high$qval[pval.high<0.05]),2)),cex=1.2)
#qval.high$q[pval.high<0.05] gives FDR for p<0.05
```

# FDR problem 1: q values increased by low count genes
```{r,echo=FALSE, fig.height=4,fig.cap=""}
plot(qval$qval[xx>1000],qval.high$qval,xlab="q values from all gene analysis", ylab="q values from high-count gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
abline(a=0,b=1,col=2)
```

# Problem 2: The Zero Assumption (ZA)

- The standard `qvalue` 
approach assumes that all the $p$ values near 1 are null.

- Analogously, one can assume that all Z scores near 0 are null. Efron refers to this as the ``Zero Assumption".

- This allows us to estimate $\pi_0$ ``conservatively" using the density of $p$ values near 1.

# Problem 2: The ZA

- The ZA seems initially natural.

- However, it turns out to imply unrealistic assumptions about the distribution of non-zero effects.

```{r, echo=FALSE, include=FALSE}
require(fdrtool)
  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue",plot=FALSE)
require(locfdr)
  hh.locfdr = locfdr(hh.zscore,nulltype=0,plot=0)
require(mixfdr)
  hh.mixfdr = mixFdr(hh.zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
```

# Implied distribution of $p$ values under $H_1$

```{r, echo=FALSE,fig.height=3.5,fig.width=5,fig.cap=""}
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

  hh.q = qvalue(hh.pval)
  abline(h=hh.q$pi0,col=nullcol,lwd=2)

  hh.hist$density=rep(hh.q$pi0,length(hh.hist$density))  
  #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
  plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)

  abline(v=0.1,lwd=2,col=2)

  text(0.05,1.2,labels="A",col=2,cex=1.2)  
  text(0.05,0.4,labels="B",col=2,cex=1.2)  
  text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(hh.q$pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
```

# Implied distribution of Z scores under alternative (fdrtool)

```{r, echo=FALSE,fig.height=4,fig.cap=""}
#plot a histogram of z scores, highlighting the alternative distribution
#of z scores that is implied by localfdr values lfdr.
  nullalthist = function(z,lfdr,...){
    h=hist(z, freq=FALSE,col=nullcol,nclass=ncz,...)
    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
    h$density = (1-avlfdr) * h$density
    plot(h,add=TRUE,col=altcol,freq=FALSE)
  }

  nullalthist(hh.zscore,hh.fdrtool$lfdr,main="fdrtool")  
```

# Implied distribution of Z scores under alternative (locfdr)

```{r, echo=FALSE,fig.height=4,fig.cap=""}

  nullalthist(hh.zscore,hh.locfdr$fdr,main="locfdr")
```

# Implied distribution of Z scores under alternative (mixfdr)

```{r, echo=FALSE,fig.height=4,fig.cap=""}

  nullalthist(hh.zscore,hh.mixfdr$fdr,main="mixfdr")
  dev.off()
```

# Problems: Summary

- By summarizing each observation by a $Z$ score or $p$ value, 
standard fdr tools ignore precision of different measurements

- Standard tools make the ZA, which implies actual effects have a (probably unrealistic) bimodal distribution. [and tends to overestimate $\pi_0$, losing power]

- Also standard tools focus only on zero vs non-zero effects. (eg what if we would
like to identify genes that have at least a 2-fold change?)

# Proposed Alternative

- Instead of working with $z$ scores or $p$ values, work with two numbers 
$(\hat\beta_j,s_j)$ for each test.

- *Incorporate the precision* of the observations $\hat\beta$ into the likelihood.
Specifically, we approximate likelihood for $\beta_j$ by a normal (``Laplace") approximation: 
$$L(\beta_j) \propto \exp(-0.5 (\beta_j - \hat\beta_j)^2/s_j^2).$$
[From $\hat\beta_j \sim N(\beta_j, s_j)$]

- *Directly model the underlying distribution of $\beta$*, using a unimodal family of distributions $g$; estimate $g$ from the data.

# Proposed Alternative: More details

- A convenient way to model $g$ is by a mixture of 0-centered
normal distributions: 
$$g(\beta; \pi) = \sum_{k=1}^K \pi_k N(\beta; 0, \sigma^2_k)$$

- The estimating $g$ comes down to estimating $\pi$ - easily done
by maximum likelihood (EM algorithm) or variational Bayes.

- By allowing $K$ large, and $\sigma_k$ to span a dense grid of values,
we get a fairly flexible unimodal symmetric distribution.

- Alternatively, a mixture of uniforms, with 0 as one end-point of the range,
provides still more flexibility, and in particular allows for asymmetry. (Grenander 1953 shows this provides the non-parametric mle in the ``no error" case; Campy + Thomas do the equal-errors case)

# Illustration: $g$ a mixture of 0-centered normals

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, dnorm(x,0,1),type="l",ylim=c(0,2),ylab="density")
lines(x, dnorm(x,0,0.1))
lines(x, dnorm(x,0,0.5))
```

# Illustration: $g$ a mixture of 0-centered normals

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, 0.5*dnorm(x,0,1)+0.5*dnorm(x,0,0.1),type="l",ylim=c(0,2),ylab="density")
```

# Illustration: $g$ a mixture of 0-anchored uniforms

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, dunif(x,0,1),type="s",ylim=c(0,5),ylab="density")
lines(x, dunif(x,0,0.2),type="s")
lines(x, dunif(x,0,0.5),type="s")
lines(x, dunif(x,-0.3,0),type="s",col=2)
lines(x, dunif(x,-0.4,0),type="s",col=2)
lines(x, dunif(x,-2,0),type="s",col=2)
```

# Illustration: $g$ a mixture of 0-anchored uniforms

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, 0.1*dunif(x,0,3)+ 0.3*dunif(x,0,0.2) + 0.2*dunif(x,0,0.5) + 0.1* dunif(x,-0.3,0) + 0.1*dunif(x,-0.3,0)+0.2*dunif(x,-0.4,0),type="s",ylim=c(0,2),ylab="density")
```

# Illustration: BRCA data

```{r}
hh.ash = ash(hh.betahat,hh.sebetahat)
hh.ash.hu = ash(hh.betahat, hh.sebetahat,mixcompdist="halfuniform")
```

```{r, echo=FALSE,fig.height=4,fig.cap=""}
h=hist(hh.betahat,freq=FALSE,main="BRCA, effect sizes",ylim=c(0,2),nclass=nc)
x = seq(-3,3,length=100)
lines(x,density(hh.ash,x),col=2,lwd=2)
lines(x,density(hh.ash.hu,x),col=3,lwd=2)
```

# Illustration: BRCA data
```{r, echo=FALSE,fig.height=4,fig.cap=""}
nullalthist(hh.zscore,hh.ashz$ZeroProb,main="ash")
```

# Illustration: BRCA data
```{r, echo=FALSE,fig.height=4,fig.cap=""}
nullalthist(hh.zscore,hh.mixfdr$fdr,main="mixfdr")
```

# Issue: identifiability of $\pi_0$

- For estimating False Discoveries, we are asking whether $\beta_j = 0$.

- However, the data cannot distinguish between $\beta_j = 0$ and $\beta_j$ "very
small"

- Similarly, the data can't tell us what proportion of g should be exactly on zero, vs near zero

- Of course this is true regardless of the method used!

# Issue: identifiability of $\pi_0$

- The Zero assumption (ZA) solves the identifiability problem by assuming that
there *are* no $\beta_j$ near zero!

- The ZA makes $\pi_0$ identifiable.

- Another view is that the estimate of $\pi_0$ under ZA will systematically tend to overestimate $\pi_0$, and so is ``conservative".

- That is it provides an ``upper bound" on $\pi_0$

# Identifiability of $\pi_0$: Solution 1

- We replaced the ZA with the unimodal assumption on g.

- This does not make $\pi_0$ identifiable, but it does effectively provide an upper bound on $\pi_0$.

- Indeed, we saw that when we estimated $\pi_0$ under the ZA the data
then contradicted the unimodal assumption on g. Thus the upper bound is
more conservative than under ZA.

- In practice, implement upper bound by putting prior on $\pi_0$ that encourages it to be big, then estimate $\pi$ by posterior mean (VB).

# Identifiability of $\pi_0$: Solution 2

- Identifiability of $\pi_0$ is primarily an issue if we insist on asking question is $\beta_j=0$?

- How about we change focus: assume *none* of the $\beta_j$ are zero ("one group approach"), and ask for which $\beta_j$ are we confident about the sign (Gelman et al, 2012).

- Positive and negative effects are often treated differently in practice anyway.

- That is we replace fdr with False Sign Rate (fsr), the probability that if we say an effect is positive (negative), it is not.

- Example: suppose we estimate that $\Pr(\beta_j<0)=0.975$ and $\Pr(\beta_j>0)=0.025$. Then we report $\beta_j$ as a ``(negative) discovery", and estimate its fsr as 0.025.

# A technicality

- Suppose you estimate $\Pr(\beta_j<0)=0.98$,  $\Pr(\beta_j>0)=0.01$, $\Pr(\beta_j=0) = 0.01$.

- Should you declare an fdr of 0.01 or 0.02?

- Maybe fsr makes more sense anyway?

# Unification of one-group and two-group solutions

- A possible concern with one-group model: if we assume all $\beta$ are non-zero, but really some are actually zero, might we not underestimate the fsr?

- Consider our example, with $\Pr(\beta_j>0)=0.025$. If we actually allowed
$\beta_j=0$ then possibly all of this probability might actually land at $\beta_j=0$.

- I argue that, assuming symmetry of $g$ near 0, this also provides an upper
bound of how much of the $\Pr(\beta_j<0)=0.975$ might also move to 0.

- Therefore a more conservative estimate of the fsr might be 0.05 (or, more generally, double what you get allowing for point mass)

# Estimation

- This approach also provides a full posterior distribution for each $\beta_j$.

- So for example we can easily compute fdrs for discoveries other than ``non-zero" (eg genes with at least 2-fold difference between conditions

# Example: BRCA data

Compare fit of $g$ both allowing $\beta_j=0$ and not allowing it.
```{r, echo=FALSE, includ=FALSE}
  hh.ash.fdr = ash(hh.betahat,hh.sebetahat,usePointMass = TRUE,prior="nullbiased",VB=TRUE)
  hh.ash.shrink = ash(hh.betahat,hh.sebetahat)
```

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot(x,cdf.ash(hh.ash.fdr,x),col=2,type="l",lwd=3,ylab="cdf")
  lines(x,cdf.ash(hh.ash.shrink,x),col=3,lwd=3,lty=2)
```

# Example: BRCA data

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot(hh.ash.fdr$localfsr,2*hh.ash.shrink$localfsr,xlab="fsr with point mass", ylab="2 * fsr without point mass", xlim=c(0,0.1),ylim=c(0,0.1))
  abline(a=0,b=1,col=2,lwd=2)
```

- Because $\pi$ is estimated from the data, the amount
of shrinkage is adaptive to the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each gene.

# Example: ASH applied to mouse data

```{r, fig.height=4,echo=FALSE,fig.cap=""}
hist(zdat[3,],main="Raw effect size estimates",xlab="betahat",xlim=c(-2,2),prob=T,ylim=c(0,3))
```

# Example: ASH applied to mouse data

```{r, fig.height=4,echo=FALSE,fig.cap=""}
hist(zdat[3,],main="Raw effect size estimates",xlab="betahat",xlim=c(-2,2),prob=T,ylim=c(0,3))
t=seq(-2,2,length=201)
lines(t,dnorm(t,0,sd=0.032), col=2, lwd=2)
```

# Example: ASH applied to mouse data

```{r, fig.height=4,echo=FALSE,fig.cap=""}
hist(zdat.ash$PosteriorMean,main="Shrunken estimates",xlab="shrunk betahat",xlim=c(-0.1,0.1),prob=T)
lines(t,dnorm(t,0,sd=0.032), col=2, lwd=2)
```

# Shrinkage is adaptive to information

```{r, echo=FALSE,fig.height=4}
#hist(zdat[3,])
#plot(zdat[3,],zdat.ash$PosteriorMean,xlim=c(-0.2,0.2))
#points(zdat[3,16677],zdat.ash$PosteriorMean[16677],col=2,pch=16)
#points(zdat[3,16079],zdat.ash$PosteriorMean[16079],col=2,pch=16)
#x[16677,]
#x[16079,]
```

```{r, echo=FALSE}
#temp = counts.associate(cc,c(-1,1,-1,1),1)
#temp.tscore = temp$zdat[3,]/temp$zdat[4,]
#temp.pval = ttest.pval(temp.tscore,df=2)
#plot(temp$zdat.ash$localfdr,temp.pval)
#identify(temp$zdat.ash$localfdr,temp.pval)
```

```{r, echo=FALSE,fig.height=3,fig.width=3,fig.cap=""}
plot(zdat.ash$localfdr,pval,ylab="p value", xlab="ASH local fdr")
```

# Shrinkage is adaptive to information

```{r, echo=FALSE,fig.height=3,fig.width=3,fig.cap=""}
plot(zdat.ash$localfdr,pval,ylab="p value", xlab="ASH local fdr")
points(zdat.ash$localfdr[15325],pval[15325],col=2,pch=16)
points(zdat.ash$localfdr[16123],pval[16123],col=2,pch=16)
```

# Shrinkage is adaptive to information

```{r, echo=FALSE}
cbind(x[,1:5],pval,zdat.ash$localfdr)[c(15325,16123),]
```

# Compare q values for high count genes, with and without low count genes

```{r, echo=FALSE,fig.height=3,fig.width=3,fig.cap=""}
plot(zdat.ash$qvalue[xx>1000],zdat.ash.high$qvalue,xlab="q values from all data",ylab="q values from high only")
abline(a=0,b=1,col=2,lwd=2)
```

# BRCA1: Compare $\pi_0$ estimates
```{r}
  c(hh.fdrtool$param[3],
  hh.locfdr$fp0[1,3],
  hh.mixfdr$pi[1],
  hh.ashz$fitted.g$pi[1])
```

# BRCA1: Compare number significant at fdr<0.05

```{r}
  c(sum(hh.fdrtool$lfdr<0.05),
  sum(hh.locfdr$fdr<0.05),
  sum(hh.mixfdr$fdr<0.05),
  sum(hh.ashz$ZeroProb<0.05))
```

# Summary: FDR via Adapative Shrinkage

- Both provide a rational approach to identifying ``significant" findings.

- Both are generic and modular: once you have the summary data, you can forget where they came from.

- But by using two numbers ($\hat\beta,s$) instead of one ($p$ values) precision of different measurementscan be better accounted for.

- ASH borrows information for estimation, as well as testing.

# Other Applications

- Widely applicable: perhaps anywhere (?) 
where shrinkage is appropriate, requiring only an estimated
effect size and standard error for each object.

- Could also use effect size estimate and $p$ value for each variable, by converting to effect size estimate and (pseudo-) standard error.

- Currently applying it to wavelet shrinkage applications.

# Guarantees?

- ``I think you have some nice ideas. How will you convince
people to use them?" (C Morris)

# Next steps?

- Extend to allow $g(\cdot;\pi)$ to depend on covariates $X$.

- Extend to allow for correlations in the measured $\hat\beta_j$.

# Thanks

- to the several postdoctoral researchers and students
who have worked with me on related topics.

- Especially Mengyin Lu who coded the VB algorithm.

# Reproducible research

- This document is produced with **knitr**, **Rstudio** and **Pandoc**.

- For more details see my `stephens999/ash` repository at `http://www.github.com/stephens999/ash`

- Website: `http://stephenslab.uchicago.edu`

# Pandoc Command used

`pandoc -s -S -i --template=my.beamer -t beamer -V theme:CambridgeUS -V colortheme:beaver  ilike-slides.md -o ilike-slides.pdf`

(alternative to produce html slides; but figures would need reworking)
`pandoc -s -S -i -t dzslides --mathjax NSmeet2013.md -o NSmeet2013.html`

Here is my session info:

```{r session-info}
print(sessionInfo(), locale=FALSE)
```

# Some odd things in the data

```{r, echo=FALSE,fig.height=3,fig.width=3}
  plot(zdat[3,],zdat[4,],ylab="standard error",xlab="beta-hat")
  x[tail(order(zdat[3,])),]
```

### 2013-11-01 talks/ilike-slides-plus.Rmd
added work on talk for NS lab meeting

--- 2013-05-15 talks/ilike-slides.Rmd
@@
 
+```
+
+```{r}
+library("DSS")
+names(g)=colnames(cc)
+designs=c(0,0,1,1)
+names(designs)=colnames(cc)
+seqData=newSeqCountSet(as.matrix(cc), designs)
+seqData=estNormFactors(seqData)
+seqData=estDispersion(seqData)
+result=waldTest(seqData)
+#note I edited waldTest to output std
+z.dss=result$difExpr/result$std
+z.dss[result$geneIndex]=result$difExpr/result$std
 ```

### 2013-11-01 Rcode/ash.R

@@ -1,3 +1,8 @@
+#todo
+# provide a way to get estimate of pi0
+# provide a "method" that sets defaults for shrinkage vs fdr
+# think about what output should be for localfdr vs localfsr with and without point mass
+#
 #' @title Main Adaptive SHrinkage function
 #'
 #' @description Takes vectors of estimates (betahat) and their standard errors (sebetahat), and applies
@@ -19,7 +24,8 @@
 #' @param prior: string, or numeric vector indicating Dirichlet prior on mixture proportions (defaults to "uniform", or 1,1...,1; also can be "nullbiased" 1,1/k-1,...,1/k-1 to put more weight on first component)
 #' @param mixsd=NULL: vector of sds for underlying mixture components 
 #' @param VB=FALSE: whether to use Variational Bayes to estimate mixture proportions (instead of EM to find MAP estimate)
-#' 
+#' @param lambda1: multiplicative "inflation factor" for standard errors (like Genomic Control)
+#' @param lambda2: additive "inflation factor" for standard errors (like Genomic Control)
 #' 
 #' @return a list with elements fitted.g is fitted mixture
 #' logLR : logP(D|mle(pi)) - logP(D|null)
@@ -38,7 +44,7 @@
 #Things to do:
 # check sampling routine
 # check number of iterations
-ash = function(betahat,sebetahat,mixcompdist = "normal",nullcheck=TRUE,df=NULL,randomstart=FALSE, usePointMass = FALSE, onlylogLR = FALSE, localfdr = TRUE, prior="uniform", mixsd=NULL, VB=FALSE){
+ash = function(betahat,sebetahat,mixcompdist = "normal",lambda1=1,lambda2=0,nullcheck=TRUE,df=NULL,randomstart=FALSE, usePointMass = FALSE, onlylogLR = FALSE, localfdr = TRUE, prior="uniform", mixsd=NULL, VB=FALSE){
   
   #if df specified, convert betahat so that bethata/sebetahat gives the same p value
   #from a z test as the original effects would give under a t test with df=df
@@ -91,7 +97,7 @@ ash = function(betahat,sebetahat,mixcompdist = "normal",nullcheck=TRUE,df=NULL,r
[…]

### 2013-11-07 paper/main.tex
add some R code for paper, and update some of the options

@@
 
+\def\lfdr{\text{lfdr}}
+\def\lfsr{\text{lfsr}}
 \def\x{\mbf{x}}
@@
 thought of as being independent and identically distributed from some unknown distribution $g(\beta)$. 
-* these effects are symmetrically distributed about 0, with density monotonically decaying as we move away from 0.
+* these effects are symmetric and unimodal about 0.
 
-Specifically we will assume a parameteric form for $g$
+Specifically we will assume a parametric form for $g$
 as a mixture of 0-centered normal distributions
@@
 
-\subsection{The False Sign Rate}
+\subsection{The local False Sign Rate}
 
-The usual definition of the local False Discovery Rate (lfdr) for observation $j$ is
-$\text{lfdr}_j = p(\beta_j = 0 | \hat\beta, s)$.
-The lfdr terminology comes from using "discovery" to refer to rejecting the null ($H_j:\beta_j=0$), so lfdr 
-gives the probability that we make a mistake if we reject
-$H_j$, that is the probability that, if we reject $H_j$,
+The local False Discovery Rate (lfdr) for observation $j$ is
+\begin{equation}
+\lfdr_j = p(\beta_j = 0 | \hat\beta, s).
+\end{equation}
+The lfdr terminology comes from using "discovery" to refer to rejecting the null ($H_j:\beta_j=0$).
+Specifically $\lfdr_j$ is the probability that, if we reject $H_j$,
 it is a  "false discovery".
@@
 a useful concept: if every $\beta_j$ is non-zero then there
-is no such thing as a false discovery and the lfdr will be 0. Gelman et al suggest that in such settings we
+is no such thing as a false discovery and the lfdr will be identically 0 for all methods. 
+Gelman et al suggest that in such settings we
 might replace the concept of a false discovery with the
@@
 fundamentally more interesting and useful than identifying which are non-zero.
-However, in such settings this usually comes down to something similar in practice: the observations that one can tell are non-zero are also the ones whose
-sign can be reliably inferred. In this sense the
-local false sign rate is a natural generalization of
-the local false discovery rate.
 
-Following these ideas of \cite{gelman.etal}, we define the local False Sign Rate (lfsr) as 
-$$\text{lfsr}_j = \min[ p(\beta_j \geq 0| \hat\beta, s), p(\beta_j \leq 0| \hat\beta, s) ].$$
-To give the intuition behind this, suppose for concreteness
-that the minimum is achieved by the first term, $p(\beta_j \geq 0| \hat\beta, s)=0.05$ say. Then, if we were to declare that $\beta_j$ is negative, the probability that we have
+Following these ideas of \cite{gelman.etal}, we define the local False Sign Rate (lfsr) for observation $j$ as 
+\begin{equation}
+\lfsr_j := \min[ p(\beta_j \geq 0| \hat\beta, s), p(\beta_j \leq 0| \hat\beta, s) ].
+\end{equation}
+Thus $\lfsr_j$ gives the probability that we would get the sign of $\beta_j$ wrong if we
+were to make our best guess. (Note that we count it as an error to state that $\beta$ is positive or negative when it is truly zero.)
+To illustrate, suppose for concreteness
+that the minimum is achieved by the first term, $p(\beta_j \geq 0| \hat\beta, s)=0.05$ say. Then
+our best guess would be that $\beta$ is negative, and the probability that we have
 made an error in sign would be 0.05.
@@
 and $\beta_j \leq 0$ include the event $\beta_j=0$.
-Thus, lfsr gives an upper bound for the lfdr,
-and so can be used as a conservative estimate of the lfdr
-if an lfdr is desired. (This may be helpful when comparing
-with methods that compute an lfdr.)
-In addition, we intuitively expect the difference $\text{lfsr}_j - \text{lfdr}_j$ to be small, at least when $\text{lfdr}_j$ is small, because when we are confident that a particular effect is non-zero we would expect to be similarly confident about the sign of the effect.
+Thus, lfsr gives an upper bound for lfdr.
 

### 2013-11-07 talks/NSmeet2013.Rmd
add some R code for paper, and update some of the options

@@
 
-# Example: BRCA1 vs BRCA2 expression
+# FDR, local fdr, and q values
+
+Although precise definitions vary depending on whether one
+takes a Bayesian or Frequentist approach to the problem, roughly
+
+- The FDR at a threshold $P$ is 
+$$\text{FDR}(P)=\Pr(\beta_j = 0 |  p_j<P).$$
+
+- The q value for observation $j$ is $q_j=\text{FDR}(p_j)$.
+
+- The local false discovery rate, fdr, at threshold $P$ is 
+$$\text{fdr}(P) = \Pr(\beta_j =0 | p_j=P ).$$
+
+- The fdr is more relevant, but slightly harder to estimate than 
+FDR because it involves density estimation rather than tail-area estimation.
+
+# Example: FDR estimation
 
@@
 
-```{r, echo=FALSE,fig.height=3.5,fig.width=5,fig.cap=""}
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
   nc=40 #number of bins in histogram
@@
 
-# Example: BRCA1 vs BRCA2 expression
+# Example:  FDR estimation
 
-```{r, echo=FALSE,fig.height=3.5,fig.width=5,fig.cap=""}
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
   nc=40 #number of bins in histogram
@@
   plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)
-
 ```
 
-# Example: BRCA1 vs BRCA2 expression
+Data from Hedenfalk et al, comparing BRCA1 vs BRCA2 expression.
 
-```{r, echo=FALSE,fig.height=3.5,fig.width=5,fig.cap=""}
+# Example:  FDR estimation
+
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
   nc=40 #number of bins in histogram
@@
 
-# Example 2: Mouse Heart Data
+# Example: fdr estimation
+
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+  nc=40 #number of bins in histogram
+  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)
+
+```
+
+# Example: fdr estimation
+
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+  nc=40 #number of bins in histogram
+  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)
+
+  require(fdrtool)
+  hh.gren = grenander(ecdf(hh.pval))
+  abline(h=min(hh.gren$f.knots),col=nullcol,lwd=2)  
+  lines(hh.gren$x.knots,hh.gren$f.knots,lwd=2)
+```
+
+# Example: fdr estimation
+
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+  nc=40 #number of bins in histogram
+  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)
+
+  require(fdrtool)
+  hh.gren = grenander(ecdf(hh.pval))
+  abline(h=min(hh.gren$f.knots),col=nullcol,lwd=2)  
+  lines(hh.gren$x.knots,hh.gren$f.knots,lwd=2)
+  abline(v=0.1,lwd=2,col=2)
+  text(0.1,0.9,labels="a",col=2,cex=1)  
+  text(0.1,0.34,labels="b",col=2,cex=1.2)  
+  text(0.6,3,labels=paste0("fdr = b/(a+b) =  ",round(min(hh.gren$f.knots)/approx(hh.gren$x.knots,hh.gren$f.knots,0.1)$y,2)),cex=1.2)
+```
+
+# FDR problem 1: different measurement precision
+
+- If some effects are measured very imprecisely, those tests ``lack power"
+and simply add noise
+
+- In particular, such tests increase the estimated number of nulls, and increase
+the FDR for other tests
+
+- It would seem preferable to simply ignore the tests with very low precision. Summarizing each test by a $p$ value (or $Z$ score) loses the information about precision.
+
+# Example: Mouse Heart Data
 
@@
 
-# Example 2: Mouse Heart Data
+# Example: Mouse Heart Data
 
@@
 
-# FDR problem 1: different genes have different precision/power
+# Mouse Data: Counts vary considerably across genes
 
 ```{r, echo=FALSE,fig.height=4,fig.cap=""}
-hist(log10(xx),main="Counts vary considerably across genes", xlab="log10(counts)")
+hist(log10(xx),main="Distribution of total counts", xlab="log10(counts)")
 ```
 
-# FDR problem 1: lower count genes, less power, add noise
+# Lower count genes, less power
 
@@
 
-# FDR problem 1: higher count genes, more power
+# Higher count genes, more power
 ```{r, echo=FALSE, fig.height=4,fig.cap=""}
@@
 abline(v=0.05,lwd=3)
-text(0.6,3,labels=paste0("FDR = ",round(max(qval.high$qval[pval.high<0.05]),2)),cex=1.2)
+#text(0.6,3,labels=paste0("FDR = ",round(max(qval.high$qval[pval.high<0.05]),2)),cex=1.2)
 #qval.high$q[pval.high<0.05] gives FDR for p<0.05
@@
 
-# FDR problem 1: q values increased by low count genes
+# FDR problem 1: low count genes add noise, increase q values
 ```{r,echo=FALSE, fig.height=4,fig.cap=""}
-plot(qval$qval[xx>1000],qval.high$qval,xlab="q values from all gene analysis", ylab="q values from high-count gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
+plot(qval.high$qval,qval$qval[xx>1000],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
 abline(a=0,b=1,col=2)
 ```
+
+# FDR problem 1: Summary
+
+- Analyzing $p$ values or $Z$ scores doesn't fully account 
+for measurement precision.
 
@@
 
-- This allows us to estimate $\pi_0$ ``conservatively" using the density of $p$ values near 1.
+- The ZA allows us to estimate the null proportion, $\pi_0$, using the density of $p$ values near 1 (or $Z$ scores near 0).
 
@@
 
-# Proposed Alternative
+# FDR via Empirical Bayes
 
-- Instead of working with $z$ scores or $p$ values, work with two numbers 
-$(\hat\beta_j,s_j)$ for each test.
+- Following previous work (e.g. Newton, Efron, Muralidharan) we take an empirical Bayes approach to FDR.
 
-- *Incorporate the precision* of the observations $\hat\beta$ into the likelihood.
-Specifically, we approximate likelihood for $\beta_j$ by a normal (``Laplace") approximation: 
+- Eg Efron assumes that the $Z$ scores come from a mixture of null, and alternative:
+$$Z_j \sim f_Z(.) = \pi_0 N(.;0,1) + (1-\pi_0) f_1(.)$$
+where $f_1$ is to be estimated from the data.
+
+- Various semi-parametric approaches taken to estimating $f_1$. For example,
+Efron uses Poisson regression; Muralidharan uses mixture of normal distributions.
+
+- $\text{fdr}(Z) \approx \pi_0 N(Z; 0,1)/ f_Z(Z)$
+
+# FDR: The New Deal
+
+- Instead of modelling $Z$ scores, model the effects $\beta$,
+$$\beta_j \sim \pi_0 \delta_0(.) + (1-\pi_0) g(.)$$
+
+- Constrain $g$ to be unimodal about 0.
+
+- *Incorporate precision* of each observation $\hat\beta$ into the likelihood.
+Specifically, approximate likelihood for $\beta_j$ by a normal: 
 $$L(\beta_j) \propto \exp(-0.5 (\beta_j - \hat\beta_j)^2/s_j^2).$$
@@
 
-- *Directly model the underlying distribution of $\beta$*, using a unimodal family of distributions $g$; estimate $g$ from the data.
+- fdr given by $$p(\beta_j =0 | \hat\beta_j) = \pi_0 p(\hat\beta_j | \beta_j=0)/p(\hat\beta_j)$$
 
-# Proposed Alternative: More details
+# FDR - The New Deal
 
@@
 
-- The estimating $g$ comes down to estimating $\pi$ - easily done
-by maximum likelihood (EM algorithm) or variational Bayes.
+- Estimating $g$ comes down to estimating $\pi$. Joint estimation of $\pi_0,\pi$ easy by maximum likelihood (EM algorithm) or variational Bayes.
 
@@
 - Alternatively, a mixture of uniforms, with 0 as one end-point of the range,
-provides still more flexibility, and in particular allows for asymmetry. (Grenander 1953 shows this provides the non-parametric mle in the ``no error" case; Campy + Thomas do the equal-errors case)
+provides still more flexibility, and in particular allows for asymmetry.
+
+- If allow a very large number of uniforms this provides the non-parametric mle for $g$; cf Grenander 1953; Campy + Thomas.
 
@@
 
-# Illustration: BRCA data
-
-```{r}
-hh.ash = ash(hh.betahat,hh.sebetahat)
-hh.ash.hu = ash(hh.betahat, hh.sebetahat,mixcompdist="halfuniform")
-```
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-h=hist(hh.betahat,freq=FALSE,main="BRCA, effect sizes",ylim=c(0,2),nclass=nc)
-x = seq(-3,3,length=100)
-lines(x,density(hh.ash,x),col=2,lwd=2)
-lines(x,density(hh.ash.hu,x),col=3,lwd=2)
-```
-
-# Illustration: BRCA data
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-nullalthist(hh.zscore,hh.ashz$ZeroProb,main="ash")
-```
-
-# Illustration: BRCA data
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-nullalthist(hh.zscore,hh.mixfdr$fdr,main="mixfdr")
-```
-
 # Issue: identifiability of $\pi_0$
@@
 
-- Similarly, the data can't tell us what proportion of g should be exactly on zero, vs near zero
-
-- Of course this is true regardless of the method used!
+- As a result $\pi_0$ is formally unidentifiable. Eg data can never rule out $\pi_0=0$.
 
@@
 - In practice, implement upper bound by putting prior on $\pi_0$ that encourages it to be big, then estimate $\pi$ by posterior mean (VB).
+
+# Illustration: BRCA data
+
+```{r, echo=FALSE,include=FALSE}
+hh.ash = ash(hh.betahat,hh.sebetahat)
+hh.ashz = ash(hh.zscore,1, pointmass=TRUE, prior="nullbiased")
+hh.ash.hu = ash(hh.betahat, hh.sebetahat,mixcompdist="halfuniform")
+```
+
+# Example: BRCA data
+
+Compare fitted $f(\beta)$, both estimating $\pi_0$ and fixing $\pi_0=0$.
+```{r, echo=FALSE, includ=FALSE}
+  hh.ash.fdr = ash(hh.zscore,1,pointmass = TRUE,prior="nullbiased")
+  hh.ash.shrink = ash(hh.zscore,1)
+```
+
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+  plot(x,cdf.ash(hh.ash.fdr,x),col=2,type="l",lwd=3,ylab="cdf")
+  lines(x,cdf.ash(hh.ash.shrink,x),col=3,lwd=3,lty=2)
+```
+
+# Recall Problem: distribution of alternative Z values multimodal
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+nullalthist(hh.zscore,hh.mixfdr$fdr,main="mixfdr")
+```
+
+# Problem Fixed: distribution of alternative Z values unimodal
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+nullalthist(hh.zscore,hh.ashz$ZeroProb,main="ash")
+```
+
+# BRCA1: Compare $\pi_0$ estimates
+```{r}
+  round(c(hh.fdrtool$param[3],
+  hh.locfdr$fp0[1,3],
+  hh.mixfdr$pi[1],
+  hh.ashz$fitted.g$pi[1]),2)
+```
+
+# BRCA1: Compare number significant at fdr<0.05
+
+```{r}
+  c(sum(hh.fdrtool$lfdr<0.05),
+  sum(hh.locfdr$fdr<0.05),
+  sum(hh.mixfdr$fdr<0.05),
+  sum(hh.ashz$ZeroProb<0.05))
+```
 
@@
 
-# Estimation
-
-- This approach also provides a full posterior distribution for each $\beta_j$.
-
-- So for example we can easily compute fdrs for discoveries other than ``non-zero" (eg genes with at least 2-fold difference between conditions
-
-# Example: BRCA data
-
-Compare fit of $g$ both allowing $\beta_j=0$ and not allowing it.
-```{r, echo=FALSE, includ=FALSE}
-  hh.ash.fdr = ash(hh.betahat,hh.sebetahat,usePointMass = TRUE,prior="nullbiased",VB=TRUE)
-  hh.ash.shrink = ash(hh.betahat,hh.sebetahat)
-```
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot(x,cdf.ash(hh.ash.fdr,x),col=2,type="l",lwd=3,ylab="cdf")
-  lines(x,cdf.ash(hh.ash.shrink,x),col=3,lwd=3,lty=2)
-```
-
 # Example: BRCA data
@@
 
-- Because $\pi$ is estimated from the data, the amount
+# Estimation and Shrinkage
+
+- Besides allowing one to estimate fdr and fsr, 
+this approach also provides a full posterior distribution for each $\beta_j$.
+
+- So for example we can easily compute fdrs for discoveries other than ``non-zero" (eg compute $\Pr(|\beta_j| > 2 \| \hat\beta_j)$).
+
+- And use it to obtain point estimates and credible intervals for each $\beta_j$, taking account of information from all the other $\beta_j$.
+
+- Because $f(\beta)$ is unimodal, the point estimates will tend to be ``shrunk" towards the overall mean (0).
+
+- Because $f(\beta)$ is estimated from the data, the amount
 of shrinkage is adaptive to the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each gene.
+
+- So we call the approach ``Adaptive Shrinkage" (ASH).
 
@@
 
-```{r, echo=FALSE,fig.height=3,fig.width=3,fig.cap=""}
+```{r, echo=FALSE,fig.height=3,fig.cap=""}
 plot(zdat.ash$localfdr,pval,ylab="p value", xlab="ASH local fdr")
@@
 ```{r, echo=FALSE}
+x = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
+xx = rowSums(x[,2:5])
+x = x[xx>0,]
+xx = xx[xx>0]
 cbind(x[,1:5],pval,zdat.ash$localfdr)[c(15325,16123),]
@@
 
-# Compare q values for high count genes, with and without low count genes
-
-```{r, echo=FALSE,fig.height=3,fig.width=3,fig.cap=""}
-plot(zdat.ash$qvalue[xx>1000],zdat.ash.high$qvalue,xlab="q values from all data",ylab="q values from high only")
-abline(a=0,b=1,col=2,lwd=2)
+# Recall FDR problem 1: q values increased by low count genes
+```{r,echo=FALSE, fig.height=4,fig.cap=""}
+plot(qval.high$qval,qval$qval[xx>1000],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
+abline(a=0,b=1,col=2)
 ```
 
-# BRCA1: Compare $\pi_0$ estimates
-```{r}
-  c(hh.fdrtool$param[3],
-  hh.locfdr$fp0[1,3],
-  hh.mixfdr$pi[1],
-  hh.ashz$fitted.g$pi[1])
-```
+# ASH q values more robust to inclusion of low count genes
 
-# BRCA1: Compare number significant at fdr<0.05
-
-```{r}
-  c(sum(hh.fdrtool$lfdr<0.05),
-  sum(hh.locfdr$fdr<0.05),
-  sum(hh.mixfdr$fdr<0.05),
-  sum(hh.ashz$ZeroProb<0.05))
+```{r, echo=FALSE,fig.height=3,fig.cap=""}
+plot(zdat.ash.high$qvalue,zdat.ash$qvalue[xx>1000],xlab="q values from high only",ylab="q values from all data")
+abline(a=0,b=1,col=2,lwd=2)
 ```
@@
 
-```{r, echo=FALSE,fig.height=3,fig.width=3}
+```{r, echo=FALSE,fig.height=3}
   plot(zdat[3,],zdat[4,],ylab="standard error",xlab="beta-hat")

### 2013-11-07 Rcode/ash.R

@@ -1,7 +1,4 @@
 #todo
-# provide a way to get estimate of pi0
-# provide a "method" that sets defaults for shrinkage vs fdr
-# think about what output should be for localfdr vs localfsr with and without point mass
 #
 #' @title Main Adaptive SHrinkage function
 #'
@@ -12,20 +9,23 @@
 #' 
 #' @param betahat, a p vector of estimates 
 #' @param sebetahat, a p vector of corresponding standard errors
+#' @param method: specifies how ash is to be run. Can be "shrinkage" (if main aim is shrinkage) or "fdr" (if main aim is to assess fdr or fsr)
+#' This is simply a convenient way to specify certain combinations of parameters: "shrinkage" sets pointmass=FALSE and prior="uniform";
+#' "fdr" sets pointmass=TRUE and prior="nullbiased".
 #' @param mixcompdist: distribution of components in mixture ("normal", "uniform" or "halfuniform")
-#' @param nullcheck: whether to check that any fitted model exceeds the "null" likelihood
-#' in which all weight is on the first component
-
-#' @param df: appropriate degrees of freedom for (t) distribution of betahat/sebetahat
-#' @param randomstart: bool, indicating whether to initialize EM randomly. If FALSE, then initializes to prior mean (for EM algorithm) or prior (for VBEM)
-#' @param usePointMass: bool, indicating whether to use a point mass at zero as one of components for a mixture distribution
-#' @param onlylogLR (= FALSE) : bool, indicating whether to use this function to get logLR. Skip posterior prob, posterior mean, localfdr...
-#' @param localfdr (=TRUE) : bool,  indicating whether to compute localfdr, localfsr, and q-value
-#' @param prior: string, or numeric vector indicating Dirichlet prior on mixture proportions (defaults to "uniform", or 1,1...,1; also can be "nullbiased" 1,1/k-1,...,1/k-1 to put more weight on first component)
-#' @param mixsd=NULL: vector of sds for underlying mixture components 
-#' @param VB=FALSE: whether to use Variational Bayes to estimate mixture proportions (instead of EM to find MAP estimate)
+#'
[…]

### 2013-11-10 Rcode/conservativepm.Rmd
updated conservative.pm

#look into methods for estimating pi0 conservatively

Do some simulations

```{r}
basicsim=function(mixsd,mixpi_alt,seedval = 100,nsamp=1000,niter=50){
  require(ashr)
  require(qvalue)
  require(fdrtool)
  require(mixfdr)
  require(locfdr)
  set.seed(seedval)  
  beta =list()
  betahatsd=list()
  betahat = list()
  zscore = list()
  pval = list()
  betahat.ash.n = list()
  betahat.ash.u = list()
  betahat.ash.npm = list()
  betahat.qval = list()
  betahat.fdrtool = list()
  betahat.locfdr = list()
  betahat.mixfdr = list()
  pi0 = rep(0,niter)
  for(i in 1:niter){
    pi0[i]=runif(1,0,1)
    mixpi = c(pi0[i],(1-pi0[i])*mixpi_alt)
    sd = sample(mixsd,nsamp,prob=mixpi,replace=TRUE )
    beta[[i]] = rnorm(nsamp,0,sd)
    betahatsd[[i]] = 1
    betahat[[i]] = beta[[i]]+rnorm(nsamp,0,betahatsd[[i]])
    zscore[[i]] = betahat[[i]]/betahatsd[[i]]
    pval[[i]] = pchisq(zscore[[i]]^2,df=1,lower.tail=F)
    betahat.ash.n[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=TRUE,prior="nullbiased",gridmult=2)
    betahat.ash.u[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=TRUE,prior="uniform",gridmult=2)
    betahat.ash.npm[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=FALSE,prior="uniform",gridmult=2)
    betahat.qval[[i]] = qvalue(pval[[i]])
    betahat.fdrtool[[i]] = fdrtool(pval[[i]],statistic="pvalue",plot=FALSE)
    betahat.locfdr[[i]] = locfdr(zscore[[i]],nulltype=0,plot=0)
    betahat.mixfdr[[i]] = mixFdr(zscore[[i]],noiseSD=1,theonull=TRUE,plot=FALSE)
  }
  return(list(beta =beta,
  betahatsd=betahatsd,
  betahat = betahat,
  zscore = zscore,
  pval = pval,
  betahat.ash.n = betahat.ash.n,
  betahat.ash.u = betahat.ash.u,
  betahat.ash.npm = betahat.ash.npm,
  betahat.qval = betahat.qval,
  betahat.fdrtool = betahat.fdrtool,
  betahat.locfdr = betahat.locfdr,
  betahat.mixfdr = betahat.mixfdr,
    pi0=pi0))
}    
mixsd = c(0,0.25,0.5,1,2)
mixpi_alt = c(0.4,0.2,0.2,0.2) #mixture proportions under the alternative
simres1 = basicsim(mixsd,mixpi_alt)
simres2= basicsim(c(0,4),c(1))
```

Show illustrative example
```{r}
plot_FDR_hist=function(sim,iter=1){
  hh.pval = sim$pval[[iter]]
  hh.zscore = sim$zscore[[iter]]
  altcol="cyan" #colors to use
  nullcol="blue" 
  nc=40 #number of bins in histograms
  ncz = 100 # number of bins in z score histograms

  hh.hist =hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

  hh.q = qvalue(hh.pval)
  abline(h=hh.q$pi0,col=nullcol,lwd=2)

  hh.hist$density=rep(hh.q$pi0,length(hh.hist$density))  
  plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)

  abline(v=0.1,lwd=2,col=2)

  text(0.05,1.2,labels="A",col=2,cex=1.2)  
  text(0.05,0.4,labels="B",col=2,cex=1.2)  
  text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(hh.q$pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
}
plot_FDR_hist(simres1,1)

```

```{r}
plot_lfdr_hist=function(sim,iter=1){
  require(fdrtool)
  hh.pval=sim$pval[[iter]]
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)

  hh.gren = grenander(ecdf(hh.pval))
  abline(h=min(hh.gren$f.knots),col=nullcol,lwd=2)  
  lines(hh.gren$x.knots,hh.gren$f.knots,lwd=2)
  abline(v=0.1,lwd=2,col=2)
  text(0.1,0.9,labels="a",col=2,cex=1)  
  text(0.1,0.34,labels="b",col=2,cex=1.2)  
  text(0.6,3,labels=paste0("lfdr = b/(a+b) =  ",round(min(hh.gren$f.knots)/approx(hh.gren$x.knots,hh.gren$f.knots,0.1)$y,2)),cex=1.2)
}
plot_lfdr_hist(simres1,1)
```

```{r}
#plot a histogram of z scores, highlighting the alternative distribution
#of z scores that is implied by localfdr values lfdr.
  nullalthist = function(z,lfdr,...){
    h=hist(z, freq=FALSE,col=nullcol,nclass=ncz,...)
    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
    h$density = (1-avlfdr) * h$density
    plot(h,add=TRUE,col=altcol,freq=FALSE)
  }

#this one puts the null on the bottom
  altnullhist = function(z,lfdr,...){
    h=hist(z, freq=FALSE,col=altcol,nclass=ncz,...)
    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
    h$density = avlfdr * h$density
    plot(h,add=TRUE,col=nullcol,freq=FALSE)
  }

  plotall_hist=function(sim,iter=1,histfun=nullalthist){
    hh.zscore=sim$zscore[[iter]]    
    par(mfcol=c(2,2))
    histfun(hh.zscore,sim$betahat.fdrtool[[iter]]$lfdr,main="fdrtool")  
    histfun(hh.zscore,sim$betahat.locfdr[[iter]]$fdr,main="locfdr")
    histfun(hh.zscore,sim$betahat.mixfdr[[iter]]$fdr,main="mixfdr")
    histfun(hh.zscore,sim$betahat.ash.n[[iter]]$lfdr,main="ash")
    par(mfcol=c(1,1))
  }

 # pdf("figures/nullalthist.pdf")
  plotall_hist(simres1,1,nullalthist)
#  dev.off()

 # pdf("figures/altnullhist.pdf")
  plotall_hist(simres1,1,altnullhist)
#  dev.off()

```

```{r}
#  par(mfcol=c(3,3))
  plot_ecdf=function(sims){
    for(i in 1:length(sims$beta)){
       plot(ecdf(sims$beta[[i]]),xlim=c(-6,6),main=paste0("iteration ",i))
       x = seq(-6,6,length=1000)
       lines(cdf.ash(sims$betahat.ash.n[[i]],x),col=2,lwd=2)
       lines(cdf.ash(sims$betahat.ash.u[[i]],x),col=3,lwd=2)
    }
  }
  plot_ecdf(simres1)
  plot_ecdf(simres2)
```

```{r}
#Plot pi0 from each method
  get_pi0.fdrtool = function(f){f$param[3]}
  get_pi0.locfdr = function(f){f$fp0[1,3]}
  get_pi0.mixfdr = function(f){f$pi[1]}

  plot_pi0 = function(sims){
    pi0=sims$pi0  
    pi0_ash.n=unlist(lapply(sims$betahat.ash.n,get_pi0))
    pi0_ash.u=unlist(lapply(sims$betahat.ash.u,get_pi0)) 
    pi0_fdrtool = unlist(lapply(sims$betahat.fdrtool,get_pi0.fdrtool))
    pi0_locfdr=unlist(lapply(sims$betahat.locfdr,get_pi0.locfdr))
    pi0_mixfdr = unlist(lapply(sims$betahat.mixfdr,get_pi0.mixfdr))
    pi0_qval = unlist(lapply(sims$betahat.qval,"[[","pi0"))

    res = data.frame(pi0=pi0,qvalue=pi0_qval,mixfdr=pi0_mixfdr, locfdr=pi0_locfdr, fdrtool=pi0_fdrtool,ash.nullbiased=pi0_ash.n,ash.uniform=pi0_ash.u)
    require(reshape2)
    res.melt = melt(res, id.vars=c("pi0"),variable.name="Method")
    p=ggplot(data=res.melt,aes(pi0,value,colour=Method)) +geom_point(shape=16) +
        geom_abline(colour = "black") +
        xlab("True pi0") +
        ylab("Estimated pi0")
    print(p +scale_y_continuous(limits=c(0,1)) +
          scale_x_continuous(limits=c(0,1)) +
          coord_equal(ratio=1))

  }

  pdf("figures/estpi0_sim1.pdf")
    plot_pi0(simres1)
  dev.off()
  pdf("figures/estpi0_sim2.pdf")
    plot_pi0(simres2)
  dev.off()

```

Figure to show that estimated betahats are not so different
```{r}
  plot(betahat.ash.u[[1]]$PosteriorMean,betahat.ash.n[[1]]$PosteriorMean)
  abline(a=0,b=1,lwd=2,col=2)
  plot(betahat.ash.u[[1]]$PosteriorSD,betahat.ash.n[[1]]$PosteriorSD)
  abline(a=0,b=1,lwd=2,col=2)
```

```{r}
  rmse = function(x,y){sqrt(mean((x-y)^2))}
  get_rmse.ash = function(a,b){rmse(a$PosteriorMean,b)}
  get_rmse.mixfdr = function(a,b){rmse(a$effectSize,b)}
  plot_rmse = function(sims,inczero=FALSE,incbetahat=FALSE){
    err.ash.n= mapply(get_rmse.ash,sims$betahat.ash.n,sims$beta)
    err.ash.u = mapply(get_rmse.ash,sims$betahat.ash.u,sims$beta)
    err.mixfdr=mapply(get_rmse.mixfdr,sims$betahat.mixfdr,sims$beta)
    err.betahat = mapply(rmse,sims$betahat,sims$beta)
    err.zero = unlist(lapply(sims$beta,rmse,y=0))

     res = data.frame(mixfdr=err.mixfdr,ash.nullbiased=err.ash.n,ash.uniform=err.ash.u)
    if(inczero){
      res=data.frame(res,zero=err.zero)
    }
    if(incbetahat){
      res=data.frame(res,betahat=err.betahat)
    }
    require(reshape2)
    res.melt = melt(res, id.vars=c("ash.uniform"),variable.name="Method")

    p=ggplot(data=res.melt,aes(ash.uniform,value,colour=Method)) +geom_point(shape=16) +
        geom_abline(colour = "black") +
        xlab("RMSE (ash.uniform)") +
        ylab("RMSE (other method)")
    print(p +scale_y_continuous(limits=c(0,max(res))) +
          scale_x_continuous(limits=c(0,max(res))) +
          coord_equal(ratio=1))
  }
  pdf("figures/rmse_sim1.pdf")
    plot_rmse(simres1)
  dev.off()
  pdf("figures/rmse_sim2.pdf")
    plot_rmse(simres2)
  dev.off()
```

```{r}
  plot_LR=function(sims){
    hist(unlist(lapply(sims$betahat.ash.u,get_loglik))-unlist(lapply(sims$betahat.ash.n,get_loglik)), xlab="loglik difference", main="loglik differences for nullbiased prior vs mle",nclass=10)
  }

  pdf("figures/logLR.pdf")
    plot_LR(simres1)
    plot_LR(simres2)
  dev.off()
```

```{r}
  plot(betahat.ash.n[[1]]$lfdr,betahat.ash.u[[1]]$lfdr,xlab="null-biased prior",ylab="uniform prior", main="fdr is not robust to estimated pi0",xlim=c(0,0.2),ylim=c(0,0.2))
  abline(a=0,b=1,col=2,lwd=2)
```

```{r}
  plot(betahat.ash.n[[1]]$lfsr,betahat.ash.u[[1]]$lfsr,xlab="null-biased prior",ylab="uniform prior", main="fsr is more robust to estimated pi0",xlim=c(0,0.2),ylim=c(0,0.2))
  points(betahat.ash.n[[1]]$lfsr,betahat.ash.npm[[1]]$lfsr,col=3)
  abline(a=0,b=1,col=2,lwd=2)
```

```{r}
  plot(betahat.ash.n[[1]]$lfsra,betahat.ash.u[[1]]$lfsra,xlab="null-biased prior",ylab="uniform prior", main="fsra is still more robust to estimated pi0",xlim=c(0,0.2),ylim=c(0,0.2))
  points(betahat.ash.n[[1]]$lfsra,betahat.ash.npm[[1]]$lfsra,col=3)
  abline(a=0,b=1,col=2,lwd=2)
```

```{r}
  plot(betahat.ash.n[[1]]$lfsr,2*betahat.ash.npm[[1]]$lfsr,xlab="null-biased prior",ylab="uniform prior", main="fsra is still more robust to estimated pi0",xlim=c(0,0.2),ylim=c(0,0.2))
  points(betahat.ash.n[[1]]$lfsra,betahat.ash.npm[[1]]$lfsra,col=3)
  abline(a=0,b=1,col=2,lwd=2)
```

QUestion: is the null-biased prior maybe a little too conservative?
Answer: log likelihoods don't suggest they are

```{r}
  hist(pval[[1]])
```

```{r}
hh.ashtrue = hh.ashz
hh.ashtrue$fitted.g$pi = c(2/3,1/15,1/15,1/15,1/15,1/15)
hh.ashtrue$fitted.g$mean = c(0,0,0,0,0,0)
hh.ashtrue$fitted.g$sd = sqrt(c(0,1,0.2,0.4,0.8,3))

loglik(hh.ashtrue,betahat,sebetahat)
loglik(hh.ashz,betahat,sebetahat)
```

### 2013-11-12 Rcode/conservativepm.Rmd
update ggplot2 figures

@@
 ```{r}
-basicsim=function(mixsd,mixpi_alt,seedval = 100,nsamp=1000,niter=50){
+opts_knit$set(progress = TRUE, verbose = TRUE,root.dir="~/Documents/git/ash/paper/Rcode")
 require(ashr)
@@
 require(locfdr)
+require(ggplot2)
+#bsd gives standard deviation of beta
+#pi0 is simulated to be uniform on [minpi0,1]
+basicsim=function(mixsd,mixpi_alt,bsd=1,minpi0=0,seedval = 100,nsamp=1000,niter=50){  
   set.seed(seedval)  
@@
   betahat.ash.npm = list()
+  betahat.ash.true = list()
   betahat.qval = list()
@@
   for(i in 1:niter){
-    pi0[i]=runif(1,0,1)
+    pi0[i]=runif(1,minpi0,1)
     mixpi = c(pi0[i],(1-pi0[i])*mixpi_alt)
@@
     beta[[i]] = rnorm(nsamp,0,sd)
-    betahatsd[[i]] = 1
+    betahatsd[[i]] = bsd
     betahat[[i]] = beta[[i]]+rnorm(nsamp,0,betahatsd[[i]])
@@
     betahat.ash.npm[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=FALSE,prior="uniform",gridmult=2)
+    betahat.ash.true[[i]] = ash(betahat[[i]],betahatsd[[i]],g=normalmix(mixpi,rep(0,length(mixpi)),mixsd))
+
     betahat.qval[[i]] = qvalue(pval[[i]])
@@
   betahat.ash.npm = betahat.ash.npm,
+  betahat.ash.true=betahat.ash.true,
   betahat.qval = betahat.qval,
@@
 }    
+```
+
+```{r, cache=TRUE}
 mixsd = c(0,0.25,0.5,1,2)
 mixpi_alt = c(0.4,0.2,0.2,0.2) #mixture proportions under the alternative
-simres1 = basicsim(mixsd,mixpi_alt)
+
+simres1 = basicsim(mixsd,mixpi_alt,niter=200,nsamp=1000)
+simres1a = basicsim(mixsd,mixpi_alt,niter=200,nsamp=10000)
 simres2= basicsim(c(0,4),c(1))
+simres3=basicsim(c(0,4),c(1),bsd=c(rep(1,500),rep(10,500)))
+simres4=basicsim(c(0,4),c(1),bsd=c(rep(1,500),rep(10,500)),minpi0=0.9,seed=200)
 ```
@@
 ```{r}
-plot_FDR_hist=function(sim,iter=1){
-  hh.pval = sim$pval[[iter]]
-  hh.zscore = sim$zscore[[iter]]
 altcol="cyan" #colors to use
@@
 
+plot_FDR_hist=function(sim,iter=1){
+  hh.pval = sim$pval[[iter]]
+  hh.zscore = sim$zscore[[iter]]
   hh.hist =hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)
@@
        lines(cdf.ash(sims$betahat.ash.u[[i]],x),col=3,lwd=2)
+       lines(cdf.ash(sims$betahat.ash.true[[i]],x),col=4,lwd=2)
     }
@@
           coord_equal(ratio=1))
+
+  }
+
+  plot_pi1 = function(sims){
+    pi0=sims$pi0  
+    pi0_ash.n=unlist(lapply(sims$betahat.ash.n,get_pi0))
+    pi0_ash.u=unlist(lapply(sims$betahat.ash.u,get_pi0)) 
+    pi0_fdrtool = unlist(lapply(sims$betahat.fdrtool,get_pi0.fdrtool))
+    pi0_locfdr=unlist(lapply(sims$betahat.locfdr,get_pi0.locfdr))
+    pi0_mixfdr = unlist(lapply(sims$betahat.mixfdr,get_pi0.mixfdr))
+    pi0_qval = unlist(lapply(sims$betahat.qval,"[[","pi0"))
+
+    res = data.frame(pi0=pi0,qvalue=pi0_qval,mixfdr=pi0_mixfdr, locfdr=pi0_locfdr, fdrtool=pi0_fdrtool,ash.nullbiased=pi0_ash.n,ash.uniform=pi0_ash.u)
+    require(reshape2)
+    res.melt = melt(res, id.vars=c("pi0"),variable.name="Method")
+    p=ggplot(data=res.melt,aes(1-pi0,log2((1-value)/(1-pi0)),colour=Method)) +geom_point(shape=16) +
+#        geom_abline(colour = "black") +
+        xlab("True pi1") +
+        ylab("log2(Estimated pi1/True pi1)")
+    print(p +scale_y_continuous(limits=c(-4,4)) +
+          scale_x_continuous(limits=c(0,1)))
 
@@
 ```{r}
-  plot(betahat.ash.n[[1]]$lfdr,betahat.ash.u[[1]]$lfdr,xlab="null-biased prior",ylab="uniform prior", main="fdr is not robust to estimated pi0",xlim=c(0,0.2),ylim=c(0,0.2))
-  abline(a=0,b=1,col=2,lwd=2)
+mean_quant=function (x, mult = 1){
+      x <- na.omit(x)
+      sd <- mult * sqrt(var(x))
+      mean <- mean(x)
+      data.frame(y = median(x), ymin = quantile(x,0.25) , ymax = quantile(x,0.75))
+}
+
+#ptype indicates what type of plot to do
+#maxlfsr controls maximum x axis value
+#maxy controls maximum y axis value
+  plot_lfsr=function(sims,maxlfsr=0.1,ptype=c("lfsr","lfsra","lfdr"),maxy=1){
+    ptype=match.arg(ptype)
+    xlabtype = ifelse(ptype=="lfdr","lfdr","lfsr")
+    res=list()
+    for(i in 1:length(sims)){
+      lfsr.ash.n= unlist(lapply(sims[[i]]$betahat.ash.n,"[[",ptype))
+      lfsr.ash.u= unlist(lapply(sims[[i]]$betahat.ash.u,"[[",ptype))
+      if(ptype=="lfdr"){
+        lfsr.ash.true= unlist(lapply(sims[[i]]$betahat.ash.true,"[[","lfdr"))
+        lfdr.mixfdr = unlist(lapply(sims[[i]]$betahat.mixfdr,"[[","fdr"))
+        lfdr.locfdr = unlist(lapply(sims[[i]]$betahat.locfdr,"[[","fdr"))
+      } else {
+         lfsr.ash.true= unlist(lapply(sims[[i]]$betahat.ash.true,"[[","lfsr"))
+      }
+
+      subset = lfsr.ash.true<maxlfsr
+
+      if(length(subset)>0){  
+        res[[i]] = data.frame(Scenario=i,ash.nullbiased=lfsr.ash.n[subset],
+                      ash.uniform=lfsr.ash.u[subset],
+                      Bayes=0.1*maxlfsr*findInterval(lfsr.ash.true[subset],seq(0,maxlfsr,length=11))-0.05*maxlfsr)
+        if(ptype=="lfdr"){
+          res[[i]]=data.frame(res[[i]],mixfdr = lfdr.mixfdr[subset])
+        }
+      }
+    }
+
+    require(reshape2)
+
+    res.melt = melt(res, id.vars=c("Bayes","Scenario"),variable.name="Method")
+
+    cbbPalette <- c("#000000","#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+    labels = c('ash (null-biased)','ash (uniform)')
+    breaks = c("ash.nullbiased","ash.uniform")
+    if(ptype=="lfdr"){
+      labels = c('mixfdr', labels)
+      breaks = c("mixfdr", breaks)
+    }
+
+    p=ggplot(data=res.melt,aes(Bayes,value,colour=Method)) + 
+        facet_grid(. ~ Scenario) +
+        #scale_fill_manual(values=cbbPalette) +
+        #scale_colour_manual(values=cbbPalette) +
+        geom_point(size=1,alpha=0.1) + 
+        stat_smooth(se=FALSE,size=2) +
+        stat_summary(fun.data="mean_quant", geom="ribbon", alpha=0.25) +
+        geom_abline(colour = "black") +
+        xlab(paste0(xlabtype," (Bayes)")) +
+        ylab(paste0(ptype," (Method)"))
+
+    print(p +scale_y_continuous(limits=c(0,maxy)) +
+          scale_x_continuous(limits=c(0,maxlfsr)) +
+            scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels))         
+  }
+
 ```
@@
 ```{r}
-  plot(betahat.ash.n[[1]]$lfsr,betahat.ash.u[[1]]$lfsr,xlab="null-biased prior",ylab="uniform prior", main="fsr is more robust to estimated pi0",xlim=c(0,0.2),ylim=c(0,0.2))
-  points(betahat.ash.n[[1]]$lfsr,betahat.ash.npm[[1]]$lfsr,col=3)
-  abline(a=0,b=1,col=2,lwd=2)
+#plots estimated lfdr and lfsr against truth.
+#ptype indicates whether to use lfsr or lfsra
+#maxx controls maximum x axis value
+#maxy controls maximum y axis value
+plot_lfsdr=function(sims,maxx=0.1,ptype=c("lfsr","lfsra"),maxy=1){
+    ptype=match.arg(ptype)
+    res=list()
+    res.lfsr=list()
+    res.lfdr=list()
+
+    for(i in 1:length(sims)){
+      lfsr.ash.true= unlist(lapply(sims[[i]]$betahat.ash.true,"[[","lfsr"))
+      lfsr.ash.n= unlist(lapply(sims[[i]]$betahat.ash.n,"[[",ptype))
+      lfsr.ash.u= unlist(lapply(sims[[i]]$betahat.ash.u,"[[",ptype))
+
+      lfdr.ash.n= unlist(lapply(sims[[i]]$betahat.ash.n,"[[","lfdr"))
+      lfdr.ash.u= unlist(lapply(sims[[i]]$betahat.ash.u,"[[","lfdr"))
+      lfdr.ash.true= unlist(lapply(sims[[i]]$betahat.ash.true,"[[","lfdr"))
+      lfdr.mixfdr = unlist(lapply(sims[[i]]$betahat.mixfdr,"[[","fdr"))
+      lfdr.locfdr = unlist(lapply(sims[[i]]$betahat.locfdr,"[[","fdr"))
+
+      subset = lfsr.ash.true<maxx
+
+      res.lfsr[[i]] = data.frame(Scenario=i,Measure='lfsr', ash.nullbiased=lfsr.ash.n[subset], ash.uniform=lfsr.ash.u[subset],Bayes=0.1*maxx*findInterval(lfsr.ash.true[subset],seq(0,maxx,length=11))-0.05*maxx,mixfdr = NA)
+
+      subset = lfdr.ash.true<maxx
+      res.lfdr[[i]] = data.frame(Scenario=i,Measure='lfdr', ash.nullbiased=lfdr.ash.n[subset], ash.uniform=lfdr.ash.u[subset],Bayes=0.1*maxx*findInterval(lfdr.ash.true[subset],seq(0,maxx,length=11))-0.05*maxx,mixfdr = lfdr.mixfdr[subset])
+
+      res[[i]]=rbind(res.lfdr[[i]],res.lfsr[[i]])
+    }
+
+    require(reshape2)
+
+    res.melt = melt(res, id.vars=c("Bayes","Scenario","Measure"),variable.name="Method")
+
+    cbbPalette <- c("#000000","#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+    labels = c('mixfdr','ash (null-biased)','ash (uniform)')
+    breaks = c("mixfdr","ash.nullbiased","ash.uniform")
+
+    p=ggplot(data=res.melt,aes(Bayes,value,colour=Method)) + 
+        facet_grid(Measure ~ Scenario ) +
+        #scale_fill_manual(values=cbbPalette) +
+        #scale_colour_manual(values=cbbPalette) +
+        geom_point(size=1,alpha=0.1) + 
+        #stat_smooth(se=FALSE,size=2) +
+        stat_summary(fun.data="mean_quant", geom="ribbon", alpha=0.25) +
+        geom_abline(colour = "red",size=1) +
+        xlab("Truth") +
+        ylab("Estimate")
+
+    print(p +scale_y_continuous(limits=c(0,maxy)) +
+          scale_x_continuous(limits=c(0,maxx)) +
+            scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels))        
+  }
+
 ```
@@
 ```{r}
-  plot(betahat.ash.n[[1]]$lfsra,betahat.ash.u[[1]]$lfsra,xlab="null-biased prior",ylab="uniform prior", main="fsra is still more robust to estimated pi0",xlim=c(0,0.2),ylim=c(0,0.2))
-  points(betahat.ash.n[[1]]$lfsra,betahat.ash.npm[[1]]$lfsra,col=3)
-  abline(a=0,b=1,col=2,lwd=2)
+png("figures/lfsdr_sim1sim2_blowup.png",height=427,width=720)
+  plot_lfsdr(list(simres1,simres1a,simres2),0.1,ptype="lfsra")
+dev.off()
 ```
@@
 ```{r}
-  plot(betahat.ash.n[[1]]$lfsr,2*betahat.ash.npm[[1]]$lfsr,xlab="null-biased prior",ylab="uniform prior", main="fsra is still more robust to estimated pi0",xlim=c(0,0.2),ylim=c(0,0.2))
-  points(betahat.ash.n[[1]]$lfsra,betahat.ash.npm[[1]]$lfsra,col=3)
-  abline(a=0,b=1,col=2,lwd=2)
+png("figures/lfdr_sim1sim2_blowup.png",height=160,width=540)
+  plot_lfsr(list(simres1,simres1a,simres2),0.1,ptype="lfdr")
+dev.off()
+```
+
+```{r}
+png("figures/lfsra_sim1sim2_blowup.png",height=160,width=540)
+  plot_lfsr(list(simres1,simres1a,simres2),0.1,ptype="lfsra")
+dev.off()
+```
+
+```{r}
+png("figures/lfsra_sim1sim2_blowup.png",height=160,width=540)
+  plot_lfsr(list(simres1,simres1a,simres2),0.1,ptype="lfsr")
+dev.off()
+```
+
+```{r}
+    plot_pi0(simres3)
+  plot_pi1(simres3)
 ```
@@
 ```{r}
-  hist(pval[[1]])
+#hh.ashtrue = hh.ashz
+#hh.ashtrue$fitted.g$pi = c(2/3,1/15,1/15,1/15,1/15,1/15)
+#hh.ashtrue$fitted.g$mean = c(0,0,0,0,0,0)
+#hh.ashtrue$fitted.g$sd = sqrt(c(0,1,0.2,0.4,0.8,3))
+
+#loglik(hh.ashtrue,betahat,sebetahat)
+#loglik(hh.ashz,betahat,sebetahat)
 ```
-
-```{r}
-hh.ashtrue = hh.ashz
-hh.ashtrue$fitted.g$pi = c(2/3,1/15,1/15,1/15,1/15,1/15)
-hh.ashtrue$fitted.g$mean = c(0,0,0,0,0,0)
-hh.ashtrue$fitted.g$sd = sqrt(c(0,1,0.2,0.4,0.8,3))
-
-loglik(hh.ashtrue,betahat,sebetahat)
-loglik(hh.ashz,betahat,sebetahat)
-```

### 2013-11-12 paper/main.tex
update ggplot2 figures

@@
 {\Large
-\textbf{Shrinkage, False Discovery Rate Estimation, and an alternative to the ``Zero assumption" when precision varies across units}
+\textbf{Shrinkage, False Discovery Rate, and False Sign Rate Estimation when precision varies across units}
 }
@@
 A common goal, particularly in genomic studies, is to identify which $\beta_j$ differ from zero. This is commonly tackled
-by first computing a $p$ value for each $j$, testing $H_j:\beta_j=0$, and then applying methods based on False Discovery Rates, such as the qvalue package.
-For example, we could compute $Z$ scores $Z_j = \beta_j/s_j$, and translate $Z_j$ to a $p$ value,
-$p_j$, and then use now-standard methods to select a $p$ value threshold to control the FDR.
+by first computing an effect size estimate ($\hat\beta_j$) and its standard error ($s_j$), converting this to a $Z$ score ($Z_j=\hat\beta_j/s_j$)
+and a corresponding $p$ value ($p_j$), testing $H_j:\beta_j=0$.  Then standard methods (e.g. the qvalue package) can be used
+to estimate False Discovery Rates at any given threshold.
 
@@
 
-Following these ideas of \cite{gelman.etal}, we define the local False Sign Rate (lfsr) for observation $j$ as 
+Following these ideas, we define the local False Sign Rate (lfsr) for observation $j$ as 
 \begin{equation}
@@
 our best guess would be that $\beta$ is negative, and the probability that we have
-made an error in sign would be 0.05.
+made an error in sign would be 0.05. We note that the idea of focussing on tail areas, rather than point null hypotheses,
+has a long history (e.g. Altham? others?).
 
@@
 
-- Illustration that 2 local false sign rate is approximately local fdr.
+\subsection{Improved conservative estimation of $\pi_0$}
+
+To illustrate, we provide simulation results for two scenarios, the first being a ``difficult" case where many non-zero $\beta$ are
+too close to  zero to be reliably detected, and the second being an ``easier" case where most non-zero $\beta$ are
+sufficiently different from zero that they can be reliably detected.
+
+Specifically we use the following distributions for the non-zero $\beta$:
+\begin{itemize}
+\item[Simulation 1:]
+\begin{equation}
+f_1(\cdot) =  0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2)
+\end{equation}
+\item[Simulation 2:]
+\begin{equation}
+f_1(\cdot) = N(0,4^2)
+\end{equation}
+\end{itemize}
+
+Emphasize that although pi0 is unidentifiable, the underlying distribution g is estimable
+quite accurately.
+
+\subsection{Local False Sign Rate}
+
+There are two reasons to use the lfsr instead of the lfdr: it is more generally meaningful (e.g. it applies
+whether or not zero effects exist),  and estimation of lfsr is more robust to modeling assumptions.
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[height=320px,width=540px]{Rcode/figures/lfsdr_sim1sim2_blowup.png}
+\end{center}
+\caption{Figure showing lfsr is more robust than lfdr}
+\end{figure}
+
+- Illustration that estimated lfsr is more robust.
+
+- Illustration that adjusted estimate of lfsr is still more robust..
+
+- Compare estimated lfsr with true lfsr for difficult simulation case, to show nearly achieves Bayes Risk?
+
+\subsection{Accounting for measurement precision improves estimates of FDR/fsr}
 
@@
 comparison with ash.
-
-- Illustration of the difference between the zero assumption and no zero assumption on FDR
-in case z is N(0,2) vs N(0,1). [no difference in precision]
 

### 2013-11-14 Rcode/conservativepm.Rmd
updated figures and text

@@
 
+    cat("applying q value\n")
     betahat.qval[[i]] = qvalue(pval[[i]])
+    cat("applying fdrtool\n")
     betahat.fdrtool[[i]] = fdrtool(pval[[i]],statistic="pvalue",plot=FALSE)
-    betahat.locfdr[[i]] = locfdr(zscore[[i]],nulltype=0,plot=0)
+    cat("applying locfdr\n")
+    betahat.locfdr[[i]] = try(locfdr(zscore[[i]],nulltype=0,plot=0))
+    cat("applying mixfdr\n")
     betahat.mixfdr[[i]] = mixFdr(zscore[[i]],noiseSD=1,theonull=TRUE,plot=FALSE)
@@
 simres2= basicsim(c(0,4),c(1))
-simres3=basicsim(c(0,4),c(1),bsd=c(rep(1,500),rep(10,500)))
-simres4=basicsim(c(0,4),c(1),bsd=c(rep(1,500),rep(10,500)),minpi0=0.9,seed=200)
+simres3=basicsim(c(0,2),c(1),bsd=c(rep(1,500),rep(10,500)))
+simres4=basicsim(c(0,2),c(1),bsd=c(rep(1,500),rep(10,500)),minpi0=0.9,seed=200)
 ```
@@
   plot_pi0 = function(sims){
-    pi0=sims$pi0  
-    pi0_ash.n=unlist(lapply(sims$betahat.ash.n,get_pi0))
-    pi0_ash.u=unlist(lapply(sims$betahat.ash.u,get_pi0)) 
-    pi0_fdrtool = unlist(lapply(sims$betahat.fdrtool,get_pi0.fdrtool))
-    pi0_locfdr=unlist(lapply(sims$betahat.locfdr,get_pi0.locfdr))
-    pi0_mixfdr = unlist(lapply(sims$betahat.mixfdr,get_pi0.mixfdr))
-    pi0_qval = unlist(lapply(sims$betahat.qval,"[[","pi0"))
+    res=list()
+    for(i in 1:length(sims)){
+      pi0=sims[[i]]$pi0  
+      pi0_ash.n=unlist(lapply(sims[[i]]$betahat.ash.n,get_pi0))
+      pi0_ash.u=unlist(lapply(sims[[i]]$betahat.ash.u,get_pi0)) 
+      pi0_fdrtool = unlist(lapply(sims[[i]]$betahat.fdrtool,get_pi0.fdrtool))
+      pi0_locfdr=unlist(lapply(sims[[i]]$betahat.locfdr,get_pi0.locfdr))
+      pi0_mixfdr = unlist(lapply(sims[[i]]$betahat.mixfdr,get_pi0.mixfdr))
+      pi0_qval = unlist(lapply(sims[[i]]$betahat.qval,"[[","pi0"))
 
-    res = data.frame(pi0=pi0,qvalue=pi0_qval,mixfdr=pi0_mixfdr, locfdr=pi0_locfdr, fdrtool=pi0_fdrtool,ash.nullbiased=pi0_ash.n,ash.uniform=pi0_ash.u)
+      res[[i]] = data.frame(Scenario=i,pi0=pi0,qvalue=pi0_qval,mixfdr=pi0_mixfdr, locfdr=pi0_locfdr, fdrtool=pi0_fdrtool,ash.nullbiased=pi0_ash.n,ash.uniform=pi0_ash.u)
+    }
+    cbbPalette <- c("#56B4E9", "#009E73", "#F0E442", "#0072B2", "#E69F00", "#000000", "#D55E00", "#CC79A7")
+    labels = c('mixfdr','qvalue','locfdr','fdrtool','ash (null-biased)','ash (mle)')
+    breaks = c("mixfdr","qvalue","locfdr","fdrtool","ash.nullbiased","ash.uniform")
+
     require(reshape2)
-    res.melt = melt(res, id.vars=c("pi0"),variable.name="Method")
-    p=ggplot(data=res.melt,aes(pi0,value,colour=Method)) +geom_point(shape=16) +
+    res.melt = melt(res, id.vars=c("pi0","Scenario"),variable.name="Method")
+    res.melt = res.melt[sample(1:nrow(res.melt)),]  
+    res.melt$Scenario = as.factor(res.melt$Scenario)
+    levels(res.melt$Scenario) = c("Scenario 1a","Scenario 1b","Scenario 2")
+    p=ggplot(data=res.melt,aes(pi0,value,colour=Method)) +geom_point(shape=1) +
+        facet_grid(. ~ Scenario) +
       geom_abline(colour = "black") +
@@
           scale_x_continuous(limits=c(0,1)) +
+          scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
           coord_equal(ratio=1))
@@
 
-  pdf("figures/estpi0_sim1.pdf")
-    plot_pi0(simres1)
-  dev.off()
-  pdf("figures/estpi0_sim2.pdf")
-    plot_pi0(simres2)
+  png("figures/estpi0_sim1sim2.png",height=160,width=540)
+    plot_pi0(list(simres1,simres1a,simres2))
   dev.off()
@@
 ```{r}
-mean_quant=function (x, mult = 1){
+mean_quant=function(x, mult = 1){
       x <- na.omit(x)
@@
     cbbPalette <- c("#000000","#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-    labels = c('ash (null-biased)','ash (uniform)')
+    labels = c('ash (null-biased)','ash (mle)')
     breaks = c("ash.nullbiased","ash.uniform")
@@
     res.melt = melt(res, id.vars=c("Bayes","Scenario","Measure"),variable.name="Method")
+    res.melt$Scenario = as.factor(res.melt$Scenario)
+    levels(res.melt$Scenario) = c("Scenario 1a","Scenario 1b","Scenario 2")
 
-    cbbPalette <- c("#000000","#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-    labels = c('mixfdr','ash (null-biased)','ash (uniform)')
+    cbbPalette <- c("#E69F00","#000000", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+    labels = c('mixfdr','ash (null-biased)','ash (mle)')
     breaks = c("mixfdr","ash.nullbiased","ash.uniform")
@@
   plot_lfsdr(list(simres1,simres1a,simres2),0.1,ptype="lfsra")
+dev.off()
+```
+
+```{r}
+png("figures/lfsdr_sim1sim2_blowup.png",height=427,width=720)
+  plot_lfsdr(list(simres1,simres1a,simres2),0.1,ymax=0.2,ptype="lfsra")
 dev.off()

### 2013-11-14 paper/Rcode/NoisyIllustration.rmd
updated figures and text

@@
 #biocLite("qvalue")
-setwd("~/Documents/git/ash/Rcode/")
 library("qvalue")
@@
 p= data.frame(pval=pval,type=c(rep("GOOD",500),rep("POOR",500)))
-histogram(~pval|type,p,layout=c(1,2),xlim=c(0,1),breaks=seq(from=0,to=1,length=20),type="count")
+require(ggplot2)
+pdf("figures/good_poor_hist.pdf",height=3,width=6)
+ggplot(p, aes(x=pval)) + facet_grid(.~type)  +
+    geom_histogram(aes(y=..density..),binwidth=.05,
+                   colour="black", fill="white")
+
+  #old version, using trellis
+  #histogram(~pval|type,p,layout=c(1,2),xlim=c(0,1),breaks=seq(from=0,to=1,length=20),type="count")
+dev.off()
 ```
@@
 library("ashr")
-beta.ash.all = ash(betahat,s)
-beta.ash.good = ash(betahat[p$type=="GOOD"],s[p$type=="GOOD"])
-plot(beta.ash.good$PosteriorMean,beta.ash.all$PosteriorMean[p$type=="GOOD"])
-plot(beta.ash.good$PositiveProb,beta.ash.all$PositiveProb[p$type=="GOOD"])
+library("locfdr")
+beta.ash.all = ash(betahat,s,method="fdr")
+beta.ash.good = ash(betahat[p$type=="GOOD"],s[p$type=="GOOD"],method="fdr")
+beta.ash.all.z = ash(betahat/s,1,method="fdr")
+beta.ash.good.z = ash(betahat[p$type=="GOOD"]/s[p$type=="GOOD"],1,method="fdr")
+beta.locfdr.all.z = locfdr(betahat/s,nulltype=0,plot=0)
+beta.locfdr.good.z = locfdr(betahat[p$type=="GOOD"]/s[p$type=="GOOD"],nulltype=0,plot=0)
+res = rbind(data.frame(x=beta.ash.good$lfsr,y=beta.ash.all$lfsr[p$type=="GOOD"],type="ash"), 
+            #data.frame(x=beta.ash.good.z$lfsr,y=beta.ash.all.z$lfsr[p$type=="GOOD"],type='ash, z scores'),
+            data.frame(x=beta.locfdr.good.z$fdr,y=beta.locfdr.all.z$fdr[p$type=="GOOD"],type='locfdr'), 
+            data.frame(x= qq.good$qvalues, y = qq.all$qvalues[p$type=="GOOD"],type="qvalue") )
+
+pdf("figures/good_vs_all.pdf",height=3,width=6.5)
+  pp= ggplot(data=res,aes(x,y)) +geom_point(shape=1) +
+        facet_grid(. ~ type) +
+      geom_abline(colour = "black") +
+        xlab("lfsr/lfdr/qvalue (Good data only)") +
+        ylab("lfsr/lfdr/qvalue (All data)")
+
+  print(pp +scale_y_continuous(limits=c(0,1)) +
+          scale_x_continuous(limits=c(0,1))  +
+          coord_equal(ratio=1))
+
+dev.off()
 ```

### 2013-11-14 paper/main.tex
updated figures and text

@@
 {\Large
-\textbf{Shrinkage, False Discovery Rate, and False Sign Rate Estimation when precision varies across units}
+\textbf{Adaptive shrinkage and the false sign rate}
 }
@@
 
- \section*{Methods}
-
- The two main assumptions we will make here are that 
-* these effects are exchangeable: that is (by de Finetti's theorem) they can be
-thought of as being independent and identically distributed from some unknown distribution $g(\beta)$. 
-* these effects are symmetric and unimodal about 0.
-
-Specifically we will assume a parametric form for $g$
-as a mixture of 0-centered normal distributions
-
 A key issue I want to address here is that if these measurements are made with different precisions, then we want to
@@
 
-The methods are based on treating the vectors $\hat\beta$
-and $s$ as ``observed data", and then performing inference for $\beta$ from these observed data, using a standard hierarchical modelling framework
-to combine information across $j=1,\dots,J$.
+ \section*{Methods}
 
-Specifically, 
-we assume that the true 
-$\beta_j$ values are independent
-and identically distributed from some distribution $g(\cdot ; \pi)$, where $\pi$ is a hyper-parameter to be estimated. 
-Then, given $\beta$, we assume that $(\hat\beta_j,s_j)$ are independent across $j$, and depend only on $\beta_j$.
+ Suppose that we are interested in the values of $n$ ``effects" $\beta_j$ ($j=1,\dots,n$).
+In some contexts our interest may focus on which of the $\beta_j$ are ``significantly" different from zero, whereas in other contexts our interest may focus on estimating their values; the methods described here are suited to both these contexts.
+ We assume that we have obtained data $D_1,\dots,D_n$ that provide independent estimates $\hat\beta_1,\dots,\hat\beta_n$ of these effects,
+ with corresponding (estimated) standard errors $s_1,\dots,s_n$. (See Discussion for ideas on how our methods may be applicable more generally; for example
+ if, in place of $s_j$, we have a $p$ value testing $\beta_j=0$.)
+
+ In outline, we use a hierarchical model to combine information across measurements, with the aim of improving both accuracy and precision of estimates. 
+ Specifically we assume that the effects $\beta_j$ are independent and identically distributed from some unknown distribution $g(\cdot)$.  
+ Our methods are based on the following key assumptions:
+   \begin{enumerate}
+  \item The distribution $g(\cdot)$ is unimodal. (We assume here that the mode is at 0, although 
+this could be relaxed.)  
+ \item The likelihood $L(\beta_j) = p(D_j | \beta_j)$ can be approximated by a normal likelihood,
+ \begin{equation} \label{eqn:lik}
+ L(\beta_j) \propto \exp[-0.5(\beta_j-\hat\beta_j)^2/s_j^2].
+ \end{equation}
+ \end{enumerate}
+Although these assumptions will not apply to all situations, we argue that both will be reasonable in many practical contexts. 
+For example, in contexts where interest focusses on which $\beta_j$ differ from zero, and ``$\beta_j=0$" is a plausible null hypothesis,  
+it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible, and that the distribution of the effects will be unimodal about 0. 
+Alternatively, in an estimation context, we can motivate Assumption 1 by considering its effect on point estimates, which is to ``shrink" the estimates towards the mode;
+this kind of shrinkage is widely regarded as useful, and provides an alternative motivation for our methods.
+Assumption 2 can be motivated either as a direct approximation to the likelihood (e.g.~by Taylor series expansion of the log-likelihood about $\hat\beta_j$),
+or by assuming that $\hat\beta_j, s_j$ are sufficient statistics for $\beta_j$ [so $L(\beta_j) \propto p(\hat\beta_j, s_j | \beta)$], and assuming that  $\hat\beta_j | s_j, \beta_j \sim N(\beta_j,s_j)$ and that $p(s_j|\beta_j)$ does not depend on $\beta_j$.
+[In some settings a $t$ distribution assumption for $\hat\beta_j$ may be more appropriate; see Discussion.]
+Assumption 2 has obvious parallels with approaches like \cite{efron, cai?} that model the $Z$ scores $Z_j = \beta_j/s_j$ as normally distributed under the null. However,
+there is an important difference: in our approach observations with a large $s_j$ have an essentially flat likelihood \ref{eqn:lik}, and so do not affect inference for
+other parameters, whereas when modeling the $Z$ scores directly, observations with large $s_j$ produce $N(0,1)$ $Z$ scores that do affect inference. In
+this way the likelihood-based approach can take better account of the informativeness of each measurement, as illustrated in Results below.
+
+%The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
+For example, we can write $p(\hat\beta_j, s_j | \beta_j) = p(\hat\beta_j | s_j, \beta_j)p(s_j | \beta_j)$.
+Now, if we are willing to assume that $s_j$ alone contains no information about $\beta_j$, or equivalently that $p(s_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\hat\beta_j | s_j, \beta_j)$, and if $\hat\beta_j | s_j, \beta_j \sim N(\beta_j, s_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.
+
+Interestingly, although the unimodal assumption for $g$ seems natural to us, much previous related work on estimating False Discovery Rates  
+has tended to assume, either explicitly or implicitly, that the effect sizes are multimodal. (In contrast, almost all analogous work in large-scale regression assumes
+unimodal distributions for the effect sizes, such as the spike and slab, Laplace, NEG, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors.)
+We discuss the effects of these assumptions later.
+For now, it is important to note that Assumption 1 relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
+It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and Assumption 1 does not contradict this.
+
+In practice, we implement our approach by assuming a parametric finite mixture  for $g$, 
+\begin{equation}
+g(\cdot; \pi) = \pi_0 \delta_0(\cdot) + (1-\pi_0) \sum_{k=1}^K \pi_k f_k(\cdot) 
+\end{equation}
+[Or possibly remove $(1-\pi_0)$ from this?]
+where $\delta_0(\cdot)$ denotes a point mass at 0, $\pi=(\pi_0,\dots,pi_K)$ are mixture proportions to be estimated, and $f_k$ are pre-specified 
+component distributions with one of the following forms: 
+\begin{enumerate}[i)]
+\item $f_k(\cdot) = N(0, \sigma^2_k)$
+\item $f_k(\cdot) = U[-a_k,a_k]$
+\item $f_k(\cdot) = U[-a_k,0] \text{ or } U[0,a_k]$,
+\end{enumerate}
+where $N(0,\sigma^2)$ denotes the normal distribution with mean 0 and variance $\sigma^2$ and
+$U[a,b]$ denotes the unform distribution on $[a,b]$.
+That is, the non-zero effects are modeled as either i) a mixture of zero-centered normal distributions; ii) a mixture of zero-centered uniform distributions;
+or iii) a mixture of ``zero-anchored" uniform distributions.
+
+Pre-specifying the distributions $f_k$, rather than estimating them, may seem both inelegant and restrictive. However,
+ any inelegance is, in our opinion, more than off-set by the very considerable computational simplifications it produces when fitting these models (see below). 
+ And any apparent restrictiveness
+is illusory: by making
+$K$ sufficiently large one can approximate any symmetric unimodal distribution arbitrarily closely using the uniforms in (ii),
+and approximate any (asymmetric) unimodal distribution using the uniforms in (iii); and even the more restrictive (i) is sufficiently flexible
+to be useful in practice. Indeed, (i) can approximate arbitrarily closely any scale mixture of normals, which includes as a special case
+the double exponential distribution, any $t$ distribution, and other distributions used in high-dimensional regression settings such as the horseshoe prior \ref{},
+all of which are more restrictive. Furthermore, the unimodal constraint is sufficiently strong to prevent serious problems with overfitting when using large $K$. Indeed, by 
+increasing $K$ one can ultimately approach the non-parametric maximum likelihood estimate for $g$.
+Related to this, see \cite{grenander}, which provides the (extremely elegant) solution to the non-parametric maximum likelihood estimate for $g$ in the limit as the errors $s_j$ all tend to zero;
+and see \cite{campy.thomas} who suggest similar methods to those used here in the case where $s_j$ are all equal.
+
+% $g() = \sum_{k=1}^K (\pi^+_k U[0,a_k] + \pi^-_k U[-a_k,0])$
+
+\subsection*{Likelihood for $\pi$}
+
+Assuming independence of the data leading to $\hat\beta_1,\dots,\hat\beta_n$, the likelihood for $\pi$ can be written as 
+\begin{equation}
+p(\hat\beta, s | \pi) = \prod_{j=1}^n p(\hat\beta_j, s_j | \pi).
+\end{equation}
+Further, the assumptions  $\hat\beta_j | \beta, s_j \sim N(\beta,s_j)$ and $\beta_j | s_j, \pi \sim g(\cdot; \pi)$ yield a mixture form for $\hat\beta_j$: 
+\begin{equation}
+p(\hat\beta_j, s_j | \pi) = \pi_0 N(0,s^2_j) + \sum_k \pi_k \tilde{f}_{kj}(\hat\beta_j)
+\end{equation}
+where $\tilde{f}_{kj}$ denotes the density of a convolution of $f_k$ with an $N(0,s^2_j)$ density.
+For example, if $f_k$ is $N(0,\sigma^2_k)$ then the convolution $\tilde{f}_{kj}$ is also normal, $\tilde{f}_{kj} = N(0,\sigma^2_k + s^2_j)$. If
+$f_k$ is $U[a_k,b_k]$ then 
+\begin{equation}
+\tilde{f}_{kj}(x) = 
+\end{equation}
+by standard results on the convolution of a uniform with a normal.
+
+\begin{equation}
+\end{equation}
 Putting these together, the joint model for the unobserved $\beta$ and the observed $\hat\beta, s$ is:
@@
 \end{align}
-The specific choices of $g$ and $L$ are described below.
 
 We fit this hierarchical model using the following "Empirical Bayes" approach. First we estimate the hyper-parameters $\pi$ by maximizing the likelihood
-$$L(\pi; \hat\beta, s) := p(\hat\beta, s | \pi) = \int p(\hat\beta, s, \beta | \pi) d\beta.$$
-Then, given this estimate $\hat\pi$, we compute the conditional distributions $$p(\beta_j | \hat\pi, \hat\beta, s) \propto g(\beta_j; \pi) L(\beta_j; \hat\beta_j, s_j).$$ 
+\begin{equation}
+L(\pi; \hat\beta, s) := p(\hat\beta, s | \pi) = \int p(\hat\beta, s, \beta | \pi) d\beta.
+\end{equation}
+This can be done very easily using an EM algorithm.
+Then, given this estimate $\hat\pi$, we compute the conditional distributions 
+\begin{equation}
+p(\beta_j | \hat\pi, \hat\beta, s) \propto g(\beta_j; \pi) L(\beta_j; \hat\beta_j, s_j).
+\end{equation} 
 In principle we would
-prefer to take a full Bayes approach that accounts for uncertainty in $\pi$, but, at least for now, we compromise this principle for the simplicity of the EB approach.
-[Note: a Variational Bayes version of this is
-also implemented, and may become our preferred approach
-after testing]
+prefer to take a full Bayes approach that accounts for uncertainty in $g$ (by accounting for uncertainty in $\pi$); however,
+we believe that in most practical applications uncertainty in $g$ will not be the most important concern, and so 
+we compromise this principle for the simplicity of the EB approach.
+[Note: a Variational Bayes approach to estimating $\pi$ is also implemented in our R package]
 
-[put picture of hierarchical model here]
+%[maybe put picture of hierarchical model here]
 
@@
 $j=1,\dots,J$. The combining of the information occurs through estimation of
-$\pi$, which involves all of the data, and it is
-
+$\pi$, which involves all of the data.
 These conditional distributions can be conveniently summarized
@@
 
-Efron (2008) states ``most of the z-values near zero come from null genes"
+Efron (2008) states the Zero Assumption as the assumption that ``most of the z-values near zero come from null genes".
 His main aim in making this assumption is to estimate an empirical null though (not assume N(0,1) for the null)
@@
 
-The key components of this hierarchical model
-are the distribution $g$ and the likelihood $L(\beta_j; \hat\beta_j, s_j)$. We make the following choices for these.
+In practice, we currently fix the number of components $K$ to be large, and take the variances $\sigma_1<\sigma_2<\dots<\sigma_K$ to be fixed, and vary from very small to very large --  sufficiently large that typically $\hat\pi_K \approx 0$.
 
-1. The likelihood for $\beta_j$ is normal, centered on $\hat\beta_j$, with standard deviation $s_j$.
-That is, 
-$$L(\beta_j; \hat\beta_j, s_j) \propto \exp[-0.5(\beta_j-\hat\beta_j)^2/s_j^2]. \quad (**)$$
-
-2. The distribution $g(\cdot; \pi)$ is a mixture of zero-centered normal distributions, 
-$$g(\cdot; \pi) = \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma^2_k).$$
-In practice, we currently fix the number of components $K$ to be large, and take the variances $\sigma_1<\sigma_2<\dots<\sigma_K$ to be fixed, and vary from very small (possibly 0), to very large --  sufficiently large that typically $\hat\pi_K=0$.
-
-The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
-For example, we can write $p(\hat\beta_j, s_j | \beta_j) = p(\hat\beta_j | s_j, \beta_j)p(s_j | \beta_j)$.
-Now, if we are willing to assume that
-$s_j$ alone contains no information about $\beta_j$, or equivalently that $p(s_j | \beta_j)$ does not depend on $\beta_j$, then
-$L(\beta_j) \propto p(\hat\beta_j | s_j, \beta_j)$,
-and if $\hat\beta_j | s_j, \beta_j \sim N(\beta_j, s_j^2)$,
-as is often asymptotically the case, then we obtain the likelihood (**) above.
-
-An alternative motivation is to think of this as a normal
-approximation to the likelihood from the 
-raw data $D_j$ that were used to compute
-$\hat\beta_j$ and $s_j$. Then if we observed these data
-the likelihood for $\beta$ would be 
-$p(D_j | \beta_j)$, and a Taylor series expansion of the log likelihood around the maximum likelihood estimate $\hat\beta_j$  yields $$l(\beta_j) \approx l(\hat\beta_j) + 0.5* (\beta_j - \hat\beta_j)^2 l''(\hat\beta_j).$$ [Fill in details?]
+%A Taylor series expansion of the log likelihood around the maximum likelihood estimate $\hat\beta_j$  yields $$l(\beta_j) \approx l(\hat\beta_j) + 0.5* (\beta_j - \hat\beta_j)^2 l''(\hat\beta_j).$$ [Fill in details?]
 
 Using a mixture of normal distributions for $g$ 
-also seems very natural: mixtures of normals provide a flexible family of distributions able to provide a good approximation to any true underlying $g$; and 
-when combined with the normal likelihood they give
+has the advantage that,
+when combined with the normal likelihood, they give
 an analytic form for the conditional distribution $p(\beta_j | \hat\pi, \hat\beta_j, s_j)$ (also a mixture of normals).
-The constraint that these normals be centered at zero may seem initially less natural. 
-Certainly this constraint could be relaxed in principle.
-However, we view it as a convenient way to impose an assumption 
-that $g$ is unimodal with its mode at 0,
-which we view as a plausible assumption in many 
-settings, and one which may be helpful to avoid
-"overfitting" of $g$. (Using normal distributions centered at 0
-also imposes an assumption
-that $g$ is symmetric about zero, which we view as less
-plausible, but represents a compromise between simplicity
-and flexibility.
-In cases where this assumption seems wildly inappropriate one could perhaps
-improve results by applying the
-model separately to positive and negative values of $\hat\beta$.)
 
-Finally, using a large number of normal components with a wide range of variances,
-rather than, say, a smaller number of components with the variances
-to be estimated, is simply for computational convenience. With fixed
-variances there exists a very simple EM algorithm to maximise the likelihood
-in $\pi$. Obtaining maximum likelihood estimates for the variances could certainly
-be implemented with a little
-more work, but it is unclear whether this would result in practically-important gains in many situations of interest.
-
-\subsection{The local False Sign Rate}
+\subsection*{The local False Sign Rate}
 
@@
 As pointed out by \cite{gelman.etal}, there are settings where
-$\beta_j=0$ is implausible, in which case the lfdr is not
-a useful concept: if every $\beta_j$ is non-zero then there
-is no such thing as a false discovery and the lfdr will be identically 0 for all methods. 
+$\beta_j=0$ is implausible, in which case the lfdr is not useful: if every $\beta_j$ is non-zero then there
+is no such thing as a false discovery and the lfdr will always be identically 0. 
 Gelman et al suggest that in such settings we
 might replace the concept of a false discovery with the
-concept of an "error in sign". The idea is that, in settings where $\beta_j=0$ is implausible, the most fundamental
+concept of an ``error in sign". The idea is that, in settings where $\beta_j=0$ is implausible, the most fundamental
 inference objective is to ask which $\beta_j$ are positive and which are negative. Indeed, even in settings where some
 $\beta_j$ are exactly zero, it could be argued that identifying which are positive and which negative is
-fundamentally more interesting and useful than identifying which are non-zero.
+fundamentally more interesting and useful than identifying which are non-zero. For example, when identifying differentially expressed genes,
+analysts will often separate the genes that are ``upregulated" in one condition from those that ``downregulated" in that condition.
 
-Following these ideas, we define the local False Sign Rate (lfsr) for observation $j$ as 
+Motivated by this, we define the local False Sign Rate (lfsr) for observation $j$ as 
 \begin{equation}
@@
 
-\subsection{Improved conservative estimation of $\pi_0$}
+\begin{figure}[!ht] \label{fig:good_poor_hist}
+\begin{center}
+\includegraphics[height=3in]{Rcode/figures/good_poor_hist.pdf}
+\includegraphics[width=6.5in]{Rcode/figures/good_vs_all.pdf}
+\end{center}
+\caption{}
+\end{figure}
 
-To illustrate, we provide simulation results for two scenarios, the first being a ``difficult" case where many non-zero $\beta$ are
-too close to  zero to be reliably detected, and the second being an ``easier" case where most non-zero $\beta$ are
-sufficiently different from zero that they can be reliably detected.
+\subsection*{Improved conservative estimation of $\pi_0$}
 
-Specifically we use the following distributions for the non-zero $\beta$:
+To illustrate estimation of $\pi_0$ we provide simulation results for two scenarios:
 \begin{itemize}
-\item[Simulation 1:]
+\item[Scenario 1:]
 \begin{equation}
@@
 \end{equation}
-\item[Simulation 2:]
+\item[Scenario 2:]
 \begin{equation}
@@
 \end{itemize}
+Scenario 1 represents a ``difficult" case where many non-zero $\beta$ are
+too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible; 
+Scenario 2 represents an ``easier" case where most non-zero $\beta$ are
+sufficiently different from zero that they can be reliably detected. making estimation of $\pi_0$ easier.
+For Scenario 1 we considered datasets of size $n=1000$ (Scenario 1a) and $n=10,000$ (Scenario 1b);
+for Scenario 2 we used $n=1000$.
 
-Emphasize that although pi0 is unidentifiable, the underlying distribution g is estimable
-quite accurately.
+For each simulation scenario we simulated 200 independent data sets. For each data set we simulated
+the true value of $\pi_0  \sim U[0,1]$, and then simulated $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) f_1(\cdot)$
+and $\hat\beta_j | \beta_j = N(\beta_j,1)$ for $j=1,\dots,n$. Thus these simulations assume the same
+precision for each measurement. We applied the methods implemented in the R packages qvalue, fdrtool, locfdr, and mixfdr
+for estimating $\pi_0$. We also applied our shrinkage-based method to estimate $\pi_0$ in two ways, one using a ``prior" distribution to encourage
+$\pi_0$ to be as large as possible (``null-biased"), and the other using maximum likelihood estimation for $\pi$.
 
-\subsection{Local False Sign Rate}
+Figure \ref{fig:pi0} compares estimated and true values of $\pi_0$ under each Scenario. 
+For Scenario 1, none of the method reliably estimate $\pi_0$, as expected since the scenario was designed to make estimation
+of $\pi_0$ impossible. However, we see that all the methods except for ash (mle) are able to provide a conservative
+estimate for $\pi_0$. Further, among these the estimates from ash (null-biased) are least conservative. Scenario 2 produces similar patterns,
+except that for this scenario the ash estimates of $\pi_0$ are quite accurate (made possible by the fact that most non-zero effects
+are very different from zero).
+
+These simulations illustrate two key points: i) although $\pi_0$ is not identifiable, the penalized likelihood approach
+provides conservative estimates for $\pi_0$; ii) replacing the ``zero assumption" with the unimodal assumption can produce
+less conservative, more accurate, estimates of $\pi_0$ (provided of course that the unimodal assumption holds, as it does in these simulations).
+
+\begin{figure}[!ht] \label{fig:pi0}
+\begin{center}
+\includegraphics[height=160px,width=540px]{Rcode/figures/estpi0_sim1sim2.png}
+\end{center}
+\caption{Comparison of true and estimated values of $\pi_0$ for simulation scenarios.
+Scenarios 1a and 1b represent ``difficult" scenarios where $\pi_0$ is impossible to estimate accurately.
+However, all methods except the ash mle method are successful in providing conservative estimates for $\pi_0$.
+The ash (null-biased) method is least conservative, and hence most accurate, due to its additional assumption that the effect size distribution $g$ is unimodal.}
+\end{figure}
+
+Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
+can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
+by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
+
+\subsection*{Local False Sign Rate}
 
 There are two reasons to use the lfsr instead of the lfdr: it is more generally meaningful (e.g. it applies
-whether or not zero effects exist),  and estimation of lfsr is more robust to modeling assumptions.
+whether or not zero effects exist),  and estimation of lfsr is more robust to modeling assumptions and estimation of $\pi_0$. 
+To illustrate this, we compared estimated and true values of both lfdr and lfsr for the simulated data (where the true
+values are computed by Bayes Theorem using the true value of $g$ FILL IN THESE DETAILS?).
 
@@
 with model in which effects
-
-- unimodal distributions could be argued to be a primary motivation for shrinkage
 

### 2013-11-14 Rcode/ash.R

@@ -159,7 +159,8 @@ ash = function(betahat,sebetahat,method = c("shrink","fdr"),
     if(mixcompdist=="halfuniform") g=unimix(c(pi,pi),c(-mixsd,rep(0,k)),c(rep(0,k),mixsd))
     maxiter = 5000
   } else {
-    maxiter = 1; # if g is specified, don't iterate the EM 
+    maxiter = 1 # if g is specified, don't iterate the EM 
+    prior = rep(1,ncomp(g)) #prior is not actually used if g specified, but required to make sure EM doesn't produce warning
   }
   
   pi.fit=EMest(betahat[completeobs],lambda1*sebetahat[completeobs]+lambda2,g,prior,null.comp=null.comp,nullcheck=nullcheck,VB=VB,maxiter = maxiter)  

### 2013-11-19 Rcode/conservativepm.Rmd
worked on ashsim.rmd

@@
 ```{r}
-  rmse = function(x,y){sqrt(mean((x-y)^2))}
-  get_rmse.ash = function(a,b){rmse(a$PosteriorMean,b)}
-  get_rmse.mixfdr = function(a,b){rmse(a$effectSize,b)}
-  plot_rmse = function(sims,inczero=FALSE,incbetahat=FALSE){
-    err.ash.n= mapply(get_rmse.ash,sims$betahat.ash.n,sims$beta)
-    err.ash.u = mapply(get_rmse.ash,sims$betahat.ash.u,sims$beta)
-    err.mixfdr=mapply(get_rmse.mixfdr,sims$betahat.mixfdr,sims$beta)
-    err.betahat = mapply(rmse,sims$betahat,sims$beta)
-    err.zero = unlist(lapply(sims$beta,rmse,y=0))
-
-     res = data.frame(mixfdr=err.mixfdr,ash.nullbiased=err.ash.n,ash.uniform=err.ash.u)
-    if(inczero){
-      res=data.frame(res,zero=err.zero)
-    }
-    if(incbetahat){
-      res=data.frame(res,betahat=err.betahat)
-    }
-    require(reshape2)
-    res.melt = melt(res, id.vars=c("ash.uniform"),variable.name="Method")
-
-    p=ggplot(data=res.melt,aes(ash.uniform,value,colour=Method)) +geom_point(shape=16) +
-        geom_abline(colour = "black") +
-        xlab("RMSE (ash.uniform)") +
-        ylab("RMSE (other method)")
-    print(p +scale_y_continuous(limits=c(0,max(res))) +
-          scale_x_continuous(limits=c(0,max(res))) +
-          coord_equal(ratio=1))
-  }
-  pdf("figures/rmse_sim1.pdf")
-    plot_rmse(simres1)
-  dev.off()
-  pdf("figures/rmse_sim2.pdf")
-    plot_rmse(simres2)
-  dev.off()
-```
-
-```{r}
-  plot_LR=function(sims){
-    hist(unlist(lapply(sims$betahat.ash.u,get_loglik))-unlist(lapply(sims$betahat.ash.n,get_loglik)), xlab="loglik difference", main="loglik differences for nullbiased prior vs mle",nclass=10)
-  }
-
-  pdf("figures/logLR.pdf")
-    plot_LR(simres1)
-    plot_LR(simres2)
-  dev.off()
-```
-
-```{r}
 mean_quant=function(x, mult = 1){

### 2013-11-19 paper/Rcode/ashsim.rmd
worked on ashsim.rmd

```{r}
  require("ashr")
  ```{r}
opts_knit$set(progress = TRUE, verbose = TRUE,root.dir="~/Documents/git/ash/paper/Rcode")
require(ashr)
require(qvalue)
require(fdrtool)
require(mixfdr)
require(locfdr)
require(ggplot2)
#simulate from mixture of normals
#compare normal, uniform and half-uniform mixtures
#bsd gives standard deviation of beta
ashsim=function(mixmean,mixsd,mixpi,bsd=1,seedval = 100,nsamp=1000,niter=50){  
  set.seed(seedval)  
  beta =list()
  betahatsd=list()
  betahat = list()
  betahat.ash.n = list()
  betahat.ash.u = list()
  betahat.ash.hu = list()
  k = length(mixmean)
  for(i in 1:niter){
    comp = sample(1:k,nsamp,prob=mixpi,replace=TRUE)
    sd = mixsd[comp]
    mean = mixmean[comp]
    beta[[i]] = rnorm(nsamp,mean,sd)
    betahatsd[[i]] = bsd
    betahat[[i]] = beta[[i]]+rnorm(nsamp,0,betahatsd[[i]])
    betahat.ash.n[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="normal",method="shrink")
    betahat.ash.u[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="uniform", method="shrink")
    betahat.ash.hu[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform", method="shrink")

  }
  return(list(beta =beta,
  betahatsd=betahatsd,
  betahat = betahat,
  betahat.ash.n = betahat.ash.n,
  betahat.ash.u = betahat.ash.u,
  betahat.ash.hu = betahat.ash.hu))
}    
```

```{r}
sim1= ashsim(c(0,0,0),c(1,1,2),c(1/3,1/3,1/3),niter=20,nsamp=1000)
sim2= ashsim(c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7),rep(1/7,7),niter=20,nsamp=1000)
sim3= ashsim(c(-2,-1,0,1),c(2,1.5,1,1),c(1/4,1/4,1/3,1/6),niter=20,nsamp=1000)

```

```{r}
len=length(sim1$beta[[1]])
res = data.frame(beta = c(sim1$beta[[1]],sim2$beta[[1]],sim3$beta[[1]]), Scenario = c(rep(1,len),rep(2,len),rep(3,len)))
x = seq(-8,6,length=100)
y = dnorm(x)
dens1 = data.frame(x=x,dn= t(density(sim1$betahat.ash.n[[1]],x)$y),du=t(density(sim1$betahat.ash.u[[1]],x)$y),dhu=t(density(sim1$betahat.ash.hu[[1]],x)$y), Scenario=1)
dens2 = data.frame(x=x,dn = t(density(sim2$betahat.ash.n[[1]],x)$y),du=t(density(sim2$betahat.ash.u[[1]],x)$y), dhu=t(density(sim2$betahat.ash.hu[[1]],x)$y), Scenario=2)
dens3 = data.frame(x=x,dn = t(density(sim3$betahat.ash.n[[1]],x)$y),du=t(density(sim3$betahat.ash.u[[1]],x)$y), dhu=t(density(sim3$betahat.ash.hu[[1]],x)$y), Scenario=3)
dens=rbind(dens1,dens2,dens3)
ggplot(res, aes(x=beta)) + facet_grid(.~Scenario)  +
    geom_histogram(aes(y=..density..),binwidth=0.5,
                   colour="black", fill="white") +  
  geom_line(data = dens, aes(x,du,color='Uniform'),size=1.2,alpha=0.8) +
  geom_line(data = dens, aes(x,dhu,color='Half Uniform'),size=1.2,alpha=0.8) +
    geom_line(data = dens, aes(x,dn,color='Normal'),size=1.2,alpha=0.8) +
  scale_colour_manual(name = 'Method', values = c('red', 'blue','green'))

```

```{r}
  rmse = function(x,y){sqrt(mean((x-y)^2))}
  get_rmse.ash = function(a,b){rmse(a$PosteriorMean,b)}
  get_rmse.mixfdr = function(a,b){rmse(a$effectSize,b)}
  plot_rmse = function(sims,inczero=FALSE,incbetahat=FALSE){
    res=list()
    res2=list()
    for(i in 1:length(sims)){
      err.ash.n= mapply(get_rmse.ash,sims[[i]]$betahat.ash.n,sims[[i]]$beta)
      err.ash.u = mapply(get_rmse.ash,sims[[i]]$betahat.ash.u,sims[[i]]$beta)
      err.ash.hu = mapply(get_rmse.ash,sims[[i]]$betahat.ash.hu,sims[[i]]$beta)
    #err.mixfdr=mapply(get_rmse.mixfdr,sims$betahat.mixfdr,sims$beta)
      err.betahat = mapply(rmse,sims[[i]]$betahat,sims[[i]]$beta)
      err.zero = unlist(lapply(sims[[i]]$beta,rmse,y=0))
      loglik.ash.n = unlist(lapply(sims[[i]]$betahat.ash.n,get_loglik))
      loglik.ash.u = unlist(lapply(sims[[i]]$betahat.ash.u,get_loglik))
      loglik.ash.hu = unlist(lapply(sims[[i]]$betahat.ash.hu,get_loglik))
     res[[i]] = data.frame(Scenario=i,ash.normal=err.ash.n,ash.uniform=err.ash.u,ash.halfuniform=err.ash.hu)
      res2[[i]] = data.frame(Scenario=i,ash.normal=loglik.ash.n,ash.uniform=loglik.ash.u,ash.halfuniform=loglik.ash.hu)

      if(inczero){
        res[[i]]=data.frame(res,zero=err.zero)
      }
      if(incbetahat){
        res[[i]]=data.frame(res,betahat=err.betahat)
      }
    }
    require(reshape2)
    res.melt = melt(res, id.vars=c("Scenario"),variable.name="Method")
    res2.melt = melt(res2, id.vars=c("Scenario"),variable.name="Method")
    ggplot(data=res.melt,aes(Method,value)) +geom_boxplot()+         facet_grid(. ~ Scenario)
     ggplot(data=res2.melt,aes(Method,value)) +geom_boxplot()+         facet_grid(. ~ Scenario)

    p=ggplot(data=res.melt,aes(ash.halfuniform,value,colour=Method)) +geom_point(shape=16) +
         facet_grid(. ~ Scenario) +
        geom_abline(colour = "black") +
        xlab("RMSE (ash.halfuniform)") +
        ylab("RMSE (other method)")
    print(p +
          coord_equal(ratio=1))
  }
  #pdf("figures/rmse_sim1.pdf")
    plot_rmse(list(sim1,sim2,sim3))
 #               dev.off()
  #pdf("figures/rmse_sim2.pdf")
  #  plot_rmse(simres2)
  #dev.off()
```

Try plotting boxplots of rmse and fit impovement compared with most general model
```{r}
  rmse = function(x,y){sqrt(mean((x-y)^2))}
  get_rmse.ash = function(a,b){rmse(a$PosteriorMean,b)}
  get_rmse.mixfdr = function(a,b){rmse(a$effectSize,b)}
  boxplot_rmse_loglik = function(sims,inczero=FALSE,incbetahat=FALSE){
    res=list()
    res2=list()
    for(i in 1:length(sims)){
      err.ash.hu = mapply(get_rmse.ash,sims[[i]]$betahat.ash.hu,sims[[i]]$beta)
      err.ash.n= mapply(get_rmse.ash,sims[[i]]$betahat.ash.n,sims[[i]]$beta)-err.ash.hu
      err.ash.u = mapply(get_rmse.ash,sims[[i]]$betahat.ash.u,sims[[i]]$beta)-err.ash.hu
      #err.mixfdr=mapply(get_rmse.mixfdr,sims$betahat.mixfdr,sims$beta)
      loglik.ash.hu = unlist(lapply(sims[[i]]$betahat.ash.hu,get_loglik))
      loglik.ash.n = unlist(lapply(sims[[i]]$betahat.ash.n,get_loglik))-loglik.ash.hu
      loglik.ash.u = unlist(lapply(sims[[i]]$betahat.ash.u,get_loglik))-loglik.ash.hu

     res[[i]] = data.frame(Scenario=i,ash.normal.err=err.ash.n,ash.uniform.err=err.ash.u)
       res2[[i]] = data.frame(Scenario=i,ash.normal=loglik.ash.n,ash.uniform=loglik.ash.u)

    }
    require(reshape2)
    res.melt = melt(res, id.vars=c("Scenario"))

    res2.melt=melt(res2,id.vars=c("Scenario"))
    ggplot(res.melt,aes(variable,value)) + geom_boxplot() + facet_grid(.~Scenario)
    ggplot(res2.melt,aes(variable,value)) + geom_boxplot() + facet_grid(.~Scenario)
  }

  pdf("figures/boxplot_rmse_loglik.pdf")
    boxplot_rmse_loglik(list(sim1,sim2,sim3))
  dev.off()

```

```{r}
  plot_LR=function(sims){
    hist(unlist(lapply(sims$betahat.ash.u,get_loglik))-unlist(lapply(sims$betahat.ash.n,get_loglik)), xlab="loglik difference", main="loglik differences for nullbiased prior vs mle",nclass=10)
  }

  pdf("figures/logLR.pdf")
    plot_LR(simres1)
    plot_LR(simres2)
  dev.off()
```

```{r}
  lapply()
```

### 2013-12-15 paper/main.tex
updated paper and code

@@
 
+It is common in statistics that you measure many ``similar" things imperfectly, and wish to estimate their values.
+The situation arises commonly in the kinds of genomics applications I am often involved in, but also in other areas of statistics.
+In genomics, for example, a very common goal
+is to compare the mean
+expression (activity) level of many genes in two conditions.
+Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
+of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
+measurements are made on only a small number of
+samples in each condition - sometimes as few as one
+sample in each condition. Thus the error in estimates
+of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
+in estimates of $\beta_j$ still greater.
+
+A fundamental idea is
+that the measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.
+
 Suppose that we measure, with error, a series of ``effects", $\beta_1,\dots,\beta_J$. To take just one concrete 
@@
 
-Given that shrinkage estimation is widely recognized as a powerful tool, there are suprisingly few software packages for performing the simplest type of shrinkage estimation considered here.
+Given that shrinkage estimation is widely recognized as a powerful tool, there are surprisingly few software packages for performing the simplest type of shrinkage estimation considered here.
 (There are more packages for the more complex setting of covariance estimation, where shrinkage is perhaps still more important.)
@@
 to the functionality provided here is mixfdr (Muralidharan). Compared with mixfdr, the key features of ashr are that it
-i) focuses on allowing for variation in the standard deviation of each observation; ii) constrains the underlying density to be unimodal (and possibly symettric).
-NOTE: should emphasise these differences in the examples.
+i) focuses on allowing for variation in the standard deviation of each observation; ii) constrains the underlying density to be unimodal (and possibly symmetric).
+NOTE: should emphasize these differences in the examples.
 
@@
 
-It is common in statistics that you measure many ``similar" things imperfectly, and wish to estimate their values.
-The situation arises commonly in the kinds of genomics applications I am often involved in, but also in other areas of statistics.
-In genomics, for example, a very common goal
-is to compare the mean
-expression (activity) level of many genes in two conditions.
-Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-measurements are made on only a small number of
-samples in each condition - sometimes as few as one
-sample in each condition. Thus the error in estimates
-of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-in estimates of $\beta_j$ still greater.
-
-A fundamental idea is
-that the measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.
-
-A key issue I want to address here is that if these measurements are made with different precisions, then we want to
-take this into account in our analysis.
-We can do this by assuming that the likelihood for $\beta$ is a normal, with mean $\hat\beta_p$ and standard deviation $s_p$. (Note that this is equivalent to the likelihood we would get
-if we ``observed" data $\hat\beta_p \sim N(\beta_p, s_p)$.)
-
  \section*{Methods}
@@
  We assume that we have obtained data $D_1,\dots,D_n$ that provide independent estimates $\hat\beta_1,\dots,\hat\beta_n$ of these effects,
- with corresponding (estimated) standard errors $s_1,\dots,s_n$. (See Discussion for ideas on how our methods may be applicable more generally; for example
- if, in place of $s_j$, we have a $p$ value testing $\beta_j=0$.)
+ with corresponding (estimated) standard errors $s_1,\dots,s_n$. For our purposes, these estimates and standard errors could be obtained from the data
+ using any method, provided they (approximately) satisfy 
+ \begin{equation} \label{eqn:betahat}
+ \hat\beta_j | \beta_j, s_j \sim N(\beta_j, s_j).
+ \end{equation}
+ [In some settings a $t$ distribution assumption for $\hat\beta_j | s_j, \beta_j$ may be more appropriate, and could also be incorporated into our methods; see Discussion.]
+ %(See Discussion for ideas on how our methods may be applicable more generally; for example if, in place of $s_j$, we have a $p$ value testing $\beta_j=0$.)
 
  In outline, we use a hierarchical model to combine information across measurements, with the aim of improving both accuracy and precision of estimates. 
- Specifically we assume that the effects $\beta_j$ are independent and identically distributed from some unknown distribution $g(\cdot)$.  
- Our methods are based on the following key assumptions:
+ Specifically we assume that the effects $\beta_j$ are independent and identically distributed from some unknown distribution $g(\cdot)$,
+ \begin{equation} \label{eqn:beta}
+ \beta_j \sim g(\cdot),
+ \end{equation}
+ where $g$ will be estimated from
+ the data.  Our methods make the following key assumptions, which distinguish them from most previous work in this area:
    \begin{enumerate}
-  \item The distribution $g(\cdot)$ is unimodal. (We assume here that the mode is at 0, although 
+  \item[A1:] The distribution $g(\cdot)$ is unimodal. (We assume here that the mode is at 0, although 
 this could be relaxed.)  
- \item The likelihood $L(\beta_j) = p(D_j | \beta_j)$ can be approximated by a normal likelihood,
+ \item[A2:] The likelihood $L(\beta_j) := p(D_j | \beta_j)$ can be approximated by a normal likelihood,
  \begin{equation} \label{eqn:lik}
- L(\beta_j) \propto \exp[-0.5(\beta_j-\hat\beta_j)^2/s_j^2].
+ \hat{L}_j(\beta_j) \propto \exp[-0.5(\beta_j-\hat\beta_j)^2/s_j^2].
  \end{equation}
+%which, in light of (\ref{eqn:betahat}), is equivalent to
+ % \begin{equation} \label{eqn:lik2}
+% L(\beta_j) \propto p(\hat\beta_j | \beta_j, s_j).
+ %\end{equation}
  \end{enumerate}
-Although these assumptions will not apply to all situations, we argue that both will be reasonable in many practical contexts. 
-For example, in contexts where interest focusses on which $\beta_j$ differ from zero, and ``$\beta_j=0$" is a plausible null hypothesis,  
+
+ \subsubsection*{Discussion on Assumptions}
+
+Although the unimodal assumption, A1, will not apply to all situations, we argue that it will be reasonable in many practical contexts. 
+For example, in contexts where interest focusses on which $\beta_j$ differ from zero, which suggests that ``$\beta_j=0$" is a plausible null hypothesis,  
 it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible, and that the distribution of the effects will be unimodal about 0. 
-Alternatively, in an estimation context, we can motivate Assumption 1 by considering its effect on point estimates, which is to ``shrink" the estimates towards the mode;
-this kind of shrinkage is widely regarded as useful, and provides an alternative motivation for our methods.
-Assumption 2 can be motivated either as a direct approximation to the likelihood (e.g.~by Taylor series expansion of the log-likelihood about $\hat\beta_j$),
-or by assuming that $\hat\beta_j, s_j$ are sufficient statistics for $\beta_j$ [so $L(\beta_j) \propto p(\hat\beta_j, s_j | \beta)$], and assuming that  $\hat\beta_j | s_j, \beta_j \sim N(\beta_j,s_j)$ and that $p(s_j|\beta_j)$ does not depend on $\beta_j$.
-[In some settings a $t$ distribution assumption for $\hat\beta_j$ may be more appropriate; see Discussion.]
-Assumption 2 has obvious parallels with approaches like \cite{efron, cai?} that model the $Z$ scores $Z_j = \beta_j/s_j$ as normally distributed under the null. However,
-there is an important difference: in our approach observations with a large $s_j$ have an essentially flat likelihood \ref{eqn:lik}, and so do not affect inference for
+Alternatively, we can motivate A1 by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
+such shrinkage is desirable from several standpoints for improving estimation accuracy. 
+Interestingly, although the unimodal assumption for $g$ seems natural to us, much previous related work on estimating False Discovery Rates  
+has assumed, either explicitly or implicitly, that the effect sizes are multimodal. (In contrast, almost all analogous work in large-scale regression assumes
+unimodal distributions for the effect sizes, such as the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors.)
+We discuss the effects of these assumptions later.
+For now, it is important to note that assumption A1 relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
+It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and A1 does not contradict this.
+
+Assumption A2 can be motivated by assuming that a) $\hat\beta_j, s_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e. they are
+``approximately sufficient" for $D_j$) and that b) $s_j$ alone contains little information about $\beta$.
+The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\hat\beta_j, s_j | \beta) = p(\hat\beta | \beta, s_j) p(s_j | \beta)$, while the latter (b) suggests assuming that $p(s_j|\beta_j)$ does not depend on $\beta_j$.
+Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\hat\beta_j | \beta, s_j)$, which, from (\ref{eqn:betahat}),  implies A2.
+(If $\hat\beta_j, s_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\hat\beta_j,s_j$, rather than conditional on $D_j$.)
+
+While Assumption A2 has parallels with approaches like \cite{efron, cai?} that model the $Z$ scores $Z_j = \hat\beta_j/s_j$ as normally distributed under the null, 
+there is an important difference: here, observations with a large $s_j$ have an essentially flat likelihood (\ref{eqn:lik}), and so do not affect inference for
 other parameters, whereas when modeling the $Z$ scores directly, observations with large $s_j$ produce $N(0,1)$ $Z$ scores that do affect inference. In
@@
 
-Interestingly, although the unimodal assumption for $g$ seems natural to us, much previous related work on estimating False Discovery Rates  
-has tended to assume, either explicitly or implicitly, that the effect sizes are multimodal. (In contrast, almost all analogous work in large-scale regression assumes
-unimodal distributions for the effect sizes, such as the spike and slab, Laplace, NEG, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors.)
-We discuss the effects of these assumptions later.
-For now, it is important to note that Assumption 1 relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
-It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and Assumption 1 does not contradict this.
+\subsubsection*{Implementing the Unimodal assumption}
 
-In practice, we implement our approach by assuming a parametric finite mixture  for $g$, 
-\begin{equation}
-g(\cdot; \pi) = \pi_0 \delta_0(\cdot) + (1-\pi_0) \sum_{k=1}^K \pi_k f_k(\cdot) 
+In practice, we implement the unimodal assumption (A1) by assuming a parametric finite mixture for $g$, 
+\begin{equation} \label{eqn:g}
+g(\cdot; \pi) = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k f_k(\cdot) 
 \end{equation}
-[Or possibly remove $(1-\pi_0)$ from this?]
-where $\delta_0(\cdot)$ denotes a point mass at 0, $\pi=(\pi_0,\dots,pi_K)$ are mixture proportions to be estimated, and $f_k$ are pre-specified 
+where $\delta_0(\cdot)$ denotes a point mass at 0, $\pi=(\pi_0,\dots,\pi_K)$ are mixture proportions to be estimated (with $\pi_j \geq 0, \sum_j \pi_j=1$), and $f_k$ are pre-specified 
 component distributions with one of the following forms: 
 \begin{enumerate}[i)]
-\item $f_k(\cdot) = N(0, \sigma^2_k)$
-\item $f_k(\cdot) = U[-a_k,a_k]$
-\item $f_k(\cdot) = U[-a_k,0] \text{ or } U[0,a_k]$,
+\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$
+\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$
+\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ or } U[\cdot; 0,a_k]$,
 \end{enumerate}
-where $N(0,\sigma^2)$ denotes the normal distribution with mean 0 and variance $\sigma^2$ and
-$U[a,b]$ denotes the unform distribution on $[a,b]$.
+where $N(\cdot; 0,\sigma^2)$ denotes the density of a normal distribution with mean 0 and variance $\sigma^2$ and
+$U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
 That is, the non-zero effects are modeled as either i) a mixture of zero-centered normal distributions; ii) a mixture of zero-centered uniform distributions;
-or iii) a mixture of ``zero-anchored" uniform distributions.
+or iii) a mixture of ``zero-anchored" uniform distributions, which have one end of their range at 0. This approach closely mirrors approaches in \cite{campy.thomas}.
 
-Pre-specifying the distributions $f_k$, rather than estimating them, may seem both inelegant and restrictive. However,
- any inelegance is, in our opinion, more than off-set by the very considerable computational simplifications it produces when fitting these models (see below). 
- And any apparent restrictiveness
-is illusory: by making
-$K$ sufficiently large one can approximate any symmetric unimodal distribution arbitrarily closely using the uniforms in (ii),
-and approximate any (asymmetric) unimodal distribution using the uniforms in (iii); and even the more restrictive (i) is sufficiently flexible
-to be useful in practice. Indeed, (i) can approximate arbitrarily closely any scale mixture of normals, which includes as a special case
-the double exponential distribution, any $t$ distribution, and other distributions used in high-dimensional regression settings such as the horseshoe prior \ref{},
-all of which are more restrictive. Furthermore, the unimodal constraint is sufficiently strong to prevent serious problems with overfitting when using large $K$. Indeed, by 
-increasing $K$ one can ultimately approach the non-parametric maximum likelihood estimate for $g$.
-Related to this, see \cite{grenander}, which provides the (extremely elegant) solution to the non-parametric maximum likelihood estimate for $g$ in the limit as the errors $s_j$ all tend to zero;
-and see \cite{campy.thomas} who suggest similar methods to those used here in the case where $s_j$ are all equal.
+To see the flexibility of the representation (\ref{eqn:g})
+think of $K$ as being ``large", with the values of $\sigma^2_k$ or $a_k$ 
+spanning a wide range of values, from
+essentially zero to unnecessarily large. In this way, as $K$ increases, the three different choices for $f_k$ can allow $g$ to approximate,
+with arbitrary accuracy,
+\begin{enumerate}[i)]
+\item any scale mixture of normals, which includes as special cases
+the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in 
+high-dimensional regression settings such as the spike and slab prior, the horseshoe prior \ref{}, or the more general three-parameter beta prior \ref{}.
+\item any symmetric unimodal distribution about 0.
+\item any (asymmetric) unimodal distribution about 0.
+\end{enumerate}
+The latter two claims are related to characterizations of unimodal distributions due to Khintchine (1938) and  Shepp (1962); see Feller, 1971, page  158. 
+Furthermore, although it is natural for statisticians to immediately worry about ``overfitting" if $K$ is large, 
+in fact the unimodal constraint is sufficient to prevent serious problems with overfitting. For example, note that the unimodal
+constraint is sufficiently strong that there will exist a non-parametric
+maximum likelihood estimate (NPMLE) for $g$ under this constraint.
+(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $s_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander}.)
+Thus allowing $K$ to be large in ii) or iii) can be thought of as approximating the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
+
+Having emphasized the flexibility that can be achieved with large $K$, we note that in practice even quite small values of $K$ (e.g. 2-10) provide highly flexible
+and effective methods. See Section \ref{} for our software default choices of $K$ and $\sigma^2_k,a_k$.
 
@@
 
-Assuming independence of the data leading to $\hat\beta_1,\dots,\hat\beta_n$, the likelihood for $\pi$ can be written as 
+Assuming independence across units, the likelihood for $\pi$ is 
 \begin{equation}
-p(\hat\beta, s | \pi) = \prod_{j=1}^n p(\hat\beta_j, s_j | \pi).
+L(\pi) := p(D_1,\dots,D_J | \pi) = \prod_{j=1}^n p(D_j | \pi)
+= \prod_j \int p(D_j | \pi, \beta_j) p(\beta_j | \pi) d\beta_j \\
+= \prod_j \int p(D_j | \beta_j) g(\beta_j; \pi) d\beta_j
 \end{equation}
-Further, the assumptions  $\hat\beta_j | \beta, s_j \sim N(\beta,s_j)$ and $\beta_j | s_j, \pi \sim g(\cdot; \pi)$ yield a mixture form for $\hat\beta_j$: 
+
+Further, the assumptions  $\hat\beta_j | \beta_j, s_j \sim N(\beta_j,s_j)$ and $\beta_j | s_j, \pi \sim g(\cdot; \pi)$ (given by \ref{eqn:g}) yield: 
 \begin{equation}
-p(\hat\beta_j, s_j | \pi) = \pi_0 N(0,s^2_j) + \sum_k \pi_k \tilde{f}_{kj}(\hat\beta_j)
+p(\hat\beta_j, s_j | \pi) = \pi_0 N(\hat\beta_j; 0,s^2_j) + \sum_k \pi_k \tilde{f}_{kj}(\hat\beta_j)
 \end{equation}
 where $\tilde{f}_{kj}$ denotes the density of a convolution of $f_k$ with an $N(0,s^2_j)$ density.
-For example, if $f_k$ is $N(0,\sigma^2_k)$ then the convolution $\tilde{f}_{kj}$ is also normal, $\tilde{f}_{kj} = N(0,\sigma^2_k + s^2_j)$. If
-$f_k$ is $U[a_k,b_k]$ then 
+For example, if $f_k(\cdot) = N(\cdot; 0,\sigma^2_k)$ then the convolution $\tilde{f}_{kj}$ is also normal, $\tilde{f}_{kj}(\cdot) = N(\cdot; 0,\sigma^2_k + s^2_j)$. If
+$f_k(\cdot) = U[\cdot; a_k,b_k]$ then $\tilde{f}_{kj}$ is the convolution of a normal with a uniform, which has density
 \begin{equation}
-\tilde{f}_{kj}(x) = 
+\tilde{f}_{kj}(x) = (\Phi(b_k/s_j) - \Phi(a_k/s_j))/(b-a),
 \end{equation}
-by standard results on the convolution of a uniform with a normal.
+where $\Phi$ is the cumulative distribution function of the standard normal $N(0,1)$ distribution.
 
-\begin{equation}
-\end{equation}
 Putting these together, the joint model for the unobserved $\beta$ and the observed $\hat\beta, s$ is:
@@
 To illustrate estimation of $\pi_0$ we provide simulation results for two scenarios:
-\begin{itemize}
+\begin{enumerate}
 \item[Scenario 1:]
@@
 \end{equation}
-\end{itemize}
+\end{enumerate}
 Scenario 1 represents a ``difficult" case where many non-zero $\beta$ are
@@
 precision for each measurement. We applied the methods implemented in the R packages qvalue, fdrtool, locfdr, and mixfdr
-for estimating $\pi_0$. We also applied our shrinkage-based method to estimate $\pi_0$ in two ways, one using a ``prior" distribution to encourage
-$\pi_0$ to be as large as possible (``null-biased"), and the other using maximum likelihood estimation for $\pi$.
+for estimating $\pi_0$. We also applied our method to estimate $\pi_0$ in two ways, one using a penalty encourage
+$\pi_0$ to be as large as possible (``null-biased"), and the other using the maximum likelihood estimate for $\pi$.
 
 Figure \ref{fig:pi0} compares estimated and true values of $\pi_0$ under each Scenario. 
-For Scenario 1, none of the method reliably estimate $\pi_0$, as expected since the scenario was designed to make estimation
+For Scenario 1, no method reliably estimates $\pi_0$. This is expected since, the scenario was designed to make accurate estimation
 of $\pi_0$ impossible. However, we see that all the methods except for ash (mle) are able to provide a conservative
@@
 comparison with ash.
+
+\subsection{Comparison of different underlying component distributions}
+
+We simulated data under three scenarios for the underlying effect size distribution $g$:
+\begin{enumerate}
+\item[Scenario A:]
+A mixture of zero-centered normals,
+\begin{equation}
+g(\cdot) =  (2/3) N(\cdot; 0,1) + (1/3) N(\cdot; 0, 2^2)
+\end{equation}
+\item[Scenario B:]
+A symmetric ``flat-topped" distribution, designed to be difficult to capture by a mixture of zero-centered normals 
+\begin{equation}
+g(\cdot) = (1/7) [N(\cdot; -1.5, 0.5^2) + N(\cdot; -1, 0.5^2) + N(\cdot; 0.5, 0.5^2) + N(\cdot; 0, 0.5^2) + N(\cdot; 0.5, 0.5^2) + N(\cdot; 1, 0.5^2) + N(\cdot; 1.5, 0.5^2)]
+\end{equation}
+\item[Scenario C:]
+An asymmetric distribution, 
+\begin{equation}
+g(\cdot) = (1/4) N(\cdot; -2 ,2^2) + (1/4) N(\cdot; -1 ,1.5^2) + (1/3) N(\cdot; 0 ,1) + (1/6) N(\cdot; 1,1) 
+\end{equation}
+\end{enumerate}
+In each case we simulated $\hat\beta_j  = \beta_j + N(0,s^2_j=1)$.
+
+Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
+than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
+likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
+For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
+increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
+expect $L_{hu}$ to be significantly larger than the others.
+
+Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
+
+\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
+\begin{center}
+\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
+%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
+\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
+\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
+\end{center}
+\caption{}
+\end{figure}
+
+- how does use of the VB approach to estimate pi affect RMSE?
 

### 2013-12-15 paper/Rcode/ashsim.rmd
updated paper and code

@@
   betahat = list()
-  betahat.ash.n = list()
-  betahat.ash.u = list()
-  betahat.ash.hu = list()
+  fit.ash.n = list()
+  fit.ash.u = list()
+  fit.ash.hu = list()
+  fit.ash.hu.vb=list()
+  fit.ash.true = list()
+  fit.mixfdr= list()
+  fit.ash.fdr.n = list()
+
+  fit.mixfdr.enull= list()
+  fit.mixfdr.J10= list()
+  fit.mixfdr.J10P0= list()
+  fit.mixfdr.J100= list()
   k = length(mixmean)
@@
     betahat[[i]] = beta[[i]]+rnorm(nsamp,0,betahatsd[[i]])
-    betahat.ash.n[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="normal",method="shrink")
-    betahat.ash.u[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="uniform", method="shrink")
-    betahat.ash.hu[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform", method="shrink")
+    fit.ash.n[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="normal",method="shrink")
+    fit.ash.u[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="uniform", method="shrink")
+    fit.ash.hu[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform", method="shrink")
+    fit.ash.hu.vb[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform",method="shrink",VB=TRUE)
+    fit.ash.true[[i]] = ash(betahat[[i]],betahatsd[[i]],g=normalmix(mixpi,mixmean,mixsd))
+    fit.mixfdr[[i]] = mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=TRUE,plot=FALSE)
 
+    fit.ash.fdr.n[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="normal",method="fdr")
+    #fit.ash.fdr.u[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="uniform", method="fdr")
+    #fit.ash.fdr.hu[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform", method="fdr")
+    fit.mixfdr.enull[[i]] = mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=FALSE,plot=FALSE)
+    fit.mixfdr.J10[[i]] = mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=TRUE,plot=FALSE,J=10)
+     #fit.mixfdr.J100[[i]] = mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=TRUE,plot=FALSE,J=100)
+    fit.mixfdr.J10P0[[i]] = try(mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=TRUE,plot=FALSE,J=10,P=0))
   }
@@
   betahat = betahat,
-  betahat.ash.n = betahat.ash.n,
-  betahat.ash.u = betahat.ash.u,
-  betahat.ash.hu = betahat.ash.hu))
+  fit.ash.n = fit.ash.n,
+  fit.ash.u = fit.ash.u,
+  fit.ash.hu = fit.ash.hu,
+  fit.ash.hu.vb=fit.ash.hu.vb,
+  fit.mixfdr = fit.mixfdr,
+  fit.ash.fdr.n = fit.ash.fdr.n,
+  fit.ash.true=fit.ash.true,
+ # fit.ash.fdr.u = fit.ash.fdr.u,
+ # fit.ash.fdr.hu = fit.ash.fdr.hu,
+  fit.mixfdr.enull = fit.mixfdr.enull,
+  fit.mixfdr.J10 = fit.mixfdr.J10,
+  #fit.mixfdr.J100= fit.mixfdr.J100,
+  fit.mixfdr.J10P0=fit.mixfdr.J10P0))
 }    
@@
 
-```{r}
+```{r, cache=TRUE}
 sim1= ashsim(c(0,0,0),c(1,1,2),c(1/3,1/3,1/3),niter=20,nsamp=1000)
@@
 
+Experiment with mixFdr to see how penalization effects fit. 
 ```{r}
+  betahat = sim1$betahat[[1]]
+  betahatsd = sim1$betahatsd[[1]]
+  set.seed(111) #to ensure reproducibility
+  fit.mixfdr.P0 = mixFdr(betahat/betahatsd,noiseSD=1,theonull=TRUE,P=0,plot=FALSE)
+  cdf.mixfdr = function(a,x){mixcdf(normalmix(a$pi,a$mu,a$sigma-1),x)}
+  x = seq(-4,4,length=100)
+
+  res = data.frame(beta=x,
+                  mixfdr.P0 = t(cdf.mixfdr(fit.mixfdr.P0,x)),
+                  mixfdr = t(cdf.mixfdr(sim1$fit.mixfdr[[1]],x)),
+                  ash.n = t(cdf.ash(sim1$fit.ash.n[[1]],x)$y),
+                  ash.fdr.n=t(cdf.ash(sim1$fit.ash.fdr.n[[1]],x)$y))
+
+  truth = t(cdf.ash(sim1$fit.ash.true[[1]],x)$y)
+
+  res.melt=melt(res, id.vars=c("beta"),variable.name="Method")   
+  res.melt$Penalization= ifelse((res.melt$Method=="mixfdr" | res.melt$Method=="ash.fdr.n"),"Default","Minimal")
+  res.melt$Methodtype=as.character(res.melt$Method)
+  res.melt$Methodtype[res.melt$Method=="mixfdr.P0"]="mixfdr"
+  res.melt$Methodtype[res.melt$Method=="ash.fdr.n"]="ash"
+  res.melt$Methodtype[res.melt$Method=="ash.n"]="ash"
+
+cbbPalette <- c("#56B4E9", "#009E73", "#000000","#F0E442", "#0072B2", "#E69F00",  "#D55E00", "#CC79A7")
+pdf("figures/penaltycompare.pdf")
+ggplot(res.melt, aes(x=beta)) +
+  geom_line(aes(x=beta,y=value,color=Methodtype,linetype=Penalization,group=Method),size=1.5,alpha=0.8) +
+  geom_line(data=res, aes(y=truth,color="truth"),alpha=1,size=0.5) +
+  scale_colour_manual(name = 'Method', values = cbbPalette)
+dev.off()
+```
+
+```{r}
+plot_examples = function(sims){
+  sim1= sims[[1]]
+  sim2= sims[[2]]
+  sim3= sims[[3]]
 len=length(sim1$beta[[1]])
@@
 y = dnorm(x)
-dens1 = data.frame(x=x,dn= t(density(sim1$betahat.ash.n[[1]],x)$y),du=t(density(sim1$betahat.ash.u[[1]],x)$y),dhu=t(density(sim1$betahat.ash.hu[[1]],x)$y), Scenario=1)
-dens2 = data.frame(x=x,dn = t(density(sim2$betahat.ash.n[[1]],x)$y),du=t(density(sim2$betahat.ash.u[[1]],x)$y), dhu=t(density(sim2$betahat.ash.hu[[1]],x)$y), Scenario=2)
-dens3 = data.frame(x=x,dn = t(density(sim3$betahat.ash.n[[1]],x)$y),du=t(density(sim3$betahat.ash.u[[1]],x)$y), dhu=t(density(sim3$betahat.ash.hu[[1]],x)$y), Scenario=3)
+dens1 = data.frame(x=x,dtrue=t(density(sim1$fit.ash.true[[1]],x)$y), Scenario=1)
+dens2 = data.frame(x=x,dtrue=t(density(sim2$fit.ash.true[[1]],x)$y), Scenario=2)
+dens3 = data.frame(x=x,dtrue=t(density(sim3$fit.ash.true[[1]],x)$y), Scenario=3)
 dens=rbind(dens1,dens2,dens3)
-ggplot(res, aes(x=beta)) + facet_grid(.~Scenario)  +
-    geom_histogram(aes(y=..density..),binwidth=0.5,
-                   colour="black", fill="white") +  
-  geom_line(data = dens, aes(x,du,color='Uniform'),size=1.2,alpha=0.8) +
-  geom_line(data = dens, aes(x,dhu,color='Half Uniform'),size=1.2,alpha=0.8) +
-    geom_line(data = dens, aes(x,dn,color='Normal'),size=1.2,alpha=0.8) +
-  scale_colour_manual(name = 'Method', values = c('red', 'blue','green'))
+ggplot(res) + facet_grid(.~Scenario)  + 
+    geom_line(data = dens, aes(x,dtrue),size=1.2,alpha=0.9,linetype=1) + scale_x_continuous(limits=c(-6,6))
+}
 
+pdf("figures/simABC_egdens.pdf",width=6.5,height=2)
+plot_examples(list(sim1,sim2,sim3))
+dev.off()
 ```
@@
     res=list()
-    res2=list()
+
     for(i in 1:length(sims)){
-      err.ash.n= mapply(get_rmse.ash,sims[[i]]$betahat.ash.n,sims[[i]]$beta)
-      err.ash.u = mapply(get_rmse.ash,sims[[i]]$betahat.ash.u,sims[[i]]$beta)
-      err.ash.hu = mapply(get_rmse.ash,sims[[i]]$betahat.ash.hu,sims[[i]]$beta)
-    #err.mixfdr=mapply(get_rmse.mixfdr,sims$betahat.mixfdr,sims$beta)
+      err.bayes = mapply(get_rmse.ash,sims[[i]]$fit.ash.true,sims[[i]]$beta)
+      err.ash.n= mapply(get_rmse.ash,sims[[i]]$fit.ash.n,sims[[i]]$beta)
+      err.ash.u = mapply(get_rmse.ash,sims[[i]]$fit.ash.u,sims[[i]]$beta)
+      err.ash.hu = mapply(get_rmse.ash,sims[[i]]$fit.ash.hu,sims[[i]]$beta)
+      err.mixfdr=mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr,sims[[i]]$beta)
+      err.mixfdr.enull = mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr,sims[[i]]$beta)
+       err.mixfdr.J10 = mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr.J10,sims[[i]]$beta)
+       #err.mixfdr.J100 = mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr.J100,sims[[i]]$beta)
+      err.ash.fdr.n = mapply(get_rmse.ash,sims[[i]]$fit.ash.fdr.n,sims[[i]]$beta)
+
       err.betahat = mapply(rmse,sims[[i]]$betahat,sims[[i]]$beta)
       err.zero = unlist(lapply(sims[[i]]$beta,rmse,y=0))        
-      loglik.ash.n = unlist(lapply(sims[[i]]$betahat.ash.n,get_loglik))
-      loglik.ash.u = unlist(lapply(sims[[i]]$betahat.ash.u,get_loglik))
-      loglik.ash.hu = unlist(lapply(sims[[i]]$betahat.ash.hu,get_loglik))
-     res[[i]] = data.frame(Scenario=i,ash.normal=err.ash.n,ash.uniform=err.ash.u,ash.halfuniform=err.ash.hu)
-      res2[[i]] = data.frame(Scenario=i,ash.normal=loglik.ash.n,ash.uniform=loglik.ash.u,ash.halfuniform=loglik.ash.hu)
-
+      res[[i]] = data.frame(Scenario=i,bayes=err.bayes,ash.normal=err.ash.n,ash.uniform=err.ash.u,ash.halfuniform=err.ash.hu,mixfdr=err.mixfdr,
+                            mixfdr.enull = err.mixfdr.enull,
+                            mixfdr.J10 = err.mixfdr.J10,
+                            #mixfdr.J100 = err.mixfdr.J100,
+                            ash.normal.fdr = err.ash.fdr.n)
       if(inczero){
-        res[[i]]=data.frame(res,zero=err.zero)
+        res[[i]]=data.frame(res[[i]],zero=err.zero)
       }
       if(incbetahat){
-        res[[i]]=data.frame(res,betahat=err.betahat)
+        res[[i]]=data.frame(res[[i]],betahat=err.betahat)
       }
@@
     require(reshape2)
-    res.melt = melt(res, id.vars=c("Scenario"),variable.name="Method")
-    res2.melt = melt(res2, id.vars=c("Scenario"),variable.name="Method")
-    ggplot(data=res.melt,aes(Method,value)) +geom_boxplot()+         facet_grid(. ~ Scenario)
-     ggplot(data=res2.melt,aes(Method,value)) +geom_boxplot()+         facet_grid(. ~ Scenario)
-
-    p=ggplot(data=res.melt,aes(ash.halfuniform,value,colour=Method)) +geom_point(shape=16) +
-         facet_grid(. ~ Scenario) +
+    res.melt = melt(res, id.vars=c("Scenario","bayes"),variable.name="Method")    
+    p=ggplot(data=res.melt,aes(bayes,value,colour=Method)) +geom_point(shape=16) +
+         facet_grid(. ~ Scenario,scale="free_x") +
         geom_abline(colour = "black") +
-        xlab("RMSE (ash.halfuniform)") +
+        xlab("RMSE (Optimal Bayes Rule)") +
         ylab("RMSE (other method)")
@@
   }
-  #pdf("figures/rmse_sim1.pdf")
+  pdf("figures/rmse_biplot.pdf",width=6.5,height=2)
     plot_rmse(list(sim1,sim2,sim3))
- #               dev.off()
-  #pdf("figures/rmse_sim2.pdf")
-  #  plot_rmse(simres2)
-  #dev.off()
-```
+  dev.off()
 
-Try plotting boxplots of rmse and fit impovement compared with most general model
-```{r}
-  rmse = function(x,y){sqrt(mean((x-y)^2))}
-  get_rmse.ash = function(a,b){rmse(a$PosteriorMean,b)}
-  get_rmse.mixfdr = function(a,b){rmse(a$effectSize,b)}
-  boxplot_rmse_loglik = function(sims,inczero=FALSE,incbetahat=FALSE){
-    res=list()
-    res2=list()
-    for(i in 1:length(sims)){
-      err.ash.hu = mapply(get_rmse.ash,sims[[i]]$betahat.ash.hu,sims[[i]]$beta)
-      err.ash.n= mapply(get_rmse.ash,sims[[i]]$betahat.ash.n,sims[[i]]$beta)-err.ash.hu
-      err.ash.u = mapply(get_rmse.ash,sims[[i]]$betahat.ash.u,sims[[i]]$beta)-err.ash.hu
-      #err.mixfdr=mapply(get_rmse.mixfdr,sims$betahat.mixfdr,sims$beta)
-      loglik.ash.hu = unlist(lapply(sims[[i]]$betahat.ash.hu,get_loglik))
-      loglik.ash.n = unlist(lapply(sims[[i]]$betahat.ash.n,get_loglik))-loglik.ash.hu
-      loglik.ash.u = unlist(lapply(sims[[i]]$betahat.ash.u,get_loglik))-loglik.ash.hu
-
-     res[[i]] = data.frame(Scenario=i,ash.normal.err=err.ash.n,ash.uniform.err=err.ash.u)
-       res2[[i]] = data.frame(Scenario=i,ash.normal=loglik.ash.n,ash.uniform=loglik.ash.u)
-
-    }
-    require(reshape2)
-    res.melt = melt(res, id.vars=c("Scenario"))
-
-    res2.melt=melt(res2,id.vars=c("Scenario"))
-    ggplot(res.melt,aes(variable,value)) + geom_boxplot() + facet_grid(.~Scenario)
-    ggplot(res2.melt,aes(variable,value)) + geom_boxplot() + facet_grid(.~Scenario)
-  }
-
-  pdf("figures/boxplot_rmse_loglik.pdf")
-    boxplot_rmse_loglik(list(sim1,sim2,sim3))
+#This figure not used in paper?
+pdf("figures/rmse_biplot_withzerobetahat.pdf")
+    plot_rmse(list(sim1,sim2,sim3),inczero=TRUE,incbetahat=TRUE)
   dev.off()
@@
 ```{r}
+  get_rmse.ash = function(a,b){rmse(a$PosteriorMean,b)}
+  get_rmse.mixfdr = function(a,b){rmse(a$effectSize,b)}
+  plot_rmse_boxplot = function(sims,inczero=FALSE,incbetahat=FALSE,incmixfdr=FALSE){
+    res=list()
+
+    for(i in 1:length(sims)){
+      err.bayes = mapply(get_rmse.ash,sims[[i]]$fit.ash.true,sims[[i]]$beta)
+      err.ash.n= mapply(get_rmse.ash,sims[[i]]$fit.ash.n,sims[[i]]$beta)
+      err.ash.u = mapply(get_rmse.ash,sims[[i]]$fit.ash.u,sims[[i]]$beta)
+      err.ash.hu = mapply(get_rmse.ash,sims[[i]]$fit.ash.hu,sims[[i]]$beta)
+      err.mixfdr=mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr,sims[[i]]$beta)
+      err.mixfdr.enull = mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr,sims[[i]]$beta)
+       err.mixfdr.J10 = mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr.J10,sims[[i]]$beta)
+#       err.mixfdr.J100 = mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr.J100,sims[[i]]$beta)
+      err.ash.fdr.n = mapply(get_rmse.ash,sims[[i]]$fit.ash.fdr.n,sims[[i]]$beta)         
+      err.betahat = mapply(rmse,sims[[i]]$betahat,sims[[i]]$beta)
+      err.zero = unlist(lapply(sims[[i]]$beta,rmse,y=0))        
+      res[[i]] = data.frame(Scenario=i,
+                            ash.normal=err.ash.n/err.bayes,
+                            ash.uniform=err.ash.u/err.bayes,
+                            ash.halfuniform=err.ash.hu/err.bayes,
+                            ash.normal.fdr = err.ash.fdr.n/err.bayes
+                            )
+      if(incmixfdr){
+        res[[i]]=data.frame(res[[i]],mixfdr=err.mixfdr/err.bayes,
+                            mixfdr.enull = err.mixfdr.enull/err.bayes,
+                            mixfdr.J10 = err.mixfdr.J10/err.bayes)
+                            #mixfdr.J100 = err.mixfdr.J100/err.bayes)
+      }
+      if(inczero){
+        res[[i]]=data.frame(res[[i]],zero=err.zero/err.bayes)
+      }
+      if(incbetahat){
+        res[[i]]=data.frame(res[[i]],betahat=err.betahat/err.bayes)
+      }
+    }
+    require(reshape2)
+    res.melt = melt(res, id.vars=c("Scenario"),variable.name="Method")    
+    ggplot(res.melt,aes(Method,value,color=Method)) + geom_boxplot() + facet_grid(.~Scenario)
+
+  }
+  pdf("figures/rmse_boxplot.pdf",width=6.5,height=2)
+    plot_rmse_boxplot(list(sim1,sim2,sim3))
+  dev.off()
+
+    pdf("figures/rmse_boxplot_extended.pdf",width=6.5,height=2)
+    plot_rmse_boxplot(list(sim1,sim2,sim3),TRUE,TRUE,TRUE)
+  dev.off()
+
+```
+
+```{r}
+  plot_loglik_boxplot = function(sims){
+    res=list()
+    for(i in 1:length(sims)){
+      loglik.bayes = mapply(get_loglik,sims[[i]]$fit.ash.true)
+      loglik.ash.n= mapply(get_loglik,sims[[i]]$fit.ash.n)
+      loglik.ash.u = mapply(get_loglik,sims[[i]]$fit.ash.u)
+      loglik.ash.hu = mapply(get_loglik,sims[[i]]$fit.ash.hu)
+
+      loglik.ash.fdr.n = mapply(get_loglik,sims[[i]]$fit.ash.fdr.n)         
+      res[[i]] = data.frame(Scenario=i,
+                            ash.normal=loglik.ash.n-loglik.bayes,
+                            ash.uniform=loglik.ash.u-loglik.bayes,
+                            ash.halfuniform=loglik.ash.hu-loglik.bayes,
+                            ash.normal.fdr = loglik.ash.fdr.n-loglik.bayes)
+
+    }
+    require(reshape2)
+    res.melt = melt(res, id.vars=c("Scenario"),variable.name="Method")    
+    ggplot(res.melt,aes(Method,value,color=Method)) + geom_boxplot() + facet_grid(.~Scenario)
+
+  }
+  pdf("figures/loglik_boxplot.pdf",width=6.5,height=2)
+    plot_loglik_boxplot(list(sim1,sim2,sim3))
+  dev.off()
+
+```
+
+```{r}
+  get_rmse.ash = function(a,b){rmse(a$PosteriorMean,b)}
+  get_rmse.mixfdr = function(a,b){rmse(a$effectSize,b)}
+  get_loglik.mixfdr = function(a,betahat,betahatsd){loglik_conv(normalmix(a$pi,a$mu,a$sigma-1),betahat,betahatsd)}
+  cdf.mixfdr = function(a,x){mixcdf(normalmix(a$pi,a$mu,a$sigma-1),x)}
+  plot_rmse_loglik_boxplot = function(sims){
+    res=list()
+
+    for(i in 1:length(sims)){
+      err.bayes = mapply(get_rmse.ash,sims[[i]]$fit.ash.true,sims[[i]]$beta)
+      err.ash.n= mapply(get_rmse.ash,sims[[i]]$fit.ash.n,sims[[i]]$beta)
+      err.ash.u = mapply(get_rmse.ash,sims[[i]]$fit.ash.u,sims[[i]]$beta)
+      err.ash.hu = mapply(get_rmse.ash,sims[[i]]$fit.ash.hu,sims[[i]]$beta)
+      err.mixfdr=mapply(get_rmse.mixfdr,sims[[i]]$fit.mixfdr,sims[[i]]$beta)
+      err.ash.fdr.n = mapply(get_rmse.ash,sims[[i]]$fit.ash.fdr.n,sims[[i]]$beta)
+
+      res[[i]] = data.frame(Scenario=i,
+                            ash.normal=err.ash.n/err.bayes,
+                            ash.uniform=err.ash.u/err.bayes,
+                            ash.halfuniform=err.ash.hu/err.bayes,
+                            ash.normal.fdr = err.ash.fdr.n/err.bayes)
+
+    }
+
+     res2=list()
+    for(i in 1:length(sims)){
+      loglik.bayes = mapply(get_loglik,sims[[i]]$fit.ash.true)
+      loglik.ash.n= mapply(get_loglik,sims[[i]]$fit.ash.n)
+      loglik.ash.u = mapply(get_loglik,sims[[i]]$fit.ash.u)
+      loglik.ash.hu = mapply(get_loglik,sims[[i]]$fit.ash.hu)
+      loglik.mixfdr=mapply(get_loglik.mixfdr,sims[[i]]$fit.mixfdr,sims[[i]]$betahat,sims[[i]]$betahatsd)
+      loglik.ash.fdr.n = mapply(get_loglik,sims[[i]]$fit.ash.fdr.n)         
+      res2[[i]] = data.frame(Scenario=i,
+                            ash.normal=loglik.ash.n-loglik.bayes,
+                            ash.uniform=loglik.ash.u-loglik.bayes,
+                            ash.halfuniform=loglik.ash.hu-loglik.bayes,
+                            ash.normal.fdr = loglik.ash.fdr.n-loglik.bayes)
+
+      }
+
+    require(reshape2)
+    res.melt = melt(res, id.vars=c("Scenario"),variable.name="Method")    
+    res2.melt = melt(res2, id.vars=c("Scenario"),variable.name="Method")    
+    res.melt$type='RMSE (vs Bayes Rule)'
+    res2.melt$type='log(likelihood) (vs Bayes Rule)'
+  ggplot(rbind(res.melt,res2.melt),aes(Method,value)) + geom_boxplot() + facet_grid(type~Scenario,scale="free_y")
+
+  }
+  pdf("figures/rmse_loglik_boxplot.pdf",width=6.5,height=5)
+    plot_rmse_loglik_boxplot(list(sim1,sim2,sim3))
+  dev.off()
+
+```
+
+looking to see why mixfdr gives worse rmse, even though loglik looks ok.
+Perhaps it overshrinks?
+```{r}
+ plot(sim3$fit.ash.hu[[1]]$PosteriorMean, sim3$fit.mixfdr[[1]]$effectSize)
+ plot(sim3$fit.ash.fdr.n[[1]]$PosteriorMean, sim3$fit.mixfdr[[1]]$effectSize)
+```
+
+problem seems to be that it overestimates pi0
+```{r}
+ sim3$fit.mixfdr[[1]]$pi
+```
+[1] 0.71660654 0.08298064 0.20041282
+
+```{r}
+temp=mixFdr(sim3$betahat[[1]]/sim3$betahatsd[[1]],noiseSD=1,theonull=TRUE,calibrate=TRUE)
+temp2=mixFdr(sim3$betahat[[1]]/sim3$betahatsd[[1]],noiseSD=1,theonull=TRUE,J=10,plot=FALSE)
+plot(ecdf(sim3$beta[[1]]),xlim=c(-6,6))
+lines(x,cdf.mixfdr(temp,x),col=3)
+lines(cdf.ash(sim3$fit.ash.n[[1]],x),col=2)
+lines(cdf.ash(sim3$fit.ash.hu[[1]],x),col=2,lty=2)
+```
+
+```{r}
   plot_LR=function(sims){
-    hist(unlist(lapply(sims$betahat.ash.u,get_loglik))-unlist(lapply(sims$betahat.ash.n,get_loglik)), xlab="loglik difference", main="loglik differences for nullbiased prior vs mle",nclass=10)
+    hist(unlist(lapply(sims$fit.ash.u,get_loglik))-unlist(lapply(sims$fit.ash.n,get_loglik)), xlab="loglik difference", main="loglik differences for nullbiased prior vs mle",nclass=10)
   }
@@
   pdf("figures/logLR.pdf")
-    plot_LR(simres1)
-    plot_LR(simres2)
+    plot_LR(sim1)
+    plot_LR(sim2)
+  dev.off()
+```
+
+## Unused figures?
+
+```{r}
+plot_examples_withfit = function(sims){
+  sim1= sims[[1]]
+  sim2= sims[[2]]
+  sim3= sims[[3]]
+len=length(sim1$beta[[1]])
+res = data.frame(beta = c(sim1$beta[[1]],sim2$beta[[1]],sim3$beta[[1]]), Scenario = c(rep(1,len),rep(2,len),rep(3,len)))
+x = seq(-8,6,length=100)
+y = dnorm(x)
+dens1 = data.frame(x=x,dn= t(density(sim1$fit.ash.n[[1]],x)$y),du=t(density(sim1$fit.ash.u[[1]],x)$y),dhu=t(density(sim1$fit.ash.hu[[1]],x)$y), Scenario=1)
+dens2 = data.frame(x=x,dn = t(density(sim2$fit.ash.n[[1]],x)$y),du=t(density(sim2$fit.ash.u[[1]],x)$y), dhu=t(density(sim2$fit.ash.hu[[1]],x)$y), Scenario=2)
+dens3 = data.frame(x=x,dn = t(density(sim3$fit.ash.n[[1]],x)$y),du=t(density(sim3$fit.ash.u[[1]],x)$y), dhu=t(density(sim3$fit.ash.hu[[1]],x)$y), Scenario=3)
+dens=rbind(dens1,dens2,dens3)
+cbbPalette <- c("#56B4E9", "#009E73", "#F0E442", "#0072B2", "#E69F00", "#000000", "#D55E00", "#CC79A7")
+ggplot(res, aes(x=beta)) + facet_grid(.~Scenario)  +
+    geom_histogram(aes(y=..density..),binwidth=0.5,
+                   colour="black", fill="white") +  
+  geom_line(data = dens, aes(x,du,color='Uniform'),size=1.2,alpha=0.9,linetype=1) +
+  geom_line(data = dens, aes(x,dhu,color='Half Uniform'),size=1.2,alpha=0.9,linetype=1) +
+    geom_line(data = dens, aes(x,dn,color='Normal'),size=1.2,alpha=0.9,linetype=1) +
+  scale_colour_manual(name = 'Method', values = cbbPalette) + scale_x_continuous(limits=c(-4,4))
+}
+
+pdf("figures/simABC_eg_withfit.pdf",width=6.5,height=2)
+plot_examples_withfit(list(sim1,sim2,sim3))
 dev.off()
@@
 ```{r}
-  lapply()
+cdf.mixfdr = function(a,x){mixcdf(normalmix(a$pi,a$mu,a$sigma-1),x)}
+plot_examples_cdf_withfit = function(sims){
+  sim1= sims[[1]]
+  sim2= sims[[2]]
+  sim3= sims[[3]]
+len=length(sim1$beta[[1]])
+res = data.frame(beta = c(sim1$beta[[1]],sim2$beta[[1]],sim3$beta[[1]]), Scenario = c(rep(1,len),rep(2,len),rep(3,len)))
+x = seq(-6,6,length=100)
+y = dnorm(x)
+cdf1 = data.frame(x=x,dn=t(cdf.ash(sim1$fit.ash.n[[1]],x)$y),du=t(cdf.ash(sim1$fit.ash.u[[1]],x)$y),dhu=t(cdf.ash(sim1$fit.ash.hu[[1]],x)$y), dtrue=t(cdf.ash(sim1$fit.ash.true[[1]],x)$y), dn.fdr=t(cdf.ash(sim1$fit.ash.fdr.n[[1]],x)$y),dm = t(cdf.mixfdr(sim1$fit.mixfdr[[1]],x)), Scenario=1)
+cdf2 = data.frame(x=x,dn=t(cdf.ash(sim2$fit.ash.n[[1]],x)$y),du=t(cdf.ash(sim2$fit.ash.u[[1]],x)$y),dhu=t(cdf.ash(sim2$fit.ash.hu[[1]],x)$y), dtrue=t(cdf.ash(sim2$fit.ash.true[[1]],x)$y),dn.fdr=t(cdf.ash(sim2$fit.ash.fdr.n[[1]],x)$y),dm = t(cdf.mixfdr(sim2$fit.mixfdr[[1]],x)),Scenario=2)
+  cdf3 = data.frame(x=x,dn=t(cdf.ash(sim3$fit.ash.n[[1]],x)$y),du=t(cdf.ash(sim3$fit.ash.u[[1]],x)$y),dhu=t(cdf.ash(sim3$fit.ash.hu[[1]],x)$y), dtrue=t(cdf.ash(sim3$fit.ash.true[[1]],x)$y),dn.fdr=t(cdf.ash(sim3$fit.ash.fdr.n[[1]],x)$y),dm = t(cdf.mixfdr(sim3$fit.mixfdr[[1]],x)),Scenario=3)
+
+cdf=rbind(cdf1,cdf2,cdf3)
+cbbPalette <- c("#56B4E9", "#009E73","#F0E442",  "#000000","#0072B2", "#E69F00", "#000000", "#D55E00", "#CC79A7")
+ggplot(res, aes(x=beta)) + facet_grid(.~Scenario)  +
+  geom_line(data = cdf, aes(x,dtrue,color='True'),size=1.5,alpha=0.9,linetype=1) +
+  geom_line(data = cdf, aes(x,du,color='Uniform'),size=1,alpha=0.9,linetype=2) +
+  geom_line(data = cdf, aes(x,dhu,color='Half Uniform'),size=1,alpha=0.9,linetype=2) +
+    geom_line(data = cdf, aes(x,dn,color='Normal'),size=1,alpha=0.9,linetype=2) +
+    geom_line(data = cdf, aes(x,dn.fdr,color='Normal, null-biased'),size=1,alpha=0.9,linetype=2) +
+    geom_line(data = cdf, aes(x,dm,color='mixfdr'),size=1,alpha=0.9,linetype=2) +
+  scale_colour_manual(name = 'Method', values = cbbPalette) + scale_x_continuous(limits=c(-6,6))
+}
+
+pdf("figures/simABC_eg_cdf_withfit.pdf",width=6.5,height=2)
+plot_examples_cdf_withfit(list(sim1,sim2,sim3))
+dev.off()
 ```

### 2014-02-14 paper/main.tex
minor updates

@@
 
-Efron (2008) states the Zero Assumption as the assumption that ``most of the z-values near zero come from null genes".
-His main aim in making this assumption is to estimate an empirical null though (not assume N(0,1) for the null)
-rather than to impose identifiability.
-
-By modeling the z scores directly under the alternative, rather than the z scores as being the truth plus noise,
-maybe that is what is most problematic? Because the resulting distribution of z scores is just not possible.
-
-Note that \cite{muralidharan:2012} models $z$ scores as something plus noise under both $H_0$ and $H_1$, which
-avoids this problem. (DOes the same maybe apply to modeling beta, rather than z scores, when the errors vary?)
-
-Rice and Spiegelhalter - BRCA data?
-
-In practice, we currently fix the number of components $K$ to be large, and take the variances $\sigma_1<\sigma_2<\dots<\sigma_K$ to be fixed, and vary from very small to very large --  sufficiently large that typically $\hat\pi_K \approx 0$.
-
 %A Taylor series expansion of the log likelihood around the maximum likelihood estimate $\hat\beta_j$  yields $$l(\beta_j) \approx l(\hat\beta_j) + 0.5* (\beta_j - \hat\beta_j)^2 l''(\hat\beta_j).$$ [Fill in details?]
-
-Using a mixture of normal distributions for $g$ 
-has the advantage that,
-when combined with the normal likelihood, they give
-an analytic form for the conditional distribution $p(\beta_j | \hat\pi, \hat\beta_j, s_j)$ (also a mixture of normals).
 
@@
 
-The local False Discovery Rate (lfdr) for observation $j$ is
+The local False Discovery Rate (lfdr) for observation $j$ is defined \cite{efron} as
 \begin{equation}
-\lfdr_j = p(\beta_j = 0 | \hat\beta, s).
+\lfdr_j = p(\beta_j = 0 | \hat\pi, \hat\beta, s).
 \end{equation}
@@
 inference objective is to ask which $\beta_j$ are positive and which are negative. Indeed, even in settings where some
-$\beta_j$ are exactly zero, it could be argued that identifying which are positive and which negative is
-fundamentally more interesting and useful than identifying which are non-zero. For example, when identifying differentially expressed genes,
+$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
 analysts will often separate the genes that are ``upregulated" in one condition from those that ``downregulated" in that condition.
@@
 \begin{equation}
-\lfsr_j := \min[ p(\beta_j \geq 0| \hat\beta, s), p(\beta_j \leq 0| \hat\beta, s) ].
+\lfsr_j := \min[ p(\beta_j \geq 0| \hat\beta, s), p(\beta_j \leq 0| \hat\pi, \hat\beta, s) ].
 \end{equation}
@@
 
+\subsection{Discussion}
+
+Efron (2008) states the Zero Assumption as the assumption that ``most of the z-values near zero come from null genes".
+His main aim in making this assumption is to estimate an empirical null though (not assume N(0,1) for the null)
+rather than to impose identifiability.
+
+By modeling the z scores directly under the alternative, rather than the z scores as being the truth plus noise,
+maybe that is what is most problematic? Because the resulting distribution of z scores is just not possible.
+
+Note that \cite{muralidharan:2012} models $z$ scores as something plus noise under both $H_0$ and $H_1$, which
+avoids this problem. (DOes the same maybe apply to modeling beta, rather than z scores, when the errors vary?)
+
+Rice and Spiegelhalter - BRCA data?
+
 % Do NOT remove this, even if you are not including acknowledgments

### 2014-02-15 paper/main.tex
minor updates

--- 2013-12-15 paper/main.tex
@@
 such shrinkage is desirable from several standpoints for improving estimation accuracy. 
-Interestingly, although the unimodal assumption for $g$ seems natural to us, much previous related work on estimating False Discovery Rates  
-has assumed, either explicitly or implicitly, that the effect sizes are multimodal. (In contrast, almost all analogous work in large-scale regression assumes
-unimodal distributions for the effect sizes, such as the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors.)
-We discuss the effects of these assumptions later.
-For now, it is important to note that assumption A1 relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
+Note that assumption A1 relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
 It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and A1 does not contradict this.
 
-Assumption A2 can be motivated by assuming that a) $\hat\beta_j, s_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e. they are
-``approximately sufficient" for $D_j$) and that b) $s_j$ alone contains little information about $\beta$.
+Interestingly, although the unimodal assumption A1 seems natural, much previous related work on estimating False Discovery Rates  
+has assumed, either explicitly or implicitly, that the effect sizes are multimodal (see Results for more details). 
+In contrast, almost all analogous work in large-scale regression assumes
+a unimodal prior distribution for the effect sizes, common choices being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors.
+One of our aims here is to emphasize potential benefits of also making the unimodal assumption in the FDR context.
+
+Assumption A2 can derived from (\ref{eqn:beta hat}) plus a couple of other assumptions.
+Specifically, assume that a) $\hat\beta_j, s_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
+they are ``approximately sufficient" for $D_j$) and that b) $s_j$ alone contains little information about $\beta_j$.
 The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\hat\beta_j, s_j | \beta) = p(\hat\beta | \beta, s_j) p(s_j | \beta)$, while the latter (b) suggests assuming that $p(s_j|\beta_j)$ does not depend on $\beta_j$.
-Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\hat\beta_j | \beta, s_j)$, which, from (\ref{eqn:betahat}),  implies A2.
-(If $\hat\beta_j, s_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\hat\beta_j,s_j$, rather than conditional on $D_j$.)
+Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\hat\beta_j | \beta, s_j)$, which, with (\ref{eqn:betahat}),  implies A2.
+If $\hat\beta_j, s_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\hat\beta_j,s_j$, rather than conditional on $D_j$.
+See \cite{johnson.xx} and \cite{wakefield.xx} for more general discussion and examples of Bayesian inference based on summary statistics.
+
+Alternatively A2 could be considered definitional for $\hat\beta_j,s_j$: that is, choose $\hat\beta_j,s_j$ such that A2 holds.
 

### 2014-02-15 paper/main.tex
update paper, minor edits

@@
 
-Assumption A2 can derived from (\ref{eqn:beta hat}) plus a couple of other assumptions.
+Assumption A2 can derived from (\ref{eqn:betahat}) plus a couple of other assumptions.
 Specifically, assume that a) $\hat\beta_j, s_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
@@
 there is an important difference: here, observations with a large $s_j$ have an essentially flat likelihood (\ref{eqn:lik}), and so do not affect inference for
-other parameters, whereas when modeling the $Z$ scores directly, observations with large $s_j$ produce $N(0,1)$ $Z$ scores that do affect inference. In
-this way the likelihood-based approach can take better account of the informativeness of each measurement, as illustrated in Results below.
+other parameters, whereas when modeling the $Z$ scores directly, observations with large $s_j$ produce $Z$ scores with an $N(0,1)$ distribution that {\it do} affect inference. 
+Because of this the likelihood-based approach can take better account of the informativeness of each measurement, as illustrated in Results below.
 
@@
 That is, the non-zero effects are modeled as either i) a mixture of zero-centered normal distributions; ii) a mixture of zero-centered uniform distributions;
-or iii) a mixture of ``zero-anchored" uniform distributions, which have one end of their range at 0. This approach closely mirrors approaches in \cite{campy.thomas}.
+or iii) a mixture of ``zero-anchored" uniform distributions, which have one end of their range at 0. These approaches closely mirror approaches in \cite{campy.thomas}.
 
@@
 The latter two claims are related to characterizations of unimodal distributions due to Khintchine (1938) and  Shepp (1962); see Feller, 1971, page  158. 
-Furthermore, although it is natural for statisticians to immediately worry about ``overfitting" if $K$ is large, 
+Furthermore, although it is natural for statisticians to worry about ``overfitting" if $K$ is large, 
 in fact the unimodal constraint is sufficient to prevent serious problems with overfitting. For example, note that the unimodal
@@
 
+Note; this needs some tidying.
+
 Assuming independence across units, the likelihood for $\pi$ is 
-\begin{equation}
+\begin{align}
 L(\pi) := p(D_1,\dots,D_J | \pi) = \prod_{j=1}^n p(D_j | \pi)
-= \prod_j \int p(D_j | \pi, \beta_j) p(\beta_j | \pi) d\beta_j \\
-= \prod_j \int p(D_j | \beta_j) g(\beta_j; \pi) d\beta_j
-\end{equation}
+& = \prod_j \int p(D_j | \pi, \beta_j) p(\beta_j | \pi) d\beta_j \\
+& = \prod_j \int p(D_j | \beta_j) g(\beta_j; \pi) d\beta_j \\
+& = \prod_j \int L(\beta_j) g(\beta_j; \pi) d\beta_j \\
+& \approx \prod_j \int \hat{L}(\beta_j) g(\beta_j; \pi) d\beta_j
+\end{align}
 

### 2014-02-24 paper/Rcode/UCtest.Rmd
added uc talk

% False Discovery Rates, A New Deal
% Matthew Stephens
% 2014/2/24

```{r, include=FALSE}
require("qvalue")
require("ashr")
require("ggplot2")
```

```{r setup, include=FALSE}
# set global chunk options
opts_chunk$set(cache=TRUE,autodep=TRUE,warning=FALSE)
dep_auto()
```

# Before we get started: Getting Organized

- Over ~10 years of working with graduate students + postdocs,
I've noticed something.
- Organized researchers get more done (and better!).
- Many of them are more organized than I am!
- Thought: I should get organized; I should help others get organized.

### 2014-02-24 paper/Rcode/conservativepm.rmd
added uc talk

--- 2013-11-19 Rcode/conservativepm.Rmd
@@
 #Plot pi0 from each method
-  get_pi0.fdrtool = function(f){f$param[3]}
-  get_pi0.locfdr = function(f){f$fp0[1,3]}
-  get_pi0.mixfdr = function(f){f$pi[1]}
-
-  plot_pi0 = function(sims){
-    res=list()
-    for(i in 1:length(sims)){
-      pi0=sims[[i]]$pi0  
-      pi0_ash.n=unlist(lapply(sims[[i]]$betahat.ash.n,get_pi0))
-      pi0_ash.u=unlist(lapply(sims[[i]]$betahat.ash.u,get_pi0)) 
-      pi0_fdrtool = unlist(lapply(sims[[i]]$betahat.fdrtool,get_pi0.fdrtool))
-      pi0_locfdr=unlist(lapply(sims[[i]]$betahat.locfdr,get_pi0.locfdr))
-      pi0_mixfdr = unlist(lapply(sims[[i]]$betahat.mixfdr,get_pi0.mixfdr))
-      pi0_qval = unlist(lapply(sims[[i]]$betahat.qval,"[[","pi0"))
-
-      res[[i]] = data.frame(Scenario=i,pi0=pi0,qvalue=pi0_qval,mixfdr=pi0_mixfdr, locfdr=pi0_locfdr, fdrtool=pi0_fdrtool,ash.nullbiased=pi0_ash.n,ash.uniform=pi0_ash.u)
-    }
-    cbbPalette <- c("#56B4E9", "#009E73", "#F0E442", "#0072B2", "#E69F00", "#000000", "#D55E00", "#CC79A7")
-    labels = c('mixfdr','qvalue','locfdr','fdrtool','ash (null-biased)','ash (mle)')
-    breaks = c("mixfdr","qvalue","locfdr","fdrtool","ash.nullbiased","ash.uniform")
-
-    require(reshape2)
-    res.melt = melt(res, id.vars=c("pi0","Scenario"),variable.name="Method")
-    res.melt = res.melt[sample(1:nrow(res.melt)),]  
-    res.melt$Scenario = as.factor(res.melt$Scenario)
-    levels(res.melt$Scenario) = c("Scenario 1a","Scenario 1b","Scenario 2")
-    p=ggplot(data=res.melt,aes(pi0,value,colour=Method)) +geom_point(shape=1) +
-        facet_grid(. ~ Scenario) +
-      geom_abline(colour = "black") +
-        xlab("True pi0") +
-        ylab("Estimated pi0")
-    print(p +scale_y_continuous(limits=c(0,1)) +
-          scale_x_continuous(limits=c(0,1)) +
-          scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
-          coord_equal(ratio=1))
-
-  }
+source("plotpi0.R")
 
@@
   png("figures/estpi0_sim1sim2.png",height=160,width=540)
-    plot_pi0(list(simres1,simres1a,simres2))
+    plot_pi0(list(simres1,simres2))
   dev.off()
@@
 ```{r}
-mean_quant=function(x, mult = 1){
-      x <- na.omit(x)
-      sd <- mult * sqrt(var(x))
-      mean <- mean(x)
-      data.frame(y = median(x), ymin = quantile(x,0.25) , ymax = quantile(x,0.75))
-}
-
-#ptype indicates what type of plot to do
-#maxlfsr controls maximum x axis value
-#maxy controls maximum y axis value
-  plot_lfsr=function(sims,maxlfsr=0.1,ptype=c("lfsr","lfsra","lfdr"),maxy=1){
-    ptype=match.arg(ptype)
-    xlabtype = ifelse(ptype=="lfdr","lfdr","lfsr")
-    res=list()
-    for(i in 1:length(sims)){
-      lfsr.ash.n= unlist(lapply(sims[[i]]$betahat.ash.n,"[[",ptype))
-      lfsr.ash.u= unlist(lapply(sims[[i]]$betahat.ash.u,"[[",ptype))
-      if(ptype=="lfdr"){
-        lfsr.ash.true= unlist(lapply(sims[[i]]$betahat.ash.true,"[[","lfdr"))
-        lfdr.mixfdr = unlist(lapply(sims[[i]]$betahat.mixfdr,"[[","fdr"))
-        lfdr.locfdr = unlist(lapply(sims[[i]]$betahat.locfdr,"[[","fdr"))
-      } else {
-         lfsr.ash.true= unlist(lapply(sims[[i]]$betahat.ash.true,"[[","lfsr"))
-      }
-
-      subset = lfsr.ash.true<maxlfsr
-
-      if(length(subset)>0){  
-        res[[i]] = data.frame(Scenario=i,ash.nullbiased=lfsr.ash.n[subset],
-                      ash.uniform=lfsr.ash.u[subset],
-                      Bayes=0.1*maxlfsr*findInterval(lfsr.ash.true[subset],seq(0,maxlfsr,length=11))-0.05*maxlfsr)
-        if(ptype=="lfdr"){
-          res[[i]]=data.frame(res[[i]],mixfdr = lfdr.mixfdr[subset])
-        }
-      }
-    }
-
-    require(reshape2)
-
-    res.melt = melt(res, id.vars=c("Bayes","Scenario"),variable.name="Method")
-
-    cbbPalette <- c("#000000","#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-    labels = c('ash (null-biased)','ash (mle)')
-    breaks = c("ash.nullbiased","ash.uniform")
-    if(ptype=="lfdr"){
-      labels = c('mixfdr', labels)
-      breaks = c("mixfdr", breaks)
-    }
-
-    p=ggplot(data=res.melt,aes(Bayes,value,colour=Method)) + 
-        facet_grid(. ~ Scenario) +
-        #scale_fill_manual(values=cbbPalette) +
-        #scale_colour_manual(values=cbbPalette) +
-        geom_point(size=1,alpha=0.1) + 
-        stat_smooth(se=FALSE,size=2) +
-        stat_summary(fun.data="mean_quant", geom="ribbon", alpha=0.25) +
-        geom_abline(colour = "black") +
-        xlab(paste0(xlabtype," (Bayes)")) +
-        ylab(paste0(ptype," (Method)"))
-
-    print(p +scale_y_continuous(limits=c(0,maxy)) +
-          scale_x_continuous(limits=c(0,maxlfsr)) +
-            scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels))         
-  }
-
-```
-
-```{r}
-#plots estimated lfdr and lfsr against truth.
-#ptype indicates whether to use lfsr or lfsra
-#maxx controls maximum x axis value
-#maxy controls maximum y axis value
-plot_lfsdr=function(sims,maxx=0.1,ptype=c("lfsr","lfsra"),maxy=1){
-    ptype=match.arg(ptype)
-    res=list()
-    res.lfsr=list()
-    res.lfdr=list()
-
-    for(i in 1:length(sims)){
-      lfsr.ash.true= unlist(lapply(sims[[i]]$betahat.ash.true,"[[","lfsr"))
-      lfsr.ash.n= unlist(lapply(sims[[i]]$betahat.ash.n,"[[",ptype))
-      lfsr.ash.u= unlist(lapply(sims[[i]]$betahat.ash.u,"[[",ptype))
-
-      lfdr.ash.n= unlist(lapply(sims[[i]]$betahat.ash.n,"[[","lfdr"))
-      lfdr.ash.u= unlist(lapply(sims[[i]]$betahat.ash.u,"[[","lfdr"))
-      lfdr.ash.true= unlist(lapply(sims[[i]]$betahat.ash.true,"[[","lfdr"))
-      lfdr.mixfdr = unlist(lapply(sims[[i]]$betahat.mixfdr,"[[","fdr"))
-      lfdr.locfdr = unlist(lapply(sims[[i]]$betahat.locfdr,"[[","fdr"))
-
-      subset = lfsr.ash.true<maxx
-
-      res.lfsr[[i]] = data.frame(Scenario=i,Measure='lfsr', ash.nullbiased=lfsr.ash.n[subset], ash.uniform=lfsr.ash.u[subset],Bayes=0.1*maxx*findInterval(lfsr.ash.true[subset],seq(0,maxx,length=11))-0.05*maxx,mixfdr = NA)
-
-      subset = lfdr.ash.true<maxx
-      res.lfdr[[i]] = data.frame(Scenario=i,Measure='lfdr', ash.nullbiased=lfdr.ash.n[subset], ash.uniform=lfdr.ash.u[subset],Bayes=0.1*maxx*findInterval(lfdr.ash.true[subset],seq(0,maxx,length=11))-0.05*maxx,mixfdr = lfdr.mixfdr[subset])
-
-      res[[i]]=rbind(res.lfdr[[i]],res.lfsr[[i]])
-    }
-
-    require(reshape2)
-
-    res.melt = melt(res, id.vars=c("Bayes","Scenario","Measure"),variable.name="Method")
-    res.melt$Scenario = as.factor(res.melt$Scenario)
-    levels(res.melt$Scenario) = c("Scenario 1a","Scenario 1b","Scenario 2")
-
-    cbbPalette <- c("#E69F00","#000000", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-    labels = c('mixfdr','ash (null-biased)','ash (mle)')
-    breaks = c("mixfdr","ash.nullbiased","ash.uniform")
-
-    p=ggplot(data=res.melt,aes(Bayes,value,colour=Method)) + 
-        facet_grid(Measure ~ Scenario ) +
-        #scale_fill_manual(values=cbbPalette) +
-        #scale_colour_manual(values=cbbPalette) +
-        geom_point(size=1,alpha=0.1) + 
-        #stat_smooth(se=FALSE,size=2) +
-        stat_summary(fun.data="mean_quant", geom="ribbon", alpha=0.25) +
-        geom_abline(colour = "red",size=1) +
-        xlab("Truth") +
-        ylab("Estimate")
-
-    print(p +scale_y_continuous(limits=c(0,maxy)) +
-          scale_x_continuous(limits=c(0,maxx)) +
-            scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels))        
-  }
-
+source("plot_lfsr.R")
 ```

### 2014-02-24 talks/UCstat2014.Rmd
added uc talk

--- 2013-11-07 talks/NSmeet2013.Rmd
@@
-% Shrinkage, False Discovery Rates, and an Alternative to the Zero Assumption
+% False Discovery Rates, A New Deal
 % Matthew Stephens
-% 2013/11/1
+% 2014/2/24
+
+```{r, include=FALSE}
+require("qvalue")
+require("ashr")
+require("ggplot2")
+```
 
@@
 # set global chunk options
-opts_chunk$set(cache=TRUE,autodep=TRUE)
+opts_chunk$set(cache=TRUE,autodep=TRUE,warning=FALSE)
 dep_auto()
-library("qvalue")
-library("ashr")
 ```
+
+# Before we get started: Getting Organized
+
+- Over ~10 years of working with graduate students + postdocs,
+I've noticed something.
+- Organized researchers get more done (and better!).
+- Many of them are more organized than I am!
+- Thought: I should get organized; I should help others get organized.
+
+# So what can you do?
+
+- Buy a notebook; bring it to meetings; make notes! 
+- Come to meetings with a written agenda. 
+- While doing research, record what you did and what the outcome was.
+- Use version control [git](http://git-scm.com) and internet repositories ([bitbucket](http://www.bitbucket.org), [github](http://www.github.com)) to organize notes, code, etc.
+- Use knitr to help make your research reproducible.
+
+# What are these repository things?
+
+- A repository: a central place in which an aggregation of data is kept and maintained in an organized way (searcharticle.com)
+- Great for sharing material across multiple people (eg student and advisor!)
+- An amateur example: [http://github.com/stephens999/ash](http://github.com/stephens999/ash)
+
+# What is knitr?
+
+- An R package
+- A tool for literate programming
+- Text, and R code are interleaved
+- When you compile the document, the code is run, and output inserted into the text.
+- Great for writing reports, and keeping a track of what you did and what the result was!
+- This talk was written with knitr (with RStudio)!
+
+# What is Reproducible Research?
+
+- Principle: when publishing results of computational procedures, we should
+publish the code that produced the results.
+- "publishing figures or results without the complete software environment could
+be compared to a mathematician publishing an announcement of a mathematical theorem without giving the proof" (Buckheit and Donohoe)
+- “an article about a computational result is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result.” [Claerbout]
+
+# Why is reproducibility important?
+
+- Not only because people are forgetful, error-prone, or dishonest!
+- Reproducing work is also the first step to extending it.
+- Helps communications among researchers (eg student + advisor).
+- If you do not publish code implementing your methods, your methods will likely go unused.
+
+# More on git, github, knitr, reproducibility
+
+- Google "The git book", to get started on git.
+- Google "Karl Broman github tutorial" for statistics-oriented intro to github.
+- Google "donohoe buckheit" for "Wavelab and reproducible research"
 
@@
 
-- Constrain $g$ to be unimodal about 0.
+- Constrain $g$ to be unimodal about 0; estimate $g$ from data.
 
@@
 
-# FDR - The New Deal
+# FDR - A New Deal
 
@@
 we get a fairly flexible unimodal symmetric distribution.
+
+- Can approximate, arbitrarily closely, any scale mixture of normals.
+Includes almost all priors used for sparse regression problems (spike-and-slab, double exponential/Laplace/Bayesian Lasso, horseshoe).
+
+# FDR - A New Deal
 
@@
 
-- In practice, implement upper bound by putting prior on $\pi_0$ that encourages it to be big, then estimate $\pi$ by posterior mean (VB).
+- In practice, implement upper bound by using penalized likelihood that encourages $\pi_0$ to be as big as possible.
 
-# Illustration: BRCA data
+# Illustration: Simulated Example
+
+```{r,echo=FALSE,include=FALSE}
+load("../paper/Rcode/sim1_out.RData") #load simulation results
+source("../paper/Rcode/plot_pi0.R")
+```
+
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+plot_pi0(list(simres1))
+```
+
+# Example: BRCA data
 
@@
 
-# Example: BRCA data
-
-Compare fitted $f(\beta)$, both estimating $\pi_0$ and fixing $\pi_0=0$.
-```{r, echo=FALSE, includ=FALSE}
+```{r, echo=FALSE, include=FALSE}
   hh.ash.fdr = ash(hh.zscore,1,pointmass = TRUE,prior="nullbiased")
   hh.ash.shrink = ash(hh.zscore,1)
-```
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot(x,cdf.ash(hh.ash.fdr,x),col=2,type="l",lwd=3,ylab="cdf")
-  lines(x,cdf.ash(hh.ash.shrink,x),col=3,lwd=3,lty=2)
 ```
@@
 
-# Identifiability of $\pi_0$: Solution 2
+# Identifiability of $\pi_0$ and the False Sign Rate
 
@@
 
-# A technicality
+# The fsr is more robust than fdr
 
-- Suppose you estimate $\Pr(\beta_j<0)=0.98$,  $\Pr(\beta_j>0)=0.01$, $\Pr(\beta_j=0) = 0.01$.
+```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
+source("../paper/Rcode/plot_lfsr.R")
+  plot_lfsr(list(simres1),0.1,ptype="lfdr")
+```
 
-- Should you declare an fdr of 0.01 or 0.02?
+# The fsr is more robust than fdr
 
-- Maybe fsr makes more sense anyway?
-
-# Unification of one-group and two-group solutions
-
-- A possible concern with one-group model: if we assume all $\beta$ are non-zero, but really some are actually zero, might we not underestimate the fsr?
-
-- Consider our example, with $\Pr(\beta_j>0)=0.025$. If we actually allowed
-$\beta_j=0$ then possibly all of this probability might actually land at $\beta_j=0$.
-
-- I argue that, assuming symmetry of $g$ near 0, this also provides an upper
-bound of how much of the $\Pr(\beta_j<0)=0.975$ might also move to 0.
-
-- Therefore a more conservative estimate of the fsr might be 0.05 (or, more generally, double what you get allowing for point mass)
-
-# Example: BRCA data
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot(hh.ash.fdr$localfsr,2*hh.ash.shrink$localfsr,xlab="fsr with point mass", ylab="2 * fsr without point mass", xlim=c(0,0.1),ylim=c(0,0.1))
-  abline(a=0,b=1,col=2,lwd=2)
+```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
+  plot_lfsr(list(simres1),0.1,ptype="lfsr")
 ```
@@
 
-- So for example we can easily compute fdrs for discoveries other than ``non-zero" (eg compute $\Pr(|\beta_j| > 2 \| \hat\beta_j)$).
+- So for example we can easily compute fdrs for discoveries other than ``non-zero" (eg compute $\Pr(\beta_j > 2 | \hat\beta_j)$).
 
@@
 
+# Example: ASH applied to mouse data
+
+```{r, fig.height=4,echo=FALSE,fig.cap=""}
+plot(zdat[3,],zdat.ash$PosteriorMean,main="Estimates vs Shrunken estimates",xlab="betahat",ylab="shrunk betahat",xlim=c(-2,2),ylim=c(-0.1,.1),prob=T)
+abline(a=0,b=1,col=2)
+```
+
+# Summary
+
+- ASH provides a generic approach to shrinkage estimation, as well as
+false discovery (sign) rates.
+
+- But by using two numbers ($\hat\beta,s$) instead of one ($p$ values or $z$ scores) precision of different measurementscan be better accounted for.
+
+- Unimodal assumption for effects reduces conservatism
+
+- False Sign Rate is more robust to assumptions, and perhaps
+therefore preferable, than False Discovery Rate.
+
+# Other Applications
+
+- Widely applicable: requiring only an estimated
+effect size and standard error for each object.
+
+- Currently applying it to wavelet shrinkage applications.
+
+# Guarantees?
+
+- ``I think you have some nice ideas. How will you convince
+people to use them?" (C Morris)
+
+- Theory anyone?
+
+# Next steps?
+
+- Incorporate $t$ likelihood as well as normal.
+
+- Incorporate shrinkage of variances and not just means.
+
+- Extend to allow $g(\cdot;\pi)$ to depend on covariates $X$.
+
+- Extend to allow for correlations in the measured $\hat\beta_j$.
+
+# Thanks
+
+- to the several postdoctoral researchers and students
+who have worked with me on related topics.
+- Including Scott Powers, Mengyin Lu, Tian Sen, Wei Wang, Zhengrong Xing.
+
+# Reproducible research
+
+- This document is produced with **knitr**, **Rstudio** and **Pandoc**.
+
+- For more details see my `stephens999/ash` repository at `http://www.github.com/stephens999/ash`
+
+- Website: `http://stephenslab.uchicago.edu`
+
+# Pandoc Command used
+
+`pandoc -s -S -i --template=my.beamer -t beamer -V theme:CambridgeUS -V colortheme:beaver  slides.md -o slides.pdf`
+
+(alternative to produce html slides; but figures would need reworking)
+`pandoc -s -S -i -t dzslides --mathjax slides.md -o slides.html`
+
+Here is my session info:
+
+```{r session-info}
+print(sessionInfo(), locale=FALSE)
+```
+
+# Some odd things in the data
+
+```{r, echo=FALSE,fig.height=3}
+  plot(zdat[3,],zdat[4,],ylab="standard error",xlab="beta-hat")
+  x[tail(order(zdat[3,])),]
+```
+
+# A technicality
+
+- Suppose you estimate $\Pr(\beta_j<0)=0.98$,  $\Pr(\beta_j>0)=0.01$, $\Pr(\beta_j=0) = 0.01$. 
+- Should you declare an fdr of 0.01 or 0.02?
+- Maybe fsr makes more sense anyway?
+
 # Shrinkage is adaptive to information
+
+Need to fix counts.associate to use fdr method in ash
 
@@
 ```{r, echo=FALSE,fig.height=3,fig.cap=""}
-plot(zdat.ash$localfdr,pval,ylab="p value", xlab="ASH local fdr")
+plot(zdat.ash$lfdr,pval,ylab="p value", xlab="ASH local fdr")
 ```
@@
 ```{r, echo=FALSE,fig.height=3,fig.cap=""}
-plot(zdat.ash$localfdr,pval,ylab="p value", xlab="ASH local fdr")
-points(zdat.ash$localfdr[15325],pval[15325],col=2,pch=16)
-points(zdat.ash$localfdr[16123],pval[16123],col=2,pch=16)
+plot(zdat.ash$lfdr,pval,ylab="p value", xlab="ASH local fdr")
+points(zdat.ash$lfdr[15325],pval[15325],col=2,pch=16)
+points(zdat.ash$lfdr[16123],pval[16123],col=2,pch=16)
 ```
@@
 xx = xx[xx>0]
-cbind(x[,1:5],pval,zdat.ash$localfdr)[c(15325,16123),]
+cbind(x[,1:5],pval,zdat.ash$lfdr)[c(15325,16123),]
 ```
@@
 
-# Summary: FDR via Adapative Shrinkage
-
-- Both provide a rational approach to identifying ``significant" findings.
-
-- Both are generic and modular: once you have the summary data, you can forget where they came from.
-
-- But by using two numbers ($\hat\beta,s$) instead of one ($p$ values) precision of different measurementscan be better accounted for.
-
-- ASH borrows information for estimation, as well as testing.
-
-# Other Applications
-
-- Widely applicable: perhaps anywhere (?) 
-where shrinkage is appropriate, requiring only an estimated
-effect size and standard error for each object.
-
-- Could also use effect size estimate and $p$ value for each variable, by converting to effect size estimate and (pseudo-) standard error.
-
-- Currently applying it to wavelet shrinkage applications.
-
-# Guarantees?
-
-- ``I think you have some nice ideas. How will you convince
-people to use them?" (C Morris)
-
-# Next steps?
-
-- Extend to allow $g(\cdot;\pi)$ to depend on covariates $X$.
-
-- Extend to allow for correlations in the measured $\hat\beta_j$.
-
-# Thanks
-
-- to the several postdoctoral researchers and students
-who have worked with me on related topics.
-
-- Especially Mengyin Lu who coded the VB algorithm.
-
-# Reproducible research
-
-- This document is produced with **knitr**, **Rstudio** and **Pandoc**.
-
-- For more details see my `stephens999/ash` repository at `http://www.github.com/stephens999/ash`
-
-- Website: `http://stephenslab.uchicago.edu`
-
-# Pandoc Command used
-
-`pandoc -s -S -i --template=my.beamer -t beamer -V theme:CambridgeUS -V colortheme:beaver  ilike-slides.md -o ilike-slides.pdf`
-
-(alternative to produce html slides; but figures would need reworking)
-`pandoc -s -S -i -t dzslides --mathjax NSmeet2013.md -o NSmeet2013.html`
-
-Here is my session info:
-
-```{r session-info}
-print(sessionInfo(), locale=FALSE)
+Compare fitted $f(\beta)$, both estimating $\pi_0$ and fixing $\pi_0=0$.
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+  x=seq(-4,4,length=100)
+  plot(cdf.ash(hh.ash.fdr,x),col=2,type="l",lwd=3,ylab="cdf")
+  lines(x,cdf.ash(hh.ash.shrink,x)$y,col=3,lwd=3,lty=2)
 ```
-
-# Some odd things in the data
-
-```{r, echo=FALSE,fig.height=3}
-  plot(zdat[3,],zdat[4,],ylab="standard error",xlab="beta-hat")
-  x[tail(order(zdat[3,])),]
-```

### 2014-02-24 talks/UCtest.Rmd
added uc talk

--- 2014-02-24 talks/UCstat2014.Rmd
@@
 
-- Can approximate, arbitrarily closely, any scale mixture of normals.
-Includes almost all priors used for sparse regression problems (spike-and-slab, double exponential/Laplace/Bayesian Lasso, horseshoe).
-
-# FDR - A New Deal
+- Can approximate, arbitrarily closely, any ``scale mixture of normals" which
+includes almost all the priors people hoave used for sparse regression problems (spike-and-slab, ``horseshoe", double exponential/Laplace/Lasso).
 
@@
 
+Compare fitted $f(\beta)$, both estimating $\pi_0$ and fixing $\pi_0=0$.
 ```{r, echo=FALSE, include=FALSE}
@@
   hh.ash.shrink = ash(hh.zscore,1)
+```
+
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+  x=seq(-4,4,length=100)
+  plot(cdf.ash(hh.ash.fdr,x),col=2,type="l",lwd=3,ylab="cdf")
+  lines(x,cdf.ash(hh.ash.shrink,x)$y,col=3,lwd=3,lty=2)
 ```
@@
 
-# Identifiability of $\pi_0$ and the False Sign Rate
+# Identifiability of $\pi_0$: Solution 2
 
@@
 
-```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-source("../paper/Rcode/plot_lfsr.R")
-  plot_lfsr(list(simres1),0.1,ptype="lfdr")
-```
+- but still need to encourage point mass on zero if want to be ``conservative".
 
-# The fsr is more robust than fdr
+- blah
 
-```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-  plot_lfsr(list(simres1),0.1,ptype="lfsr")
+# Unification of one-group and two-group solutions
+
+- A possible concern with one-group model: if we assume all $\beta$ are non-zero, but really some are actually zero, might we not underestimate the fsr?
+
+- Consider our example, with $\Pr(\beta_j>0)=0.025$. If we actually allowed
+$\beta_j=0$ then possibly all of this probability might actually land at $\beta_j=0$.
+
+- I argue that, assuming symmetry of $g$ near 0, this also provides an upper
+bound of how much of the $\Pr(\beta_j<0)=0.975$ might also move to 0.
+
+- Therefore a more conservative estimate of the fsr might be 0.05 (or, more generally, double what you get allowing for point mass)
+
+# Example: BRCA data
+
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+  plot(hh.ash.fdr$lfsr,2*hh.ash.shrink$lfsr,xlab="fsr with point mass", ylab="2 * fsr without point mass", xlim=c(0,0.1),ylim=c(0,0.1))
+  abline(a=0,b=1,col=2,lwd=2)
 ```
@@
 
-- So for example we can easily compute fdrs for discoveries other than ``non-zero" (eg compute $\Pr(\beta_j > 2 | \hat\beta_j)$).
-
-- And use it to obtain point estimates and credible intervals for each $\beta_j$, taking account of information from all the other $\beta_j$.
-
-- Because $f(\beta)$ is unimodal, the point estimates will tend to be ``shrunk" towards the overall mean (0).
-
-- Because $f(\beta)$ is estimated from the data, the amount
-of shrinkage is adaptive to the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each gene.
-
-- So we call the approach ``Adaptive Shrinkage" (ASH).
-
-# Example: ASH applied to mouse data
-
-```{r, fig.height=4,echo=FALSE,fig.cap=""}
-hist(zdat[3,],main="Raw effect size estimates",xlab="betahat",xlim=c(-2,2),prob=T,ylim=c(0,3))
-```
-
-# Example: ASH applied to mouse data
-
-```{r, fig.height=4,echo=FALSE,fig.cap=""}
-hist(zdat[3,],main="Raw effect size estimates",xlab="betahat",xlim=c(-2,2),prob=T,ylim=c(0,3))
-t=seq(-2,2,length=201)
-lines(t,dnorm(t,0,sd=0.032), col=2, lwd=2)
-```
-
-# Example: ASH applied to mouse data
-
-```{r, fig.height=4,echo=FALSE,fig.cap=""}
-hist(zdat.ash$PosteriorMean,main="Shrunken estimates",xlab="shrunk betahat",xlim=c(-0.1,0.1),prob=T)
-lines(t,dnorm(t,0,sd=0.032), col=2, lwd=2)
-```
-
-# Example: ASH applied to mouse data
-
-```{r, fig.height=4,echo=FALSE,fig.cap=""}
-plot(zdat[3,],zdat.ash$PosteriorMean,main="Estimates vs Shrunken estimates",xlab="betahat",ylab="shrunk betahat",xlim=c(-2,2),ylim=c(-0.1,.1),prob=T)
-abline(a=0,b=1,col=2)
-```
-
-# Summary
-
-- ASH provides a generic approach to shrinkage estimation, as well as
-false discovery (sign) rates.
-
-- But by using two numbers ($\hat\beta,s$) instead of one ($p$ values or $z$ scores) precision of different measurementscan be better accounted for.
-
-- Unimodal assumption for effects reduces conservatism
-
-- False Sign Rate is more robust to assumptions, and perhaps
-therefore preferable, than False Discovery Rate.
-
-# Other Applications
-
-- Widely applicable: requiring only an estimated
-effect size and standard error for each object.
-
-- Currently applying it to wavelet shrinkage applications.
-
-# Guarantees?
-
-- ``I think you have some nice ideas. How will you convince
-people to use them?" (C Morris)
-
-- Theory anyone?
-
-# Next steps?
-
-- Incorporate $t$ likelihood as well as normal.
-
-- Incorporate shrinkage of variances and not just means.
-
-- Extend to allow $g(\cdot;\pi)$ to depend on covariates $X$.
-
-- Extend to allow for correlations in the measured $\hat\beta_j$.
-
-# Thanks
-
-- to the several postdoctoral researchers and students
-who have worked with me on related topics.
-- Including Scott Powers, Mengyin Lu, Tian Sen, Wei Wang, Zhengrong Xing.
-
-# Reproducible research
-
-- This document is produced with **knitr**, **Rstudio** and **Pandoc**.
-
-- For more details see my `stephens999/ash` repository at `http://www.github.com/stephens999/ash`
-
-- Website: `http://stephenslab.uchicago.edu`
-
-# Pandoc Command used
-
-`pandoc -s -S -i --template=my.beamer -t beamer -V theme:CambridgeUS -V colortheme:beaver  slides.md -o slides.pdf`
-
-(alternative to produce html slides; but figures would need reworking)
-`pandoc -s -S -i -t dzslides --mathjax slides.md -o slides.html`
-
-Here is my session info:
-
-```{r session-info}
-print(sessionInfo(), locale=FALSE)
-```
-
-# Some odd things in the data
-
-```{r, echo=FALSE,fig.height=3}
-  plot(zdat[3,],zdat[4,],ylab="standard error",xlab="beta-hat")
-  x[tail(order(zdat[3,])),]
-```
-
-# A technicality
-
-- Suppose you estimate $\Pr(\beta_j<0)=0.98$,  $\Pr(\beta_j>0)=0.01$, $\Pr(\beta_j=0) = 0.01$. 
-- Should you declare an fdr of 0.01 or 0.02?
-- Maybe fsr makes more sense anyway?
-
-# Shrinkage is adaptive to information
-
-Need to fix counts.associate to use fdr method in ash
-
-```{r, echo=FALSE,fig.height=4}
-#hist(zdat[3,])
-#plot(zdat[3,],zdat.ash$PosteriorMean,xlim=c(-0.2,0.2))
-#points(zdat[3,16677],zdat.ash$PosteriorMean[16677],col=2,pch=16)
-#points(zdat[3,16079],zdat.ash$PosteriorMean[16079],col=2,pch=16)
-#x[16677,]
-#x[16079,]
-```
-
-```{r, echo=FALSE}
-#temp = counts.associate(cc,c(-1,1,-1,1),1)
-#temp.tscore = temp$zdat[3,]/temp$zdat[4,]
-#temp.pval = ttest.pval(temp.tscore,df=2)
-#plot(temp$zdat.ash$localfdr,temp.pval)
-#identify(temp$zdat.ash$localfdr,temp.pval)
-```
-
-```{r, echo=FALSE,fig.height=3,fig.cap=""}
-plot(zdat.ash$lfdr,pval,ylab="p value", xlab="ASH local fdr")
-```
-
-# Shrinkage is adaptive to information
-
-```{r, echo=FALSE,fig.height=3,fig.cap=""}
-plot(zdat.ash$lfdr,pval,ylab="p value", xlab="ASH local fdr")
-points(zdat.ash$lfdr[15325],pval[15325],col=2,pch=16)
-points(zdat.ash$lfdr[16123],pval[16123],col=2,pch=16)
-```
-
-# Shrinkage is adaptive to information
-
-```{r, echo=FALSE}
-x = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
-xx = rowSums(x[,2:5])
-x = x[xx>0,]
-xx = xx[xx>0]
-cbind(x[,1:5],pval,zdat.ash$lfdr)[c(15325,16123),]
-```
-
-# Recall FDR problem 1: q values increased by low count genes
-```{r,echo=FALSE, fig.height=4,fig.cap=""}
-plot(qval.high$qval,qval$qval[xx>1000],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
-abline(a=0,b=1,col=2)
-```
-
-# ASH q values more robust to inclusion of low count genes
-
-```{r, echo=FALSE,fig.height=3,fig.cap=""}
-plot(zdat.ash.high$qvalue,zdat.ash$qvalue[xx>1000],xlab="q values from high only",ylab="q values from all data")
-abline(a=0,b=1,col=2,lwd=2)
-```
-
-Compare fitted $f(\beta)$, both estimating $\pi_0$ and fixing $\pi_0=0$.
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  x=seq(-4,4,length=100)
-  plot(cdf.ash(hh.ash.fdr,x),col=2,type="l",lwd=3,ylab="cdf")
-  lines(x,cdf.ash(hh.ash.shrink,x)$y,col=3,lwd=3,lty=2)
-```
+- So for example we can easily compute fdrs for discoveries other than ``non-zero" (eg compute $\Pr(|\beta_j| > 2 | \hat\beta_j)$).

### 2014-02-24 paper/Rcode/plot_lfsr.R

@@ -1,0 +1,126 @@
+require(reshape2)
+require(ggplot2)
+
+mean_quant=function(x, mult = 1){
+  x <- na.omit(x)
+  sd <- mult * sqrt(var(x))
+  mean <- mean(x)
+  data.frame(y = median(x), ymin = quantile(x,0.25) , ymax = quantile(x,0.75))
+}
+
+#ptype indicates what type of plot to do
+#maxlfsr controls maximum x axis value
+#maxy controls maximum y axis value
+plot_lfsr=function(sims,maxlfsr=0.1,ptype=c("lfsr","lfsra","lfdr"),maxy=1){
+  ptype=match.arg(ptype)
+  xlabtype = ifelse(ptype=="lfdr","lfdr","lfsr")
+  res=list()
+  for(i in 1:length(sims)){
+    lfsr.ash.n= unlist(lapply(sims[[i]]$betahat.ash.n,"[[",ptype))
+    lfsr.ash.u= unlist(lapply(sims[[i]]$betahat.ash.u,"[[",ptype))
+    if(ptype=="lfdr"){
+      lfsr.ash.true= unlist(lapply(sims[[i]]$betahat.ash.true,"[[","lfdr"))
+      lfdr.mixfdr = unlist(lapply(sims[[i]]$betahat.mixfdr,"[[","fdr"))
+      lfdr.locfdr = unlist(lapply(sims[[i]]$betahat.locfdr,"[[","fdr"))
+    } else {
+      lfsr.ash.true= unlist(lapply(sims[[i]]$betahat.ash.true,"[[","lfsr"))
+    }
+    
+    subset = lfsr.ash.true<maxlfsr
+    
+    if(length(subset)>0){  
+      res[[i]] = data.frame(Scenario=i,ash.nullbiased=lfsr.ash.n[subset],
+                            ash.uniform=lfsr.ash.u[subset],
+                            Bayes=0.1*maxlfsr*findInterval(lfsr.ash.true[subset],seq(0,maxlfsr,length=11))-0.05*maxlfsr)
+      if(ptype=="lfdr"){
+        res[[i]]=data.frame(res[[i]],mixfdr = lfdr.mixfdr[subset])
+      }
+    }
+  }
+  
+
+  
+  res.melt = melt(res, id.vars=c("Bayes","Scenario"),variable.name="Method")
+  
+  
[…]

### 2014-02-24 paper/Rcode/plot_pi0.R

@@ -1,0 +1,40 @@
+#functions to plot estimates of pi0 from simulations
+require(ggplot2)
+require(reshape2)
+
+get_pi0.fdrtool = function(f){f$param[3]}
+get_pi0.locfdr = function(f){f$fp0[1,3]}
+get_pi0.mixfdr = function(f){f$pi[1]}
+
+plot_pi0 = function(sims){
+  res=list()
+  for(i in 1:length(sims)){
+    pi0=sims[[i]]$pi0  
+    pi0_ash.n=unlist(lapply(sims[[i]]$betahat.ash.n,get_pi0))
+    pi0_ash.u=unlist(lapply(sims[[i]]$betahat.ash.u,get_pi0)) 
+    pi0_fdrtool = unlist(lapply(sims[[i]]$betahat.fdrtool,get_pi0.fdrtool))
+    pi0_locfdr=unlist(lapply(sims[[i]]$betahat.locfdr,get_pi0.locfdr))
+    pi0_mixfdr = unlist(lapply(sims[[i]]$betahat.mixfdr,get_pi0.mixfdr))
+    pi0_qval = unlist(lapply(sims[[i]]$betahat.qval,"[[","pi0"))
+    
+    res[[i]] = data.frame(Scenario=i,pi0=pi0,qvalue=pi0_qval,mixfdr=pi0_mixfdr, locfdr=pi0_locfdr, fdrtool=pi0_fdrtool,ash.nullbiased=pi0_ash.n,ash.uniform=pi0_ash.u)
+  }
+  cbbPalette <- c("#56B4E9", "#009E73", "#F0E442", "#0072B2", "#E69F00", "#000000", "#D55E00", "#CC79A7")
+  labels = c('mixfdr','qvalue','locfdr','fdrtool','ash (null-biased)','ash (mle)')
+  breaks = c("mixfdr","qvalue","locfdr","fdrtool","ash.nullbiased","ash.uniform")
+  
+  res.melt = melt(res, id.vars=c("pi0","Scenario"),variable.name="Method")
+  res.melt = res.melt[sample(1:nrow(res.melt)),]  
+  res.melt$Scenario = as.factor(res.melt$Scenario)
+  levels(res.melt$Scenario) = c("Scenario 1a","Scenario 1b","Scenario 2")
+  p=ggplot(data=res.melt,aes(pi0,value,colour=Method)) +geom_point(shape=1) +
+    facet_grid(. ~ Scenario) +
+    geom_abline(colour = "black") +
[…]

### 2014-02-24 paper/Rcode/sim1.R

@@ -1,0 +1,76 @@
+#R code to run some simulations
+
+require(ashr)
+require(qvalue)
+require(fdrtool)
+require(mixfdr)
+require(locfdr)
+require(ggplot2)
+#bsd gives standard deviation of beta
+#pi0 is simulated to be uniform on [minpi0,1]
+basicsim=function(mixsd,mixpi_alt,bsd=1,minpi0=0,seedval = 100,nsamp=1000,niter=50){  
+  set.seed(seedval)  
+  beta =list()
+  betahatsd=list()
+  betahat = list()
+  zscore = list()
+  pval = list()
+  betahat.ash.n = list()
+  betahat.ash.u = list()
+  betahat.ash.npm = list()
+  betahat.ash.true = list()
+  betahat.qval = list()
+  betahat.fdrtool = list()
+  betahat.locfdr = list()
+  betahat.mixfdr = list()
+  pi0 = rep(0,niter)
+  for(i in 1:niter){
+    pi0[i]=runif(1,minpi0,1)
+    mixpi = c(pi0[i],(1-pi0[i])*mixpi_alt)
+    sd = sample(mixsd,nsamp,prob=mixpi,replace=TRUE )
+    beta[[i]] = rnorm(nsamp,0,sd)
+    betahatsd[[i]] = bsd
+    betahat[[i]] = beta[[i]]+rnorm(nsamp,0,betahatsd[[i]])
+    zscore[[i]] = betahat[[i]]/betahatsd[[i]]
+    pval[[i]] = pchisq(zscore[[i]]^2,df=1,lower.tail=F)
+    betahat.ash.n[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=TRUE,prior="nullbiased",gridmult=2)
+    betahat.ash.u[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=TRUE,prior="uniform",gridmult=2)
+    betahat.ash.npm[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=FALSE,prior="uniform",gridmult=2)
+    betahat.ash.true[[i]] = ash(betahat[[i]],betahatsd[[i]],g=normalmix(mixpi,rep(0,length(mixpi)),mixsd))
+    
+    cat("applying q value\n")
+    betahat.qval[[i]] = qvalue(pval[[i]])
+    cat("applying fdrtool\n")
+    betahat.fdrtool[[i]] = fdrtool(pval[[i]],statistic="pvalue",plot=FALSE)
[…]

### 2014-02-25 talks/UCstat2014.Rmd
updated and reran longer simulations after talk

@@
 - While doing research, record what you did and what the outcome was.
-- Use version control [git](http://git-scm.com) and internet repositories ([bitbucket](http://www.bitbucket.org), [github](http://www.github.com)) to organize notes, code, etc.
-- Use knitr to help make your research reproducible.
+- Use version control ([git](http://git-scm.com)) and internet repositories ([bitbucket](http://www.bitbucket.org), [github](http://www.github.com)) to organize notes, code, etc.
+- Use *knitr* to help make your research reproducible.
 
@@
   plot(zdat[3,],zdat[4,],ylab="standard error",xlab="beta-hat")
-  x[tail(order(zdat[3,])),]
+  dd[tail(order(zdat[3,])),]
 ```
@@
 ```{r, echo=FALSE}
-x = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
-xx = rowSums(x[,2:5])
-x = x[xx>0,]
+dd = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
+xx = rowSums(dd[,2:5])
+dd = dd[xx>0,]
 xx = xx[xx>0]
-cbind(x[,1:5],pval,zdat.ash$lfdr)[c(15325,16123),]
+cbind(dd[,1:5],pval,zdat.ash$lfdr)[c(15325,16123),]
 ```

### 2014-02-25 paper/Rcode/sim1.R

@@ -66,9 +66,9 @@ basicsim=function(mixsd,mixpi_alt,bsd=1,minpi0=0,seedval = 100,nsamp=1000,niter=
 mixsd = c(0,0.25,0.5,1,2)
 mixpi_alt = c(0.4,0.2,0.2,0.2) #mixture proportions under the alternative
 
-simres1 = basicsim(mixsd,mixpi_alt,niter=50,nsamp=1000)
-#simres1 = basicsim(mixsd,mixpi_alt,niter=200,nsamp=1000)
-#simres1a = basicsim(mixsd,mixpi_alt,niter=200,nsamp=10000)
+#simres1 = basicsim(mixsd,mixpi_alt,niter=50,nsamp=1000)
+simres1 = basicsim(mixsd,mixpi_alt,niter=200,nsamp=1000)
+simres1a = basicsim(mixsd,mixpi_alt,niter=200,nsamp=10000)
 simres2= basicsim(c(0,4),c(1))
 simres3=basicsim(c(0,2),c(1),bsd=c(rep(1,500),rep(10,500)))
 simres4=basicsim(c(0,2),c(1),bsd=c(rep(1,500),rep(10,500)),minpi0=0.9,seed=200)

### 2014-02-26 paper/Rcode/ashsim.rmd
rearranged simulations for paper some

--- 2013-11-19 paper/Rcode/ashsim.rmd
@@
-```{r}
-  require("ashr")
 ```{r}
@@
 require(ggplot2)   
-#simulate from mixture of normals
-#compare normal, uniform and half-uniform mixtures
-#bsd gives standard deviation of beta
-ashsim=function(mixmean,mixsd,mixpi,bsd=1,seedval = 100,nsamp=1000,niter=50){  
-  set.seed(seedval)  
-  beta =list()
-  betahatsd=list()
-  betahat = list()
-  betahat.ash.n = list()
-  betahat.ash.u = list()
-  betahat.ash.hu = list()
-  k = length(mixmean)
-  for(i in 1:niter){
-    comp = sample(1:k,nsamp,prob=mixpi,replace=TRUE)
-    sd = mixsd[comp]
-    mean = mixmean[comp]
-    beta[[i]] = rnorm(nsamp,mean,sd)
-    betahatsd[[i]] = bsd
-    betahat[[i]] = beta[[i]]+rnorm(nsamp,0,betahatsd[[i]])
-    betahat.ash.n[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="normal",method="shrink")
-    betahat.ash.u[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="uniform", method="shrink")
-    betahat.ash.hu[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform", method="shrink")
-
-  }
-  return(list(beta =beta,
-  betahatsd=betahatsd,
-  betahat = betahat,
-  betahat.ash.n = betahat.ash.n,
-  betahat.ash.u = betahat.ash.u,
-  betahat.ash.hu = betahat.ash.hu))
-}    
 ```
@@
 ```{r}
-sim1= ashsim(c(0,0,0),c(1,1,2),c(1/3,1/3,1/3),niter=20,nsamp=1000)
-sim2= ashsim(c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7),rep(1/7,7),niter=20,nsamp=1000)
-sim3= ashsim(c(-2,-1,0,1),c(2,1.5,1,1),c(1/4,1/4,1/3,1/6),niter=20,nsamp=1000)
-
+  load("ashsim_out.RData")
+  source("plot_examples.R")
+  source("plot_rmse.R")
 ```
@@
 ```{r}
-len=length(sim1$beta[[1]])
-res = data.frame(beta = c(sim1$beta[[1]],sim2$beta[[1]],sim3$beta[[1]]), Scenario = c(rep(1,len),rep(2,len),rep(3,len)))
-x = seq(-8,6,length=100)
-y = dnorm(x)
-dens1 = data.frame(x=x,dn= t(density(sim1$betahat.ash.n[[1]],x)$y),du=t(density(sim1$betahat.ash.u[[1]],x)$y),dhu=t(density(sim1$betahat.ash.hu[[1]],x)$y), Scenario=1)
-dens2 = data.frame(x=x,dn = t(density(sim2$betahat.ash.n[[1]],x)$y),du=t(density(sim2$betahat.ash.u[[1]],x)$y), dhu=t(density(sim2$betahat.ash.hu[[1]],x)$y), Scenario=2)
-dens3 = data.frame(x=x,dn = t(density(sim3$betahat.ash.n[[1]],x)$y),du=t(density(sim3$betahat.ash.u[[1]],x)$y), dhu=t(density(sim3$betahat.ash.hu[[1]],x)$y), Scenario=3)
-dens=rbind(dens1,dens2,dens3)
-ggplot(res, aes(x=beta)) + facet_grid(.~Scenario)  +
-    geom_histogram(aes(y=..density..),binwidth=0.5,
-                   colour="black", fill="white") +  
-  geom_line(data = dens, aes(x,du,color='Uniform'),size=1.2,alpha=0.8) +
-  geom_line(data = dens, aes(x,dhu,color='Half Uniform'),size=1.2,alpha=0.8) +
-    geom_line(data = dens, aes(x,dn,color='Normal'),size=1.2,alpha=0.8) +
-  scale_colour_manual(name = 'Method', values = c('red', 'blue','green'))
-
+pdf("figures/simABC_egdens.pdf",width=6.5,height=2)
+plot_examples(list(simA,simB,simC))
+dev.off()
 ```
@@
 ```{r}
-  rmse = function(x,y){sqrt(mean((x-y)^2))}
-  get_rmse.ash = function(a,b){rmse(a$PosteriorMean,b)}
-  get_rmse.mixfdr = function(a,b){rmse(a$effectSize,b)}
-  plot_rmse = function(sims,inczero=FALSE,incbetahat=FALSE){
-    res=list()
-    res2=list()
-    for(i in 1:length(sims)){
-      err.ash.n= mapply(get_rmse.ash,sims[[i]]$betahat.ash.n,sims[[i]]$beta)
-      err.ash.u = mapply(get_rmse.ash,sims[[i]]$betahat.ash.u,sims[[i]]$beta)
-      err.ash.hu = mapply(get_rmse.ash,sims[[i]]$betahat.ash.hu,sims[[i]]$beta)
-    #err.mixfdr=mapply(get_rmse.mixfdr,sims$betahat.mixfdr,sims$beta)
-      err.betahat = mapply(rmse,sims[[i]]$betahat,sims[[i]]$beta)
-      err.zero = unlist(lapply(sims[[i]]$beta,rmse,y=0))
-      loglik.ash.n = unlist(lapply(sims[[i]]$betahat.ash.n,get_loglik))
-      loglik.ash.u = unlist(lapply(sims[[i]]$betahat.ash.u,get_loglik))
-      loglik.ash.hu = unlist(lapply(sims[[i]]$betahat.ash.hu,get_loglik))
-     res[[i]] = data.frame(Scenario=i,ash.normal=err.ash.n,ash.uniform=err.ash.u,ash.halfuniform=err.ash.hu)
-      res2[[i]] = data.frame(Scenario=i,ash.normal=loglik.ash.n,ash.uniform=loglik.ash.u,ash.halfuniform=loglik.ash.hu)
+  pdf("figures/rmse_biplot.pdf",width=6.5,height=2)
+    plot_rmse(list(simA,simB,simC))
+  dev.off()
 
-      if(inczero){
-        res[[i]]=data.frame(res,zero=err.zero)
-      }
-      if(incbetahat){
-        res[[i]]=data.frame(res,betahat=err.betahat)
-      }
-    }
-    require(reshape2)
-    res.melt = melt(res, id.vars=c("Scenario"),variable.name="Method")
-    res2.melt = melt(res2, id.vars=c("Scenario"),variable.name="Method")
-    ggplot(data=res.melt,aes(Method,value)) +geom_boxplot()+         facet_grid(. ~ Scenario)
-     ggplot(data=res2.melt,aes(Method,value)) +geom_boxplot()+         facet_grid(. ~ Scenario)
-
-    p=ggplot(data=res.melt,aes(ash.halfuniform,value,colour=Method)) +geom_point(shape=16) +
-         facet_grid(. ~ Scenario) +
-        geom_abline(colour = "black") +
-        xlab("RMSE (ash.halfuniform)") +
-        ylab("RMSE (other method)")
-    print(p +
-          coord_equal(ratio=1))
-  }
-  #pdf("figures/rmse_sim1.pdf")
-    plot_rmse(list(sim1,sim2,sim3))
- #               dev.off()
-  #pdf("figures/rmse_sim2.pdf")
-  #  plot_rmse(simres2)
-  #dev.off()
-```
-
-Try plotting boxplots of rmse and fit impovement compared with most general model
-```{r}
-  rmse = function(x,y){sqrt(mean((x-y)^2))}
-  get_rmse.ash = function(a,b){rmse(a$PosteriorMean,b)}
-  get_rmse.mixfdr = function(a,b){rmse(a$effectSize,b)}
-  boxplot_rmse_loglik = function(sims,inczero=FALSE,incbetahat=FALSE){
-    res=list()
-    res2=list()
-    for(i in 1:length(sims)){
-      err.ash.hu = mapply(get_rmse.ash,sims[[i]]$betahat.ash.hu,sims[[i]]$beta)
-      err.ash.n= mapply(get_rmse.ash,sims[[i]]$betahat.ash.n,sims[[i]]$beta)-err.ash.hu
-      err.ash.u = mapply(get_rmse.ash,sims[[i]]$betahat.ash.u,sims[[i]]$beta)-err.ash.hu
-      #err.mixfdr=mapply(get_rmse.mixfdr,sims$betahat.mixfdr,sims$beta)
-      loglik.ash.hu = unlist(lapply(sims[[i]]$betahat.ash.hu,get_loglik))
-      loglik.ash.n = unlist(lapply(sims[[i]]$betahat.ash.n,get_loglik))-loglik.ash.hu
-      loglik.ash.u = unlist(lapply(sims[[i]]$betahat.ash.u,get_loglik))-loglik.ash.hu
-
-     res[[i]] = data.frame(Scenario=i,ash.normal.err=err.ash.n,ash.uniform.err=err.ash.u)
-       res2[[i]] = data.frame(Scenario=i,ash.normal=loglik.ash.n,ash.uniform=loglik.ash.u)
-
-    }
-    require(reshape2)
-    res.melt = melt(res, id.vars=c("Scenario"))
-
-    res2.melt=melt(res2,id.vars=c("Scenario"))
-    ggplot(res.melt,aes(variable,value)) + geom_boxplot() + facet_grid(.~Scenario)
-    ggplot(res2.melt,aes(variable,value)) + geom_boxplot() + facet_grid(.~Scenario)
-  }
-
-  pdf("figures/boxplot_rmse_loglik.pdf")
-    boxplot_rmse_loglik(list(sim1,sim2,sim3))
+#This figure not used in paper?
+pdf("figures/rmse_biplot_withzerobetahat.pdf")
+    plot_rmse(list(simA,simB,simC),inczero=TRUE,incbetahat=TRUE)
 dev.off()
@@
 ```{r}
+
+  pdf("figures/rmse_boxplot.pdf",width=6.5,height=2)
+    plot_rmse_boxplot(list(simA,simB,simC))
+  dev.off()
+
+    pdf("figures/rmse_boxplot_extended.pdf",width=6.5,height=2)
+    plot_rmse_boxplot(list(simA,simB,simC),TRUE,TRUE,TRUE)
+  dev.off()
+
+```
+
+```{r}
+
+  pdf("figures/loglik_boxplot.pdf",width=6.5,height=2)
+    plot_loglik_boxplot(list(simA,simB,simC))
+  dev.off()
+
+```
+
+```{r}
+
+  pdf("figures/rmse_loglik_boxplot.pdf",width=6.5,height=5)
+    plot_rmse_loglik_boxplot(list(simA,simB,simC))
+  dev.off()
+
+```
+
+looking to see why mixfdr gives worse rmse, even though loglik looks ok.
+Perhaps it overshrinks?
+```{r}
+ plot(simC$fit.ash.hu[[1]]$PosteriorMean, simC$fit.mixfdr[[1]]$effectSize)
+ plot(simC$fit.ash.fdr.n[[1]]$PosteriorMean, simC$fit.mixfdr[[1]]$effectSize)
+```
+
+problem seems to be that it overestimates pi0
+```{r}
+ simC$fit.mixfdr[[1]]$pi
+```
+[1] 0.71660654 0.08298064 0.20041282
+
+```{r}
+temp=mixFdr(simC$betahat[[1]]/simC$betahatsd[[1]],noiseSD=1,theonull=TRUE,calibrate=TRUE)
+temp2=mixFdr(simC$betahat[[1]]/simC$betahatsd[[1]],noiseSD=1,theonull=TRUE,J=10,plot=FALSE)
+plot(ecdf(simC$beta[[1]]),xlim=c(-6,6))
+lines(x,cdf.mixfdr(temp,x),col=3)
+lines(cdf.ash(simC$fit.ash.n[[1]],x),col=2)
+lines(cdf.ash(simC$fit.ash.hu[[1]],x),col=2,lty=2)
+```
+
+```{r}
   plot_LR=function(sims){
-    hist(unlist(lapply(sims$betahat.ash.u,get_loglik))-unlist(lapply(sims$betahat.ash.n,get_loglik)), xlab="loglik difference", main="loglik differences for nullbiased prior vs mle",nclass=10)
+    hist(unlist(lapply(sims$fit.ash.u,get_loglik))-unlist(lapply(sims$fit.ash.n,get_loglik)), xlab="loglik difference", main="loglik differences for nullbiased prior vs mle",nclass=10)
   }
@@
   pdf("figures/logLR.pdf")
-    plot_LR(simres1)
-    plot_LR(simres2)
+    plot_LR(simA)
+    plot_LR(simB)
+  dev.off()
+```
+
+## Unused figures?
+
+```{r}
+
+pdf("figures/simABC_eg_withfit.pdf",width=6.5,height=2)
+plot_examples_withfit(list(simA,simB,simC))
 dev.off()
@@
 ```{r}
-  lapply()
+pdf("figures/simABC_eg_cdf_withfit.pdf",width=6.5,height=2)
+plot_examples_cdf_withfit(list(simA,simB,simC))
+dev.off()
 ```
+
+Experiment with mixFdr to see how penalization effects fit. 
+```{r}
+  betahat = simA$betahat[[1]]
+  betahatsd = simA$betahatsd[[1]]
+  set.seed(111) #to ensure reproducibility
+  fit.mixfdr.P0 = mixFdr(betahat/betahatsd,noiseSD=1,theonull=TRUE,P=0,plot=FALSE)
+  cdf.mixfdr = function(a,x){mixcdf(normalmix(a$pi,a$mu,a$sigma-1),x)}
+  x = seq(-4,4,length=100)
+
+  res = data.frame(beta=x,
+                  mixfdr.P0 = t(cdf.mixfdr(fit.mixfdr.P0,x)),
+                  mixfdr = t(cdf.mixfdr(simA$fit.mixfdr[[1]],x)),
+                  ash.n = t(cdf.ash(simA$fit.ash.n[[1]],x)$y),
+                  ash.fdr.n=t(cdf.ash(simA$fit.ash.fdr.n[[1]],x)$y))
+
+  truth = t(cdf.ash(simA$fit.ash.true[[1]],x)$y)
+
+  res.melt=melt(res, id.vars=c("beta"),variable.name="Method")   
+  res.melt$Penalization= ifelse((res.melt$Method=="mixfdr" | res.melt$Method=="ash.fdr.n"),"Default","Minimal")
+  res.melt$Methodtype=as.character(res.melt$Method)
+  res.melt$Methodtype[res.melt$Method=="mixfdr.P0"]="mixfdr"
+  res.melt$Methodtype[res.melt$Method=="ash.fdr.n"]="ash"
+  res.melt$Methodtype[res.melt$Method=="ash.n"]="ash"
+
+cbbPalette <- c("#56B4E9", "#009E73", "#000000","#F0E442", "#0072B2", "#E69F00",  "#D55E00", "#CC79A7")
+
+pdf("figures/penaltycompare.pdf")
+ggplot(res.melt, aes(x=beta)) +
+  geom_line(aes(x=beta,y=value,color=Methodtype,linetype=Penalization,group=Method),size=1.5,alpha=0.8) +
+  geom_line(data=res, aes(y=truth,color="truth"),alpha=1,size=0.5) +
+  scale_colour_manual(name = 'Method', values = cbbPalette)
+dev.off()
+```

### 2014-02-26 paper/Rcode/ashsim.R

@@ -1,0 +1,69 @@
+#simulate from mixture of normals
+#compare normal, uniform and half-uniform mixtures
+#bsd gives standard deviation of beta
+require(ashr)
+require(mixfdr)
+
+ashsim=function(mixmean,mixsd,mixpi,bsd=1,seedval = 100,nsamp=1000,niter=50){  
+  set.seed(seedval)  
+  beta =list()
+  betahatsd=list()
+  betahat = list()
+  fit.ash.n = list()
+  fit.ash.u = list()
+  fit.ash.hu = list()
+  fit.ash.hu.vb=list()
+  fit.ash.true = list()
+  fit.mixfdr= list()
+  fit.ash.fdr.n = list()
+  
+  fit.mixfdr.enull= list()
+  fit.mixfdr.J10= list()
+  fit.mixfdr.J10P0= list()
+  fit.mixfdr.J100= list()
+  k = length(mixmean)
+  for(i in 1:niter){
+    comp = sample(1:k,nsamp,prob=mixpi,replace=TRUE)
+    sd = mixsd[comp]
+    mean = mixmean[comp]
+    beta[[i]] = rnorm(nsamp,mean,sd)
+    betahatsd[[i]] = bsd
+    betahat[[i]] = beta[[i]]+rnorm(nsamp,0,betahatsd[[i]])
+    fit.ash.n[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="normal",method="shrink")
+    fit.ash.u[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="uniform", method="shrink")
+    fit.ash.hu[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform", method="shrink")
+    fit.ash.hu.vb[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform",method="shrink",VB=TRUE)
+    fit.ash.true[[i]] = ash(betahat[[i]],betahatsd[[i]],g=normalmix(mixpi,mixmean,mixsd))
+    fit.mixfdr[[i]] = mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=TRUE,plot=FALSE)
+    
+    fit.ash.fdr.n[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="normal",method="fdr")
[…]

### 2014-02-26 paper/Rcode/plot_examples.R

@@ -1,0 +1,68 @@
+require(ggplot2)
+
+cdf.mixfdr = function(a,x){mixcdf(normalmix(a$pi,a$mu,a$sigma-1),x)}
+
+plot_examples = function(sims){
+  sim1= sims[[1]]
+  sim2= sims[[2]]
+  sim3= sims[[3]]
+  len=length(sim1$beta[[1]])
+  res = data.frame(beta = c(sim1$beta[[1]],sim2$beta[[1]],sim3$beta[[1]]), Scenario = c(rep(1,len),rep(2,len),rep(3,len)))
+  x = seq(-8,6,length=100)
+  y = dnorm(x)
+  dens1 = data.frame(x=x,dtrue=t(density(sim1$fit.ash.true[[1]],x)$y), Scenario=1)
+  dens2 = data.frame(x=x,dtrue=t(density(sim2$fit.ash.true[[1]],x)$y), Scenario=2)
+  dens3 = data.frame(x=x,dtrue=t(density(sim3$fit.ash.true[[1]],x)$y), Scenario=3)
+  dens=rbind(dens1,dens2,dens3)
+  ggplot(res) + facet_grid(.~Scenario)  + 
+    geom_line(data = dens, aes(x,dtrue),size=1.2,alpha=0.9,linetype=1) + scale_x_continuous(limits=c(-6,6))
+}
+
+plot_examples_cdf_withfit = function(sims){
+  sim1= sims[[1]]
+  sim2= sims[[2]]
+  sim3= sims[[3]]
+  len=length(sim1$beta[[1]])
+  res = data.frame(beta = c(sim1$beta[[1]],sim2$beta[[1]],sim3$beta[[1]]), Scenario = c(rep(1,len),rep(2,len),rep(3,len)))
+  x = seq(-6,6,length=100)
+  y = dnorm(x)
+  cdf1 = data.frame(x=x,dn=t(cdf.ash(sim1$fit.ash.n[[1]],x)$y),du=t(cdf.ash(sim1$fit.ash.u[[1]],x)$y),dhu=t(cdf.ash(sim1$fit.ash.hu[[1]],x)$y), dtrue=t(cdf.ash(sim1$fit.ash.true[[1]],x)$y), dn.fdr=t(cdf.ash(sim1$fit.ash.fdr.n[[1]],x)$y),dm = t(cdf.mixfdr(sim1$fit.mixfdr[[1]],x)), Scenario=1)
[…]

### 2014-02-26 paper/Rcode/sim1.R

@@ -66,7 +66,7 @@ basicsim=function(mixsd,mixpi_alt,bsd=1,minpi0=0,seedval = 100,nsamp=1000,niter=
 mixsd = c(0,0.25,0.5,1,2)
 mixpi_alt = c(0.4,0.2,0.2,0.2) #mixture proportions under the alternative
 
-#simres1 = basicsim(mixsd,mixpi_alt,niter=50,nsamp=1000)
+set.seed(100)
 simres1 = basicsim(mixsd,mixpi_alt,niter=200,nsamp=1000)
 simres1a = basicsim(mixsd,mixpi_alt,niter=200,nsamp=10000)
 simres2= basicsim(c(0,4),c(1))

### 2014-02-26 paper/Rcode/ashsim.rmd
updated mix.R to export loglik_conv.default

@@
   source("plot_rmse.R")
-```
-
-```{r}
-pdf("figures/simABC_egdens.pdf",width=6.5,height=2)
-plot_examples(list(simA,simB,simC))
-dev.off()
-```
-
-```{r}
-  pdf("figures/rmse_biplot.pdf",width=6.5,height=2)
-    plot_rmse(list(simA,simB,simC))
-  dev.off()
-
-#This figure not used in paper?
-pdf("figures/rmse_biplot_withzerobetahat.pdf")
-    plot_rmse(list(simA,simB,simC),inczero=TRUE,incbetahat=TRUE)
-dev.off()
-
-```
-
-```{r}
-
-  pdf("figures/rmse_boxplot.pdf",width=6.5,height=2)
-    plot_rmse_boxplot(list(simA,simB,simC))
-  dev.off()
-
-    pdf("figures/rmse_boxplot_extended.pdf",width=6.5,height=2)
-    plot_rmse_boxplot(list(simA,simB,simC),TRUE,TRUE,TRUE)
-  dev.off()
-
-```
-
-```{r}
-
-  pdf("figures/loglik_boxplot.pdf",width=6.5,height=2)
-    plot_loglik_boxplot(list(simA,simB,simC))
-  dev.off()
-
-```
-
-```{r}
-
-  pdf("figures/rmse_loglik_boxplot.pdf",width=6.5,height=5)
-    plot_rmse_loglik_boxplot(list(simA,simB,simC))
-  dev.off()
-
 ```
@@
 plot(ecdf(simC$beta[[1]]),xlim=c(-6,6))
+x= seq(-6,6,length=100)
 lines(x,cdf.mixfdr(temp,x),col=3)
@@
 lines(cdf.ash(simC$fit.ash.hu[[1]],x),col=2,lty=2)
-```
-
-```{r}
-  plot_LR=function(sims){
-    hist(unlist(lapply(sims$fit.ash.u,get_loglik))-unlist(lapply(sims$fit.ash.n,get_loglik)), xlab="loglik difference", main="loglik differences for nullbiased prior vs mle",nclass=10)
-  }
-
-  pdf("figures/logLR.pdf")
-    plot_LR(simA)
-    plot_LR(simB)
-  dev.off()
-```
-
-## Unused figures?
-
-```{r}
-
-pdf("figures/simABC_eg_withfit.pdf",width=6.5,height=2)
-plot_examples_withfit(list(simA,simB,simC))
-dev.off()
-```
-
-```{r}
-pdf("figures/simABC_eg_cdf_withfit.pdf",width=6.5,height=2)
-plot_examples_cdf_withfit(list(simA,simB,simC))
-dev.off()
 ```
@@
 
-pdf("figures/penaltycompare.pdf")
+#pdf("figures/penaltycompare.pdf")
 ggplot(res.melt, aes(x=beta)) +
@@
   scale_colour_manual(name = 'Method', values = cbbPalette)
-dev.off()
+#dev.off()
 ```

### 2014-02-26 paper/Rcode/fdr.examples.rmd
rearranged simulations; updated makefigs.R

--- 2014-02-24 paper/Rcode/conservativepm.rmd
@@
-#look into methods for estimating pi0 conservatively
-
-Do some simulations
-
 ```{r}
@@
 require(ggplot2)
-#bsd gives standard deviation of beta
-#pi0 is simulated to be uniform on [minpi0,1]
-basicsim=function(mixsd,mixpi_alt,bsd=1,minpi0=0,seedval = 100,nsamp=1000,niter=50){  
-  set.seed(seedval)  
-  beta =list()
-  betahatsd=list()
-  betahat = list()
-  zscore = list()
-  pval = list()
-  betahat.ash.n = list()
-  betahat.ash.u = list()
-  betahat.ash.npm = list()
-  betahat.ash.true = list()
-  betahat.qval = list()
-  betahat.fdrtool = list()
-  betahat.locfdr = list()
-  betahat.mixfdr = list()
-  pi0 = rep(0,niter)
-  for(i in 1:niter){
-    pi0[i]=runif(1,minpi0,1)
-    mixpi = c(pi0[i],(1-pi0[i])*mixpi_alt)
-    sd = sample(mixsd,nsamp,prob=mixpi,replace=TRUE )
-    beta[[i]] = rnorm(nsamp,0,sd)
-    betahatsd[[i]] = bsd
-    betahat[[i]] = beta[[i]]+rnorm(nsamp,0,betahatsd[[i]])
-    zscore[[i]] = betahat[[i]]/betahatsd[[i]]
-    pval[[i]] = pchisq(zscore[[i]]^2,df=1,lower.tail=F)
-    betahat.ash.n[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=TRUE,prior="nullbiased",gridmult=2)
-    betahat.ash.u[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=TRUE,prior="uniform",gridmult=2)
-    betahat.ash.npm[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=FALSE,prior="uniform",gridmult=2)
-    betahat.ash.true[[i]] = ash(betahat[[i]],betahatsd[[i]],g=normalmix(mixpi,rep(0,length(mixpi)),mixsd))
-
-    cat("applying q value\n")
-    betahat.qval[[i]] = qvalue(pval[[i]])
-    cat("applying fdrtool\n")
-    betahat.fdrtool[[i]] = fdrtool(pval[[i]],statistic="pvalue",plot=FALSE)
-    cat("applying locfdr\n")
-    betahat.locfdr[[i]] = try(locfdr(zscore[[i]],nulltype=0,plot=0))
-    cat("applying mixfdr\n")
-    betahat.mixfdr[[i]] = mixFdr(zscore[[i]],noiseSD=1,theonull=TRUE,plot=FALSE)
-  }
-  return(list(beta =beta,
-  betahatsd=betahatsd,
-  betahat = betahat,
-  zscore = zscore,
-  pval = pval,
-  betahat.ash.n = betahat.ash.n,
-  betahat.ash.u = betahat.ash.u,
-  betahat.ash.npm = betahat.ash.npm,
-  betahat.ash.true=betahat.ash.true,
-  betahat.qval = betahat.qval,
-  betahat.fdrtool = betahat.fdrtool,
-  betahat.locfdr = betahat.locfdr,
-  betahat.mixfdr = betahat.mixfdr,
-    pi0=pi0))
-}    
 ```
 
-```{r, cache=TRUE}
-mixsd = c(0,0.25,0.5,1,2)
-mixpi_alt = c(0.4,0.2,0.2,0.2) #mixture proportions under the alternative
-
-simres1 = basicsim(mixsd,mixpi_alt,niter=200,nsamp=1000)
-simres1a = basicsim(mixsd,mixpi_alt,niter=200,nsamp=10000)
-simres2= basicsim(c(0,4),c(1))
-simres3=basicsim(c(0,2),c(1),bsd=c(rep(1,500),rep(10,500)))
-simres4=basicsim(c(0,2),c(1),bsd=c(rep(1,500),rep(10,500)),minpi0=0.9,seed=200)
+```{r}
+load("sim1.RData")
 ```
@@
 ```
-
-```{r}
-#  par(mfcol=c(3,3))
-  plot_ecdf=function(sims){
-    for(i in 1:length(sims$beta)){
-       plot(ecdf(sims$beta[[i]]),xlim=c(-6,6),main=paste0("iteration ",i))
-       x = seq(-6,6,length=1000)
-       lines(cdf.ash(sims$betahat.ash.n[[i]],x),col=2,lwd=2)
-       lines(cdf.ash(sims$betahat.ash.u[[i]],x),col=3,lwd=2)
-       lines(cdf.ash(sims$betahat.ash.true[[i]],x),col=4,lwd=2)
-    }
-  }
-  plot_ecdf(simres1)
-  plot_ecdf(simres2)
-```
-
-```{r}
-#Plot pi0 from each method
-source("plotpi0.R")
-
-  plot_pi1 = function(sims){
-    pi0=sims$pi0  
-    pi0_ash.n=unlist(lapply(sims$betahat.ash.n,get_pi0))
-    pi0_ash.u=unlist(lapply(sims$betahat.ash.u,get_pi0)) 
-    pi0_fdrtool = unlist(lapply(sims$betahat.fdrtool,get_pi0.fdrtool))
-    pi0_locfdr=unlist(lapply(sims$betahat.locfdr,get_pi0.locfdr))
-    pi0_mixfdr = unlist(lapply(sims$betahat.mixfdr,get_pi0.mixfdr))
-    pi0_qval = unlist(lapply(sims$betahat.qval,"[[","pi0"))
-
-    res = data.frame(pi0=pi0,qvalue=pi0_qval,mixfdr=pi0_mixfdr, locfdr=pi0_locfdr, fdrtool=pi0_fdrtool,ash.nullbiased=pi0_ash.n,ash.uniform=pi0_ash.u)
-    require(reshape2)
-    res.melt = melt(res, id.vars=c("pi0"),variable.name="Method")
-    p=ggplot(data=res.melt,aes(1-pi0,log2((1-value)/(1-pi0)),colour=Method)) +geom_point(shape=16) +
-#        geom_abline(colour = "black") +
-        xlab("True pi1") +
-        ylab("log2(Estimated pi1/True pi1)")
-    print(p +scale_y_continuous(limits=c(-4,4)) +
-          scale_x_continuous(limits=c(0,1)))
-
-  }
-
-  png("figures/estpi0_sim1sim2.png",height=160,width=540)
-    plot_pi0(list(simres1,simres2))
-  dev.off()
-
-```
-
-Figure to show that estimated betahats are not so different
-```{r}
-  plot(betahat.ash.u[[1]]$PosteriorMean,betahat.ash.n[[1]]$PosteriorMean)
-  abline(a=0,b=1,lwd=2,col=2)
-  plot(betahat.ash.u[[1]]$PosteriorSD,betahat.ash.n[[1]]$PosteriorSD)
-  abline(a=0,b=1,lwd=2,col=2)
-```
-
-```{r}
-source("plot_lfsr.R")
-```
-
-```{r}
-png("figures/lfsdr_sim1sim2_blowup.png",height=427,width=720)
-  plot_lfsdr(list(simres1,simres1a,simres2),0.1,ptype="lfsra")
-dev.off()
-```
-
-```{r}
-png("figures/lfsdr_sim1sim2_blowup.png",height=427,width=720)
-  plot_lfsdr(list(simres1,simres1a,simres2),0.1,ymax=0.2,ptype="lfsra")
-dev.off()
-```
-
-```{r}
-png("figures/lfdr_sim1sim2_blowup.png",height=160,width=540)
-  plot_lfsr(list(simres1,simres1a,simres2),0.1,ptype="lfdr")
-dev.off()
-```
-
-```{r}
-png("figures/lfsra_sim1sim2_blowup.png",height=160,width=540)
-  plot_lfsr(list(simres1,simres1a,simres2),0.1,ptype="lfsra")
-dev.off()
-```
-
-```{r}
-png("figures/lfsra_sim1sim2_blowup.png",height=160,width=540)
-  plot_lfsr(list(simres1,simres1a,simres2),0.1,ptype="lfsr")
-dev.off()
-```
-
-```{r}
-    plot_pi0(simres3)
-  plot_pi1(simres3)
-```
-
-QUestion: is the null-biased prior maybe a little too conservative?
-Answer: log likelihoods don't suggest they are
-
-```{r}
-#hh.ashtrue = hh.ashz
-#hh.ashtrue$fitted.g$pi = c(2/3,1/15,1/15,1/15,1/15,1/15)
-#hh.ashtrue$fitted.g$mean = c(0,0,0,0,0,0)
-#hh.ashtrue$fitted.g$sd = sqrt(c(0,1,0.2,0.4,0.8,3))
-
-#loglik(hh.ashtrue,betahat,sebetahat)
-#loglik(hh.ashz,betahat,sebetahat)
-```

### 2014-02-26 paper/Rcode/ashsim.R

@@ -1,69 +1,0 @@
-#simulate from mixture of normals
-#compare normal, uniform and half-uniform mixtures
-#bsd gives standard deviation of beta
-require(ashr)
-require(mixfdr)
-
-ashsim=function(mixmean,mixsd,mixpi,bsd=1,seedval = 100,nsamp=1000,niter=50){  
-  set.seed(seedval)  
-  beta =list()
-  betahatsd=list()
-  betahat = list()
-  fit.ash.n = list()
-  fit.ash.u = list()
-  fit.ash.hu = list()
-  fit.ash.hu.vb=list()
-  fit.ash.true = list()
-  fit.mixfdr= list()
-  fit.ash.fdr.n = list()
-  
-  fit.mixfdr.enull= list()
-  fit.mixfdr.J10= list()
-  fit.mixfdr.J10P0= list()
-  fit.mixfdr.J100= list()
-  k = length(mixmean)
-  for(i in 1:niter){
-    comp = sample(1:k,nsamp,prob=mixpi,replace=TRUE)
-    sd = mixsd[comp]
-    mean = mixmean[comp]
-    beta[[i]] = rnorm(nsamp,mean,sd)
-    betahatsd[[i]] = bsd
-    betahat[[i]] = beta[[i]]+rnorm(nsamp,0,betahatsd[[i]])
-    fit.ash.n[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="normal",method="shrink")
-    fit.ash.u[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="uniform", method="shrink")
-    fit.ash.hu[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform", method="shrink")
-    fit.ash.hu.vb[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform",method="shrink",VB=TRUE)
-    fit.ash.true[[i]] = ash(betahat[[i]],betahatsd[[i]],g=normalmix(mixpi,mixmean,mixsd))
-    fit.mixfdr[[i]] = mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=TRUE,plot=FALSE)
-    
-    fit.ash.fdr.n[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="normal",method="fdr")
[…]

### 2014-02-26 paper/Rcode/makefigs.R

@@ -1,0 +1,94 @@
+require(ashr)
+require(qvalue)
+require(fdrtool)
+require(mixfdr)
+require(locfdr)
+require(ggplot2)   
+
+load("sim1_out.RData")
+load("ashsim_out.RData")
+source("plot_examples.R")
+source("plot_rmse.R")
+source("plot_pi0.R")
+source("plot_lfsr.R")
+
+pdf("figures/simABC_egdens.pdf",width=6.5,height=2)
+plot_examples(list(simA,simB,simC))
+dev.off()
+
+pdf("figures/rmse_biplot.pdf",width=6.5,height=2)
+plot_rmse(list(simA,simB,simC))
+dev.off()
+
+#This figure not used in paper?
+pdf("figures/rmse_biplot_withzerobetahat.pdf")
+plot_rmse(list(simA,simB,simC),inczero=TRUE,incbetahat=TRUE)
+dev.off()
+
+pdf("figures/rmse_boxplot.pdf",width=6.5,height=2)
+plot_rmse_boxplot(list(simA,simB,simC))
+dev.off()
+
+pdf("figures/rmse_boxplot_extended.pdf",width=6.5,height=2)
+plot_rmse_boxplot(list(simA,simB,simC),TRUE,TRUE,TRUE)
+dev.off()
+
+pdf("figures/loglik_boxplot.pdf",width=6.5,height=2)
+plot_loglik_boxplot(list(simA,simB,simC))
+dev.off()
+
+pdf("figures/rmse_loglik_boxplot.pdf",width=6.5,height=5)
+plot_rmse_loglik_boxplot(list(simA,simB,simC))
+dev.off()
+
+
+pdf("figures/simABC_eg_withfit.pdf",width=6.5,height=2)
+plot_examples_withfit(list(simA,simB,simC))
+dev.off()
+
+pdf("figures/simABC_eg_cdf_withfit.pdf",width=6.5,height=2)
+plot_examples_cdf_withfit(list(simA,simB,simC))
+dev.off()
+
+plot_LR=function(sims){
+  hist(unlist(lapply(sims$fit.ash.u,get_loglik))-unlist(lapply(sims$fit.ash.n,get_loglik)), xlab="loglik difference", main="loglik differences for nullbiased prior vs mle",nclass=10)
+}
+
+pdf("figures/logLR.pdf")
+plot_LR(simA)
+plot_LR(simB)
+dev.off()
+
+
+
+
[…]

### 2014-02-26 paper/Rcode/plot_pi0.R

@@ -38,3 +38,24 @@ plot_pi0 = function(sims){
           coord_equal(ratio=1))
   
 }
+
+plot_pi1 = function(sims){
+  pi0=sims$pi0  
+  pi0_ash.n=unlist(lapply(sims$betahat.ash.n,get_pi0))
+  pi0_ash.u=unlist(lapply(sims$betahat.ash.u,get_pi0)) 
+  pi0_fdrtool = unlist(lapply(sims$betahat.fdrtool,get_pi0.fdrtool))
+  pi0_locfdr=unlist(lapply(sims$betahat.locfdr,get_pi0.locfdr))
+  pi0_mixfdr = unlist(lapply(sims$betahat.mixfdr,get_pi0.mixfdr))
+  pi0_qval = unlist(lapply(sims$betahat.qval,"[[","pi0"))
+  
+  res = data.frame(pi0=pi0,qvalue=pi0_qval,mixfdr=pi0_mixfdr, locfdr=pi0_locfdr, fdrtool=pi0_fdrtool,ash.nullbiased=pi0_ash.n,ash.uniform=pi0_ash.u)
+  require(reshape2)
+  res.melt = melt(res, id.vars=c("pi0"),variable.name="Method")
+  p=ggplot(data=res.melt,aes(1-pi0,log2((1-value)/(1-pi0)),colour=Method)) +geom_point(shape=16) +
+    #        geom_abline(colour = "black") +
+    xlab("True pi1") +
+    ylab("log2(Estimated pi1/True pi1)")
+  print(p +scale_y_continuous(limits=c(-4,4)) +
+          scale_x_continuous(limits=c(0,1)))
+  
+}

### 2014-02-26 paper/Rcode/sim1.R

@@ -67,10 +67,13 @@ mixsd = c(0,0.25,0.5,1,2)
 mixpi_alt = c(0.4,0.2,0.2,0.2) #mixture proportions under the alternative
 
 set.seed(100)
-simres1 = basicsim(mixsd,mixpi_alt,niter=200,nsamp=1000)
-simres1a = basicsim(mixsd,mixpi_alt,niter=200,nsamp=10000)
+#these are the simulations for scenarios 1a, 1b and 2 from the paper
+simres1a = basicsim(mixsd,mixpi_alt,niter=200,nsamp=1000)
+simres1b = basicsim(mixsd,mixpi_alt,niter=200,nsamp=10000)
 simres2= basicsim(c(0,4),c(1))
+
+#these do a situation where precision varies across measurements
 simres3=basicsim(c(0,2),c(1),bsd=c(rep(1,500),rep(10,500)))
 simres4=basicsim(c(0,2),c(1),bsd=c(rep(1,500),rep(10,500)),minpi0=0.9,seed=200)
 
-save.image(file="sim1_out.RData")
+save.image(file="sim1.RData")

### 2014-02-26 paper/Rcode/simABC.R

@@ -1,0 +1,69 @@
+#simulate from mixture of normals
+#compare normal, uniform and half-uniform mixtures
+#bsd gives standard deviation of beta
+require(ashr)
+require(mixfdr)
+
+ashsim=function(mixmean,mixsd,mixpi,bsd=1,seedval = 100,nsamp=1000,niter=50){  
+  set.seed(seedval)  
+  beta =list()
+  betahatsd=list()
+  betahat = list()
+  fit.ash.n = list()
+  fit.ash.u = list()
+  fit.ash.hu = list()
+  fit.ash.hu.vb=list()
+  fit.ash.true = list()
+  fit.mixfdr= list()
+  fit.ash.fdr.n = list()
+  
+  fit.mixfdr.enull= list()
+  fit.mixfdr.J10= list()
+  fit.mixfdr.J10P0= list()
+  fit.mixfdr.J100= list()
+  k = length(mixmean)
+  for(i in 1:niter){
+    comp = sample(1:k,nsamp,prob=mixpi,replace=TRUE)
+    sd = mixsd[comp]
+    mean = mixmean[comp]
+    beta[[i]] = rnorm(nsamp,mean,sd)
+    betahatsd[[i]] = bsd
+    betahat[[i]] = beta[[i]]+rnorm(nsamp,0,betahatsd[[i]])
+    fit.ash.n[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="normal",method="shrink")
+    fit.ash.u[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="uniform", method="shrink")
+    fit.ash.hu[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform", method="shrink")
+    fit.ash.hu.vb[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform",method="shrink",VB=TRUE)
+    fit.ash.true[[i]] = ash(betahat[[i]],betahatsd[[i]],g=normalmix(mixpi,mixmean,mixsd))
+    fit.mixfdr[[i]] = mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=TRUE,plot=FALSE)
+    
+    fit.ash.fdr.n[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="normal",method="fdr")
[…]

### 2014-02-27 package/build.R
added rcpp-package directory

@@ -2,8 +2,13 @@
 #in R:
 library(Rcpp)
 Rcpp.package.skeleton("ashr", path=".", code_files=c("ash.oldversion.R", "ash.R", "mix.R"), cpp_files="MixEM.cpp", example_code = FALSE, attributes = TRUE)
+
+#require(roxygen2)
+#require(devtools)
+#roxygenize("~/Documents/git/ash/package/ashr/R/ashr")
+#document("~/Documents/git/ash/package/ashr/R/ashr")
+
 #in the terminal
-#cd ..
 #tar -pczf ashr.tar.gz ashr
 #in R
-install.packages("ash/package/ashr.tar.gz",repos=NULL,type="source")
+install.packages("~/Documents/git/ash/package/ashr/R/ashr.tar.gz",repos=NULL,type="source")

### 2014-02-27 paper/Rcode/makefigs.R

@@ -5,8 +5,8 @@ require(mixfdr)
 require(locfdr)
 require(ggplot2)   
 
-load("sim1_out.RData")
-load("ashsim_out.RData")
+load("sim1.RData")
+load("simABC.RData")
 source("plot_examples.R")
 source("plot_rmse.R")
 source("plot_pi0.R")

### 2014-02-27 rcpp-package/ashr/src/MixEM.cpp

@@ -1,0 +1,66 @@
+#include <Rcpp.h>
+#include <cmath> //not sure this is necessary
+#include <algorithm> 
+
+using namespace Rcpp;
+
+// [[Rcpp::export]]
+List cxxMixEM(NumericMatrix matrix_lik, NumericVector prior, NumericVector pi_init, double tol=0.0001, int maxiter=5000){//note: no default pi_init=NULL  
+  int n=matrix_lik.nrow(), k=matrix_lik.ncol(), j=0;
+  bool converged=NA_LOGICAL;
+  NumericVector pi(k);
+
+  if(Rf_isNull(pi_init))
+    std::fill(pi.begin(), pi.end(), 1./(double)k);
+  else{
+    pi=clone(pi_init);
+    for (int i=0;i<k;i++)//set any estimates that are very small to be very small    
+      pi[i]=std::max(1e-5, pi[i]); 
+    pi=pi/sum(pi); //normalize pi
+  }
+  NumericMatrix m(n,k);
+  NumericVector m_rowsum(n);
+  NumericMatrix classprob(m);
+  std::vector<double> loglik, priordens;
+  loglik.reserve(maxiter);
+  priordens.reserve(maxiter); 
+  
+  for (int i=0;i<k;i++){
+    m.column(i)=pi[i]*matrix_lik.column(i);
+    m_rowsum=m_rowsum+m.column(i);
+  }
+  for (int i=0;i<k;i++)//can vectorize this with sugar?
+    classprob.column(i)=classprob.column(i)/m_rowsum;
+  loglik.push_back(sum(log(m_rowsum)));
+  priordens.push_back(sum((prior-1.)*log(pi)));
+
+  for(j=1;j<maxiter;j++){
+    //update pi
+    //to do: can vectorize this with sugar?
+    for (int i=0;i<k;i++)//set any estimates that are less than zero, which can happen with prior<1, to 0
+      pi[i]=std::max(1e-5, sum(classprob.column(i))+prior[i]-1.);
+    pi=pi/sum(pi); //normalize pi
+    
+    //Now re-estimate pi
+    std::fill(m_rowsum.begin(), m_rowsum.end(), 0);
+    for (int i=0;i<k;i++){
+      m.column(i)=pi[i]*matrix_lik.column(i);
+      m_rowsum=m_rowsum+m.column(i);
+    }
+    for (int i=0;i<k;i++)
+      classprob.column(i)=classprob.column(i)/m_rowsum;
[…]

### 2014-02-27 tests/test_cxx/test.R

@@ -1,4 +1,4 @@
-install.packages("../../package/ashr.tar.gz",repos=NULL,type="source")
+install.packages("../../rcpp-package/ashr_0.1.tar.gz",repos=NULL,type="source")
 library(ashr)
 
 # simulate n beta-hat values, nnull under the null with altmean and altsd
@@ -26,6 +26,7 @@ Rprof(outfile)
 system.time(beta.ash <- ash(ss$betahat, ss$betasd, prior = "uniform", cxx =cxx))
 Rprof(NULL)
 
+
 # check that return identical values (up to machine error)
 all.equal(beta.ash$PosteriorMean, beta.ash.cxx$PosteriorMean)
 all.equal(beta.ash$qvalue, beta.ash.cxx$qvalue)
@@ -34,4 +35,22 @@ all.equal(beta.ash$PositiveProb, beta.ash.cxx$PositiveProb)
 #to see profiling output run this in the terminal
 #R CMD Rprof ash.out > ash.out.txt
 #R CMD Rprof ash.cxx.out > ash.cxx.out.txt
+
+# repeat using halfuniform
+cxx=TRUE
+outfile="ash.cxx.out"
+Rprof(outfile)
+system.time(beta.ash.cxx <- ash(ss$betahat, ss$betasd, prior = "uniform", mixcompdist="halfuniform",cxx =cxx))
+Rprof(NULL)
+
+cxx=FALSE
+outfile="ash.out"
+Rprof(outfile)
+system.time(beta.ash <- ash(ss$betahat, ss$betasd, prior = "uniform", mixcompdist="halfuniform", cxx =cxx))
+Rprof(NULL)
+
+# check that return identical values (up to machine error)
+all.equal(beta.ash$PosteriorMean, beta.ash.cxx$PosteriorMean)
+all.equal(beta.ash$qvalue, beta.ash.cxx$qvalue)
+all.equal(beta.ash$PositiveProb, beta.ash.cxx$PositiveProb)
 

### 2014-03-04 Rcode/readme.rmd

@@
 ```{r}
-hh.ash = ash(hh.betahat,hh.sebetahat,df=n1+n2-2)
+hh.ash = ash(hh.betahat,hh.sebetahat,df=n1+n2-2,method="fdr")
 sum(hh.ash$qvalue<0.05)
-plot(hh.ash$localfdr, log(hh.pval),main="Illustration of how ASH reorders significance of observations",ylim=c(-10,-1),xlim=c(0,0.1),cex=0.2,xlab="ASH local fdr",ylab="Original p value")
+plot(hh.ash$lfdr, log(hh.pval),main="Illustration of how ASH reorders significance of observations",ylim=c(-10,-1),xlim=c(0,0.1),cex=0.2,xlab="ASH local fdr",ylab="Original p value")
 ```
@@
 ```{r}
-hh.ash2 = ash(hh.zscore,rep(1,length(hh.zscore)),df=20)
-plot(hh.ash$localfdr,hh.ash2$localfdr,cex=0.2,ylab="ignoring sd",xlab="accounting for sd",main="Effect of accounting for measurement precision on local fdr")
+hh.ash2 = ash(hh.zscore,rep(1,length(hh.zscore)),df=20,method="fdr")
+plot(hh.ash$lfdr,hh.ash2$lfdr,cex=0.2,ylab="ignoring sd",xlab="accounting for sd",main="Effect of accounting for measurement precision on local fdr")
 abline(a=0,b=1,col=2)
 cat(sum(hh.ash$qvalue<0.05),sum(hh.ash2$qvalue<0.05))
-cat(sum(hh.ash$localfdr<0.05),sum(hh.ash2$localfdr<0.05))
-plot(hh.ash2$localfdr, log(hh.pval),main="Illustration of how ASH without precision does not reorder significance?",ylim=c(-10,-1),xlim=c(0,0.1),cex=0.2,xlab="ASH local fdr",ylab="Original p value")
+cat(sum(hh.ash$lfdr<0.05),sum(hh.ash2$lfdr<0.05))
+plot(hh.ash2$lfdr, log(hh.pval),main="Illustration of how ASH without precision does not reorder significance?",ylim=c(-10,-1),xlim=c(0,0.1),cex=0.2,xlab="ASH local fdr",ylab="Original p value")
 ```
@@
 ```{r}
-cat(LogLik_conv(hh.ash$fitted.g, hh.betahat, hh.sebetahat),mixseLoglik(hh.betahat,hh.ash$fitted.g$pi, hh.ash$fitted.g$mean, hh.ash$fitted.g$sd, hh.sebetahat),mixseLoglik(hh.betahat,hh.ash2$fitted.g$pi, hh.ash2$fitted.g$mean, sqrt(hh.ash2$fitted.g$sd^2+1), hh.sebetahat,FUN="*"))
+cat(loglik.ash(hh.ash, hh.betahat, hh.sebetahat),loglik.ash(hh.ash2,hh.betahat,hh.sebetahat,TRUE))
 ```
@@
 zsim2 = musim2+rnorm(10000)
-zsim2.ash=ash(zsim2,rep(1,10000))
+zsim2.ash=ash(zsim2,rep(1,10000),method="fdr")
 hist(musim2[1:3000])
@@
 zsim = musim+rnorm(10000)
-zsim.ash=ash(zsim,rep(1,10000))
+zsim.ash=ash(zsim,rep(1,10000),method="fdr")
 ```
@@
 
-test1.ash1 = ash(bhat.test, hh.sebetahat)
-test1.ash2 = ash(z.test, rep(1,n))
+test1.ash1 = ash(bhat.test, hh.sebetahat,method="fdr")
+test1.ash2 = ash(z.test, rep(1,n),method="fdr")
 cat(sum(test1.ash1$qval<0.05),sum(test1.ash2$qval<0.05))
 
-test2.ash1 = ash(bhat2.test, hh.sebetahat)
-test2.ash2 = ash(z2.test, rep(1,n))
+test2.ash1 = ash(bhat2.test, hh.sebetahat,method="fdr")
+test2.ash2 = ash(z2.test, rep(1,n),method="fdr")
 cat(sum(test2.ash1$qval<0.05),sum(test2.ash2$qval<0.05))

### 2014-03-04 paper/Rcode/conservativepm.rmd

@@
 require(ggplot2)
-#bsd gives standard deviation of beta
-#pi0 is simulated to be uniform on [minpi0,1]
-basicsim=function(mixsd,mixpi_alt,bsd=1,minpi0=0,seedval = 100,nsamp=1000,niter=50){  
-  set.seed(seedval)  
-  beta =list()
-  betahatsd=list()
-  betahat = list()
-  zscore = list()
-  pval = list()
-  betahat.ash.n = list()
-  betahat.ash.u = list()
-  betahat.ash.npm = list()
-  betahat.ash.true = list()
-  betahat.qval = list()
-  betahat.fdrtool = list()
-  betahat.locfdr = list()
-  betahat.mixfdr = list()
-  pi0 = rep(0,niter)
-  for(i in 1:niter){
-    pi0[i]=runif(1,minpi0,1)
-    mixpi = c(pi0[i],(1-pi0[i])*mixpi_alt)
-    sd = sample(mixsd,nsamp,prob=mixpi,replace=TRUE )
-    beta[[i]] = rnorm(nsamp,0,sd)
-    betahatsd[[i]] = bsd
-    betahat[[i]] = beta[[i]]+rnorm(nsamp,0,betahatsd[[i]])
-    zscore[[i]] = betahat[[i]]/betahatsd[[i]]
-    pval[[i]] = pchisq(zscore[[i]]^2,df=1,lower.tail=F)
-    betahat.ash.n[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=TRUE,prior="nullbiased",gridmult=2)
-    betahat.ash.u[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=TRUE,prior="uniform",gridmult=2)
-    betahat.ash.npm[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=FALSE,prior="uniform",gridmult=2)
-    betahat.ash.true[[i]] = ash(betahat[[i]],betahatsd[[i]],g=normalmix(mixpi,rep(0,length(mixpi)),mixsd))
-
-    cat("applying q value\n")
-    betahat.qval[[i]] = qvalue(pval[[i]])
-    cat("applying fdrtool\n")
-    betahat.fdrtool[[i]] = fdrtool(pval[[i]],statistic="pvalue",plot=FALSE)
-    cat("applying locfdr\n")
-    betahat.locfdr[[i]] = try(locfdr(zscore[[i]],nulltype=0,plot=0))
-    cat("applying mixfdr\n")
-    betahat.mixfdr[[i]] = mixFdr(zscore[[i]],noiseSD=1,theonull=TRUE,plot=FALSE)
-  }
-  return(list(beta =beta,
-  betahatsd=betahatsd,
-  betahat = betahat,
-  zscore = zscore,
-  pval = pval,
-  betahat.ash.n = betahat.ash.n,
-  betahat.ash.u = betahat.ash.u,
-  betahat.ash.npm = betahat.ash.npm,
-  betahat.ash.true=betahat.ash.true,
-  betahat.qval = betahat.qval,
-  betahat.fdrtool = betahat.fdrtool,
-  betahat.locfdr = betahat.locfdr,
-  betahat.mixfdr = betahat.mixfdr,
-    pi0=pi0))
-}    
-```
-
-```{r, cache=TRUE}
-mixsd = c(0,0.25,0.5,1,2)
-mixpi_alt = c(0.4,0.2,0.2,0.2) #mixture proportions under the alternative
-
-simres1 = basicsim(mixsd,mixpi_alt,niter=200,nsamp=1000)
-simres1a = basicsim(mixsd,mixpi_alt,niter=200,nsamp=10000)
-simres2= basicsim(c(0,4),c(1))
-simres3=basicsim(c(0,2),c(1),bsd=c(rep(1,500),rep(10,500)))
-simres4=basicsim(c(0,2),c(1),bsd=c(rep(1,500),rep(10,500)),minpi0=0.9,seed=200)
-```
-
-Show illustrative example
-```{r}
-altcol="cyan" #colors to use
-nullcol="blue" 
-nc=40 #number of bins in histograms
-ncz = 100 # number of bins in z score histograms
-
-plot_FDR_hist=function(sim,iter=1){
-  hh.pval = sim$pval[[iter]]
-  hh.zscore = sim$zscore[[iter]]
-  hh.hist =hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)
-
-  hh.q = qvalue(hh.pval)
-  abline(h=hh.q$pi0,col=nullcol,lwd=2)
-
-  hh.hist$density=rep(hh.q$pi0,length(hh.hist$density))  
-  plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)
-
-  abline(v=0.1,lwd=2,col=2)
-
-  text(0.05,1.2,labels="A",col=2,cex=1.2)  
-  text(0.05,0.4,labels="B",col=2,cex=1.2)  
-  text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(hh.q$pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
-}
-plot_FDR_hist(simres1,1)
-
 ```
@@
 ```{r}
-plot_lfdr_hist=function(sim,iter=1){
-  require(fdrtool)
-  hh.pval=sim$pval[[iter]]
-  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)
-
-  hh.gren = grenander(ecdf(hh.pval))
-  abline(h=min(hh.gren$f.knots),col=nullcol,lwd=2)  
-  lines(hh.gren$x.knots,hh.gren$f.knots,lwd=2)
-  abline(v=0.1,lwd=2,col=2)
-  text(0.1,0.9,labels="a",col=2,cex=1)  
-  text(0.1,0.34,labels="b",col=2,cex=1.2)  
-  text(0.6,3,labels=paste0("lfdr = b/(a+b) =  ",round(min(hh.gren$f.knots)/approx(hh.gren$x.knots,hh.gren$f.knots,0.1)$y,2)),cex=1.2)
-}
-plot_lfdr_hist(simres1,1)
+load("sim1.RData")
 ```
 
-```{r}
-#plot a histogram of z scores, highlighting the alternative distribution
-#of z scores that is implied by localfdr values lfdr.
-  nullalthist = function(z,lfdr,...){
-    h=hist(z, freq=FALSE,col=nullcol,nclass=ncz,...)
-    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
-    h$density = (1-avlfdr) * h$density
-    plot(h,add=TRUE,col=altcol,freq=FALSE)
-  }
-
-#this one puts the null on the bottom
-  altnullhist = function(z,lfdr,...){
-    h=hist(z, freq=FALSE,col=altcol,nclass=ncz,...)
-    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
-    h$density = avlfdr * h$density
-    plot(h,add=TRUE,col=nullcol,freq=FALSE)
-  }
-
-  plotall_hist=function(sim,iter=1,histfun=nullalthist){
-    hh.zscore=sim$zscore[[iter]]    
-    par(mfcol=c(2,2))
-    histfun(hh.zscore,sim$betahat.fdrtool[[iter]]$lfdr,main="fdrtool")  
-    histfun(hh.zscore,sim$betahat.locfdr[[iter]]$fdr,main="locfdr")
-    histfun(hh.zscore,sim$betahat.mixfdr[[iter]]$fdr,main="mixfdr")
-    histfun(hh.zscore,sim$betahat.ash.n[[iter]]$lfdr,main="ash")
-    par(mfcol=c(1,1))
-  }
-
- # pdf("figures/nullalthist.pdf")
-  plotall_hist(simres1,1,nullalthist)
-#  dev.off()
-
- # pdf("figures/altnullhist.pdf")
-  plotall_hist(simres1,1,altnullhist)
-#  dev.off()
-
-```
+#look into methods for estimating pi0 conservatively
 
@@
 
-```{r}
-#Plot pi0 from each method
-source("plotpi0.R")
-
-  plot_pi1 = function(sims){
-    pi0=sims$pi0  
-    pi0_ash.n=unlist(lapply(sims$betahat.ash.n,get_pi0))
-    pi0_ash.u=unlist(lapply(sims$betahat.ash.u,get_pi0)) 
-    pi0_fdrtool = unlist(lapply(sims$betahat.fdrtool,get_pi0.fdrtool))
-    pi0_locfdr=unlist(lapply(sims$betahat.locfdr,get_pi0.locfdr))
-    pi0_mixfdr = unlist(lapply(sims$betahat.mixfdr,get_pi0.mixfdr))
-    pi0_qval = unlist(lapply(sims$betahat.qval,"[[","pi0"))
-
-    res = data.frame(pi0=pi0,qvalue=pi0_qval,mixfdr=pi0_mixfdr, locfdr=pi0_locfdr, fdrtool=pi0_fdrtool,ash.nullbiased=pi0_ash.n,ash.uniform=pi0_ash.u)
-    require(reshape2)
-    res.melt = melt(res, id.vars=c("pi0"),variable.name="Method")
-    p=ggplot(data=res.melt,aes(1-pi0,log2((1-value)/(1-pi0)),colour=Method)) +geom_point(shape=16) +
-#        geom_abline(colour = "black") +
-        xlab("True pi1") +
-        ylab("log2(Estimated pi1/True pi1)")
-    print(p +scale_y_continuous(limits=c(-4,4)) +
-          scale_x_continuous(limits=c(0,1)))
-
-  }
-
-  png("figures/estpi0_sim1sim2.png",height=160,width=540)
-    plot_pi0(list(simres1,simres2))
-  dev.off()
-
-```
-
 Figure to show that estimated betahats are not so different
@@
 abline(a=0,b=1,lwd=2,col=2)
-  plot(betahat.ash.u[[1]]$PosteriorSD,betahat.ash.n[[1]]$PosteriorSD)
-  abline(a=0,b=1,lwd=2,col=2)
-```
-
-```{r}
-source("plot_lfsr.R")
-```
-
-```{r}
-png("figures/lfsdr_sim1sim2_blowup.png",height=427,width=720)
-  plot_lfsdr(list(simres1,simres1a,simres2),0.1,ptype="lfsra")
-dev.off()
-```
-
-```{r}
-png("figures/lfsdr_sim1sim2_blowup.png",height=427,width=720)
-  plot_lfsdr(list(simres1,simres1a,simres2),0.1,ymax=0.2,ptype="lfsra")
-dev.off()
-```
-
-```{r}
-png("figures/lfdr_sim1sim2_blowup.png",height=160,width=540)
-  plot_lfsr(list(simres1,simres1a,simres2),0.1,ptype="lfdr")
-dev.off()
-```
-
-```{r}
-png("figures/lfsra_sim1sim2_blowup.png",height=160,width=540)
-  plot_lfsr(list(simres1,simres1a,simres2),0.1,ptype="lfsra")
-dev.off()
-```
-
-```{r}
-png("figures/lfsra_sim1sim2_blowup.png",height=160,width=540)
-  plot_lfsr(list(simres1,simres1a,simres2),0.1,ptype="lfsr")
-dev.off()
-```
-
-```{r}
-    plot_pi0(simres3)
-  plot_pi1(simres3)
 ```

### 2014-03-28 examples/fsr.rmd
added examples

# FSR Example

This is an example of how the false sign rate and the false discovery rate can
be very different when a large proportion of the nulls are false.

In this example, we assume that the true $\beta_j$ ($j=1,\dots,n$) are normally distributed, $\beta_j \sim N(0,1)$. 
So no $\beta_j$ is exactly 0, and all the null hypotheses $H_j: \beta_j=0$ are false.
Since all the nulls are false, no discovery is a false discovery,
and the ``true" FDR is always 0, whatever threshold is used.

```{r}
set.seed(100)
library(ashr)
library(qvalue)
n = 10000
beta = rnorm(n)
sebetahat = 1
betahat = beta + rnorm(n,0,sebetahat)

z = betahat/sebetahat
p = pchisq(z^2, df=1,lower.tail=FALSE)
z.ash = ash(z,1,method="fdr")
z.qvalue = qvalue(p)
```

In this example all the nulls are false, so the actual proportion of true nulls, $\pi_0$, is equal to 0. However, the estimated value of $\pi_0$ can be very different from zero. From qvalue we get
an estimated $\pi_0$ of `r z.qvalue$pi0`. This makes sense when you look at the
distribution of p values, and recall that qvalue makes the assumption that all $p$ values near 1 are null (see histogram below). In contrast, ash produces a much smaller estimate of $\pi_0$ 
($\pi0=$ `r get_pi0(z.ash)`), because it allows that many $p$ values near 1 may actually arise
even when $\beta_j \neq 0$. Of course, in this case both qvalue and ash provide "conservative"
estimates of $\pi_0$, but the qvalue estimate is much more conservative.

```{r}
  hist(p,prob=TRUE)
  abline(h=z.qvalue$pi0,lwd=2)
  text(0.95, z.qvalue$pi0+0.06, "pi0 (q value)")
  abline(h=get_pi0(z.ash),col=2,lwd=2)
  text(0.95, get_pi0(z.ash)+0.06,"pi0 (ash)",col=2)
```

Because of its very conservative estimate of $\pi_0$, qvalue also
provides much more conservative estimates of the FDR. For example, at a $p$ value threshold of 0.05, qvalue estimates the FDR to be `r max(z.qvalue$q[p<0.05])` whereas ash estimates the FDR
as `r max(z.ash$qvalue[p<0.05])`. Of course, the true value of the FDR is zero, so both are conservative.

Even though the true FDR is zero at any threshold, it seems, intuitively, that "discoveries" with
a $p$ value near 1 are not very meaningful. One way to formalize this is to consider the ``false sign rate" instead of the false discovery rate. The local false sign rate (lfsr) for each discovery is the probability that the true $\beta_j$ has the opposite sign to the corresponding $z$ score, $z_j$. The tail false sign rate (FSR) at a given threshold is the average of the lfsr values for discoveries that exceed that threshold. In this example, discoveries with a $p$ value near 1 will have a high lfsr, even though they have a local false discovery rate (lfdr) of 0. When, as here, $\pi_0$ is large, the lfsr is much more meaningful than
the lfdr. When $\pi_0$ is small both measures are meaningful, 
and indeed are generally quite similar. Consequently we recommend lfsr for general usage.

In this example, at a $p$ value threshold of 0.05, $\ash$ estimates the local false sign rate (lfsr) to be `r max(z.ash$lfsr[p<0.05])` and the tail false sign rate (FSR) to be `r mean(z.ash$lfsr[p<0.05])`. The actual FSR at that threshold is `r mean(sign(z[p<0.05])!=sign(beta[p<0.05]))`.

### 2014-03-28 examples/qval.nonp.rmd
added examples

# When measurement precision is constant, differences in $q$-values between methods should be driven by estimates of $\pi_0$

Here we explore the root cause of differences between $q$-values
produced by different methods, in the simple case where measurement precisions $(se(\hat\beta_j))$
are the same for all observations. We claim that these differences
should be driven primarily by the fact that different methods produce 
different estimates for $\pi_0$, rather
than other differences in methodology.

First, we outline the intuition behind why differences should be driven primarily by differences in $\pi_0$. Suppose we are given a set of $n$ $p$-values corresponding to $n$ distinct null hypotheses.
Given an estimate of the proportion of true nulls, $\pi_0$, the false discovery
rate at threshold $p$, FDR($p$), is easily estimated "non-parametrically" using the fact that the expected number of false discoveries at threshold $p$ is 
$n p \pi_0$. The $q$-value for observation $j$ is then, by definition, $q_j = \text{FDR}(p_j)$.
This simple "non-parametric" approach to computing $q$-values is implemented in the following function:
```{r}
  qval.nonp = function(p,pi0){
    n =length(p)
    efp = n*p*pi0 # expected number of false positives at threshold p
    np =  rank(p) #number of positives
    return(efp/np)
  }
```

This approach is sufficiently straightforward and natural that one might make the following claim:
any method that, using the same value $\pi_0$, produces appreciably different $q$-values
from the non-parametric approach, should be considered with suspicion (or at least with caution).
This would in turn imply that differences between (sensible) methods should be driven by differences in $\pi_0$, since if they have the same $\pi_0$ then they should both produce results similar to the simple nonparametric approach.

This said, the method for computing $q$-values in `ash` is, on the face of it, very different from the non-parametric approach, and so it is not clear in advance that the two will match closely. In `ash` $q$-values are computed by first computing the local false discovery rate (lfdr) for each observation, and then, at any given threhold, 
averaging the lfdr values that exceed this threshold to estimate the $q$-value at that threshold. 
This approach is effectively estimating a tail-area (the $q$ value) by first estimating
a density (the lfdr) and then approximating an integral (by averaging). 
From this viewpoint, `ash` tackles an easy problem (estimating a tail area) by first solving a harder problem (estimating a density), and one might worry that this could introduce
errors.

We examine this issue by simulation. First we define a simple simulation function that simulates
a mixture of 0 and non-zero (normally-distributed) beta values.
```{r}
  # simulate n beta-hat values, nnull under the null with altmean and altsd
  # being the mean and sd of beta under the alternative
  simdata = function(n, nnull, altmean, altsd, betahatsd) {
    null = c(rep(1, nnull), rep(0, n - nnull))
    beta = c(rep(0, nnull), rnorm(n - nnull, altmean, altsd))
    betahat = rnorm(n, beta, betahatsd)
    return(list(null = null, beta = beta, betahat = betahat, betahatsd = betahatsd))
  }
  library(ashr)
  library(qvalue)
  library(locfdr)
  library(mixfdr)
```

In our first simulation, half the $\beta_j$ are null, and the other half are $\beta_j \sim N(0,2^2)$. Then $\hat\beta_j \sim N(\beta_j, 1)$. We apply `qvalue`,`ash`,`locfdr` and `mixFdr` to estimate the $q$-values. (The last three methods all estimate local fdr values; we convert these to corresponding estimated $q$ values using the `ashr` function `qval.from.lfdr`, which, for each observation $j$, computes the $q$ value as the average of the lfdr values more significant than observation $j$.)
```{r results='hide'}
  set.seed(100)
  test = simdata(10000, 5000, 0, 2, 1) 
  test.ash = ash(test$betahat,test$betahatsd,method="fdr")
  z = test$betahat/test$betahatsd
  p = pchisq(z^2, df=1,lower.tail=FALSE)
  test.qv = qvalue(p)
  test.locfdr = locfdr(z,nulltype=0,plot=0)
  test.locfdr.qval = qval.from.lfdr(test.locfdr$fdr)
  test.mixfdr = mixFdr(z,theonull=TRUE,noiseSD=1,plots=FALSE)
  test.mixfdr.qval = qval.from.lfdr(test.mixfdr$fdr)
```

The estimates of $\pi_0$ from the four methods are given in the following table:
```{r}
  kable(data.frame(method=c("qvalue","ash","locfdr","mixfdr"),pi0=round(c(test.qv$pi0,get_pi0(test.ash),test.locfdr$fp0[1,3],test.mixfdr$pi[1]),digits=2)))
```

Now we perform a "sanity check" for each method, comparing its $q$-values against the
nonparametric $q$ values obtained using the same estimate of $\pi_0$. Both `ash` and `qvalue`
clearly pass this sanity check (for `qvalue`, this is almost tautological, since it effectively uses the non-parametric procedure). For `locfdr`, it also passes the sanity check except for
some artifacts near $p=1$ that we do not fully understand. For `mixfdr` there seems to be a systematic deviation from the non-parametric method that suggests.
```{r}
  par(mfcol=c(2,2))
  plot(qval.nonp(p,get_pi0(test.ash)),test.ash$qvalue, xlab="nonparametric q-values (with pi0 estimated by ash)", ylab = "ash q-values",main="ash")
  abline(a=0,b=1,col=2,lwd=2)

  plot(qval.nonp(p,test.qv$pi0),test.qv$qvalues, xlab="nonparametric q-values (with pi0 estimated by qvalue)", ylab = "qvalue q-values",main="qvalue")
  abline(a=0,b=1,col=2,lwd=2)

 plot(qval.nonp(p,test.mixfdr$pi[1]),test.mixfdr.qval, xlab="nonparametric q-values (with pi0 estimated by mixfdr)", ylab = "mixfdr q-values",main="mixfdr")
  abline(a=0,b=1,col=2,lwd=2)

 plot(qval.nonp(p,test.locfdr$fp0[1,3]),test.locfdr.qval, xlab="nonparametric q-values (with pi0 estimated by locfdr)", ylab = "locfdr q-values", main="locfdr")
  abline(a=0,b=1,col=2,lwd=2) 
  par(mfcol=c(1,1))
```

In this second simulation, half the $\beta_j$ are null, and the alternatives
are $\beta_j \sim N(2,1)$. Note that this deviates quite a bit from the `ash` assumption that
the non-zero $\beta_j$ will have a unimodal distribution centered on 0. This presumably
accounts for the slightly worse correspondance between the `ash` and non-parametric $q$-value compared with the previous simulation.
```{r}
  set.seed(100)
  test = simdata(10000, 5000, 2, 1, 1) 
  test.ash = ash(test$betahat,test$betahatsd,method="fdr")
  z = test$betahat/test$betahatsd
  p = pchisq(z^2, df=1,lower.tail=FALSE)
  plot(qval.nonp(p,get_pi0(test.ash)),test.ash$qvalue, xlab="nonparametric q-values (ash estimate of pi0)", ylab = "ash q-values")
  abline(a=0,b=1,col=2,lwd=2)
```

Although these simulations are by no-means comprehensive, they show that, for these the primary
difference

### 2014-04-30 examples/qval.nonp.rmd
update README for Rcpp and non-Rcpp versions

@@
 clearly pass this sanity check (for `qvalue`, this is almost tautological, since it effectively uses the non-parametric procedure). For `locfdr`, it also passes the sanity check except for
-some artifacts near $p=1$ that we do not fully understand. For `mixfdr` there seems to be a systematic deviation from the non-parametric method that suggests.
+some artifacts near $p=1$ that we do not fully understand. For `mixfdr` there seems to be a systematic deviation from the non-parametric method that suggests...
 ```{r}
@@
 Although these simulations are by no-means comprehensive, they show that, for these the primary
-difference
+difference between ash $q$ values and qvalue $q$ values is
+the different estimates of $\pi_0$.

### 2014-04-30 Rcode/intro.rmd
corrected bug in intro.rmd

@@
 m=beta.mfdr.J15
-mTrue = list(pi=c(0.8,0.2),mu=c(0,0),sig=c(1,2))
+mTrue = list(pi=c(0.8,0.2),mu=c(0,0),sig=c(1,sqrt(5)))
 par(mfcol=c(1,1))

### 2014-04-30 Rcode/intro.rmd
update intro.rmd to replace localfdr with lfdr

--- 2013-10-12 Rcode/intro.rmd
@@
  lines(s, fdrCurve, col = 2)
-points(ss$betahat,beta.ash$localfdr,col=3)
-points(ss$betahat,beta.ash.u$localfdr,col=4)
-points(ss$betahat,beta.ash.hu$localfdr,col=5)
+points(ss$betahat,beta.ash$lfdr,col=3)
+points(ss$betahat,beta.ash.u$lfdr,col=4)
+points(ss$betahat,beta.ash.hu$lfdr,col=5)
 
@@
 z.ash = ash(z,1,mixcompdist="halfuniform")
-points(z,z.ash$localfdr,col=3)
+points(z,z.ash$lfdr,col=3)
 ```
@@
 abline(a=0,b=1)
-plot(beta.ash.auto$localfdr, beta.ash$localfdr, xlab="lfdr from auto method", ylab="ldfr from fixed method")
+plot(beta.ash.auto$lfdr, beta.ash$lfdr, xlab="lfdr from auto method", ylab="ldfr from fixed method")
 abline(a=0,b=1)
@@
 abline(a=0,b=1)
-plot(beta.ash.auto$localfdr, beta.ash.vb.uniform$localfdr, xlab="lfdr from auto method", ylab="ldfr from vb method")
-points(beta.ash.auto$localfdr, beta.ash.vb.null$localfdr, col=2)
+plot(beta.ash.auto$lfdr, beta.ash.vb.uniform$lfdr, xlab="lfdr from auto method", ylab="ldfr from vb method")
+points(beta.ash.auto$lfdr, beta.ash.vb.null$lfdr, col=2)
 abline(a=0,b=1)
@@
 lfdr = rep(0,length(betahat))
-lfdr[pos] = betapos.ash$localfdr
-lfdr[!pos] = betaneg.ash$localfdr
+lfdr[pos] = betapos.ash$lfdr
+lfdr[!pos] = betaneg.ash$lfdr
 qv = qval.from.localfdr(lfdr)

### 2014-05-09 talks/BOG2014.Rmd
added BOG slides

% False Discovery Rates, A New Deal
% Matthew Stephens
% 2014/5/8

```{r, include=FALSE}
require("qvalue")
require("ashr")
require("ggplot2")
```

```{r setup, include=FALSE}
# set global chunk options
opts_chunk$set(cache=TRUE,autodep=TRUE,warning=FALSE,dev='pdf')
dep_auto()
```

# What you can expect from this talk

- No experimental data!

- No fancy heatmaps!

- 100\% Money-back guarantee to half the False Discovery Rate (FDR) in your data!

# The Canonical Genomics Experiment

- Measure lots of things, with error

- Get estimates of effects $\beta_j$ ($\hat\beta_j$) and their standard errors $s_j$

- Turn these into Z-scores, $z_j = \hat\beta_j/s_j$

- Turn these into $p$ values, $p_j$

- Apply `qvalue` 
to identify findings ``significant" at a given FDR.

- ...?

```{r, echo=FALSE}
  #simple simulated example
  ncz = 100 # number of bins in z score histograms
  set.seed(111)
  hh.betahat = rnorm(1000,0,2)
  hh.sebetahat = 1
  hh.zscore = hh.betahat/hh.sebetahat
  hh.pval = pchisq(hh.zscore^2,df=1,lower.tail=F)
  hh.ash = ash(hh.betahat,hh.sebetahat, method="fdr")
  hh.q = qvalue(hh.pval)
```

```{r, echo=FALSE}
  plot_FDReg_hist = function(hh.pval,pi0,nc=40,nullcol="blue",altcol="cyan",type=4,title="Distribution of p values",...){
    hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main=title,nclass=nc,col=altcol,...)
    if(type>1){
      abline(h=pi0,col=nullcol,lwd=2)

      hh.hist$density=rep(pi0,length(hh.hist$density))  
      #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
      plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)
    }
    if(type>2){
    abline(v=0.1,lwd=2,col=2)
    }
    if(type>3){
      text(0.05,1.2,labels="A",col=2,cex=1.2)  
      text(0.05,0.4,labels="B",col=2,cex=1.2)  
      text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
    }
  }
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=1)                 
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=2)                 
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=3)                 
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4)                 
```

# Problem 1: The Zero Assumption (ZA)

- The standard `qvalue` 
approach assumes that all the $p$ values near 1 are null.

- Analogously, one can assume that all Z scores near 0 are null. Efron refers to this as the ``Zero Assumption".

- Seems initially natural.

```{r, echo=FALSE, include=FALSE}
require(fdrtool)
  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue",plot=FALSE)
require(locfdr)
  hh.locfdr = locfdr(hh.zscore,nulltype=0,plot=0)
require(mixfdr)
  hh.mixfdr = mixFdr(hh.zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
```

# Implied distribution of $p$ values under $H_1$

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4)                 
```

# Implied distribution of Z scores under alternative

```{r, echo=FALSE,fig.height=4,fig.cap=""}
#plot a histogram of z scores, highlighting the alternative distribution
#of z scores that is implied by localfdr values lfdr.
  nullalthist = function(z,lfdr,nullcol="blue",altcol="cyan",...){
    h=hist(z, freq=FALSE,col=nullcol,nclass=ncz,...)
    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
    h$density = (1-avlfdr) * h$density
    plot(h,add=TRUE,col=altcol,freq=FALSE)
  }

  nullalthist(hh.zscore,hh.fdrtool$lfdr)  
```

# FDR problem 2: different measurement precision

- In some cases the measurement precisions differ among units

- Eg effect sizes of rare SNPs have larger standard error than those of common SNPs

- Eg Expression levels of low-expressed genes have less precision than high-expressed genes

- If some effects are measured less precisely than others, those tests ``lack power"
and dilute signal, increasing FDR

```{r, echo=FALSE}
#install q value package
#source("http://bioconductor.org/biocLite.R")
#biocLite("qvalue")
library("qvalue")
library("lattice") #library for some of the plots

#set up some data with mixture of values of s
set.seed(100)
s.good = 0.5
s.poor = 10
J.good= 500
J.poor = 500
J=J.good+J.poor
beta = c(rnorm(J,0,1))
s = c(rep(s.good,J.good),rep(s.poor,J.poor))
betahat = beta + rnorm(J,0,s)
#compute the usual zscore and corresponding p value
zscore = betahat/s
pval = pchisq(zscore^2,df=1,lower.tail=F)
```

# FDR problem 2: different measurement precision

- Simulation: effects $\beta_j \sim N(0,1)$
- 500 "good" observations with low standard error ($s_j=0.5$)
- 500 "poor" observations with very high standard error ($s_j=10$)

# FDR problem 2: different measurement precision

```{r, echo=FALSE,fig.height=4,fig.cap=""}
plot_FDReg_hist(pval[1:500],0,title="distribution of GOOD p values",type=1,ylim=c(0,15))
```

# FDR problem 2: different measurement precision

```{r, echo=FALSE,fig.height=4,fig.cap=""}
plot_FDReg_hist(pval[501:1000],0,title="distribution of POOR p values",type=1,ylim=c(0,15))
```

# FDR problem 2: different measurement precision

```{r, echo=FALSE,fig.height=4,fig.cap=""}
plot_FDReg_hist(pval[1:1000],0,title="distribution of ALL p values",type=1,ylim=c(0,15))
```
```{r, include=FALSE}
qq.all = qvalue(pval)
qq.good = qvalue(pval[1:500])
```

# FDR problem 2: different measurement precision

```{r, echo=FALSE,fig.height=4,fig.cap=""}
plot_FDReg_hist(pval[1:500],qq.good$pi0,title="GOOD p values",type=4,ylim=c(0,15))
```

# FDR problem 2: different measurement precision

```{r, echo=FALSE,fig.height=4,fig.cap=""}
plot_FDReg_hist(pval[1:1000],qq.all$pi0,title="ALL p values",type=4,ylim=c(0,15))
```

# Problems: Summary

Standard tools are highly conservative.

- The ZA, which implies actual effects have a (probably unrealistic) bimodal distribution; causes overestimate of $\pi_0$, losing power.

- By focussing on $p$ values, low-precision measurements can dilute high-precision measurements.

# FDR via Empirical Bayes

- Following previous work (e.g. Newton, Efron, Muralidharan) we take an empirical Bayes approach to FDR.

- Eg Efron assumes that the $Z$ scores come from a mixture of null, and alternative:
$$Z_j \sim f_Z(.) = \pi_0 N(.;0,1) + (1-\pi_0) f_1(.)$$
where $f_1, \pi_0$ are to be estimated from the data.

- Once $f_1$ and $\pi_0$ estimated, FDR calculations are straightforward.

# FDR: A New Deal

- Instead of modelling $Z$ scores, model the effects $\beta$,
$$\beta_j \sim \pi_0 \delta_0(.) + (1-\pi_0) g(.)$$

- Constrain $g$ to be unimodal about 0; estimate $g$ from data.

- *Incorporate precision* of each observation $\hat\beta$ into the likelihood.
Specifically, approximate likelihood for $\beta_j$ by assuming 
$$\hat\beta_j \sim N(\beta_j, s_j)$$

# FDR - A New Deal

- A convenient way to model $g$: mixture of 0-centered
normal distributions: 
$$g(\beta; \pi) = \sum_{k=1}^K \pi_k N(\beta; 0, \sigma^2_k)$$

- Estimating $g$ comes down to estimating $\pi$. Joint estimation of $\pi_0,\pi$ easy by maximum likelihood (EM algorithm).

- By allowing $K$ large, and $\sigma_k$ to span a dense grid of values,
we get a flexible unimodal symmetric distribution.

- Can approximate, arbitrarily closely, any scale mixture of normals.
Includes almost all priors used for sparse regression problems (spike-and-slab, double exponential/Laplace/Bayesian Lasso, horseshoe).

# FDR - A New Deal

- Alternatively, a mixture of uniforms, with 0 as one end-point of the range,
provides still more flexibility, and in particular allows for asymmetry.

- If allow a very large number of uniforms this provides the non-parametric mle for $g$; cf Grenander 1953; Cordy + Thomas 1997.

# Illustration: $g$ a mixture of 0-centered normals

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, dnorm(x,0,1),type="l",ylim=c(0,2),ylab="density")
lines(x, dnorm(x,0,0.1))
lines(x, dnorm(x,0,0.5))
```

# Illustration: $g$ a mixture of 0-centered normals

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, 0.5*dnorm(x,0,1)+0.5*dnorm(x,0,0.1),type="l",ylim=c(0,2),ylab="density")
```

# Illustration: $g$ a mixture of 0-anchored uniforms

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, dunif(x,0,1),type="s",ylim=c(0,5),ylab="density")
lines(x, dunif(x,0,0.2),type="s")
lines(x, dunif(x,0,0.5),type="s")
lines(x, dunif(x,-0.3,0),type="s",col=2)
lines(x, dunif(x,-0.4,0),type="s",col=2)
lines(x, dunif(x,-2,0),type="s",col=2)
```

# Illustration: $g$ a mixture of 0-anchored uniforms

```{r, echo=FALSE,fig.height=4,fig.cap=""}
x=seq(-4,4,length=100)
plot(x, 0.1*dunif(x,0,3)+ 0.3*dunif(x,0,0.2) + 0.2*dunif(x,0,0.5) + 0.1* dunif(x,-0.3,0) + 0.1*dunif(x,-0.3,0)+0.2*dunif(x,-0.4,0),type="s",ylim=c(0,2),ylab="density")
```

# Recall Problem 1: distribution of alternative Z values multimodal
```{r, echo=FALSE,fig.height=4,fig.cap=""}
nullalthist(hh.zscore,hh.fdrtool$lfdr,main="qvalue")  
```

# Problem Fixed: distribution of alternative Z values unimodal
```{r, echo=FALSE,fig.height=4,fig.cap=""}
nullalthist(hh.zscore,hh.ash$lfdr,main="ash")
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4,title="qvalue")                 
```

# Example: FDR estimation

```{r, echo=FALSE,fig.height=4,fig.cap=""}
  plot_FDReg_hist(hh.pval,get_pi0(hh.ash),type=4,title="ash")                 
```

```{r, include=FALSE}
  beta.ash.good = ash(betahat[1:500],s[1:500],method="fdr")
  beta.ash.all = ash(betahat,s,method="fdr")
  print(c(sum(beta.ash.good$qvalue<0.05),sum(beta.ash.all$qvalue<0.05)))
```

# Recall Problem 2: poor measurements dilute good measurments

Number of findings at (estimated) FDR=0.05:
```{r, echo=FALSE}
  res.tab = data.frame(qvalue = c(sum(qq.all$qvalues<0.05),sum(qq.good$qvalues<0.05)),
                       ash =c(sum(beta.ash.all$qvalue<0.05),sum(beta.ash.good$qvalue<0.05)))
  row.names(res.tab)=c("ALL","GOOD")
  kable(res.tab,row.names=TRUE,format="pandoc")
```

# A new problem: an embarrassment of riches

- If the null is mostly false, the new approach can  provide unsettling results

- The FDR can be small for all observations, even those with $p \approx 1$!

- In the illustrative example, the maximum $q$ value is `r round(max(hh.ash$qvalue),2)`

# Perhaps we didn't really understand the question?

- Problem arises only if we insist on asking question ``is $\beta_j=0$?"

- Given enough signal, we become convinced that very few of the $\beta_j=0$

- But for some $\beta_j$ we still may have little information about actual value

- Suggests a change of focus: assume *none* of the $\beta_j$ are zero ("one group approach"), and ask for which $\beta_j$ are we confident about the sign (Gelman et al, 2012).

# The False Sign Rate

- Suggestion: replace FDR with local false sign rate (lfsr), the probability that if we say an effect is positive (negative), it is not.

- Example: suppose we estimate that $\Pr(\beta_j<0)=0.975$ and $\Pr(\beta_j>0)=0.025$. Then we report $\beta_j$ as a ``(negative) discovery", and estimate its fsr as 0.025.

# Even with many signals, large $p$ values have high lfsr

```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
  res=data.frame(p=hh.pval, lfsr=hh.ash$lfsr, qvalue = hh.ash$qvalue)
  res.melt = melt(res, id.vars=c("p"),variable.name="Measure")
cbbPalette <- c("#000000", "#D55E00", "#CC79A7")
 labels = c('qvalue','lfsr')
  breaks = c("qvalue","lfsr")

  pp= ggplot(data=res.melt,aes(p,value,color=Measure)) +geom_point(shape=1) +
      geom_abline(colour = "black") +
        xlab("p-value") +
        ylab("lfsr/qvalue")

  print(pp +scale_y_continuous(limits=c(0,0.1)) +
          scale_x_continuous(limits=c(0,0.1))  +
           scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
          coord_equal(ratio=1))

```

# Summary

- ASH provides a generic approach to shrinkage estimation, as well as
false discovery (sign) rates.

- Unimodal assumption for effects reduces conservatism

- But by using two numbers ($\hat\beta,s$) instead of one ($p$ values or $z$ scores) precision of different measurement is better accounted for.

- In high-signal contexts, False Sign Rate is preferable to False Discovery Rate.

# Reproducible Research?

- Principle: when publishing results of computational procedures, we should
publish the code that produced the results.
- "publishing figures or results without the complete software environment could
be compared to a mathematician publishing an announcement of a mathematical theorem without giving the proof" (Buckheit and Donohoe)
- “an article about a computational result is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result.” [Claerbout]

# This talk is reproducible!

- `http://www.github.com/stephens999/ash`

# Why is reproducibility important?

- Not only because people are forgetful, error-prone, or dishonest!
- Reproducing work is also the first step to extending it.
- Helps communications among researchers (eg student + advisor).

# Thanks

- to the developers of **R**, **knitr**, **Rstudio** and **Pandoc**.

- to the several postdoctoral researchers and students
who have worked with me on related topics.

- Including Ester Pantaleo, Scott Powers, Mengyin Lu, Sen Tian, Wei Wang, Zhengrong Xing.

- NHGRI for funding.

- `ashr` package: `http://www.github.com/stephens999/ash`

# Pandoc Command used

`pandoc -s -S -i --template=my.beamer -t beamer -V theme:CambridgeUS -V colortheme:beaver  slides.md -o slides.pdf`

(alternative to produce html slides; but figures would need reworking)
`pandoc -s -S -i -t dzslides --mathjax slides.md -o slides.html`

Here is my session info:

```{r session-info}
print(sessionInfo(), locale=FALSE)
```

# Issue: you can't actually estimate $\pi_0$!

- Data cannot distinguish between $\beta_j = 0$ and $\beta_j$ "very
small"

- As a result $\pi_0$ cannot be estimated: the data can never rule out $\pi_0=0$.

- But the unimodal constraint bounds how big $\pi_0$ can be.

- Use penalized likelihood to make $\pi_0$ ``as big as possible",
subject to the unimodal constraint.

# Approach remains conservative (if unimodal assumption holds)

```{r,echo=FALSE,include=FALSE}
load("../paper/Rcode/sim1.RData") #load simulation results
source("../paper/Rcode/plot_pi0.R")
```

```{r, echo=FALSE,fig.height=4,fig.cap=""}
plot_pi0_onlyqval(list(simres1a,simres2))
```

# Next steps?

- Incorporate $t$ likelihood as well as normal.

- Incorporate shrinkage of variances and not just means.

- Extend to allow $g(\cdot;\pi)$ to depend on covariates $X$.

- Extend to allow for correlations in the measured $\hat\beta_j$.

# Is this an important problem?

- The original paper introducing FDR (Benjamini and Hochberg, 1995) has
been cited 21,787 times according to Google Scholar.

- That is three times a day for the last 19 years!

# FDR and q values

Although precise definitions vary, roughly

- The FDR at a threshold $P$ is 
$$\text{FDR}(P)=\Pr(\beta_j = 0 |  p_j<P).$$

- The q value for observation $j$ is $q_j=\text{FDR}(p_j)$.

# Estimation and Shrinkage

- Besides allowing one to estimate fdr and fsr, 
this approach also provides a posterior distribution for each $\beta_j$.

- So for example we can easily compute fdrs for discoveries other than ``non-zero" (eg compute $\Pr(\beta_j > 2 | \hat\beta_j)$).

- And use it to obtain point estimates and credible intervals for each $\beta_j$, taking account of information from all the other $\beta_j$.

- Because $f(\beta)$ is unimodal, the point estimates will tend to be ``shrunk" towards the overall mean (0).

- Because $f(\beta)$ is estimated from the data, the amount
of shrinkage is adaptive to the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each gene.

- So we call the approach ``Adaptive Shrinkage" (ASH).

### 2014-05-22 Rcode/jash.rmd
add jash.rmd

```{r setup, include=FALSE}
library(limma)
library(qvalue)
library(ashr)
source('jash.R')
set.seed(54321)
```

Simulations
===========

Simulate data: N genes, 2 conditions, nA/nB samples for each condition.
Observations are normal distributed.
```{r}
simfull=function(N,Nnull,nA,nB,mean,betasd,diffmean,a,lambda,c){
  null = c(rep(1, Nnull), rep(0, N - Nnull))
  truetau=rgamma(N,shape=a,rate=a/lambda)
  truebetaA=rnorm(N,mean=mean,sd=betasd)
  truediff=c(rep(0,Nnull),rnorm(N-Nnull,mean=diffmean,sd=1/sqrt(c*truetau[Nnull+1:N]))) # true effect
  truebetaB=truebetaA+truediff
  YA=matrix(rnorm(N*nA,mean=rep(truebetaA,nA),sd=rep(1/sqrt(truetau),nA)),ncol=nA)
  YB=matrix(rnorm(N*nB,mean=rep(truebetaB,nB),sd=rep(1/sqrt(truetau),nB)),ncol=nB)
  return(list(YA=cbind(YA,YB),null=null,truetau=truetau,truebetaA=truebetaA,
              truebetaB=truebetaB,truediff=truediff))
}
```

Shrinkage
=========

An example of effects/variances shrinkage using jash & limma:

```{r}
# Simulate data
d.test=simfull(N=10000,Nnull=7000,nA=2,nB=2,mean=5,betasd=0.5,diffmean=0,a=10,lambda=1,c=0.25)
fac=factor(rep(c(1,2),each=2))
design=model.matrix(~fac)
```

The histogram of true effects/variances of this simulated dataset:

```{r,echo=FALSE}
hist(d.test$truediff,prob=TRUE,breaks=seq(-10.5,10.5,by=1),xlab=expression(paste('true ',beta[g2])),main=expression(paste('Histogram of true ',beta[g2])))
hist(d.test$truetau,breaks=40,xlab=expression(paste('true ',tau[g])),main=expression(paste('Histogram of true ',tau[g])))
```

Use jash and limma to shrink gene-specific variances:

```{r}
# jash
jash.test=jash(d.test$Y,fac,usePointMass=TRUE)

# limma
fit.test <- lmFit(d.test$Y, design=design)
fit.test <- eBayes(fit.test)
```

Compare the original $\hat{\beta}_g$'s and $1/\hat{\tau_g}$'s with their shrinkage estimates (posterior mean) by jash:

```{r,echo=FALSE}
betahat=fit.test$coefficients[,2]
sebetahat=fit.test$sigma
plot(betahat,jash.test$PosteriorMean,xlim=c(-10,10),ylim=c(-10,10),xlab=expression(paste(hat(beta)[g])),ylab=expression(paste('E(',beta[g],'|Y)')),main='(a)')
abline(0,1,lty=2)
plot(log(sebetahat^2),log(1/jash.test$PosteriorPrec),xlim=c(-6,6),ylim=c(-2,2),xlab=expression(paste('log(1/',hat(tau)[g],')')),ylab=expression(paste('log(1/E(',tau[g],'|Y))')),main='(b)')
abline(0,1,lty=2)
```

Compare the true $\tau_g$'s with the estimated $\tau_g$'s by jash or limma:

```{r, echo=FALSE}
plot(fit.test$s2.post,1/jash.test$PosteriorPrec,xlim=c(0,3),ylim=c(0,3),xlab='limma',ylab='jash',
     main='Estimated gene-specific variances by limma and jash')
abline(0,1,lty=2)
plot(log(1/d.test$truetau),log(1/jash.test$PosteriorPrec),xlim=c(-1,1),ylim=c(-1,1),xlab=expression(paste('-log(true ',tau,')')),ylab=expression(paste('-log(estimated ',tau,')')))
points(log(1/d.test$truetau),log(fit.test$s2.post),col=2)
abline(0,1,lty=2)
legend('topleft',pch=c(1,1),col=c('black','red'),legend=c('jash','limma'))
```

```{r}
# MSE of true tau and estimated tau (jash)
mean((log(1/d.test$truetau)-log(1/jash.test$PosteriorPrec))^2)
# MSE of true tau and estimated tau (limma)
mean((log(1/d.test$truetau)-log(fit.test$s2.post))^2)
```

Estimate FDR
============

Compare the true/estimated FDR by jash and limma for 8 simulation scenarios:

```{r,echo=FALSE}
# For each case, simulate 20 times
nsim=20

# Case 1
jash.e.FDR.1=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.1=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.1=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.1=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(c(1,1,2,2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.1=simfull(N=10000,Nnull=5000,nA=2,nB=2,mean=5,betasd=0.5,diffmean=0,a=10,lambda=1,c=1)
  jash.1=jash(d.1$Y,fac,usePointMass=TRUE)
  fit.1 <- lmFit(d.1$Y, design=design)
  fit.1 <- eBayes(fit.1)
  o.jash = order(jash.1$qvalue)
  o.limma = order(fit.1$p.value[,2])
  jash.FDR.1[,i]=cumsum(d.1$null[o.jash])/(1:10000)
  limma.FDR.1[,i]=cumsum(d.1$null[o.limma])/(1:10000)
  jash.e.FDR.1[,i]=sort(jash.1$qvalue)
  limma.e.FDR.1[,i]=sort(qvalue(topTable(fit.1,coef='fac2',n=10000)$P.Value)$qval)
}

ng=150
plot(apply(jash.FDR.1,1,mean)[1:ng],type='l',ylim=c(0,0.2),ylab='FDR' )
lines(apply(limma.FDR.1,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.1,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.1,1,mean)[1:ng],lty=2,col=2)
title('Case 1')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))

# Case 2
jash.e.FDR.2=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.2=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.2=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.2=matrix(rep(NA,nsim*10000),ncol=nsim)
ash.FDR.2=matrix(rep(NA,nsim*10000),ncol=nsim)
ash.e.FDR.2=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(c(1,1,2,2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.2=simfull(N=10000,Nnull=5000,nA=2,nB=2,mean=5,betasd=0.5,diffmean=0,a=10,lambda=1,c=0.1)
  jash.2=jash(d.2$Y,fac,usePointMass=TRUE)
  fit.2 <- lmFit(d.2$Y, design=design)
  fit.2 <- eBayes(fit.2)
  o.jash = order(jash.2$qvalue)
  o.limma = order(fit.2$p.value[,2])

  jash.FDR.2[,i]=cumsum(d.2$null[o.jash])/(1:10000)
  limma.FDR.2[,i]=cumsum(d.2$null[o.limma])/(1:10000)
  jash.e.FDR.2[,i]=sort(jash.2$qvalue)
  limma.e.FDR.2[,i]=sort(qvalue(topTable(fit.2,coef='fac2',n=10000)$P.Value)$qval)

  ash.2 = ash(fit.2$coef[,2]/(sqrt(fit.2$s2.post)*fit.2$stdev.unscaled[,2]),1,method='fdr')
  o.ash=order(ash.2$qvalue)
  ash.FDR.2[,i]=cumsum(d.2$null[o.ash])/(1:10000)
  ash.e.FDR.2[,i]=sort(ash.2$qvalue)
}

ng=2000
plot(apply(limma.FDR.2,1,mean)[1:ng],col=2,type='l',ylim=c(0,0.05),ylab='FDR',xlab='Number of genes')
lines(apply(jash.FDR.2,1,mean)[1:ng])
lines(apply(jash.e.FDR.2,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.2,1,mean)[1:ng],lty=2,col=2)
title('Case 2')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))

# Case 3
jash.e.FDR.3=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.3=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.3=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.3=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.ef.FDR.3=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(c(1,1,1,2,2,2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.3=simfull(N=10000,Nnull=5000,nA=3,nB=3,mean=5,betasd=0.5,diffmean=0,a=10,lambda=1,c=0.25)
  jash.3=jash(d.3$Y,fac,usePointMass=TRUE)
  fit.3 <- lmFit(d.3$Y, design=design)
  fit.3 <- eBayes(fit.3)
  o.jash = order(jash.3$qvalue)
  o.limma = order(fit.3$p.value[,2])

  jash.FDR.3[,i]=cumsum(d.3$null[o.jash])/(1:10000)
  limma.FDR.3[,i]=cumsum(d.3$null[o.limma])/(1:10000)
  jash.e.FDR.3[,i]=sort(jash.3$qvalue)
  limma.e.FDR.3[,i]=sort(qvalue(topTable(fit.3,coef='fac2',n=10000)$P.Value)$qval)
}

ng=2500
plot(apply(limma.FDR.3,1,mean)[1:ng],col=2,type='l',ylim=c(0,0.1),ylab='FDR' )
lines(apply(jash.FDR.3,1,mean)[1:ng])
lines(apply(jash.e.FDR.3,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.3,1,mean)[1:ng],lty=2,col=2)
title('Case 3')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))

# Case 4
jash.e.FDR.4=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.4=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.4=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.4=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(rep(c(1,2),each=50))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.4=simfull(N=10000,Nnull=5000,nA=50,nB=50,mean=7,betasd=0.5,diffmean=0,a=1,lambda=1,c=0.25)
  jash.4=jash(d.4$Y,fac,usePointMass=TRUE)
  fit.4 <- lmFit(d.4$Y, design=design)
  fit.4 <- eBayes(fit.4)
  o.jash = order(jash.4$qvalue)
  o.limma = order(fit.4$p.value[,2])

  jash.FDR.4[,i]=cumsum(d.4$null[o.jash])/(1:10000)
  limma.FDR.4[,i]=cumsum(d.4$null[o.limma])/(1:10000)
  jash.e.FDR.4[,i]=sort(jash.4$qvalue)
  limma.e.FDR.4[,i]=sort(qvalue(topTable(fit.4,coef='fac2',n=10000)$P.Value)$qval)
}

ng=10000
plot(apply(jash.FDR.4,1,mean)[1:ng],type='l',ylim=c(0,1),ylab='FDR' )
lines(apply(limma.FDR.4,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.4,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.4,1,mean)[1:ng],lty=2,col=2)
title('Case 4')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))

# Case 5
jash.e.FDR.5=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.5=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.5=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.5=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(rep(c(1,2),each=2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.5=simfull(N=10000,Nnull=7000,nA=2,nB=2,mean=7,betasd=1,diffmean=0,a=5,lambda=1,c=0.25)
  jash.5=jash(d.5$Y,fac,usePointMass=TRUE)
  fit.5 <- lmFit(d.5$Y, design=design)
  fit.5 <- eBayes(fit.5)
  o.jash = order(jash.5$qvalue)
  o.limma = order(fit.5$p.value[,2])

  jash.FDR.5[,i]=cumsum(d.5$null[o.jash])/(1:10000)
  limma.FDR.5[,i]=cumsum(d.5$null[o.limma])/(1:10000)
  jash.e.FDR.5[,i]=sort(jash.5$qvalue)
  limma.e.FDR.5[,i]=sort(qvalue(topTable(fit.5,coef='fac2',n=10000)$P.Value)$qval)
}

ng=2000
plot(apply(jash.FDR.5,1,mean)[1:ng],type='l',ylim=c(0,0.5),ylab='FDR' )
lines(apply(limma.FDR.5,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.5,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.5,1,mean)[1:ng],lty=2,col=2)
title('Case 5')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))

# Case 6
jash.e.FDR.6=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.6=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.6=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.6=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(rep(c(1,2),each=2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.6=simfull(N=10000,Nnull=9000,nA=2,nB=2,mean=7,betasd=4,diffmean=0,a=10,lambda=1,c=0.1)
  jash.6=jash(d.6$Y,fac,usePointMass=TRUE)
  fit.6 <- lmFit(d.6$Y, design=design)
  fit.6 <- eBayes(fit.6)
  o.jash = order(jash.6$qvalue)
  o.limma = order(fit.6$p.value[,2])

  jash.FDR.6[,i]=cumsum(d.6$null[o.jash])/(1:10000)
  limma.FDR.6[,i]=cumsum(d.6$null[o.limma])/(1:10000)
  jash.e.FDR.6[,i]=sort(jash.6$qvalue)
  limma.e.FDR.6[,i]=sort(qvalue(topTable(fit.6,coef='fac2',n=10000)$P.Value)$qval)
}

ng=10000
plot(apply(jash.FDR.6,1,mean)[1:ng],type='l',ylim=c(0,1),ylab='FDR' )
lines(apply(limma.FDR.6,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.6,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.6,1,mean)[1:ng],lty=2,col=2)
title('Case 6')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))

# Case 7
jash.e.FDR.7=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.7=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.7=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.7=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(rep(c(1,2),each=2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.7=simfull(N=10000,Nnull=7000,nA=2,nB=2,mean=7,betasd=2,diffmean=0,a=1,lambda=0.25,c=0.1)
  jash.7=jash(d.7$Y,fac,usePointMass=TRUE)
  fit.7 <- lmFit(d.7$Y, design=design)
  fit.7 <- eBayes(fit.7)
  o.jash = order(jash.7$qvalue)
  o.limma = order(fit.7$p.value[,2])

  jash.FDR.7[,i]=cumsum(d.7$null[o.jash])/(1:10000)
  limma.FDR.7[,i]=cumsum(d.7$null[o.limma])/(1:10000)
  jash.e.FDR.7[,i]=sort(jash.7$qvalue)
  limma.e.FDR.7[,i]=sort(qvalue(topTable(fit.7,coef='fac2',n=10000)$P.Value)$qval)
}

ng=1000
plot(apply(jash.FDR.7,1,mean)[1:ng],type='l',ylim=c(0,0.2),ylab='FDR' )
lines(apply(limma.FDR.7,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.7,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.7,1,mean)[1:ng],lty=2,col=2)
title('Case 7')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))

# Case 8
jash.e.FDR.8=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.8=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.8=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.8=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(rep(c(1,2),each=2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.8=simfull(N=10000,Nnull=7000,nA=2,nB=2,mean=7,betasd=2,diffmean=0,a=1,lambda=4,c=0.25)
  jash.8=jash(d.8$Y,fac,usePointMass=TRUE)
  fit.8 <- lmFit(d.8$Y, design=design)
  fit.8 <- eBayes(fit.8)
  o.jash = order(jash.8$qvalue)
  o.limma = order(fit.8$p.value[,2])

  jash.FDR.8[,i]=cumsum(d.8$null[o.jash])/(1:10000)
  limma.FDR.8[,i]=cumsum(d.8$null[o.limma])/(1:10000)
  jash.e.FDR.8[,i]=sort(jash.8$qvalue)
  limma.e.FDR.8[,i]=sort(qvalue(topTable(fit.8,coef='fac2',n=10000)$P.Value)$qval)
}

ng=250
plot(apply(jash.FDR.8,1,mean)[1:ng],type='l',ylim=c(0,0.2),ylab='FDR' )
lines(apply(limma.FDR.8,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.8,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.8,1,mean)[1:ng],lty=2,col=2)
title('Case 8')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))
```

Compare with limma, jash, limma+ash (Case 2 for example):

```{r,echo=FALSE}
ng=2000
plot(apply(limma.FDR.2,1,mean)[1:ng],col=2,type='l',ylim=c(0,0.05),ylab='FDR',xlab='Number of genes')
lines(apply(jash.FDR.2,1,mean)[1:ng])
lines(apply(ash.FDR.2,1,mean)[1:ng],col='blue')
lines(apply(jash.e.FDR.2,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.2,1,mean)[1:ng],lty=2,col=2)
lines(apply(ash.e.FDR.2,1,mean)[1:ng], col='blue',lty=2)
legend('topleft',col=c('red','black','blue','red','black','blue'),lty=c(2,2,2,1,1,1),legend=c('limma','jash','ash+limma','true FDR (jash)','true FDR (limma)','true FDR (ash+limma)'))
```

Robustness
==========

Simulate data from the following 3 alternative models:

Case 1: $\beta_g \sim \pi_0\delta_0+\pi_1N(0, \sigma^2)$, $\tau_g \sim Gamma(a,a/\lambda)$, $\hat{\beta}_g\sim N(\beta_g,1/\tau_g)$.
```{r}
simrob=function(N,Nnull,nA,nB,mean,betasd,diffmean,diffsd,a,lambda){
  null = c(rep(1, Nnull), rep(0, N - Nnull))
  truetau=rgamma(N,shape=a,rate=a/lambda)
  truebetaA=rnorm(N,mean=mean,sd=betasd)
  truediff=c(rep(0,Nnull),rnorm(N-Nnull,mean=diffmean,sd=diffsd))
  truebetaB=truebetaA+truediff
  YA=matrix(rnorm(N*nA,mean=rep(truebetaA,nA),sd=rep(1/sqrt(truetau),nA)),ncol=nA)
  YB=matrix(rnorm(N*nB,mean=rep(truebetaB,nB),sd=rep(1/sqrt(truetau),nB)),ncol=nB)
  return(list(YA=cbind(YA,YB),null=null,truetau=truetau,truebetaA=truebetaA,
              truebetaB=truebetaB,truediff=truediff))
}
```
```{r,echo=FALSE}
# Case 1
jash.e.FDR.rob=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.rob=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.rob=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.rob=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(rep(c(1,2),each=2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.rob=simrob(N=10000,Nnull=5000,nA=2,nB=2,mean=5,betasd=0.5,diffmean=0,diffsd=1,a=5,lambda=2)
  fac=factor(rep(c(1,2),each=2))
  design=model.matrix(~fac)

  jash.rob=jash(d.rob$Y,fac,usePointMass=TRUE)
  fit.rob <- lmFit(d.rob$Y, design=design)
  fit.rob <- eBayes(fit.rob)
  o.jash = order(jash.rob$qvalue)
  o.limma = order(fit.rob$p.value[,2])

  jash.FDR.rob[,i]=cumsum(d.rob$null[o.jash])/(1:10000)
  limma.FDR.rob[,i]=cumsum(d.rob$null[o.limma])/(1:10000)
  jash.e.FDR.rob[,i]=sort(jash.rob$qvalue)
  limma.e.FDR.rob[,i]=sort(qvalue(topTable(fit.rob,coef='fac2',n=10000)$P.Value)$qval)
}

ng=2000
plot(apply(jash.FDR.rob,1,mean)[1:ng],type='l',ylim=c(0,0.3),ylab='FDR' )
lines(apply(limma.FDR.rob,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.rob,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.rob,1,mean)[1:ng],lty=2,col=2)
title('Case 1')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))
```

Case 2: $\beta_g \sim \pi_0\delta_0+\pi_1N(0, 1/(c\tau_g))$, $\tau_g \sim N(\mu,\sigma^2)$, $\hat{\beta}_g\sim N(\beta_g,1/\tau_g)$.
```{r,echo=FALSE}
simrob2=function(N,Nnull,nA,nB,mean,betasd,diffmean,lambda,tausd,c){
  null = c(rep(1, Nnull), rep(0, N - Nnull))
  truetau=rnorm(N,mean=lambda,sd=tausd)
  truebetaA=rnorm(N,mean=mean,sd=betasd)
  truediff=c(rep(0,Nnull),rnorm(N-Nnull,mean=diffmean,sd=1/sqrt(c*truetau[Nnull+1:N])))
  truebetaB=truebetaA+truediff
  YA=matrix(rnorm(N*nA,mean=rep(truebetaA,nA),sd=rep(1/sqrt(truetau),nA)),ncol=nA)
  YB=matrix(rnorm(N*nB,mean=rep(truebetaB,nB),sd=rep(1/sqrt(truetau),nB)),ncol=nB)
  return(list(YA=cbind(YA,YB),null=null,truetau=truetau,truebetaA=truebetaA,
              truebetaB=truebetaB,truediff=truediff))
}
```
```{r,echo=FALSE}
# Case 2
jash.e.FDR.rob2=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.rob2=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.rob2=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.rob2=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(rep(c(1,2),each=2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.rob2=simrob2(N=10000,Nnull=5000,nA=2,nB=2,mean=5,betasd=0.5,diffmean=0,lambda=3,tausd=0.25,c=1)
  fac=factor(rep(c(1,2),each=2))
  design=model.matrix(~fac)

  jash.rob2=jash(d.rob2$Y,fac,usePointMass=TRUE)
  fit.rob2 <- lmFit(d.rob2$Y, design=design)
  fit.rob2 <- eBayes(fit.rob2)
  o.jash = order(jash.rob2$qvalue)
  o.limma = order(fit.rob2$p.value[,2])

  jash.FDR.rob2[,i]=cumsum(d.rob2$null[o.jash])/(1:10000)
  limma.FDR.rob2[,i]=cumsum(d.rob2$null[o.limma])/(1:10000)
  jash.e.FDR.rob2[,i]=sort(jash.rob2$qvalue)
  limma.e.FDR.rob2[,i]=sort(qvalue(topTable(fit.rob2,coef='fac2',n=10000)$P.Value)$qval)
  }

ng=1000
plot(apply(jash.FDR.rob2,1,mean)[1:ng],type='l',ylim=c(0,0.6),ylab='FDR' )
lines(apply(limma.FDR.rob2,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.rob2,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.rob2,1,mean)[1:ng],lty=2,col=2)
title('Case 2')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))
```

Case 3: $\beta_g \sim \pi_0\delta_0+\pi_1N(0, 1/(c\tau_g))$, $log(\tau_g) \sim N(\mu,\sigma^2)$, $\hat{\beta}_g\sim N(\beta_g,1/\tau_g)$.
```{r,echo=FALSE}
simrob3=function(N,Nnull,nA,nB,mean,betasd,diffmean,lambda,logtausd,c){
  null = c(rep(1, Nnull), rep(0, N - Nnull))
  truetau=rlnorm(N,mean=log(lambda),sd=logtausd)
  truebetaA=rnorm(N,mean=mean,sd=betasd)
  truediff=c(rep(0,Nnull),rnorm(N-Nnull,mean=diffmean,sd=1/sqrt(c*truetau[Nnull+1:N])))
  truebetaB=truebetaA+truediff
  YA=matrix(rnorm(N*nA,mean=rep(truebetaA,nA),sd=rep(1/sqrt(truetau),nA)),ncol=nA)
  YB=matrix(rnorm(N*nB,mean=rep(truebetaB,nB),sd=rep(1/sqrt(truetau),nB)),ncol=nB)
  return(list(YA=cbind(YA,YB),null=null,truetau=truetau,truebetaA=truebetaA,
              truebetaB=truebetaB,truediff=truediff))
}
```
```{r,echo=FALSE}
# Case 3
jash.e.FDR.rob3=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.e.FDR.rob3=matrix(rep(NA,nsim*10000),ncol=nsim)
jash.FDR.rob3=matrix(rep(NA,nsim*10000),ncol=nsim)
limma.FDR.rob3=matrix(rep(NA,nsim*10000),ncol=nsim)

fac=factor(rep(c(1,2),each=2))
design=model.matrix(~fac)
for (i in 1:nsim){
  d.rob3=simrob3(N=10000,Nnull=5000,nA=2,nB=2,mean=5,betasd=0.5,diffmean=0,lambda=3,logtausd=1,c=1)
  fac=factor(rep(c(1,2),each=2))
  design=model.matrix(~fac)

  jash.rob3=jash(d.rob3$Y,fac,usePointMass=TRUE)
  fit.rob3 <- lmFit(d.rob3$Y, design=design)
  fit.rob3 <- eBayes(fit.rob3)
  o.jash = order(jash.rob3$qvalue)
  o.limma = order(fit.rob3$p.value[,2])

  jash.FDR.rob3[,i]=cumsum(d.rob3$null[o.jash])/(1:10000)
  limma.FDR.rob3[,i]=cumsum(d.rob3$null[o.limma])/(1:10000)
  jash.e.FDR.rob3[,i]=sort(jash.rob3$qvalue)
  limma.e.FDR.rob3[,i]=sort(qvalue(topTable(fit.rob3,coef='fac2',n=10000)$P.Value)$qval)
}

ng=1000
plot(apply(jash.FDR.rob3,1,mean)[1:ng],type='l',ylim=c(0,0.5),ylab='FDR' )
lines(apply(limma.FDR.rob3,1,mean)[1:ng], col=2)
lines(apply(jash.e.FDR.rob3,1,mean)[1:ng],lty=2)
lines(apply(limma.e.FDR.rob3,1,mean)[1:ng],lty=2,col=2)
title('Case 3')
legend('topleft',col=c('red','red','black','black'),lty=c(1,2,1,2),
       legend=c('true FDR (jash)','est FDR (jash)','true FDR (limma)','est FDR (limma)'))
```

Application
===========

BRCA dataset:
```{r,echo=FALSE}
hh = read.table("nejm_brca_release_edit.csv",sep=",",skip=3)
subset = apply(hh, 1, max)<20
hh = hh[subset,]

labs = read.table("nejm_brca_release_edit.csv",sep=",",skip=1,nrows=1)
labs = 1*(labs=="BRCA1") + 2 * (labs=="BRCA2")

hh.betahat = apply(hh[,labs==1],1,mean) - apply(hh[,labs==2],1,mean)
n1 = sum(labs==1)
n2 = sum(labs==2)
hh.sebetahat = sqrt(apply(hh[,labs==1],1,var)/n1 + apply(hh[,labs==2],1,var)/n2)

hh.jash=jasha(hh.betahat,hh.sebetahat,df=n1+n2-2)

library(limma)
hh.new=cbind(hh[,labs==1],hh[,labs==2])
fac=factor(c(rep(1,n1),rep(2,n2)))
design=model.matrix(~fac)
fit <- lmFit(hh.new, design=design)
fit <- eBayes(fit)
```

```{r}
# Number of genes called significant (jash, FDR<0.05)
sum(hh.jash$qvalue<0.05)

# Number of genes called significant (limma, FDR<0.05)
sum(qvalue(topTable(fit,coef='fac2',n=3169)$P.Val)$qval<0.05)
```

### 2014-05-22 Rcode/jash_test.R

@@ -1,132 +1,0 @@
-set.seed(22222)
-require(ashr)
-source("jash.r")
-
-# N: num of obs
-# n: num of replicates
-simdata = function(N, Nnull, altmean, altsd, betasd, n) {
-  null = c(rep(1, Nnull), rep(0, N - Nnull))
-  beta = c(rep(0, Nnull), rnorm(N - Nnull, altmean, altsd))
-  Y = matrix(rnorm(N*n, rep(beta,n), rep(betasd,n)), ncol=n)
-  return(list(null = null, beta = beta, Y = Y))
-}
-
-ss = simdata(10000, 8000, 0, 3, 1, 3)
-beta.jash = jash(ss$Y)
-beta.jash.auto = jash(ss$Y,auto=TRUE)
-
-betahat = apply(ss$Y,1,mean)
-sebetahat = apply(ss$Y,1,sd)/sqrt(dim(ss$Y)[2])
-beta.ash.auto = ash(betahat,sebetahat,method="fdr")
-
-# Histogram of true beta
-hist(ss$beta, prob = TRUE, breaks = seq(-11, 11, length = 20))
-
-plot(betahat,beta.jash$PosteriorMean,xlab='Observed betahat',ylab='Posterior mean',cex=0.5,pch=20)
-points(betahat,beta.ash.auto$PosteriorMean, col=2, cex=0.5,pch=20)
-title('Black: jash; Red: ash.auto')
-
-# Compare sampled distribution P(beta|Y) to the normal mixture distribution P(beta|Y,hat(tau))
-# Red lines are the conditional posterior distribution P(beta|Y,hat(tau)) 
-par(mfrow=c(1,2))
-sample.jash = posterior_sample_jash(beta.jash$post,1000,1:2)
-hist(sample.jash$beta[1,],prob = TRUE, breaks = seq(-2, 2, length = 20),ylim=c(0,2.5),main='First unit')
-x = seq(-2, 2, length = 1000)
-normdens = dnorm(outer(x,rep(1,length(beta.jash$a.vec))),mean=beta.jash$condpost$normmean[rep(1,1000),],sd=beta.jash$condpost$normsd[rep(1,1000),])
-mixnormdens = rowSums(beta.jash$condpost$condpi[rep(1,1000),]*normdens)
-lines(x,mixnormdens,type='l',col=2) 
-
[…]

### 2014-05-27 examples/fsr.rmd
add Sen's code for t likelihood

@@
 ```{r}
+opts_chunk$set(fig.path='figure/lfsr/')
 set.seed(100)
@@
 
-```{r}
+```{r, }
   hist(p,prob=TRUE)

### 2014-05-29 examples/tlike.Rmd
updated EM algorithm (noncxx version) to run with accelerated EM using SQUAREM pacakge; added some .Rmd files in tests and examples that I forgot to add in previous commit

% ash with student-t likelihood
% Sen Tian
% 05/26/2014

```{r, include=FALSE}
library(ashr)
```

```{r setup, include=FALSE}
# set global chunk options
opts_chunk$set(warning=FALSE)
dep_auto()
```

# Examples of using the student-t likelihood
- If 'df' option is specified in function 'ash', the student-t likelihood is applied.
- student-t likelihood with normal mixture is not yet implemented

### Example1.1: density estimations with two different likelihoods
```{r, echo=TRUE}
  simdata = function(n, nnull, altmean, altsd, betasd, df) {
    null = c(rep(1, nnull), rep(0, n - nnull))
    beta = c(rep(0, nnull), rnorm(n - nnull, altmean, altsd))
    betahat =  beta + betasd * rt(n,df)
    return(list(null = null, beta = beta, betahat = betahat, betasd = betasd))
  }

  ss = simdata(1000, 800, 0, 2, 1, 1.5)
  beta.ash.u_n = ash(ss$betahat, ss$betasd, mixcompdist="uniform",method="shrink", cxx=FALSE)
  beta.ash.u_t1 = ash(ss$betahat, ss$betasd, mixcompdist="uniform",method="shrink", cxx=FALSE,df=1.5)
  beta.ash.u_t2 = ash(ss$betahat, ss$betasd, mixcompdist="uniform",method="shrink", cxx=FALSE,df=5)
  beta.ash.u_t1n = ash(ss$betahat, ss$betasd, mixcompdist="uniform",method="fdr", cxx=FALSE,df=1.5)

  # plot the cdf
  x = seq(-6, 6, length = 1000)
  plot(ecdf(ss$beta), main="CDF, Uniform mixture",xlim=c(-6,6))
  lines(x,cdf.ash(beta.ash.u_n,x)$y,col=2,lwd=2)
  lines(x,cdf.ash(beta.ash.u_t1,x)$y,col=3,lwd=2)
  lines(x,cdf.ash(beta.ash.u_t2,x)$y,col=4,lwd=2)
  lines(x,cdf.ash(beta.ash.u_t1n,x)$y,col=5,lwd=2)
  legend("topleft",legend = c("normal-lik","t-lik,df=1.5","t-lik, df=5","t-lik, df=1.5, fdr"), 
         col = c(2,3,4,5), lwd = c(2,2,2,2))
```

### Example1.2: density estimation with student-t likelihood and normal likelihood with quasi df
```{r, echo=TRUE}
  effective.effect=function(betahat,se,df){
    p = pt(betahat/se,df)
    qnorm(p,sd=se)
  }
  ss$betahat = effective.effect(ss$betahat,ss$betasd,1.5)
  beta.ash.u_n1 =  ash(ss$betahat, ss$betasd, mixcompdist="uniform",method="shrink", cxx=FALSE)
  plot(ecdf(ss$beta), main="CDF, Uniform mixture",xlim=c(-6,6))
  lines(x,cdf.ash(beta.ash.u_n1,x)$y,col=2,lwd=2)
  lines(x,cdf.ash(beta.ash.u_t1,x)$y,col=3,lwd=2)
  legend("topleft",legend = c("normal-lik with quasi df=1.5","t-lik,df=1.5"), 
         col = c(2,3), lwd = c(2,2))
```

### 2014-05-29 scratchwork/makenew.mixEM.Rmd
updated EM algorithm (noncxx version) to run with accelerated EM using SQUAREM pacakge; added some .Rmd files in tests and examples that I forgot to add in previous commit

The aim here is to use the SQUAREM package to improve the speed of convergence of the EM algorithm.

First here is our existing implementation.

```{r}
mixEM = function(matrix_lik, prior, pi.init = NULL,tol=0.0001, maxiter=5000){
  n=nrow(matrix_lik)
  k=ncol(matrix_lik)
  B = rep(0,maxiter)
  pi = pi.init
  if(is.null(pi.init)){
    pi = rep(1/k,k)# Use as starting point for pi
  } 
  pi = ifelse(pi<1e-5,1e-5,pi) #set any estimates that are too small to be just very small
  pi = normalize(pi)

  loglik = rep(0,maxiter)
  priordens= rep(0,maxiter)
  m  = t(pi * t(matrix_lik)) # matrix_lik is n by k; so this is also n by k
  m.rowsum = rowSums(m)
  loglik[1] = sum(log(m.rowsum))
  priordens[1] = sum((prior-1)*log(pi)) 
  classprob = m/m.rowsum #an n by k matrix
  i=1
  if(maxiter >= 2){
    for(i in 2:maxiter){  
      pi = colSums(classprob) + prior-1
      pi = ifelse(pi<1e-5,1e-5,pi) #set any estimates that are less than zero, which can happen with prior<1, to 0
      pi = normalize(pi)

      #Now re-estimate pi
      m  = t(pi * t(matrix_lik)) 
      m.rowsum = rowSums(m)
      loglik[i] = sum(log(m.rowsum))
      priordens[i] = sum((prior-1)*log(pi)) 
      classprob = m/m.rowsum

      if(abs(loglik[i]+priordens[i]-loglik[i-1]-priordens[i-1])<tol) break;
    }
  }
  converged=(abs(loglik[i]+priordens[i]-loglik[i-1]-priordens[i-1])<tol)
  if(!converged){
      warning("EM algorithm in function mixEM failed to converge. Results may be unreliable. Try increasing maxiter and rerunning.")
  }
  return(list(pihat = pi, B=loglik[1:i], 
              niter = i, converged=converged))
}
```

Now a new implementation based on SQUAREM
```{r}
normalize=function(x){return(x/sum(x))}
library(SQUAREM)

fixpoint = function(pi, matrix_lik, prior){  
  m  = t(pi * t(matrix_lik)) # matrix_lik is n by k; so this is also n by k
  m.rowsum = rowSums(m)
  classprob = m/m.rowsum #an n by k matrix
  pinew = normalize(colSums(classprob) + prior-1)
  return(pinew)
}

penloglik = function(pi, matrix_lik, prior){
  m  = t(pi * t(matrix_lik)) # matrix_lik is n by k; so this is also n by k
  m.rowsum = rowSums(m)
  loglik = sum(log(m.rowsum))
  priordens = sum((prior-1)*log(pi))
  return(loglik+priordens)
}

mixEM2 = function(matrix_lik, prior, pi.init = NULL,tol=1e-7, maxiter=5000){
  if(is.null(pi.init)){
    pi = rep(1/k,k)# Use as starting point for pi
  } 
  res = squarem(par=pi.init,fixptfn=fixpoint, objfn=penloglik,matrix_lik=matrix_lik, prior=prior, control=list(maxiter=maxiter,tol=tol))
  return(list(pihat = res$par, B=res$value.objfn, 
              niter = res$iter, converged=res$convergence))
}

```

Now simulate some data
```{r}
set.seed(100)
  sampsize=100000
  sd = c(1,1.1,1.2)
  z = rnorm(sampsize,0,sample(sd,sampsize,replace=TRUE))
  lik = t(vapply(z,dnorm,sd,sd=sd))
  prior=c(1,1,1)
  pi.init=c(0.3,0.2,0.5)
```

```{r}
  system.time(res1<-mixEM(lik,prior,pi.init))
  system.time(res2<-mixEM2(lik, prior, pi.init))
  options(digits=13)
  res1
  res2
```

### 2014-05-29 scratchwork/test.turboem.Rmd
updated EM algorithm (noncxx version) to run with accelerated EM using SQUAREM pacakge; added some .Rmd files in tests and examples that I forgot to add in previous commit

Test turboEM
========================================================

First we have to define a fixpoint function and an objective function
for turboEM to work with
```{r}
normalize=function(x){return(x/sum(x))}

fixpoint = function(pi, matrix_lik, prior){  
  m  = t(pi * t(matrix_lik)) # matrix_lik is n by k; so this is also n by k
  m.rowsum = rowSums(m)
  classprob = m/m.rowsum #an n by k matrix
  pinew = normalize(colSums(classprob) + prior-1)
  return(pinew)
}

penloglik = function(pi, matrix_lik, prior){
  m  = t(pi * t(matrix_lik)) # matrix_lik is n by k; so this is also n by k
  m.rowsum = rowSums(m)
  loglik = sum(log(m.rowsum))
  priordens = sum((prior-1)*log(pi))
  return(loglik+priordens)
}

#Note this doesn't impose the constraint that they have to sum to 1... might be a problem?
pconstr <- function(par) {
  return(all(0 <= par & par <= 1))
}
```

Now simulate some test data
```{r}
  set.seed(100)
  sampsize=10000
  sd = c(1,1.1,1.2)
  z = rnorm(sampsize,0,sample(sd,sampsize,replace=TRUE))
  lik = t(vapply(z,dnorm,sd,sd=sd))
  prior=c(1,1,1)
  pi.init=c(0.3,0.2,0.5)
```

```{r}
  library(turboEM)
  res = turboem(par=pi.init, control.run=list(convtype = "objfn",tol=1e-5),fixptfn=fixpoint, objfn=penloglik, pconstr=pconstr,method=c("em", "squarem", "pem","decme","qn"), matrix_lik=lik,prior=prior)
  options(digits=13)
  res
  library(devtools)
  devtools::load_all("../package/ashr")
  system.time(res2<-mixEM(lik, prior, pi.init))
  length(res2$B)
  res2$B[length(res2$B)]
```

The number of iterations for the ash implementation is `r length(res2$B)` and the objective achieved
is `r res2$B[length(res2$B)]`.

Now try a bigger sample, just comparing EM and squareEM
```{r}
  set.seed(100)
  sampsize=500000
  sd = c(1,1.1,1.2)
  z = rnorm(sampsize,0,sample(sd,sampsize,replace=TRUE))
  lik = t(vapply(z,dnorm,sd,sd=sd))
  prior=c(1,1,1)
  pi.init=c(0.3,0.2,0.5)
  res = turboem(par=pi.init, control.run=list(convtype = "objfn",tol=1e-5),fixptfn=fixpoint, objfn=penloglik, pconstr=pconstr,method=c("em", "squarem"), matrix_lik=lik,prior=prior)
  options(digits=13)
  res
```

Now run the squarem package directly
```{r}
  library(SQUAREM)
  system.time(res3<-squarem(par=pi.init,fixptfn=fixpoint, objfn=penloglik,matrix_lik=lik, prior=prior))
```

It seems from this limited assessment that i) the squareEM approach is the most effective,
and ii) the SQUAREM package implementation is faster then the turboem implementation.

### 2014-05-29 tests/largedata.Rmd
updated EM algorithm (noncxx version) to run with accelerated EM using SQUAREM pacakge; added some .Rmd files in tests and examples that I forgot to add in previous commit

# Test ash on large data

```{r}
  library(ashr)
  set.seed(111)
  z = rnorm(100000)
  z.ash = ash(z,1)
  z.ash2 = ash(z,1,method="fdr")
```

### 2014-06-02 scratchwork/makenew.mixEM.Rmd
updated variational em

@@
   loglik = sum(log(m.rowsum))
-  priordens = sum((prior-1)*log(pi))
+  subset = (prior-1) != 0
+  priordens = sum((prior-1)[subset]*log(pi[subset]))
   return(loglik+priordens)
@@
 
+#as above but with initial iterations done with smaller sample sizes
+#and tolerance set according to sample size
+mixEM3 = function(matrix_lik, prior, pi.init = NULL,tol=1e-7, maxiter=5000){
+  if(is.null(pi.init)){
+    pi = rep(1/k,k)# Use as starting point for pi
+  } 
+  n = nrow(matrix_lik)
+  if(n>1000){       
+    res = squarem(par=pi.init,fixptfn=fixpoint, objfn=penloglik,matrix_lik=matrix_lik[1:1000,], prior=prior, control=list(maxiter=maxiter,tol=1e-3,trace=TRUE))
+    pi.init = res$par
+    print(pi.init)
+  }
+  res = squarem(par=pi.init,fixptfn=fixpoint, objfn=penloglik,matrix_lik=matrix_lik, prior=prior, control=list(maxiter=maxiter,tol=tol,trace=TRUE))
+  return(list(pihat = res$par, B=res$value.objfn, 
+              niter = res$iter, converged=res$convergence))
+}
 ```
@@
   system.time(res2<-mixEM2(lik, prior, pi.init))
+   system.time(res3<-mixEM3(lik, prior, pi.init))
   options(digits=13)
@@
   res2
+  res3
 ```
+
+What we found: subsetting by initially running on only 1000 data points
+did not work so well - it converged to a value of pi that was pretty different from
+the mle for all 100,000 so didn't really save anything.
+
+This section just to run 100 times to see if any warnings results
+```{r}
+#   options(warn=2)
+#    for(i in 1:100){
+#      set.seed(i)
+#     z0 = rnorm(sampsize,0,1)
+#     sd=c(1,2,4,8,16,32)
+#     prior=c(1,1,1,1,1,1)
+#     pi.init=c(1,1,1,1,1,1)/6
+#     lik = t(vapply(z0,dnorm,sd,sd=sd))
+#     print(i)
+#     res2[[i]]<-mixEM2(lik, prior, pi.init)
+#     }
+```

### 2014-06-02 tests/largedata.Rmd
updated variational em

@@
 # Test ash on large data
+
+```{r}
+  require(ashr)
+  eg5 = matrix(scan("~/Dropbox/eg.j5.txt"), byrow=TRUE, nrow=2)
+  temp1=ash(eg5[1,],eg5[2,],cxx=FALSE)
+  temp2=fast.ash(eg5[1,],eg5[2,],prior="nullbiased",pointmass=TRUE,gridmult=2,cxx=FALSE)
+
+```
+
+```{r}
+  eg9 = matrix(scan("~/Dropbox/eg.j9.txt"), byrow=TRUE, nrow=2)
+  temp1=ash(eg9[1,],eg9[2,],cxx=FALSE)
+  temp2=fast.ash(eg9[1,],eg9[2,],prior="nullbiased",pointmass=TRUE,gridmult=2,cxx=FALSE)
+
+```
 
@@
   z = rnorm(100000)
-  z.ash = ash(z,1)
+  z.ash = ash(z,1,cxx=FALSE,method="fdr")
   z.ash2 = ash(z,1,method="fdr")

### 2014-06-02 paper/Rcode/sim1.R

@@ -17,6 +17,7 @@ basicsim=function(mixsd,mixpi_alt,bsd=1,minpi0=0,seedval = 100,nsamp=1000,niter=
   pval = list()
   betahat.ash.n = list()
   betahat.ash.u = list()
+  betahat.ash.ur = list()
   betahat.ash.npm = list()
   betahat.ash.true = list()
   betahat.qval = list()
@@ -35,6 +36,7 @@ basicsim=function(mixsd,mixpi_alt,bsd=1,minpi0=0,seedval = 100,nsamp=1000,niter=
     pval[[i]] = pchisq(zscore[[i]]^2,df=1,lower.tail=F)
     betahat.ash.n[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=TRUE,prior="nullbiased",gridmult=2)
     betahat.ash.u[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=TRUE,prior="uniform",gridmult=2)
+    betahat.ash.ur[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=TRUE,prior="uniform",randomstart=TRUE,gridmult=2)
     betahat.ash.npm[[i]] = ash(betahat[[i]],betahatsd[[i]],pointmass=FALSE,prior="uniform",gridmult=2)
     betahat.ash.true[[i]] = ash(betahat[[i]],betahatsd[[i]],g=normalmix(mixpi,rep(0,length(mixpi)),mixsd))
     

### 2014-06-02 paper/Rcode/simABC.R

@@ -16,6 +16,9 @@ ashsim=function(mixmean,mixsd,mixpi,bsd=1,seedval = 100,nsamp=1000,niter=50){
   fit.ash.true = list()
   fit.mixfdr= list()
   fit.ash.fdr.n = list()
+  fit.ash.fdr.u = list()
+  fit.ash.fdr.hu = list()
+  
   
   fit.mixfdr.enull= list()
   fit.mixfdr.J10= list()
@@ -37,8 +40,8 @@ ashsim=function(mixmean,mixsd,mixpi,bsd=1,seedval = 100,nsamp=1000,niter=50){
     fit.mixfdr[[i]] = mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=TRUE,plot=FALSE)
     
     fit.ash.fdr.n[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="normal",method="fdr")
-    #fit.ash.fdr.u[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="uniform", method="fdr")
-    #fit.ash.fdr.hu[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform", method="fdr")
+    fit.ash.fdr.u[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="uniform", method="fdr")
+    fit.ash.fdr.hu[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform", method="fdr")
     fit.mixfdr.enull[[i]] = mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=FALSE,plot=FALSE)
     fit.mixfdr.J10[[i]] = mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=TRUE,plot=FALSE,J=10)
     #fit.mixfdr.J100[[i]] = mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=TRUE,plot=FALSE,J=100)
@@ -63,7 +66,7 @@ ashsim=function(mixmean,mixsd,mixpi,bsd=1,seedval = 100,nsamp=1000,niter=50){
 } 
 
 set.seed(111)
[…]

### 2014-06-13 tests/largedata.Rmd
updated package with check for negative pis

--- 2014-05-29 tests/largedata.Rmd
@@
 # Test ash on large data
+
+```{r}
+  require(ashr)
+  eg5 = matrix(scan("~/Dropbox/eg.j5.txt"), byrow=TRUE, nrow=2)
+  temp1=ash(eg5[1,],eg5[2,],cxx=FALSE)
+  #temp2=fast.ash(eg5[1,],eg5[2,],prior="nullbiased",pointmass=TRUE,gridmult=2,cxx=FALSE)
+
+```
+
+```{r}
+  eg9 = matrix(scan("~/Dropbox/eg.j9.txt"), byrow=TRUE, nrow=2)
+  temp1=ash(eg9[1,],eg9[2,],cxx=FALSE)
+  #temp2=fast.ash(eg9[1,],eg9[2,],prior="nullbiased",pointmass=TRUE,gridmult=2,cxx=FALSE)
+
+```
 
@@
   z = rnorm(100000)
-  z.ash = ash(z,1)
+  z.ash = ash(z,1,cxx=FALSE,method="fdr")
   z.ash2 = ash(z,1,method="fdr")
+  z.ash$fitted.g
+  z.ash2$fitted.g
 ```

### 2014-06-13 scratchwork/makenew.mixEM.Rmd
fixing dropbox merge issue

@@
 
-penloglik = function(pi, matrix_lik, prior){
+negpenloglik = function(pi, matrix_lik, prior){
   m  = t(pi * t(matrix_lik)) # matrix_lik is n by k; so this is also n by k
@@
   priordens = sum((prior-1)[subset]*log(pi[subset]))
-  return(loglik+priordens)
+  return(-(loglik+priordens))
 }
@@
   } 
-  res = squarem(par=pi.init,fixptfn=fixpoint, objfn=penloglik,matrix_lik=matrix_lik, prior=prior, control=list(maxiter=maxiter,tol=tol))
+  res = squarem(par=pi.init,fixptfn=fixpoint, objfn=negpenloglik,matrix_lik=matrix_lik, prior=prior, control=list(maxiter=maxiter,tol=tol))
   return(list(pihat = res$par, B=res$value.objfn, 
@@
   if(n>1000){       
-    res = squarem(par=pi.init,fixptfn=fixpoint, objfn=penloglik,matrix_lik=matrix_lik[1:1000,], prior=prior, control=list(maxiter=maxiter,tol=1e-3,trace=TRUE))
+    res = squarem(par=pi.init,fixptfn=fixpoint, objfn=negpenloglik,matrix_lik=matrix_lik[1:1000,], prior=prior, control=list(maxiter=maxiter,tol=1e-3,trace=TRUE))
     pi.init = res$par
@@
   }
-  res = squarem(par=pi.init,fixptfn=fixpoint, objfn=penloglik,matrix_lik=matrix_lik, prior=prior, control=list(maxiter=maxiter,tol=tol,trace=TRUE))
+  res = squarem(par=pi.init,fixptfn=fixpoint, objfn=negpenloglik,matrix_lik=matrix_lik, prior=prior, control=list(maxiter=maxiter,tol=tol,trace=TRUE))
   return(list(pihat = res$par, B=res$value.objfn, 

### 2014-06-13 scratchwork/test.turboem.Rmd
fixing dropbox merge issue

@@
   priordens = sum((prior-1)*log(pi))
-  return(loglik+priordens)
+  return(-(loglik+priordens))
 }
@@
   library(turboEM)
-  res = turboem(par=pi.init, control.run=list(convtype = "objfn",tol=1e-5),fixptfn=fixpoint, objfn=penloglik, pconstr=pconstr,method=c("em", "squarem", "pem","decme","qn"), matrix_lik=lik,prior=prior)
+  res = turboem(par=pi.init, control.run=list(convtype = "objfn",tol=1e-5),fixptfn=fixpoint, objfn=negpenloglik, pconstr=pconstr,method=c("em", "squarem", "pem","decme","qn"), matrix_lik=lik,prior=prior)
   options(digits=13)

### 2014-06-13 scratchwork/test.turboem.Rmd
corrected error in sign of objective function calculation for SQUAREM

@@
   pi.init=c(0.3,0.2,0.5)
-  res = turboem(par=pi.init, control.run=list(convtype = "objfn",tol=1e-5),fixptfn=fixpoint, objfn=penloglik, pconstr=pconstr,method=c("em", "squarem"), matrix_lik=lik,prior=prior)
+  res = turboem(par=pi.init, control.run=list(convtype = "objfn",tol=1e-5),fixptfn=fixpoint, objfn=negpenloglik, pconstr=pconstr,method=c("em", "squarem"), matrix_lik=lik,prior=prior)
   options(digits=13)
@@
   library(SQUAREM)
-  system.time(res3<-squarem(par=pi.init,fixptfn=fixpoint, objfn=penloglik,matrix_lik=lik, prior=prior))
+  system.time(res3<-squarem(par=pi.init,fixptfn=fixpoint, objfn=negpenloglik,matrix_lik=lik, prior=prior))
 ```

### 2014-07-13 talks/ASC2014.Rmd
add ASC talk

--- 2014-02-25 talks/UCstat2014.Rmd
@@
 % Matthew Stephens
-% 2014/2/24
+% 2014/7/8
 
@@
 ```{r setup, include=FALSE}
-# set global chunk options
+  #set global chunk options
   opts_chunk$set(cache=TRUE,autodep=TRUE,warning=FALSE)
@@
 
-# Before we get started: Getting Organized
+# Before we get started: Let's get Organized!
 
@@
 - Use *knitr* to help make your research reproducible.
+- Talk about the tools you find useful!
 
@@
 - Apply `qvalue` 
-to identify findings ``significant" at a given False Discovery Rate.
+to identify findings ``significant" at a given FDR.
 
@@
 
-# Example: FDR estimation
+```{r, echo=FALSE}
+  #simple simulated example
+  ncz = 100 # number of bins in z score histograms
+  set.seed(111)
+  hh.betahat = rnorm(1000,0,2)
+  hh.sebetahat = 1
+  hh.zscore = hh.betahat/hh.sebetahat
+  hh.pval = pchisq(hh.zscore^2,df=1,lower.tail=F)
+  hh.ash = ash(hh.betahat,hh.sebetahat, method="fdr")
+  hh.q = qvalue(hh.pval)
+```
 
 ```{r, echo=FALSE}
-  #Set up Analysis of Hedenfalk BRCA data
-  setwd("~/Documents/git/ash/paper/Rcode")
+  plot_FDReg_hist = function(hh.pval,pi0,nc=40,nullcol="blue",altcol="cyan",type=4,title="Distribution of p values",...){
+    hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main=title,nclass=nc,col=altcol,...)
+    if(type>1){
+      abline(h=pi0,col=nullcol,lwd=2)
 
-  altcol="cyan" #colors to use
-  nullcol="blue" 
-  nc=40 #number of bins in histograms
-  ncz = 100 # number of bins in z score histograms
-
-  hh = read.table("../../data/nejm_brca_release_edit.csv",sep=",",skip=3)
-  subset = apply(hh, 1, max)<20
-  hh = hh[subset,]
-
-  labs = read.table("../../data/nejm_brca_release_edit.csv",sep=",",skip=1,nrows=1)
-  labs = 1*(labs=="BRCA1") + 2 * (labs=="BRCA2")
-
-  hh.betahat = apply(hh[,labs==1],1,mean) - apply(hh[,labs==2],1,mean)
-  n1 = sum(labs==1)
-  n2 = sum(labs==2)
-  hh.sebetahat = sqrt(apply(hh[,labs==1],1,var)/n1 + apply(hh[,labs==2],1,var)/n2)
-  hh.zscore = hh.betahat/hh.sebetahat
-  hh.pval = pchisq(hh.zscore^2,df=1,lower.tail=F)
-
-```
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  nc=40 #number of bins in histogram
-  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)
-
+      hh.hist$density=rep(pi0,length(hh.hist$density))  
+      #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
+      plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)
+    }
+    if(type>2){
+    abline(v=0.1,lwd=2,col=2)
+    }
+    if(type>3){
+      text(0.05,1.2,labels="A",col=2,cex=1.2)  
+      text(0.05,0.4,labels="B",col=2,cex=1.2)  
+      text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
+    }
+  }
 ```
@@
 ```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  nc=40 #number of bins in histogram
-  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)
-
-  hh.q = qvalue(hh.pval)
-  abline(h=hh.q$pi0,col=nullcol,lwd=2)
-
-  hh.hist$density=rep(hh.q$pi0,length(hh.hist$density))  
-  #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
-  plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)
+  plot_FDReg_hist(hh.pval,hh.q$pi0,type=1)                 
 ```
-
-Data from Hedenfalk et al, comparing BRCA1 vs BRCA2 expression.
 
@@
 ```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  nc=40 #number of bins in histogram
-  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)
-
-  hh.q = qvalue(hh.pval)
-  abline(h=hh.q$pi0,col=nullcol,lwd=2)
-
-  hh.hist$density=rep(hh.q$pi0,length(hh.hist$density))  
-  #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
-  plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)
-
-  abline(v=0.1,lwd=2,col=2)
-
-  text(0.05,1.2,labels="A",col=2,cex=1.2)  
-  text(0.05,0.4,labels="B",col=2,cex=1.2)  
-  text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(hh.q$pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
+  plot_FDReg_hist(hh.pval,hh.q$pi0,type=2)                 
 ```
 
-# Example: fdr estimation
+# Example: FDR estimation
 
 ```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  nc=40 #number of bins in histogram
-  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)
-
+  plot_FDReg_hist(hh.pval,hh.q$pi0,type=3)                 
 ```
 
-# Example: fdr estimation
+# Example: FDR estimation
 
 ```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  nc=40 #number of bins in histogram
-  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)
-
-  require(fdrtool)
-  hh.gren = grenander(ecdf(hh.pval))
-  abline(h=min(hh.gren$f.knots),col=nullcol,lwd=2)  
-  lines(hh.gren$x.knots,hh.gren$f.knots,lwd=2)
+  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4)                 
 ```
 
-# Example: fdr estimation
+# Is this an important problem?
+
+- The original paper introducing FDR (Benjamini and Hochberg, 1995) has
+been cited 21,787 times (May 2014) according to Google Scholar.
+
+- That is three times a day for the last 19 years!
+
+# Problem 1: The Zero Assumption (ZA)
+
+- The standard `qvalue` 
+approach assumes that all the $p$ values near 1 are null.
+
+- Analogously, one can assume that all Z scores near 0 are null. Efron refers to this as the ``Zero Assumption".
+
+- Seems initially natural.
+
+```{r, echo=FALSE, include=FALSE}
+require(fdrtool)
+  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue",plot=FALSE)
+require(locfdr)
+  hh.locfdr = locfdr(hh.zscore,nulltype=0,plot=0)
+require(mixfdr)
+  hh.mixfdr = mixFdr(hh.zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
+```
+
+# Implied distribution of $p$ values under $H_1$
 
 ```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  nc=40 #number of bins in histogram
-  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)
-
-  require(fdrtool)
-  hh.gren = grenander(ecdf(hh.pval))
-  abline(h=min(hh.gren$f.knots),col=nullcol,lwd=2)  
-  lines(hh.gren$x.knots,hh.gren$f.knots,lwd=2)
-  abline(v=0.1,lwd=2,col=2)
-  text(0.1,0.9,labels="a",col=2,cex=1)  
-  text(0.1,0.34,labels="b",col=2,cex=1.2)  
-  text(0.6,3,labels=paste0("fdr = b/(a+b) =  ",round(min(hh.gren$f.knots)/approx(hh.gren$x.knots,hh.gren$f.knots,0.1)$y,2)),cex=1.2)
+  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4)                 
 ```
 
-# FDR problem 1: different measurement precision
+# Implied distribution of Z scores under alternative
 
-- If some effects are measured very imprecisely, those tests ``lack power"
-and simply add noise
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+#plot a histogram of z scores, highlighting the alternative distribution
+#of z scores that is implied by localfdr values lfdr.
+  nullalthist = function(z,lfdr,nullcol="blue",altcol="cyan",...){
+    h=hist(z, freq=FALSE,col=nullcol,nclass=ncz,...)
+    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
+    h$density = (1-avlfdr) * h$density
+    plot(h,add=TRUE,col=altcol,freq=FALSE)
+  }
 
-- In particular, such tests increase the estimated number of nulls, and increase
-the FDR for other tests
+  nullalthist(hh.zscore,hh.fdrtool$lfdr)  
+```
 
-- It would seem preferable to simply ignore the tests with very low precision. Summarizing each test by a $p$ value (or $Z$ score) loses the information about precision.
+# FDR problem 2: different measurement precision
+
+- In some cases the measurement precisions differ among units
+
+- Eg Expression levels of low-expressed genes have less precision than high-expressed genes
+
+- If some effects are measured less precisely than others, those tests ``lack power"
+and dilute signal, increasing FDR
 
@@
 ```{r, echo=FALSE}
-setwd("~/Documents/git/ash/talks/")
-## load Poisson_binomial and ash functions 
-source("../../stat45800/code/Rcode/PoissonBinomial.funcs.R")  
-#source("../Rcode/ash.R")
-
 x = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE,include=FALSE}
 
-cc = x[, 2:5]
-
-g = c(-1, -1, 1, 1)
-ngene = dim(x)[1]
-
-cc.assoc = counts.associate(cc, g, 1)
-zdat.ash = cc.assoc$zdat.ash
-zdat = cc.assoc$zdat
-
+load("../scratchwork/Smemo_qbinom_zdat.RData")
 # two-sided test
@@
 qval.high = qvalue(pval.high)
-cc.assoc.high = counts.associate(cc[highxp,],g,1)
-zdat.ash.high = cc.assoc.high$zdat.ash
-zdat.high = cc.assoc.high$zdat
+
+zdat.ash = ash(zdat[3,],zdat[4,],df=2,method="fdr",mixcompdist="halfuniform")
+zdat.ash.high = ash(zdat[3,highxp],zdat[4,highxp],df=2,method="fdr",mixcompdist="halfuniform")
 
@@
 ```{r, echo=FALSE,fig.height=4,fig.cap=""}
-h=hist(pval,prob=TRUE,ylim=c(0,4),main="p values, all genes",xlab="p value",breaks=seq(0,1,length=21),col=altcol)
-abline(h=qval$pi0,col=nullcol,lwd=2)
-h$density=rep(qval$pi0,length(h$density))  
-plot(h,col=nullcol,add=TRUE,freq=FALSE)
-abline(v=0.05,lwd=3)
-text(0.6,3,labels=paste0("FDR = ",round(max(qval$qval[pval<0.05]),2)),cex=1.2)
+qval= qvalue(pval)
+plot_FDReg_hist(pval,pi0=qval$pi0,ylim=c(0,4))
 ```
@@
 ```{r, echo=FALSE,fig.height=4,fig.cap=""}
-h=hist(pval[xx<1000],prob=TRUE,ylim=c(0,4),main="p values, low count genes",xlab="p value",breaks=seq(0,1,length=21),col=altcol)
-qval.low=qvalue(pval[xx<1000])
-pval.low=pval[xx<1000]
-abline(h=qval.low$pi0,col=nullcol,lwd=2)
-h$density=rep(qval.low$pi0,length(h$density))  
-plot(h,col=nullcol,add=TRUE,freq=FALSE)
-abline(v=0.05,lwd=3)
-#text(0.6,3,labels=paste0("FDR = ",round(max(qval.low$qval[pval.low<0.05]),2)),cex=1.2)
+qval.low=qvalue(pval[!highxp])
+pval.low=pval[!highxp]
+plot_FDReg_hist(pval.low,pi0=qval.low$pi0,ylim=c(0,4))
 ```
@@
 ```{r, echo=FALSE, fig.height=4,fig.cap=""}
-
-h=hist(pval.high,prob=TRUE,ylim=c(0,4),main="p values, high count genes",xlab="p value",col=altcol)
-
-abline(h=qval.high$pi0,col=nullcol,lwd=2)
-h$density=rep(qval.high$pi0,length(h$density))  
-plot(h,col=nullcol,add=TRUE,freq=FALSE)
-abline(v=0.05,lwd=3)
-#text(0.6,3,labels=paste0("FDR = ",round(max(qval.high$qval[pval.high<0.05]),2)),cex=1.2)
-#qval.high$q[pval.high<0.05] gives FDR for p<0.05
+plot_FDReg_hist(pval.high,pi0=qval.high$pi0,ylim=c(0,4))
 ```
 
-# FDR problem 1: low count genes add noise, increase q values
-```{r,echo=FALSE, fig.height=4,fig.cap=""}
-plot(qval.high$qval,qval$qval[xx>1000],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
-abline(a=0,b=1,col=2)
+# Low-count genes dilute signal at high-count genes
+```{r, echo=FALSE, fig.height=4,fig.cap=""}
+plot_FDReg_hist(pval,pi0=qval$pi0,ylim=c(0,4))
 ```
 
-# FDR problem 1: Summary
-
-- Analyzing $p$ values or $Z$ scores doesn't fully account 
-for measurement precision.
-
-# Problem 2: The Zero Assumption (ZA)
-
-- The standard `qvalue` 
-approach assumes that all the $p$ values near 1 are null.
-
-- Analogously, one can assume that all Z scores near 0 are null. Efron refers to this as the ``Zero Assumption".
-
-- The ZA allows us to estimate the null proportion, $\pi_0$, using the density of $p$ values near 1 (or $Z$ scores near 0).
-
-# Problem 2: The ZA
-
-- The ZA seems initially natural.
-
-- However, it turns out to imply unrealistic assumptions about the distribution of non-zero effects.
-
-```{r, echo=FALSE, include=FALSE}
-require(fdrtool)
-  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue",plot=FALSE)
-require(locfdr)
-  hh.locfdr = locfdr(hh.zscore,nulltype=0,plot=0)
-require(mixfdr)
-  hh.mixfdr = mixFdr(hh.zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
-```
-
-# Implied distribution of $p$ values under $H_1$
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  nc=40 #number of bins in histogram
-  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)
-
-  hh.q = qvalue(hh.pval)
-  abline(h=hh.q$pi0,col=nullcol,lwd=2)
-
-  hh.hist$density=rep(hh.q$pi0,length(hh.hist$density))  
-  #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
-  plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)
-
-  abline(v=0.1,lwd=2,col=2)
-
-  text(0.05,1.2,labels="A",col=2,cex=1.2)  
-  text(0.05,0.4,labels="B",col=2,cex=1.2)  
-  text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(hh.q$pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
-```
-
-# Implied distribution of Z scores under alternative (fdrtool)
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-#plot a histogram of z scores, highlighting the alternative distribution
-#of z scores that is implied by localfdr values lfdr.
-  nullalthist = function(z,lfdr,...){
-    h=hist(z, freq=FALSE,col=nullcol,nclass=ncz,...)
-    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
-    h$density = (1-avlfdr) * h$density
-    plot(h,add=TRUE,col=altcol,freq=FALSE)
-  }
-
-  nullalthist(hh.zscore,hh.fdrtool$lfdr,main="fdrtool")  
-```
-
-# Implied distribution of Z scores under alternative (locfdr)
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-
-  nullalthist(hh.zscore,hh.locfdr$fdr,main="locfdr")
-```
-
-# Implied distribution of Z scores under alternative (mixfdr)
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-
-  nullalthist(hh.zscore,hh.mixfdr$fdr,main="mixfdr")
-  dev.off()
+# FDR problem 2: low count genes add noise, increase q values
+```{r,echo=FALSE, fig.height=4,fig.cap=""}
+plot(qval.high$qval,qval$qval[highxp],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
+abline(a=0,b=1,col=2)
 ```
@@
 
-- By summarizing each observation by a $Z$ score or $p$ value, 
-standard fdr tools ignore precision of different measurements
+Standard tools are unduly conservative.
 
-- Standard tools make the ZA, which implies actual effects have a (probably unrealistic) bimodal distribution. [and tends to overestimate $\pi_0$, losing power]
+- The ZA, which implies actual effects have a (probably unrealistic) bimodal distribution; causes overestimate of $\pi_0$, losing power.
 
-- Also standard tools focus only on zero vs non-zero effects. (eg what if we would
-like to identify genes that have at least a 2-fold change?)
+- By focussing on $p$ values, low-precision measurements can dilute high-precision measurements.
 
@@
 $$Z_j \sim f_Z(.) = \pi_0 N(.;0,1) + (1-\pi_0) f_1(.)$$
-where $f_1$ is to be estimated from the data.
+where $f_1, \pi_0$ are to be estimated from the data.
 
@@
 
-- $\text{fdr}(Z) \approx \pi_0 N(Z; 0,1)/ f_Z(Z)$
+- Once $f_1$ and $\pi_0$ estimated, FDR calculations are straightforward.
 
-# FDR: The New Deal
+# FDR: A New Deal
 
@@
 [From $\hat\beta_j \sim N(\beta_j, s_j)$]
-
-- fdr given by $$p(\beta_j =0 | \hat\beta_j) = \pi_0 p(\hat\beta_j | \beta_j=0)/p(\hat\beta_j)$$
+Or, better, use a $t$ likelihood if $s_j$ estimated using few observations.
 
@@
 
-- A convenient way to model $g$ is by a mixture of 0-centered
+- A convenient way to model $g$: mixture of 0-centered
 normal distributions: 
@@
 
-- Estimating $g$ comes down to estimating $\pi$. Joint estimation of $\pi_0,\pi$ easy by maximum likelihood (EM algorithm) or variational Bayes.
+- Estimating $g$ comes down to estimating $\pi$. Joint estimation of $\pi_0,\pi$ easy by maximum likelihood (EM algorithm).
 
 - By allowing $K$ large, and $\sigma_k$ to span a dense grid of values,
-we get a fairly flexible unimodal symmetric distribution.
+we get a flexible unimodal symmetric distribution.
 
@@
 
-- If allow a very large number of uniforms this provides the non-parametric mle for $g$; cf Grenander 1953; Campy + Thomas.
+- If allow a very large number of uniforms this provides the non-parametric mle for $g$; cf Grenander 1953; Cordy + Thomas 1997.
 
@@
 
-# Issue: identifiability of $\pi_0$
+# Adaptive Shrinkage
 
-- For estimating False Discoveries, we are asking whether $\beta_j = 0$.
+- This approach actually provides a full posterior distribution for each $\beta_j$.
 
-- However, the data cannot distinguish between $\beta_j = 0$ and $\beta_j$ "very
-small"
+- So easy to obtain point estimates and credible intervals.
 
-- As a result $\pi_0$ is formally unidentifiable. Eg data can never rule out $\pi_0=0$.
+- Because $g(\beta)$ is unimodal, the point estimates (and CIs) will tend to be ``shrunk" towards the overall mean (0).
 
-# Issue: identifiability of $\pi_0$
+- Because $g(\beta)$ is estimated from the data, the amount
+of shrinkage is adaptive to signal in the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each observation.
 
-- The Zero assumption (ZA) solves the identifiability problem by assuming that
-there *are* no $\beta_j$ near zero!
+- So we call the approach ``Adaptive Shrinkage" (ASH).
 
-- The ZA makes $\pi_0$ identifiable.
+# Recall Problem 1: distribution of alternative Z values multimodal
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+nullalthist(hh.zscore,hh.fdrtool$lfdr,main="qvalue")  
+```
 
-- Another view is that the estimate of $\pi_0$ under ZA will systematically tend to overestimate $\pi_0$, and so is ``conservative".
+# Problem Fixed: distribution of alternative Z values unimodal
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+nullalthist(hh.zscore,hh.ash$lfdr,main="ash")
+```
 
-- That is it provides an ``upper bound" on $\pi_0$
+# Example: FDR estimation
 
-# Identifiability of $\pi_0$: Solution 1
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4,title="qvalue")                 
+```
 
-- We replaced the ZA with the unimodal assumption on g.
+# Example: FDR estimation
 
-- This does not make $\pi_0$ identifiable, but it does effectively provide an upper bound on $\pi_0$.
+```{r, echo=FALSE,fig.height=4,fig.cap=""}
+  plot_FDReg_hist(hh.pval,get_pi0(hh.ash),type=4,title="ash")                 
+```
 
-- Indeed, we saw that when we estimated $\pi_0$ under the ZA the data
-then contradicted the unimodal assumption on g. Thus the upper bound is
-more conservative than under ZA.
+# Recall Problem 2: low count genes add noise, increase q values
+```{r,echo=FALSE, fig.height=4,fig.cap=""}
+plot(qval.high$qval,qval$qval[highxp],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
+abline(a=0,b=1,col=2)
+```
 
-- In practice, implement upper bound by using penalized likelihood that encourages $\pi_0$ to be as big as possible.
+# Problem Fixed: incorporating precision reduces influence of low-count genes
+```{r,echo=FALSE, fig.height=4,fig.cap=""}
+plot(zdat.ash.high$qvalue,zdat.ash$qvalue[highxp],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
+abline(a=0,b=1,col=2)
+```
 
-# Illustration: Simulated Example
+# A new problem: an embarrassment of riches
+
+- If the null is mostly false, the new approach can  provide unsettling results
+
+- The FDR can be small for all observations, even those with $p \approx 1$!
+
+- In the illustrative example, the maximum $q$ value is `r round(max(hh.ash$qvalue),2)`
+
+# Perhaps we didn't really understand the question?
+
+- Problem arises only if we insist on asking question ``is $\beta_j=0$?"
+
+- Given enough signal, we become convinced that very few of the $\beta_j=0$
+
+- But for some $\beta_j$ we still may have little information about actual value
+
+- Suggests a change of focus: ask for which $\beta_j$ are we confident about the sign (cf Gelman et al, 2012).
+
+# The False Sign Rate
+
+- Suggestion: replace FDR with local false sign rate (lfsr), the probability that if we say an effect is positive (negative), it is not.
+
+- Example: suppose we estimate that $\Pr(\beta_j<0)=0.95, \Pr(\beta_j=0) = 0.025$ and $\Pr(\beta_j>0)=0.025$. Then we report $\beta_j$ as a ``(negative) discovery", and estimate its lfsr as 0.05.
+
+# Even with many signals, large $p$ values have high lfsr
+
+```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
+  #res=data.frame(p=hh.pval, lfsr=hh.ash$lfsr, qvalue = hh.ash$qvalue)
+  library("reshape2")
+  res=data.frame(p=pval, lfsr=zdat.ash$lfsr, qvalue = zdat.ash$qvalue)
+  res.melt = melt(res, id.vars=c("p"),variable.name="Measure")
+  cbbPalette <- c("#000000", "#D55E00", "#CC79A7")
+  labels = c('qvalue','lfsr')
+  breaks = c("qvalue","lfsr")
+
+  pp= ggplot(data=res.melt,aes(p,value,color=Measure)) +geom_point(shape=1) +
+      geom_abline(colour = "black") +
+        xlab("p-value") +
+        ylab("lfsr/qvalue")
+
+  print(pp +scale_y_continuous(limits=c(0,1)) +
+          scale_x_continuous(limits=c(0,1))  +
+           scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
+          coord_equal(ratio=1))
+
+```
+
+# Advantages of False Sign Rate
+
+- The FDR is typically not actually identifiable from data.
+
+- This is because data cannot distinguish between $\beta_j = 0$ and $\beta_j$ "very
+small". So $\pi_0$ is not identifiable, and FDR is very sensitive to $\pi_0$.
+
+- So methods for estimating $\pi_0$ and FDR, including those presented here, are designed to be ``conservative" (i.e. overestimate the FDR).
+
+- The False Sign Rate is much less senstive to $\pi_0$, and hence more identifiable from data!
+
+# Simulated example: $\pi_0$ not identifiable.
 
 ```{r,echo=FALSE,include=FALSE}
-load("../paper/Rcode/sim1_out.RData") #load simulation results
+load("../paper/Rcode/sim1.RData") #load simulation results
 source("../paper/Rcode/plot_pi0.R")
@@
 ```{r, echo=FALSE,fig.height=4,fig.cap=""}
-plot_pi0(list(simres1))
+plot_pi0(list(simres1a))
 ```
 
-# Example: BRCA data
-
-```{r, echo=FALSE,include=FALSE}
-hh.ash = ash(hh.betahat,hh.sebetahat)
-hh.ashz = ash(hh.zscore,1, pointmass=TRUE, prior="nullbiased")
-hh.ash.hu = ash(hh.betahat, hh.sebetahat,mixcompdist="halfuniform")
-```
-
-```{r, echo=FALSE, include=FALSE}
-  hh.ash.fdr = ash(hh.zscore,1,pointmass = TRUE,prior="nullbiased")
-  hh.ash.shrink = ash(hh.zscore,1)
-```
-
-# Recall Problem: distribution of alternative Z values multimodal
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-nullalthist(hh.zscore,hh.mixfdr$fdr,main="mixfdr")
-```
-
-# Problem Fixed: distribution of alternative Z values unimodal
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-nullalthist(hh.zscore,hh.ashz$ZeroProb,main="ash")
-```
-
-# BRCA1: Compare $\pi_0$ estimates
-```{r}
-  round(c(hh.fdrtool$param[3],
-  hh.locfdr$fp0[1,3],
-  hh.mixfdr$pi[1],
-  hh.ashz$fitted.g$pi[1]),2)
-```
-
-# BRCA1: Compare number significant at fdr<0.05
-
-```{r}
-  c(sum(hh.fdrtool$lfdr<0.05),
-  sum(hh.locfdr$fdr<0.05),
-  sum(hh.mixfdr$fdr<0.05),
-  sum(hh.ashz$ZeroProb<0.05))
-```
-
-# Identifiability of $\pi_0$ and the False Sign Rate
-
-- Identifiability of $\pi_0$ is primarily an issue if we insist on asking question is $\beta_j=0$?
-
-- How about we change focus: assume *none* of the $\beta_j$ are zero ("one group approach"), and ask for which $\beta_j$ are we confident about the sign (Gelman et al, 2012).
-
-- Positive and negative effects are often treated differently in practice anyway.
-
-- That is we replace fdr with False Sign Rate (fsr), the probability that if we say an effect is positive (negative), it is not.
-
-- Example: suppose we estimate that $\Pr(\beta_j<0)=0.975$ and $\Pr(\beta_j>0)=0.025$. Then we report $\beta_j$ as a ``(negative) discovery", and estimate its fsr as 0.025.
-
-# The fsr is more robust than fdr
+# Simulated Example: so fdr not identifiable
 
@@
 source("../paper/Rcode/plot_lfsr.R")
-  plot_lfsr(list(simres1),0.1,ptype="lfdr")
+  plot_lfsr(list(simres1a),0.1,ptype="lfdr")
 ```
 
-# The fsr is more robust than fdr
+# Simulated Example: fsr much more identifiable
 
 ```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-  plot_lfsr(list(simres1),0.1,ptype="lfsr")
-```
-
-# Estimation and Shrinkage
-
-- Besides allowing one to estimate fdr and fsr, 
-this approach also provides a full posterior distribution for each $\beta_j$.
-
-- So for example we can easily compute fdrs for discoveries other than ``non-zero" (eg compute $\Pr(\beta_j > 2 | \hat\beta_j)$).
-
-- And use it to obtain point estimates and credible intervals for each $\beta_j$, taking account of information from all the other $\beta_j$.
-
-- Because $f(\beta)$ is unimodal, the point estimates will tend to be ``shrunk" towards the overall mean (0).
-
-- Because $f(\beta)$ is estimated from the data, the amount
-of shrinkage is adaptive to the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each gene.
-
-- So we call the approach ``Adaptive Shrinkage" (ASH).
-
-# Example: ASH applied to mouse data
-
-```{r, fig.height=4,echo=FALSE,fig.cap=""}
-hist(zdat[3,],main="Raw effect size estimates",xlab="betahat",xlim=c(-2,2),prob=T,ylim=c(0,3))
-```
-
-# Example: ASH applied to mouse data
-
-```{r, fig.height=4,echo=FALSE,fig.cap=""}
-hist(zdat[3,],main="Raw effect size estimates",xlab="betahat",xlim=c(-2,2),prob=T,ylim=c(0,3))
-t=seq(-2,2,length=201)
-lines(t,dnorm(t,0,sd=0.032), col=2, lwd=2)
-```
-
-# Example: ASH applied to mouse data
-
-```{r, fig.height=4,echo=FALSE,fig.cap=""}
-hist(zdat.ash$PosteriorMean,main="Shrunken estimates",xlab="shrunk betahat",xlim=c(-0.1,0.1),prob=T)
-lines(t,dnorm(t,0,sd=0.032), col=2, lwd=2)
-```
-
-# Example: ASH applied to mouse data
-
-```{r, fig.height=4,echo=FALSE,fig.cap=""}
-plot(zdat[3,],zdat.ash$PosteriorMean,main="Estimates vs Shrunken estimates",xlab="betahat",ylab="shrunk betahat",xlim=c(-2,2),ylim=c(-0.1,.1),prob=T)
-abline(a=0,b=1,col=2)
+  plot_lfsr(list(simres1a),0.1,ptype="lfsr")
 ```
@@
 
-- But by using two numbers ($\hat\beta,s$) instead of one ($p$ values or $z$ scores) precision of different measurementscan be better accounted for.
+- But by using two numbers ($\hat\beta,s$) instead of one ($p$ values or $z$ scores) varying precision of 
+measurements is better accounted for.
 
@@
 
-- False Sign Rate is more robust to assumptions, and perhaps
-therefore preferable, than False Discovery Rate.
+- False Sign Rate preferable to False Discovery Rate: more identifiable, and
+better representation of information in data for ``high-signal" situations.
 
@@
 
-- Currently applying it to wavelet shrinkage applications.
-
-# Guarantees?
-
-- ``I think you have some nice ideas. How will you convince
-people to use them?" (C Morris)
-
-- Theory anyone?
+- E.g. Currently applying it to wavelet shrinkage applications.
 
@@
 
-- Incorporate $t$ likelihood as well as normal.
+- Incorporate shrinkage of variances and not just means. (e.g. ``moderated $t$ test", Smyth et al)
 
-- Incorporate shrinkage of variances and not just means.
+- Allow $g(\cdot;\pi)$ to depend on covariates $X$.
 
-- Extend to allow $g(\cdot;\pi)$ to depend on covariates $X$.
-
-- Extend to allow for correlations in the measured $\hat\beta_j$.
+- Allow for correlations in the measured $\hat\beta_j$.
 
@@
 
+- to the developers of **R**, **knitr**, **Rstudio** and **Pandoc**.
+
 - to the several postdoctoral researchers and students
 who have worked with me on related topics.
-- Including Scott Powers, Mengyin Lu, Tian Sen, Wei Wang, Zhengrong Xing.
 
-# Reproducible research
+- Including Ester Pantaleo, Scott Powers, Mengyin Lu, Sen Tian, Wei Wang, Zhengrong Xing.
 
-- This document is produced with **knitr**, **Rstudio** and **Pandoc**.
+- NHGRI for funding.
 
-- For more details see my `stephens999/ash` repository at `http://www.github.com/stephens999/ash`
-
-- Website: `http://stephenslab.uchicago.edu`
+- `ashr` package: `http://www.github.com/stephens999/ash`
 
@@
 - Maybe fsr makes more sense anyway?
+
+# Adaptive Shrinkage of point estimates
+
+- Recall idea: amount of shrinkage depends on measurement precision, $s_j$.
+
+# Adaptive Shrinkage of point estimates
+
+```{r, fig.height=4,echo=FALSE,fig.cap=""}
+  res=data.frame(betahat =zdat[3,], Estimate=zdat.ash$PosteriorMean,highxp=highxp)
+  cbbPalette <- c("#000000", "#D55E00", "#CC79A7")
+  labels = c('betahat','Estimate')
+  breaks = c("betahat","Estimate")
+
+  pp= ggplot(data=res,aes(betahat,Estimate,col=as.factor(highxp))) +geom_point(shape=1) +
+      geom_abline(colour = "black") +
+        xlab("betahat") +
+        ylab("shrunk betahat")
+
+  print(pp +scale_y_continuous(limits=c(-2,2)) +
+          scale_x_continuous(limits=c(-2,2))  +
+           scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
+          coord_equal(ratio=1))
+
+#plot(zdat[3,],zdat.ash$PosteriorMean,main="Less shrinkage for high-count genes (red)",xlab="betahat",ylab="shrunk betahat",xlim=c(-2,2),ylim=c(-2,2), col=1+highxp)
+#abline(a=0,b=1,col=2)
+```
 

### 2014-07-18 paper/main.tex
added uc talk
fixed conflict

@@
 
-\documentclass[10pt]{article}
+\documentclass[11pt]{article}
 
@@
 {\Large
-\textbf{Adaptive shrinkage and the false sign rate}
+\textbf{False Discovery Rates: A New Deal}
 }
@@
 
-It is common in statistics that you measure many ``similar" things imperfectly, and wish to estimate their values.
-The situation arises commonly in the kinds of genomics applications I am often involved in, but also in other areas of statistics.
-In genomics, for example, a very common goal
-is to compare the mean
-expression (activity) level of many genes in two conditions.
-Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-measurements are made on only a small number of
-samples in each condition - sometimes as few as one
-sample in each condition. Thus the error in estimates
-of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-in estimates of $\beta_j$ still greater.
+Since its introduction in 19xx by Benjamini and Hochberg 
+\cite{bh:xx}, the ``False Discovery Rate" (FDR) has quickly established itself
+as a key concept in modern statistics. 
+%, and the primary tool by which most practitioners handle large-scale multiple testing.
+Here we consider one setting in which FDR-based methods are widely used:
+an experimenter has measured a large numbers of ``effects", each with an associated
+measurement precision (standard error), 
+and the goal is to identify which of the effects we can declare to be non-zero (with some prescribed level of confidence).
+This setting arises commonly in many areas of statistics, and particularly in genomics applications.
+For example, in genomics
+a common goal is to identify ``differentially-expressed genes" -- that is, genes
+whose mean expression (activity) level differs in two conditions.
 
-A fundamental idea is
-that the measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.
+%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
+%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
+%measurements are made on only a small number of
+%samples in each condition - sometimes as few as one
+%sample in each condition. Thus the error in estimates
+%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
+%in estimates of $\beta_j$ still greater.
+%
 
-Suppose that we measure, with error, a series of ``effects", $\beta_1,\dots,\beta_J$. To take just one concrete 
-example, $\beta_j$ could be 
-the difference in the mean (log) expression levels of gene $j$ ($j=1,\dots,J$) between 2 conditions. In this case, a measurement of
-the effect might be the difference in sample means obtained in the two conditions. We will let $\hat\beta_j$ denote the measured value of $\beta_j$, and assume that
-each measurement comes with an associated standard error, 
-$s_j$. A key aim here will be to take proper account of the fact that some measurements may be more precise than others: that is, to take proper account of variation in $s_j$ across $j$.
+Many tools have been developed for this problem \cite{}, and several are widely used.
+Here we argue that these existing tools have sub-optimal features that can be improved on, at least in some circumstances.
+In particular we argue that existing tools i) fail to take proper account of variations in the measurement precision of different effects;
+ii) make, implicitly, unrealistic assumptions about the distribution of the non-zero effects. We provide methods
+to address these problems, implemented in an R package {\tt ashr} (short for ``Adaptive Shrinkage") and demonstrate that these methods can lead to a substantial reduction in estimated FDRs (or, equivalently,
+a substantial increase in the number of discoveries at a given FDR).
 
-A common goal, particularly in genomic studies, is to identify which $\beta_j$ differ from zero. This is commonly tackled
-by first computing an effect size estimate ($\hat\beta_j$) and its standard error ($s_j$), converting this to a $Z$ score ($Z_j=\hat\beta_j/s_j$)
-and a corresponding $p$ value ($p_j$), testing $H_j:\beta_j=0$.  Then standard methods (e.g. the qvalue package) can be used
-to estimate False Discovery Rates at any given threshold.
-
-There are two issues with this approach that I would like to address here.
-The first is that it really does not take proper account of the measurement errors.
-To see this, consider an example where half the measurements are quite precise, and the other half are really, really, poor.
-Intuitively, the poor measurements tell us nothing, and any sane analysis should effectively ignore them.
-However, in a standard FDR-type analysis, these poor measurements add ``noise" and affect estimated FDRs.
-This is because the $p$ values from the poor measurements will be effectively uniformly distributed, and some will be significant at any given threshold.
-
-The second issue is that directly modeling the p values, say via non-parametric methods, without taking account of their precision, 
-can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-The solution is to model beta hat as a convolution of some distribution g and an error component.
+Although we motivate our tools by FDR, we are providing generic tools for shrinkage, and
+these could be useful in settings such as wavelets.
+The potential for shrinkage-based estimation to
+address the multiple comparisons problems tackled by the FDR
+has been highlighted
+previously, including Greenland and Robins (1991),
+Efron (2008) and Gelman et al (2012). [Note, check also Louis, JASA, 1984]
 
@@
 
-The potential for shrinkage-based estimation to
-address the multiple comparisons setting 
-has been highlighted
-previously, including Greenland and Robins (1991),
-Efron (2008) and Gelman et al (2012). [Note, check also Louis, JASA, 1984]
-
  \section*{Methods}
 
- Suppose that we are interested in the values of $n$ ``effects" $\beta_j$ ($j=1,\dots,n$).
+ \subsection*{Notation, and existing methods}
+
+Suppose that we are interested in the values of $J$ ``effects" $\beta_j$ ($j=1,\dots,J$).
 In some contexts our interest may focus on which of the $\beta_j$ are ``significantly" different from zero, whereas in other contexts our interest may focus on estimating their values; the methods described here are suited to both these contexts.
- We assume that we have obtained data $D_1,\dots,D_n$ that provide independent estimates $\hat\beta_1,\dots,\hat\beta_n$ of these effects,
- with corresponding (estimated) standard errors $s_1,\dots,s_n$. For our purposes, these estimates and standard errors could be obtained from the data
+ We assume that we have obtained data $D_1,\dots,D_J$ that provide independent estimates $\hat\beta_1,\dots,\hat\beta_J$ of these effects,
+ with corresponding (estimated) standard errors $s_1,\dots,s_J$. For our purposes, these estimates and standard errors could be obtained from the data
  using any method, provided they (approximately) satisfy 
@@
  [In some settings a $t$ distribution assumption for $\hat\beta_j | s_j, \beta_j$ may be more appropriate, and could also be incorporated into our methods; see Discussion.]
+
+A standard genomics analysis pipeline would first convert these numbers to a $Z$ score ($Z_j=\hat\beta_j/s_j$)
+and a corresponding $p$ value ($p_j$), testing $H_j:\beta_j=0$.
+
+Definition of FDR and lfdr (Bayesian); connection to frequentist definitions (Storey).
+
+Then standard methods (e.g. the qvalue package) can be used
+to estimate False Discovery Rates at any given threshold.
+
+\subsection*{Problems with existing methods}
+
+ There are two issues with this approach that I would like to address here.
+The first is that it really does not take proper account of the measurement errors.
+To see this, consider an example where half the measurements are quite precise, and the other half are really, really, poor.
+Intuitively, the poor measurements tell us nothing, and any sane analysis should effectively ignore them.
+However, in a standard FDR-type analysis, these poor measurements add ``noise" and affect estimated FDRs.
+This is because the $p$ values from the poor measurements will be effectively uniformly distributed, and some will be significant at any given threshold.
+
+Illustration (Figure).
+
+Note that the problem could be reduced by discarding measurements that are too noisy. However,
+note limitation of this.
+
+The second issue is that directly modeling the p values, or z scores, say via non-parametric methods, without taking account of their precision, 
+can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
+distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
+The solution is to model beta hat as a convolution of some distribution g and an error component.
+
+Illustration (Figure).
+
  %(See Discussion for ideas on how our methods may be applicable more generally; for example if, in place of $s_j$, we have a $p$ value testing $\beta_j=0$.)
@@
 \section*{Results}
+
+Note: we'd like to make sure it behaves ok in the case where most observations are null.
+Perhaps simulate with 999 null and 1 alternative?
+Also maybe look at performance for small sample sizes - e.g. n=1?
 
@@
 
-By modeling the z scores directly under the alternative, rather than the z scores as being the truth plus noise,
-maybe that is what is most problematic? Because the resulting distribution of z scores is just not possible.
-
 Note that \cite{muralidharan:2012} models $z$ scores as something plus noise under both $H_0$ and $H_1$, which
@@
 Rice and Spiegelhalter - BRCA data?
+
+A fundamental idea is
+that the measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.
 

### 2014-09-03 issues/issue21/Ash Nonzeromean.rmd
Create Ash Nonzeromean.rmd

---
title: "Nonzero Mean ASH"
author: "Chaoxing Dai"
date: "2014-Sep-03"
output: pdf_document
---

#Likelihood

By assuming a unimodal distribution, we have the hierarchical model being the following:  
$$\beta _{j}\sim g(.) \sim \sum_{k=1}^{K} \pi _{k}f_{k}(.) \sim\sum_{k=1}^{K} \pi _{k}N(\cdot ;\mu,\sigma_{k}^{2})$$

$$\widehat{\beta}_{j}\mid \beta_{j},s_{j}^{2} \sim N(\cdot ;\beta_{j},s_{j}^{2})$$

Then we would have the likelihood for $\pi$ and $\mu$ being (Here we assume iid or exchangable data)

$$L(\pi,\mu)=\prod_{j=1}^{n} p(D_{j}\mid \pi,\mu)=\prod_{j=1}^{n} \int p(\widehat{\beta}_{j} \mid s_{j},\pi,\mu,\beta_{j}) p(\beta_{j} \mid \pi, \mu) d\beta_{j}$$

$$=\prod_{j=1}^{n} \int 
\frac{1}{\sqrt{2 \pi s_{j}^{2}}} exp\left \{ -\frac{1}{2s_{j}^{2}}\left ( \widehat{\beta}_{j}-\beta_{j} \right )^{2} \right \}
\sum_{k=1}^{K} \pi_{k}
\frac{1}{\sqrt{2 \pi \sigma _{k}^{2}}} exp\left \{ -\frac{1}{2\sigma_{k}^{2}}\left ( \beta_{j}-\mu \right )^{2} \right \}
d\beta_{j}$$

$$=\prod_{j=1}^{n} \sum_{k=1}^{K} \pi_{k}\int 
\frac{1}{\sqrt{2 \pi s_{j}^{2}\times 2 \pi \sigma _{k}^{2}}}
exp\left \{ -\frac{1}{2s_{j}^{2}}\left ( \widehat{\beta}_{j}-\beta_{j} \right )^{2}  -\frac{1}{2\sigma_{k}^{2}}\left ( \beta_{j}-\mu \right )^{2} \right \}
d\beta_{j}$$

$$=\prod_{j=1}^{n} \sum_{k=1}^{K} \pi_{k}
\frac{1}{\sqrt{2 \pi \left ( s_{j}^{2}+\sigma_{k}^{2} \right )}}
exp\left [ 
-\frac{\left ( \widehat{\beta}_{j}-\mu \right )^{2}}
{2\left ( s_{j}^{2}+\sigma_{k}^{2}  \right )}
\right ]$$

Where the term after $\pi_{k}$ is the density of $N(\cdot ;\mu,s_{j}^{2}+\sigma_{k}^{2})$ at $\widehat{\beta}_{j}$ , i.e.

$$L(\pi,\mu)=\prod_{j=1}^{n} \sum_{k=1}^{K} \pi_{k} N(\widehat{\beta}_{j} ;\mu,s_{j}^{2}+\sigma_{k}^{2})$$

#EM Algorithm

For convenience we denote $x_{j}\equiv \widehat{\beta}_{j}$, and introduce a latent variable $z_{j}$ specifying the mixture component that $x_{j}$ belongs to. We then have

$$p(x_{j},z_{j} \mid \mu,\pi )=\prod_{k=1}^{K} [\pi_{k} N(x_{j};\mu,s_{j}^{2}+\sigma_{k}^{2})]^{I_{z_{j}}=k}$$

$$log\,p(x_{j},z_{j} \mid \mu,\pi )=\sum_{k=1}^{K} [I_{z_{j}}=k]
 [log(\pi_{k}) - \frac{1}{2}log(2\pi)- \frac{1}{2}log(s_{j}^{2}+\sigma_{k}^{2})-\frac{\left ( x_{j}-\mu \right )^{2}}
{2\left ( s_{j}^{2}+\sigma_{k}^{2}  \right )}]$$

$$log\,p(X,Z \mid \mu,\pi )=\sum_{j=1}^{n}\sum_{k=1}^{K} [I_{z_{j}}=k]
 [log(\pi_{k}) - \frac{1}{2}log(2\pi)- \frac{1}{2}log(s_{j}^{2}+\sigma_{k}^{2})-\frac{\left ( x_{j}-\mu \right )^{2}}
{2\left ( s_{j}^{2}+\sigma_{k}^{2}  \right )}]$$

Then we will have the E-step, given the estimate of $\pi$ and $mu$, we calculate the responsibiity of j-th data point to the k-th mixture component.

$$\omega_{jk}=p(z_{j}=k \mid x_{j},\pi,\mu)=\frac{\pi_{k}N(x_{j} ;\mu,s_{j}^{2}+\sigma_{k}^{2})}{\sum_{l=1}^{K}\pi_{l}N(x_{j} ;\mu,s_{j}^{2}+\sigma_{l}^{2})}$$

For the M-step, we take $\omega_{jk}$ in the previous step, and maximize the expected loglikelihood over the latent variable $Z$.

$$Q=E_{Z\mid X,\mu, \pi}[log\,p(X,Z \mid \mu,\pi )]=\sum_{j=1}^{n}\sum_{k=1}^{K} [\omega_{jk}]
 [log(\pi_{k}) - \frac{1}{2}log(2\pi)- \frac{1}{2}log(s_{j}^{2}+\sigma_{k}^{2})-\frac{\left ( x_{j}-\mu \right )^{2}}
{2\left ( s_{j}^{2}+\sigma_{k}^{2}  \right )}]$$

We take partial derivatives of Q with respect to $\pi$ and $\mu$ and set them to be 0 to get the next estimates of our parameters, subject to the constraint that $\sum_{k=1}^{K} \pi_{k}=1$. Then we have

$$\widehat{\mu}=\frac{\sum_{j=1}^{n}\sum_{k=1}^{K}\frac{\omega_{jk}x_{j}}{s_{j}^{2}+\sigma_{k}^{2}}}
{\sum_{j=1}^{n}\sum_{k=1}^{K}\frac{\omega_{jk}}{s_{j}^{2}+\sigma_{k}^{2}}}$$

$$\widehat{\pi_{k}}=\frac{1}{n}\sum_{j=1}^{n}\omega_{jk}$$

Then we go back to the E-step and repeat with the new estimate, until we get a convergence.

### 2014-09-03 paper/main.tex
just a test to illustrate commit to Rick

@@
 \def \fp {fastPHASE\ }
+\def\bhat{\hat{\beta}}
+\def\shat{\hat{s}}
 \def\var{\rm Var}
@@
 
-The initial goal of the ASH (Adaptive SHrinkage) project is to provide simple, generic, and flexible methods to derive ``shrinkage-based" estimates and credible intervals for unknown quantities $\beta=(\beta_1,\dots,\beta_J)$, given only estimates of those quantities ($\hat\beta=(\hat\beta_1,\dots, \hat\beta_J)$) and their corresponding estimated standard errors ($s=(s_1,\dots,s_J)$).
+The initial goal of the ASH (Adaptive SHrinkage) project is to provide simple, generic, and flexible methods to derive ``shrinkage-based" estimates and credible intervals for unknown quantities $\beta=(\beta_1,\dots,\beta_J)$, given only estimates of those quantities ($\bhat=(\bhat_1,\dots, \bhat_J)$) and their corresponding estimated standard errors ($s=(s_1,\dots,s_J)$).
 
@@
 First, the appropriate amount of shrinkage is determined from the data, rather than being pre-specified.
-Second, the amount of shrinkage undergone by each $\hat\beta_j$ will depend on the standard error $s_j$: measurements with high standard error will undergo more shrinkage than measurements with low standard error.
+Second, the amount of shrinkage undergone by each $\bhat_j$ will depend on the standard error $s_j$: measurements with high standard error will undergo more shrinkage than measurements with low standard error.
 
@@
 As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
-Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\hat\beta_j/s_j$,
+Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
 and then applying a generic procedure, such as that of Benjamini 
@@
 generic methods that work directly with two numbers for each 
-measurement $(\hat\beta_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
+measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
 Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
-uncertainty in each measurement $\hat\beta_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
+uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
+
+\def\df{df}
+\def\FDR{\text{FDR}}
+\def\fdr{\text{lfdr}}
 
@@
 In some contexts our interest may focus on which of the $\beta_j$ are ``significantly" different from zero, whereas in other contexts our interest may focus on estimating their values; the methods described here are suited to both these contexts.
- We assume that we have obtained data $D_1,\dots,D_J$ that provide independent estimates $\hat\beta_1,\dots,\hat\beta_J$ of these effects,
- with corresponding (estimated) standard errors $s_1,\dots,s_J$. For our purposes, these estimates and standard errors could be obtained from the data
+ We assume that we have obtained data $D_1,\dots,D_J$ that provide independent estimates $\bhat_1,\dots,\bhat_J$ of these effects,
+ with corresponding (estimated) standard errors $\shat_1,\dots,\shat_J$. For our purposes, these estimates and standard errors could be obtained from the data
  using any method, provided they (approximately) satisfy 
- \begin{equation} \label{eqn:betahat}
- \hat\beta_j | \beta_j, s_j \sim N(\beta_j, s_j).
+ \begin{equation} \label{eqn:betahat-t}
+ \bhat_j \big{|} \, \beta_j, \shat_j \sim t_\nu(\beta_j,\shat_j),
  \end{equation}
- [In some settings a $t$ distribution assumption for $\hat\beta_j | s_j, \beta_j$ may be more appropriate, and could also be incorporated into our methods; see Discussion.]
+for some known degrees of freedom (\df) $\nu$, where $t_\nu(\mu,\sigma)$ denotes the generalized $t$ distribution on $\nu$ \df, with mean $\mu$ and scale parameter $\sigma$.
+(Equivalently, $(\bhat_j-\beta_j)/\shat_j \big{|} \, \beta_j \sim t_\nu$, where $t_\nu$ denotes the standard $t$ distribution on $\nu$ \df.)
+In the special case $\nu \rightarrow \infty$ this includes the situation where the $\bhat$ can be considered normally distributed:
+  \begin{equation} \label{eqn:betahat-norm}
+ \bhat_j  \big{|} \, \beta_j, \shat_j \sim N(\beta_j,\shat^2_j).
+ \end{equation}
+ Assuming (\ref{eqn:betahat-t}), the $t$-statistics defined by
+ \begin{equation} 
+ t_j:=\bhat_j/\shat_j
+ \end{equation}
+ have a $t_\nu$ distribution under the $j$th null hypothesis $H_j:\beta_j=0$, and this fact can be used to compute a $p$ value $p_j$ testing $H_j$.
 
-A standard genomics analysis pipeline would first convert these numbers to a $Z$ score ($Z_j=\hat\beta_j/s_j$)
-and a corresponding $p$ value ($p_j$), testing $H_j:\beta_j=0$.
+One nice feature of the False Discovery Rate (FDR) is that Frequentist and Bayesian definitions are closely aligned.
+Here we provide Bayesian definitions, and refer the reader to \cite{storey.03}, particularly his Theorem 1,  for connections with frequentist definitions.
+The (tail) FDR at threshold $T$ is defined as
+\begin{equation}
+\FDR(T) := \Pr(\text{$H_j$ false} | t_j \geq T).
+\end{equation}
+The local FDR at $t$ is defined \cite{efron} as
+\begin{equation}
+\fdr(t) := \Pr(\text{$H_j$ is false} | t_j = t).
+\end{equation}
+The local FDR is perhaps more relevant than the tail FDR, but less widely used because
+it is harder to estimate (density estimation is harder than tail area estimation). 
+The $q$ value for observation $j$ is defined \cite{storey.03} as:
+\begin{equation}
+q_j := \FDR(t_j).
+\end{equation}
+That is, the $q$ value is the FDR if we reject the null for all observations that are at least as significant as $j$.
 
-Definition of FDR and lfdr (Bayesian); connection to frequentist definitions (Storey).
-
-Then standard methods (e.g. the qvalue package) can be used
-to estimate False Discovery Rates at any given threshold.
+One standard genomics analysis pipeline is to take the $p$ values $p_1,\dots,p_J$ and convert them to (estimated) $q$ values $q_1,\dots,q_J$
+ using the R package {\tt qvalue}. A graphical description of how {\tt qvalue} works is given in Figure \ref{}.
 
@@
 
-The second issue is that directly modeling the p values, or z scores, say via non-parametric methods, without taking account of their precision, 
+The second issue is that directly modeling the p values, or z scores, say via non-parametric methods, 
 can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
@@
 
- %(See Discussion for ideas on how our methods may be applicable more generally; for example if, in place of $s_j$, we have a $p$ value testing $\beta_j=0$.)
+ %(See Discussion for ideas on how our methods may be applicable more generally; for example if, in place of $\shat_j$, we have a $p$ value testing $\beta_j=0$.)
 
@@
  \begin{equation} \label{eqn:lik}
- \hat{L}_j(\beta_j) \propto \exp[-0.5(\beta_j-\hat\beta_j)^2/s_j^2].
+ \hat{L}_j(\beta_j) \propto \exp[-0.5(\beta_j-\bhat_j)^2/\shat_j^2].
  \end{equation}
@@
  % \begin{equation} \label{eqn:lik2}
-% L(\beta_j) \propto p(\hat\beta_j | \beta_j, s_j).
+% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
  %\end{equation}
@@
 Assumption A2 can derived from (\ref{eqn:betahat}) plus a couple of other assumptions.
-Specifically, assume that a) $\hat\beta_j, s_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
-they are ``approximately sufficient" for $D_j$) and that b) $s_j$ alone contains little information about $\beta_j$.
-The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\hat\beta_j, s_j | \beta) = p(\hat\beta | \beta, s_j) p(s_j | \beta)$, while the latter (b) suggests assuming that $p(s_j|\beta_j)$ does not depend on $\beta_j$.
-Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\hat\beta_j | \beta, s_j)$, which, with (\ref{eqn:betahat}),  implies A2.
-If $\hat\beta_j, s_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\hat\beta_j,s_j$, rather than conditional on $D_j$.
+Specifically, assume that a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
+they are ``approximately sufficient" for $D_j$) and that b) $\shat_j$ alone contains little information about $\beta_j$.
+The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$.
+Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$, which, with (\ref{eqn:betahat}),  implies A2.
+If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$.
 See \cite{johnson.xx} and \cite{wakefield.xx} for more general discussion and examples of Bayesian inference based on summary statistics.
 
-Alternatively A2 could be considered definitional for $\hat\beta_j,s_j$: that is, choose $\hat\beta_j,s_j$ such that A2 holds.
+Alternatively A2 could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j$ such that A2 holds.
 
-While Assumption A2 has parallels with approaches like \cite{efron, cai?} that model the $Z$ scores $Z_j = \hat\beta_j/s_j$ as normally distributed under the null, 
-there is an important difference: here, observations with a large $s_j$ have an essentially flat likelihood (\ref{eqn:lik}), and so do not affect inference for
-other parameters, whereas when modeling the $Z$ scores directly, observations with large $s_j$ produce $Z$ scores with an $N(0,1)$ distribution that {\it do} affect inference. 
+While Assumption A2 has parallels with approaches like \cite{efron, cai?} that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
+there is an important difference: here, observations with a large $\shat_j$ have an essentially flat likelihood (\ref{eqn:lik}), and so do not affect inference for
+other parameters, whereas when modeling the $Z$ scores directly, observations with large $\shat_j$ produce $Z$ scores with an $N(0,1)$ distribution that {\it do} affect inference. 
 Because of this the likelihood-based approach can take better account of the informativeness of each measurement, as illustrated in Results below.
@@
 %The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
-For example, we can write $p(\hat\beta_j, s_j | \beta_j) = p(\hat\beta_j | s_j, \beta_j)p(s_j | \beta_j)$.
-Now, if we are willing to assume that $s_j$ alone contains no information about $\beta_j$, or equivalently that $p(s_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\hat\beta_j | s_j, \beta_j)$, and if $\hat\beta_j | s_j, \beta_j \sim N(\beta_j, s_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.
+For example, we can write $p(\bhat_j, \shat_j | \beta_j) = p(\bhat_j | \shat_j, \beta_j)p(\shat_j | \beta_j)$.
+Now, if we are willing to assume that $\shat_j$ alone contains no information about $\beta_j$, or equivalently that $p(\shat_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\bhat_j | \shat_j, \beta_j)$, and if $\bhat_j | \shat_j, \beta_j \sim N(\beta_j, \shat_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.
 
@@
 maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $s_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander}.)
+(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander}.)
 Thus allowing $K$ to be large in ii) or iii) can be thought of as approximating the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
@@
 
-Further, the assumptions  $\hat\beta_j | \beta_j, s_j \sim N(\beta_j,s_j)$ and $\beta_j | s_j, \pi \sim g(\cdot; \pi)$ (given by \ref{eqn:g}) yield: 
+Further, the assumptions  $\bhat_j | \beta_j, \shat_j \sim N(\beta_j,\shat_j)$ and $\beta_j | \shat_j, \pi \sim g(\cdot; \pi)$ (given by \ref{eqn:g}) yield: 
 \begin{equation}
-p(\hat\beta_j, s_j | \pi) = \pi_0 N(\hat\beta_j; 0,s^2_j) + \sum_k \pi_k \tilde{f}_{kj}(\hat\beta_j)
+p(\bhat_j, \shat_j | \pi) = \pi_0 N(\bhat_j; 0,s^2_j) + \sum_k \pi_k \tilde{f}_{kj}(\bhat_j)
 \end{equation}
@@
 \begin{equation}
-\tilde{f}_{kj}(x) = (\Phi(b_k/s_j) - \Phi(a_k/s_j))/(b-a),
+\tilde{f}_{kj}(x) = (\Phi(b_k/\shat_j) - \Phi(a_k/\shat_j))/(b-a),
 \end{equation}
@@
 
-Putting these together, the joint model for the unobserved $\beta$ and the observed $\hat\beta, s$ is:
+Putting these together, the joint model for the unobserved $\beta$ and the observed $\bhat, s$ is:
 \begin{align}
-p(\hat\beta, s, \beta | \pi) & = \prod_j g(\beta_j ; \pi) p(\hat\beta_j, s_j | \beta_j) \\
-& = \prod_j g(\beta_j ; \pi) L(\beta_j; \hat\beta_j,s_j).
+p(\bhat, s, \beta | \pi) & = \prod_j g(\beta_j ; \pi) p(\bhat_j, \shat_j | \beta_j) \\
+& = \prod_j g(\beta_j ; \pi) L(\beta_j; \bhat_j,\shat_j).
 \end{align}
@@
 \begin{equation}
-L(\pi; \hat\beta, s) := p(\hat\beta, s | \pi) = \int p(\hat\beta, s, \beta | \pi) d\beta.
+L(\pi; \bhat, s) := p(\bhat, s | \pi) = \int p(\bhat, s, \beta | \pi) d\beta.
 \end{equation}
@@
 \begin{equation}
-p(\beta_j | \hat\pi, \hat\beta, s) \propto g(\beta_j; \pi) L(\beta_j; \hat\beta_j, s_j).
+p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
 \end{equation} 
@@
 
-The conditional distributions $p(\beta_j | \hat\pi, \hat\beta, s)$ 
+The conditional distributions $p(\beta_j | \hat\pi, \bhat, s)$ 
 encapsulate uncertainty in the values for $\beta_j$, combining information across
@@
 
-%A Taylor series expansion of the log likelihood around the maximum likelihood estimate $\hat\beta_j$  yields $$l(\beta_j) \approx l(\hat\beta_j) + 0.5* (\beta_j - \hat\beta_j)^2 l''(\hat\beta_j).$$ [Fill in details?]
+%A Taylor series expansion of the log likelihood around the maximum likelihood estimate $\bhat_j$  yields $$l(\beta_j) \approx l(\bhat_j) + 0.5* (\beta_j - \bhat_j)^2 l''(\bhat_j).$$ [Fill in details?]
 
@@
 \begin{equation}
-\lfdr_j = p(\beta_j = 0 | \hat\pi, \hat\beta, s).
+\lfdr_j = p(\beta_j = 0 | \hat\pi, \bhat, s).
 \end{equation}
@@
 \begin{equation}
-\lfsr_j := \min[ p(\beta_j \geq 0| \hat\beta, s), p(\beta_j \leq 0| \hat\pi, \hat\beta, s) ].
+\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
 \end{equation}
@@
 To illustrate, suppose for concreteness
-that the minimum is achieved by the first term, $p(\beta_j \geq 0| \hat\beta, s)=0.05$ say. Then
+that the minimum is achieved by the first term, $p(\beta_j \geq 0| \bhat, s)=0.05$ say. Then
 our best guess would be that $\beta$ is negative, and the probability that we have
@@
 1. Estimate $\pi$ by maximizing the likelihood $L(\pi)$.
-2. Compute the conditional distributions $p(\beta_j | \hat\beta_j, s_j, \hat\pi)$.
+2. Compute the conditional distributions $p(\beta_j | \bhat_j, \shat_j, \hat\pi)$.
 
@@
 the true value of $\pi_0  \sim U[0,1]$, and then simulated $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) f_1(\cdot)$
-and $\hat\beta_j | \beta_j = N(\beta_j,1)$ for $j=1,\dots,n$. Thus these simulations assume the same
+and $\bhat_j | \beta_j = N(\beta_j,1)$ for $j=1,\dots,n$. Thus these simulations assume the same
 precision for each measurement. We applied the methods implemented in the R packages qvalue, fdrtool, locfdr, and mixfdr
@@
 \end{enumerate}
-In each case we simulated $\hat\beta_j  = \beta_j + N(0,s^2_j=1)$.
+In each case we simulated $\bhat_j  = \beta_j + N(0,s^2_j=1)$.
 

### 2014-09-10 issues/issue21/Ash Nonzeromean.rmd
Documentation updated,nonzeroMean method updated

@@
 author: "Chaoxing Dai"
-date: "2014-Sep-03"
+date: "2014-Sep-10"
 output: pdf_document
@@
 
-#Likelihood
+#Normal mixture with normal likelihood
+```
+df=NULL,mixcompdist="normal"
+```
+##Likelihood
 
@@
 
-#EM Algorithm
+##EM Algorithm
 
@@
 Then we go back to the E-step and repeat with the new estimate, until we get a convergence.
+
+#Uniform mixture with Normal likelihood
+```
+df=NULL,mixcompdist="uniform"
+```
+##Likelihood
+We know that the convolution of a uniform distribution and a normal would be of the following form  
+$f(x)\sim N(x;\mu,\sigma^{2})$  
+$g(y) \sim U(y;a,b)$  
+$(f*g)(z)=\frac{1}{b-a}[\Psi_{0}(\frac{z-a-\mu}{\sigma})-\Psi_{0}(\frac{z-b-\mu}{\sigma})]$  
+Where $\Psi_{0}(\cdot)$ is the cdf for standard normal distribution.
+
+By assuming a unimodal distribution, we have the hierarchical model being the following:  
+$$\beta _{j}\sim g(.) \sim \sum_{k=1}^{K} \pi _{k}f_{k}(.) \sim\sum_{k=1}^{K} \pi _{k} U(\cdot ;-\sigma_{k}+\mu,\sigma_{k}+\mu)$$
+
+$$\widehat{\beta}_{j}\mid \beta_{j},s_{j}^{2} \sim N(\cdot ;\beta_{j},s_{j}^{2})$$
+
+Then we would have the likelihood for $\pi$ and $\mu$ being (Here we assume iid or exchangable data)
+
+$$L(\pi,\mu)=\prod_{j=1}^{n} p(D_{j}\mid \pi,\mu)=\prod_{j=1}^{n} \int p(\widehat{\beta}_{j} \mid s_{j},\pi,\mu,\beta_{j}) p(\beta_{j} \mid \pi, \mu) d\beta_{j}$$
+
+$$=\prod_{j=1}^{n} \sum_{k=1}^{K} \pi_{k}
+\frac{1}{2\sigma_{k}}[\Psi_{0}(\frac{\widehat{\beta}_{j}+\sigma_{k}-\mu}{s_{j}})-\Psi_{0}(\frac{\widehat{\beta}_{j}-\sigma_{k}-\mu}{s_{j}})]$$
+
+##EM Algorithm
+
+For convenience we denote $x_{j}\equiv \widehat{\beta}_{j}$, and introduce a latent variable $z_{j}$ specifying the mixture component that $x_{j}$ belongs to. We then have
+
+$$p(x_{j},z_{j} \mid \mu,\pi )=\prod_{k=1}^{K} [\pi_{k} \frac{1}{2\sigma_{k}}[\Psi_{0}(\frac{\widehat{\beta}_{j}+\sigma_{k}-\mu}{s_{j}})-\Psi_{0}(\frac{\widehat{\beta}_{j}-\sigma_{k}-\mu}{s_{j}})]]^{I_{z_{j}}=k}$$
+
+$$log p(x_{j},z_{j} \mid \mu,\pi )=\sum_{k=1}^{K} [I_{z_{j}}=k] log [\pi_{k} \frac{1}{2\sigma_{k}}[\Psi_{0}(\frac{\widehat{\beta}_{j}+\sigma_{k}-\mu}{s_{j}})-\Psi_{0}(\frac{\widehat{\beta}_{j}-\sigma_{k}-\mu}{s_{j}})]]$$
+
+$$=\sum_{k=1}^{K} [I_{z_{j}}=k] [log \pi_{k}-log(2\sigma_{k}) +log [\Psi_{0}(\frac{\widehat{\beta}_{j}+\sigma_{k}-\mu}{s_{j}})-\Psi_{0}(\frac{\widehat{\beta}_{j}-\sigma_{k}-\mu}{s_{j}})]] $$
+
+Then we will have the E-step, given the estimate of $\pi$ and $mu$, we calculate the responsibiity of j-th data point to the k-th mixture component.
+
+$$\omega_{jk}=p(z_{j}=k \mid x_{j},\pi,\mu)=\frac{\frac{1}{2\sigma_{k}}[\Psi_{0}(\frac{\widehat{\beta}_{j}+\sigma_{k}-\mu}{s_{j}})-\Psi_{0}(\frac{\widehat{\beta}_{j}-\sigma_{k}-\mu}{s_{j}})]}
+{\sum_{l=1}^{K}\frac{1}{2\sigma_{l}}[\Psi_{0}(\frac{\widehat{\beta}_{j}+\sigma_{l}-\mu}{s_{j}})-\Psi_{0}(\frac{\widehat{\beta}_{j}-\sigma_{l}-\mu}{s_{j}})]}$$
+
+For the M-step, we take $\omega_{jk}$ in the previous step, and maximize the expected loglikelihood over the latent variable $Z$.
+
+$$Q=E_{Z\mid X,\mu, \pi}[log\,p(X,Z \mid \mu,\pi )]=\sum_{j=1}^{n}\sum_{k=1}^{K} [\omega_{jk}]
+ [log \pi_{k}-log(2\sigma_{k}) +log [\Psi_{0}(\frac{\widehat{\beta}_{j}+\sigma_{k}-\mu}{s_{j}})-\Psi_{0}(\frac{\widehat{\beta}_{j}-\sigma_{k}-\mu}{s_{j}})]] $$
+
+The maximization over $\mu$ no longer has a closed form, and we will use the one-dimensional R optimization function $optim()$ to perform the M step.
+
+For $\pi$, we have the same expression as before.
+
+$$\widehat{\pi_{k}}=\frac{1}{n}\sum_{j=1}^{n}\omega_{jk}$$
+
+Then we go back to the E-step and repeat with the new estimate, until we get a convergence.
+
+#Other cases
+```
+df=NULL,mixcompdist="halfuniform"
+df=scalar,mixcompdist="normal"
+df=scalar,mixcompdist="uniform"
+df=scalar,mixcompdist="halfuniform"
+```
+
+Similar to the second one, as we do not have closed form solution for $\mu$ in the M step of EM algorithm, would use optimization function with SQUAREM, and a lot of functions in EMest will be re-used.

### 2014-10-14 paper/Rcode/CurrentMethodsIllustration.rmd
updated paper

@@
 ```{r}
-#plot a histogram of z scores, highlighting the alternative distribution
-#of z scores that is implied by localfdr values lfdr.
-  nullalthist = function(z,lfdr,...){
-    h=hist(z, freq=FALSE,col=nullcol,nclass=ncz,...)
-    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
-    h$density = (1-avlfdr) * h$density
-    plot(h,add=TRUE,col=altcol,freq=FALSE)
-  }
+  source("nullalthist.R")
 

### 2014-10-14 paper/Rcode/NoisyIllustration.rmd
updated paper

@@
 p= data.frame(pval=pval,type=c(rep("GOOD",500),rep("POOR",500)))
+p2 = data.frame(pval=pval,type=rep("ALL",1000))
 require(ggplot2)
 pdf("figures/good_poor_hist.pdf",height=3,width=6)
-ggplot(p, aes(x=pval)) + facet_grid(.~type)  +
-    geom_histogram(aes(y=..density..),binwidth=.05,
-                   colour="black", fill="white")
-
-  #old version, using trellis
+ggplot(rbind(p2,p), aes(x=pval)) + facet_grid(.~type)  +
+    #geom_histogram(aes(y=..density..),binwidth=.05,colour="black", fill="white") 
+  geom_histogram(binwidth=.05,colour="black", fill="white")
+  #old version, using lattice
   #histogram(~pval|type,p,layout=c(1,2),xlim=c(0,1),breaks=seq(from=0,to=1,length=20),type="count")

### 2014-10-14 paper/Rcode/fdr.examples.rmd
updated paper

@@
 ```{r}
-opts_knit$set(progress = TRUE, verbose = TRUE,root.dir="~/Documents/git/ash/paper/Rcode")
+#opts_knit$set(progress = TRUE, verbose = TRUE,root.dir="~/Documents/git/ash/paper/Rcode")
 require(ashr)
@@
 require(ggplot2)
+source("nullalthist.R")
+source("plot_FDReg_hist.R")
 ```
@@
 
-plot_FDR_hist=function(sim,iter=1){
-  hh.pval = sim$pval[[iter]]
-  hh.zscore = sim$zscore[[iter]]
-  hh.hist =hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values",nclass=nc,col=altcol)
-
-  hh.q = qvalue(hh.pval)
-  abline(h=hh.q$pi0,col=nullcol,lwd=2)
-
-  hh.hist$density=rep(hh.q$pi0,length(hh.hist$density))  
-  plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)
-
-  abline(v=0.1,lwd=2,col=2)
-
-  text(0.05,1.2,labels="A",col=2,cex=1.2)  
-  text(0.05,0.4,labels="B",col=2,cex=1.2)  
-  text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(hh.q$pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
-}
-plot_FDR_hist(simres1,1)
+source("plot_FDReg_hist.R")
+plot_FDReg_hist(simres1a,1)
 
@@
 }
-plot_lfdr_hist(simres1,1)
+plot_lfdr_hist(simres1a,1)
 ```
@@
 ```{r}
-#plot a histogram of z scores, highlighting the alternative distribution
-#of z scores that is implied by localfdr values lfdr.
-  nullalthist = function(z,lfdr,...){
-    h=hist(z, freq=FALSE,col=nullcol,nclass=ncz,...)
-    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
-    h$density = (1-avlfdr) * h$density
-    plot(h,add=TRUE,col=altcol,freq=FALSE)
-  }
-
-#this one puts the null on the bottom
-  altnullhist = function(z,lfdr,...){
-    h=hist(z, freq=FALSE,col=altcol,nclass=ncz,...)
-    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
-    h$density = avlfdr * h$density
-    plot(h,add=TRUE,col=nullcol,freq=FALSE)
-  }
-
-  plotall_hist=function(sim,iter=1,histfun=nullalthist){
-    hh.zscore=sim$zscore[[iter]]    
-    par(mfcol=c(2,2))
-    histfun(hh.zscore,sim$betahat.fdrtool[[iter]]$lfdr,main="fdrtool")  
-    histfun(hh.zscore,sim$betahat.locfdr[[iter]]$fdr,main="locfdr")
-    histfun(hh.zscore,sim$betahat.mixfdr[[iter]]$fdr,main="mixfdr")
-    histfun(hh.zscore,sim$betahat.ash.n[[iter]]$lfdr,main="ash")
-    par(mfcol=c(1,1))
-  }
 
  # pdf("figures/nullalthist.pdf")
-  plotall_hist(simres1,1,nullalthist)
+  plotall_hist(simres1a,1,nullalthist)
 #  dev.off()
@@
  # pdf("figures/altnullhist.pdf")
-  plotall_hist(simres1,1,altnullhist)
+  plotall_hist(simres1a,1,altnullhist)
 #  dev.off()

### 2014-10-14 paper/main.tex
updated paper

@@
 \usepackage{amsmath}
+%\usepackage{natbib}
 % amssymb package, useful for mathematical symbols
@@
 \usepackage{booktabs}
+\usepackage{xspace}
 % graphicx package, useful for including eps and pdf graphics
@@
 % Use the PLoS provided bibtex style
-\bibliographystyle{plos2009}
+\bibliographystyle{/Users/stephens/Dropbox/Documents/stylefiles/plos2009}
 
@@
 \def\indep{\perp}
+\def\qvalue{{\tt qvalue}\xspace}
+\def\locfdr{{\tt locfdr}\xspace}
+\def\mixfdr{{\tt mixfdr}\xspace}
+\def\ashr{{\tt ashr}\xspace}
+
 \newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
@@
 %% END MACROS SECTION
-\externaldocument{plos_one_rev1_suppinfo}
+%\externaldocument{plos_one_rev1_suppinfo}
 
@@
 
-Since its introduction in 19xx by Benjamini and Hochberg 
-\cite{bh:xx}, the ``False Discovery Rate" (FDR) has quickly established itself
-as a key concept in modern statistics. 
-%, and the primary tool by which most practitioners handle large-scale multiple testing.
+Since its introduction in 1995 by Benjamini and Hochberg 
+\cite{benjamini1995controlling}, the ``False Discovery Rate" (FDR) has quickly established itself
+as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing.
+
 Here we consider one setting in which FDR-based methods are widely used:
@@
 has been highlighted
-previously, including Greenland and Robins (1991),
-Efron (2008) and Gelman et al (2012). [Note, check also Louis, JASA, 1984]
+previously, including \cite{greenland1991empirical},
+\cite{efron2008microarrays} and \cite{gelman2012we}. [Note, check also Louis, JASA, 1984]
 
@@
 The only package we have found that provides anything similar
-to the functionality provided here is mixfdr (Muralidharan). Compared with mixfdr, the key features of ashr are that it
+to the functionality provided here is \mixfdr \cite{muralidharan2010empirical}. Compared with \mixfdr, the key features of ashr are that it
 i) focuses on allowing for variation in the standard deviation of each observation; ii) constrains the underlying density to be unimodal (and possibly symmetric).
@@
 Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
-and then applying a generic procedure, such as that of Benjamini 
-and Hochberg (1995?) or Storey (2001?), designed to control or
-estimate the false discovery rate (FDR) or the positive FDR (Storey, 2001?). In essence we aim to provide analagous
+and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
+estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
 generic methods that work directly with two numbers for each 
@@
  have a $t_\nu$ distribution under the $j$th null hypothesis $H_j:\beta_j=0$, and this fact can be used to compute a $p$ value $p_j$ testing $H_j$.
+Some methods for FDR analysis require $z$ scores instead of $t$ statistics or $p$ values; these methods can be applied by translating each $t_j$ into a corresponding
+$z$ score, $z_j$.
 
@@
 \end{equation}
-The local FDR at $t$ is defined \cite{efron} as
+The local FDR at $t$ is defined \cite{efron2008microarrays} as
 \begin{equation}
@@
 \end{equation}
-The local FDR is perhaps more relevant than the tail FDR, but less widely used because
+The local FDR is perhaps more relevant than the tail FDR, but less widely used, perhaps because
 it is harder to estimate (density estimation is harder than tail area estimation). 
@@
 \end{equation}
-That is, the $q$ value is the FDR if we reject the null for all observations that are at least as significant as $j$.
+That is, the $q$ value for observation $j$ is the FDR if we reject the null for all observations that are at least as significant as observation $j$.
 
 One standard genomics analysis pipeline is to take the $p$ values $p_1,\dots,p_J$ and convert them to (estimated) $q$ values $q_1,\dots,q_J$
- using the R package {\tt qvalue}. A graphical description of how {\tt qvalue} works is given in Figure \ref{}.
+ using the R package \qvalue. Figure \ref{fig:qvalue} shows a graphical illustration of the procedure implemented in \qvalue. 
+ In brief, the idea is that the overall distribution of $p$ values can be decomposed into two constitutive
+ components: one component corresponding to the ``null" tests, and a second component corresponding to the ``alternative" tests. 
+To perform this decomposition, \qvalue makes two key assumptions:
+  i) the null $p$ values have a uniform distribution, and ii) the $p$ values near 1 are all from
+ tests where the null is true. Under these assumptions performing the decomposition boils down to 
+ drawing the horizontal line on the $p$ value histogram that intersects with the distribution at $p=1$: the area under this horizontal
+ line defines the uniform component that is due to the null $p$ values, with the remainder necessarily corresponding to
+ the alternative $p$ values. Now, for any given threshold, although we do not know {\it which} $p$ values correspond
+ to  null tests  and which to alternative tests, the areas of the two components indicate {\it how many} $p$ values correspond to each,
+ allowing the FDR to be estimated (Figure \ref{fig:qvalue}).
+
+ Of course \qvalue is not the only approach to estimating False Discovery Rates. Indeed, there is such a large
+ literature on FDR methods that we do not attempt a comprehensive review. Instead we highlight the two methods that
+ seem most relevant to our work here. The first is due to \cite{efron.xx}, implemented in the R package \locfdr.
+ This method differs from \qvalue in that, instead of working with $p$ values, it works with $z$ scores.
+ Analagous to \qvalue, \locfdr attempts to decompose the overall distribution of $z$ scores into two constitutive components,
+ one corresponding to the null tests, and the other corresponding to the alternative tests. 
+ In the simplest case
+ \locfdr  makes assumptions analogous to the two key assumptions of \qvalue: i) the null $z$ scores have a $N(0,1)$ distribution,
+ and ii) the $z$ scores near 0 are from the null, which Efron
+calls the ``Zero Assumption" (ZA). Under these assumptions, \locfdr can decompose the
+overall distribution of $z$ scores into a null and alternative component, and estimate the local FDR for each observation. 
+ (An important focus of Efron's work \cite{}, also implemented in \locfdr, 
+ is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
+theoretical null distribution. Although important, this issue is largely orthogonal to the issues we focus on here, which arise
+even if the theoretical null distribution holds precisely in the empirical data, and so we consider it just briefly in the Discussion.)
+
+Another method for FDR estimation, and perhaps the one most closely related to our
+work here, is due to \cite{muralidharan.xx}, implemented in the R package \mixfdr.
+Like \locfdr, \mixfdr works with $z$ scores, but it differs in that it {\it explicitly} models the alternative
+distribution of $z$ scores (using a mixture of normal distributions). In principle this approach also differs from \qvalue and \locfdr in
+that it does not explicitly make the Zero Assumption. However, the fitting procedure used means that the approach often comes close to making the ZA in practice.
+
+Figure \ref{fig:qvalue} illustrates these three methods in a simple simulated setting.
+
+Note that incorporating differential errors messes up the EM algorithm?
+
+Efron models the overall distribution of $z$ scores using a non-parametric regression approach, and then ``subtracts out" the
+null component, leaving the remainder to be the alternative component.
+
+Mention correspondance between fdr and FDR?
+
+ %The details of how this decomposition is performed vary among methods, but they share one important element in common with \qvalue: just as \qvalue assumes that the $p$ values near 1 are from the null, so \locfdr assumes that the $z$ scores near 0 are from the null; Efron calls this assumption the ``Zero Assumption" (ZA).
+
+ Before proceeding, we feel obliged to correct the common misconception
+ that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
+ not depend on the number of tests. This is because, as the number of tests increases, both the true positives and false positives increase linearly,
+ and the FDR remains the same. (If this intuitve argument does not convince, see \cite{storey.xx}, and note that the FDR does not depend on the number of tests $m$.)
+  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
+ if there are lots of strong signals then the FDR at a given threshold may be low, even if a large number of tests were performed; and conversely
+ if there are no strong signals then the FDR at the same threshold may be high, even if relatively few tests were performed. 
+ In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
+
+Possibly show two situations in figure - a high signal and low signal situation?
+
+\begin{figure}
+\includegraphics{Rcode/figures/FDReg_hist.pdf}
+\caption{Illustration of how FDR is estimated by the \qvalue \, package, by decomposing the overall distribution of $p$ values (left) into a null component and alternative component (right).
+This is done by estimating the density of $p$ values near 1, and drawing a horizontal line at that height to represent the density of the $p$ values from the null distribution (based on the assumption that all $p$ values near 1 are null).
+At a given threshold (red vertical line at 0.1), the area $A$ corresponds to the true discoveries, and the area $B$ corresponds to the true discoveries,
+so the FDR is estimated by $B/(A+B)$ which in this case is 0.28.} \label{fig:qvalue}
+\end{figure}
+
+\begin{figure}
+\center \includegraphics[height=6in]{Rcode/figures/decomp_ZA.pdf}
+\caption{Comparison of the way that different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (light blue).
+In the $z$ score space the alternative distribution is placed on the bottom to make its shape easy to see.
+The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
+In contrast the method introduced here (\ashr) assumes that the effect sizes (and thus the $z$ scores) have a unimodal distribution about 0, resulting in a very different decomposition that in this case is much closer to the truth.
+(These data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything light blue.)} \label{fig:ZA}
+\end{figure}
 
@@
 
- There are two issues with this approach that I would like to address here.
-The first is that it really does not take proper account of the measurement errors.
-To see this, consider an example where half the measurements are quite precise, and the other half are really, really, poor.
+ There are two issues with existing approaches to FDR estimation and control that I would like to address here. 
+ The first is that they do not take proper account of differences in measurement precision across units (i.e. differences in $\shat_j$). 
+ The second is that the Zero Assumption, although initially appealing as a ``conservative" assumption,
+  is often an unrealistic and unnecessarily conservative assumption. As we shall see, both factors can cause the FDR to be overestimated.
+
+ To illustrate the first issue - differing measurement precision across units - consider a simplistic
+ situation where half the measurements are quite precise, and the other half are really, really, poor. 
  Intuitively, the poor measurements tell us nothing, and any sane analysis should effectively ignore them.
@@
 This is because the $p$ values from the poor measurements will be effectively uniformly distributed, and some will be significant at any given threshold.
+
+To give a concrete example, we performed a simple simulation in which half the observations have
+good precision ($s_j = 0.5$) and half the observations have very poor precision ($s_j=10$). In both cases,
+we assume that half the effects are null ($\beta_j=0$)
+and the other half are normally distributed with standard deviation 2 ($\beta_j \sim N(0,2)$). 
+That is, we simulate $\beta_j$ from a mixture 
+\begin{equation}
+p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,2)
+\end{equation}
+where $\delta_0(\cdot)$ denotes a point mass at 0.
 
@@
 note limitation of this.
+
+To illustrate the second issue, it is helpful to go back from the $p$ values in Figure \ref{}
+
+we start by noting that the $p$ values in Figure \ref{} were actually generated under a scenario where
+{\it no tests were null}! Thus the qvalue procedure, which estimates the proportion of nulls to be xx\%, is acting very conservatively in this case.
+However, it could be argued that some degree of conservativeness is inevitable, and that nonetheless the procedure is acting sensibly.
+Therefore, to illustrate this point we really want to focus on the fact that, at least in some settings, the assumption is {\it unrealistic}.
+Specifically, we want to argue that the implied distribution of the alternative $p$ values is unrealistic. In fact, this is hard to see in $p$ value space,
+and so instead we translate these $p$ values back to the $z$ scores that lead to these $p$ values.
 
@@
 Illustration (Figure).
+
+It should be noted that existing methods were perhaps developed with the expectation that they would typically 
+be applied in settings where the vast majority of all tests are null, and under such circumstances
+the problems caused by this assumption are less severe. However, it seems desirable to have methods that
+also perform well in settings where  this assumptions does not hold, and this motivates the methods developed here.
+
+ \subsection*{Adaptive Shrinkage}
 
@@
 this could be relaxed.)  
- \item[A2:] The likelihood $L(\beta_j) := p(D_j | \beta_j)$ can be approximated by a normal likelihood,
+ \item[A2:] The likelihood $L(\beta_j) := p(D_j | \beta_j)$ can be approximated by $p(\bhat_j | \beta_j, \shat_j)$, given by \ref{eqn:betahat-t}.
+Equivalently
   \begin{equation} \label{eqn:lik}
- \hat{L}_j(\beta_j) \propto \exp[-0.5(\beta_j-\bhat_j)^2/\shat_j^2].
+ \hat{L}_j(\beta_j) \propto t_\nu((\beta_j-\bhat_j)/\shat_j)
  \end{equation}
+ where $t_\nu$ denotes the density of the standard $t$ distribution on $\nu$ degrees of freedom.
 %which, in light of (\ref{eqn:betahat}), is equivalent to
@@
 
-Assumption A2 can derived from (\ref{eqn:betahat}) plus a couple of other assumptions.
-Specifically, assume that a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
-they are ``approximately sufficient" for $D_j$) and that b) $\shat_j$ alone contains little information about $\beta_j$.
+Assumption A2 can derived from two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
+they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
 The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$.
-Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$, which, with (\ref{eqn:betahat}),  implies A2.
+Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$.
 If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$.
-See \cite{johnson.xx} and \cite{wakefield.xx} for more general discussion and examples of Bayesian inference based on summary statistics.
+See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.
 
-Alternatively A2 could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j$ such that A2 holds.
+Alternatively A2 could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that A2 holds.
 
-While Assumption A2 has parallels with approaches like \cite{efron, cai?} that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
+While Assumption A2 has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
+ that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
 there is an important difference: here, observations with a large $\shat_j$ have an essentially flat likelihood (\ref{eqn:lik}), and so do not affect inference for
@@
 That is, the non-zero effects are modeled as either i) a mixture of zero-centered normal distributions; ii) a mixture of zero-centered uniform distributions;
-or iii) a mixture of ``zero-anchored" uniform distributions, which have one end of their range at 0. These approaches closely mirror approaches in \cite{campy.thomas}.
+or iii) a mixture of ``zero-anchored" uniform distributions, which have one end of their range at 0. These approaches closely mirror approaches in \cite{cordy1997deconvolution}.
 
@@
 maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander}.)
+(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
 Thus allowing $K$ to be large in ii) or iii) can be thought of as approximating the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
@@
 
-The local False Discovery Rate (lfdr) for observation $j$ is defined \cite{efron} as
+The local False Discovery Rate (lfdr) for observation $j$ is defined \cite{efron2001empirical} as
 \begin{equation}
@@
 
-As pointed out by \cite{gelman.etal}, there are settings where
+As pointed out by \cite{gelman2012we}, there are settings where
 $\beta_j=0$ is implausible, in which case the lfdr is not useful: if every $\beta_j$ is non-zero then there
 is no such thing as a false discovery and the lfdr will always be identically 0. 
-Gelman et al suggest that in such settings we
+Gelman et.~al.~suggest that in such settings we
 might replace the concept of a false discovery with the
@@
 and $\bhat_j | \beta_j = N(\beta_j,1)$ for $j=1,\dots,n$. Thus these simulations assume the same
-precision for each measurement. We applied the methods implemented in the R packages qvalue, fdrtool, locfdr, and mixfdr
+precision for each measurement. We applied the methods implemented in the R packages qvalue, fdrtool, locfdr, and \mixfdr
 for estimating $\pi_0$. We also applied our method to estimate $\pi_0$ in two ways, one using a penalty encourage
@@
 
-Note on multiple comparisons: it isn't really a "problem" but an "opportunity". This viewpoint also espoused by Greenland and Robins. It isn't the number of tests that is relevant (the false
+Note on multiple comparisons: it isn't really a "problem" but an "opportunity". This viewpoint also espoused by \cite{greenland1991empirical}. It isn't the number of tests that is relevant (the false
 discovery rate at a given threshold does not depend on the number of tests). It is the *results* of the tests that are relevant.
@@
 
-Imagine that you are preparing to perform 1 million tests,
-each based on a $Z$ score that is assumed to be $N(0,1)$ under the null.
-You first order these tests randomly, and begin by 
-performing the first test, which returns a $Z$ score of 4.
-At this point you are interrupted by a friend, who asks how the analysis is going.
-"It's early days, but looking promising" you reply.
-Well, who wouldn't?
-If the aim is to find lots of significant differences, a strong first result is surely a good outcome.
-
-At this point you have reason to expect that many of the
-subsequent tests also output strong results.
-
-Now consider two alternative scenarios for the remaining 999,999 tests. In the first scenario, the remaining tests
-produce $Z$ values that fit very well with the null, closely following a standard normal distribution; in the second
-scenario a large proportion of the remaining tests, say 50 percent, show outcomes that lie outside of $[-4,4]$.
-
-If your friend enquired after your analysis again, your
-response would surely differ in the first scenario ("Oh, it didn't pan out so well after all") vs the second ("It went great"). Further,
-in the first scenario, if your friend pressed you further about the results of the first test,
-you would likely, I think, be inclined to put them down to chance.
-In contrast, in the second scenario, the first test turned out to be, as you hoped, a harbinger of good things to come, and in this scenario you would likely regard that test as likely corresponding to a true discovery.
-
-The key point is that it is the *outcomes* of the tests, not the *number* of tests, that impacts interpretation of that first test.
-
-(Some may by pondering whether the fact that you are about to perform another
-999,999 tests should be considered pertinent in responding to your friend. Our view is that it is  irrelevant.
-The standard frequentist framework would disagree, because
-it requires the analyst to consider hypothetical repetitions
-of the "experiment", and so the fact that the experiment
-consists of a million tests is pertinent. However, this
-argument is a distraction from the main point.)
-
-Indeed, we believe that the practice of focussing on the *number* of tests performed is
-
 Focussing on the number of tests performed can be
@@
 
-The standard argument is that,
-when performing multiple tests, some will be significant
-just by chance.
-
 \subsection{Discussion}
 
-Efron (2008) states the Zero Assumption as the assumption that ``most of the z-values near zero come from null genes".
+\cite{efron2008microarrays} states the Zero Assumption as the assumption that ``most of the z-values near zero come from null genes".
 His main aim in making this assumption is to estimate an empirical null though (not assume N(0,1) for the null)
@@
 
-Note that \cite{muralidharan:2012} models $z$ scores as something plus noise under both $H_0$ and $H_1$, which
+Note that \cite{muralidharan2010empirical} models $z$ scores as something plus noise under both $H_0$ and $H_1$, which
 avoids this problem. (DOes the same maybe apply to modeling beta, rather than z scores, when the errors vary?)
@@
 % Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgments}
+\section*{Acknowledgements}
 
 Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
-prepared using \LaTeX.
+prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the \cite{knitr} package within the RStudio GUI, and 
+git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ash} , and includes contributions 
+from Chaixing (Rick) Dai, Mengyin Lu, and Tian Sen.
+
+This work was supported by NIH grant .
 
@@
 % The bibtex filename
-\bibliography{/Users/stephens/Documents/mainbib}
+\bibliography{/Users/stephens/Dropbox/Documents/mainbib}
 

### 2014-10-14 paper/Rcode/makefigs.R

@@ -11,6 +11,9 @@ source("plot_examples.R")
 source("plot_rmse.R")
 source("plot_pi0.R")
 source("plot_lfsr.R")
+source("plot_FDReg_hist.R")
+
+
 
 pdf("figures/simABC_egdens.pdf",width=6.5,height=2)
 plot_examples(list(simA,simB,simC))

### 2014-10-14 paper/Rcode/CompareBetahatvsZscoreAnalysis.rmd
added example rmd for comparing the betahat and z score analyses

---
title: "CompareBetahatvsZscoreAnalysis"
output: html_document
---

The ashr package implements a shrinkage-based Empirical Bayes
method for estimating the values of effects $\beta_j$ based
on estimates of the effects ($\hat\beta_j$) and their standard errors ($s_j$).

A key modelling assumption is that the effects $\beta_j$ are identically distributed
from a unimodal distribution $g$. 
In particular, ashr assumes that $\beta_j$ is independent of the standard error $s_j$.
That is,
$$\beta_j | s_j \sim g(\cdot).$$

Here we consider an alternative modelling assumption:
$$\beta_j/s_j | s_j \sim g(\cdot).$$
Under this alternative assumption
the *standardized effects* $\beta_j/s_j$ are identically distributed from a unimodal
distribution. Under this assumption the expected size of the *unstandardized* effect $\beta_j$ depends on the standard error $s_j$: the larger $s_j$ is, the larger (in absolute value) $\beta_j$ is expected to be.

What might motivate this alternative assumption? 
We can provide two distinct motivations. First, suppose for concreteness that the effects
$\beta_j$ reflect differences in gene expression between two conditions. One factor that affects
$s_j$ is the variance of gene $j$ within each condition. One could imagine that genes with a larger variance within each condition are less tightly regulated, and therefore more likely to show a large difference between conditions (i.e. large $\beta_j$) than genes with a small variance. This provides
a biological motivation for the possibility that larger $s_j$ might 
correlate with larger $\beta_j$ (although of course not for the exact functional form above).
A second motivation is more statistical: it turns out that this assumption is in some sense
the implicit assumption made by existing methods to fdr analysis based on p values. 
More specifically, under this alternative assumption, when 
attempting to identify ``significant" effects, the empirical Bayes approach will rank the genes 
in the same way as the usual $p$ values computed from $\hat\beta_j/s_j$.

It is straightforward to use the ashr package to perform analysis under this alternative model:
one simply replaces betahat with the standardized betahat, $\hat\beta_j$, and the standard errors
for these standarized betahat values with 1. It is also straightforward to compare
the two competing modelling assumptions by computing the log likelihood ratio, $log[p_1(\hat\beta | s, \hat g_1)/p_2(\hat\beta | s, \hat g_2)]$, where $p_1$ denotes the default ash model,
and $p_2$ denotes the alternative model discussed here.

Let's illustrate by a simulated example. We assume that the standard errors come
from a gamma distribution, and then generate the effects $\beta_j$ under the alternative assumption so that genes with bigger standard errors tend to have bigger effects.
```{r}
  set.seed(1234)
  nsamp=1000
  betahat.se = rgamma(nsamp,1,1)
  beta = betahat.se * rnorm(nsamp) #simulate effects under the alternative assumption 
  betahat = rnorm(nsamp,beta,betahat.se)
  zscore = betahat/betahat.se
  pval = pchisq(zscore^2,df=1,lower.tail=FALSE)
  plot(betahat, -log(pval))
```

Here is the standard ash analysis.
```{r}
  library(ashr)
  ash1.res = ash(betahat, betahat.se,method="fdr")
```

And here is how we can perform the analysis under the alternative model,
by replacing betahat with the standardized betahat, and replacing the standard error with 1:
```{r}
  ash2.res = ash(betahat/betahat.se, 1,method="fdr")
```

```{r}
  loglik1=ashr:::loglik.ash(ash1.res, betahat, betahat.se, zscore=FALSE) 
  loglik2=ashr:::loglik.ash(ash2.res,betahat, betahat.se, zscore=TRUE)
  loglik1-loglik2
```

Then the log likelihood ratio is loglik1-loglik2=-93.9. This highly negative
loglikelihood indicate that the data strongly favor 
the alternative model 2, which is expected because the data were generated under this model.
(One might be tempted to ask whether the log likelihood ratio is ``significant". We don't 
know how to address this question, but suggest in practice it doesn't matter: if the loglikelihood ratio is positive then use the default analysis, if it is negative then use the alternative analysis.)

To finish, we illustrate that if we generate data under the original model then that model
is preferred by the log likelihood:
```{r}
  set.seed(1234)
  nsamp=1000
  betahat.se = rgamma(nsamp,1,1)
  beta = rnorm(nsamp) #simulate effects under the original assumption 
  betahat = rnorm(nsamp,beta,betahat.se)
  zscore = betahat/betahat.se
  pval = pchisq(zscore^2,df=1,lower.tail=FALSE)
  ash1.res = ash(betahat, betahat.se,method="fdr")
  ash2.res = ash(betahat/betahat.se, 1,method="fdr")
  loglik1=ashr:::loglik.ash(ash1.res, betahat, betahat.se, zscore=FALSE) 
  loglik2=ashr:::loglik.ash(ash2.res,betahat, betahat.se, zscore=TRUE)
  loglik1-loglik2
```

### 2014-10-15 paper/Rcode/CompareBetahatvsZscoreAnalysis.rmd
implemented computation of loglik for uniform components and zscores=TRUE

@@
 ---
-title: "CompareBetahatvsZscoreAnalysis"
+title: "Compare Betahat vs Zscore Analysis"
 output: html_document
@@
 That is,
-$$\beta_j | s_j \sim g(\cdot).$$
+$$\beta_j | s_j \sim g(\cdot)     (*).$$
 
 Here we consider an alternative modelling assumption:
-$$\beta_j/s_j | s_j \sim g(\cdot).$$
+$$\beta_j/s_j | s_j \sim g(\cdot)    (**).$$
 Under this alternative assumption
@@
 
-What might motivate this alternative assumption? 
+By analogy with Wen and Stephens (AoAS), we refer to the model (*) as the EE model ("exchangeable effects"), and the model (**)
+as the ES model ("exchangeable standardized effects")
+
+What might motivate the ES model? 
 We can provide two distinct motivations. First, suppose for concreteness that the effects
@@
 
-It is straightforward to use the ashr package to perform analysis under this alternative model:
+It is straightforward to use the ashr package to perform analysis under the ES model:
 one simply replaces betahat with the standardized betahat, $\hat\beta_j$, and the standard errors
 for these standarized betahat values with 1. It is also straightforward to compare
-the two competing modelling assumptions by computing the log likelihood ratio, $log[p_1(\hat\beta | s, \hat g_1)/p_2(\hat\beta | s, \hat g_2)]$, where $p_1$ denotes the default ash model,
-and $p_2$ denotes the alternative model discussed here.
+the two competing modelling assumptions (EE vs ES) by computing the log likelihood ratio, $log[p_{EE}(\hat\beta | s, \hat g_{EE})/p_{ES}(\hat\beta | s, \hat g_{ES})]$.
 
-Let's illustrate by a simulated example. We assume that the standard errors come
-from a gamma distribution, and then generate the effects $\beta_j$ under the alternative assumption so that genes with bigger standard errors tend to have bigger effects.
+We now illustrate by a simulated example. We assume that the standard errors come
+from a gamma distribution, and then generate the effects $\beta_j$ under the ES model so that genes with bigger standard errors tend to have bigger effects.
 ```{r}
@@
 
-Here is the standard ash analysis.
+Here is the EE analysis.
 ```{r}
@@
 
-And here is how we can perform the analysis under the alternative model,
+And here is how we can perform the ES analysis,
 by replacing betahat with the standardized betahat, and replacing the standard error with 1:
@@
 ```
+Note that the posterior distributions computed by ash will now be for the standardized
+effects, and not for the effects $\beta_j$. However, the local false sign rate (lfsr) and
+related quantities are unaffected by this because the standardized effect is positive if and only if the effect is positive.
 
+Now we compare the EE vs ES models:
 ```{r}
-  loglik1=ashr:::loglik.ash(ash1.res, betahat, betahat.se, zscore=FALSE) 
-  loglik2=ashr:::loglik.ash(ash2.res,betahat, betahat.se, zscore=TRUE)
-  loglik1-loglik2
+  loglikEE=ashr:::loglik.ash(ash1.res, betahat, betahat.se,df=NULL, zscore=FALSE) 
+  loglikES=ashr:::loglik.ash(ash2.res,betahat, betahat.se, df=NULL, zscore=TRUE)
+  loglikEE-loglikES
 ```
 
-Then the log likelihood ratio is loglik1-loglik2=-93.9. This highly negative
+Then the log likelihood ratio is loglikEE-loglikES=`r loglikEE-loglikES`. This highly negative
 loglikelihood indicate that the data strongly favor 
-the alternative model 2, which is expected because the data were generated under this model.
+the ES model, which is expected because the data were generated under this model.
 (One might be tempted to ask whether the log likelihood ratio is ``significant". We don't 
-know how to address this question, but suggest in practice it doesn't matter: if the loglikelihood ratio is positive then use the default analysis, if it is negative then use the alternative analysis.)
+know how to address this question, but suggest in practice it doesn't matter: if the loglikelihood ratio is positive then use EE, if it is negative then use ES.)
 
-To finish, we illustrate that if we generate data under the original model then that model
+To finish, we illustrate that if we generate data under the EE model then that model
 is preferred by the log likelihood:
@@
   ash2.res = ash(betahat/betahat.se, 1,method="fdr")
-  loglik1=ashr:::loglik.ash(ash1.res, betahat, betahat.se, zscore=FALSE) 
-  loglik2=ashr:::loglik.ash(ash2.res,betahat, betahat.se, zscore=TRUE)
-  loglik1-loglik2
+  loglikEE=ashr:::loglik.ash(ash1.res, betahat, betahat.se, df=NULL,zscore=FALSE) 
+  loglikES=ashr:::loglik.ash(ash2.res,betahat, betahat.se, df=NULL,zscore=TRUE)
+  loglikEE-loglikES
 ```
+So here the log likelihood ratio is positive, indicating that the EE model is preferred.

### 2014-10-15 paper/Rcode/CompareBetahatvsZscoreAnalysis.rmd
updated user interface for the way the EE and ES models are fitted

@@
 It is straightforward to use the ashr package to perform analysis under the ES model:
-one simply replaces betahat with the standardized betahat, $\hat\beta_j$, and the standard errors
-for these standarized betahat values with 1. It is also straightforward to compare
+simply specify model="ES". (Internally, this causes ash to 
+simply replaces betahat with the standardized betahat, $\hat\beta_j$, and the standard errors
+for these standarized betahat values with 1.) It is also straightforward to compare
 the two competing modelling assumptions (EE vs ES) by computing the log likelihood ratio, $log[p_{EE}(\hat\beta | s, \hat g_{EE})/p_{ES}(\hat\beta | s, \hat g_{ES})]$.
@@
   library(ashr)
-  ash1.res = ash(betahat, betahat.se,method="fdr")
+  ash1.res = ash(betahat, betahat.se,method="fdr",model="EE")
 ```
@@
 ```{r}
-  ash2.res = ash(betahat/betahat.se, 1,method="fdr")
+  ash2.res = ash(betahat, betahat.se,method="fdr", model="ES")
 ```
@@
 ```{r}
-  loglikEE=ashr:::loglik.ash(ash1.res, betahat, betahat.se,df=NULL, zscore=FALSE) 
-  loglikES=ashr:::loglik.ash(ash2.res,betahat, betahat.se, df=NULL, zscore=TRUE)
+  loglikEE=ashr:::loglik.ash(ash1.res, betahat, betahat.se,df=NULL, model="EE") 
+  loglikES=ashr:::loglik.ash(ash2.res,betahat, betahat.se, df=NULL, model="ES")
   loglikEE-loglikES
@@
 
-To finish, we illustrate that if we generate data under the EE model then that model
-is preferred by the log likelihood:
+The above illustrates these ideas on simulations from the ES model. For comparison,
+we now provide results for simulations under the EE model.
 ```{r}
@@
   pval = pchisq(zscore^2,df=1,lower.tail=FALSE)
-  ash1.res = ash(betahat, betahat.se,method="fdr")
-  ash2.res = ash(betahat/betahat.se, 1,method="fdr")
-  loglikEE=ashr:::loglik.ash(ash1.res, betahat, betahat.se, df=NULL,zscore=FALSE) 
-  loglikES=ashr:::loglik.ash(ash2.res,betahat, betahat.se, df=NULL,zscore=TRUE)
+  ash1.res = ash(betahat, betahat.se,method="fdr",model="EE")
+  ash2.res = ash(betahat, betahat.se, method="fdr",model="ES")
+  loglikEE=ashr:::loglik.ash(ash1.res, betahat, betahat.se, df=NULL,model = "EE") 
+  loglikES=ashr:::loglik.ash(ash2.res,betahat, betahat.se, df=NULL,model = "ES")
   loglikEE-loglikES
 ```
-So here the log likelihood ratio is positive, indicating that the EE model is preferred.
+So here the log likelihood ratio is positive, indicating that the EE model is preferred
+(which is expected since the data were generated under that model).
+
+Finally, note that the likelihood should always be computed under the same
+model as was used when applying ash. You'll get a warning if you do this wrong, although it will still run:
+```{r}
+  ashr:::loglik.ash(ash2.res,betahat, betahat.se, df=NULL,model = "EE")
+```

### 2014-10-16 paper/Rcode/CompareBetahatvsZscoreAnalysis.rmd
added loglik for returned values in ash so that user doesn't have to do this separately

@@
 ---
-title: "Compare Betahat vs Zscore Analysis"
+title: "Comparing Exchangeable Effects (EE) and Exchangeable Standardized Effects (ES) models"
 output: html_document
@@
 That is,
-$$\beta_j | s_j \sim g(\cdot)     (*).$$
+$$\beta_j | s_j \sim g(\cdot)     \quad (*).$$
 
 Here we consider an alternative modelling assumption:
-$$\beta_j/s_j | s_j \sim g(\cdot)    (**).$$
+$$\beta_j/s_j | s_j \sim g(\cdot)   \quad (**).$$
 Under this alternative assumption
@@
 $s_j$ is the variance of gene $j$ within each condition.
-One could imagine that genes with a larger variance within each condition are less tightly regulated, and therefore more likely to show a large difference between conditions (i.e.
+One could imagine that, perhaps, genes with a larger variance within each condition are less tightly regulated, and therefore more likely to show a large difference between conditions (i.e.
 large $\beta_j$) than genes with a small variance.
@@
 simply specify model="ES". (Internally, this causes ash to 
-simply replaces betahat with the standardized betahat, $\hat\beta_j$, and the standard errors
+replaces betahat with the standardized betahat, $\hat\beta_j/s_j$, and the standard errors
 for these standarized betahat values with 1.) It is also straightforward to compare
@@
   library(ashr)
-  ash1.res = ash(betahat, betahat.se,method="fdr",model="EE")
+  ashEE.res = ash(betahat, betahat.se,method="fdr",model="EE")
 ```
@@
 ```{r}
-  ash2.res = ash(betahat, betahat.se,method="fdr", model="ES")
+  ashES.res = ash(betahat, betahat.se,method="fdr", model="ES")
 ```
@@
 ```{r}
-  loglikEE=ashr:::loglik.ash(ash1.res, betahat, betahat.se,df=NULL, model="EE") 
-  loglikES=ashr:::loglik.ash(ash2.res,betahat, betahat.se, df=NULL, model="ES")
-  loglikEE-loglikES
+  ashEE.res$loglik-ashES.res$loglik
 ```
 
-Then the log likelihood ratio is loglikEE-loglikES=`r loglikEE-loglikES`. This highly negative
+Then the log likelihood ratio is loglikEE-loglikES=`r ashEE.res$loglik-ashES.res$loglik`. This highly negative
 loglikelihood indicate that the data strongly favor 
@@
   pval = pchisq(zscore^2,df=1,lower.tail=FALSE)
-  ash1.res = ash(betahat, betahat.se,method="fdr",model="EE")
-  ash2.res = ash(betahat, betahat.se, method="fdr",model="ES")
-  loglikEE=ashr:::loglik.ash(ash1.res, betahat, betahat.se, df=NULL,model = "EE") 
-  loglikES=ashr:::loglik.ash(ash2.res,betahat, betahat.se, df=NULL,model = "ES")
-  loglikEE-loglikES
+  ashEE.res = ash(betahat, betahat.se,method="fdr",model="EE")
+  ashES.res = ash(betahat, betahat.se, method="fdr",model="ES")
+  ashEE.res$loglik-ashES.res$loglik
 ```
@@
 (which is expected since the data were generated under that model).
-
-Finally, note that the likelihood should always be computed under the same
-model as was used when applying ash. You'll get a warning if you do this wrong, although it will still run:
-```{r}
-  ashr:::loglik.ash(ash2.res,betahat, betahat.se, df=NULL,model = "EE")
-```

### 2014-10-29 talks/UNC2014.Rmd
added UNC talk draft

--- 2014-07-13 talks/ASC2014.Rmd
@@
-% False Discovery Rates, A New Deal
-% Matthew Stephens
-% 2014/7/8
+---
+title: "False Discovery Rates, A New Deal"
+author: Matthew Stephens
+date: October 30, 2014
+output: ioslides_presentation
+---
 
@@
 require("ggplot2")
+require("knitr")
 ```
@@
 
-# Before we get started: Let's get Organized!
+## Before we get started: Let's get Organized!
 
@@
 
-# So what can you do?
+## So what can you do?
 
@@
 
-# What are these repository things?
+## What are these repository things?
 
@@
 
-# What is knitr?
+## What is knitr?
 
@@
 
-# What is Reproducible Research?
+## What is Reproducible Research?
 
@@
 
-# Why is reproducibility important?
+## Why is reproducibility important?
 
@@
 
-# More on git, github, knitr, reproducibility
+## More on git, github, knitr, reproducibility
 
@@
 
-# The Canonical Genomics Experiment
+## The Canonical Genomics Experiment
 
@@
 
-# FDR, local fdr, and q values
+## FDR, local fdr, and q values
 
@@
 ```{r, echo=FALSE}
-  #simple simulated example
-  ncz = 100 # number of bins in z score histograms
+  library("ashr") #simple simulated example
+  ncz = 100 ## number of bins in z score histograms
   set.seed(111)
@@
 
-# Example: FDR estimation
+## Example: FDR estimation
 
@@
 
-# Is this an important problem?
+## Is this an important problem?
 
@@
 
-# Problem 1: The Zero Assumption (ZA)
+## Problem 1: The Zero Assumption (ZA)
 
@@
 
-# Implied distribution of $p$ values under $H_1$
+## Implied distribution of $p$ values under $H_1$
 
@@
 
-# Implied distribution of Z scores under alternative
+## Implied distribution of Z scores under alternative
 
@@
 
-# FDR problem 2: different measurement precision
+## FDR problem 2: different measurement precision
 
@@
 
-# Example: Mouse Heart Data
+## Example: Mouse Heart Data
 
@@
 load("../scratchwork/Smemo_qbinom_zdat.RData")
-# two-sided test
+## two-sided test
   ttest.pval = function(t, df) {
@@
 
-highxp = xx>1000 # select high expressed genes
+highxp = xx>1000 ## select high expressed genes
 pval.high = pval[highxp]
@@
 zdat.ash.high = ash(zdat[3,highxp],zdat[4,highxp],df=2,method="fdr",mixcompdist="halfuniform")
+zdat.ash.ES = ash(zdat[3,],zdat[4,],df=2,method="fdr",mixcompdist="halfuniform",model="ES")
 
@@
 
-# Mouse Data: Counts vary considerably across genes
+## Mouse Data: Counts vary considerably across genes
 
@@
 
-# Lower count genes, less power
+## Lower count genes, less power
 
@@
 
-# Higher count genes, more power
+## Higher count genes, more power
 ```{r, echo=FALSE, fig.height=4,fig.cap=""}
@@
 
-# Low-count genes dilute signal at high-count genes
+## Low-count genes dilute signal at high-count genes
 ```{r, echo=FALSE, fig.height=4,fig.cap=""}
@@
 
-# FDR problem 2: low count genes add noise, increase q values
+## FDR problem 2: low count genes add noise, increase q values
 ```{r,echo=FALSE, fig.height=4,fig.cap=""}
@@
 
-# Problems: Summary
+## Problems: Summary
 
@@
 
-# FDR via Empirical Bayes
+## FDR via Empirical Bayes
 
@@
 
-# FDR: A New Deal
+## FDR: A New Deal
 
@@
 
-# FDR - A New Deal
+## FDR - A New Deal
 
@@
 
-# Illustration: $g$ a mixture of 0-centered normals
+## Illustration: $g$ a mixture of 0-centered normals
 
@@
 
-# Illustration: $g$ a mixture of 0-anchored uniforms
+## Illustration: $g$ a mixture of 0-anchored uniforms
 
@@
 
-# Adaptive Shrinkage
+## Adaptive Shrinkage
 
@@
 
-# Recall Problem 1: distribution of alternative Z values multimodal
+## Recall Problem 1: distribution of alternative Z values multimodal
 ```{r, echo=FALSE,fig.height=4,fig.cap=""}
@@
 
-# Problem Fixed: distribution of alternative Z values unimodal
+## Problem Fixed: distribution of alternative Z values unimodal
 ```{r, echo=FALSE,fig.height=4,fig.cap=""}
@@
 
-# Recall Problem 2: low count genes add noise, increase q values
+## Recall Problem 2: low count genes add noise, increase q values
 ```{r,echo=FALSE, fig.height=4,fig.cap=""}
@@
 
-# Problem Fixed: incorporating precision reduces influence of low-count genes
+## Problem Fixed: incorporating precision reduces influence of low-count genes
 ```{r,echo=FALSE, fig.height=4,fig.cap=""}
@@
 
-# A new problem: an embarrassment of riches
+## A new problem: an embarrassment of riches
 
@@
 
-# Perhaps we didn't really understand the question?
+## Perhaps we didn't really understand the question?
 
@@
 
-# The False Sign Rate
+## The False Sign Rate
 
@@
 
-# Even with many signals, large $p$ values have high lfsr
+## Even with many signals, large $p$ values have high lfsr
 
@@
 
-# Advantages of False Sign Rate
+## Advantages of False Sign Rate
 
@@
 
-# Simulated example: $\pi_0$ not identifiable.
+## Simulated example: $\pi_0$ not identifiable.
 
@@
 
-# Simulated Example: so fdr not identifiable
+## Simulated Example: so fdr not identifiable
 
@@
 
-# Simulated Example: fsr much more identifiable
+## Simulated Example: fsr much more identifiable
 
@@
 
-# Summary
+## Summary
 
@@
 
-# Other Applications
+## Other Applications
 
@@
 
-# Next steps?
+## Next steps?
 
@@
 
-# Thanks
+## Thanks
 
@@
 
-- Including Ester Pantaleo, Scott Powers, Mengyin Lu, Sen Tian, Wei Wang, Zhengrong Xing.
+- Including  Chaoxing Dai, Mengyin Lu, Ester Pantaleo, Scott Powers, Sen Tian, Wei Wang, Zhengrong Xing.
 
@@
 
-# Pandoc Command used
+## Pandoc Command used
 
@@
 ```
-
-# Some odd things in the data
-
-```{r, echo=FALSE,fig.height=3}
-  plot(zdat[3,],zdat[4,],ylab="standard error",xlab="beta-hat")
-  dd[tail(order(zdat[3,])),]
-```
-
-# A technicality
-
-- Suppose you estimate $\Pr(\beta_j<0)=0.98$,  $\Pr(\beta_j>0)=0.01$, $\Pr(\beta_j=0) = 0.01$. 
-- Should you declare an fdr of 0.01 or 0.02?
-- Maybe fsr makes more sense anyway?
-
-# Adaptive Shrinkage of point estimates
-
-- Recall idea: amount of shrinkage depends on measurement precision, $s_j$.
-
-# Adaptive Shrinkage of point estimates
-
-```{r, fig.height=4,echo=FALSE,fig.cap=""}
-  res=data.frame(betahat =zdat[3,], Estimate=zdat.ash$PosteriorMean,highxp=highxp)
-  cbbPalette <- c("#000000", "#D55E00", "#CC79A7")
-  labels = c('betahat','Estimate')
-  breaks = c("betahat","Estimate")
-
-  pp= ggplot(data=res,aes(betahat,Estimate,col=as.factor(highxp))) +geom_point(shape=1) +
-      geom_abline(colour = "black") +
-        xlab("betahat") +
-        ylab("shrunk betahat")
-
-  print(pp +scale_y_continuous(limits=c(-2,2)) +
-          scale_x_continuous(limits=c(-2,2))  +
-           scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
-          coord_equal(ratio=1))
-
-#plot(zdat[3,],zdat.ash$PosteriorMean,main="Less shrinkage for high-count genes (red)",xlab="betahat",ylab="shrunk betahat",xlim=c(-2,2),ylim=c(-2,2), col=1+highxp)
-#abline(a=0,b=1,col=2)
-```
-
-# Shrinkage is adaptive to information
-
-Need to fix counts.associate to use fdr method in ash
-
-```{r, echo=FALSE,fig.height=4}
-#hist(zdat[3,])
-#plot(zdat[3,],zdat.ash$PosteriorMean,xlim=c(-0.2,0.2))
-#points(zdat[3,16677],zdat.ash$PosteriorMean[16677],col=2,pch=16)
-#points(zdat[3,16079],zdat.ash$PosteriorMean[16079],col=2,pch=16)
-#x[16677,]
-#x[16079,]
-```
-
-```{r, echo=FALSE}
-#temp = counts.associate(cc,c(-1,1,-1,1),1)
-#temp.tscore = temp$zdat[3,]/temp$zdat[4,]
-#temp.pval = ttest.pval(temp.tscore,df=2)
-#plot(temp$zdat.ash$localfdr,temp.pval)
-#identify(temp$zdat.ash$localfdr,temp.pval)
-```
-
-```{r, echo=FALSE,fig.height=3,fig.cap=""}
-plot(zdat.ash$lfdr,pval,ylab="p value", xlab="ASH local fdr")
-```
-
-# Shrinkage is adaptive to information
-
-```{r, echo=FALSE,fig.height=3,fig.cap=""}
-plot(zdat.ash$lfdr,pval,ylab="p value", xlab="ASH local fdr")
-points(zdat.ash$lfdr[15325],pval[15325],col=2,pch=16)
-points(zdat.ash$lfdr[16123],pval[16123],col=2,pch=16)
-```
-
-# Shrinkage is adaptive to information
-
-```{r, echo=FALSE}
-dd = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
-xx = rowSums(dd[,2:5])
-dd = dd[xx>0,]
-xx = xx[xx>0]
-cbind(dd[,1:5],pval,zdat.ash$lfdr)[c(15325,16123),]
-```
-
-# Recall FDR problem 1: q values increased by low count genes
-```{r,echo=FALSE, fig.height=4,fig.cap=""}
-plot(qval.high$qval,qval$qval[xx>1000],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
-abline(a=0,b=1,col=2)
-```
-
-# ASH q values more robust to inclusion of low count genes
-
-```{r, echo=FALSE,fig.height=3,fig.cap=""}
-plot(zdat.ash.high$qvalue,zdat.ash$qvalue[xx>1000],xlab="q values from high only",ylab="q values from all data")
-abline(a=0,b=1,col=2,lwd=2)
-```
-
-Compare fitted $f(\beta)$, both estimating $\pi_0$ and fixing $\pi_0=0$.
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  x=seq(-4,4,length=100)
-  plot(cdf.ash(hh.ash.fdr,x),col=2,type="l",lwd=3,ylab="cdf")
-  lines(x,cdf.ash(hh.ash.shrink,x)$y,col=3,lwd=3,lty=2)
-```

### 2014-11-03 paper/main.tex
added figure on good and poor measurements to paper

@@
 
+Mention correspondance between fdr and FDR?
+Tidy to deal with fact that definitions I give here implicitly assume $\fdr(t)$ etc don't depend on $j$.
+Mention that Frequentist and Bayesian alignment depends on assumptions?
+
 One standard genomics analysis pipeline is to take the $p$ values $p_1,\dots,p_J$ and convert them to (estimated) $q$ values $q_1,\dots,q_J$
@@
 distribution of $z$ scores (using a mixture of normal distributions). In principle this approach also differs from \qvalue and \locfdr in
-that it does not explicitly make the Zero Assumption. However, the fitting procedure used means that the approach often comes close to making the ZA in practice.
+that it does not explicitly make the Zero Assumption. However, we have found that in practice the results often approximately satisfy the ZA,
+which we believe to be due to details in the fitting procedure used.
 
@@
 null component, leaving the remainder to be the alternative component.
-
-Mention correspondance between fdr and FDR?
 
@@
 \caption{Comparison of the way that different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (light blue).
-In the $z$ score space the alternative distribution is placed on the bottom to make its shape easy to see.
+In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
 The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
@@
 To give a concrete example, we performed a simple simulation in which half the observations have
-good precision ($s_j = 0.5$) and half the observations have very poor precision ($s_j=10$). In both cases,
+reasonably good precision ($s_j = 1$) and half the observations have very poor precision ($s_j=10$). In both cases,
 we assume that half the effects are null ($\beta_j=0$)
-and the other half are normally distributed with standard deviation 2 ($\beta_j \sim N(0,2)$). 
+and the other half are normally distributed with standard deviation 1 ($\beta_j \sim N(0,1)$). 
 That is, we simulate $\beta_j$ from a mixture 
 \begin{equation}
-p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,2)
+p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1)
 \end{equation}
@@
 
-Illustration (Figure).
+\begin{figure}
+\center \includegraphics[width=6.5in]{Rcode/figures/GOODPOOReg_hist.pdf}
+\caption{Illustration of the way that high-noise observations can contaminate signal from low-noise observations.} \label{fig:GOODPOOR}
+\end{figure}
 

### 2014-11-03 paper/Rcode/makefig_FDReg_hist.R
added code to produce good/poor figure

@@ -88,7 +88,7 @@ axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
 dev.off()
 
 ################################### GOODPOOReg_hist.pdf ###############################
-#
+
 #simple simulated example to illustrate high and low signal
 ntest = 10000
 set.seed(112)
@@ -115,5 +115,47 @@ plot_FDReg_hist(pval,1,type=1,title="Combined",yaxt='n',ylab="",nc=20,cex.axis=1
 axis(side=2, labels=FALSE,tck=-0.01)
 dev.off()
 
+################################### GOODPOOReg_scatterplot.pdf ###############################
+
+#apply the different methods
+res.qvalue = qvalue(pval)
+res.locfdr = locfdr(zscore,nulltype=0,plot=0)
+res.ash = ash(betahat,sebetahat,method="fdr")
+
+res.qvalue.good = qvalue(pval[GOOD])
+res.locfdr.good = locfdr(zscore[GOOD],nulltype=0,plot=0)
+res.ash.good = ash(betahat[GOOD],sebetahat[GOOD],method="fdr")
+
+pdf("figures/GOODPOOReg_scatter.pdf",width=6.5,height=3)
+par(mai=c(0.3,0.3,0.2,0.2),mgp = c(3, 0.5, 0))
+layout(matrix(1:3,ncol=3,byrow=TRUE))
+plot(res.qvalue.good$q,res.qvalue$q[GOOD],main="qvalue",xlim=c(0,1),ylim=c(0,1),axes=F)
+axis(side=2)
+axis(side=1)
+abline(a=0,b=1,col=2)
+plot(res.locfdr.good$fdr,res.locfdr$fdr[GOOD],main="locfdr",xlim=c(0,1),ylim=c(0,1),axes=F)
+axis(side=1)
+abline(a=0,b=1,col=2)
+plot(res.ash.good$lfsr,res.ash$lfsr[GOOD],main="ash",xlim=c(0,1),ylim=c(0,1),axes=F)
+axis(side=1)
+abline(a=0,b=1,col=2)
+dev.off()
+
+res = rbind(data.frame(x=res.ash.good$lfsr,y=res.ash$lfsr[GOOD],type="ash"), 
+            data.frame(x=res.locfdr.good$fdr,y=res.locfdr$fdr[GOOD],type='locfdr'), 
+            data.frame(x= res.qvalue.good$qvalues, y = res.qvalue$qvalues[GOOD],type="qvalue") )
+
[…]

### 2014-11-04 paper/main.tex
added code to produce good/poor figure
slight update of text and figures for good-poor illustration

@@
 \usepackage{enumerate}
+\usepackage{subfig}
 \usepackage{multirow} 
@@
 ii) make, implicitly, unrealistic assumptions about the distribution of the non-zero effects. We provide methods
-to address these problems, implemented in an R package {\tt ashr} (short for ``Adaptive Shrinkage") and demonstrate that these methods can lead to a substantial reduction in estimated FDRs (or, equivalently,
+to address these problems, implemented in an R package \ashr (short for ``Adaptive Shrinkage") and demonstrate that these methods can lead to a substantial reduction in estimated FDRs (or, equivalently,
 a substantial increase in the number of discoveries at a given FDR).
@@
 The only package we have found that provides anything similar
-to the functionality provided here is \mixfdr \cite{muralidharan2010empirical}. Compared with \mixfdr, the key features of ashr are that it
+to the functionality provided here is \mixfdr \cite{muralidharan2010empirical}. Compared with \mixfdr, the key features of \ashr are that it
 i) focuses on allowing for variation in the standard deviation of each observation; ii) constrains the underlying density to be unimodal (and possibly symmetric).
@@
 
+we start by noting that the $p$ values in Figure \ref{} were actually generated under a scenario where
+{\it no tests were null}! Thus the qvalue procedure, which estimates the proportion of nulls to be xx\%, is acting very conservatively in this case.
+However, it could be argued that some degree of conservativeness is inevitable, and that nonetheless the procedure is acting sensibly.
+Therefore, to illustrate this point we really want to focus on the fact that, at least in some settings, the assumption is {\it unrealistic}.
+Specifically, we want to argue that the implied distribution of the alternative $p$ values is unrealistic. In fact, this is hard to see in $p$ value space,
+and so instead we translate these $p$ values back to the $z$ scores that lead to these $p$ values.
+
+The second issue is that directly modeling the p values, or z scores, say via non-parametric methods, 
+can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
+distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
+The solution is to model beta hat as a convolution of some distribution g and an error component.
+
+It should be noted that existing methods were perhaps developed with the expectation that they would typically 
+be applied in settings where the vast majority of all tests are null, and under such circumstances
+the problems caused by this assumption are less severe. However, it seems desirable to have methods that
+also perform well in settings where  this assumptions does not hold, and this motivates the methods developed here.
+
  %The details of how this decomposition is performed vary among methods, but they share one important element in common with \qvalue: just as \qvalue assumes that the $p$ values near 1 are from the null, so \locfdr assumes that the $z$ scores near 0 are from the null; Efron calls this assumption the ``Zero Assumption" (ZA).
@@
  In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
-
-Possibly show two situations in figure - a high signal and low signal situation?
 
@@
  There are two issues with existing approaches to FDR estimation and control that I would like to address here. 
- The first is that they do not take proper account of differences in measurement precision across units (i.e. differences in $\shat_j$). 
+ The first is that they may do not take proper account of differences in measurement precision across units (i.e. differences in $\shat_j$). 
  The second is that the Zero Assumption, although initially appealing as a ``conservative" assumption,
@@
 To give a concrete example, we performed a simple simulation in which half the observations have
-reasonably good precision ($s_j = 1$) and half the observations have very poor precision ($s_j=10$). In both cases,
+reasonably good precision (standard error $s_j = 1$) and half the observations have very poor precision ($s_j=10$). In both cases,
 we assume that half the effects are null ($\beta_j=0$)
@@
 
+Figure \ref{fig:goodpoor-hist} shows the distribution of the resulting $p$ values for the good-precision and poor-precision measurements, as
+well as the combined distribution. As expected, the poor-precision measurements lead to almost-uniform $p$ values, and when combined
+with the good-precision measurements they dilute the overall signal in the $p$ values (e.g. they reduce the density of $p$ values near 0).
+This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$, which work with the $p$ values and $Z$ scores respectively:
+the estimated error rates ($q$-values, or lfdr values) for the good-precision observations increase when the low-precision observations are included in the analysis
+(Figure \ref{fig:goodpoor-scatter}). In contrast, the methods described here ($\ashr$) take account of the precision of each measurement, and so
+the estimated error rates for the good-precision observations are unaffected by including the low-precision observations in the analysis.
+(A key assumption of this simulation is that the distribution of the effects $\beta$ is the same for the good-precision and poor-precision measurements;
+we return to discuss this assumption in more detail in section \ref{xx})
+
 \begin{figure}
-\center \includegraphics[width=6.5in]{Rcode/figures/GOODPOOReg_hist.pdf}
-\caption{Illustration of the way that high-noise observations can contaminate signal from low-noise observations.} \label{fig:GOODPOOR}
+\centering
+\subfloat[][Histograms of $p$ values for good-precision, poor-precision, and combined observations]{
+\includegraphics[width=6.5in]{Rcode/figures/GOODPOOReg_hist.pdf} 
+    \label{fig:goodpoor-hist}
+} \\
+\subfloat[][Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
+Each point shows the estimated error rate ($q$ value, lfdr, or lfsr) for a good-precision observation under the two different analyses.]{
+\includegraphics[width=6.5in]{Rcode/figures/GOODPOOReg_scatter.pdf}
+    \label{fig:goodpoor-scatter}
+}
+\caption{Simulated illustration how poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
+$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
+a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
+The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the lfdr estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
+In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
 \end{figure}
-
-Note that the problem could be reduced by discarding measurements that are too noisy. However,
-note limitation of this.
-
-To illustrate the second issue, it is helpful to go back from the $p$ values in Figure \ref{}
-
-we start by noting that the $p$ values in Figure \ref{} were actually generated under a scenario where
-{\it no tests were null}! Thus the qvalue procedure, which estimates the proportion of nulls to be xx\%, is acting very conservatively in this case.
-However, it could be argued that some degree of conservativeness is inevitable, and that nonetheless the procedure is acting sensibly.
-Therefore, to illustrate this point we really want to focus on the fact that, at least in some settings, the assumption is {\it unrealistic}.
-Specifically, we want to argue that the implied distribution of the alternative $p$ values is unrealistic. In fact, this is hard to see in $p$ value space,
-and so instead we translate these $p$ values back to the $z$ scores that lead to these $p$ values.
-
-The second issue is that directly modeling the p values, or z scores, say via non-parametric methods, 
-can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-Illustration (Figure).
-
-It should be noted that existing methods were perhaps developed with the expectation that they would typically 
-be applied in settings where the vast majority of all tests are null, and under such circumstances
-the problems caused by this assumption are less severe. However, it seems desirable to have methods that
-also perform well in settings where  this assumptions does not hold, and this motivates the methods developed here.
 

### 2014-11-04 paper/Rcode/makefig_FDReg_hist.R

@@ -108,8 +108,8 @@ pval = pchisq(zscore^2,df=1,lower.tail=F)
 pdf("figures/GOODPOOReg_hist.pdf",width=6.5,height=3)
 par(mai=c(0.3,0.3,0.2,0.2),mgp = c(3, 0.5, 0))
 layout(matrix(1:3,ncol=3,byrow=TRUE))
-plot_FDReg_hist(pval[GOOD],1,type=1,title="Low-noise observations",ylab="",nc=20,cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
-plot_FDReg_hist(pval[-GOOD],1,type=1,title="High-noise observations",ylab="",nc=20,yaxt='n',cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
+plot_FDReg_hist(pval[GOOD],1,type=1,title="Good-precision observations",ylab="",nc=20,cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
+plot_FDReg_hist(pval[-GOOD],1,type=1,title="Poor-precision observations",ylab="",nc=20,yaxt='n',cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
 axis(side=2, labels=FALSE,tck=-0.01)
 plot_FDReg_hist(pval,1,type=1,title="Combined",yaxt='n',ylab="",nc=20,cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
 axis(side=2, labels=FALSE,tck=-0.01)
@@ -136,21 +136,21 @@ abline(a=0,b=1,col=2)
 plot(res.locfdr.good$fdr,res.locfdr$fdr[GOOD],main="locfdr",xlim=c(0,1),ylim=c(0,1),axes=F)
 axis(side=1)
 abline(a=0,b=1,col=2)
-plot(res.ash.good$lfsr,res.ash$lfsr[GOOD],main="ash",xlim=c(0,1),ylim=c(0,1),axes=F)
+plot(res.ash.good$lfsr,res.ash$lfsr[GOOD],main="ashr",xlim=c(0,1),ylim=c(0,1),axes=F)
 axis(side=1)
 abline(a=0,b=1,col=2)
 dev.off()
 
-res = rbind(data.frame(x=res.ash.good$lfsr,y=res.ash$lfsr[GOOD],type="ash"), 
+res = rbind(data.frame(x= res.qvalue.good$qvalues, y = res.qvalue$qvalues[GOOD],type="qvalue"),
[…]

### 2014-11-08 Rcode/efron.fcr.rmd
added rmd file looking at FCR in Efron's example

---
title: "FSR in non-unimodal case"
output: html_document
---

This example comes from Efron (2008) p16 when examining the false coverage rate (FCR).

```{r}
  set.seed(100)
  nsamp=10000
  mu0 = rep(0,nsamp)
  mu1 = rnorm(nsamp,-3,1)
  comp = rbinom(nsamp,1,0.1)
  mu = ifelse(comp==0,mu0,mu1)
  z = rnorm(nsamp,mu,1)
```

```{r}
  library(ashr)
  res.ash = ash(z,1,mixcompdist="halfuniform",method="fdr")
  CImatrix= ashci(res.ash,level=0.95)
```

Compare results with Bayes rule
```{r}
  BayesComparePlot=function(){
    plot(z,mu,xlim=c(-8,0))
    points(z[CImatrix[,1]],CImatrix[,5],col="red")
    points(z[CImatrix[,1]],CImatrix[,4],col="red")

    fdr = 0.9*dnorm(z)/(0.9*dnorm(z)+0.1*dnorm(z,-3,sqrt(2)))
    o=order(z)
    upper = ifelse(fdr[o]<0.025,(z[o]-3)/2+qnorm(0.975+fdr[o])/sqrt(2),0)
    lines(z[o],upper,col="green",lwd=2)
    lines(z[o],(z[o]-3)/2-qnorm(0.975)/sqrt(2),col="green",lwd=2)
    abline(v=max(z[fdr<0.05]))
  }
  BayesComparePlot()
```

It seems the lower 97.5% point is pretty accurate.

Try whole thing again to see if it was a fluke
```{r}
  set.seed(101)
  nsamp=10000
  mu0 = rep(0,nsamp)
  mu1 = rnorm(nsamp,-3,1)
  comp = rbinom(nsamp,1,0.1)
  mu = ifelse(comp==0,mu0,mu1)
  z = rnorm(nsamp,mu,1)
```

```{r}
   res.ash = ash(z,1,mixcompdist="halfuniform",method="fdr")
  CImatrix= ashci(res.ash,level=0.95)
  BayesComparePlot()
```

And try with normal mixture in ash:

```{r}
   res.ash = ash(z,1,mixcompdist="normal",method="fdr")
  CImatrix= ashci(res.ash,level=0.95)
  BayesComparePlot()
```

And uniform mixture:
```{r}
   res.ash = ash(z,1,mixcompdist="uniform",method="fdr")
  CImatrix= ashci(res.ash,level=0.95)
  BayesComparePlot()
```

Things to try:
- simulation study for coverage?
- compare with EB FCR methods in JRSS B?

### 2014-11-08 paper/main.tex
added rmd file looking at FCR in Efron's example

@@
 measurement precision (standard error), 
-and the goal is to identify which of the effects we can declare to be non-zero (with some prescribed level of confidence).
+and the goal is to identify which of the effects we can declare to be non-zero (with some prescribed level of confidence), and also
+to provide estimates -- both point estimates and interval estimates -- for these effects.
 This setting arises commonly in many areas of statistics, and particularly in genomics applications.
@@
 a common goal is to identify ``differentially-expressed genes" -- that is, genes
-whose mean expression (activity) level differs in two conditions.
+whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
+expression level between the two conditions, often referred to as the ``log-fold change". 
+%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.
 
@@
 
-Many tools have been developed for this problem \cite{}, and several are widely used.
-Here we argue that these existing tools have sub-optimal features that can be improved on, at least in some circumstances.
-In particular we argue that existing tools i) fail to take proper account of variations in the measurement precision of different effects;
-ii) make, implicitly, unrealistic assumptions about the distribution of the non-zero effects. We provide methods
-to address these problems, implemented in an R package \ashr (short for ``Adaptive Shrinkage") and demonstrate that these methods can lead to a substantial reduction in estimated FDRs (or, equivalently,
+Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and several are widely used.
+Here we present new tools that bring two important new ideas to bear on the problem.
+First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
+For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
+may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
+imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
+Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
+is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
+Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
+demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
 a substantial increase in the number of discoveries at a given FDR).
 
-Although we motivate our tools by FDR, we are providing generic tools for shrinkage, and
-these could be useful in settings such as wavelets.
+Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests, 
+provides a general tool for shrinkage estimation. Shrinkage estimation is a powerful tool, with many potential applications, including for example wavelet denoising, 
+where shrinkage of wavelet coefficients is used to smooth signals. 
+However, in contrast to FDR methods, shrinkage methods are not so widely used in practice, at least in the kinds
+of genomics settings I am familiar with. One possible reason for this is that most work on shrinkage-based estimation has focussed
+on point estimates, rather than on interval estimates which are often of more direct practical relevance. Another contributing
+factor may be the paucity of general-purpose software packages for this kind of shrinkage estimation.
+ (There are, perhaps, more packages for shrinkage esitmation in the more complex setting of covariance estimation, where shrinkage is perhaps still more important.)
+Our work here addresses both these issues, providing a general-purpose software tool that uses shrinkage ideas to improve both
+point and interval estimates.
+
+By ``adaptive" shrinkage we 
+have two properties in mind. First, the appropriate amount of shrinkage is determined from the data, rather than being pre-specified.
+Second, the amount of shrinkage undergone by each measurement will depend on its precision, or standard error: measurements with high standard error will undergo more shrinkage than measurements with low standard error.
+Although shrinkage-based estimation can be motivated in various ways, our key goal here is to combine information across the multiple measurements to improve inference for each individual effect.
+By improved inference, we mean both
+improved average accuracy of point estimates, which is
+the traditional focus of shrinkage-based methods, \emph{and} improved assessments of uncertainty, such as credible intervals.
+
 The potential for shrinkage-based estimation to
-address the multiple comparisons problems tackled by the FDR
+address the multiple comparisons problems tackled by FDR-based methods
 has been highlighted
@@
 \cite{efron2008microarrays} and \cite{gelman2012we}. [Note, check also Louis, JASA, 1984] 
+Furthermore,
 
@@
 
-Although shrinkage-based estimation can be motivated in various ways, our key goal here is to combine information across the multiple measurements $j=1,\dots,J$ to improve inference for each individual $\beta_j$.
-By improved inference, we mean both
-improved average accuracy of point estimates, which is
-the traditional focus of shrinkage-based methods, \emph{and} improved assessments of uncertainty.
+Given that shrinkage estimation is widely recognized as a powerful tool, there are
 
-By ``adaptive" shrinkage we 
-have two key properties in mind.
-First, the appropriate amount of shrinkage is determined from the data, rather than being pre-specified.
-Second, the amount of shrinkage undergone by each $\bhat_j$ will depend on the standard error $s_j$: measurements with high standard error will undergo more shrinkage than measurements with low standard error.
-
-Given that shrinkage estimation is widely recognized as a powerful tool, there are surprisingly few software packages for performing the simplest type of shrinkage estimation considered here.
-(There are more packages for the more complex setting of covariance estimation, where shrinkage is perhaps still more important.)
 The only package we have found that provides anything similar
@@
 i) focuses on allowing for variation in the standard deviation of each observation; ii) constrains the underlying density to be unimodal (and possibly symmetric).
-NOTE: should emphasize these differences in the examples.
 
@@
 
+Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
+
  \subsection*{Adaptive Shrinkage}
@@
 
- In outline, we use a hierarchical model to combine information across measurements, with the aim of improving both accuracy and precision of estimates. 
+ In outline, we use an ``Empirical Bayes" hierarchical model to combine information across measurements and perform inference.
+ Here performing inference means obtaining an (approximate) posterior distribution for each effect $\beta_j$,
+which can be used to estimate False Discovery Rates, and to produce both point and interval estimates for $\beta_j$.
+ Although ``multiple comparisons" is usually presented as a ``problem", our emphasis here is that the multiple measurements
+ actually provide an opportunity: the hierarchical model combines information across measurements to improve both accuracy and precision of estimates.
+
  Specifically we assume that the effects $\beta_j$ are independent and identically distributed from some unknown distribution $g(\cdot)$,
@@
  where $g$ will be estimated from
- the data.  Our methods make the following key assumptions, which distinguish them from most previous work in this area:
+ the data.   
+  Our methods further make the following key assumptions, which distinguish them from most previous work in this area:
    \begin{enumerate}
@@
 this could be relaxed.)  
- \item[A2:] The likelihood $L(\beta_j) := p(D_j | \beta_j)$ can be approximated by $p(\bhat_j | \beta_j, \shat_j)$, given by \ref{eqn:betahat-t}.
+ \item[A2:] The likelihood $L(\beta) := p(D | \beta)$ can be approximated by $\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
+ $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by (\ref{eqn:betahat-t}).
 Equivalently
   \begin{equation} \label{eqn:lik}
- \hat{L}_j(\beta_j) \propto t_\nu((\beta_j-\bhat_j)/\shat_j)
+ \hat{L}(\beta) \propto \prod_j t_\nu((\beta_j-\bhat_j)/\shat_j)
  \end{equation}
@@
 Interestingly, although the unimodal assumption A1 seems natural, much previous related work on estimating False Discovery Rates  
-has assumed, either explicitly or implicitly, that the effect sizes are multimodal (see Results for more details). 
-In contrast, almost all analogous work in large-scale regression assumes
-a unimodal prior distribution for the effect sizes, common choices being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors.
-One of our aims here is to emphasize potential benefits of also making the unimodal assumption in the FDR context.
+has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}. 
+In contrast, almost all analogous work in large-scale regression assumes the effects (regression coefficents) to have a unimodal prior distribution - 
+common choices being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors.
+One of our aims here is to emphasize the considerable practical benefits of making the unimodal assumption in the FDR context.
 
-Assumption A2 can derived from two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
+Assumption A2 can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
 they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
@@
 there is an important difference: here, observations with a large $\shat_j$ have an essentially flat likelihood (\ref{eqn:lik}), and so do not affect inference for
-other parameters, whereas when modeling the $Z$ scores directly, observations with large $\shat_j$ produce $Z$ scores with an $N(0,1)$ distribution that {\it do} affect inference. 
-Because of this the likelihood-based approach can take better account of the informativeness of each measurement, as illustrated in Results below.
+$g$ or the other effects, whereas when modeling the $Z$ scores directly, observations with large $\shat_j$ produce $Z$ scores with an $N(0,1)$ distribution that {\it do} affect inference.
+Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
 
@@
 our best guess would be that $\beta$ is negative, and the probability that we have
-made an error in sign would be 0.05. We note that the idea of focussing on tail areas, rather than point null hypotheses,
-has a long history (e.g. Altham? others?).
+made an error in sign would be 0.05.
+
+We note that the idea of focussing on tail areas, rather than point null hypotheses,
+has a long history (e.g. Altham? others?). For example, 
+Tukey, p32, "the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?""
+More recently Gelman.
 
@@
 
+Comparisons with Efron, whose 1-group model is effectively the approach we have taken here.
+He says ``deconvolution
+calculations are notoriously tricky and I have not
+been able to produce a stable general algorithm." Here our claim is that the unimodal constraint is enough to produce that
+general stable algorithm.
+
 \subsection{The number of components is not critical}

### 2014-11-09 Rcode/ash.withoneobs.rmd
add example with one observation

---
title: "ash with one observation"
output: html_document
---

At http://andrewgelman.com/2013/11/21/hidden-dangers-noninformative-priors/
Andrew Gelman considers some of the danters of so-called ``non-informative" priors.
Here's a quote:

``Finally, the simplest example yet, and my new favorite: we assign a flat noninformative prior to a continuous parameter theta. We now observe data, y ~ N(theta,1), and the observation is y=1. This is of course completely consistent with being pure noise, but the posterior probability is 84% that theta>0. I don’t believe that 84%. I think (in general) that it is too high."

I agree with him: although there is no single ``right" answer to the posterior probability
that theta>0 in this situation, 84% seems too high in general.

Here we examine what ash would do, under various settings, with this observation.

```{r}
library(ashr)
res.ash=list()
res.ash$ns = ash(1,1,mixcompdist="normal",method="shrink")$lfsr[1]
res.ash$nf = ash(1,1,mixcompdist="normal",method="fdr")$lfsr[1]
res.ash$us = ash(1,1,mixcompdist="uniform",method="shrink")$lfsr[1]
res.ash$uf = ash(1,1,mixcompdist="uniform",method="fdr")$lfsr[1]
res.ash$hus = ash(1,1,mixcompdist="halfuniform",method="shrink")$lfsr[1]
res.ash$huf = ash(1,1,mixcompdist="halfuniform",method="fdr")$lfsr[1]
res.ash
```

So what is going on here? With the method="fdr" ash uses a penalized likelihood to
encourages the estimate of g() to be ``null" if that is (somewhat) consistent with the data. 
In this case, we have a single observation, z=1, which is consistent with the null. Ash then
estimates g to be a point mass on 0, and the lfsr is 1.

For method=``shrink" there is no penalty on the likelihood, and the symmetric models (uniform and normal) end up producing false sign rates of about 50%. The asymmetric model (halfuniform) decides,
on the basis of this single observation, that everything is positive, and so lfsr is 0. This is clearly wrong!

I don't think we really learn a whole lot from this example, but I do think the results for
the symmetric models with method="shrink" are at least somewhat sensible.

### 2014-11-10 Rcode/assess.null.rmd
added assess.null.rmd example

---
title: "Assess Null Behavior"
output: html_document
---

```{r}
  library(ashr)
  library(qvalue)
  set.seed(112)
  res.qvalue=list()
  res.ash.hu=list()
  res.ash.u=list()
  res.ash.n=list()
  res.ash.s.hu=list()
  res.ash.s.u=list()
  res.ash.s.n=list()
  nsamp = 1000
  nset=100
  for(i in 1:nset){
    z=rnorm(nsamp)
    pval = pchisq(z^2,df=1,lower.tail=F)
    res.qvalue[[i]]=qvalue(pval)
    res.ash.hu[[i]]=ash(z,1,method="fdr",mixcompdist="halfuniform")
    res.ash.u[[i]]=ash(z,1,method="fdr",mixcompdist="uniform")
    res.ash.n[[i]]=ash(z,1,method="fdr",mixcompdist="normal")

    res.ash.s.hu[[i]]=ash(z,1,method="shrink",mixcompdist="halfuniform")
    res.ash.s.u[[i]]=ash(z,1,method="shrink",mixcompdist="uniform")
    res.ash.s.n[[i]]=ash(z,1,method="shrink",mixcompdist="normal")

  }

```

```{r}
  getpi0.qvalue=function(q){return(q$pi0)}
  pi0.qvalue=lapply(res.qvalue,getpi0.qvalue)
  pi0.ash.hu =lapply(res.ash.hu,get_pi0)
  pi0.ash.u =lapply(res.ash.u,get_pi0)
  pi0.ash.n =lapply(res.ash.n,get_pi0)
  par(mfcol=c(4,1))
  hist(as.numeric(pi0.qvalue),xlim=c(0.8,1.0),breaks=seq(0.8,1,length=21))
  hist(as.numeric(pi0.ash.hu),xlim=c(0.8,1.0),breaks=seq(0.8,1,length=21))
  hist(as.numeric(pi0.ash.n),xlim=c(0.8,1.0),breaks=seq(0.8,1,length=21))
  hist(as.numeric(pi0.ash.u),xlim=c(0.8,1.0),breaks=seq(0.8,1,length=21))

```

```{r}
  getnd.qvalue=function(q){return(sum(q$qvalue<0.05))}
  getnd.ash = function(a){return(sum(a$lfsr<0.05))} #number of discoveries at 5% lfsr

#note: should perhaps in principle switch to tail FSR rather than lfsr for this assessment?
  nd.qvalue = lapply(res.qvalue,getnd.qvalue)
  nd.ash.hu=lapply(res.ash.hu,getnd.ash)
  nd.ash.u=lapply(res.ash.u,getnd.ash)  
  nd.ash.n=lapply(res.ash.n,getnd.ash)
  nd.ash.s.hu=lapply(res.ash.s.hu,getnd.ash)
  nd.ash.s.u=lapply(res.ash.s.u,getnd.ash)  
  nd.ash.s.n=lapply(res.ash.s.n,getnd.ash)

  summary(as.numeric(nd.qvalue))
  summary(as.numeric(nd.ash.hu))
  summary(as.numeric(nd.ash.u))
  summary(as.numeric(nd.ash.n))
   summary(as.numeric(nd.ash.s.hu))
  summary(as.numeric(nd.ash.s.u))
  summary(as.numeric(nd.ash.s.n))
```

This shows that, except for the combination "shrink,halfuniform" there are very few observations with lfsr>0.05. For the shrink,halfuniform combination what is going on? Remember that in the shrink
method there is no point mass at 0. So all the components are positive or negative. Turns out that in this case it is quite common for the method to converge to a situation where nearly all the weight is on one (nearly-null) component. When that happens the lfsr is estimated near 0 for all the observations... so this combination should probably be avoided.

### 2014-12-17 talks/ASC2014.newtext.Rmd
added some talks

--- 2014-07-13 talks/ASC2014.Rmd
@@
+---
+title: "Test-Presentation"
+author: "Matthew Stephens"
+date: "September 23, 2014"
+output: ioslides_presentation
+---
+
 % False Discovery Rates, A New Deal
@@
   #set global chunk options
-  opts_chunk$set(cache=TRUE,autodep=TRUE,warning=FALSE)
-  dep_auto()
+  #opts_chunk$set(cache=TRUE,autodep=TRUE,warning=FALSE)
+  #dep_auto()
 ```
@@
 ```
+
+## R Markdown
+
+This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+
+When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
+
+## Slide with Bullets
+
+- Bullet 1
+- Bullet 2
+- Bullet 3
+
+## Slide with R Code and Output
+
+```{r}
+summary(cars)
+```
+
+## Slide with Plot
+
+```{r, echo=FALSE}
+plot(cars)
+```

### 2014-12-17 talks/UNC2014.dev.Rmd
added some talks

--- 2014-10-29 talks/UNC2014.Rmd
@@
 date: October 30, 2014
+runtime: shiny
 output: ioslides_presentation
@@
 
+```{r, include=FALSE}
+ #plot a histogram of z scores, highlighting the alternative distribution
+#of z scores that is implied by pi0
+nullalthist_z = function(z,pi0,nullcol="blue",altcol="cyan",...){
+  h=hist(z, freq=FALSE,col=nullcol,nclass=40,ylim=c(-0.5,0.5),...)
+  nb = length(h$breaks)
+  nullexp = pi0 * (pnorm(h$breaks)[-1] - pnorm(h$breaks[-nb]))/(h$breaks[-1]-h$breaks[-nb])
+  h$density = h$density - nullexp
+  plot(h,add=TRUE,col=altcol,freq=FALSE)
+}
+
+ set.seed(100)
+z = c(rnorm(5000,0,2),rnorm(5000,0,1))
+```
+
+## Interactive Illustration (shiny package)
+```{r, echo=FALSE, cache=FALSE}
+ library("shiny")
+inputPanel(  
+  sliderInput("pi0", label = "Proportion of nulls:",
+              min = 0, max = 1, value = 0.5, step = 0.1)
+)
+
+renderPlot({
+  nullalthist_z(z, input$pi0)
+})
+```
+
 ## Recall Problem 1: distribution of alternative Z values multimodal
@@
 
+## Coda: Modelling effects vs Standardized effects
+
+- Recall we assumed $\beta_j | s_j \sim g(\cdot)$ independent of $s_j$.
+
+- But it is plausible that distribution of $\beta_j$ depends on $s_j$.
+
+- Might motivate assuming $z_j = \beta_j/s_j | s_j \sim g_1(\cdot)$.
+
+- Indeed, this is the implicit assumption underlying standard procedures that rank by the $z$ scores.
+
+- This model is equally easy to fit with same software! (And can compute log-likelihood of the different models.)
+
+## Illustration: Modelling effects vs Standardized effects
+
+```{r, echo=FALSE,fig.height=4,fig.cap="Model Effects",warning=FALSE,message=FALSE}
+  plot(zdat[3,],abs(zdat[3,]/zdat[4,]),col=1+(zdat.ash$lfsr<0.05),xlim=c(-2,2),xlab="beta-hat",ylab="zscore")
+```
+
+## Illustration: Modelling effects vs Standardized effects
+
+```{r, echo=FALSE,fig.height=4,fig.cap="Model Standardized Effects",warning=FALSE,message=FALSE}
+  plot(zdat[3,],abs(zdat[3,]/zdat[4,]),col=1+(zdat.ash.ES$lfsr<0.05),xlim=c(-2,2),xlab="beta-hat",ylab="zscore")
+```
+
+Loglikelihood difference = `r zdat.ash$loglik-zdat.ash.ES$loglik` in favor of modelling effects.
+
 ## Other Applications
@@
 
-- to the developers of **R**, **knitr**, **Rstudio** and **Pandoc**.
+- to the developers of **R**, **knitr**, **shiny**, **Rstudio** and **Pandoc**.
 
@@
 
-- `ashr` package: `http://www.github.com/stephens999/ash`
+- `ashr` R package: `http://www.github.com/stephens999/ash`
 
-## Pandoc Command used
-
-`pandoc -s -S -i --template=my.beamer -t beamer -V theme:CambridgeUS -V colortheme:beaver  slides.md -o slides.pdf`
-
-(alternative to produce html slides; but figures would need reworking)
-`pandoc -s -S -i -t dzslides --mathjax slides.md -o slides.html`
-
-Here is my session info:
+## Session Info
 

### 2014-12-17 talks/gtex.Rmd
added some talks

--- 2014-10-29 talks/UNC2014.Rmd
@@
 title: "False Discovery Rates, A New Deal"
-author: Matthew Stephens
-date: October 30, 2014
+author: "Matthew Stephens"
+date: "November 24, 2014"
 output: ioslides_presentation
@@
 
-## Before we get started: Let's get Organized!
-
-- Over ~10 years of working with graduate students + postdocs,
-I've noticed something.
-- Organized researchers get more done (and better!).
-- Many of them are more organized than I am!
-- Thought: I should get organized; I should help others get organized.
-
-## So what can you do?
-
-- Buy a notebook; bring it to meetings; make notes! 
-- Come to meetings with a written agenda. 
-- While doing research, record what you did and what the outcome was.
-- Use version control ([git](http://git-scm.com)) and internet repositories ([bitbucket](http://www.bitbucket.org), [github](http://www.github.com)) to organize notes, code, etc.
-- Use *knitr* to help make your research reproducible.
-- Talk about the tools you find useful!
-
-## What are these repository things?
-
-- A repository: a central place in which an aggregation of data is kept and maintained in an organized way (searcharticle.com)
-- Great for sharing material across multiple people (eg student and advisor!)
-- An amateur example: [http://github.com/stephens999/ash](http://github.com/stephens999/ash)
-
-## What is knitr?
-
-- An R package
-- A tool for literate programming
-- Text, and R code are interleaved
-- When you compile the document, the code is run, and output inserted into the text.
-- Great for writing reports, and keeping a track of what you did and what the result was!
-- This talk was written with knitr (with RStudio)!
-
-## What is Reproducible Research?
-
-- Principle: when publishing results of computational procedures, we should
-publish the code that produced the results.
-- "publishing figures or results without the complete software environment could
-be compared to a mathematician publishing an announcement of a mathematical theorem without giving the proof" (Buckheit and Donohoe)
-- “an article about a computational result is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result.” [Claerbout]
-
-## Why is reproducibility important?
-
-- Not only because people are forgetful, error-prone, or dishonest!
-- Reproducing work is also the first step to extending it.
-- Helps communications among researchers (eg student + advisor).
-- If you do not publish code implementing your methods, your methods will likely go unused.
-
-## More on git, github, knitr, reproducibility
-
-- Google "The git book", to get started on git.
-- Google "Karl Broman github tutorial" for statistics-oriented intro to github.
-- Google "donohoe buckheit" for "Wavelab and reproducible research"
-
 ## The Canonical Genomics Experiment
@@
 - ...?
-
-## FDR, local fdr, and q values
-
-Although precise definitions vary depending on whether one
-takes a Bayesian or Frequentist approach to the problem, roughly
-
-- The FDR at a threshold $P$ is 
-$$\text{FDR}(P)=\Pr(\beta_j = 0 |  p_j<P).$$
-
-- The q value for observation $j$ is $q_j=\text{FDR}(p_j)$.
-
-- The local false discovery rate, fdr, at threshold $P$ is 
-$$\text{fdr}(P) = \Pr(\beta_j =0 | p_j=P ).$$
-
-- The fdr is more relevant, but slightly harder to estimate than 
-FDR because it involves density estimation rather than tail-area estimation.
 
@@
 ```
-
-## Is this an important problem?
-
-- The original paper introducing FDR (Benjamini and Hochberg, 1995) has
-been cited 21,787 times (May 2014) according to Google Scholar.
-
-- That is three times a day for the last 19 years!
 
@@
 plot_FDReg_hist(pval,pi0=qval$pi0,ylim=c(0,4))
-```
-
-## Mouse Data: Counts vary considerably across genes
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-hist(log10(xx),main="Distribution of total counts", xlab="log10(counts)")
 ```
@@
 
+```{r, include=FALSE}
+ #plot a histogram of z scores, highlighting the alternative distribution
+#of z scores that is implied by pi0
+nullalthist_z = function(z,pi0,nullcol="blue",altcol="cyan",...){
+  h=hist(z, freq=FALSE,col=nullcol,nclass=40,ylim=c(-0.5,0.5),...)
+  nb = length(h$breaks)
+  nullexp = pi0 * (pnorm(h$breaks)[-1] - pnorm(h$breaks[-nb]))/(h$breaks[-1]-h$breaks[-nb])
+  h$density = h$density - nullexp
+  plot(h,add=TRUE,col=altcol,freq=FALSE)
+}
+
+ set.seed(100)
+z = c(rnorm(5000,0,2),rnorm(5000,0,1))
+```
+
 ## Recall Problem 1: distribution of alternative Z values multimodal
@@
 
-- In the illustrative example, the maximum $q$ value is `r round(max(hh.ash$qvalue),2)`
+- Actually this can be true for existing `qvalue` approach too!
 
@@
 
-## Advantages of False Sign Rate
-
-- The FDR is typically not actually identifiable from data.
-
-- This is because data cannot distinguish between $\beta_j = 0$ and $\beta_j$ "very
-small". So $\pi_0$ is not identifiable, and FDR is very sensitive to $\pi_0$.
-
-- So methods for estimating $\pi_0$ and FDR, including those presented here, are designed to be ``conservative" (i.e. overestimate the FDR).
-
-- The False Sign Rate is much less senstive to $\pi_0$, and hence more identifiable from data!
-
-## Simulated example: $\pi_0$ not identifiable.
-
-```{r,echo=FALSE,include=FALSE}
-load("../paper/Rcode/sim1.RData") #load simulation results
-source("../paper/Rcode/plot_pi0.R")
-```
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-plot_pi0(list(simres1a))
-```
-
-## Simulated Example: so fdr not identifiable
-
-```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-source("../paper/Rcode/plot_lfsr.R")
-  plot_lfsr(list(simres1a),0.1,ptype="lfdr")
-```
-
-## Simulated Example: fsr much more identifiable
-
-```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-  plot_lfsr(list(simres1a),0.1,ptype="lfsr")
-```
-
 ## Summary
@@
 
-## Other Applications
-
-- Widely applicable: requiring only an estimated
-effect size and standard error for each object.
-
-- E.g. Currently applying it to wavelet shrinkage applications.
-
 ## Next steps?
@@
 
-- NHGRI for funding.
+- NHGRI and GTEX for funding.
 
-- `ashr` package: `http://www.github.com/stephens999/ash`
+- `ashr` R package: `http://www.github.com/stephens999/ash`
 
-## Pandoc Command used
-
-`pandoc -s -S -i --template=my.beamer -t beamer -V theme:CambridgeUS -V colortheme:beaver  slides.md -o slides.pdf`
-
-(alternative to produce html slides; but figures would need reworking)
-`pandoc -s -S -i -t dzslides --mathjax slides.md -o slides.html`
-
-Here is my session info:
+## Session Info
 

### 2014-12-17 talks/test.Rmd
added some talks

--- 2014-12-17 talks/ASC2014.newtext.Rmd
@@
 ---
-title: "Test-Presentation"
-author: "Matthew Stephens"
-date: "September 23, 2014"
+title: "test"
 output: ioslides_presentation
 ---
-
-% False Discovery Rates, A New Deal
-% Matthew Stephens
-% 2014/7/8
-
-```{r, include=FALSE}
-require("qvalue")
-require("ashr")
-require("ggplot2")
-```
-
-```{r setup, include=FALSE}
-  #set global chunk options
-  #opts_chunk$set(cache=TRUE,autodep=TRUE,warning=FALSE)
-  #dep_auto()
-```
-
-# Before we get started: Let's get Organized!
-
-- Over ~10 years of working with graduate students + postdocs,
-I've noticed something.
-- Organized researchers get more done (and better!).
-- Many of them are more organized than I am!
-- Thought: I should get organized; I should help others get organized.
-
-# So what can you do?
-
-- Buy a notebook; bring it to meetings; make notes! 
-- Come to meetings with a written agenda. 
-- While doing research, record what you did and what the outcome was.
-- Use version control ([git](http://git-scm.com)) and internet repositories ([bitbucket](http://www.bitbucket.org), [github](http://www.github.com)) to organize notes, code, etc.
-- Use *knitr* to help make your research reproducible.
-- Talk about the tools you find useful!
-
-# What are these repository things?
-
-- A repository: a central place in which an aggregation of data is kept and maintained in an organized way (searcharticle.com)
-- Great for sharing material across multiple people (eg student and advisor!)
-- An amateur example: [http://github.com/stephens999/ash](http://github.com/stephens999/ash)
-
-# What is knitr?
-
-- An R package
-- A tool for literate programming
-- Text, and R code are interleaved
-- When you compile the document, the code is run, and output inserted into the text.
-- Great for writing reports, and keeping a track of what you did and what the result was!
-- This talk was written with knitr (with RStudio)!
-
-# What is Reproducible Research?
-
-- Principle: when publishing results of computational procedures, we should
-publish the code that produced the results.
-- "publishing figures or results without the complete software environment could
-be compared to a mathematician publishing an announcement of a mathematical theorem without giving the proof" (Buckheit and Donohoe)
-- “an article about a computational result is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result.” [Claerbout]
-
-# Why is reproducibility important?
-
-- Not only because people are forgetful, error-prone, or dishonest!
-- Reproducing work is also the first step to extending it.
-- Helps communications among researchers (eg student + advisor).
-- If you do not publish code implementing your methods, your methods will likely go unused.
-
-# More on git, github, knitr, reproducibility
-
-- Google "The git book", to get started on git.
-- Google "Karl Broman github tutorial" for statistics-oriented intro to github.
-- Google "donohoe buckheit" for "Wavelab and reproducible research"
-
-# The Canonical Genomics Experiment
-
-- Measure lots of things, with error
-
-- Get estimates of effects $\beta_j$ ($\hat\beta_j$) and their standard errors $s_j$
-
-- Turn these into Z-scores, $z_j = \hat\beta_j/s_j$
-
-- Turn these into $p$ values, $p_j$
-
-- Apply `qvalue` 
-to identify findings ``significant" at a given FDR.
-
-- ...?
-
-# FDR, local fdr, and q values
-
-Although precise definitions vary depending on whether one
-takes a Bayesian or Frequentist approach to the problem, roughly
-
-- The FDR at a threshold $P$ is 
-$$\text{FDR}(P)=\Pr(\beta_j = 0 |  p_j<P).$$
-
-- The q value for observation $j$ is $q_j=\text{FDR}(p_j)$.
-
-- The local false discovery rate, fdr, at threshold $P$ is 
-$$\text{fdr}(P) = \Pr(\beta_j =0 | p_j=P ).$$
-
-- The fdr is more relevant, but slightly harder to estimate than 
-FDR because it involves density estimation rather than tail-area estimation.
-
-```{r, echo=FALSE}
-  #simple simulated example
-  ncz = 100 # number of bins in z score histograms
-  set.seed(111)
-  hh.betahat = rnorm(1000,0,2)
-  hh.sebetahat = 1
-  hh.zscore = hh.betahat/hh.sebetahat
-  hh.pval = pchisq(hh.zscore^2,df=1,lower.tail=F)
-  hh.ash = ash(hh.betahat,hh.sebetahat, method="fdr")
-  hh.q = qvalue(hh.pval)
-```
-
-```{r, echo=FALSE}
-  plot_FDReg_hist = function(hh.pval,pi0,nc=40,nullcol="blue",altcol="cyan",type=4,title="Distribution of p values",...){
-    hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main=title,nclass=nc,col=altcol,...)
-    if(type>1){
-      abline(h=pi0,col=nullcol,lwd=2)
-
-      hh.hist$density=rep(pi0,length(hh.hist$density))  
-      #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
-      plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)
-    }
-    if(type>2){
-    abline(v=0.1,lwd=2,col=2)
-    }
-    if(type>3){
-      text(0.05,1.2,labels="A",col=2,cex=1.2)  
-      text(0.05,0.4,labels="B",col=2,cex=1.2)  
-      text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
-    }
-  }
-```
-
-# Example: FDR estimation
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=1)                 
-```
-
-# Example: FDR estimation
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=2)                 
-```
-
-# Example: FDR estimation
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=3)                 
-```
-
-# Example: FDR estimation
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4)                 
-```
-
-# Is this an important problem?
-
-- The original paper introducing FDR (Benjamini and Hochberg, 1995) has
-been cited 21,787 times (May 2014) according to Google Scholar.
-
-- That is three times a day for the last 19 years!
-
-# Problem 1: The Zero Assumption (ZA)
-
-- The standard `qvalue` 
-approach assumes that all the $p$ values near 1 are null.
-
-- Analogously, one can assume that all Z scores near 0 are null. Efron refers to this as the ``Zero Assumption".
-
-- Seems initially natural.
-
-```{r, echo=FALSE, include=FALSE}
-require(fdrtool)
-  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue",plot=FALSE)
-require(locfdr)
-  hh.locfdr = locfdr(hh.zscore,nulltype=0,plot=0)
-require(mixfdr)
-  hh.mixfdr = mixFdr(hh.zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
-```
-
-# Implied distribution of $p$ values under $H_1$
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4)                 
-```
-
-# Implied distribution of Z scores under alternative
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-#plot a histogram of z scores, highlighting the alternative distribution
-#of z scores that is implied by localfdr values lfdr.
-  nullalthist = function(z,lfdr,nullcol="blue",altcol="cyan",...){
-    h=hist(z, freq=FALSE,col=nullcol,nclass=ncz,...)
-    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
-    h$density = (1-avlfdr) * h$density
-    plot(h,add=TRUE,col=altcol,freq=FALSE)
-  }
-
-  nullalthist(hh.zscore,hh.fdrtool$lfdr)  
-```
-
-# FDR problem 2: different measurement precision
-
-- In some cases the measurement precisions differ among units
-
-- Eg Expression levels of low-expressed genes have less precision than high-expressed genes
-
-- If some effects are measured less precisely than others, those tests ``lack power"
-and dilute signal, increasing FDR
-
-# Example: Mouse Heart Data
-
-```{r, echo=FALSE}
-x = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
-xx = rowSums(x[,2:5])
-x = x[xx>0,]
-xx = xx[xx>0]
-```
-
-```{r echo=FALSE}
-head(x)
-```
-
-- Data on 150 mouse hearts, dissected into left and right ventricle
-(courtesy Scott Schmemo, Marcelo Nobrega)
-
-```{r, echo=FALSE,include=FALSE}
-
-load("../scratchwork/Smemo_qbinom_zdat.RData")
-# two-sided test
-  ttest.pval = function(t, df) {
-    pval = pt(t, df = df, lower.tail = T)
-    pval = 2 * ifelse(pval < 0.5, pval, 1 - pval)
-    return(pval)
-}
-
-tscore = zdat[3, ]/zdat[4, ]
-pval = ttest.pval(tscore, df = 2)
-qval = qvalue(pval)
-
-highxp = xx>1000 # select high expressed genes
-pval.high = pval[highxp]
-qval.high = qvalue(pval.high)
-
-zdat.ash = ash(zdat[3,],zdat[4,],df=2,method="fdr",mixcompdist="halfuniform")
-zdat.ash.high = ash(zdat[3,highxp],zdat[4,highxp],df=2,method="fdr",mixcompdist="halfuniform")
-
-```
-
-# Example: Mouse Heart Data
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-qval= qvalue(pval)
-plot_FDReg_hist(pval,pi0=qval$pi0,ylim=c(0,4))
-```
-
-# Mouse Data: Counts vary considerably across genes
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-hist(log10(xx),main="Distribution of total counts", xlab="log10(counts)")
-```
-
-# Lower count genes, less power
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-qval.low=qvalue(pval[!highxp])
-pval.low=pval[!highxp]
-plot_FDReg_hist(pval.low,pi0=qval.low$pi0,ylim=c(0,4))
-```
-
-# Higher count genes, more power
-```{r, echo=FALSE, fig.height=4,fig.cap=""}
-plot_FDReg_hist(pval.high,pi0=qval.high$pi0,ylim=c(0,4))
-```
-
-# Low-count genes dilute signal at high-count genes
-```{r, echo=FALSE, fig.height=4,fig.cap=""}
-plot_FDReg_hist(pval,pi0=qval$pi0,ylim=c(0,4))
-```
-
-# FDR problem 2: low count genes add noise, increase q values
-```{r,echo=FALSE, fig.height=4,fig.cap=""}
-plot(qval.high$qval,qval$qval[highxp],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
-abline(a=0,b=1,col=2)
-```
-
-# Problems: Summary
-
-Standard tools are unduly conservative.
-
-- The ZA, which implies actual effects have a (probably unrealistic) bimodal distribution; causes overestimate of $\pi_0$, losing power.
-
-- By focussing on $p$ values, low-precision measurements can dilute high-precision measurements.
-
-# FDR via Empirical Bayes
-
-- Following previous work (e.g. Newton, Efron, Muralidharan) we take an empirical Bayes approach to FDR.
-
-- Eg Efron assumes that the $Z$ scores come from a mixture of null, and alternative:
-$$Z_j \sim f_Z(.) = \pi_0 N(.;0,1) + (1-\pi_0) f_1(.)$$
-where $f_1, \pi_0$ are to be estimated from the data.
-
-- Various semi-parametric approaches taken to estimating $f_1$. For example,
-Efron uses Poisson regression; Muralidharan uses mixture of normal distributions.
-
-- Once $f_1$ and $\pi_0$ estimated, FDR calculations are straightforward.
-
-# FDR: A New Deal
-
-- Instead of modelling $Z$ scores, model the effects $\beta$,
-$$\beta_j \sim \pi_0 \delta_0(.) + (1-\pi_0) g(.)$$
-
-- Constrain $g$ to be unimodal about 0; estimate $g$ from data.
-
-- *Incorporate precision* of each observation $\hat\beta$ into the likelihood.
-Specifically, approximate likelihood for $\beta_j$ by a normal: 
-$$L(\beta_j) \propto \exp(-0.5 (\beta_j - \hat\beta_j)^2/s_j^2).$$
-[From $\hat\beta_j \sim N(\beta_j, s_j)$]
-Or, better, use a $t$ likelihood if $s_j$ estimated using few observations.
-
-# FDR - A New Deal
-
-- A convenient way to model $g$: mixture of 0-centered
-normal distributions: 
-$$g(\beta; \pi) = \sum_{k=1}^K \pi_k N(\beta; 0, \sigma^2_k)$$
-
-- Estimating $g$ comes down to estimating $\pi$. Joint estimation of $\pi_0,\pi$ easy by maximum likelihood (EM algorithm).
-
-- By allowing $K$ large, and $\sigma_k$ to span a dense grid of values,
-we get a flexible unimodal symmetric distribution.
-
-- Can approximate, arbitrarily closely, any scale mixture of normals.
-Includes almost all priors used for sparse regression problems (spike-and-slab, double exponential/Laplace/Bayesian Lasso, horseshoe).
-
-# FDR - A New Deal
-
-- Alternatively, a mixture of uniforms, with 0 as one end-point of the range,
-provides still more flexibility, and in particular allows for asymmetry.
-
-- If allow a very large number of uniforms this provides the non-parametric mle for $g$; cf Grenander 1953; Cordy + Thomas 1997.
-
-# Illustration: $g$ a mixture of 0-centered normals
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-x=seq(-4,4,length=100)
-plot(x, dnorm(x,0,1),type="l",ylim=c(0,2),ylab="density")
-lines(x, dnorm(x,0,0.1))
-lines(x, dnorm(x,0,0.5))
-```
-
-# Illustration: $g$ a mixture of 0-centered normals
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-x=seq(-4,4,length=100)
-plot(x, 0.5*dnorm(x,0,1)+0.5*dnorm(x,0,0.1),type="l",ylim=c(0,2),ylab="density")
-```
-
-# Illustration: $g$ a mixture of 0-anchored uniforms
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-x=seq(-4,4,length=100)
-plot(x, dunif(x,0,1),type="s",ylim=c(0,5),ylab="density")
-lines(x, dunif(x,0,0.2),type="s")
-lines(x, dunif(x,0,0.5),type="s")
-lines(x, dunif(x,-0.3,0),type="s",col=2)
-lines(x, dunif(x,-0.4,0),type="s",col=2)
-lines(x, dunif(x,-2,0),type="s",col=2)
-```
-
-# Illustration: $g$ a mixture of 0-anchored uniforms
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-x=seq(-4,4,length=100)
-plot(x, 0.1*dunif(x,0,3)+ 0.3*dunif(x,0,0.2) + 0.2*dunif(x,0,0.5) + 0.1* dunif(x,-0.3,0) + 0.1*dunif(x,-0.3,0)+0.2*dunif(x,-0.4,0),type="s",ylim=c(0,2),ylab="density")
-```
-
-# Adaptive Shrinkage
-
-- This approach actually provides a full posterior distribution for each $\beta_j$.
-
-- So easy to obtain point estimates and credible intervals.
-
-- Because $g(\beta)$ is unimodal, the point estimates (and CIs) will tend to be ``shrunk" towards the overall mean (0).
-
-- Because $g(\beta)$ is estimated from the data, the amount
-of shrinkage is adaptive to signal in the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each observation.
-
-- So we call the approach ``Adaptive Shrinkage" (ASH).
-
-# Recall Problem 1: distribution of alternative Z values multimodal
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-nullalthist(hh.zscore,hh.fdrtool$lfdr,main="qvalue")  
-```
-
-# Problem Fixed: distribution of alternative Z values unimodal
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-nullalthist(hh.zscore,hh.ash$lfdr,main="ash")
-```
-
-# Example: FDR estimation
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4,title="qvalue")                 
-```
-
-# Example: FDR estimation
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,get_pi0(hh.ash),type=4,title="ash")                 
-```
-
-# Recall Problem 2: low count genes add noise, increase q values
-```{r,echo=FALSE, fig.height=4,fig.cap=""}
-plot(qval.high$qval,qval$qval[highxp],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
-abline(a=0,b=1,col=2)
-```
-
-# Problem Fixed: incorporating precision reduces influence of low-count genes
-```{r,echo=FALSE, fig.height=4,fig.cap=""}
-plot(zdat.ash.high$qvalue,zdat.ash$qvalue[highxp],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
-abline(a=0,b=1,col=2)
-```
-
-# A new problem: an embarrassment of riches
-
-- If the null is mostly false, the new approach can  provide unsettling results
-
-- The FDR can be small for all observations, even those with $p \approx 1$!
-
-- In the illustrative example, the maximum $q$ value is `r round(max(hh.ash$qvalue),2)`
-
-# Perhaps we didn't really understand the question?
-
-- Problem arises only if we insist on asking question ``is $\beta_j=0$?"
-
-- Given enough signal, we become convinced that very few of the $\beta_j=0$
-
-- But for some $\beta_j$ we still may have little information about actual value
-
-- Suggests a change of focus: ask for which $\beta_j$ are we confident about the sign (cf Gelman et al, 2012).
-
-# The False Sign Rate
-
-- Suggestion: replace FDR with local false sign rate (lfsr), the probability that if we say an effect is positive (negative), it is not.
-
-- Example: suppose we estimate that $\Pr(\beta_j<0)=0.95, \Pr(\beta_j=0) = 0.025$ and $\Pr(\beta_j>0)=0.025$. Then we report $\beta_j$ as a ``(negative) discovery", and estimate its lfsr as 0.05.
-
-# Even with many signals, large $p$ values have high lfsr
-
-```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-  #res=data.frame(p=hh.pval, lfsr=hh.ash$lfsr, qvalue = hh.ash$qvalue)
-  library("reshape2")
-  res=data.frame(p=pval, lfsr=zdat.ash$lfsr, qvalue = zdat.ash$qvalue)
-  res.melt = melt(res, id.vars=c("p"),variable.name="Measure")
-  cbbPalette <- c("#000000", "#D55E00", "#CC79A7")
-  labels = c('qvalue','lfsr')
-  breaks = c("qvalue","lfsr")
-
-  pp= ggplot(data=res.melt,aes(p,value,color=Measure)) +geom_point(shape=1) +
-      geom_abline(colour = "black") +
-        xlab("p-value") +
-        ylab("lfsr/qvalue")
-
-  print(pp +scale_y_continuous(limits=c(0,1)) +
-          scale_x_continuous(limits=c(0,1))  +
-           scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
-          coord_equal(ratio=1))
-
-```
-
-# Advantages of False Sign Rate
-
-- The FDR is typically not actually identifiable from data.
-
-- This is because data cannot distinguish between $\beta_j = 0$ and $\beta_j$ "very
-small". So $\pi_0$ is not identifiable, and FDR is very sensitive to $\pi_0$.
-
-- So methods for estimating $\pi_0$ and FDR, including those presented here, are designed to be ``conservative" (i.e. overestimate the FDR).
-
-- The False Sign Rate is much less senstive to $\pi_0$, and hence more identifiable from data!
-
-# Simulated example: $\pi_0$ not identifiable.
-
-```{r,echo=FALSE,include=FALSE}
-load("../paper/Rcode/sim1.RData") #load simulation results
-source("../paper/Rcode/plot_pi0.R")
-```
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-plot_pi0(list(simres1a))
-```
-
-# Simulated Example: so fdr not identifiable
-
-```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-source("../paper/Rcode/plot_lfsr.R")
-  plot_lfsr(list(simres1a),0.1,ptype="lfdr")
-```
-
-# Simulated Example: fsr much more identifiable
-
-```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-  plot_lfsr(list(simres1a),0.1,ptype="lfsr")
-```
-
-# Summary
-
-- ASH provides a generic approach to shrinkage estimation, as well as
-false discovery (sign) rates.
-
-- But by using two numbers ($\hat\beta,s$) instead of one ($p$ values or $z$ scores) varying precision of 
-measurements is better accounted for.
-
-- Unimodal assumption for effects reduces conservatism
-
-- False Sign Rate preferable to False Discovery Rate: more identifiable, and
-better representation of information in data for ``high-signal" situations.
-
-# Other Applications
-
-- Widely applicable: requiring only an estimated
-effect size and standard error for each object.
-
-- E.g. Currently applying it to wavelet shrinkage applications.
-
-# Next steps?
-
-- Incorporate shrinkage of variances and not just means. (e.g. ``moderated $t$ test", Smyth et al)
-
-- Allow $g(\cdot;\pi)$ to depend on covariates $X$.
-
-- Allow for correlations in the measured $\hat\beta_j$.
-
-# Thanks
-
-- to the developers of **R**, **knitr**, **Rstudio** and **Pandoc**.
-
-- to the several postdoctoral researchers and students
-who have worked with me on related topics.
-
-- Including Ester Pantaleo, Scott Powers, Mengyin Lu, Sen Tian, Wei Wang, Zhengrong Xing.
-
-- NHGRI for funding.
-
-- `ashr` package: `http://www.github.com/stephens999/ash`
-
-# Pandoc Command used
-
-`pandoc -s -S -i --template=my.beamer -t beamer -V theme:CambridgeUS -V colortheme:beaver  slides.md -o slides.pdf`
-
-(alternative to produce html slides; but figures would need reworking)
-`pandoc -s -S -i -t dzslides --mathjax slides.md -o slides.html`
-
-Here is my session info:
-
-```{r session-info}
-print(sessionInfo(), locale=FALSE)
-```
-
-# Some odd things in the data
-
-```{r, echo=FALSE,fig.height=3}
-  plot(zdat[3,],zdat[4,],ylab="standard error",xlab="beta-hat")
-  dd[tail(order(zdat[3,])),]
-```
-
-# A technicality
-
-- Suppose you estimate $\Pr(\beta_j<0)=0.98$,  $\Pr(\beta_j>0)=0.01$, $\Pr(\beta_j=0) = 0.01$. 
-- Should you declare an fdr of 0.01 or 0.02?
-- Maybe fsr makes more sense anyway?
-
-# Adaptive Shrinkage of point estimates
-
-- Recall idea: amount of shrinkage depends on measurement precision, $s_j$.
-
-# Adaptive Shrinkage of point estimates
-
-```{r, fig.height=4,echo=FALSE,fig.cap=""}
-  res=data.frame(betahat =zdat[3,], Estimate=zdat.ash$PosteriorMean,highxp=highxp)
-  cbbPalette <- c("#000000", "#D55E00", "#CC79A7")
-  labels = c('betahat','Estimate')
-  breaks = c("betahat","Estimate")
-
-  pp= ggplot(data=res,aes(betahat,Estimate,col=as.factor(highxp))) +geom_point(shape=1) +
-      geom_abline(colour = "black") +
-        xlab("betahat") +
-        ylab("shrunk betahat")
-
-  print(pp +scale_y_continuous(limits=c(-2,2)) +
-          scale_x_continuous(limits=c(-2,2))  +
-           scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
-          coord_equal(ratio=1))
-
-#plot(zdat[3,],zdat.ash$PosteriorMean,main="Less shrinkage for high-count genes (red)",xlab="betahat",ylab="shrunk betahat",xlim=c(-2,2),ylim=c(-2,2), col=1+highxp)
-#abline(a=0,b=1,col=2)
-```
-
-# Shrinkage is adaptive to information
-
-Need to fix counts.associate to use fdr method in ash
-
-```{r, echo=FALSE,fig.height=4}
-#hist(zdat[3,])
-#plot(zdat[3,],zdat.ash$PosteriorMean,xlim=c(-0.2,0.2))
-#points(zdat[3,16677],zdat.ash$PosteriorMean[16677],col=2,pch=16)
-#points(zdat[3,16079],zdat.ash$PosteriorMean[16079],col=2,pch=16)
-#x[16677,]
-#x[16079,]
-```
-
-```{r, echo=FALSE}
-#temp = counts.associate(cc,c(-1,1,-1,1),1)
-#temp.tscore = temp$zdat[3,]/temp$zdat[4,]
-#temp.pval = ttest.pval(temp.tscore,df=2)
-#plot(temp$zdat.ash$localfdr,temp.pval)
-#identify(temp$zdat.ash$localfdr,temp.pval)
-```
-
-```{r, echo=FALSE,fig.height=3,fig.cap=""}
-plot(zdat.ash$lfdr,pval,ylab="p value", xlab="ASH local fdr")
-```
-
-# Shrinkage is adaptive to information
-
-```{r, echo=FALSE,fig.height=3,fig.cap=""}
-plot(zdat.ash$lfdr,pval,ylab="p value", xlab="ASH local fdr")
-points(zdat.ash$lfdr[15325],pval[15325],col=2,pch=16)
-points(zdat.ash$lfdr[16123],pval[16123],col=2,pch=16)
-```
-
-# Shrinkage is adaptive to information
-
-```{r, echo=FALSE}
-dd = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
-xx = rowSums(dd[,2:5])
-dd = dd[xx>0,]
-xx = xx[xx>0]
-cbind(dd[,1:5],pval,zdat.ash$lfdr)[c(15325,16123),]
-```
-
-# Recall FDR problem 1: q values increased by low count genes
-```{r,echo=FALSE, fig.height=4,fig.cap=""}
-plot(qval.high$qval,qval$qval[xx>1000],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
-abline(a=0,b=1,col=2)
-```
-
-# ASH q values more robust to inclusion of low count genes
-
-```{r, echo=FALSE,fig.height=3,fig.cap=""}
-plot(zdat.ash.high$qvalue,zdat.ash$qvalue[xx>1000],xlab="q values from high only",ylab="q values from all data")
-abline(a=0,b=1,col=2,lwd=2)
-```
-
-Compare fitted $f(\beta)$, both estimating $\pi_0$ and fixing $\pi_0=0$.
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  x=seq(-4,4,length=100)
-  plot(cdf.ash(hh.ash.fdr,x),col=2,type="l",lwd=3,ylab="cdf")
-  lines(x,cdf.ash(hh.ash.shrink,x)$y,col=3,lwd=3,lty=2)
-```
 

### 2015-02-05 paper/main.tex
updated dsc to new format

@@
 measurement precision (standard error), 
-and the goal is to identify which of the effects we can declare to be non-zero (with some prescribed level of confidence), and also
-to provide estimates -- both point estimates and interval estimates -- for these effects.
+and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
+Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
+here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
 This setting arises commonly in many areas of statistics, and particularly in genomics applications.
@@
 whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
-expression level between the two conditions, often referred to as the ``log-fold change". 
+mean expression level between the two conditions, often referred to as the ``log-fold change". 
 %Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.
@@
 
-Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and several are widely used.
-Here we present new tools that bring two important new ideas to bear on the problem.
+Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
+Here we present new tools that bring two important ideas to bear on the problem.
 First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
@@
 factor may be the paucity of general-purpose software packages for this kind of shrinkage estimation.
- (There are, perhaps, more packages for shrinkage esitmation in the more complex setting of covariance estimation, where shrinkage is perhaps still more important.)
+% (There are, perhaps, more packages for shrinkage estimation in the more complex setting of covariance estimation, where shrinkage is perhaps still more important.)
 Our work here addresses both these issues, providing a general-purpose software tool that uses shrinkage ideas to improve both
@@
 In contrast the method introduced here (\ashr) assumes that the effect sizes (and thus the $z$ scores) have a unimodal distribution about 0, resulting in a very different decomposition that in this case is much closer to the truth.
-(These data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything light blue.)} \label{fig:ZA}
+(These data were simulated under a model where all of the effects are non-zero -- specifically, $\beta_j \sim N(0,1)$, $s_j=1$, resulting in $z_j \sim N(0,2)$ -- so the ``true" decomposition would make everything light blue.)} \label{fig:ZA}
 \end{figure}
@@
 \centering
-\subfloat[][Histograms of $p$ values for good-precision, poor-precision, and combined observations]{
+\subfloat[][Density histograms of $p$ values for good-precision, poor-precision, and combined observations]{
 \includegraphics[width=6.5in]{Rcode/figures/GOODPOOReg_hist.pdf} 
@@
 }
-\caption{Simulated illustration how poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
+\caption{Simulation illustrating how poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
 $p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
@@
 
-\begin{figure}[!ht] \label{fig:good_poor_hist}
-\begin{center}
-\includegraphics[height=3in]{Rcode/figures/good_poor_hist.pdf}
-\includegraphics[width=6.5in]{Rcode/figures/good_vs_all.pdf}
-\end{center}
-\caption{}
-\end{figure}
+PENN Note: need to fix bug in dscr where when an element of result is null, the data.frame command doesn't work.
+Want to convert it to missing - maybe in the scoring function.
 
@@
 \begin{equation}
-g(\cdot) = (1/7) [N(\cdot; -1.5, 0.5^2) + N(\cdot; -1, 0.5^2) + N(\cdot; 0.5, 0.5^2) + N(\cdot; 0, 0.5^2) + N(\cdot; 0.5, 0.5^2) + N(\cdot; 1, 0.5^2) + N(\cdot; 1.5, 0.5^2)]
+g(\cdot) = (1/7) [N(\cdot; -1.5, 0.5^2) + N(\cdot; -1, 0.5^2) + N(\cdot; -0.5, 0.5^2) + N(\cdot; 0, 0.5^2) + N(\cdot; 0.5, 0.5^2) + N(\cdot; 1, 0.5^2) + N(\cdot; 1.5, 0.5^2)]
 \end{equation}
@@
 
-This work was supported by NIH grant .
+This work was supported by NIH grant xxx and a grant from the Gordon and Betty Moore Foundation.
 

### 2015-02-05 dsc/run.R

@@ -1,21 +1,0 @@
-#install.packages("/Users/stephens/Dropbox/Documents/git/dscr_0.1.tar.gz",repos=NULL,type="source")
-
-library("plyr")
-library("ashr")
-library("reshape2")
-library(dscr)
-
-source("datamaker.R")
-source("method.R")
-source("scenario.R")
-source("score.R")
-
-res=run_dsc(scenarios,methods,score)
-
-
-head(res)
-
-aggregate(td~method+scenario,res,mean)
-aggregate(fd~method+scenario,res,mean)
-res$fdr = res$fd/(res$fd+res$td)
-aggregate(fdr~method+scenario,res,mean)

### 2015-02-06 dsc-shrink/graphs.Rmd
added shiny graphic for showing boxplot of dsc-shrink results

---
title: "Graphs"
runtime: shiny
output: html_document
---

Here's a boxplot of results. 
```{r, echo=FALSE}
library(ggplot2)
library(dplyr)
load("res.RData")
scenario_names = as.character(unique(res$scenario))
method_names = as.character(unique(res$method))
numeric_criteria = names(res)[unlist(lapply(res,is.numeric))]

inputPanel(    
    # Create the checkboxes and select them all by default
    checkboxGroupInput("scen.subset", "Choose Scenarios", 
                        choices  = scenario_names,
                        selected = scenario_names),

    checkboxGroupInput("method.subset", "Choose Methods", 
                        choices  = method_names,
                        selected = method_names),

    selectInput("criteria", "Choose Criteria", 
                        choices  = numeric_criteria,
                        selected = "RMSE")
)

renderPlot({
  res.filter = filter(res,scenario %in% input$scen.subset & method %in% input$method.subset)
  print(input)
  res.filter$value = res.filter[[input$criteria]]
  ggplot(res.filter,aes(method,value,color=method)) + geom_boxplot() + facet_grid(.~scenario)
})
```

### 2015-02-06 dsc-shrink/summarize.R

@@ -1,8 +1,19 @@
 aggregate(RMSE~method+scenario,res, mean)
 library(ggplot2)
+
+### Plot RMSE boxplot
 ggplot(res,aes(method,RMSE,color=method)) + geom_boxplot() + facet_grid(.~scenario)
 
-p=ggplot(data=res,aes(pi0,pi0_est,colour=method)) +geom_point(shape=1) +
+### Plot RMSE boxplot for subset of methods
+ggplot(filter(res,scenario %in% c("An","Bn","Cn")),aes(method,RMSE,color=method)) + geom_boxplot() + facet_grid(.~scenario)
+
+### or boxplot of elapsed time
+ggplot(filter(res,!(scenario %in% c("hard-b"))),aes(method,elapsed,color=method)) + geom_boxplot() + facet_grid(.~scenario)
+
+
+## plot true pi0 vs estimated pi0
+
+p=ggplot(data=filter(res,scenario %in% c("hard","hard-b","easy")),aes(pi0,pi0_est,colour=method)) +geom_point(shape=1) +
   facet_grid(. ~ scenario) +
   geom_abline(colour = "black") +
   xlab("True pi0") +

### 2015-02-06 scratchwork/Smemo.Rmd
added some scratchwork files

DESeq analysis of the Smemo et al heart data
========================================================

```{r}
  set.seed(100)
  opts_chunk$set(cache=TRUE,autodep=TRUE,warning=FALSE)
  dep_auto()
```

Read in data
```{r}
  x = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
  x.counts = x[,2:5]
  x = x[rowSums(x.counts)>0,]
  rownames(x)=1:nrow(x)
  x.counts = x[,2:5]
```

Run DESeq on the data, following template in the DESeq vignette.
```{r}
  library("DESeq")

  condition=factor(c("left","left","right","right"))
  cds = newCountDataSet(x.counts, condition )
  cds = estimateSizeFactors(cds)
  sizeFactors( cds )
  cds = estimateDispersions( cds )
  plotDispEsts( cds )
  res = nbinomTest( cds, "left", "right" )
  save("res", file="Smemo_DESeq_res.RData")
```

Also run it on "null" data created by permuting conditions
```{r}
  cds.null = newCountDataSet(x.counts, condition=factor(c("left","right","left","right")))
  cds.null = estimateSizeFactors(cds.null)
  sizeFactors( cds.null )
  cds.null = estimateDispersions( cds.null )
  plotDispEsts( cds.null )
  res.null = nbinomTest( cds.null, "left", "right" )
  save("res.null", file="Smemo_DESeq_res.null.RData")
```

Now examine distribution of p values. Note that it seems that for very low-count genes a lot of 
p values are equal to 1, so I might choose to remove these.
```{r}
  tots = rowSums(x.counts)
  hist(res$pval[tots<100])
  hist(res$pval[tots<1000 & tots >=100])
  hist(res$pval[tots>=1000])
  hist(tots[res$pval==1],nclass=1000,xlim=c(0,1000))
```

Compare with simple quasi-binomial glm
```{r}
  source("../../stat45800/code/Rcode/PoissonBinomial.funcs.R")  
  g = c(-0.5, -0.5, 0.5, 0.5)
  zdat = counts.associate(x.counts, g, 1)
  save("zdat",file="Smemo_qbinom_zdat.RData")
```

Also run simple quasi-binomial glm on permuted dat
```{r}
  zdat.null = counts.associate(x.counts, c(-0.5,0.5,-0.5,0.5), 1)
  save("zdat.null",file="Smemo_qbinom_zdat.null.RData")
```

```{r}
  # two-sided test
  ttest.pval = function(t, df) {
    pval = pt(t, df = df, lower.tail = T)
    pval = 2 * ifelse(pval < 0.5, pval, 1 - pval)
    return(pval)
  }

  tscore = zdat[3, ]/zdat[4, ]
  pval = ttest.pval(tscore, df = 2)
  #qval = qvalue(pval)

  #highxp = xx>1000 # select high expressed genes
  #pval.high = pval[highxp]
  #qval.high = qvalue(pval.high)
  #cc.assoc.high = counts.associate(cc[highxp,],g,1)
  #zdat.ash.high = cc.assoc.high$zdat.ash
  #zdat.high = cc.assoc.high$zdat
```

```{r}
  plot(zdat[3,],res$log2FoldChange)
  plot(pval,res$pval,xlim=c(0,0.1),ylim=c(0,0.1))

```

```{r}
  tscore.null = zdat.null[3, ]/zdat.null[4, ]
  pval.null = ttest.pval(tscore.null, df = 2)
```

# Example: ASH applied to mouse data

```{r, fig.height=4,echo=FALSE,fig.cap=""}
hist(zdat[3,],main="Raw effect size estimates",xlab="betahat",xlim=c(-2,2),prob=T,ylim=c(0,3))
```

# Example: ASH applied to mouse data

```{r, fig.height=4,echo=FALSE,fig.cap=""}
hist(zdat[3,],main="Raw effect size estimates",xlab="betahat",xlim=c(-2,2),prob=T,ylim=c(0,3))
t=seq(-2,2,length=201)
lines(density(zdat.ash,t),type="l")
```

# Example: ASH applied to mouse data

```{r, fig.height=4,echo=FALSE,fig.cap=""}
hist(zdat.ash$PosteriorMean,main="Shrunken estimates",xlab="shrunk betahat",prob=T)
lines(density(zdat.ash,t),type="l")
```

# Example: ASH applied to mouse data

```{r, fig.height=4,echo=FALSE,fig.cap=""}
plot(zdat[3,],zdat.ash$PosteriorMean,main="Estimates vs Shrunken estimates",xlab="betahat",ylab="shrunk betahat",xlim=c(-2,2),prob=T)
abline(a=0,b=1,col=2)
```

```{r}
  subset = res$pval<1
  betahat = res$log2FoldChange
  #replace the infinite estimates of log2FoldChange with the largest finite fold change
  betahat[is.infinite(betahat)] = sign(betahat[is.infinite(betahat)]) * max(abs(betahat[is.finite(betahat)]))
  zscore = sign(betahat) * sqrt(qchisq(res$pval,df=1))
  betahatsd = betahat/zscore
  library("ashr")
  beta.ash = ash(betahat[subset],betahatsd[subset])
  beta.ash.fdr = ash(betahat[subset],betahatsd[subset],method="fdr")
```

### 2015-02-06 scratchwork/checknull.Rmd
added some scratchwork files

Check performance of ash under null
========================================================

Here we simulate all null data sets and look at the 
estimated values of pi0

```{r}
  library(ashr)
  z.ash=list()
  for(i in 1:100){
    set.seed(i)
    z = rnorm(1000) #c(rnorm(950),rnorm(50,0,3))
    z.ash[[i]] = ash(z,1,cxx=FALSE,method="fdr")
    print(i)
    print(get_pi0(z.ash[[i]]))
  }

  nfp = function(a,t=0.05){return(sum(a$lfsr[1:950]<t))} #find number of false positives at threshold t
  mean(unlist(lapply(z.ash,nfp)))
```

### 2015-02-06 scratchwork/makenew.mixVBEM.Rmd
added some scratchwork files

make new mixVBEM
========================================================

First our existing implementation

```{r}
#return the KL-divergence between 2 dirichlet distributions
#p,q are the vectors of dirichlet parameters of same lengths
diriKL = function(p,q){
  p.sum = sum(p)
  q.sum = sum(q)
  k = length(q)
  KL = lgamma(q.sum)-lgamma(p.sum)+sum((q-p)*(digamma(q)-digamma(rep(q.sum,k))))+sum(lgamma(p)-lgamma(q))
  return(KL)
}

#helper function for VBEM
VB.update = function(matrix_lik, pipost){
  avgpipost = matrix(exp(rep(digamma(pipost),n)-rep(digamma(sum(pipost)),k*n)),ncol=k,byrow=TRUE)
  classprob = avgpipost * matrix_lik
  classprob = classprob/rowSums(classprob) # n by k matrix
  B = sum(classprob*log(avgpipost*matrix_lik),na.rm=TRUE) - diriKL(prior,pipost) - sum(classprob*log(classprob)) #negative free energy
  return(list(classprob=classprob,B=B))
}

mixVBEM = function(matrix_lik, prior, post.init=NULL, tol=0.0001, maxiter=5000){
  n=nrow(matrix_lik)
  k=ncol(matrix_lik)
  B = rep(0,maxiter)
  pipost = post.init
  if(is.null(post.init)){
    pipost = prior # Dirichlet posterior on pi
  }
  avgpipost = matrix(exp(rep(digamma(pipost),n)-rep(digamma(sum(pipost)),k*n)),ncol=k,byrow=TRUE)
  classprob = avgpipost * matrix_lik
  classprob = classprob/rowSums(classprob) # n by k matrix  
  B[1] = sum(classprob*log(avgpipost*matrix_lik),na.rm=TRUE) - diriKL(prior,pipost) - sum(classprob*log(classprob))  #negative free energy
  i=1

  if(maxiter>=2){
    for(i in 2:maxiter){  
      pipost = colSums(classprob) + prior

      #Now re-estimate pipost
      avgpipost = matrix(exp(rep(digamma(pipost),n)-rep(digamma(sum(pipost)),k*n)),ncol=k,byrow=TRUE)
      classprob = avgpipost*matrix_lik
      classprob = classprob/rowSums(classprob) # n by k matrix

      B[i] = sum(classprob*log(avgpipost*matrix_lik),na.rm=TRUE) - diriKL(prior,pipost) - sum(classprob*log(classprob))

      if(abs(B[i]-B[i-1])<tol) break;
    } 
    if(i>maxiter){i=maxiter}
  }

  converged=(abs(B[i]-B[i-1])<tol)
  if(!converged){
      warning("EM algorithm in function mixVBEM failed to converge. Results may be unreliable. Try increasing maxiter and rerunning.")
  }

  return(list(pihat = pipost/sum(pipost), B=B[1:i], niter = i, converged=converged,post=pipost))
}

```

Now the squarem version
```{r}
require(SQUAREM)

VBfixpoint = function(pipost, matrix_lik, prior){  
  n=nrow(matrix_lik)
  k=ncol(matrix_lik)
  avgpipost = matrix(exp(rep(digamma(pipost),n)-rep(digamma(sum(pipost)),k*n)),ncol=k,byrow=TRUE)
  classprob = avgpipost*matrix_lik
  classprob = classprob/rowSums(classprob) # n by k matrix
  pipostnew = colSums(classprob) + prior
  return(pipostnew)
}

VBnegpenloglik = function(pipost, matrix_lik, prior){
  n=nrow(matrix_lik)
  k=ncol(matrix_lik)
  avgpipost = matrix(exp(rep(digamma(pipost),n)-rep(digamma(sum(pipost)),k*n)),ncol=k,byrow=TRUE)
  classprob = avgpipost*matrix_lik
  classprob = classprob/rowSums(classprob) # n by k matrix

  B= sum(classprob*log(avgpipost*matrix_lik),na.rm=TRUE) - diriKL(prior,pipost) - sum(classprob*log(classprob)) 
  return(-B)
}

mixVBEM2 = function(matrix_lik, prior, pi.init = NULL,tol=1e-7, maxiter=5000){
  if(is.null(pi.init)){
    pi = rep(1/k,k)# Use as starting point for pi
  } 
  res = squarem(par=pi.init,fixptfn=VBfixpoint, objfn=VBnegpenloglik,matrix_lik=matrix_lik, prior=prior, control=list(maxiter=maxiter,tol=tol))
  return(list(pihat = res$par, B=res$value.objfn, 
              niter = res$iter, converged=res$convergence))
}
```

Now simulate some data
```{r}
  set.seed(100)
  sampsize=100000
  sd = c(1,1.1,1.2)
  z = rnorm(sampsize,0,sample(sd,sampsize,replace=TRUE))
  lik = t(vapply(z,dnorm,sd,sd=sd))
  prior=c(1,1,1)
  pi.init=c(0.3,0.2,0.5)
```

```{r}
  system.time(res1<-mixVBEM(lik,prior,pi.init))
  system.time(res2<-mixVBEM2(lik,prior,pi.init))
  options(digits=13)
  res1
  res2
```

### 2015-02-06 scratchwork/test.gradient.Rmd
added some scratchwork files

Try testing whether gradient of pi0 is negative at pi0=1 (null)
========================================================

The kth element of this vector is the derivative 
of the loglik for $\pi=(\pi_0,...,1-\pi_0,...)$ with respect to $\pi_0$ at $\pi_0=1$.
```{r}
gradient = function(matrix_lik){
  n = nrow(matrix_lik)
  grad = n - colSums(matrix_lik/matrix_lik[,1]) 
  return(grad)
}
```

Simulate some data
```{r}
 set.seed(100)
  sampsize=10000
  sd = c(1,1.1,1.2)
  z0 = rnorm(sampsize,0,1)
  lik0 = t(vapply(z0,dnorm,sd,sd=sd))
  z1 = rnorm(sampsize,0,sample(sd,sampsize,replace=TRUE))
  lik1 = t(vapply(z1,dnorm,sd,sd=sd))

```

```{r}
  gradient(lik0)
  gradient(lik1)
```

```{r}
  for(i in 1:100){
    set.seed(i)
    z0 = rnorm(sampsize,0,1)
  lik0 = t(vapply(z0,dnorm,sd,sd=sd))
    print(i)
    print(gradient(lik0))
  }
```

```{r}
  set.seed(96)
  z0 = rnorm(sampsize,0,1)
  lik0 = t(vapply(z0,dnorm,sd,sd=sd))
  print(gradient(lik0))
  temp=lik0/lik0[,1]

```

### 2015-02-06 Rcode/testnonzeromode.asymmetric.Rmd
added a bunch of files

--- 2014-11-08 Rcode/efron.fcr.rmd
@@
 ---
-title: "FSR in non-unimodal case"
+title: "Test non-zero mode in asymmetric case"
 output: html_document
@@
 This example comes from Efron (2008) p16 when examining the false coverage rate (FCR).
+It is clearly asymmetric. However, the strongest mode is at 0 (there's a point mass there). I'm hoping the nonzeromode option picks that up.
 
@@
   library(ashr)
-  res.ash = ash(z,1,mixcompdist="halfuniform",method="fdr")
-  CImatrix= ashci(res.ash,level=0.95)
+  res.ash = ash(z,1,mixcompdist="halfuniform",method="fdr",nonzeromode=TRUE)
+  res.ash
 ```
-
-Compare results with Bayes rule
-```{r}
-  BayesComparePlot=function(){
-    plot(z,mu,xlim=c(-8,0))
-    points(z[CImatrix[,1]],CImatrix[,5],col="red")
-    points(z[CImatrix[,1]],CImatrix[,4],col="red")
-
-    fdr = 0.9*dnorm(z)/(0.9*dnorm(z)+0.1*dnorm(z,-3,sqrt(2)))
-    o=order(z)
-    upper = ifelse(fdr[o]<0.025,(z[o]-3)/2+qnorm(0.975+fdr[o])/sqrt(2),0)
-    lines(z[o],upper,col="green",lwd=2)
-    lines(z[o],(z[o]-3)/2-qnorm(0.975)/sqrt(2),col="green",lwd=2)
-    abline(v=max(z[fdr<0.05]))
-  }
-  BayesComparePlot()
-```
-
-It seems the lower 97.5% point is pretty accurate.
-
-Try whole thing again to see if it was a fluke
-```{r}
-  set.seed(101)
-  nsamp=10000
-  mu0 = rep(0,nsamp)
-  mu1 = rnorm(nsamp,-3,1)
-  comp = rbinom(nsamp,1,0.1)
-  mu = ifelse(comp==0,mu0,mu1)
-  z = rnorm(nsamp,mu,1)
-```
-
-```{r}
-   res.ash = ash(z,1,mixcompdist="halfuniform",method="fdr")
-  CImatrix= ashci(res.ash,level=0.95)
-  BayesComparePlot()
-```
-
-And try with normal mixture in ash:
-
-```{r}
-   res.ash = ash(z,1,mixcompdist="normal",method="fdr")
-  CImatrix= ashci(res.ash,level=0.95)
-  BayesComparePlot()
-```
-
-And uniform mixture:
-```{r}
-   res.ash = ash(z,1,mixcompdist="uniform",method="fdr")
-  CImatrix= ashci(res.ash,level=0.95)
-  BayesComparePlot()
-```
-
-Things to try:
-- simulation study for coverage?
-- compare with EB FCR methods in JRSS B?

### 2015-02-06 talks/one_hour_seminar.Rmd
added a bunch of files

--- 2014-10-29 talks/UNC2014.Rmd
@@
 author: Matthew Stephens
-date: October 30, 2014
+date: February 3, 2015
 output: ioslides_presentation
@@
 
-## More on git, github, knitr, reproducibility
+## Commercial Break:
+
+A quick ad for a recently-started project I call "Dynamic Statistical Comparisons":
+
+- Overall Goal: Improve how the statistics and genomics (and other) communities make comparisons among methods.
+
+- How? Centralize reproducible statistical comparisons in easily-extensible public repositories.
+
+## Dynamic Statistical Comparisons: advantages
+
+- Authors can run their own methods. 
+- Ease of adding new data can quickly uncover over-tuning to specific datasets.
+- Work is divided, and not duplicated.
+- Promotes collaboration through competition. 
+- Ensures results are reproducible, and extendable.
+
+## More on git, github, knitr, reproducibility, DSCs
 
@@
 - Google "donohoe buckheit" for "Wavelab and reproducible research"
+- [http://stephens999.github.io/blog/](http://stephens999.github.io/blog/) for more on DSC
 
@@
 
-Although precise definitions vary depending on whether one
-takes a Bayesian or Frequentist approach to the problem, roughly
+We take a Bayesian perspective and use the following definitions
+(see Storey, 2003, for connections with Frequentist definitions):
 
@@
 - to the developers of **R**, **knitr**, **Rstudio** and **Pandoc**.
+Also the very cool **SQUAREM** package.
 
@@
 
-- NHGRI for funding.
+- NHGRI, Moore Foundation for funding.
 
@@
 
-## Pandoc Command used
-
-`pandoc -s -S -i --template=my.beamer -t beamer -V theme:CambridgeUS -V colortheme:beaver  slides.md -o slides.pdf`
-
-(alternative to produce html slides; but figures would need reworking)
-`pandoc -s -S -i -t dzslides --mathjax slides.md -o slides.html`
-
-Here is my session info:
-
-```{r session-info}
-print(sessionInfo(), locale=FALSE)
-```
+- `dscr` package: `http://www.github.com/stephens999/dscr`

### 2015-02-06 talks/shiny/shinytest.Rmd
added a bunch of files

--- 2014-10-29 talks/UNC2014.Rmd
@@
 ---
-title: "False Discovery Rates, A New Deal"
-author: Matthew Stephens
-date: October 30, 2014
+title: Interactive illustration of unimodal assumption
+runtime: shiny
 output: ioslides_presentation
@@
 ```{r, include=FALSE}
-require("qvalue")
-require("ashr")
-require("ggplot2")
-require("knitr")
-```
-
-```{r setup, include=FALSE}
-  #set global chunk options
-  opts_chunk$set(cache=TRUE,autodep=TRUE,warning=FALSE)
-  dep_auto()
-```
-
-## Before we get started: Let's get Organized!
-
-- Over ~10 years of working with graduate students + postdocs,
-I've noticed something.
-- Organized researchers get more done (and better!).
-- Many of them are more organized than I am!
-- Thought: I should get organized; I should help others get organized.
-
-## So what can you do?
-
-- Buy a notebook; bring it to meetings; make notes! 
-- Come to meetings with a written agenda. 
-- While doing research, record what you did and what the outcome was.
-- Use version control ([git](http://git-scm.com)) and internet repositories ([bitbucket](http://www.bitbucket.org), [github](http://www.github.com)) to organize notes, code, etc.
-- Use *knitr* to help make your research reproducible.
-- Talk about the tools you find useful!
-
-## What are these repository things?
-
-- A repository: a central place in which an aggregation of data is kept and maintained in an organized way (searcharticle.com)
-- Great for sharing material across multiple people (eg student and advisor!)
-- An amateur example: [http://github.com/stephens999/ash](http://github.com/stephens999/ash)
-
-## What is knitr?
-
-- An R package
-- A tool for literate programming
-- Text, and R code are interleaved
-- When you compile the document, the code is run, and output inserted into the text.
-- Great for writing reports, and keeping a track of what you did and what the result was!
-- This talk was written with knitr (with RStudio)!
-
-## What is Reproducible Research?
-
-- Principle: when publishing results of computational procedures, we should
-publish the code that produced the results.
-- "publishing figures or results without the complete software environment could
-be compared to a mathematician publishing an announcement of a mathematical theorem without giving the proof" (Buckheit and Donohoe)
-- “an article about a computational result is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result.” [Claerbout]
-
-## Why is reproducibility important?
-
-- Not only because people are forgetful, error-prone, or dishonest!
-- Reproducing work is also the first step to extending it.
-- Helps communications among researchers (eg student + advisor).
-- If you do not publish code implementing your methods, your methods will likely go unused.
-
-## More on git, github, knitr, reproducibility
-
-- Google "The git book", to get started on git.
-- Google "Karl Broman github tutorial" for statistics-oriented intro to github.
-- Google "donohoe buckheit" for "Wavelab and reproducible research"
-
-## The Canonical Genomics Experiment
-
-- Measure lots of things, with error
-
-- Get estimates of effects $\beta_j$ ($\hat\beta_j$) and their standard errors $s_j$
-
-- Turn these into Z-scores, $z_j = \hat\beta_j/s_j$
-
-- Turn these into $p$ values, $p_j$
-
-- Apply `qvalue` 
-to identify findings ``significant" at a given FDR.
-
-- ...?
-
-## FDR, local fdr, and q values
-
-Although precise definitions vary depending on whether one
-takes a Bayesian or Frequentist approach to the problem, roughly
-
-- The FDR at a threshold $P$ is 
-$$\text{FDR}(P)=\Pr(\beta_j = 0 |  p_j<P).$$
-
-- The q value for observation $j$ is $q_j=\text{FDR}(p_j)$.
-
-- The local false discovery rate, fdr, at threshold $P$ is 
-$$\text{fdr}(P) = \Pr(\beta_j =0 | p_j=P ).$$
-
-- The fdr is more relevant, but slightly harder to estimate than 
-FDR because it involves density estimation rather than tail-area estimation.
-
-```{r, echo=FALSE}
-  library("ashr") #simple simulated example
-  ncz = 100 ## number of bins in z score histograms
-  set.seed(111)
-  hh.betahat = rnorm(1000,0,2)
-  hh.sebetahat = 1
-  hh.zscore = hh.betahat/hh.sebetahat
-  hh.pval = pchisq(hh.zscore^2,df=1,lower.tail=F)
-  hh.ash = ash(hh.betahat,hh.sebetahat, method="fdr")
-  hh.q = qvalue(hh.pval)
-```
-
-```{r, echo=FALSE}
-  plot_FDReg_hist = function(hh.pval,pi0,nc=40,nullcol="blue",altcol="cyan",type=4,title="Distribution of p values",...){
-    hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main=title,nclass=nc,col=altcol,...)
-    if(type>1){
-      abline(h=pi0,col=nullcol,lwd=2)
-
-      hh.hist$density=rep(pi0,length(hh.hist$density))  
-      #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
-      plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)
-    }
-    if(type>2){
-    abline(v=0.1,lwd=2,col=2)
-    }
-    if(type>3){
-      text(0.05,1.2,labels="A",col=2,cex=1.2)  
-      text(0.05,0.4,labels="B",col=2,cex=1.2)  
-      text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
-    }
-  }
-```
-
-## Example: FDR estimation
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=1)                 
-```
-
-## Example: FDR estimation
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=2)                 
-```
-
-## Example: FDR estimation
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=3)                 
-```
-
-## Example: FDR estimation
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4)                 
-```
-
-## Is this an important problem?
-
-- The original paper introducing FDR (Benjamini and Hochberg, 1995) has
-been cited 21,787 times (May 2014) according to Google Scholar.
-
-- That is three times a day for the last 19 years!
-
-## Problem 1: The Zero Assumption (ZA)
-
-- The standard `qvalue` 
-approach assumes that all the $p$ values near 1 are null.
-
-- Analogously, one can assume that all Z scores near 0 are null. Efron refers to this as the ``Zero Assumption".
-
-- Seems initially natural.
-
-```{r, echo=FALSE, include=FALSE}
-require(fdrtool)
-  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue",plot=FALSE)
-require(locfdr)
-  hh.locfdr = locfdr(hh.zscore,nulltype=0,plot=0)
-require(mixfdr)
-  hh.mixfdr = mixFdr(hh.zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
-```
-
-## Implied distribution of $p$ values under $H_1$
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4)                 
-```
-
-## Implied distribution of Z scores under alternative
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
  #plot a histogram of z scores, highlighting the alternative distribution
-#of z scores that is implied by localfdr values lfdr.
-  nullalthist = function(z,lfdr,nullcol="blue",altcol="cyan",...){
-    h=hist(z, freq=FALSE,col=nullcol,nclass=ncz,...)
-    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
-    h$density = (1-avlfdr) * h$density
+#of z scores that is implied by pi0
+nullalthist_z = function(z,pi0,nullcol="blue",altcol="cyan",...){
+  h=hist(z, freq=FALSE,col=nullcol,nclass=40,ylim=c(-0.5,0.5),...)
+  nb = length(h$breaks)
+  nullexp = pi0 * (pnorm(h$breaks)[-1] - pnorm(h$breaks[-nb]))/(h$breaks[-1]-h$breaks[-nb])
+  h$density = h$density - nullexp
   plot(h,add=TRUE,col=altcol,freq=FALSE)
@@
 
-  nullalthist(hh.zscore,hh.fdrtool$lfdr)  
+ set.seed(100)
+z = c(rnorm(5000,0,2),rnorm(5000,0,1))
 ```
 
-## FDR problem 2: different measurement precision
+## Illustration of unimodal assumption
+Increase the proportion of nulls until the alternative (cyan) looks unimodal.
+```{r, echo=FALSE}
 
-- In some cases the measurement precisions differ among units
+inputPanel(  
+  sliderInput("pi0", label = "Proportion of nulls:",
+              min = 0, max = 1, value = 0, step = 0.1)
+)
 
-- Eg Expression levels of low-expressed genes have less precision than high-expressed genes
-
-- If some effects are measured less precisely than others, those tests ``lack power"
-and dilute signal, increasing FDR
-
-## Example: Mouse Heart Data
-
-```{r, echo=FALSE}
-x = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
-xx = rowSums(x[,2:5])
-x = x[xx>0,]
-xx = xx[xx>0]
+renderPlot({
+  nullalthist_z(z, input$pi0)
+})
 ```
-
-```{r echo=FALSE}
-head(x)
-```
-
-- Data on 150 mouse hearts, dissected into left and right ventricle
-(courtesy Scott Schmemo, Marcelo Nobrega)
-
-```{r, echo=FALSE,include=FALSE}
-
-load("../scratchwork/Smemo_qbinom_zdat.RData")
-## two-sided test
-  ttest.pval = function(t, df) {
-    pval = pt(t, df = df, lower.tail = T)
-    pval = 2 * ifelse(pval < 0.5, pval, 1 - pval)
-    return(pval)
-}
-
-tscore = zdat[3, ]/zdat[4, ]
-pval = ttest.pval(tscore, df = 2)
-qval = qvalue(pval)
-
-highxp = xx>1000 ## select high expressed genes
-pval.high = pval[highxp]
-qval.high = qvalue(pval.high)
-
-zdat.ash = ash(zdat[3,],zdat[4,],df=2,method="fdr",mixcompdist="halfuniform")
-zdat.ash.high = ash(zdat[3,highxp],zdat[4,highxp],df=2,method="fdr",mixcompdist="halfuniform")
-zdat.ash.ES = ash(zdat[3,],zdat[4,],df=2,method="fdr",mixcompdist="halfuniform",model="ES")
-
-```
-
-## Example: Mouse Heart Data
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-qval= qvalue(pval)
-plot_FDReg_hist(pval,pi0=qval$pi0,ylim=c(0,4))
-```
-
-## Mouse Data: Counts vary considerably across genes
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-hist(log10(xx),main="Distribution of total counts", xlab="log10(counts)")
-```
-
-## Lower count genes, less power
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-qval.low=qvalue(pval[!highxp])
-pval.low=pval[!highxp]
-plot_FDReg_hist(pval.low,pi0=qval.low$pi0,ylim=c(0,4))
-```
-
-## Higher count genes, more power
-```{r, echo=FALSE, fig.height=4,fig.cap=""}
-plot_FDReg_hist(pval.high,pi0=qval.high$pi0,ylim=c(0,4))
-```
-
-## Low-count genes dilute signal at high-count genes
-```{r, echo=FALSE, fig.height=4,fig.cap=""}
-plot_FDReg_hist(pval,pi0=qval$pi0,ylim=c(0,4))
-```
-
-## FDR problem 2: low count genes add noise, increase q values
-```{r,echo=FALSE, fig.height=4,fig.cap=""}
-plot(qval.high$qval,qval$qval[highxp],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
-abline(a=0,b=1,col=2)
-```
-
-## Problems: Summary
-
-Standard tools are unduly conservative.
-
-- The ZA, which implies actual effects have a (probably unrealistic) bimodal distribution; causes overestimate of $\pi_0$, losing power.
-
-- By focussing on $p$ values, low-precision measurements can dilute high-precision measurements.
-
-## FDR via Empirical Bayes
-
-- Following previous work (e.g. Newton, Efron, Muralidharan) we take an empirical Bayes approach to FDR.
-
-- Eg Efron assumes that the $Z$ scores come from a mixture of null, and alternative:
-$$Z_j \sim f_Z(.) = \pi_0 N(.;0,1) + (1-\pi_0) f_1(.)$$
-where $f_1, \pi_0$ are to be estimated from the data.
-
-- Various semi-parametric approaches taken to estimating $f_1$. For example,
-Efron uses Poisson regression; Muralidharan uses mixture of normal distributions.
-
-- Once $f_1$ and $\pi_0$ estimated, FDR calculations are straightforward.
-
-## FDR: A New Deal
-
-- Instead of modelling $Z$ scores, model the effects $\beta$,
-$$\beta_j \sim \pi_0 \delta_0(.) + (1-\pi_0) g(.)$$
-
-- Constrain $g$ to be unimodal about 0; estimate $g$ from data.
-
-- *Incorporate precision* of each observation $\hat\beta$ into the likelihood.
-Specifically, approximate likelihood for $\beta_j$ by a normal: 
-$$L(\beta_j) \propto \exp(-0.5 (\beta_j - \hat\beta_j)^2/s_j^2).$$
-[From $\hat\beta_j \sim N(\beta_j, s_j)$]
-Or, better, use a $t$ likelihood if $s_j$ estimated using few observations.
-
-## FDR - A New Deal
-
-- A convenient way to model $g$: mixture of 0-centered
-normal distributions: 
-$$g(\beta; \pi) = \sum_{k=1}^K \pi_k N(\beta; 0, \sigma^2_k)$$
-
-- Estimating $g$ comes down to estimating $\pi$. Joint estimation of $\pi_0,\pi$ easy by maximum likelihood (EM algorithm).
-
-- By allowing $K$ large, and $\sigma_k$ to span a dense grid of values,
-we get a flexible unimodal symmetric distribution.
-
-- Can approximate, arbitrarily closely, any scale mixture of normals.
-Includes almost all priors used for sparse regression problems (spike-and-slab, double exponential/Laplace/Bayesian Lasso, horseshoe).
-
-## FDR - A New Deal
-
-- Alternatively, a mixture of uniforms, with 0 as one end-point of the range,
-provides still more flexibility, and in particular allows for asymmetry.
-
-- If allow a very large number of uniforms this provides the non-parametric mle for $g$; cf Grenander 1953; Cordy + Thomas 1997.
-
-## Illustration: $g$ a mixture of 0-centered normals
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-x=seq(-4,4,length=100)
-plot(x, dnorm(x,0,1),type="l",ylim=c(0,2),ylab="density")
-lines(x, dnorm(x,0,0.1))
-lines(x, dnorm(x,0,0.5))
-```
-
-## Illustration: $g$ a mixture of 0-centered normals
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-x=seq(-4,4,length=100)
-plot(x, 0.5*dnorm(x,0,1)+0.5*dnorm(x,0,0.1),type="l",ylim=c(0,2),ylab="density")
-```
-
-## Illustration: $g$ a mixture of 0-anchored uniforms
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-x=seq(-4,4,length=100)
-plot(x, dunif(x,0,1),type="s",ylim=c(0,5),ylab="density")
-lines(x, dunif(x,0,0.2),type="s")
-lines(x, dunif(x,0,0.5),type="s")
-lines(x, dunif(x,-0.3,0),type="s",col=2)
-lines(x, dunif(x,-0.4,0),type="s",col=2)
-lines(x, dunif(x,-2,0),type="s",col=2)
-```
-
-## Illustration: $g$ a mixture of 0-anchored uniforms
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-x=seq(-4,4,length=100)
-plot(x, 0.1*dunif(x,0,3)+ 0.3*dunif(x,0,0.2) + 0.2*dunif(x,0,0.5) + 0.1* dunif(x,-0.3,0) + 0.1*dunif(x,-0.3,0)+0.2*dunif(x,-0.4,0),type="s",ylim=c(0,2),ylab="density")
-```
-
-## Adaptive Shrinkage
-
-- This approach actually provides a full posterior distribution for each $\beta_j$.
-
-- So easy to obtain point estimates and credible intervals.
-
-- Because $g(\beta)$ is unimodal, the point estimates (and CIs) will tend to be ``shrunk" towards the overall mean (0).
-
-- Because $g(\beta)$ is estimated from the data, the amount
-of shrinkage is adaptive to signal in the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each observation.
-
-- So we call the approach ``Adaptive Shrinkage" (ASH).
-
-## Recall Problem 1: distribution of alternative Z values multimodal
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-nullalthist(hh.zscore,hh.fdrtool$lfdr,main="qvalue")  
-```
-
-## Problem Fixed: distribution of alternative Z values unimodal
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-nullalthist(hh.zscore,hh.ash$lfdr,main="ash")
-```
-
-## Example: FDR estimation
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4,title="qvalue")                 
-```
-
-## Example: FDR estimation
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,get_pi0(hh.ash),type=4,title="ash")                 
-```
-
-## Recall Problem 2: low count genes add noise, increase q values
-```{r,echo=FALSE, fig.height=4,fig.cap=""}
-plot(qval.high$qval,qval$qval[highxp],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
-abline(a=0,b=1,col=2)
-```
-
-## Problem Fixed: incorporating precision reduces influence of low-count genes
-```{r,echo=FALSE, fig.height=4,fig.cap=""}
-plot(zdat.ash.high$qvalue,zdat.ash$qvalue[highxp],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
-abline(a=0,b=1,col=2)
-```
-
-## A new problem: an embarrassment of riches
-
-- If the null is mostly false, the new approach can  provide unsettling results
-
-- The FDR can be small for all observations, even those with $p \approx 1$!
-
-- In the illustrative example, the maximum $q$ value is `r round(max(hh.ash$qvalue),2)`
-
-## Perhaps we didn't really understand the question?
-
-- Problem arises only if we insist on asking question ``is $\beta_j=0$?"
-
-- Given enough signal, we become convinced that very few of the $\beta_j=0$
-
-- But for some $\beta_j$ we still may have little information about actual value
-
-- Suggests a change of focus: ask for which $\beta_j$ are we confident about the sign (cf Gelman et al, 2012).
-
-## The False Sign Rate
-
-- Suggestion: replace FDR with local false sign rate (lfsr), the probability that if we say an effect is positive (negative), it is not.
-
-- Example: suppose we estimate that $\Pr(\beta_j<0)=0.95, \Pr(\beta_j=0) = 0.025$ and $\Pr(\beta_j>0)=0.025$. Then we report $\beta_j$ as a ``(negative) discovery", and estimate its lfsr as 0.05.
-
-## Even with many signals, large $p$ values have high lfsr
-
-```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-  #res=data.frame(p=hh.pval, lfsr=hh.ash$lfsr, qvalue = hh.ash$qvalue)
-  library("reshape2")
-  res=data.frame(p=pval, lfsr=zdat.ash$lfsr, qvalue = zdat.ash$qvalue)
-  res.melt = melt(res, id.vars=c("p"),variable.name="Measure")
-  cbbPalette <- c("#000000", "#D55E00", "#CC79A7")
-  labels = c('qvalue','lfsr')
-  breaks = c("qvalue","lfsr")
-
-  pp= ggplot(data=res.melt,aes(p,value,color=Measure)) +geom_point(shape=1) +
-      geom_abline(colour = "black") +
-        xlab("p-value") +
-        ylab("lfsr/qvalue")
-
-  print(pp +scale_y_continuous(limits=c(0,1)) +
-          scale_x_continuous(limits=c(0,1))  +
-           scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
-          coord_equal(ratio=1))
-
-```
-
-## Advantages of False Sign Rate
-
-- The FDR is typically not actually identifiable from data.
-
-- This is because data cannot distinguish between $\beta_j = 0$ and $\beta_j$ "very
-small". So $\pi_0$ is not identifiable, and FDR is very sensitive to $\pi_0$.
-
-- So methods for estimating $\pi_0$ and FDR, including those presented here, are designed to be ``conservative" (i.e. overestimate the FDR).
-
-- The False Sign Rate is much less senstive to $\pi_0$, and hence more identifiable from data!
-
-## Simulated example: $\pi_0$ not identifiable.
-
-```{r,echo=FALSE,include=FALSE}
-load("../paper/Rcode/sim1.RData") #load simulation results
-source("../paper/Rcode/plot_pi0.R")
-```
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-plot_pi0(list(simres1a))
-```
-
-## Simulated Example: so fdr not identifiable
-
-```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-source("../paper/Rcode/plot_lfsr.R")
-  plot_lfsr(list(simres1a),0.1,ptype="lfdr")
-```
-
-## Simulated Example: fsr much more identifiable
-
-```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-  plot_lfsr(list(simres1a),0.1,ptype="lfsr")
-```
-
-## Summary
-
-- ASH provides a generic approach to shrinkage estimation, as well as
-false discovery (sign) rates.
-
-- But by using two numbers ($\hat\beta,s$) instead of one ($p$ values or $z$ scores) varying precision of 
-measurements is better accounted for.
-
-- Unimodal assumption for effects reduces conservatism
-
-- False Sign Rate preferable to False Discovery Rate: more identifiable, and
-better representation of information in data for ``high-signal" situations.
-
-## Other Applications
-
-- Widely applicable: requiring only an estimated
-effect size and standard error for each object.
-
-- E.g. Currently applying it to wavelet shrinkage applications.
-
-## Next steps?
-
-- Incorporate shrinkage of variances and not just means. (e.g. ``moderated $t$ test", Smyth et al)
-
-- Allow $g(\cdot;\pi)$ to depend on covariates $X$.
-
-- Allow for correlations in the measured $\hat\beta_j$.
-
-## Thanks
-
-- to the developers of **R**, **knitr**, **Rstudio** and **Pandoc**.
-
-- to the several postdoctoral researchers and students
-who have worked with me on related topics.
-
-- Including  Chaoxing Dai, Mengyin Lu, Ester Pantaleo, Scott Powers, Sen Tian, Wei Wang, Zhengrong Xing.
-
-- NHGRI for funding.
-
-- `ashr` package: `http://www.github.com/stephens999/ash`
-
-## Pandoc Command used
-
-`pandoc -s -S -i --template=my.beamer -t beamer -V theme:CambridgeUS -V colortheme:beaver  slides.md -o slides.pdf`
-
-(alternative to produce html slides; but figures would need reworking)
-`pandoc -s -S -i -t dzslides --mathjax slides.md -o slides.html`
-
-Here is my session info:
-
-```{r session-info}
-print(sessionInfo(), locale=FALSE)
-```

### 2015-02-19 tests/test_nonparam.Rmd
introduced ability to specify non-parametric g

---
title: "test nonparam"
output: html_document
---

```{r}
#Following tests nonparam estimation
library("ashr")
set.seed(100)
# First, try fixing g to a true g (but not unimodal)
beta = c(rnorm(1000,0,1),rnorm(1000,5,1))
betahatsd = 1
betahat = rnorm(length(beta),mean=beta,sd=betahatsd)
g = normalmix(c(0.5,0.5),c(0,5),c(1,1))
res.1=ash(betahat,betahatsd,g=g,control=list(maxiter=0)) # use the mixture components in g, and also the mixture proportions since don't iterate!
res.2=ash(betahat,betahatsd,g=g) # use the mixture components in g, but iterate!
```

Now try non-parametric version by setting guni to be piecewise uniform
```{r}
K=100
bins=seq(-10,10,length=K+1)
binc = (bins[-1]+bins[-(K+1)])/2 # bincenters
guni = unimix(rep(1,K)/K, bins[-(K+1)],bins[-1])

res.nonp=ash(betahat,betahatsd,g=guni) # use the mixture components in g, but iterate!
res.nonp2=ash(betahat,betahatsd,g=guni,randomstart=TRUE)

plot(binc,get_density(res.1,binc)$y,type="l")
lines(binc,get_density(res.2,binc)$y,col=2)
lines(binc,get_density(res.nonp,binc)$y,type="s",col=3)
lines(binc,get_density(res.nonp2,binc)$y,type="s",col=4)
```

Check likelihoods
```{r}
get_loglik(res.1)
get_loglik(res.2)
get_loglik(res.nonp)
get_loglik(res.nonp2)
```

Now try VB to avoid overfitting?
```{r}
res.VB=ash(betahat,betahatsd,g=guni,randomstart=TRUE,VB = TRUE,prior=rep(1,K)) 
res.VB2=ash(betahat,betahatsd,g=guni,randomstart=TRUE,VB = TRUE,prior=rep(1/2,K)) 
res.VB3=ash(betahat,betahatsd,g=guni,randomstart=TRUE,VB = TRUE,prior=rep(1/K,K))

plot(binc,get_density(res.1,binc)$y,type="l")
lines(binc,get_density(res.VB,binc)$y,col=2)
lines(binc,get_density(res.VB2,binc)$y,col=3)
lines(binc,get_density(res.VB3,binc)$y,col=4)
```

compare evidence
```{r}
  get_loglik(res.VB)
  get_loglik(res.VB2)
  get_loglik(res.VB3)
```

### 2015-02-19 paper/Rcode/simABC.R

@@ -36,7 +36,7 @@ ashsim=function(mixmean,mixsd,mixpi,bsd=1,seedval = 100,nsamp=1000,niter=50){
     fit.ash.u[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="uniform", method="shrink")
     fit.ash.hu[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform", method="shrink")
     fit.ash.hu.vb[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform",method="shrink",VB=TRUE)
-    fit.ash.true[[i]] = ash(betahat[[i]],betahatsd[[i]],g=normalmix(mixpi,mixmean,mixsd))
+    fit.ash.true[[i]] = ash(betahat[[i]],betahatsd[[i]],g=normalmix(mixpi,mixmean,mixsd),control=list(maxiter=0))
     fit.mixfdr[[i]] = mixFdr(betahat[[i]]/betahatsd[[i]],noiseSD=1,theonull=TRUE,plot=FALSE)
     
     fit.ash.fdr.n[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="normal",method="fdr")

### 2015-02-20 tests/test_nonparam.Rmd
update example

@@
 Now try VB to avoid overfitting?
-```{r}
+```{r cache=TRUE}
 res.VB=ash(betahat,betahatsd,g=guni,randomstart=TRUE,VB = TRUE,prior=rep(1,K)) 
@@
 ```
+
+Try same thing with less noise
+```{r}
+set.seed(100)
+# First, try fixing g to a true g (but not unimodal)
+beta = c(rnorm(1000,0,1),rnorm(1000,5,1))
+betahatsd = 0.05
+betahat = rnorm(length(beta),mean=beta,sd=betahatsd)
+g = normalmix(c(0.5,0.5),c(0,5),c(1,1))
+res.1=ash(betahat,betahatsd,g=g,control=list(maxiter=0)) # use the mixture components in g, and also the mixture proportions since don't iterate!
+res.2=ash(betahat,betahatsd,g=g) # use the mixture components in g, but iterate!
+```
+
+Now try non-parametric version by setting guni to be piecewise uniform
+```{r}
+K=100
+bins=seq(-10,10,length=K+1)
+binc = (bins[-1]+bins[-(K+1)])/2 # bincenters
+guni = unimix(rep(1,K)/K, bins[-(K+1)],bins[-1])
+
+res.nonp=ash(betahat,betahatsd,g=guni) # use the mixture components in g, but iterate!
+res.nonp2=ash(betahat,betahatsd,g=guni,randomstart=TRUE)
+
+plot(binc,get_density(res.1,binc)$y,type="l")
+lines(binc,get_density(res.2,binc)$y,col=2)
+lines(binc,get_density(res.nonp,binc)$y,type="s",col=3)
+lines(binc,get_density(res.nonp2,binc)$y,type="s",col=4)
+```

### 2015-02-26 examples/qval.nonp.rmd
updated paper

@@
   library(mixfdr)
+  library(knitr)
 ```

### 2015-02-26 paper/main.tex
added shiny graphic for showing boxplot of dsc-shrink results
updated paper

@@
 \cite{benjamini1995controlling}, the ``False Discovery Rate" (FDR) has quickly established itself
-as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing.
+as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing
+in which the goal is to identify the non-zero ``effects" among a large number of imprecisely-measured effects.
+
+Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
+EB approaches to this problem have been quite extensively studied by several authors, 
+especially B.~Efron and co-authors 
+\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
+see also \cite{kendziorski2003parametric,newton2004detecting, 
+datta2005empirical,muralidharan2010empirical} for example.
+So what is the ``New Deal" here? We introduce two simple new ideas that can substantially improve
+inference. The first idea is to {\it assume that the distribution of the non-zero effects is unimodal}, and to perform
+non-parametric inference under this assumption. This idea has the twin benefits of both simplifying and
+improving inference, at least when the assumption is correct. And even when the assumption is incorrect, we show
+that the approach can be relatively robust to this. The second idea is to use two numbers (effect sizes, and their standard errors), rather than just one ($p$ values, or $z$ scores)
+ to summarize each measurement. This idea allows any differences in the precision of different measurements to be taken into account,
+ and in particular avoids the problem that poor-precision measurements can contaminate high-precision measurements, resulting in inflated FDR values.
+
+In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
+the idea that it may be preferable to focus on estimation rather than on testing.
+In principle Bayesian approaches, including EB, can naturally unify testing and estimation into a single framework -- testing is
+simply estimation but with a point mass of some weight on the null hypothesis that an effect is exactly zero.
+However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
+perspectives, in practice, large-scale studies that assess many effects almost invariably focus on testing significance, and
+controlling the FDR, rather than on estimating the sizes of the effect. 
+The two new ideas highlighted above 
+help facilitate estimation as well as testing, and by providing simple software to implement our methods
+we hope to promote and accelerate the shift in this direction.
 
@@
 
- \subsection*{Notation, and existing methods}
+ \subsection*{Notation}
 
@@
 In some contexts our interest may focus on which of the $\beta_j$ are ``significantly" different from zero, whereas in other contexts our interest may focus on estimating their values; the methods described here are suited to both these contexts.
- We assume that we have obtained data $D_1,\dots,D_J$ that provide independent estimates $\bhat_1,\dots,\bhat_J$ of these effects,
- with corresponding (estimated) standard errors $\shat_1,\dots,\shat_J$. For our purposes, these estimates and standard errors could be obtained from the data
- using any method, provided they (approximately) satisfy 
- \begin{equation} \label{eqn:betahat-t}
- \bhat_j \big{|} \, \beta_j, \shat_j \sim t_\nu(\beta_j,\shat_j),
+ We assume that we have obtained data $D=(D_1,\dots,D_J)$ that provide independent estimates $\bhat_1,\dots,\bhat_J$ of these effects,
+ with corresponding (estimated) standard errors $\shat_1,\dots,\shat_J$. Let $H_j$ denote the event that the $j$th null hypothesis holds, and $t_j$ denote
+ the $t$ statistic testing this null:
+ \begin{gather}
+ H_j: \beta_j = 0 \\  
+ t_j :=\bhat_j/\shat_j. 
+ \end{gather}
+ The $t$ statistics, of course, have a $t$ distribution under the null:
+ \begin{equation} \label{eqn:tstat}
+ t_j | H_j \sim T_\nu
  \end{equation}
-for some known degrees of freedom (\df) $\nu$, where $t_\nu(\mu,\sigma)$ denotes the generalized $t$ distribution on $\nu$ \df, with mean $\mu$ and scale parameter $\sigma$.
+ where $T_\nu$ denotes the $t$ distribution on $\nu$ \df, with $\nu$ assumed known. 
+ Let $p_j$ denote the $p$ value obtained by comparing $t_j$ with its null distribution,
+ \begin{equation}
+ p_j := \Pr(|T_\nu| > |t_j| ).
+ \end{equation}
+Some methods for FDR analysis require $z$ scores instead of $t$ statistics or $p$ values;
+ these methods can be applied by translating each $t_j$ into a corresponding
+$z$ score, which we denote $z_j$.
+
+ %For our purposes, these estimates and standard errors could be obtained from the data
+%using any method, provided they (approximately) satisfy 
+% \begin{equation} \label{eqn:betahat-t}
+% \bhat_j \big{|} \, \beta_j, \shat_j \sim t_\nu(\beta_j,\shat_j),
+% \end{equation}
+%for some known degrees of freedom (\df) $\nu$, where $t_\nu(\mu,\sigma)$ denotes the generalized $t$ distribution on $\nu$ \df, with mean $\mu$ and scale parameter $\sigma$.
 (Equivalently, $(\bhat_j-\beta_j)/\shat_j \big{|} \, \beta_j \sim t_\nu$, where $t_\nu$ denotes the standard $t$ distribution on $\nu$ \df.)
-In the special case $\nu \rightarrow \infty$ this includes the situation where the $\bhat$ can be considered normally distributed:
-  \begin{equation} \label{eqn:betahat-norm}
- \bhat_j  \big{|} \, \beta_j, \shat_j \sim N(\beta_j,\shat^2_j).
+%In the special case $\nu \rightarrow \infty$ this includes the situation where the $\bhat$ can be considered normally distributed:
+%  \begin{equation} \label{eqn:betahat-norm}
+% \bhat_j  \big{|} \, \beta_j, \shat_j \sim N(\beta_j,\shat^2_j).
+% \end{equation}
+% Assuming (\ref{eqn:betahat-t}), the $t$-statistics defined by
+
+Let $\Gamma$ denote a subset of the tests $1,\dots,J$ declared ``significant" by some procedure. 
+We define the FDR for $\Gamma$ as the proportion of these tests that are ``false discoveries" in that $H_j$ is true:
+\begin{equation} \label{eqn:FDR}
+\FDR(\Gamma) := \frac{\#\{j: H_j \cap j \in \Gamma\}}{\#\{j: j \in \Gamma\}},
 \end{equation}
- Assuming (\ref{eqn:betahat-t}), the $t$-statistics defined by
+taken to be 0 if both numerator and denominator are 0. Frequentist approaches to controlling the FDR attempt to control
+the expectation of $\FDR(\Gamma)$ where the expecation is taken over the sampling distribution of $\Gamma$ (which is a function of the procedure
+used, considered fixed, and the data $D$, considered random). Bayesian approaches, in contrast, condition on the observed data, and compute
+the posterior distribution for the $H_j$ given the data, which induces a posterior distribution on $\FDR(\Gamma)$. This posterior
+distribution can be used to obtain a point estimate for $\FDR(\Gamma)$, for example using the posterior mean. 
+To be more explicit, Bayesian (and EB) approaches give the posterior probability that $H_j$ holds, which
+analagous to \cite{efron2008microarrays} we refer to as the ``local FDR", and denote $\fdr_j$:
+%\begin{equation} \label{eqn:FDR}
+%\FDR(T) := \Pr(\text{$H_j$ false} | t_j \geq T).
+%\end{equation}
+\begin{equation} \label{eqn:fdr}
+\fdr_j := \Pr(H_j | D).
+\end{equation}
+The FDR can be estimated by the posterior mean for $\FDR(\Gamma)$, given by
 \begin{equation} 
- t_j:=\bhat_j/\shat_j
+\widehat\FDR(\Gamma) := \frac{\sum_{j \in \Gamma} \fdr_j }{\#\{j: j \in \Gamma\}}.
 \end{equation}
- have a $t_\nu$ distribution under the $j$th null hypothesis $H_j:\beta_j=0$, and this fact can be used to compute a $p$ value $p_j$ testing $H_j$.
-Some methods for FDR analysis require $z$ scores instead of $t$ statistics or $p$ values; these methods can be applied by translating each $t_j$ into a corresponding
-$z$ score, $z_j$.
 
-One nice feature of the False Discovery Rate (FDR) is that Frequentist and Bayesian definitions are closely aligned.
-Here we provide Bayesian definitions, and refer the reader to \cite{storey.03}, particularly his Theorem 1,  for connections with frequentist definitions.
-The (tail) FDR at threshold $T$ is defined as
+The $q$ value for observation $j$ is defined \cite{storey.03} as, roughly, the FDR for the set of observations that are at least as significant as observation $j$.
+In a Bayesian analysis it is natural to order the significance of the hypotheses by $\fdr_j$, so we define
 \begin{equation}
-\FDR(T) := \Pr(\text{$H_j$ false} | t_j \geq T).
+q_j := \FDR(\{k:\fdr_k \leq \fdr_j\}).
 \end{equation}
-The local FDR at $t$ is defined \cite{efron2008microarrays} as
+For existing EB methods (but not for our method), typically $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
 \begin{equation}
-\fdr(t) := \Pr(\text{$H_j$ is false} | t_j = t).
+q_j = \FDR(\{k:p_k \leq p_j\})
 \end{equation}
-The local FDR is perhaps more relevant than the tail FDR, but less widely used, perhaps because
-it is harder to estimate (density estimation is harder than tail area estimation). 
-The $q$ value for observation $j$ is defined \cite{storey.03} as:
-\begin{equation}
-q_j := \FDR(t_j).
-\end{equation}
-That is, the $q$ value for observation $j$ is the FDR if we reject the null for all observations that are at least as significant as observation $j$.
+which is close to the definition in \cite{storey.03}.
 
-Mention correspondance between fdr and FDR?
-Tidy to deal with fact that definitions I give here implicitly assume $\fdr(t)$ etc don't depend on $j$.
-Mention that Frequentist and Bayesian alignment depends on assumptions?
+One nice feature of FDR procedures is that Frequentist and Bayesian procedures are often closely aligned, at least under certain assumptions; see
+for example \cite{storey.03}, particularly his Theorem 1.
 
-One standard genomics analysis pipeline is to take the $p$ values $p_1,\dots,p_J$ and convert them to (estimated) $q$ values $q_1,\dots,q_J$
- using the R package \qvalue. Figure \ref{fig:qvalue} shows a graphical illustration of the procedure implemented in \qvalue. 
+\subsection{Existing methods}
+
+\subsubsection{qvalue}
+
+One standard genomics analysis pipeline is to take the $p$ values $p_1,\dots,p_J$ and use them to estimate $q$ values $q_1,\dots,q_J$
+ using the R package \qvalue.
+  Figure \ref{fig:qvalue} shows a graphical illustration of the procedure implemented in \qvalue. 
  In brief, the idea is that the overall distribution of $p$ values can be decomposed into two constitutive
@@
 
- Of course \qvalue is not the only approach to estimating False Discovery Rates. Indeed, there is such a large
- literature on FDR methods that we do not attempt a comprehensive review. Instead we highlight the two methods that
- seem most relevant to our work here. The first is due to \cite{efron.xx}, implemented in the R package \locfdr.
+\subsubsection{EB approaches: \locfdr and \mixfdr}
+
+The fundamental
+ idea of EB procedures is similar to that of \qvalue: to attempt to partition the observations into null and alternative components.
+Here we highlight two EB methods that
+ seem most relevant to our work here and which had good software implementations at the time of writing. 
+ The first is due to \cite{efron2008microarrays}, implemented in the R package \locfdr.
  This method differs from \qvalue in that, instead of working with $p$ values, it works with $z$ scores.
- Analagous to \qvalue, \locfdr attempts to decompose the overall distribution of $z$ scores into two constitutive components,
- one corresponding to the null tests, and the other corresponding to the alternative tests. 
+ %Analagous to \qvalue, \locfdr attempts to decompose the overall distribution of $z$ scores into two constitutive components,
+ %one corresponding to the null tests, and the other corresponding to the alternative tests. 
  In the simplest case
  \locfdr  makes assumptions analogous to the two key assumptions of \qvalue: i) the null $z$ scores have a $N(0,1)$ distribution,
- and ii) the $z$ scores near 0 are from the null, which Efron
-calls the ``Zero Assumption" (ZA). Under these assumptions, \locfdr can decompose the
-overall distribution of $z$ scores into a null and alternative component, and estimate the local FDR for each observation. 
- (An important focus of Efron's work \cite{}, also implemented in \locfdr, 
+ and ii) all the $z$ scores near 0 are from the null, which Efron
+calls the ``Zero Assumption" (ZA).  
+The \locfdr package first estimates the overall distribution of $z$ scores using a non-parametric regression approach, and then, using the two assumptions,  ``subtracts out" the
+null component, leaving the remainder to be the alternative component. In this way \locfdr, like \qvalue, avoids explicit modelling of the alternative component.
+ (An important focus of Efron's work, also implemented in \locfdr, 
  is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
@@
 
-Another method for FDR estimation, and perhaps the one most closely related to our
-work here, is due to \cite{muralidharan.xx}, implemented in the R package \mixfdr.
+Another EB method, and perhaps the one most closely related to our
+work here, is due to \cite{muralidharan2010empirical}, implemented in the R package \mixfdr.
 Like \locfdr, \mixfdr works with $z$ scores, but it differs in that it {\it explicitly} models the alternative
-distribution of $z$ scores (using a mixture of normal distributions). In principle this approach also differs from \qvalue and \locfdr in
+distribution of $z$ scores, using a mixture of normal distributions. In principle this approach also differs from \qvalue and \locfdr in
 that it does not explicitly make the Zero Assumption. However, we have found that in practice the results often approximately satisfy the ZA,
@@
 
-Figure \ref{fig:qvalue} illustrates these three methods in a simple simulated setting.
-
-Note that incorporating differential errors messes up the EM algorithm?
-
-Efron models the overall distribution of $z$ scores using a non-parametric regression approach, and then ``subtracts out" the
-null component, leaving the remainder to be the alternative component.
-
-we start by noting that the $p$ values in Figure \ref{} were actually generated under a scenario where
-{\it no tests were null}! Thus the qvalue procedure, which estimates the proportion of nulls to be xx\%, is acting very conservatively in this case.
-However, it could be argued that some degree of conservativeness is inevitable, and that nonetheless the procedure is acting sensibly.
-Therefore, to illustrate this point we really want to focus on the fact that, at least in some settings, the assumption is {\it unrealistic}.
-Specifically, we want to argue that the implied distribution of the alternative $p$ values is unrealistic. In fact, this is hard to see in $p$ value space,
-and so instead we translate these $p$ values back to the $z$ scores that lead to these $p$ values.
-
-The second issue is that directly modeling the p values, or z scores, say via non-parametric methods, 
-can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-It should be noted that existing methods were perhaps developed with the expectation that they would typically 
-be applied in settings where the vast majority of all tests are null, and under such circumstances
-the problems caused by this assumption are less severe. However, it seems desirable to have methods that
-also perform well in settings where  this assumptions does not hold, and this motivates the methods developed here.
-
  %The details of how this decomposition is performed vary among methods, but they share one important element in common with \qvalue: just as \qvalue assumes that the $p$ values near 1 are from the null, so \locfdr assumes that the $z$ scores near 0 are from the null; Efron calls this assumption the ``Zero Assumption" (ZA).
-
- Before proceeding, we feel obliged to correct the common misconception
- that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
- not depend on the number of tests. This is because, as the number of tests increases, both the true positives and false positives increase linearly,
- and the FDR remains the same. (If this intuitve argument does not convince, see \cite{storey.xx}, and note that the FDR does not depend on the number of tests $m$.)
-  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
- if there are lots of strong signals then the FDR at a given threshold may be low, even if a large number of tests were performed; and conversely
- if there are no strong signals then the FDR at the same threshold may be high, even if relatively few tests were performed. 
- In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
 
@@
 
-\begin{figure}
-\center \includegraphics[height=6in]{Rcode/figures/decomp_ZA.pdf}
-\caption{Comparison of the way that different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (light blue).
-In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
-The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ashr) assumes that the effect sizes (and thus the $z$ scores) have a unimodal distribution about 0, resulting in a very different decomposition that in this case is much closer to the truth.
-(These data were simulated under a model where all of the effects are non-zero -- specifically, $\beta_j \sim N(0,1)$, $s_j=1$, resulting in $z_j \sim N(0,2)$ -- so the ``true" decomposition would make everything light blue.)} \label{fig:ZA}
-\end{figure}
-
-\subsection*{Problems with existing methods}
-
- There are two issues with existing approaches to FDR estimation and control that I would like to address here. 
- The first is that they may do not take proper account of differences in measurement precision across units (i.e. differences in $\shat_j$). 
- The second is that the Zero Assumption, although initially appealing as a ``conservative" assumption,
-  is often an unrealistic and unnecessarily conservative assumption. As we shall see, both factors can cause the FDR to be overestimated.
-
- To illustrate the first issue - differing measurement precision across units - consider a simplistic
- situation where half the measurements are quite precise, and the other half are really, really, poor. 
- Intuitively, the poor measurements tell us nothing, and any sane analysis should effectively ignore them.
-However, in a standard FDR-type analysis, these poor measurements add ``noise" and affect estimated FDRs.
-This is because the $p$ values from the poor measurements will be effectively uniformly distributed, and some will be significant at any given threshold.
-
-To give a concrete example, we performed a simple simulation in which half the observations have
-reasonably good precision (standard error $s_j = 1$) and half the observations have very poor precision ($s_j=10$). In both cases,
-we assume that half the effects are null ($\beta_j=0$)
-and the other half are normally distributed with standard deviation 1 ($\beta_j \sim N(0,1)$). 
-That is, we simulate $\beta_j$ from a mixture 
-\begin{equation}
-p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1)
-\end{equation}
-where $\delta_0(\cdot)$ denotes a point mass at 0.
-
-Figure \ref{fig:goodpoor-hist} shows the distribution of the resulting $p$ values for the good-precision and poor-precision measurements, as
-well as the combined distribution. As expected, the poor-precision measurements lead to almost-uniform $p$ values, and when combined
-with the good-precision measurements they dilute the overall signal in the $p$ values (e.g. they reduce the density of $p$ values near 0).
-This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$, which work with the $p$ values and $Z$ scores respectively:
-the estimated error rates ($q$-values, or lfdr values) for the good-precision observations increase when the low-precision observations are included in the analysis
-(Figure \ref{fig:goodpoor-scatter}). In contrast, the methods described here ($\ashr$) take account of the precision of each measurement, and so
-the estimated error rates for the good-precision observations are unaffected by including the low-precision observations in the analysis.
-(A key assumption of this simulation is that the distribution of the effects $\beta$ is the same for the good-precision and poor-precision measurements;
-we return to discuss this assumption in more detail in section \ref{xx})
-
-\begin{figure}
-\centering
-\subfloat[][Density histograms of $p$ values for good-precision, poor-precision, and combined observations]{
-\includegraphics[width=6.5in]{Rcode/figures/GOODPOOReg_hist.pdf} 
-    \label{fig:goodpoor-hist}
-} \\
-\subfloat[][Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the estimated error rate ($q$ value, lfdr, or lfsr) for a good-precision observation under the two different analyses.]{
-\includegraphics[width=6.5in]{Rcode/figures/GOODPOOReg_scatter.pdf}
-    \label{fig:goodpoor-scatter}
-}
-\caption{Simulation illustrating how poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
-$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
-a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the lfdr estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
-In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
-\end{figure}
-
-Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
-
  \subsection*{Adaptive Shrinkage}
@@
 
- In outline, we use an ``Empirical Bayes" hierarchical model to combine information across measurements and perform inference.
- Here performing inference means obtaining an (approximate) posterior distribution for each effect $\beta_j$,
-which can be used to estimate False Discovery Rates, and to produce both point and interval estimates for $\beta_j$.
- Although ``multiple comparisons" is usually presented as a ``problem", our emphasis here is that the multiple measurements
- actually provide an opportunity: the hierarchical model combines information across measurements to improve both accuracy and precision of estimates.
+ Here we introduce an EB method, which, for reasons discussed below, we call ``adaptive shrinkage" or ``ash".
+ The method provides not only estimated $\fdr_j$ and q values, but also shrinkage point estimates, shrinkage credible intervals,
+ and estimates of the local false sign rate, defined below. It is implemented in the R package \ashr available at \url{http://github.com/stephens999/ashr}.
 
+  Rather than directly modelling the distributions of the $p$ values, or $z$ scores, our starting point is to model the distribution of the effects $\beta_j$.
  Specifically we assume that the effects $\beta_j$ are independent and identically distributed from some unknown distribution $g(\cdot)$,
   \begin{equation} \label{eqn:beta}
- \beta_j \sim g(\cdot),
+ \beta_j \sim g(\cdot).
  \end{equation}
- where $g$ will be estimated from
- the data.   
-  Our methods further make the following key assumptions, which distinguish them from most previous work in this area:
-   \begin{enumerate}
-  \item[A1:] The distribution $g(\cdot)$ is unimodal. (We assume here that the mode is at 0, although 
-this could be relaxed.)  
- \item[A2:] The likelihood $L(\beta) := p(D | \beta)$ can be approximated by $\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
- $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by (\ref{eqn:betahat-t}).
-Equivalently
+ A key assumption, which distinguishes our approach from previous work, is that $g$ is unimodal. We assume here that the mode is at 0, and allow
+ that $g$ can have a point mass at 0, corresponding to the ``null" effects. 
+ (An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data.)
+
+ \def\dt{\text{dt}}
+
+To perform likelihood-based inference for $g,\beta$, we assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
+$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
+ $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
   \begin{equation} \label{eqn:lik}
- \hat{L}(\beta) \propto \prod_j t_\nu((\beta_j-\bhat_j)/\shat_j)
+ \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
  \end{equation}
- where $t_\nu$ denotes the density of the standard $t$ distribution on $\nu$ degrees of freedom.
+ where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e.
+the density of $\mu+\sigma T_\nu$).
+In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
+  \begin{equation} \label{eqn:normlik}
+ \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
+ \end{equation}
 %which, in light of (\ref{eqn:betahat}), is equivalent to
@@
  %\end{equation}
- \end{enumerate}
 
@@
 
-Although the unimodal assumption, A1, will not apply to all situations, we argue that it will be reasonable in many practical contexts. 
+Although the unimodal assumption (UA) will not apply to all situations, we argue that it will be reasonable in many practical contexts. 
 For example, in contexts where interest focusses on which $\beta_j$ differ from zero, which suggests that ``$\beta_j=0$" is a plausible null hypothesis,  
 it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible, and that the distribution of the effects will be unimodal about 0. 
-Alternatively, we can motivate A1 by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
+Alternatively, we can motivate  the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
 such shrinkage is desirable from several standpoints for improving estimation accuracy. 
-Note that assumption A1 relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
-It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and A1 does not contradict this.
+Note that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
+It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
 
-Interestingly, although the unimodal assumption A1 seems natural, much previous related work on estimating False Discovery Rates  
-has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}. 
-In contrast, almost all analogous work in large-scale regression assumes the effects (regression coefficents) to have a unimodal prior distribution - 
-common choices being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors.
-One of our aims here is to emphasize the considerable practical benefits of making the unimodal assumption in the FDR context.
+%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
+%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}. 
+Note that, although the UA appears novel in this context,
+almost all analogous work in large-scale regression make the UA for the regression coefficents - 
+common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{}.
+One of our aims here is to emphasize the considerable practical benefits of making the UA in the FDR context.
 
-Assumption A2 can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
+The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
 they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
@@
 
-Alternatively A2 could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that A2 holds.
+This likelihood approximation can be motivated by assuming that i) $\bhat$
+ $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
+ $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$.
 
-While Assumption A2 has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
- that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
-there is an important difference: here, observations with a large $\shat_j$ have an essentially flat likelihood (\ref{eqn:lik}), and so do not affect inference for
+Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.
+
+A key property of the likelihood (\ref{eqn:lik}) is that
+observations with a large $\shat_j$ have an essentially flat likelihood (\ref{eqn:lik}), and so do not affect inference for
 $g$ or the other effects, whereas when modeling the $Z$ scores directly, observations with large $\shat_j$ produce $Z$ scores with an $N(0,1)$ distribution that {\it do} affect inference.
 Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
+%While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
+ %that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
+%there is an important difference: here,
 
@@
 
-In practice, we implement the unimodal assumption (A1) by assuming a parametric finite mixture for $g$, 
+In practice, we implement the UA by assuming a parametric finite mixture for $g$, 
 \begin{equation} \label{eqn:g}
@@
 \begin{enumerate}[i)]
-\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$
-\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$
+\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$,
+\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,
 \item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ or } U[\cdot; 0,a_k]$,
@@
 or iii) a mixture of ``zero-anchored" uniform distributions, which have one end of their range at 0. These approaches closely mirror approaches in \cite{cordy1997deconvolution}.
+[Case i) is implemented only for the special case $\nu \rightarrow \infty$ in (\ref{eqn:like}).]
 
@@
 \item any symmetric unimodal distribution about 0.
-\item any (asymmetric) unimodal distribution about 0.
+\item any unimodal distribution about 0.
 \end{enumerate}
@@
 
+Note use of SQUAREM.
+
+Note penalty preferring $\pi_0$.
+
+\subsection{various notes...}
+
+Here performing inference means obtaining an (approximate) posterior distribution for each effect $\beta_j$,
+which can be used to estimate False Discovery Rates, and to produce both point and interval estimates for $\beta_j$.
+ Although ``multiple comparisons" is usually presented as a ``problem", our emphasis here is that the multiple measurements
+ actually provide an opportunity: the hierarchical model combines information across measurements to improve both accuracy and precision of estimates.
+
+Note that incorporating differential errors messes up the EM algorithm?
+
+we start by noting that the $p$ values in Figure \ref{} were actually generated under a scenario where
+{\it no tests were null}! Thus the qvalue procedure, which estimates the proportion of nulls to be xx\%, is acting very conservatively in this case.
+However, it could be argued that some degree of conservativeness is inevitable, and that nonetheless the procedure is acting sensibly.
+Therefore, to illustrate this point we really want to focus on the fact that, at least in some settings, the assumption is {\it unrealistic}.
+Specifically, we want to argue that the implied distribution of the alternative $p$ values is unrealistic. In fact, this is hard to see in $p$ value space,
+and so instead we translate these $p$ values back to the $z$ scores that lead to these $p$ values.
+
+The second issue is that directly modeling the p values, or z scores, say via non-parametric methods, 
+can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
+distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
+The solution is to model beta hat as a convolution of some distribution g and an error component.
+
+It should be noted that existing methods were perhaps developed with the expectation that they would typically 
+be applied in settings where the vast majority of all tests are null, and under such circumstances
+the problems caused by this assumption are less severe. However, it seems desirable to have methods that
+also perform well in settings where  this assumptions does not hold, and this motivates the methods developed here.
+
 \section*{Results}
+
+\subsection*{Effects of the Unimodal Assumption}
+
+Here we consider several effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
+standard error, $s_j=1$, and we consider the normal limit (\ref{eqn:normallik}). That
+is,
+\begin{equation} \label{eqn:bhatgivenbz}
+\bhat_j | \beta_j \sim N(0,s^2_j=1)
+\end{equation}
+so $z_j:=\bhat_j/s_j$ are standard normal if $H_j$ holds.
+
+To briefly summarize the results in this section:
+\begin{enumerate}
+\item The UA can produce very different inferences compared with existing methods, and is generally less conservative.
+\item The UA can still yield conservative estimates of the proportion of nulls that hold, $\pi_0$, and hence conservative estimates of FDR and fdr.
+\item The UA results in a numerically highly-stable procedure, especially compared with a fully non-parametric approach.
+\item The UA is somewhat robust to deviations from unimodality.
+\end{enumerate}
+
+\subsubsection{The UA can produce very different inferences compared with existing methods}
+
+To illustrate that the UA can have a big impact on inference 
+we simulate data where the true $\beta_j$ are distributed as $N(0,1)$ (and then $\bhat_j | \beta_j \sim N(\beta_j,1)$.
+We feed the resulting $z$ scores to the R packages \locfdr and \mixfdr, and to our package \ashr, and the corresponding $p$ values to the R package \qvalue.
+
+TODO: put FDRs on graph? Simulate with 50\% null?
+
+\begin{figure}
+\center \includegraphics[height=6in]{Rcode/figures/decomp_ZA.pdf}
+\caption{Comparison of the way that different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (light blue).
+In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
+The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
+In contrast the method introduced here (\ashr) assumes that the effect sizes (and thus the $z$ scores) have a unimodal distribution about 0, resulting in a very different decomposition that in this case is much closer to the truth.
+(These data were simulated under a model where all of the effects are non-zero -- specifically, $\beta_j \sim N(0,1)$, $s_j=1$, resulting in $z_j \sim N(0,2^2)$ -- so the ``true" decomposition would make everything light blue.)} \label{fig:ZA}
+\end{figure}
+
+\subsubsection*{Conservative estimates of $\pi_0$}
+
+One concern, given the
+
+\subsubsection*{Robustness to deviations from UA}
+
+Show that the q values are driven by changes in pi_0; even when not unimodal pretty close...
+
+\subsubsection*{The UA produces stable estimates}
+
+The EM algorithm is notorious for convergence to local optima. 
+But in this case, over hundreds of applications of the procedure, we have observed no serious problems 
+caused by potential multimodality.
+
+\subsection*{Differing measurement precision across units}
+
+ To illustrate the first issue - differing measurement precision across units - consider a simplistic
+ situation where half the measurements are quite precise, and the other half are really, really, poor. 
+ Intuitively, the poor measurements tell us nothing, and any sane analysis should effectively ignore them.
+However, in a standard FDR-type analysis, these poor measurements add ``noise" and affect estimated FDRs.
+This is because the $p$ values from the poor measurements will be effectively uniformly distributed, and some will be significant at any given threshold.
+
+To give a concrete example, we performed a simple simulation in which half the observations have
+reasonably good precision (standard error $s_j = 1$) and half the observations have very poor precision ($s_j=10$). In both cases,
+we assume that half the effects are null ($\beta_j=0$)
+and the other half are normally distributed with standard deviation 1 ($\beta_j \sim N(0,1)$). 
+That is, we simulate $\beta_j$ from a mixture 
+\begin{equation}
+p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1)
+\end{equation}
+where $\delta_0(\cdot)$ denotes a point mass at 0.
+
+Figure \ref{fig:goodpoor-hist} shows the distribution of the resulting $p$ values for the good-precision and poor-precision measurements, as
+well as the combined distribution. As expected, the poor-precision measurements lead to almost-uniform $p$ values, and when combined
+with the good-precision measurements they dilute the overall signal in the $p$ values (e.g. they reduce the density of $p$ values near 0).
+This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$, which work with the $p$ values and $Z$ scores respectively:
+the estimated error rates ($q$-values, or lfdr values) for the good-precision observations increase when the low-precision observations are included in the analysis
+(Figure \ref{fig:goodpoor-scatter}). In contrast, the methods described here ($\ashr$) take account of the precision of each measurement, and so
+the estimated error rates for the good-precision observations are unaffected by including the low-precision observations in the analysis.
+(A key assumption of this simulation is that the distribution of the effects $\beta$ is the same for the good-precision and poor-precision measurements;
+we return to discuss this assumption in more detail in section \ref{xx})
+
+\begin{figure}
+\centering
+\subfloat[][Density histograms of $p$ values for good-precision, poor-precision, and combined observations]{
+\includegraphics[width=6.5in]{Rcode/figures/GOODPOOReg_hist.pdf} 
+    \label{fig:goodpoor-hist}
+} \\
+\subfloat[][Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
+Each point shows the estimated error rate ($q$ value, lfdr, or lfsr) for a good-precision observation under the two different analyses.]{
+\includegraphics[width=6.5in]{Rcode/figures/GOODPOOReg_scatter.pdf}
+    \label{fig:goodpoor-scatter}
+}
+\caption{Simulation illustrating how poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
+$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
+a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
+The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the lfdr estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
+In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
+\end{figure}
+
+Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
 
@@
 Also maybe look at performance for small sample sizes - e.g. n=1?
-
-PENN Note: need to fix bug in dscr where when an element of result is null, the data.frame command doesn't work.
-Want to convert it to missing - maybe in the scoring function.
 
@@
 These simulations illustrate two key points: i) although $\pi_0$ is not identifiable, the penalized likelihood approach
-provides conservative estimates for $\pi_0$; ii) replacing the ``zero assumption" with the unimodal assumption can produce
-less conservative, more accurate, estimates of $\pi_0$ (provided of course that the unimodal assumption holds, as it does in these simulations).
+provides conservative estimates for $\pi_0$; ii) replacing the ``zero assumption" with the UA can produce
+less conservative, more accurate, estimates of $\pi_0$ (provided of course that the UA holds, as it does in these simulations).
 
@@
 
+Before proceeding, we feel obliged to correct the common misconception
+ that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
+ not depend on the number of tests. This is because, as the number of tests increases, both the true positives and false positives increase linearly,
+ and the FDR remains the same. (If this intuitve argument does not convince, see \cite{storey.xx}, and note that the FDR does not depend on the number of tests $m$.)
+  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
+ if there are lots of strong signals then the FDR at a given threshold may be low, even if a large number of tests were performed; and conversely
+ if there are no strong signals then the FDR at the same threshold may be high, even if relatively few tests were performed. 
+ In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
+
+ There are two issues with existing approaches to FDR estimation and control that I would like to address here. 
+ The first is that they may do not take proper account of differences in measurement precision across units (i.e. differences in $\shat_j$). 
+ The second is that the Zero Assumption, although initially appealing as a ``conservative" assumption,
+  is often an unrealistic and unnecessarily conservative assumption. As we shall see, both factors can cause the FDR to be overestimated.
+
 \cite{efron2008microarrays} states the Zero Assumption as the assumption that ``most of the z-values near zero come from null genes".

### 2015-02-27 examples/qval.nonp.rmd
split ash into ash and ash.workhorse

@@
   test = simdata(10000, 5000, 0, 2, 1) 
-  test.ash = ash(test$betahat,test$betahatsd,method="fdr")
+  test.ash = ash(test$betahat,test$betahatsd)
+  test.ash.hu = ash(test$betahat,test$betahatsd,mixcompdist="halfuniform")
   z = test$betahat/test$betahatsd
@@
 ```{r}
-  par(mfcol=c(2,2))
-  plot(qval.nonp(p,get_pi0(test.ash)),test.ash$qvalue, xlab="nonparametric q-values (with pi0 estimated by ash)", ylab = "ash q-values",main="ash")
-  abline(a=0,b=1,col=2,lwd=2)
 
-  plot(qval.nonp(p,test.qv$pi0),test.qv$qvalues, xlab="nonparametric q-values (with pi0 estimated by qvalue)", ylab = "qvalue q-values",main="qvalue")
-  abline(a=0,b=1,col=2,lwd=2)
+  res = rbind(data.frame(method="qvalue",q = test.qv$qvalues, qnonp = qval.nonp(p,test.qv$pi0)),
+         data.frame(method="mixfdr",q = test.mixfdr.qval, qnonp = qval.nonp(p,test.mixfdr$pi[1])),
+         data.frame(method="locfdr",q = test.locfdr.qval, qnonp = qval.nonp(p,test.locfdr$fp0[1,3])),
 
- plot(qval.nonp(p,test.mixfdr$pi[1]),test.mixfdr.qval, xlab="nonparametric q-values (with pi0 estimated by mixfdr)", ylab = "mixfdr q-values",main="mixfdr")
-  abline(a=0,b=1,col=2,lwd=2)
+         data.frame(method="ashr",q = test.ash$qvalue, qnonp = qval.nonp(p,get_pi0(test.ash)))
+         )
+  library(ggplot2)
+  ggplot(res, aes(x=q, y=qnonp)) + facet_grid(.~method) + geom_point(shape=1) +xlim(0,1) +ylim(0,1) + xlab("qvalue computed by method") +ylab("qvalue computed \n non-parametrically") +geom_abline(slope=1,intercept=0,color=2) + coord_equal() + theme(axis.text.x  = element_text(size=8), axis.text.y  = element_text(size=8))
+  ggsave("qval_comparison.png", width=8, height=2, dpi=100)
 
- plot(qval.nonp(p,test.locfdr$fp0[1,3]),test.locfdr.qval, xlab="nonparametric q-values (with pi0 estimated by locfdr)", ylab = "locfdr q-values", main="locfdr")
-  abline(a=0,b=1,col=2,lwd=2) 
-  par(mfcol=c(1,1))
 ```
@@
 In this second simulation, half the $\beta_j$ are null, and the alternatives
-are $\beta_j \sim N(2,1)$. Note that this deviates quite a bit from the `ash` assumption that
+are $\beta_j \sim N(+-2,1)$. Note that this deviates quite a bit from the `ash` assumption that
 the non-zero $\beta_j$ will have a unimodal distribution centered on 0. This presumably
@@
   set.seed(100)
-  test = simdata(10000, 5000, 2, 1, 1) 
-  test.ash = ash(test$betahat,test$betahatsd,method="fdr")
+  test = rbind(data.frame(simdata(5000, 2500, 2, 1, 1)), data.frame(simdata(5000,2500,-2,1,1))) 
+  test.ash = ash(test$betahat,test$betahatsd)
+
   z = test$betahat/test$betahatsd
   p = pchisq(z^2, df=1,lower.tail=FALSE)
-  plot(qval.nonp(p,get_pi0(test.ash)),test.ash$qvalue, xlab="nonparametric q-values (ash estimate of pi0)", ylab = "ash q-values")
-  abline(a=0,b=1,col=2,lwd=2)
+  test.qv = qvalue(p)
+  test.locfdr = locfdr(z,nulltype=0,plot=0)
+  test.locfdr.qval = qval.from.lfdr(test.locfdr$fdr)
+  test.mixfdr = mixFdr(z,theonull=TRUE,noiseSD=1,plots=FALSE)
+  test.mixfdr.qval = qval.from.lfdr(test.mixfdr$fdr)
+  kable(data.frame(method=c("qvalue","ash","locfdr","mixfdr"),pi0=round(c(test.qv$pi0,get_pi0(test.ash),test.locfdr$fp0[1,3],test.mixfdr$pi[1]),digits=2)))
+
+  res = rbind(data.frame(method="qvalue",q = test.qv$qvalues, qnonp = qval.nonp(p,test.qv$pi0)),
+         data.frame(method="mixfdr",q = test.mixfdr.qval, qnonp = qval.nonp(p,test.mixfdr$pi[1])),
+         data.frame(method="locfdr",q = test.locfdr.qval, qnonp = qval.nonp(p,test.locfdr$fp0[1,3])),
+
+         data.frame(method="ashr",q = test.ash$qvalue, qnonp = qval.nonp(p,get_pi0(test.ash)))
+         )
+  library(ggplot2)
+  ggplot(res, aes(x=q, y=qnonp)) + facet_grid(.~method) + geom_point(shape=1) +xlim(0,1) +ylim(0,1) + xlab("qvalue computed by method") +ylab("qvalue computed \n non-parametrically") +geom_abline(slope=1,intercept=0,color=2) + coord_equal() + theme(axis.text.x  = element_text(size=8), axis.text.y  = element_text(size=8))
+  ggsave("qval_comparison_nonUA.png", width=8, height=2, dpi=100)
+
 ```

### 2015-03-05 dsc-shrink/methods/ash.wrapper.R
updated dsc-shrink

@@ -15,6 +15,16 @@ ash.wrapper=function(input,args=NULL){
   if(is.null(args)){
     args=list(mixcompdist="halfuniform",method="fdr")
   }
-  res = do.call(ash, args= c(list(betahat=input$betahat,sebetahat=input$sebetahat),args))
-  return(list(beta_est=res$PosteriorMean,pi0_est=get_pi0(res)))
+  res = do.call(ash, args= c(list(betahat=input$betahat,sebetahat=input$sebetahat,df=input$df),args))
+  return(res)
 }
+
+ash2beta_est =function(output){
+  return (list(beta_est=output$PosteriorMean))
+} 
+
+ash2pi0_est =function(output){
+  return (list(pi0_est=get_pi0(output)))
+} 
+
+

### 2015-03-05 dsc-shrink/methods/mixfdr.wrapper.R

@@ -15,6 +15,15 @@ library(mixfdr)
 mixfdr.wrapper=function(input,args=NULL){
   res = try(do.call(mixFdr, args= c(list(x=input$betahat/input$sebetahat, noiseSD=1, plots=FALSE),args)))
   if(inherits(res,"try-error")){res=list(effectSize=rep(NA,length(input$betahat)),pi0_est=NA)}
-  return(list(beta_est=res$effectSize*input$sebetahat,pi0_est=res$pi[1]))
+  return(list(res=res,input=input))
 }
+
+mixfdr2beta_est = function(output){
+  return (list(beta_est=output$res$effectSize*output$input$sebetahat))
+}
+
+mixfdr2pi0_est = function(output){
+  return (list(pi0_est=output$res$pi[1]))
+}
+
 

### 2015-03-05 dsc-shrink/run_dsc.R

@@ -1,9 +1,22 @@
 library(dscr)
 
-source("scenarios.R")
-source("methods.R")
+dsc_shrink=new.dsc("shrink","dsc-shrink-files")
+source("addScenarios.R")
+source("addMethods.R")
+
+addOutputParser(dsc_shrink,"ash2beta",ash2beta_est,"ash_output","beta_est_output")
+addOutputParser(dsc_shrink,"mixfdr2beta",mixfdr2beta_est,"mixfdr_output","beta_est_output")
+
+addOutputParser(dsc_shrink,"ash2pi0",ash2pi0_est,"ash_output","pi0_est_output")
+addOutputParser(dsc_shrink,"mixfdr2pi0",mixfdr2pi0_est,"mixfdr_output","pi0_est_output")
+
+
 source("score.R")
-res=run_dsc(scenarios,methods,score)
+addScore(dsc_shrink,score,"beta_err","beta_est_output")
+addScore(dsc_shrink,score2,"pi0_score","pi0_est_output")
+
+
+res=run_dsc(dsc_shrink)
 save(res,file="res.RData")
 
 

### 2015-03-05 dsc-shrink/scenarios.R

@@ -1,114 +1,0 @@
-sourceDir("datamakers")
-scenarios=list()
-
-#Now, for each scenario create an element of scenarios of the following form
-scenarios[[1]]=list(name="A",
-                    fn=rnormmix_datamaker,
-                    args=list(
-                      g=normalmix(c(2/3,1/3),c(0,0),c(1,2)),
-                      min_pi0=0,
-                      max_pi0=0,
-                      nsamp=1000,
-                      betahatsd=1
-                      ),
-                    seed=1:100)
-
-scenarios[[2]]=list(name="B",
-                    fn=rnormmix_datamaker,
-                    args=list(
-                      g=normalmix(rep(1/7,7),c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7)),
-                      min_pi0=0,
-                      max_pi0=0,
-                      nsamp=1000,
-                      betahatsd=1
-                    ),
-                    seed=1:100)
-
-
-scenarios[[3]] = list(name="C",
-                      fn=rnormmix_datamaker,
-                      args=list(
-                        g=normalmix(c(1/4,1/4,1/3,1/6),c(-2,-1,0,1),c(2,1.5,1,1)),
-                        min_pi0=0,
-                        max_pi0=0,
-                        nsamp=1000,
-                        betahatsd=1
-                      ),
-                      seed=1:100)
-
-#scenarios An, Bn, Cn are the same as A,B,C but with nulls included (pi0 uniform on [0,1])
-scenarios[[4]]=list(name="An",
-                    fn=rnormmix_datamaker,
-                    args=list(
-                      g=normalmix(c(2/3,1/3),c(0,0),c(1,2)),
-                      min_pi0=0,
-                      max_pi0=1,
-                      nsamp=1000,
-                      betahatsd=1
-                    ),
-                    seed=1:100)
-
-scenarios[[5]]=list(name="Bn",
-                    fn=rnormmix_datamaker,
-                    args=list(
-                      g=normalmix(rep(1/7,7),c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7)),
-                      min_pi0=0,
-                      max_pi0=1,
-                      nsamp=1000,
-                      betahatsd=1
-                    ),
-                    seed=1:100)
-
-
-scenarios[[6]] = list(name="Cn",
-                      fn=rnormmix_datamaker,
[…]

### 2015-03-05 dsc-shrink/score.R

@@ -13,8 +13,10 @@
 #'
 score = function(data, output){
   return(list(RMSE=sqrt(mean((data$meta$beta-output$beta_est)^2)),
-              MAE = median(abs(data$meta$beta-output$beta_est)),
-              pi0 = data$meta$pi0,
-              pi0_est = output$pi0_est)
-  )
+              MAE = median(abs(data$meta$beta-output$beta_est))))
+}
+
+score2 = function(data, output){
+  return(list(pi0 = data$meta$pi0,
+              pi0_est = output$pi0_est))
 }

### 2015-03-16 Rcode/efron.fcr.rmd
some updates to fcr and paper

@@
   nsamp=10000
+  altmean=-3
   mu0 = rep(0,nsamp)
-  mu1 = rnorm(nsamp,-3,1)
+  mu1 = rnorm(nsamp,altmean,1)
   comp = rbinom(nsamp,1,0.1)
@@
   library(ashr)
-  res.ash = ash(z,1,mixcompdist="halfuniform",method="fdr")
+  res.ash = ash(z,1,mixcompdist="halfuniform")
   CImatrix= ashci(res.ash,level=0.95)
@@
 ```{r}
-  BayesComparePlot=function(){
+  BayesComparePlot=function(altmean=-3){
     plot(z,mu,xlim=c(-8,0))
@@
 
-    fdr = 0.9*dnorm(z)/(0.9*dnorm(z)+0.1*dnorm(z,-3,sqrt(2)))
+    fdr = 0.9*dnorm(z)/(0.9*dnorm(z)+0.1*dnorm(z,altmean,sqrt(2)))
     o=order(z)
-    upper = ifelse(fdr[o]<0.025,(z[o]-3)/2+qnorm(0.975+fdr[o])/sqrt(2),0)
+    upper = ifelse(fdr[o]<0.025,(z[o]+altmean)/2+qnorm(0.975+fdr[o])/sqrt(2),0)
     lines(z[o],upper,col="green",lwd=2)
-    lines(z[o],(z[o]-3)/2-qnorm(0.975)/sqrt(2),col="green",lwd=2)
+    lines(z[o],(z[o]+altmean)/2-qnorm(0.975)/sqrt(2),col="green",lwd=2)
     abline(v=max(z[fdr<0.05]))
   }
-  BayesComparePlot()
+  BayesComparePlot(altmean)
 ```
@@
 ```{r}
-   res.ash = ash(z,1,mixcompdist="halfuniform",method="fdr")
+   res.ash = ash(z,1,mixcompdist="halfuniform")
   CImatrix= ashci(res.ash,level=0.95)
@@
 ```{r}
-   res.ash = ash(z,1,mixcompdist="normal",method="fdr")
+   res.ash = ash(z,1,mixcompdist="normal")
   CImatrix= ashci(res.ash,level=0.95)
@@
 ```{r}
-   res.ash = ash(z,1,mixcompdist="uniform",method="fdr")
+   res.ash = ash(z,1,mixcompdist="uniform")
   CImatrix= ashci(res.ash,level=0.95)

### 2015-03-16 paper/main.tex
split ash into ash and ash.workhorse
updated dsc-shrink
some updates to fcr and paper

@@
 
-Show that the q values are driven by changes in pi_0; even when not unimodal pretty close...
+Show that the q values are driven by changes in $\pi_0$; even when not unimodal pretty close...
 
@@
 caused by potential multimodality.
+
+\subsubsection*{Reverse asymptotics}
 
@@
 
+\subsection*{False Coverage Rates}
+
+To summarize the False Coverage Rate problem, we can do no better than quote Efron, who says
+``An insightful and ingenious paper by Benjamini and
+Yekutieli (2005) discusses the following problem in simultaneous
+significance testing: having applied false
+discovery rate methods to select a set of nonnull cases,
+how can confidence intervals be assigned to the true
+effect size for each selected case?" \cite{efron2008} notes the potential for EB approaches to tackle this problem,
+but also notes the difficulties of implementing a stable general algorithm.
+%deconvolution
+%calculations are notoriously tricky and I have not
+%been able to produce a stable general algorithm. ; see also 
+Zhao+Hwang consider in detail the case where the non-null effects are normally distributed.
+
+Results in efron.fcr.rmd suggests that lower CI is OK; upper is more problematic because of difficulty in accurately estimating tail?
+
+To do: implement methods to compute posterior cdf of each observation at given points. 
+Then compute these for the simulation studies. Check whether issue with upper bound is due to non-unimodal distribution.
+
 \subsection*{Improved conservative estimation of $\pi_0$}

### 2015-03-16 talks/one_hour_seminar.Rmd
some updates to fcr and paper

@@
 author: Matthew Stephens
-date: February 3, 2015
+date: March 9, 2015
 output: ioslides_presentation
@@
 
-- `ashr` package: `http://www.github.com/stephens999/ash`
+- `ashr` package: `http://www.github.com/stephens999/ashr`
 

### 2015-03-16 dsc-shrink/addMethods.R

@@ -1,4 +1,6 @@
 sourceDir("methods")
+
+addMethod(dsc_shrink,name="ash.hu.cxx",fn =ash.wrapper,args=list(mixcompdist="halfunif",cxx=TRUE),outputtype = "ash_output")
 addMethod(dsc_shrink,name="ash.hu",fn =ash.wrapper,args=list(mixcompdist="halfunif"),outputtype = "ash_output")
 addMethod(dsc_shrink,name="ash.u",fn =ash.wrapper,args=list(mixcompdist="unif"),outputtype = "ash_output")
 addMethod(dsc_shrink,name="ash.n",fn =ash.wrapper,args=list(mixcompdist="normal"),outputtype = "ash_output")

### 2015-03-16 dsc-shrink/run_dsc.R

@@ -20,4 +20,5 @@ res=run_dsc(dsc_shrink)
 save(res,file="res.RData")
 
 
-
+xtabs(pi0_est ~ method+scenario,res$pi0_score)
+xtabs(user.self ~ method+scenario,res$pi0_score)

### 2015-04-08 Rcode/symm.fcr.rmd
lots of changes

--- 2015-03-16 Rcode/efron.fcr.rmd
@@
 ---
-title: "FSR in non-unimodal case"
+title: "FCR in unimodal case"
 output: html_document
@@
 
-This example comes from Efron (2008) p16 when examining the false coverage rate (FCR).
+Take a look at the calibration of posterior for unimodal case.
+Our exp
 
@@
   set.seed(100)
-  nsamp=10000
-  altmean=-3
+  nsamp=1000
+  altmean=0
   mu0 = rep(0,nsamp)
-  mu1 = rnorm(nsamp,altmean,1)
-  comp = rbinom(nsamp,1,0.1)
+  mu1 = rnorm(nsamp,altmean,2)
+  comp = rbinom(nsamp,1,0.8)
   mu = ifelse(comp==0,mu0,mu1)
@@
   CImatrix= ashci(res.ash,level=0.95)
+  cobs=vcdf_post(res.ash$fitted.g,mu,z,rep(1,nsamp),NULL)
+  cd = diag(cobs)
+  hist(cd[res.ash$NegativeProb>0.95])
+  hist(cd[res.ash$PositiveProb>0.95])
 ```

### 2015-04-08 dsc-shrink/investigate_cdf.rmd
lots of changes

---
title: "investigate_cdf"
output: html_document
---

```{r}
load("res.RData")
#load("res10.RData")
library("dscr")
library("ashr")
library(reshape2)
library(ggplot2)
library(dplyr)

#res=res10
neglong = 
  res$negprob %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars=c("method","scenario","seed",".id"),value.name="negprob") %>%
    filter(negprob > 0.9)

poslong = 
  res$posprob %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars=c("method","scenario","seed",".id"),value.name="posprob") %>%
    filter(posprob > 0.9)

reslong = 
  res$cdf_score %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%    
    melt(id.vars=c("method","scenario","seed",".id"))

reslong.pos = inner_join(reslong,poslong)
reslong.neg = inner_join(reslong,neglong)

```

```{r}

cdfplot=function(df,title=""){
  p=ggplot(df,aes(x=value,color=method)) + stat_ecdf() + facet_grid(.~scenario) + ggtitle(title)
  print(p +coord_equal(ratio = 1)+geom_abline())
  #print(p+coord_cartesian(xlim = c(0,1),ylim=c(0,1)) +coord_equal(ratio = 1))
}
pdf("cdfplots_coverage.pdf")
cdfplot(reslong.pos %>% 
          filter(method %in% c("ash.n","ash.hu","ash.u","ash.hu.s") & scenario %in% c("A","B","C","efron_FCR")), "Positive Discoveries"
        )

cdfplot(reslong.neg %>% 
          filter(method %in% c("ash.n","ash.hu","ash.u","ash.hu.s") & scenario %in% c("A","B","C","efron_FCR")), "Negative Discoveries"
        )

cdfplot(reslong %>% 
          filter(method %in% c("ash.n","ash.hu","ash.u","ash.hu.s") & scenario %in% c("A","B","C","efron_FCR")), "All Discoveries"
        )

dev.off()
```

Table of upper tail for all observations
```{r}
xtabs(ut~method+scenario,reslong %>% group_by(scenario,method) %>% summarize(ut = mean(value>0.95))) %>% round(2)
```

Table of lower tail for all observations
```{r}
xtabs(lt~method+scenario,reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05))) %>% round(2)
```

Table of 0.90 CI coverage for all observations
```{r}
xtabs(bt~method+scenario,reslong %>% group_by(scenario,method) %>% summarize(bt = mean(value>0.05 & value <0.95))) %>% round(2)
```

This is the table for the upper tail of the negative findings
and lower tail of positive findings. This should generally perform well.
```{r}
xtabs(ut~method+scenario,reslong.neg %>% group_by(scenario,method) %>% summarize(ut = mean(value>0.95))) %>% round(2)

xtabs(lt~method+scenario,reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05))) %>% round(2)
```

This is the table for the upper tail of the negative findings
and lower tail of positive findings. The uniform methods
tend to over-shrink.
```{r}
xtabs(lt~method+scenario,reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05))) %>% round(2)

xtabs(ut~method+scenario,reslong.pos %>% group_by(scenario,method) %>% summarize(ut = mean(value>0.95))) %>% round(2)

# > xtabs(lt~method+scenario,reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05))) %>% round(2)
#             scenario
# method          A   An    B   Bn    C   Cn easy efron_FCR hard
#   ash.hu     0.12 0.12 0.07 0.08 0.07 0.07 0.06      0.06 0.13
#   ash.hu.cxx 0.12 0.12 0.07 0.08 0.07 0.07 0.06      0.06 0.13
#   ash.hu.nw2 0.11 0.10 0.08 0.10 0.07 0.07 0.06      0.07 0.13
#   ash.hu.s   0.08 0.08 0.08 0.08 0.06 0.06 0.05      0.05 0.11
#   ash.n      0.06 0.05 0.00 0.00 0.06 0.06 0.05      0.03 0.06
#   ash.n.nw2  0.06 0.05 0.01 0.02 0.07 0.07 0.05      0.03 0.05
#   ash.n.s    0.06 0.05 0.01 0.03 0.07 0.07 0.05      0.04 0.05
#   ash.u      0.10 0.10 0.07 0.07 0.09 0.09 0.05      0.07 0.12
#   ash.u.nw2  0.09 0.09 0.08 0.10 0.09 0.09 0.06      0.07 0.12
#   ash.u.s    0.07 0.07 0.07 0.10 0.09 0.09 0.05      0.07 0.09
# > xtabs(ut~method+scenario,reslong.neg %>% group_by(scenario,method) %>% summarize(ut = mean(value>0.95))) %>% round(2)
#             scenario
# method          A   An    B   Bn    C   Cn easy efron_FCR hard
#   ash.hu     0.05 0.07 0.04 0.07 0.05 0.05 0.05      0.05 0.06
#   ash.hu.cxx 0.05 0.07 0.04 0.07 0.05 0.05 0.05      0.05 0.06
#   ash.hu.nw2 0.05 0.06 0.04 0.07 0.05 0.05 0.05      0.05 0.07
#   ash.hu.s   0.06 0.15 0.05 0.16 0.05 0.18 0.19      0.26 0.26
#   ash.n      0.04 0.06 0.04 0.06 0.04 0.04 0.05      0.03 0.05
#   ash.n.nw2  0.04 0.06 0.04 0.06 0.04 0.04 0.05      0.03 0.05
#   ash.n.s    0.05 0.05 0.04 0.05 0.04 0.04 0.05      0.03 0.06
#   ash.u      0.05 0.06 0.04 0.06 0.04 0.04 0.05      0.04 0.06
#   ash.u.nw2  0.04 0.06 0.04 0.06 0.04 0.04 0.05      0.04 0.06
#   ash.u.s    0.05 0.05 0.05 0.05 0.04 0.04 0.05      0.03 0.06
# > xtabs(lt~method+scenario,reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05))) %>% round(2)
#             scenario
# method          A   An    B   Bn    C   Cn easy efron_FCR hard
#   ash.hu     0.05 0.06 0.04 0.04 0.06 0.05 0.05      0.00 0.06
#   ash.hu.cxx 0.05 0.06 0.04 0.04 0.06 0.05 0.05      0.00 0.06
#   ash.hu.nw2 0.05 0.05 0.04 0.04 0.05 0.05 0.05      0.33 0.05
#   ash.hu.s   0.06 0.24 0.06 0.26 0.06 0.25 0.17      0.56 0.42
#   ash.n      0.05 0.05 0.03 0.03 0.09 0.10 0.04      0.34 0.04
#   ash.n.nw2  0.04 0.04 0.04 0.04 0.08 0.09 0.04      0.31 0.04
#   ash.n.s    0.05 0.05 0.04 0.05 0.07 0.08 0.05      0.41 0.05
#   ash.u      0.05 0.06 0.04 0.04 0.10 0.11 0.05      0.33 0.05
#   ash.u.nw2  0.05 0.05 0.04 0.04 0.08 0.09 0.05      0.33 0.04
#   ash.u.s    0.05 0.05 0.05 0.06 0.08 0.08 0.05      0.48 0.06
# > xtabs(ut~method+scenario,reslong.pos %>% group_by(scenario,method) %>% summarize(ut = mean(value>0.95))) %>% round(2)
#             scenario
# method          A   An    B   Bn    C   Cn easy efron_FCR hard
#   ash.hu     0.10 0.11 0.06 0.07 0.10 0.12 0.06      0.00 0.17
#   ash.hu.cxx 0.10 0.11 0.06 0.07 0.10 0.12 0.06      0.00 0.17
#   ash.hu.nw2 0.09 0.09 0.07 0.11 0.09 0.11 0.06      0.00 0.17
#   ash.hu.s   0.07 0.07 0.07 0.08 0.08 0.07 0.05      0.00 0.09
#   ash.n      0.05 0.05 0.00 0.00 0.00 0.00 0.05      0.00 0.06
#   ash.n.nw2  0.05 0.06 0.01 0.01 0.01 0.01 0.05      0.00 0.07
#   ash.n.s    0.05 0.05 0.01 0.03 0.01 0.01 0.05      0.00 0.07
#   ash.u      0.09 0.11 0.07 0.08 0.01 0.01 0.06      0.00 0.13
#   ash.u.nw2  0.08 0.09 0.08 0.11 0.01 0.01 0.06      0.00 0.12
#   ash.u.s    0.07 0.07 0.07 0.11 0.01 0.02 0.06      0.00 0.10
# > xtabs(bt~method+scenario,reslong.neg %>% group_by(scenario,method) %>% summarize(bt = mean(value>0.05 & value <0.95))) %>% round(2)
#             scenario
# method          A   An    B   Bn    C   Cn easy efron_FCR hard
#   ash.hu     0.83 0.82 0.89 0.85 0.88 0.87 0.89      0.89 0.81
#   ash.hu.cxx 0.83 0.82 0.89 0.85 0.88 0.87 0.89      0.89 0.81
#   ash.hu.nw2 0.85 0.83 0.88 0.84 0.88 0.88 0.89      0.88 0.80
#   ash.hu.s   0.86 0.77 0.87 0.76 0.88 0.76 0.75      0.69 0.63
#   ash.n      0.90 0.89 0.96 0.94 0.90 0.90 0.90      0.94 0.89
#   ash.n.nw2  0.90 0.89 0.95 0.92 0.90 0.90 0.90      0.94 0.90
#   ash.n.s    0.89 0.91 0.94 0.92 0.89 0.90 0.90      0.93 0.89
#   ash.u      0.85 0.84 0.89 0.86 0.87 0.86 0.89      0.89 0.82
#   ash.u.nw2  0.87 0.85 0.88 0.83 0.87 0.87 0.89      0.89 0.82
#   ash.u.s    0.87 0.88 0.88 0.85 0.87 0.88 0.90      0.90 0.85
# > xtabs(bt~method+scenario,reslong.pos %>% group_by(scenario,method) %>% summarize(bt = mean(value>0.05 & value <0.95))) %>% round(2)
#             scenario
# method          A   An    B   Bn    C   Cn easy efron_FCR hard
#   ash.hu     0.84 0.83 0.90 0.89 0.84 0.83 0.89      1.00 0.77
#   ash.hu.cxx 0.84 0.83 0.90 0.89 0.84 0.83 0.89      1.00 0.77
#   ash.hu.nw2 0.86 0.85 0.88 0.85 0.86 0.84 0.89      0.67 0.78
#   ash.hu.s   0.87 0.68 0.87 0.65 0.87 0.67 0.77      0.44 0.49
#   ash.n      0.90 0.90 0.96 0.96 0.91 0.90 0.91      0.66 0.89
#   ash.n.nw2  0.90 0.90 0.95 0.94 0.92 0.91 0.90      0.69 0.89
#   ash.n.s    0.90 0.90 0.94 0.92 0.92 0.91 0.90      0.59 0.88
#   ash.u      0.86 0.84 0.90 0.88 0.89 0.88 0.90      0.67 0.82
#   ash.u.nw2  0.88 0.86 0.88 0.85 0.90 0.90 0.90      0.67 0.83
#   ash.u.s    0.88 0.88 0.87 0.84 0.91 0.90 0.89      0.52 0.84
```

```{r}
# res1 = as.matrix(res1[,5:1004])
# neg = as.matrix(neg[,5:1004])
# pos = as.matrix(pos[,5:1004])
# 
# hist(res1[pos>0.9])
# hist(res1[neg>0.9])
# 
# res_cdf.mix = as.matrix(res$cdf_score[grepl("mixfdr",res$cdf_score$method),5:1004])
# fdr.mix = as.matrix(res$fdr[grepl("mixfdr",res$fdr$method),5:1004])
# 
# hist(res_cdf.mix[fdr.mix<0.1])
# 
# 
# 
# res_ecdf = apply(res_cdf,1,ecdf)
# 
# #plot for scenario A, histogram of cdf values for observations tha are
# #significant, separately for positive and negative findings
# pnhistA=function(method,seed){
#   temp=dscr::loadExample(dsc_shrink,seed,"A",method)
#   neg=(temp$output$NegativeProb>0.9)
#   pos=(temp$output$PositiveProb>0.9)
#   par(mfcol=c(2,1))
#   hist(as.numeric(res_cdf[seed,neg]),nclass=20)
#   hist(as.numeric(res_cdf[seed,pos]),nclass=20)
# }
# 
# plotcdfA=function(method,seed){
#   temp=dscr::loadExample(dsc_shrink,seed,"A",method)
#   par(mfcol=c(1,1))
#   plot(temp$output)
#   g=normalmix(c(2/3,1/3),c(0,0),c(1,2)) # the g for scenario A
#   lines(seq(-5,5,length=100),mixcdf(g,seq(-5,5,length=100)),col=2,lwd=2)
# }
# 
# temp=dscr::loadExample(dsc_shrink,4,"A","ash.n")
# neg = temp$output$NegativeProb>0.9
# pnhistA("ash.n",4)
# plotcdfA("ash.n",4)
# temp=loadExample(dsc_shrink,4,"A","ash.n")
# plot(temp$meta$beta,temp$input$betahat,col=neg+1)

```

Table for both tails - coverage of 0.9 symetric intervals.
```{r}
xtabs(bt~method+scenario,reslong.neg %>% group_by(scenario,method) %>% summarize(bt = mean(value>0.05 & value <0.95))) %>% round(2)
xtabs(bt~method+scenario,reslong.pos %>% group_by(scenario,method) %>% summarize(bt = mean(value>0.05 & value <0.95))) %>% round(2)

```

### 2015-04-08 dsc-stability/check_convergence.rmd
lots of changes

---
title: "check_convergence_problems"
output: html_document
---

Here I tried to check whether an example that did not converge well 
actually affected results.

```{r}
  library(dscr)
  library(ashr)
  library(ggplot2)
  library(reshape2)
  load("dsc_ar.RData")
  res=dsc_ar$res
#check out which one showed biggest convergence problem
  prob = res[which.max(res$diff1),]
  prob_eg = loadExample(dsc_ar,seed=prob$seed,scenario=as.character(prob$scenario),method=as.character(prob$method))
```

Now run ash 
```{r}
  set.seed(prob$seed+1)
  ash.res=list()
  ash.res[[1]] = ash(prob_eg$input$betahat,prob_eg$input$sebetahat,mixcompdist="halfuniform")
  for(i in 2:11){ash.res[[i]] = ash(prob_eg$input$betahat,prob_eg$input$sebetahat,mixcompdist="halfuniform",randomstart=TRUE)}
```

```{r}
  loglik = mapply(get_loglik,ash.res)
  loglik
```

```{r}
  best=which.max(loglik)
  plot(ash.res[[best]]$lfsr,ash.res[[1]]$lfsr,xlab="lfsr (best run)",ylab="lfsr (default run)",main="Comparison of lfsr (best vs default run)")
  abline(a=0,b=1,col=2)
  plot(ash.res[[best]]$PosteriorMean,ash.res[[1]]$PosteriorMean,xlab="Posterior Mean (best run)",ylab="Posterior Mean (default run)",main="Comparison of Posterior Means (best vs default run)")
  abline(a=0,b=1,col=2)

```

Compare Cdfs:
```{r}
  x = seq(-5,5,length=100)
  cdf_res = data.frame(x=x,
                       best=as.numeric(cdf.ash(ash.res[[best]],x)$y),
                       default=as.numeric(cdf.ash(ash.res[[1]],x)$y))
  cdf_res=melt(cdf_res,id.vars="x",value.name="y",variable.name="run")
  qplot(x=x,y=y,data=cdf_res,geom="line",colour=run,main="comparison of estimated cdf (best run and default run)")
```

### 2015-04-08 paper/main.tex
lots of changes

--- 2015-02-26 paper/main.tex
@@
  line defines the uniform component that is due to the null $p$ values, with the remainder necessarily corresponding to
- the alternative $p$ values. Now, for any given threshold, although we do not know {\it which} $p$ values correspond
+ the alternative $p$ values. In particular, the height of the horizontal line is an estimate of $\pi_0$, the proportion of $p$ values that come from the null.
+  Now, for any given threshold, although we do not know {\it which} $p$ values correspond
  to  null tests  and which to alternative tests, the areas of the two components indicate {\it how many} $p$ values correspond to each,
- allowing the FDR to be estimated (Figure \ref{fig:qvalue}).
+ allowing the FDR to be estimated (Figure \ref{fig:qvalue}). Specifically, at threshold $\gamma$, and given an estimate $\hat\pi_0$ for $\pi_0$,  a
+ natural non-parametric estimate of the FDR is
+ \begin{equation} \label{eqn:FDRest}
+\hat{\FDR}(\gamma) = J \hat\pi_0 \gamma / \#\{ j: p_j \leq \gamma \}.
+ \end{equation}
+ See \cite{storey:2002direct}, equation (8).
 
@@
 distribution of $z$ scores, using a mixture of normal distributions. In principle this approach also differs from \qvalue and \locfdr in
-that it does not explicitly make the Zero Assumption. However, we have found that in practice the results often approximately satisfy the ZA,
+that it does not explicitly make the ZA. However, we have found that in practice the results often approximately satisfy the ZA,
 which we believe to be due to details in the fitting procedure used.
@@
 \end{equation}
-where $\delta_0(\cdot)$ denotes a point mass at 0, $\pi=(\pi_0,\dots,\pi_K)$ are mixture proportions to be estimated (with $\pi_j \geq 0, \sum_j \pi_j=1$), and $f_k$ are pre-specified 
+where $\delta_0(\cdot)$ denotes a point mass at 0, $\pi=(\pi_0,\dots,\pi_K)$ are mixture proportions to be estimated (with $\pi_k \geq 0, \sum_k \pi_k=1$), and $f_k$ are pre-specified 
 component distributions with one of the following forms: 
@@
 
-Here we consider several effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
+Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
 standard error, $s_j=1$, and we consider the normal limit (\ref{eqn:normallik}). That
@@
 \begin{equation} \label{eqn:bhatgivenbz}
-\bhat_j | \beta_j \sim N(0,s^2_j=1)
-\end{equation}
-so $z_j:=\bhat_j/s_j$ are standard normal if $H_j$ holds.
+\bhat_j | \beta_j \sim N(\beta_j,s^2_j=1)
+\end{equation}
+so $z_j:=\bhat_j/s_j=\bhat_j$ are standard normal if $H_j$ holds.
 
@@
 \begin{enumerate}
-\item The UA can produce very different inferences compared with existing methods, and is generally less conservative.
-\item The UA can still yield conservative estimates of the proportion of nulls that hold, $\pi_0$, and hence conservative estimates of FDR and fdr.
-\item The UA results in a numerically highly-stable procedure, especially compared with a fully non-parametric approach.
+\item The UA can produce very different inferences compared with the ZA.
+\item The UA can yield conservative estimates of the proportion of nulls that hold, $\pi_0$, and hence conservative estimates of FDR and fdr.
+\item The UA results in a stable procedure, both numerically and statistically.
 \item The UA is somewhat robust to deviations from unimodality.
@@
 
-\subsubsection{The UA can produce very different inferences compared with existing methods}
-
-To illustrate that the UA can have a big impact on inference 
-we simulate data where the true $\beta_j$ are distributed as $N(0,1)$ (and then $\bhat_j | \beta_j \sim N(\beta_j,1)$.
-We feed the resulting $z$ scores to the R packages \locfdr and \mixfdr, and to our package \ashr, and the corresponding $p$ values to the R package \qvalue.
-
-TODO: put FDRs on graph? Simulate with 50\% null?
+\subsubsection{The UA compared with the ZA}
 
@@
 The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ashr) assumes that the effect sizes (and thus the $z$ scores) have a unimodal distribution about 0, resulting in a very different decomposition that in this case is much closer to the truth.
-(These data were simulated under a model where all of the effects are non-zero -- specifically, $\beta_j \sim N(0,1)$, $s_j=1$, resulting in $z_j \sim N(0,2^2)$ -- so the ``true" decomposition would make everything light blue.)} \label{fig:ZA}
+In contrast the method introduced here (\ashr) assumes that the effect sizes (and thus the $z$ scores) have a unimodal distribution about 0, resulting in a very different decomposition.
+(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero -- specifically, $\beta_j \sim N(0,1^2)$, $s_j=1$, resulting in $z_j \sim N(0,2^2)$ -- so the ``true" decomposition would make everything light blue.)} \label{fig:ZA}
 \end{figure}
 
-\subsubsection*{Conservative estimates of $\pi_0$}
-
-One concern, given the
-
-\subsubsection*{Robustness to deviations from UA}
-
-Show that the q values are driven by changes in pi_0; even when not unimodal pretty close...
-
-\subsubsection*{The UA produces stable estimates}
-
-The EM algorithm is notorious for convergence to local optima. 
-But in this case, over hundreds of applications of the procedure, we have observed no serious problems 
-caused by potential multimodality.
+Figure \ref{fig:ZA} graphically illustrates the difference between the ZA made by existing methods,
+and the UA made by our method. The effects of the ZA are visually clear, producing a ``hole" 
+ in the alternative $z$ score distribution around 0 for \qvalue, \locfdr and \mixfdr; in contrast, the UA 
+ fills in this hole to produce a peak at 0.  Of course the null distribution also has a peak at 0, and the local fdr 
+ under the UA is still smallest for $z$ scores that are far from zero (i.e. large $z$ scores remain the ``most significant")
+ as one would hope.
+
+ How do these differences influence estimates of the FDR? To address this, we note that 
+ the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
+ of $\pi_0$}, the proportion of nulls (at least in this simple case where the 
+ $s_i$ are all identical). To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
+ strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
+ although the methods considered here differ from one another in many ways, having estimated $\pi_0$
+ they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, our limited
+ empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
+
+Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
+ and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
+Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
+At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
+overestimates of $\pi_0$, which will yield overestimates of the FDR).
+
+To confirm this we performed comprehensive simulation experiments across a range of alternative scenarios (i.e. values for the distribution of non-zero effects,
+which we denote $g_1$), and values
+for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:pi0sims}a, with details in Table \ref{table:scenarios}.
+They range from a ``spiky" distribution -- where many non-zero $\beta$ are
+too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
+flatter distribution, which is a normal distribution with large variance (``bignorm") -- where most non-zero $\beta$ are easily detected
+making reliable estimation of $\pi_0$ much easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
+which, although we view as generally unrealistic, we include to assess robustness of our methods to deviations from the UA.
+
+\begin{figure} \label{fig:pi0sims}
+\begin{center}
+\includegraphics[height=2in]{figures/scenarios_density.pdf} 
+\includegraphics[height=2in]{figures/pi0_est.pdf}
+\end{center}
+\caption{a) Summary of the alternative distributions used for different simulation scenarios.
+b) Comparison of true and estimated values of $\pi_0$ in these simulated scenarios.
+In general $\pi_0$ is impossible to estimate accurately, and all methods yield conservative (over-)estimates for $\pi_0$ (deliberately so, to ensure conservative estimates of FDR).
+As expected, when the UA holds, the \ashr method is least conservative, and hence most accurate.
+When the UA assumption does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative, but quite accurate, particularly when using a mixture of uniforms to model the unimodal distribution.}
+\end{figure}
+
+\begin{table}[!ht]
+\begin{tabular}{c c } \toprule
+Scenario & Alternative distribution, $g_1$  \\ \midrule
+spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
+near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
+flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
+ &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
+skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
+bignormal & $N(0,4^2)$ \\ 
+bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
+\end{tabular}
+\caption{Summary of simulation scenarios considered} \label{table:scenarios}
+\end{table}
+
+%In the results section we sometimes find it notationally convenient to rewrite (\ref{eqn:g}) as a mixture of a ``null" component $\delta_0$, and a non-null component, which we denote $g_1$:
+%\begin{equation} \label{eqn:g1}
+%g(\cdot; \pi) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot).
+%\end{equation}
+
+For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
+data as follows:
+\begin{enumerate}
+\item Simulate $\pi_0  \sim U[0,1]$.
+\item For $j=1,\dots,J$, simulate $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)$.
+\item For $j=1,\dots,J$, simulate $\bhat_j | \beta_j \sim N(\beta_j,1)$.
+\end{enumerate}
+Thus these simulations assume the same
+precision $(s_j=1)$ for each measurement.
+
+We applied the methods implemented in the R packages \qvalue, \locfdr, \mixfdr and \ashr to
+estimate $\pi_0$ for each data set. For \ashr we show results for $g$ modelled as a mixture of normal components (``ash.n") 
+and a mixture of uniform components (``ash.u"). (Results using a mixture of ``half-uniforms" are generally similar to the uniforms and omitted for clarity.)
+
+Figure \ref{fig:pi0} compares estimated and true values of $\pi_0$ under each Scenario. 
+For most scenarios no method reliably estimates $\pi_0$. This is expected because most scenarios include
+an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
+estimation of$\pi_0$ impossible. This is particularly true for the ``spiky" scenario, for example.  The
+exception is the ``bignorm" scenario, where the \ashr estimates of $\pi_0$ are quite accurate - made possible by the fact that, in this scenario, most non-zero effects
+are very different from zero.
+
+Although accurate estimation of $\pi_0$ is impossible in general, the simulation results demonstrate
+that, when the UA holds, \ashr consistently provides substantially smaller, more accurate, estimates of $\pi_0$, than existing methods. 
+As noted above, this means that estimates of FDR from \ashr will also be smaller and more accurate under these scenarios (while
+remaining conservative).
+When the UA assumption does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative, although quite accurate.
+We view such a strong bimodal scenario as rather unlikely in most applications where FDR methods
+are used; nonetheless, users may find it reassuring that even in this case the results from \ashr should not be horribly misleading.
+
+%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
+%
+%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
+%and so all effects are non-zero and no nulls hold. This scenario is rather different
+%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
+%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
+%``mostly non-null"
+%scenario, and we would agree! On the other hand, the distribution of $p$ values
+%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
+%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
+%Thus it seems desirable to have methods that perform well in both cases, and this is
+%an important feature of the methods we introduce here.
+%
+%TODO: put FDRs on graph? Simulate with 50\% null?
+
+\subsubsection*{Numerical Stability}
+
+The EM algorithm, which we use here to fit our model, is notorious for convergence to local optima. 
+However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
+caused by this behaviour. To quantify this, we ran \ashr 10 times on each of XX simulated datasets using a random initialization for $\pi$,
+in addition running it using our default initialization procedure (XX need to specify in methods?). We then compared the largest
+log-likelihood achieved across all 11 runs with the log-likelihood achieved by our default run.
+When using a mixture of normals (ash.n) the results were extremely stable: 95\% showed a negligible log-likelihood difference (< 0.02), 
+and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 80\% showed a negligible log-likelihood difference (<0.02),
+and 9\% of runs showed an appreciable log-likelihood difference (>1), with the largest difference being 5.2. However, perhaps suprisingly, 
+even for this largest difference results from the default run (e.g.
+the lfsrs, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood \url{github.com/stephens999/ash/dsc-stability/check_convergence.rmd}.
+
+\subsubsection*{The UA helps provide reliable estimates of $g$}
+
+Although we have focussed up to now on estimating the FDR, an important advantage of
+an empirical Bayes approach like ours, compared with less-model-based alternatives like \qvalue, is that the EB 
+approach can provide
+measures of the uncertainty in the {\it size} of each effect $\beta_j$.
+Specifically, the
+EB approach provides a posterior distribution for each $\beta_j$, which can be used
+to address any question about the effect sizes, including, for example, ``which effects likely exceed $T$'', where
+$T$ is some effect threshold of interest (and ``likely" can be quantified by whatever probability threshold is desired). 
+Further, because the posterior distribution is, by definition,
+conditional on the observed data, inferences based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
+been selected based on the observed data. %(This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.)
+In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
+\cite{benjamini.yekutieli:2005} which \cite{efron2008} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
+how can confidence intervals be assigned to the true
+effect size for each selected case?". \cite{efron2008} notes the potential for EB approaches to tackle this problem.
+%but also notes the difficulties of implementing a stable general algorithm.
+
+These features of the EB approach are extremely attractive in principle.
+But their usefulness in practice
+depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
+which is notoriously difficult to solve in general. Indeed, Efron emphasises
+the difficulties of implementing a stable general algorithm, noting in his rejoinder
+``the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying
+to be overly nonparametric in constructing the empirical
+Bayes Fdr estimates, but it is hard to imagine a
+generally satisfactory parametric formulation..." (Efron, rejoinder, p46).
+The key point we want to emphasise here is that a stable non-parametric
+approach is not only possible, but in fact quite straightforward, {\it provided we restrict ourselves
+to settings where we can assume that $g$ is unimodal}.
+While not meeting Efron's desire for an entirely general nonparametric approach, we nonetheless believe 
+that these settings include many cases of practical interest.
+
+To illustrate this, Figure \ref{fig:egcdf} compares
+the estimated $g$ from our method for each of the simulation scenarios above with the 
+estimated $g$ from mixFdr which does not make the UA (and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
+Even with only one simulated example for each scenario, the greater reliability of estimates afforded
+by the UA is immediately apparent.
+
+%making the UA provides an attractive and simple way to achieve estimating $g$  large
+%subset of cases of interest (those where the UA can reasonably be assumed to hold).
+
+%In practice, the utility of the posterior distributions noted above depends primarily on the accuracy
+%of the estimate of the distribution $g(\cdot)$.
+
+ %quote from efron rejoinder , p46:
+%MorrisÕ Section 3 is especially pertinent. His formula
+%for p(?|z) is related to my discussion of the
+%BenjaminiÐYekutieli False Coverage Rate method in
+%Section 7, particularly (7.2)Ð(7.4). Originally I had
+%hoped to develop an empirical Bayes method for estimating
+%such models, but the effort foundered on practical
+%difficulties involving the perils of deconvolution.
+%Section 6 on the Òone-group modelÓ is the ugly duckling
+%of the current paper, but it bears on some important
+%points raised in the discussion. Figure 7 concerns
+%a fuzzy version of simultaneous hypothesis testing,
+%where, as in MorrisÕ hospital example, the usual singlepoint
+%null hypotheses seem unequal to the task. The
+%development from (6.6) onwards, particularly (6.12),
+%bears on the possibility of nonnormal null distributions,
+%and is about as far as I can go in answering Professors
+%Rice and SpiegelhalterÕs penultimate question.
+%With g(?) a normal distribution, model (6.1) returns
+%us to the realm of Stein estimation, the scene of my
+%happy collaborations with Carl Morris. I continue to
+%be surprised at how much harder bumpy, nonnormal
+%models like (7.1) are to deal with. JamesÐStein estimation
+%works fine with, say, N = 10 component problems,
+%while the RobbinsÕ form of empirical Bayes appropriate
+%to (7.1) seems to require hundreds or thousands.
+%The information calculations in Efron (2008)
+%reinforce this gloomy assessment. Maybe I am trying
+%to be overly nonparametric in constructing the empirical
+%Bayes Fdr estimates, but it is hard to imagine a
+%generally satisfactory parametric formulation for (6.1).
+%Or perhaps it is just that hypothesis testing is more demanding
+%than estimation.
+%My Note: (6.1) is the general model (mu \sim g()).
+%7.1 is the case where the effects are a mixture of zero and N(-3,1).
+
+%deconvolution
+%calculations are notoriously tricky and I have not
+%been able to produce a stable general algorithm. ; see also 
+Zhao+Hwang consider in detail the case where the non-null effects are normally distributed.
+
+Results in efron.fcr.rmd suggests that lower CI is OK; upper is more problematic because of difficulty in accurately estimating tail?
+
+To do: implement methods to compute posterior cdf of each observation at given points. 
+Then compute these for the simulation studies. Check whether issue with upper bound is due to non-unimodal distribution.
+
+\begin{figure} \label{fig:egcdf}
+\includegraphics[height=2in]{figures/egcdf.pdf}
+\includegraphics[height=2in]{figures/mean_cdf.pdf}
+\end{figure}
+
+\subsubsection*{Reverse asymptotics}
+
+Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
+
+Note: we'd like to make sure it behaves ok in the case where most observations are null.
+Perhaps simulate with 999 null and 1 alternative?
+Also maybe look at performance for small sample sizes - e.g. n=1?
+
+Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
+can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
+by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
+
+\subsection*{Improved conservative estimation of $\pi_0$}
+\subsection*{Local False Sign Rate}
+
+There are two reasons to use the lfsr instead of the lfdr: it is more generally meaningful (e.g. it applies
+whether or not zero effects exist),  and estimation of lfsr is more robust to modeling assumptions and estimation of $\pi_0$. 
+To illustrate this, we compared estimated and true values of both lfdr and lfsr for the simulated data (where the true
+values are computed by Bayes Theorem using the true value of $g$ FILL IN THESE DETAILS?).
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[height=320px,width=540px]{Rcode/figures/lfsdr_sim1sim2_blowup.png}
+\end{center}
+\caption{Figure showing lfsr is more robust than lfdr}
+\end{figure}
+
+- Illustration that estimated lfsr is more robust.
+
+- Illustration that adjusted estimate of lfsr is still more robust..
+
+- Compare estimated lfsr with true lfsr for difficult simulation case, to show nearly achieves Bayes Risk?
+
+%\subsection{Accounting for measurement precision improves estimates of FDR/fsr}
 
@@
 
-Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
-
-Note: we'd like to make sure it behaves ok in the case where most observations are null.
-Perhaps simulate with 999 null and 1 alternative?
-Also maybe look at performance for small sample sizes - e.g. n=1?
-
-\subsection*{Improved conservative estimation of $\pi_0$}
-
-To illustrate estimation of $\pi_0$ we provide simulation results for two scenarios:
-\begin{enumerate}
-\item[Scenario 1:]
-\begin{equation}
-f_1(\cdot) =  0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2)
-\end{equation}
-\item[Scenario 2:]
-\begin{equation}
-f_1(\cdot) = N(0,4^2)
-\end{equation}
-\end{enumerate}
-Scenario 1 represents a ``difficult" case where many non-zero $\beta$ are
-too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible; 
-Scenario 2 represents an ``easier" case where most non-zero $\beta$ are
-sufficiently different from zero that they can be reliably detected. making estimation of $\pi_0$ easier.
-For Scenario 1 we considered datasets of size $n=1000$ (Scenario 1a) and $n=10,000$ (Scenario 1b);
-for Scenario 2 we used $n=1000$.
-
-For each simulation scenario we simulated 200 independent data sets. For each data set we simulated
-the true value of $\pi_0  \sim U[0,1]$, and then simulated $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) f_1(\cdot)$
-and $\bhat_j | \beta_j = N(\beta_j,1)$ for $j=1,\dots,n$. Thus these simulations assume the same
-precision for each measurement. We applied the methods implemented in the R packages qvalue, fdrtool, locfdr, and \mixfdr
-for estimating $\pi_0$. We also applied our method to estimate $\pi_0$ in two ways, one using a penalty encourage
-$\pi_0$ to be as large as possible (``null-biased"), and the other using the maximum likelihood estimate for $\pi$.
-
-Figure \ref{fig:pi0} compares estimated and true values of $\pi_0$ under each Scenario. 
-For Scenario 1, no method reliably estimates $\pi_0$. This is expected since, the scenario was designed to make accurate estimation
-of $\pi_0$ impossible. However, we see that all the methods except for ash (mle) are able to provide a conservative
-estimate for $\pi_0$. Further, among these the estimates from ash (null-biased) are least conservative. Scenario 2 produces similar patterns,
-except that for this scenario the ash estimates of $\pi_0$ are quite accurate (made possible by the fact that most non-zero effects
-are very different from zero).
-
-These simulations illustrate two key points: i) although $\pi_0$ is not identifiable, the penalized likelihood approach
-provides conservative estimates for $\pi_0$; ii) replacing the ``zero assumption" with the UA can produce
-less conservative, more accurate, estimates of $\pi_0$ (provided of course that the UA holds, as it does in these simulations).
-
-\begin{figure}[!ht] \label{fig:pi0}
-\begin{center}
-\includegraphics[height=160px,width=540px]{Rcode/figures/estpi0_sim1sim2.png}
-\end{center}
-\caption{Comparison of true and estimated values of $\pi_0$ for simulation scenarios.
-Scenarios 1a and 1b represent ``difficult" scenarios where $\pi_0$ is impossible to estimate accurately.
-However, all methods except the ash mle method are successful in providing conservative estimates for $\pi_0$.
-The ash (null-biased) method is least conservative, and hence most accurate, due to its additional assumption that the effect size distribution $g$ is unimodal.}
-\end{figure}
-
-Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
-can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
-by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
-
-\subsection*{Local False Sign Rate}
-
-There are two reasons to use the lfsr instead of the lfdr: it is more generally meaningful (e.g. it applies
-whether or not zero effects exist),  and estimation of lfsr is more robust to modeling assumptions and estimation of $\pi_0$. 
-To illustrate this, we compared estimated and true values of both lfdr and lfsr for the simulated data (where the true
-values are computed by Bayes Theorem using the true value of $g$ FILL IN THESE DETAILS?).
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[height=320px,width=540px]{Rcode/figures/lfsdr_sim1sim2_blowup.png}
-\end{center}
-\caption{Figure showing lfsr is more robust than lfdr}
-\end{figure}
-
-- Illustration that estimated lfsr is more robust.
-
-- Illustration that adjusted estimate of lfsr is still more robust..
-
-- Compare estimated lfsr with true lfsr for difficult simulation case, to show nearly achieves Bayes Risk?
-
-\subsection{Accounting for measurement precision improves estimates of FDR/fsr}
-
-- Illustration of effects of contamination on FDR by low-quality data
-comparison with ash.
-
 \subsection{Comparison of different underlying component distributions}

### 2015-04-08 dsc-shrink/addMethods.R

@@ -5,8 +5,22 @@ addMethod(dsc_shrink,name="ash.hu",fn =ash.wrapper,args=list(mixcompdist="halfun
 addMethod(dsc_shrink,name="ash.u",fn =ash.wrapper,args=list(mixcompdist="unif"),outputtype = "ash_output")
 addMethod(dsc_shrink,name="ash.n",fn =ash.wrapper,args=list(mixcompdist="normal"),outputtype = "ash_output")
 
+addMethod(dsc_shrink,name="ash.hu.nw2",fn =ash.wrapper,args=list(mixcompdist="halfunif",nullweight=2,cxx=TRUE),outputtype = "ash_output")
+addMethod(dsc_shrink,name="ash.u.nw2",fn =ash.wrapper,args=list(mixcompdist="unif",nullweight=2,cxx=TRUE),outputtype = "ash_output")
+addMethod(dsc_shrink,name="ash.n.nw2",fn =ash.wrapper,args=list(mixcompdist="normal",nullweight=2,cxx=TRUE),outputtype = "ash_output")
+
+
+addMethod(dsc_shrink,name="ash.hu.s",fn =ash.wrapper,args=list(mixcompdist="halfunif",method="shrink"),outputtype = "ash_output")
+addMethod(dsc_shrink,name="ash.u.s",fn =ash.wrapper,args=list(mixcompdist="unif",method="shrink"),outputtype = "ash_output")
+addMethod(dsc_shrink,name="ash.n.s",fn =ash.wrapper,args=list(mixcompdist="normal",method="shrink"),outputtype = "ash_output")
+
+
 addMethod(dsc_shrink,name="mixfdr.tnull", fn=mixfdr.wrapper, args = list(theonull=TRUE),outputtype = "mixfdr_output")
 addMethod(dsc_shrink,name="mixfdr.enull", fn=mixfdr.wrapper, args = list(theonull=FALSE),outputtype = "mixfdr_output")
+
+addMethod(dsc_shrink,name="locfdr", fn=locfdr.wrapper,outputtype = "locfdr_output")
+addMethod(dsc_shrink,name="qvalue", fn=qvalue.wrapper,outputtype = "qvalue_output")
+
 #addMethod(dsc_shrink,name="mixfdr.tnull.J10", fn=mixfdr.wrapper, args = list(theonull=TRUE,J=10),outputtype = "mixfdr_output")
[…]

### 2015-04-08 dsc-shrink/addMethods.null.R

@@ -1,0 +1,20 @@
+sourceDir("methods")
+
+addMethod(dsc_znull,name="ash.hu",fn =ash.wrapper,args=list(mixcompdist="halfunif"),outputtype = "ash_output")
+addMethod(dsc_znull,name="ash.u",fn =ash.wrapper,args=list(mixcompdist="unif"),outputtype = "ash_output")
+addMethod(dsc_znull,name="ash.n",fn =ash.wrapper,args=list(mixcompdist="normal"),outputtype = "ash_output")
+
+addMethod(dsc_znull,name="ash.hu.s",fn =ash.wrapper,args=list(mixcompdist="halfunif",method="shrink"),outputtype = "ash_output")
+addMethod(dsc_znull,name="ash.u.s",fn =ash.wrapper,args=list(mixcompdist="unif",method="shrink"),outputtype = "ash_output")
+addMethod(dsc_znull,name="ash.n.s",fn =ash.wrapper,args=list(mixcompdist="normal",method="shrink"),outputtype = "ash_output")
+
+addMethod(dsc_znull,name="ash.hu.nw2",fn =ash.wrapper,args=list(mixcompdist="halfunif",nullweight=2),outputtype = "ash_output")
+addMethod(dsc_znull,name="ash.u.nw2",fn =ash.wrapper,args=list(mixcompdist="unif",nullweight=2),outputtype = "ash_output")
+addMethod(dsc_znull,name="ash.n.nw2",fn =ash.wrapper,args=list(mixcompdist="normal",nullweight=2),outputtype = "ash_output")
+
+
+addMethod(dsc_znull,name="ash.hu.s.gmfine",fn =ash.wrapper,args=list(mixcompdist="halfunif",method="shrink",gridmult=2^.25),outputtype = "ash_output")
+addMethod(dsc_znull,name="ash.u.s.gmfine",fn =ash.wrapper,args=list(mixcompdist="unif",method="shrink",gridmult=2^.25),outputtype = "ash_output")
+addMethod(dsc_znull,name="ash.n.s.gmfine",fn =ash.wrapper,args=list(mixcompdist="normal",method="shrink",gridmult=2^.25),outputtype = "ash_output")
+
+

### 2015-04-08 dsc-shrink/addScenarios.null.R

@@ -1,0 +1,17 @@
+sourceDir("datamakers")
+
+
+#Now, for each scenario create an element of scenarios of the following form
+addScenario(dsc_znull,name="znull.1000",
+            fn=null_z_datamaker,
+            args=list(
+              nsamp=1000
+            ),
+            seed=1:500)
+
+addScenario(dsc_znull,name="znull.100",
+            fn=null_z_datamaker,
+            args=list(
+              nsamp=100
+            ),
+            seed=1:500)

### 2015-04-08 dsc-shrink/addScenarios10.R

@@ -8,7 +8,7 @@ addScenario(dsc_shrink,name="A",
                       g=normalmix(c(2/3,1/3),c(0,0),c(1,2)),
                       min_pi0=0,
                       max_pi0=0,
-                      nsamp=1000,
+                      nsamp=10,
                       betahatsd=1
                       ),
                     seed=1:100)
@@ -19,7 +19,7 @@ addScenario(dsc_shrink,name="B",
                       g=normalmix(rep(1/7,7),c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7)),
                       min_pi0=0,
                       max_pi0=0,
-                      nsamp=1000,
+                      nsamp=10,
                       betahatsd=1
                     ),
                     seed=1:100)
@@ -31,7 +31,7 @@ addScenario(dsc_shrink,name="C",
                         g=normalmix(c(1/4,1/4,1/3,1/6),c(-2,-1,0,1),c(2,1.5,1,1)),
                         min_pi0=0,
                         max_pi0=0,
-                        nsamp=1000,
+                        nsamp=10,
                         betahatsd=1
                       ),
                       seed=1:100)
@@ -43,7 +43,7 @@ addScenario(dsc_shrink,name="An",
                       g=normalmix(c(2/3,1/3),c(0,0),c(1,2)),
                       min_pi0=0,
                       max_pi0=1,
-                      nsamp=1000,
+                      nsamp=10,
                       betahatsd=1
                     ),
                     seed=1:100)
@@ -54,7 +54,7 @@ addScenario(dsc_shrink,name="Bn",
                       g=normalmix(rep(1/7,7),c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7)),
                       min_pi0=0,
                       max_pi0=1,
-                      nsamp=1000,
+                      nsamp=10,
                       betahatsd=1
                     ),
                     seed=1:100)
@@ -66,7 +66,7 @@ addScenario(dsc_shrink,name="Cn",
                         g=normalmix(c(1/4,1/4,1/3,1/6),c(-2,-1,0,1),c(2,1.5,1,1)),
                         min_pi0=0,
                         max_pi0=1,
-                        nsamp=1000,
+                        nsamp=10,
                         betahatsd=1
                       ),
                       seed=1:100)
@@ -77,7 +77,7 @@ addScenario(dsc_shrink,name="hard",
                         g=normalmix(c(.4,.2,.2,.2),c(0,0,0,0),c(.25,.5,1,2)),
[…]

### 2015-04-08 dsc-shrink/addScenarios100.R

@@ -8,7 +8,7 @@ addScenario(dsc_shrink,name="A",
                       g=normalmix(c(2/3,1/3),c(0,0),c(1,2)),
                       min_pi0=0,
                       max_pi0=0,
-                      nsamp=1000,
+                      nsamp=100,
                       betahatsd=1
                       ),
                     seed=1:100)
@@ -19,7 +19,7 @@ addScenario(dsc_shrink,name="B",
                       g=normalmix(rep(1/7,7),c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7)),
                       min_pi0=0,
                       max_pi0=0,
-                      nsamp=1000,
+                      nsamp=100,
                       betahatsd=1
                     ),
                     seed=1:100)
@@ -31,7 +31,7 @@ addScenario(dsc_shrink,name="C",
                         g=normalmix(c(1/4,1/4,1/3,1/6),c(-2,-1,0,1),c(2,1.5,1,1)),
                         min_pi0=0,
                         max_pi0=0,
-                        nsamp=1000,
+                        nsamp=100,
                         betahatsd=1
                       ),
                       seed=1:100)
@@ -43,7 +43,7 @@ addScenario(dsc_shrink,name="An",
                       g=normalmix(c(2/3,1/3),c(0,0),c(1,2)),
                       min_pi0=0,
                       max_pi0=1,
-                      nsamp=1000,
+                      nsamp=100,
                       betahatsd=1
                     ),
                     seed=1:100)
@@ -54,7 +54,7 @@ addScenario(dsc_shrink,name="Bn",
                       g=normalmix(rep(1/7,7),c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7)),
                       min_pi0=0,
                       max_pi0=1,
-                      nsamp=1000,
+                      nsamp=100,
                       betahatsd=1
                     ),
                     seed=1:100)
@@ -66,7 +66,7 @@ addScenario(dsc_shrink,name="Cn",
                         g=normalmix(c(1/4,1/4,1/3,1/6),c(-2,-1,0,1),c(2,1.5,1,1)),
                         min_pi0=0,
                         max_pi0=1,
-                        nsamp=1000,
+                        nsamp=100,
                         betahatsd=1
                       ),
                       seed=1:100)
@@ -77,7 +77,7 @@ addScenario(dsc_shrink,name="hard",
                         g=normalmix(c(.4,.2,.2,.2),c(0,0,0,0),c(.25,.5,1,2)),
[…]

### 2015-04-08 dsc-shrink/methods/locfdr.wrapper.R

@@ -1,0 +1,24 @@
+#' @title wrapper for locfdr 
+#'
+#' @description Runs locfdr on z scores
+#' @details None
+#' 
+#' @param input a list with elements betahat and sebetahat
+#' @param args a list containing other additional arguments to locfdr
+#' 
+#' @return output a list 
+#'
+library(locfdr)
+
+locfdr.wrapper=function(input,args=NULL){
+  res = try(do.call(locfdr, args= c(list(zz=input$betahat/input$sebetahat, nulltype=0, plot=0),args)))
+  if(inherits(res,"try-error")){res=list(fdr=rep(NA,length(input$betahat)),fp0=matrix(NA,nrow=6,ncol=3))}
+  return(res)
+}
+
+
+locfdr2pi0_est = function(output){
+  return (list(pi0_est=output$fp0[1,3]))
+}
+
+

### 2015-04-08 dsc-shrink/methods/mixfdr.wrapper.R

@@ -26,4 +26,9 @@ mixfdr2pi0_est = function(output){
   return (list(pi0_est=output$res$pi[1]))
 }
 
+mixfdr2fitted.g = function(output){
+  return (list(fitted.g=
+    normalmix(output$res$pi,output$res$mu,
+              output$res$sigma-output$res$noiseSD)))
+}
 

### 2015-04-08 dsc-shrink/methods/qvalue.wrapper.R

@@ -1,0 +1,27 @@
+#' @title wrapper for qvalue for FDR DSC
+#'
+#' @description Runs qvalue to compute FDR/q values  
+#' @details None
+#' 
+#' @param input a list with elements betahat and sebetahat
+#' @param add.args a list with additional arguments to qvalue 
+#' 
+#' @return output a list with the following elements
+#' \item{qvalue}{vector of qvalues, with jth element being the q value corresponding to (betahat_j,sebetahat_j)
+#'
+#'
+library(qvalue)
+
+qvalue.wrapper = function(input,args=NULL){
+  zscore = input$betahat/input$sebetahat
+  pvalue = 2*pnorm(-abs(zscore),lower.tail=TRUE)
+  res = qvalue(pvalue)
+  return(res)
+}
+
+qvalue2pi0_est = function(output){
+  if(!is.list(output)){return(list(pi0_est=NA))} #deal with case where ERROR thrown
+  else{
+    return (list(pi0_est=output$pi0))
+  }
+}

### 2015-04-08 dsc-shrink/plot_cdf.R

@@ -1,0 +1,67 @@
+library(plyr)
+library(dplyr)
+library(dscr)
+library(magrittr)
+library(ashr)
+library(ggplot2)
+source("methods/mixfdr.wrapper.R")
+load("res.RData")
+
+MEANPLOTFILE = "../paper/figures/meancdf.pdf"
+PLOTSCENARIOS = c("hard","An","Bn","Cn","easy")
+PLOTNAMES = c("spiky","near-normal","flat-top","skew","big-normal")
+PLOTMETHODS = c("ash.n","ash.hu","truth","mixfdr.tnull")
+SEEDS = 1
+
+#df is a list with components method,seed, scenario
+#cdf is evaluated at x
+#returns list of x,y,pi0, where cdf values are in y
+get_cdf = function(df,dsc=dsc_shrink,x=seq(-6,6,length=100)){
+  m=df$method
+  if(m=="truth"){m="ash.n"}
+  temp=loadExample(dsc,df$seed,df$scenario,m)
+  pi0=temp$meta$pi0
+  
+  if(df$method=="truth"){
+    s=dsc$scenarios[[df$scenario]]
+    galt=s$args$g
+    g=normalmix(c(pi0,(1-pi0)*mixprop(galt)),
+                c(0,comp_mean(galt)),
+                c(0,comp_sd(galt)))
+  } else {
+    if(grepl("mixfdr",df$method)){
+      temp$output= mixfdr2fitted.g(temp$output)
+    }
+    g=temp$output$fitted.g
+  }
+  data.frame(x=x,y=as.numeric(mixcdf(g,x)),pi0=pi0)
+}
+
+#set up dataframe with cdf for all methods and all datasets
+df= expand.grid(seed=SEEDS,scenario=PLOTSCENARIOS, method=PLOTMETHODS,stringsAsFactors = FALSE)
+
+df.cdf=ddply(df,.(seed,scenario,method), get_cdf)
+df.cdf.mean = df.cdf %>% group_by(x,method,scenario) %>% summarise(y=mean(y))
+df.midpi0.cdf.mean = df.cdf %>% filter(pi0<0.55 & pi0>0.45) %>% group_by(x,method,scenario) %>% summarise(y=mean(y))
+
+pdf(MEANPLOTFILE)
+
+df.cdf$scenario=factor(df.cdf$scenario,levels=PLOTSCENARIOS)
+levels(df.cdf$scenario)=PLOTNAMES
[…]

### 2015-04-08 dsc-shrink/plot_egdens.R

@@ -1,0 +1,30 @@
+library("ashr")
+library("ggplot2")
+library("dplyr")
+library("dscr")
+load("res.RData")
+PLOTSCENARIOS = c("hard","A","B","C","easy","bimodal")
+PLOTNAMES = c("spiky","near-normal","flat-top","skew","big-normal","bimodal")
+PLOTFILE = "../paper/figures/scenarios_density.pdf"
+
+df=data.frame()
+
+for(i in PLOTSCENARIOS){
+  s=dsc_shrink$scenarios[[i]]
+  g=s$args$g
+  x = seq(-6,6,length=100)
+  y = as.numeric(dens(g,x))
+  df = rbind(df,data.frame(x=x,y=y,scenario=i))
+}
+
+
+df$scenario = factor(df$scenario,levels=PLOTSCENARIOS)
+levels(df$scenario)= PLOTNAMES
+  
+pdf(PLOTFILE,height=3,width=9)
+ggplot(df, aes(x=x,y=y)) + geom_line(size=1.2,linetype=1) + facet_grid(.~scenario) + ylab("density")
+dev.off()
+
+
+  
+ 

### 2015-04-08 dsc-shrink/run_dsc100.R

@@ -1,7 +1,7 @@
 library(dscr)
 
-dsc_shrink=new.dsc("shrink","dsc-shrink-files")
-source("addScenarios.R")
+dsc_shrink=new.dsc("shrink100","dsc-shrink100-files")
+source("addScenarios100.R")
 source("addMethods.R")
 
 addOutputParser(dsc_shrink,"ash2beta",ash2beta_est,"ash_output","beta_est_output")
@@ -9,11 +9,9 @@ addOutputParser(dsc_shrink,"mixfdr2beta",mixfdr2beta_est,"mixfdr_output","beta_e
 
 addOutputParser(dsc_shrink,"ash2pi0",ash2pi0_est,"ash_output","pi0_est_output")
 addOutputParser(dsc_shrink,"mixfdr2pi0",mixfdr2pi0_est,"mixfdr_output","pi0_est_output")
-addOutputParser(dsc_shrink,"locfdr2pi0",locfdr2pi0_est,"locfdr_output","pi0_est_output")
-addOutputParser(dsc_shrink,"qvalue2pi0",qvalue2pi0_est,"qvalue_output","pi0_est_output")
 
 addOutputParser(dsc_shrink,"ash2fitted.g",ash2fitted.g,"ash_output","g_output")
-addOutputParser(dsc_shrink,"mixfdr2fitted.g",mixfdr2fitted.g,"mixfdr_output","g_output")
+#addOutputParser(dsc_shrink,"mixfdr2fitted.g",mixfdr2fitted.g,"mixfdr_output","g_output")
 
 
 
@@ -27,7 +25,12 @@ addScore(dsc_shrink,score_fdr,"fdr","mixfdr_output") #just extracts the fdr
 addScore(dsc_shrink,score_betahat,"betahat","mixfdr_output") #just extracts the fdr
 
 
-res=run_dsc(dsc_shrink)
-save(res,dsc_shrink,file="res.RData")
+
+res100=run_dsc(dsc_shrink)
+save(res100,dsc_shrink,file="res100.RData")
+
+
+xtabs(pi0_est ~ method+scenario,res100$pi0_score)
+xtabs(user.self ~ method+scenario,res100$pi0_score)
 
 

### 2015-04-08 dsc-shrink/run_dsc_znull.R

@@ -1,0 +1,60 @@
+#the aim of this dsc is to assess the distribution of the logLR 
+#value for ash under the null that all effects are 0
+
+library(dscr)
+
+dsc_znull=new.dsc("znull","dsc-znull-files")
+source("addScenarios.null.R")
+source("addMethods.null.R")
+
+source("score.R")
+addScore(dsc_znull,score_logLR,"logLR","ash_output")
+
+
+res=run_dsc(dsc_znull)
+save(res,dsc_znull,file="res.znull.RData")
+
+
+
+library(ggplot2)
+library(dplyr)
+temp=ungroup(res %>% group_by(scenario,method) %>% summarise(gt0=mean(logLR>0)))
+xtabs(gt0~method+scenario,temp)
+
+# scenario
+# method            znull.100 znull.1000
+# ash.hu              0.438      0.742
+# ash.hu.nw2          0.814      0.916
+# ash.hu.s            0.988      0.998
+# ash.hu.s.gmfine     0.978      0.996
+# ash.n               0.170      0.352
+# ash.n.nw2           0.330      0.496
+# ash.n.s             0.476      0.490
+# ash.n.s.gmfine      0.472      0.502
+# ash.u               0.210      0.424
+# ash.u.nw2           0.356      0.534
+# ash.u.s             0.482      0.570
+# ash.u.s.gmfine      0.484      0.562
+# 
+# Note that the ash.n.s basically conforms exactly to the null expectation
+# of 0.5 chi2_0 +0.5 chi2_1 from Stram and Lee, Biometrics
+# conjecture: the uniform has the same asymptotic behaviour
+# and the half uniform is like the sum of two of these?
+
+#from this
+#> qchisq(0.9,df=1)
+#[1] 2.705543
+# so a 95% procedure would need to check for logLR > 2.7/2 = 1.35
+
+
+pdf("qplots.znull.pdf")
+ggplot(res %>% filter(logLR!=0) %>% filter(grepl("ash.hu",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
+  stat_qq(dist = qchisq, dparams=list(df=1)) + geom_abline() + ggtitle("qqplot under null vs chisq-1; line slope=1")
+ggplot(res %>% filter(logLR!=0) %>% filter(grepl("ash.u",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
[…]

### 2015-04-08 dsc-shrink/set_plot_colors.R

@@ -1,0 +1,6 @@
+library("RColorBrewer")
+library("ggplot2")
+myColors <- brewer.pal(7,"Set1")
+names(myColors) <- c("mixfdr.tnull","ash.hu","ash.n","ash.u","qvalue","locfdr","truth")
+colScale <- scale_colour_manual(name = "method",values = myColors)
+

### 2015-04-08 dsc-stability/datamakers/datamaker.R

@@ -1,0 +1,42 @@
+library(ashr)
+
+#' @title datamaker for FDR DSC
+#'
+#' @description Simulates data for a DSC for methods to compute FDR/q values  
+#' @details None
+#' 
+#' @param seed The seed for the pseudo-rng set before generating the parameters
+#' @param args A list of the remaining arguments, which in this case is
+#' \item{nsamp}{The number of samples to create}
+#' \item{g}{An object of class normalmix specifying the mixture distribution from which non-null beta values 
+#' are to be simulated}
+#' \item{min_pi0}{The minimum value of pi0, the proportion of true nulls}
+#' \item{max_pi0}{The maximum value of pi0, the proportion of true null}
+#' \item{betahatsd}{The standard deviation of betahat to use}
+#'
+#' @return a list with the following elements
+#' \item{meta}{A list containing the meta data. In this case beta}
+#' \item{input}{A list containing the input data; in this case the set of betahat values and their standard errors}
+#' 
+rnormmix_datamaker = function(args){
+  #here is the meat of the function that needs to be defined for each dsc to be done
+  pi0 = runif(1,args$min_pi0,args$max_pi0) #generate the proportion of true nulls randomly
+  
+  k = ncomp(args$g)
+  comp = sample(1:k,args$nsamp,mixprop(args$g),replace=TRUE) #randomly draw a component
+  isnull = (runif(args$nsamp,0,1) < pi0)
+  beta = ifelse(isnull, 0,rnorm(args$nsamp,comp_mean(args$g)[comp],comp_sd(args$g)[comp]))
+  sebetahat = args$betahatsd
+  betahat = beta + rnorm(args$nsamp,0,sebetahat)
+  meta=list(beta=beta,pi0=pi0)
+  input=list(betahat=betahat,sebetahat=sebetahat)
+  
+  #end of meat of function
+  
+  data = list(meta=meta,input=input)
+  
+  return(data)
+  
+}
+
+

### 2015-04-08 dsc-stability/methods/ash.wrapper.R

@@ -1,0 +1,28 @@
+#' @title wrapper for ash for shrinkage DSC
+#'
+#' @description Runs ash to compute betahat values  
+#' @details None
+#' 
+#' @param input a list with elements betahat and sebetahat
+#' @param args a list containing other additional arguments to ash
+#' 
+#' @return output a list containing a vector of loglikelihoods from 10 independent runs of ash
+#'
+library(ashr)
+
+ash.wrapper=function(input,args=NULL){
+  if(is.null(args)){
+    args=list(mixcompdist="halfuniform")
+  }
+  loglik=rep(0,11)
+  
+  #first run is with non-random start (ie default start)
+  res=do.call(ash, args= c(list(betahat=input$betahat,sebetahat=input$sebetahat,randomstart=FALSE),args))
+  loglik[1]= get_loglik(res)
+    
+  for(i in 2:11){
+    res = do.call(ash, args= c(list(betahat=input$betahat,sebetahat=input$sebetahat,randomstart=TRUE),args))
+    loglik[i]= get_loglik(res)
+  }
+  return(list(loglik=loglik))
+}

### 2015-04-08 dsc-stability/run_dsc.R

@@ -1,8 +1,16 @@
+library(dscr)
+sourceDir("methods")
 sourceDir("datamakers")
+
+###### Initialize #######
+
+dsc_ar=new.dsc("ash-robust","ash-robust-files")
+
+###### Add Scenarios #####
 
 
 #Now, for each scenario create an element of scenarios of the following form
-addScenario(dsc_shrink,name="A",
+addScenario(dsc_ar,name="A",
                     fn=rnormmix_datamaker,
                     args=list(
                       g=normalmix(c(2/3,1/3),c(0,0),c(1,2)),
@@ -10,10 +18,10 @@ addScenario(dsc_shrink,name="A",
                       max_pi0=0,
                       nsamp=1000,
                       betahatsd=1
-                      ),
+                    ),
                     seed=1:100)
 
-addScenario(dsc_shrink,name="B",
+addScenario(dsc_ar,name="B",
                     fn=rnormmix_datamaker,
                     args=list(
                       g=normalmix(rep(1/7,7),c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7)),
@@ -25,7 +33,7 @@ addScenario(dsc_shrink,name="B",
                     seed=1:100)
 
 
-addScenario(dsc_shrink,name="C",
+addScenario(dsc_ar,name="C",
                       fn=rnormmix_datamaker,
                       args=list(
                         g=normalmix(c(1/4,1/4,1/3,1/6),c(-2,-1,0,1),c(2,1.5,1,1)),
@@ -37,7 +45,7 @@ addScenario(dsc_shrink,name="C",
                       seed=1:100)
 
 #scenarios An, Bn, Cn are the same as A,B,C but with nulls included (pi0 uniform on [0,1])
-addScenario(dsc_shrink,name="An",
+addScenario(dsc_ar,name="An",
                     fn=rnormmix_datamaker,
                     args=list(
                       g=normalmix(c(2/3,1/3),c(0,0),c(1,2)),
@@ -48,7 +56,7 @@ addScenario(dsc_shrink,name="An",
                     ),
                     seed=1:100)
 
-addScenario(dsc_shrink,name="Bn",
+addScenario(dsc_ar,name="Bn",
                     fn=rnormmix_datamaker,
                     args=list(
[…]

### 2015-04-14 dsc-shrink/investigate_cdf.rmd
fixed bug in mixfdr2fitted.g

@@
 #load("res10.RData")
+#res=res10
 library("dscr")
@@
 
-#res=res10
 neglong = 
@@
 cdfplot(reslong %>% 
-          filter(method %in% c("ash.n","ash.hu","ash.u","ash.hu.s") & scenario %in% c("A","B","C","efron_FCR")), "All Discoveries"
+          filter(method %in% c("ash.n","ash.hu","ash.u","ash.hu.s") & scenario %in% c("A","B","C","efron_FCR")), "All observations"
         )

### 2015-04-14 paper/main.tex
fixed bug in mixfdr2fitted.g

@@
 Even with only one simulated example for each scenario, the greater reliability of estimates afforded
-by the UA is immediately apparent.
+by the UA is immediately apparent. In particular the greater tendency for mixFdr to overestimate $\pi_0$ is readily visible in its behaviour at 0.
+And the estimated cdf from mixFdr often has a vertical (or almost vertical)
+segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. This corresponds to
+one of the 3 mixture components being estimated to be centered at that location with very small variance. 
+The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
+While the UA is not the only way to achieve this, it is a simple and effective approach.
+
+Interestingly, even in the ``bimodal" scenario where the UA is severely contradicted, the estimated $g$ from $\ashr$ 
+is visually more accurate than that from mixFdr: although mixFdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. We believe this
+is mostly because the noise level is sufficiently large to make reliable estimation of the multiple modes difficult, rather than reflecting
+deficiencies in the specific choices made by mixFdr, and indeed in simulations where the modes are sufficiently well-spaced
+to be clearly visible in the observed $\betahat$, mixFdr is able to fit this modes (Supplementary Information; 
+\url{dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
+clearly visible in the observed $\betahat$.
 
@@
 
-%deconvolution
-%calculations are notoriously tricky and I have not
-%been able to produce a stable general algorithm. ; see also 
+\subsubsection*{Reliable estimates of $g$ lead to reasonable coverage behaviour}
+
+Coverage rates are given in:
+\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
+
 Zhao+Hwang consider in detail the case where the non-null effects are normally distributed.
@@
 Results in efron.fcr.rmd suggests that lower CI is OK; upper is more problematic because of difficulty in accurately estimating tail?
-
-To do: implement methods to compute posterior cdf of each observation at given points. 
-Then compute these for the simulation studies. Check whether issue with upper bound is due to non-unimodal distribution.
 
@@
 
+Notes from Adrian Raftery -
+
+Newcombe reference (1880s?)
+Dean and Raftery
+Uniform is super-efficient
+
 \subsection{The number of components is not critical}

### 2015-04-14 dsc-shrink/plot_cdf_eg.R

@@ -5,7 +5,7 @@ library(magrittr)
 library(ashr)
 library(ggplot2)
 source("methods/mixfdr.wrapper.R")
-source("../paper/Rcode2/Rhelperfiles/set_plot_colors.R")
+source("set_plot_colors.R")
 load("res.RData")
 
 #df is a list with components method,seed, scenario

### 2015-04-15 dsc-robust/summarize_dsc_robust.rmd

---
title: "Summarize dsc_robust"
output: html_document
---

Here we summarize the results of the simulation study to look at convergence of ash.
```{r}
  library(dscr)
  library(ashr)
  library(ggplot2)
  library(reshape2)
  library("dplyr")
  load("dsc_robust.RData") 
  res = dsc_robust$res
```

```{r}
  quantile((res %>% filter(method=="ash.n"))$score.diff1,c(0.8,0.9,0.95,0.99,1))
  quantile((res %>% filter(method=="ash.hu"))$score.diff1,c(0.8,0.9,0.95,0.99,1))
  quantile((res %>% filter(method=="ash.u"))$score.diff1,c(0.8,0.9,0.95,0.99,1))

  plot(ecdf((res %>% filter(method %in% c("ash.u","ash.hu")))$score.diff1),ylim=c(0.8,1))
  plot(ecdf((res %>% filter(method %in% c("ash.n")))$score.diff1),ylim=c(0.8,1))
```

```{r}
  mean((res %>% filter(method %in% c("ash.u","ash.hu")))$score.diff1<0.02)
  mean((res %>% filter(method %in% c("ash.u","ash.hu")))$score.diff1>1)
  mean((res %>% filter(method %in% c("ash.n")))$score.diff1<0.02)  
```

Now check out the example that showed the biggest convergence problem
```{r}
#check out which one showed biggest convergence problem
prob = res[which.max(res$score.diff1),]
prob_eg = loadExample(dsc_robust,seed=prob$seed,scenario=as.character(prob$scenario),method=as.character(prob$method))
```

Run ash on the example to check we can recapitulate the convergence problem.
```{r}
set.seed(prob$seed+1) #dscr sets the seed for the method to 1+seed 
ash.res=list()
ash.res[[1]] = ash(prob_eg$input$betahat,prob_eg$input$sebetahat,mixcompdist="uniform",cxx=TRUE)
for(i in 2:11){ash.res[[i]] = ash(prob_eg$input$betahat,prob_eg$input$sebetahat,mixcompdist="uniform",randomstart=TRUE,cxx=TRUE)}
```

```{r}
loglik = mapply(get_loglik,ash.res)
loglik
```

```{r}
best=which.max(loglik)
plot(ash.res[[best]]$lfsr,ash.res[[1]]$lfsr,xlab="lfsr (best run)",ylab="lfsr (default run)",main="Comparison of lfsr (best vs default run)")
abline(a=0,b=1,col=2)
plot(ash.res[[best]]$PosteriorMean,ash.res[[1]]$PosteriorMean,xlab="Posterior Mean (best run)",ylab="Posterior Mean (default run)",main="Comparison of Posterior Means (best vs default run)")
abline(a=0,b=1,col=2)

```

Compare Cdfs:
```{r}
  x = seq(-5,5,length=100)
  cdf_res = data.frame(x=x,
                     best=as.numeric(cdf.ash(ash.res[[best]],x)$y),
                     default=as.numeric(cdf.ash(ash.res[[1]],x)$y))
  cdf_res=melt(cdf_res,id.vars="x",value.name="y",variable.name="run")
  qplot(x=x,y=y,data=cdf_res,geom="line",colour=run,main="comparison of estimated cdf (best run and default run)")
```

### 2015-04-15 paper/main.tex

@@
 However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
-caused by this behaviour. To quantify this, we ran \ashr 10 times on each of XX simulated datasets using a random initialization for $\pi$,
+caused by this behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
 in addition running it using our default initialization procedure (XX need to specify in methods?). We then compared the largest
 log-likelihood achieved across all 11 runs with the log-likelihood achieved by our default run.
-When using a mixture of normals (ash.n) the results were extremely stable: 95\% showed a negligible log-likelihood difference (< 0.02), 
-and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 80\% showed a negligible log-likelihood difference (<0.02),
-and 9\% of runs showed an appreciable log-likelihood difference (>1), with the largest difference being 5.2. However, perhaps suprisingly, 
+When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference (< 0.02), 
+and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference (<0.02),
+and 6\% of runs showed an appreciable log-likelihood difference (>1), with the largest difference being 5.0. However, perhaps suprisingly, 
 even for this largest difference results from the default run (e.g.
-the lfsrs, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood \url{github.com/stephens999/ash/dsc-stability/check_convergence.rmd}.
+the lfsrs, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood \url{github.com/stephens999/ash/dsc-robust/summarize_dsc_robust.rmd}.
 

### 2015-04-15 dsc-shrink/check_mixfdr_lownoise.rmd
automate table generation

---
title: "check_mixfdr_lownoise"
output: html_document
---

Here we check that mixfdr actually does OK at fitting the bimodal distribution when 
noise is small compared with bimodality.

First simulate bimodal data with well-separated peaks. The histogram
shows the multi-modal nature of the underlying g is clear.
```{r}
  source("datamakers/datamaker.R")
  sim.bimodal=rnormmix_datamaker(
              args=list(g=normalmix(c(0.5,0.5),c(-5,5),c(1,1)),
              min_pi0=0.4,
              max_pi0=0.4,
              nsamp=1000,
              betahatsd=1)
            )
  hist(sim.bimodal$input$betahat,nclass=20,xlim=c(-10,10),prob=TRUE)
```

Now run mixfdr on it. Note that the fit captures the multi-modal nature of g, 
although mixfdr continues to overestimate pi0.
```{r}
  source("methods/mixfdr.wrapper.R")
  sim.bimodal.mixfdroutput = mixfdr.wrapper(sim.bimodal$input,args=list(theonull=TRUE))
  g = mixfdr2fitted.g(sim.bimodal.mixfdroutput)$fitted.g
  x=seq(-10,10,length=100)
  plot(ecdf(sim.bimodal$meta$beta),main="Cdf of true beta: empirical (black) and fitted (red)")
  lines(x,mixcdf(g,x),col=2)
```

### 2015-04-15 dsc-shrink/summarize_coverage.rmd
automate table generation

---
title: "Summarize Coverage"
output: html_document
---

```{r}
load("res.RData")

library("dscr")
library("ashr")
library(reshape2)
library(ggplot2)
library(dplyr)

neglong = 
  res$negprob %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars=c("method","scenario","seed",".id"),value.name="negprob") %>%
    filter(negprob > 0.9)

poslong = 
  res$posprob %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars=c("method","scenario","seed",".id"),value.name="posprob") %>%
    filter(posprob > 0.9)

reslong = 
  res$cdf_score %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%    
    melt(id.vars=c("method","scenario","seed",".id"))

reslong.pos = inner_join(reslong,poslong)
reslong.neg = inner_join(reslong,neglong)

```

Table of lower 5\% tail for all observations
```{r}
xtabs(lt~method+scenario,reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05))) %>% round(2)
```

Table of lower 5\% tail of positive findings. Because of the unimodal assumption
and the favoritism toward the null this should assess problems
with``over shrinkage" toward 0.
```{r}
xtabs(lt~method+scenario,reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05))) %>% round(2)
```

Table of lower 5\% tail of negative findings. This should indicate problems with tail behaviour of $g$. The uniform methods tend to over-shrink.
```{r}
xtabs(lt~method+scenario,reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05))) %>% round(2)
```

```{r}
save_latex_coverage_table=function(df,methodnames,filename,SCENARIONAMES=c("spiky","near-normal","flat-top","skew","big-normal","bimodal")){
  library(xtable)
  df$method=factor(df$method,levels=methodnames)
  df$scenario=factor(df$scenario,levels=SCENARIONAMES)
  mat <- as.matrix(xtabs(lt~method+scenario,df))
  mat <- xtable(mat,digits=rep(2,ncol(mat)+1))
  write(print(mat, 
        sanitize.text.function = function(x){x},
        floating=FALSE, 
        hline.after=NULL, 
        add.to.row=list(pos=list(-1,0, nrow(mat)), 
                        command=c('\\toprule ',
                                  '\\midrule ',
                                  '\\bottomrule '))
      ),file=filename)

}

save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05)),c("ash.n","ash.u","ash.hu"),"../paper/tables/coverage_neg.tex")
save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05)),c("ash.n","ash.u","ash.hu"),"../paper/tables/coverage_pos.tex")
save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05)),c("ash.n","ash.u","ash.hu"),"../paper/tables/coverage_all.tex")
```

### 2015-04-15 paper/main.tex
automate table generation

@@
 \usepackage{cite}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 is visually more accurate than that from mixFdr: although mixFdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. We believe this
-is mostly because the noise level is sufficiently large to make reliable estimation of the multiple modes difficult, rather than reflecting
-deficiencies in the specific choices made by mixFdr, and indeed in simulations where the modes are sufficiently well-spaced
-to be clearly visible in the observed $\betahat$, mixFdr is able to fit this modes (Supplementary Information; 
+mostly reflects the fact that the noise level is sufficiently large to make reliable estimation of the multiple modes difficult, rather than 
+fundamental deficiencies of mixFdr. Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
+to be clearly visible in the observed $\bhat$, mixFdr fits these modes (Supplementary Information; 
 \url{dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
-clearly visible in the observed $\betahat$.
+clearly visible in the observed $\bhat$.
 
@@
 
+\begin{table}[!ht]
+\begin{subtable}{\textwidth}
+	\input{tables/coverage_all.tex}
+	\caption{All}
+\end{subtable}
+
+\begin{subtable}{\textwidth}
+\input{tables/coverage_neg.tex}
+\caption{Negative findings}
+\end{subtable}
+
+\begin{subtable}{\textwidth}
+\input{tables/coverage_pos.tex}
+\caption{Positive findings}
+\end{subtable}
+
+\caption{Table of 5\% lower tail coverage rates by scenario and method}
+\end{table}
+
 \subsubsection*{Reverse asymptotics}

### 2015-05-01 dsc-shrink/summarize_coverage.rmd
updated paper

@@
 load("res.RData")
-
+coverthresh = 0.05 # threshold at which we look at coverage
+findthresh=0.95 #threshold at we define a discovery significant
 library("dscr")
@@
     melt(id.vars=c("method","scenario","seed",".id"),value.name="negprob") %>%
-    filter(negprob > 0.9)
+    filter(negprob > findthresh)
 
@@
     melt(id.vars=c("method","scenario","seed",".id"),value.name="posprob") %>%
-    filter(posprob > 0.9)
+    filter(posprob > findthresh)
 
@@
 reslong.neg = inner_join(reslong,neglong)
+dim(reslong.pos)
+dim(reslong.neg)
 
@@
 
-Table of lower 5\% tail for all observations
+Overall proportion of negative findings is `r nrow(reslong.neg)/nrow(reslong)`.
+
+Overall proportion of positive findings is `r nrow(reslong.pos)/nrow(reslong)`.
+
+Table of lower tail for all observations
 ```{r}
-xtabs(lt~method+scenario,reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05))) %>% round(2)
+xtabs(lt~method+scenario,reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh))) %>% round(2)
 ```
 
-Table of lower 5\% tail of positive findings. Because of the unimodal assumption
+Table of lower tail of positive findings. Because of the unimodal assumption
 and the favoritism toward the null this should assess problems
@@
 ```{r}
-xtabs(lt~method+scenario,reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05))) %>% round(2)
+xtabs(lt~method+scenario,reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh))) %>% round(2)
 ```
 
-Table of lower 5\% tail of negative findings. This should indicate problems with tail behaviour of $g$. The uniform methods tend to over-shrink.
+Table of lower tail of negative findings. This should indicate problems with tail behaviour of $g$. The uniform methods tend to over-shrink.
 ```{r}
-xtabs(lt~method+scenario,reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05))) %>% round(2)
+xtabs(lt~method+scenario,reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh))) %>% round(2)
 ```
@@
 ```{r}
-save_latex_coverage_table=function(df,methodnames,filename,SCENARIONAMES=c("spiky","near-normal","flat-top","skew","big-normal","bimodal")){
+save_latex_coverage_table=function(df,methodnames,filename,SCENARIONAMES=c("spiky","near-normal","flat-top","skew","big-normal","bimodal"),switch=FALSE){
   library(xtable)
@@
   mat <- as.matrix(xtabs(lt~method+scenario,df))
+  if(switch){
+    mat=1-mat
+  }
   mat <- xtable(mat,digits=rep(2,ncol(mat)+1))
@@
 
-save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05)),c("ash.n","ash.u","ash.hu"),"../paper/tables/coverage_neg.tex")
-save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05)),c("ash.n","ash.u","ash.hu"),"../paper/tables/coverage_pos.tex")
-save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<0.05)),c("ash.n","ash.u","ash.hu"),"../paper/tables/coverage_all.tex")
+save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/coverage_neg.tex")
+save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/coverage_pos.tex")
+save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/coverage_all.tex")
+
+save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/coverage_neg_nopen.tex")
+save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/coverage_pos_nopen.tex")
+save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/coverage_all_nopen.tex")
 ```
+
+These tables show 1-the left tail instead of the left tail.
+```{r}
+save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/scoverage_neg.tex",switch=TRUE)
+save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/scoverage_pos.tex",switch=TRUE)
+save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/scoverage_all.tex",switch=TRUE)
+
+save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/scoverage_neg_nopen.tex",switch=TRUE)
+save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/scoverage_pos_nopen.tex",switch=TRUE)
+save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/scoverage_all_nopen.tex",switch=TRUE)
+```

### 2015-05-01 paper/main.tex
updated paper

@@
 \usepackage{enumerate}
-\usepackage{subfig}
 \usepackage{multirow} 
@@
 Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
-EB approaches to this problem have been quite extensively studied by several authors, 
+indeed, the notion that EB approaches could be helpful in handling multiple comparisons predates introduction of
+the FDR  (e.g.~\cite{greenland1991empirical}). More recently, EB approaches to the FDR
+have been extensively studied by several authors, 
 especially B.~Efron and co-authors 
@@
 inference. The first idea is to {\it assume that the distribution of the non-zero effects is unimodal}, and to perform
-non-parametric inference under this assumption. This idea has the twin benefits of both simplifying and
-improving inference, at least when the assumption is correct. And even when the assumption is incorrect, we show
-that the approach can be relatively robust to this. The second idea is to use two numbers (effect sizes, and their standard errors), rather than just one ($p$ values, or $z$ scores)
+non-parametric inference under this assumption. This idea has the twin benefits of simplifying and improving inference,
+at least when the unimodal assumption is correct.  And even when the assumption is incorrect, the resulting errors in inference are not too egregious. 
+The second idea is to use two numbers (effect sizes, and their standard errors), rather than just one ($p$ values, or $z$ scores)
  to summarize each measurement. This idea allows any differences in the precision of different measurements to be taken into account,
- and in particular avoids the problem that poor-precision measurements can contaminate high-precision measurements, resulting in inflated FDR values.
+ and in particular avoids the problem that poor-precision measurements can contaminate high-precision measurements and inflate the FDR.
 
@@
 controlling the FDR, rather than on estimating the sizes of the effect. 
-The two new ideas highlighted above 
-help facilitate estimation as well as testing, and by providing simple software to implement our methods
+The two new ideas highlighted above facilitate estimation as well as testing, and by providing easy-to-use software implementing our methods
 we hope to promote and accelerate the shift in this direction.
 
-Here we consider one setting in which FDR-based methods are widely used:
-an experimenter has measured a large numbers of ``effects", each with an associated
-measurement precision (standard error), 
-and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
-Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
-here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
-This setting arises commonly in many areas of statistics, and particularly in genomics applications.
-For example, in genomics
-a common goal is to identify ``differentially-expressed genes" -- that is, genes
-whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
-mean expression level between the two conditions, often referred to as the ``log-fold change". 
+The use of EB methods for estimation also has a long and close relationship with another key concept in modern statistics: shrinkage estimation \cite{efron.morris.73.jasa}.
+This close relationship becomes especially intimate here due to our assumption that the distribution of the underlying effects is
+unimodal, since this ensures that estimates are shrunk towards a common point -- the mode of the distribution.
+%Indeed, within our EB framework, multiple comparisons correction can be seen as a natural limit of shrinkage estimation:
+%if effects are distributed uni-modally about zero then it is natural to shrink estimates towards zero. 
+% are essentially two sides of the same coin, one related to testing and the other to estimation. 
+Thus, although we focus here primarily on FDR-related issues, another important contribution of our work is to provide {\it generic} and {\it adaptive} shrinkage estimation procedures.
+By generic, we mean that these methods can be applied in any setting where a series of effect estimates and corresponding standard errors are available.
+By adaptive, we 
+have two properties in mind. First, the appropriate amount of shrinkage is determined from the data: when the data indicate that large effects are relatively common then
+shrinkage of large observed effects is less than when the data indicate that large effects are rare.
+Second, the amount of shrinkage undergone by each measurement will depend on its precision, or standard error: measurements with big standard errors will undergo more shrinkage than measurements with small standard errors.
+Shrinkage estimation is a powerful tool, with many potential applications, including for example wavelet denoising, 
+where shrinkage of wavelet coefficients is used to smooth signals. Our methods provide an attractive and competitive alternative
+to existing EB shrinkage methods for that context, as is explored more fully elsewhere (Xing and Stephens).
+
+Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
+
 %Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.
@@
 
-Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
-Here we present new tools that bring two important ideas to bear on the problem.
-First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
-For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
-may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
-imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
-Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
-is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
-Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
-demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
-a substantial increase in the number of discoveries at a given FDR).
-
-Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests, 
-provides a general tool for shrinkage estimation. Shrinkage estimation is a powerful tool, with many potential applications, including for example wavelet denoising, 
-where shrinkage of wavelet coefficients is used to smooth signals. 
-However, in contrast to FDR methods, shrinkage methods are not so widely used in practice, at least in the kinds
-of genomics settings I am familiar with. One possible reason for this is that most work on shrinkage-based estimation has focussed
-on point estimates, rather than on interval estimates which are often of more direct practical relevance. Another contributing
-factor may be the paucity of general-purpose software packages for this kind of shrinkage estimation.
-% (There are, perhaps, more packages for shrinkage estimation in the more complex setting of covariance estimation, where shrinkage is perhaps still more important.)
-Our work here addresses both these issues, providing a general-purpose software tool that uses shrinkage ideas to improve both
-point and interval estimates.
-
-By ``adaptive" shrinkage we 
-have two properties in mind. First, the appropriate amount of shrinkage is determined from the data, rather than being pre-specified.
-Second, the amount of shrinkage undergone by each measurement will depend on its precision, or standard error: measurements with high standard error will undergo more shrinkage than measurements with low standard error.
-Although shrinkage-based estimation can be motivated in various ways, our key goal here is to combine information across the multiple measurements to improve inference for each individual effect.
-By improved inference, we mean both
-improved average accuracy of point estimates, which is
-the traditional focus of shrinkage-based methods, \emph{and} improved assessments of uncertainty, such as credible intervals.
-
-The potential for shrinkage-based estimation to
-address the multiple comparisons problems tackled by FDR-based methods
-has been highlighted
-previously, including \cite{greenland1991empirical},
-\cite{efron2008microarrays} and \cite{gelman2012we}. [Note, check also Louis, JASA, 1984] 
-Furthermore,
-
-The initial goal of the ASH (Adaptive SHrinkage) project is to provide simple, generic, and flexible methods to derive ``shrinkage-based" estimates and credible intervals for unknown quantities $\beta=(\beta_1,\dots,\beta_J)$, given only estimates of those quantities ($\bhat=(\bhat_1,\dots, \bhat_J)$) and their corresponding estimated standard errors ($s=(s_1,\dots,s_J)$).
-
-Given that shrinkage estimation is widely recognized as a powerful tool, there are
-
-The only package we have found that provides anything similar
-to the functionality provided here is \mixfdr \cite{muralidharan2010empirical}. Compared with \mixfdr, the key features of \ashr are that it
-i) focuses on allowing for variation in the standard deviation of each observation; ii) constrains the underlying density to be unimodal (and possibly symmetric).
-
-As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
+%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
+%Here we present new tools that bring two important ideas to bear on the problem.
+%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
+%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
+%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
+%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
+%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
+%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
+%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
+%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
+%a substantial increase in the number of discoveries at a given FDR).
+
+%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
+%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
 Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
-and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
-estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
-generic methods that work directly with two numbers for each 
-measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
+%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
+%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
+%generic methods that work directly with two numbers for each 
+%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
 Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
-uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
+%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
 
@@
 
-Suppose that we are interested in the values of $J$ ``effects" $\beta_j$ ($j=1,\dots,J$).
-In some contexts our interest may focus on which of the $\beta_j$ are ``significantly" different from zero, whereas in other contexts our interest may focus on estimating their values; the methods described here are suited to both these contexts.
- We assume that we have obtained data $D=(D_1,\dots,D_J)$ that provide independent estimates $\bhat_1,\dots,\bhat_J$ of these effects,
- with corresponding (estimated) standard errors $\shat_1,\dots,\shat_J$. Let $H_j$ denote the event that the $j$th null hypothesis holds, and $t_j$ denote
+% Here we consider one setting in which FDR-based methods are widely used:
+%an experimenter has measured a large numbers of ``effects", each with an associated
+%measurement precision (standard error), 
+%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
+%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
+%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
+%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
+%For example, in genomics
+%a common goal is to identify ``differentially-expressed genes" -- that is, genes
+%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
+%mean expression level between the two conditions, often referred to as the ``log-fold change".
+
+Suppose that we are interested in the values of $J$ ``effects" $\beta_j$ ($j=1,\dots,J$). For example, in a canonical genomics application
+$\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
+In some contexts interest may focus on which of the $\beta_j$ are ``significantly" different from zero, whereas in other contexts interest may focus on estimating their values; the methods described here tackle both these problems.
+ We assume that we have obtained data $D$ that consist of estimates $\bhat_1,\dots,\bhat_J$ of these effects,
+ with corresponding (estimated) standard errors $\shat_1,\dots,\shat_J$. Let $H_j$ denote the event that the $j$th null hypothesis holds ($H_j:\beta_j=0$), and $t_j$ denote
  the $t$ statistic testing this null:
- \begin{gather}
- H_j: \beta_j = 0 \\  
+ \begin{equation}  
  t_j :=\bhat_j/\shat_j. 
- \end{gather}
+ \end{equation}
  The $t$ statistics, of course, have a $t$ distribution under the null:
@@
 \end{equation}
+Although our approach here is (Empirical) Bayesian, one nice feature of FDR procedures is that 
+Frequentist and Bayesian procedures are often closely aligned, at least under certain assumptions; see
+for example \cite{storey.03}, particularly his Theorem 1.
 
@@
 \end{equation}
-For existing EB methods (but not for our method), typically $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
+In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
 \begin{equation}
@@
 
-One nice feature of FDR procedures is that Frequentist and Bayesian procedures are often closely aligned, at least under certain assumptions; see
-for example \cite{storey.03}, particularly his Theorem 1.
-
-\subsection{Existing methods}
-
-\subsubsection{qvalue}
+\subsection*{Existing methods for FDR analysis}
+
+\subsubsection*{qvalue}
 
@@
 
-\subsubsection{EB approaches: \locfdr and \mixfdr}
+\subsubsection*{EB approaches: \locfdr and \mixfdr}
 
 The fundamental
- idea of EB procedures is similar to that of \qvalue: to attempt to partition the observations into null and alternative components.
+ idea of EB procedures is similar to that of \qvalue: partition the observations into null and alternative components.
 Here we highlight two EB methods that
@@
 
- Here we introduce an EB method, which, for reasons discussed below, we call ``adaptive shrinkage" or ``ash".
- The method provides not only estimated $\fdr_j$ and q values, but also shrinkage point estimates, shrinkage credible intervals,
- and estimates of the local false sign rate, defined below. It is implemented in the R package \ashr available at \url{http://github.com/stephens999/ashr}.
-
-  Rather than directly modelling the distributions of the $p$ values, or $z$ scores, our starting point is to model the distribution of the effects $\beta_j$.
- Specifically we assume that the effects $\beta_j$ are independent and identically distributed from some unknown distribution $g(\cdot)$,
+ Here we introduce an EB method, which, for reasons discussed above, we call ``adaptive shrinkage" or ``ash".
+ The method provides not only estimated $\fdr_j$ and q values, but also shrinkage point estimates and credible intervals for $\beta_j$,
+ and estimates of the local false sign rate, defined below.
+
+We begin by presenting the simplest version of our model, before embellishing it.
+Instead of attempting to partition the (observed) $z$ scores or $p$ values into null and alternative components,
+we use a model that partitions the (unobserved) effects $\beta_j$  into null and alternative components:
+\begin{equation} \label{eqn:betaj}
+\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)
+\end{equation}
+where $\delta_0$ denotes a point mass at 0, and $g_1$ denotes the (unknown) distribution of the non-zero $\beta_j$.
+Thus $\pi_0$ denotes the proportion of effects that are null.
+
+At this point we introduce a key assumption, which distinguishes our approach from previous work: we assume that $g_1$ is unimodal,
+with a mode at 0. One simple way to achieve this ``Unimodal Assumption" (UA)  is
+to assume that $g_1$ is a mixture of zero-mean normal distributions
+\begin{equation} \label{eqn:g1}
+g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
+\end{equation}
+where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
+Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
+represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
+ The symmetry of the normal distribution implies that (\ref{eqn:g1}) is not only unimodal, but also symmetric about zero. We discuss 
+ more flexible alternatives based on using mixtures of uniform distributions, and details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$), as well as the UA more generally, below.
+
+Since the $\beta_j$ are unobserved, we need a model to connect the $\beta_j$ to observations. Here we
+treat the estimates $\bhat_j$ as the observed data, and assume that they are noisy measurements of the true $\beta_j$ with given standard error
+$\shat_j$:
+\begin{equation} \label{eqn:bhat}
+\bhat_j | \beta_j, \shat_j \sim N(\beta_j, \shat^2_j).
+\end{equation}
+This approach, which follows \cite{wakefield:2009}, 
+should be a good approximation if $\bhat_j$ and $\shat_j$ are based on sufficiently many observations. We discuss below an improvement 
+for moderate numbers of observations, based on replacing this normal likelihood with a $t$ likelihood.
+
+Combining (\ref{eqn:betaj})-(\ref{eqn:bhat}), and assuming independence of the observations and effects for different $j$ (given $\pi_0,w_1,\dots,w_k$), yields
+a hierarchical model for $\beta,\bhat$ given $\shat$, with unknown hyper-parameters $(\pi_0,w_1,\dots,w_k)$. To simplify notation we define $\pi_k := (1-\pi_0)w_k$ 
+so that the unknown hyper-parameters can be written as a vector $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1.
+The usual empirical Bayes approach to fitting this model would involve two steps:
+\begin{enumerate}
+\item Estimate the hyperparameters $\pi$ by maximum likelihood, yielding a maximum likelihood estimate $\hat{\pi}$.
+\item Compute quantities of interest (e.g. local fdr values, $q$ values, credible intervals etc.) from the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$.
+\end{enumerate}
+We make one modification to this procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we instead maximise a penalized
+likelihood, where the penalty (detailed below) encourages $\hat\pi_0$ to be as big as possible (within the constrains of the model and observed data). We use
+this penalty here because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR. 
+Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{EM} to maximize the penalized likelihood, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
+(Note that the simplicity of the EM algorithm in step 1 is due to our using a fixed grid for $\sigma_k$ in (\ref{eqn:g1}), instead of estimating $\sigma_k$,
+which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
+
+\subsubsection*{Embellishments}
+
+\subsubsection*{More flexible prior component distributions}
+
+\subsubsection*{Replace normal likelihood with $t$ likelihood}
+
+\subsubsection*{}
+
+DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
+To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
+Because the prior distribution is unimodal about zero,
+the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
+amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
+For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
+than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
+small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
+
+The posterior distribution $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ then reflects a trade-off
+between the information in the prior distribution, which is more-or-less concentrated near zero,
+
+There are two key points to appreciate here. First, by estimating the hyperparameters $\pi$ from the data, 
+the method adapts to the amount of signal in the data: if most $\beta_j$ are truly at or near zero then $\pi$ will reflect this by
+concentrating on zero and very small values of $\sigma_k$, and the posterior distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$
+will be tend also to be concentrated around zero (strong shrinkage). In contrast, if most $\beta_j$ are large then $\pi$ will be put most weight
+on components with large $\sigma_k$, and the posterior distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ will be approximately centered  on $\bhat_j$
+(weak shrinkage). Indeed, in the limit as $\pi$ puts all it's weight on very large $\sigma_k$ the prior becomes flatter and flatter
+and the posterior on $\beta_j$ becomes $N(\bhat_j,\shat_j)$. Thus, in this limit, the Bayesian credible intervals match standard confidence intervals.
+
+ and only very strong observations will have a small fdr; if most $\beta_j$ are large
+then this will also be reflected in the estimates of $\pi$ and even borderline signals might have a small fdr. 
+This idea is also used by existing FDR methods, but here, by modelling the effects, it extends to our estimates of the effects:
+
+The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ encapsulate uncertainty in the values for $\beta_j$, combining information across
+$j=1,\dots,J$. The combining of the information occurs through estimation of
+$\pi$, which involves all of the data.
+END OF DISCUSSION
+
+--- NOte 
+In principle we might prefer a full Bayes approach that accounts for uncertainty in $g$; however,
+we believe that in most practical applications uncertainty in $g$ will not be the most important concern, and 
+compromise this principle for the simplicity of the EB approach.
+
+\begin{equation} \label{eqn:Lpi}
+\bhat_j | \shat_j, \pi \sim \pi_0 N(\cdot; 0, \shat^2_j) + \sum_k \pi_k N(\cdot; 0,\sigma_k^2 + \shat^2_j) \quad \text{independently for $j=1,\dots,J$}.
+\end{equation}
+
+Our EB approach can now be simply stated. 
+\begin{enumerate}
+\item Estimate the parameters $\pi$ (and hence also the distribution $g_1$) by maximising the likelihood 
+\begin{equation}
+L(\pi) := \prod_j p(\bhat_j | \shat_j, \pi),
+\end{equation}
+where the terms of the likelihood are mixtures of the form (\ref{eqn:Lpi}). 
+\end{enumerate}
+
+ In outline, our goal is to perform Bayesian inference for the true effects $\beta$ using information in the summary statistics $\bhat,\shat$.
+ That is we aim to compute the joint posterior distribution $p(\beta | \bhat,\shat)$.
+This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
+from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
+with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
+ By working with summary statistics $\bhat, \shat$, rather than more refined data, we aim to produce generic methods that 
+ can be applied whenever such summary data are available - just as \qvalue can be applied to any set of $p$ values for example.
+Any attempt to produce generic methods is likely to involve a compromise between functionality and generality; 
+we believe that by working with two numbers ($\bhat_j,\shat_j$) for each observation, rather than one ($p_j$ or $z_j$), 
+we can gain substantially in functionality (e.g.~ we can estimate effect sizes, as well as testing, and we can take better account of variations in
+measurement precision across units $j$) while losing only a little in generality.
+
+We assume that (given $\shat$) the
+effects $\beta_j$ are independent draws from some common ``prior" distribution $g$:
  \begin{equation} \label{eqn:beta}
- \beta_j \sim g(\cdot).
- \end{equation}
- A key assumption, which distinguishes our approach from previous work, is that $g$ is unimodal. We assume here that the mode is at 0, and allow
+ \beta_j \, | \, g, \shat \sim g(\cdot) \quad \text{independently for $j=1,\dots,J$.}
+ \end{equation}
+ A key assumption, which distinguishes our approach from previous work, is that $g$ is unimodal (and, optionally, symmetric). 
+ We assume here that the mode is at 0, and allow
  that $g$ can have a point mass at 0, corresponding to the ``null" effects.
-(An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data.)
+(An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood.)
+Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
+This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
+ \begin{equation} \label{eqn:beta-alpha}
+ \frac{\beta_j}{\shat_j^\alpha} \, \big| \, g, \shat \sim g(\cdot) \quad \text{independently for $j=1,\dots,J$}
+ \end{equation}
+for any $\alpha \geq 0$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
+assuming that the ``standardized effects" $\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
+it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
+the assumption made by existing methods that rank tests by their $p$ values (or $z$ scores). See Results for further discussion.
+
+ Second, we assume that the observations $\bhat$ are independent given $\beta$,
+ with $\bhat_j$ depending on $\beta,\shat$ only through $\beta_j$ and $\shat_j$:
+ \begin{equation} \label{eqn:lik}
+ \bhat_j \, | \, \beta, \shat \sim p(\bhat_j | \beta_j, \shat_j), \quad \text{independently for $j=1,\dots,J$.}
+ \end{equation}
+A natural assumption, made by \cite{wakefield:2009},  is that 
+ \begin{equation} \label{eqn:normlik}
+ \bhat_j \, | \, \beta_j, \shat_j \sim N(\beta_j, \shat_j),
+ \end{equation}
+ which should be a good approximation if the estimates $\bhat_j,\shat_j$ are based on a large number of observations.
+ To try to improve this approximation when there are fewer observations we generalize this to
+  \begin{equation} \label{eqn:lik}
+ \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
+ \end{equation}
+where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
+and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
+
+Equation (\ref{eqn:lik}) is, of course, motivated by the standard asymptotic result
+\begin{equation} \label{eqn:tdist}
+(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
+\end{equation}
+However (\ref{eqn:tdist}) does not imply (\ref{eqn:lik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:lik}) it is conditioned on.
+In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
+currently pursuing this improved approach in work with Mengyin Lu, and results will be published elsewhere. We note here that
+modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is small (e.g. $<10$),
+where our simpler approach can produce unsatisfactory results (see later).
+
+Given the distribution $g$, the independence assumptions in (\ref{eqn:beta}) and (\ref{eqn:lik}) imply that the posterior distribution $\beta | g,\bhat,\shat$ conveniently 
+factorizes into independent components $p(\beta | g,\bhat,\shat) = \prod_j p(\beta_j | g, \bhat_j, \shat_j)$.
+The Empirical Bayes approach involves first estimating $g$ using all the data, and then computing quantities of interest conditional on the estimated $g$. That is:
+ \begin{itemize}
+ \item Estimate $g$ by maximising the likelihood $\hat{g} = \arg \max L(g)$ where $L(g) := p(\bhat | g, \shat)$ and the
+ maximisation is subject to the constraint that $g$ is unimodal.
+ \item Compute the conditional distributions $p(\beta_j | \hat{g} \bhat, \shat)$ ,
+ from which we can compute all quantities of interest such as lfdr, lfsr, $q$-values, posterior credible intervals/regions etc.
+  \end{itemize}
+
+ \begin{note}
+The conditional distributions $p(\beta_j | \hat{g}, \bhat, s)$ 
+encapsulate uncertainty in the values for $\beta_j$, combining information across
+$j=1,\dots,J$. The combining of the information occurs through estimation of
+$g$, which involves all of the data.
+In principle we might prefer a full Bayes approach that accounts for uncertainty in $g$; however,
+we believe that in most practical applications uncertainty in $g$ will not be the most important concern, and 
+compromise this principle for the simplicity of the EB approach.
+\end{note}
 
 \def\dt{\text{dt}}
-
-To perform likelihood-based inference for $g,\beta$, we assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
-$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
- $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
-  \begin{equation} \label{eqn:lik}
- \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
- \end{equation}
- where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e.
+\begin{note}
+ A key property of the likelihood (\ref{eqn:lik}) [and indeed (\ref{eqn:normlik})] is that
+observations with a large $\shat_j$ have an essentially flat likelihood, and so do not affect inference for
+$g$ or the other effects. This contrasts with existing approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
+$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, even under the alternative hypothesis (assuming
+that the effect $\beta_j$ is small compared with $\shat_j$). These observations therefore do affect inference; specifically they add noise
+and inflate estimates of the FDR. Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
+\end{note}
+
+%
+%assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
+%$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
+% $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
+%  \begin{equation} \label{eqn:lik}
+% \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
+% \end{equation}
+% where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e.
 the density of $\mu+\sigma T_\nu$).
 In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
-  \begin{equation} \label{eqn:normlik}
- \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
- \end{equation}
+%  \begin{equation} \label{eqn:normlik}
+% \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
+% \end{equation}
 %which, in light of (\ref{eqn:betahat}), is equivalent to
@@
 
- \subsubsection*{Discussion on Assumptions}
-
+ \begin{note}
 Although the unimodal assumption (UA) will not apply to all situations, we argue that it will be reasonable in many practical contexts. 
@@
 One of our aims here is to emphasize the considerable practical benefits of making the UA in the FDR context.
-
-The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
+\end{note}
+%The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
 they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
@@
 
-This likelihood approximation can be motivated by assuming that i) $\bhat$
- $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
- $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$.
-
-Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.
-
-A key property of the likelihood (\ref{eqn:lik}) is that
-observations with a large $\shat_j$ have an essentially flat likelihood (\ref{eqn:lik}), and so do not affect inference for
-$g$ or the other effects, whereas when modeling the $Z$ scores directly, observations with large $\shat_j$ produce $Z$ scores with an $N(0,1)$ distribution that {\it do} affect inference.
-Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
+%This likelihood approximation can be motivated by assuming that i) $\bhat$
+% $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
+% $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$. 
+%
+
+%Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.
+
 %While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
@@
 
-\subsubsection*{Implementing the Unimodal assumption}
+\subsubsection*{Implementation outline}
 
@@
 \end{equation}
-where $\delta_0(\cdot)$ denotes a point mass at 0, $\pi=(\pi_0,\dots,\pi_K)$ are mixture proportions to be estimated (with $\pi_k \geq 0, \sum_k \pi_k=1$), and $f_k$ are pre-specified 
-component distributions with one of the following forms: 
+where $\delta_0(\cdot)$ denotes a point mass at 0, $\pi=(\pi_0,\dots,\pi_K)$ are mixture proportions to be estimated (with $\pi_k \geq 0, \sum_k \pi_k=1$), and $f_k$ are pre-specified component distributions with one of the following forms:
 \begin{enumerate}[i)]
@@
 
-To see the flexibility of the representation (\ref{eqn:g})
-think of $K$ as being ``large", with the values of $\sigma^2_k$ or $a_k$ 
+To see the flexibility of (\ref{eqn:g}) think of $K$ as ``large", and consider a dense grid of values of $\sigma^2_k$ or $a_k$ 
 spanning a wide range of values, from
-essentially zero to unnecessarily large. In this way, as $K$ increases, the three different choices for $f_k$ can allow $g$ to approximate,
+essentially zero to unnecessarily large. As $K$ increases, and the grid becomes denser, the three different choices for $f_k$ can allow $g$ to approximate,
 with arbitrary accuracy,
@@
 \end{enumerate}
-The latter two claims are related to characterizations of unimodal distributions due to Khintchine (1938) and  Shepp (1962); see Feller, 1971, page  158. 
-Furthermore, although it is natural for statisticians to worry about ``overfitting" if $K$ is large, 
+The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine:1938} and  \cite{shepp:1962}; see \cite{feller:1971}, page  158. 
+Although it is natural to worry about ``overfitting" if $K$ is large, 
 in fact the unimodal constraint is sufficient to prevent serious problems with overfitting. For example, note that the unimodal
@@
 (Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
-Thus allowing $K$ to be large in ii) or iii) can be thought of as approximating the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
-
-Having emphasized the flexibility that can be achieved with large $K$, we note that in practice even quite small values of $K$ (e.g. 2-10) provide highly flexible
-and effective methods. See Section \ref{} for our software default choices of $K$ and $\sigma^2_k,a_k$.
+Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
+This said, in practice even quite small values of $K$ (e.g.
+2-10) can provide reasonable performance.
+See Section \ref{details} for our software default choices of $K,\sigma^2_k,a_k$.
 
 % $g() = \sum_{k=1}^K (\pi^+_k U[0,a_k] + \pi^-_k U[-a_k,0])$
+
+Using the normal likelihood (\ref{eqn:normlik}) and any of the component forms (i)-(iii) for $f_k$ 
+yields a simple EM algorithm for estimating $\pi$ in Step 1, and simple analytic forms for the conditional
+distributions in Step 2. Using the $t$ likelihood (\ref{eqn:lik}) is equally simple for the uniform forms (ii) and (iii) for $f_k$ (but 
+tricky for the normal $f_k$ and so we have not implemented that case).
+We used the elegant R package {\tt SQUAREM} \cite{squarem}) to accelerate convergence of our EM algorithm.
 
@@
 \end{equation}
-In principle we would
-prefer to take a full Bayes approach that accounts for uncertainty in $g$ (by accounting for uncertainty in $\pi$); however,
-we believe that in most practical applications uncertainty in $g$ will not be the most important concern, and so 
-we compromise this principle for the simplicity of the EB approach.
-[Note: a Variational Bayes approach to estimating $\pi$ is also implemented in our R package]
+
+[Note: a Variational Bayes approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
 
@@
 
-The conditional distributions $p(\beta_j | \hat\pi, \bhat, s)$ 
-encapsulate uncertainty in the values for $\beta_j$, combining information across
-$j=1,\dots,J$. The combining of the information occurs through estimation of
-$\pi$, which involves all of the data.
-These conditional distributions can be conveniently summarized
-in various ways, including point estimates (e.g. the posterior means or medians),
-and credible intervals/regions.
-
 %A Taylor series expansion of the log likelihood around the maximum likelihood estimate $\bhat_j$  yields $$l(\beta_j) \approx l(\bhat_j) + 0.5* (\beta_j - \bhat_j)^2 l''(\bhat_j).$$ [Fill in details?]
@@
 \subsection*{The local False Sign Rate}
-
-The local False Discovery Rate (lfdr) for observation $j$ is defined \cite{efron2001empirical} as
-\begin{equation}
-\lfdr_j = p(\beta_j = 0 | \hat\pi, \bhat, s).
-\end{equation}
-The lfdr terminology comes from using "discovery" to refer to rejecting the null ($H_j:\beta_j=0$).
-Specifically $\lfdr_j$ is the probability that, if we reject $H_j$,
-it is a  "false discovery".
 
@@
 We note that the idea of focussing on tail areas, rather than point null hypotheses,
-has a long history (e.g. Altham? others?). For example, 
+has a long history. For example, 
 Tukey, p32, "the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?""
-More recently Gelman.
 
@@
 
-Note use of SQUAREM.
-
-Note penalty preferring $\pi_0$.
-
-\subsection{various notes...}
-
-Here performing inference means obtaining an (approximate) posterior distribution for each effect $\beta_j$,
-which can be used to estimate False Discovery Rates, and to produce both point and interval estimates for $\beta_j$.
- Although ``multiple comparisons" is usually presented as a ``problem", our emphasis here is that the multiple measurements
- actually provide an opportunity: the hierarchical model combines information across measurements to improve both accuracy and precision of estimates.
-
-Note that incorporating differential errors messes up the EM algorithm?
-
-we start by noting that the $p$ values in Figure \ref{} were actually generated under a scenario where
-{\it no tests were null}! Thus the qvalue procedure, which estimates the proportion of nulls to be xx\%, is acting very conservatively in this case.
-However, it could be argued that some degree of conservativeness is inevitable, and that nonetheless the procedure is acting sensibly.
-Therefore, to illustrate this point we really want to focus on the fact that, at least in some settings, the assumption is {\it unrealistic}.
-Specifically, we want to argue that the implied distribution of the alternative $p$ values is unrealistic. In fact, this is hard to see in $p$ value space,
-and so instead we translate these $p$ values back to the $z$ scores that lead to these $p$ values.
-
-The second issue is that directly modeling the p values, or z scores, say via non-parametric methods, 
-can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-It should be noted that existing methods were perhaps developed with the expectation that they would typically 
-be applied in settings where the vast majority of all tests are null, and under such circumstances
-the problems caused by this assumption are less severe. However, it seems desirable to have methods that
-also perform well in settings where  this assumptions does not hold, and this motivates the methods developed here.
+We used the elegant R package {\tt SQUAREM} \cite{squarem}) to accelerate convergence of our EM algorithm.
+
+\subsection*{Some Details}
+
+\subsubsection*{Choice of grid for $\sigma_k, a_k$}
+
+\def\sigmamax{\sigma_{text{max}}}
+\def\sigmamin{\sigma_{text{min}}}
+When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
+in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
+as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
+based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
+is negative we set $\sigmamax = 8\sigmamin$.)
+
+When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.
+
+Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
+a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.
+
+\subsubsection*{Penalty term on $\pi$}
+
+In practice data do not distinguish between effects that are ``very small" and ``exactly zero".
+Thus, if a mixture component $f_k$ has most mass very near zero then
+this introduces non-identifiability between $\pi_k$ and $\pi_0$. In practice this is a major problem
+only if one focuses on false discovery rates rather than false sign rates: the fdr is sensitive to $\pi_0$ whereas
+the fsr is not. However, to make fdr (and fsr) estimates from our method ``conservative" we add a penalty 
+term $h(\pi;\alpha)$ to the likelihood to encourage over-estimation of $\pi_0$:
+\begin{equation}
+h(\pi;\alpha) = \prod_{k=0}^K \pi_k^{\alpha_k-1}
+\end{equation}
+where $\alpha_k \geq 1 \, \forall k$. The default is $\alpha_0=10$ and $\alpha_k=1$, which yielded
+consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0}).
+
+Although this penalty is based on a Dirichlet density, we don't consider this to be a ``prior distribution" for $\pi$.
+
+\subsubsection*{Initialization}
+
+We initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. 
+Our rationale to initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
+but weak signal in the data cannot quickly draw the algorithm towards the null.
 
@@
 However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
-caused by this behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
+caused by such behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
 in addition running it using our default initialization procedure (XX need to specify in methods?). We then compared the largest
@@
 $T$ is some effect threshold of interest (and ``likely" can be quantified by whatever probability threshold is desired). 
+Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
+this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
+In our case, because $g$ is unimodal about zero, the effect of this prior is always to ``shrink" the posterior distributions
+towards zero, with the strength of the shrinkage being estimated from the data.
 Further, because the posterior distribution is, by definition,
 conditional on the observed data, inferences based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
-been selected based on the observed data. %(This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.)
+been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
 In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
@@
 how can confidence intervals be assigned to the true
-effect size for each selected case?". \cite{efron2008} notes the potential for EB approaches to tackle this problem.
+effect size for each selected case?". \cite{efron2008} notes the potential for EB approaches to tackle this problem,
+and \cite{zhao.hwang} consider in detail the case where the non-null effects are normally distributed.
 %but also notes the difficulties of implementing a stable general algorithm.
 
-These features of the EB approach are extremely attractive in principle.
-But their usefulness in practice
+The ability of the EB approach to provide shrunken posterior distributions and interval estimates for parameters
+that are valid ``post-selection" is extremely attractive in principle.
+But its usefulness in practice
 depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
-which is notoriously difficult to solve in general. Indeed, Efron emphasises
+which are notoriously difficult to solve in general. Indeed, Efron emphasises
 the difficulties of implementing a stable general algorithm, noting in his rejoinder
@@
 generally satisfactory parametric formulation..." (Efron, rejoinder, p46).
-The key point we want to emphasise here is that a stable non-parametric
-approach is not only possible, but in fact quite straightforward, {\it provided we restrict ourselves
-to settings where we can assume that $g$ is unimodal}.
+The key point we want to emphasise here is that {\it if we restrict ourselves
+to settings where we can assume that $g$ is unimodal}
+then a stable non-parametric approach becomes not only possible, but indeed quite straightforward.
 While not meeting Efron's desire for an entirely general nonparametric approach, we nonetheless believe 
@@
 The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
-While the UA is not the only way to achieve this, it is a simple and effective approach.
+While the UA is not the only way to achieve this, we find it an attractive, simple and effective approach.
 
@@
 clearly visible in the observed $\bhat$.
+
+We note one caveat on the accuracy of estimated $g$: we tend to systematically
+overestimate the mass of $g$ near zero, due to our deliberate use of a penalty term (\ref{}) to avoid under-estimation of $\pi_0$.
+This is visually apparent in Figure \ref{fig:egcdf} (on careful inspection): the estimated cdf is generally below the true cdf just to the left of zero,
+and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect,
+and applying our methods without the penalty term removes this systematic effect (although at the cost of sometimes 
+under-estimating $\pi_0$).
+
+\begin{figure}[h!] \label{fig:egcdf}
+\begin{subfigure}{\textwidth}
+	\includegraphics[height=2in]{figures/egcdf.pdf}
+	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with mixFdr.}
+\end{subfigure}
+\begin{subfigure}{\textwidth}
+\includegraphics[height=2in]{figures/mean_cdf.pdf}
+\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{}) so $\pi_0$ is systematically overestimated.}
+\end{subfigure}
+\begin{subfigure}{\textwidth}
+\includegraphics[height=2in]{figures/mean_cdf_nopen.pdf}
+\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{}) so $\pi_0$ is not systematically overestimated.
+Modest systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the modest effects of model mis-specification.}
+\end{subfigure}
+\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.}
+\end{figure}
 
@@
 
-\subsubsection*{Reliable estimates of $g$ lead to reasonable coverage behaviour}
-
-Coverage rates are given in:
-\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
-
-Zhao+Hwang consider in detail the case where the non-null effects are normally distributed.
-
-Results in efron.fcr.rmd suggests that lower CI is OK; upper is more problematic because of difficulty in accurately estimating tail?
-
-\begin{figure} \label{fig:egcdf}
-\includegraphics[height=2in]{figures/egcdf.pdf}
-\includegraphics[height=2in]{figures/mean_cdf.pdf}
-\end{figure}
+\subsubsection*{Calibration of posterior distributions}
+
+To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over all 100 simulations in each Scenario). 
+Specifically we examine the empirical coverage of nominal lower 95\% credible bounds for a) all observations; b) significant negative discoveries; c) significant positive discoveries.
+We examine only lower bounds because the results for upper bounds follow by
+symmetry (except for the one asymetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
+in each case: for negative discoveries the lower bound is typically large and negative and limits how big (in absolute value) 
+the effect could be; for positive discoveries the lower bound 
+is positive, and limits how small (in absolute value) the effect could be. Intuitively, the lower bound for negative discoveries depends on the accuracy of $g$ in its tail,
+whereas for positive discoveries it is more dependent on the accuracy of $g$ in the center.
+
+The results are shown in Table \ref{tab:coverage}.
+
+As noted above, the use of a penalty term to avoid underestimating $\pi_0$ means that the estimated $g$s tends to have, systematically, slightly too much mass near 0.
+If the prior has too much mass near 0 then so will the posterior, so we should expect posterior intervals may also be ``over-shrunk" towards zero. 
+The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound will be negative and overshrinking will move the bound towards zero, making the left tail too big; for positive discoveries, the lower confidence bound will be positive, and
+overshrinking will move it toward zero, making the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
+and be prepared to put up with some overshrinkage to avoid it.
+
+%Coverage rates are given in:
+%\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
+
+\section*{Problems with removing the penalty term in the half-uniform case}
+
+When we tried using no penalty term in the half-uniform case we came across an unanticipated problem: when the data
+are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
+sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
+for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
+and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
+
+Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.
 
@@
 \begin{subtable}{\textwidth}
-	\input{tables/coverage_all.tex}
-	\caption{All}
+	\input{tables/scoverage_all.tex}
+	\caption{All observations.
+Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario.
+This is due to the penalty term \ref{} which tends to cause over-shrinking towards zero; removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods.
+(Removing the penalty in the half-uniform case can lead to unexpected problems explained in section \ref{sec:unexpected}).}
 \end{subtable}
@@
 \begin{subtable}{\textwidth}
-\input{tables/coverage_neg.tex}
-\caption{Negative findings}
+\input{tables/scoverage_neg.tex}
+\caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
+and the normal-based method in the flat-top scenario.
+These results likely reflect inaccurate estimates of the tails of $g$ due to a disconnect between the tail of $g$ and the component distributions in these cases.
+For example, the uniform methods sometimes substantially underestimate the length of the tail of $g$ in these long-tailed scenarios,
+causing over-shrinkage of the tail toward 0.}
 \end{subtable}
@@
 \begin{subtable}{\textwidth}
-\input{tables/coverage_pos.tex}
-\caption{Positive findings}
+\input{tables/scoverage_pos.tex}
+\caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
 \end{subtable}
 
-\caption{Table of 5\% lower tail coverage rates by scenario and method}
+\caption{Table of empirical coverage for nominal 95\% lower credible bounds. } \label{tab:coverage}
+\end{table}
+
+\begin{table}[!ht]
+\begin{subtable}{\textwidth}
+	\input{tables/scoverage_all_nopen.tex}
+	\caption{All observations}
+\end{subtable}
+
+\begin{subtable}{\textwidth}
+\input{tables/scoverage_neg_nopen.tex}
+\caption{``Significant" negative discoveries.}
+\end{subtable}
+
+\begin{subtable}{\textwidth}
+\input{tables/scoverage_pos_nopen.tex}
+\caption{``Significant" positive discoveries.}
+\end{subtable}
+
+\caption{Table of empirical coverage for nominal 95\% lower credible bounds (methods without the penalty term. }
 \end{table}
@@
 \centering
-\subfloat[][Density histograms of $p$ values for good-precision, poor-precision, and combined observations]{
+\begin{subfigure}{\textwidth}
 \includegraphics[width=6.5in]{Rcode/figures/GOODPOOReg_hist.pdf} 
     \label{fig:goodpoor-hist}
-} \\
-\subfloat[][Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the estimated error rate ($q$ value, lfdr, or lfsr) for a good-precision observation under the two different analyses.]{
+    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations}
+ \end{subfigure}
+\begin{subfigure}{\textwidth}
 \includegraphics[width=6.5in]{Rcode/figures/GOODPOOReg_scatter.pdf}
-    \label{fig:goodpoor-scatter}
-}
+    \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
+Each point shows the estimated error rate ($q$ value, lfdr, or lfsr) for a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
+\end{subfigure}
 \caption{Simulation illustrating how poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
@@
 seen as an approximation.
+
+\subsection{various notes...}
+
+Here performing inference means obtaining an (approximate) posterior distribution for each effect $\beta_j$,
+which can be used to estimate False Discovery Rates, and to produce both point and interval estimates for $\beta_j$.
+ Although ``multiple comparisons" is usually presented as a ``problem", our emphasis here is that the multiple measurements
+ actually provide an opportunity: the hierarchical model combines information across measurements to improve both accuracy and precision of estimates.
+
+Note that incorporating differential errors messes up the EM algorithm?
+
+we start by noting that the $p$ values in Figure \ref{} were actually generated under a scenario where
+{\it no tests were null}! Thus the qvalue procedure, which estimates the proportion of nulls to be xx\%, is acting very conservatively in this case.
+However, it could be argued that some degree of conservativeness is inevitable, and that nonetheless the procedure is acting sensibly.
+Therefore, to illustrate this point we really want to focus on the fact that, at least in some settings, the assumption is {\it unrealistic}.
+Specifically, we want to argue that the implied distribution of the alternative $p$ values is unrealistic. In fact, this is hard to see in $p$ value space,
+and so instead we translate these $p$ values back to the $z$ scores that lead to these $p$ values.
+
+The second issue is that directly modeling the p values, or z scores, say via non-parametric methods, 
+can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
+distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
+The solution is to model beta hat as a convolution of some distribution g and an error component.
 

### 2015-05-01 dsc-shrink/plot_cdf_eg.R

@@ -60,4 +60,8 @@ pdf("../paper/figures/mean_cdf.pdf",height=3,width=9)
 plot_mean_cdf(1:100,PLOTMETHODS=c("ash.n","ash.u","ash.hu","truth"),pi0filter=TRUE)
 dev.off()
 
-
+pdf("../paper/figures/mean_cdf_nopen.pdf",height=3,width=9)
+names(myColors) <- c("mixfdr.tnull","ash.hu.s","ash.n.s","ash.u.s","qvalue","locfdr","truth")
+colScale <- scale_colour_manual(name = "method",values = myColors)
+plot_mean_cdf(1:100,PLOTMETHODS=c("ash.n.s","ash.u.s","ash.hu.s","truth"),pi0filter=TRUE)
+dev.off()

### 2015-05-24 talks/ilike2015.rmd
add ilike talk on dsc

--- 2015-03-16 talks/one_hour_seminar.Rmd
@@
 ---
-title: "False Discovery Rates, A New Deal"
+title: "Dynamic Statistical Comparisons, and other adventures in getting organized"
 author: Matthew Stephens
-date: March 9, 2015
+date: May 22, 2015
 output: ioslides_presentation
@@
 
-## Before we get started: Let's get Organized!
+## Getting Organized!
 
@@
 - Organized researchers get more done (and better!).
-- Many of them are more organized than I am!
+- Many of them are more organized than I am.
 - Thought: I should get organized; I should help others get organized.
@@
 
+## What is knitr?
+
+ ![Image](screenshot.png)
+
 ## What is Reproducible Research?
@@
 
-## Commercial Break:
+## More on git, github, knitr, reproducibility
 
-A quick ad for a recently-started project I call "Dynamic Statistical Comparisons":
+- Google "The git book", to get started on git.
+- Google "Karl Broman github tutorial" for statistics-oriented intro to github.
+- Google "donohoe buckheit" for "Wavelab and reproducible research"
 
-- Overall Goal: Improve how the statistics and genomics (and other) communities make comparisons among methods.
+## More ridiculously-personal tips
+
+- "Don't try to show them how clever you are" (BD Ripley)
+- Minimise email for research communications 
+- Chunk time
+- Relax! (Exercise, Vacation)
+- Admin on Fridays
+- Learn some programming principles
+- Benchmark early and often
+
+## Minimise email
+
+![Image](slack.screenshot.png)
+
+## Benchmarking:
+
+By which I mean, any quantitative assessment and comparisons of performance (accuracy; speed).
+
+## Benchmarking:
+
+Benefits include:
+
+- Communication
+- Tweaking/honing
+- Basic quality assurance (cf "testing")
+
+## Benchmarking:
+
+Even basic benchmarks have benefits (...Done is better than perfect)
+
+## Dynamic Statistical Comparisons
+
+- Overall Goal: Improve how we, in the statistics community, make comparisons among methods.
+
+- Reduce work and redundancy; make comparisons more transparent
 
 - How? Centralize reproducible statistical comparisons in easily-extensible public repositories.
+
+## Inspiration: Public Benchmarking
+
+- Protein folding (CASP)
+
+- Image analysis (Pascal)
+
+- The Neflix Prize
+
+- Genetics Analysis Workshop
+
+- RECOMB/DREAM challenge
+
+- Sage Bionetworks
 
@@
 
-## More on git, github, knitr, reproducibility, DSCs
-
-- Google "The git book", to get started on git.
-- Google "Karl Broman github tutorial" for statistics-oriented intro to github.
-- Google "donohoe buckheit" for "Wavelab and reproducible research"
 - [http://stephens999.github.io/blog/](http://stephens999.github.io/blog/) for more on DSC
 
-## The Canonical Genomics Experiment
+## DSC Requirements
 
-- Measure lots of things, with error
+- Well-defined input format for methods
+- Well-defined output format
+- Scoring function (output, input, meta-data)
+- Ways to produce input, meta-data
+- Methods to turn input -> output
 
-- Get estimates of effects $\beta_j$ ($\hat\beta_j$) and their standard errors $s_j$
+## DSCR - an R package
 
-- Turn these into Z-scores, $z_j = \hat\beta_j/s_j$
+- Define Scenarios (produce input, meta-data)
+- Define Methods (turn input -> output)
+- Define Scores
 
-- Turn these into $p$ values, $p_j$
+A simple function call then runs all methods on all scenarios, outputs scores
 
-- Apply `qvalue` 
-to identify findings ``significant" at a given FDR.
+[Simple Vignette](http://htmlpreview.github.com/?https://github.com/stephens999/dscr/blob/master/vignettes/one_sample_location.html)
 
-- ...?
+## DSC examples we're working on
 
-## FDR, local fdr, and q values
+- Covariance Estimation
+- Large-scale linear regression
+- 1-D Smoothing with Wavelets [after Antoniadis et al, 2001]
 
-We take a Bayesian perspective and use the following definitions
-(see Storey, 2003, for connections with Frequentist definitions):
-
-- The FDR at a threshold $P$ is 
-$$\text{FDR}(P)=\Pr(\beta_j = 0 |  p_j<P).$$
-
-- The q value for observation $j$ is $q_j=\text{FDR}(p_j)$.
-
-- The local false discovery rate, fdr, at threshold $P$ is 
-$$\text{fdr}(P) = \Pr(\beta_j =0 | p_j=P ).$$
-
-- The fdr is more relevant, but slightly harder to estimate than 
-FDR because it involves density estimation rather than tail-area estimation.
+## Wavelet Smoothing Example: The "Blocks" Function
 
 ```{r, echo=FALSE}
-  library("ashr") #simple simulated example
-  ncz = 100 ## number of bins in z score histograms
-  set.seed(111)
-  hh.betahat = rnorm(1000,0,2)
-  hh.sebetahat = 1
-  hh.zscore = hh.betahat/hh.sebetahat
-  hh.pval = pchisq(hh.zscore^2,df=1,lower.tail=F)
-  hh.ash = ash(hh.betahat,hh.sebetahat, method="fdr")
-  hh.q = qvalue(hh.pval)
+  blocks = function(n){
+     t=1:n/n
+    pos=c(.1, .13, .15, .23, .25, .40, .44, .65, .76, .78, .81)
+    hgt = 2.88/5*c(4, (-5), 3, (-4), 5, (-4.2), 2.1, 4.3, (-3.1), 2.1, (-4.2))
+    mu.blk = rep(0,n)
+    for(j in 1:length(pos)){
+      mu.blk = mu.blk + (1 + sign(t-pos[j]))*(hgt[j]/2)
+    }
+    mu=0.2+0.6*(mu.blk-min(mu.blk))/max(mu.blk-min(mu.blk))
+    return(mu)
+  }
+len = 1024
+x = seq(0,1,length=len)
+mu=blocks(len)
+plot(x,mu,type="l",col=2,lwd=2)
+y=mu+rnorm(len,0,sd=sd(mu))
+points(x,y,cex=0.5)
 ```
 
-```{r, echo=FALSE}
-  plot_FDReg_hist = function(hh.pval,pi0,nc=40,nullcol="blue",altcol="cyan",type=4,title="Distribution of p values",...){
-    hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main=title,nclass=nc,col=altcol,...)
-    if(type>1){
-      abline(h=pi0,col=nullcol,lwd=2)
+## Wavelet Transform of Noisy Data
 
-      hh.hist$density=rep(pi0,length(hh.hist$density))  
-      #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
-      plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)
-    }
-    if(type>2){
-    abline(v=0.1,lwd=2,col=2)
-    }
-    if(type>3){
-      text(0.05,1.2,labels="A",col=2,cex=1.2)  
-      text(0.05,0.4,labels="B",col=2,cex=1.2)  
-      text(0.6,3,labels=paste0("FDR = B/(A+B) =  ",round(pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.2)
-    }
-  }
+```{r,echo=FALSE}
+  y.wd = wavethresh::wd(y,filter=1,family="DaubExPhase")
+  plot(y.wd,first.level=5) #Haar transform
 ```
 
-## Example: FDR estimation
+## Wavelet Transform of True Mean
 
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=1)                 
+```{r, echo=FALSE}
+mu.wd = wavethresh::wd(mu,filter=1,family="DaubExPhase")
+plot(mu.wd,first.level=5) #Haar transform
 ```
 
-## Example: FDR estimation
+## Wavelet Shrinkage
 
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=2)                 
+- WCs of true mean ($$\beta$$) are "sparse" (or at least, have spiky density about 0)
+
+- WCs for observed data ($$\betahat$$) are noisy observations of true WCs.
+
+- If we shrink the $$\betahat$$ toward 0, they will give a better estimate for $$\beta$$
+
+- And inverting the wavelet transform will give better estimate for $\mu$
+
+## Shrink + Backtransform = Smooth
+
+```{r, echo=FALSE}
+y.wd.thresh = wavethresh::threshold(y.wd, policy="BayesThresh")
+y.est = wavethresh::wr(y.wd.thresh)
+plot(x,mu,type="l",col=2,lwd=2)
+lines(x,y.est,type="l")
 ```
 
-## Example: FDR estimation
+## How to Shrink?
 
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=3)                 
-```
+We've been developing generic EB approaches to shrinkage estimation:
 
-## Example: FDR estimation
+- Assume $\beta_j \sim g()$ where $g$ has some concentration about 0 ("sparse").
 
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4)                 
-```
+- Assume we have corresponding estimates $\hat\beta_j$, with standard error $s_j$.
 
-## Is this an important problem?
+- 
+## How to Shrink?
 
-- The original paper introducing FDR (Benjamini and Hochberg, 1995) has
-been cited 21,787 times (May 2014) according to Google Scholar.
+- *Incorporate precision* of each observation $\hat\beta$ by assuming
+$$\hat\beta_j | \beta_j \sim N(\beta_j, s_j^2)$$
 
-- That is three times a day for the last 19 years!
+- Constrain $g$ to be *unimodal* about 0;
 
-## Problem 1: The Zero Assumption (ZA)
+- Shrink via two-stage procedure: first estimate $g$ "non-parametrically" from data; then
+compute $p(\beta_j | \hat{g}, \hat\beta_j,s_j)$.
 
-- The standard `qvalue` 
-approach assumes that all the $p$ values near 1 are null.
-
-- Analogously, one can assume that all Z scores near 0 are null. Efron refers to this as the ``Zero Assumption".
-
-- Seems initially natural.
-
-```{r, echo=FALSE, include=FALSE}
-require(fdrtool)
-  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue",plot=FALSE)
-require(locfdr)
-  hh.locfdr = locfdr(hh.zscore,nulltype=0,plot=0)
-require(mixfdr)
-  hh.mixfdr = mixFdr(hh.zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
-```
-
-## Implied distribution of $p$ values under $H_1$
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4)                 
-```
-
-## Implied distribution of Z scores under alternative
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-#plot a histogram of z scores, highlighting the alternative distribution
-#of z scores that is implied by localfdr values lfdr.
-  nullalthist = function(z,lfdr,nullcol="blue",altcol="cyan",...){
-    h=hist(z, freq=FALSE,col=nullcol,nclass=ncz,...)
-    avlfdr = unlist(lapply(split(lfdr,cut(z,h$breaks),drop=FALSE),mean))
-    h$density = (1-avlfdr) * h$density
-    plot(h,add=TRUE,col=altcol,freq=FALSE)
-  }
-
-  nullalthist(hh.zscore,hh.fdrtool$lfdr)  
-```
-
-## FDR problem 2: different measurement precision
-
-- In some cases the measurement precisions differ among units
-
-- Eg Expression levels of low-expressed genes have less precision than high-expressed genes
-
-- If some effects are measured less precisely than others, those tests ``lack power"
-and dilute signal, increasing FDR
-
-## Example: Mouse Heart Data
-
-```{r, echo=FALSE}
-x = read.table(paste0("../../stat45800/data/nobrega/expression/counts.txt"), header = TRUE)
-xx = rowSums(x[,2:5])
-x = x[xx>0,]
-xx = xx[xx>0]
-```
-
-```{r echo=FALSE}
-head(x)
-```
-
-- Data on 150 mouse hearts, dissected into left and right ventricle
-(courtesy Scott Schmemo, Marcelo Nobrega)
-
-```{r, echo=FALSE,include=FALSE}
-
-load("../scratchwork/Smemo_qbinom_zdat.RData")
-## two-sided test
-  ttest.pval = function(t, df) {
-    pval = pt(t, df = df, lower.tail = T)
-    pval = 2 * ifelse(pval < 0.5, pval, 1 - pval)
-    return(pval)
-}
-
-tscore = zdat[3, ]/zdat[4, ]
-pval = ttest.pval(tscore, df = 2)
-qval = qvalue(pval)
-
-highxp = xx>1000 ## select high expressed genes
-pval.high = pval[highxp]
-qval.high = qvalue(pval.high)
-
-zdat.ash = ash(zdat[3,],zdat[4,],df=2,method="fdr",mixcompdist="halfuniform")
-zdat.ash.high = ash(zdat[3,highxp],zdat[4,highxp],df=2,method="fdr",mixcompdist="halfuniform")
-zdat.ash.ES = ash(zdat[3,],zdat[4,],df=2,method="fdr",mixcompdist="halfuniform",model="ES")
-
-```
-
-## Example: Mouse Heart Data
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-qval= qvalue(pval)
-plot_FDReg_hist(pval,pi0=qval$pi0,ylim=c(0,4))
-```
-
-## Mouse Data: Counts vary considerably across genes
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-hist(log10(xx),main="Distribution of total counts", xlab="log10(counts)")
-```
-
-## Lower count genes, less power
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-qval.low=qvalue(pval[!highxp])
-pval.low=pval[!highxp]
-plot_FDReg_hist(pval.low,pi0=qval.low$pi0,ylim=c(0,4))
-```
-
-## Higher count genes, more power
-```{r, echo=FALSE, fig.height=4,fig.cap=""}
-plot_FDReg_hist(pval.high,pi0=qval.high$pi0,ylim=c(0,4))
-```
-
-## Low-count genes dilute signal at high-count genes
-```{r, echo=FALSE, fig.height=4,fig.cap=""}
-plot_FDReg_hist(pval,pi0=qval$pi0,ylim=c(0,4))
-```
-
-## FDR problem 2: low count genes add noise, increase q values
-```{r,echo=FALSE, fig.height=4,fig.cap=""}
-plot(qval.high$qval,qval$qval[highxp],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
-abline(a=0,b=1,col=2)
-```
-
-## Problems: Summary
-
-Standard tools are unduly conservative.
-
-- The ZA, which implies actual effects have a (probably unrealistic) bimodal distribution; causes overestimate of $\pi_0$, losing power.
-
-- By focussing on $p$ values, low-precision measurements can dilute high-precision measurements.
-
-## FDR via Empirical Bayes
-
-- Following previous work (e.g. Newton, Efron, Muralidharan) we take an empirical Bayes approach to FDR.
-
-- Eg Efron assumes that the $Z$ scores come from a mixture of null, and alternative:
-$$Z_j \sim f_Z(.) = \pi_0 N(.;0,1) + (1-\pi_0) f_1(.)$$
-where $f_1, \pi_0$ are to be estimated from the data.
-
-- Various semi-parametric approaches taken to estimating $f_1$. For example,
-Efron uses Poisson regression; Muralidharan uses mixture of normal distributions.
-
-- Once $f_1$ and $\pi_0$ estimated, FDR calculations are straightforward.
-
-## FDR: A New Deal
-
-- Instead of modelling $Z$ scores, model the effects $\beta$,
-$$\beta_j \sim \pi_0 \delta_0(.) + (1-\pi_0) g(.)$$
-
-- Constrain $g$ to be unimodal about 0; estimate $g$ from data.
-
-- *Incorporate precision* of each observation $\hat\beta$ into the likelihood.
-Specifically, approximate likelihood for $\beta_j$ by a normal: 
-$$L(\beta_j) \propto \exp(-0.5 (\beta_j - \hat\beta_j)^2/s_j^2).$$
-[From $\hat\beta_j \sim N(\beta_j, s_j)$]
-Or, better, use a $t$ likelihood if $s_j$ estimated using few observations.
-
-## FDR - A New Deal
+## Details
 
@@
 Includes almost all priors used for sparse regression problems (spike-and-slab, double exponential/Laplace/Bayesian Lasso, horseshoe).
-
-## FDR - A New Deal
-
-- Alternatively, a mixture of uniforms, with 0 as one end-point of the range,
-provides still more flexibility, and in particular allows for asymmetry.
-
-- If allow a very large number of uniforms this provides the non-parametric mle for $g$; cf Grenander 1953; Cordy + Thomas 1997.
 
@@
 
-- This approach actually provides a full posterior distribution for each $\beta_j$.
+- Approach provides posterior distribution for each $\beta_j$.
 
-- So easy to obtain point estimates and credible intervals.
+- For wavelet shrinkage we just use posterior mean.
 
-- Because $g(\beta)$ is unimodal, the point estimates (and CIs) will tend to be ``shrunk" towards the overall mean (0).
+- Because $g(\beta)$ is unimodal, the point estimates (and CIs) will tend to be "shrunk" towards the overall mean (0).
 
@@
 
-- So we call the approach ``Adaptive Shrinkage" (ASH).
+- So we call the approach "Adaptive Shrinkage" (ash).
 
-## Recall Problem 1: distribution of alternative Z values multimodal
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-nullalthist(hh.zscore,hh.fdrtool$lfdr,main="qvalue")  
+## Smoothing with Adaptive Shrinkage (smash)
+
+```{r, echo=FALSE, results='hide', message=FALSE}
+load("../../dscr-smash/res.RData")
+library("dplyr")
+library("ggplot2")
+res.1.v1 = res %>% filter(grepl(".1.v1",scenario)) %>% filter(method %in% c("smash.haar","smash.homo.s8","ebayesthresh","sure.homo.s8"))
+PLOTNAMES=c("angles","blip","blocks","bump","corner","doppler","spike")
+res.1.v1$scenario=factor(res.1.v1$scenario,levels=paste0(c("ang","blip","blk","bump","cor","dop","sp"),".1.v1"))
+levels(res.1.v1$scenario)=PLOTNAMES
+ggplot(res.1.v1,aes(x=method,y=mise,fill=method)) + geom_violin() + facet_grid(.~scenario) + theme(axis.text.x=element_blank())
 ```
 
-## Problem Fixed: distribution of alternative Z values unimodal
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-nullalthist(hh.zscore,hh.ash$lfdr,main="ash")
+## Homoskedastic errors
+
+```{r, echo=FALSE, results='hide', message=FALSE}
+res.1.v1 = res %>% filter(grepl(".1.v1",scenario)) %>% filter(method %in% c("smash.haar","smash.homo.s8","ebayesthresh"))
+
+PLOTNAMES=c("angles","blip","blocks","bump","corner","doppler","spike")
+res.1.v1$scenario=factor(res.1.v1$scenario,levels=paste0(c("ang","blip","blk","bump","cor","dop","sp"),".1.v1"))
+levels(res.1.v1$scenario)=PLOTNAMES
+
+ggplot(res.1.v1,aes(x=method,y=mise,fill=method)) + geom_violin() + facet_grid(.~scenario) +theme(axis.text.x=element_blank())
 ```
 
-## Example: FDR estimation
+## Heteroskedastic errors
 
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4,title="qvalue")                 
+```{r, echo=FALSE, results='hide', message=FALSE}
+res.1.v4 = res %>% filter(grepl(".1.v4",scenario)) %>% filter(method %in% c("smash.haar","smash.homo.s8","ebayesthresh"))
+res.1.v4$scenario=factor(res.1.v4$scenario,levels=paste0(c("ang","blip","blk","bump","cor","dop","sp"),".1.v4"))
+levels(res.1.v4$scenario)=PLOTNAMES
+ggplot(res.1.v4,aes(x=method,y=mise,fill=method)) + geom_violin() + facet_grid(.~scenario) + theme(axis.text.x=element_blank())
 ```
 
-## Example: FDR estimation
+## Question
 
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,get_pi0(hh.ash),type=4,title="ash")                 
-```
-
-## Recall Problem 2: low count genes add noise, increase q values
-```{r,echo=FALSE, fig.height=4,fig.cap=""}
-plot(qval.high$qval,qval$qval[highxp],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
-abline(a=0,b=1,col=2)
-```
-
-## Problem Fixed: incorporating precision reduces influence of low-count genes
-```{r,echo=FALSE, fig.height=4,fig.cap=""}
-plot(zdat.ash.high$qvalue,zdat.ash$qvalue[highxp],xlab="q values from high-count gene analysis", ylab="q values from all gene analysis", main="q values for high count genes",xlim=c(0,.5),ylim=c(0,.5))
-abline(a=0,b=1,col=2)
-```
-
-## A new problem: an embarrassment of riches
-
-- If the null is mostly false, the new approach can  provide unsettling results
-
-- The FDR can be small for all observations, even those with $p \approx 1$!
-
-- In the illustrative example, the maximum $q$ value is `r round(max(hh.ash$qvalue),2)`
-
-## Perhaps we didn't really understand the question?
-
-- Problem arises only if we insist on asking question ``is $\beta_j=0$?"
-
-- Given enough signal, we become convinced that very few of the $\beta_j=0$
-
-- But for some $\beta_j$ we still may have little information about actual value
-
-- Suggests a change of focus: ask for which $\beta_j$ are we confident about the sign (cf Gelman et al, 2012).
-
-## The False Sign Rate
-
-- Suggestion: replace FDR with local false sign rate (lfsr), the probability that if we say an effect is positive (negative), it is not.
-
-- Example: suppose we estimate that $\Pr(\beta_j<0)=0.95, \Pr(\beta_j=0) = 0.025$ and $\Pr(\beta_j>0)=0.025$. Then we report $\beta_j$ as a ``(negative) discovery", and estimate its lfsr as 0.05.
-
-## Even with many signals, large $p$ values have high lfsr
-
-```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-  #res=data.frame(p=hh.pval, lfsr=hh.ash$lfsr, qvalue = hh.ash$qvalue)
-  library("reshape2")
-  res=data.frame(p=pval, lfsr=zdat.ash$lfsr, qvalue = zdat.ash$qvalue)
-  res.melt = melt(res, id.vars=c("p"),variable.name="Measure")
-  cbbPalette <- c("#000000", "#D55E00", "#CC79A7")
-  labels = c('qvalue','lfsr')
-  breaks = c("qvalue","lfsr")
-
-  pp= ggplot(data=res.melt,aes(p,value,color=Measure)) +geom_point(shape=1) +
-      geom_abline(colour = "black") +
-        xlab("p-value") +
-        ylab("lfsr/qvalue")
-
-  print(pp +scale_y_continuous(limits=c(0,1)) +
-          scale_x_continuous(limits=c(0,1))  +
-           scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
-          coord_equal(ratio=1))
-
-```
-
-## Advantages of False Sign Rate
-
-- The FDR is typically not actually identifiable from data.
-
-- This is because data cannot distinguish between $\beta_j = 0$ and $\beta_j$ "very
-small". So $\pi_0$ is not identifiable, and FDR is very sensitive to $\pi_0$.
-
-- So methods for estimating $\pi_0$ and FDR, including those presented here, are designed to be ``conservative" (i.e. overestimate the FDR).
-
-- The False Sign Rate is much less senstive to $\pi_0$, and hence more identifiable from data!
-
-## Simulated example: $\pi_0$ not identifiable.
-
-```{r,echo=FALSE,include=FALSE}
-load("../paper/Rcode/sim1.RData") #load simulation results
-source("../paper/Rcode/plot_pi0.R")
-```
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-plot_pi0(list(simres1a))
-```
-
-## Simulated Example: so fdr not identifiable
-
-```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-source("../paper/Rcode/plot_lfsr.R")
-  plot_lfsr(list(simres1a),0.1,ptype="lfdr")
-```
-
-## Simulated Example: fsr much more identifiable
-
-```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-  plot_lfsr(list(simres1a),0.1,ptype="lfsr")
-```
+- Does anyone have applications where allowing for heteroskedastic errors might be useful?
 
@@
 
-- ASH provides a generic approach to shrinkage estimation, as well as
-false discovery (sign) rates.
-
-- But by using two numbers ($\hat\beta,s$) instead of one ($p$ values or $z$ scores) varying precision of 
-measurements is better accounted for.
-
-- Unimodal assumption for effects reduces conservatism
-
-- False Sign Rate preferable to False Discovery Rate: more identifiable, and
-better representation of information in data for ``high-signal" situations.
-
-## Other Applications
-
-- Widely applicable: requiring only an estimated
-effect size and standard error for each object.
-
-- E.g. Currently applying it to wavelet shrinkage applications.
-
-## Next steps?
-
-- Incorporate shrinkage of variances and not just means. (e.g. ``moderated $t$ test", Smyth et al)
-
-- Allow $g(\cdot;\pi)$ to depend on covariates $X$.
-
-- Allow for correlations in the measured $\hat\beta_j$.
+- We're looking for conspirators interested in organizing benchmarking!
 
@@
 
+- Zhengrong Xing
+
 - to the developers of **R**, **knitr**, **Rstudio** and **Pandoc**.
-Also the very cool **SQUAREM** package.
+Also the very cool **SQUAREM** and **wavethresh** packages.
 
@@
 
-- Including  Chaoxing Dai, Mengyin Lu, Ester Pantaleo, Scott Powers, Sen Tian, Wei Wang, Zhengrong Xing.
+- Including Chaoxing Dai, Mengyin Lu, Ester Pantaleo, Scott Powers, Sen Tian, Wei Wang, .
 

### 2015-06-23 examples/adaptive_shrinkage.rmd
update

---
title: "Adaptive Shrinkage Illustration"
output: html_document
---

The goal here is to illustrate the "adaptive" nature of the adaptive shrinkage. The shrinkage is adaptive in two senses. First,
the amount of shrinkage depends on the distribution $g$ of the true effects, which is learned from the data: when $g$ is very peaked about zero
then ash learns this and deduces that signals should be more strongly shrunk towards zero than when g is less peaked about zero.
Second, the amount of shrinkage of each observation depends on its standard error: the smaller the standard error, the more informative
the data, and so the less shrinkage that occurs. From an Empirical Bayesian point of view both of these 
points are entirely natural: the posterior depends on both the prior
and the likelihood; the prior, $g$, is learned from the data, and the likelihood incorporates the standard error of each observation.

We simulate from two scenarios, the first with effects more peaked about zero, the second with effects less peaked.
```{r}
set.seed(100)
library(ashr)
beta1 = c(rnorm(500,0,0.01),rnorm(500,0,0.5))
beta2 = rnorm(1000,0,1)
s = 1/rgamma(1000,10,10)
e = rnorm(1000,0,s)
beta1hat = beta1 + e
beta2hat = beta2 + e
```

Now run ash on both:
```{r}
beta1.ash = ash(beta1hat, s)
beta2.ash = ash(beta2hat, s)
```

```{r}
df = data.frame(betahat = c(beta1hat,beta2hat), beta_est = c(beta1.ash$PosteriorMean, beta2.ash$PosteriorMean),s=c(s,s),scenario=c(rep("peaked",1000),rep("less peaked",1000)))
library(ggplot2)
ggplot(df,aes(x=betahat,y=beta_est)) + geom_point(col=s) + facet_grid(.~scenario) 
#need to add equal scales, and color gradient

```

### 2015-06-23 examples/fsr.rmd
update

@@
+---
+title: "False Sign Rate Illustration"
+output: html_document
+---
+
 # FSR Example
@@
 ```{r}
-opts_chunk$set(fig.path='figure/lfsr/')
+#opts_chunk$set(fig.path='figure/lfsr/')
 set.seed(100)
@@
 The actual FSR at that threshold is `r mean(sign(z[p<0.05])!=sign(beta[p<0.05]))`.
+(The conservative estimates from $\ash$ are likely partly due to our deliberate use of a penalty term to provide conservative estimates of $\pi_0$.)

### 2015-06-23 paper/main.tex
update

@@
 the idea that it may be preferable to focus on estimation rather than on testing.
-In principle Bayesian approaches, including EB, can naturally unify testing and estimation into a single framework -- testing is
-simply estimation but with a point mass of some weight on the null hypothesis that an effect is exactly zero.
+In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
+simply estimation with some positive prior probability that the effect is exactly zero.
 However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
@@
 \begin{equation}
-q_j := \FDR(\{k:\fdr_k \leq \fdr_j\}).
-\end{equation}
-In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
+q_j := \widehat\FDR(\{k:\fdr_k \leq \fdr_j\}).
+\end{equation}
+%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
+%\begin{equation}
+%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
+%\end{equation}
+%which is close to the definition in \cite{storey.03}.
+
+\subsection*{The local False Sign Rate}
+
+\cite{gelman2012we} emphasise that there are settings where
+$\beta_j=0$ is implausible, in which case the FDR is not useful: if every $\beta_j$ is non-zero then there
+is no such thing as a false discovery and the FDR will always be identically 0. 
+Gelman et.~al.~suggest that in such settings we should replace the concept of a false discovery with the
+concept of an ``error in sign". The idea is that, in settings where $\beta_j=0$ is implausible, the most fundamental
+inference objective is to ask which $\beta_j$ are positive and which are negative.
+Indeed, even in settings where some
+$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
+analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
+Further, this idea has a long history: for example, \cite{tukey:xx}, p32, suggests one should address
+\begin{quote}
+...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
+\end{quote}
+
+Motivated by this, we define the local False Sign Rate (lfsr) for observation $j$ as 
 \begin{equation}
-q_j = \FDR(\{k:p_k \leq p_j\})
-\end{equation}
-which is close to the definition in \cite{storey.03}.
+\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
+\end{equation}
+Thus $\lfsr_j$ gives the probability that we would get the sign of $\beta_j$ wrong if we
+were to make our best guess. (Naturally, we consider it an error to state that $\beta$ is positive or negative when it is truly zero.)
+To illustrate, suppose for concreteness
+that the minimum is achieved by the first term, $p(\beta_j \geq 0| \bhat, s)=0.05$ say. Then
+our best guess would be that $\beta$ is negative, and the probability that we have
+made an error in sign would be 0.05.
+
+Note that $\lfsr_j \geq \lfdr_j$ 
+because both the events $\beta_j \geq 0$
+and $\beta_j \leq 0$ include the event $\beta_j=0$.
+Thus, lfsr gives an upper bound for lfdr.
+
+Because our methods involve explicitly modelling effects $\beta_j$, rather than 
+$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
+As show later, these estimates are much more robust to underlying assumptions than estimates
+of the FDR or $\lfdr$. Thus, although for comparison with existing methods we have
+focussed on FDR computations, in practice we strongly recommend using $\lfsr$ and not $\lfdr$.
 
@@
  Here we introduce an EB method, which, for reasons discussed above, we call ``adaptive shrinkage" or ``ash".
- The method provides not only estimated $\fdr_j$ and q values, but also shrinkage point estimates and credible intervals for $\beta_j$,
- and estimates of the local false sign rate, defined below.
+ The method provides not only estimated $\fdr_j$ and $q$ values, but also shrinkage point estimates and credible intervals for $\beta_j$,
+ and estimates of the local false sign rates.
 
 We begin by presenting the simplest version of our model, before embellishing it.
-Instead of attempting to partition the (observed) $z$ scores or $p$ values into null and alternative components,
-we use a model that partitions the (unobserved) effects $\beta_j$  into null and alternative components:
+Instead of partitioning the (observed) $z$ scores or $p$ values into null and alternative components,
+we use a model that partitions the (unobserved) effects $\beta_j$  into null and alternative components.
+That is, letting $g$ denote the distribution of the effects $\beta_j$, we write
 \begin{equation} \label{eqn:betaj}
-\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)
+g(\beta_j) = \pi_0 \delta_0 + (1-\pi_0) g_1(\beta_j)
 \end{equation}
@@
 $\shat_j$:
-\begin{equation} \label{eqn:bhat}
+\begin{equation} \label{eqn:normlik}
 \bhat_j | \beta_j, \shat_j \sim N(\beta_j, \shat^2_j).
@@
 
-Combining (\ref{eqn:betaj})-(\ref{eqn:bhat}), and assuming independence of the observations and effects for different $j$ (given $\pi_0,w_1,\dots,w_k$), yields
+Combining (\ref{eqn:betaj})-(\ref{eqn:normlik}), and assuming independence of the observations and effects for different $j$ (given $\shat, \pi_0,w_1,\dots,w_k$), yields
 a hierarchical model for $\beta,\bhat$ given $\shat$, with unknown hyper-parameters $(\pi_0,w_1,\dots,w_k)$. To simplify notation we define $\pi_k := (1-\pi_0)w_k$ 
@@
 
+The conditional distributions $p(\beta_j | \hat{\pi}, \bhat_j, \shat_j)$ 
+encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
+$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
+$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
+the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
+observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
+corresondingly little effect on the estimated FDR of other observations. 
+We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
+$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
+that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
+estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
+Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
+
+\subsubsection*{Discussion on UA}
+
+ Since we view the UA as a key assumption we briefly discuss it here.
+Although the UA will not apply to all situations, we argue that it will be reasonable in many practical contexts, especially those
+involving FDR where interest focusses on which $\beta_j$ differ from zero. This is because whenever ``$\beta_j=0$" is a plausible null hypothesis,  
+it seems reasonable to imagine that ``$\beta_j$ very near 0" is also plausible. From there it seems reasonable to imagine that larger effects become
+decreasingly plausible, and so the distribution of the effects will be unimodal about 0. 
+Alternatively, we can motivate  the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
+such shrinkage is desirable from several standpoints for improving estimation accuracy. 
+Note that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
+It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
+
+%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
+%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}. 
+In addition, we note that although the UA is not made by any commonly-used existing FDR methods,
+almost all analogous work in sparse regression models make the UA for the regression coefficents - 
+common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{}.
+(These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
+it may be fruitful to apply our methods to the regression context; indeed see \cite{goddard} for work in this vein.) 
+The UA assumption on regression coeffients is directly analagous to our UA for effects, and so we view its
+widespread use in the regression context as supporting its use here.
+
+In addition to these arguments, the UA also has a considerable practical benefit: it produces very simple procedures that are both computationally
+and statistically stable. We illustrate these features in Results.
+
 \subsubsection*{Embellishments}
 
-\subsubsection*{More flexible prior component distributions}
+\subsubsection*{More flexible unimodal distributions}
+
+To provide a more general approach to defining the distribution of effects $g(\beta_j)$, consider 
+\begin{equation} \label{eqn:g}
+g(\cdot; \pi) = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k f_k(\cdot) 
+\end{equation}
+where $f_k$ are pre-specified component distributions with one of the following forms: 
+\begin{enumerate}[(i)]
+\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$,
+\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,
+\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ or } U[\cdot; 0,a_k]$,
+\end{enumerate}
+where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
+(In (iii) we include both components in the mixture (\ref{eqn:g}), so for a grid of values $a_1,\dots,a_K$ there
+are actually $2K+1$ components in the mixture.)
+The version of our model described above corresponds to (i).
+Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
+under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
+(\ref{eqn:tlik}) introduced below.
+Our use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
+
+Moving from (i) to (iii) the representation (\ref{eqn:g}) becomes increasingly flexible. Indeed,
+using a large dense grid of $\sigma^2_k$ or $a_k$, we can allow $g$ to approximate,
+with arbitrary accuracy,
+\begin{enumerate}[(i)]
+\item any scale mixture of normals, which includes as special cases
+the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
+\item any symmetric unimodal distribution about 0.
+\item any unimodal distribution about 0.
+\end{enumerate}
+The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine:1938} and  \cite{shepp:1962}; see \cite{feller:1971}, page  158. 
+In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.
+
+Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
+The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
+See Section \ref{details} for our software defaults.
+
+%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting.
+For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint.
+(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
+%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
 
@@
 
-\subsubsection*{}
-
-DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
-To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
-Because the prior distribution is unimodal about zero,
-the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
-amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
-For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
-than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
-small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
-
-The posterior distribution $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ then reflects a trade-off
-between the information in the prior distribution, which is more-or-less concentrated near zero,
-
-There are two key points to appreciate here. First, by estimating the hyperparameters $\pi$ from the data, 
-the method adapts to the amount of signal in the data: if most $\beta_j$ are truly at or near zero then $\pi$ will reflect this by
-concentrating on zero and very small values of $\sigma_k$, and the posterior distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$
-will be tend also to be concentrated around zero (strong shrinkage). In contrast, if most $\beta_j$ are large then $\pi$ will be put most weight
-on components with large $\sigma_k$, and the posterior distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ will be approximately centered  on $\bhat_j$
-(weak shrinkage). Indeed, in the limit as $\pi$ puts all it's weight on very large $\sigma_k$ the prior becomes flatter and flatter
-and the posterior on $\beta_j$ becomes $N(\bhat_j,\shat_j)$. Thus, in this limit, the Bayesian credible intervals match standard confidence intervals.
-
- and only very strong observations will have a small fdr; if most $\beta_j$ are large
-then this will also be reflected in the estimates of $\pi$ and even borderline signals might have a small fdr. 
-This idea is also used by existing FDR methods, but here, by modelling the effects, it extends to our estimates of the effects:
-
-The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ encapsulate uncertainty in the values for $\beta_j$, combining information across
-$j=1,\dots,J$. The combining of the information occurs through estimation of
-$\pi$, which involves all of the data.
-END OF DISCUSSION
-
---- NOte 
-In principle we might prefer a full Bayes approach that accounts for uncertainty in $g$; however,
-we believe that in most practical applications uncertainty in $g$ will not be the most important concern, and 
-compromise this principle for the simplicity of the EB approach.
-
-\begin{equation} \label{eqn:Lpi}
-\bhat_j | \shat_j, \pi \sim \pi_0 N(\cdot; 0, \shat^2_j) + \sum_k \pi_k N(\cdot; 0,\sigma_k^2 + \shat^2_j) \quad \text{independently for $j=1,\dots,J$}.
-\end{equation}
-
-Our EB approach can now be simply stated. 
-\begin{enumerate}
-\item Estimate the parameters $\pi$ (and hence also the distribution $g_1$) by maximising the likelihood 
-\begin{equation}
-L(\pi) := \prod_j p(\bhat_j | \shat_j, \pi),
-\end{equation}
-where the terms of the likelihood are mixtures of the form (\ref{eqn:Lpi}). 
-\end{enumerate}
-
- In outline, our goal is to perform Bayesian inference for the true effects $\beta$ using information in the summary statistics $\bhat,\shat$.
- That is we aim to compute the joint posterior distribution $p(\beta | \bhat,\shat)$.
-This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
-from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
-with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
- By working with summary statistics $\bhat, \shat$, rather than more refined data, we aim to produce generic methods that 
- can be applied whenever such summary data are available - just as \qvalue can be applied to any set of $p$ values for example.
-Any attempt to produce generic methods is likely to involve a compromise between functionality and generality; 
-we believe that by working with two numbers ($\bhat_j,\shat_j$) for each observation, rather than one ($p_j$ or $z_j$), 
-we can gain substantially in functionality (e.g.~ we can estimate effect sizes, as well as testing, and we can take better account of variations in
-measurement precision across units $j$) while losing only a little in generality.
-
-We assume that (given $\shat$) the
-effects $\beta_j$ are independent draws from some common ``prior" distribution $g$:
- \begin{equation} \label{eqn:beta}
- \beta_j \, | \, g, \shat \sim g(\cdot) \quad \text{independently for $j=1,\dots,J$.}
- \end{equation}
- A key assumption, which distinguishes our approach from previous work, is that $g$ is unimodal (and, optionally, symmetric). 
- We assume here that the mode is at 0, and allow
- that $g$ can have a point mass at 0, corresponding to the ``null" effects.
-(An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood.)
-Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
+We can generalize the normal likelihood assumption (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
+  \begin{equation} \label{eqn:tlik}
+ \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
+ \end{equation}
+where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
+and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
+The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
+This generalization does not complicate inference when used with mixtures of uniform components for $g$ in (\ref{eqn:g}).
+For normal components the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
+
+Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
+\begin{equation} \label{eqn:tdist}
+(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
+\end{equation}
+However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
+In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
+currently pursuing this improved approach in work with Mengyin Lu, and results will be published elsewhere. We note here that
+modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
+where our simpler approach can produce unsatisfactory results (see later).
+
+\subsubsection*{Dependence of effects on standard errors}
+
+Equation (\ref{eqn:betaj}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
 This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
  \begin{equation} \label{eqn:beta-alpha}
- \frac{\beta_j}{\shat_j^\alpha} \, \big| \, g, \shat \sim g(\cdot) \quad \text{independently for $j=1,\dots,J$}
+ \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot)
  \end{equation}
 for any $\alpha \geq 0$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
-assuming that the ``standardized effects" $\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
+assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
 it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
-the assumption made by existing methods that rank tests by their $p$ values (or $z$ scores). See Results for further discussion.
-
- Second, we assume that the observations $\bhat$ are independent given $\beta$,
- with $\bhat_j$ depending on $\beta,\shat$ only through $\beta_j$ and $\shat_j$:
- \begin{equation} \label{eqn:lik}
- \bhat_j \, | \, \beta, \shat \sim p(\bhat_j | \beta_j, \shat_j), \quad \text{independently for $j=1,\dots,J$.}
- \end{equation}
-A natural assumption, made by \cite{wakefield:2009},  is that 
- \begin{equation} \label{eqn:normlik}
- \bhat_j \, | \, \beta_j, \shat_j \sim N(\beta_j, \shat_j),
- \end{equation}
- which should be a good approximation if the estimates $\bhat_j,\shat_j$ are based on a large number of observations.
- To try to improve this approximation when there are fewer observations we generalize this to
-  \begin{equation} \label{eqn:lik}
- \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
- \end{equation}
-where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
-and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
-
-Equation (\ref{eqn:lik}) is, of course, motivated by the standard asymptotic result
-\begin{equation} \label{eqn:tdist}
-(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
-\end{equation}
-However (\ref{eqn:tdist}) does not imply (\ref{eqn:lik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:lik}) it is conditioned on.
-In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
-currently pursuing this improved approach in work with Mengyin Lu, and results will be published elsewhere. We note here that
-modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is small (e.g. $<10$),
-where our simpler approach can produce unsatisfactory results (see later).
-
-Given the distribution $g$, the independence assumptions in (\ref{eqn:beta}) and (\ref{eqn:lik}) imply that the posterior distribution $\beta | g,\bhat,\shat$ conveniently 
-factorizes into independent components $p(\beta | g,\bhat,\shat) = \prod_j p(\beta_j | g, \bhat_j, \shat_j)$.
-The Empirical Bayes approach involves first estimating $g$ using all the data, and then computing quantities of interest conditional on the estimated $g$. That is:
- \begin{itemize}
- \item Estimate $g$ by maximising the likelihood $\hat{g} = \arg \max L(g)$ where $L(g) := p(\bhat | g, \shat)$ and the
- maximisation is subject to the constraint that $g$ is unimodal.
- \item Compute the conditional distributions $p(\beta_j | \hat{g} \bhat, \shat)$ ,
- from which we can compute all quantities of interest such as lfdr, lfsr, $q$-values, posterior credible intervals/regions etc.
-  \end{itemize}
-
- \begin{note}
-The conditional distributions $p(\beta_j | \hat{g}, \bhat, s)$ 
-encapsulate uncertainty in the values for $\beta_j$, combining information across
-$j=1,\dots,J$. The combining of the information occurs through estimation of
-$g$, which involves all of the data.
-In principle we might prefer a full Bayes approach that accounts for uncertainty in $g$; however,
-we believe that in most practical applications uncertainty in $g$ will not be the most important concern, and 
-compromise this principle for the simplicity of the EB approach.
-\end{note}
+the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.
+
+The model \ref{eqn:beta-alpha} for general $\alpha$ is actually easily fitted using the algorithm for $\alpha=0$. 
+To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
+ standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
+ Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
+yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
+In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
+treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
+the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
+
+Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
+how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
+it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
+rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
+and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
+this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. For brevity we
+do not investigate this idea any further here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
+
+\subsubsection*{Non-zero mode}
+
+An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood.
+
+\subsection*{Implementation Details}
+
+See Section \ref{sec:implement} for further implementation details.
 
 \def\dt{\text{dt}}
-\begin{note}
- A key property of the likelihood (\ref{eqn:lik}) [and indeed (\ref{eqn:normlik})] is that
-observations with a large $\shat_j$ have an essentially flat likelihood, and so do not affect inference for
-$g$ or the other effects. This contrasts with existing approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, even under the alternative hypothesis (assuming
-that the effect $\beta_j$ is small compared with $\shat_j$). These observations therefore do affect inference; specifically they add noise
-and inflate estimates of the FDR. Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-\end{note}
 
@@
 
- \begin{note}
-Although the unimodal assumption (UA) will not apply to all situations, we argue that it will be reasonable in many practical contexts. 
-For example, in contexts where interest focusses on which $\beta_j$ differ from zero, which suggests that ``$\beta_j=0$" is a plausible null hypothesis,  
-it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible, and that the distribution of the effects will be unimodal about 0. 
-Alternatively, we can motivate  the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
-such shrinkage is desirable from several standpoints for improving estimation accuracy. 
-Note that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
-It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
-
-%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
-%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}. 
-Note that, although the UA appears novel in this context,
-almost all analogous work in large-scale regression make the UA for the regression coefficents - 
-common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{}.
-One of our aims here is to emphasize the considerable practical benefits of making the UA in the FDR context.
-\end{note}
 %The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
@@
 
-\subsubsection*{Implementation outline}
-
-In practice, we implement the UA by assuming a parametric finite mixture for $g$, 
-\begin{equation} \label{eqn:g}
-g(\cdot; \pi) = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k f_k(\cdot) 
-\end{equation}
-where $\delta_0(\cdot)$ denotes a point mass at 0, $\pi=(\pi_0,\dots,\pi_K)$ are mixture proportions to be estimated (with $\pi_k \geq 0, \sum_k \pi_k=1$), and $f_k$ are pre-specified component distributions with one of the following forms:
-\begin{enumerate}[i)]
-\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$,
-\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,
-\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ or } U[\cdot; 0,a_k]$,
-\end{enumerate}
-where $N(\cdot; 0,\sigma^2)$ denotes the density of a normal distribution with mean 0 and variance $\sigma^2$ and
-$U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
-That is, the non-zero effects are modeled as either i) a mixture of zero-centered normal distributions; ii) a mixture of zero-centered uniform distributions;
-or iii) a mixture of ``zero-anchored" uniform distributions, which have one end of their range at 0. These approaches closely mirror approaches in \cite{cordy1997deconvolution}.
-[Case i) is implemented only for the special case $\nu \rightarrow \infty$ in (\ref{eqn:like}).]
-
-To see the flexibility of (\ref{eqn:g}) think of $K$ as ``large", and consider a dense grid of values of $\sigma^2_k$ or $a_k$ 
-spanning a wide range of values, from
-essentially zero to unnecessarily large. As $K$ increases, and the grid becomes denser, the three different choices for $f_k$ can allow $g$ to approximate,
-with arbitrary accuracy,
-\begin{enumerate}[i)]
-\item any scale mixture of normals, which includes as special cases
-the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in 
-high-dimensional regression settings such as the spike and slab prior, the horseshoe prior \ref{}, or the more general three-parameter beta prior \ref{}.
-\item any symmetric unimodal distribution about 0.
-\item any unimodal distribution about 0.
-\end{enumerate}
-The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine:1938} and  \cite{shepp:1962}; see \cite{feller:1971}, page  158. 
-Although it is natural to worry about ``overfitting" if $K$ is large, 
-in fact the unimodal constraint is sufficient to prevent serious problems with overfitting. For example, note that the unimodal
-constraint is sufficiently strong that there will exist a non-parametric
-maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
-Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
-This said, in practice even quite small values of $K$ (e.g.
-2-10) can provide reasonable performance.
-See Section \ref{details} for our software default choices of $K,\sigma^2_k,a_k$.
-
-% $g() = \sum_{k=1}^K (\pi^+_k U[0,a_k] + \pi^-_k U[-a_k,0])$
-
-Using the normal likelihood (\ref{eqn:normlik}) and any of the component forms (i)-(iii) for $f_k$ 
-yields a simple EM algorithm for estimating $\pi$ in Step 1, and simple analytic forms for the conditional
-distributions in Step 2. Using the $t$ likelihood (\ref{eqn:lik}) is equally simple for the uniform forms (ii) and (iii) for $f_k$ (but 
-tricky for the normal $f_k$ and so we have not implemented that case).
-We used the elegant R package {\tt SQUAREM} \cite{squarem}) to accelerate convergence of our EM algorithm.
-
-\subsection*{Likelihood for $\pi$}
-
-Note; this needs some tidying.
-
-Assuming independence across units, the likelihood for $\pi$ is 
-\begin{align}
-L(\pi) := p(D_1,\dots,D_J | \pi) = \prod_{j=1}^n p(D_j | \pi)
-& = \prod_j \int p(D_j | \pi, \beta_j) p(\beta_j | \pi) d\beta_j \\
-& = \prod_j \int p(D_j | \beta_j) g(\beta_j; \pi) d\beta_j \\
-& = \prod_j \int L(\beta_j) g(\beta_j; \pi) d\beta_j \\
-& \approx \prod_j \int \hat{L}(\beta_j) g(\beta_j; \pi) d\beta_j
-\end{align}
-
-Further, the assumptions  $\bhat_j | \beta_j, \shat_j \sim N(\beta_j,\shat_j)$ and $\beta_j | \shat_j, \pi \sim g(\cdot; \pi)$ (given by \ref{eqn:g}) yield: 
-\begin{equation}
-p(\bhat_j, \shat_j | \pi) = \pi_0 N(\bhat_j; 0,s^2_j) + \sum_k \pi_k \tilde{f}_{kj}(\bhat_j)
-\end{equation}
-where $\tilde{f}_{kj}$ denotes the density of a convolution of $f_k$ with an $N(0,s^2_j)$ density.
-For example, if $f_k(\cdot) = N(\cdot; 0,\sigma^2_k)$ then the convolution $\tilde{f}_{kj}$ is also normal, $\tilde{f}_{kj}(\cdot) = N(\cdot; 0,\sigma^2_k + s^2_j)$. If
-$f_k(\cdot) = U[\cdot; a_k,b_k]$ then $\tilde{f}_{kj}$ is the convolution of a normal with a uniform, which has density
-\begin{equation}
-\tilde{f}_{kj}(x) = (\Phi(b_k/\shat_j) - \Phi(a_k/\shat_j))/(b-a),
-\end{equation}
-where $\Phi$ is the cumulative distribution function of the standard normal $N(0,1)$ distribution.
-
-Putting these together, the joint model for the unobserved $\beta$ and the observed $\bhat, s$ is:
-\begin{align}
-p(\bhat, s, \beta | \pi) & = \prod_j g(\beta_j ; \pi) p(\bhat_j, \shat_j | \beta_j) \\
-& = \prod_j g(\beta_j ; \pi) L(\beta_j; \bhat_j,\shat_j).
-\end{align}
-
-We fit this hierarchical model using the following "Empirical Bayes" approach. First we estimate the hyper-parameters $\pi$ by maximizing the likelihood
-\begin{equation}
-L(\pi; \bhat, s) := p(\bhat, s | \pi) = \int p(\bhat, s, \beta | \pi) d\beta.
-\end{equation}
-This can be done very easily using an EM algorithm.
-Then, given this estimate $\hat\pi$, we compute the conditional distributions 
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
-\end{equation}
-
-[Note: a Variational Bayes approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
-
-%[maybe put picture of hierarchical model here]
-
-%A Taylor series expansion of the log likelihood around the maximum likelihood estimate $\bhat_j$  yields $$l(\beta_j) \approx l(\bhat_j) + 0.5* (\beta_j - \bhat_j)^2 l''(\bhat_j).$$ [Fill in details?]
-
-\subsection*{The local False Sign Rate}
-
-As pointed out by \cite{gelman2012we}, there are settings where
-$\beta_j=0$ is implausible, in which case the lfdr is not useful: if every $\beta_j$ is non-zero then there
-is no such thing as a false discovery and the lfdr will always be identically 0. 
-Gelman et.~al.~suggest that in such settings we
-might replace the concept of a false discovery with the
-concept of an ``error in sign". The idea is that, in settings where $\beta_j=0$ is implausible, the most fundamental
-inference objective is to ask which $\beta_j$ are positive and which are negative. Indeed, even in settings where some
-$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
-analysts will often separate the genes that are ``upregulated" in one condition from those that ``downregulated" in that condition.
-
-Motivated by this, we define the local False Sign Rate (lfsr) for observation $j$ as 
-\begin{equation}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
-\end{equation}
-Thus $\lfsr_j$ gives the probability that we would get the sign of $\beta_j$ wrong if we
-were to make our best guess. (Note that we count it as an error to state that $\beta$ is positive or negative when it is truly zero.)
-To illustrate, suppose for concreteness
-that the minimum is achieved by the first term, $p(\beta_j \geq 0| \bhat, s)=0.05$ say. Then
-our best guess would be that $\beta$ is negative, and the probability that we have
-made an error in sign would be 0.05.
-
-We note that the idea of focussing on tail areas, rather than point null hypotheses,
-has a long history. For example, 
-Tukey, p32, "the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?""
-
-Note that $\text{lfsr}_j \geq \text{lfdr}_j$ 
-because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ include the event $\beta_j=0$.
-Thus, lfsr gives an upper bound for lfdr.
-
-\subsection{Computation Outline}
-
-As outlined above, we fit the model using the following Empirical Bayes procedure:
-1. Estimate $\pi$ by maximizing the likelihood $L(\pi)$.
-2. Compute the conditional distributions $p(\beta_j | \bhat_j, \shat_j, \hat\pi)$.
-
-Using a normal likelihood $L(\beta_j)$, and assuming
-$g$ to be a mixture of normal distributions with fixed variances, 
-yields a simple EM algorithm
-for estimating $\pi$ in Step 1, and simple analytic forms for the conditional
-distributions in Step 2.
-
-We used the elegant R package {\tt SQUAREM} \cite{squarem}) to accelerate convergence of our EM algorithm.
-
-\subsection*{Some Details}
-
-\subsubsection*{Choice of grid for $\sigma_k, a_k$}
-
-\def\sigmamax{\sigma_{text{max}}}
-\def\sigmamin{\sigma_{text{min}}}
-When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
-in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
-as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
-based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
-is negative we set $\sigmamax = 8\sigmamin$.)
-
-When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.
-
-Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
-a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.
-
-\subsubsection*{Penalty term on $\pi$}
-
-In practice data do not distinguish between effects that are ``very small" and ``exactly zero".
-Thus, if a mixture component $f_k$ has most mass very near zero then
-this introduces non-identifiability between $\pi_k$ and $\pi_0$. In practice this is a major problem
-only if one focuses on false discovery rates rather than false sign rates: the fdr is sensitive to $\pi_0$ whereas
-the fsr is not. However, to make fdr (and fsr) estimates from our method ``conservative" we add a penalty 
-term $h(\pi;\alpha)$ to the likelihood to encourage over-estimation of $\pi_0$:
-\begin{equation}
-h(\pi;\alpha) = \prod_{k=0}^K \pi_k^{\alpha_k-1}
-\end{equation}
-where $\alpha_k \geq 1 \, \forall k$. The default is $\alpha_0=10$ and $\alpha_k=1$, which yielded
-consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0}).
-
-Although this penalty is based on a Dirichlet density, we don't consider this to be a ``prior distribution" for $\pi$.
-
-\subsubsection*{Initialization}
-
-We initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. 
-Our rationale to initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
-but weak signal in the data cannot quickly draw the algorithm towards the null.
-
 \section*{Results}
@@
 \end{equation}
-so $z_j:=\bhat_j/s_j=\bhat_j$ are standard normal if $H_j$ holds.
+so $z_j:=\bhat_j/s_j=\bhat_j$ are standard normal if null $H_j$ holds.
 
@@
 \item The UA can produce very different inferences compared with the ZA.
-\item The UA can yield conservative estimates of the proportion of nulls that hold, $\pi_0$, and hence conservative estimates of FDR and fdr.
+\item The UA can yield conservative estimates of the proportion of nulls that hold, $\pi_0$, and hence conservative estimates of FDR and \lfdr.
 \item The UA results in a stable procedure, both numerically and statistically.
@@
 estimate $\pi_0$ for each data set. For \ashr we show results for $g$ modelled as a mixture of normal components (``ash.n") 
-and a mixture of uniform components (``ash.u"). (Results using a mixture of ``half-uniforms" are generally similar to the uniforms and omitted for clarity.)
+and a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
+as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted for clarity.)
 
@@
 even for this largest difference results from the default run (e.g.
-the lfsrs, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood \url{github.com/stephens999/ash/dsc-robust/summarize_dsc_robust.rmd}.
+the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood \url{github.com/stephens999/ash/dsc-robust/summarize_dsc_robust.rmd}.
 
@@
 
-Although we have focussed up to now on estimating the FDR, an important advantage of
-an empirical Bayes approach like ours, compared with less-model-based alternatives like \qvalue, is that the EB 
-approach can provide
-measures of the uncertainty in the {\it size} of each effect $\beta_j$.
-Specifically, the
-EB approach provides a posterior distribution for each $\beta_j$, which can be used
+Although we have focussed on estimating the FDR, an important advantage of
+our EB approach based on modelling the effects $\beta_j$ (rather than $p$ values or $z$ scores)  is that it
+can estimate the {\it size} of each effect $\beta_j$.
+Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
 to address any question about the effect sizes, including, for example, ``which effects likely exceed $T$'', where
@@
 this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
-In our case, because $g$ is unimodal about zero, the effect of this prior is always to ``shrink" the posterior distributions
+In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
 towards zero, with the strength of the shrinkage being estimated from the data.
@@
 then a stable non-parametric approach becomes not only possible, but indeed quite straightforward.
-While not meeting Efron's desire for an entirely general nonparametric approach, we nonetheless believe 
+While not meeting Efron's desire for an entirely general nonparametric approach, we believe 
 that these settings include many cases of practical interest.
@@
 We note one caveat on the accuracy of estimated $g$: we tend to systematically
-overestimate the mass of $g$ near zero, due to our deliberate use of a penalty term (\ref{}) to avoid under-estimation of $\pi_0$.
-This is visually apparent in Figure \ref{fig:egcdf} (on careful inspection): the estimated cdf is generally below the true cdf just to the left of zero,
-and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect,
-and applying our methods without the penalty term removes this systematic effect (although at the cost of sometimes 
-under-estimating $\pi_0$).
+overestimate the mass of $g$ near zero, due to our deliberate use of a penalty term (\ref{eqn:penalty}) to avoid under-estimation of $\pi_0$.
+On careful inspection, this is visually apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
+and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf}b),
+and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
+under-estimating $\pi_0$ (Figure \ref{fig:egcdf}c).
 
@@
 The results are shown in Table \ref{tab:coverage}.  
-
-As noted above, the use of a penalty term to avoid underestimating $\pi_0$ means that the estimated $g$s tends to have, systematically, slightly too much mass near 0.
+As noted above, our use of a penalty term to avoid underestimating $\pi_0$ means that the estimated $g$s tends to have, systematically, slightly too much mass near 0.
 If the prior has too much mass near 0 then so will the posterior, so we should expect posterior intervals may also be ``over-shrunk" towards zero. 
@@
 
-\section*{Problems with removing the penalty term in the half-uniform case}
+\subsubsection*{Problems with removing the penalty term in the half-uniform case}
 
@@
 
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds. } \label{tab:coverage}
+\caption{Table of empirical coverage for nominal 95\% lower credible bounds (methods with penalty term). } \label{tab:coverage}
 \end{table}
@@
 
-\subsubsection*{Reverse asymptotics}
-
-Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
-
-Note: we'd like to make sure it behaves ok in the case where most observations are null.
-Perhaps simulate with 999 null and 1 alternative?
-Also maybe look at performance for small sample sizes - e.g. n=1?
-
-Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
-can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
-by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
-
-\subsection*{Improved conservative estimation of $\pi_0$}
-\subsection*{Local False Sign Rate}
-
-There are two reasons to use the lfsr instead of the lfdr: it is more generally meaningful (e.g. it applies
-whether or not zero effects exist),  and estimation of lfsr is more robust to modeling assumptions and estimation of $\pi_0$. 
-To illustrate this, we compared estimated and true values of both lfdr and lfsr for the simulated data (where the true
-values are computed by Bayes Theorem using the true value of $g$ FILL IN THESE DETAILS?).
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[height=320px,width=540px]{Rcode/figures/lfsdr_sim1sim2_blowup.png}
-\end{center}
-\caption{Figure showing lfsr is more robust than lfdr}
-\end{figure}
-
-- Illustration that estimated lfsr is more robust.
-
-- Illustration that adjusted estimate of lfsr is still more robust..
-
-- Compare estimated lfsr with true lfsr for difficult simulation case, to show nearly achieves Bayes Risk?
+%\subsubsection*{Reverse asymptotics}
+%
+%Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
+%
+%Note: we'd like to make sure it behaves ok in the case where most observations are null.
+%Perhaps simulate with 999 null and 1 alternative?
+%Also maybe look at performance for small sample sizes - e.g. n=1?
+%
+%Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
+%can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
+%by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
+%
 
@@
 
- To illustrate the first issue - differing measurement precision across units - consider a simplistic
- situation where half the measurements are quite precise, and the other half are really, really, poor. 
- Intuitively, the poor measurements tell us nothing, and any sane analysis should effectively ignore them.
-However, in a standard FDR-type analysis, these poor measurements add ``noise" and affect estimated FDRs.
-This is because the $p$ values from the poor measurements will be effectively uniformly distributed, and some will be significant at any given threshold.
-
-To give a concrete example, we performed a simple simulation in which half the observations have
-reasonably good precision (standard error $s_j = 1$) and half the observations have very poor precision ($s_j=10$). In both cases,
-we assume that half the effects are null ($\beta_j=0$)
-and the other half are normally distributed with standard deviation 1 ($\beta_j \sim N(0,1)$). 
-That is, we simulate $\beta_j$ from a mixture 
+ We turn now to the second important component of our work: allowing for varying
+ measurement precision across units. The key to this is our use of the likelihood
+ (\ref{eqn:lik}), which explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
+
+  To illustrate, we begin with a simple
+ simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
+ ($s_j=10$).  In both cases,
+we assume that half the effects are null
+and the other half are normally distributed with standard deviation 1:
 \begin{equation}
-p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1)
-\end{equation}
-where $\delta_0(\cdot)$ denotes a point mass at 0.
-
-Figure \ref{fig:goodpoor-hist} shows the distribution of the resulting $p$ values for the good-precision and poor-precision measurements, as
-well as the combined distribution. As expected, the poor-precision measurements lead to almost-uniform $p$ values, and when combined
-with the good-precision measurements they dilute the overall signal in the $p$ values (e.g. they reduce the density of $p$ values near 0).
-This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$, which work with the $p$ values and $Z$ scores respectively:
+p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1).
+\end{equation}
+In this setting, the poor-precision measurements tell us very little, and any sane analysis should effectively ignore them. 
+ However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor-hist}). This is because the poor measurements 
+ produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
+ which, when combined with the good-precision measurements, dilute the overall signal (e.g. they reduce the density of $p$ values near 0).
+ This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
 the estimated error rates ($q$-values, or lfdr values) for the good-precision observations increase when the low-precision observations are included in the analysis
-(Figure \ref{fig:goodpoor-scatter}). In contrast, the methods described here ($\ashr$) take account of the precision of each measurement, and so
-the estimated error rates for the good-precision observations are unaffected by including the low-precision observations in the analysis.
-(A key assumption of this simulation is that the distribution of the effects $\beta$ is the same for the good-precision and poor-precision measurements;
-we return to discuss this assumption in more detail in section \ref{xx})
+(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
+for the good-precision observations are unaffected by including the low-precision observations in the analysis.
 
@@
 
-\subsection{Comparison of different underlying component distributions}
-
-We simulated data under three scenarios for the underlying effect size distribution $g$:
-\begin{enumerate}
-\item[Scenario A:]
-A mixture of zero-centered normals,
-\begin{equation}
-g(\cdot) =  (2/3) N(\cdot; 0,1) + (1/3) N(\cdot; 0, 2^2)
-\end{equation}
-\item[Scenario B:]
-A symmetric ``flat-topped" distribution, designed to be difficult to capture by a mixture of zero-centered normals 
-\begin{equation}
-g(\cdot) = (1/7) [N(\cdot; -1.5, 0.5^2) + N(\cdot; -1, 0.5^2) + N(\cdot; -0.5, 0.5^2) + N(\cdot; 0, 0.5^2) + N(\cdot; 0.5, 0.5^2) + N(\cdot; 1, 0.5^2) + N(\cdot; 1.5, 0.5^2)]
-\end{equation}
-\item[Scenario C:]
-An asymmetric distribution, 
-\begin{equation}
-g(\cdot) = (1/4) N(\cdot; -2 ,2^2) + (1/4) N(\cdot; -1 ,1.5^2) + (1/3) N(\cdot; 0 ,1) + (1/6) N(\cdot; 1,1) 
-\end{equation}
-\end{enumerate}
-In each case we simulated $\bhat_j  = \beta_j + N(0,s^2_j=1)$.
-
-Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
-than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
-likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
-For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
-increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
-expect $L_{hu}$ to be significantly larger than the others.
-
-Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
-
-\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
+\subsection*{Adaptive Shrinkage}
+
+Example of how shrinkage is adaptive to signal and to standard error?
+
+DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
+To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
+Because the prior distribution is unimodal about zero,
+the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
+amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
+For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
+than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
+small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
+
+The posterior distribution $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ then reflects a trade-off
+between the information in the prior distribution, which is more-or-less concentrated near zero,
+
+There are two key points to appreciate here. First, by estimating the hyperparameters $\pi$ from the data, 
+the method adapts to the amount of signal in the data: if most $\beta_j$ are truly at or near zero then $\pi$ will reflect this by
+concentrating on zero and very small values of $\sigma_k$, and the posterior distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$
+will be tend also to be concentrated around zero (strong shrinkage). In contrast, if most $\beta_j$ are large then $\pi$ will be put most weight
+on components with large $\sigma_k$, and the posterior distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ will be approximately centered  on $\bhat_j$
+(weak shrinkage). Indeed, in the limit as $\pi$ puts all it's weight on very large $\sigma_k$ the prior becomes flatter and flatter
+and the posterior on $\beta_j$ becomes $N(\bhat_j,\shat_j)$. Thus, in this limit, the Bayesian credible intervals match standard confidence intervals.
+
+ and only very strong observations will have a small fdr; if most $\beta_j$ are large
+then this will also be reflected in the estimates of $\pi$ and even borderline signals might have a small fdr. 
+This idea is also used by existing FDR methods, but here, by modelling the effects, it extends to our estimates of the effects:
+
+The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ encapsulate uncertainty in the values for $\beta_j$, combining information across
+$j=1,\dots,J$. The combining of the information occurs through estimation of
+$\pi$, which involves all of the data.
+
+\subsection*{Local False Sign Rate}
+
+There are two reasons to use the lfsr instead of the lfdr: it is more generally meaningful (e.g. it applies
+whether or not zero effects exist),  and estimation of lfsr is more robust to modeling assumptions and estimation of $\pi_0$. 
+To illustrate this, we compared estimated and true values of both lfdr and lfsr for the simulated data (where the true
+values are computed by Bayes Theorem using the true value of $g$).
+
+\begin{figure}[!ht]
 \begin{center}
-\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
-\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
-\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
+\includegraphics[height=320px,width=540px]{Rcode/figures/lfsdr_sim1sim2_blowup.png}
 \end{center}
-\caption{}
+\caption{Figure showing lfsr is more robust than lfdr}
 \end{figure}
 
-- how does use of the VB approach to estimate pi affect RMSE?
-
-- Comparison between exchangeable effects vs Exchangeable standardized effects; Hedenberg data.
-
-- Possibly difference between CIs obtained by ash and usual CIs (e.g. ``of intervals where local fsr is <0.05, what proportion are the sign correct"?).
-
-- Asymmetric case (Half uniforms)
-
-Things to emphasise for paper:
-- the number of components is not critical; the unimodal constraint is enough. You can increase the number of components arbitrarily and the likelihood remains bounded.
-- the ability to incorporate item-specific measurement error
-- the ability to compare a model in which effects are proportional to error
-with model in which effects
-
-- bayesian coverage intervals (attempt to) give much stronger guarantess than standard confidence intervals. Eg among the 0.95 intervals excluding 0,
-less than 5% should actually be 0?
-
-Some advantages of the unimodal constraint:
--more statsitically efficient (if true); show with small sample sizes.
-- less sensitive to number of components? Can allow number to tend to infinity?
-- may be less sensitive to local optima?
-could demonstrate this by looking at convergence more carefully, and comparing
-results with random starting points.
-
-- it allows us to easily just vary sigma on a grid, and fit pi, which
-makes allowing different noise levels really easy!
-
-Comparisons with Efron, whose 1-group model is effectively the approach we have taken here.
-He says ``deconvolution
-calculations are notoriously tricky and I have not
-been able to produce a stable general algorithm." Here our claim is that the unimodal constraint is enough to produce that
-general stable algorithm.
-
-Notes from Adrian Raftery -
-
-Newcombe reference (1880s?)
+% Compare estimated lfsr with true lfsr for difficult simulation case, to show nearly achieves Bayes Risk?
+
+\subsection{Null behaviour}
+
+Results of simulations under the null.
+%\subsection{Comparison of different underlying component distributions}
+%
+%Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
+%than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
+%likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
+%For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
+%increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
+%expect $L_{hu}$ to be significantly larger than the others.
+%
+%Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
+%
+%\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
+%\begin{center}
+%\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
+%%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
+%\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
+%\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
+%\end{center}
+%\caption{}
+%\end{figure}
+
+%Notes from Adrian Raftery - 
+Newcombe (1886)
 Dean and Raftery
-Uniform is super-efficient
-
-\subsection{The number of components is not critical}
-OR - this section could be more generally about selecting between
-models (normal, uniforms etc) using log-likelihood.
-
-Because of the uni-modal constraint, the number of mixture components
-is generally not critical, at least provided it is ``sufficiently large". 
-Indeed, as the number of components tends to infinity, the likelihood
-is bounded above, and even for large numbers of components and small amounts of data the inferred underlying distributions tend not to be too crazy, showing few signs of ``overfitting".
-
-However, it is true that the normal distributions tend to yield smoother estimated underlying distributions. Similarly, using the posterior mean for $\pi$, rather
-than the maximum likelihood estimate, tends to lead to somewhat smoother fitted distributions. In addition, with small amounts of data the underlying
-distributions are inevitably not well determined by the data, and the fits may vary depending on the underlying assumptions made (e.g.
-number of components or distribution of the components); however, even then shrinkage-based estimates
-of the betas can be relatively robust.
+%Uniform is super-efficient
 
@@
 
-\subsection{Side notes on Multiple comparisons}
-
-Note on multiple comparisons: it isn't really a "problem" but an "opportunity". This viewpoint also espoused by \cite{greenland1991empirical}. It isn't the number of tests that is relevant (the false
-discovery rate at a given threshold does not depend on the number of tests). It is the *results* of the tests that are relevant.
-
-Performing multiple comparisons, or multiple tests, is often
-regarded as a "problem". However, here we regard it instead as an opportunity - an opportunity to combine (or "pool") information across tests or comparisons.
-
-Focussing on the number of tests performed can be
-seen as an approximation.
-
-\subsection{various notes...}
-
-Here performing inference means obtaining an (approximate) posterior distribution for each effect $\beta_j$,
-which can be used to estimate False Discovery Rates, and to produce both point and interval estimates for $\beta_j$.
- Although ``multiple comparisons" is usually presented as a ``problem", our emphasis here is that the multiple measurements
- actually provide an opportunity: the hierarchical model combines information across measurements to improve both accuracy and precision of estimates.
-
-Note that incorporating differential errors messes up the EM algorithm?
-
-we start by noting that the $p$ values in Figure \ref{} were actually generated under a scenario where
+\subsection{Discussion}
+
+Cite johnston and silverman; raykar and zhao, sinica.
+
+Choice of mixture components. Use of likelihood to compare different mixture components?
+
+ noting that the $p$ values in Figure \ref{} were actually generated under a scenario where
 {\it no tests were null}! Thus the qvalue procedure, which estimates the proportion of nulls to be xx\%, is acting very conservatively in this case.
@@
 
-The second issue is that directly modeling the p values, or z scores, say via non-parametric methods, 
+- it allows us to easily just vary sigma on a grid, and fit pi, which
+makes allowing different noise levels really easy! In contrast, incorporating differential errors messes up the EM algorithm if
+we try to estimate $\sigma_k$.
+
+Directly modeling the p values, or z scores, say via non-parametric methods, 
 can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
@@
 
-\subsection{Discussion}
+About benefits of generic approach.
+ In outline, our goal is to perform Bayesian inference for the true effects $\beta$ using information in the summary statistics $\bhat,\shat$.
+ That is we aim to compute the joint posterior distribution $p(\beta | \bhat,\shat)$.
+This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
+from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
+with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
+ By working with summary statistics $\bhat, \shat$, rather than more refined data, we aim to produce generic methods that 
+ can be applied whenever such summary data are available - just as \qvalue can be applied to any set of $p$ values for example.
+Any attempt to produce generic methods is likely to involve a compromise between functionality and generality; 
+we believe that by working with two numbers ($\bhat_j,\shat_j$) for each observation, rather than one ($p_j$ or $z_j$), 
+we can gain substantially in functionality (e.g.~ we can estimate effect sizes, as well as testing, and we can take better account of variations in
+measurement precision across units $j$) while losing only a little in generality.
 
@@
 
+ Note on multiple comparisons: it isn't really a "problem" but an "opportunity". This viewpoint also espoused by \cite{greenland1991empirical}. It isn't the number of tests that is relevant (the false
+discovery rate at a given threshold does not depend on the number of tests). It is the *results* of the tests that are relevant.
+Performing multiple comparisons, or multiple tests, is often
+regarded as a "problem". However, here we regard it instead as an opportunity - an opportunity to combine (or "pool") information across tests or comparisons.
+Focussing on the number of tests performed can be
+seen as an approximation.
+
+Here performing inference means obtaining an (approximate) posterior distribution for each effect $\beta_j$,
+which can be used to estimate False Discovery Rates, and to produce both point and interval estimates for $\beta_j$.
+ Although ``multiple comparisons" is usually presented as a ``problem", our emphasis here is that the multiple measurements
+ actually provide an opportunity: the hierarchical model combines information across measurements to improve both accuracy and precision of estimates.
+
  There are two issues with existing approaches to FDR estimation and control that I would like to address here. 
@@
 that the measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.
+
+--- NOte 
+In principle we might prefer a full Bayes approach that accounts for uncertainty in $\pi$ (or, more generally, in $g$); however,
+we believe that in most practical applications uncertainty in $g$ will not be the most important concern, and 
+compromise this principle for the simplicity of the EB approach.
+
+\section*{Implementation details}
+
+\subsubsection*{Choice of grid for $\sigma_k, a_k$}
+
+\def\sigmamax{\sigma_{\text{max}}}
+\def\sigmamin{\sigma_{\text{min}}}
+When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
+in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
+as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
+based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
+is negative we set $\sigmamax = 8\sigmamin$.)
+
+When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.
+
+Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
+a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.
+
+\subsubsection*{Penalty term on $\pi$}
+
+In practice data do not distinguish between effects that are ``very small" and ``exactly zero".
+Thus, if a mixture component $f_k$ has most mass very near zero then
+this introduces non-identifiability between $\pi_k$ and $\pi_0$. In practice this is a major problem
+only if one focuses on false discovery rates rather than false sign rates: the \lfdr is sensitive to $\pi_0$ whereas
+the \lfsr is not. However, to make \lfdr (and \lfsr) estimates from our method ``conservative" we add a penalty 
+term $h(\pi;\lambda)$ to the likelihood to encourage over-estimation of $\pi_0$:
+\begin{equation}
+h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
+\end{equation}
+where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
+consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0}).
+
+Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
+we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
+
+\subsubsection*{Initialization}
+
+We initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. 
+Our rationale to initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
+but weak signal in the data cannot quickly draw the algorithm towards the null.
+
+\subsubsection*{Likelihood for $\pi$ and EM algorithm}
+
+To formally derive the likelihood we restate our modelling assumptions more formally.
+We provide details first for the normal likelihood and then briefly describe modifications for the $t$ likelihood.
+
+We treat the standard errors as $\shat=(\shat_1,\dots,\shat_J)$ as given, and perform all inference conditional on $\shat$.
+We assume that given $\shat$ and the hyper-parameters $\pi$, the $(\beta_j,\bhat_j)$ pairs are independent 
+\begin{align} \label{eqn:b1}
+p(\beta_j \, | \, \shat,\pi) & = \sum_{k=0}^K \pi_k f_k(\beta_j) \\ \label{eqn:b2}
+p(\bhat_j \, | \, \beta_j, \shat, \pi)  & = N(\bhat_j; \beta_j, \shat^2_j),
+\end{align}
+where for notational convenience we have replaced the point mass $\delta_0$ with $f_0$.
+
+Multiplying (\ref{eqn:b1}) and (\ref{eqn:b2}) together gives $p(\bhat_j, \beta_j | \shat,\pi)$, which integrating over $\beta_j$ yields
+\begin{equation}
+p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
+\end{equation}
+where 
+\begin{equation}
+\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
+\end{equation}
+denotes the convolution of $f_k$ with the normal density.
+These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
+Specifically, 
+\begin{equation}
+\tilde{f}_k(\bhat_j)  = 
+\begin{cases}
+N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
+\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
+\end{cases}
+\end{equation}
+where $\Psi$ denotes the distribution function of the standard normal distribution.
+If we use a $t_\nu$ likelihood instead of a normal likelihood then the convolution is tricky for $f_k$ normal and
+we have not implemented it; for $f_k$ uniform the result (\ref{eqn:uconv}) holds but with the standard normal distribution function replaced
+with the $t_\nu$ distribution function.
+
+With this in place, the likelihood for $\pi$ is obtained by multiplying across $j$:
+\begin{equation}
+L(\pi) = \prod_j \sum_k \pi_k w_{kj}
+\end{equation}
+where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. 
+We add to this likelihood a penalty term $h(\pi;\lambda)$ above,
+and use an EM algorithm to maximize $L(\pi)+h(\pi;\lambda)$. The one-step updates for this EM algorithm are:
+\begin{align} \label{eqn:w}
+w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
+n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{E Step}
+\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{M step}
+\end{align}
+Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
+and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
+probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$.
+Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
+We used the elegant R package {\tt SQUAREM} \cite{squarem}) to accelerate convergence of this EM algorithm.
+
+[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
+
+\subsubsection*{Conditional distributions}
+
+Given $\hat\pi$, we compute the conditional distributions 
+\begin{equation}
+p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
+\end{equation} 
+Each posterior is a mixture on $K+1$ components:
+\begin{equation}
+p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
+\end{equation}
+where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
+and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
+$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
+All these posterior distributions are easily available.
+For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
+If $f_k$ is normal and $L$ is normal, then this is a normal distribution.
 

### 2015-06-23 talks/twentymin_seminar.Rmd
update

--- 2014-05-09 talks/BOG2014.Rmd
@@
-% False Discovery Rates, A New Deal
-% Matthew Stephens
-% 2014/5/8
+---
+title: "False Discovery Rates, A New Deal"
+author: Matthew Stephens
+date: June 12, 2015
+output: ioslides_presentation
+---
 
@@
 ```{r setup, include=FALSE}
-# set global chunk options
-opts_chunk$set(cache=TRUE,autodep=TRUE,warning=FALSE,dev='pdf')
+  #set global chunk options
+  opts_chunk$set(cache=FALSE,autodep=TRUE,warning=FALSE)
   dep_auto()
@@
 
-# What you can expect from this talk
+## How can we make more progress together?
 
-- No experimental data!
+- Abstract our problems
+- Develop benchmarks
+- Build generic tools
 
-- No fancy heatmaps!
+## The power of generic tools
 
-- 100\% Money-back guarantee to half the False Discovery Rate (FDR) in your data!
+- Linear regression
+- k-means clustering
+- Principal Components Analysis
+- qvalue for FDR
 
-# The Canonical Genomics Experiment
+Goal: generic tools for shrinkage estimation
+
+## The Canonical Genomics Experiment
 
@@
 - Apply `qvalue` 
-to identify findings ``significant" at a given FDR.
+to estimate FDR at any given threshold.
 
-- ...?
+$$\text{FDR}(\gamma)=\Pr(\beta_j = 0 |  p_j< \gamma).$$
+
+- ...
 
@@
 
-# Problem 1: The Zero Assumption (ZA)
+## The outsize impact of generic procedures
+
+- The original paper introducing FDR (Benjamini and Hochberg, 1995) has
+been cited 27,331 times (June 2015) according to Google Scholar.
+
+- That is three times a day for the last 20 years!
+
+- More than ten times a day for the last year!
+
+## Problem 1: The Zero Assumption (ZA)
 
@@
 
-# Problems: Summary
-
-Standard tools are highly conservative.
+## Problems: Summary
 
@@
 
-# FDR via Empirical Bayes
+(Both tend to lead to conservative behaviour: overestimating FDR)
+
+## Can we do better?
+
+Two Ideas:
+
+- Use two numbers ($\hat\beta_j,s_j$) instead of one number ($p_j$ or $z_j$).
+
+- Introduce "modelling assumption": distribution of $\beta_j$ is unimodal about 0.
+
+## FDR via Empirical Bayes
 
@@
 
+- Various semi-parametric approaches taken to estimating $f_1$. For example,
+Efron uses Poisson regression; Muralidharan uses mixture of normal distributions.
+
 - Once $f_1$ and $\pi_0$ estimated, FDR calculations are straightforward.
 
-# FDR: A New Deal
+## FDR: A New Deal
 
@@
 
-- *Incorporate precision* of each observation $\hat\beta$ into the likelihood.
-Specifically, approximate likelihood for $\beta_j$ by assuming 
-$$\hat\beta_j \sim N(\beta_j, s_j)$$
+- *Incorporate precision* of each observation $\hat\beta$ into the likelihood:
+$$\hat\beta_j | s_j \sim N(\beta_j, s_j)$$
 
-# FDR - A New Deal
+(cf Bayes inference from summary statistics; V.Johnson, J.Wakefield)
+
+## FDR - A New Deal
 
@@
 
-# Recall Problem 1: distribution of alternative Z values multimodal
+## Adaptive Shrinkage
+
+- This approach actually provides a full posterior distribution for each $\beta_j$.
+
+- So easy to obtain point estimates and credible intervals.
+
+- Because $g(\beta)$ is unimodal, the point estimates (and CIs) will tend to be ``shrunk" towards the overall mean (0).
+
+- Because $g(\beta)$ is estimated from the data, the amount
+of shrinkage is adaptive to signal in the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each observation.
+
+- So we call the approach ``Adaptive Shrinkage" (ASH).
+
+## Recall Problem 1: distribution of alternative Z values multimodal
 ```{r, echo=FALSE,fig.height=4,fig.cap=""}
@@
 
-# Example: FDR estimation
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,hh.q$pi0,type=4,title="qvalue")                 
-```
-
-# Example: FDR estimation
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-  plot_FDReg_hist(hh.pval,get_pi0(hh.ash),type=4,title="ash")                 
-```
+## Recall Problem 2: poor measurements dilute good measurments
 
@@
 ```
-
-# Recall Problem 2: poor measurements dilute good measurments
 
@@
 
-# A new problem: an embarrassment of riches
+## Some more extensive simulations
 
-- If the null is mostly false, the new approach can  provide unsettling results
+![Image](../paper/figures/scenarios_density.pdf)
+
+$$\beta_j \sim \pi_0 \delta_0(.) + (1-\pi_0) g(.)$$
+
+Goals: Estimate $\pi_0$? Estimate $g$?
+
+## Estimation of $\pi_0$
+
+![Image](../paper/figures/pi0_est.pdf)
+
+All methods are, generally, "conservative" (deliberately).
+
+## Estimation of $g$
+
+- Estimating $g$ is a “deconvolution problem”.
+- Notoriously difficult in general. 
+- Efron emphasises the difficulties of implementing a stable general algorithm:
+
+<div class="red">
+“the effort foundered on practical difficulties involving the perils of deconvolution...
+Maybe I am trying to be overly nonparametric...
+but it is hard to imagine a generally satisfactory parametric formulation...”
+</div>
+
+## Unimodal assumption stabilizes estimation of $g$
+
+![Image](../paper/figures/egcdf.pdf)
+
+## A new problem: an embarrassment of riches
+
+- If the null is mostly false, this approach can  provide unsettling results
 
@@
 
-# Perhaps we didn't really understand the question?
+## Perhaps we didn't really understand the question?
 
-- Problem arises only if we insist on asking question ``is $\beta_j=0$?"
+- Problem arises only if we insist on asking question "is $\beta_j=0$?"
 
@@
 
-- Suggests a change of focus: assume *none* of the $\beta_j$ are zero ("one group approach"), and ask for which $\beta_j$ are we confident about the sign (Gelman et al, 2012).
+- Suggests a change of focus: ask for which $\beta_j$ are we confident about the sign (cf Gelman et al, 2012; Tukey 1961).
 
-# The False Sign Rate
+## Generic procedure: summary
 
-- Suggestion: replace FDR with local false sign rate (lfsr), the probability that if we say an effect is positive (negative), it is not.
+- Inputs: $(\hat{\beta}_1,s_1), \dots, (\hat{\beta}_p,s_p)$
+- Assumptions:
 
-- Example: suppose we estimate that $\Pr(\beta_j<0)=0.975$ and $\Pr(\beta_j>0)=0.025$. Then we report $\beta_j$ as a ``(negative) discovery", and estimate its fsr as 0.025.
+$$\hat{\beta}_j | \beta_j, s_j \sim N(\beta_j,s_j)$$
+$$\beta_j | s_j \sim g(), \text{unimodal}.$$
 
-# Even with many signals, large $p$ values have high lfsr
+- Output: estimate of $g$; "shrunken" posterior distribution for $\beta_j$.
 
-```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-  res=data.frame(p=hh.pval, lfsr=hh.ash$lfsr, qvalue = hh.ash$qvalue)
-  res.melt = melt(res, id.vars=c("p"),variable.name="Measure")
-cbbPalette <- c("#000000", "#D55E00", "#CC79A7")
- labels = c('qvalue','lfsr')
-  breaks = c("qvalue","lfsr")
+## Generic procedure: embellishments
 
-  pp= ggplot(data=res.melt,aes(p,value,color=Measure)) +geom_point(shape=1) +
-      geom_abline(colour = "black") +
-        xlab("p-value") +
-        ylab("lfsr/qvalue")
+- Inputs: $(\hat{\beta}_1,s_1), \dots, (\hat{\beta}_p,s_p)$
+- Assumptions:
 
-  print(pp +scale_y_continuous(limits=c(0,0.1)) +
-          scale_x_continuous(limits=c(0,0.1))  +
-           scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
-          coord_equal(ratio=1))
+$$\hat{\beta}_j | \beta_j, s_j \sim t(\beta_j,s_j)$$
+$$\beta_j | s_j \sim s_j^\alpha g(), \text{unimodal}.$$
 
-```
+- Output: estimate of $g$; "shrunken" posterior distribution for $\beta_j$.
 
-# Summary
+## Other Applications
 
-- ASH provides a generic approach to shrinkage estimation, as well as
-false discovery (sign) rates.
+- Widely applicable: requiring only an estimated
+effect size and standard error for each object.
 
-- Unimodal assumption for effects reduces conservatism
+- E.g. Currently applying it to wavelet shrinkage applications.
 
-- But by using two numbers ($\hat\beta,s$) instead of one ($p$ values or $z$ scores) precision of different measurement is better accounted for.
+## Next steps?
 
-- In high-signal contexts, False Sign Rate is preferable to False Discovery Rate.
+- Incorporate shrinkage of variances and not just means. (e.g. ``moderated $t$ test", Smyth et al).
 
-# Reproducible Research?
+- Allow $g(\cdot;\pi)$ to depend on covariates $X$.
 
-- Principle: when publishing results of computational procedures, we should
-publish the code that produced the results.
-- "publishing figures or results without the complete software environment could
-be compared to a mathematician publishing an announcement of a mathematical theorem without giving the proof" (Buckheit and Donohoe)
-- “an article about a computational result is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result.” [Claerbout]
+- Allow for correlations in the measured $\hat\beta_j$.
 
-# This talk is reproducible!
+- Multivariate version: we observe a matrix of $\hat\beta_j$ and $s_j$ values.
 
-- `http://www.github.com/stephens999/ash`
-
-# Why is reproducibility important?
-
-- Not only because people are forgetful, error-prone, or dishonest!
-- Reproducing work is also the first step to extending it.
-- Helps communications among researchers (eg student + advisor).
-
-# Thanks
+## Thanks
 
 - to the developers of **R**, **knitr**, **Rstudio** and **Pandoc**.
+Also the very cool **SQUAREM** package.
 
@@
 
-- Including Ester Pantaleo, Scott Powers, Mengyin Lu, Sen Tian, Wei Wang, Zhengrong Xing.
+- Including  Chaoxing Dai, Mengyin Lu, Ester Pantaleo, Scott Powers, Sen Tian, Wei Wang, Zhengrong Xing.
 
-- NHGRI for funding.
+- NHGRI, Moore Foundation for funding.
 
-- `ashr` package: `http://www.github.com/stephens999/ash`
-
-# Pandoc Command used
-
-`pandoc -s -S -i --template=my.beamer -t beamer -V theme:CambridgeUS -V colortheme:beaver  slides.md -o slides.pdf`
-
-(alternative to produce html slides; but figures would need reworking)
-`pandoc -s -S -i -t dzslides --mathjax slides.md -o slides.html`
-
-Here is my session info:
-
-```{r session-info}
-print(sessionInfo(), locale=FALSE)
-```
-
-# Issue: you can't actually estimate $\pi_0$!
-
-- Data cannot distinguish between $\beta_j = 0$ and $\beta_j$ "very
-small"
-
-- As a result $\pi_0$ cannot be estimated: the data can never rule out $\pi_0=0$.
-
-- But the unimodal constraint bounds how big $\pi_0$ can be.
-
-- Use penalized likelihood to make $\pi_0$ ``as big as possible",
-subject to the unimodal constraint.
-
-# Approach remains conservative (if unimodal assumption holds)
-
-```{r,echo=FALSE,include=FALSE}
-load("../paper/Rcode/sim1.RData") #load simulation results
-source("../paper/Rcode/plot_pi0.R")
-```
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-plot_pi0_onlyqval(list(simres1a,simres2))
-```
-
-# Next steps?
-
-- Incorporate $t$ likelihood as well as normal.
-
-- Incorporate shrinkage of variances and not just means.
-
-- Extend to allow $g(\cdot;\pi)$ to depend on covariates $X$.
-
-- Extend to allow for correlations in the measured $\hat\beta_j$.
-
-# Is this an important problem?
-
-- The original paper introducing FDR (Benjamini and Hochberg, 1995) has
-been cited 21,787 times according to Google Scholar.
-
-- That is three times a day for the last 19 years!
-
-# FDR and q values
-
-Although precise definitions vary, roughly
-
-- The FDR at a threshold $P$ is 
-$$\text{FDR}(P)=\Pr(\beta_j = 0 |  p_j<P).$$
-
-- The q value for observation $j$ is $q_j=\text{FDR}(p_j)$.
-
-# Estimation and Shrinkage
-
-- Besides allowing one to estimate fdr and fsr, 
-this approach also provides a posterior distribution for each $\beta_j$.
-
-- So for example we can easily compute fdrs for discoveries other than ``non-zero" (eg compute $\Pr(\beta_j > 2 | \hat\beta_j)$).
-
-- And use it to obtain point estimates and credible intervals for each $\beta_j$, taking account of information from all the other $\beta_j$.
-
-- Because $f(\beta)$ is unimodal, the point estimates will tend to be ``shrunk" towards the overall mean (0).
-
-- Because $f(\beta)$ is estimated from the data, the amount
-of shrinkage is adaptive to the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each gene.
-
-- So we call the approach ``Adaptive Shrinkage" (ASH).
+- `ashr` package: `http://www.github.com/stephens999/ashr`

### 2015-06-23 talks/wavelet_vignette.rmd
update

--- 2015-05-24 talks/ilike2015.rmd
@@
 ---
-title: "Dynamic Statistical Comparisons, and other adventures in getting organized"
+title: "Wavelet Vignette"
 author: Matthew Stephens
@@
 
-```{r, include=FALSE}
-require("qvalue")
-require("ashr")
-require("ggplot2")
-require("knitr")
-```
-
-```{r setup, include=FALSE}
-  #set global chunk options
-  opts_chunk$set(cache=TRUE,autodep=TRUE,warning=FALSE)
-  dep_auto()
-```
-
-## Getting Organized!
-
-- Over ~10 years of working with graduate students + postdocs,
-I've noticed something.
-- Organized researchers get more done (and better!).
-- Many of them are more organized than I am.
-- Thought: I should get organized; I should help others get organized.
-
-## So what can you do?
-
-- Buy a notebook; bring it to meetings; make notes! 
-- Come to meetings with a written agenda. 
-- While doing research, record what you did and what the outcome was.
-- Use version control ([git](http://git-scm.com)) and internet repositories ([bitbucket](http://www.bitbucket.org), [github](http://www.github.com)) to organize notes, code, etc.
-- Use *knitr* to help make your research reproducible.
-- Talk about the tools you find useful!
-
-## What are these repository things?
-
-- A repository: a central place in which an aggregation of data is kept and maintained in an organized way (searcharticle.com)
-- Great for sharing material across multiple people (eg student and advisor!)
-- An amateur example: [http://github.com/stephens999/ash](http://github.com/stephens999/ash)
-
-## What is knitr?
-
-- An R package
-- A tool for literate programming
-- Text, and R code are interleaved
-- When you compile the document, the code is run, and output inserted into the text.
-- Great for writing reports, and keeping a track of what you did and what the result was!
-- This talk was written with knitr (with RStudio)!
-
-## What is knitr?
-
- ![Image](screenshot.png)
-
-## What is Reproducible Research?
-
-- Principle: when publishing results of computational procedures, we should
-publish the code that produced the results.
-- "publishing figures or results without the complete software environment could
-be compared to a mathematician publishing an announcement of a mathematical theorem without giving the proof" (Buckheit and Donohoe)
-- “an article about a computational result is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result.” [Claerbout]
-
-## Why is reproducibility important?
-
-- Not only because people are forgetful, error-prone, or dishonest!
-- Reproducing work is also the first step to extending it.
-- Helps communications among researchers (eg student + advisor).
-- If you do not publish code implementing your methods, your methods will likely go unused.
-
-## More on git, github, knitr, reproducibility
-
-- Google "The git book", to get started on git.
-- Google "Karl Broman github tutorial" for statistics-oriented intro to github.
-- Google "donohoe buckheit" for "Wavelab and reproducible research"
-
-## More ridiculously-personal tips
-
-- "Don't try to show them how clever you are" (BD Ripley)
-- Minimise email for research communications 
-- Chunk time
-- Relax! (Exercise, Vacation)
-- Admin on Fridays
-- Learn some programming principles
-- Benchmark early and often
-
-## Minimise email
-
-![Image](slack.screenshot.png)
-
-## Benchmarking:
-
-By which I mean, any quantitative assessment and comparisons of performance (accuracy; speed).
-
-## Benchmarking:
-
-Benefits include:
-
-- Communication
-- Tweaking/honing
-- Basic quality assurance (cf "testing")
-
-## Benchmarking:
-
-Even basic benchmarks have benefits (...Done is better than perfect)
-
-## Dynamic Statistical Comparisons
-
-- Overall Goal: Improve how we, in the statistics community, make comparisons among methods.
-
-- Reduce work and redundancy; make comparisons more transparent
-
-- How? Centralize reproducible statistical comparisons in easily-extensible public repositories.
-
-## Inspiration: Public Benchmarking
-
-- Protein folding (CASP)
-
-- Image analysis (Pascal)
-
-- The Neflix Prize
-
-- Genetics Analysis Workshop
-
-- RECOMB/DREAM challenge
-
-- Sage Bionetworks
-
-## Dynamic Statistical Comparisons: advantages
-
-- Authors can run their own methods. 
-- Ease of adding new data can quickly uncover over-tuning to specific datasets.
-- Work is divided, and not duplicated.
-- Promotes collaboration through competition. 
-- Ensures results are reproducible, and extendable.
-
-- [http://stephens999.github.io/blog/](http://stephens999.github.io/blog/) for more on DSC
-
-## DSC Requirements
-
-- Well-defined input format for methods
-- Well-defined output format
-- Scoring function (output, input, meta-data)
-- Ways to produce input, meta-data
-- Methods to turn input -> output
-
-## DSCR - an R package
-
-- Define Scenarios (produce input, meta-data)
-- Define Methods (turn input -> output)
-- Define Scores
-
-A simple function call then runs all methods on all scenarios, outputs scores
-
-[Simple Vignette](http://htmlpreview.github.com/?https://github.com/stephens999/dscr/blob/master/vignettes/one_sample_location.html)
-
-## DSC examples we're working on
-
-- Covariance Estimation
-- Large-scale linear regression
-- 1-D Smoothing with Wavelets [after Antoniadis et al, 2001]
-
-## Wavelet Smoothing Example: The "Blocks" Function
+## Example: The "Blocks"" Function
 
@@
 
-## Wavelet Shrinkage
-
-- WCs of true mean ($$\beta$$) are "sparse" (or at least, have spiky density about 0)
-
-- WCs for observed data ($$\betahat$$) are noisy observations of true WCs.
-
-- If we shrink the $$\betahat$$ toward 0, they will give a better estimate for $$\beta$$
-
-- And inverting the wavelet transform will give better estimate for $\mu$
-
 ## Shrink + Backtransform = Smooth
-
 ```{r, echo=FALSE}
@@
 ```
-
-## How to Shrink?
-
-We've been developing generic EB approaches to shrinkage estimation:
-
-- Assume $\beta_j \sim g()$ where $g$ has some concentration about 0 ("sparse").
-
-- Assume we have corresponding estimates $\hat\beta_j$, with standard error $s_j$.
-
-- 
-## How to Shrink?
-
-- *Incorporate precision* of each observation $\hat\beta$ by assuming
-$$\hat\beta_j | \beta_j \sim N(\beta_j, s_j^2)$$
-
-- Constrain $g$ to be *unimodal* about 0;
-
-- Shrink via two-stage procedure: first estimate $g$ "non-parametrically" from data; then
-compute $p(\beta_j | \hat{g}, \hat\beta_j,s_j)$.
-
-## Details
-
-- A convenient way to model $g$: mixture of 0-centered
-normal distributions: 
-$$g(\beta; \pi) = \sum_{k=1}^K \pi_k N(\beta; 0, \sigma^2_k)$$
-
-- Estimating $g$ comes down to estimating $\pi$. Joint estimation of $\pi_0,\pi$ easy by maximum likelihood (EM algorithm).
-
-- By allowing $K$ large, and $\sigma_k$ to span a dense grid of values,
-we get a flexible unimodal symmetric distribution.
-
-- Can approximate, arbitrarily closely, any scale mixture of normals.
-Includes almost all priors used for sparse regression problems (spike-and-slab, double exponential/Laplace/Bayesian Lasso, horseshoe).
-
-## Illustration: $g$ a mixture of 0-centered normals
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-x=seq(-4,4,length=100)
-plot(x, dnorm(x,0,1),type="l",ylim=c(0,2),ylab="density")
-lines(x, dnorm(x,0,0.1))
-lines(x, dnorm(x,0,0.5))
-```
-
-## Illustration: $g$ a mixture of 0-centered normals
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-x=seq(-4,4,length=100)
-plot(x, 0.5*dnorm(x,0,1)+0.5*dnorm(x,0,0.1),type="l",ylim=c(0,2),ylab="density")
-```
-
-## Illustration: $g$ a mixture of 0-anchored uniforms
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-x=seq(-4,4,length=100)
-plot(x, dunif(x,0,1),type="s",ylim=c(0,5),ylab="density")
-lines(x, dunif(x,0,0.2),type="s")
-lines(x, dunif(x,0,0.5),type="s")
-lines(x, dunif(x,-0.3,0),type="s",col=2)
-lines(x, dunif(x,-0.4,0),type="s",col=2)
-lines(x, dunif(x,-2,0),type="s",col=2)
-```
-
-## Illustration: $g$ a mixture of 0-anchored uniforms
-
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-x=seq(-4,4,length=100)
-plot(x, 0.1*dunif(x,0,3)+ 0.3*dunif(x,0,0.2) + 0.2*dunif(x,0,0.5) + 0.1* dunif(x,-0.3,0) + 0.1*dunif(x,-0.3,0)+0.2*dunif(x,-0.4,0),type="s",ylim=c(0,2),ylab="density")
-```
-
-## Adaptive Shrinkage
-
-- Approach provides posterior distribution for each $\beta_j$.
-
-- For wavelet shrinkage we just use posterior mean.
-
-- Because $g(\beta)$ is unimodal, the point estimates (and CIs) will tend to be "shrunk" towards the overall mean (0).
-
-- Because $g(\beta)$ is estimated from the data, the amount
-of shrinkage is adaptive to signal in the data. And because of the role of $s_j$, the amount of shrinkage adapts to the information on each observation.
-
-- So we call the approach "Adaptive Shrinkage" (ash).
-
-## Smoothing with Adaptive Shrinkage (smash)
-
-```{r, echo=FALSE, results='hide', message=FALSE}
-load("../../dscr-smash/res.RData")
-library("dplyr")
-library("ggplot2")
-res.1.v1 = res %>% filter(grepl(".1.v1",scenario)) %>% filter(method %in% c("smash.haar","smash.homo.s8","ebayesthresh","sure.homo.s8"))
-PLOTNAMES=c("angles","blip","blocks","bump","corner","doppler","spike")
-res.1.v1$scenario=factor(res.1.v1$scenario,levels=paste0(c("ang","blip","blk","bump","cor","dop","sp"),".1.v1"))
-levels(res.1.v1$scenario)=PLOTNAMES
-ggplot(res.1.v1,aes(x=method,y=mise,fill=method)) + geom_violin() + facet_grid(.~scenario) + theme(axis.text.x=element_blank())
-```
-
-## Homoskedastic errors
-
-```{r, echo=FALSE, results='hide', message=FALSE}
-res.1.v1 = res %>% filter(grepl(".1.v1",scenario)) %>% filter(method %in% c("smash.haar","smash.homo.s8","ebayesthresh"))
-
-PLOTNAMES=c("angles","blip","blocks","bump","corner","doppler","spike")
-res.1.v1$scenario=factor(res.1.v1$scenario,levels=paste0(c("ang","blip","blk","bump","cor","dop","sp"),".1.v1"))
-levels(res.1.v1$scenario)=PLOTNAMES
-
-ggplot(res.1.v1,aes(x=method,y=mise,fill=method)) + geom_violin() + facet_grid(.~scenario) +theme(axis.text.x=element_blank())
-```
-
-## Heteroskedastic errors
-
-```{r, echo=FALSE, results='hide', message=FALSE}
-res.1.v4 = res %>% filter(grepl(".1.v4",scenario)) %>% filter(method %in% c("smash.haar","smash.homo.s8","ebayesthresh"))
-res.1.v4$scenario=factor(res.1.v4$scenario,levels=paste0(c("ang","blip","blk","bump","cor","dop","sp"),".1.v4"))
-levels(res.1.v4$scenario)=PLOTNAMES
-ggplot(res.1.v4,aes(x=method,y=mise,fill=method)) + geom_violin() + facet_grid(.~scenario) + theme(axis.text.x=element_blank())
-```
-
-## Question
-
-- Does anyone have applications where allowing for heteroskedastic errors might be useful?
-
-## Summary
-
-- We're looking for conspirators interested in organizing benchmarking!
-
-## Thanks
-
-- Zhengrong Xing
-
-- to the developers of **R**, **knitr**, **Rstudio** and **Pandoc**.
-Also the very cool **SQUAREM** and **wavethresh** packages.
-
-- to the several postdoctoral researchers and students
-who have worked with me on related topics.
-
-- Including Chaoxing Dai, Mengyin Lu, Ester Pantaleo, Scott Powers, Sen Tian, Wei Wang, .
-
-- NHGRI, Moore Foundation for funding.
-
-- `ashr` package: `http://www.github.com/stephens999/ashr`
-
-- `dscr` package: `http://www.github.com/stephens999/dscr`

### 2015-09-03 examples/adaptive_shrinkage.rmd
minor updates

@@
 Second, the amount of shrinkage of each observation depends on its standard error: the smaller the standard error, the more informative
-the data, and so the less shrinkage that occurs. From an Empirical Bayesian point of view both of these 
+the data, and so the less shrinkage that occurs. From an Empirical Bayesian perspective both of these 
 points are entirely natural: the posterior depends on both the prior
@@
 We simulate from two scenarios, the first with effects more peaked about zero, the second with effects less peaked.
-```{r}
+```{r initialize}
 set.seed(100)
 library(ashr)
-beta1 = c(rnorm(500,0,0.01),rnorm(500,0,0.5))
-beta2 = rnorm(1000,0,1)
-s = 1/rgamma(1000,10,10)
-e = rnorm(1000,0,s)
-beta1hat = beta1 + e
-beta2hat = beta2 + e
+
+#load the simulation function, (rnormmix_datamaker)
+source("../dsc-shrink/datamakers/datamaker.R")
+
+NSAMP=1000
+s = 1/rgamma(NSAMP,5,5)
+sim.bignormal = rnormmix_datamaker(args=list(g=normalmix(c(1),c(0),c(4)),
+                        min_pi0=0,
+                        max_pi0=0,
+                        nsamp=NSAMP,
+                        betahatsd=s
+                      ))
+
+sim.spiky = rnormmix_datamaker(args=list(g=normalmix(c(.4,.2,.2,.2),c(0,0,0,0),c(.25,.5,1,2)),
+                        min_pi0=0,
+                        max_pi0=0,
+                        nsamp=NSAMP,
+                        betahatsd=s
+                      ))
+
 ```
@@
 Now run ash on both:
-```{r}
-beta1.ash = ash(beta1hat, s)
-beta2.ash = ash(beta2hat, s)
+```{r run-ash}
+beta.bignormal.ash = ash(sim.bignormal$input$betahat, s)
+beta.spiky.ash = ash(sim.spiky$input$betahat, s)
+```
+
+Here we plot the shrunken estimates against the observed values, colored
+according to the (square root of) precision: precise esitmates being colored red, and less precise estimates being blue. Two key features of the plots
+illustrate the ideas of adaptive shrinkage: i) the estimates under the spiky scenario are shrunk more strongly, illustrating that shrinkage adapts to the underlying distribution of beta; ii) in both cases, estimates with large standard error (blue) are shrunk more than estimates with small standard error (red) illustrating that shrinkage adapts to measurement precision.
+
+```{r plotting}
+df = data.frame(betahat = c(sim.spiky$input$betahat,sim.bignormal$input$betahat),
+                beta_est = c(beta.spiky.ash$PosteriorMean, beta.bignormal.ash$PosteriorMean),
+                lfsr = c(beta.spiky.ash$lfsr, beta.bignormal.ash$lfsr),
+                s=c(s,s),
+                scenario=c(rep("spiky",NSAMP),rep("big-normal",NSAMP)))
+df$signif = df$lfsr<0.05
+library(ggplot2)
+library(scales)
+ggplot(df,aes(x=betahat,y=beta_est,color=1/s)) + 
+  xlab("Observed beta-hat") +
+  ylab("Shrunken beta estimate") + geom_point() +  facet_grid(.~scenario) + geom_abline(intercept=0,slope=1) +
+scale_colour_gradient2(midpoint = median(1/s),low="blue", mid="white",high="red")  +coord_fixed(ratio=1)
+ggsave("ash.pdf")
 ```
@@
 ```{r}
-df = data.frame(betahat = c(beta1hat,beta2hat), beta_est = c(beta1.ash$PosteriorMean, beta2.ash$PosteriorMean),s=c(s,s),scenario=c(rep("peaked",1000),rep("less peaked",1000)))
-library(ggplot2)
-ggplot(df,aes(x=betahat,y=beta_est)) + geom_point(col=s) + facet_grid(.~scenario) 
-#need to add equal scales, and color gradient
+  ggplot(df,aes(x=pnorm(-abs(betahat/s)),y=lfsr,color=log(s))) + geom_point() +  facet_grid(.~scenario) +xlim(c(0, 0.025)) + xlab("p value") + ylab("lfsr") +
+  scale_colour_gradient2(midpoint = 0,low="blue", mid="white",high="red")  
+```
 
+```{r}
+hist(s,main="histogram of standard errors")
+summary(s)
 ```
+
+This is a "volcano plot" showing effect size against p value.
+The blue points are "significant" in that they have lfsr<0.05.
+In this case the significance by lfsr is not quite the same as cutting off at a given $p$ value (you can see that the decision boundary is not quite the same as drawing a horizontal line), but also not that different, presumably because the standard errors, although varying across observations, do not vary greatly.
+```{r volcano}
+  ggplot(df,aes(x=betahat,y=-log10(2*pnorm(-abs(betahat/s))),col=signif)) + geom_point(alpha=1,size=1.75) +  facet_grid(.~scenario) + theme(legend.position = "none") +
+  xlim(c(-10, 10)) + ylim(c(0, 15)) +
+  xlab("Effect (beta)") + ylab("-log10 p-value")
+```

### 2015-09-03 paper/main.tex
minor updates

@@
 datta2005empirical,muralidharan2010empirical} for example.
-So what is the ``New Deal" here? We introduce two simple new ideas that can substantially improve
-inference. The first idea is to {\it assume that the distribution of the non-zero effects is unimodal}, and to perform
-non-parametric inference under this assumption. This idea has the twin benefits of simplifying and improving inference,
-at least when the unimodal assumption is correct.  And even when the assumption is incorrect, the resulting errors in inference are not too egregious. 
-The second idea is to use two numbers (effect sizes, and their standard errors), rather than just one ($p$ values, or $z$ scores)
- to summarize each measurement. This idea allows any differences in the precision of different measurements to be taken into account,
- and in particular avoids the problem that poor-precision measurements can contaminate high-precision measurements and inflate the FDR.
+So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
+existing widely-used FDR pipelines) and can substantially improve
+inference. The first idea is to {\it assume that the distribution of effects is unimodal}, and to perform
+non-parametric inference under this assumption. This idea leads to a very simple, fast, and stable computer implementation,
+ and improves inference of FDR, at least when the unimodal assumption is correct. 
+ (And when the assumption is incorrect, errors in inference are not too egregious.) 
+The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
+to summarize each measurement. This idea allows variations in measurement precision to be accounted for,
+ and avoids the problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
 
@@
 However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
-perspectives, in practice, large-scale studies that assess many effects almost invariably focus on testing significance, and
-controlling the FDR, rather than on estimating the sizes of the effect. 
-The two new ideas highlighted above facilitate estimation as well as testing, and by providing easy-to-use software implementing our methods
-we hope to promote and accelerate the shift in this direction.
-
-The use of EB methods for estimation also has a long and close relationship with another key concept in modern statistics: shrinkage estimation \cite{efron.morris.73.jasa}.
-This close relationship becomes especially intimate here due to our assumption that the distribution of the underlying effects is
-unimodal, since this ensures that estimates are shrunk towards a common point -- the mode of the distribution.
+perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
+controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the terminology
+``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
+and correspondingly define the ``false signed discovery rate" (FSDR) in a natural way. We show that in some settings, particularly those with many discoveries, 
+the FSDR and FDR can be quite different, and emphasise benefits of the FSDR, particularly its increased robustness to modelling assumptions.
+
+%The methods we provide here can address both problems, and we provide results that highlight important benefits of focussing on estimation, and particularly on assessing the correctness of the {\it sign} of the estimated effects, rather than on testing for non-zero effects.
+
+The use of EB methods for estimation also has a long and close relationship with another key concept in modern statistics: shrinkage estimation \cite{efron1973}.
+This close relationship becomes especially intimate here due to our unimodal assumption on the underlying effects, 
+which encourages shrinkage towards the mode of the distribution.
 %Indeed, within our EB framework, multiple comparisons correction can be seen as a natural limit of shrinkage estimation:
@@
 shrinkage of large observed effects is less than when the data indicate that large effects are rare.
-Second, the amount of shrinkage undergone by each measurement will depend on its precision, or standard error: measurements with big standard errors will undergo more shrinkage than measurements with small standard errors.
+Second, the amount of shrinkage undergone by each measurement will depend on its precision, or standard error: measurements with big standard errors undergo more shrinkage than measurements with small standard errors.
 Shrinkage estimation is a powerful tool, with many potential applications, including for example wavelet denoising, 
 where shrinkage of wavelet coefficients is used to smooth signals. Our methods provide an attractive and competitive alternative
-to existing EB shrinkage methods for that context, as is explored more fully elsewhere (Xing and Stephens).
+to existing EB shrinkage methods for that context, as will be explored more fully elsewhere (Xing and Stephens, in preparation).
 
@@
 \def\fdr{\text{lfdr}}
+\def\FSDR{\text{FSDR}}
+\def\fsdr{\text{lfsdr}}
 
  \section*{Methods}
+ \includegraphics[]{main.pdf}
 
@@
 analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
-Further, this idea has a long history: for example, \cite{tukey:xx}, p32, suggests one should address
+Further, this idea has a long history: for example, \cite{tukey1962future}, p32, suggests one should address
 \begin{quote}
@@
  \end{equation}
- See \cite{storey:2002direct}, equation (8).
+ See \cite{storey.02}, equation (8).
 
@@
 this penalty here because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR. 
-Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{EM} to maximize the penalized likelihood, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
+Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77} to maximize the penalized likelihood, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
 (Note that the simplicity of the EM algorithm in step 1 is due to our using a fixed grid for $\sigma_k$ in (\ref{eqn:g1}), instead of estimating $\sigma_k$,
@@
 almost all analogous work in sparse regression models make the UA for the regression coefficents - 
-common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{}.
+common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
 (These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
-it may be fruitful to apply our methods to the regression context; indeed see \cite{goddard} for work in this vein.) 
+it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein.) 
 The UA assumption on regression coeffients is directly analagous to our UA for effects, and so we view its
@@
 \end{enumerate}
-The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine:1938} and  \cite{shepp:1962}; see \cite{feller:1971}, page  158. 
+The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
 In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.
@@
 \end{equation}
-so $z_j:=\bhat_j/s_j=\bhat_j$ are standard normal if null $H_j$ holds.
+so $z_j:=\bhat_j/s_j=\bhat_j$ is standard normal if null $H_j$ holds.
 
@@
 too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
-flatter distribution, which is a normal distribution with large variance (``bignorm") -- where most non-zero $\beta$ are easily detected
+flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
 making reliable estimation of $\pi_0$ much easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
@@
 skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
-bignormal & $N(0,4^2)$ \\ 
+big-normal & $N(0,4^2)$ \\ 
 bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
@@
 and a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
-as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted for clarity.)
+as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
 
@@
 estimation of$\pi_0$ impossible. This is particularly true for the ``spiky" scenario, for example.  The
-exception is the ``bignorm" scenario, where the \ashr estimates of $\pi_0$ are quite accurate - made possible by the fact that, in this scenario, most non-zero effects
+exception is the ``big-normal" scenario, where the \ashr estimates of $\pi_0$ are quite accurate - made possible by the fact that, in this scenario, most non-zero effects
 are very different from zero.
@@
 In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
-\cite{benjamini.yekutieli:2005} which \cite{efron2008} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
+\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
 how can confidence intervals be assigned to the true
-effect size for each selected case?". \cite{efron2008} notes the potential for EB approaches to tackle this problem,
-and \cite{zhao.hwang} consider in detail the case where the non-null effects are normally distributed.
+effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
+and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
 %but also notes the difficulties of implementing a stable general algorithm.
@@
 Bayes Fdr estimates, but it is hard to imagine a
-generally satisfactory parametric formulation..." (Efron, rejoinder, p46).
+generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
 The key point we want to emphasise here is that {\it if we restrict ourselves
@@
 the estimated $g$ from our method for each of the simulation scenarios above with the 
-estimated $g$ from mixFdr which does not make the UA (and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
+estimated $g$ from \mixfdr which does not make the UA (and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
 Even with only one simulated example for each scenario, the greater reliability of estimates afforded
-by the UA is immediately apparent. In particular the greater tendency for mixFdr to overestimate $\pi_0$ is readily visible in its behaviour at 0.
-And the estimated cdf from mixFdr often has a vertical (or almost vertical)
+by the UA is immediately apparent. In particular the greater tendency for \mixfdr to overestimate $\pi_0$ is readily visible in its behaviour at 0.
+And the estimated cdf from \mixfdr often has a vertical (or almost vertical)
 segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. This corresponds to
@@
 Interestingly, even in the ``bimodal" scenario where the UA is severely contradicted, the estimated $g$ from $\ashr$ 
-is visually more accurate than that from mixFdr: although mixFdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. We believe this
+is visually more accurate than that from \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. We believe this
 mostly reflects the fact that the noise level is sufficiently large to make reliable estimation of the multiple modes difficult, rather than 
-fundamental deficiencies of mixFdr. Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
-to be clearly visible in the observed $\bhat$, mixFdr fits these modes (Supplementary Information; 
+fundamental deficiencies of \mixfdr. Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
+to be clearly visible in the observed $\bhat$, \mixfdr fits these modes (Supplementary Information; 
 \url{dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
@@
 	\includegraphics[height=2in]{figures/egcdf.pdf}
-	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with mixFdr.}
+	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
 \end{subfigure}
@@
  We turn now to the second important component of our work: allowing for varying
- measurement precision across units. The key to this is our use of the likelihood
- (\ref{eqn:lik}), which explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
+ measurement precision across units. The key to this is our use of a likelihood,
+ (\ref{eqn:normlik}) or (\ref{eqn:tlik}), which explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
 
@@
     \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the estimated error rate ($q$ value, lfdr, or lfsr) for a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
+Each point shows the estimated error rate ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) for a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
 \end{subfigure}
@@
 
-Example of how shrinkage is adaptive to signal and to standard error?
-
-DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
-To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
+\cite{johnstone2004needles} note the ability of Empirical Bayes methods to adapt to overall signal strength
+when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then, given enough data, 
+the estimated distribution of effects ($g$ in (\ref{eqn:prior})) will reflect this by being concentrated near 0, and the posterior distributions
+ $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
+ In contrast, if most $\beta_j$ are large then $\hat{g}$ will reflect this by being flatter, resulting in less shrinkage.
+ Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
+ and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
+
+Here we build on this idea in two ways. First,
+by using flexible non-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
+Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
+adapts to measurement precision: more precise measurements undergo
+ less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
+than observations with
+small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
+To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
+Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=400px]{../examples/ash.pdf} 
+\end{center}
+\caption{Figure illustrating adaptive shrinkage.
+Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0.
+The standard error of each observation, $s_j$, was simulated from Inverse Gamma(5,5), which resulted in $s_j$ varying from 0.34 to 6.7.
+The shrunken estimates ($y$ axis) are plotted against the observed estimates ($x$ axis), with color indicating the (square root of the) precision of each measurement (blue=lower-precision; red=higher-precision).
+The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
+\end{figure}
+
+\subsection*{Model comparison, and the $p$ value prior}
+
+Thus far we have focussed on the model (\ref{eqn:beta}), where the distribution of $\beta_j$ is independent of the standard error $s_j$. 
+This model corresponds to the case $\alpha=0$ in the more general model (\ref{eqn:beta-alpha}).
+In the previous section we emphasised that, under this model, imprecise measurements are shrunk more strongly than
+precise measurements, and that the $\lfsr$ and $\lfdr$ are not a monotonic function of $p$ value. 
+In this section we emphasise that this behaviour {\it depends on the assumption that $\beta_j$ is independent of the standard error $s_j$.}
+Specifically, if we make a different assumption, setting $\alpha=1$ in (\ref{eqn:beta-alpha}) so $\beta_j$ scales linearly with $s_j$,
+then we get different behaviour, as summarized by the following result.
+\begin{result}
+Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and under any prior distribution of the form 
+(\ref{eqn:beta-alpha}) with $\alpha=1$, 
+the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$.
+Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$)
+is montonic with $|t_j|$. 
+\end{result}
+
+This result is a translation and extension of the result of \cite{wakefield:2009}, who showed that, when testing a series of null hypotheses, 
+with a normal likelihood, Bayes Factors and $p$ values produce the same ranking of tests {\it if} the Bayes Factor for the $j$th test 
+is computed using a normal prior with mean 0 and variance $K s^2_j$ for some constant $K$. 
+%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
+%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.
+
+Most previous authors of empirical Bayes approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
+have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
+In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
+This is analogous
+to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
+standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
+that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.
+In this sense, setting $\alpha=1$ in our approach provides the closest correspondance with existing methods (although the unimodal assumption
+still distinguishes our approach from existing methods)
+
+Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
+which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
+Framing the problem in this way - as comparing different modelling assumptions for $\beta_j$, rather than as 
+comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
+likelihood-based methods can be used to select $\alpha$. For example, following the logic of the Empirical Bayes approach it
+would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
+which has been implemented in our software by C. Dai. Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
+may prefer to choose whichever of $\alpha=0,1$ yields the highest likelihood, and this is also straightforward.
+
+%DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
+%To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
 Because the prior distribution is unimodal about zero,
-the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
-amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
-For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
-than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
-small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
-
-The posterior distribution $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ then reflects a trade-off
-between the information in the prior distribution, which is more-or-less concentrated near zero,
-
-There are two key points to appreciate here. First, by estimating the hyperparameters $\pi$ from the data, 
-the method adapts to the amount of signal in the data: if most $\beta_j$ are truly at or near zero then $\pi$ will reflect this by
-concentrating on zero and very small values of $\sigma_k$, and the posterior distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$
-will be tend also to be concentrated around zero (strong shrinkage). In contrast, if most $\beta_j$ are large then $\pi$ will be put most weight
-on components with large $\sigma_k$, and the posterior distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ will be approximately centered  on $\bhat_j$
-(weak shrinkage). Indeed, in the limit as $\pi$ puts all it's weight on very large $\sigma_k$ the prior becomes flatter and flatter
-and the posterior on $\beta_j$ becomes $N(\bhat_j,\shat_j)$. Thus, in this limit, the Bayesian credible intervals match standard confidence intervals.
-
- and only very strong observations will have a small fdr; if most $\beta_j$ are large
-then this will also be reflected in the estimates of $\pi$ and even borderline signals might have a small fdr. 
-This idea is also used by existing FDR methods, but here, by modelling the effects, it extends to our estimates of the effects:
-
-The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ encapsulate uncertainty in the values for $\beta_j$, combining information across
-$j=1,\dots,J$. The combining of the information occurs through estimation of
-$\pi$, which involves all of the data.
+%the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
+%amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
+%For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
+%than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
+%small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
+
+previous work based on normal means:
+http://link.springer.com/article/10.1007/BF02562623
 
@@
  not depend on the number of tests. This is because, as the number of tests increases, both the true positives and false positives increase linearly,
- and the FDR remains the same. (If this intuitve argument does not convince, see \cite{storey.xx}, and note that the FDR does not depend on the number of tests $m$.)
+ and the FDR remains the same. (If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.)
   A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
@@
 Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
-We used the elegant R package {\tt SQUAREM} \cite{squarem}) to accelerate convergence of this EM algorithm.
-
-[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
+We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple}) to accelerate convergence of this EM algorithm.
+
+%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
 
@@
 Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
-prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the \cite{knitr} package within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ash} , and includes contributions 
+prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
+git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} , and includes contributions 
 from Chaixing (Rick) Dai, Mengyin Lu, and Tian Sen.

### 2015-09-14 paper/main.tex
add Tukey poster, and other updates

@@
  \section*{Methods}
- \includegraphics[]{main.pdf}
-
- \subsection*{Notation}
+
+ \subsection*{Outline}
 
@@
 %mean expression level between the two conditions, often referred to as the ``log-fold change". 
-
-Suppose that we are interested in the values of $J$ ``effects" $\beta_j$ ($j=1,\dots,J$). For example, in a canonical genomics application
-$\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
-In some contexts interest may focus on which of the $\beta_j$ are ``significantly" different from zero, whereas in other contexts interest may focus on estimating their values; the methods described here tackle both these problems.
- We assume that we have obtained data $D$ that consist of estimates $\bhat_1,\dots,\bhat_J$ of these effects,
- with corresponding (estimated) standard errors $\shat_1,\dots,\shat_J$. Let $H_j$ denote the event that the $j$th null hypothesis holds ($H_j:\beta_j=0$), and $t_j$ denote
- the $t$ statistic testing this null:
+We begin by presenting the simplest version of method, and comparing it with existing work, before adding further embellishments.
+
+Suppose that we are interested in the values of $J$ ``effects" $\beta=(\beta_1,\dots,\beta_J)$. For example, in a typical genomics application
+that aims to identify differentially expressed genes,  $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
+In contexts where FDR methods are applied, interest often focuses on identifying significant non-zero effects;
+that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we will also tackle
+the more general problem of estimating, and assessing uncertainty in, the effect sizes $\beta_j$.
+
+We assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of these effects,
+and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  
+Initially we assume that the number of observations used to compute these estimates are sufficiently large to justify a normal approximation,
+\begin{equation} \label{eqn:normlik}
+p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j) 
+\end{equation}
+where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
+
+We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
+as a mixture of null and alternative components:
+\begin{equation} \label{eqn:betaj0}
+g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
+\end{equation}
+where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
+We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat_j$ (a point we return to later), so
+\begin{equation}
+p(\beta | \shat) = \prod_j g(\beta_j).
+\end{equation}
+
+Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
+with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
+to use a mixture of zero-mean normal distributions:
+\begin{equation} \label{eqn:g1}
+g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
+\end{equation}
+where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
+Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
+represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
+ The symmetry of the normal distribution implies that (\ref{eqn:g1}) is not only unimodal, but also symmetric about zero. We discuss 
+ more flexible alternatives based on using mixtures of uniform distributions, and details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below.
+
+ Substituting (\ref{eqn:g1}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
+ yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
+ \begin{equation} \label{eqn:g}
+ g(\cdot; \pi) & = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
+ \end{equation}
+
+ Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, our joint model for $(\beta,\bhat)$ is:
+ \begin{align} \label{eqn:betaj}
+ p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\
+ p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2),
+ \end{align}
+ with $g$ given by the mixture of normals $(\ref{eqn:g})$.
+ This implies that the marginal model $p(\bhat | \shat,\pi)$ is also a mixture of normals:
+\begin{equation}
+p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)].
+\end{equation}
+
+The usual EB approach to fitting this model would involve two simple steps:
+\begin{enumerate}
+\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, yielding $\hat{\pi} := \arg \max L(\pi)$.
+\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
+against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | | \bhat, \shat,\hat{\pi})$.
+\end{enumerate}
+Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
+(Note that the simplicity of the EM algorithm in step 1 is due to our using a fixed grid for $\sigma_k$ in (\ref{eqn:g1}), instead of estimating $\sigma_k$,
+which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
+In practice we make a small modification to this procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we instead maximise a penalized
+likelihood, where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data (see details below). 
+We introduce this penalty because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
+
+The conditional distributions $p(\beta_j | \hat{\pi}, \bhat_j, \shat_j)$ 
+encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
+$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
+$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
+the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
+observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
+corresondingly little effect on the estimated FDR of other observations. 
+We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
+$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
+that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
+estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
+Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
+
+This approach, which follows \cite{wakefield:2009}, 
+should be a good approximation if $\bhat_j$ and $\shat_j$ are based on sufficiently many observations. We discuss below an improvement 
+for moderate numbers of observations, based on replacing this normal likelihood with a $t$ likelihood.
+
+What is new here?
+
+This approach, which follows  \cite{wakefield:2009},
+should be a good approximation if $\bhat_j$ and $\shat_j$ are based on sufficiently many observations.
+
+%Given these data, a typical FDR analysis pipeline in genomics would compute $z$ statistics $z_j = \bhat_j/\shat_j$, and a corresponding
+$p$ value $p_j$, and then feed these $p$ values to a software package like \qvalue, which will estimate the false discovery rate
+
+ Initially we assume
+ \begin{equation} \label{eqn:normlik}
+\bhat_j | \beta_j, \shat_j \sim N(\beta_j, \shat^2_j).
+\end{equation}
+This approach, which follows \cite{wakefield:2009}, 
+should be a good approximation if $\bhat_j$ and $\shat_j$ are based on sufficiently many observations. We discuss below an improvement 
+for moderate numbers of observations, based on replacing this normal likelihood with a $t$ likelihood.
+
+ Let $t_j$ denote
+ the $t$ statistic testing $H_j$:
  \begin{equation}  
@@
  \end{equation}
- where $T_\nu$ denotes the $t$ distribution on $\nu$ \df, with $\nu$ assumed known. 
+ where $T_\nu$ denotes the $t$ distribution on $\nu$ \df, with $\nu$ assumed known. (For simplicity we assume the degrees of freedom are the same for every test,
+ but that would be straightforward to relax).
  Let $p_j$ denote the $p$ value obtained by comparing $t_j$ with its null distribution,
@@
 \begin{equation} \label{eqn:FDR}
-\FDR(\Gamma) := \frac{\#\{j: H_j \cap j \in \Gamma\}}{\#\{j: j \in \Gamma\}},
+\FDR(\Gamma; H) := \frac{\#\{j: H_j \cap j \in \Gamma\}}{\#\{j: j \in \Gamma\}},
 \end{equation}
 taken to be 0 if both numerator and denominator are 0. Frequentist approaches to controlling the FDR attempt to control
-the expectation of $\FDR(\Gamma)$ where the expecation is taken over the sampling distribution of $\Gamma$ (which is a function of the procedure
+the expectation of $\FDR(\Gamma;H)$ where the expecation is taken over the sampling distribution of $\Gamma$ (which is a function of the procedure
 used, considered fixed, and the data $D$, considered random). Bayesian approaches, in contrast, condition on the observed data, and compute

### 2015-09-14 talks/Tukey_poster.tex
add Tukey poster, and other updates

\documentclass[a0paper,landscape]{tikzposter} % See Section 3

\usepackage[backend=bibtex,style=numeric,sorting=none]{biblatex}
\addbibresource{~/Dropbox/Documents/mainbib.bib}

\usepackage{amsmath}
%\usepackage{natbib}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{xspace}
% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

\def\qvalue{{\tt qvalue}\xspace}
\def\locfdr{{\tt locfdr}\xspace}
\def\mixfdr{{\tt mixfdr}\xspace}
\def\ashr{{\tt ashr}\xspace}
\def\bhat{\hat{\beta}}
\def\shat{\hat{s}}

\title{False Discovery Rates - A New Deal} \institute{Department of Statistics and Department of Human Genetics, University of Chicago} % See Section 4.1
\author{Matthew Stephens} 
%\titlegraphic{Logos}
\usetheme{Default} % See Section 5
\usecolorstyle{Spain}
\begin{document}
\maketitle % See Section 4.1
\begin{columns} % See Section 4.4
\column{0.38}
\block{The Set-up}{
For $j=1,\dots,J$, let 
\begin{align*}
\beta_j & \text{ denote the (unobserved) $j$th effect of interest.} \\
\bhat_j & \text{  denote a noisy measurement of $\beta_j$.} \\
\shat_j & \text{ denote the (estimated) standard error of $\bhat_j$.} \\
z_j & := \bhat_j/\shat_j \text{, and $p_j$ be the corresponding $p$ value testing $H_j: \beta_j=0$}
\end{align*}

A typical genomics pipeline analyses the $p$ values with the goal of estimating or controlling the false discovery rate (FDR) \cite{benjamini1995controlling}.

\innerblock{Examples}{
\qvalue \cite{storey.02} essentially attempts to decompose the distribution of $p$ values into two components, one null (uniform distribution), and the other alternative.

\locfdr \cite{efron2008microarrays} attempts to decompose the distribution of $z$ scores into two components, one null (standard normal), and the other alternative.

\mixfdr \cite{muralidharan2010empirical} attempts to decompose the distribution of $z$ scores into two components, one null (standard normal), and the other alternative (mixture of normals).% See Section 4.2
}

For example, \mixfdr assumes that 
\begin{equation}
p(z_j) = \pi_0 N(z_j; 0,1) + (1-\pi_0) f_1(z_j)
\end{equation}
where $f_1$ is assumed to follow a mixture of normal distributions.
It then estimates $\pi_0$ and $f_1$ by (penalized) maximum likelihood, from which estimates of (local) FDR follow.
}

\block{A New Deal}{
Similar to \cite{efron2008microarrays,muralidharan2010empirical} we take an Empirical Bayes (EB) approach to the problem, but with some key differences:
\begin{enumerate}
\item Decompose distribution of $\beta_j$ (instead of $p$ values or $z$ score): \qquad
$$p(\beta_j | s_j)  = \pi_0 \delta_0 + (1-\pi_0) g_1(\beta_j)$$
\item {\bf Unimodal Assumption (UA):} Assume $g_1$ to be unimodal about 0. 
\item Connect the observations with the model via a normal likelihood assumption:
$$\bhat_j | \beta_j, \shat_j \sim N(\beta_j, \shat_j).$$
\end{enumerate}

\innerblock{Details}{
\begin{enumerate}
\item Simple way to implement the UA is via a mixture of 0-centered normals:
$g_1(\beta_j) = \sum_{k=1}^K \pi_k N(\beta_j; 0,\sigma_k^2)$
where $K$ is large, and $\sigma_1,\dots,\sigma_K$ are a fixed fine grid spanning values from very small to very large.
Mixture proportions estimated by a simple EM algorithm.
%\item The UA provides very computationally stable and statistically stable estimates. (It regularizes $g_1$.)
%\item Using two numbers $\beta_j,\shat_j$, instead of one ($z_j$ or $p_j$) for each observation, allows differences in measument precision to be better accounted for. Specifically, large $\shat_j$ implies a flat likelihood, and so little information, something that is lost in modelling $z_j$ or $p_j$ directly.
\item The mixture of normals (above) also assumes $g_1$ is symmetric; can allow asymmetry using mixures of uniforms that end or start at 0.
\item Can generalize normal likelihood to $t$ likelihood to help allow for estimation of $\shat_j$.
\item Add penalty term to the likelihood to encourage $\pi_0$ to be as big as possible given data (avoids underestimating FDR).
\end{enumerate}
}
The methods are implemented in an R package \ashr (``Adaptive SHrinkage") available at \url{http://github.com/stephens999/ashr}. Code for this poster, draft paper in progress, and other info at \url{http://github.com/stephens999/ash}.
}

\column{0.31} % See Section 4.4

\block{UA changes $p$ value and $z$ score decomposition (dark-blue=null; cyan = alt)}{
Results from this approach can be strikingly different from existing methods
\includegraphics[width=13in]{../paper/figures/decomp_ZA_poster.pdf}
}

\block{Illustrative Simulations}{

%\begin{tikzfigure}[We conducted simulations under a variety of true $g_1$] \label{fig:sims}
\innerblock{We conducted simulations using various effect distributions $g_1$}{
\includegraphics[width=12.5in]{../paper/figures/scenarios_density.pdf}
}
%\end{tikzfigure}

\innerblock{The UA (if it holds) allows less conservative estimates of $\pi_0$}{
All methods over-estimate $\pi_0$ (deliberately); $\ashr$ is least conservative.
%which will lead to less conservative estimates of FDR). If UA doesn't hold, estimates are anti-conservative, but not too egregious.
\includegraphics[width=12.5in]{../paper/figures/pi0_est.pdf}
}

\innerblock{The UA leads to stable estimates of $g$ (vs \mixfdr)}{
Figures compare true cdf with estimate from different methods.
\includegraphics[width=12.5in,height=3.5in]{../paper/figures/egcdf.pdf} \\
See also \cite{efron2008microarrays} which discusses challenges of getting stable non-parametric estimates of $g$; the UA provides one simple solution to this problem. 
}

}

\column{0.31}

\block{Problem: low precision measurements have less power, diluting signal}{
The $p$ values from high-precision measurements (left) show more signal than those from low-precision measurements (center) due to higher power of former. Combining them (right) dilutes the signal.
\center\includegraphics[width=12in,height=3.2in]{../paper/figures/GOODPOOReg_hist.pdf}
\innerblock{Taking precision into account avoids dilution}{
Figure shows $q$ values (FDR estimates) for good-precision measurements. For \qvalue and \locfdr, the $q$ values are inflated by including low-precision measurements in the analysis. For \ashr they are not, because the $N(\beta,\shat_j)$ likelihood is essentially flat when $\shat_j$ is large, so low-precision measurements effectively ignored.
\includegraphics[width=12.5in,height=3.5in]{../paper/figures/GOODPOOReg_scatter.pdf}
}
}

\block{Adaptive Shrinkage}{
Methods are applicable not only to FDR, but also shrinkage more generally. Eg as a competitor to EB approaches like \cite{johnstone2004needles}, which
make more restrictive parametric assumptions than the UA. Because \ashr provides posterior distribution on each effect given observed data, can also tackle the ``post-selection" problem of providing interval estimates for ``significant" effects.
}

%\note{Notetext} % See Section 4.3
\renewcommand*{\bibfont}{\small}
\block{References}{
\begin{center}\mbox{}\vspace{-\baselineskip}
    \printbibliography[heading=none]   
    \end{center} 
}

\end{columns}

\end{document}

### 2015-09-14 dsc-shrink/plot_pi0_est.R

@@ -25,6 +25,6 @@ p=ggplot(df,
 print(p +scale_y_continuous(limits=c(0,1)) +
         scale_x_continuous(limits=c(0,1)) +
         #scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
-        coord_equal(ratio=1) + colScale)
+        coord_equal(ratio=1) + colScale + theme(legend.position = "bottom"))
 
 dev.off()

### 2015-09-14 paper/Rcode/makefig_FDReg_hist.R

@@ -37,7 +37,7 @@ bin_fdr = res.qvalue$pi0/temp$density #compute local fdr in each bin of histogra
 qval_fdr = bin_fdr[as.numeric(cut(pval,temp$breaks))]
 qval_fdr = pmin(1,qval_fdr) #threshold at 1, so fdr never bigger than 1
 
-pdf("figures/FDReg_hist.pdf",width=6,height=3)
+pdf("../figures/FDReg_hist.pdf",width=6,height=3)
 par(mai=c(0.5,0.5,0.2,0.2),mgp = c(3, 0.5, 0))
 layout(matrix(1:2,ncol=2,byrow=TRUE))
 plot_FDReg_hist(pval,res.qvalue$pi0,type=1,title="p values",cex.axis=0.8,cex.main=0.8)
@@ -47,7 +47,7 @@ dev.off()
 
 ############################## decomp_ZA.pdf #############################################
 
-pdf("figures/decomp_ZA.pdf",width=6,height=6)
+pdf("../figures/decomp_ZA.pdf",width=6,height=6)
 layout(matrix(1:12,ncol=3,byrow=FALSE))
 plotlabel=function(label,cex=1.5){
   plot(0,0,type="n",axes=FALSE,xlab="",ylab="")
@@ -87,6 +87,42 @@ axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
 
 dev.off()
 
+#a different layout (4 columns, 2 rows) for my Tukey poster
+pdf("../figures/decomp_ZA_poster.pdf",width=6,height=2)
+layout(matrix(1:8,ncol=4,byrow=TRUE))
+par(mar = c(1.3,2,1.5,0.2),mgp = c(3, 0.5, 0))
+ncz=25
+
+#p value histograms
+altnullhist(pval,qval_fdr,main="p values: qvalue",ncz=ncz,xaxt='n',cex.axis=0.8)
+#plot_FDReg_hist(pval,res.qvalue$pi0,type=2,title="p values",xaxt='n',cex.axis=0.8)
+axis(side=1, labels=FALSE,tck=-0.01)
+#mtext(side=3,"p values",line=1)
+altnullhist(pval,res.locfdr$fdr,main="locfdr",ncz=ncz,xaxt='n',cex.axis=0.8)
+axis(side=1, labels=FALSE,tck=-0.01)
[…]

### 2015-09-16 paper/main.tex
more reorganization of paper

@@
 %mean expression level between the two conditions, often referred to as the ``log-fold change". 
-We begin by presenting the simplest version of method, and comparing it with existing work, before adding further embellishments.
+We begin by presenting the simplest version of method, and comparing it with existing work, before detailing further embellishments.
 
@@
 We assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of these effects,
-and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  
-Initially we assume that the number of observations used to compute these estimates are sufficiently large to justify a normal approximation,
+and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
+which by Bayes theorem can be written as 
+\begin{equation}
+p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
+\end{equation}
+
+We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
+$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
+\begin{equation}
+p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
+\end{equation}
+Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
 \begin{equation} \label{eqn:normlik}
-p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j) 
+p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
 \end{equation}
@@
 where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
-We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat_j$ (a point we return to later), so
+We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
 \begin{equation}
@@
  \begin{equation} \label{eqn:g}
- g(\cdot; \pi) & = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
- \end{equation}
-
- Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, our joint model for $(\beta,\bhat)$ is:
+ g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
+ \end{equation}
+
+ Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
  \begin{align} \label{eqn:betaj}
@@
  with $g$ given by the mixture of normals $(\ref{eqn:g})$.
- This implies that the marginal model $p(\bhat | \shat,\pi)$ is also a mixture of normals:
-\begin{equation}
+
+In words, the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
+each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
+ Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
+\begin{equation} \label{eqn:marginallik}
 p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)].
@@
 \begin{enumerate}
-\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, yielding $\hat{\pi} := \arg \max L(\pi)$.
+\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
 \item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
@@
 which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
+
 In practice we make a small modification to this procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we instead maximise a penalized
@@
 
-The conditional distributions $p(\beta_j | \hat{\pi}, \bhat_j, \shat_j)$ 
-encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-corresondingly little effect on the estimated FDR of other observations. 
-We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-This approach, which follows \cite{wakefield:2009}, 
-should be a good approximation if $\bhat_j$ and $\shat_j$ are based on sufficiently many observations. We discuss below an improvement 
-for moderate numbers of observations, based on replacing this normal likelihood with a $t$ likelihood.
-
-What is new here?
-
-This approach, which follows  \cite{wakefield:2009},
-should be a good approximation if $\bhat_j$ and $\shat_j$ are based on sufficiently many observations.
-
-%Given these data, a typical FDR analysis pipeline in genomics would compute $z$ statistics $z_j = \bhat_j/\shat_j$, and a corresponding
-$p$ value $p_j$, and then feed these $p$ values to a software package like \qvalue, which will estimate the false discovery rate
-
- Initially we assume
- \begin{equation} \label{eqn:normlik}
-\bhat_j | \beta_j, \shat_j \sim N(\beta_j, \shat^2_j).
-\end{equation}
-This approach, which follows \cite{wakefield:2009}, 
-should be a good approximation if $\bhat_j$ and $\shat_j$ are based on sufficiently many observations. We discuss below an improvement 
-for moderate numbers of observations, based on replacing this normal likelihood with a $t$ likelihood.
-
- Let $t_j$ denote
- the $t$ statistic testing $H_j$:
+\subsection*{Quantities of interest, and the local false sign rate}
+
+As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
+In practice it is common, and desirable, to summarize this distribution to convey the ``significance" of observation $j$.
+
+For exampe, previous EB approaches to FDR analysis have focussed on the ``local FDR" 
+\cite{efron2008microarrays}
+\begin{equation} \label{eqn:fdr}
+\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
+\end{equation}
+The ``local" in ``local FDR" refers to the fact that $\fdr_j$ is the probability that the {\it particular}
+effect $j$ would be a false discovery, if we were to declare it a discovery.
+Storey's $q$ values  \cite{storey.03} provide an alternative way to summarize significance,
+and both $\fdr_j$ and a $q$ value for each observation ($q_j$) can be easily computed using our approach (see Appendix \ref{} for details on $q$ value calculations).
+
+Here however, we argue that an alternative quantity, which we refer to as the
+``local false sign rate", provides a more practically meaningful measure of significance than either $\fdr_j$ or $q_j$.
+Motivated by this, we define the local False Sign Rate (lfsr) for observation $j$ as 
 \begin{equation}
- t_j :=\bhat_j/\shat_j. 
- \end{equation}
- The $t$ statistics, of course, have a $t$ distribution under the null:
- \begin{equation} \label{eqn:tstat}
- t_j | H_j \sim T_\nu
- \end{equation}
- where $T_\nu$ denotes the $t$ distribution on $\nu$ \df, with $\nu$ assumed known. (For simplicity we assume the degrees of freedom are the same for every test,
- but that would be straightforward to relax).
- Let $p_j$ denote the $p$ value obtained by comparing $t_j$ with its null distribution,
- \begin{equation}
- p_j := \Pr(|T_\nu| > |t_j| ).
- \end{equation}
-Some methods for FDR analysis require $z$ scores instead of $t$ statistics or $p$ values;
- these methods can be applied by translating each $t_j$ into a corresponding
-$z$ score, which we denote $z_j$.
-
- %For our purposes, these estimates and standard errors could be obtained from the data
-%using any method, provided they (approximately) satisfy 
-% \begin{equation} \label{eqn:betahat-t}
-% \bhat_j \big{|} \, \beta_j, \shat_j \sim t_\nu(\beta_j,\shat_j),
-% \end{equation}
-%for some known degrees of freedom (\df) $\nu$, where $t_\nu(\mu,\sigma)$ denotes the generalized $t$ distribution on $\nu$ \df, with mean $\mu$ and scale parameter $\sigma$.
-(Equivalently, $(\bhat_j-\beta_j)/\shat_j \big{|} \, \beta_j \sim t_\nu$, where $t_\nu$ denotes the standard $t$ distribution on $\nu$ \df.)
-%In the special case $\nu \rightarrow \infty$ this includes the situation where the $\bhat$ can be considered normally distributed:
-%  \begin{equation} \label{eqn:betahat-norm}
-% \bhat_j  \big{|} \, \beta_j, \shat_j \sim N(\beta_j,\shat^2_j).
-% \end{equation}
-% Assuming (\ref{eqn:betahat-t}), the $t$-statistics defined by
-
-Let $\Gamma$ denote a subset of the tests $1,\dots,J$ declared ``significant" by some procedure. 
-We define the FDR for $\Gamma$ as the proportion of these tests that are ``false discoveries" in that $H_j$ is true:
-\begin{equation} \label{eqn:FDR}
-\FDR(\Gamma; H) := \frac{\#\{j: H_j \cap j \in \Gamma\}}{\#\{j: j \in \Gamma\}},
-\end{equation}
-taken to be 0 if both numerator and denominator are 0. Frequentist approaches to controlling the FDR attempt to control
-the expectation of $\FDR(\Gamma;H)$ where the expecation is taken over the sampling distribution of $\Gamma$ (which is a function of the procedure
-used, considered fixed, and the data $D$, considered random). Bayesian approaches, in contrast, condition on the observed data, and compute
-the posterior distribution for the $H_j$ given the data, which induces a posterior distribution on $\FDR(\Gamma)$. This posterior
-distribution can be used to obtain a point estimate for $\FDR(\Gamma)$, for example using the posterior mean. 
-To be more explicit, Bayesian (and EB) approaches give the posterior probability that $H_j$ holds, which
-analagous to \cite{efron2008microarrays} we refer to as the ``local FDR", and denote $\fdr_j$:
-%\begin{equation} \label{eqn:FDR}
-%\FDR(T) := \Pr(\text{$H_j$ false} | t_j \geq T).
-%\end{equation}
-\begin{equation} \label{eqn:fdr}
-\fdr_j := \Pr(H_j | D).
-\end{equation}
-The FDR can be estimated by the posterior mean for $\FDR(\Gamma)$, given by
-\begin{equation} 
-\widehat\FDR(\Gamma) := \frac{\sum_{j \in \Gamma} \fdr_j }{\#\{j: j \in \Gamma\}}.
-\end{equation}
-Although our approach here is (Empirical) Bayesian, one nice feature of FDR procedures is that 
-Frequentist and Bayesian procedures are often closely aligned, at least under certain assumptions; see
-for example \cite{storey.03}, particularly his Theorem 1.
-
-The $q$ value for observation $j$ is defined \cite{storey.03} as, roughly, the FDR for the set of observations that are at least as significant as observation $j$.
-In a Bayesian analysis it is natural to order the significance of the hypotheses by $\fdr_j$, so we define
-\begin{equation}
-q_j := \widehat\FDR(\{k:\fdr_k \leq \fdr_j\}).
-\end{equation}
-%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
-%\begin{equation}
-%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
-%\end{equation}
-%which is close to the definition in \cite{storey.03}.
-
-\subsection*{The local False Sign Rate}
+\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
+\end{equation}
+Thus $\lfsr_j$ gives the probability that we would get the sign of $\beta_j$ wrong if we
+were to make our best guess. (Naturally, we consider it an error to state that $\beta$ is positive or negative when it is truly zero.)
+To illustrate, suppose for concreteness
+that the minimum is achieved by the first term, $p(\beta_j \geq 0| \bhat, s)=0.05$ say. Then
+our best guess would be that $\beta$ is negative, and the probability that we have
+made an error in sign would be 0.05.
+
+Note that $\lfsr_j \geq \lfdr_j$ 
+because both the events $\beta_j \geq 0$
+and $\beta_j \leq 0$ include the event $\beta_j=0$.
+Thus, lfsr gives an upper bound for lfdr.
+
+Because our methods involve explicitly modelling effects $\beta_j$, rather than 
+$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
+As show later, these estimates are much more robust to underlying assumptions than estimates
+of the FDR or $\lfdr$. Thus, although for comparison with existing methods we have
+focussed on FDR computations, in practice we strongly recommend using $\lfsr$ and not $\lfdr$.
 
@@
 
-Motivated by this, we define the local False Sign Rate (lfsr) for observation $j$ as 
-\begin{equation}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
-\end{equation}
-Thus $\lfsr_j$ gives the probability that we would get the sign of $\beta_j$ wrong if we
-were to make our best guess. (Naturally, we consider it an error to state that $\beta$ is positive or negative when it is truly zero.)
-To illustrate, suppose for concreteness
-that the minimum is achieved by the first term, $p(\beta_j \geq 0| \bhat, s)=0.05$ say. Then
-our best guess would be that $\beta$ is negative, and the probability that we have
-made an error in sign would be 0.05.
-
-Note that $\lfsr_j \geq \lfdr_j$ 
-because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ include the event $\beta_j=0$.
-Thus, lfsr gives an upper bound for lfdr.
-
-Because our methods involve explicitly modelling effects $\beta_j$, rather than 
-$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
-As show later, these estimates are much more robust to underlying assumptions than estimates
-of the FDR or $\lfdr$. Thus, although for comparison with existing methods we have
-focussed on FDR computations, in practice we strongly recommend using $\lfsr$ and not $\lfdr$.
-
 \subsection*{Existing methods for FDR analysis}
+
+\subsection{Comparisons and contrasts with previous work}
+
+Note use of (\ref{eqn:normlik}) in \cite{wakefield:2009}.
+
+The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
+encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
+$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
+$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
+the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
+observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
+corresondingly little effect on the estimated FDR of other observations. 
+We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
+$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
+that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
+estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
+Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
 
@@
 
+\appendix
+
+\section{Definitions of FDR-related concepts used here}
+
+Let $\Gamma$ denote a subset of the tests $1,\dots,J$ declared ``significant" by some procedure. 
+We define the FDR for $\Gamma$ as the proportion of these tests that are ``false discoveries" in that $H_j$ is true:
+\begin{equation} \label{eqn:FDR}
+\FDR(\Gamma; H) := \frac{\#\{j: H_j \cap j \in \Gamma\}}{\#\{j: j \in \Gamma\}},
+\end{equation}
+taken to be 0 if both numerator and denominator are 0. Frequentist approaches to controlling the FDR attempt to control
+the expectation of $\FDR(\Gamma;H)$ where the expecation is taken over the sampling distribution of $\Gamma$ (which is a function of the procedure
+used, considered fixed, and the data $D$, considered random). Bayesian approaches, in contrast, condition on the observed data, and compute
+the posterior distribution for the $H_j$ given the data, which induces a posterior distribution on $\FDR(\Gamma)$. This posterior
+distribution can be used to obtain a point estimate for $\FDR(\Gamma)$, for example using the posterior mean. 
+To be more explicit, Bayesian (and EB) approaches give the posterior probability that $H_j$ holds, which
+analagous to \cite{efron2008microarrays} we refer to as the ``local FDR", and denote $\fdr_j$:
+%\begin{equation} \label{eqn:FDR}
+%\FDR(T) := \Pr(\text{$H_j$ false} | t_j \geq T).
+%\end{equation}
+\begin{equation} \label{eqn:fdr}
+\fdr_j := \Pr(H_j | D).
+\end{equation}
+The FDR can be estimated by the posterior mean for $\FDR(\Gamma)$, given by
+\begin{equation} 
+\widehat\FDR(\Gamma) := \frac{\sum_{j \in \Gamma} \fdr_j }{\#\{j: j \in \Gamma\}}.
+\end{equation}
+Although our approach here is (Empirical) Bayesian, one nice feature of FDR procedures is that 
+Frequentist and Bayesian procedures are often closely aligned, at least under certain assumptions; see
+for example \cite{storey.03}, particularly his Theorem 1.
+
+The $q$ value for observation $j$ is defined \cite{storey.03} as, roughly, the FDR for the set of observations that are at least as significant as observation $j$.
+In a Bayesian analysis it is natural to order the significance of the hypotheses by $\fdr_j$, so we define
+\begin{equation}
+q_j := \widehat\FDR(\{k:\fdr_k \leq \fdr_j\}).
+\end{equation}
+%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
+%\begin{equation}
+%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
+%\end{equation}
+%which is close to the definition in \cite{storey.03}.
+
 % Do NOT remove this, even if you are not including acknowledgments

### 2015-09-17 paper/main.tex
tidied section on lfsr to include only minimal reference to quantities not actually highlighted in results

@@
 perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
-controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the terminology
-``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
-and correspondingly define the ``false signed discovery rate" (FSDR) in a natural way. We show that in some settings, particularly those with many discoveries, 
-the FSDR and FDR can be quite different, and emphasise benefits of the FSDR, particularly its increased robustness to modelling assumptions.
+controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
+``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (lfdr) \cite{efron2008microarrays}, but which measures confidence  
+in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
+the lfsr and lfdr can be quite different, and emphasise benefits of the lfsr, particularly its increased robustness to modelling assumptions.
+
+%terminology
+%``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
+%and correspondingly define the ``false signed discovery rate", or simply ``false sign rate" (FSR), in a natural way.
 
@@
 \def\fdr{\text{lfdr}}
-\def\FSDR{\text{FSDR}}
-\def\fsdr{\text{lfsdr}}
+\def\FSR{\text{FSR}}
+\def\fsr{\text{lfsr}}
 
@@
 represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
- The symmetry of the normal distribution implies that (\ref{eqn:g1}) is not only unimodal, but also symmetric about zero. We discuss 
- more flexible alternatives based on using mixtures of uniform distributions, and details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below.
+ The symmetry of the normal distribution implies that (\ref{eqn:g1}) is not only unimodal, but also symmetric about zero. We introduce
+ more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below.
 
@@
 
-\subsection*{Quantities of interest, and the local false sign rate}
+\subsection*{The local False Discovery Rate and local False Sign Rate}
 
 As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
-In practice it is common, and desirable, to summarize this distribution to convey the ``significance" of observation $j$.
-
-For exampe, previous EB approaches to FDR analysis have focussed on the ``local FDR" 
-\cite{efron2008microarrays}
+In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
+One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
+the probability that effect $j$ would be a false discovery, if we were to declare it a discovery.
+In other words it is the probability that $\beta_j$ is actually zero:
 \begin{equation} \label{eqn:fdr}
@@
 \end{equation}
-The ``local" in ``local FDR" refers to the fact that $\fdr_j$ is the probability that the {\it particular}
-effect $j$ would be a false discovery, if we were to declare it a discovery.
-Storey's $q$ values  \cite{storey.03} provide an alternative way to summarize significance,
-and both $\fdr_j$ and a $q$ value for each observation ($q_j$) can be easily computed using our approach (see Appendix \ref{} for details on $q$ value calculations).
-
-Here however, we argue that an alternative quantity, which we refer to as the
-``local false sign rate", provides a more practically meaningful measure of significance than either $\fdr_j$ or $q_j$.
-Motivated by this, we define the local False Sign Rate (lfsr) for observation $j$ as 
-\begin{equation}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
-\end{equation}
-Thus $\lfsr_j$ gives the probability that we would get the sign of $\beta_j$ wrong if we
-were to make our best guess. (Naturally, we consider it an error to state that $\beta$ is positive or negative when it is truly zero.)
-To illustrate, suppose for concreteness
-that the minimum is achieved by the first term, $p(\beta_j \geq 0| \bhat, s)=0.05$ say. Then
-our best guess would be that $\beta$ is negative, and the probability that we have
-made an error in sign would be 0.05.
-
-Note that $\lfsr_j \geq \lfdr_j$ 
-because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ include the event $\beta_j=0$.
-Thus, lfsr gives an upper bound for lfdr.
-
-Because our methods involve explicitly modelling effects $\beta_j$, rather than 
-$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
-As show later, these estimates are much more robust to underlying assumptions than estimates
-of the FDR or $\lfdr$. Thus, although for comparison with existing methods we have
-focussed on FDR computations, in practice we strongly recommend using $\lfsr$ and not $\lfdr$.
-
-\cite{gelman2012we} emphasise that there are settings where
-$\beta_j=0$ is implausible, in which case the FDR is not useful: if every $\beta_j$ is non-zero then there
-is no such thing as a false discovery and the FDR will always be identically 0. 
-Gelman et.~al.~suggest that in such settings we should replace the concept of a false discovery with the
-concept of an ``error in sign". The idea is that, in settings where $\beta_j=0$ is implausible, the most fundamental
-inference objective is to ask which $\beta_j$ are positive and which are negative.
-Indeed, even in settings where some
-$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
-analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
-Further, this idea has a long history: for example, \cite{tukey1962future}, p32, suggests one should address
+
+The local FDR, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
+whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
+the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey.91}) argued that
+``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$.
+Thus asking 'Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
 \begin{quote}
@@
 \end{quote}
-
-\subsection*{Existing methods for FDR analysis}
-
-\subsection{Comparisons and contrasts with previous work}
+Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
+moving focus away from ``type I errors" towards ``type S errors", meaning errors in sign.
+
+Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
+would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
+\begin{equation} \label{eqn:lfsr}
+\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
+\end{equation}
+To illustrate, suppose for concreteness
+that $p(\beta_j < 0| \bhat, s, \hat\pi))=0.95,  p(\beta_j =0| \bhat, s, \hat\pi)) = 0.03, p(\beta_j >0| \bhat, s, \hat\pi) = 0.02$. Then 
+from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:lfdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
+our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.
+
+As our notation suggests, $\lfsr_j$ is intended to be compared and contrasted with $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
+small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
+Of coure, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
+this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
+(this follows from the definition because both the events $\beta_j \geq 0$
+and $\beta_j \leq 0$ include the event $\beta_j=0$).
+In this sense, as a measure of ``significance", $\lfsr$ is more ``conservative" than $\lfdr$. More importantly, as we illustrate in Section \ref{sec:lfsr},
+$\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.
+
+In addition to these ``local" measures of significance for each observation, we can also compute average error rates over subsets of observations,
+which can lead to measures of significance analogous to Storey's $q$ values for example. See Appendix \ref{} for details, which also outlines 
+frequentist versions of the false sign rate.
+
+%Because our methods involve explicitly modelling effects $\beta_j$, rather than 
+%$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
+
+%Indeed, even in settings where some
+%$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
+%analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
+
+\subsection*{Comparisons and contrasts with previous work}
 
@@
 
- \subsection*{Adaptive Shrinkage}
-
- %(See Discussion for ideas on how our methods may be applicable more generally; for example if, in place of $\shat_j$, we have a $p$ value testing $\beta_j=0$.)
-
- Here we introduce an EB method, which, for reasons discussed above, we call ``adaptive shrinkage" or ``ash".
- The method provides not only estimated $\fdr_j$ and $q$ values, but also shrinkage point estimates and credible intervals for $\beta_j$,
- and estimates of the local false sign rates.
-
-We begin by presenting the simplest version of our model, before embellishing it.
-Instead of partitioning the (observed) $z$ scores or $p$ values into null and alternative components,
-we use a model that partitions the (unobserved) effects $\beta_j$  into null and alternative components.
-That is, letting $g$ denote the distribution of the effects $\beta_j$, we write
-\begin{equation} \label{eqn:betaj}
-g(\beta_j) = \pi_0 \delta_0 + (1-\pi_0) g_1(\beta_j)
-\end{equation}
-where $\delta_0$ denotes a point mass at 0, and $g_1$ denotes the (unknown) distribution of the non-zero $\beta_j$.
-Thus $\pi_0$ denotes the proportion of effects that are null.
-
-At this point we introduce a key assumption, which distinguishes our approach from previous work: we assume that $g_1$ is unimodal,
-with a mode at 0. One simple way to achieve this ``Unimodal Assumption" (UA)  is
-to assume that $g_1$ is a mixture of zero-mean normal distributions
-\begin{equation} \label{eqn:g1}
-g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
-\end{equation}
-where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
-Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
-represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
- The symmetry of the normal distribution implies that (\ref{eqn:g1}) is not only unimodal, but also symmetric about zero. We discuss 
- more flexible alternatives based on using mixtures of uniform distributions, and details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$), as well as the UA more generally, below.
-
-Since the $\beta_j$ are unobserved, we need a model to connect the $\beta_j$ to observations. Here we
-treat the estimates $\bhat_j$ as the observed data, and assume that they are noisy measurements of the true $\beta_j$ with given standard error
-$\shat_j$:
-\begin{equation} \label{eqn:normlik}
-\bhat_j | \beta_j, \shat_j \sim N(\beta_j, \shat^2_j).
-\end{equation}
-This approach, which follows \cite{wakefield:2009}, 
-should be a good approximation if $\bhat_j$ and $\shat_j$ are based on sufficiently many observations. We discuss below an improvement 
-for moderate numbers of observations, based on replacing this normal likelihood with a $t$ likelihood.
-
-Combining (\ref{eqn:betaj})-(\ref{eqn:normlik}), and assuming independence of the observations and effects for different $j$ (given $\shat, \pi_0,w_1,\dots,w_k$), yields
-a hierarchical model for $\beta,\bhat$ given $\shat$, with unknown hyper-parameters $(\pi_0,w_1,\dots,w_k)$. To simplify notation we define $\pi_k := (1-\pi_0)w_k$ 
-so that the unknown hyper-parameters can be written as a vector $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1.
-The usual empirical Bayes approach to fitting this model would involve two steps:
-\begin{enumerate}
-\item Estimate the hyperparameters $\pi$ by maximum likelihood, yielding a maximum likelihood estimate $\hat{\pi}$.
-\item Compute quantities of interest (e.g. local fdr values, $q$ values, credible intervals etc.) from the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$.
-\end{enumerate}
-We make one modification to this procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we instead maximise a penalized
-likelihood, where the penalty (detailed below) encourages $\hat\pi_0$ to be as big as possible (within the constrains of the model and observed data). We use
-this penalty here because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR. 
-Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77} to maximize the penalized likelihood, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(Note that the simplicity of the EM algorithm in step 1 is due to our using a fixed grid for $\sigma_k$ in (\ref{eqn:g1}), instead of estimating $\sigma_k$,
-which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
-
-The conditional distributions $p(\beta_j | \hat{\pi}, \bhat_j, \shat_j)$ 
-encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-corresondingly little effect on the estimated FDR of other observations. 
-We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
 \subsubsection*{Discussion on UA}
@@
 To provide a more general approach to defining the distribution of effects $g(\beta_j)$, consider 
-\begin{equation} \label{eqn:g}
+\begin{equation} \label{eqn:g-general}
 g(\cdot; \pi) = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k f_k(\cdot) 
@@
 In a Bayesian analysis it is natural to order the significance of the hypotheses by $\fdr_j$, so we define
-\begin{equation}
+
+Frequentist approaches to FDR often concentrate on the {\it average} error rate for a subset of observations
+whose significance exceeds some threshold. Bayesian versions of such quantities can also be computed.
+For example, the FDR for any subset of effects $\Gamma \subset \{1,\dots,J\}$ can be estimated by
+\begin{equation} \label{eqn:FDRhat}
+\widehat\FDR(\Gamma) := \frac{\sum_{j \in \Gamma} \fdr_j }{\#\{j: j \in \Gamma\}}.
+\end{equation}
+And, analogous to \cite{storey.03}, we can define the $q$ value 
+for observation $j$ ($q_j$) as an estimate of the FDR for all observations that are at least as significant as $j$:
+\begin{equation} \label{eqn:qj}
 q_j := \widehat\FDR(\{k:\fdr_k \leq \fdr_j\}).
 \end{equation}
+
+Since the FDR terminology has become so widespread, we here introduce terminology for these ideas that parallels FDR terminology.
+Recall that, in FDR parlance, a ``discovery" is an effect that is declared to be non-zero; and a ``false discovery" refers to any such declaration that is
+made in error. The FDR is the (expected) proportion of false discoveries among all discoveries.
+Consider now insisting that for every discovery, we {\it also declare the sign of the effect} (positive or negative).
+We call such a declaration a ``signed discovery"; and correspondingly a ``false signed discovery" is any such declaration that is
+made in error - that is, any signed discovery that is either 0, or whose sign is opposite to that declared.
+Based on these definitions we can define the ``false signed discovery rate" (or ``false sign rate", FSR, for short) as the (expected) proportion
+of false signed discoveries among all signed discoveries. Similarly, we can define the local false sign rate, $\lfsr_j$ to be
+the probability that observation $j$ would be a false signed discovery, were we to declare it a signed discovery. Or, equivalently, but perhaps
+more simply, the $\lfsr_j$ is the probability we would get the sign of $\beta_j$ wrong if we
+were to be forced to declare it either positive or negative.  Explicitly:
+
+%(Naturally, we consider it an error to declare $\beta$ to be positive or negative when it is truly zero.)
+By analogy with (\ref{eqn:FDRhat}) we can also define an estimate of the average False Sign Rate for any subset of effects,
+\begin{equation}  \label{eqn:FSRhat}
+\widehat\FSR(\Gamma) := \frac{\sum_{j \in \Gamma} \fsr_j }{\#\{j: j \in \Gamma\}}.
+\end{equation}
+And, by analogy with (\ref{eqn:qj}) we can define the ``$s$-value" for each observation
+\begin{equation} \label{eqn:svalue}
+s_j := \widehat\FSR(\{k:\fsr_k \leq \fsr_j\})
+\end{equation}
+being the average error rate for all observations that are at least as significant as $j$.
+
 %In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so

### 2015-09-18 paper/main.tex
added paragraph on differences between existing methods and new method

@@
  \begin{align} \label{eqn:betaj}
- p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\
+ p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
  p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
@@
 Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
-moving focus away from ``type I errors" towards ``type S errors", meaning errors in sign.
+focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.
 
@@
 (this follows from the definition because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ include the event $\beta_j=0$).
+and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
 In this sense, as a measure of ``significance", $\lfsr$ is more ``conservative" than $\lfdr$. More importantly, as we illustrate in Section \ref{sec:lfsr},
@@
 In addition to these ``local" measures of significance for each observation, we can also compute average error rates over subsets of observations,
-which can lead to measures of significance analogous to Storey's $q$ values for example. See Appendix \ref{} for details, which also outlines 
-frequentist versions of the false sign rate.
+leading to measures of significance analogous to Storey's $q$ values. See Appendix \ref{} for details. The appendix also provides a more
+thorough discussion of the connection between the Bayesian quantities we compute here and frequentist quantities, including  
+a brief discussion of the False Sign Rate from a frequentist perspective.
 
@@
 
-Note use of (\ref{eqn:normlik}) in \cite{wakefield:2009}.
-
-The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
-encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-corresondingly little effect on the estimated FDR of other observations. 
-We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-\subsubsection*{qvalue}
-
-One standard genomics analysis pipeline is to take the $p$ values $p_1,\dots,p_J$ and use them to estimate $q$ values $q_1,\dots,q_J$
- using the R package \qvalue.
-  Figure \ref{fig:qvalue} shows a graphical illustration of the procedure implemented in \qvalue. 
- In brief, the idea is that the overall distribution of $p$ values can be decomposed into two constitutive
- components: one component corresponding to the ``null" tests, and a second component corresponding to the ``alternative" tests. 
-To perform this decomposition, \qvalue makes two key assumptions:
+Among previous methods that explicitly consider the estimation of False Discovery Rates and related quantities,
+our work here seems most naturally compared with the Empirical Bayes methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
+ (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
+  %Here we highlight the key differences between our approach and these three existing approaches.
+
+There are two key differences between the approach implemented here and all of these three existing methods. First, our approach 
+directly models the distribution of the {\it unobserved} effects $\beta_j$ as a mixture of null and alternative components (\ref{eqn:betaj}),
+and links these unobserved effects to
+the observations ($\bhat_j,\shat_j$) through a likelihood (\ref{eqn:normlik2}). In contrast $\locfdr, \mixfdr$ and $\qvalue$ all work directly with the observed test statistics 
+-- either $z$ scores (for \locfdr and \mixfdr), or $p$ values (\qvalue) -- and attempt to decompose them into null and alternative components. 
+While this may seem a trivial difference, it is not.  For example, as we illustrate in our results (Section \ref{good-poor-hist}), our approach
+can take better account of variations in measurement precision among observations, essentially because the measurement precision is reflected in the
+curvature of the likelihood \ref{eqn:normlik2}. More generally, we might summarize this difference by saying that whereas
+existing methods summarize the information on $\beta_j$ by a single number ($p_j$ or $z_j$), we instead work with two numbers
+($\bhat_j,\shat_j$), which, clearly, have the potential to be more informative.
+
+%allows for different modelling assumptions to be compared through likelihood comparisons.
+
+accounts for
+the different measurement precision of different observations through the likelihood \ref{eqn:normlik}, rather than
+
+directly models
+the  and whose
+
+. This may seem
+a trivial difference, but i
+
+works with two numbers
+$\bhat_j,\shat_j$ for each observation, rather than
+
+we assume that the distribution
+of the effects
+
+The fundamental idea underlying \locfdr, \mixfdr and \qvalue is to decompose the overall distribution of test statistics -- either $z$ scores (for \locfdr and \mixfdr), or $p$ values (\qvalue) -- into null and alternative components.
+
+Our approach differs from this in modelling the decomposition at the level of the {\it effect}s, rather than
+
+Second, our approach differs by making the Unimodal assumption on the
+
+ Once this decomposition is done computations of FDR-related quantities is essentially straightforward,
+and the primary difference between the methods is how they perform the decomposition. Starting with \qvalue, it makes two key assumptions:
   i) the null $p$ values have a uniform distribution, and ii) the $p$ values near 1 are all from
@@
  the alternative $p$ values. In particular, the height of the horizontal line is an estimate of $\pi_0$, the proportion of $p$ values that come from the null.
-  Now, for any given threshold, although we do not know {\it which} $p$ values correspond
- to  null tests  and which to alternative tests, the areas of the two components indicate {\it how many} $p$ values correspond to each,
- allowing the FDR to be estimated (Figure \ref{fig:qvalue}). Specifically, at threshold $\gamma$, and given an estimate $\hat\pi_0$ for $\pi_0$,  a
- natural non-parametric estimate of the FDR is
- \begin{equation} \label{eqn:FDRest}
-\hat{\FDR}(\gamma) = J \hat\pi_0 \gamma / \#\{ j: p_j \leq \gamma \}.
- \end{equation}
- See \cite{storey.02}, equation (8).
-
-\subsubsection*{EB approaches: \locfdr and \mixfdr}
-
-The fundamental
- idea of EB procedures is similar to that of \qvalue: partition the observations into null and alternative components.
-Here we highlight two EB methods that
- seem most relevant to our work here and which had good software implementations at the time of writing. 
- The first is due to \cite{efron2008microarrays}, implemented in the R package \locfdr.
+
+Starting with \locfdr,
+
+%two EB methods that
+% seem most relevant to our work here and which had good software implementations at the time of writing. 
+ The first is due to 
  This method differs from \qvalue in that, instead of working with $p$ values, it works with $z$ scores.
@@
 
-Another EB method, and perhaps the one most closely related to our
-work here, is due to \cite{muralidharan2010empirical}, implemented in the R package \mixfdr.
 Like \locfdr, \mixfdr works with $z$ scores, but it differs in that it {\it explicitly} models the alternative
@@
 which we believe to be due to details in the fitting procedure used.
+
+Note use of (\ref{eqn:normlik}) in \cite{wakefield:2009}.
+
+The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
+encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
+$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
+$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
+the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
+observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
+corresondingly little effect on the estimated FDR of other observations. 
+We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
+$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
+that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
+estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
+Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
+
+\subsubsection*{qvalue}
+
+%One standard genomics analysis pipeline is to take the $p$ values $p_1,\dots,p_J$ and use them to estimate $q$ values $q_1,\dots,q_J$
+%using the R package \qvalue.
+ % Figure \ref{fig:qvalue} shows a graphical illustration of the procedure implemented in \qvalue. 
+ %In brief, the idea is that the overall distribution of $p$ values can be decomposed into two constitutive
+ %components: one component corresponding to the ``null" tests, and a second component corresponding to the ``alternative" tests. 
+To perform this decomposition, \qvalue makes two key assumptions:
+  i) the null $p$ values have a uniform distribution, and ii) the $p$ values near 1 are all from
+ tests where the null is true. Under these assumptions performing the decomposition boils down to 
+ drawing the horizontal line on the $p$ value histogram that intersects with the distribution at $p=1$: the area under this horizontal
+ line defines the uniform component that is due to the null $p$ values, with the remainder necessarily corresponding to
+ the alternative $p$ values. In particular, the height of the horizontal line is an estimate of $\pi_0$, the proportion of $p$ values that come from the null.
+  Now, for any given threshold, although we do not know {\it which} $p$ values correspond
+ to  null tests  and which to alternative tests, the areas of the two components indicate {\it how many} $p$ values correspond to each,
+ allowing the FDR to be estimated (Figure \ref{fig:qvalue}). Specifically, at threshold $\gamma$, and given an estimate $\hat\pi_0$ for $\pi_0$,  a
+ natural non-parametric estimate of the FDR is
+ \begin{equation} \label{eqn:FDRest}
+\hat{\FDR}(\gamma) = J \hat\pi_0 \gamma / \#\{ j: p_j \leq \gamma \}.
+ \end{equation}
+ See \cite{storey.02}, equation (8).
+
+\subsubsection*{EB approaches: \locfdr and \mixfdr}
 

### 2015-09-21 paper/main.tex
update paper

@@
 One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
-the probability that effect $j$ would be a false discovery, if we were to declare it a discovery.
-In other words it is the probability that $\beta_j$ is actually zero:
+the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
+In other words it is the posterior probability that $\beta_j$ is actually zero:
 \begin{equation} \label{eqn:fdr}
@@
 whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
-the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey.91}) argued that
+the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey1991philosophy}) argued that
 ``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$.
@@
 
-\subsection*{Comparisons and contrasts with previous work}
+\subsection*{Related work}
+
+\subsubsection*{Previous approaches focussed on FDR}
 
@@
 
-There are two key differences between the approach implemented here and all of these three existing methods. First, our approach 
-directly models the distribution of the {\it unobserved} effects $\beta_j$ as a mixture of null and alternative components (\ref{eqn:betaj}),
-and links these unobserved effects to
-the observations ($\bhat_j,\shat_j$) through a likelihood (\ref{eqn:normlik2}). In contrast $\locfdr, \mixfdr$ and $\qvalue$ all work directly with the observed test statistics 
--- either $z$ scores (for \locfdr and \mixfdr), or $p$ values (\qvalue) -- and attempt to decompose them into null and alternative components. 
-While this may seem a trivial difference, it is not.  For example, as we illustrate in our results (Section \ref{good-poor-hist}), our approach
-can take better account of variations in measurement precision among observations, essentially because the measurement precision is reflected in the
-curvature of the likelihood \ref{eqn:normlik2}. More generally, we might summarize this difference by saying that whereas
-existing methods summarize the information on $\beta_j$ by a single number ($p_j$ or $z_j$), we instead work with two numbers
-($\bhat_j,\shat_j$), which, clearly, have the potential to be more informative.
+There are two key differences between the approach implemented here and all of these three existing methods. First, 
+whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
+we instead work with two numbers ($\bhat_j,\shat_j$).
+Here we are building on \cite{wakefield:2009}, who develops Bayesian tests (Bayes Factors) for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
+Using two numbers instead of one clearly has the potential to be more informative, and indeed,
+as an example of this, Section \ref{good-poor-hist} illustrates how it can improve performance by taking better account of variation in measurement precision among observations.
+(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
+% Or more generally Section xx illustrates the potential practical effects
 
 %allows for different modelling assumptions to be compared through likelihood comparisons.
-
-accounts for
-the different measurement precision of different observations through the likelihood \ref{eqn:normlik}, rather than
-
-directly models
-the  and whose
-
-. This may seem
-a trivial difference, but i
-
-works with two numbers
-$\bhat_j,\shat_j$ for each observation, rather than
-
-we assume that the distribution
-of the effects
-
-The fundamental idea underlying \locfdr, \mixfdr and \qvalue is to decompose the overall distribution of test statistics -- either $z$ scores (for \locfdr and \mixfdr), or $p$ values (\qvalue) -- into null and alternative components.
-
-Our approach differs from this in modelling the decomposition at the level of the {\it effect}s, rather than
-
-Second, our approach differs by making the Unimodal assumption on the
-
- Once this decomposition is done computations of FDR-related quantities is essentially straightforward,
-and the primary difference between the methods is how they perform the decomposition. Starting with \qvalue, it makes two key assumptions:
-  i) the null $p$ values have a uniform distribution, and ii) the $p$ values near 1 are all from
- tests where the null is true. Under these assumptions performing the decomposition boils down to 
- drawing the horizontal line on the $p$ value histogram that intersects with the distribution at $p=1$: the area under this horizontal
- line defines the uniform component that is due to the null $p$ values, with the remainder necessarily corresponding to
- the alternative $p$ values. In particular, the height of the horizontal line is an estimate of $\pi_0$, the proportion of $p$ values that come from the null.
-
-Starting with \locfdr,
-
-%two EB methods that
-% seem most relevant to our work here and which had good software implementations at the time of writing. 
- The first is due to 
- This method differs from \qvalue in that, instead of working with $p$ values, it works with $z$ scores.
- %Analagous to \qvalue, \locfdr attempts to decompose the overall distribution of $z$ scores into two constitutive components,
- %one corresponding to the null tests, and the other corresponding to the alternative tests. 
- In the simplest case
- \locfdr  makes assumptions analogous to the two key assumptions of \qvalue: i) the null $z$ scores have a $N(0,1)$ distribution,
- and ii) all the $z$ scores near 0 are from the null, which Efron
-calls the ``Zero Assumption" (ZA).  
-The \locfdr package first estimates the overall distribution of $z$ scores using a non-parametric regression approach, and then, using the two assumptions,  ``subtracts out" the
-null component, leaving the remainder to be the alternative component. In this way \locfdr, like \qvalue, avoids explicit modelling of the alternative component.
- (An important focus of Efron's work, also implemented in \locfdr, 
- is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
-theoretical null distribution. Although important, this issue is largely orthogonal to the issues we focus on here, which arise
-even if the theoretical null distribution holds precisely in the empirical data, and so we consider it just briefly in the Discussion.)
-
-Like \locfdr, \mixfdr works with $z$ scores, but it differs in that it {\it explicitly} models the alternative
-distribution of $z$ scores, using a mixture of normal distributions. In principle this approach also differs from \qvalue and \locfdr in
-that it does not explicitly make the ZA. However, we have found that in practice the results often approximately satisfy the ZA,
-which we believe to be due to details in the fitting procedure used.
-
-Note use of (\ref{eqn:normlik}) in \cite{wakefield:2009}.
-
-The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
-encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-corresondingly little effect on the estimated FDR of other observations. 
-We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-\subsubsection*{qvalue}
-
-%One standard genomics analysis pipeline is to take the $p$ values $p_1,\dots,p_J$ and use them to estimate $q$ values $q_1,\dots,q_J$
-%using the R package \qvalue.
- % Figure \ref{fig:qvalue} shows a graphical illustration of the procedure implemented in \qvalue. 
- %In brief, the idea is that the overall distribution of $p$ values can be decomposed into two constitutive
- %components: one component corresponding to the ``null" tests, and a second component corresponding to the ``alternative" tests. 
-To perform this decomposition, \qvalue makes two key assumptions:
-  i) the null $p$ values have a uniform distribution, and ii) the $p$ values near 1 are all from
- tests where the null is true. Under these assumptions performing the decomposition boils down to 
- drawing the horizontal line on the $p$ value histogram that intersects with the distribution at $p=1$: the area under this horizontal
- line defines the uniform component that is due to the null $p$ values, with the remainder necessarily corresponding to
- the alternative $p$ values. In particular, the height of the horizontal line is an estimate of $\pi_0$, the proportion of $p$ values that come from the null.
-  Now, for any given threshold, although we do not know {\it which} $p$ values correspond
- to  null tests  and which to alternative tests, the areas of the two components indicate {\it how many} $p$ values correspond to each,
- allowing the FDR to be estimated (Figure \ref{fig:qvalue}). Specifically, at threshold $\gamma$, and given an estimate $\hat\pi_0$ for $\pi_0$,  a
- natural non-parametric estimate of the FDR is
- \begin{equation} \label{eqn:FDRest}
-\hat{\FDR}(\gamma) = J \hat\pi_0 \gamma / \#\{ j: p_j \leq \gamma \}.
- \end{equation}
- See \cite{storey.02}, equation (8).
-
-\subsubsection*{EB approaches: \locfdr and \mixfdr}
-
- %The details of how this decomposition is performed vary among methods, but they share one important element in common with \qvalue: just as \qvalue assumes that the $p$ values near 1 are from the null, so \locfdr assumes that the $z$ scores near 0 are from the null; Efron calls this assumption the ``Zero Assumption" (ZA).
-
-\begin{figure}
-\includegraphics{Rcode/figures/FDReg_hist.pdf}
-\caption{Illustration of how FDR is estimated by the \qvalue \, package, by decomposing the overall distribution of $p$ values (left) into a null component and alternative component (right).
-This is done by estimating the density of $p$ values near 1, and drawing a horizontal line at that height to represent the density of the $p$ values from the null distribution (based on the assumption that all $p$ values near 1 are null).
-At a given threshold (red vertical line at 0.1), the area $A$ corresponds to the true discoveries, and the area $B$ corresponds to the true discoveries,
-so the FDR is estimated by $B/(A+B)$ which in this case is 0.28.} \label{fig:qvalue}
-\end{figure}
-
-\subsubsection*{Discussion on UA}
-
- Since we view the UA as a key assumption we briefly discuss it here.
-Although the UA will not apply to all situations, we argue that it will be reasonable in many practical contexts, especially those
-involving FDR where interest focusses on which $\beta_j$ differ from zero. This is because whenever ``$\beta_j=0$" is a plausible null hypothesis,  
-it seems reasonable to imagine that ``$\beta_j$ very near 0" is also plausible. From there it seems reasonable to imagine that larger effects become
+Second, our unimodal assumption (UA) that the effects are unimodal about zero is quite different from assumptions made by $\qvalue, \locfdr$ or $\mixfdr$.
+Indeed, $\locfdr$ assumes that all $z$ scores near 0 are null (Efron calls this the Zero Assumption; ZA), which implies that 
+under the alternative hypothesis the distribution of $z$ scores has {\it no mass at 0}; this contrasts strikingly with the UA, which implies that this distribution has its peak at 0! 
+Similarly, $\qvalue$ assumes that all $p$ values near 1 are null,
+which is the same as the ZA because $p$ values near 1 correspond to $z$ scores near 0. And although \mixfdr does not formally make the ZA,
+we have found that in practice, with default settings, the results often approximately satisfy the ZA (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}).
+Thus, not only do these existing methods not make the UA, they actually make assumptions that are, in some sense, as different
+from the UA as they can be.
+
+Given that the UA and ZA are so different, it seems worth discussing why we favor the UA. 
+Although the UA will not apply to all situations, we believe that it will often be reasonable, especially in FDR-related contexts
+that have traditionally focussed on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis,  
+it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible. Further, it seems reasonable to expect that larger effects become
 decreasingly plausible, and so the distribution of the effects will be unimodal about 0.
-Alternatively, we can motivate  the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
-such shrinkage is desirable from several standpoints for improving estimation accuracy. 
-Note that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
+To paraphrase Tukey, ``All we know about the world teaches us that large effects are rare, whereas small effects abound."
+We emphasise that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
 It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
+
+%Tukey "peri-null"
 
@@
 %has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}.
-In addition, we note that although the UA is not made by any commonly-used existing FDR methods,
-almost all analogous work in sparse regression models make the UA for the regression coefficents - 
+
+Alternatively, we could motivate the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
+such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
+make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).
+Further, almost all analogous work in sparse regression models make the UA for the regression coefficents - 
 common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
-(These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
-it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein.) 
-The UA assumption on regression coeffients is directly analagous to our UA for effects, and so we view its
+These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
+it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
+The UA assumption on regression coefficients is directly analagous to our UA for effects, and so we view its
 widespread use in the regression context as supporting its use here.
@@
 
-\subsubsection*{Embellishments}
+\subsubsection*{Other work}
+
+There is also a very considerable literature that does not directly focus on the FDR problem, but which involves essentially the same models,
+and deals with closely-related issues.  Among these, \cite{cordy1997deconvolution} is perhaps most similar our work:
+indeed all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term, appear in their paper.
+However, they focus entirely on the estimation of $g$, in contrast to our focus on estimating $\beta_j$; 
+and they provide no software implementation.
+More generally, the literature is too large to do a comprehensive review here, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
+Some pointers to recent papers in which other relevant
+citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}. Much (but not all) of the literature focusses on the homoskedastic case (i.e. $\shat_j$ all equal)
+whereas we allow for heteroskedasticity. Much of the recent literature involves more flexible non-parametric assumptions on 
+$g$ than the UA assumption we make here, and these methods provide
+alternatives to our approach in settings where the UA assumption is considered too restrictive.
+
+%The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
+%encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
+%$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
+%$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
+%the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
+%observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
+%corresondingly little effect on the estimated FDR of other observations. 
+%We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
+%$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
+%that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
+%estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
+%Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
+
+\subsection*{Embellishments}
 
@@
 
-To provide a more general approach to defining the distribution of effects $g(\beta_j)$, consider 
+Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:g}) implies that $g$ is not only unimodal, but also symmetric.
+Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
+ Therefore, we have implemented a more general approach, based on 
 \begin{equation} \label{eqn:g-general}
-g(\cdot; \pi) = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k f_k(\cdot) 
-\end{equation}
-where $f_k$ are pre-specified component distributions with one of the following forms: 
+g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) 
+\end{equation}
+where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
 \begin{enumerate}[(i)]
@@
 \item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,
-\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ or } U[\cdot; 0,a_k]$,
+\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,
 \end{enumerate}
 where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
-(In (iii) we include both components in the mixture (\ref{eqn:g}), so for a grid of values $a_1,\dots,a_K$ there
-are actually $2K+1$ components in the mixture.)
-The version of our model described above corresponds to (i).
+(In (iii) we include both components in the mixture (\ref{eqn:g}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
+and $\pi$ is a $2K+1$ vector that sums to 1.)
+The simplest version (\ref{eqn:g}) corresponds to (i).
 Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
 under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
-(\ref{eqn:tlik}) introduced below.
-Our use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
+(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
+(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
+however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)
 
@@
 The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
-See Section \ref{details} for our software defaults.
+See Section \ref{sec:grid} for our software defaults.
 
@@
 
-We can generalize the normal likelihood assumption (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
+We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
   \begin{equation} \label{eqn:tlik}
@@
 The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
-This generalization does not complicate inference when used with mixtures of uniform components for $g$ in (\ref{eqn:g}).
-For normal components the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
+This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms.
+When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
 
@@
  \begin{equation} \label{eqn:beta-alpha}
- \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot)
- \end{equation}
-for any $\alpha \geq 0$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
+ \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
+ \end{equation}
+for any $\alpha \geq 0$. Setting $\alpha=0$ yields (\ref{eqn:betaj}), and setting $\alpha=1$ corresponds to
 assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
@@
 
-The model \ref{eqn:beta-alpha} for general $\alpha$ is actually easily fitted using the algorithm for $\alpha=0$. 
+The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
 To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
@@
 yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
-In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
-treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
-the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
+%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
+%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
+%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
 
@@
 
+add details of software parameters and package versions used?
+
 \subsection*{Effects of the Unimodal Assumption}
@@
 \item The UA can produce very different inferences compared with the ZA.
-\item The UA can yield conservative estimates of the proportion of nulls that hold, $\pi_0$, and hence conservative estimates of FDR and \lfdr.
+\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of \lfdr.
 \item The UA results in a stable procedure, both numerically and statistically.
@@
 
-\subsubsection{The UA compared with the ZA}
+\subsubsection*{The UA and ZA can produce different inferences}
 
@@
 In contrast the method introduced here (\ashr) assumes that the effect sizes (and thus the $z$ scores) have a unimodal distribution about 0, resulting in a very different decomposition.
-(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero -- specifically, $\beta_j \sim N(0,1^2)$, $s_j=1$, resulting in $z_j \sim N(0,2^2)$ -- so the ``true" decomposition would make everything light blue.)} \label{fig:ZA}
+(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero -- specifically, $\beta_j \sim N(0,1)$, $s_j=1$, resulting in $z_j =\bhat_j\sim N(0,2)$ -- so the ``true" decomposition would make everything light blue.)} \label{fig:ZA}
 \end{figure}
 
-Figure \ref{fig:ZA} graphically illustrates the difference between the ZA made by existing methods,
-and the UA made by our method. The effects of the ZA are visually clear, producing a ``hole" 
- in the alternative $z$ score distribution around 0 for \qvalue, \locfdr and \mixfdr; in contrast, the UA 
- fills in this hole to produce a peak at 0.  Of course the null distribution also has a peak at 0, and the local fdr 
- under the UA is still smallest for $z$ scores that are far from zero (i.e. large $z$ scores remain the ``most significant")
- as one would hope.
-
- How do these differences influence estimates of the FDR? To address this, we note that 
- the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
- of $\pi_0$}, the proportion of nulls (at least in this simple case where the 
- $s_i$ are all identical). To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
+To illustrate the different inferences from the UA and ZA we show results for a single
+dataset simulated with $\beta_j \sim N(0,1)$ and $\shat_j=1$, assumed known (so $\bhat_j  \sim N(0,2^2)$). We used each of the methods
+\qvalue, \locfdr, \mixfdr and \ashr (all implemented in R packages) 
+to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
+The results (Figure \ref{fig:ZA}) illustrate the clear difference between the existing methods and our method. The effects of the ZA made by \qvalue and \locfdr
+are visually clear, producing a ``hole"  in the alternative $z$ score distribution around 0. Although \mixfdr does not formally make the ZA, its decomposition
+exhibits a similar hole. In contrast, due to the UA, the alterantive $z$ score distribution for \ashr is required to have a mode at 0, 
+effectively ``filling in" the hole.  (Of course the null distribution also has a peak at 0, and the local fdr 
+ under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)
+
+Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA.
+Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
+This would increase the null component (dark blue) at the expense of
+the alternative component (light blue).
+Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0.
+The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, perhaps just past) the point where this dip appears.
+In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
+
+ \subsubsection*{The UA produces conservative estimates of $\pi_0$ (if it holds)}
+
+ How do these differences influence estimates of the FDR? To address this, we note that  -- in this simple case where the 
+ $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
+ of $\pi_0$. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
  strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
  although the methods considered here differ from one another in many ways, having estimated $\pi_0$
- they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, our limited
+ they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
  empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
@@
 overestimates of $\pi_0$, which will yield overestimates of the FDR).
-
 To confirm this we performed comprehensive simulation experiments across a range of alternative scenarios (i.e. values for the distribution of non-zero effects,
@@
 flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
-making reliable estimation of $\pi_0$ much easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
-which, although we view as generally unrealistic, we include to assess robustness of our methods to deviations from the UA.
+making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
+which, although we view as often unrealistic, we include to assess robustness of our methods to deviations from the UA.
 
@@
 Thus these simulations assume the same
-precision $(s_j=1)$ for each measurement.
-
-We applied the methods implemented in the R packages \qvalue, \locfdr, \mixfdr and \ashr to
+precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).
+
+We applied  \qvalue, \locfdr, \mixfdr and \ashr to
 estimate $\pi_0$ for each data set. For \ashr we show results for $g$ modelled as a mixture of normal components (``ash.n") 
@@
 
+ An important focus of Efron's work, also implemented in \locfdr, 
+ is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
+theoretical null distribution. Although important, this issue is largely orthogonal to the issues we focus on here, which arise
+even if the theoretical null distribution holds precisely in the empirical data, and so we consider it just briefly in the Discussion.
+
  noting that the $p$ values in Figure \ref{} were actually generated under a scenario where
@@
 
-\section*{Implementation details}
-
-\subsubsection*{Choice of grid for $\sigma_k, a_k$}
+\section*{Implementation details} \label{sec:implement}
+
+\subsubsection*{Choice of grid for $\sigma_k, a_k$} \label{sec:grid}
 
@@
 term $h(\pi;\lambda)$ to the likelihood to encourage over-estimation of $\pi_0$:
-\begin{equation}
+\begin{equation} \label{eqn:penalty}
 h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
@@
 
+\subsubsection*{qvalue}
+
+%One standard genomics analysis pipeline is to take the $p$ values $p_1,\dots,p_J$ and use them to estimate $q$ values $q_1,\dots,q_J$
+%using the R package \qvalue.
+ % Figure \ref{fig:qvalue} shows a graphical illustration of the procedure implemented in \qvalue. 
+ %In brief, the idea is that the overall distribution of $p$ values can be decomposed into two constitutive
+ %components: one component corresponding to the ``null" tests, and a second component corresponding to the ``alternative" tests. 
+To perform this decomposition, \qvalue makes two key assumptions:
+  i) the null $p$ values have a uniform distribution, and ii) the $p$ values near 1 are all from
+ tests where the null is true. Under these assumptions performing the decomposition boils down to 
+ drawing the horizontal line on the $p$ value histogram that intersects with the distribution at $p=1$: the area under this horizontal
+ line defines the uniform component that is due to the null $p$ values, with the remainder necessarily corresponding to
+ the alternative $p$ values. In particular, the height of the horizontal line is an estimate of $\pi_0$, the proportion of $p$ values that come from the null.
+  Now, for any given threshold, although we do not know {\it which} $p$ values correspond
+ to  null tests  and which to alternative tests, the areas of the two components indicate {\it how many} $p$ values correspond to each,
+ allowing the FDR to be estimated (Figure \ref{fig:qvalue}). Specifically, at threshold $\gamma$, and given an estimate $\hat\pi_0$ for $\pi_0$,  a
+ natural non-parametric estimate of the FDR is
+ \begin{equation} \label{eqn:FDRest}
+\hat{\FDR}(\gamma) = J \hat\pi_0 \gamma / \#\{ j: p_j \leq \gamma \}.
+ \end{equation}
+ See \cite{storey.02}, equation (8).
+
 Frequentist approaches to FDR often concentrate on the {\it average} error rate for a subset of observations

### 2015-09-21 paper/main.tex
latest update to paper

@@
 
-add details of software parameters and package versions used?
+We compare results of \ashr with existing FDR-based methods implemented
+in the R packages \qvalue (v1.99.1 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
+In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
+to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.
 
@@
 
-Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
-standard error, $s_j=1$, and we consider the normal limit (\ref{eqn:normallik}). That
-is,
+Here we consider the effects of making the UA. In this section, to isolate these effects we consider the simplest case, where every observation has the same
+standard error, $s_j=1$. We also assume that the sample size is sufficient that the estimated standard error for each observation is accurate, $\shat_j=1$,
+and we consider the normal limit (\ref{eqn:normlik}). That is,
 \begin{equation} \label{eqn:bhatgivenbz}
-\bhat_j | \beta_j \sim N(\beta_j,s^2_j=1)
-\end{equation}
-so $z_j:=\bhat_j/s_j=\bhat_j$ is standard normal if null $H_j$ holds.
+\bhat_j | \beta_j \sim N(\beta_j,1),
+\end{equation}
+so $z_j:=\bhat_j/\shat_j=\bhat_j$ is standard normal if null $H_j$ holds.
+in this way we attempt to minimize the differences between our method (which works with $\bhat_j,\shat_j$) and
+the other methods that work directly with the $z$ scores or their corresponding $p$ values.
 
@@
  $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
- of $\pi_0$. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
+ of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
  strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
@@
 overestimates of $\pi_0$, which will yield overestimates of the FDR).
-To confirm this we performed comprehensive simulation experiments across a range of alternative scenarios (i.e. values for the distribution of non-zero effects,
-which we denote $g_1$), and values
+To confirm this we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects,
+$g_1$), and values
 for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:pi0sims}a, with details in Table \ref{table:scenarios}.
@@
 
-\begin{figure} \label{fig:pi0sims}
+\begin{figure} 
 \begin{center}
@@
 As expected, when the UA holds, the \ashr method is least conservative, and hence most accurate.
-When the UA assumption does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative, but quite accurate, particularly when using a mixture of uniforms to model the unimodal distribution.}
+When the UA assumption does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative (less so for ash.u than for ash.n in this case)} \label{fig:pi0sims}
 \end{figure}
@@
 
-Figure \ref{fig:pi0} compares estimated and true values of $\pi_0$ under each Scenario. 
+Figure \ref{fig:pi0sims}b) compares estimated and true values of $\pi_0$ under each Scenario. 
 For most scenarios no method reliably estimates $\pi_0$. This is expected because most scenarios include
 an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
-estimation of$\pi_0$ impossible. This is particularly true for the ``spiky" scenario, for example.  The
+estimation of $\pi_0$ impossible. This is particularly true for the ``spiky" scenario, for example.  The
 exception is the ``big-normal" scenario, where the \ashr estimates of $\pi_0$ are quite accurate - made possible by the fact that, in this scenario, most non-zero effects
@@
 remaining conservative).
-When the UA assumption does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative, although quite accurate.
+When the UA assumption does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.
 We view such a strong bimodal scenario as rather unlikely in most applications where FDR methods
@@
 caused by such behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
-in addition running it using our default initialization procedure (XX need to specify in methods?). We then compared the largest
-log-likelihood achieved across all 11 runs with the log-likelihood achieved by our default run.
+in addition running it using our default initialization procedure (see Implementation details). We then compared the largest
+log-likelihood achieved across all 11 runs with the log-likelihood achieved by the default run.
 When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference (< 0.02), 
@@
 
-We note one caveat on the accuracy of estimated $g$: we tend to systematically
+We note one caveat on the accuracy of estimated $g$: \ashr tends to systematically
 overestimate the mass of $g$ near zero, due to our deliberate use of a penalty term (\ref{eqn:penalty}) to avoid under-estimation of $\pi_0$.
@@
 
-\begin{figure}[h!] \label{fig:egcdf}
+\begin{figure}[h!] 
 \begin{subfigure}{\textwidth}
@@
 \includegraphics[height=2in]{figures/mean_cdf.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{}) so $\pi_0$ is systematically overestimated.}
+\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
 \end{subfigure}
@@
 \includegraphics[height=2in]{figures/mean_cdf_nopen.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{}) so $\pi_0$ is not systematically overestimated.
+\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
 Modest systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the modest effects of model mis-specification.}
 \end{subfigure}
-\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.}
+\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.} \label{fig:egcdf}
 \end{figure}
@@
 
-Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.
+%Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.
 
@@
 \begin{subtable}{\textwidth}
-	\input{tables/scoverage_all.tex}
+	\centering\input{tables/scoverage_all.tex}
 	\caption{All observations.
 Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario.
-This is due to the penalty term \ref{} which tends to cause over-shrinking towards zero; removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods.
-(Removing the penalty in the half-uniform case can lead to unexpected problems explained in section \ref{sec:unexpected}).}
+This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero; removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods.
+(Removing the penalty in the half-uniform case can lead to unexpected problems explained in Section \ref{sec:unexpected}).}
 \end{subtable}
@@
 \begin{subtable}{\textwidth}
-\input{tables/scoverage_neg.tex}
+\centering\input{tables/scoverage_neg.tex}
 \caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
@@
 \begin{subtable}{\textwidth}
-\input{tables/scoverage_pos.tex}
+\centering\input{tables/scoverage_pos.tex}
 \caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
@@
 
-previous work based on normal means:
-http://link.springer.com/article/10.1007/BF02562623
+%previous work based on normal means:
+%http://link.springer.com/article/10.1007/BF02562623
 
@@
 
-We initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. 
+By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
 Our rationale to initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),

### 2015-09-22 paper/main.tex
update paper

@@
 
-The local FDR, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
+The lfdr, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
 whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
@@
 ``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$.
-Thus asking 'Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
+Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
 \begin{quote}
@@
 that $p(\beta_j < 0| \bhat, s, \hat\pi))=0.95,  p(\beta_j =0| \bhat, s, \hat\pi)) = 0.03, p(\beta_j >0| \bhat, s, \hat\pi) = 0.02$. Then 
-from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:lfdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
+from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
 our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.
@@
 and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
-In this sense, as a measure of ``significance", $\lfsr$ is more ``conservative" than $\lfdr$. More importantly, as we illustrate in Section \ref{sec:lfsr},
+In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Section \ref{sec:lfsr},
 $\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.
 
-In addition to these ``local" measures of significance for each observation, we can also compute average error rates over subsets of observations,
-leading to measures of significance analogous to Storey's $q$ values. See Appendix \ref{} for details. The appendix also provides a more
+From these ``local" measures of significance, we can also compute average error rates over subsets of observations
+$\Gamma \subset \{1,\dots,J\}$. For example,   
+\begin{equation} \label{eqn:FDR}
+\widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
+\end{equation}
+estimates the FDR we would obtain if we were to declare all tests in $\Gamma$ significant.
+And 
+\begin{equation} \label{eqn:qvalue}
+q_j := \widehat \FDR(\{k: \fdr_k \leq \fdr_j\}) 
+\end{equation}
+provides a measure of significance analogous to Storey's $q$ value \cite{storey.03}.
+
+See Appendix for a more
 thorough discussion of the connection between the Bayesian quantities we compute here and frequentist quantities, including  
@@
 There is also a very considerable literature that does not directly focus on the FDR problem, but which involves essentially the same models,
-and deals with closely-related issues.  Among these, \cite{cordy1997deconvolution} is perhaps most similar our work:
+and deals with closely-related issues.  Among these, \cite{cordy1997deconvolution} is perhaps most similar to our work:
 indeed all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term, appear in their paper.
@@
 Some pointers to recent papers in which other relevant
-citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}. Much (but not all) of the literature focusses on the homoskedastic case (i.e. $\shat_j$ all equal)
-whereas we allow for heteroskedasticity. Much of the recent literature involves more flexible non-parametric assumptions on 
-$g$ than the UA assumption we make here, and these methods provide
-alternatives to our approach in settings where the UA assumption is considered too restrictive.
+citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}.
+Much (but not all) of the literature focusses on the homoskedastic case (i.e.
+$\shat_j$ all equal) whereas we allow for heteroskedasticity.
+And much (but not all) of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential.
+Several recent papers consider more flexible non-parametric assumptions on
+$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
+estimate (NPMLE) for $g$.
+These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive.
+However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
+may perform well for point estimation, it seems possible it will not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
 
@@
 Here we consider the effects of making the UA. In this section, to isolate these effects we consider the simplest case, where every observation has the same
-standard error, $s_j=1$. We also assume that the sample size is sufficient that the estimated standard error for each observation is accurate, $\shat_j=1$,
-and we consider the normal limit (\ref{eqn:normlik}). That is,
-\begin{equation} \label{eqn:bhatgivenbz}
-\bhat_j | \beta_j \sim N(\beta_j,1),
-\end{equation}
-so $z_j:=\bhat_j/\shat_j=\bhat_j$ is standard normal if null $H_j$ holds.
-in this way we attempt to minimize the differences between our method (which works with $\bhat_j,\shat_j$) and
-the other methods that work directly with the $z$ scores or their corresponding $p$ values.
+standard error, $s_j=1$ and all methods are provided that information. That is,
+$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$.
+In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the only difference between our method and methods like $\locfdr$ and $\mixfdr$ are in how they estimate $g$.
 
@@
 To illustrate the different inferences from the UA and ZA we show results for a single
-dataset simulated with $\beta_j \sim N(0,1)$ and $\shat_j=1$, assumed known (so $\bhat_j  \sim N(0,2^2)$). We used each of the methods
-\qvalue, \locfdr, \mixfdr and \ashr (all implemented in R packages) 
-to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
+dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2^2)$). We used each of the methods
+\qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
 The results (Figure \ref{fig:ZA}) illustrate the clear difference between the existing methods and our method. The effects of the ZA made by \qvalue and \locfdr
 are visually clear, producing a ``hole"  in the alternative $z$ score distribution around 0. Although \mixfdr does not formally make the ZA, its decomposition
-exhibits a similar hole. In contrast, due to the UA, the alterantive $z$ score distribution for \ashr is required to have a mode at 0, 
+exhibits a similar hole. In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
 effectively ``filling in" the hole.  (Of course the null distribution also has a peak at 0, and the local fdr 

### 2015-09-22 talks/shiny/shinytest.Rmd
update paper

@@
 ---
-title: Interactive illustration of unimodal assumption
+title: Illustration of unimodal assumption and role of penalty in ashr
 runtime: shiny
-output: ioslides_presentation
+output: 
+  ioslides_presentation:
+    smaller: true
 ---
@@
 nullalthist_z = function(z,pi0,nullcol="blue",altcol="cyan",...){
-  h=hist(z, freq=FALSE,col=nullcol,nclass=40,ylim=c(-0.5,0.5),...)
+  h=hist(z, freq=FALSE,col=nullcol,nclass=50,ylim=c(-0.5,0.5),...)
   nb = length(h$breaks)
@@
 
-## Illustration of unimodal assumption
-Increase the proportion of nulls until the alternative (cyan) looks unimodal.
+##
+
+Slide the slider to see how the proportion of nulls (pi0) changes the decomposition of $z$ scores into null (dark blue) and alternative (cyan).
+
+If you increase pi0 until the alternative (cyan) just stops looking unimodal, this mimics the UA of ashr. The ashr penalty pushes pi0 as far as possible in this way.
+
+If you increase pi0 until cyan hits 0 then you are making the ZA of locfdr and qvalue.
 ```{r, echo=FALSE}

### 2015-09-24 dsc-shrink/investigate_hu_badcoverage.rmd
rejigged dsc-shrink dsc for new version of dscr; added plot of lfsr and relevant text to paper

---
title: "Investigate coverage problem with half uniform"
output: html_document
---

```{r}
load("res.RData")

library("dscr")
library("ashr")
library(reshape2)
library(ggplot2)
library(dplyr)

neglong = 
  res$negprob %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars=c("method","scenario","seed",".id"),value.name="negprob") %>%
    filter(negprob > 0.95)

poslong = 
  res$posprob %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars=c("method","scenario","seed",".id"),value.name="posprob") %>%
    filter(posprob > 0.95)

reslong = 
  res$cdf_score %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%    
    melt(id.vars=c("method","scenario","seed",".id"))

reslong.pos = inner_join(reslong,poslong)
reslong.neg = inner_join(reslong,neglong)

```

```{r}
ash.hu.ft = reslong.pos %>% dplyr::filter(method=="ash.hu.s" & scenario=="flat-top")
head(ash.hu.ft)
```

```{r}
  eg=loadExample(dsc_shrink,42,"flat-top",method="ash.hu.s")
  eg$output$fitted.g
```

Notice how almost all the inferred weight is on a small positive component. As a result false sign rate will be small, and there will be a strong tendency
to overestimate 0 effects. This leads to coverage problems observed.

Now let's look at an example with u and spiky which seems to be somewhat badly calibrated for the negative discoveries. 
```{r}
ash.u.s.spiky = reslong.neg %>% dplyr::filter(method=="ash.u.s" & scenario=="spiky")
ash.n.s.spiky = reslong.neg %>% dplyr::filter(method=="ash.n.s" & scenario=="spiky")
hist(ash.u.s.spiky$value,nclass=100,main="histogram of quantile where observation falls in its CI; significant negative discoveries")
hist(ash.n.s.spiky$value,nclass=100,main="histogram of quantile where observation falls in its CI; significant negative discoveries")
```

So what seems to be happening here is that the uniform tail is too short; when observation falls outside of this tail it gets a zero quantile of posterior interval.

Can we find an example?
```{r}
  table((ash.u.s.spiky %>% filter(value==0))$seed)
  eg =loadExample(dsc_shrink,seed = 39,method="ash.u.s",scenario="spiky")
```

For comparison, here are the positive discoveries; here there is not too bad.
```{r}
ash.u.s.spiky = reslong.pos %>% dplyr::filter(method=="ash.u.s" & scenario=="spiky")
ash.n.s.spiky = reslong.pos %>% dplyr::filter(method=="ash.n.s" & scenario=="spiky")
hist(ash.n.s.spiky$value)
hist(ash.u.s.spiky$value)
```

### 2015-09-24 paper/main.tex
rejigged dsc-shrink dsc for new version of dscr; added plot of lfsr and relevant text to paper

@@
 \usepackage{graphicx}
+
+\usepackage{sectsty}
+\subsectionfont{\normalsize}
+\subsubsectionfont{\mdseries\itshape}
 
@@
 \item The UA can produce very different inferences compared with the ZA.
-\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of \lfdr.
-\item The UA results in a stable procedure, both numerically and statistically.
-\item The UA is somewhat robust to deviations from unimodality.
+\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of \lfdr and \FDR.
+\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
 \end{enumerate}
@@
 Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0.
-The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, perhaps just past) the point where this dip appears.
+The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears.
 In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
-
- \subsubsection*{The UA produces conservative estimates of $\pi_0$ (if it holds)}
-
- How do these differences influence estimates of the FDR? To address this, we note that  -- in this simple case where the 
- $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
- of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
- strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
- although the methods considered here differ from one another in many ways, having estimated $\pi_0$
- they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
- empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
-
-Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
- and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
-Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
-At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
-overestimates of $\pi_0$, which will yield overestimates of the FDR).
-To confirm this we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects,
+See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
+demonstration.
+
+ \subsubsection*{The UA can produce conservative estimates of $\pi_0$}
+
+The illustrative example in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
+Consequently \ashr will estimate smaller $\lfdr$s and FDRs than existing methods that make the ZA. 
+This is desirable, provided that these estimates remain conservative: that is,
+that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
+The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behaviour. To check its effectiveness
+we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects,
 $g_1$), and values
-for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:pi0sims}a, with details in Table \ref{table:scenarios}.
+for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
 They range from a ``spiky" distribution -- where many non-zero $\beta$ are
@@
 making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
-which, although we view as often unrealistic, we include to assess robustness of our methods to deviations from the UA.
+which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.
+
+% How do these differences influence estimates of the FDR? To address this, note that  -- in this simple case where the 
+% $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
+% of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
+% strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
+% although the methods considered here differ from one another in many ways, having estimated $\pi_0$
+% they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
+% empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
+%
+
+%Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
+% and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
+%Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
+%At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
+%overestimates of $\pi_0$, which will yield overestimates of the FDR).
 
@@
 \begin{center}
-\includegraphics[height=2in]{figures/scenarios_density.pdf} 
-\includegraphics[height=2in]{figures/pi0_est.pdf} 
+\begin{subfigure}{\textwidth}
+	\includegraphics[height=1.8in]{figures/scenarios_density.pdf} 
+	\caption{Densities $g_1$ used in simulations.} \label{fig:altdens}
+\end{subfigure}
+\begin{subfigure}{\textwidth}
+\includegraphics[height=1.8in]{figures/pi0_est.pdf} 
+\caption{Comparison of true and estimated values of $\pi_0$.
+When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ashr being least conservative, and hence most accurate.
+When the UA does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.} \label{fig:pi0sims}
+\end{subfigure}
+\begin{subfigure}{\textwidth}
+\includegraphics[height=1.8in]{figures/lfdr_est.png} 
+\caption{Comparison of true and estimated \lfdr~from \ashr (ash.n).
+Black line is $y=x$ and red line is $y=2x$.
+Estimates of \lfdr~are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
+\end{subfigure}
+\begin{subfigure}{\textwidth}
+\includegraphics[height=1.8in]{figures/lfsr_est.png} 
+\caption{As in c), but for \lfsr~instead of \lfdr. Estimates of \lfsr~are consistently less conservative than \lfdr~when UA holds, and also less anti-conservative in bimodal scenario.}
+\end{subfigure}
 \end{center}
-\caption{a) Summary of the alternative distributions used for different simulation scenarios.
-b) Comparison of true and estimated values of $\pi_0$ in these simulated scenarios.
-In general $\pi_0$ is impossible to estimate accurately, and all methods yield conservative (over-)estimates for $\pi_0$ (deliberately so, to ensure conservative estimates of FDR).
-As expected, when the UA holds, the \ashr method is least conservative, and hence most accurate.
-When the UA assumption does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative (less so for ash.u than for ash.n in this case)} \label{fig:pi0sims}
+\caption{Results of simulation studies (constant precision $s_j=1$).} \label{fig:lfsr}
 \end{figure}
-
-\begin{table}[!ht]
-\begin{tabular}{c c } \toprule
-Scenario & Alternative distribution, $g_1$  \\ \midrule
-spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
-near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
-flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
- &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
-skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
-big-normal & $N(0,4^2)$ \\ 
-bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
-\end{tabular}
-\caption{Summary of simulation scenarios considered} \label{table:scenarios}
-\end{table}
 
@@
 
-We applied  \qvalue, \locfdr, \mixfdr and \ashr to
-estimate $\pi_0$ for each data set. For \ashr we show results for $g$ modelled as a mixture of normal components (``ash.n") 
-and a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
+Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ashr ($y$ axis) with the true values ($x$ axis). 
+For \ashr we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
+and as a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
 as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
-
-Figure \ref{fig:pi0sims}b) compares estimated and true values of $\pi_0$ under each Scenario. 
-For most scenarios no method reliably estimates $\pi_0$. This is expected because most scenarios include
-an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
-estimation of $\pi_0$ impossible. This is particularly true for the ``spiky" scenario, for example.  The
-exception is the ``big-normal" scenario, where the \ashr estimates of $\pi_0$ are quite accurate - made possible by the fact that, in this scenario, most non-zero effects
-are very different from zero.
-
-Although accurate estimation of $\pi_0$ is impossible in general, the simulation results demonstrate
-that, when the UA holds, \ashr consistently provides substantially smaller, more accurate, estimates of $\pi_0$, than existing methods. 
-As noted above, this means that estimates of FDR from \ashr will also be smaller and more accurate under these scenarios (while
-remaining conservative).
-When the UA assumption does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.
-We view such a strong bimodal scenario as rather unlikely in most applications where FDR methods
-are used; nonetheless, users may find it reassuring that even in this case the results from \ashr should not be horribly misleading.
+The results show that \ashr provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
+in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ashr estimates can be slightly anti-conservative.
+We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
+are used. (In addition, the effects on \lfsr estimates turn out to be relatively modest; see below).
 
@@
 
+\subsubsection*{The lfsr is more robust than lfdr}
+
+The results above show that \ashr can improve on existing methods in producing smaller, more accurate, 
+estimates of $\pi_0$ (and, consequently, more accurate estimates of False discovery rates).
+Nonetheless, in many scenarios \ashr continues
+to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
+The explanation for this is that these scenarios include 
+an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
+estimation of $\pi_0$ impossible. Put another way, and as is well known, $\pi_0$ is not identifiable:
+ the data can effectively provide an upper bound on plausible values of $\pi_0$,
+but not a lower bound (because the data cannot rule out that everything is non-null, but with miniscule effects).
+If we desire conservative behaviour we have to estimate $\pi_0$ by its upper bound, which can
+be substantially larger than the true value. 
+Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
+overestimation of $\pi_0$ is corresponding overestimation of FDR (and \lfdr, and $q$ values).
+To illustrate, Figure \ref{fig:lfdr} compares the estimated \lfdr~from ash.n with the true value (computed using
+Bayes rule from the true $g_1$ and $\pi0$). As predicted, \lfdr~is overestimated, especially in scenarios where 
+$\pi_0$ is overestimated. All methods are similarly affected by this, and the ones that most grossly overestimate $\pi_0$ also most grossly overestimated \lfdr and FDR/$q$-values (not shown).
+
+The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
+ become substantially less troublesome if we use the local false sign rate \lfsr (Section \ref{}), rather than \lfdr, to measure significance. 
+ This is essentially because \lfsr is less sensitive to the estimate of $\pi_0$.
+To illustrate, Figure \ref{fig:lfsr} compares the estimated \lfsr~from ash.n with the true value: although the estimated \lfsr continue to
+be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the \lfdr, especially for
+the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in \lfsr than \lfdr.
+
+We emphasise that, compared with previous debates regarding estimation vs testing, 
+this section advances an additional reason for favoring focussing on the sign of an effect
+rather than testing whether it is 0. As noted above, many authors have argued to focus on estimation 
+because it is implausible that effects are {\it exactly} 0. Here we add that {\it even if one believes
+that some effects may be exactly zero}, it is still better to focus on the sign, because generally {\it the data are more informative about that question}
+and so inferences are more robust to, say, the inevitable mis-estimation of $\pi_0$.
+To provide some intuition, consider an observation with a $z$ score of 0. The \lfdr of this observation can range from 0 (if $\pi_0=0$)
+to 1 (if $\pi_0=1$). But, assuming a symmetric $g$, the $\lfsr>0.5$ whatever the value of $\pi_0$, because the observation $z=0$ says
+nothing about the sign of the effect.
+Thus, are two reasons to use the lfsr instead of the lfdr: it is more generally meaningful (e.g. it applies
+whether or not zero effects truly exist),  and estimation of lfsr is more robust.
+
 \subsubsection*{Numerical Stability}
@@
 log-likelihood achieved across all 11 runs with the log-likelihood achieved by the default run.
-When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference (< 0.02), 
-and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference (<0.02),
-and 6\% of runs showed an appreciable log-likelihood difference (>1), with the largest difference being 5.0. However, perhaps suprisingly, 
+When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference ($< 0.02$), 
+and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference ($<0.02$),
+and 6\% of runs showed an appreciable log-likelihood difference ($>1$), with the largest difference being 5.0. However, perhaps suprisingly, 
 even for this largest difference results from the default run (e.g.
@@
 Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
-to address any question about the effect sizes, including, for example, ``which effects likely exceed $T$'', where
-$T$ is some effect threshold of interest (and ``likely" can be quantified by whatever probability threshold is desired). 
-Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
-this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
-In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
-towards zero, with the strength of the shrinkage being estimated from the data.
+to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
+threshold $T$.
+%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
+%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
+%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
+%towards zero, with the strength of the shrinkage being estimated from the data.
 Further, because the posterior distribution is, by definition,
-conditional on the observed data, inferences based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
+conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
 been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
@@
 
-The ability of the EB approach to provide shrunken posterior distributions and interval estimates for parameters
-that are valid ``post-selection" is extremely attractive in principle.
+The ability of the EB approach to provide valid ``post-selection" interval estimates is extremely attractive in principle.
 But its usefulness in practice
@@
 ``the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying
-to be overly nonparametric in constructing the empirical
-Bayes Fdr estimates, but it is hard to imagine a
+to be overly nonparametric ... but it is hard to imagine a
 generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
-The key point we want to emphasise here is that {\it if we restrict ourselves
-to settings where we can assume that $g$ is unimodal}
-then a stable non-parametric approach becomes not only possible, but indeed quite straightforward.
+Our key point here is that the UA greatly simplifies the deconvolution problem.
 While not meeting Efron's desire for an entirely general nonparametric approach, we believe 
-that these settings include many cases of practical interest.
-
-To illustrate this, Figure \ref{fig:egcdf} compares
-the estimated $g$ from our method for each of the simulation scenarios above with the 
-estimated $g$ from \mixfdr which does not make the UA (and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
-Even with only one simulated example for each scenario, the greater reliability of estimates afforded
-by the UA is immediately apparent. In particular the greater tendency for \mixfdr to overestimate $\pi_0$ is readily visible in its behaviour at 0.
-And the estimated cdf from \mixfdr often has a vertical (or almost vertical)
-segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. This corresponds to
-one of the 3 mixture components being estimated to be centered at that location with very small variance. 
+that the UA can handle many cases of practical interest.
+
+To illustrate this, Figure \ref{fig:egcdf} compares the
+estimated $g$ from \ashr  with that from \mixfdr which does not make the UA 
+(and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
+The greater reliability of estimates afforded
+by the UA is immediately apparent. In particular the estimated cdf from \mixfdr often has an almost-vertical
+segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. 
 The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
@@
 
-Interestingly, even in the ``bimodal" scenario where the UA is severely contradicted, the estimated $g$ from $\ashr$ 
-is visually more accurate than that from \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. We believe this
-mostly reflects the fact that the noise level is sufficiently large to make reliable estimation of the multiple modes difficult, rather than 
-fundamental deficiencies of \mixfdr. Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
+Interestingly, even in the ``bimodal" scenario $\ashr$ 
+is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
+Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
+Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
 to be clearly visible in the observed $\bhat$, \mixfdr fits these modes (Supplementary Information; 
@@
 
-We note one caveat on the accuracy of estimated $g$: \ashr tends to systematically
-overestimate the mass of $g$ near zero, due to our deliberate use of a penalty term (\ref{eqn:penalty}) to avoid under-estimation of $\pi_0$.
-On careful inspection, this is visually apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
+We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ashr tends to systematically
+overestimate the mass of $g$ near zero. On careful inspection, this is apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
 and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf}b),
@@
 
-\subsubsection*{Calibration of posterior distributions}
-
-To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over all 100 simulations in each Scenario). 
+\subsubsection*{Calibration of posterior intervals}
+
+To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over 100 simulations in each Scenario). 
 Specifically we examine the empirical coverage of nominal lower 95\% credible bounds for a) all observations; b) significant negative discoveries; c) significant positive discoveries.
 We examine only lower bounds because the results for upper bounds follow by
-symmetry (except for the one asymetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
+symmetry (except for the one asymmetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
 in each case: for negative discoveries the lower bound is typically large and negative and limits how big (in absolute value) 
@@
 
-The results are shown in Table \ref{tab:coverage}.  
-As noted above, our use of a penalty term to avoid underestimating $\pi_0$ means that the estimated $g$s tends to have, systematically, slightly too much mass near 0.
-If the prior has too much mass near 0 then so will the posterior, so we should expect posterior intervals may also be ``over-shrunk" towards zero. 
-The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound will be negative and overshrinking will move the bound towards zero, making the left tail too big; for positive discoveries, the lower confidence bound will be positive, and
-overshrinking will move it toward zero, making the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
-and be prepared to put up with some overshrinkage to avoid it.
+The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
+as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
+
+% As noted above, the penalty term (\ref{eqn:penalty}) causes systematic over-estimation of the mass of $g$ near 0,  which one would expect to overshrink interval estimates towards zero. 
+%The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound is negative and overshrinking (moving it too far towards zero) makes the left tail too big; conversely, for positive discoveries, the lower confidence bound is positive, and
+%overshrinking (moving it too far towards zero) makes the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
+%and be prepared to put up with some overshrinkage to avoid it.
 
@@
 %\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
-
-\subsubsection*{Problems with removing the penalty term in the half-uniform case}
-
-When we tried using no penalty term in the half-uniform case we came across an unanticipated problem: when the data
-are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
-sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
-for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
-and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
 
@@
 Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario.
-This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero; removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods.
-(Removing the penalty in the half-uniform case can lead to unexpected problems explained in Section \ref{sec:unexpected}).}
+This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero.
+Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (not shown).
+Removing the penalty in the half-uniform case is not recommended (see Appendix).}
 \end{subtable}
@@
 
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds (methods with penalty term). } \label{tab:coverage}
+\caption{Table of empirical coverage for nominal 95\% lower credible bounds} \label{tab:coverage}
 \end{table}
 
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\input{tables/scoverage_all_nopen.tex}
-	\caption{All observations}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\input{tables/scoverage_neg_nopen.tex}
-\caption{``Significant" negative discoveries.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\input{tables/scoverage_pos_nopen.tex}
-\caption{``Significant" positive discoveries.}
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds (methods without the penalty term. }
-\end{table}
-
+%\begin{table}[!ht]
+%\begin{subtable}{\textwidth}
+%	\input{tables/scoverage_all_nopen.tex}
+%	\caption{All observations}
+%\end{subtable}
+%
+%\begin{subtable}{\textwidth}
+%\input{tables/scoverage_neg_nopen.tex}
+%\caption{``Significant" negative discoveries.}
+%\end{subtable}
+%
+%\begin{subtable}{\textwidth}
+%\input{tables/scoverage_pos_nopen.tex}
+%\caption{``Significant" positive discoveries.}
+%\end{subtable}
+%
+%\caption{Table of empirical coverage for nominal 95\% lower credible bounds (methods without the penalty term. }
+%\end{table}
+%
 %\subsubsection*{Reverse asymptotics}
@@
  We turn now to the second important component of our work: allowing for varying
- measurement precision across units. The key to this is our use of a likelihood,
- (\ref{eqn:normlik}) or (\ref{eqn:tlik}), which explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
-
-  To illustrate, we begin with a simple
- simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
+ measurement precision across units. The key to this is the use of a likelihood,
+ (\ref{eqn:normlik}) or (\ref{eqn:tlik}), that explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
+
+  To illustrate, we conduct a simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
  ($s_j=10$).  In both cases,
@@
 In this setting, the poor-precision measurements tell us very little, and any sane analysis should effectively ignore them. 
- However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor-hist}). This is because the poor measurements 
+ However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor}). This is because the poor measurements 
  produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
@@
 (Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
-for the good-precision observations are unaffected by including the low-precision observations in the analysis.
+for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).
 
 \begin{figure}
-\centering
 \begin{subfigure}{\textwidth}
-\includegraphics[width=6.5in]{Rcode/figures/GOODPOOReg_hist.pdf} 
+\centering\includegraphics[width=4in]{Rcode/figures/GOODPOOReg_hist.pdf} 
     \label{fig:goodpoor-hist}
@@
 \begin{subfigure}{\textwidth}
-\includegraphics[width=6.5in]{Rcode/figures/GOODPOOReg_scatter.pdf}
+\centering\includegraphics[width=4in]{Rcode/figures/GOODPOOReg_scatter.pdf}
     \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
@@
 \end{subfigure}
-\caption{Simulation illustrating how poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
+\caption{Simulation illustrating how, for existing FDR methods, 
+poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
 $p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
@@
 
-\subsection*{Adaptive Shrinkage}
-
-\cite{johnstone2004needles} note the ability of Empirical Bayes methods to adapt to overall signal strength
-when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then, given enough data, 
-the estimated distribution of effects ($g$ in (\ref{eqn:prior})) will reflect this by being concentrated near 0, and the posterior distributions
+\subsubsection*{Adaptive Shrinkage}
+
+Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
+when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
+the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
  $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
- In contrast, if most $\beta_j$ are large then $\hat{g}$ will reflect this by being flatter, resulting in less shrinkage.
+ In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
  Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
@@
 Here we build on this idea in two ways. First,
-by using flexible non-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
+by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
 Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
@@
 
-\subsection*{Local False Sign Rate}
-
-There are two reasons to use the lfsr instead of the lfdr: it is more generally meaningful (e.g. it applies
-whether or not zero effects exist),  and estimation of lfsr is more robust to modeling assumptions and estimation of $\pi_0$. 
-To illustrate this, we compared estimated and true values of both lfdr and lfsr for the simulated data (where the true
-values are computed by Bayes Theorem using the true value of $g$).
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[height=320px,width=540px]{Rcode/figures/lfsdr_sim1sim2_blowup.png}
-\end{center}
-\caption{Figure showing lfsr is more robust than lfdr}
-\end{figure}
-
-% Compare estimated lfsr with true lfsr for difficult simulation case, to show nearly achieves Bayes Risk?
-
 \subsection{Null behaviour}
@@
 Results of simulations under the null.
+
 %\subsection{Comparison of different underlying component distributions}
@@
 %Notes from Adrian Raftery - 
-Newcombe (1886)
-Dean and Raftery
+%Newcombe (1886)
+%Dean and Raftery
 %Uniform is super-efficient
@@
 \subsection{Discussion}
-
-Cite johnston and silverman; raykar and zhao, sinica.
 
@@
 \appendix
+
+\begin{table}[!ht]
+\centering\begin{tabular}{c c } \toprule
+Scenario & Alternative distribution, $g_1$  \\ \midrule
+spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
+near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
+flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
+ &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
+skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
+big-normal & $N(0,4^2)$ \\ 
+bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
+\end{tabular}
+\caption{Summary of simulation scenarios considered} \label{table:scenarios}
+\end{table}
 
@@
 %which is close to the definition in \cite{storey.03}.
+\section*{Problems with removing the penalty term in the half-uniform case}
+
+We note here an unanticipated problem we came across when using no penalty term in the half-uniform case: when the data
+are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
+sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
+for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
+and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
 

### 2015-09-24 dsc-shrink/addMethods.R

@@ -1,26 +1,0 @@
-sourceDir("methods")
-
-addMethod(dsc_shrink,name="ash.hu.nocxx",fn =ash.wrapper,args=list(mixcompdist="halfunif",cxx=FALSE),outputtype = "ash_output")
-addMethod(dsc_shrink,name="ash.hu",fn =ash.wrapper,args=list(mixcompdist="halfunif",cxx=TRUE),outputtype = "ash_output")
-addMethod(dsc_shrink,name="ash.u",fn =ash.wrapper,args=list(mixcompdist="unif",cxx=TRUE),outputtype = "ash_output")
-addMethod(dsc_shrink,name="ash.n",fn =ash.wrapper,args=list(mixcompdist="normal",cxx=TRUE),outputtype = "ash_output")
-
-addMethod(dsc_shrink,name="ash.hu.nw2",fn =ash.wrapper,args=list(mixcompdist="halfunif",nullweight=2,cxx=TRUE),outputtype = "ash_output")
-addMethod(dsc_shrink,name="ash.u.nw2",fn =ash.wrapper,args=list(mixcompdist="unif",nullweight=2,cxx=TRUE),outputtype = "ash_output")
-addMethod(dsc_shrink,name="ash.n.nw2",fn =ash.wrapper,args=list(mixcompdist="normal",nullweight=2,cxx=TRUE),outputtype = "ash_output")
-
-
-addMethod(dsc_shrink,name="ash.hu.s",fn =ash.wrapper,args=list(mixcompdist="halfunif",method="shrink",cxx=TRUE),outputtype = "ash_output")
-addMethod(dsc_shrink,name="ash.u.s",fn =ash.wrapper,args=list(mixcompdist="unif",method="shrink",cxx=TRUE),outputtype = "ash_output")
-addMethod(dsc_shrink,name="ash.n.s",fn =ash.wrapper,args=list(mixcompdist="normal",method="shrink",cxx=TRUE),outputtype = "ash_output")
-
-
-addMethod(dsc_shrink,name="mixfdr.tnull", fn=mixfdr.wrapper, args = list(theonull=TRUE),outputtype = "mixfdr_output")
-addMethod(dsc_shrink,name="mixfdr.enull", fn=mixfdr.wrapper, args = list(theonull=FALSE),outputtype = "mixfdr_output")
-
[…]

### 2015-09-24 dsc-shrink/addMethods.null.R

@@ -1,20 +1,20 @@
-sourceDir("methods")
+source_dir("methods")
 
-addMethod(dsc_znull,name="ash.hu",fn =ash.wrapper,args=list(mixcompdist="halfunif"),outputtype = "ash_output")
-addMethod(dsc_znull,name="ash.u",fn =ash.wrapper,args=list(mixcompdist="unif"),outputtype = "ash_output")
-addMethod(dsc_znull,name="ash.n",fn =ash.wrapper,args=list(mixcompdist="normal"),outputtype = "ash_output")
+add_method(dsc_znull,name="ash.hu",fn =ash.wrapper,args=list(mixcompdist="halfunif"),outputtype = "ash_output")
+add_method(dsc_znull,name="ash.u",fn =ash.wrapper,args=list(mixcompdist="unif"),outputtype = "ash_output")
+add_method(dsc_znull,name="ash.n",fn =ash.wrapper,args=list(mixcompdist="normal"),outputtype = "ash_output")
 
-addMethod(dsc_znull,name="ash.hu.s",fn =ash.wrapper,args=list(mixcompdist="halfunif",method="shrink"),outputtype = "ash_output")
-addMethod(dsc_znull,name="ash.u.s",fn =ash.wrapper,args=list(mixcompdist="unif",method="shrink"),outputtype = "ash_output")
-addMethod(dsc_znull,name="ash.n.s",fn =ash.wrapper,args=list(mixcompdist="normal",method="shrink"),outputtype = "ash_output")
+add_method(dsc_znull,name="ash.hu.s",fn =ash.wrapper,args=list(mixcompdist="halfunif",method="shrink"),outputtype = "ash_output")
+add_method(dsc_znull,name="ash.u.s",fn =ash.wrapper,args=list(mixcompdist="unif",method="shrink"),outputtype = "ash_output")
+add_method(dsc_znull,name="ash.n.s",fn =ash.wrapper,args=list(mixcompdist="normal",method="shrink"),outputtype = "ash_output")
 
-addMethod(dsc_znull,name="ash.hu.nw2",fn =ash.wrapper,args=list(mixcompdist="halfunif",nullweight=2),outputtype = "ash_output")
-addMethod(dsc_znull,name="ash.u.nw2",fn =ash.wrapper,args=list(mixcompdist="unif",nullweight=2),outputtype = "ash_output")
[…]

### 2015-09-24 dsc-shrink/addScenarios.R

@@ -1,4 +1,4 @@
-sourceDir("datamakers")
+source_dir("datamakers")
 
 
 addNamedScenarios = function(dsc,names,nsamp=1000){
@@ -6,7 +6,7 @@ addNamedScenarios = function(dsc,names,nsamp=1000){
 #the -nn suffix indicates no nulls (min_pi0 = max_pi0 = 0)
 
 if("near-normal-nn" %in% names){
-	addScenario(dsc,name="near-normal-nn",
+	add_scenario(dsc,name="near-normal-nn",
                     fn=rnormmix_datamaker,
                     args=list(
                       g=normalmix(c(2/3,1/3),c(0,0),c(1,2)),
@@ -19,7 +19,7 @@ if("near-normal-nn" %in% names){
 }
 
 if("flat-top-nn" %in% names){
-	addScenario(dsc,name="flat-top-nn",
+	add_scenario(dsc,name="flat-top-nn",
                     fn=rnormmix_datamaker,
                     args=list(
                       g=normalmix(rep(1/7,7),c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7)),
@@ -31,7 +31,7 @@ if("flat-top-nn" %in% names){
                     seed=1:100)
 }
 
-addScenario(dsc,name="skew-nn",
+add_scenario(dsc,name="skew-nn",
                       fn=rnormmix_datamaker,
                       args=list(
                         g=normalmix(c(1/4,1/4,1/3,1/6),c(-2,-1,0,1),c(2,1.5,1,1)),
@@ -42,7 +42,7 @@ addScenario(dsc,name="skew-nn",
                       ),
                       seed=1:100)
 
-addScenario(dsc,name="near-normal",
+add_scenario(dsc,name="near-normal",
                     fn=rnormmix_datamaker,
                     args=list(
                       g=normalmix(c(2/3,1/3),c(0,0),c(1,2)),
@@ -53,7 +53,7 @@ addScenario(dsc,name="near-normal",
                     ),
                     seed=1:100)
 
-addScenario(dsc,name="flat-top",
+add_scenario(dsc,name="flat-top",
                     fn=rnormmix_datamaker,
                     args=list(
                       g=normalmix(rep(1/7,7),c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7)),
@@ -65,7 +65,7 @@ addScenario(dsc,name="flat-top",
                     seed=1:100)
 
 
-addScenario(dsc,name="skew",
[…]

### 2015-09-24 dsc-shrink/addScenarios.null.R

@@ -1,15 +1,15 @@
-sourceDir("datamakers")
+source_dir("datamakers")
 
 
 #Now, for each scenario create an element of scenarios of the following form
-addScenario(dsc_znull,name="znull.1000",
+add_scenario(dsc_znull,name="znull.1000",
             fn=null_z_datamaker,
             args=list(
               nsamp=1000
             ),
             seed=1:500)
 
-addScenario(dsc_znull,name="znull.100",
+add_scenario(dsc_znull,name="znull.100",
             fn=null_z_datamaker,
             args=list(
               nsamp=100

### 2015-09-24 dsc-shrink/addScenarios100.R

@@ -1,8 +1,8 @@
-sourceDir("datamakers")
+source_dir("datamakers")
 
 
 #Now, for each scenario create an element of scenarios of the following form
-addScenario(dsc_shrink,name="A",
+add_scenario(dsc_shrink,name="A",
                     fn=rnormmix_datamaker,
                     args=list(
                       g=normalmix(c(2/3,1/3),c(0,0),c(1,2)),
@@ -13,7 +13,7 @@ addScenario(dsc_shrink,name="A",
                       ),
                     seed=1:100)
 
-addScenario(dsc_shrink,name="B",
+add_scenario(dsc_shrink,name="B",
                     fn=rnormmix_datamaker,
                     args=list(
                       g=normalmix(rep(1/7,7),c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7)),
@@ -25,7 +25,7 @@ addScenario(dsc_shrink,name="B",
                     seed=1:100)
 
 
-addScenario(dsc_shrink,name="C",
+add_scenario(dsc_shrink,name="C",
                       fn=rnormmix_datamaker,
                       args=list(
                         g=normalmix(c(1/4,1/4,1/3,1/6),c(-2,-1,0,1),c(2,1.5,1,1)),
@@ -37,7 +37,7 @@ addScenario(dsc_shrink,name="C",
                       seed=1:100)
 
 #scenarios An, Bn, Cn are the same as A,B,C but with nulls included (pi0 uniform on [0,1])
-addScenario(dsc_shrink,name="An",
+add_scenario(dsc_shrink,name="An",
                     fn=rnormmix_datamaker,
                     args=list(
                       g=normalmix(c(2/3,1/3),c(0,0),c(1,2)),
@@ -48,7 +48,7 @@ addScenario(dsc_shrink,name="An",
                     ),
                     seed=1:100)
 
-addScenario(dsc_shrink,name="Bn",
+add_scenario(dsc_shrink,name="Bn",
                     fn=rnormmix_datamaker,
                     args=list(
                       g=normalmix(rep(1/7,7),c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7)),
@@ -60,7 +60,7 @@ addScenario(dsc_shrink,name="Bn",
                     seed=1:100)
 
 
-addScenario(dsc_shrink,name="Cn",
+add_scenario(dsc_shrink,name="Cn",
                       fn=rnormmix_datamaker,
                       args=list(
[…]

### 2015-09-24 dsc-shrink/datamakers/datamaker.R

@@ -27,7 +27,7 @@ rnormmix_datamaker = function(args){
   beta = ifelse(isnull, 0,rnorm(args$nsamp,comp_mean(args$g)[comp],comp_sd(args$g)[comp]))
   sebetahat = args$betahatsd
   betahat = beta + rnorm(args$nsamp,0,sebetahat)
-  meta=list(beta=beta,pi0=pi0)
+  meta=list(g1=args$g,beta=beta,pi0=pi0)
   input=list(betahat=betahat,sebetahat=sebetahat,df=NULL)
   
   #end of meat of function

### 2015-09-24 dsc-shrink/methods/ash.wrapper.R

@@ -19,6 +19,20 @@ ash.wrapper=function(input,args=NULL){
   return(res)
 }
 
+#uses ash function to compute bayes rule by running it with true g
+bayes.wrapper = function(input, meta, args=NULL){
+  pi0 = meta$pi0
+  g1 = meta$g1 
+  g=g1
+  #create g by adding null component to g1
+  g$pi= c(pi0, (1-pi0)*g1$pi)
+  g$mean = c(0,g1$mean)
+  g$sd = c(0,g1$sd)
+  #do computations for bayes rule by running ash with g, with 0 iterations
+  res = do.call(ash, args= list(betahat=input$betahat,sebetahat=input$sebetahat,df=input$df,g=g,control=list(maxiter=0)))
+  return(res)
+}
+
 ash2beta_est =function(output){
   return (list(beta_est=output$PosteriorMean))
 } 

### 2015-09-24 dsc-shrink/plot_pi0_est.R

@@ -15,16 +15,16 @@ df = res$pi0_score %>% filter(scenario %in% PLOTSCENARIOS) %>% filter(method %in
 df$scenario = factor(df$scenario,levels=PLOTSCENARIOS)
 levels(df$scenario)= PLOTNAMES
 
-p=ggplot(df,
-         aes(score.pi0,score.pi0_est,colour=method,alpha=ALPHALEVEL)) +geom_point(shape=1) +
+pi0_plot=ggplot(df,
+         aes(pi0,pi0_est,colour=method,alpha=ALPHALEVEL)) +geom_point(shape=1) +
   facet_grid(. ~ scenario) + 
   guides(alpha=FALSE) +
   geom_abline(colour = "black") +
   xlab("True pi0") +
   ylab("Estimated pi0") 
-print(p +scale_y_continuous(limits=c(0,1)) +
+print(pi0_plot +scale_y_continuous(limits=c(0,1)) +
         scale_x_continuous(limits=c(0,1)) +
         #scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
-        coord_equal(ratio=1) + colScale + theme(legend.position = "bottom"))
+        coord_equal(ratio=1) + colScale + theme(legend.position = "top",axis.text.x = element_text(size = 8,angle=45)))
 
 dev.off()

### 2015-09-24 dsc-shrink/run_dsc.R

@@ -1,32 +1,34 @@
 library(dscr)
 
-dsc_shrink=new.dsc("shrink","dsc-shrink-files")
-source("addNamedScenarios.R")
-addNamedScenarios(dsc_shrink,c("spiky","near-normal","flat-top","skew","big-normal","bimodal"))
+dsc_shrink=new_dsc("shrink","dsc-shrink-files")
+source("add_named_scenarios.R")
+add_named_scenarios(dsc_shrink,c("spiky","near-normal","flat-top","skew","big-normal","bimodal"))
 
-source("addMethods.R")
+source("add_methods.R")
 
-addOutputParser(dsc_shrink,"ash2beta",ash2beta_est,"ash_output","beta_est_output")
-addOutputParser(dsc_shrink,"mixfdr2beta",mixfdr2beta_est,"mixfdr_output","beta_est_output")
+add_output_parser(dsc_shrink,"ash2beta",ash2beta_est,"ash_output","beta_est_output")
+add_output_parser(dsc_shrink,"mixfdr2beta",mixfdr2beta_est,"mixfdr_output","beta_est_output")
 
-addOutputParser(dsc_shrink,"ash2pi0",ash2pi0_est,"ash_output","pi0_est_output")
-addOutputParser(dsc_shrink,"mixfdr2pi0",mixfdr2pi0_est,"mixfdr_output","pi0_est_output")
-addOutputParser(dsc_shrink,"locfdr2pi0",locfdr2pi0_est,"locfdr_output","pi0_est_output")
-addOutputParser(dsc_shrink,"qvalue2pi0",qvalue2pi0_est,"qvalue_output","pi0_est_output")
+add_output_parser(dsc_shrink,"ash2pi0",ash2pi0_est,"ash_output","pi0_est_output")
+add_output_parser(dsc_shrink,"mixfdr2pi0",mixfdr2pi0_est,"mixfdr_output","pi0_est_output")
+add_output_parser(dsc_shrink,"locfdr2pi0",locfdr2pi0_est,"locfdr_output","pi0_est_output")
+add_output_parser(dsc_shrink,"qvalue2pi0",qvalue2pi0_est,"qvalue_output","pi0_est_output")
 
-addOutputParser(dsc_shrink,"ash2fitted.g",ash2fitted.g,"ash_output","g_output")
-addOutputParser(dsc_shrink,"mixfdr2fitted.g",mixfdr2fitted.g,"mixfdr_output","g_output")
[…]

### 2015-09-24 dsc-shrink/run_dsc_znull.R

@@ -4,11 +4,11 @@
 library(dscr)
 
 dsc_znull=new.dsc("znull","dsc-znull-files")
-source("addScenarios.null.R")
-source("addMethods.null.R")
+source("add_scenarios.null.R")
+source("add_methods.null.R")
 
 source("score.R")
-addScore(dsc_znull,score_logLR,"logLR","ash_output")
+add_score(dsc_znull,score_logLR,"logLR","ash_output")
 
 
 res=run_dsc(dsc_znull)

### 2015-09-24 dsc-shrink/score.R

@@ -38,6 +38,14 @@ score_fdr = function(data, output){
   return(c(S=output$res$fdr))
 }
 
+score_lfsr = function(data, output){
+  return(c(S=output$lfsr))
+}
+
+score_lfdr = function(data, output){
+  return(c(S=output$lfdr))
+}
+
 score_betahat = function(data, output){
   return(c(S=data$input$betahat))
 }

### 2015-09-30 examples/adaptive_shrinkage.rmd
update paper and dsc-shrink

@@
 library(ashr)
-
+library(ggplot2)
+library(scales)
 #load the simulation function, (rnormmix_datamaker)
@@
 ```{r plotting}
-df = data.frame(betahat = c(sim.spiky$input$betahat,sim.bignormal$input$betahat),
-                beta_est = c(beta.spiky.ash$PosteriorMean, beta.bignormal.ash$PosteriorMean),
-                lfsr = c(beta.spiky.ash$lfsr, beta.bignormal.ash$lfsr),
-                s=c(s,s),
-                scenario=c(rep("spiky",NSAMP),rep("big-normal",NSAMP)))
+make_df_for_ashplot = function(sim1,sim2,ash1,ash2,name1="spiky",name2="big-normal"){
+  df = data.frame(betahat = c(sim1$input$betahat,sim2$input$betahat),
+                beta_est = c(ash1$PosteriorMean, ash2$PosteriorMean),
+                lfsr = c(ash1$lfsr, ash2$lfsr),
+                s=c(sim1$input$sebetahat,sim2$input$sebetahat),
+                scenario=c(rep(name1,NSAMP),rep(name2,NSAMP)))
   df$signif = df$lfsr<0.05
-library(ggplot2)
-library(scales)
+  return(df)
+}
+
+ashplot=function(df,xlab="Observed beta-hat",ylab="Shrunken beta estimate"){
   ggplot(df,aes(x=betahat,y=beta_est,color=1/s)) + 
-  xlab("Observed beta-hat") +
-  ylab("Shrunken beta estimate") + geom_point() +  facet_grid(.~scenario) + geom_abline(intercept=0,slope=1) +
-scale_colour_gradient2(midpoint = median(1/s),low="blue", mid="white",high="red")  +coord_fixed(ratio=1)
+    xlab(xlab) +
+    ylab(ylab) + geom_point() +  facet_grid(.~scenario) +   geom_abline(intercept=0,slope=1) +
+  scale_colour_gradient2(midpoint = median(1/s),low="blue", mid="white",high="red",space="Lab")  +coord_fixed(ratio=1)
+}
+df=make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash,beta.bignormal.ash)
+ashplot(df)
 ggsave("ash.pdf")
@@
 
+A related consequence is that significance of each observation is no longer monotonic with $p$ value.
+
 ```{r}
+pval_plot=function(df){
   ggplot(df,aes(x=pnorm(-abs(betahat/s)),y=lfsr,color=log(s))) + geom_point() +  facet_grid(.~scenario) +xlim(c(0, 0.025)) + xlab("p value") + ylab("lfsr") +
   scale_colour_gradient2(midpoint = 0,low="blue", mid="white",high="red")  
+}
+pval_plot(df)
 ```
 
-```{r}
-hist(s,main="histogram of standard errors")
-summary(s)
+Let's see how these are affected by changing the modelling assumptions so that
+the standardized beta are exchangeable (rather than the beta being exchangeable).
+```{r run-ash-ES}
+beta.bignormal.ash.ES = ash(sim.bignormal$input$betahat, s,model="ES",mixcompdist="normal")
+beta.spiky.ash.ES = ash(sim.spiky$input$betahat, s,model="ES",mixcompdist="normal")
+df.ES = make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash.ES,beta.bignormal.ash.ES)
+ashplot(df.ES,ylab="Shrunken beta estimate (ES model)")
+pval_plot(df.ES)
 ```
@@
 ```
+
+```{r sehist}
+hist(s,main="histogram of standard errors")
+summary(s)
+```

### 2015-09-30 paper/main.tex
update paper and dsc-shrink

@@
 So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
-existing widely-used FDR pipelines) and can substantially improve
+existing widely-used FDR pipelines) and can substantially affect
 inference. The first idea is to {\it assume that the distribution of effects is unimodal}, and to perform
-non-parametric inference under this assumption. This idea leads to a very simple, fast, and stable computer implementation,
+non-parametric inference under this assumption. This yields a very simple, fast, and stable computer implementation,
  and improves inference of FDR, at least when the unimodal assumption is correct.  
- (And when the assumption is incorrect, errors in inference are not too egregious.) 
 The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
@@
 We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
-\begin{equation}
+\begin{equation} \label{eqn:beta-iid}
 p(\beta | \shat) = \prod_j g(\beta_j).
@@
 to use a mixture of zero-mean normal distributions:
-\begin{equation} \label{eqn:g1}
+\begin{equation} \label{eqn:mixnorm}
 g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
@@
 represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
- The symmetry of the normal distribution implies that (\ref{eqn:g1}) is not only unimodal, but also symmetric about zero. We introduce
+ The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
  more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below.
 
- Substituting (\ref{eqn:g1}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
+ Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
  yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
@@
 Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(Note that the simplicity of the EM algorithm in step 1 is due to our using a fixed grid for $\sigma_k$ in (\ref{eqn:g1}), instead of estimating $\sigma_k$,
+(Note that the simplicity of the EM algorithm in step 1 is due to our using a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
 which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
@@
 $\Gamma \subset \{1,\dots,J\}$. For example,   
-\begin{equation} \label{eqn:FDR}
+\begin{equation} \label{eqn:FDRhat}
 \widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
@@
 and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
-this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. For brevity we
-do not investigate this idea any further here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
+this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
+do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
 
@@
 
-%In the results section we sometimes find it notationally convenient to rewrite (\ref{eqn:g}) as a mixture of a ``null" component $\delta_0$, and a non-null component, which we denote $g_1$:
-%\begin{equation} \label{eqn:g1}
-%g(\cdot; \pi) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot).
-%\end{equation}
-
 For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
@@
 nothing about the sign of the effect.
-Thus, are two reasons to use the lfsr instead of the lfdr: it is more generally meaningful (e.g. it applies
+Thus, are two reasons to use the lfsr instead of the lfdr: it answers a question that is more generally meaningful (e.g. it applies
 whether or not zero effects truly exist),  and estimation of lfsr is more robust.
+
+Finally for this section, we briefly consider the question of whether we could obtain better estimates of $\lfsr$ than are provided here. 
+We speculate that to do this we would either have to i) relax our relatively stringent requirement for conservative behaviour, or ii) be willing to make the assumption
+that no effects are exactly 0. In  support of i), note that under all scenarios the cloud of estimates for $\lfsr$ roughly stop at the line $y=x$, suggesting that
+a reduction in stringency (e.g.~reducing the penalty term \ref{eqn:penalty}) would risk crossing this line more frequently.
+
+discussion on whether we really need a point mass.
 
@@
 
-%\subsection{Accounting for measurement precision improves estimates of FDR/fsr}
-
 \subsection*{Differing measurement precision across units}
@@
 \begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{Rcode/figures/GOODPOOReg_hist.pdf} 
+\centering\includegraphics[width=4in]{figures/GOODPOOReg_hist.pdf} 
     \label{fig:goodpoor-hist}
@@
 \begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{Rcode/figures/GOODPOOReg_scatter.pdf}
+\centering\includegraphics[width=4in]{figures/GOODPOOReg_scatter.pdf}
     \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the estimated error rate ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) for a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
+Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) of a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
 \end{subfigure}
@@
 In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
+\end{figure}
+
+\subsubsection*{Reordering of significance, and the ``$p$ value prior"}
+
+Another important consequence of accounting for varying
+ measurement precision across units is that \ashr re-orders the significance of the observations compared
+ with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
+ We see that poor precision measurements are assigned a higher lfsr than good precision measuments that have the same $p$ value.
+ The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
+ and so the $\lfsr$ for these poor-precision measurements is always high.
+
+This tendency to reorder the significance of observations, and specifically to downweight the significance of observations with low precision,
+is a consequence of an assumption
+that we have made up to now: that the distribution of the effect $\beta_j$ is independent of its standard error $s_j$ (equation (\ref{eqn:betaj})). 
+One way to relax this assumption is to allow $\beta_j$ to scale with $s_j^\alpha$ for some
+power $\alpha$; that is, to assume model (\ref{eqn:beta-alpha}). Setting $\alpha=0$ implies that $\beta_j$ is
+independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
+effects (in absolute value). This latter assumption may often be qualitatively plausible: 
+for example, in gene expression studies the standard error for gene $j$ depends partly on
+the variance of its expression among samples, and genes with a larger variance may tend to be less tightly regulated and
+so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
+The special case $\alpha=1$ corresponds to assuming that the $z$ scores $z_j=\beta_j/s_j$, and hence the corresponding $p$ values, 
+are identically distributed, independent of $s_j$. This is essentially the assumption made (implicitly or explicitly) by existing methods like \locfdr, \mixfdr and \qvalue,
+which model the $z_j$ or $p_j$ directly.  Under this assumption, and with the normal mixture prior for $\beta_j/s_j$ (\ref{eqn:mixnorm}),
+the \lfsr computed by \ashr provides the same {\it ranking} of observations as the $z$ scores and $p$ values (see Figure \ref{fig:lfsr_pval}, right panel).
+
+This observation extends a result in
+\cite{wakefield:2009}, who showed that, when testing a series of null hypotheses, 
+with a normal likelihood, Bayes Factors and $p$ values produce the same ranking of tests {\it if} the Bayes Factor for the $j$th test 
+is computed using a normal prior for the alternative effects with mean 0 and variance $K s^2_j$ for some constant $K$.
+Because of this correspondance in ranking, Wakefield refers to this as the ``$p$-value prior". 
+Our result extends Wakefield's results to the mixture of normals prior, from a single normal. As an aside we note
+that the result does depend on the normal assumptions: for example, it does not hold in general for mixtures of uniforms as a prior; nor
+if we use a $t$ likelihood instead of a normal likelihood. That is, under these models, the EB
+analysis may reorder the significance of the observations (compared with the $p$ values) even when $\alpha=1$.
+
+%\begin{result}
+%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
+%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
+%\end{result}
+
+%Indeed, any FDR-based approach that ranks the significance of observations 
+%by their $p$ values, like the Benjamini--Hochberg method, could be viewed as making the same implicit assumption.
+
+To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
+Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
+In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
+Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
+which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
+Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
+comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
+likelihood-based methods can be used to select $\alpha$. For example, following the logic of the Empirical Bayes approach it
+would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
+which has been implemented in our software by C. Dai. 
+(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
+may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)
+
+%\begin{result}
+%Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and using a mixture of normals for the prior distribution 
+%(\ref{eqn:beta-alpha}) with $\alpha=1$, 
+%the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$.
+Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$)
+%is montonic with $|t_j|$. 
+%\end{result}
+
+%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
+%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.
+
+%Most previous authors of empirical Bayes approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
+%have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
+%In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
+%This is analogous
+%to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
+%standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
+%that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.
+
+\begin{figure}
+\centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
+\caption{} \label{fig:lfsr_pval}
 \end{figure}
@@
 
-\subsection*{Model comparison, and the $p$ value prior}
-
-Thus far we have focussed on the model (\ref{eqn:beta}), where the distribution of $\beta_j$ is independent of the standard error $s_j$. 
-This model corresponds to the case $\alpha=0$ in the more general model (\ref{eqn:beta-alpha}).
-In the previous section we emphasised that, under this model, imprecise measurements are shrunk more strongly than
-precise measurements, and that the $\lfsr$ and $\lfdr$ are not a monotonic function of $p$ value. 
-In this section we emphasise that this behaviour {\it depends on the assumption that $\beta_j$ is independent of the standard error $s_j$.}
-Specifically, if we make a different assumption, setting $\alpha=1$ in (\ref{eqn:beta-alpha}) so $\beta_j$ scales linearly with $s_j$,
-then we get different behaviour, as summarized by the following result.
-\begin{result}
-Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and under any prior distribution of the form 
-(\ref{eqn:beta-alpha}) with $\alpha=1$, 
-the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$.
-Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$)
-is montonic with $|t_j|$. 
-\end{result}
-
-This result is a translation and extension of the result of \cite{wakefield:2009}, who showed that, when testing a series of null hypotheses, 
-with a normal likelihood, Bayes Factors and $p$ values produce the same ranking of tests {\it if} the Bayes Factor for the $j$th test 
-is computed using a normal prior with mean 0 and variance $K s^2_j$ for some constant $K$. 
-%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
-%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.
-
-Most previous authors of empirical Bayes approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
-have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
-In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
-This is analogous
-to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
-standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
-that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.
-In this sense, setting $\alpha=1$ in our approach provides the closest correspondance with existing methods (although the unimodal assumption
-still distinguishes our approach from existing methods)
-
-Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
-which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
-Framing the problem in this way - as comparing different modelling assumptions for $\beta_j$, rather than as 
-comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
-likelihood-based methods can be used to select $\alpha$. For example, following the logic of the Empirical Bayes approach it
-would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
-which has been implemented in our software by C. Dai. Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
-may prefer to choose whichever of $\alpha=0,1$ yields the highest likelihood, and this is also straightforward.
-
 %DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
@@
 
-\subsection{Null behaviour}
-
-Results of simulations under the null.
+%\subsection{Null behaviour}
+
+%Results of simulations under the null.
 
@@
 
-\subsection{Do we need a point mass at zero?}
-
-In some settings it is the convention to focus on testing whether $\beta_j=0$.
+%\subsection{Do we need a point mass at zero?}
+%
+%In some settings it is the convention to focus on testing whether $\beta_j=0$.
 However some dislike this focus, objecting that it is unlikely to be the case that $\beta_j=0$ exactly.
 For example, when comparing the average expression of a gene in human samples vs chimp samples, it might be considered unlikely that the expression
-is *exactly* the same in both. Whether or not $\beta_j=0$
-is considered unlikely may depend on the context.
-However, in most contexts, finite data cannot
-distinguish between $\beta_j=0$ and $\beta_j$ being very close to zero.
+%is *exactly* the same in both. Whether or not $\beta_j=0$
+%is considered unlikely may depend on the context.
+%However, in most contexts, finite data cannot
+%distinguish between $\beta_j=0$ and $\beta_j$ being very close to zero.
 Thus finite data cannot usually convince a skeptic that $\beta_j$ is exactly zero, rather than just very small.
@@
 In this sense there is an assymetry between the inferences "$\beta_j$ is zero" and "$\beta_j$ is non-zero", an assymetry that is reflected in the admonition "failure to reject the null does not imply it to be true".
-
-Thus any analysis that purports to distinguish between these cases must be making an assumption.
-
-Consider two analyses of the same data, using two different "priors" $g$ for $\beta_j$, that effectively differ only in their assumptions about whether or not $\beta_j$ can be exactly zero.
+%
+%Thus any analysis that purports to distinguish between these cases must be making an assumption. 
+%
+%Consider two analyses of the same data, using two different "priors" $g$ for $\beta_j$, that effectively differ only in their assumptions about whether or not $\beta_j$ can be exactly zero.
 For concreteness, consider
-\[ g_1(\cdot) = \pi \delta_0(\cdot) + (1-\pi) N(\cdot; 0,\sigma^2) \]
-and
-\[g_2(\cdot) = \pi N(\cdot; 0, \epsilon^2) + (1-\pi) N(\cdot; 0, \sigma^2).\]
-If $\epsilon^2$ is sufficiently small, then these 
-priors are "approximately the same", and will lead to "approximately the same" posteriors and inferences in many senses.
+%\[ g_1(\cdot) = \pi \delta_0(\cdot) + (1-\pi) N(\cdot; 0,\sigma^2) \]
+%and
+%\[g_2(\cdot) = \pi N(\cdot; 0, \epsilon^2) + (1-\pi) N(\cdot; 0, \sigma^2).\]
+%If $\epsilon^2$ is sufficiently small, then these 
+%priors are "approximately the same", and will lead to "approximately the same" posteriors and inferences in many senses.
 To discuss these, let $p_j$ denote the posterior under prior $g_j$.
@@
 However, we will not have $p_1(\beta_j=0) \approx p_2(\beta_j=0)$: the latter will always be zero, while the former could be appreciable.
-
- What if, instead, we examine $p_1(\beta_j >0)$ and $p_2(\beta_j >0)$?
+%
+% What if, instead, we examine $p_1(\beta_j >0)$ and $p_2(\beta_j >0)$?
 Again, these will differ.
@@
 This is because if $p_1(\beta_j>0)=1-\alpha$, then $p_1(\beta_j=0)$ will often be close to $\alpha$, so for small $\epsilon$ $p_2(\beta_j)$ will have mass $\alpha$ near 0, of which half will be positive and half will be negative.
-Thus if we do an analysis without a point mass, but allow
-for mass near 0, then we may predict what the results would have been if we had used a point mass.
-
-Let's try: 
-```{r}
-beta.ash.pm = ash(ss$betahat, ss$betasd, usePointMass=TRUE)
-print(beta.ash.pm)
-print(beta.ash.auto)
-plot(beta.ash.auto$localfsr,beta.ash.pm$localfsr,main="comparison of ash localfsr, with and without point mass",xlab="no point mass", ylab="with point mass",xlim=c(0,1),ylim=c(0,1))
-abline(a=0,b=1)
-abline(a=0,b=2)
-```
-
-Our conclusion: if we simulate data with a point mass,
-and we analyze it without a point mass, we may underestimate
-the lfsr by a factor of 2.
+%Thus if we do an analysis without a point mass, but allow
+%for mass near 0, then we may predict what the results would have been if we had used a point mass.
+%
+%Let's try: 
+%```{r}
+%beta.ash.pm = ash(ss$betahat, ss$betasd, usePointMass=TRUE)
+%print(beta.ash.pm)
+%print(beta.ash.auto)
+%plot(beta.ash.auto$localfsr,beta.ash.pm$localfsr,main="comparison of ash localfsr, with and without point mass",xlab="no point mass", ylab="with point mass",xlim=c(0,1),ylim=c(0,1))
+%abline(a=0,b=1)
+%abline(a=0,b=2)
+%```
+%
+%Our conclusion: if we simulate data with a point mass,
+%and we analyze it without a point mass, we may underestimate
+%the lfsr by a factor of 2.
 Therefore, to be conservative, we might prefer to analyze the data allowing for the point mass, or, if analyzed without a point mass, multiply estimated false sign rates by 2.
-In fact the latter might be preferable: even if we analyze the data with a point mass, there is going to be some unidentifiability
-that means estimating the pi value on the point mass will be somewhat unreliable, and we still might underestimate the false sign rate if we rely on that estimate.  
-TO THINK ABOUT: does multiplying the smaller of Pr(<0) and Pr(>0) by 2, and adding to Pr(=0) solve the problem in either case?
+%
 
@@
 we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
-
-\subsubsection*{Initialization}
-
-By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
-Our rationale to initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
-but weak signal in the data cannot quickly draw the algorithm towards the null.
 
@@
 %[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
+\subsubsection*{Initialization}
+
+By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
+Our rationale to initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
+but weak signal in the data cannot quickly draw the algorithm towards the null.
+
+In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
+$\pi_0=1$. If not then we replace the EM solution with the global null solution. (This helps guard against errors due to convergence to a local optimum when the
+data are consistent with the global null.)
 
@@
 We define the FDR for $\Gamma$ as the proportion of these tests that are ``false discoveries" in that $H_j$ is true:
-\begin{equation} \label{eqn:FDR}
+\begin{equation}
 \FDR(\Gamma; H) := \frac{\#\{j: H_j \cap j \in \Gamma\}}{\#\{j: j \in \Gamma\}},
@@
 %\end{equation}
-\begin{equation} \label{eqn:fdr}
+\begin{equation} 
 \fdr_j := \Pr(H_j | D).
@@
 
-%One standard genomics analysis pipeline is to take the $p$ values $p_1,\dots,p_J$ and use them to estimate $q$ values $q_1,\dots,q_J$
-%using the R package \qvalue.
- % Figure \ref{fig:qvalue} shows a graphical illustration of the procedure implemented in \qvalue. 
- %In brief, the idea is that the overall distribution of $p$ values can be decomposed into two constitutive
- %components: one component corresponding to the ``null" tests, and a second component corresponding to the ``alternative" tests. 
 To perform this decomposition, \qvalue makes two key assumptions:
@@
  natural non-parametric estimate of the FDR is
- \begin{equation} \label{eqn:FDRest}
+ \begin{equation} 
 \hat{\FDR}(\gamma) = J \hat\pi_0 \gamma / \#\{ j: p_j \leq \gamma \}.
@@
 for observation $j$ ($q_j$) as an estimate of the FDR for all observations that are at least as significant as $j$:
-\begin{equation} \label{eqn:qj}
+\begin{equation}
 q_j := \widehat\FDR(\{k:\fdr_k \leq \fdr_j\}).

### 2015-09-30 dsc-shrink/add_named_scenarios.R

@@ -8,56 +8,15 @@ gdef = list(spiky=normalmix(c(.4,.2,.2,.2),c(0,0,0,0),c(.25,.5,1,2)),
             near_normal=normalmix(c(2/3,1/3),c(0,0),c(1,2))
             )
 
-add_named_scenarios = function(dsc,names,nsamp=1000){
-
-#the -nn suffix indicates no nulls (min_pi0 = max_pi0 = 0)
-
-if("near-normal-nn" %in% names){
-	add_scenario(dsc,name="near-normal-nn",
-                    fn=rnormmix_datamaker,
-                    args=list(
-                      g=normalmix(c(2/3,1/3),c(0,0),c(1,2)),
-                      min_pi0=0,
-                      max_pi0=0,
-                      nsamp=nsamp,
-                      betahatsd=1
-                      ),
-                    seed=1:100)
-}
-
-if("flat-top-nn" %in% names){
-	add_scenario(dsc,name="flat-top-nn",
-                    fn=rnormmix_datamaker,
-                    args=list(
-                      g=normalmix(rep(1/7,7),c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7)),
-                      min_pi0=0,
-                      max_pi0=0,
-                      nsamp=nsamp,
-                      betahatsd=1
-                    ),
-                    seed=1:100)
-}
-
-if("skew-nn" %in% names){
-  add_scenario(dsc,name="skew-nn",
-                      fn=rnormmix_datamaker,
-                      args=list(
-                        g=normalmix(c(1/4,1/4,1/3,1/6),c(-2,-1,0,1),c(2,1.5,1,1)),
-                        min_pi0=0,
-                        max_pi0=0,
-                        nsamp=nsamp,
-                        betahatsd=1
-                      ),
-                      seed=1:100)
-}
+add_named_scenarios = function(dsc,names,nsamp=1000,min_pi0=0,max_pi0=1,suffix=""){
 
 if("near-normal" %in% names){
-  add_scenario(dsc,name="near-normal",
+  add_scenario(dsc,name=paste0("near-normal",suffix),
                     fn=rnormmix_datamaker,
                     args=list(
                       g=gdef$near_normal,
-                      min_pi0=0,
-                      max_pi0=1,
+                      min_pi0=min_pi0,
+                      max_pi0=max_pi0,
                       nsamp=nsamp,
                       betahatsd=1
[…]

### 2015-09-30 dsc-shrink/plot_lfsr.R

@@ -70,4 +70,12 @@ print(p1.s+theme(legend.position="none",axis.text.x = element_text(size = 8,angl
 ggsave("../paper/figures/lfsr_est_s.png",height=3,width=9)
 ggsave("../paper/figures/lfsr_est_s.pdf",height=3,width=9)
 
+lfsr.s.nn = process_score_for_plotting_against_gold(
+    res$lfsr,PLOTSEEDS=1:100,PLOTMETHODS="ash.n.s",
+    PLOTSCENARIOS=paste0(c("spiky","near-normal","flat-top","skew","big-normal","bimodal"),"-nn"))
+p1.s.nn=plot_lfsr(lfsr.s.nn,ylim=c(0,1),xlim=c(0,0.2))
+
+print(p1.s.nn+theme(legend.position="none",axis.text.x = element_text(size = 8,angle=45))
+      +coord_equal(ratio=1/5))
+ggsave("../paper/figures/lfsr_est_s.nn.png",height=3,width=9)
 

### 2015-09-30 dsc-shrink/run_dsc.R

@@ -2,7 +2,8 @@ library(dscr)
 
 dsc_shrink=new_dsc("shrink","dsc-shrink-files")
 source("add_named_scenarios.R")
-add_named_scenarios(dsc_shrink,c("spiky","near-normal","flat-top","skew","big-normal","bimodal"))
+add_named_scenarios(dsc_shrink,c("spiky","near-normal","flat-top","skew","big-normal","bimodal","near-normal-nn"))
+add_named_scenarios(dsc_shrink,c("spiky","near-normal","flat-top","skew","big-normal","bimodal"),min_pi0=0,max_pi0=0,suffix="-nn")
 
 source("add_methods.R")
 

### 2015-09-30 paper/Rcode/makefig_FDReg_hist.R

@@ -123,75 +123,3 @@ dev.off()
 
 
 
-################################### GOODPOOReg_hist.pdf ###############################
-
-#simple simulated example to illustrate high and low signal
-ntest = 10000
-set.seed(112)
-
-#simulate (0.5 N(0,1) + 0.5 delta_0)
-null_alt = rbinom(ntest,1,0.5)
-beta = rnorm(ntest)
-beta = ifelse(null_alt==1,beta,0)
-GOOD= 1:(ntest/2)
-sebetahat = rep(1,ntest)
-sebetahat[-GOOD] = 10
-
-betahat = beta + rnorm(ntest, 0, sebetahat)
-zscore = betahat/sebetahat
-pval = pchisq(zscore^2,df=1,lower.tail=F)
-
-pdf("../figures/GOODPOOReg_hist.pdf",width=6.5,height=3)
-par(mai=c(0.3,0.3,0.2,0.2),mgp = c(3, 0.5, 0))
-layout(matrix(1:3,ncol=3,byrow=TRUE))
-plot_FDReg_hist(pval[GOOD],1,type=1,title="Good-precision observations",ylab="",nc=20,cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
-plot_FDReg_hist(pval[-GOOD],1,type=1,title="Poor-precision observations",ylab="",nc=20,yaxt='n',cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
-axis(side=2, labels=FALSE,tck=-0.01)
-plot_FDReg_hist(pval,1,type=1,title="Combined",yaxt='n',ylab="",nc=20,cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
-axis(side=2, labels=FALSE,tck=-0.01)
-dev.off()
-
-################################### GOODPOOReg_scatterplot.pdf ###############################
-
-#apply the different methods
-res.qvalue = qvalue(pval)
-res.locfdr = locfdr(zscore,nulltype=0,plot=0)
-res.ash = ash(betahat,sebetahat,method="fdr")
-
-res.qvalue.good = qvalue(pval[GOOD])
-res.locfdr.good = locfdr(zscore[GOOD],nulltype=0,plot=0)
-res.ash.good = ash(betahat[GOOD],sebetahat[GOOD],method="fdr")
-
[…]

### 2015-09-30 paper/main.tex
simplify model description

@@
 existing widely-used FDR pipelines) and can substantially affect
-inference. The first idea is to {\it assume that the distribution of effects is unimodal}, and to perform
-non-parametric inference under this assumption. This yields a very simple, fast, and stable computer implementation,
- and improves inference of FDR, at least when the unimodal assumption is correct.  
+inference. The first idea is to {\it assume that the distribution of effects is unimodal}. This yields a very simple, fast, and stable computer 
+implementation, as well as improving inference of FDR when the unimodal assumption is correct.  
 The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
-to summarize each measurement. This idea allows variations in measurement precision to be accounted for,
- and avoids the problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
+to summarize each measurement. This idea allows variations in measurement precision to be better accounted for,
+ and avoids a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
 
@@
 
-The use of EB methods for estimation also has a long and close relationship with another key concept in modern statistics: shrinkage estimation \cite{efron1973}.
-This close relationship becomes especially intimate here due to our unimodal assumption on the underlying effects, 
-which encourages shrinkage towards the mode of the distribution.
-%Indeed, within our EB framework, multiple comparisons correction can be seen as a natural limit of shrinkage estimation:
-%if effects are distributed uni-modally about zero then it is natural to shrink estimates towards zero. 
-% are essentially two sides of the same coin, one related to testing and the other to estimation. 
-Thus, although we focus here primarily on FDR-related issues, another important contribution of our work is to provide {\it generic} and {\it adaptive} shrinkage estimation procedures.
-By generic, we mean that these methods can be applied in any setting where a series of effect estimates and corresponding standard errors are available.
-By adaptive, we 
-have two properties in mind. First, the appropriate amount of shrinkage is determined from the data: when the data indicate that large effects are relatively common then
-shrinkage of large observed effects is less than when the data indicate that large effects are rare.
-Second, the amount of shrinkage undergone by each measurement will depend on its precision, or standard error: measurements with big standard errors undergo more shrinkage than measurements with small standard errors.
-Shrinkage estimation is a powerful tool, with many potential applications, including for example wavelet denoising, 
-where shrinkage of wavelet coefficients is used to smooth signals. Our methods provide an attractive and competitive alternative
-to existing EB shrinkage methods for that context, as will be explored more fully elsewhere (Xing and Stephens, in preparation).
-
 Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
+(We address the reasons for this name, and connections with shrinkage analysis, in the Discussion.)
 
@@
 
- \subsection*{Outline}
+ \subsection*{Model Outine}
 
@@
 %mean expression level between the two conditions, often referred to as the ``log-fold change". 
-We begin by presenting the simplest version of method, and comparing it with existing work, before detailing further embellishments.
+We begin with the simplest version of our method, and compare it with existing work, before detailing embellishments.
 
@@
 that aims to identify differentially expressed genes,  $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
-In contexts where FDR methods are applied, interest often focuses on identifying significant non-zero effects;
-that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we will also tackle
-the more general problem of estimating, and assessing uncertainty in, the effect sizes $\beta_j$.
-
-We assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of these effects,
+In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
+that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
+more general problem of estimating, and assessing uncertainty in, $\beta_j$.
+
+Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
 and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
@@
 
-We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
-$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
-\begin{equation}
-p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
-\end{equation}
-Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
+For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
+This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
+A simple way to implement the UA is to assume that $g$ is
+a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
+\begin{align} \label{eqn:beta}
+p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
+g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
+\end{align}
+where $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
+Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyperparameters, which are non-negative and sum to one, and are to be estimated,
+while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
+represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
+(We encourate the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
+although of course in practice we use a finite grid -- see Implementation Details.)
+
+For the likelihood $p(\bhat | \beta, \shat)$ we assume  
 \begin{equation} \label{eqn:normlik}
-p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
-\end{equation}
-where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
-
-We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
-as a mixture of null and alternative components:
-\begin{equation} \label{eqn:betaj0}
-g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
-\end{equation}
-where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
-We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
-\begin{equation} \label{eqn:beta-iid}
-p(\beta | \shat) = \prod_j g(\beta_j).
-\end{equation}
-
-Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
-with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
-to use a mixture of zero-mean normal distributions:
-\begin{equation} \label{eqn:mixnorm}
-g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
-\end{equation}
-where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
-Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
-represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
- The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
- more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below.
-
- Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
- yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
- \begin{equation} \label{eqn:g}
- g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
- \end{equation}
-
- Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
- \begin{align} \label{eqn:betaj}
- p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
- p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
- \end{align}
- with $g$ given by the mixture of normals $(\ref{eqn:g})$.
-
-In words, the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
+p(\bhat | \beta, \shat) = \prod_j N(\bhat_j; \beta_j, \shat_j^2).
+\end{equation}
+Here, in additional to some conditional independence assumptions, we are 
+effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
+
+This simple model features both the key ideas we want to emphasise in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
+the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
+with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
+However, this simple model also has several additional assumptions that can be relaxed. Specifically,
+\begin{enumerate}
+\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) also implies that $g$ is symmetric about 0; more flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see (\ref{eqn:g-general})].
+\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
+We can relax this to allow for a relationship between these quantities [see (\ref{eqn:beta-alpha})].
+\item The likelihood (\ref{eqn:normlik}) effectively assumes that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
+We can generalize this likelihood using a $t$ likelihood [see (\ref{eqn:tlik})].
+\end{enumerate}
+Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
+(which are also made by most existing EB approaches to this problem). Correlations among tests certainly arise in practice, either due to genuine correlations
+in the system of study, or due to unmeasured confounders, and their potential to impact results of an FDR analysis is important to consider
+whatever analysis methods are used: see \cite{efron2007correlation,leek:2007} for relevant discussion.
+
+%We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
+%$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
+%\begin{equation}
+%p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
+%\end{equation}
+%Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
+%\begin{equation} \label{eqn:normlik}
+%p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
+%\end{equation}
+%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
+%
+%We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
+%as a mixture of null and alternative components:
+%\begin{equation} \label{eqn:betaj0}
+%g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
+%\end{equation}
+%where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
+%We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
+%\begin{equation} \label{eqn:beta-iid}
+%p(\beta | \shat) = \prod_j g(\beta_j).
+%\end{equation}
+%
+%Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
+%with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
+%to use a mixture of zero-mean normal distributions:
+%\begin{equation} \label{eqn:mixnorm}
+%g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
+%\end{equation}
+%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
+%Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
+%represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
+% The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
+% more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below. 
+% 
+% Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
+% yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
+% \begin{equation} \label{eqn:g}
+% g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
+% \end{equation}
+% 
+% Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
+% \begin{align} \label{eqn:betaj}
+% p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
+% p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
+% \end{align}
+% with $g$ given by the mixture of normals $(\ref{eqn:g})$.
+
+\subsection*{Fitting the model}
+
+In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
 each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
@@
 \begin{equation} \label{eqn:marginallik}
-p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)].
-\end{equation}
+p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
+\end{equation}
+where we define $\sigma_0:=0$.
 
@@
 In practice we make a small modification to this procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we instead maximise a penalized
-likelihood, where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data (see details below). 
+likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
 We introduce this penalty because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
@@
 
-Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:g}) implies that $g$ is not only unimodal, but also symmetric.
+Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
 Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
@@
 where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
-(In (iii) we include both components in the mixture (\ref{eqn:g}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
+(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
 and $\pi$ is a $2K+1$ vector that sums to 1.)
-The simplest version (\ref{eqn:g}) corresponds to (i).
+The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
 Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
@@
 
-Moving from (i) to (iii) the representation (\ref{eqn:g}) becomes increasingly flexible. Indeed,
+Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
 using a large dense grid of $\sigma^2_k$ or $a_k$, we can allow $g$ to approximate,
@@
 
-\subsubsection*{Replace normal likelihood with $t$ likelihood}
-
-We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
-  \begin{equation} \label{eqn:tlik}
- \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
- \end{equation}
-where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
-and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
-The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
-This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms.
-When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
-
-Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
-\begin{equation} \label{eqn:tdist}
-(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
-\end{equation}
-However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
-In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
-currently pursuing this improved approach in work with Mengyin Lu, and results will be published elsewhere. We note here that
-modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
-where our simpler approach can produce unsatisfactory results (see later).
-
 \subsubsection*{Dependence of effects on standard errors}
@@
 do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
+
+\subsubsection*{Replace normal likelihood with $t$ likelihood}
+
+We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
+  \begin{equation} \label{eqn:tlik}
+ \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
+ \end{equation}
+where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
+and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
+The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
+This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms.
+When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
+
+Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
+\begin{equation} \label{eqn:tdist}
+(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
+\end{equation}
+However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
+In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
+currently pursuing this improved approach in work with Mengyin Lu, and results will be published elsewhere. We note here that
+modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
+where our simpler approach can produce unsatisfactory results (see later).
 
@@
 
+The use of EB methods for estimation also has a long and close relationship with another key concept in modern statistics: shrinkage estimation \cite{efron1973}.
+This close relationship becomes especially intimate here due to our unimodal assumption on the underlying effects, 
+which encourages shrinkage towards the mode of the distribution.
+%Indeed, within our EB framework, multiple comparisons correction can be seen as a natural limit of shrinkage estimation:
+%if effects are distributed uni-modally about zero then it is natural to shrink estimates towards zero. 
+% are essentially two sides of the same coin, one related to testing and the other to estimation. 
+Thus, although we focus here primarily on FDR-related issues, another important contribution of our work is to provide {\it generic} and {\it adaptive} shrinkage estimation procedures.
+By generic, we mean that these methods can be applied in any setting where a series of effect estimates and corresponding standard errors are available.
+By adaptive, we 
+have two properties in mind. First, the appropriate amount of shrinkage is determined from the data: when the data indicate that large effects are relatively common then
+shrinkage of large observed effects is less than when the data indicate that large effects are rare.
+Second, the amount of shrinkage undergone by each measurement will depend on its precision, or standard error: measurements with big standard errors undergo more shrinkage than measurements with small standard errors.
+Shrinkage estimation is a powerful tool, with many potential applications, including for example wavelet denoising, 
+where shrinkage of wavelet coefficients is used to smooth signals. Our methods provide an attractive and competitive alternative
+to existing EB shrinkage methods for that context, as will be explored more fully elsewhere (Xing and Stephens, in preparation).
+
 Choice of mixture components. Use of likelihood to compare different mixture components?

### 2015-10-01 paper/main.tex
update paper and figure to v2 of qvalue

@@
 
-\def\lfdr{\text{lfdr}}
-\def\lfsr{\text{lfsr}}
+\def\lfdr{\textit{lfdr}}
+\def\lfsr{\textit{lfsr}}
 \def\x{\mbf{x}}
@@
 controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
-``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (lfdr) \cite{efron2008microarrays}, but which measures confidence  
+``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
 in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
-the lfsr and lfdr can be quite different, and emphasise benefits of the lfsr, particularly its increased robustness to modelling assumptions.
+the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the lfsr, particularly its increased robustness to modelling assumptions.
 
@@
 \def\FDR{\text{FDR}}
-\def\fdr{\text{lfdr}}
+\def\fdr{\text{\lfdr}}
 \def\FSR{\text{FSR}}
@@
 %mean expression level between the two conditions, often referred to as the ``log-fold change". 
-We begin with the simplest version of our method, and compare it with existing work, before detailing embellishments.
+Here we describe the simplest version of our method, and briefly discuss embellishments we have also implemented.
 
@@
 \end{enumerate}
+These embellishments are detailed in Embellishments and Implementation Details.
 Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
@@
 Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(Note that the simplicity of the EM algorithm in step 1 is due to our using a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
+(The simplicity of the EM algorithm in step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
 which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
@@
 
-The lfdr, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
+The \lfdr, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
 whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
@@
 \end{equation}
-To illustrate, suppose for concreteness
-that $p(\beta_j < 0| \bhat, s, \hat\pi))=0.95,  p(\beta_j =0| \bhat, s, \hat\pi)) = 0.03, p(\beta_j >0| \bhat, s, \hat\pi) = 0.02$. Then 
+To illustrate, suppose that 
+\begin{gather*}
+p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
+p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
+p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
+\end{gather*} 
+Then 
 from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
@@
 and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
-In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Section \ref{sec:lfsr},
+In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
 $\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.
@@
 
-Among previous methods that explicitly consider the estimation of False Discovery Rates and related quantities,
-our work here seems most naturally compared with the Empirical Bayes methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
+Among previous methods that explicitly consider the FDR and related quantities,
+our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
  (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
@@
 
-There are two key differences between the approach implemented here and all of these three existing methods. First, 
+There are two key differences between our approach and all of these three existing methods. First, 
 whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
-we instead work with two numbers ($\bhat_j,\shat_j$).
-Here we are building on \cite{wakefield:2009}, who develops Bayesian tests (Bayes Factors) for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
-Using two numbers instead of one clearly has the potential to be more informative, and indeed,
-as an example of this, Section \ref{good-poor-hist} illustrates how it can improve performance by taking better account of variation in measurement precision among observations.
-(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
+we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
+for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
+Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 
+how it can improve performance by taking better account of variation in measurement precision among observations.
+%(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
 % Or more generally Section xx illustrates the potential practical effects
@@
 
-Given that the UA and ZA are so different, it seems worth discussing why we favor the UA. 
+Given that the UA and ZA are so different, it seems worth discussing why we generally favor the UA. 
 Although the UA will not apply to all situations, we believe that it will often be reasonable, especially in FDR-related contexts
@@
 
+In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficents - 
+common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
+These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
+it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
+The UA assumption on regression coefficients is directly analagous to our UA here, and so its
+widespread use in the regression context supports its use here.
+
 Alternatively, we could motivate the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
@@
 make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).
-Further, almost all analogous work in sparse regression models make the UA for the regression coefficents - 
-common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
-These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
-it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
-The UA assumption on regression coefficients is directly analagous to our UA for effects, and so we view its
-widespread use in the regression context as supporting its use here.
-
-In addition to these arguments, the UA also has a considerable practical benefit: it produces very simple procedures that are both computationally
+
+Finally, the UA also has a considerable practical benefit: it yields simple algorithms that are both computationally
 and statistically stable. We illustrate these features in Results.
@@
 
-There is also a very considerable literature that does not directly focus on the FDR problem, but which involves essentially the same models,
-and deals with closely-related issues.  Among these, \cite{cordy1997deconvolution} is perhaps most similar to our work:
-indeed all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term, appear in their paper.
-However, they focus entirely on the estimation of $g$, in contrast to our focus on estimating $\beta_j$; 
-and they provide no software implementation.
-More generally, the literature is too large to do a comprehensive review here, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
+There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
+ Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
+indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
+This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
+Also, they provide no software implementation.
+More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
 Some pointers to recent papers in which other relevant
 citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}.
-Much (but not all) of the literature focusses on the homoskedastic case (i.e.
+Much of the literature focusses on the homoskedastic case (i.e.
 $\shat_j$ all equal) whereas we allow for heteroskedasticity.
-And much (but not all) of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential.
+And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential.
 Several recent papers consider more flexible non-parametric assumptions on
@@
 may perform well for point estimation, it seems possible it will not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
+Indeed, one way of thinking about the UA is as a way to regularize $g$.
 
@@
 %Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-\subsection*{Embellishments}
-
-\subsubsection*{More flexible unimodal distributions}
-
-Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
-Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
- Therefore, we have implemented a more general approach, based on 
-\begin{equation} \label{eqn:g-general}
-g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) 
-\end{equation}
-where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
-\begin{enumerate}[(i)]
-\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$,
-\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,
-\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,
-\end{enumerate}
-where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
-(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
-and $\pi$ is a $2K+1$ vector that sums to 1.)
-The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
-Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
-under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
-(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
-(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
-however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)
-
-Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
-using a large dense grid of $\sigma^2_k$ or $a_k$, we can allow $g$ to approximate,
-with arbitrary accuracy,
-\begin{enumerate}[(i)]
-\item any scale mixture of normals, which includes as special cases
-the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
-\item any symmetric unimodal distribution about 0.
-\item any unimodal distribution about 0.
-\end{enumerate}
-The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
-In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.
-
-Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
-The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
-See Section \ref{sec:grid} for our software defaults.
-
-%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting.
-For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
-%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
-
-\subsubsection*{Dependence of effects on standard errors}
-
-Equation (\ref{eqn:betaj}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
-This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
- \begin{equation} \label{eqn:beta-alpha}
- \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
- \end{equation}
-for any $\alpha \geq 0$. Setting $\alpha=0$ yields (\ref{eqn:betaj}), and setting $\alpha=1$ corresponds to
-assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
-it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
-the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.
-
-The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
-To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
- standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
- Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
-yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
-%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
-%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
-%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
-
-Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
-how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
-it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
-rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
-and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
-this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
-do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
-
-\subsubsection*{Replace normal likelihood with $t$ likelihood}
-
-We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
-  \begin{equation} \label{eqn:tlik}
- \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
- \end{equation}
-where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
-and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
-The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
-This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms.
-When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
-
-Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
-\begin{equation} \label{eqn:tdist}
-(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
-\end{equation}
-However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
-In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
-currently pursuing this improved approach in work with Mengyin Lu, and results will be published elsewhere. We note here that
-modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
-where our simpler approach can produce unsatisfactory results (see later).
-
-\subsubsection*{Non-zero mode}
-
-An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood.
-
-\subsection*{Implementation Details}
-
-See Section \ref{sec:implement} for further implementation details.
 
@@
 We compare results of \ashr with existing FDR-based methods implemented
-in the R packages \qvalue (v1.99.1 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
+in the R packages \qvalue (v2.0 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
 In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
@@
 
-Here we consider the effects of making the UA. In this section, to isolate these effects we consider the simplest case, where every observation has the same
+Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
 standard error, $s_j=1$ and all methods are provided that information. That is,
@@
 \begin{enumerate}
-\item The UA can produce very different inferences compared with the ZA.
-\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of \lfdr and \FDR.
+\item The UA can produce very different inferences compared with the ZA made by existing methods.
+\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
 \item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
@@
 \begin{figure}
-\center \includegraphics[height=6in]{Rcode/figures/decomp_ZA.pdf}
-\caption{Comparison of the way that different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (light blue).
+\center \includegraphics[height=6in]{figures/decomp_ZA.pdf}
+\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
+The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
 In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
 The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ashr) assumes that the effect sizes (and thus the $z$ scores) have a unimodal distribution about 0, resulting in a very different decomposition.
-(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero -- specifically, $\beta_j \sim N(0,1)$, $s_j=1$, resulting in $z_j =\bhat_j\sim N(0,2)$ -- so the ``true" decomposition would make everything light blue.)} \label{fig:ZA}
+In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
+(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
 \end{figure}
@@
 To illustrate the different inferences from the UA and ZA we show results for a single
-dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2^2)$). We used each of the methods
+dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
+but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
 \qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
@@
 
-The illustrative example in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
+The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
 Consequently \ashr will estimate smaller $\lfdr$s and FDRs than existing methods that make the ZA. 
 This is desirable, provided that these estimates remain conservative: that is,
-that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
+provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
 The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behaviour. To check its effectiveness
-we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects,
-$g_1$), and values
+we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
 for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
@@
 	\includegraphics[height=1.8in]{figures/scenarios_density.pdf} 
-	\caption{Densities $g_1$ used in simulations.} \label{fig:altdens}
+	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
 \end{subfigure}
@@
 \includegraphics[height=1.8in]{figures/lfdr_est.png} 
-\caption{Comparison of true and estimated \lfdr~from \ashr (ash.n).
+\caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
 Black line is $y=x$ and red line is $y=2x$.
-Estimates of \lfdr~are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
+Estimates of $\lfdr$ are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
 \end{subfigure}
@@
 \includegraphics[height=1.8in]{figures/lfsr_est.png} 
-\caption{As in c), but for \lfsr~instead of \lfdr. Estimates of \lfsr~are consistently less conservative than \lfdr~when UA holds, and also less anti-conservative in bimodal scenario.}
+\caption{As in c), but for $\lfsr$ instead of $\lfdr$.
+Estimates of $\lfsr$ are consistently less conservative than $\lfdr$ when UA holds, and also less anti-conservative in bimodal scenario.}  \label{fig:lfsr}
 \end{subfigure}
 \end{center}
-\caption{Results of simulation studies (constant precision $s_j=1$).} \label{fig:lfsr}
+\caption{Results of simulation studies (constant precision $s_j=1$).}
 \end{figure}
@@
 \end{enumerate}
-Thus these simulations assume the same
-precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).
+%Thus these simulations assume the same
+%precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).
 
@@
 We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
-are used. (In addition, the effects on \lfsr estimates turn out to be relatively modest; see below).
+are used. (In addition, the effects on $\lfsr$ estimates turn out to be relatively modest; see below).
 
@@
 
-\subsubsection*{The lfsr is more robust than lfdr}
+\subsubsection*{The $\lfsr$ is more robust than \lfdr}
 
 The results above show that \ashr can improve on existing methods in producing smaller, more accurate, 
-estimates of $\pi_0$ (and, consequently, more accurate estimates of False discovery rates).
+estimates of $\pi_0$, which will lead to more accurate estimates of FDR.
 Nonetheless, in many scenarios \ashr continues
 to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
-The explanation for this is that these scenarios include 
+This is because these scenarios include 
 an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
@@
 but not a lower bound (because the data cannot rule out that everything is non-null, but with miniscule effects).
-If we desire conservative behaviour we have to estimate $\pi_0$ by its upper bound, which can
+To obtain conservative behaviour we must estimate $\pi_0$ by this upper bound, which can
 be substantially larger than the true value.
+
 Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
 overestimation of $\pi_0$ is corresponding overestimation of FDR (and \lfdr, and $q$ values).
-To illustrate, Figure \ref{fig:lfdr} compares the estimated \lfdr~from ash.n with the true value (computed using
-Bayes rule from the true $g_1$ and $\pi0$). As predicted, \lfdr~is overestimated, especially in scenarios where 
-$\pi_0$ is overestimated. All methods are similarly affected by this, and the ones that most grossly overestimate $\pi_0$ also most grossly overestimated \lfdr and FDR/$q$-values (not shown).
+To illustrate, Figure \ref{fig:lfdr} compares the estimated $\lfdr$ from ash.n with the true value (computed using
+Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios where 
+$\pi_0$ is overestimated. All methods are similarly affected by this, and the ones that most grossly overestimate $\pi_0$ also most grossly overestimated $\lfdr$ and FDR/$q$-values (XXX).
 
 The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
- become substantially less troublesome if we use the local false sign rate \lfsr (Section \ref{}), rather than \lfdr, to measure significance. 
- This is essentially because \lfsr is less sensitive to the estimate of $\pi_0$.
-To illustrate, Figure \ref{fig:lfsr} compares the estimated \lfsr~from ash.n with the true value: although the estimated \lfsr continue to
-be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the \lfdr, especially for
-the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in \lfsr than \lfdr.
+ become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
+ This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
+To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
+be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
+the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in $\lfsr$ than $\lfdr$.
 
@@
 and so inferences are more robust to, say, the inevitable mis-estimation of $\pi_0$.
-To provide some intuition, consider an observation with a $z$ score of 0. The \lfdr of this observation can range from 0 (if $\pi_0=0$)
+To provide some intuition, consider an observation with a $z$ score of 0. The $\lfdr$ of this observation can range from 0 (if $\pi_0=0$)
 to 1 (if $\pi_0=1$). But, assuming a symmetric $g$, the $\lfsr>0.5$ whatever the value of $\pi_0$, because the observation $z=0$ says
 nothing about the sign of the effect.
-Thus, are two reasons to use the lfsr instead of the lfdr: it answers a question that is more generally meaningful (e.g. it applies
-whether or not zero effects truly exist),  and estimation of lfsr is more robust.
+Thus, are two reasons to use the $\lfsr$ instead of the \lfdr: it answers a question that is more generally meaningful (e.g. it applies
+whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.
 
@@
 that no effects are exactly 0. In  support of i), note that under all scenarios the cloud of estimates for $\lfsr$ roughly stop at the line $y=x$, suggesting that
-a reduction in stringency (e.g.~reducing the penalty term \ref{eqn:penalty}) would risk crossing this line more frequently.
+a reduction in stringency (e.g.~reducing the penalty term (\ref{eqn:penalty})) would risk crossing this line more frequently.
 
@@
 even for this largest difference results from the default run (e.g.
-the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood \url{github.com/stephens999/ash/dsc-robust/summarize_dsc_robust.rmd}.
+the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood.
+See \url{http://github.com/stephens999/ash/blob/master/dsc-robust/summarize_dsc_robust.rmd} for further details.
 
@@
 
-Although we have focussed on estimating the FDR, an important advantage of
-our EB approach based on modelling the effects $\beta_j$ (rather than $p$ values or $z$ scores)  is that it
+An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
 can estimate the {\it size} of each effect $\beta_j$.
@@
 depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
-which are notoriously difficult to solve in general. Indeed, Efron emphasises
+which are notoriously difficult in general. Indeed, Efron emphasises
 the difficulties of implementing a stable general algorithm, noting in his rejoinder
@@
 Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
-to be clearly visible in the observed $\bhat$, \mixfdr fits these modes (Supplementary Information; 
-\url{dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
+to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
+(\url{http://github.com/stephens999/ash/blob/master/dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
 clearly visible in the observed $\bhat$.
@@
 \caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
-Modest systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the modest effects of model mis-specification.}
+Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
 \end{subfigure}
@@
 This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero.
-Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (not shown).
+Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (Table \ref{tab:nopen}).
 Removing the penalty in the half-uniform case is not recommended (see Appendix).}
@@
 
-%\begin{table}[!ht]
-%\begin{subtable}{\textwidth}
-%	\input{tables/scoverage_all_nopen.tex}
-%	\caption{All observations}
-%\end{subtable}
-%
-%\begin{subtable}{\textwidth}
-%\input{tables/scoverage_neg_nopen.tex}
-%\caption{``Significant" negative discoveries.}
-%\end{subtable}
-%
-%\begin{subtable}{\textwidth}
-%\input{tables/scoverage_pos_nopen.tex}
-%\caption{``Significant" positive discoveries.}
-%\end{subtable}
-%
-%\caption{Table of empirical coverage for nominal 95\% lower credible bounds (methods without the penalty term. }
-%\end{table}
 %
@@
  This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
-the estimated error rates ($q$-values, or lfdr values) for the good-precision observations increase when the low-precision observations are included in the analysis
+the estimated error rates ($q$-values, or $\lfdr$ values) for the good-precision observations increase when the low-precision observations are included in the analysis
 (Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
@@
 a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the lfdr estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
+The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
 In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
@@
  with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
- We see that poor precision measurements are assigned a higher lfsr than good precision measuments that have the same $p$ value.
+ We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measuments that have the same $p$ value.
  The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
@@
 which model the $z_j$ or $p_j$ directly.  Under this assumption, and with the normal mixture prior for $\beta_j/s_j$ (\ref{eqn:mixnorm}),
-the \lfsr computed by \ashr provides the same {\it ranking} of observations as the $z$ scores and $p$ values (see Figure \ref{fig:lfsr_pval}, right panel).
+the $\lfsr$ computed by \ashr provides the same {\it ranking} of observations as the $z$ scores and $p$ values (see Figure \ref{fig:lfsr_pval}, right panel).
 
@@
 comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
-likelihood-based methods can be used to select $\alpha$. For example, following the logic of the Empirical Bayes approach it
+likelihood-based methods can be used to select $\alpha$. For example, following the logic of the EB approach it
 would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
@@
 
-%Most previous authors of empirical Bayes approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
+%Most previous authors of EB approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
 %have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
@@
 %and we analyze it without a point mass, we may underestimate
-%the lfsr by a factor of 2.
+%the $\lfsr$ by a factor of 2.
 Therefore, to be conservative, we might prefer to analyze the data allowing for the point mass, or, if analyzed without a point mass, multiply estimated false sign rates by 2.
@@
 
-\section*{Implementation details} \label{sec:implement}
+\subsection*{Embellishments and Implementation details} \label{sec:implement}
+
+\subsubsection*{More flexible unimodal distributions}
+
+Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
+Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
+ Therefore, we have implemented a more general approach, based on 
+\begin{equation} \label{eqn:g-general}
+g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) 
+\end{equation}
+where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
+\begin{enumerate}[(i)]
+\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$,
+\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,
+\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,
+\end{enumerate}
+where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
+(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
+and $\pi$ is a $2K+1$ vector that sums to 1.)
+The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
+Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
+under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
+(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
+(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
+however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)
+
+Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
+using a large dense grid of $\sigma^2_k$ or $a_k$, we can allow $g$ to approximate,
+with arbitrary accuracy,
+\begin{enumerate}[(i)]
+\item any scale mixture of normals, which includes as special cases
+the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
+\item any symmetric unimodal distribution about 0.
+\item any unimodal distribution about 0.
+\end{enumerate}
+The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
+In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.
+
+Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
+The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
+See Section \ref{sec:grid} for our software defaults.
+
+%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting.
+For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint.
+(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
+%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
+
+\subsubsection*{Dependence of effects on standard errors}
+
+Equation (\ref{eqn:betaj}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
+This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
+ \begin{equation} \label{eqn:beta-alpha}
+ \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
+ \end{equation}
+for any $\alpha \geq 0$. Setting $\alpha=0$ yields (\ref{eqn:betaj}), and setting $\alpha=1$ corresponds to
+assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
+it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
+the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.
+
+The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
+To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
+ standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
+ Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
+yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
+%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
+%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
+%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
+
+Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
+how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
+it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
+rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
+and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
+this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
+do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
+
+\subsubsection*{Replace normal likelihood with $t$ likelihood}
+
+We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
+  \begin{equation} \label{eqn:tlik}
+ \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
+ \end{equation}
+where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
+and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
+The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
+This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms.
+When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
+
+Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
+\begin{equation} \label{eqn:tdist}
+(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
+\end{equation}
+However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
+In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
+currently pursuing this improved approach in work with Mengyin Lu, and results will be published elsewhere. We note here that
+modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
+where our simpler approach can produce unsatisfactory results (see later).
+
+\subsubsection*{Non-zero mode}
+
+An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood.
 
@@
 this introduces non-identifiability between $\pi_k$ and $\pi_0$. In practice this is a major problem
-only if one focuses on false discovery rates rather than false sign rates: the \lfdr is sensitive to $\pi_0$ whereas
-the \lfsr is not. However, to make \lfdr (and \lfsr) estimates from our method ``conservative" we add a penalty 
-term $h(\pi;\lambda)$ to the likelihood to encourage over-estimation of $\pi_0$:
+only if one focuses on false discovery rates rather than false sign rates: the $\lfdr$ is sensitive to $\pi_0$ whereas
+the $\lfsr$ is not. However, to make $\lfdr$ (and $\lfsr$) estimates from our method ``conservative" we add a penalty 
+term $log(h(\pi;\lambda))$ to the log-likelihood $\log(L(\pi))$ to encourage over-estimation of $\pi_0$:
 \begin{equation} \label{eqn:penalty}
@@
 
+\begin{table}[!ht]
+\begin{subtable}{\textwidth}
+	\input{tables/scoverage_all_nopen.tex}
+	\caption{All observations}
+\end{subtable}
+
+\begin{subtable}{\textwidth}
+\input{tables/scoverage_neg_nopen.tex}
+\caption{``Significant" negative discoveries.}
+\end{subtable}
+
+\begin{subtable}{\textwidth}
+\input{tables/scoverage_pos_nopen.tex}
+\caption{``Significant" positive discoveries.}
+\end{subtable}
+
+\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{table:nopen}
+\end{table}
 %In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so

### 2015-10-01 paper/Rcode/makefig_FDReg_hist.R

@@ -36,6 +36,7 @@ temp=hist(pval,nclass=50)
 bin_fdr = res.qvalue$pi0/temp$density #compute local fdr in each bin of histogram
 qval_fdr = bin_fdr[as.numeric(cut(pval,temp$breaks))]
 qval_fdr = pmin(1,qval_fdr) #threshold at 1, so fdr never bigger than 1
+qval_fdr = qvalue::lfdr(pval)
 
 pdf("../figures/FDReg_hist.pdf",width=6,height=3)
 par(mai=c(0.5,0.5,0.2,0.2),mgp = c(3, 0.5, 0))

### 2015-10-05 paper/main.tex
refined discussion of p value prior

@@
 
-The \lfdr, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
+The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
 whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
@@
 Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
-overestimation of $\pi_0$ is corresponding overestimation of FDR (and \lfdr, and $q$ values).
+overestimation of $\pi_0$ is corresponding overestimation of FDR (and $\lfdr$, and $q$ values).
 To illustrate, Figure \ref{fig:lfdr} compares the estimated $\lfdr$ from ash.n with the true value (computed using
-Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios where 
-$\pi_0$ is overestimated. All methods are similarly affected by this, and the ones that most grossly overestimate $\pi_0$ also most grossly overestimated $\lfdr$ and FDR/$q$-values (XXX).
+Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios which involve many
+non-zero effects that are very near 0 (e.g.~the spiky scenario with $\pi_0$ small) where
+$\pi_0$ can be grossly overestimated.  
+(Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.)
 
@@
 
-We emphasise that, compared with previous debates regarding estimation vs testing, 
-this section advances an additional reason for favoring focussing on the sign of an effect
-rather than testing whether it is 0. As noted above, many authors have argued to focus on estimation 
-because it is implausible that effects are {\it exactly} 0. Here we add that {\it even if one believes
+Note that, compared with previous debates regarding testing, 
+this section advances an additional reason for focussing on the sign of the effect, rather than just testing whether it is 0. 
+In previous debates authors have argued against testing whether an effect is 0
+because it is {\it implausible that effects are exactly 0}. Here we add that {\it even if one believes
 that some effects may be exactly zero}, it is still better to focus on the sign, because generally {\it the data are more informative about that question}
@@
 nothing about the sign of the effect.
-Thus, are two reasons to use the $\lfsr$ instead of the \lfdr: it answers a question that is more generally meaningful (e.g. it applies
+Thus, are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
 whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.
 
-Finally for this section, we briefly consider the question of whether we could obtain better estimates of $\lfsr$ than are provided here. 
-We speculate that to do this we would either have to i) relax our relatively stringent requirement for conservative behaviour, or ii) be willing to make the assumption
-that no effects are exactly 0. In  support of i), note that under all scenarios the cloud of estimates for $\lfsr$ roughly stop at the line $y=x$, suggesting that
-a reduction in stringency (e.g.~reducing the penalty term (\ref{eqn:penalty})) would risk crossing this line more frequently.
-
-discussion on whether we really need a point mass.
+Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
+Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
+if we are prepared to assume that no effects are exactly zero, then removing the point mass 
+yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
+However, there is ``no free lunch":  if in fact some effects are exactly zero
+then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
+We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.
+
+\begin{figure}
+\begin{center}
+\begin{subfigure}{\textwidth}
+\includegraphics[height=1.8in]{figures/lfsr_est_s_nn.png} 
+\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
+Black line is $y=x$ and red line is $y=2x$.
+The results illustrate how estimates of $\lfsr$ can be more accurate in this case.
+That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
+\end{subfigure}
+\begin{subfigure}{\textwidth}
+\includegraphics[height=1.8in]{figures/lfsr_est_s.png} 
+\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
+Black line is $y=x$ and red line is $y=2x$.
+The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
+\end{subfigure}
+\end{center}
+\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
+\end{figure}
 
@@
 \end{equation}
-In this setting, the poor-precision measurements tell us very little, and any sane analysis should effectively ignore them. 
+In this setting, the poor-precision measurements $(s_j=10)$ tell us very little, and any sane analysis should effectively ignore them. 
  However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor}). This is because the poor measurements 
@@
 
-Another important consequence of accounting for varying
- measurement precision across units is that \ashr re-orders the significance of the observations compared
+Another consequence of accounting for differences in measurement precision is that \ashr may re-order the significance of the observations compared
  with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
@@
 
-This tendency to reorder the significance of observations, and specifically to downweight the significance of observations with low precision,
-is a consequence of an assumption
-that we have made up to now: that the distribution of the effect $\beta_j$ is independent of its standard error $s_j$ (equation (\ref{eqn:betaj})). 
-One way to relax this assumption is to allow $\beta_j$ to scale with $s_j^\alpha$ for some
-power $\alpha$; that is, to assume model (\ref{eqn:beta-alpha}). Setting $\alpha=0$ implies that $\beta_j$ is
+The potential for Bayesian analyses to re-order the significance of observations, and specifically to downweight imprecise observations,
+was previously discussed in \cite{guan.stephens.08}.
+However, Wakefield \cite{wakefield:2009} showed that, under a certain prior assumption, the Bayesian analysis produces the same ranking of significance as $p$ values (or their $z$ scores).
+He named this prior the ``$p$-value prior" because it can be thought of as the
+implicit prior assumption that is being made if we rank the significance of observations by their $p$ value.
+Wakefield's $p$-value prior assumes that the less precise effect estimates correspond to larger true effects, and specifically that they scale proportional
+to the standard errors $s_j$. More specifically still, it assumes a normal prior for the non-zero $\beta_j$ with mean 0 and variance $K s^2_j$ for some constant $K$.
+Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. Specifically, if, instead of assuming that $\beta_j$ is independent
+of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
+then the $\lfsr$ computed by ash.n provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
+as is illustrated in Figure \ref{fig:lfsr_pval}, right panel. (This result does not hold when using the mixtures of uniforms prior, ash.u.)
+
+This $p$-value prior assumes that the $z$ scores $z_j=\beta_j/s_j$
+are identically distributed, independent of $s_j$. This is essentially the assumption made, implicitly or explicitly, by existing methods -- like \locfdr, \mixfdr and \qvalue --
+that model the $z_j$ or $p_j$ directly.  In contrast, we have assumed up to now that the $\beta_j$ are independent of $s_j$. 
+We can set both these assumptions within a more general framework, which allows that $\beta_j/s^\alpha$ is independent of $s_j$ for some $\alpha$ [Equation (\ref{eqn:beta-alpha}))].
+Setting $\alpha=0$ implies that $\beta_j$ is
 independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
@@
 so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
-The special case $\alpha=1$ corresponds to assuming that the $z$ scores $z_j=\beta_j/s_j$, and hence the corresponding $p$ values, 
-are identically distributed, independent of $s_j$. This is essentially the assumption made (implicitly or explicitly) by existing methods like \locfdr, \mixfdr and \qvalue,
-which model the $z_j$ or $p_j$ directly.  Under this assumption, and with the normal mixture prior for $\beta_j/s_j$ (\ref{eqn:mixnorm}),
-the $\lfsr$ computed by \ashr provides the same {\it ranking} of observations as the $z$ scores and $p$ values (see Figure \ref{fig:lfsr_pval}, right panel).
-
-This observation extends a result in
-\cite{wakefield:2009}, who showed that, when testing a series of null hypotheses, 
-with a normal likelihood, Bayes Factors and $p$ values produce the same ranking of tests {\it if} the Bayes Factor for the $j$th test 
-is computed using a normal prior for the alternative effects with mean 0 and variance $K s^2_j$ for some constant $K$.
-Because of this correspondance in ranking, Wakefield refers to this as the ``$p$-value prior". 
-Our result extends Wakefield's results to the mixture of normals prior, from a single normal. As an aside we note
-that the result does depend on the normal assumptions: for example, it does not hold in general for mixtures of uniforms as a prior; nor
-if we use a $t$ likelihood instead of a normal likelihood. That is, under these models, the EB
-analysis may reorder the significance of the observations (compared with the $p$ values) even when $\alpha=1$.
-
-%\begin{result}
-%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
-%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
-%\end{result}
-
-%Indeed, any FDR-based approach that ranks the significance of observations 
-%by their $p$ values, like the Benjamini--Hochberg method, could be viewed as making the same implicit assumption.
-
-To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
-Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
-In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
-Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
-which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
+On the other hand, there is no particular reason to expect that either $\alpha=1$ or $\alpha=0$ will be the optimal choice.
+Indeed, optimal choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$,
+which will be dataset-specific. 
 Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
@@
 which has been implemented in our software by C. Dai.
-(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
-may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)
+
+%\begin{result}
+%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
+%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
+%\end{result}
+
+%To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
+%Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
+%In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
+%Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
+%which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
+%(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
+%may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)
 
@@
 
-Here we build on this idea in two ways. First,
+Our methods we build on this idea in two ways. First,
 by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
@@
 
-%\subsection{Do we need a point mass at zero?}
-%
-%In some settings it is the convention to focus on testing whether $\beta_j=0$.
-However some dislike this focus, objecting that it is unlikely to be the case that $\beta_j=0$ exactly.
-For example, when comparing the average expression of a gene in human samples vs chimp samples, it might be considered unlikely that the expression
-%is *exactly* the same in both. Whether or not $\beta_j=0$
-%is considered unlikely may depend on the context.
-%However, in most contexts, finite data cannot
-%distinguish between $\beta_j=0$ and $\beta_j$ being very close to zero.
-Thus finite data cannot usually convince a skeptic that $\beta_j$ is exactly zero, rather than just very small.
-In contrast it is easy to imagine data that would convince a doubter that $\beta_j$ is truly non-zero.
-In this sense there is an assymetry between the inferences "$\beta_j$ is zero" and "$\beta_j$ is non-zero", an assymetry that is reflected in the admonition "failure to reject the null does not imply it to be true".
-%
-%Thus any analysis that purports to distinguish between these cases must be making an assumption. 
-%
-%Consider two analyses of the same data, using two different "priors" $g$ for $\beta_j$, that effectively differ only in their assumptions about whether or not $\beta_j$ can be exactly zero.
-For concreteness, consider
-%\[ g_1(\cdot) = \pi \delta_0(\cdot) + (1-\pi) N(\cdot; 0,\sigma^2) \]
-%and
-%\[g_2(\cdot) = \pi N(\cdot; 0, \epsilon^2) + (1-\pi) N(\cdot; 0, \sigma^2).\]
-%If $\epsilon^2$ is sufficiently small, then these 
-%priors are "approximately the same", and will lead to "approximately the same" posteriors and inferences in many senses.
-To discuss these, let $p_j$ denote the posterior under prior $g_j$.
-Then, for any given (small) $\delta$, we will have $p_1(|\beta_j|<\delta) \approx p_2(|\beta_j|< \delta)$.
-However, we will not have $p_1(\beta_j=0) \approx p_2(\beta_j=0)$: the latter will always be zero, while the former could be appreciable.
-%
-% What if, instead, we examine $p_1(\beta_j >0)$ and $p_2(\beta_j >0)$?
-Again, these will differ.
-If this probability is big in the first analysis, say $1-\alpha$ with $\alpha$ small, then it could be as big as $1-\alpha/2$ in the second analysis.
-This is because if $p_1(\beta_j>0)=1-\alpha$, then $p_1(\beta_j=0)$ will often be close to $\alpha$, so for small $\epsilon$ $p_2(\beta_j)$ will have mass $\alpha$ near 0, of which half will be positive and half will be negative.
-%Thus if we do an analysis without a point mass, but allow
-%for mass near 0, then we may predict what the results would have been if we had used a point mass.
-%
-%Let's try: 
-%```{r}
-%beta.ash.pm = ash(ss$betahat, ss$betasd, usePointMass=TRUE)
-%print(beta.ash.pm)
-%print(beta.ash.auto)
-%plot(beta.ash.auto$localfsr,beta.ash.pm$localfsr,main="comparison of ash localfsr, with and without point mass",xlab="no point mass", ylab="with point mass",xlim=c(0,1),ylim=c(0,1))
-%abline(a=0,b=1)
-%abline(a=0,b=2)
-%```
-%
-%Our conclusion: if we simulate data with a point mass,
-%and we analyze it without a point mass, we may underestimate
-%the $\lfsr$ by a factor of 2.
-Therefore, to be conservative, we might prefer to analyze the data allowing for the point mass, or, if analyzed without a point mass, multiply estimated false sign rates by 2.
-%
-
-\subsection{Discussion}
+\section*{Discussion}
+
+Benefits of two numbers instead of one.  A general idea.
 
@@
 
-Equation (\ref{eqn:betaj}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
+Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
 This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
@@
  \end{equation}
-for any $\alpha \geq 0$. Setting $\alpha=0$ yields (\ref{eqn:betaj}), and setting $\alpha=1$ corresponds to
+for any $\alpha \geq 0$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
 assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:

### 2015-10-05 dsc-shrink/plot_lfsr.R

@@ -77,5 +77,5 @@ p1.s.nn=plot_lfsr(lfsr.s.nn,ylim=c(0,1),xlim=c(0,0.2))
 
 print(p1.s.nn+theme(legend.position="none",axis.text.x = element_text(size = 8,angle=45))
       +coord_equal(ratio=1/5))
-ggsave("../paper/figures/lfsr_est_s.nn.png",height=3,width=9)
+ggsave("../paper/figures/lfsr_est_s_nn.png",height=3,width=9)
 

### 2015-10-06 examples/adaptive_shrinkage.rmd
finish refining discussion and implementation

@@
 
+Now plot lfsr against z scores, colored
+according to the (square root of) precision
+
+```{r plotting lfsr}
+
+z_lfsr_plot=function(df,ylab="Observed Z score",xlab="lfsr"){
+  ggplot(df,aes(x=lfsr,y=betahat/s,color=1/s)) + 
+    xlab(xlab) +
+    ylab(ylab) + geom_point() +  facet_grid(.~scenario) + scale_colour_gradient2(midpoint = median(1/s),low="blue", mid="white",high="red",space="Lab") 
+}
+
+z_lfsr_plot(df)
+ggsave("z_vs_lfsr.pdf")
+```
+
 A related consequence is that significance of each observation is no longer monotonic with $p$ value.
@@
 the standardized beta are exchangeable (rather than the beta being exchangeable).
-```{r run-ash-ES}
-beta.bignormal.ash.ES = ash(sim.bignormal$input$betahat, s,model="ES",mixcompdist="normal")
-beta.spiky.ash.ES = ash(sim.spiky$input$betahat, s,model="ES",mixcompdist="normal")
-df.ES = make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash.ES,beta.bignormal.ash.ES)
-ashplot(df.ES,ylab="Shrunken beta estimate (ES model)")
-pval_plot(df.ES)
+```{r run-ash-ET}
+beta.bignormal.ash.ET = ash(sim.bignormal$input$betahat, s,model="ET",mixcompdist="normal")
+beta.spiky.ash.ET = ash(sim.spiky$input$betahat, s,model="ET",mixcompdist="normal")
+df.ET = make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash.ET,beta.bignormal.ash.ET)
+ashplot(df.ET,ylab="Shrunken beta estimate (ET model)")
+pval_plot(df.ET)
 ```

### 2015-10-06 paper/main.tex
finish refining discussion and implementation

@@
 
-In practice we make a small modification to this procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we instead maximise a penalized
+Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximise a penalized
 likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
@@
 \centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
-\caption{} \label{fig:lfsr_pval}
+\caption{Figure illustrating affects of prior assumptions on re-ordering of significance.
+Left panel shows results under our ``default prior" which assumes that effects $\beta_j$ are identically distributed, independent of $s_j$.
+Right panel shows results under the ``$p$-value prior", which assumes that $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$.} \label{fig:lfsr_pval}
 \end{figure}
 
-\subsubsection*{Adaptive Shrinkage}
-
-Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
-when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
-the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
- $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
- In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
- Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
- and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
-
-Our methods we build on this idea in two ways. First,
-by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
-Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
-adapts to measurement precision: more precise measurements undergo
- less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
-than observations with
-small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
-To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
-Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=400px]{../examples/ash.pdf} 
-\end{center}
-\caption{Figure illustrating adaptive shrinkage.
+%\subsubsection*{Adaptive Shrinkage}
+%
+%Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
+%when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
+%the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
+% $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
+% In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
+% Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
+% and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
+%
+%Our methods we build on this idea in two ways. First,
+%by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
+%Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
+%adapts to measurement precision: more precise measurements undergo
+% less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
+%than observations with
+%small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
+%To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
+%Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
+%
+%\begin{figure}[!ht]
+%\begin{center}
+%\includegraphics[width=400px]{../examples/ash.pdf} 
+%\end{center}
+%\caption{Figure illustrating adaptive shrinkage.
 Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0.
@@
 The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
-\end{figure}
+%\end{figure}
 
@@
 
-Benefits of two numbers instead of one.  A general idea.
-
-The use of EB methods for estimation also has a long and close relationship with another key concept in modern statistics: shrinkage estimation \cite{efron1973}.
-This close relationship becomes especially intimate here due to our unimodal assumption on the underlying effects, 
-which encourages shrinkage towards the mode of the distribution.
-%Indeed, within our EB framework, multiple comparisons correction can be seen as a natural limit of shrinkage estimation:
-%if effects are distributed uni-modally about zero then it is natural to shrink estimates towards zero. 
-% are essentially two sides of the same coin, one related to testing and the other to estimation. 
-Thus, although we focus here primarily on FDR-related issues, another important contribution of our work is to provide {\it generic} and {\it adaptive} shrinkage estimation procedures.
-By generic, we mean that these methods can be applied in any setting where a series of effect estimates and corresponding standard errors are available.
-By adaptive, we 
-have two properties in mind. First, the appropriate amount of shrinkage is determined from the data: when the data indicate that large effects are relatively common then
-shrinkage of large observed effects is less than when the data indicate that large effects are rare.
-Second, the amount of shrinkage undergone by each measurement will depend on its precision, or standard error: measurements with big standard errors undergo more shrinkage than measurements with small standard errors.
-Shrinkage estimation is a powerful tool, with many potential applications, including for example wavelet denoising, 
-where shrinkage of wavelet coefficients is used to smooth signals. Our methods provide an attractive and competitive alternative
-to existing EB shrinkage methods for that context, as will be explored more fully elsewhere (Xing and Stephens, in preparation).
-
-Choice of mixture components. Use of likelihood to compare different mixture components?
-
- An important focus of Efron's work, also implemented in \locfdr, 
- is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
-theoretical null distribution. Although important, this issue is largely orthogonal to the issues we focus on here, which arise
-even if the theoretical null distribution holds precisely in the empirical data, and so we consider it just briefly in the Discussion.
-
- noting that the $p$ values in Figure \ref{} were actually generated under a scenario where
-{\it no tests were null}! Thus the qvalue procedure, which estimates the proportion of nulls to be xx\%, is acting very conservatively in this case.
-However, it could be argued that some degree of conservativeness is inevitable, and that nonetheless the procedure is acting sensibly.
-Therefore, to illustrate this point we really want to focus on the fact that, at least in some settings, the assumption is {\it unrealistic}.
-Specifically, we want to argue that the implied distribution of the alternative $p$ values is unrealistic. In fact, this is hard to see in $p$ value space,
-and so instead we translate these $p$ values back to the $z$ scores that lead to these $p$ values.
-
-- it allows us to easily just vary sigma on a grid, and fit pi, which
-makes allowing different noise levels really easy! In contrast, incorporating differential errors messes up the EM algorithm if
-we try to estimate $\sigma_k$.
-
-Directly modeling the p values, or z scores, say via non-parametric methods, 
-can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-About benefits of generic approach.
- In outline, our goal is to perform Bayesian inference for the true effects $\beta$ using information in the summary statistics $\bhat,\shat$.
- That is we aim to compute the joint posterior distribution $p(\beta | \bhat,\shat)$.
-This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
-from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
-with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
- By working with summary statistics $\bhat, \shat$, rather than more refined data, we aim to produce generic methods that 
- can be applied whenever such summary data are available - just as \qvalue can be applied to any set of $p$ values for example.
-Any attempt to produce generic methods is likely to involve a compromise between functionality and generality; 
-we believe that by working with two numbers ($\bhat_j,\shat_j$) for each observation, rather than one ($p_j$ or $z_j$), 
-we can gain substantially in functionality (e.g.~ we can estimate effect sizes, as well as testing, and we can take better account of variations in
-measurement precision across units $j$) while losing only a little in generality.
-
-Before proceeding, we feel obliged to correct the common misconception
- that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
- not depend on the number of tests. This is because, as the number of tests increases, both the true positives and false positives increase linearly,
- and the FDR remains the same. (If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.)
-  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
- if there are lots of strong signals then the FDR at a given threshold may be low, even if a large number of tests were performed; and conversely
- if there are no strong signals then the FDR at the same threshold may be high, even if relatively few tests were performed. 
- In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
-
- Note on multiple comparisons: it isn't really a "problem" but an "opportunity". This viewpoint also espoused by \cite{greenland1991empirical}. It isn't the number of tests that is relevant (the false
-discovery rate at a given threshold does not depend on the number of tests). It is the *results* of the tests that are relevant.
-Performing multiple comparisons, or multiple tests, is often
-regarded as a "problem". However, here we regard it instead as an opportunity - an opportunity to combine (or "pool") information across tests or comparisons.
-Focussing on the number of tests performed can be
-seen as an approximation.
-
-Here performing inference means obtaining an (approximate) posterior distribution for each effect $\beta_j$,
-which can be used to estimate False Discovery Rates, and to produce both point and interval estimates for $\beta_j$.
- Although ``multiple comparisons" is usually presented as a ``problem", our emphasis here is that the multiple measurements
- actually provide an opportunity: the hierarchical model combines information across measurements to improve both accuracy and precision of estimates.
-
- There are two issues with existing approaches to FDR estimation and control that I would like to address here. 
- The first is that they may do not take proper account of differences in measurement precision across units (i.e. differences in $\shat_j$). 
- The second is that the Zero Assumption, although initially appealing as a ``conservative" assumption,
-  is often an unrealistic and unnecessarily conservative assumption. As we shall see, both factors can cause the FDR to be overestimated.
-
-\cite{efron2008microarrays} states the Zero Assumption as the assumption that ``most of the z-values near zero come from null genes".
-His main aim in making this assumption is to estimate an empirical null though (not assume N(0,1) for the null)
-rather than to impose identifiability.
-
-Note that \cite{muralidharan2010empirical} models $z$ scores as something plus noise under both $H_0$ and $H_1$, which
-avoids this problem. (DOes the same maybe apply to modeling beta, rather than z scores, when the errors vary?)
-
-Rice and Spiegelhalter - BRCA data?
-
-A fundamental idea is
-that the measurements of $\beta_j$ for each gene can be used to improve inference for the values of $\beta$ for other genes.
-
---- NOte 
-In principle we might prefer a full Bayes approach that accounts for uncertainty in $\pi$ (or, more generally, in $g$); however,
-we believe that in most practical applications uncertainty in $g$ will not be the most important concern, and 
-compromise this principle for the simplicity of the EB approach.
-
-\subsection*{Embellishments and Implementation details} \label{sec:implement}
+We have presented an Empirical Bayes approach to large-scale multiple testing that emphasises two ideas.
+First, we emphasise the potential benefits of using two numbers ($\bhat$, and its standard error)
+rather than just one number (a $p$ value or $z$ score)  to summarize the information on each test.
+While requiring two numbers is slightly more onorous than requiring one, in many settings these
+numbers are easily available and if so we argue it makes sense to use them.
+Second, we note the potential benefits -- both statistical and computational -- of assuming that the effects come from a unimodal distribution, and provide
+flexible implementations for performing inference under this assumption. We also introduce the ``false sign rate" as
+an alternative measure of error to the FDR, and illustrate its improved robustness to errors in model fit, particularly mis-estimation
+of the proportion of null tests, $\pi_0$.
+
+Multiple testing is often referred to as a ``problem" or a ``burden". In our opinion, EB approaches turn this idea on its head,
+treating multiple testing as an {\it opportunity}: specifically, an opportunity to learn about the prior distributions, and other modelling assumptions,
+to improve inference and make informed decisions about significance (see also \cite{greenland1991empirical}). This view
+also emphasises that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
+Indeed, the FDR at a given fixed threshold does not depend on the number of tests: as the number of tests increases, both the true positives and false positives increase linearly,
+and the FDR remains the same.
+(If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example.
+The EB approach captures this dependence in an intuitive way:
+if there are lots of strong signals then we infer $\pi_0$ to be small, and the estimated FDR (or $\lfdr$, or $\lfsr$) at a given threshold may be low, even if a large number of tests were performed; and conversely
+if there are no strong signals then we infer $\pi_0$ to be large and the FDR at the same threshold may be high, even if relatively few tests were performed. 
+More generally, overall signal strength is reflected in the estimated $g$, which in turn influences the estimated FDR.
+
+Two important practical issues that we have not addressed here are correlations among tests,
+and the potential for deviations from the theoretical null distributions of test statistics. These two
+issues are connected: specifically, unmeasured confounding factors can cause both correlations among tests
+and deviations from the theoretical null \cite{efron2007correlation,leek:2007}. And although there are certainly other
+factors that could cause dependence among tests, unmeasured confounders are perhaps the most worrisome in practice
+because they can induce strong correlations among large numbers of tests and profoundly impact results.
+Approaches to deal with unmeasured confounders can be largely divided into two types: those
+that simply attempt to correct for the resulting inflation of test statistics \cite{devlin1999genomic,efron2004large}, and those
+that attempt to infer counfounders using clustering, principal components analysis, or factor models 
+\cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
+and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable,
+they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alterntively, our methods
+could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:lik}) using
+$\bhat_j | \shat_j \sim N(0,\lambda_1 \shat_j + \lambda_2)$, where $\lambda_1,\lambda_2$ are to be estimated. However,
+this immediately raises issues of identifiability and we do not pursue the idea further here.
+
+Another important practical issue is the challenge of small sample sizes. For example, in genomics applications researchers sometimes
+attempt to identify differences between two conditions based on only a handful of samples in each. In such settings the
+normal likelihood approximation (\ref{eqn:normlik}) will be inadequate. And, although the $t$ likelihood (\ref{eqn:tlik}) partially
+addresses this issue, it is also, it turns out, not entirely satisfactory.  The root of the problem is that, with small sample sizes, raw estimated standard
+errors $\shat_j$ can be horribly variable. In genomics it is routine to address this issue by applying EB methods \cite{smyth:2004} to ``moderate" (i.e.~shrink) variance estimates,
+before computing $p$ values from ``moderated" test statistics. We are currently investigating how
+our methods should incorporate such ``moderated" variance estimates to make it applicable to small sample settings.
+
+Our approach involves compromises between flexibility, generality, and simplicity on the one hand, and statistical efficiency and principle on the other.
+For example, in using an EB approach that uses a point estimate for $g$, rather than a fully Bayes approach that accounts for uncertainty in $g$,
+we have opted for simplicity over statistical principle. And in summarizing every test by
+two numbers and making a normal or $t$ approximation to the likelihood, we have aimed to produce generic methods that 
+ can be applied whenever such summary data are available -- just as \qvalue can be applied to any set of $p$ values for example -- although
+ possibly at the expense of statistical efficiency compared with developing multiple tailored approaches based on context-specific likelihoods. 
+ Any attempt to produce generic methods will involve compromise between generality and efficiency.
+ In genomics, many analyses -- not only FDR-based analyses -- involve first computing a series of $p$ values before
+ subjecting them to some further downstream analysis. An important message here is 
+ that working with two numbers ($\bhat_j,\shat_j$), rather than one ($p_j$ or $z_j$), 
+ can yield substantial gains in functionality (e.g.~ estimating effect sizes, as well as testing; accounting for variations in
+measurement precision across units) while losing only a little in generality.  We hope that our work will
+encourate development of methods that exploit this idea in other contexts.
+
+%This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
+%from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
+%with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
+
+%
+%
+%Both \locfdr and \mixfdr allow for deviations from
+%the theoretical null distribution by allowing that, under the null, the $z$ scores are $N(0,\sigma^2)$ with $\sigma$ estimated from the data rather
+%than fixed to 1. Within our approach this idea could be incorporated by changing the likelihood - for example, to assume $\bhat_j \sim N(0,s_ 
+%
+%These are tricky issues that, as far as we are aware, can affect
+%all approaches to FDR analysis. An important focus of Efron's work, also implemented in \locfdr, 
+% is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
+%theoretical null distribution. 
+%
+%For example, we and others have worked on Bayesian approaches to genetic association testing, using subjective prior specification of
+%effect sizes based on external data (e.g.~\cite{stephens:2009}). However, the large numbers of tests brings the opportunity to learn about
+%priors, and here we aim to exploit this
+
+%Choice of mixture components. Use of likelihood to compare different mixture components?
+
+%  
+%Directly modeling the p values, or z scores, say via non-parametric methods, 
+%can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
+%distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
+%The solution is to model beta hat as a convolution of some distribution g and an error component.
+
+%Before proceeding, we feel obliged to correct the common misconception
+% that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
+% not depend on the number of tests. 
+%  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
+% 
+% In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
+
+\section*{Detailed Methods} \label{sec:implement}
+
+\subsection*{Embellishments}
 
@@
 Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
- Therefore, we have implemented a more general approach, based on 
+ Therefore, we have implemented a more general approach based on 
 \begin{equation} \label{eqn:g-general}
-g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) 
+g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) ,
 \end{equation}
@@
 Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
-using a large dense grid of $\sigma^2_k$ or $a_k$, we can allow $g$ to approximate,
+using a large dense grid of $\sigma^2_k$ or $a_k$, (i)-(iii) can respectively approximate,
 with arbitrary accuracy,
@@
 The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
-See Section \ref{sec:grid} for our software defaults.
+See Implementation for details of our software defaults.
 
@@
  \end{equation}
-for any $\alpha \geq 0$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
+for any $\alpha$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
 assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
@@
 
-Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
-how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
-it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
-rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
-and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
-this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
-do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
+%Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
+%how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
+%it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
+%rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
+%and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
+%this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
+%do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
 
@@
 The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
-This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms.
+This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms; see Implementation below.
 When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
@@
 In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
-currently pursuing this improved approach in work with Mengyin Lu, and results will be published elsewhere. We note here that
-modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
-where our simpler approach can produce unsatisfactory results (see later).
+currently pursuing this improved approach (joint work with M.Lu) and results will be published elsewhere. 
+%We note here that
+%modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
+%where our simpler approach can produce unsatisfactory results.
 
@@
 
-An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood.
+An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood, rather than fixed to 0.
+
+\subsection*{Implementation Details}
+
+\subsubsection*{Likelihood for $\pi$}
+
+We define the likelihood for $\pi$ to be the probability of the observed data $\bhat$ conditional on $\shat$: $L(\pi) := p(\bhat | \shat,\pi),$ which
+by our conditional indendence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$.
+[One might prefer to define the likelihood as $p(\bhat, \shat | \pi) = p(\bhat | \shat, \pi) p(\shat | \pi)$, in which case our definition comes down to assuming that the term $p(\shat | \pi)$ does not depend on $\pi$.]
+
+Using the prior $\beta_j \sim \sum_{k=0}^K \pi_k f_k(\beta_j)$ given by (\ref{eqn:g-general}), and the normal likelihood (\ref{eqn:normlik}), integrating over $\beta_j$ yields
+\begin{equation}
+p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
+\end{equation}
+where 
+\begin{equation}
+\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
+\end{equation}
+denotes the convolution of $f_k$ with a normal density.
+These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
+Specifically, 
+\begin{equation} \label{eqn:uconv}
+\tilde{f}_k(\bhat_j)  = 
+\begin{cases}
+N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
+\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
+\end{cases}
+\end{equation}
+where $\Psi$ denotes the cumulative distribution function (c.d.f.) of the standard normal distribution.
+If we replace the normal likelihood with the $t_\nu$ likelihood (\ref{eqn:tlik}) then the convolution 
+for $f_k$ uniform the convolution is still given by (\ref{eqn:uconv}) but with $\Psi$ the c.d.f. 
+of the $t_\nu$ distribution function. (The convolution for $f_k$ normal is tricky and we have not implemented it.)
+
+\subsubsection*{Penalty term on $\pi$}
+
+To make $\lfdr$ and $\lfsr$ estimates from our method ``conservative" we add a penalty 
+term $log(h(\pi;\lambda))$ to the log-likelihood $\log L(\pi)$ to encourage over-estimation of $\pi_0$:
+\begin{equation} \label{eqn:penalty}
+h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
+\end{equation}
+where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
+consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0sims}).
+
+Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
+we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
+
+\subsubsection*{EM algorithm}
+
+With this in place, the penalized log-likelihood for $\pi$ is given by:
+\begin{equation}
+\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
+\end{equation}
+where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This can be maximized using an EM algorithm \cite{dempster77}, whose
+one-step updates are:
+\begin{align} \label{eqn:w}
+w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
+n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
+\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}
+\end{align}
+Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
+and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
+probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$. (The $w_{kj}$ are sometimes referred to as the ``responsibilities".)
+Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
+We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple} to accelerate convergence of this EM algorithm.
+
+%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
+\subsubsection*{Initialization}
+
+By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
+Our rationale for initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
+but weak signal in the data cannot quickly draw the algorithm towards the null.
+
+In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
+$\pi_0=1; \pi_k=0 (k>0)$. If not then we replace the EM solution with the global null solution. This aims to guard against errors due to convergence to a local optimum when the
+data are consistent with the global null.
+
+\subsubsection*{Conditional distributions}
+
+Given $\hat\pi$, we compute the conditional distributions 
+\begin{equation}
+p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
+\end{equation} 
+Each posterior is a mixture on $K+1$ components:
+\begin{equation}
+p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
+\end{equation}
+where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
+and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
+$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
+All these posterior distributions are easily available.
+For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
+If $f_k$ is normal and $L$ is normal, then this is a normal distribution.
 
@@
 a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.
-
-\subsubsection*{Penalty term on $\pi$}
-
-In practice data do not distinguish between effects that are ``very small" and ``exactly zero".
-Thus, if a mixture component $f_k$ has most mass very near zero then
-this introduces non-identifiability between $\pi_k$ and $\pi_0$. In practice this is a major problem
-only if one focuses on false discovery rates rather than false sign rates: the $\lfdr$ is sensitive to $\pi_0$ whereas
-the $\lfsr$ is not. However, to make $\lfdr$ (and $\lfsr$) estimates from our method ``conservative" we add a penalty 
-term $log(h(\pi;\lambda))$ to the log-likelihood $\log(L(\pi))$ to encourage over-estimation of $\pi_0$:
-\begin{equation} \label{eqn:penalty}
-h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
-\end{equation}
-where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
-consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0}).
-
-Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
-we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
-
-\subsubsection*{Likelihood for $\pi$ and EM algorithm}
-
-To formally derive the likelihood we restate our modelling assumptions more formally.
-We provide details first for the normal likelihood and then briefly describe modifications for the $t$ likelihood.
-
-We treat the standard errors as $\shat=(\shat_1,\dots,\shat_J)$ as given, and perform all inference conditional on $\shat$.
-We assume that given $\shat$ and the hyper-parameters $\pi$, the $(\beta_j,\bhat_j)$ pairs are independent 
-\begin{align} \label{eqn:b1}
-p(\beta_j \, | \, \shat,\pi) & = \sum_{k=0}^K \pi_k f_k(\beta_j) \\ \label{eqn:b2}
-p(\bhat_j \, | \, \beta_j, \shat, \pi)  & = N(\bhat_j; \beta_j, \shat^2_j),
-\end{align}
-where for notational convenience we have replaced the point mass $\delta_0$ with $f_0$.
-
-Multiplying (\ref{eqn:b1}) and (\ref{eqn:b2}) together gives $p(\bhat_j, \beta_j | \shat,\pi)$, which integrating over $\beta_j$ yields
-\begin{equation}
-p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
-\end{equation}
-where 
-\begin{equation}
-\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
-\end{equation}
-denotes the convolution of $f_k$ with the normal density.
-These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
-Specifically, 
-\begin{equation}
-\tilde{f}_k(\bhat_j)  = 
-\begin{cases}
-N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
-\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
-\end{cases}
-\end{equation}
-where $\Psi$ denotes the distribution function of the standard normal distribution.
-If we use a $t_\nu$ likelihood instead of a normal likelihood then the convolution is tricky for $f_k$ normal and
-we have not implemented it; for $f_k$ uniform the result (\ref{eqn:uconv}) holds but with the standard normal distribution function replaced
-with the $t_\nu$ distribution function.
-
-With this in place, the likelihood for $\pi$ is obtained by multiplying across $j$:
-\begin{equation}
-L(\pi) = \prod_j \sum_k \pi_k w_{kj}
-\end{equation}
-where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. 
-We add to this likelihood a penalty term $h(\pi;\lambda)$ above,
-and use an EM algorithm to maximize $L(\pi)+h(\pi;\lambda)$. The one-step updates for this EM algorithm are:
-\begin{align} \label{eqn:w}
-w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
-n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{E Step}
-\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{M step}
-\end{align}
-Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
-and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
-probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$.
-Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
-We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple}) to accelerate convergence of this EM algorithm.
-
-%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
-\subsubsection*{Initialization}
-
-By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
-Our rationale to initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
-but weak signal in the data cannot quickly draw the algorithm towards the null.
-
-In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
-$\pi_0=1$. If not then we replace the EM solution with the global null solution. (This helps guard against errors due to convergence to a local optimum when the
-data are consistent with the global null.)
-
-\subsubsection*{Conditional distributions}
-
-Given $\hat\pi$, we compute the conditional distributions 
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
-\end{equation} 
-Each posterior is a mixture on $K+1$ components:
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
-\end{equation}
-where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
-and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
-$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
-All these posterior distributions are easily available.
-For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
-If $f_k$ is normal and $L$ is normal, then this is a normal distribution.
 
@@
 
-\section{Definitions of FDR-related concepts used here}
+\section{Summary of FDR-related concepts}
 
 Let $\Gamma$ denote a subset of the tests $1,\dots,J$ declared ``significant" by some procedure. 
-We define the FDR for $\Gamma$ as the proportion of these tests that are ``false discoveries" in that $H_j$ is true:
+Define the FDR for $\Gamma$ as the proportion of these tests that are ``false discoveries" in that $H_j$ is true:
 \begin{equation}
@@
 the expectation of $\FDR(\Gamma;H)$ where the expecation is taken over the sampling distribution of $\Gamma$ (which is a function of the procedure
-used, considered fixed, and the data $D$, considered random). Bayesian approaches, in contrast, condition on the observed data, and compute
+used, considered fixed, and the data $D$, considered random). This expectation may of course depend on $H$ which is unknown, and the goal
+of a frequentist procedure is usually to control the $\FDR$ for all possible values of $H$ (``strong control") or specifically when all the $H$ are null (``weak control").
+
+Bayesian approaches, in contrast, condition on the observed data, and compute
 the posterior distribution for the $H_j$ given the data, which induces a posterior distribution on $\FDR(\Gamma)$. This posterior

### 2015-10-06 paper/main.tex
remove appendix on FDR

@@
 
-See Appendix for a more
-thorough discussion of the connection between the Bayesian quantities we compute here and frequentist quantities, including  
-a brief discussion of the False Sign Rate from a frequentist perspective.
-
 %Because our methods involve explicitly modelling effects $\beta_j$, rather than 
@@
 \begin{enumerate}[(i)]
-\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$,
-\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,
-\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,
+\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$, \qquad (``ash.n")
+\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,  \qquad (``ash.u")
+\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,  \qquad (``ash.hu")
 \end{enumerate}
@@
 we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
+
+\subsubsection*{Problems with removing the penalty term in the half-uniform case}
+
+It is straightforward to remove the penalty term by setting $\lambda_k=1$ in (\ref{eqn:penalty}).
+We note here an unanticipated problem we came across when using no penalty term in the half-uniform case 
+(i.e.~ $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$ in (\ref{eqn:g-general})): when the data
+are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
+sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
+for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
+and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
 
@@
 
-\section{Summary of FDR-related concepts}
-
-Let $\Gamma$ denote a subset of the tests $1,\dots,J$ declared ``significant" by some procedure. 
-Define the FDR for $\Gamma$ as the proportion of these tests that are ``false discoveries" in that $H_j$ is true:
-\begin{equation}
-\FDR(\Gamma; H) := \frac{\#\{j: H_j \cap j \in \Gamma\}}{\#\{j: j \in \Gamma\}},
-\end{equation}
-taken to be 0 if both numerator and denominator are 0. Frequentist approaches to controlling the FDR attempt to control
-the expectation of $\FDR(\Gamma;H)$ where the expecation is taken over the sampling distribution of $\Gamma$ (which is a function of the procedure
-used, considered fixed, and the data $D$, considered random). This expectation may of course depend on $H$ which is unknown, and the goal
-of a frequentist procedure is usually to control the $\FDR$ for all possible values of $H$ (``strong control") or specifically when all the $H$ are null (``weak control").
-
-Bayesian approaches, in contrast, condition on the observed data, and compute
-the posterior distribution for the $H_j$ given the data, which induces a posterior distribution on $\FDR(\Gamma)$. This posterior
-distribution can be used to obtain a point estimate for $\FDR(\Gamma)$, for example using the posterior mean. 
-To be more explicit, Bayesian (and EB) approaches give the posterior probability that $H_j$ holds, which
-analagous to \cite{efron2008microarrays} we refer to as the ``local FDR", and denote $\fdr_j$:
-%\begin{equation} \label{eqn:FDR}
-%\FDR(T) := \Pr(\text{$H_j$ false} | t_j \geq T).
-%\end{equation}
-\begin{equation} 
-\fdr_j := \Pr(H_j | D).
-\end{equation}
-The FDR can be estimated by the posterior mean for $\FDR(\Gamma)$, given by
-\begin{equation} 
-\widehat\FDR(\Gamma) := \frac{\sum_{j \in \Gamma} \fdr_j }{\#\{j: j \in \Gamma\}}.
-\end{equation}
-Although our approach here is (Empirical) Bayesian, one nice feature of FDR procedures is that 
-Frequentist and Bayesian procedures are often closely aligned, at least under certain assumptions; see
-for example \cite{storey.03}, particularly his Theorem 1.
-
-The $q$ value for observation $j$ is defined \cite{storey.03} as, roughly, the FDR for the set of observations that are at least as significant as observation $j$.
-In a Bayesian analysis it is natural to order the significance of the hypotheses by $\fdr_j$, so we define
-
-\subsubsection*{qvalue}
-
-To perform this decomposition, \qvalue makes two key assumptions:
-  i) the null $p$ values have a uniform distribution, and ii) the $p$ values near 1 are all from
- tests where the null is true. Under these assumptions performing the decomposition boils down to 
- drawing the horizontal line on the $p$ value histogram that intersects with the distribution at $p=1$: the area under this horizontal
- line defines the uniform component that is due to the null $p$ values, with the remainder necessarily corresponding to
- the alternative $p$ values. In particular, the height of the horizontal line is an estimate of $\pi_0$, the proportion of $p$ values that come from the null.
-  Now, for any given threshold, although we do not know {\it which} $p$ values correspond
- to  null tests  and which to alternative tests, the areas of the two components indicate {\it how many} $p$ values correspond to each,
- allowing the FDR to be estimated (Figure \ref{fig:qvalue}). Specifically, at threshold $\gamma$, and given an estimate $\hat\pi_0$ for $\pi_0$,  a
- natural non-parametric estimate of the FDR is
- \begin{equation} 
-\hat{\FDR}(\gamma) = J \hat\pi_0 \gamma / \#\{ j: p_j \leq \gamma \}.
- \end{equation}
- See \cite{storey.02}, equation (8).
-
-Frequentist approaches to FDR often concentrate on the {\it average} error rate for a subset of observations
-whose significance exceeds some threshold. Bayesian versions of such quantities can also be computed.
-For example, the FDR for any subset of effects $\Gamma \subset \{1,\dots,J\}$ can be estimated by
-\begin{equation} \label{eqn:FDRhat}
-\widehat\FDR(\Gamma) := \frac{\sum_{j \in \Gamma} \fdr_j }{\#\{j: j \in \Gamma\}}.
-\end{equation}
-And, analogous to \cite{storey.03}, we can define the $q$ value 
-for observation $j$ ($q_j$) as an estimate of the FDR for all observations that are at least as significant as $j$:
-\begin{equation}
-q_j := \widehat\FDR(\{k:\fdr_k \leq \fdr_j\}).
-\end{equation}
-
-Since the FDR terminology has become so widespread, we here introduce terminology for these ideas that parallels FDR terminology.
-Recall that, in FDR parlance, a ``discovery" is an effect that is declared to be non-zero; and a ``false discovery" refers to any such declaration that is
-made in error. The FDR is the (expected) proportion of false discoveries among all discoveries.
-Consider now insisting that for every discovery, we {\it also declare the sign of the effect} (positive or negative).
-We call such a declaration a ``signed discovery"; and correspondingly a ``false signed discovery" is any such declaration that is
-made in error - that is, any signed discovery that is either 0, or whose sign is opposite to that declared.
-Based on these definitions we can define the ``false signed discovery rate" (or ``false sign rate", FSR, for short) as the (expected) proportion
-of false signed discoveries among all signed discoveries. Similarly, we can define the local false sign rate, $\lfsr_j$ to be
-the probability that observation $j$ would be a false signed discovery, were we to declare it a signed discovery. Or, equivalently, but perhaps
-more simply, the $\lfsr_j$ is the probability we would get the sign of $\beta_j$ wrong if we
-were to be forced to declare it either positive or negative.  Explicitly:
-
-%(Naturally, we consider it an error to declare $\beta$ to be positive or negative when it is truly zero.)
-By analogy with (\ref{eqn:FDRhat}) we can also define an estimate of the average False Sign Rate for any subset of effects,
-\begin{equation}  \label{eqn:FSRhat}
-\widehat\FSR(\Gamma) := \frac{\sum_{j \in \Gamma} \fsr_j }{\#\{j: j \in \Gamma\}}.
-\end{equation}
-And, by analogy with (\ref{eqn:qj}) we can define the ``$s$-value" for each observation
-\begin{equation} \label{eqn:svalue}
-s_j := \widehat\FSR(\{k:\fsr_k \leq \fsr_j\})
-\end{equation}
-being the average error rate for all observations that are at least as significant as $j$.
-
 \begin{table}[!ht]
 \begin{subtable}{\textwidth}
-	\input{tables/scoverage_all_nopen.tex}
+	\centering\input{tables/scoverage_all_nopen.tex}
 	\caption{All observations}
@@
 \begin{subtable}{\textwidth}
-\input{tables/scoverage_neg_nopen.tex}
+\centering\input{tables/scoverage_neg_nopen.tex}
 \caption{``Significant" negative discoveries.}
@@
 \begin{subtable}{\textwidth}
-\input{tables/scoverage_pos_nopen.tex}
+\centering\input{tables/scoverage_pos_nopen.tex}
 \caption{``Significant" positive discoveries.}
@@
 %which is close to the definition in \cite{storey.03}.
-\section*{Problems with removing the penalty term in the half-uniform case}
-
-We note here an unanticipated problem we came across when using no penalty term in the half-uniform case: when the data
-are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
-sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
-for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
-and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
 
@@
 git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} , and includes contributions 
-from Chaixing (Rick) Dai, Mengyin Lu, and Tian Sen.
-
-This work was supported by NIH grant xxx and a grant from the Gordon and Betty Moore Foundation.
+from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.
+
+This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.
 

### 2015-10-06 For_each_simulation_.tex
New Article!

--- 2015-10-06 paper/main.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
-\documentclass[11pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-%\usepackage{natbib}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-\usepackage{booktabs}
-\usepackage{xspace}
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-\usepackage{sectsty}
-\subsectionfont{\normalsize}
-\subsubsectionfont{\mdseries\itshape}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\usepackage{color}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-% Use the PLoS provided bibtex style
-\bibliographystyle{/Users/stephens/Dropbox/Documents/stylefiles/plos2009}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
-\pagestyle{myheadings}
-%% ** EDIT HERE **
-\usepackage{enumerate}
-\usepackage{multirow} 
-\usepackage{url}
-\usepackage{xr} %for cross-referencing
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-\newtheorem{algorithm}{Algorithm}
-\newtheorem{proposition}{Proposition}
-\newtheorem{restateproposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{result}{Result}
-\newtheorem{note}{Note}
-\newtheorem{definition}{Definition}
-
-\def\lfdr{\textit{lfdr}}
-\def\lfsr{\textit{lfsr}}
-\def\x{\mbf{x}}
-\def\y{\mbf{y}}
-\def\e{\mbf{e}}
-\def\g{\mbf{g}}
-\def\p{p}
-\def\Yb{\hat{Y}_1^{\text{BAYES}}}
-\def\Xperp{X_{1 \perp 0}}
-\def\Pperp{P_{1 \perp 0}}
-\def\Pperpb{\Pperp^{\text{B}}}
-\def \hatd { \widehat{ d } }
-\def \dhatin {\hatd^{\mbox{\tiny IN}}}
-\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
-\def \mvn{\text{MVN}}
-\def\mn{\text{MN}}
-\def \iw{\text{W}^{-1}}
-\def \fp {fastPHASE\ }
-\def\bhat{\hat{\beta}}
-\def\shat{\hat{s}}
-\def\var{\rm Var}
-\def\mvr{\text{BMVR}}
-\def\bfuni{\text{BF}_\text{uni}}
-\def\one{\mbf{1}}
-\def\zero{\mbf{0}}
-\def\hall{H_{\text{all}}}
-\def\pall{\p_{\text{all}}}
-\def\bfall{\text{BF}_\text{all}}
-\def\bfav{\text{BF}_\text{av}}
-\def\gall{\gamma_\text{all}}
-\def\limbfall{\bfall^\rightarrow}
-\def\limbfav{\bfav^\rightarrow}
-\def\limbfgamma{\BF_\gamma^\rightarrow}
-\def\sxx{S_{xx}}
-\def\syy{S_{yy}}
-\def\vyy{V_{yy}}
-\def\vyx{V_{yx}}
-\def\vxx{{V_{xx}}}
-\def\rss{\text{\rm RSS}}
-\def\dss{\text{\rm $\Delta$SS}}
-\def\U{U}
-\def\D{D}
-\def\I{I}
-\def\luu{L_{UU}}
-\def\indep{\perp}
-\def\qvalue{{\tt qvalue}\xspace}
-\def\locfdr{{\tt locfdr}\xspace}
-\def\mixfdr{{\tt mixfdr}\xspace}
-\def\ashr{{\tt ashr}\xspace}
-
-\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
-                                % mode too
-\newcommand\height{{\it height}}
-\newcommand\weight{{\it weight}}
-\def\BF{\text{BF}}
-
-%% END MACROS SECTION
-%\externaldocument{plos_one_rev1_suppinfo}
-
-\begin{document}
-
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{False Discovery Rates: A New Deal}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Matthew Stephens$^{1*}$, 
-\\
-\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
-\\
-$\ast$ E-mail: Corresponding mstephens@uchicago.edu
-\end{flushleft}
-
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-%\section*{Author Summary}
-
-\section*{Introduction}
-
-Since its introduction in 1995 by Benjamini and Hochberg 
-\cite{benjamini1995controlling}, the ``False Discovery Rate" (FDR) has quickly established itself
-as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing
-in which the goal is to identify the non-zero ``effects" among a large number of imprecisely-measured effects.
-
-Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
-indeed, the notion that EB approaches could be helpful in handling multiple comparisons predates introduction of
-the FDR  (e.g.~\cite{greenland1991empirical}). More recently, EB approaches to the FDR
-have been extensively studied by several authors, 
-especially B.~Efron and co-authors 
-\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
-see also \cite{kendziorski2003parametric,newton2004detecting, 
-datta2005empirical,muralidharan2010empirical} for example.
-So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
-existing widely-used FDR pipelines) and can substantially affect
-inference. The first idea is to {\it assume that the distribution of effects is unimodal}. This yields a very simple, fast, and stable computer 
-implementation, as well as improving inference of FDR when the unimodal assumption is correct.  
-The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
-to summarize each measurement. This idea allows variations in measurement precision to be better accounted for,
- and avoids a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
-
-In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
-the idea that it may be preferable to focus on estimation rather than on testing.
-In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
-simply estimation with some positive prior probability that the effect is exactly zero.
-However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
-perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
-controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
-``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
-in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
-the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the lfsr, particularly its increased robustness to modelling assumptions.
-
-%terminology
-%``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
-%and correspondingly define the ``false signed discovery rate", or simply ``false sign rate" (FSR), in a natural way.
-
-%The methods we provide here can address both problems, and we provide results that highlight important benefits of focussing on estimation, and particularly on assessing the correctness of the {\it sign} of the estimated effects, rather than on testing for non-zero effects.
-
-Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
-(We address the reasons for this name, and connections with shrinkage analysis, in the Discussion.)
-
-%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.
-
-%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-%measurements are made on only a small number of
-%samples in each condition - sometimes as few as one
-%sample in each condition. Thus the error in estimates
-%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-%in estimates of $\beta_j$ still greater.
-%
-
-%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
-%Here we present new tools that bring two important ideas to bear on the problem.
-%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
-%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
-%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
-%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
-%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
-%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
-%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
-%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
-%a substantial increase in the number of discoveries at a given FDR).
-
-%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
-%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
-Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
-%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
-%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
-%generic methods that work directly with two numbers for each 
-%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
-Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
-%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
-
-\def\df{df}
-\def\FDR{\text{FDR}}
-\def\fdr{\text{\lfdr}}
-\def\FSR{\text{FSR}}
-\def\fsr{\text{lfsr}}
-
- \section*{Methods}
-
- \subsection*{Model Outine}
-
-% Here we consider one setting in which FDR-based methods are widely used:
-%an experimenter has measured a large numbers of ``effects", each with an associated
-%measurement precision (standard error), 
-%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
-%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
-%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
-%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
-%For example, in genomics
-%a common goal is to identify ``differentially-expressed genes" -- that is, genes
-%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
-%mean expression level between the two conditions, often referred to as the ``log-fold change". 
-Here we describe the simplest version of our method, and briefly discuss embellishments we have also implemented.
-
-Suppose that we are interested in the values of $J$ ``effects" $\beta=(\beta_1,\dots,\beta_J)$. For example, in a typical genomics application
-that aims to identify differentially expressed genes,  $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
-In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
-that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
-more general problem of estimating, and assessing uncertainty in, $\beta_j$.
-
-Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
-and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
-which by Bayes theorem can be written as 
-\begin{equation}
-p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
-\end{equation}
-
-For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
-This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
-A simple way to implement the UA is to assume that $g$ is
-a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
-\begin{align} \label{eqn:beta}
-p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
-g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
-\end{align}
-where $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyperparameters, which are non-negative and sum to one, and are to be estimated,
-while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
-represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
-(We encourate the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
-although of course in practice we use a finite grid -- see Implementation Details.)
-
-For the likelihood $p(\bhat | \beta, \shat)$ we assume  
-\begin{equation} \label{eqn:normlik}
-p(\bhat | \beta, \shat) = \prod_j N(\bhat_j; \beta_j, \shat_j^2).
-\end{equation}
-Here, in additional to some conditional independence assumptions, we are 
-effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-
-This simple model features both the key ideas we want to emphasise in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
-the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
-with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
-However, this simple model also has several additional assumptions that can be relaxed. Specifically,
-\begin{enumerate}
-\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) also implies that $g$ is symmetric about 0; more flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see (\ref{eqn:g-general})].
-\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
-We can relax this to allow for a relationship between these quantities [see (\ref{eqn:beta-alpha})].
-\item The likelihood (\ref{eqn:normlik}) effectively assumes that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-We can generalize this likelihood using a $t$ likelihood [see (\ref{eqn:tlik})].
-\end{enumerate}
-These embellishments are detailed in Embellishments and Implementation Details.
-Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
-(which are also made by most existing EB approaches to this problem). Correlations among tests certainly arise in practice, either due to genuine correlations
-in the system of study, or due to unmeasured confounders, and their potential to impact results of an FDR analysis is important to consider
-whatever analysis methods are used: see \cite{efron2007correlation,leek:2007} for relevant discussion.
-
-%We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
-%$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
-%\begin{equation}
-%p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
-%\end{equation}
-%Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
-%\begin{equation} \label{eqn:normlik}
-%p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
-%
-%We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
-%as a mixture of null and alternative components:
-%\begin{equation} \label{eqn:betaj0}
-%g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
-%\end{equation}
-%where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
-%We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
-%\begin{equation} \label{eqn:beta-iid}
-%p(\beta | \shat) = \prod_j g(\beta_j).
-%\end{equation}
-%
-%Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
-%with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
-%to use a mixture of zero-mean normal distributions:
-%\begin{equation} \label{eqn:mixnorm}
-%g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
-%Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
-%represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
-% The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
-% more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below. 
-% 
-% Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
-% yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
-% \begin{equation} \label{eqn:g}
-% g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
-% \end{equation}
-% 
-% Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
-% \begin{align} \label{eqn:betaj}
-% p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
-% p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
-% \end{align}
-% with $g$ given by the mixture of normals $(\ref{eqn:g})$.
-
-\subsection*{Fitting the model}
-
-In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
-each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
-Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
-\begin{equation} \label{eqn:marginallik}
-p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
-\end{equation}
-where we define $\sigma_0:=0$.
-
-The usual EB approach to fitting this model would involve two simple steps:
-\begin{enumerate}
-\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
-\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
-against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | | \bhat, \shat,\hat{\pi})$.
-\end{enumerate}
-Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(The simplicity of the EM algorithm in step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
-which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
-
-Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximise a penalized
-likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
-We introduce this penalty because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
-
-\subsection*{The local False Discovery Rate and local False Sign Rate}
-
-As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
-In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
-One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
-the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
-In other words it is the posterior probability that $\beta_j$ is actually zero:
-\begin{equation} \label{eqn:fdr}
-\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
-\end{equation}
-
-The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
-whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
-the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey1991philosophy}) argued that
-``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$.
-Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
-\begin{quote}
-...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
-\end{quote}
-Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
-focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.
-
-Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
-would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
-\begin{equation} \label{eqn:lfsr}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
-\end{equation}
-To illustrate, suppose that 
-\begin{gather*}
-p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
-p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
-p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
-\end{gather*} 
-Then 
-from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
-our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.
-
-As our notation suggests, $\lfsr_j$ is intended to be compared and contrasted with $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
-small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
-Of coure, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
-this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
-(this follows from the definition because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
-In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
-$\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.
-
-From these ``local" measures of significance, we can also compute average error rates over subsets of observations
-$\Gamma \subset \{1,\dots,J\}$. For example,   
-\begin{equation} \label{eqn:FDRhat}
-\widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
-\end{equation}
-estimates the FDR we would obtain if we were to declare all tests in $\Gamma$ significant.
-And 
-\begin{equation} \label{eqn:qvalue}
-q_j := \widehat \FDR(\{k: \fdr_k \leq \fdr_j\}) 
-\end{equation}
-provides a measure of significance analogous to Storey's $q$ value \cite{storey.03}.
-
-%Because our methods involve explicitly modelling effects $\beta_j$, rather than 
-%$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
-
-%Indeed, even in settings where some
-%$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
-%analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
-
-\subsection*{Related work}
-
-\subsubsection*{Previous approaches focussed on FDR}
-
-Among previous methods that explicitly consider the FDR and related quantities,
-our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
- (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
-  %Here we highlight the key differences between our approach and these three existing approaches.
-
-There are two key differences between our approach and all of these three existing methods. First, 
-whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
-we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
-for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
-Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 
-how it can improve performance by taking better account of variation in measurement precision among observations.
-%(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
-% Or more generally Section xx illustrates the potential practical effects
-
-%allows for different modelling assumptions to be compared through likelihood comparisons.
-Second, our unimodal assumption (UA) that the effects are unimodal about zero is quite different from assumptions made by $\qvalue, \locfdr$ or $\mixfdr$.
-Indeed, $\locfdr$ assumes that all $z$ scores near 0 are null (Efron calls this the Zero Assumption; ZA), which implies that 
-under the alternative hypothesis the distribution of $z$ scores has {\it no mass at 0}; this contrasts strikingly with the UA, which implies that this distribution has its peak at 0! 
-Similarly, $\qvalue$ assumes that all $p$ values near 1 are null,
-which is the same as the ZA because $p$ values near 1 correspond to $z$ scores near 0. And although \mixfdr does not formally make the ZA,
-we have found that in practice, with default settings, the results often approximately satisfy the ZA (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}).
-Thus, not only do these existing methods not make the UA, they actually make assumptions that are, in some sense, as different
-from the UA as they can be.
-
-Given that the UA and ZA are so different, it seems worth discussing why we generally favor the UA. 
-Although the UA will not apply to all situations, we believe that it will often be reasonable, especially in FDR-related contexts
-that have traditionally focussed on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis,  
-it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible. Further, it seems reasonable to expect that larger effects become
-decreasingly plausible, and so the distribution of the effects will be unimodal about 0.
-To paraphrase Tukey, ``All we know about the world teaches us that large effects are rare, whereas small effects abound."
-We emphasise that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
-It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
-
-%Tukey "peri-null"
-
-%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
-%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}.
-
-In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficents - 
-common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
-These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
-it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
-The UA assumption on regression coefficients is directly analagous to our UA here, and so its
-widespread use in the regression context supports its use here.
-
-Alternatively, we could motivate the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
-such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
-make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).
-
-Finally, the UA also has a considerable practical benefit: it yields simple algorithms that are both computationally
-and statistically stable. We illustrate these features in Results.
-
-\subsubsection*{Other work}
-
-There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
- Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
-indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
-This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
-Also, they provide no software implementation.
-More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
-Some pointers to recent papers in which other relevant
-citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}.
-Much of the literature focusses on the homoskedastic case (i.e.
-$\shat_j$ all equal) whereas we allow for heteroskedasticity.
-And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential.
-Several recent papers consider more flexible non-parametric assumptions on
-$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
-estimate (NPMLE) for $g$.
-These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive.
-However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
-may perform well for point estimation, it seems possible it will not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
-Indeed, one way of thinking about the UA is as a way to regularize $g$.
-
-%The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
-%encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-%$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-%$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-%the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-%observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-%corresondingly little effect on the estimated FDR of other observations. 
-%We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-%$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-%that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-%estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-%Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-\def\dt{\text{dt}}
-
-%
-%assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
-%$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
-% $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
-%  \begin{equation} \label{eqn:lik}
-% \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
-% \end{equation}
-% where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e.
-the density of $\mu+\sigma T_\nu$).
-In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
-%  \begin{equation} \label{eqn:normlik}
-% \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
-% \end{equation}
-%which, in light of (\ref{eqn:betahat}), is equivalent to
- % \begin{equation} \label{eqn:lik2}
-% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
- %\end{equation}
-
-%The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
-they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
-The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$.
-Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$.
-If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$.
-See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.
-
-%This likelihood approximation can be motivated by assuming that i) $\bhat$
-% $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
-% $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$. 
-%
-
-%Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.
-
-%While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
- %that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
-%there is an important difference: here,
-
-%The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
-For example, we can write $p(\bhat_j, \shat_j | \beta_j) = p(\bhat_j | \shat_j, \beta_j)p(\shat_j | \beta_j)$.
-Now, if we are willing to assume that $\shat_j$ alone contains no information about $\beta_j$, or equivalently that $p(\shat_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\bhat_j | \shat_j, \beta_j)$, and if $\bhat_j | \shat_j, \beta_j \sim N(\beta_j, \shat_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.
-
-\section*{Results}
-
-We compare results of \ashr with existing FDR-based methods implemented
-in the R packages \qvalue (v2.0 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
-In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
-to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.
-
-\subsection*{Effects of the Unimodal Assumption}
-
-Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
-standard error, $s_j=1$ and all methods are provided that information. That is,
-$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$.
-In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the only difference between our method and methods like $\locfdr$ and $\mixfdr$ are in how they estimate $g$.
-
-To briefly summarize the results in this section:
-\begin{enumerate}
-\item The UA can produce very different inferences compared with the ZA made by existing methods.
-\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
-\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
-\end{enumerate}
-
-\subsubsection*{The UA and ZA can produce different inferences}
-
-\begin{figure}
-\center \includegraphics[height=6in]{figures/decomp_ZA.pdf}
-\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
-The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
-In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
-The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
-(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
-\end{figure}
-
-To illustrate the different inferences from the UA and ZA we show results for a single
-dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
-but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
-\qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
-The results (Figure \ref{fig:ZA}) illustrate the clear difference between the existing methods and our method. The effects of the ZA made by \qvalue and \locfdr
-are visually clear, producing a ``hole"  in the alternative $z$ score distribution around 0. Although \mixfdr does not formally make the ZA, its decomposition
-exhibits a similar hole. In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
-effectively ``filling in" the hole.  (Of course the null distribution also has a peak at 0, and the local fdr 
- under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)
-
-Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA.
-Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
-This would increase the null component (dark blue) at the expense of
-the alternative component (light blue).
-Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0.
-The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears.
-In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
-See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
-demonstration.
-
- \subsubsection*{The UA can produce conservative estimates of $\pi_0$}
-
-The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
-Consequently \ashr will estimate smaller $\lfdr$s and FDRs than existing methods that make the ZA. 
-This is desirable, provided that these estimates remain conservative: that is,
-provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
-The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behaviour. To check its effectiveness
-we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
-for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
-They range from a ``spiky" distribution -- where many non-zero $\beta$ are
-too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
-flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
-making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
-which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.
-
-% How do these differences influence estimates of the FDR? To address this, note that  -- in this simple case where the 
-% $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
-% of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
-% strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
-% although the methods considered here differ from one another in many ways, having estimated $\pi_0$
-% they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
-% empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
-%
-
-%Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
-% and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
-%Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
-%At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
-%overestimates of $\pi_0$, which will yield overestimates of the FDR).
-
-\begin{figure} 
-\begin{center}
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=1.8in]{figures/scenarios_density.pdf} 
-	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/pi0_est.pdf} 
-\caption{Comparison of true and estimated values of $\pi_0$.
-When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ashr being least conservative, and hence most accurate.
-When the UA does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.} \label{fig:pi0sims}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfdr_est.png} 
-\caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
-Black line is $y=x$ and red line is $y=2x$.
-Estimates of $\lfdr$ are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est.png} 
-\caption{As in c), but for $\lfsr$ instead of $\lfdr$.
-Estimates of $\lfsr$ are consistently less conservative than $\lfdr$ when UA holds, and also less anti-conservative in bimodal scenario.}  \label{fig:lfsr}
-\end{subfigure}
-\end{center}
-\caption{Results of simulation studies (constant precision $s_j=1$).}
-\end{figure}
-
 For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
 data as follows:
+
 \begin{enumerate}
@@
 \end{enumerate}
-%Thus these simulations assume the same
-%precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).
 
@@
 are used. (In addition, the effects on $\lfsr$ estimates turn out to be relatively modest; see below).
-
-%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
-%
-%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
-%and so all effects are non-zero and no nulls hold. This scenario is rather different
-%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
-%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
-%``mostly non-null"
-%scenario, and we would agree! On the other hand, the distribution of $p$ values
-%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
-%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
-%Thus it seems desirable to have methods that perform well in both cases, and this is
-%an important feature of the methods we introduce here.
-%
-%TODO: put FDRs on graph? Simulate with 50\% null?
 
@@
 
-\begin{figure}
-\begin{center}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s_nn.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be more accurate in this case.
-That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
-\end{subfigure}
-\end{center}
-\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
-\end{figure}
-
 \subsubsection*{Numerical Stability}
@@
 threshold $T$.
-%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
-%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
-%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
-%towards zero, with the strength of the shrinkage being estimated from the data.
 Further, because the posterior distribution is, by definition,
@@
 and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
-%but also notes the difficulties of implementing a stable general algorithm.
 
@@
 under-estimating $\pi_0$ (Figure \ref{fig:egcdf}c).
-
-\begin{figure}[h!] 
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=2in]{figures/egcdf.pdf}
-	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf_nopen.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
-Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
-\end{subfigure}
-\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.} \label{fig:egcdf}
-\end{figure}
-
-%making the UA provides an attractive and simple way to achieve estimating $g$  large
-%subset of cases of interest (those where the UA can reasonably be assumed to hold).
-
-%In practice, the utility of the posterior distributions noted above depends primarily on the accuracy
-%of the estimate of the distribution $g(\cdot)$.
-
- %quote from efron rejoinder , p46:
-%MorrisÕ Section 3 is especially pertinent. His formula
-%for p(?|z) is related to my discussion of the
-%BenjaminiÐYekutieli False Coverage Rate method in
-%Section 7, particularly (7.2)Ð(7.4). Originally I had
-%hoped to develop an empirical Bayes method for estimating
-%such models, but the effort foundered on practical
-%difficulties involving the perils of deconvolution.
-%Section 6 on the Òone-group modelÓ is the ugly duckling
-%of the current paper, but it bears on some important
-%points raised in the discussion. Figure 7 concerns
-%a fuzzy version of simultaneous hypothesis testing,
-%where, as in MorrisÕ hospital example, the usual singlepoint
-%null hypotheses seem unequal to the task. The
-%development from (6.6) onwards, particularly (6.12),
-%bears on the possibility of nonnormal null distributions,
-%and is about as far as I can go in answering Professors
-%Rice and SpiegelhalterÕs penultimate question.
-%With g(?) a normal distribution, model (6.1) returns
-%us to the realm of Stein estimation, the scene of my
-%happy collaborations with Carl Morris. I continue to
-%be surprised at how much harder bumpy, nonnormal
-%models like (7.1) are to deal with. JamesÐStein estimation
-%works fine with, say, N = 10 component problems,
-%while the RobbinsÕ form of empirical Bayes appropriate
-%to (7.1) seems to require hundreds or thousands.
-%The information calculations in Efron (2008)
-%reinforce this gloomy assessment. Maybe I am trying
-%to be overly nonparametric in constructing the empirical
-%Bayes Fdr estimates, but it is hard to imagine a
-%generally satisfactory parametric formulation for (6.1).
-%Or perhaps it is just that hypothesis testing is more demanding
-%than estimation.
-%My Note: (6.1) is the general model (mu \sim g()).
-%7.1 is the case where the effects are a mixture of zero and N(-3,1).
-
-\subsubsection*{Calibration of posterior intervals}
-
-To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over 100 simulations in each Scenario). 
-Specifically we examine the empirical coverage of nominal lower 95\% credible bounds for a) all observations; b) significant negative discoveries; c) significant positive discoveries.
-We examine only lower bounds because the results for upper bounds follow by
-symmetry (except for the one asymmetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
-in each case: for negative discoveries the lower bound is typically large and negative and limits how big (in absolute value) 
-the effect could be; for positive discoveries the lower bound 
-is positive, and limits how small (in absolute value) the effect could be. Intuitively, the lower bound for negative discoveries depends on the accuracy of $g$ in its tail,
-whereas for positive discoveries it is more dependent on the accuracy of $g$ in the center.
-
-The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
-as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
-
-% As noted above, the penalty term (\ref{eqn:penalty}) causes systematic over-estimation of the mass of $g$ near 0,  which one would expect to overshrink interval estimates towards zero. 
-%The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound is negative and overshrinking (moving it too far towards zero) makes the left tail too big; conversely, for positive discoveries, the lower confidence bound is positive, and
-%overshrinking (moving it too far towards zero) makes the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
-%and be prepared to put up with some overshrinkage to avoid it.
-
-%Coverage rates are given in:
-%\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
-
-%Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all.tex}
-	\caption{All observations.
-Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario.
-This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero.
-Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (Table \ref{tab:nopen}).
-Removing the penalty in the half-uniform case is not recommended (see Appendix).}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg.tex}
-\caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
-and the normal-based method in the flat-top scenario.
-These results likely reflect inaccurate estimates of the tails of $g$ due to a disconnect between the tail of $g$ and the component distributions in these cases.
-For example, the uniform methods sometimes substantially underestimate the length of the tail of $g$ in these long-tailed scenarios,
-causing over-shrinkage of the tail toward 0.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos.tex}
-\caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds} \label{tab:coverage}
-\end{table}
-
-%
-%\subsubsection*{Reverse asymptotics}
-%
-%Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
-%
-%Note: we'd like to make sure it behaves ok in the case where most observations are null.
-%Perhaps simulate with 999 null and 1 alternative?
-%Also maybe look at performance for small sample sizes - e.g. n=1?
-%
-%Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
-%can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
-%by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
-%
-
-\subsection*{Differing measurement precision across units}
-
- We turn now to the second important component of our work: allowing for varying
- measurement precision across units. The key to this is the use of a likelihood,
- (\ref{eqn:normlik}) or (\ref{eqn:tlik}), that explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
-
-  To illustrate, we conduct a simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
- ($s_j=10$).  In both cases,
-we assume that half the effects are null
-and the other half are normally distributed with standard deviation 1:
-\begin{equation}
-p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1).
-\end{equation}
-In this setting, the poor-precision measurements $(s_j=10)$ tell us very little, and any sane analysis should effectively ignore them. 
- However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor}). This is because the poor measurements 
- produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
- which, when combined with the good-precision measurements, dilute the overall signal (e.g. they reduce the density of $p$ values near 0).
- This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
-the estimated error rates ($q$-values, or $\lfdr$ values) for the good-precision observations increase when the low-precision observations are included in the analysis
-(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
-for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).
-
-\begin{figure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_hist.pdf} 
-    \label{fig:goodpoor-hist}
-    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations}
- \end{subfigure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_scatter.pdf}
-    \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) of a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
-\end{subfigure}
-\caption{Simulation illustrating how, for existing FDR methods, 
-poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
-$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
-a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
-In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
-\end{figure}
-
- \subsubsection*{Reordering of significance, and the ``$p$ value prior"}
-
-Another consequence of accounting for differences in measurement precision is that \ashr may re-order the significance of the observations compared
- with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
- We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measuments that have the same $p$ value.
- The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
- and so the $\lfsr$ for these poor-precision measurements is always high.
-
-The potential for Bayesian analyses to re-order the significance of observations, and specifically to downweight imprecise observations,
-was previously discussed in \cite{guan.stephens.08}.
-However, Wakefield \cite{wakefield:2009} showed that, under a certain prior assumption, the Bayesian analysis produces the same ranking of significance as $p$ values (or their $z$ scores).
-He named this prior the ``$p$-value prior" because it can be thought of as the
-implicit prior assumption that is being made if we rank the significance of observations by their $p$ value.
-Wakefield's $p$-value prior assumes that the less precise effect estimates correspond to larger true effects, and specifically that they scale proportional
-to the standard errors $s_j$. More specifically still, it assumes a normal prior for the non-zero $\beta_j$ with mean 0 and variance $K s^2_j$ for some constant $K$.
-Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. Specifically, if, instead of assuming that $\beta_j$ is independent
-of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
-then the $\lfsr$ computed by ash.n provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
-as is illustrated in Figure \ref{fig:lfsr_pval}, right panel. (This result does not hold when using the mixtures of uniforms prior, ash.u.)
-
-This $p$-value prior assumes that the $z$ scores $z_j=\beta_j/s_j$
-are identically distributed, independent of $s_j$. This is essentially the assumption made, implicitly or explicitly, by existing methods -- like \locfdr, \mixfdr and \qvalue --
-that model the $z_j$ or $p_j$ directly.  In contrast, we have assumed up to now that the $\beta_j$ are independent of $s_j$. 
-We can set both these assumptions within a more general framework, which allows that $\beta_j/s^\alpha$ is independent of $s_j$ for some $\alpha$ [Equation (\ref{eqn:beta-alpha}))].
-Setting $\alpha=0$ implies that $\beta_j$ is
-independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
-effects (in absolute value). This latter assumption may often be qualitatively plausible: 
-for example, in gene expression studies the standard error for gene $j$ depends partly on
-the variance of its expression among samples, and genes with a larger variance may tend to be less tightly regulated and
-so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
-On the other hand, there is no particular reason to expect that either $\alpha=1$ or $\alpha=0$ will be the optimal choice.
-Indeed, optimal choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$,
-which will be dataset-specific. 
-Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
-comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
-likelihood-based methods can be used to select $\alpha$. For example, following the logic of the EB approach it
-would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
-which has been implemented in our software by C. Dai.
-
-%\begin{result}
-%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
-%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
-%\end{result}
-
-%To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
-%Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
-%In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
-%Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
-%which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
-%(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
-%may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)
-
-%\begin{result}
-%Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and using a mixture of normals for the prior distribution 
-%(\ref{eqn:beta-alpha}) with $\alpha=1$, 
-%the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$.
-Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$)
-%is montonic with $|t_j|$. 
-%\end{result}
-
-%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
-%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.
-
-%Most previous authors of EB approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
-%have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
-%In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
-%This is analogous
-%to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
-%standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
-%that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.
-
-\begin{figure}
-\centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
-\caption{Figure illustrating affects of prior assumptions on re-ordering of significance.
-Left panel shows results under our ``default prior" which assumes that effects $\beta_j$ are identically distributed, independent of $s_j$.
-Right panel shows results under the ``$p$-value prior", which assumes that $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$.} \label{fig:lfsr_pval}
-\end{figure}
-
-%\subsubsection*{Adaptive Shrinkage}
-%
-%Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
-%when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
-%the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
-% $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
-% In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
-% Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
-% and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
-%
-%Our methods we build on this idea in two ways. First,
-%by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
-%Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
-%adapts to measurement precision: more precise measurements undergo
-% less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
-%than observations with
-%small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
-%To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
-%Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
-%
-%\begin{figure}[!ht]
-%\begin{center}
-%\includegraphics[width=400px]{../examples/ash.pdf} 
-%\end{center}
-%\caption{Figure illustrating adaptive shrinkage.
-Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0.
-The standard error of each observation, $s_j$, was simulated from Inverse Gamma(5,5), which resulted in $s_j$ varying from 0.34 to 6.7.
-The shrunken estimates ($y$ axis) are plotted against the observed estimates ($x$ axis), with color indicating the (square root of the) precision of each measurement (blue=lower-precision; red=higher-precision).
-The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
-%\end{figure}
-
-%DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
-%To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
-Because the prior distribution is unimodal about zero,
-%the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
-%amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
-%For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
-%than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
-%small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
-
-%previous work based on normal means:
-%http://link.springer.com/article/10.1007/BF02562623
-
-%\subsection{Null behaviour}
-
-%Results of simulations under the null.
-
-%\subsection{Comparison of different underlying component distributions}
-%
-%Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
-%than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
-%likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
-%For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
-%increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
-%expect $L_{hu}$ to be significantly larger than the others.
-%
-%Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
-%
-%\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
-%\begin{center}
-%\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
-%%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
-%\end{center}
-%\caption{}
-%\end{figure}
-
-%Notes from Adrian Raftery - 
-%Newcombe (1886)
-%Dean and Raftery
-%Uniform is super-efficient
-
-\section*{Discussion}
-
-We have presented an Empirical Bayes approach to large-scale multiple testing that emphasises two ideas.
-First, we emphasise the potential benefits of using two numbers ($\bhat$, and its standard error)
-rather than just one number (a $p$ value or $z$ score)  to summarize the information on each test.
-While requiring two numbers is slightly more onorous than requiring one, in many settings these
-numbers are easily available and if so we argue it makes sense to use them.
-Second, we note the potential benefits -- both statistical and computational -- of assuming that the effects come from a unimodal distribution, and provide
-flexible implementations for performing inference under this assumption. We also introduce the ``false sign rate" as
-an alternative measure of error to the FDR, and illustrate its improved robustness to errors in model fit, particularly mis-estimation
-of the proportion of null tests, $\pi_0$.
-
-Multiple testing is often referred to as a ``problem" or a ``burden". In our opinion, EB approaches turn this idea on its head,
-treating multiple testing as an {\it opportunity}: specifically, an opportunity to learn about the prior distributions, and other modelling assumptions,
-to improve inference and make informed decisions about significance (see also \cite{greenland1991empirical}). This view
-also emphasises that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
-Indeed, the FDR at a given fixed threshold does not depend on the number of tests: as the number of tests increases, both the true positives and false positives increase linearly,
-and the FDR remains the same.
-(If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example.
-The EB approach captures this dependence in an intuitive way:
-if there are lots of strong signals then we infer $\pi_0$ to be small, and the estimated FDR (or $\lfdr$, or $\lfsr$) at a given threshold may be low, even if a large number of tests were performed; and conversely
-if there are no strong signals then we infer $\pi_0$ to be large and the FDR at the same threshold may be high, even if relatively few tests were performed. 
-More generally, overall signal strength is reflected in the estimated $g$, which in turn influences the estimated FDR.
-
-Two important practical issues that we have not addressed here are correlations among tests,
-and the potential for deviations from the theoretical null distributions of test statistics. These two
-issues are connected: specifically, unmeasured confounding factors can cause both correlations among tests
-and deviations from the theoretical null \cite{efron2007correlation,leek:2007}. And although there are certainly other
-factors that could cause dependence among tests, unmeasured confounders are perhaps the most worrisome in practice
-because they can induce strong correlations among large numbers of tests and profoundly impact results.
-Approaches to deal with unmeasured confounders can be largely divided into two types: those
-that simply attempt to correct for the resulting inflation of test statistics \cite{devlin1999genomic,efron2004large}, and those
-that attempt to infer counfounders using clustering, principal components analysis, or factor models 
-\cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
-and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable,
-they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alterntively, our methods
-could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:lik}) using
-$\bhat_j | \shat_j \sim N(0,\lambda_1 \shat_j + \lambda_2)$, where $\lambda_1,\lambda_2$ are to be estimated. However,
-this immediately raises issues of identifiability and we do not pursue the idea further here.
-
-Another important practical issue is the challenge of small sample sizes. For example, in genomics applications researchers sometimes
-attempt to identify differences between two conditions based on only a handful of samples in each. In such settings the
-normal likelihood approximation (\ref{eqn:normlik}) will be inadequate. And, although the $t$ likelihood (\ref{eqn:tlik}) partially
-addresses this issue, it is also, it turns out, not entirely satisfactory.  The root of the problem is that, with small sample sizes, raw estimated standard
-errors $\shat_j$ can be horribly variable. In genomics it is routine to address this issue by applying EB methods \cite{smyth:2004} to ``moderate" (i.e.~shrink) variance estimates,
-before computing $p$ values from ``moderated" test statistics. We are currently investigating how
-our methods should incorporate such ``moderated" variance estimates to make it applicable to small sample settings.
-
-Our approach involves compromises between flexibility, generality, and simplicity on the one hand, and statistical efficiency and principle on the other.
-For example, in using an EB approach that uses a point estimate for $g$, rather than a fully Bayes approach that accounts for uncertainty in $g$,
-we have opted for simplicity over statistical principle. And in summarizing every test by
-two numbers and making a normal or $t$ approximation to the likelihood, we have aimed to produce generic methods that 
- can be applied whenever such summary data are available -- just as \qvalue can be applied to any set of $p$ values for example -- although
- possibly at the expense of statistical efficiency compared with developing multiple tailored approaches based on context-specific likelihoods. 
- Any attempt to produce generic methods will involve compromise between generality and efficiency.
- In genomics, many analyses -- not only FDR-based analyses -- involve first computing a series of $p$ values before
- subjecting them to some further downstream analysis. An important message here is 
- that working with two numbers ($\bhat_j,\shat_j$), rather than one ($p_j$ or $z_j$), 
- can yield substantial gains in functionality (e.g.~ estimating effect sizes, as well as testing; accounting for variations in
-measurement precision across units) while losing only a little in generality.  We hope that our work will
-encourate development of methods that exploit this idea in other contexts.
-
-%This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
-%from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
-%with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
-
-%
-%
-%Both \locfdr and \mixfdr allow for deviations from
-%the theoretical null distribution by allowing that, under the null, the $z$ scores are $N(0,\sigma^2)$ with $\sigma$ estimated from the data rather
-%than fixed to 1. Within our approach this idea could be incorporated by changing the likelihood - for example, to assume $\bhat_j \sim N(0,s_ 
-%
-%These are tricky issues that, as far as we are aware, can affect
-%all approaches to FDR analysis. An important focus of Efron's work, also implemented in \locfdr, 
-% is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
-%theoretical null distribution. 
-%
-%For example, we and others have worked on Bayesian approaches to genetic association testing, using subjective prior specification of
-%effect sizes based on external data (e.g.~\cite{stephens:2009}). However, the large numbers of tests brings the opportunity to learn about
-%priors, and here we aim to exploit this
-
-%Choice of mixture components. Use of likelihood to compare different mixture components?
-
-%  
-%Directly modeling the p values, or z scores, say via non-parametric methods, 
-%can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-%distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-%The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-%Before proceeding, we feel obliged to correct the common misconception
-% that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
-% not depend on the number of tests. 
-%  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
-% 
-% In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
-
-\section*{Detailed Methods} \label{sec:implement}
-
-\subsection*{Embellishments}
-
-\subsubsection*{More flexible unimodal distributions}
-
-Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
-Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
- Therefore, we have implemented a more general approach based on 
-\begin{equation} \label{eqn:g-general}
-g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) ,
-\end{equation}
-where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
-\begin{enumerate}[(i)]
-\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$, \qquad (``ash.n")
-\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,  \qquad (``ash.u")
-\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,  \qquad (``ash.hu")
-\end{enumerate}
-where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
-(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
-and $\pi$ is a $2K+1$ vector that sums to 1.)
-The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
-Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
-under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
-(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
-(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
-however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)
-
-Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
-using a large dense grid of $\sigma^2_k$ or $a_k$, (i)-(iii) can respectively approximate,
-with arbitrary accuracy,
-\begin{enumerate}[(i)]
-\item any scale mixture of normals, which includes as special cases
-the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
-\item any symmetric unimodal distribution about 0.
-\item any unimodal distribution about 0.
-\end{enumerate}
-The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
-In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.
-
-Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
-The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
-See Implementation for details of our software defaults.
-
-%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting.
-For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
-%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
-
-\subsubsection*{Dependence of effects on standard errors}
-
-Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
-This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
- \begin{equation} \label{eqn:beta-alpha}
- \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
- \end{equation}
-for any $\alpha$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
-assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
-it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
-the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.
-
-The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
-To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
- standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
- Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
-yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
-%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
-%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
-%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
-
-%Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
-%how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
-%it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
-%rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
-%and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
-%this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
-%do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
-
-\subsubsection*{Replace normal likelihood with $t$ likelihood}
-
-We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
-  \begin{equation} \label{eqn:tlik}
- \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
- \end{equation}
-where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
-and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
-The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
-This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms; see Implementation below.
-When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
-
-Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
-\begin{equation} \label{eqn:tdist}
-(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
-\end{equation}
-However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
-In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
-currently pursuing this improved approach (joint work with M.Lu) and results will be published elsewhere. 
-%We note here that
-%modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
-%where our simpler approach can produce unsatisfactory results.
-
-\subsubsection*{Non-zero mode}
-
-An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood, rather than fixed to 0.
-
-\subsection*{Implementation Details}
-
-\subsubsection*{Likelihood for $\pi$}
-
-We define the likelihood for $\pi$ to be the probability of the observed data $\bhat$ conditional on $\shat$: $L(\pi) := p(\bhat | \shat,\pi),$ which
-by our conditional indendence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$.
-[One might prefer to define the likelihood as $p(\bhat, \shat | \pi) = p(\bhat | \shat, \pi) p(\shat | \pi)$, in which case our definition comes down to assuming that the term $p(\shat | \pi)$ does not depend on $\pi$.]
-
-Using the prior $\beta_j \sim \sum_{k=0}^K \pi_k f_k(\beta_j)$ given by (\ref{eqn:g-general}), and the normal likelihood (\ref{eqn:normlik}), integrating over $\beta_j$ yields
-\begin{equation}
-p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
-\end{equation}
-where 
-\begin{equation}
-\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
-\end{equation}
-denotes the convolution of $f_k$ with a normal density.
-These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
-Specifically, 
-\begin{equation} \label{eqn:uconv}
-\tilde{f}_k(\bhat_j)  = 
-\begin{cases}
-N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
-\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
-\end{cases}
-\end{equation}
-where $\Psi$ denotes the cumulative distribution function (c.d.f.) of the standard normal distribution.
-If we replace the normal likelihood with the $t_\nu$ likelihood (\ref{eqn:tlik}) then the convolution 
-for $f_k$ uniform the convolution is still given by (\ref{eqn:uconv}) but with $\Psi$ the c.d.f. 
-of the $t_\nu$ distribution function. (The convolution for $f_k$ normal is tricky and we have not implemented it.)
-
-\subsubsection*{Penalty term on $\pi$}
-
-To make $\lfdr$ and $\lfsr$ estimates from our method ``conservative" we add a penalty 
-term $log(h(\pi;\lambda))$ to the log-likelihood $\log L(\pi)$ to encourage over-estimation of $\pi_0$:
-\begin{equation} \label{eqn:penalty}
-h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
-\end{equation}
-where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
-consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0sims}).
-
-Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
-we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
-
-\subsubsection*{Problems with removing the penalty term in the half-uniform case}
-
-It is straightforward to remove the penalty term by setting $\lambda_k=1$ in (\ref{eqn:penalty}).
-We note here an unanticipated problem we came across when using no penalty term in the half-uniform case 
-(i.e.~ $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$ in (\ref{eqn:g-general})): when the data
-are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
-sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
-for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
-and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
-
-\subsubsection*{EM algorithm}
-
-With this in place, the penalized log-likelihood for $\pi$ is given by:
-\begin{equation}
-\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
-\end{equation}
-where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This can be maximized using an EM algorithm \cite{dempster77}, whose
-one-step updates are:
-\begin{align} \label{eqn:w}
-w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
-n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
-\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}
-\end{align}
-Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
-and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
-probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$. (The $w_{kj}$ are sometimes referred to as the ``responsibilities".)
-Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
-We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple} to accelerate convergence of this EM algorithm.
-
-%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
-\subsubsection*{Initialization}
-
-By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
-Our rationale for initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
-but weak signal in the data cannot quickly draw the algorithm towards the null.
-
-In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
-$\pi_0=1; \pi_k=0 (k>0)$. If not then we replace the EM solution with the global null solution. This aims to guard against errors due to convergence to a local optimum when the
-data are consistent with the global null.
-
-\subsubsection*{Conditional distributions}
-
-Given $\hat\pi$, we compute the conditional distributions 
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
-\end{equation} 
-Each posterior is a mixture on $K+1$ components:
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
-\end{equation}
-where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
-and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
-$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
-All these posterior distributions are easily available.
-For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
-If $f_k$ is normal and $L$ is normal, then this is a normal distribution.
-
-\subsubsection*{Choice of grid for $\sigma_k, a_k$} \label{sec:grid}
-
-\def\sigmamax{\sigma_{\text{max}}}
-\def\sigmamin{\sigma_{\text{min}}}
-When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
-in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
-as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
-based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
-is negative we set $\sigmamax = 8\sigmamin$.)
-
-When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.
-
-Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
-a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.
-
-\appendix
-
-\begin{table}[!ht]
-\centering\begin{tabular}{c c } \toprule
-Scenario & Alternative distribution, $g_1$  \\ \midrule
-spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
-near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
-flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
- &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
-skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
-big-normal & $N(0,4^2)$ \\ 
-bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
-\end{tabular}
-\caption{Summary of simulation scenarios considered} \label{table:scenarios}
-\end{table}
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all_nopen.tex}
-	\caption{All observations}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg_nopen.tex}
-\caption{``Significant" negative discoveries.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos_nopen.tex}
-\caption{``Significant" positive discoveries.}
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{table:nopen}
-\end{table}
-%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
-%\begin{equation}
-%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
-%\end{equation}
-%which is close to the definition in \cite{storey.03}.
-
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgements}
-
-Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
-prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} , and includes contributions 
-from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.
-
-This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.
-
-%\section*{References}
-% The bibtex filename
-\bibliography{/Users/stephens/Dropbox/Documents/mainbib}
-
-\section*{Figure Legends}
-
-\clearpage
-
-\section*{Tables}
-
-\section*{Supporting Information Legends}
-
-Supplementary material can be found in {\bf Supplementary Information S1.}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
-\end{document}

### 2015-10-06 To_illustrate_the_di.tex
New Article!

--- 2015-10-06 paper/main.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
-\documentclass[11pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-%\usepackage{natbib}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-\usepackage{booktabs}
-\usepackage{xspace}
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-\usepackage{sectsty}
-\subsectionfont{\normalsize}
-\subsubsectionfont{\mdseries\itshape}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\usepackage{color}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-% Use the PLoS provided bibtex style
-\bibliographystyle{/Users/stephens/Dropbox/Documents/stylefiles/plos2009}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
-\pagestyle{myheadings}
-%% ** EDIT HERE **
-\usepackage{enumerate}
-\usepackage{multirow} 
-\usepackage{url}
-\usepackage{xr} %for cross-referencing
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-\newtheorem{algorithm}{Algorithm}
-\newtheorem{proposition}{Proposition}
-\newtheorem{restateproposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{result}{Result}
-\newtheorem{note}{Note}
-\newtheorem{definition}{Definition}
-
-\def\lfdr{\textit{lfdr}}
-\def\lfsr{\textit{lfsr}}
-\def\x{\mbf{x}}
-\def\y{\mbf{y}}
-\def\e{\mbf{e}}
-\def\g{\mbf{g}}
-\def\p{p}
-\def\Yb{\hat{Y}_1^{\text{BAYES}}}
-\def\Xperp{X_{1 \perp 0}}
-\def\Pperp{P_{1 \perp 0}}
-\def\Pperpb{\Pperp^{\text{B}}}
-\def \hatd { \widehat{ d } }
-\def \dhatin {\hatd^{\mbox{\tiny IN}}}
-\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
-\def \mvn{\text{MVN}}
-\def\mn{\text{MN}}
-\def \iw{\text{W}^{-1}}
-\def \fp {fastPHASE\ }
-\def\bhat{\hat{\beta}}
-\def\shat{\hat{s}}
-\def\var{\rm Var}
-\def\mvr{\text{BMVR}}
-\def\bfuni{\text{BF}_\text{uni}}
-\def\one{\mbf{1}}
-\def\zero{\mbf{0}}
-\def\hall{H_{\text{all}}}
-\def\pall{\p_{\text{all}}}
-\def\bfall{\text{BF}_\text{all}}
-\def\bfav{\text{BF}_\text{av}}
-\def\gall{\gamma_\text{all}}
-\def\limbfall{\bfall^\rightarrow}
-\def\limbfav{\bfav^\rightarrow}
-\def\limbfgamma{\BF_\gamma^\rightarrow}
-\def\sxx{S_{xx}}
-\def\syy{S_{yy}}
-\def\vyy{V_{yy}}
-\def\vyx{V_{yx}}
-\def\vxx{{V_{xx}}}
-\def\rss{\text{\rm RSS}}
-\def\dss{\text{\rm $\Delta$SS}}
-\def\U{U}
-\def\D{D}
-\def\I{I}
-\def\luu{L_{UU}}
-\def\indep{\perp}
-\def\qvalue{{\tt qvalue}\xspace}
-\def\locfdr{{\tt locfdr}\xspace}
-\def\mixfdr{{\tt mixfdr}\xspace}
-\def\ashr{{\tt ashr}\xspace}
-
-\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
-                                % mode too
-\newcommand\height{{\it height}}
-\newcommand\weight{{\it weight}}
-\def\BF{\text{BF}}
-
-%% END MACROS SECTION
-%\externaldocument{plos_one_rev1_suppinfo}
-
-\begin{document}
-
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{False Discovery Rates: A New Deal}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Matthew Stephens$^{1*}$, 
-\\
-\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
-\\
-$\ast$ E-mail: Corresponding mstephens@uchicago.edu
-\end{flushleft}
-
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-%\section*{Author Summary}
-
-\section*{Introduction}
-
-Since its introduction in 1995 by Benjamini and Hochberg 
-\cite{benjamini1995controlling}, the ``False Discovery Rate" (FDR) has quickly established itself
-as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing
-in which the goal is to identify the non-zero ``effects" among a large number of imprecisely-measured effects.
-
-Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
-indeed, the notion that EB approaches could be helpful in handling multiple comparisons predates introduction of
-the FDR  (e.g.~\cite{greenland1991empirical}). More recently, EB approaches to the FDR
-have been extensively studied by several authors, 
-especially B.~Efron and co-authors 
-\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
-see also \cite{kendziorski2003parametric,newton2004detecting, 
-datta2005empirical,muralidharan2010empirical} for example.
-So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
-existing widely-used FDR pipelines) and can substantially affect
-inference. The first idea is to {\it assume that the distribution of effects is unimodal}. This yields a very simple, fast, and stable computer 
-implementation, as well as improving inference of FDR when the unimodal assumption is correct.  
-The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
-to summarize each measurement. This idea allows variations in measurement precision to be better accounted for,
- and avoids a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
-
-In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
-the idea that it may be preferable to focus on estimation rather than on testing.
-In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
-simply estimation with some positive prior probability that the effect is exactly zero.
-However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
-perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
-controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
-``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
-in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
-the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the lfsr, particularly its increased robustness to modelling assumptions.
-
-%terminology
-%``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
-%and correspondingly define the ``false signed discovery rate", or simply ``false sign rate" (FSR), in a natural way.
-
-%The methods we provide here can address both problems, and we provide results that highlight important benefits of focussing on estimation, and particularly on assessing the correctness of the {\it sign} of the estimated effects, rather than on testing for non-zero effects.
-
-Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
-(We address the reasons for this name, and connections with shrinkage analysis, in the Discussion.)
-
-%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.
-
-%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-%measurements are made on only a small number of
-%samples in each condition - sometimes as few as one
-%sample in each condition. Thus the error in estimates
-%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-%in estimates of $\beta_j$ still greater.
-%
-
-%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
-%Here we present new tools that bring two important ideas to bear on the problem.
-%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
-%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
-%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
-%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
-%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
-%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
-%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
-%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
-%a substantial increase in the number of discoveries at a given FDR).
-
-%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
-%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
-Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
-%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
-%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
-%generic methods that work directly with two numbers for each 
-%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
-Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
-%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
-
-\def\df{df}
-\def\FDR{\text{FDR}}
-\def\fdr{\text{\lfdr}}
-\def\FSR{\text{FSR}}
-\def\fsr{\text{lfsr}}
-
- \section*{Methods}
-
- \subsection*{Model Outine}
-
-% Here we consider one setting in which FDR-based methods are widely used:
-%an experimenter has measured a large numbers of ``effects", each with an associated
-%measurement precision (standard error), 
-%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
-%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
-%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
-%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
-%For example, in genomics
-%a common goal is to identify ``differentially-expressed genes" -- that is, genes
-%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
-%mean expression level between the two conditions, often referred to as the ``log-fold change". 
-Here we describe the simplest version of our method, and briefly discuss embellishments we have also implemented.
-
-Suppose that we are interested in the values of $J$ ``effects" $\beta=(\beta_1,\dots,\beta_J)$. For example, in a typical genomics application
-that aims to identify differentially expressed genes,  $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
-In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
-that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
-more general problem of estimating, and assessing uncertainty in, $\beta_j$.
-
-Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
-and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
-which by Bayes theorem can be written as 
-\begin{equation}
-p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
-\end{equation}
-
-For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
-This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
-A simple way to implement the UA is to assume that $g$ is
-a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
-\begin{align} \label{eqn:beta}
-p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
-g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
-\end{align}
-where $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyperparameters, which are non-negative and sum to one, and are to be estimated,
-while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
-represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
-(We encourate the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
-although of course in practice we use a finite grid -- see Implementation Details.)
-
-For the likelihood $p(\bhat | \beta, \shat)$ we assume  
-\begin{equation} \label{eqn:normlik}
-p(\bhat | \beta, \shat) = \prod_j N(\bhat_j; \beta_j, \shat_j^2).
-\end{equation}
-Here, in additional to some conditional independence assumptions, we are 
-effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-
-This simple model features both the key ideas we want to emphasise in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
-the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
-with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
-However, this simple model also has several additional assumptions that can be relaxed. Specifically,
-\begin{enumerate}
-\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) also implies that $g$ is symmetric about 0; more flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see (\ref{eqn:g-general})].
-\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
-We can relax this to allow for a relationship between these quantities [see (\ref{eqn:beta-alpha})].
-\item The likelihood (\ref{eqn:normlik}) effectively assumes that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-We can generalize this likelihood using a $t$ likelihood [see (\ref{eqn:tlik})].
-\end{enumerate}
-These embellishments are detailed in Embellishments and Implementation Details.
-Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
-(which are also made by most existing EB approaches to this problem). Correlations among tests certainly arise in practice, either due to genuine correlations
-in the system of study, or due to unmeasured confounders, and their potential to impact results of an FDR analysis is important to consider
-whatever analysis methods are used: see \cite{efron2007correlation,leek:2007} for relevant discussion.
-
-%We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
-%$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
-%\begin{equation}
-%p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
-%\end{equation}
-%Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
-%\begin{equation} \label{eqn:normlik}
-%p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
-%
-%We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
-%as a mixture of null and alternative components:
-%\begin{equation} \label{eqn:betaj0}
-%g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
-%\end{equation}
-%where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
-%We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
-%\begin{equation} \label{eqn:beta-iid}
-%p(\beta | \shat) = \prod_j g(\beta_j).
-%\end{equation}
-%
-%Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
-%with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
-%to use a mixture of zero-mean normal distributions:
-%\begin{equation} \label{eqn:mixnorm}
-%g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
-%Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
-%represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
-% The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
-% more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below. 
-% 
-% Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
-% yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
-% \begin{equation} \label{eqn:g}
-% g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
-% \end{equation}
-% 
-% Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
-% \begin{align} \label{eqn:betaj}
-% p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
-% p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
-% \end{align}
-% with $g$ given by the mixture of normals $(\ref{eqn:g})$.
-
-\subsection*{Fitting the model}
-
-In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
-each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
-Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
-\begin{equation} \label{eqn:marginallik}
-p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
-\end{equation}
-where we define $\sigma_0:=0$.
-
-The usual EB approach to fitting this model would involve two simple steps:
-\begin{enumerate}
-\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
-\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
-against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | | \bhat, \shat,\hat{\pi})$.
-\end{enumerate}
-Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(The simplicity of the EM algorithm in step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
-which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
-
-Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximise a penalized
-likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
-We introduce this penalty because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
-
-\subsection*{The local False Discovery Rate and local False Sign Rate}
-
-As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
-In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
-One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
-the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
-In other words it is the posterior probability that $\beta_j$ is actually zero:
-\begin{equation} \label{eqn:fdr}
-\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
-\end{equation}
-
-The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
-whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
-the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey1991philosophy}) argued that
-``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$.
-Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
-\begin{quote}
-...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
-\end{quote}
-Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
-focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.
-
-Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
-would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
-\begin{equation} \label{eqn:lfsr}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
-\end{equation}
-To illustrate, suppose that 
-\begin{gather*}
-p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
-p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
-p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
-\end{gather*} 
-Then 
-from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
-our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.
-
-As our notation suggests, $\lfsr_j$ is intended to be compared and contrasted with $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
-small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
-Of coure, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
-this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
-(this follows from the definition because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
-In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
-$\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.
-
-From these ``local" measures of significance, we can also compute average error rates over subsets of observations
-$\Gamma \subset \{1,\dots,J\}$. For example,   
-\begin{equation} \label{eqn:FDRhat}
-\widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
-\end{equation}
-estimates the FDR we would obtain if we were to declare all tests in $\Gamma$ significant.
-And 
-\begin{equation} \label{eqn:qvalue}
-q_j := \widehat \FDR(\{k: \fdr_k \leq \fdr_j\}) 
-\end{equation}
-provides a measure of significance analogous to Storey's $q$ value \cite{storey.03}.
-
-%Because our methods involve explicitly modelling effects $\beta_j$, rather than 
-%$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
-
-%Indeed, even in settings where some
-%$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
-%analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
-
-\subsection*{Related work}
-
-\subsubsection*{Previous approaches focussed on FDR}
-
-Among previous methods that explicitly consider the FDR and related quantities,
-our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
- (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
-  %Here we highlight the key differences between our approach and these three existing approaches.
-
-There are two key differences between our approach and all of these three existing methods. First, 
-whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
-we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
-for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
-Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 
-how it can improve performance by taking better account of variation in measurement precision among observations.
-%(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
-% Or more generally Section xx illustrates the potential practical effects
-
-%allows for different modelling assumptions to be compared through likelihood comparisons.
-Second, our unimodal assumption (UA) that the effects are unimodal about zero is quite different from assumptions made by $\qvalue, \locfdr$ or $\mixfdr$.
-Indeed, $\locfdr$ assumes that all $z$ scores near 0 are null (Efron calls this the Zero Assumption; ZA), which implies that 
-under the alternative hypothesis the distribution of $z$ scores has {\it no mass at 0}; this contrasts strikingly with the UA, which implies that this distribution has its peak at 0! 
-Similarly, $\qvalue$ assumes that all $p$ values near 1 are null,
-which is the same as the ZA because $p$ values near 1 correspond to $z$ scores near 0. And although \mixfdr does not formally make the ZA,
-we have found that in practice, with default settings, the results often approximately satisfy the ZA (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}).
-Thus, not only do these existing methods not make the UA, they actually make assumptions that are, in some sense, as different
-from the UA as they can be.
-
-Given that the UA and ZA are so different, it seems worth discussing why we generally favor the UA. 
-Although the UA will not apply to all situations, we believe that it will often be reasonable, especially in FDR-related contexts
-that have traditionally focussed on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis,  
-it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible. Further, it seems reasonable to expect that larger effects become
-decreasingly plausible, and so the distribution of the effects will be unimodal about 0.
-To paraphrase Tukey, ``All we know about the world teaches us that large effects are rare, whereas small effects abound."
-We emphasise that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
-It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
-
-%Tukey "peri-null"
-
-%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
-%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}.
-
-In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficents - 
-common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
-These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
-it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
-The UA assumption on regression coefficients is directly analagous to our UA here, and so its
-widespread use in the regression context supports its use here.
-
-Alternatively, we could motivate the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
-such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
-make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).
-
-Finally, the UA also has a considerable practical benefit: it yields simple algorithms that are both computationally
-and statistically stable. We illustrate these features in Results.
-
-\subsubsection*{Other work}
-
-There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
- Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
-indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
-This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
-Also, they provide no software implementation.
-More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
-Some pointers to recent papers in which other relevant
-citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}.
-Much of the literature focusses on the homoskedastic case (i.e.
-$\shat_j$ all equal) whereas we allow for heteroskedasticity.
-And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential.
-Several recent papers consider more flexible non-parametric assumptions on
-$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
-estimate (NPMLE) for $g$.
-These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive.
-However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
-may perform well for point estimation, it seems possible it will not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
-Indeed, one way of thinking about the UA is as a way to regularize $g$.
-
-%The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
-%encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-%$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-%$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-%the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-%observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-%corresondingly little effect on the estimated FDR of other observations. 
-%We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-%$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-%that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-%estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-%Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-\def\dt{\text{dt}}
-
-%
-%assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
-%$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
-% $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
-%  \begin{equation} \label{eqn:lik}
-% \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
-% \end{equation}
-% where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e.
-the density of $\mu+\sigma T_\nu$).
-In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
-%  \begin{equation} \label{eqn:normlik}
-% \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
-% \end{equation}
-%which, in light of (\ref{eqn:betahat}), is equivalent to
- % \begin{equation} \label{eqn:lik2}
-% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
- %\end{equation}
-
-%The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
-they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
-The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$.
-Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$.
-If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$.
-See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.
-
-%This likelihood approximation can be motivated by assuming that i) $\bhat$
-% $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
-% $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$. 
-%
-
-%Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.
-
-%While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
- %that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
-%there is an important difference: here,
-
-%The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
-For example, we can write $p(\bhat_j, \shat_j | \beta_j) = p(\bhat_j | \shat_j, \beta_j)p(\shat_j | \beta_j)$.
-Now, if we are willing to assume that $\shat_j$ alone contains no information about $\beta_j$, or equivalently that $p(\shat_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\bhat_j | \shat_j, \beta_j)$, and if $\bhat_j | \shat_j, \beta_j \sim N(\beta_j, \shat_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.
-
-\section*{Results}
-
-We compare results of \ashr with existing FDR-based methods implemented
-in the R packages \qvalue (v2.0 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
-In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
-to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.
-
-\subsection*{Effects of the Unimodal Assumption}
-
-Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
-standard error, $s_j=1$ and all methods are provided that information. That is,
-$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$.
-In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the only difference between our method and methods like $\locfdr$ and $\mixfdr$ are in how they estimate $g$.
-
-To briefly summarize the results in this section:
-\begin{enumerate}
-\item The UA can produce very different inferences compared with the ZA made by existing methods.
-\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
-\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
-\end{enumerate}
-
-\subsubsection*{The UA and ZA can produce different inferences}
-
-\begin{figure}
-\center \includegraphics[height=6in]{figures/decomp_ZA.pdf}
-\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
-The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
-In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
-The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
-(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
-\end{figure}
-
 To illustrate the different inferences from the UA and ZA we show results for a single
@@
 which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.
-
-% How do these differences influence estimates of the FDR? To address this, note that  -- in this simple case where the 
-% $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
-% of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
-% strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
-% although the methods considered here differ from one another in many ways, having estimated $\pi_0$
-% they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
-% empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
-%
-
-%Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
-% and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
-%Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
-%At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
-%overestimates of $\pi_0$, which will yield overestimates of the FDR).
-
-\begin{figure} 
-\begin{center}
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=1.8in]{figures/scenarios_density.pdf} 
-	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/pi0_est.pdf} 
-\caption{Comparison of true and estimated values of $\pi_0$.
-When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ashr being least conservative, and hence most accurate.
-When the UA does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.} \label{fig:pi0sims}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfdr_est.png} 
-\caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
-Black line is $y=x$ and red line is $y=2x$.
-Estimates of $\lfdr$ are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est.png} 
-\caption{As in c), but for $\lfsr$ instead of $\lfdr$.
-Estimates of $\lfsr$ are consistently less conservative than $\lfdr$ when UA holds, and also less anti-conservative in bimodal scenario.}  \label{fig:lfsr}
-\end{subfigure}
-\end{center}
-\caption{Results of simulation studies (constant precision $s_j=1$).}
-\end{figure}
-
-For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
-data as follows:
-\begin{enumerate}
-\item Simulate $\pi_0  \sim U[0,1]$.
-\item For $j=1,\dots,J$, simulate $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)$.
-\item For $j=1,\dots,J$, simulate $\bhat_j | \beta_j \sim N(\beta_j,1)$.
-\end{enumerate}
-%Thus these simulations assume the same
-%precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).
-
-Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ashr ($y$ axis) with the true values ($x$ axis). 
-For \ashr we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
-and as a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
-as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
-The results show that \ashr provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
-in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ashr estimates can be slightly anti-conservative.
-We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
-are used. (In addition, the effects on $\lfsr$ estimates turn out to be relatively modest; see below).
-
-%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
-%
-%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
-%and so all effects are non-zero and no nulls hold. This scenario is rather different
-%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
-%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
-%``mostly non-null"
-%scenario, and we would agree! On the other hand, the distribution of $p$ values
-%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
-%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
-%Thus it seems desirable to have methods that perform well in both cases, and this is
-%an important feature of the methods we introduce here.
-%
-%TODO: put FDRs on graph? Simulate with 50\% null?
-
-\subsubsection*{The $\lfsr$ is more robust than \lfdr}
-
-The results above show that \ashr can improve on existing methods in producing smaller, more accurate, 
-estimates of $\pi_0$, which will lead to more accurate estimates of FDR.
-Nonetheless, in many scenarios \ashr continues
-to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
-This is because these scenarios include 
-an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
-estimation of $\pi_0$ impossible. Put another way, and as is well known, $\pi_0$ is not identifiable:
- the data can effectively provide an upper bound on plausible values of $\pi_0$,
-but not a lower bound (because the data cannot rule out that everything is non-null, but with miniscule effects).
-To obtain conservative behaviour we must estimate $\pi_0$ by this upper bound, which can
-be substantially larger than the true value.
-
-Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
-overestimation of $\pi_0$ is corresponding overestimation of FDR (and $\lfdr$, and $q$ values).
-To illustrate, Figure \ref{fig:lfdr} compares the estimated $\lfdr$ from ash.n with the true value (computed using
-Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios which involve many
-non-zero effects that are very near 0 (e.g.~the spiky scenario with $\pi_0$ small) where
-$\pi_0$ can be grossly overestimated.  
-(Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.)
-
-The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
- become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
- This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
-To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
-be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
-the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in $\lfsr$ than $\lfdr$.
-
-Note that, compared with previous debates regarding testing, 
-this section advances an additional reason for focussing on the sign of the effect, rather than just testing whether it is 0. 
-In previous debates authors have argued against testing whether an effect is 0
-because it is {\it implausible that effects are exactly 0}. Here we add that {\it even if one believes
-that some effects may be exactly zero}, it is still better to focus on the sign, because generally {\it the data are more informative about that question}
-and so inferences are more robust to, say, the inevitable mis-estimation of $\pi_0$.
-To provide some intuition, consider an observation with a $z$ score of 0. The $\lfdr$ of this observation can range from 0 (if $\pi_0=0$)
-to 1 (if $\pi_0=1$). But, assuming a symmetric $g$, the $\lfsr>0.5$ whatever the value of $\pi_0$, because the observation $z=0$ says
-nothing about the sign of the effect.
-Thus, are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
-whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.
-
-Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
-Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
-if we are prepared to assume that no effects are exactly zero, then removing the point mass 
-yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
-However, there is ``no free lunch":  if in fact some effects are exactly zero
-then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
-We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.
-
-\begin{figure}
-\begin{center}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s_nn.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be more accurate in this case.
-That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
-\end{subfigure}
-\end{center}
-\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
-\end{figure}
-
-\subsubsection*{Numerical Stability}
-
-The EM algorithm, which we use here to fit our model, is notorious for convergence to local optima. 
-However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
-caused by such behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
-in addition running it using our default initialization procedure (see Implementation details). We then compared the largest
-log-likelihood achieved across all 11 runs with the log-likelihood achieved by the default run.
-When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference ($< 0.02$), 
-and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference ($<0.02$),
-and 6\% of runs showed an appreciable log-likelihood difference ($>1$), with the largest difference being 5.0. However, perhaps suprisingly, 
-even for this largest difference results from the default run (e.g.
-the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood.
-See \url{http://github.com/stephens999/ash/blob/master/dsc-robust/summarize_dsc_robust.rmd} for further details.
-
-\subsubsection*{The UA helps provide reliable estimates of $g$}
-
-An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
-can estimate the {\it size} of each effect $\beta_j$.
-Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
-to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
-threshold $T$.
-%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
-%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
-%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
-%towards zero, with the strength of the shrinkage being estimated from the data.
-Further, because the posterior distribution is, by definition,
-conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
-been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
-In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
-\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
-how can confidence intervals be assigned to the true
-effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
-and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
-%but also notes the difficulties of implementing a stable general algorithm.
-
-The ability of the EB approach to provide valid ``post-selection" interval estimates is extremely attractive in principle.
-But its usefulness in practice
-depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
-which are notoriously difficult in general. Indeed, Efron emphasises
-the difficulties of implementing a stable general algorithm, noting in his rejoinder
-``the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying
-to be overly nonparametric ... but it is hard to imagine a
-generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
-Our key point here is that the UA greatly simplifies the deconvolution problem.
-While not meeting Efron's desire for an entirely general nonparametric approach, we believe 
-that the UA can handle many cases of practical interest.
-
-To illustrate this, Figure \ref{fig:egcdf} compares the
-estimated $g$ from \ashr  with that from \mixfdr which does not make the UA 
-(and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
-The greater reliability of estimates afforded
-by the UA is immediately apparent. In particular the estimated cdf from \mixfdr often has an almost-vertical
-segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. 
-The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
-While the UA is not the only way to achieve this, we find it an attractive, simple and effective approach.
-
-Interestingly, even in the ``bimodal" scenario $\ashr$ 
-is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
-Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
-Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
-to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
-(\url{http://github.com/stephens999/ash/blob/master/dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
-clearly visible in the observed $\bhat$.
-
-We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ashr tends to systematically
-overestimate the mass of $g$ near zero. On careful inspection, this is apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
-and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf}b),
-and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
-under-estimating $\pi_0$ (Figure \ref{fig:egcdf}c).
-
-\begin{figure}[h!] 
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=2in]{figures/egcdf.pdf}
-	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf_nopen.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
-Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
-\end{subfigure}
-\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.} \label{fig:egcdf}
-\end{figure}
-
-%making the UA provides an attractive and simple way to achieve estimating $g$  large
-%subset of cases of interest (those where the UA can reasonably be assumed to hold).
-
-%In practice, the utility of the posterior distributions noted above depends primarily on the accuracy
-%of the estimate of the distribution $g(\cdot)$.
-
- %quote from efron rejoinder , p46:
-%MorrisÕ Section 3 is especially pertinent. His formula
-%for p(?|z) is related to my discussion of the
-%BenjaminiÐYekutieli False Coverage Rate method in
-%Section 7, particularly (7.2)Ð(7.4). Originally I had
-%hoped to develop an empirical Bayes method for estimating
-%such models, but the effort foundered on practical
-%difficulties involving the perils of deconvolution.
-%Section 6 on the Òone-group modelÓ is the ugly duckling
-%of the current paper, but it bears on some important
-%points raised in the discussion. Figure 7 concerns
-%a fuzzy version of simultaneous hypothesis testing,
-%where, as in MorrisÕ hospital example, the usual singlepoint
-%null hypotheses seem unequal to the task. The
-%development from (6.6) onwards, particularly (6.12),
-%bears on the possibility of nonnormal null distributions,
-%and is about as far as I can go in answering Professors
-%Rice and SpiegelhalterÕs penultimate question.
-%With g(?) a normal distribution, model (6.1) returns
-%us to the realm of Stein estimation, the scene of my
-%happy collaborations with Carl Morris. I continue to
-%be surprised at how much harder bumpy, nonnormal
-%models like (7.1) are to deal with. JamesÐStein estimation
-%works fine with, say, N = 10 component problems,
-%while the RobbinsÕ form of empirical Bayes appropriate
-%to (7.1) seems to require hundreds or thousands.
-%The information calculations in Efron (2008)
-%reinforce this gloomy assessment. Maybe I am trying
-%to be overly nonparametric in constructing the empirical
-%Bayes Fdr estimates, but it is hard to imagine a
-%generally satisfactory parametric formulation for (6.1).
-%Or perhaps it is just that hypothesis testing is more demanding
-%than estimation.
-%My Note: (6.1) is the general model (mu \sim g()).
-%7.1 is the case where the effects are a mixture of zero and N(-3,1).
-
-\subsubsection*{Calibration of posterior intervals}
-
-To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over 100 simulations in each Scenario). 
-Specifically we examine the empirical coverage of nominal lower 95\% credible bounds for a) all observations; b) significant negative discoveries; c) significant positive discoveries.
-We examine only lower bounds because the results for upper bounds follow by
-symmetry (except for the one asymmetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
-in each case: for negative discoveries the lower bound is typically large and negative and limits how big (in absolute value) 
-the effect could be; for positive discoveries the lower bound 
-is positive, and limits how small (in absolute value) the effect could be. Intuitively, the lower bound for negative discoveries depends on the accuracy of $g$ in its tail,
-whereas for positive discoveries it is more dependent on the accuracy of $g$ in the center.
-
-The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
-as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
-
-% As noted above, the penalty term (\ref{eqn:penalty}) causes systematic over-estimation of the mass of $g$ near 0,  which one would expect to overshrink interval estimates towards zero. 
-%The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound is negative and overshrinking (moving it too far towards zero) makes the left tail too big; conversely, for positive discoveries, the lower confidence bound is positive, and
-%overshrinking (moving it too far towards zero) makes the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
-%and be prepared to put up with some overshrinkage to avoid it.
-
-%Coverage rates are given in:
-%\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
-
-%Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all.tex}
-	\caption{All observations.
-Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario.
-This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero.
-Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (Table \ref{tab:nopen}).
-Removing the penalty in the half-uniform case is not recommended (see Appendix).}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg.tex}
-\caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
-and the normal-based method in the flat-top scenario.
-These results likely reflect inaccurate estimates of the tails of $g$ due to a disconnect between the tail of $g$ and the component distributions in these cases.
-For example, the uniform methods sometimes substantially underestimate the length of the tail of $g$ in these long-tailed scenarios,
-causing over-shrinkage of the tail toward 0.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos.tex}
-\caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds} \label{tab:coverage}
-\end{table}
-
-%
-%\subsubsection*{Reverse asymptotics}
-%
-%Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
-%
-%Note: we'd like to make sure it behaves ok in the case where most observations are null.
-%Perhaps simulate with 999 null and 1 alternative?
-%Also maybe look at performance for small sample sizes - e.g. n=1?
-%
-%Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
-%can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
-%by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
-%
-
-\subsection*{Differing measurement precision across units}
-
- We turn now to the second important component of our work: allowing for varying
- measurement precision across units. The key to this is the use of a likelihood,
- (\ref{eqn:normlik}) or (\ref{eqn:tlik}), that explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
-
-  To illustrate, we conduct a simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
- ($s_j=10$).  In both cases,
-we assume that half the effects are null
-and the other half are normally distributed with standard deviation 1:
-\begin{equation}
-p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1).
-\end{equation}
-In this setting, the poor-precision measurements $(s_j=10)$ tell us very little, and any sane analysis should effectively ignore them. 
- However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor}). This is because the poor measurements 
- produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
- which, when combined with the good-precision measurements, dilute the overall signal (e.g. they reduce the density of $p$ values near 0).
- This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
-the estimated error rates ($q$-values, or $\lfdr$ values) for the good-precision observations increase when the low-precision observations are included in the analysis
-(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
-for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).
-
-\begin{figure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_hist.pdf} 
-    \label{fig:goodpoor-hist}
-    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations}
- \end{subfigure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_scatter.pdf}
-    \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) of a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
-\end{subfigure}
-\caption{Simulation illustrating how, for existing FDR methods, 
-poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
-$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
-a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
-In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
-\end{figure}
-
- \subsubsection*{Reordering of significance, and the ``$p$ value prior"}
-
-Another consequence of accounting for differences in measurement precision is that \ashr may re-order the significance of the observations compared
- with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
- We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measuments that have the same $p$ value.
- The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
- and so the $\lfsr$ for these poor-precision measurements is always high.
-
-The potential for Bayesian analyses to re-order the significance of observations, and specifically to downweight imprecise observations,
-was previously discussed in \cite{guan.stephens.08}.
-However, Wakefield \cite{wakefield:2009} showed that, under a certain prior assumption, the Bayesian analysis produces the same ranking of significance as $p$ values (or their $z$ scores).
-He named this prior the ``$p$-value prior" because it can be thought of as the
-implicit prior assumption that is being made if we rank the significance of observations by their $p$ value.
-Wakefield's $p$-value prior assumes that the less precise effect estimates correspond to larger true effects, and specifically that they scale proportional
-to the standard errors $s_j$. More specifically still, it assumes a normal prior for the non-zero $\beta_j$ with mean 0 and variance $K s^2_j$ for some constant $K$.
-Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. Specifically, if, instead of assuming that $\beta_j$ is independent
-of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
-then the $\lfsr$ computed by ash.n provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
-as is illustrated in Figure \ref{fig:lfsr_pval}, right panel. (This result does not hold when using the mixtures of uniforms prior, ash.u.)
-
-This $p$-value prior assumes that the $z$ scores $z_j=\beta_j/s_j$
-are identically distributed, independent of $s_j$. This is essentially the assumption made, implicitly or explicitly, by existing methods -- like \locfdr, \mixfdr and \qvalue --
-that model the $z_j$ or $p_j$ directly.  In contrast, we have assumed up to now that the $\beta_j$ are independent of $s_j$. 
-We can set both these assumptions within a more general framework, which allows that $\beta_j/s^\alpha$ is independent of $s_j$ for some $\alpha$ [Equation (\ref{eqn:beta-alpha}))].
-Setting $\alpha=0$ implies that $\beta_j$ is
-independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
-effects (in absolute value). This latter assumption may often be qualitatively plausible: 
-for example, in gene expression studies the standard error for gene $j$ depends partly on
-the variance of its expression among samples, and genes with a larger variance may tend to be less tightly regulated and
-so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
-On the other hand, there is no particular reason to expect that either $\alpha=1$ or $\alpha=0$ will be the optimal choice.
-Indeed, optimal choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$,
-which will be dataset-specific. 
-Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
-comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
-likelihood-based methods can be used to select $\alpha$. For example, following the logic of the EB approach it
-would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
-which has been implemented in our software by C. Dai.
-
-%\begin{result}
-%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
-%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
-%\end{result}
-
-%To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
-%Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
-%In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
-%Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
-%which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
-%(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
-%may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)
-
-%\begin{result}
-%Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and using a mixture of normals for the prior distribution 
-%(\ref{eqn:beta-alpha}) with $\alpha=1$, 
-%the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$.
-Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$)
-%is montonic with $|t_j|$. 
-%\end{result}
-
-%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
-%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.
-
-%Most previous authors of EB approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
-%have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
-%In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
-%This is analogous
-%to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
-%standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
-%that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.
-
-\begin{figure}
-\centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
-\caption{Figure illustrating affects of prior assumptions on re-ordering of significance.
-Left panel shows results under our ``default prior" which assumes that effects $\beta_j$ are identically distributed, independent of $s_j$.
-Right panel shows results under the ``$p$-value prior", which assumes that $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$.} \label{fig:lfsr_pval}
-\end{figure}
-
-%\subsubsection*{Adaptive Shrinkage}
-%
-%Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
-%when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
-%the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
-% $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
-% In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
-% Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
-% and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
-%
-%Our methods we build on this idea in two ways. First,
-%by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
-%Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
-%adapts to measurement precision: more precise measurements undergo
-% less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
-%than observations with
-%small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
-%To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
-%Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
-%
-%\begin{figure}[!ht]
-%\begin{center}
-%\includegraphics[width=400px]{../examples/ash.pdf} 
-%\end{center}
-%\caption{Figure illustrating adaptive shrinkage.
-Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0.
-The standard error of each observation, $s_j$, was simulated from Inverse Gamma(5,5), which resulted in $s_j$ varying from 0.34 to 6.7.
-The shrunken estimates ($y$ axis) are plotted against the observed estimates ($x$ axis), with color indicating the (square root of the) precision of each measurement (blue=lower-precision; red=higher-precision).
-The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
-%\end{figure}
-
-%DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
-%To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
-Because the prior distribution is unimodal about zero,
-%the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
-%amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
-%For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
-%than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
-%small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
-
-%previous work based on normal means:
-%http://link.springer.com/article/10.1007/BF02562623
-
-%\subsection{Null behaviour}
-
-%Results of simulations under the null.
-
-%\subsection{Comparison of different underlying component distributions}
-%
-%Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
-%than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
-%likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
-%For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
-%increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
-%expect $L_{hu}$ to be significantly larger than the others.
-%
-%Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
-%
-%\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
-%\begin{center}
-%\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
-%%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
-%\end{center}
-%\caption{}
-%\end{figure}
-
-%Notes from Adrian Raftery - 
-%Newcombe (1886)
-%Dean and Raftery
-%Uniform is super-efficient
-
-\section*{Discussion}
-
-We have presented an Empirical Bayes approach to large-scale multiple testing that emphasises two ideas.
-First, we emphasise the potential benefits of using two numbers ($\bhat$, and its standard error)
-rather than just one number (a $p$ value or $z$ score)  to summarize the information on each test.
-While requiring two numbers is slightly more onorous than requiring one, in many settings these
-numbers are easily available and if so we argue it makes sense to use them.
-Second, we note the potential benefits -- both statistical and computational -- of assuming that the effects come from a unimodal distribution, and provide
-flexible implementations for performing inference under this assumption. We also introduce the ``false sign rate" as
-an alternative measure of error to the FDR, and illustrate its improved robustness to errors in model fit, particularly mis-estimation
-of the proportion of null tests, $\pi_0$.
-
-Multiple testing is often referred to as a ``problem" or a ``burden". In our opinion, EB approaches turn this idea on its head,
-treating multiple testing as an {\it opportunity}: specifically, an opportunity to learn about the prior distributions, and other modelling assumptions,
-to improve inference and make informed decisions about significance (see also \cite{greenland1991empirical}). This view
-also emphasises that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
-Indeed, the FDR at a given fixed threshold does not depend on the number of tests: as the number of tests increases, both the true positives and false positives increase linearly,
-and the FDR remains the same.
-(If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example.
-The EB approach captures this dependence in an intuitive way:
-if there are lots of strong signals then we infer $\pi_0$ to be small, and the estimated FDR (or $\lfdr$, or $\lfsr$) at a given threshold may be low, even if a large number of tests were performed; and conversely
-if there are no strong signals then we infer $\pi_0$ to be large and the FDR at the same threshold may be high, even if relatively few tests were performed. 
-More generally, overall signal strength is reflected in the estimated $g$, which in turn influences the estimated FDR.
-
-Two important practical issues that we have not addressed here are correlations among tests,
-and the potential for deviations from the theoretical null distributions of test statistics. These two
-issues are connected: specifically, unmeasured confounding factors can cause both correlations among tests
-and deviations from the theoretical null \cite{efron2007correlation,leek:2007}. And although there are certainly other
-factors that could cause dependence among tests, unmeasured confounders are perhaps the most worrisome in practice
-because they can induce strong correlations among large numbers of tests and profoundly impact results.
-Approaches to deal with unmeasured confounders can be largely divided into two types: those
-that simply attempt to correct for the resulting inflation of test statistics \cite{devlin1999genomic,efron2004large}, and those
-that attempt to infer counfounders using clustering, principal components analysis, or factor models 
-\cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
-and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable,
-they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alterntively, our methods
-could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:lik}) using
-$\bhat_j | \shat_j \sim N(0,\lambda_1 \shat_j + \lambda_2)$, where $\lambda_1,\lambda_2$ are to be estimated. However,
-this immediately raises issues of identifiability and we do not pursue the idea further here.
-
-Another important practical issue is the challenge of small sample sizes. For example, in genomics applications researchers sometimes
-attempt to identify differences between two conditions based on only a handful of samples in each. In such settings the
-normal likelihood approximation (\ref{eqn:normlik}) will be inadequate. And, although the $t$ likelihood (\ref{eqn:tlik}) partially
-addresses this issue, it is also, it turns out, not entirely satisfactory.  The root of the problem is that, with small sample sizes, raw estimated standard
-errors $\shat_j$ can be horribly variable. In genomics it is routine to address this issue by applying EB methods \cite{smyth:2004} to ``moderate" (i.e.~shrink) variance estimates,
-before computing $p$ values from ``moderated" test statistics. We are currently investigating how
-our methods should incorporate such ``moderated" variance estimates to make it applicable to small sample settings.
-
-Our approach involves compromises between flexibility, generality, and simplicity on the one hand, and statistical efficiency and principle on the other.
-For example, in using an EB approach that uses a point estimate for $g$, rather than a fully Bayes approach that accounts for uncertainty in $g$,
-we have opted for simplicity over statistical principle. And in summarizing every test by
-two numbers and making a normal or $t$ approximation to the likelihood, we have aimed to produce generic methods that 
- can be applied whenever such summary data are available -- just as \qvalue can be applied to any set of $p$ values for example -- although
- possibly at the expense of statistical efficiency compared with developing multiple tailored approaches based on context-specific likelihoods. 
- Any attempt to produce generic methods will involve compromise between generality and efficiency.
- In genomics, many analyses -- not only FDR-based analyses -- involve first computing a series of $p$ values before
- subjecting them to some further downstream analysis. An important message here is 
- that working with two numbers ($\bhat_j,\shat_j$), rather than one ($p_j$ or $z_j$), 
- can yield substantial gains in functionality (e.g.~ estimating effect sizes, as well as testing; accounting for variations in
-measurement precision across units) while losing only a little in generality.  We hope that our work will
-encourate development of methods that exploit this idea in other contexts.
-
-%This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
-%from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
-%with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
-
-%
-%
-%Both \locfdr and \mixfdr allow for deviations from
-%the theoretical null distribution by allowing that, under the null, the $z$ scores are $N(0,\sigma^2)$ with $\sigma$ estimated from the data rather
-%than fixed to 1. Within our approach this idea could be incorporated by changing the likelihood - for example, to assume $\bhat_j \sim N(0,s_ 
-%
-%These are tricky issues that, as far as we are aware, can affect
-%all approaches to FDR analysis. An important focus of Efron's work, also implemented in \locfdr, 
-% is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
-%theoretical null distribution. 
-%
-%For example, we and others have worked on Bayesian approaches to genetic association testing, using subjective prior specification of
-%effect sizes based on external data (e.g.~\cite{stephens:2009}). However, the large numbers of tests brings the opportunity to learn about
-%priors, and here we aim to exploit this
-
-%Choice of mixture components. Use of likelihood to compare different mixture components?
-
-%  
-%Directly modeling the p values, or z scores, say via non-parametric methods, 
-%can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-%distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-%The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-%Before proceeding, we feel obliged to correct the common misconception
-% that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
-% not depend on the number of tests. 
-%  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
-% 
-% In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
-
-\section*{Detailed Methods} \label{sec:implement}
-
-\subsection*{Embellishments}
-
-\subsubsection*{More flexible unimodal distributions}
-
-Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
-Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
- Therefore, we have implemented a more general approach based on 
-\begin{equation} \label{eqn:g-general}
-g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) ,
-\end{equation}
-where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
-\begin{enumerate}[(i)]
-\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$, \qquad (``ash.n")
-\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,  \qquad (``ash.u")
-\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,  \qquad (``ash.hu")
-\end{enumerate}
-where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
-(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
-and $\pi$ is a $2K+1$ vector that sums to 1.)
-The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
-Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
-under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
-(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
-(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
-however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)
-
-Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
-using a large dense grid of $\sigma^2_k$ or $a_k$, (i)-(iii) can respectively approximate,
-with arbitrary accuracy,
-\begin{enumerate}[(i)]
-\item any scale mixture of normals, which includes as special cases
-the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
-\item any symmetric unimodal distribution about 0.
-\item any unimodal distribution about 0.
-\end{enumerate}
-The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
-In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.
-
-Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
-The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
-See Implementation for details of our software defaults.
-
-%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting.
-For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
-%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
-
-\subsubsection*{Dependence of effects on standard errors}
-
-Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
-This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
- \begin{equation} \label{eqn:beta-alpha}
- \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
- \end{equation}
-for any $\alpha$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
-assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
-it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
-the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.
-
-The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
-To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
- standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
- Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
-yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
-%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
-%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
-%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
-
-%Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
-%how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
-%it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
-%rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
-%and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
-%this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
-%do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
-
-\subsubsection*{Replace normal likelihood with $t$ likelihood}
-
-We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
-  \begin{equation} \label{eqn:tlik}
- \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
- \end{equation}
-where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
-and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
-The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
-This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms; see Implementation below.
-When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
-
-Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
-\begin{equation} \label{eqn:tdist}
-(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
-\end{equation}
-However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
-In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
-currently pursuing this improved approach (joint work with M.Lu) and results will be published elsewhere. 
-%We note here that
-%modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
-%where our simpler approach can produce unsatisfactory results.
-
-\subsubsection*{Non-zero mode}
-
-An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood, rather than fixed to 0.
-
-\subsection*{Implementation Details}
-
-\subsubsection*{Likelihood for $\pi$}
-
-We define the likelihood for $\pi$ to be the probability of the observed data $\bhat$ conditional on $\shat$: $L(\pi) := p(\bhat | \shat,\pi),$ which
-by our conditional indendence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$.
-[One might prefer to define the likelihood as $p(\bhat, \shat | \pi) = p(\bhat | \shat, \pi) p(\shat | \pi)$, in which case our definition comes down to assuming that the term $p(\shat | \pi)$ does not depend on $\pi$.]
-
-Using the prior $\beta_j \sim \sum_{k=0}^K \pi_k f_k(\beta_j)$ given by (\ref{eqn:g-general}), and the normal likelihood (\ref{eqn:normlik}), integrating over $\beta_j$ yields
-\begin{equation}
-p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
-\end{equation}
-where 
-\begin{equation}
-\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
-\end{equation}
-denotes the convolution of $f_k$ with a normal density.
-These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
-Specifically, 
-\begin{equation} \label{eqn:uconv}
-\tilde{f}_k(\bhat_j)  = 
-\begin{cases}
-N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
-\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
-\end{cases}
-\end{equation}
-where $\Psi$ denotes the cumulative distribution function (c.d.f.) of the standard normal distribution.
-If we replace the normal likelihood with the $t_\nu$ likelihood (\ref{eqn:tlik}) then the convolution 
-for $f_k$ uniform the convolution is still given by (\ref{eqn:uconv}) but with $\Psi$ the c.d.f. 
-of the $t_\nu$ distribution function. (The convolution for $f_k$ normal is tricky and we have not implemented it.)
-
-\subsubsection*{Penalty term on $\pi$}
-
-To make $\lfdr$ and $\lfsr$ estimates from our method ``conservative" we add a penalty 
-term $log(h(\pi;\lambda))$ to the log-likelihood $\log L(\pi)$ to encourage over-estimation of $\pi_0$:
-\begin{equation} \label{eqn:penalty}
-h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
-\end{equation}
-where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
-consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0sims}).
-
-Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
-we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
-
-\subsubsection*{Problems with removing the penalty term in the half-uniform case}
-
-It is straightforward to remove the penalty term by setting $\lambda_k=1$ in (\ref{eqn:penalty}).
-We note here an unanticipated problem we came across when using no penalty term in the half-uniform case 
-(i.e.~ $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$ in (\ref{eqn:g-general})): when the data
-are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
-sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
-for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
-and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
-
-\subsubsection*{EM algorithm}
-
-With this in place, the penalized log-likelihood for $\pi$ is given by:
-\begin{equation}
-\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
-\end{equation}
-where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This can be maximized using an EM algorithm \cite{dempster77}, whose
-one-step updates are:
-\begin{align} \label{eqn:w}
-w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
-n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
-\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}
-\end{align}
-Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
-and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
-probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$. (The $w_{kj}$ are sometimes referred to as the ``responsibilities".)
-Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
-We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple} to accelerate convergence of this EM algorithm.
-
-%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
-\subsubsection*{Initialization}
-
-By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
-Our rationale for initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
-but weak signal in the data cannot quickly draw the algorithm towards the null.
-
-In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
-$\pi_0=1; \pi_k=0 (k>0)$. If not then we replace the EM solution with the global null solution. This aims to guard against errors due to convergence to a local optimum when the
-data are consistent with the global null.
-
-\subsubsection*{Conditional distributions}
-
-Given $\hat\pi$, we compute the conditional distributions 
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
-\end{equation} 
-Each posterior is a mixture on $K+1$ components:
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
-\end{equation}
-where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
-and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
-$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
-All these posterior distributions are easily available.
-For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
-If $f_k$ is normal and $L$ is normal, then this is a normal distribution.
-
-\subsubsection*{Choice of grid for $\sigma_k, a_k$} \label{sec:grid}
-
-\def\sigmamax{\sigma_{\text{max}}}
-\def\sigmamin{\sigma_{\text{min}}}
-When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
-in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
-as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
-based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
-is negative we set $\sigmamax = 8\sigmamin$.)
-
-When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.
-
-Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
-a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.
-
-\appendix
-
-\begin{table}[!ht]
-\centering\begin{tabular}{c c } \toprule
-Scenario & Alternative distribution, $g_1$  \\ \midrule
-spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
-near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
-flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
- &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
-skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
-big-normal & $N(0,4^2)$ \\ 
-bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
-\end{tabular}
-\caption{Summary of simulation scenarios considered} \label{table:scenarios}
-\end{table}
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all_nopen.tex}
-	\caption{All observations}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg_nopen.tex}
-\caption{``Significant" negative discoveries.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos_nopen.tex}
-\caption{``Significant" positive discoveries.}
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{table:nopen}
-\end{table}
-%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
-%\begin{equation}
-%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
-%\end{equation}
-%which is close to the definition in \cite{storey.03}.
-
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgements}
-
-Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
-prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} , and includes contributions 
-from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.
-
-This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.
-
-%\section*{References}
-% The bibtex filename
-\bibliography{/Users/stephens/Dropbox/Documents/mainbib}
-
-\section*{Figure Legends}
-
-\clearpage
-
-\section*{Tables}
-
-\section*{Supporting Information Legends}
-
-Supplementary material can be found in {\bf Supplementary Information S1.}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
-\end{document}

### 2015-10-06 figures/decomp_ZA/caption.tex
New Article!

--- 2015-10-06 paper/main.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
-\documentclass[11pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-%\usepackage{natbib}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-\usepackage{booktabs}
-\usepackage{xspace}
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-\usepackage{sectsty}
-\subsectionfont{\normalsize}
-\subsubsectionfont{\mdseries\itshape}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\usepackage{color}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-% Use the PLoS provided bibtex style
-\bibliographystyle{/Users/stephens/Dropbox/Documents/stylefiles/plos2009}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
-\pagestyle{myheadings}
-%% ** EDIT HERE **
-\usepackage{enumerate}
-\usepackage{multirow} 
-\usepackage{url}
-\usepackage{xr} %for cross-referencing
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-\newtheorem{algorithm}{Algorithm}
-\newtheorem{proposition}{Proposition}
-\newtheorem{restateproposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{result}{Result}
-\newtheorem{note}{Note}
-\newtheorem{definition}{Definition}
-
-\def\lfdr{\textit{lfdr}}
-\def\lfsr{\textit{lfsr}}
-\def\x{\mbf{x}}
-\def\y{\mbf{y}}
-\def\e{\mbf{e}}
-\def\g{\mbf{g}}
-\def\p{p}
-\def\Yb{\hat{Y}_1^{\text{BAYES}}}
-\def\Xperp{X_{1 \perp 0}}
-\def\Pperp{P_{1 \perp 0}}
-\def\Pperpb{\Pperp^{\text{B}}}
-\def \hatd { \widehat{ d } }
-\def \dhatin {\hatd^{\mbox{\tiny IN}}}
-\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
-\def \mvn{\text{MVN}}
-\def\mn{\text{MN}}
-\def \iw{\text{W}^{-1}}
-\def \fp {fastPHASE\ }
-\def\bhat{\hat{\beta}}
-\def\shat{\hat{s}}
-\def\var{\rm Var}
-\def\mvr{\text{BMVR}}
-\def\bfuni{\text{BF}_\text{uni}}
-\def\one{\mbf{1}}
-\def\zero{\mbf{0}}
-\def\hall{H_{\text{all}}}
-\def\pall{\p_{\text{all}}}
-\def\bfall{\text{BF}_\text{all}}
-\def\bfav{\text{BF}_\text{av}}
-\def\gall{\gamma_\text{all}}
-\def\limbfall{\bfall^\rightarrow}
-\def\limbfav{\bfav^\rightarrow}
-\def\limbfgamma{\BF_\gamma^\rightarrow}
-\def\sxx{S_{xx}}
-\def\syy{S_{yy}}
-\def\vyy{V_{yy}}
-\def\vyx{V_{yx}}
-\def\vxx{{V_{xx}}}
-\def\rss{\text{\rm RSS}}
-\def\dss{\text{\rm $\Delta$SS}}
-\def\U{U}
-\def\D{D}
-\def\I{I}
-\def\luu{L_{UU}}
-\def\indep{\perp}
-\def\qvalue{{\tt qvalue}\xspace}
-\def\locfdr{{\tt locfdr}\xspace}
-\def\mixfdr{{\tt mixfdr}\xspace}
-\def\ashr{{\tt ashr}\xspace}
-
-\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
-                                % mode too
-\newcommand\height{{\it height}}
-\newcommand\weight{{\it weight}}
-\def\BF{\text{BF}}
-
-%% END MACROS SECTION
-%\externaldocument{plos_one_rev1_suppinfo}
-
-\begin{document}
-
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{False Discovery Rates: A New Deal}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Matthew Stephens$^{1*}$, 
-\\
-\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
-\\
-$\ast$ E-mail: Corresponding mstephens@uchicago.edu
-\end{flushleft}
-
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-%\section*{Author Summary}
-
-\section*{Introduction}
-
-Since its introduction in 1995 by Benjamini and Hochberg 
-\cite{benjamini1995controlling}, the ``False Discovery Rate" (FDR) has quickly established itself
-as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing
-in which the goal is to identify the non-zero ``effects" among a large number of imprecisely-measured effects.
-
-Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
-indeed, the notion that EB approaches could be helpful in handling multiple comparisons predates introduction of
-the FDR  (e.g.~\cite{greenland1991empirical}). More recently, EB approaches to the FDR
-have been extensively studied by several authors, 
-especially B.~Efron and co-authors 
-\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
-see also \cite{kendziorski2003parametric,newton2004detecting, 
-datta2005empirical,muralidharan2010empirical} for example.
-So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
-existing widely-used FDR pipelines) and can substantially affect
-inference. The first idea is to {\it assume that the distribution of effects is unimodal}. This yields a very simple, fast, and stable computer 
-implementation, as well as improving inference of FDR when the unimodal assumption is correct.  
-The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
-to summarize each measurement. This idea allows variations in measurement precision to be better accounted for,
- and avoids a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
-
-In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
-the idea that it may be preferable to focus on estimation rather than on testing.
-In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
-simply estimation with some positive prior probability that the effect is exactly zero.
-However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
-perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
-controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
-``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
-in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
-the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the lfsr, particularly its increased robustness to modelling assumptions.
-
-%terminology
-%``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
-%and correspondingly define the ``false signed discovery rate", or simply ``false sign rate" (FSR), in a natural way.
-
-%The methods we provide here can address both problems, and we provide results that highlight important benefits of focussing on estimation, and particularly on assessing the correctness of the {\it sign} of the estimated effects, rather than on testing for non-zero effects.
-
-Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
-(We address the reasons for this name, and connections with shrinkage analysis, in the Discussion.)
-
-%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.
-
-%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-%measurements are made on only a small number of
-%samples in each condition - sometimes as few as one
-%sample in each condition. Thus the error in estimates
-%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-%in estimates of $\beta_j$ still greater.
-%
-
-%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
-%Here we present new tools that bring two important ideas to bear on the problem.
-%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
-%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
-%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
-%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
-%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
-%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
-%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
-%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
-%a substantial increase in the number of discoveries at a given FDR).
-
-%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
-%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
-Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
-%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
-%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
-%generic methods that work directly with two numbers for each 
-%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
-Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
-%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
-
-\def\df{df}
-\def\FDR{\text{FDR}}
-\def\fdr{\text{\lfdr}}
-\def\FSR{\text{FSR}}
-\def\fsr{\text{lfsr}}
-
- \section*{Methods}
-
- \subsection*{Model Outine}
-
-% Here we consider one setting in which FDR-based methods are widely used:
-%an experimenter has measured a large numbers of ``effects", each with an associated
-%measurement precision (standard error), 
-%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
-%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
-%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
-%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
-%For example, in genomics
-%a common goal is to identify ``differentially-expressed genes" -- that is, genes
-%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
-%mean expression level between the two conditions, often referred to as the ``log-fold change". 
-Here we describe the simplest version of our method, and briefly discuss embellishments we have also implemented.
-
-Suppose that we are interested in the values of $J$ ``effects" $\beta=(\beta_1,\dots,\beta_J)$. For example, in a typical genomics application
-that aims to identify differentially expressed genes,  $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
-In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
-that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
-more general problem of estimating, and assessing uncertainty in, $\beta_j$.
-
-Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
-and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
-which by Bayes theorem can be written as 
-\begin{equation}
-p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
-\end{equation}
-
-For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
-This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
-A simple way to implement the UA is to assume that $g$ is
-a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
-\begin{align} \label{eqn:beta}
-p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
-g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
-\end{align}
-where $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyperparameters, which are non-negative and sum to one, and are to be estimated,
-while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
-represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
-(We encourate the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
-although of course in practice we use a finite grid -- see Implementation Details.)
-
-For the likelihood $p(\bhat | \beta, \shat)$ we assume  
-\begin{equation} \label{eqn:normlik}
-p(\bhat | \beta, \shat) = \prod_j N(\bhat_j; \beta_j, \shat_j^2).
-\end{equation}
-Here, in additional to some conditional independence assumptions, we are 
-effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-
-This simple model features both the key ideas we want to emphasise in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
-the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
-with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
-However, this simple model also has several additional assumptions that can be relaxed. Specifically,
-\begin{enumerate}
-\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) also implies that $g$ is symmetric about 0; more flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see (\ref{eqn:g-general})].
-\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
-We can relax this to allow for a relationship between these quantities [see (\ref{eqn:beta-alpha})].
-\item The likelihood (\ref{eqn:normlik}) effectively assumes that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-We can generalize this likelihood using a $t$ likelihood [see (\ref{eqn:tlik})].
-\end{enumerate}
-These embellishments are detailed in Embellishments and Implementation Details.
-Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
-(which are also made by most existing EB approaches to this problem). Correlations among tests certainly arise in practice, either due to genuine correlations
-in the system of study, or due to unmeasured confounders, and their potential to impact results of an FDR analysis is important to consider
-whatever analysis methods are used: see \cite{efron2007correlation,leek:2007} for relevant discussion.
-
-%We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
-%$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
-%\begin{equation}
-%p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
-%\end{equation}
-%Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
-%\begin{equation} \label{eqn:normlik}
-%p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
-%
-%We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
-%as a mixture of null and alternative components:
-%\begin{equation} \label{eqn:betaj0}
-%g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
-%\end{equation}
-%where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
-%We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
-%\begin{equation} \label{eqn:beta-iid}
-%p(\beta | \shat) = \prod_j g(\beta_j).
-%\end{equation}
-%
-%Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
-%with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
-%to use a mixture of zero-mean normal distributions:
-%\begin{equation} \label{eqn:mixnorm}
-%g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
-%Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
-%represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
-% The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
-% more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below. 
-% 
-% Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
-% yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
-% \begin{equation} \label{eqn:g}
-% g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
-% \end{equation}
-% 
-% Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
-% \begin{align} \label{eqn:betaj}
-% p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
-% p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
-% \end{align}
-% with $g$ given by the mixture of normals $(\ref{eqn:g})$.
-
-\subsection*{Fitting the model}
-
-In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
-each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
-Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
-\begin{equation} \label{eqn:marginallik}
-p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
-\end{equation}
-where we define $\sigma_0:=0$.
-
-The usual EB approach to fitting this model would involve two simple steps:
-\begin{enumerate}
-\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
-\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
-against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | | \bhat, \shat,\hat{\pi})$.
-\end{enumerate}
-Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(The simplicity of the EM algorithm in step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
-which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
-
-Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximise a penalized
-likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
-We introduce this penalty because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
-
-\subsection*{The local False Discovery Rate and local False Sign Rate}
-
-As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
-In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
-One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
-the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
-In other words it is the posterior probability that $\beta_j$ is actually zero:
-\begin{equation} \label{eqn:fdr}
-\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
-\end{equation}
-
-The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
-whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
-the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey1991philosophy}) argued that
-``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$.
-Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
-\begin{quote}
-...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
-\end{quote}
-Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
-focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.
-
-Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
-would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
-\begin{equation} \label{eqn:lfsr}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
-\end{equation}
-To illustrate, suppose that 
-\begin{gather*}
-p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
-p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
-p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
-\end{gather*} 
-Then 
-from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
-our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.
-
-As our notation suggests, $\lfsr_j$ is intended to be compared and contrasted with $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
-small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
-Of coure, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
-this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
-(this follows from the definition because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
-In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
-$\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.
-
-From these ``local" measures of significance, we can also compute average error rates over subsets of observations
-$\Gamma \subset \{1,\dots,J\}$. For example,   
-\begin{equation} \label{eqn:FDRhat}
-\widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
-\end{equation}
-estimates the FDR we would obtain if we were to declare all tests in $\Gamma$ significant.
-And 
-\begin{equation} \label{eqn:qvalue}
-q_j := \widehat \FDR(\{k: \fdr_k \leq \fdr_j\}) 
-\end{equation}
-provides a measure of significance analogous to Storey's $q$ value \cite{storey.03}.
-
-%Because our methods involve explicitly modelling effects $\beta_j$, rather than 
-%$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
-
-%Indeed, even in settings where some
-%$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
-%analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
-
-\subsection*{Related work}
-
-\subsubsection*{Previous approaches focussed on FDR}
-
-Among previous methods that explicitly consider the FDR and related quantities,
-our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
- (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
-  %Here we highlight the key differences between our approach and these three existing approaches.
-
-There are two key differences between our approach and all of these three existing methods. First, 
-whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
-we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
-for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
-Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 
-how it can improve performance by taking better account of variation in measurement precision among observations.
-%(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
-% Or more generally Section xx illustrates the potential practical effects
-
-%allows for different modelling assumptions to be compared through likelihood comparisons.
-Second, our unimodal assumption (UA) that the effects are unimodal about zero is quite different from assumptions made by $\qvalue, \locfdr$ or $\mixfdr$.
-Indeed, $\locfdr$ assumes that all $z$ scores near 0 are null (Efron calls this the Zero Assumption; ZA), which implies that 
-under the alternative hypothesis the distribution of $z$ scores has {\it no mass at 0}; this contrasts strikingly with the UA, which implies that this distribution has its peak at 0! 
-Similarly, $\qvalue$ assumes that all $p$ values near 1 are null,
-which is the same as the ZA because $p$ values near 1 correspond to $z$ scores near 0. And although \mixfdr does not formally make the ZA,
-we have found that in practice, with default settings, the results often approximately satisfy the ZA (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}).
-Thus, not only do these existing methods not make the UA, they actually make assumptions that are, in some sense, as different
-from the UA as they can be.
-
-Given that the UA and ZA are so different, it seems worth discussing why we generally favor the UA. 
-Although the UA will not apply to all situations, we believe that it will often be reasonable, especially in FDR-related contexts
-that have traditionally focussed on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis,  
-it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible. Further, it seems reasonable to expect that larger effects become
-decreasingly plausible, and so the distribution of the effects will be unimodal about 0.
-To paraphrase Tukey, ``All we know about the world teaches us that large effects are rare, whereas small effects abound."
-We emphasise that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
-It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
-
-%Tukey "peri-null"
-
-%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
-%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}.
-
-In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficents - 
-common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
-These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
-it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
-The UA assumption on regression coefficients is directly analagous to our UA here, and so its
-widespread use in the regression context supports its use here.
-
-Alternatively, we could motivate the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
-such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
-make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).
-
-Finally, the UA also has a considerable practical benefit: it yields simple algorithms that are both computationally
-and statistically stable. We illustrate these features in Results.
-
-\subsubsection*{Other work}
-
-There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
- Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
-indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
-This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
-Also, they provide no software implementation.
-More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
-Some pointers to recent papers in which other relevant
-citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}.
-Much of the literature focusses on the homoskedastic case (i.e.
-$\shat_j$ all equal) whereas we allow for heteroskedasticity.
-And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential.
-Several recent papers consider more flexible non-parametric assumptions on
-$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
-estimate (NPMLE) for $g$.
-These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive.
-However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
-may perform well for point estimation, it seems possible it will not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
-Indeed, one way of thinking about the UA is as a way to regularize $g$.
-
-%The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
-%encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-%$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-%$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-%the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-%observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-%corresondingly little effect on the estimated FDR of other observations. 
-%We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-%$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-%that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-%estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-%Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-\def\dt{\text{dt}}
-
-%
-%assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
-%$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
-% $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
-%  \begin{equation} \label{eqn:lik}
-% \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
-% \end{equation}
-% where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e.
-the density of $\mu+\sigma T_\nu$).
-In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
-%  \begin{equation} \label{eqn:normlik}
-% \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
-% \end{equation}
-%which, in light of (\ref{eqn:betahat}), is equivalent to
- % \begin{equation} \label{eqn:lik2}
-% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
- %\end{equation}
-
-%The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
-they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
-The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$.
-Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$.
-If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$.
-See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.
-
-%This likelihood approximation can be motivated by assuming that i) $\bhat$
-% $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
-% $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$. 
-%
-
-%Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.
-
-%While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
- %that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
-%there is an important difference: here,
-
-%The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
-For example, we can write $p(\bhat_j, \shat_j | \beta_j) = p(\bhat_j | \shat_j, \beta_j)p(\shat_j | \beta_j)$.
-Now, if we are willing to assume that $\shat_j$ alone contains no information about $\beta_j$, or equivalently that $p(\shat_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\bhat_j | \shat_j, \beta_j)$, and if $\bhat_j | \shat_j, \beta_j \sim N(\beta_j, \shat_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.
-
-\section*{Results}
-
-We compare results of \ashr with existing FDR-based methods implemented
-in the R packages \qvalue (v2.0 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
-In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
-to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.
-
-\subsection*{Effects of the Unimodal Assumption}
-
-Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
-standard error, $s_j=1$ and all methods are provided that information. That is,
-$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$.
-In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the only difference between our method and methods like $\locfdr$ and $\mixfdr$ are in how they estimate $g$.
-
-To briefly summarize the results in this section:
-\begin{enumerate}
-\item The UA can produce very different inferences compared with the ZA made by existing methods.
-\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
-\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
-\end{enumerate}
-
-\subsubsection*{The UA and ZA can produce different inferences}
-
-\begin{figure}
-\center \includegraphics[height=6in]{figures/decomp_ZA.pdf}
-\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
+\label{fig:ZA} Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
 The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
@@
 In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
-(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
-\end{figure}
-
-To illustrate the different inferences from the UA and ZA we show results for a single
-dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
-but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
-\qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
-The results (Figure \ref{fig:ZA}) illustrate the clear difference between the existing methods and our method. The effects of the ZA made by \qvalue and \locfdr
-are visually clear, producing a ``hole"  in the alternative $z$ score distribution around 0. Although \mixfdr does not formally make the ZA, its decomposition
-exhibits a similar hole. In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
-effectively ``filling in" the hole.  (Of course the null distribution also has a peak at 0, and the local fdr 
- under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)
-
-Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA.
-Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
-This would increase the null component (dark blue) at the expense of
-the alternative component (light blue).
-Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0.
-The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears.
-In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
-See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
-demonstration.
-
- \subsubsection*{The UA can produce conservative estimates of $\pi_0$}
-
-The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
-Consequently \ashr will estimate smaller $\lfdr$s and FDRs than existing methods that make the ZA. 
-This is desirable, provided that these estimates remain conservative: that is,
-provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
-The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behaviour. To check its effectiveness
-we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
-for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
-They range from a ``spiky" distribution -- where many non-zero $\beta$ are
-too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
-flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
-making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
-which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.
-
-% How do these differences influence estimates of the FDR? To address this, note that  -- in this simple case where the 
-% $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
-% of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
-% strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
-% although the methods considered here differ from one another in many ways, having estimated $\pi_0$
-% they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
-% empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
-%
-
-%Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
-% and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
-%Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
-%At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
-%overestimates of $\pi_0$, which will yield overestimates of the FDR).
-
-\begin{figure} 
-\begin{center}
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=1.8in]{figures/scenarios_density.pdf} 
-	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/pi0_est.pdf} 
-\caption{Comparison of true and estimated values of $\pi_0$.
-When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ashr being least conservative, and hence most accurate.
-When the UA does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.} \label{fig:pi0sims}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfdr_est.png} 
-\caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
-Black line is $y=x$ and red line is $y=2x$.
-Estimates of $\lfdr$ are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est.png} 
-\caption{As in c), but for $\lfsr$ instead of $\lfdr$.
-Estimates of $\lfsr$ are consistently less conservative than $\lfdr$ when UA holds, and also less anti-conservative in bimodal scenario.}  \label{fig:lfsr}
-\end{subfigure}
-\end{center}
-\caption{Results of simulation studies (constant precision $s_j=1$).}
-\end{figure}
-
-For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
-data as follows:
-\begin{enumerate}
-\item Simulate $\pi_0  \sim U[0,1]$.
-\item For $j=1,\dots,J$, simulate $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)$.
-\item For $j=1,\dots,J$, simulate $\bhat_j | \beta_j \sim N(\beta_j,1)$.
-\end{enumerate}
-%Thus these simulations assume the same
-%precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).
-
-Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ashr ($y$ axis) with the true values ($x$ axis). 
-For \ashr we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
-and as a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
-as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
-The results show that \ashr provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
-in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ashr estimates can be slightly anti-conservative.
-We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
-are used. (In addition, the effects on $\lfsr$ estimates turn out to be relatively modest; see below).
-
-%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
-%
-%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
-%and so all effects are non-zero and no nulls hold. This scenario is rather different
-%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
-%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
-%``mostly non-null"
-%scenario, and we would agree! On the other hand, the distribution of $p$ values
-%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
-%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
-%Thus it seems desirable to have methods that perform well in both cases, and this is
-%an important feature of the methods we introduce here.
-%
-%TODO: put FDRs on graph? Simulate with 50\% null?
-
-\subsubsection*{The $\lfsr$ is more robust than \lfdr}
-
-The results above show that \ashr can improve on existing methods in producing smaller, more accurate, 
-estimates of $\pi_0$, which will lead to more accurate estimates of FDR.
-Nonetheless, in many scenarios \ashr continues
-to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
-This is because these scenarios include 
-an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
-estimation of $\pi_0$ impossible. Put another way, and as is well known, $\pi_0$ is not identifiable:
- the data can effectively provide an upper bound on plausible values of $\pi_0$,
-but not a lower bound (because the data cannot rule out that everything is non-null, but with miniscule effects).
-To obtain conservative behaviour we must estimate $\pi_0$ by this upper bound, which can
-be substantially larger than the true value.
-
-Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
-overestimation of $\pi_0$ is corresponding overestimation of FDR (and $\lfdr$, and $q$ values).
-To illustrate, Figure \ref{fig:lfdr} compares the estimated $\lfdr$ from ash.n with the true value (computed using
-Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios which involve many
-non-zero effects that are very near 0 (e.g.~the spiky scenario with $\pi_0$ small) where
-$\pi_0$ can be grossly overestimated.  
-(Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.)
-
-The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
- become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
- This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
-To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
-be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
-the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in $\lfsr$ than $\lfdr$.
-
-Note that, compared with previous debates regarding testing, 
-this section advances an additional reason for focussing on the sign of the effect, rather than just testing whether it is 0. 
-In previous debates authors have argued against testing whether an effect is 0
-because it is {\it implausible that effects are exactly 0}. Here we add that {\it even if one believes
-that some effects may be exactly zero}, it is still better to focus on the sign, because generally {\it the data are more informative about that question}
-and so inferences are more robust to, say, the inevitable mis-estimation of $\pi_0$.
-To provide some intuition, consider an observation with a $z$ score of 0. The $\lfdr$ of this observation can range from 0 (if $\pi_0=0$)
-to 1 (if $\pi_0=1$). But, assuming a symmetric $g$, the $\lfsr>0.5$ whatever the value of $\pi_0$, because the observation $z=0$ says
-nothing about the sign of the effect.
-Thus, are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
-whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.
-
-Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
-Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
-if we are prepared to assume that no effects are exactly zero, then removing the point mass 
-yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
-However, there is ``no free lunch":  if in fact some effects are exactly zero
-then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
-We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.
-
-\begin{figure}
-\begin{center}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s_nn.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be more accurate in this case.
-That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
-\end{subfigure}
-\end{center}
-\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
-\end{figure}
-
-\subsubsection*{Numerical Stability}
-
-The EM algorithm, which we use here to fit our model, is notorious for convergence to local optima. 
-However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
-caused by such behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
-in addition running it using our default initialization procedure (see Implementation details). We then compared the largest
-log-likelihood achieved across all 11 runs with the log-likelihood achieved by the default run.
-When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference ($< 0.02$), 
-and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference ($<0.02$),
-and 6\% of runs showed an appreciable log-likelihood difference ($>1$), with the largest difference being 5.0. However, perhaps suprisingly, 
-even for this largest difference results from the default run (e.g.
-the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood.
-See \url{http://github.com/stephens999/ash/blob/master/dsc-robust/summarize_dsc_robust.rmd} for further details.
-
-\subsubsection*{The UA helps provide reliable estimates of $g$}
-
-An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
-can estimate the {\it size} of each effect $\beta_j$.
-Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
-to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
-threshold $T$.
-%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
-%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
-%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
-%towards zero, with the strength of the shrinkage being estimated from the data.
-Further, because the posterior distribution is, by definition,
-conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
-been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
-In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
-\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
-how can confidence intervals be assigned to the true
-effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
-and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
-%but also notes the difficulties of implementing a stable general algorithm.
-
-The ability of the EB approach to provide valid ``post-selection" interval estimates is extremely attractive in principle.
-But its usefulness in practice
-depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
-which are notoriously difficult in general. Indeed, Efron emphasises
-the difficulties of implementing a stable general algorithm, noting in his rejoinder
-``the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying
-to be overly nonparametric ... but it is hard to imagine a
-generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
-Our key point here is that the UA greatly simplifies the deconvolution problem.
-While not meeting Efron's desire for an entirely general nonparametric approach, we believe 
-that the UA can handle many cases of practical interest.
-
-To illustrate this, Figure \ref{fig:egcdf} compares the
-estimated $g$ from \ashr  with that from \mixfdr which does not make the UA 
-(and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
-The greater reliability of estimates afforded
-by the UA is immediately apparent. In particular the estimated cdf from \mixfdr often has an almost-vertical
-segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. 
-The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
-While the UA is not the only way to achieve this, we find it an attractive, simple and effective approach.
-
-Interestingly, even in the ``bimodal" scenario $\ashr$ 
-is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
-Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
-Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
-to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
-(\url{http://github.com/stephens999/ash/blob/master/dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
-clearly visible in the observed $\bhat$.
-
-We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ashr tends to systematically
-overestimate the mass of $g$ near zero. On careful inspection, this is apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
-and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf}b),
-and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
-under-estimating $\pi_0$ (Figure \ref{fig:egcdf}c).
-
-\begin{figure}[h!] 
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=2in]{figures/egcdf.pdf}
-	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf_nopen.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
-Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
-\end{subfigure}
-\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.} \label{fig:egcdf}
-\end{figure}
-
-%making the UA provides an attractive and simple way to achieve estimating $g$  large
-%subset of cases of interest (those where the UA can reasonably be assumed to hold).
-
-%In practice, the utility of the posterior distributions noted above depends primarily on the accuracy
-%of the estimate of the distribution $g(\cdot)$.
-
- %quote from efron rejoinder , p46:
-%MorrisÕ Section 3 is especially pertinent. His formula
-%for p(?|z) is related to my discussion of the
-%BenjaminiÐYekutieli False Coverage Rate method in
-%Section 7, particularly (7.2)Ð(7.4). Originally I had
-%hoped to develop an empirical Bayes method for estimating
-%such models, but the effort foundered on practical
-%difficulties involving the perils of deconvolution.
-%Section 6 on the Òone-group modelÓ is the ugly duckling
-%of the current paper, but it bears on some important
-%points raised in the discussion. Figure 7 concerns
-%a fuzzy version of simultaneous hypothesis testing,
-%where, as in MorrisÕ hospital example, the usual singlepoint
-%null hypotheses seem unequal to the task. The
-%development from (6.6) onwards, particularly (6.12),
-%bears on the possibility of nonnormal null distributions,
-%and is about as far as I can go in answering Professors
-%Rice and SpiegelhalterÕs penultimate question.
-%With g(?) a normal distribution, model (6.1) returns
-%us to the realm of Stein estimation, the scene of my
-%happy collaborations with Carl Morris. I continue to
-%be surprised at how much harder bumpy, nonnormal
-%models like (7.1) are to deal with. JamesÐStein estimation
-%works fine with, say, N = 10 component problems,
-%while the RobbinsÕ form of empirical Bayes appropriate
-%to (7.1) seems to require hundreds or thousands.
-%The information calculations in Efron (2008)
-%reinforce this gloomy assessment. Maybe I am trying
-%to be overly nonparametric in constructing the empirical
-%Bayes Fdr estimates, but it is hard to imagine a
-%generally satisfactory parametric formulation for (6.1).
-%Or perhaps it is just that hypothesis testing is more demanding
-%than estimation.
-%My Note: (6.1) is the general model (mu \sim g()).
-%7.1 is the case where the effects are a mixture of zero and N(-3,1).
-
-\subsubsection*{Calibration of posterior intervals}
-
-To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over 100 simulations in each Scenario). 
-Specifically we examine the empirical coverage of nominal lower 95\% credible bounds for a) all observations; b) significant negative discoveries; c) significant positive discoveries.
-We examine only lower bounds because the results for upper bounds follow by
-symmetry (except for the one asymmetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
-in each case: for negative discoveries the lower bound is typically large and negative and limits how big (in absolute value) 
-the effect could be; for positive discoveries the lower bound 
-is positive, and limits how small (in absolute value) the effect could be. Intuitively, the lower bound for negative discoveries depends on the accuracy of $g$ in its tail,
-whereas for positive discoveries it is more dependent on the accuracy of $g$ in the center.
-
-The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
-as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
-
-% As noted above, the penalty term (\ref{eqn:penalty}) causes systematic over-estimation of the mass of $g$ near 0,  which one would expect to overshrink interval estimates towards zero. 
-%The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound is negative and overshrinking (moving it too far towards zero) makes the left tail too big; conversely, for positive discoveries, the lower confidence bound is positive, and
-%overshrinking (moving it too far towards zero) makes the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
-%and be prepared to put up with some overshrinkage to avoid it.
-
-%Coverage rates are given in:
-%\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
-
-%Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all.tex}
-	\caption{All observations.
-Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario.
-This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero.
-Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (Table \ref{tab:nopen}).
-Removing the penalty in the half-uniform case is not recommended (see Appendix).}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg.tex}
-\caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
-and the normal-based method in the flat-top scenario.
-These results likely reflect inaccurate estimates of the tails of $g$ due to a disconnect between the tail of $g$ and the component distributions in these cases.
-For example, the uniform methods sometimes substantially underestimate the length of the tail of $g$ in these long-tailed scenarios,
-causing over-shrinkage of the tail toward 0.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos.tex}
-\caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds} \label{tab:coverage}
-\end{table}
-
-%
-%\subsubsection*{Reverse asymptotics}
-%
-%Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
-%
-%Note: we'd like to make sure it behaves ok in the case where most observations are null.
-%Perhaps simulate with 999 null and 1 alternative?
-%Also maybe look at performance for small sample sizes - e.g. n=1?
-%
-%Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
-%can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
-%by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
-%
-
-\subsection*{Differing measurement precision across units}
-
- We turn now to the second important component of our work: allowing for varying
- measurement precision across units. The key to this is the use of a likelihood,
- (\ref{eqn:normlik}) or (\ref{eqn:tlik}), that explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
-
-  To illustrate, we conduct a simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
- ($s_j=10$).  In both cases,
-we assume that half the effects are null
-and the other half are normally distributed with standard deviation 1:
-\begin{equation}
-p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1).
-\end{equation}
-In this setting, the poor-precision measurements $(s_j=10)$ tell us very little, and any sane analysis should effectively ignore them. 
- However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor}). This is because the poor measurements 
- produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
- which, when combined with the good-precision measurements, dilute the overall signal (e.g. they reduce the density of $p$ values near 0).
- This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
-the estimated error rates ($q$-values, or $\lfdr$ values) for the good-precision observations increase when the low-precision observations are included in the analysis
-(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
-for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).
-
-\begin{figure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_hist.pdf} 
-    \label{fig:goodpoor-hist}
-    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations}
- \end{subfigure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_scatter.pdf}
-    \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) of a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
-\end{subfigure}
-\caption{Simulation illustrating how, for existing FDR methods, 
-poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
-$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
-a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
-In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
-\end{figure}
-
- \subsubsection*{Reordering of significance, and the ``$p$ value prior"}
-
-Another consequence of accounting for differences in measurement precision is that \ashr may re-order the significance of the observations compared
- with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
- We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measuments that have the same $p$ value.
- The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
- and so the $\lfsr$ for these poor-precision measurements is always high.
-
-The potential for Bayesian analyses to re-order the significance of observations, and specifically to downweight imprecise observations,
-was previously discussed in \cite{guan.stephens.08}.
-However, Wakefield \cite{wakefield:2009} showed that, under a certain prior assumption, the Bayesian analysis produces the same ranking of significance as $p$ values (or their $z$ scores).
-He named this prior the ``$p$-value prior" because it can be thought of as the
-implicit prior assumption that is being made if we rank the significance of observations by their $p$ value.
-Wakefield's $p$-value prior assumes that the less precise effect estimates correspond to larger true effects, and specifically that they scale proportional
-to the standard errors $s_j$. More specifically still, it assumes a normal prior for the non-zero $\beta_j$ with mean 0 and variance $K s^2_j$ for some constant $K$.
-Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. Specifically, if, instead of assuming that $\beta_j$ is independent
-of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
-then the $\lfsr$ computed by ash.n provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
-as is illustrated in Figure \ref{fig:lfsr_pval}, right panel. (This result does not hold when using the mixtures of uniforms prior, ash.u.)
-
-This $p$-value prior assumes that the $z$ scores $z_j=\beta_j/s_j$
-are identically distributed, independent of $s_j$. This is essentially the assumption made, implicitly or explicitly, by existing methods -- like \locfdr, \mixfdr and \qvalue --
-that model the $z_j$ or $p_j$ directly.  In contrast, we have assumed up to now that the $\beta_j$ are independent of $s_j$. 
-We can set both these assumptions within a more general framework, which allows that $\beta_j/s^\alpha$ is independent of $s_j$ for some $\alpha$ [Equation (\ref{eqn:beta-alpha}))].
-Setting $\alpha=0$ implies that $\beta_j$ is
-independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
-effects (in absolute value). This latter assumption may often be qualitatively plausible: 
-for example, in gene expression studies the standard error for gene $j$ depends partly on
-the variance of its expression among samples, and genes with a larger variance may tend to be less tightly regulated and
-so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
-On the other hand, there is no particular reason to expect that either $\alpha=1$ or $\alpha=0$ will be the optimal choice.
-Indeed, optimal choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$,
-which will be dataset-specific. 
-Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
-comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
-likelihood-based methods can be used to select $\alpha$. For example, following the logic of the EB approach it
-would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
-which has been implemented in our software by C. Dai.
-
-%\begin{result}
-%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
-%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
-%\end{result}
-
-%To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
-%Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
-%In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
-%Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
-%which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
-%(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
-%may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)
-
-%\begin{result}
-%Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and using a mixture of normals for the prior distribution 
-%(\ref{eqn:beta-alpha}) with $\alpha=1$, 
-%the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$.
-Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$)
-%is montonic with $|t_j|$. 
-%\end{result}
-
-%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
-%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.
-
-%Most previous authors of EB approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
-%have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
-%In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
-%This is analogous
-%to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
-%standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
-%that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.
-
-\begin{figure}
-\centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
-\caption{Figure illustrating affects of prior assumptions on re-ordering of significance.
-Left panel shows results under our ``default prior" which assumes that effects $\beta_j$ are identically distributed, independent of $s_j$.
-Right panel shows results under the ``$p$-value prior", which assumes that $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$.} \label{fig:lfsr_pval}
-\end{figure}
-
-%\subsubsection*{Adaptive Shrinkage}
-%
-%Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
-%when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
-%the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
-% $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
-% In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
-% Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
-% and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
-%
-%Our methods we build on this idea in two ways. First,
-%by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
-%Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
-%adapts to measurement precision: more precise measurements undergo
-% less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
-%than observations with
-%small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
-%To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
-%Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
-%
-%\begin{figure}[!ht]
-%\begin{center}
-%\includegraphics[width=400px]{../examples/ash.pdf} 
-%\end{center}
-%\caption{Figure illustrating adaptive shrinkage.
-Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0.
-The standard error of each observation, $s_j$, was simulated from Inverse Gamma(5,5), which resulted in $s_j$ varying from 0.34 to 6.7.
-The shrunken estimates ($y$ axis) are plotted against the observed estimates ($x$ axis), with color indicating the (square root of the) precision of each measurement (blue=lower-precision; red=higher-precision).
-The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
-%\end{figure}
-
-%DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
-%To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
-Because the prior distribution is unimodal about zero,
-%the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
-%amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
-%For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
-%than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
-%small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
-
-%previous work based on normal means:
-%http://link.springer.com/article/10.1007/BF02562623
-
-%\subsection{Null behaviour}
-
-%Results of simulations under the null.
-
-%\subsection{Comparison of different underlying component distributions}
-%
-%Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
-%than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
-%likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
-%For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
-%increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
-%expect $L_{hu}$ to be significantly larger than the others.
-%
-%Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
-%
-%\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
-%\begin{center}
-%\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
-%%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
-%\end{center}
-%\caption{}
-%\end{figure}
-
-%Notes from Adrian Raftery - 
-%Newcombe (1886)
-%Dean and Raftery
-%Uniform is super-efficient
-
-\section*{Discussion}
-
-We have presented an Empirical Bayes approach to large-scale multiple testing that emphasises two ideas.
-First, we emphasise the potential benefits of using two numbers ($\bhat$, and its standard error)
-rather than just one number (a $p$ value or $z$ score)  to summarize the information on each test.
-While requiring two numbers is slightly more onorous than requiring one, in many settings these
-numbers are easily available and if so we argue it makes sense to use them.
-Second, we note the potential benefits -- both statistical and computational -- of assuming that the effects come from a unimodal distribution, and provide
-flexible implementations for performing inference under this assumption. We also introduce the ``false sign rate" as
-an alternative measure of error to the FDR, and illustrate its improved robustness to errors in model fit, particularly mis-estimation
-of the proportion of null tests, $\pi_0$.
-
-Multiple testing is often referred to as a ``problem" or a ``burden". In our opinion, EB approaches turn this idea on its head,
-treating multiple testing as an {\it opportunity}: specifically, an opportunity to learn about the prior distributions, and other modelling assumptions,
-to improve inference and make informed decisions about significance (see also \cite{greenland1991empirical}). This view
-also emphasises that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
-Indeed, the FDR at a given fixed threshold does not depend on the number of tests: as the number of tests increases, both the true positives and false positives increase linearly,
-and the FDR remains the same.
-(If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example.
-The EB approach captures this dependence in an intuitive way:
-if there are lots of strong signals then we infer $\pi_0$ to be small, and the estimated FDR (or $\lfdr$, or $\lfsr$) at a given threshold may be low, even if a large number of tests were performed; and conversely
-if there are no strong signals then we infer $\pi_0$ to be large and the FDR at the same threshold may be high, even if relatively few tests were performed. 
-More generally, overall signal strength is reflected in the estimated $g$, which in turn influences the estimated FDR.
-
-Two important practical issues that we have not addressed here are correlations among tests,
-and the potential for deviations from the theoretical null distributions of test statistics. These two
-issues are connected: specifically, unmeasured confounding factors can cause both correlations among tests
-and deviations from the theoretical null \cite{efron2007correlation,leek:2007}. And although there are certainly other
-factors that could cause dependence among tests, unmeasured confounders are perhaps the most worrisome in practice
-because they can induce strong correlations among large numbers of tests and profoundly impact results.
-Approaches to deal with unmeasured confounders can be largely divided into two types: those
-that simply attempt to correct for the resulting inflation of test statistics \cite{devlin1999genomic,efron2004large}, and those
-that attempt to infer counfounders using clustering, principal components analysis, or factor models 
-\cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
-and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable,
-they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alterntively, our methods
-could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:lik}) using
-$\bhat_j | \shat_j \sim N(0,\lambda_1 \shat_j + \lambda_2)$, where $\lambda_1,\lambda_2$ are to be estimated. However,
-this immediately raises issues of identifiability and we do not pursue the idea further here.
-
-Another important practical issue is the challenge of small sample sizes. For example, in genomics applications researchers sometimes
-attempt to identify differences between two conditions based on only a handful of samples in each. In such settings the
-normal likelihood approximation (\ref{eqn:normlik}) will be inadequate. And, although the $t$ likelihood (\ref{eqn:tlik}) partially
-addresses this issue, it is also, it turns out, not entirely satisfactory.  The root of the problem is that, with small sample sizes, raw estimated standard
-errors $\shat_j$ can be horribly variable. In genomics it is routine to address this issue by applying EB methods \cite{smyth:2004} to ``moderate" (i.e.~shrink) variance estimates,
-before computing $p$ values from ``moderated" test statistics. We are currently investigating how
-our methods should incorporate such ``moderated" variance estimates to make it applicable to small sample settings.
-
-Our approach involves compromises between flexibility, generality, and simplicity on the one hand, and statistical efficiency and principle on the other.
-For example, in using an EB approach that uses a point estimate for $g$, rather than a fully Bayes approach that accounts for uncertainty in $g$,
-we have opted for simplicity over statistical principle. And in summarizing every test by
-two numbers and making a normal or $t$ approximation to the likelihood, we have aimed to produce generic methods that 
- can be applied whenever such summary data are available -- just as \qvalue can be applied to any set of $p$ values for example -- although
- possibly at the expense of statistical efficiency compared with developing multiple tailored approaches based on context-specific likelihoods. 
- Any attempt to produce generic methods will involve compromise between generality and efficiency.
- In genomics, many analyses -- not only FDR-based analyses -- involve first computing a series of $p$ values before
- subjecting them to some further downstream analysis. An important message here is 
- that working with two numbers ($\bhat_j,\shat_j$), rather than one ($p_j$ or $z_j$), 
- can yield substantial gains in functionality (e.g.~ estimating effect sizes, as well as testing; accounting for variations in
-measurement precision across units) while losing only a little in generality.  We hope that our work will
-encourate development of methods that exploit this idea in other contexts.
-
-%This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
-%from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
-%with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
-
-%
-%
-%Both \locfdr and \mixfdr allow for deviations from
-%the theoretical null distribution by allowing that, under the null, the $z$ scores are $N(0,\sigma^2)$ with $\sigma$ estimated from the data rather
-%than fixed to 1. Within our approach this idea could be incorporated by changing the likelihood - for example, to assume $\bhat_j \sim N(0,s_ 
-%
-%These are tricky issues that, as far as we are aware, can affect
-%all approaches to FDR analysis. An important focus of Efron's work, also implemented in \locfdr, 
-% is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
-%theoretical null distribution. 
-%
-%For example, we and others have worked on Bayesian approaches to genetic association testing, using subjective prior specification of
-%effect sizes based on external data (e.g.~\cite{stephens:2009}). However, the large numbers of tests brings the opportunity to learn about
-%priors, and here we aim to exploit this
-
-%Choice of mixture components. Use of likelihood to compare different mixture components?
-
-%  
-%Directly modeling the p values, or z scores, say via non-parametric methods, 
-%can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-%distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-%The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-%Before proceeding, we feel obliged to correct the common misconception
-% that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
-% not depend on the number of tests. 
-%  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
-% 
-% In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
-
-\section*{Detailed Methods} \label{sec:implement}
-
-\subsection*{Embellishments}
-
-\subsubsection*{More flexible unimodal distributions}
-
-Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
-Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
- Therefore, we have implemented a more general approach based on 
-\begin{equation} \label{eqn:g-general}
-g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) ,
-\end{equation}
-where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
-\begin{enumerate}[(i)]
-\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$, \qquad (``ash.n")
-\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,  \qquad (``ash.u")
-\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,  \qquad (``ash.hu")
-\end{enumerate}
-where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
-(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
-and $\pi$ is a $2K+1$ vector that sums to 1.)
-The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
-Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
-under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
-(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
-(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
-however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)
-
-Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
-using a large dense grid of $\sigma^2_k$ or $a_k$, (i)-(iii) can respectively approximate,
-with arbitrary accuracy,
-\begin{enumerate}[(i)]
-\item any scale mixture of normals, which includes as special cases
-the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
-\item any symmetric unimodal distribution about 0.
-\item any unimodal distribution about 0.
-\end{enumerate}
-The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
-In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.
-
-Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
-The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
-See Implementation for details of our software defaults.
-
-%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting.
-For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
-%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
-
-\subsubsection*{Dependence of effects on standard errors}
-
-Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
-This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
- \begin{equation} \label{eqn:beta-alpha}
- \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
- \end{equation}
-for any $\alpha$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
-assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
-it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
-the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.
-
-The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
-To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
- standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
- Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
-yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
-%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
-%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
-%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
-
-%Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
-%how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
-%it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
-%rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
-%and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
-%this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
-%do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
-
-\subsubsection*{Replace normal likelihood with $t$ likelihood}
-
-We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
-  \begin{equation} \label{eqn:tlik}
- \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
- \end{equation}
-where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
-and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
-The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
-This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms; see Implementation below.
-When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
-
-Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
-\begin{equation} \label{eqn:tdist}
-(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
-\end{equation}
-However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
-In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
-currently pursuing this improved approach (joint work with M.Lu) and results will be published elsewhere. 
-%We note here that
-%modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
-%where our simpler approach can produce unsatisfactory results.
-
-\subsubsection*{Non-zero mode}
-
-An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood, rather than fixed to 0.
-
-\subsection*{Implementation Details}
-
-\subsubsection*{Likelihood for $\pi$}
-
-We define the likelihood for $\pi$ to be the probability of the observed data $\bhat$ conditional on $\shat$: $L(\pi) := p(\bhat | \shat,\pi),$ which
-by our conditional indendence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$.
-[One might prefer to define the likelihood as $p(\bhat, \shat | \pi) = p(\bhat | \shat, \pi) p(\shat | \pi)$, in which case our definition comes down to assuming that the term $p(\shat | \pi)$ does not depend on $\pi$.]
-
-Using the prior $\beta_j \sim \sum_{k=0}^K \pi_k f_k(\beta_j)$ given by (\ref{eqn:g-general}), and the normal likelihood (\ref{eqn:normlik}), integrating over $\beta_j$ yields
-\begin{equation}
-p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
-\end{equation}
-where 
-\begin{equation}
-\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
-\end{equation}
-denotes the convolution of $f_k$ with a normal density.
-These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
-Specifically, 
-\begin{equation} \label{eqn:uconv}
-\tilde{f}_k(\bhat_j)  = 
-\begin{cases}
-N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
-\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
-\end{cases}
-\end{equation}
-where $\Psi$ denotes the cumulative distribution function (c.d.f.) of the standard normal distribution.
-If we replace the normal likelihood with the $t_\nu$ likelihood (\ref{eqn:tlik}) then the convolution 
-for $f_k$ uniform the convolution is still given by (\ref{eqn:uconv}) but with $\Psi$ the c.d.f. 
-of the $t_\nu$ distribution function. (The convolution for $f_k$ normal is tricky and we have not implemented it.)
-
-\subsubsection*{Penalty term on $\pi$}
-
-To make $\lfdr$ and $\lfsr$ estimates from our method ``conservative" we add a penalty 
-term $log(h(\pi;\lambda))$ to the log-likelihood $\log L(\pi)$ to encourage over-estimation of $\pi_0$:
-\begin{equation} \label{eqn:penalty}
-h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
-\end{equation}
-where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
-consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0sims}).
-
-Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
-we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
-
-\subsubsection*{Problems with removing the penalty term in the half-uniform case}
-
-It is straightforward to remove the penalty term by setting $\lambda_k=1$ in (\ref{eqn:penalty}).
-We note here an unanticipated problem we came across when using no penalty term in the half-uniform case 
-(i.e.~ $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$ in (\ref{eqn:g-general})): when the data
-are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
-sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
-for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
-and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
-
-\subsubsection*{EM algorithm}
-
-With this in place, the penalized log-likelihood for $\pi$ is given by:
-\begin{equation}
-\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
-\end{equation}
-where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This can be maximized using an EM algorithm \cite{dempster77}, whose
-one-step updates are:
-\begin{align} \label{eqn:w}
-w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
-n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
-\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}
-\end{align}
-Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
-and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
-probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$. (The $w_{kj}$ are sometimes referred to as the ``responsibilities".)
-Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
-We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple} to accelerate convergence of this EM algorithm.
-
-%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
-\subsubsection*{Initialization}
-
-By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
-Our rationale for initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
-but weak signal in the data cannot quickly draw the algorithm towards the null.
-
-In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
-$\pi_0=1; \pi_k=0 (k>0)$. If not then we replace the EM solution with the global null solution. This aims to guard against errors due to convergence to a local optimum when the
-data are consistent with the global null.
-
-\subsubsection*{Conditional distributions}
-
-Given $\hat\pi$, we compute the conditional distributions 
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
-\end{equation} 
-Each posterior is a mixture on $K+1$ components:
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
-\end{equation}
-where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
-and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
-$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
-All these posterior distributions are easily available.
-For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
-If $f_k$ is normal and $L$ is normal, then this is a normal distribution.
-
-\subsubsection*{Choice of grid for $\sigma_k, a_k$} \label{sec:grid}
-
-\def\sigmamax{\sigma_{\text{max}}}
-\def\sigmamin{\sigma_{\text{min}}}
-When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
-in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
-as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
-based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
-is negative we set $\sigmamax = 8\sigmamin$.)
-
-When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.
-
-Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
-a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.
-
-\appendix
-
-\begin{table}[!ht]
-\centering\begin{tabular}{c c } \toprule
-Scenario & Alternative distribution, $g_1$  \\ \midrule
-spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
-near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
-flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
- &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
-skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
-big-normal & $N(0,4^2)$ \\ 
-bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
-\end{tabular}
-\caption{Summary of simulation scenarios considered} \label{table:scenarios}
-\end{table}
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all_nopen.tex}
-	\caption{All observations}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg_nopen.tex}
-\caption{``Significant" negative discoveries.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos_nopen.tex}
-\caption{``Significant" positive discoveries.}
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{table:nopen}
-\end{table}
-%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
-%\begin{equation}
-%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
-%\end{equation}
-%which is close to the definition in \cite{storey.03}.
-
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgements}
-
-Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
-prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} , and includes contributions 
-from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.
-
-This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.
-
-%\section*{References}
-% The bibtex filename
-\bibliography{/Users/stephens/Dropbox/Documents/mainbib}
-
-\section*{Figure Legends}
-
-\clearpage
-
-\section*{Tables}
-
-\section*{Supporting Information Legends}
-
-Supplementary material can be found in {\bf Supplementary Information S1.}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
-\end{document}
+(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)

### 2015-10-06 sectionAcknowledgeme.tex
New Article!

--- 2015-10-06 paper/main.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
-\documentclass[11pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-%\usepackage{natbib}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-\usepackage{booktabs}
-\usepackage{xspace}
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-\usepackage{sectsty}
-\subsectionfont{\normalsize}
-\subsubsectionfont{\mdseries\itshape}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\usepackage{color}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-% Use the PLoS provided bibtex style
-\bibliographystyle{/Users/stephens/Dropbox/Documents/stylefiles/plos2009}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
-\pagestyle{myheadings}
-%% ** EDIT HERE **
-\usepackage{enumerate}
-\usepackage{multirow} 
-\usepackage{url}
-\usepackage{xr} %for cross-referencing
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-\newtheorem{algorithm}{Algorithm}
-\newtheorem{proposition}{Proposition}
-\newtheorem{restateproposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{result}{Result}
-\newtheorem{note}{Note}
-\newtheorem{definition}{Definition}
-
-\def\lfdr{\textit{lfdr}}
-\def\lfsr{\textit{lfsr}}
-\def\x{\mbf{x}}
-\def\y{\mbf{y}}
-\def\e{\mbf{e}}
-\def\g{\mbf{g}}
-\def\p{p}
-\def\Yb{\hat{Y}_1^{\text{BAYES}}}
-\def\Xperp{X_{1 \perp 0}}
-\def\Pperp{P_{1 \perp 0}}
-\def\Pperpb{\Pperp^{\text{B}}}
-\def \hatd { \widehat{ d } }
-\def \dhatin {\hatd^{\mbox{\tiny IN}}}
-\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
-\def \mvn{\text{MVN}}
-\def\mn{\text{MN}}
-\def \iw{\text{W}^{-1}}
-\def \fp {fastPHASE\ }
-\def\bhat{\hat{\beta}}
-\def\shat{\hat{s}}
-\def\var{\rm Var}
-\def\mvr{\text{BMVR}}
-\def\bfuni{\text{BF}_\text{uni}}
-\def\one{\mbf{1}}
-\def\zero{\mbf{0}}
-\def\hall{H_{\text{all}}}
-\def\pall{\p_{\text{all}}}
-\def\bfall{\text{BF}_\text{all}}
-\def\bfav{\text{BF}_\text{av}}
-\def\gall{\gamma_\text{all}}
-\def\limbfall{\bfall^\rightarrow}
-\def\limbfav{\bfav^\rightarrow}
-\def\limbfgamma{\BF_\gamma^\rightarrow}
-\def\sxx{S_{xx}}
-\def\syy{S_{yy}}
-\def\vyy{V_{yy}}
-\def\vyx{V_{yx}}
-\def\vxx{{V_{xx}}}
-\def\rss{\text{\rm RSS}}
-\def\dss{\text{\rm $\Delta$SS}}
-\def\U{U}
-\def\D{D}
-\def\I{I}
-\def\luu{L_{UU}}
-\def\indep{\perp}
-\def\qvalue{{\tt qvalue}\xspace}
-\def\locfdr{{\tt locfdr}\xspace}
-\def\mixfdr{{\tt mixfdr}\xspace}
-\def\ashr{{\tt ashr}\xspace}
-
-\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
-                                % mode too
-\newcommand\height{{\it height}}
-\newcommand\weight{{\it weight}}
-\def\BF{\text{BF}}
-
-%% END MACROS SECTION
-%\externaldocument{plos_one_rev1_suppinfo}
-
-\begin{document}
-
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{False Discovery Rates: A New Deal}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Matthew Stephens$^{1*}$, 
-\\
-\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
-\\
-$\ast$ E-mail: Corresponding mstephens@uchicago.edu
-\end{flushleft}
-
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-%\section*{Author Summary}
-
-\section*{Introduction}
-
-Since its introduction in 1995 by Benjamini and Hochberg 
-\cite{benjamini1995controlling}, the ``False Discovery Rate" (FDR) has quickly established itself
-as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing
-in which the goal is to identify the non-zero ``effects" among a large number of imprecisely-measured effects.
-
-Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
-indeed, the notion that EB approaches could be helpful in handling multiple comparisons predates introduction of
-the FDR  (e.g.~\cite{greenland1991empirical}). More recently, EB approaches to the FDR
-have been extensively studied by several authors, 
-especially B.~Efron and co-authors 
-\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
-see also \cite{kendziorski2003parametric,newton2004detecting, 
-datta2005empirical,muralidharan2010empirical} for example.
-So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
-existing widely-used FDR pipelines) and can substantially affect
-inference. The first idea is to {\it assume that the distribution of effects is unimodal}. This yields a very simple, fast, and stable computer 
-implementation, as well as improving inference of FDR when the unimodal assumption is correct.  
-The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
-to summarize each measurement. This idea allows variations in measurement precision to be better accounted for,
- and avoids a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
-
-In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
-the idea that it may be preferable to focus on estimation rather than on testing.
-In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
-simply estimation with some positive prior probability that the effect is exactly zero.
-However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
-perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
-controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
-``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
-in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
-the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the lfsr, particularly its increased robustness to modelling assumptions.
-
-%terminology
-%``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
-%and correspondingly define the ``false signed discovery rate", or simply ``false sign rate" (FSR), in a natural way.
-
-%The methods we provide here can address both problems, and we provide results that highlight important benefits of focussing on estimation, and particularly on assessing the correctness of the {\it sign} of the estimated effects, rather than on testing for non-zero effects.
-
-Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
-(We address the reasons for this name, and connections with shrinkage analysis, in the Discussion.)
-
-%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.
-
-%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-%measurements are made on only a small number of
-%samples in each condition - sometimes as few as one
-%sample in each condition. Thus the error in estimates
-%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-%in estimates of $\beta_j$ still greater.
-%
-
-%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
-%Here we present new tools that bring two important ideas to bear on the problem.
-%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
-%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
-%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
-%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
-%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
-%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
-%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
-%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
-%a substantial increase in the number of discoveries at a given FDR).
-
-%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
-%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
-Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
-%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
-%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
-%generic methods that work directly with two numbers for each 
-%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
-Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
-%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
-
-\def\df{df}
-\def\FDR{\text{FDR}}
-\def\fdr{\text{\lfdr}}
-\def\FSR{\text{FSR}}
-\def\fsr{\text{lfsr}}
-
- \section*{Methods}
-
- \subsection*{Model Outine}
-
-% Here we consider one setting in which FDR-based methods are widely used:
-%an experimenter has measured a large numbers of ``effects", each with an associated
-%measurement precision (standard error), 
-%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
-%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
-%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
-%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
-%For example, in genomics
-%a common goal is to identify ``differentially-expressed genes" -- that is, genes
-%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
-%mean expression level between the two conditions, often referred to as the ``log-fold change". 
-Here we describe the simplest version of our method, and briefly discuss embellishments we have also implemented.
-
-Suppose that we are interested in the values of $J$ ``effects" $\beta=(\beta_1,\dots,\beta_J)$. For example, in a typical genomics application
-that aims to identify differentially expressed genes,  $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
-In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
-that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
-more general problem of estimating, and assessing uncertainty in, $\beta_j$.
-
-Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
-and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
-which by Bayes theorem can be written as 
-\begin{equation}
-p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
-\end{equation}
-
-For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
-This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
-A simple way to implement the UA is to assume that $g$ is
-a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
-\begin{align} \label{eqn:beta}
-p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
-g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
-\end{align}
-where $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyperparameters, which are non-negative and sum to one, and are to be estimated,
-while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
-represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
-(We encourate the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
-although of course in practice we use a finite grid -- see Implementation Details.)
-
-For the likelihood $p(\bhat | \beta, \shat)$ we assume  
-\begin{equation} \label{eqn:normlik}
-p(\bhat | \beta, \shat) = \prod_j N(\bhat_j; \beta_j, \shat_j^2).
-\end{equation}
-Here, in additional to some conditional independence assumptions, we are 
-effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-
-This simple model features both the key ideas we want to emphasise in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
-the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
-with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
-However, this simple model also has several additional assumptions that can be relaxed. Specifically,
-\begin{enumerate}
-\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) also implies that $g$ is symmetric about 0; more flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see (\ref{eqn:g-general})].
-\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
-We can relax this to allow for a relationship between these quantities [see (\ref{eqn:beta-alpha})].
-\item The likelihood (\ref{eqn:normlik}) effectively assumes that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-We can generalize this likelihood using a $t$ likelihood [see (\ref{eqn:tlik})].
-\end{enumerate}
-These embellishments are detailed in Embellishments and Implementation Details.
-Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
-(which are also made by most existing EB approaches to this problem). Correlations among tests certainly arise in practice, either due to genuine correlations
-in the system of study, or due to unmeasured confounders, and their potential to impact results of an FDR analysis is important to consider
-whatever analysis methods are used: see \cite{efron2007correlation,leek:2007} for relevant discussion.
-
-%We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
-%$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
-%\begin{equation}
-%p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
-%\end{equation}
-%Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
-%\begin{equation} \label{eqn:normlik}
-%p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
-%
-%We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
-%as a mixture of null and alternative components:
-%\begin{equation} \label{eqn:betaj0}
-%g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
-%\end{equation}
-%where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
-%We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
-%\begin{equation} \label{eqn:beta-iid}
-%p(\beta | \shat) = \prod_j g(\beta_j).
-%\end{equation}
-%
-%Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
-%with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
-%to use a mixture of zero-mean normal distributions:
-%\begin{equation} \label{eqn:mixnorm}
-%g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
-%Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
-%represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
-% The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
-% more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below. 
-% 
-% Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
-% yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
-% \begin{equation} \label{eqn:g}
-% g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
-% \end{equation}
-% 
-% Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
-% \begin{align} \label{eqn:betaj}
-% p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
-% p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
-% \end{align}
-% with $g$ given by the mixture of normals $(\ref{eqn:g})$.
-
-\subsection*{Fitting the model}
-
-In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
-each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
-Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
-\begin{equation} \label{eqn:marginallik}
-p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
-\end{equation}
-where we define $\sigma_0:=0$.
-
-The usual EB approach to fitting this model would involve two simple steps:
-\begin{enumerate}
-\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
-\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
-against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | | \bhat, \shat,\hat{\pi})$.
-\end{enumerate}
-Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(The simplicity of the EM algorithm in step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
-which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
-
-Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximise a penalized
-likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
-We introduce this penalty because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
-
-\subsection*{The local False Discovery Rate and local False Sign Rate}
-
-As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
-In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
-One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
-the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
-In other words it is the posterior probability that $\beta_j$ is actually zero:
-\begin{equation} \label{eqn:fdr}
-\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
-\end{equation}
-
-The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
-whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
-the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey1991philosophy}) argued that
-``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$.
-Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
-\begin{quote}
-...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
-\end{quote}
-Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
-focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.
-
-Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
-would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
-\begin{equation} \label{eqn:lfsr}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
-\end{equation}
-To illustrate, suppose that 
-\begin{gather*}
-p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
-p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
-p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
-\end{gather*} 
-Then 
-from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
-our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.
-
-As our notation suggests, $\lfsr_j$ is intended to be compared and contrasted with $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
-small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
-Of coure, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
-this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
-(this follows from the definition because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
-In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
-$\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.
-
-From these ``local" measures of significance, we can also compute average error rates over subsets of observations
-$\Gamma \subset \{1,\dots,J\}$. For example,   
-\begin{equation} \label{eqn:FDRhat}
-\widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
-\end{equation}
-estimates the FDR we would obtain if we were to declare all tests in $\Gamma$ significant.
-And 
-\begin{equation} \label{eqn:qvalue}
-q_j := \widehat \FDR(\{k: \fdr_k \leq \fdr_j\}) 
-\end{equation}
-provides a measure of significance analogous to Storey's $q$ value \cite{storey.03}.
-
-%Because our methods involve explicitly modelling effects $\beta_j$, rather than 
-%$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
-
-%Indeed, even in settings where some
-%$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
-%analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
-
-\subsection*{Related work}
-
-\subsubsection*{Previous approaches focussed on FDR}
-
-Among previous methods that explicitly consider the FDR and related quantities,
-our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
- (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
-  %Here we highlight the key differences between our approach and these three existing approaches.
-
-There are two key differences between our approach and all of these three existing methods. First, 
-whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
-we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
-for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
-Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 
-how it can improve performance by taking better account of variation in measurement precision among observations.
-%(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
-% Or more generally Section xx illustrates the potential practical effects
-
-%allows for different modelling assumptions to be compared through likelihood comparisons.
-Second, our unimodal assumption (UA) that the effects are unimodal about zero is quite different from assumptions made by $\qvalue, \locfdr$ or $\mixfdr$.
-Indeed, $\locfdr$ assumes that all $z$ scores near 0 are null (Efron calls this the Zero Assumption; ZA), which implies that 
-under the alternative hypothesis the distribution of $z$ scores has {\it no mass at 0}; this contrasts strikingly with the UA, which implies that this distribution has its peak at 0! 
-Similarly, $\qvalue$ assumes that all $p$ values near 1 are null,
-which is the same as the ZA because $p$ values near 1 correspond to $z$ scores near 0. And although \mixfdr does not formally make the ZA,
-we have found that in practice, with default settings, the results often approximately satisfy the ZA (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}).
-Thus, not only do these existing methods not make the UA, they actually make assumptions that are, in some sense, as different
-from the UA as they can be.
-
-Given that the UA and ZA are so different, it seems worth discussing why we generally favor the UA. 
-Although the UA will not apply to all situations, we believe that it will often be reasonable, especially in FDR-related contexts
-that have traditionally focussed on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis,  
-it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible. Further, it seems reasonable to expect that larger effects become
-decreasingly plausible, and so the distribution of the effects will be unimodal about 0.
-To paraphrase Tukey, ``All we know about the world teaches us that large effects are rare, whereas small effects abound."
-We emphasise that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
-It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
-
-%Tukey "peri-null"
-
-%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
-%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}.
-
-In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficents - 
-common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
-These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
-it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
-The UA assumption on regression coefficients is directly analagous to our UA here, and so its
-widespread use in the regression context supports its use here.
-
-Alternatively, we could motivate the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
-such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
-make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).
-
-Finally, the UA also has a considerable practical benefit: it yields simple algorithms that are both computationally
-and statistically stable. We illustrate these features in Results.
-
-\subsubsection*{Other work}
-
-There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
- Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
-indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
-This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
-Also, they provide no software implementation.
-More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
-Some pointers to recent papers in which other relevant
-citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}.
-Much of the literature focusses on the homoskedastic case (i.e.
-$\shat_j$ all equal) whereas we allow for heteroskedasticity.
-And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential.
-Several recent papers consider more flexible non-parametric assumptions on
-$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
-estimate (NPMLE) for $g$.
-These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive.
-However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
-may perform well for point estimation, it seems possible it will not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
-Indeed, one way of thinking about the UA is as a way to regularize $g$.
-
-%The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
-%encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-%$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-%$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-%the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-%observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-%corresondingly little effect on the estimated FDR of other observations. 
-%We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-%$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-%that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-%estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-%Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-\def\dt{\text{dt}}
-
-%
-%assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
-%$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
-% $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
-%  \begin{equation} \label{eqn:lik}
-% \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
-% \end{equation}
-% where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e.
-the density of $\mu+\sigma T_\nu$).
-In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
-%  \begin{equation} \label{eqn:normlik}
-% \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
-% \end{equation}
-%which, in light of (\ref{eqn:betahat}), is equivalent to
- % \begin{equation} \label{eqn:lik2}
-% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
- %\end{equation}
-
-%The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
-they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
-The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$.
-Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$.
-If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$.
-See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.
-
-%This likelihood approximation can be motivated by assuming that i) $\bhat$
-% $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
-% $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$. 
-%
-
-%Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.
-
-%While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
- %that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
-%there is an important difference: here,
-
-%The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
-For example, we can write $p(\bhat_j, \shat_j | \beta_j) = p(\bhat_j | \shat_j, \beta_j)p(\shat_j | \beta_j)$.
-Now, if we are willing to assume that $\shat_j$ alone contains no information about $\beta_j$, or equivalently that $p(\shat_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\bhat_j | \shat_j, \beta_j)$, and if $\bhat_j | \shat_j, \beta_j \sim N(\beta_j, \shat_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.
-
-\section*{Results}
-
-We compare results of \ashr with existing FDR-based methods implemented
-in the R packages \qvalue (v2.0 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
-In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
-to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.
-
-\subsection*{Effects of the Unimodal Assumption}
-
-Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
-standard error, $s_j=1$ and all methods are provided that information. That is,
-$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$.
-In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the only difference between our method and methods like $\locfdr$ and $\mixfdr$ are in how they estimate $g$.
-
-To briefly summarize the results in this section:
-\begin{enumerate}
-\item The UA can produce very different inferences compared with the ZA made by existing methods.
-\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
-\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
-\end{enumerate}
-
-\subsubsection*{The UA and ZA can produce different inferences}
-
-\begin{figure}
-\center \includegraphics[height=6in]{figures/decomp_ZA.pdf}
-\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
-The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
-In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
-The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
-(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
-\end{figure}
-
-To illustrate the different inferences from the UA and ZA we show results for a single
-dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
-but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
-\qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
-The results (Figure \ref{fig:ZA}) illustrate the clear difference between the existing methods and our method. The effects of the ZA made by \qvalue and \locfdr
-are visually clear, producing a ``hole"  in the alternative $z$ score distribution around 0. Although \mixfdr does not formally make the ZA, its decomposition
-exhibits a similar hole. In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
-effectively ``filling in" the hole.  (Of course the null distribution also has a peak at 0, and the local fdr 
- under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)
-
-Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA.
-Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
-This would increase the null component (dark blue) at the expense of
-the alternative component (light blue).
-Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0.
-The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears.
-In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
-See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
-demonstration.
-
- \subsubsection*{The UA can produce conservative estimates of $\pi_0$}
-
-The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
-Consequently \ashr will estimate smaller $\lfdr$s and FDRs than existing methods that make the ZA. 
-This is desirable, provided that these estimates remain conservative: that is,
-provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
-The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behaviour. To check its effectiveness
-we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
-for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
-They range from a ``spiky" distribution -- where many non-zero $\beta$ are
-too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
-flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
-making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
-which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.
-
-% How do these differences influence estimates of the FDR? To address this, note that  -- in this simple case where the 
-% $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
-% of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
-% strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
-% although the methods considered here differ from one another in many ways, having estimated $\pi_0$
-% they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
-% empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
-%
-
-%Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
-% and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
-%Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
-%At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
-%overestimates of $\pi_0$, which will yield overestimates of the FDR).
-
-\begin{figure} 
-\begin{center}
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=1.8in]{figures/scenarios_density.pdf} 
-	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/pi0_est.pdf} 
-\caption{Comparison of true and estimated values of $\pi_0$.
-When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ashr being least conservative, and hence most accurate.
-When the UA does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.} \label{fig:pi0sims}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfdr_est.png} 
-\caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
-Black line is $y=x$ and red line is $y=2x$.
-Estimates of $\lfdr$ are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est.png} 
-\caption{As in c), but for $\lfsr$ instead of $\lfdr$.
-Estimates of $\lfsr$ are consistently less conservative than $\lfdr$ when UA holds, and also less anti-conservative in bimodal scenario.}  \label{fig:lfsr}
-\end{subfigure}
-\end{center}
-\caption{Results of simulation studies (constant precision $s_j=1$).}
-\end{figure}
-
-For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
-data as follows:
-\begin{enumerate}
-\item Simulate $\pi_0  \sim U[0,1]$.
-\item For $j=1,\dots,J$, simulate $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)$.
-\item For $j=1,\dots,J$, simulate $\bhat_j | \beta_j \sim N(\beta_j,1)$.
-\end{enumerate}
-%Thus these simulations assume the same
-%precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).
-
-Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ashr ($y$ axis) with the true values ($x$ axis). 
-For \ashr we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
-and as a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
-as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
-The results show that \ashr provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
-in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ashr estimates can be slightly anti-conservative.
-We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
-are used. (In addition, the effects on $\lfsr$ estimates turn out to be relatively modest; see below).
-
-%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
-%
-%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
-%and so all effects are non-zero and no nulls hold. This scenario is rather different
-%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
-%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
-%``mostly non-null"
-%scenario, and we would agree! On the other hand, the distribution of $p$ values
-%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
-%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
-%Thus it seems desirable to have methods that perform well in both cases, and this is
-%an important feature of the methods we introduce here.
-%
-%TODO: put FDRs on graph? Simulate with 50\% null?
-
-\subsubsection*{The $\lfsr$ is more robust than \lfdr}
-
-The results above show that \ashr can improve on existing methods in producing smaller, more accurate, 
-estimates of $\pi_0$, which will lead to more accurate estimates of FDR.
-Nonetheless, in many scenarios \ashr continues
-to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
-This is because these scenarios include 
-an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
-estimation of $\pi_0$ impossible. Put another way, and as is well known, $\pi_0$ is not identifiable:
- the data can effectively provide an upper bound on plausible values of $\pi_0$,
-but not a lower bound (because the data cannot rule out that everything is non-null, but with miniscule effects).
-To obtain conservative behaviour we must estimate $\pi_0$ by this upper bound, which can
-be substantially larger than the true value.
-
-Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
-overestimation of $\pi_0$ is corresponding overestimation of FDR (and $\lfdr$, and $q$ values).
-To illustrate, Figure \ref{fig:lfdr} compares the estimated $\lfdr$ from ash.n with the true value (computed using
-Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios which involve many
-non-zero effects that are very near 0 (e.g.~the spiky scenario with $\pi_0$ small) where
-$\pi_0$ can be grossly overestimated.  
-(Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.)
-
-The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
- become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
- This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
-To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
-be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
-the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in $\lfsr$ than $\lfdr$.
-
-Note that, compared with previous debates regarding testing, 
-this section advances an additional reason for focussing on the sign of the effect, rather than just testing whether it is 0. 
-In previous debates authors have argued against testing whether an effect is 0
-because it is {\it implausible that effects are exactly 0}. Here we add that {\it even if one believes
-that some effects may be exactly zero}, it is still better to focus on the sign, because generally {\it the data are more informative about that question}
-and so inferences are more robust to, say, the inevitable mis-estimation of $\pi_0$.
-To provide some intuition, consider an observation with a $z$ score of 0. The $\lfdr$ of this observation can range from 0 (if $\pi_0=0$)
-to 1 (if $\pi_0=1$). But, assuming a symmetric $g$, the $\lfsr>0.5$ whatever the value of $\pi_0$, because the observation $z=0$ says
-nothing about the sign of the effect.
-Thus, are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
-whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.
-
-Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
-Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
-if we are prepared to assume that no effects are exactly zero, then removing the point mass 
-yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
-However, there is ``no free lunch":  if in fact some effects are exactly zero
-then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
-We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.
-
-\begin{figure}
-\begin{center}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s_nn.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be more accurate in this case.
-That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
-\end{subfigure}
-\end{center}
-\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
-\end{figure}
-
-\subsubsection*{Numerical Stability}
-
-The EM algorithm, which we use here to fit our model, is notorious for convergence to local optima. 
-However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
-caused by such behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
-in addition running it using our default initialization procedure (see Implementation details). We then compared the largest
-log-likelihood achieved across all 11 runs with the log-likelihood achieved by the default run.
-When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference ($< 0.02$), 
-and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference ($<0.02$),
-and 6\% of runs showed an appreciable log-likelihood difference ($>1$), with the largest difference being 5.0. However, perhaps suprisingly, 
-even for this largest difference results from the default run (e.g.
-the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood.
-See \url{http://github.com/stephens999/ash/blob/master/dsc-robust/summarize_dsc_robust.rmd} for further details.
-
-\subsubsection*{The UA helps provide reliable estimates of $g$}
-
-An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
-can estimate the {\it size} of each effect $\beta_j$.
-Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
-to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
-threshold $T$.
-%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
-%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
-%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
-%towards zero, with the strength of the shrinkage being estimated from the data.
-Further, because the posterior distribution is, by definition,
-conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
-been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
-In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
-\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
-how can confidence intervals be assigned to the true
-effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
-and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
-%but also notes the difficulties of implementing a stable general algorithm.
-
-The ability of the EB approach to provide valid ``post-selection" interval estimates is extremely attractive in principle.
-But its usefulness in practice
-depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
-which are notoriously difficult in general. Indeed, Efron emphasises
-the difficulties of implementing a stable general algorithm, noting in his rejoinder
-``the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying
-to be overly nonparametric ... but it is hard to imagine a
-generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
-Our key point here is that the UA greatly simplifies the deconvolution problem.
-While not meeting Efron's desire for an entirely general nonparametric approach, we believe 
-that the UA can handle many cases of practical interest.
-
-To illustrate this, Figure \ref{fig:egcdf} compares the
-estimated $g$ from \ashr  with that from \mixfdr which does not make the UA 
-(and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
-The greater reliability of estimates afforded
-by the UA is immediately apparent. In particular the estimated cdf from \mixfdr often has an almost-vertical
-segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. 
-The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
-While the UA is not the only way to achieve this, we find it an attractive, simple and effective approach.
-
-Interestingly, even in the ``bimodal" scenario $\ashr$ 
-is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
-Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
-Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
-to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
-(\url{http://github.com/stephens999/ash/blob/master/dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
-clearly visible in the observed $\bhat$.
-
-We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ashr tends to systematically
-overestimate the mass of $g$ near zero. On careful inspection, this is apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
-and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf}b),
-and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
-under-estimating $\pi_0$ (Figure \ref{fig:egcdf}c).
-
-\begin{figure}[h!] 
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=2in]{figures/egcdf.pdf}
-	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf_nopen.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
-Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
-\end{subfigure}
-\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.} \label{fig:egcdf}
-\end{figure}
-
-%making the UA provides an attractive and simple way to achieve estimating $g$  large
-%subset of cases of interest (those where the UA can reasonably be assumed to hold).
-
-%In practice, the utility of the posterior distributions noted above depends primarily on the accuracy
-%of the estimate of the distribution $g(\cdot)$.
-
- %quote from efron rejoinder , p46:
-%MorrisÕ Section 3 is especially pertinent. His formula
-%for p(?|z) is related to my discussion of the
-%BenjaminiÐYekutieli False Coverage Rate method in
-%Section 7, particularly (7.2)Ð(7.4). Originally I had
-%hoped to develop an empirical Bayes method for estimating
-%such models, but the effort foundered on practical
-%difficulties involving the perils of deconvolution.
-%Section 6 on the Òone-group modelÓ is the ugly duckling
-%of the current paper, but it bears on some important
-%points raised in the discussion. Figure 7 concerns
-%a fuzzy version of simultaneous hypothesis testing,
-%where, as in MorrisÕ hospital example, the usual singlepoint
-%null hypotheses seem unequal to the task. The
-%development from (6.6) onwards, particularly (6.12),
-%bears on the possibility of nonnormal null distributions,
-%and is about as far as I can go in answering Professors
-%Rice and SpiegelhalterÕs penultimate question.
-%With g(?) a normal distribution, model (6.1) returns
-%us to the realm of Stein estimation, the scene of my
-%happy collaborations with Carl Morris. I continue to
-%be surprised at how much harder bumpy, nonnormal
-%models like (7.1) are to deal with. JamesÐStein estimation
-%works fine with, say, N = 10 component problems,
-%while the RobbinsÕ form of empirical Bayes appropriate
-%to (7.1) seems to require hundreds or thousands.
-%The information calculations in Efron (2008)
-%reinforce this gloomy assessment. Maybe I am trying
-%to be overly nonparametric in constructing the empirical
-%Bayes Fdr estimates, but it is hard to imagine a
-%generally satisfactory parametric formulation for (6.1).
-%Or perhaps it is just that hypothesis testing is more demanding
-%than estimation.
-%My Note: (6.1) is the general model (mu \sim g()).
-%7.1 is the case where the effects are a mixture of zero and N(-3,1).
-
-\subsubsection*{Calibration of posterior intervals}
-
-To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over 100 simulations in each Scenario). 
-Specifically we examine the empirical coverage of nominal lower 95\% credible bounds for a) all observations; b) significant negative discoveries; c) significant positive discoveries.
-We examine only lower bounds because the results for upper bounds follow by
-symmetry (except for the one asymmetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
-in each case: for negative discoveries the lower bound is typically large and negative and limits how big (in absolute value) 
-the effect could be; for positive discoveries the lower bound 
-is positive, and limits how small (in absolute value) the effect could be. Intuitively, the lower bound for negative discoveries depends on the accuracy of $g$ in its tail,
-whereas for positive discoveries it is more dependent on the accuracy of $g$ in the center.
-
-The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
-as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
-
-% As noted above, the penalty term (\ref{eqn:penalty}) causes systematic over-estimation of the mass of $g$ near 0,  which one would expect to overshrink interval estimates towards zero. 
-%The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound is negative and overshrinking (moving it too far towards zero) makes the left tail too big; conversely, for positive discoveries, the lower confidence bound is positive, and
-%overshrinking (moving it too far towards zero) makes the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
-%and be prepared to put up with some overshrinkage to avoid it.
-
-%Coverage rates are given in:
-%\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
-
-%Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all.tex}
-	\caption{All observations.
-Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario.
-This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero.
-Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (Table \ref{tab:nopen}).
-Removing the penalty in the half-uniform case is not recommended (see Appendix).}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg.tex}
-\caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
-and the normal-based method in the flat-top scenario.
-These results likely reflect inaccurate estimates of the tails of $g$ due to a disconnect between the tail of $g$ and the component distributions in these cases.
-For example, the uniform methods sometimes substantially underestimate the length of the tail of $g$ in these long-tailed scenarios,
-causing over-shrinkage of the tail toward 0.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos.tex}
-\caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds} \label{tab:coverage}
-\end{table}
-
-%
-%\subsubsection*{Reverse asymptotics}
-%
-%Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
-%
-%Note: we'd like to make sure it behaves ok in the case where most observations are null.
-%Perhaps simulate with 999 null and 1 alternative?
-%Also maybe look at performance for small sample sizes - e.g. n=1?
-%
-%Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
-%can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
-%by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
-%
-
-\subsection*{Differing measurement precision across units}
-
- We turn now to the second important component of our work: allowing for varying
- measurement precision across units. The key to this is the use of a likelihood,
- (\ref{eqn:normlik}) or (\ref{eqn:tlik}), that explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
-
-  To illustrate, we conduct a simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
- ($s_j=10$).  In both cases,
-we assume that half the effects are null
-and the other half are normally distributed with standard deviation 1:
-\begin{equation}
-p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1).
-\end{equation}
-In this setting, the poor-precision measurements $(s_j=10)$ tell us very little, and any sane analysis should effectively ignore them. 
- However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor}). This is because the poor measurements 
- produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
- which, when combined with the good-precision measurements, dilute the overall signal (e.g. they reduce the density of $p$ values near 0).
- This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
-the estimated error rates ($q$-values, or $\lfdr$ values) for the good-precision observations increase when the low-precision observations are included in the analysis
-(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
-for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).
-
-\begin{figure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_hist.pdf} 
-    \label{fig:goodpoor-hist}
-    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations}
- \end{subfigure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_scatter.pdf}
-    \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) of a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
-\end{subfigure}
-\caption{Simulation illustrating how, for existing FDR methods, 
-poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
-$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
-a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
-In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
-\end{figure}
-
- \subsubsection*{Reordering of significance, and the ``$p$ value prior"}
-
-Another consequence of accounting for differences in measurement precision is that \ashr may re-order the significance of the observations compared
- with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
- We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measuments that have the same $p$ value.
- The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
- and so the $\lfsr$ for these poor-precision measurements is always high.
-
-The potential for Bayesian analyses to re-order the significance of observations, and specifically to downweight imprecise observations,
-was previously discussed in \cite{guan.stephens.08}.
-However, Wakefield \cite{wakefield:2009} showed that, under a certain prior assumption, the Bayesian analysis produces the same ranking of significance as $p$ values (or their $z$ scores).
-He named this prior the ``$p$-value prior" because it can be thought of as the
-implicit prior assumption that is being made if we rank the significance of observations by their $p$ value.
-Wakefield's $p$-value prior assumes that the less precise effect estimates correspond to larger true effects, and specifically that they scale proportional
-to the standard errors $s_j$. More specifically still, it assumes a normal prior for the non-zero $\beta_j$ with mean 0 and variance $K s^2_j$ for some constant $K$.
-Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. Specifically, if, instead of assuming that $\beta_j$ is independent
-of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
-then the $\lfsr$ computed by ash.n provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
-as is illustrated in Figure \ref{fig:lfsr_pval}, right panel. (This result does not hold when using the mixtures of uniforms prior, ash.u.)
-
-This $p$-value prior assumes that the $z$ scores $z_j=\beta_j/s_j$
-are identically distributed, independent of $s_j$. This is essentially the assumption made, implicitly or explicitly, by existing methods -- like \locfdr, \mixfdr and \qvalue --
-that model the $z_j$ or $p_j$ directly.  In contrast, we have assumed up to now that the $\beta_j$ are independent of $s_j$. 
-We can set both these assumptions within a more general framework, which allows that $\beta_j/s^\alpha$ is independent of $s_j$ for some $\alpha$ [Equation (\ref{eqn:beta-alpha}))].
-Setting $\alpha=0$ implies that $\beta_j$ is
-independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
-effects (in absolute value). This latter assumption may often be qualitatively plausible: 
-for example, in gene expression studies the standard error for gene $j$ depends partly on
-the variance of its expression among samples, and genes with a larger variance may tend to be less tightly regulated and
-so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
-On the other hand, there is no particular reason to expect that either $\alpha=1$ or $\alpha=0$ will be the optimal choice.
-Indeed, optimal choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$,
-which will be dataset-specific. 
-Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
-comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
-likelihood-based methods can be used to select $\alpha$. For example, following the logic of the EB approach it
-would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
-which has been implemented in our software by C. Dai.
-
-%\begin{result}
-%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
-%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
-%\end{result}
-
-%To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
-%Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
-%In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
-%Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
-%which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
-%(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
-%may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)
-
-%\begin{result}
-%Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and using a mixture of normals for the prior distribution 
-%(\ref{eqn:beta-alpha}) with $\alpha=1$, 
-%the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$.
-Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$)
-%is montonic with $|t_j|$. 
-%\end{result}
-
-%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
-%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.
-
-%Most previous authors of EB approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
-%have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
-%In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
-%This is analogous
-%to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
-%standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
-%that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.
-
-\begin{figure}
-\centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
-\caption{Figure illustrating affects of prior assumptions on re-ordering of significance.
-Left panel shows results under our ``default prior" which assumes that effects $\beta_j$ are identically distributed, independent of $s_j$.
-Right panel shows results under the ``$p$-value prior", which assumes that $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$.} \label{fig:lfsr_pval}
-\end{figure}
-
-%\subsubsection*{Adaptive Shrinkage}
-%
-%Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
-%when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
-%the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
-% $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
-% In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
-% Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
-% and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
-%
-%Our methods we build on this idea in two ways. First,
-%by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
-%Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
-%adapts to measurement precision: more precise measurements undergo
-% less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
-%than observations with
-%small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
-%To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
-%Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
-%
-%\begin{figure}[!ht]
-%\begin{center}
-%\includegraphics[width=400px]{../examples/ash.pdf} 
-%\end{center}
-%\caption{Figure illustrating adaptive shrinkage.
-Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0.
-The standard error of each observation, $s_j$, was simulated from Inverse Gamma(5,5), which resulted in $s_j$ varying from 0.34 to 6.7.
-The shrunken estimates ($y$ axis) are plotted against the observed estimates ($x$ axis), with color indicating the (square root of the) precision of each measurement (blue=lower-precision; red=higher-precision).
-The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
-%\end{figure}
-
-%DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
-%To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
-Because the prior distribution is unimodal about zero,
-%the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
-%amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
-%For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
-%than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
-%small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
-
-%previous work based on normal means:
-%http://link.springer.com/article/10.1007/BF02562623
-
-%\subsection{Null behaviour}
-
-%Results of simulations under the null.
-
-%\subsection{Comparison of different underlying component distributions}
-%
-%Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
-%than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
-%likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
-%For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
-%increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
-%expect $L_{hu}$ to be significantly larger than the others.
-%
-%Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
-%
-%\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
-%\begin{center}
-%\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
-%%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
-%\end{center}
-%\caption{}
-%\end{figure}
-
-%Notes from Adrian Raftery - 
-%Newcombe (1886)
-%Dean and Raftery
-%Uniform is super-efficient
-
-\section*{Discussion}
-
-We have presented an Empirical Bayes approach to large-scale multiple testing that emphasises two ideas.
-First, we emphasise the potential benefits of using two numbers ($\bhat$, and its standard error)
-rather than just one number (a $p$ value or $z$ score)  to summarize the information on each test.
-While requiring two numbers is slightly more onorous than requiring one, in many settings these
-numbers are easily available and if so we argue it makes sense to use them.
-Second, we note the potential benefits -- both statistical and computational -- of assuming that the effects come from a unimodal distribution, and provide
-flexible implementations for performing inference under this assumption. We also introduce the ``false sign rate" as
-an alternative measure of error to the FDR, and illustrate its improved robustness to errors in model fit, particularly mis-estimation
-of the proportion of null tests, $\pi_0$.
-
-Multiple testing is often referred to as a ``problem" or a ``burden". In our opinion, EB approaches turn this idea on its head,
-treating multiple testing as an {\it opportunity}: specifically, an opportunity to learn about the prior distributions, and other modelling assumptions,
-to improve inference and make informed decisions about significance (see also \cite{greenland1991empirical}). This view
-also emphasises that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
-Indeed, the FDR at a given fixed threshold does not depend on the number of tests: as the number of tests increases, both the true positives and false positives increase linearly,
-and the FDR remains the same.
-(If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example.
-The EB approach captures this dependence in an intuitive way:
-if there are lots of strong signals then we infer $\pi_0$ to be small, and the estimated FDR (or $\lfdr$, or $\lfsr$) at a given threshold may be low, even if a large number of tests were performed; and conversely
-if there are no strong signals then we infer $\pi_0$ to be large and the FDR at the same threshold may be high, even if relatively few tests were performed. 
-More generally, overall signal strength is reflected in the estimated $g$, which in turn influences the estimated FDR.
-
-Two important practical issues that we have not addressed here are correlations among tests,
-and the potential for deviations from the theoretical null distributions of test statistics. These two
-issues are connected: specifically, unmeasured confounding factors can cause both correlations among tests
-and deviations from the theoretical null \cite{efron2007correlation,leek:2007}. And although there are certainly other
-factors that could cause dependence among tests, unmeasured confounders are perhaps the most worrisome in practice
-because they can induce strong correlations among large numbers of tests and profoundly impact results.
-Approaches to deal with unmeasured confounders can be largely divided into two types: those
-that simply attempt to correct for the resulting inflation of test statistics \cite{devlin1999genomic,efron2004large}, and those
-that attempt to infer counfounders using clustering, principal components analysis, or factor models 
-\cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
-and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable,
-they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alterntively, our methods
-could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:lik}) using
-$\bhat_j | \shat_j \sim N(0,\lambda_1 \shat_j + \lambda_2)$, where $\lambda_1,\lambda_2$ are to be estimated. However,
-this immediately raises issues of identifiability and we do not pursue the idea further here.
-
-Another important practical issue is the challenge of small sample sizes. For example, in genomics applications researchers sometimes
-attempt to identify differences between two conditions based on only a handful of samples in each. In such settings the
-normal likelihood approximation (\ref{eqn:normlik}) will be inadequate. And, although the $t$ likelihood (\ref{eqn:tlik}) partially
-addresses this issue, it is also, it turns out, not entirely satisfactory.  The root of the problem is that, with small sample sizes, raw estimated standard
-errors $\shat_j$ can be horribly variable. In genomics it is routine to address this issue by applying EB methods \cite{smyth:2004} to ``moderate" (i.e.~shrink) variance estimates,
-before computing $p$ values from ``moderated" test statistics. We are currently investigating how
-our methods should incorporate such ``moderated" variance estimates to make it applicable to small sample settings.
-
-Our approach involves compromises between flexibility, generality, and simplicity on the one hand, and statistical efficiency and principle on the other.
-For example, in using an EB approach that uses a point estimate for $g$, rather than a fully Bayes approach that accounts for uncertainty in $g$,
-we have opted for simplicity over statistical principle. And in summarizing every test by
-two numbers and making a normal or $t$ approximation to the likelihood, we have aimed to produce generic methods that 
- can be applied whenever such summary data are available -- just as \qvalue can be applied to any set of $p$ values for example -- although
- possibly at the expense of statistical efficiency compared with developing multiple tailored approaches based on context-specific likelihoods. 
- Any attempt to produce generic methods will involve compromise between generality and efficiency.
- In genomics, many analyses -- not only FDR-based analyses -- involve first computing a series of $p$ values before
- subjecting them to some further downstream analysis. An important message here is 
- that working with two numbers ($\bhat_j,\shat_j$), rather than one ($p_j$ or $z_j$), 
- can yield substantial gains in functionality (e.g.~ estimating effect sizes, as well as testing; accounting for variations in
-measurement precision across units) while losing only a little in generality.  We hope that our work will
-encourate development of methods that exploit this idea in other contexts.
-
-%This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
-%from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
-%with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
-
-%
-%
-%Both \locfdr and \mixfdr allow for deviations from
-%the theoretical null distribution by allowing that, under the null, the $z$ scores are $N(0,\sigma^2)$ with $\sigma$ estimated from the data rather
-%than fixed to 1. Within our approach this idea could be incorporated by changing the likelihood - for example, to assume $\bhat_j \sim N(0,s_ 
-%
-%These are tricky issues that, as far as we are aware, can affect
-%all approaches to FDR analysis. An important focus of Efron's work, also implemented in \locfdr, 
-% is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
-%theoretical null distribution. 
-%
-%For example, we and others have worked on Bayesian approaches to genetic association testing, using subjective prior specification of
-%effect sizes based on external data (e.g.~\cite{stephens:2009}). However, the large numbers of tests brings the opportunity to learn about
-%priors, and here we aim to exploit this
-
-%Choice of mixture components. Use of likelihood to compare different mixture components?
-
-%  
-%Directly modeling the p values, or z scores, say via non-parametric methods, 
-%can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-%distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-%The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-%Before proceeding, we feel obliged to correct the common misconception
-% that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
-% not depend on the number of tests. 
-%  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
-% 
-% In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
-
-\section*{Detailed Methods} \label{sec:implement}
-
-\subsection*{Embellishments}
-
-\subsubsection*{More flexible unimodal distributions}
-
-Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
-Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
- Therefore, we have implemented a more general approach based on 
-\begin{equation} \label{eqn:g-general}
-g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) ,
-\end{equation}
-where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
-\begin{enumerate}[(i)]
-\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$, \qquad (``ash.n")
-\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,  \qquad (``ash.u")
-\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,  \qquad (``ash.hu")
-\end{enumerate}
-where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
-(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
-and $\pi$ is a $2K+1$ vector that sums to 1.)
-The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
-Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
-under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
-(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
-(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
-however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)
-
-Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
-using a large dense grid of $\sigma^2_k$ or $a_k$, (i)-(iii) can respectively approximate,
-with arbitrary accuracy,
-\begin{enumerate}[(i)]
-\item any scale mixture of normals, which includes as special cases
-the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
-\item any symmetric unimodal distribution about 0.
-\item any unimodal distribution about 0.
-\end{enumerate}
-The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
-In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.
-
-Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
-The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
-See Implementation for details of our software defaults.
-
-%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting.
-For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
-%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
-
-\subsubsection*{Dependence of effects on standard errors}
-
-Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
-This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
- \begin{equation} \label{eqn:beta-alpha}
- \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
- \end{equation}
-for any $\alpha$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
-assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
-it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
-the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.
-
-The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
-To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
- standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
- Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
-yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
-%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
-%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
-%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
-
-%Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
-%how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
-%it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
-%rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
-%and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
-%this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
-%do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
-
-\subsubsection*{Replace normal likelihood with $t$ likelihood}
-
-We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
-  \begin{equation} \label{eqn:tlik}
- \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
- \end{equation}
-where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
-and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
-The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
-This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms; see Implementation below.
-When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
-
-Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
-\begin{equation} \label{eqn:tdist}
-(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
-\end{equation}
-However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
-In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
-currently pursuing this improved approach (joint work with M.Lu) and results will be published elsewhere. 
-%We note here that
-%modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
-%where our simpler approach can produce unsatisfactory results.
-
-\subsubsection*{Non-zero mode}
-
-An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood, rather than fixed to 0.
-
-\subsection*{Implementation Details}
-
-\subsubsection*{Likelihood for $\pi$}
-
-We define the likelihood for $\pi$ to be the probability of the observed data $\bhat$ conditional on $\shat$: $L(\pi) := p(\bhat | \shat,\pi),$ which
-by our conditional indendence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$.
-[One might prefer to define the likelihood as $p(\bhat, \shat | \pi) = p(\bhat | \shat, \pi) p(\shat | \pi)$, in which case our definition comes down to assuming that the term $p(\shat | \pi)$ does not depend on $\pi$.]
-
-Using the prior $\beta_j \sim \sum_{k=0}^K \pi_k f_k(\beta_j)$ given by (\ref{eqn:g-general}), and the normal likelihood (\ref{eqn:normlik}), integrating over $\beta_j$ yields
-\begin{equation}
-p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
-\end{equation}
-where 
-\begin{equation}
-\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
-\end{equation}
-denotes the convolution of $f_k$ with a normal density.
-These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
-Specifically, 
-\begin{equation} \label{eqn:uconv}
-\tilde{f}_k(\bhat_j)  = 
-\begin{cases}
-N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
-\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
-\end{cases}
-\end{equation}
-where $\Psi$ denotes the cumulative distribution function (c.d.f.) of the standard normal distribution.
-If we replace the normal likelihood with the $t_\nu$ likelihood (\ref{eqn:tlik}) then the convolution 
-for $f_k$ uniform the convolution is still given by (\ref{eqn:uconv}) but with $\Psi$ the c.d.f. 
-of the $t_\nu$ distribution function. (The convolution for $f_k$ normal is tricky and we have not implemented it.)
-
-\subsubsection*{Penalty term on $\pi$}
-
-To make $\lfdr$ and $\lfsr$ estimates from our method ``conservative" we add a penalty 
-term $log(h(\pi;\lambda))$ to the log-likelihood $\log L(\pi)$ to encourage over-estimation of $\pi_0$:
-\begin{equation} \label{eqn:penalty}
-h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
-\end{equation}
-where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
-consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0sims}).
-
-Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
-we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
-
-\subsubsection*{Problems with removing the penalty term in the half-uniform case}
-
-It is straightforward to remove the penalty term by setting $\lambda_k=1$ in (\ref{eqn:penalty}).
-We note here an unanticipated problem we came across when using no penalty term in the half-uniform case 
-(i.e.~ $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$ in (\ref{eqn:g-general})): when the data
-are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
-sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
-for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
-and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
-
-\subsubsection*{EM algorithm}
-
-With this in place, the penalized log-likelihood for $\pi$ is given by:
-\begin{equation}
-\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
-\end{equation}
-where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This can be maximized using an EM algorithm \cite{dempster77}, whose
-one-step updates are:
-\begin{align} \label{eqn:w}
-w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
-n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
-\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}
-\end{align}
-Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
-and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
-probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$. (The $w_{kj}$ are sometimes referred to as the ``responsibilities".)
-Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
-We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple} to accelerate convergence of this EM algorithm.
-
-%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
-\subsubsection*{Initialization}
-
-By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
-Our rationale for initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
-but weak signal in the data cannot quickly draw the algorithm towards the null.
-
-In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
-$\pi_0=1; \pi_k=0 (k>0)$. If not then we replace the EM solution with the global null solution. This aims to guard against errors due to convergence to a local optimum when the
-data are consistent with the global null.
-
-\subsubsection*{Conditional distributions}
-
-Given $\hat\pi$, we compute the conditional distributions 
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
-\end{equation} 
-Each posterior is a mixture on $K+1$ components:
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
-\end{equation}
-where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
-and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
-$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
-All these posterior distributions are easily available.
-For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
-If $f_k$ is normal and $L$ is normal, then this is a normal distribution.
-
-\subsubsection*{Choice of grid for $\sigma_k, a_k$} \label{sec:grid}
-
-\def\sigmamax{\sigma_{\text{max}}}
-\def\sigmamin{\sigma_{\text{min}}}
-When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
-in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
-as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
-based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
-is negative we set $\sigmamax = 8\sigmamin$.)
-
-When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.
-
-Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
-a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.
-
-\appendix
-
-\begin{table}[!ht]
-\centering\begin{tabular}{c c } \toprule
-Scenario & Alternative distribution, $g_1$  \\ \midrule
-spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
-near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
-flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
- &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
-skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
-big-normal & $N(0,4^2)$ \\ 
-bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
-\end{tabular}
-\caption{Summary of simulation scenarios considered} \label{table:scenarios}
-\end{table}
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all_nopen.tex}
-	\caption{All observations}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg_nopen.tex}
-\caption{``Significant" negative discoveries.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos_nopen.tex}
-\caption{``Significant" positive discoveries.}
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{table:nopen}
-\end{table}
-%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
-%\begin{equation}
-%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
-%\end{equation}
-%which is close to the definition in \cite{storey.03}.
-
-% Do NOT remove this, even if you are not including acknowledgments
 \section*{Acknowledgements}
 
-Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
+Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the {\tt ggplot2} package \cite{ggplot2}, and text
 prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} , and includes contributions 
+git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} and includes contributions 
 from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.
@@
 This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.
-
-%\section*{References}
-% The bibtex filename
-\bibliography{/Users/stephens/Dropbox/Documents/mainbib}
-
-\section*{Figure Legends}
-
-\clearpage
-
-\section*{Tables}
-
-\section*{Supporting Information Legends}
-
-Supplementary material can be found in {\bf Supplementary Information S1.}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
-\end{document}

### 2015-10-06 sectionDetailed_Meth.tex
New Article!

--- 2015-10-06 paper/main.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
-\documentclass[11pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-%\usepackage{natbib}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-\usepackage{booktabs}
-\usepackage{xspace}
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-\usepackage{sectsty}
-\subsectionfont{\normalsize}
-\subsubsectionfont{\mdseries\itshape}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\usepackage{color}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-% Use the PLoS provided bibtex style
-\bibliographystyle{/Users/stephens/Dropbox/Documents/stylefiles/plos2009}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
-\pagestyle{myheadings}
-%% ** EDIT HERE **
-\usepackage{enumerate}
-\usepackage{multirow} 
-\usepackage{url}
-\usepackage{xr} %for cross-referencing
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-\newtheorem{algorithm}{Algorithm}
-\newtheorem{proposition}{Proposition}
-\newtheorem{restateproposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{result}{Result}
-\newtheorem{note}{Note}
-\newtheorem{definition}{Definition}
-
-\def\lfdr{\textit{lfdr}}
-\def\lfsr{\textit{lfsr}}
-\def\x{\mbf{x}}
-\def\y{\mbf{y}}
-\def\e{\mbf{e}}
-\def\g{\mbf{g}}
-\def\p{p}
-\def\Yb{\hat{Y}_1^{\text{BAYES}}}
-\def\Xperp{X_{1 \perp 0}}
-\def\Pperp{P_{1 \perp 0}}
-\def\Pperpb{\Pperp^{\text{B}}}
-\def \hatd { \widehat{ d } }
-\def \dhatin {\hatd^{\mbox{\tiny IN}}}
-\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
-\def \mvn{\text{MVN}}
-\def\mn{\text{MN}}
-\def \iw{\text{W}^{-1}}
-\def \fp {fastPHASE\ }
-\def\bhat{\hat{\beta}}
-\def\shat{\hat{s}}
-\def\var{\rm Var}
-\def\mvr{\text{BMVR}}
-\def\bfuni{\text{BF}_\text{uni}}
-\def\one{\mbf{1}}
-\def\zero{\mbf{0}}
-\def\hall{H_{\text{all}}}
-\def\pall{\p_{\text{all}}}
-\def\bfall{\text{BF}_\text{all}}
-\def\bfav{\text{BF}_\text{av}}
-\def\gall{\gamma_\text{all}}
-\def\limbfall{\bfall^\rightarrow}
-\def\limbfav{\bfav^\rightarrow}
-\def\limbfgamma{\BF_\gamma^\rightarrow}
-\def\sxx{S_{xx}}
-\def\syy{S_{yy}}
-\def\vyy{V_{yy}}
-\def\vyx{V_{yx}}
-\def\vxx{{V_{xx}}}
-\def\rss{\text{\rm RSS}}
-\def\dss{\text{\rm $\Delta$SS}}
-\def\U{U}
-\def\D{D}
-\def\I{I}
-\def\luu{L_{UU}}
-\def\indep{\perp}
-\def\qvalue{{\tt qvalue}\xspace}
-\def\locfdr{{\tt locfdr}\xspace}
-\def\mixfdr{{\tt mixfdr}\xspace}
-\def\ashr{{\tt ashr}\xspace}
-
-\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
-                                % mode too
-\newcommand\height{{\it height}}
-\newcommand\weight{{\it weight}}
-\def\BF{\text{BF}}
-
-%% END MACROS SECTION
-%\externaldocument{plos_one_rev1_suppinfo}
-
-\begin{document}
-
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{False Discovery Rates: A New Deal}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Matthew Stephens$^{1*}$, 
-\\
-\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
-\\
-$\ast$ E-mail: Corresponding mstephens@uchicago.edu
-\end{flushleft}
-
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-%\section*{Author Summary}
-
-\section*{Introduction}
-
-Since its introduction in 1995 by Benjamini and Hochberg 
-\cite{benjamini1995controlling}, the ``False Discovery Rate" (FDR) has quickly established itself
-as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing
-in which the goal is to identify the non-zero ``effects" among a large number of imprecisely-measured effects.
-
-Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
-indeed, the notion that EB approaches could be helpful in handling multiple comparisons predates introduction of
-the FDR  (e.g.~\cite{greenland1991empirical}). More recently, EB approaches to the FDR
-have been extensively studied by several authors, 
-especially B.~Efron and co-authors 
-\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
-see also \cite{kendziorski2003parametric,newton2004detecting, 
-datta2005empirical,muralidharan2010empirical} for example.
-So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
-existing widely-used FDR pipelines) and can substantially affect
-inference. The first idea is to {\it assume that the distribution of effects is unimodal}. This yields a very simple, fast, and stable computer 
-implementation, as well as improving inference of FDR when the unimodal assumption is correct.  
-The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
-to summarize each measurement. This idea allows variations in measurement precision to be better accounted for,
- and avoids a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
-
-In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
-the idea that it may be preferable to focus on estimation rather than on testing.
-In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
-simply estimation with some positive prior probability that the effect is exactly zero.
-However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
-perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
-controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
-``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
-in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
-the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the lfsr, particularly its increased robustness to modelling assumptions.
-
-%terminology
-%``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
-%and correspondingly define the ``false signed discovery rate", or simply ``false sign rate" (FSR), in a natural way.
-
-%The methods we provide here can address both problems, and we provide results that highlight important benefits of focussing on estimation, and particularly on assessing the correctness of the {\it sign} of the estimated effects, rather than on testing for non-zero effects.
-
-Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
-(We address the reasons for this name, and connections with shrinkage analysis, in the Discussion.)
-
-%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.
-
-%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-%measurements are made on only a small number of
-%samples in each condition - sometimes as few as one
-%sample in each condition. Thus the error in estimates
-%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-%in estimates of $\beta_j$ still greater.
-%
-
-%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
-%Here we present new tools that bring two important ideas to bear on the problem.
-%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
-%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
-%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
-%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
-%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
-%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
-%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
-%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
-%a substantial increase in the number of discoveries at a given FDR).
-
-%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
-%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
-Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
-%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
-%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
-%generic methods that work directly with two numbers for each 
-%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
-Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
-%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
-
-\def\df{df}
-\def\FDR{\text{FDR}}
-\def\fdr{\text{\lfdr}}
-\def\FSR{\text{FSR}}
-\def\fsr{\text{lfsr}}
-
- \section*{Methods}
-
- \subsection*{Model Outine}
-
-% Here we consider one setting in which FDR-based methods are widely used:
-%an experimenter has measured a large numbers of ``effects", each with an associated
-%measurement precision (standard error), 
-%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
-%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
-%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
-%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
-%For example, in genomics
-%a common goal is to identify ``differentially-expressed genes" -- that is, genes
-%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
-%mean expression level between the two conditions, often referred to as the ``log-fold change". 
-Here we describe the simplest version of our method, and briefly discuss embellishments we have also implemented.
-
-Suppose that we are interested in the values of $J$ ``effects" $\beta=(\beta_1,\dots,\beta_J)$. For example, in a typical genomics application
-that aims to identify differentially expressed genes,  $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
-In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
-that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
-more general problem of estimating, and assessing uncertainty in, $\beta_j$.
-
-Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
-and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
-which by Bayes theorem can be written as 
-\begin{equation}
-p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
-\end{equation}
-
-For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
-This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
-A simple way to implement the UA is to assume that $g$ is
-a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
-\begin{align} \label{eqn:beta}
-p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
-g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
-\end{align}
-where $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyperparameters, which are non-negative and sum to one, and are to be estimated,
-while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
-represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
-(We encourate the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
-although of course in practice we use a finite grid -- see Implementation Details.)
-
-For the likelihood $p(\bhat | \beta, \shat)$ we assume  
-\begin{equation} \label{eqn:normlik}
-p(\bhat | \beta, \shat) = \prod_j N(\bhat_j; \beta_j, \shat_j^2).
-\end{equation}
-Here, in additional to some conditional independence assumptions, we are 
-effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-
-This simple model features both the key ideas we want to emphasise in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
-the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
-with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
-However, this simple model also has several additional assumptions that can be relaxed. Specifically,
-\begin{enumerate}
-\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) also implies that $g$ is symmetric about 0; more flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see (\ref{eqn:g-general})].
-\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
-We can relax this to allow for a relationship between these quantities [see (\ref{eqn:beta-alpha})].
-\item The likelihood (\ref{eqn:normlik}) effectively assumes that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-We can generalize this likelihood using a $t$ likelihood [see (\ref{eqn:tlik})].
-\end{enumerate}
-These embellishments are detailed in Embellishments and Implementation Details.
-Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
-(which are also made by most existing EB approaches to this problem). Correlations among tests certainly arise in practice, either due to genuine correlations
-in the system of study, or due to unmeasured confounders, and their potential to impact results of an FDR analysis is important to consider
-whatever analysis methods are used: see \cite{efron2007correlation,leek:2007} for relevant discussion.
-
-%We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
-%$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
-%\begin{equation}
-%p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
-%\end{equation}
-%Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
-%\begin{equation} \label{eqn:normlik}
-%p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
-%
-%We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
-%as a mixture of null and alternative components:
-%\begin{equation} \label{eqn:betaj0}
-%g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
-%\end{equation}
-%where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
-%We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
-%\begin{equation} \label{eqn:beta-iid}
-%p(\beta | \shat) = \prod_j g(\beta_j).
-%\end{equation}
-%
-%Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
-%with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
-%to use a mixture of zero-mean normal distributions:
-%\begin{equation} \label{eqn:mixnorm}
-%g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
-%Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
-%represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
-% The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
-% more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below. 
-% 
-% Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
-% yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
-% \begin{equation} \label{eqn:g}
-% g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
-% \end{equation}
-% 
-% Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
-% \begin{align} \label{eqn:betaj}
-% p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
-% p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
-% \end{align}
-% with $g$ given by the mixture of normals $(\ref{eqn:g})$.
-
-\subsection*{Fitting the model}
-
-In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
-each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
-Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
-\begin{equation} \label{eqn:marginallik}
-p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
-\end{equation}
-where we define $\sigma_0:=0$.
-
-The usual EB approach to fitting this model would involve two simple steps:
-\begin{enumerate}
-\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
-\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
-against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | | \bhat, \shat,\hat{\pi})$.
-\end{enumerate}
-Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(The simplicity of the EM algorithm in step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
-which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
-
-Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximise a penalized
-likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
-We introduce this penalty because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
-
-\subsection*{The local False Discovery Rate and local False Sign Rate}
-
-As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
-In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
-One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
-the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
-In other words it is the posterior probability that $\beta_j$ is actually zero:
-\begin{equation} \label{eqn:fdr}
-\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
-\end{equation}
-
-The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
-whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
-the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey1991philosophy}) argued that
-``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$.
-Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
-\begin{quote}
-...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
-\end{quote}
-Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
-focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.
-
-Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
-would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
-\begin{equation} \label{eqn:lfsr}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
-\end{equation}
-To illustrate, suppose that 
-\begin{gather*}
-p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
-p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
-p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
-\end{gather*} 
-Then 
-from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
-our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.
-
-As our notation suggests, $\lfsr_j$ is intended to be compared and contrasted with $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
-small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
-Of coure, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
-this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
-(this follows from the definition because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
-In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
-$\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.
-
-From these ``local" measures of significance, we can also compute average error rates over subsets of observations
-$\Gamma \subset \{1,\dots,J\}$. For example,   
-\begin{equation} \label{eqn:FDRhat}
-\widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
-\end{equation}
-estimates the FDR we would obtain if we were to declare all tests in $\Gamma$ significant.
-And 
-\begin{equation} \label{eqn:qvalue}
-q_j := \widehat \FDR(\{k: \fdr_k \leq \fdr_j\}) 
-\end{equation}
-provides a measure of significance analogous to Storey's $q$ value \cite{storey.03}.
-
-%Because our methods involve explicitly modelling effects $\beta_j$, rather than 
-%$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
-
-%Indeed, even in settings where some
-%$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
-%analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
-
-\subsection*{Related work}
-
-\subsubsection*{Previous approaches focussed on FDR}
-
-Among previous methods that explicitly consider the FDR and related quantities,
-our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
- (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
-  %Here we highlight the key differences between our approach and these three existing approaches.
-
-There are two key differences between our approach and all of these three existing methods. First, 
-whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
-we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
-for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
-Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 
-how it can improve performance by taking better account of variation in measurement precision among observations.
-%(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
-% Or more generally Section xx illustrates the potential practical effects
-
-%allows for different modelling assumptions to be compared through likelihood comparisons.
-Second, our unimodal assumption (UA) that the effects are unimodal about zero is quite different from assumptions made by $\qvalue, \locfdr$ or $\mixfdr$.
-Indeed, $\locfdr$ assumes that all $z$ scores near 0 are null (Efron calls this the Zero Assumption; ZA), which implies that 
-under the alternative hypothesis the distribution of $z$ scores has {\it no mass at 0}; this contrasts strikingly with the UA, which implies that this distribution has its peak at 0! 
-Similarly, $\qvalue$ assumes that all $p$ values near 1 are null,
-which is the same as the ZA because $p$ values near 1 correspond to $z$ scores near 0. And although \mixfdr does not formally make the ZA,
-we have found that in practice, with default settings, the results often approximately satisfy the ZA (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}).
-Thus, not only do these existing methods not make the UA, they actually make assumptions that are, in some sense, as different
-from the UA as they can be.
-
-Given that the UA and ZA are so different, it seems worth discussing why we generally favor the UA. 
-Although the UA will not apply to all situations, we believe that it will often be reasonable, especially in FDR-related contexts
-that have traditionally focussed on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis,  
-it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible. Further, it seems reasonable to expect that larger effects become
-decreasingly plausible, and so the distribution of the effects will be unimodal about 0.
-To paraphrase Tukey, ``All we know about the world teaches us that large effects are rare, whereas small effects abound."
-We emphasise that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
-It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
-
-%Tukey "peri-null"
-
-%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
-%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}.
-
-In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficents - 
-common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
-These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
-it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
-The UA assumption on regression coefficients is directly analagous to our UA here, and so its
-widespread use in the regression context supports its use here.
-
-Alternatively, we could motivate the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
-such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
-make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).
-
-Finally, the UA also has a considerable practical benefit: it yields simple algorithms that are both computationally
-and statistically stable. We illustrate these features in Results.
-
-\subsubsection*{Other work}
-
-There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
- Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
-indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
-This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
-Also, they provide no software implementation.
-More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
-Some pointers to recent papers in which other relevant
-citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}.
-Much of the literature focusses on the homoskedastic case (i.e.
-$\shat_j$ all equal) whereas we allow for heteroskedasticity.
-And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential.
-Several recent papers consider more flexible non-parametric assumptions on
-$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
-estimate (NPMLE) for $g$.
-These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive.
-However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
-may perform well for point estimation, it seems possible it will not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
-Indeed, one way of thinking about the UA is as a way to regularize $g$.
-
-%The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
-%encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-%$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-%$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-%the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-%observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-%corresondingly little effect on the estimated FDR of other observations. 
-%We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-%$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-%that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-%estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-%Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-\def\dt{\text{dt}}
-
-%
-%assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
-%$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
-% $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
-%  \begin{equation} \label{eqn:lik}
-% \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
-% \end{equation}
-% where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e.
-the density of $\mu+\sigma T_\nu$).
-In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
-%  \begin{equation} \label{eqn:normlik}
-% \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
-% \end{equation}
-%which, in light of (\ref{eqn:betahat}), is equivalent to
- % \begin{equation} \label{eqn:lik2}
-% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
- %\end{equation}
-
-%The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
-they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
-The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$.
-Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$.
-If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$.
-See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.
-
-%This likelihood approximation can be motivated by assuming that i) $\bhat$
-% $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
-% $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$. 
-%
-
-%Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.
-
-%While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
- %that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
-%there is an important difference: here,
-
-%The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
-For example, we can write $p(\bhat_j, \shat_j | \beta_j) = p(\bhat_j | \shat_j, \beta_j)p(\shat_j | \beta_j)$.
-Now, if we are willing to assume that $\shat_j$ alone contains no information about $\beta_j$, or equivalently that $p(\shat_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\bhat_j | \shat_j, \beta_j)$, and if $\bhat_j | \shat_j, \beta_j \sim N(\beta_j, \shat_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.
-
-\section*{Results}
-
-We compare results of \ashr with existing FDR-based methods implemented
-in the R packages \qvalue (v2.0 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
-In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
-to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.
-
-\subsection*{Effects of the Unimodal Assumption}
-
-Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
-standard error, $s_j=1$ and all methods are provided that information. That is,
-$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$.
-In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the only difference between our method and methods like $\locfdr$ and $\mixfdr$ are in how they estimate $g$.
-
-To briefly summarize the results in this section:
-\begin{enumerate}
-\item The UA can produce very different inferences compared with the ZA made by existing methods.
-\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
-\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
-\end{enumerate}
-
-\subsubsection*{The UA and ZA can produce different inferences}
-
-\begin{figure}
-\center \includegraphics[height=6in]{figures/decomp_ZA.pdf}
-\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
-The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
-In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
-The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
-(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
-\end{figure}
-
-To illustrate the different inferences from the UA and ZA we show results for a single
-dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
-but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
-\qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
-The results (Figure \ref{fig:ZA}) illustrate the clear difference between the existing methods and our method. The effects of the ZA made by \qvalue and \locfdr
-are visually clear, producing a ``hole"  in the alternative $z$ score distribution around 0. Although \mixfdr does not formally make the ZA, its decomposition
-exhibits a similar hole. In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
-effectively ``filling in" the hole.  (Of course the null distribution also has a peak at 0, and the local fdr 
- under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)
-
-Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA.
-Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
-This would increase the null component (dark blue) at the expense of
-the alternative component (light blue).
-Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0.
-The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears.
-In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
-See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
-demonstration.
-
- \subsubsection*{The UA can produce conservative estimates of $\pi_0$}
-
-The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
-Consequently \ashr will estimate smaller $\lfdr$s and FDRs than existing methods that make the ZA. 
-This is desirable, provided that these estimates remain conservative: that is,
-provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
-The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behaviour. To check its effectiveness
-we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
-for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
-They range from a ``spiky" distribution -- where many non-zero $\beta$ are
-too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
-flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
-making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
-which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.
-
-% How do these differences influence estimates of the FDR? To address this, note that  -- in this simple case where the 
-% $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
-% of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
-% strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
-% although the methods considered here differ from one another in many ways, having estimated $\pi_0$
-% they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
-% empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
-%
-
-%Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
-% and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
-%Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
-%At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
-%overestimates of $\pi_0$, which will yield overestimates of the FDR).
-
-\begin{figure} 
-\begin{center}
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=1.8in]{figures/scenarios_density.pdf} 
-	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/pi0_est.pdf} 
-\caption{Comparison of true and estimated values of $\pi_0$.
-When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ashr being least conservative, and hence most accurate.
-When the UA does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.} \label{fig:pi0sims}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfdr_est.png} 
-\caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
-Black line is $y=x$ and red line is $y=2x$.
-Estimates of $\lfdr$ are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est.png} 
-\caption{As in c), but for $\lfsr$ instead of $\lfdr$.
-Estimates of $\lfsr$ are consistently less conservative than $\lfdr$ when UA holds, and also less anti-conservative in bimodal scenario.}  \label{fig:lfsr}
-\end{subfigure}
-\end{center}
-\caption{Results of simulation studies (constant precision $s_j=1$).}
-\end{figure}
-
-For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
-data as follows:
-\begin{enumerate}
-\item Simulate $\pi_0  \sim U[0,1]$.
-\item For $j=1,\dots,J$, simulate $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)$.
-\item For $j=1,\dots,J$, simulate $\bhat_j | \beta_j \sim N(\beta_j,1)$.
-\end{enumerate}
-%Thus these simulations assume the same
-%precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).
-
-Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ashr ($y$ axis) with the true values ($x$ axis). 
-For \ashr we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
-and as a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
-as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
-The results show that \ashr provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
-in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ashr estimates can be slightly anti-conservative.
-We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
-are used. (In addition, the effects on $\lfsr$ estimates turn out to be relatively modest; see below).
-
-%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
-%
-%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
-%and so all effects are non-zero and no nulls hold. This scenario is rather different
-%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
-%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
-%``mostly non-null"
-%scenario, and we would agree! On the other hand, the distribution of $p$ values
-%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
-%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
-%Thus it seems desirable to have methods that perform well in both cases, and this is
-%an important feature of the methods we introduce here.
-%
-%TODO: put FDRs on graph? Simulate with 50\% null?
-
-\subsubsection*{The $\lfsr$ is more robust than \lfdr}
-
-The results above show that \ashr can improve on existing methods in producing smaller, more accurate, 
-estimates of $\pi_0$, which will lead to more accurate estimates of FDR.
-Nonetheless, in many scenarios \ashr continues
-to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
-This is because these scenarios include 
-an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
-estimation of $\pi_0$ impossible. Put another way, and as is well known, $\pi_0$ is not identifiable:
- the data can effectively provide an upper bound on plausible values of $\pi_0$,
-but not a lower bound (because the data cannot rule out that everything is non-null, but with miniscule effects).
-To obtain conservative behaviour we must estimate $\pi_0$ by this upper bound, which can
-be substantially larger than the true value.
-
-Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
-overestimation of $\pi_0$ is corresponding overestimation of FDR (and $\lfdr$, and $q$ values).
-To illustrate, Figure \ref{fig:lfdr} compares the estimated $\lfdr$ from ash.n with the true value (computed using
-Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios which involve many
-non-zero effects that are very near 0 (e.g.~the spiky scenario with $\pi_0$ small) where
-$\pi_0$ can be grossly overestimated.  
-(Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.)
-
-The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
- become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
- This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
-To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
-be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
-the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in $\lfsr$ than $\lfdr$.
-
-Note that, compared with previous debates regarding testing, 
-this section advances an additional reason for focussing on the sign of the effect, rather than just testing whether it is 0. 
-In previous debates authors have argued against testing whether an effect is 0
-because it is {\it implausible that effects are exactly 0}. Here we add that {\it even if one believes
-that some effects may be exactly zero}, it is still better to focus on the sign, because generally {\it the data are more informative about that question}
-and so inferences are more robust to, say, the inevitable mis-estimation of $\pi_0$.
-To provide some intuition, consider an observation with a $z$ score of 0. The $\lfdr$ of this observation can range from 0 (if $\pi_0=0$)
-to 1 (if $\pi_0=1$). But, assuming a symmetric $g$, the $\lfsr>0.5$ whatever the value of $\pi_0$, because the observation $z=0$ says
-nothing about the sign of the effect.
-Thus, are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
-whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.
-
-Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
-Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
-if we are prepared to assume that no effects are exactly zero, then removing the point mass 
-yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
-However, there is ``no free lunch":  if in fact some effects are exactly zero
-then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
-We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.
-
-\begin{figure}
-\begin{center}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s_nn.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be more accurate in this case.
-That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
-\end{subfigure}
-\end{center}
-\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
-\end{figure}
-
-\subsubsection*{Numerical Stability}
-
-The EM algorithm, which we use here to fit our model, is notorious for convergence to local optima. 
-However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
-caused by such behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
-in addition running it using our default initialization procedure (see Implementation details). We then compared the largest
-log-likelihood achieved across all 11 runs with the log-likelihood achieved by the default run.
-When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference ($< 0.02$), 
-and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference ($<0.02$),
-and 6\% of runs showed an appreciable log-likelihood difference ($>1$), with the largest difference being 5.0. However, perhaps suprisingly, 
-even for this largest difference results from the default run (e.g.
-the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood.
-See \url{http://github.com/stephens999/ash/blob/master/dsc-robust/summarize_dsc_robust.rmd} for further details.
-
-\subsubsection*{The UA helps provide reliable estimates of $g$}
-
-An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
-can estimate the {\it size} of each effect $\beta_j$.
-Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
-to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
-threshold $T$.
-%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
-%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
-%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
-%towards zero, with the strength of the shrinkage being estimated from the data.
-Further, because the posterior distribution is, by definition,
-conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
-been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
-In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
-\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
-how can confidence intervals be assigned to the true
-effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
-and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
-%but also notes the difficulties of implementing a stable general algorithm.
-
-The ability of the EB approach to provide valid ``post-selection" interval estimates is extremely attractive in principle.
-But its usefulness in practice
-depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
-which are notoriously difficult in general. Indeed, Efron emphasises
-the difficulties of implementing a stable general algorithm, noting in his rejoinder
-``the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying
-to be overly nonparametric ... but it is hard to imagine a
-generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
-Our key point here is that the UA greatly simplifies the deconvolution problem.
-While not meeting Efron's desire for an entirely general nonparametric approach, we believe 
-that the UA can handle many cases of practical interest.
-
-To illustrate this, Figure \ref{fig:egcdf} compares the
-estimated $g$ from \ashr  with that from \mixfdr which does not make the UA 
-(and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
-The greater reliability of estimates afforded
-by the UA is immediately apparent. In particular the estimated cdf from \mixfdr often has an almost-vertical
-segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. 
-The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
-While the UA is not the only way to achieve this, we find it an attractive, simple and effective approach.
-
-Interestingly, even in the ``bimodal" scenario $\ashr$ 
-is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
-Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
-Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
-to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
-(\url{http://github.com/stephens999/ash/blob/master/dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
-clearly visible in the observed $\bhat$.
-
-We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ashr tends to systematically
-overestimate the mass of $g$ near zero. On careful inspection, this is apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
-and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf}b),
-and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
-under-estimating $\pi_0$ (Figure \ref{fig:egcdf}c).
-
-\begin{figure}[h!] 
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=2in]{figures/egcdf.pdf}
-	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf_nopen.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
-Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
-\end{subfigure}
-\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.} \label{fig:egcdf}
-\end{figure}
-
-%making the UA provides an attractive and simple way to achieve estimating $g$  large
-%subset of cases of interest (those where the UA can reasonably be assumed to hold).
-
-%In practice, the utility of the posterior distributions noted above depends primarily on the accuracy
-%of the estimate of the distribution $g(\cdot)$.
-
- %quote from efron rejoinder , p46:
-%MorrisÕ Section 3 is especially pertinent. His formula
-%for p(?|z) is related to my discussion of the
-%BenjaminiÐYekutieli False Coverage Rate method in
-%Section 7, particularly (7.2)Ð(7.4). Originally I had
-%hoped to develop an empirical Bayes method for estimating
-%such models, but the effort foundered on practical
-%difficulties involving the perils of deconvolution.
-%Section 6 on the Òone-group modelÓ is the ugly duckling
-%of the current paper, but it bears on some important
-%points raised in the discussion. Figure 7 concerns
-%a fuzzy version of simultaneous hypothesis testing,
-%where, as in MorrisÕ hospital example, the usual singlepoint
-%null hypotheses seem unequal to the task. The
-%development from (6.6) onwards, particularly (6.12),
-%bears on the possibility of nonnormal null distributions,
-%and is about as far as I can go in answering Professors
-%Rice and SpiegelhalterÕs penultimate question.
-%With g(?) a normal distribution, model (6.1) returns
-%us to the realm of Stein estimation, the scene of my
-%happy collaborations with Carl Morris. I continue to
-%be surprised at how much harder bumpy, nonnormal
-%models like (7.1) are to deal with. JamesÐStein estimation
-%works fine with, say, N = 10 component problems,
-%while the RobbinsÕ form of empirical Bayes appropriate
-%to (7.1) seems to require hundreds or thousands.
-%The information calculations in Efron (2008)
-%reinforce this gloomy assessment. Maybe I am trying
-%to be overly nonparametric in constructing the empirical
-%Bayes Fdr estimates, but it is hard to imagine a
-%generally satisfactory parametric formulation for (6.1).
-%Or perhaps it is just that hypothesis testing is more demanding
-%than estimation.
-%My Note: (6.1) is the general model (mu \sim g()).
-%7.1 is the case where the effects are a mixture of zero and N(-3,1).
-
-\subsubsection*{Calibration of posterior intervals}
-
-To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over 100 simulations in each Scenario). 
-Specifically we examine the empirical coverage of nominal lower 95\% credible bounds for a) all observations; b) significant negative discoveries; c) significant positive discoveries.
-We examine only lower bounds because the results for upper bounds follow by
-symmetry (except for the one asymmetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
-in each case: for negative discoveries the lower bound is typically large and negative and limits how big (in absolute value) 
-the effect could be; for positive discoveries the lower bound 
-is positive, and limits how small (in absolute value) the effect could be. Intuitively, the lower bound for negative discoveries depends on the accuracy of $g$ in its tail,
-whereas for positive discoveries it is more dependent on the accuracy of $g$ in the center.
-
-The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
-as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
-
-% As noted above, the penalty term (\ref{eqn:penalty}) causes systematic over-estimation of the mass of $g$ near 0,  which one would expect to overshrink interval estimates towards zero. 
-%The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound is negative and overshrinking (moving it too far towards zero) makes the left tail too big; conversely, for positive discoveries, the lower confidence bound is positive, and
-%overshrinking (moving it too far towards zero) makes the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
-%and be prepared to put up with some overshrinkage to avoid it.
-
-%Coverage rates are given in:
-%\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
-
-%Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all.tex}
-	\caption{All observations.
-Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario.
-This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero.
-Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (Table \ref{tab:nopen}).
-Removing the penalty in the half-uniform case is not recommended (see Appendix).}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg.tex}
-\caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
-and the normal-based method in the flat-top scenario.
-These results likely reflect inaccurate estimates of the tails of $g$ due to a disconnect between the tail of $g$ and the component distributions in these cases.
-For example, the uniform methods sometimes substantially underestimate the length of the tail of $g$ in these long-tailed scenarios,
-causing over-shrinkage of the tail toward 0.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos.tex}
-\caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds} \label{tab:coverage}
-\end{table}
-
-%
-%\subsubsection*{Reverse asymptotics}
-%
-%Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
-%
-%Note: we'd like to make sure it behaves ok in the case where most observations are null.
-%Perhaps simulate with 999 null and 1 alternative?
-%Also maybe look at performance for small sample sizes - e.g. n=1?
-%
-%Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
-%can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
-%by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
-%
-
-\subsection*{Differing measurement precision across units}
-
- We turn now to the second important component of our work: allowing for varying
- measurement precision across units. The key to this is the use of a likelihood,
- (\ref{eqn:normlik}) or (\ref{eqn:tlik}), that explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
-
-  To illustrate, we conduct a simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
- ($s_j=10$).  In both cases,
-we assume that half the effects are null
-and the other half are normally distributed with standard deviation 1:
-\begin{equation}
-p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1).
-\end{equation}
-In this setting, the poor-precision measurements $(s_j=10)$ tell us very little, and any sane analysis should effectively ignore them. 
- However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor}). This is because the poor measurements 
- produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
- which, when combined with the good-precision measurements, dilute the overall signal (e.g. they reduce the density of $p$ values near 0).
- This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
-the estimated error rates ($q$-values, or $\lfdr$ values) for the good-precision observations increase when the low-precision observations are included in the analysis
-(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
-for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).
-
-\begin{figure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_hist.pdf} 
-    \label{fig:goodpoor-hist}
-    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations}
- \end{subfigure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_scatter.pdf}
-    \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) of a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
-\end{subfigure}
-\caption{Simulation illustrating how, for existing FDR methods, 
-poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
-$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
-a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
-In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
-\end{figure}
-
- \subsubsection*{Reordering of significance, and the ``$p$ value prior"}
-
-Another consequence of accounting for differences in measurement precision is that \ashr may re-order the significance of the observations compared
- with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
- We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measuments that have the same $p$ value.
- The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
- and so the $\lfsr$ for these poor-precision measurements is always high.
-
-The potential for Bayesian analyses to re-order the significance of observations, and specifically to downweight imprecise observations,
-was previously discussed in \cite{guan.stephens.08}.
-However, Wakefield \cite{wakefield:2009} showed that, under a certain prior assumption, the Bayesian analysis produces the same ranking of significance as $p$ values (or their $z$ scores).
-He named this prior the ``$p$-value prior" because it can be thought of as the
-implicit prior assumption that is being made if we rank the significance of observations by their $p$ value.
-Wakefield's $p$-value prior assumes that the less precise effect estimates correspond to larger true effects, and specifically that they scale proportional
-to the standard errors $s_j$. More specifically still, it assumes a normal prior for the non-zero $\beta_j$ with mean 0 and variance $K s^2_j$ for some constant $K$.
-Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. Specifically, if, instead of assuming that $\beta_j$ is independent
-of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
-then the $\lfsr$ computed by ash.n provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
-as is illustrated in Figure \ref{fig:lfsr_pval}, right panel. (This result does not hold when using the mixtures of uniforms prior, ash.u.)
-
-This $p$-value prior assumes that the $z$ scores $z_j=\beta_j/s_j$
-are identically distributed, independent of $s_j$. This is essentially the assumption made, implicitly or explicitly, by existing methods -- like \locfdr, \mixfdr and \qvalue --
-that model the $z_j$ or $p_j$ directly.  In contrast, we have assumed up to now that the $\beta_j$ are independent of $s_j$. 
-We can set both these assumptions within a more general framework, which allows that $\beta_j/s^\alpha$ is independent of $s_j$ for some $\alpha$ [Equation (\ref{eqn:beta-alpha}))].
-Setting $\alpha=0$ implies that $\beta_j$ is
-independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
-effects (in absolute value). This latter assumption may often be qualitatively plausible: 
-for example, in gene expression studies the standard error for gene $j$ depends partly on
-the variance of its expression among samples, and genes with a larger variance may tend to be less tightly regulated and
-so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
-On the other hand, there is no particular reason to expect that either $\alpha=1$ or $\alpha=0$ will be the optimal choice.
-Indeed, optimal choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$,
-which will be dataset-specific. 
-Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
-comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
-likelihood-based methods can be used to select $\alpha$. For example, following the logic of the EB approach it
-would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
-which has been implemented in our software by C. Dai.
-
-%\begin{result}
-%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
-%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
-%\end{result}
-
-%To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
-%Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
-%In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
-%Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
-%which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
-%(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
-%may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)
-
-%\begin{result}
-%Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and using a mixture of normals for the prior distribution 
-%(\ref{eqn:beta-alpha}) with $\alpha=1$, 
-%the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$.
-Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$)
-%is montonic with $|t_j|$. 
-%\end{result}
-
-%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
-%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.
-
-%Most previous authors of EB approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
-%have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
-%In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
-%This is analogous
-%to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
-%standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
-%that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.
-
-\begin{figure}
-\centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
-\caption{Figure illustrating affects of prior assumptions on re-ordering of significance.
-Left panel shows results under our ``default prior" which assumes that effects $\beta_j$ are identically distributed, independent of $s_j$.
-Right panel shows results under the ``$p$-value prior", which assumes that $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$.} \label{fig:lfsr_pval}
-\end{figure}
-
-%\subsubsection*{Adaptive Shrinkage}
-%
-%Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
-%when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
-%the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
-% $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
-% In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
-% Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
-% and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
-%
-%Our methods we build on this idea in two ways. First,
-%by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
-%Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
-%adapts to measurement precision: more precise measurements undergo
-% less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
-%than observations with
-%small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
-%To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
-%Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
-%
-%\begin{figure}[!ht]
-%\begin{center}
-%\includegraphics[width=400px]{../examples/ash.pdf} 
-%\end{center}
-%\caption{Figure illustrating adaptive shrinkage.
-Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0.
-The standard error of each observation, $s_j$, was simulated from Inverse Gamma(5,5), which resulted in $s_j$ varying from 0.34 to 6.7.
-The shrunken estimates ($y$ axis) are plotted against the observed estimates ($x$ axis), with color indicating the (square root of the) precision of each measurement (blue=lower-precision; red=higher-precision).
-The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
-%\end{figure}
-
-%DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
-%To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
-Because the prior distribution is unimodal about zero,
-%the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
-%amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
-%For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
-%than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
-%small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
-
-%previous work based on normal means:
-%http://link.springer.com/article/10.1007/BF02562623
-
-%\subsection{Null behaviour}
-
-%Results of simulations under the null.
-
-%\subsection{Comparison of different underlying component distributions}
-%
-%Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
-%than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
-%likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
-%For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
-%increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
-%expect $L_{hu}$ to be significantly larger than the others.
-%
-%Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
-%
-%\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
-%\begin{center}
-%\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
-%%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
-%\end{center}
-%\caption{}
-%\end{figure}
-
-%Notes from Adrian Raftery - 
-%Newcombe (1886)
-%Dean and Raftery
-%Uniform is super-efficient
-
-\section*{Discussion}
-
-We have presented an Empirical Bayes approach to large-scale multiple testing that emphasises two ideas.
-First, we emphasise the potential benefits of using two numbers ($\bhat$, and its standard error)
-rather than just one number (a $p$ value or $z$ score)  to summarize the information on each test.
-While requiring two numbers is slightly more onorous than requiring one, in many settings these
-numbers are easily available and if so we argue it makes sense to use them.
-Second, we note the potential benefits -- both statistical and computational -- of assuming that the effects come from a unimodal distribution, and provide
-flexible implementations for performing inference under this assumption. We also introduce the ``false sign rate" as
-an alternative measure of error to the FDR, and illustrate its improved robustness to errors in model fit, particularly mis-estimation
-of the proportion of null tests, $\pi_0$.
-
-Multiple testing is often referred to as a ``problem" or a ``burden". In our opinion, EB approaches turn this idea on its head,
-treating multiple testing as an {\it opportunity}: specifically, an opportunity to learn about the prior distributions, and other modelling assumptions,
-to improve inference and make informed decisions about significance (see also \cite{greenland1991empirical}). This view
-also emphasises that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
-Indeed, the FDR at a given fixed threshold does not depend on the number of tests: as the number of tests increases, both the true positives and false positives increase linearly,
-and the FDR remains the same.
-(If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example.
-The EB approach captures this dependence in an intuitive way:
-if there are lots of strong signals then we infer $\pi_0$ to be small, and the estimated FDR (or $\lfdr$, or $\lfsr$) at a given threshold may be low, even if a large number of tests were performed; and conversely
-if there are no strong signals then we infer $\pi_0$ to be large and the FDR at the same threshold may be high, even if relatively few tests were performed. 
-More generally, overall signal strength is reflected in the estimated $g$, which in turn influences the estimated FDR.
-
-Two important practical issues that we have not addressed here are correlations among tests,
-and the potential for deviations from the theoretical null distributions of test statistics. These two
-issues are connected: specifically, unmeasured confounding factors can cause both correlations among tests
-and deviations from the theoretical null \cite{efron2007correlation,leek:2007}. And although there are certainly other
-factors that could cause dependence among tests, unmeasured confounders are perhaps the most worrisome in practice
-because they can induce strong correlations among large numbers of tests and profoundly impact results.
-Approaches to deal with unmeasured confounders can be largely divided into two types: those
-that simply attempt to correct for the resulting inflation of test statistics \cite{devlin1999genomic,efron2004large}, and those
-that attempt to infer counfounders using clustering, principal components analysis, or factor models 
-\cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
-and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable,
-they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alterntively, our methods
-could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:lik}) using
-$\bhat_j | \shat_j \sim N(0,\lambda_1 \shat_j + \lambda_2)$, where $\lambda_1,\lambda_2$ are to be estimated. However,
-this immediately raises issues of identifiability and we do not pursue the idea further here.
-
-Another important practical issue is the challenge of small sample sizes. For example, in genomics applications researchers sometimes
-attempt to identify differences between two conditions based on only a handful of samples in each. In such settings the
-normal likelihood approximation (\ref{eqn:normlik}) will be inadequate. And, although the $t$ likelihood (\ref{eqn:tlik}) partially
-addresses this issue, it is also, it turns out, not entirely satisfactory.  The root of the problem is that, with small sample sizes, raw estimated standard
-errors $\shat_j$ can be horribly variable. In genomics it is routine to address this issue by applying EB methods \cite{smyth:2004} to ``moderate" (i.e.~shrink) variance estimates,
-before computing $p$ values from ``moderated" test statistics. We are currently investigating how
-our methods should incorporate such ``moderated" variance estimates to make it applicable to small sample settings.
-
-Our approach involves compromises between flexibility, generality, and simplicity on the one hand, and statistical efficiency and principle on the other.
-For example, in using an EB approach that uses a point estimate for $g$, rather than a fully Bayes approach that accounts for uncertainty in $g$,
-we have opted for simplicity over statistical principle. And in summarizing every test by
-two numbers and making a normal or $t$ approximation to the likelihood, we have aimed to produce generic methods that 
- can be applied whenever such summary data are available -- just as \qvalue can be applied to any set of $p$ values for example -- although
- possibly at the expense of statistical efficiency compared with developing multiple tailored approaches based on context-specific likelihoods. 
- Any attempt to produce generic methods will involve compromise between generality and efficiency.
- In genomics, many analyses -- not only FDR-based analyses -- involve first computing a series of $p$ values before
- subjecting them to some further downstream analysis. An important message here is 
- that working with two numbers ($\bhat_j,\shat_j$), rather than one ($p_j$ or $z_j$), 
- can yield substantial gains in functionality (e.g.~ estimating effect sizes, as well as testing; accounting for variations in
-measurement precision across units) while losing only a little in generality.  We hope that our work will
-encourate development of methods that exploit this idea in other contexts.
-
-%This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
-%from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
-%with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
-
-%
-%
-%Both \locfdr and \mixfdr allow for deviations from
-%the theoretical null distribution by allowing that, under the null, the $z$ scores are $N(0,\sigma^2)$ with $\sigma$ estimated from the data rather
-%than fixed to 1. Within our approach this idea could be incorporated by changing the likelihood - for example, to assume $\bhat_j \sim N(0,s_ 
-%
-%These are tricky issues that, as far as we are aware, can affect
-%all approaches to FDR analysis. An important focus of Efron's work, also implemented in \locfdr, 
-% is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
-%theoretical null distribution. 
-%
-%For example, we and others have worked on Bayesian approaches to genetic association testing, using subjective prior specification of
-%effect sizes based on external data (e.g.~\cite{stephens:2009}). However, the large numbers of tests brings the opportunity to learn about
-%priors, and here we aim to exploit this
-
-%Choice of mixture components. Use of likelihood to compare different mixture components?
-
-%  
-%Directly modeling the p values, or z scores, say via non-parametric methods, 
-%can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-%distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-%The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-%Before proceeding, we feel obliged to correct the common misconception
-% that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
-% not depend on the number of tests. 
-%  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
-% 
-% In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
-
 \section*{Detailed Methods} \label{sec:implement}
@@
  Therefore, we have implemented a more general approach based on
-\begin{equation} \label{eqn:g-general}
+
+\begin{equation}
+\label{eqn:g-general}
 g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) ,
@@
 where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms:
-\begin{enumerate}[(i)]
+
+\begin{enumerate}
+[(i)]
 \item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$, \qquad (``ash.n")
@@
 with arbitrary accuracy,
-\begin{enumerate}[(i)]
+
+\begin{enumerate}
+[(i)]
 \item any scale mixture of normals, which includes as special cases
@@
 
-%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting.
-For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
-%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
-
 \subsubsection*{Dependence of effects on standard errors}
@@
 This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
- \begin{equation} \label{eqn:beta-alpha}
+
+\begin{equation}
+\label{eqn:beta-alpha}
  \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
@@
 yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
-%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
-%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
-%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
-
-%Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
-%how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
-%it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
-%rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
-%and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
-%this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
-%do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
 
@@
 We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
-  \begin{equation} \label{eqn:tlik}
+
+\begin{equation}
+\label{eqn:tlik}
  \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
@@
 Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
-\begin{equation} \label{eqn:tdist}
+
+\begin{equation}
+\label{eqn:tdist}
 (\bhat_j-\beta_j)/\shat_j \sim T_\nu.
@@
 currently pursuing this improved approach (joint work with M.Lu) and results will be published elsewhere.
-%We note here that
-%modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
-%where our simpler approach can produce unsatisfactory results.
 
@@
 Using the prior $\beta_j \sim \sum_{k=0}^K \pi_k f_k(\beta_j)$ given by (\ref{eqn:g-general}), and the normal likelihood (\ref{eqn:normlik}), integrating over $\beta_j$ yields
+
 \begin{equation}
@@
 where
+
 \begin{equation}
@@
 Specifically,
-\begin{equation} \label{eqn:uconv}
+
+\begin{equation}
+\label{eqn:uconv}
 \tilde{f}_k(\bhat_j)  = 
@@
 term $log(h(\pi;\lambda))$ to the log-likelihood $\log L(\pi)$ to encourage over-estimation of $\pi_0$:
-\begin{equation} \label{eqn:penalty}
+
+\begin{equation}
+\label{eqn:penalty}
 h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
@@
 With this in place, the penalized log-likelihood for $\pi$ is given by:
+
 \begin{equation}
@@
 one-step updates are:
-\begin{align} \label{eqn:w}
+
+\begin{align}
+\label{eqn:w}
 w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
@@
 
-%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
 \subsubsection*{Initialization}
@@
 Given $\hat\pi$, we compute the conditional distributions
+
 \begin{equation}
@@
 Each posterior is a mixture on $K+1$ components:
+
 \begin{equation}
@@
 
-\def\sigmamax{\sigma_{\text{max}}}
-\def\sigmamin{\sigma_{\text{min}}}
 When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
@@
 \appendix
-
-\begin{table}[!ht]
-\centering\begin{tabular}{c c } \toprule
-Scenario & Alternative distribution, $g_1$  \\ \midrule
-spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
-near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
-flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
- &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
-skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
-big-normal & $N(0,4^2)$ \\ 
-bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
-\end{tabular}
-\caption{Summary of simulation scenarios considered} \label{table:scenarios}
-\end{table}
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all_nopen.tex}
-	\caption{All observations}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg_nopen.tex}
-\caption{``Significant" negative discoveries.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos_nopen.tex}
-\caption{``Significant" positive discoveries.}
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{table:nopen}
-\end{table}
-%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
-%\begin{equation}
-%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
-%\end{equation}
-%which is close to the definition in \cite{storey.03}.
-
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgements}
-
-Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
-prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} , and includes contributions 
-from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.
-
-This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.
-
-%\section*{References}
-% The bibtex filename
-\bibliography{/Users/stephens/Dropbox/Documents/mainbib}
-
-\section*{Figure Legends}
-
-\clearpage
-
-\section*{Tables}
-
-\section*{Supporting Information Legends}
-
-Supplementary material can be found in {\bf Supplementary Information S1.}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
-\end{document}

### 2015-10-06 sectionDiscussion__W.tex
New Article!

--- 2015-10-06 paper/main.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
-\documentclass[11pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-%\usepackage{natbib}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-\usepackage{booktabs}
-\usepackage{xspace}
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-\usepackage{sectsty}
-\subsectionfont{\normalsize}
-\subsubsectionfont{\mdseries\itshape}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\usepackage{color}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-% Use the PLoS provided bibtex style
-\bibliographystyle{/Users/stephens/Dropbox/Documents/stylefiles/plos2009}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
-\pagestyle{myheadings}
-%% ** EDIT HERE **
-\usepackage{enumerate}
-\usepackage{multirow} 
-\usepackage{url}
-\usepackage{xr} %for cross-referencing
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-\newtheorem{algorithm}{Algorithm}
-\newtheorem{proposition}{Proposition}
-\newtheorem{restateproposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{result}{Result}
-\newtheorem{note}{Note}
-\newtheorem{definition}{Definition}
-
-\def\lfdr{\textit{lfdr}}
-\def\lfsr{\textit{lfsr}}
-\def\x{\mbf{x}}
-\def\y{\mbf{y}}
-\def\e{\mbf{e}}
-\def\g{\mbf{g}}
-\def\p{p}
-\def\Yb{\hat{Y}_1^{\text{BAYES}}}
-\def\Xperp{X_{1 \perp 0}}
-\def\Pperp{P_{1 \perp 0}}
-\def\Pperpb{\Pperp^{\text{B}}}
-\def \hatd { \widehat{ d } }
-\def \dhatin {\hatd^{\mbox{\tiny IN}}}
-\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
-\def \mvn{\text{MVN}}
-\def\mn{\text{MN}}
-\def \iw{\text{W}^{-1}}
-\def \fp {fastPHASE\ }
-\def\bhat{\hat{\beta}}
-\def\shat{\hat{s}}
-\def\var{\rm Var}
-\def\mvr{\text{BMVR}}
-\def\bfuni{\text{BF}_\text{uni}}
-\def\one{\mbf{1}}
-\def\zero{\mbf{0}}
-\def\hall{H_{\text{all}}}
-\def\pall{\p_{\text{all}}}
-\def\bfall{\text{BF}_\text{all}}
-\def\bfav{\text{BF}_\text{av}}
-\def\gall{\gamma_\text{all}}
-\def\limbfall{\bfall^\rightarrow}
-\def\limbfav{\bfav^\rightarrow}
-\def\limbfgamma{\BF_\gamma^\rightarrow}
-\def\sxx{S_{xx}}
-\def\syy{S_{yy}}
-\def\vyy{V_{yy}}
-\def\vyx{V_{yx}}
-\def\vxx{{V_{xx}}}
-\def\rss{\text{\rm RSS}}
-\def\dss{\text{\rm $\Delta$SS}}
-\def\U{U}
-\def\D{D}
-\def\I{I}
-\def\luu{L_{UU}}
-\def\indep{\perp}
-\def\qvalue{{\tt qvalue}\xspace}
-\def\locfdr{{\tt locfdr}\xspace}
-\def\mixfdr{{\tt mixfdr}\xspace}
-\def\ashr{{\tt ashr}\xspace}
-
-\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
-                                % mode too
-\newcommand\height{{\it height}}
-\newcommand\weight{{\it weight}}
-\def\BF{\text{BF}}
-
-%% END MACROS SECTION
-%\externaldocument{plos_one_rev1_suppinfo}
-
-\begin{document}
-
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{False Discovery Rates: A New Deal}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Matthew Stephens$^{1*}$, 
-\\
-\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
-\\
-$\ast$ E-mail: Corresponding mstephens@uchicago.edu
-\end{flushleft}
-
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-%\section*{Author Summary}
-
-\section*{Introduction}
-
-Since its introduction in 1995 by Benjamini and Hochberg 
-\cite{benjamini1995controlling}, the ``False Discovery Rate" (FDR) has quickly established itself
-as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing
-in which the goal is to identify the non-zero ``effects" among a large number of imprecisely-measured effects.
-
-Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
-indeed, the notion that EB approaches could be helpful in handling multiple comparisons predates introduction of
-the FDR  (e.g.~\cite{greenland1991empirical}). More recently, EB approaches to the FDR
-have been extensively studied by several authors, 
-especially B.~Efron and co-authors 
-\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
-see also \cite{kendziorski2003parametric,newton2004detecting, 
-datta2005empirical,muralidharan2010empirical} for example.
-So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
-existing widely-used FDR pipelines) and can substantially affect
-inference. The first idea is to {\it assume that the distribution of effects is unimodal}. This yields a very simple, fast, and stable computer 
-implementation, as well as improving inference of FDR when the unimodal assumption is correct.  
-The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
-to summarize each measurement. This idea allows variations in measurement precision to be better accounted for,
- and avoids a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
-
-In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
-the idea that it may be preferable to focus on estimation rather than on testing.
-In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
-simply estimation with some positive prior probability that the effect is exactly zero.
-However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
-perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
-controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
-``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
-in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
-the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the lfsr, particularly its increased robustness to modelling assumptions.
-
-%terminology
-%``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
-%and correspondingly define the ``false signed discovery rate", or simply ``false sign rate" (FSR), in a natural way.
-
-%The methods we provide here can address both problems, and we provide results that highlight important benefits of focussing on estimation, and particularly on assessing the correctness of the {\it sign} of the estimated effects, rather than on testing for non-zero effects.
-
-Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
-(We address the reasons for this name, and connections with shrinkage analysis, in the Discussion.)
-
-%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.
-
-%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-%measurements are made on only a small number of
-%samples in each condition - sometimes as few as one
-%sample in each condition. Thus the error in estimates
-%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-%in estimates of $\beta_j$ still greater.
-%
-
-%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
-%Here we present new tools that bring two important ideas to bear on the problem.
-%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
-%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
-%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
-%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
-%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
-%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
-%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
-%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
-%a substantial increase in the number of discoveries at a given FDR).
-
-%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
-%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
-Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
-%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
-%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
-%generic methods that work directly with two numbers for each 
-%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
-Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
-%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
-
-\def\df{df}
-\def\FDR{\text{FDR}}
-\def\fdr{\text{\lfdr}}
-\def\FSR{\text{FSR}}
-\def\fsr{\text{lfsr}}
-
- \section*{Methods}
-
- \subsection*{Model Outine}
-
-% Here we consider one setting in which FDR-based methods are widely used:
-%an experimenter has measured a large numbers of ``effects", each with an associated
-%measurement precision (standard error), 
-%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
-%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
-%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
-%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
-%For example, in genomics
-%a common goal is to identify ``differentially-expressed genes" -- that is, genes
-%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
-%mean expression level between the two conditions, often referred to as the ``log-fold change". 
-Here we describe the simplest version of our method, and briefly discuss embellishments we have also implemented.
-
-Suppose that we are interested in the values of $J$ ``effects" $\beta=(\beta_1,\dots,\beta_J)$. For example, in a typical genomics application
-that aims to identify differentially expressed genes,  $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
-In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
-that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
-more general problem of estimating, and assessing uncertainty in, $\beta_j$.
-
-Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
-and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
-which by Bayes theorem can be written as 
-\begin{equation}
-p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
-\end{equation}
-
-For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
-This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
-A simple way to implement the UA is to assume that $g$ is
-a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
-\begin{align} \label{eqn:beta}
-p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
-g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
-\end{align}
-where $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyperparameters, which are non-negative and sum to one, and are to be estimated,
-while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
-represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
-(We encourate the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
-although of course in practice we use a finite grid -- see Implementation Details.)
-
-For the likelihood $p(\bhat | \beta, \shat)$ we assume  
-\begin{equation} \label{eqn:normlik}
-p(\bhat | \beta, \shat) = \prod_j N(\bhat_j; \beta_j, \shat_j^2).
-\end{equation}
-Here, in additional to some conditional independence assumptions, we are 
-effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-
-This simple model features both the key ideas we want to emphasise in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
-the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
-with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
-However, this simple model also has several additional assumptions that can be relaxed. Specifically,
-\begin{enumerate}
-\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) also implies that $g$ is symmetric about 0; more flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see (\ref{eqn:g-general})].
-\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
-We can relax this to allow for a relationship between these quantities [see (\ref{eqn:beta-alpha})].
-\item The likelihood (\ref{eqn:normlik}) effectively assumes that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-We can generalize this likelihood using a $t$ likelihood [see (\ref{eqn:tlik})].
-\end{enumerate}
-These embellishments are detailed in Embellishments and Implementation Details.
-Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
-(which are also made by most existing EB approaches to this problem). Correlations among tests certainly arise in practice, either due to genuine correlations
-in the system of study, or due to unmeasured confounders, and their potential to impact results of an FDR analysis is important to consider
-whatever analysis methods are used: see \cite{efron2007correlation,leek:2007} for relevant discussion.
-
-%We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
-%$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
-%\begin{equation}
-%p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
-%\end{equation}
-%Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
-%\begin{equation} \label{eqn:normlik}
-%p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
-%
-%We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
-%as a mixture of null and alternative components:
-%\begin{equation} \label{eqn:betaj0}
-%g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
-%\end{equation}
-%where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
-%We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
-%\begin{equation} \label{eqn:beta-iid}
-%p(\beta | \shat) = \prod_j g(\beta_j).
-%\end{equation}
-%
-%Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
-%with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
-%to use a mixture of zero-mean normal distributions:
-%\begin{equation} \label{eqn:mixnorm}
-%g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
-%Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
-%represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
-% The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
-% more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below. 
-% 
-% Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
-% yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
-% \begin{equation} \label{eqn:g}
-% g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
-% \end{equation}
-% 
-% Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
-% \begin{align} \label{eqn:betaj}
-% p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
-% p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
-% \end{align}
-% with $g$ given by the mixture of normals $(\ref{eqn:g})$.
-
-\subsection*{Fitting the model}
-
-In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
-each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
-Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
-\begin{equation} \label{eqn:marginallik}
-p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
-\end{equation}
-where we define $\sigma_0:=0$.
-
-The usual EB approach to fitting this model would involve two simple steps:
-\begin{enumerate}
-\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
-\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
-against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | | \bhat, \shat,\hat{\pi})$.
-\end{enumerate}
-Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(The simplicity of the EM algorithm in step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
-which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
-
-Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximise a penalized
-likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
-We introduce this penalty because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
-
-\subsection*{The local False Discovery Rate and local False Sign Rate}
-
-As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
-In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
-One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
-the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
-In other words it is the posterior probability that $\beta_j$ is actually zero:
-\begin{equation} \label{eqn:fdr}
-\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
-\end{equation}
-
-The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
-whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
-the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey1991philosophy}) argued that
-``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$.
-Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
-\begin{quote}
-...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
-\end{quote}
-Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
-focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.
-
-Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
-would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
-\begin{equation} \label{eqn:lfsr}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
-\end{equation}
-To illustrate, suppose that 
-\begin{gather*}
-p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
-p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
-p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
-\end{gather*} 
-Then 
-from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
-our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.
-
-As our notation suggests, $\lfsr_j$ is intended to be compared and contrasted with $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
-small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
-Of coure, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
-this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
-(this follows from the definition because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
-In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
-$\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.
-
-From these ``local" measures of significance, we can also compute average error rates over subsets of observations
-$\Gamma \subset \{1,\dots,J\}$. For example,   
-\begin{equation} \label{eqn:FDRhat}
-\widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
-\end{equation}
-estimates the FDR we would obtain if we were to declare all tests in $\Gamma$ significant.
-And 
-\begin{equation} \label{eqn:qvalue}
-q_j := \widehat \FDR(\{k: \fdr_k \leq \fdr_j\}) 
-\end{equation}
-provides a measure of significance analogous to Storey's $q$ value \cite{storey.03}.
-
-%Because our methods involve explicitly modelling effects $\beta_j$, rather than 
-%$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
-
-%Indeed, even in settings where some
-%$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
-%analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
-
-\subsection*{Related work}
-
-\subsubsection*{Previous approaches focussed on FDR}
-
-Among previous methods that explicitly consider the FDR and related quantities,
-our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
- (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
-  %Here we highlight the key differences between our approach and these three existing approaches.
-
-There are two key differences between our approach and all of these three existing methods. First, 
-whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
-we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
-for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
-Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 
-how it can improve performance by taking better account of variation in measurement precision among observations.
-%(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
-% Or more generally Section xx illustrates the potential practical effects
-
-%allows for different modelling assumptions to be compared through likelihood comparisons.
-Second, our unimodal assumption (UA) that the effects are unimodal about zero is quite different from assumptions made by $\qvalue, \locfdr$ or $\mixfdr$.
-Indeed, $\locfdr$ assumes that all $z$ scores near 0 are null (Efron calls this the Zero Assumption; ZA), which implies that 
-under the alternative hypothesis the distribution of $z$ scores has {\it no mass at 0}; this contrasts strikingly with the UA, which implies that this distribution has its peak at 0! 
-Similarly, $\qvalue$ assumes that all $p$ values near 1 are null,
-which is the same as the ZA because $p$ values near 1 correspond to $z$ scores near 0. And although \mixfdr does not formally make the ZA,
-we have found that in practice, with default settings, the results often approximately satisfy the ZA (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}).
-Thus, not only do these existing methods not make the UA, they actually make assumptions that are, in some sense, as different
-from the UA as they can be.
-
-Given that the UA and ZA are so different, it seems worth discussing why we generally favor the UA. 
-Although the UA will not apply to all situations, we believe that it will often be reasonable, especially in FDR-related contexts
-that have traditionally focussed on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis,  
-it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible. Further, it seems reasonable to expect that larger effects become
-decreasingly plausible, and so the distribution of the effects will be unimodal about 0.
-To paraphrase Tukey, ``All we know about the world teaches us that large effects are rare, whereas small effects abound."
-We emphasise that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
-It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
-
-%Tukey "peri-null"
-
-%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
-%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}.
-
-In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficents - 
-common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
-These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
-it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
-The UA assumption on regression coefficients is directly analagous to our UA here, and so its
-widespread use in the regression context supports its use here.
-
-Alternatively, we could motivate the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
-such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
-make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).
-
-Finally, the UA also has a considerable practical benefit: it yields simple algorithms that are both computationally
-and statistically stable. We illustrate these features in Results.
-
-\subsubsection*{Other work}
-
-There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
- Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
-indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
-This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
-Also, they provide no software implementation.
-More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
-Some pointers to recent papers in which other relevant
-citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}.
-Much of the literature focusses on the homoskedastic case (i.e.
-$\shat_j$ all equal) whereas we allow for heteroskedasticity.
-And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential.
-Several recent papers consider more flexible non-parametric assumptions on
-$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
-estimate (NPMLE) for $g$.
-These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive.
-However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
-may perform well for point estimation, it seems possible it will not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
-Indeed, one way of thinking about the UA is as a way to regularize $g$.
-
-%The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
-%encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-%$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-%$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-%the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-%observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-%corresondingly little effect on the estimated FDR of other observations. 
-%We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-%$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-%that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-%estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-%Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-\def\dt{\text{dt}}
-
-%
-%assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
-%$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
-% $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
-%  \begin{equation} \label{eqn:lik}
-% \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
-% \end{equation}
-% where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e.
-the density of $\mu+\sigma T_\nu$).
-In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
-%  \begin{equation} \label{eqn:normlik}
-% \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
-% \end{equation}
-%which, in light of (\ref{eqn:betahat}), is equivalent to
- % \begin{equation} \label{eqn:lik2}
-% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
- %\end{equation}
-
-%The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
-they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
-The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$.
-Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$.
-If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$.
-See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.
-
-%This likelihood approximation can be motivated by assuming that i) $\bhat$
-% $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
-% $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$. 
-%
-
-%Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.
-
-%While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
- %that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
-%there is an important difference: here,
-
-%The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
-For example, we can write $p(\bhat_j, \shat_j | \beta_j) = p(\bhat_j | \shat_j, \beta_j)p(\shat_j | \beta_j)$.
-Now, if we are willing to assume that $\shat_j$ alone contains no information about $\beta_j$, or equivalently that $p(\shat_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\bhat_j | \shat_j, \beta_j)$, and if $\bhat_j | \shat_j, \beta_j \sim N(\beta_j, \shat_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.
-
-\section*{Results}
-
-We compare results of \ashr with existing FDR-based methods implemented
-in the R packages \qvalue (v2.0 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
-In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
-to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.
-
-\subsection*{Effects of the Unimodal Assumption}
-
-Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
-standard error, $s_j=1$ and all methods are provided that information. That is,
-$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$.
-In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the only difference between our method and methods like $\locfdr$ and $\mixfdr$ are in how they estimate $g$.
-
-To briefly summarize the results in this section:
-\begin{enumerate}
-\item The UA can produce very different inferences compared with the ZA made by existing methods.
-\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
-\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
-\end{enumerate}
-
-\subsubsection*{The UA and ZA can produce different inferences}
-
-\begin{figure}
-\center \includegraphics[height=6in]{figures/decomp_ZA.pdf}
-\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
-The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
-In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
-The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
-(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
-\end{figure}
-
-To illustrate the different inferences from the UA and ZA we show results for a single
-dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
-but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
-\qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
-The results (Figure \ref{fig:ZA}) illustrate the clear difference between the existing methods and our method. The effects of the ZA made by \qvalue and \locfdr
-are visually clear, producing a ``hole"  in the alternative $z$ score distribution around 0. Although \mixfdr does not formally make the ZA, its decomposition
-exhibits a similar hole. In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
-effectively ``filling in" the hole.  (Of course the null distribution also has a peak at 0, and the local fdr 
- under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)
-
-Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA.
-Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
-This would increase the null component (dark blue) at the expense of
-the alternative component (light blue).
-Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0.
-The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears.
-In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
-See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
-demonstration.
-
- \subsubsection*{The UA can produce conservative estimates of $\pi_0$}
-
-The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
-Consequently \ashr will estimate smaller $\lfdr$s and FDRs than existing methods that make the ZA. 
-This is desirable, provided that these estimates remain conservative: that is,
-provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
-The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behaviour. To check its effectiveness
-we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
-for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
-They range from a ``spiky" distribution -- where many non-zero $\beta$ are
-too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
-flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
-making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
-which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.
-
-% How do these differences influence estimates of the FDR? To address this, note that  -- in this simple case where the 
-% $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
-% of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
-% strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
-% although the methods considered here differ from one another in many ways, having estimated $\pi_0$
-% they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
-% empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
-%
-
-%Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
-% and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
-%Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
-%At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
-%overestimates of $\pi_0$, which will yield overestimates of the FDR).
-
-\begin{figure} 
-\begin{center}
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=1.8in]{figures/scenarios_density.pdf} 
-	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/pi0_est.pdf} 
-\caption{Comparison of true and estimated values of $\pi_0$.
-When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ashr being least conservative, and hence most accurate.
-When the UA does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.} \label{fig:pi0sims}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfdr_est.png} 
-\caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
-Black line is $y=x$ and red line is $y=2x$.
-Estimates of $\lfdr$ are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est.png} 
-\caption{As in c), but for $\lfsr$ instead of $\lfdr$.
-Estimates of $\lfsr$ are consistently less conservative than $\lfdr$ when UA holds, and also less anti-conservative in bimodal scenario.}  \label{fig:lfsr}
-\end{subfigure}
-\end{center}
-\caption{Results of simulation studies (constant precision $s_j=1$).}
-\end{figure}
-
-For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
-data as follows:
-\begin{enumerate}
-\item Simulate $\pi_0  \sim U[0,1]$.
-\item For $j=1,\dots,J$, simulate $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)$.
-\item For $j=1,\dots,J$, simulate $\bhat_j | \beta_j \sim N(\beta_j,1)$.
-\end{enumerate}
-%Thus these simulations assume the same
-%precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).
-
-Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ashr ($y$ axis) with the true values ($x$ axis). 
-For \ashr we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
-and as a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
-as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
-The results show that \ashr provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
-in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ashr estimates can be slightly anti-conservative.
-We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
-are used. (In addition, the effects on $\lfsr$ estimates turn out to be relatively modest; see below).
-
-%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
-%
-%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
-%and so all effects are non-zero and no nulls hold. This scenario is rather different
-%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
-%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
-%``mostly non-null"
-%scenario, and we would agree! On the other hand, the distribution of $p$ values
-%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
-%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
-%Thus it seems desirable to have methods that perform well in both cases, and this is
-%an important feature of the methods we introduce here.
-%
-%TODO: put FDRs on graph? Simulate with 50\% null?
-
-\subsubsection*{The $\lfsr$ is more robust than \lfdr}
-
-The results above show that \ashr can improve on existing methods in producing smaller, more accurate, 
-estimates of $\pi_0$, which will lead to more accurate estimates of FDR.
-Nonetheless, in many scenarios \ashr continues
-to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
-This is because these scenarios include 
-an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
-estimation of $\pi_0$ impossible. Put another way, and as is well known, $\pi_0$ is not identifiable:
- the data can effectively provide an upper bound on plausible values of $\pi_0$,
-but not a lower bound (because the data cannot rule out that everything is non-null, but with miniscule effects).
-To obtain conservative behaviour we must estimate $\pi_0$ by this upper bound, which can
-be substantially larger than the true value.
-
-Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
-overestimation of $\pi_0$ is corresponding overestimation of FDR (and $\lfdr$, and $q$ values).
-To illustrate, Figure \ref{fig:lfdr} compares the estimated $\lfdr$ from ash.n with the true value (computed using
-Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios which involve many
-non-zero effects that are very near 0 (e.g.~the spiky scenario with $\pi_0$ small) where
-$\pi_0$ can be grossly overestimated.  
-(Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.)
-
-The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
- become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
- This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
-To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
-be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
-the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in $\lfsr$ than $\lfdr$.
-
-Note that, compared with previous debates regarding testing, 
-this section advances an additional reason for focussing on the sign of the effect, rather than just testing whether it is 0. 
-In previous debates authors have argued against testing whether an effect is 0
-because it is {\it implausible that effects are exactly 0}. Here we add that {\it even if one believes
-that some effects may be exactly zero}, it is still better to focus on the sign, because generally {\it the data are more informative about that question}
-and so inferences are more robust to, say, the inevitable mis-estimation of $\pi_0$.
-To provide some intuition, consider an observation with a $z$ score of 0. The $\lfdr$ of this observation can range from 0 (if $\pi_0=0$)
-to 1 (if $\pi_0=1$). But, assuming a symmetric $g$, the $\lfsr>0.5$ whatever the value of $\pi_0$, because the observation $z=0$ says
-nothing about the sign of the effect.
-Thus, are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
-whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.
-
-Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
-Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
-if we are prepared to assume that no effects are exactly zero, then removing the point mass 
-yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
-However, there is ``no free lunch":  if in fact some effects are exactly zero
-then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
-We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.
-
-\begin{figure}
-\begin{center}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s_nn.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be more accurate in this case.
-That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
-\end{subfigure}
-\end{center}
-\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
-\end{figure}
-
-\subsubsection*{Numerical Stability}
-
-The EM algorithm, which we use here to fit our model, is notorious for convergence to local optima. 
-However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
-caused by such behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
-in addition running it using our default initialization procedure (see Implementation details). We then compared the largest
-log-likelihood achieved across all 11 runs with the log-likelihood achieved by the default run.
-When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference ($< 0.02$), 
-and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference ($<0.02$),
-and 6\% of runs showed an appreciable log-likelihood difference ($>1$), with the largest difference being 5.0. However, perhaps suprisingly, 
-even for this largest difference results from the default run (e.g.
-the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood.
-See \url{http://github.com/stephens999/ash/blob/master/dsc-robust/summarize_dsc_robust.rmd} for further details.
-
-\subsubsection*{The UA helps provide reliable estimates of $g$}
-
-An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
-can estimate the {\it size} of each effect $\beta_j$.
-Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
-to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
-threshold $T$.
-%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
-%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
-%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
-%towards zero, with the strength of the shrinkage being estimated from the data.
-Further, because the posterior distribution is, by definition,
-conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
-been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
-In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
-\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
-how can confidence intervals be assigned to the true
-effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
-and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
-%but also notes the difficulties of implementing a stable general algorithm.
-
-The ability of the EB approach to provide valid ``post-selection" interval estimates is extremely attractive in principle.
-But its usefulness in practice
-depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
-which are notoriously difficult in general. Indeed, Efron emphasises
-the difficulties of implementing a stable general algorithm, noting in his rejoinder
-``the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying
-to be overly nonparametric ... but it is hard to imagine a
-generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
-Our key point here is that the UA greatly simplifies the deconvolution problem.
-While not meeting Efron's desire for an entirely general nonparametric approach, we believe 
-that the UA can handle many cases of practical interest.
-
-To illustrate this, Figure \ref{fig:egcdf} compares the
-estimated $g$ from \ashr  with that from \mixfdr which does not make the UA 
-(and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
-The greater reliability of estimates afforded
-by the UA is immediately apparent. In particular the estimated cdf from \mixfdr often has an almost-vertical
-segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. 
-The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
-While the UA is not the only way to achieve this, we find it an attractive, simple and effective approach.
-
-Interestingly, even in the ``bimodal" scenario $\ashr$ 
-is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
-Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
-Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
-to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
-(\url{http://github.com/stephens999/ash/blob/master/dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
-clearly visible in the observed $\bhat$.
-
-We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ashr tends to systematically
-overestimate the mass of $g$ near zero. On careful inspection, this is apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
-and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf}b),
-and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
-under-estimating $\pi_0$ (Figure \ref{fig:egcdf}c).
-
-\begin{figure}[h!] 
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=2in]{figures/egcdf.pdf}
-	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf_nopen.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
-Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
-\end{subfigure}
-\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.} \label{fig:egcdf}
-\end{figure}
-
-%making the UA provides an attractive and simple way to achieve estimating $g$  large
-%subset of cases of interest (those where the UA can reasonably be assumed to hold).
-
-%In practice, the utility of the posterior distributions noted above depends primarily on the accuracy
-%of the estimate of the distribution $g(\cdot)$.
-
- %quote from efron rejoinder , p46:
-%MorrisÕ Section 3 is especially pertinent. His formula
-%for p(?|z) is related to my discussion of the
-%BenjaminiÐYekutieli False Coverage Rate method in
-%Section 7, particularly (7.2)Ð(7.4). Originally I had
-%hoped to develop an empirical Bayes method for estimating
-%such models, but the effort foundered on practical
-%difficulties involving the perils of deconvolution.
-%Section 6 on the Òone-group modelÓ is the ugly duckling
-%of the current paper, but it bears on some important
-%points raised in the discussion. Figure 7 concerns
-%a fuzzy version of simultaneous hypothesis testing,
-%where, as in MorrisÕ hospital example, the usual singlepoint
-%null hypotheses seem unequal to the task. The
-%development from (6.6) onwards, particularly (6.12),
-%bears on the possibility of nonnormal null distributions,
-%and is about as far as I can go in answering Professors
-%Rice and SpiegelhalterÕs penultimate question.
-%With g(?) a normal distribution, model (6.1) returns
-%us to the realm of Stein estimation, the scene of my
-%happy collaborations with Carl Morris. I continue to
-%be surprised at how much harder bumpy, nonnormal
-%models like (7.1) are to deal with. JamesÐStein estimation
-%works fine with, say, N = 10 component problems,
-%while the RobbinsÕ form of empirical Bayes appropriate
-%to (7.1) seems to require hundreds or thousands.
-%The information calculations in Efron (2008)
-%reinforce this gloomy assessment. Maybe I am trying
-%to be overly nonparametric in constructing the empirical
-%Bayes Fdr estimates, but it is hard to imagine a
-%generally satisfactory parametric formulation for (6.1).
-%Or perhaps it is just that hypothesis testing is more demanding
-%than estimation.
-%My Note: (6.1) is the general model (mu \sim g()).
-%7.1 is the case where the effects are a mixture of zero and N(-3,1).
-
-\subsubsection*{Calibration of posterior intervals}
-
-To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over 100 simulations in each Scenario). 
-Specifically we examine the empirical coverage of nominal lower 95\% credible bounds for a) all observations; b) significant negative discoveries; c) significant positive discoveries.
-We examine only lower bounds because the results for upper bounds follow by
-symmetry (except for the one asymmetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
-in each case: for negative discoveries the lower bound is typically large and negative and limits how big (in absolute value) 
-the effect could be; for positive discoveries the lower bound 
-is positive, and limits how small (in absolute value) the effect could be. Intuitively, the lower bound for negative discoveries depends on the accuracy of $g$ in its tail,
-whereas for positive discoveries it is more dependent on the accuracy of $g$ in the center.
-
-The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
-as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
-
-% As noted above, the penalty term (\ref{eqn:penalty}) causes systematic over-estimation of the mass of $g$ near 0,  which one would expect to overshrink interval estimates towards zero. 
-%The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound is negative and overshrinking (moving it too far towards zero) makes the left tail too big; conversely, for positive discoveries, the lower confidence bound is positive, and
-%overshrinking (moving it too far towards zero) makes the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
-%and be prepared to put up with some overshrinkage to avoid it.
-
-%Coverage rates are given in:
-%\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
-
-%Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all.tex}
-	\caption{All observations.
-Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario.
-This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero.
-Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (Table \ref{tab:nopen}).
-Removing the penalty in the half-uniform case is not recommended (see Appendix).}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg.tex}
-\caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
-and the normal-based method in the flat-top scenario.
-These results likely reflect inaccurate estimates of the tails of $g$ due to a disconnect between the tail of $g$ and the component distributions in these cases.
-For example, the uniform methods sometimes substantially underestimate the length of the tail of $g$ in these long-tailed scenarios,
-causing over-shrinkage of the tail toward 0.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos.tex}
-\caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds} \label{tab:coverage}
-\end{table}
-
-%
-%\subsubsection*{Reverse asymptotics}
-%
-%Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
-%
-%Note: we'd like to make sure it behaves ok in the case where most observations are null.
-%Perhaps simulate with 999 null and 1 alternative?
-%Also maybe look at performance for small sample sizes - e.g. n=1?
-%
-%Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
-%can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
-%by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
-%
-
-\subsection*{Differing measurement precision across units}
-
- We turn now to the second important component of our work: allowing for varying
- measurement precision across units. The key to this is the use of a likelihood,
- (\ref{eqn:normlik}) or (\ref{eqn:tlik}), that explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
-
-  To illustrate, we conduct a simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
- ($s_j=10$).  In both cases,
-we assume that half the effects are null
-and the other half are normally distributed with standard deviation 1:
-\begin{equation}
-p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1).
-\end{equation}
-In this setting, the poor-precision measurements $(s_j=10)$ tell us very little, and any sane analysis should effectively ignore them. 
- However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor}). This is because the poor measurements 
- produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
- which, when combined with the good-precision measurements, dilute the overall signal (e.g. they reduce the density of $p$ values near 0).
- This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
-the estimated error rates ($q$-values, or $\lfdr$ values) for the good-precision observations increase when the low-precision observations are included in the analysis
-(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
-for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).
-
-\begin{figure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_hist.pdf} 
-    \label{fig:goodpoor-hist}
-    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations}
- \end{subfigure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_scatter.pdf}
-    \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) of a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
-\end{subfigure}
-\caption{Simulation illustrating how, for existing FDR methods, 
-poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
-$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
-a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
-In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
-\end{figure}
-
- \subsubsection*{Reordering of significance, and the ``$p$ value prior"}
-
-Another consequence of accounting for differences in measurement precision is that \ashr may re-order the significance of the observations compared
- with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
- We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measuments that have the same $p$ value.
- The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
- and so the $\lfsr$ for these poor-precision measurements is always high.
-
-The potential for Bayesian analyses to re-order the significance of observations, and specifically to downweight imprecise observations,
-was previously discussed in \cite{guan.stephens.08}.
-However, Wakefield \cite{wakefield:2009} showed that, under a certain prior assumption, the Bayesian analysis produces the same ranking of significance as $p$ values (or their $z$ scores).
-He named this prior the ``$p$-value prior" because it can be thought of as the
-implicit prior assumption that is being made if we rank the significance of observations by their $p$ value.
-Wakefield's $p$-value prior assumes that the less precise effect estimates correspond to larger true effects, and specifically that they scale proportional
-to the standard errors $s_j$. More specifically still, it assumes a normal prior for the non-zero $\beta_j$ with mean 0 and variance $K s^2_j$ for some constant $K$.
-Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. Specifically, if, instead of assuming that $\beta_j$ is independent
-of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
-then the $\lfsr$ computed by ash.n provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
-as is illustrated in Figure \ref{fig:lfsr_pval}, right panel. (This result does not hold when using the mixtures of uniforms prior, ash.u.)
-
-This $p$-value prior assumes that the $z$ scores $z_j=\beta_j/s_j$
-are identically distributed, independent of $s_j$. This is essentially the assumption made, implicitly or explicitly, by existing methods -- like \locfdr, \mixfdr and \qvalue --
-that model the $z_j$ or $p_j$ directly.  In contrast, we have assumed up to now that the $\beta_j$ are independent of $s_j$. 
-We can set both these assumptions within a more general framework, which allows that $\beta_j/s^\alpha$ is independent of $s_j$ for some $\alpha$ [Equation (\ref{eqn:beta-alpha}))].
-Setting $\alpha=0$ implies that $\beta_j$ is
-independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
-effects (in absolute value). This latter assumption may often be qualitatively plausible: 
-for example, in gene expression studies the standard error for gene $j$ depends partly on
-the variance of its expression among samples, and genes with a larger variance may tend to be less tightly regulated and
-so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
-On the other hand, there is no particular reason to expect that either $\alpha=1$ or $\alpha=0$ will be the optimal choice.
-Indeed, optimal choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$,
-which will be dataset-specific. 
-Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
-comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
-likelihood-based methods can be used to select $\alpha$. For example, following the logic of the EB approach it
-would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
-which has been implemented in our software by C. Dai.
-
-%\begin{result}
-%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
-%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
-%\end{result}
-
-%To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
-%Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
-%In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
-%Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
-%which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
-%(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
-%may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)
-
-%\begin{result}
-%Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and using a mixture of normals for the prior distribution 
-%(\ref{eqn:beta-alpha}) with $\alpha=1$, 
-%the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$.
-Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$)
-%is montonic with $|t_j|$. 
-%\end{result}
-
-%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
-%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.
-
-%Most previous authors of EB approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
-%have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
-%In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
-%This is analogous
-%to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
-%standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
-%that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.
-
-\begin{figure}
-\centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
-\caption{Figure illustrating affects of prior assumptions on re-ordering of significance.
-Left panel shows results under our ``default prior" which assumes that effects $\beta_j$ are identically distributed, independent of $s_j$.
-Right panel shows results under the ``$p$-value prior", which assumes that $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$.} \label{fig:lfsr_pval}
-\end{figure}
-
-%\subsubsection*{Adaptive Shrinkage}
-%
-%Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
-%when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
-%the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
-% $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
-% In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
-% Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
-% and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
-%
-%Our methods we build on this idea in two ways. First,
-%by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
-%Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
-%adapts to measurement precision: more precise measurements undergo
-% less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
-%than observations with
-%small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
-%To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
-%Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
-%
-%\begin{figure}[!ht]
-%\begin{center}
-%\includegraphics[width=400px]{../examples/ash.pdf} 
-%\end{center}
-%\caption{Figure illustrating adaptive shrinkage.
-Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0.
-The standard error of each observation, $s_j$, was simulated from Inverse Gamma(5,5), which resulted in $s_j$ varying from 0.34 to 6.7.
-The shrunken estimates ($y$ axis) are plotted against the observed estimates ($x$ axis), with color indicating the (square root of the) precision of each measurement (blue=lower-precision; red=higher-precision).
-The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
-%\end{figure}
-
-%DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
-%To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
-Because the prior distribution is unimodal about zero,
-%the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
-%amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
-%For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
-%than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
-%small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
-
-%previous work based on normal means:
-%http://link.springer.com/article/10.1007/BF02562623
-
-%\subsection{Null behaviour}
-
-%Results of simulations under the null.
-
-%\subsection{Comparison of different underlying component distributions}
-%
-%Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
-%than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
-%likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
-%For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
-%increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
-%expect $L_{hu}$ to be significantly larger than the others.
-%
-%Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
-%
-%\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
-%\begin{center}
-%\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
-%%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
-%\end{center}
-%\caption{}
-%\end{figure}
-
-%Notes from Adrian Raftery - 
-%Newcombe (1886)
-%Dean and Raftery
-%Uniform is super-efficient
-
 \section*{Discussion}
@@
 they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alterntively, our methods
-could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:lik}) using
+could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:normlik}) using
 $\bhat_j | \shat_j \sim N(0,\lambda_1 \shat_j + \lambda_2)$, where $\lambda_1,\lambda_2$ are to be estimated. However,
@@
 encourate development of methods that exploit this idea in other contexts.
-
-%This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
-%from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
-%with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
-
-%
-%
-%Both \locfdr and \mixfdr allow for deviations from
-%the theoretical null distribution by allowing that, under the null, the $z$ scores are $N(0,\sigma^2)$ with $\sigma$ estimated from the data rather
-%than fixed to 1. Within our approach this idea could be incorporated by changing the likelihood - for example, to assume $\bhat_j \sim N(0,s_ 
-%
-%These are tricky issues that, as far as we are aware, can affect
-%all approaches to FDR analysis. An important focus of Efron's work, also implemented in \locfdr, 
-% is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
-%theoretical null distribution. 
-%
-%For example, we and others have worked on Bayesian approaches to genetic association testing, using subjective prior specification of
-%effect sizes based on external data (e.g.~\cite{stephens:2009}). However, the large numbers of tests brings the opportunity to learn about
-%priors, and here we aim to exploit this
-
-%Choice of mixture components. Use of likelihood to compare different mixture components?
-
-%  
-%Directly modeling the p values, or z scores, say via non-parametric methods, 
-%can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-%distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-%The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-%Before proceeding, we feel obliged to correct the common misconception
-% that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
-% not depend on the number of tests. 
-%  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
-% 
-% In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
-
-\section*{Detailed Methods} \label{sec:implement}
-
-\subsection*{Embellishments}
-
-\subsubsection*{More flexible unimodal distributions}
-
-Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
-Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
- Therefore, we have implemented a more general approach based on 
-\begin{equation} \label{eqn:g-general}
-g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) ,
-\end{equation}
-where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
-\begin{enumerate}[(i)]
-\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$, \qquad (``ash.n")
-\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,  \qquad (``ash.u")
-\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,  \qquad (``ash.hu")
-\end{enumerate}
-where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
-(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
-and $\pi$ is a $2K+1$ vector that sums to 1.)
-The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
-Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
-under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
-(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
-(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
-however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)
-
-Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
-using a large dense grid of $\sigma^2_k$ or $a_k$, (i)-(iii) can respectively approximate,
-with arbitrary accuracy,
-\begin{enumerate}[(i)]
-\item any scale mixture of normals, which includes as special cases
-the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
-\item any symmetric unimodal distribution about 0.
-\item any unimodal distribution about 0.
-\end{enumerate}
-The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
-In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.
-
-Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
-The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
-See Implementation for details of our software defaults.
-
-%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting.
-For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
-%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
-
-\subsubsection*{Dependence of effects on standard errors}
-
-Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
-This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
- \begin{equation} \label{eqn:beta-alpha}
- \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
- \end{equation}
-for any $\alpha$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
-assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
-it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
-the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.
-
-The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
-To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
- standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
- Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
-yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
-%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
-%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
-%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
-
-%Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
-%how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
-%it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
-%rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
-%and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
-%this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
-%do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
-
-\subsubsection*{Replace normal likelihood with $t$ likelihood}
-
-We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
-  \begin{equation} \label{eqn:tlik}
- \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
- \end{equation}
-where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
-and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
-The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
-This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms; see Implementation below.
-When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
-
-Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
-\begin{equation} \label{eqn:tdist}
-(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
-\end{equation}
-However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
-In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
-currently pursuing this improved approach (joint work with M.Lu) and results will be published elsewhere. 
-%We note here that
-%modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
-%where our simpler approach can produce unsatisfactory results.
-
-\subsubsection*{Non-zero mode}
-
-An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood, rather than fixed to 0.
-
-\subsection*{Implementation Details}
-
-\subsubsection*{Likelihood for $\pi$}
-
-We define the likelihood for $\pi$ to be the probability of the observed data $\bhat$ conditional on $\shat$: $L(\pi) := p(\bhat | \shat,\pi),$ which
-by our conditional indendence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$.
-[One might prefer to define the likelihood as $p(\bhat, \shat | \pi) = p(\bhat | \shat, \pi) p(\shat | \pi)$, in which case our definition comes down to assuming that the term $p(\shat | \pi)$ does not depend on $\pi$.]
-
-Using the prior $\beta_j \sim \sum_{k=0}^K \pi_k f_k(\beta_j)$ given by (\ref{eqn:g-general}), and the normal likelihood (\ref{eqn:normlik}), integrating over $\beta_j$ yields
-\begin{equation}
-p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
-\end{equation}
-where 
-\begin{equation}
-\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
-\end{equation}
-denotes the convolution of $f_k$ with a normal density.
-These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
-Specifically, 
-\begin{equation} \label{eqn:uconv}
-\tilde{f}_k(\bhat_j)  = 
-\begin{cases}
-N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
-\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
-\end{cases}
-\end{equation}
-where $\Psi$ denotes the cumulative distribution function (c.d.f.) of the standard normal distribution.
-If we replace the normal likelihood with the $t_\nu$ likelihood (\ref{eqn:tlik}) then the convolution 
-for $f_k$ uniform the convolution is still given by (\ref{eqn:uconv}) but with $\Psi$ the c.d.f. 
-of the $t_\nu$ distribution function. (The convolution for $f_k$ normal is tricky and we have not implemented it.)
-
-\subsubsection*{Penalty term on $\pi$}
-
-To make $\lfdr$ and $\lfsr$ estimates from our method ``conservative" we add a penalty 
-term $log(h(\pi;\lambda))$ to the log-likelihood $\log L(\pi)$ to encourage over-estimation of $\pi_0$:
-\begin{equation} \label{eqn:penalty}
-h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
-\end{equation}
-where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
-consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0sims}).
-
-Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
-we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
-
-\subsubsection*{Problems with removing the penalty term in the half-uniform case}
-
-It is straightforward to remove the penalty term by setting $\lambda_k=1$ in (\ref{eqn:penalty}).
-We note here an unanticipated problem we came across when using no penalty term in the half-uniform case 
-(i.e.~ $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$ in (\ref{eqn:g-general})): when the data
-are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
-sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
-for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
-and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
-
-\subsubsection*{EM algorithm}
-
-With this in place, the penalized log-likelihood for $\pi$ is given by:
-\begin{equation}
-\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
-\end{equation}
-where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This can be maximized using an EM algorithm \cite{dempster77}, whose
-one-step updates are:
-\begin{align} \label{eqn:w}
-w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
-n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
-\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}
-\end{align}
-Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
-and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
-probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$. (The $w_{kj}$ are sometimes referred to as the ``responsibilities".)
-Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
-We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple} to accelerate convergence of this EM algorithm.
-
-%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
-\subsubsection*{Initialization}
-
-By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
-Our rationale for initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
-but weak signal in the data cannot quickly draw the algorithm towards the null.
-
-In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
-$\pi_0=1; \pi_k=0 (k>0)$. If not then we replace the EM solution with the global null solution. This aims to guard against errors due to convergence to a local optimum when the
-data are consistent with the global null.
-
-\subsubsection*{Conditional distributions}
-
-Given $\hat\pi$, we compute the conditional distributions 
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
-\end{equation} 
-Each posterior is a mixture on $K+1$ components:
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
-\end{equation}
-where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
-and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
-$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
-All these posterior distributions are easily available.
-For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
-If $f_k$ is normal and $L$ is normal, then this is a normal distribution.
-
-\subsubsection*{Choice of grid for $\sigma_k, a_k$} \label{sec:grid}
-
-\def\sigmamax{\sigma_{\text{max}}}
-\def\sigmamin{\sigma_{\text{min}}}
-When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
-in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
-as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
-based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
-is negative we set $\sigmamax = 8\sigmamin$.)
-
-When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.
-
-Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
-a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.
-
-\appendix
-
-\begin{table}[!ht]
-\centering\begin{tabular}{c c } \toprule
-Scenario & Alternative distribution, $g_1$  \\ \midrule
-spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
-near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
-flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
- &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
-skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
-big-normal & $N(0,4^2)$ \\ 
-bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
-\end{tabular}
-\caption{Summary of simulation scenarios considered} \label{table:scenarios}
-\end{table}
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all_nopen.tex}
-	\caption{All observations}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg_nopen.tex}
-\caption{``Significant" negative discoveries.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos_nopen.tex}
-\caption{``Significant" positive discoveries.}
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{table:nopen}
-\end{table}
-%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
-%\begin{equation}
-%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
-%\end{equation}
-%which is close to the definition in \cite{storey.03}.
-
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgements}
-
-Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
-prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} , and includes contributions 
-from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.
-
-This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.
-
-%\section*{References}
-% The bibtex filename
-\bibliography{/Users/stephens/Dropbox/Documents/mainbib}
-
-\section*{Figure Legends}
-
-\clearpage
-
-\section*{Tables}
-
-\section*{Supporting Information Legends}
-
-Supplementary material can be found in {\bf Supplementary Information S1.}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
-\end{document}

### 2015-10-06 sectionIntroduction_.tex
New Article!

--- 2015-10-06 paper/main.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
-\documentclass[11pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-%\usepackage{natbib}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-\usepackage{booktabs}
-\usepackage{xspace}
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-\usepackage{sectsty}
-\subsectionfont{\normalsize}
-\subsubsectionfont{\mdseries\itshape}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\usepackage{color}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-% Use the PLoS provided bibtex style
-\bibliographystyle{/Users/stephens/Dropbox/Documents/stylefiles/plos2009}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
-\pagestyle{myheadings}
-%% ** EDIT HERE **
-\usepackage{enumerate}
-\usepackage{multirow} 
-\usepackage{url}
-\usepackage{xr} %for cross-referencing
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-\newtheorem{algorithm}{Algorithm}
-\newtheorem{proposition}{Proposition}
-\newtheorem{restateproposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{result}{Result}
-\newtheorem{note}{Note}
-\newtheorem{definition}{Definition}
-
-\def\lfdr{\textit{lfdr}}
-\def\lfsr{\textit{lfsr}}
-\def\x{\mbf{x}}
-\def\y{\mbf{y}}
-\def\e{\mbf{e}}
-\def\g{\mbf{g}}
-\def\p{p}
-\def\Yb{\hat{Y}_1^{\text{BAYES}}}
-\def\Xperp{X_{1 \perp 0}}
-\def\Pperp{P_{1 \perp 0}}
-\def\Pperpb{\Pperp^{\text{B}}}
-\def \hatd { \widehat{ d } }
-\def \dhatin {\hatd^{\mbox{\tiny IN}}}
-\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
-\def \mvn{\text{MVN}}
-\def\mn{\text{MN}}
-\def \iw{\text{W}^{-1}}
-\def \fp {fastPHASE\ }
-\def\bhat{\hat{\beta}}
-\def\shat{\hat{s}}
-\def\var{\rm Var}
-\def\mvr{\text{BMVR}}
-\def\bfuni{\text{BF}_\text{uni}}
-\def\one{\mbf{1}}
-\def\zero{\mbf{0}}
-\def\hall{H_{\text{all}}}
-\def\pall{\p_{\text{all}}}
-\def\bfall{\text{BF}_\text{all}}
-\def\bfav{\text{BF}_\text{av}}
-\def\gall{\gamma_\text{all}}
-\def\limbfall{\bfall^\rightarrow}
-\def\limbfav{\bfav^\rightarrow}
-\def\limbfgamma{\BF_\gamma^\rightarrow}
-\def\sxx{S_{xx}}
-\def\syy{S_{yy}}
-\def\vyy{V_{yy}}
-\def\vyx{V_{yx}}
-\def\vxx{{V_{xx}}}
-\def\rss{\text{\rm RSS}}
-\def\dss{\text{\rm $\Delta$SS}}
-\def\U{U}
-\def\D{D}
-\def\I{I}
-\def\luu{L_{UU}}
-\def\indep{\perp}
-\def\qvalue{{\tt qvalue}\xspace}
-\def\locfdr{{\tt locfdr}\xspace}
-\def\mixfdr{{\tt mixfdr}\xspace}
-\def\ashr{{\tt ashr}\xspace}
-
-\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
-                                % mode too
-\newcommand\height{{\it height}}
-\newcommand\weight{{\it weight}}
-\def\BF{\text{BF}}
-
-%% END MACROS SECTION
-%\externaldocument{plos_one_rev1_suppinfo}
-
-\begin{document}
-
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{False Discovery Rates: A New Deal}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Matthew Stephens$^{1*}$, 
-\\
-\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
-\\
-$\ast$ E-mail: Corresponding mstephens@uchicago.edu
-\end{flushleft}
-
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-%\section*{Author Summary}
-
 \section*{Introduction}
@@
 
-%terminology
-%``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
-%and correspondingly define the ``false signed discovery rate", or simply ``false sign rate" (FSR), in a natural way.
-
-%The methods we provide here can address both problems, and we provide results that highlight important benefits of focussing on estimation, and particularly on assessing the correctness of the {\it sign} of the estimated effects, rather than on testing for non-zero effects.
-
 Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
 (We address the reasons for this name, and connections with shrinkage analysis, in the Discussion.)
-
-%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.
-
-%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-%measurements are made on only a small number of
-%samples in each condition - sometimes as few as one
-%sample in each condition. Thus the error in estimates
-%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-%in estimates of $\beta_j$ still greater.
-%
-
-%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
-%Here we present new tools that bring two important ideas to bear on the problem.
-%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
-%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
-%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
-%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
-%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
-%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
-%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
-%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
-%a substantial increase in the number of discoveries at a given FDR).
-
-%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
-%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
-Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
-%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
-%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
-%generic methods that work directly with two numbers for each 
-%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
-Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
-%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
-
-\def\df{df}
-\def\FDR{\text{FDR}}
-\def\fdr{\text{\lfdr}}
-\def\FSR{\text{FSR}}
-\def\fsr{\text{lfsr}}
-
- \section*{Methods}
-
- \subsection*{Model Outine}
-
-% Here we consider one setting in which FDR-based methods are widely used:
-%an experimenter has measured a large numbers of ``effects", each with an associated
-%measurement precision (standard error), 
-%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
-%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
-%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
-%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
-%For example, in genomics
-%a common goal is to identify ``differentially-expressed genes" -- that is, genes
-%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
-%mean expression level between the two conditions, often referred to as the ``log-fold change". 
-Here we describe the simplest version of our method, and briefly discuss embellishments we have also implemented.
-
-Suppose that we are interested in the values of $J$ ``effects" $\beta=(\beta_1,\dots,\beta_J)$. For example, in a typical genomics application
-that aims to identify differentially expressed genes,  $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
-In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
-that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
-more general problem of estimating, and assessing uncertainty in, $\beta_j$.
-
-Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
-and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
-which by Bayes theorem can be written as 
-\begin{equation}
-p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
-\end{equation}
-
-For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
-This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
-A simple way to implement the UA is to assume that $g$ is
-a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
-\begin{align} \label{eqn:beta}
-p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
-g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
-\end{align}
-where $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyperparameters, which are non-negative and sum to one, and are to be estimated,
-while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
-represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
-(We encourate the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
-although of course in practice we use a finite grid -- see Implementation Details.)
-
-For the likelihood $p(\bhat | \beta, \shat)$ we assume  
-\begin{equation} \label{eqn:normlik}
-p(\bhat | \beta, \shat) = \prod_j N(\bhat_j; \beta_j, \shat_j^2).
-\end{equation}
-Here, in additional to some conditional independence assumptions, we are 
-effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-
-This simple model features both the key ideas we want to emphasise in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
-the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
-with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
-However, this simple model also has several additional assumptions that can be relaxed. Specifically,
-\begin{enumerate}
-\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) also implies that $g$ is symmetric about 0; more flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see (\ref{eqn:g-general})].
-\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
-We can relax this to allow for a relationship between these quantities [see (\ref{eqn:beta-alpha})].
-\item The likelihood (\ref{eqn:normlik}) effectively assumes that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-We can generalize this likelihood using a $t$ likelihood [see (\ref{eqn:tlik})].
-\end{enumerate}
-These embellishments are detailed in Embellishments and Implementation Details.
-Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
-(which are also made by most existing EB approaches to this problem). Correlations among tests certainly arise in practice, either due to genuine correlations
-in the system of study, or due to unmeasured confounders, and their potential to impact results of an FDR analysis is important to consider
-whatever analysis methods are used: see \cite{efron2007correlation,leek:2007} for relevant discussion.
-
-%We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
-%$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
-%\begin{equation}
-%p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
-%\end{equation}
-%Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
-%\begin{equation} \label{eqn:normlik}
-%p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
-%
-%We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
-%as a mixture of null and alternative components:
-%\begin{equation} \label{eqn:betaj0}
-%g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
-%\end{equation}
-%where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
-%We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
-%\begin{equation} \label{eqn:beta-iid}
-%p(\beta | \shat) = \prod_j g(\beta_j).
-%\end{equation}
-%
-%Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
-%with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
-%to use a mixture of zero-mean normal distributions:
-%\begin{equation} \label{eqn:mixnorm}
-%g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
-%Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
-%represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
-% The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
-% more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below. 
-% 
-% Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
-% yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
-% \begin{equation} \label{eqn:g}
-% g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
-% \end{equation}
-% 
-% Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
-% \begin{align} \label{eqn:betaj}
-% p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
-% p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
-% \end{align}
-% with $g$ given by the mixture of normals $(\ref{eqn:g})$.
-
-\subsection*{Fitting the model}
-
-In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
-each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
-Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
-\begin{equation} \label{eqn:marginallik}
-p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
-\end{equation}
-where we define $\sigma_0:=0$.
-
-The usual EB approach to fitting this model would involve two simple steps:
-\begin{enumerate}
-\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
-\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
-against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | | \bhat, \shat,\hat{\pi})$.
-\end{enumerate}
-Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(The simplicity of the EM algorithm in step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
-which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
-
-Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximise a penalized
-likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
-We introduce this penalty because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
-
-\subsection*{The local False Discovery Rate and local False Sign Rate}
-
-As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
-In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
-One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
-the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
-In other words it is the posterior probability that $\beta_j$ is actually zero:
-\begin{equation} \label{eqn:fdr}
-\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
-\end{equation}
-
-The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
-whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
-the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey1991philosophy}) argued that
-``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$.
-Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
-\begin{quote}
-...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
-\end{quote}
-Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
-focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.
-
-Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
-would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
-\begin{equation} \label{eqn:lfsr}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
-\end{equation}
-To illustrate, suppose that 
-\begin{gather*}
-p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
-p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
-p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
-\end{gather*} 
-Then 
-from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
-our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.
-
-As our notation suggests, $\lfsr_j$ is intended to be compared and contrasted with $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
-small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
-Of coure, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
-this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
-(this follows from the definition because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
-In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
-$\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.
-
-From these ``local" measures of significance, we can also compute average error rates over subsets of observations
-$\Gamma \subset \{1,\dots,J\}$. For example,   
-\begin{equation} \label{eqn:FDRhat}
-\widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
-\end{equation}
-estimates the FDR we would obtain if we were to declare all tests in $\Gamma$ significant.
-And 
-\begin{equation} \label{eqn:qvalue}
-q_j := \widehat \FDR(\{k: \fdr_k \leq \fdr_j\}) 
-\end{equation}
-provides a measure of significance analogous to Storey's $q$ value \cite{storey.03}.
-
-%Because our methods involve explicitly modelling effects $\beta_j$, rather than 
-%$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
-
-%Indeed, even in settings where some
-%$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
-%analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
-
-\subsection*{Related work}
-
-\subsubsection*{Previous approaches focussed on FDR}
-
-Among previous methods that explicitly consider the FDR and related quantities,
-our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
- (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
-  %Here we highlight the key differences between our approach and these three existing approaches.
-
-There are two key differences between our approach and all of these three existing methods. First, 
-whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
-we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
-for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
-Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 
-how it can improve performance by taking better account of variation in measurement precision among observations.
-%(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
-% Or more generally Section xx illustrates the potential practical effects
-
-%allows for different modelling assumptions to be compared through likelihood comparisons.
-Second, our unimodal assumption (UA) that the effects are unimodal about zero is quite different from assumptions made by $\qvalue, \locfdr$ or $\mixfdr$.
-Indeed, $\locfdr$ assumes that all $z$ scores near 0 are null (Efron calls this the Zero Assumption; ZA), which implies that 
-under the alternative hypothesis the distribution of $z$ scores has {\it no mass at 0}; this contrasts strikingly with the UA, which implies that this distribution has its peak at 0! 
-Similarly, $\qvalue$ assumes that all $p$ values near 1 are null,
-which is the same as the ZA because $p$ values near 1 correspond to $z$ scores near 0. And although \mixfdr does not formally make the ZA,
-we have found that in practice, with default settings, the results often approximately satisfy the ZA (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}).
-Thus, not only do these existing methods not make the UA, they actually make assumptions that are, in some sense, as different
-from the UA as they can be.
-
-Given that the UA and ZA are so different, it seems worth discussing why we generally favor the UA. 
-Although the UA will not apply to all situations, we believe that it will often be reasonable, especially in FDR-related contexts
-that have traditionally focussed on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis,  
-it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible. Further, it seems reasonable to expect that larger effects become
-decreasingly plausible, and so the distribution of the effects will be unimodal about 0.
-To paraphrase Tukey, ``All we know about the world teaches us that large effects are rare, whereas small effects abound."
-We emphasise that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
-It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
-
-%Tukey "peri-null"
-
-%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
-%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}.
-
-In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficents - 
-common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
-These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
-it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
-The UA assumption on regression coefficients is directly analagous to our UA here, and so its
-widespread use in the regression context supports its use here.
-
-Alternatively, we could motivate the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
-such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
-make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).
-
-Finally, the UA also has a considerable practical benefit: it yields simple algorithms that are both computationally
-and statistically stable. We illustrate these features in Results.
-
-\subsubsection*{Other work}
-
-There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
- Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
-indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
-This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
-Also, they provide no software implementation.
-More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
-Some pointers to recent papers in which other relevant
-citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}.
-Much of the literature focusses on the homoskedastic case (i.e.
-$\shat_j$ all equal) whereas we allow for heteroskedasticity.
-And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential.
-Several recent papers consider more flexible non-parametric assumptions on
-$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
-estimate (NPMLE) for $g$.
-These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive.
-However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
-may perform well for point estimation, it seems possible it will not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
-Indeed, one way of thinking about the UA is as a way to regularize $g$.
-
-%The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
-%encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-%$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-%$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-%the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-%observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-%corresondingly little effect on the estimated FDR of other observations. 
-%We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-%$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-%that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-%estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-%Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-\def\dt{\text{dt}}
-
-%
-%assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
-%$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
-% $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
-%  \begin{equation} \label{eqn:lik}
-% \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
-% \end{equation}
-% where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e.
-the density of $\mu+\sigma T_\nu$).
-In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
-%  \begin{equation} \label{eqn:normlik}
-% \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
-% \end{equation}
-%which, in light of (\ref{eqn:betahat}), is equivalent to
- % \begin{equation} \label{eqn:lik2}
-% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
- %\end{equation}
-
-%The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
-they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
-The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$.
-Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$.
-If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$.
-See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.
-
-%This likelihood approximation can be motivated by assuming that i) $\bhat$
-% $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
-% $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$. 
-%
-
-%Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.
-
-%While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
- %that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
-%there is an important difference: here,
-
-%The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
-For example, we can write $p(\bhat_j, \shat_j | \beta_j) = p(\bhat_j | \shat_j, \beta_j)p(\shat_j | \beta_j)$.
-Now, if we are willing to assume that $\shat_j$ alone contains no information about $\beta_j$, or equivalently that $p(\shat_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\bhat_j | \shat_j, \beta_j)$, and if $\bhat_j | \shat_j, \beta_j \sim N(\beta_j, \shat_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.
-
-\section*{Results}
-
-We compare results of \ashr with existing FDR-based methods implemented
-in the R packages \qvalue (v2.0 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
-In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
-to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.
-
-\subsection*{Effects of the Unimodal Assumption}
-
-Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
-standard error, $s_j=1$ and all methods are provided that information. That is,
-$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$.
-In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the only difference between our method and methods like $\locfdr$ and $\mixfdr$ are in how they estimate $g$.
-
-To briefly summarize the results in this section:
-\begin{enumerate}
-\item The UA can produce very different inferences compared with the ZA made by existing methods.
-\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
-\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
-\end{enumerate}
-
-\subsubsection*{The UA and ZA can produce different inferences}
-
-\begin{figure}
-\center \includegraphics[height=6in]{figures/decomp_ZA.pdf}
-\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
-The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
-In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
-The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
-(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
-\end{figure}
-
-To illustrate the different inferences from the UA and ZA we show results for a single
-dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
-but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
-\qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
-The results (Figure \ref{fig:ZA}) illustrate the clear difference between the existing methods and our method. The effects of the ZA made by \qvalue and \locfdr
-are visually clear, producing a ``hole"  in the alternative $z$ score distribution around 0. Although \mixfdr does not formally make the ZA, its decomposition
-exhibits a similar hole. In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
-effectively ``filling in" the hole.  (Of course the null distribution also has a peak at 0, and the local fdr 
- under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)
-
-Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA.
-Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
-This would increase the null component (dark blue) at the expense of
-the alternative component (light blue).
-Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0.
-The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears.
-In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
-See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
-demonstration.
-
- \subsubsection*{The UA can produce conservative estimates of $\pi_0$}
-
-The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
-Consequently \ashr will estimate smaller $\lfdr$s and FDRs than existing methods that make the ZA. 
-This is desirable, provided that these estimates remain conservative: that is,
-provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
-The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behaviour. To check its effectiveness
-we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
-for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
-They range from a ``spiky" distribution -- where many non-zero $\beta$ are
-too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
-flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
-making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
-which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.
-
-% How do these differences influence estimates of the FDR? To address this, note that  -- in this simple case where the 
-% $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
-% of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
-% strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
-% although the methods considered here differ from one another in many ways, having estimated $\pi_0$
-% they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
-% empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
-%
-
-%Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
-% and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
-%Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
-%At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
-%overestimates of $\pi_0$, which will yield overestimates of the FDR).
-
-\begin{figure} 
-\begin{center}
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=1.8in]{figures/scenarios_density.pdf} 
-	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/pi0_est.pdf} 
-\caption{Comparison of true and estimated values of $\pi_0$.
-When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ashr being least conservative, and hence most accurate.
-When the UA does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.} \label{fig:pi0sims}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfdr_est.png} 
-\caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
-Black line is $y=x$ and red line is $y=2x$.
-Estimates of $\lfdr$ are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est.png} 
-\caption{As in c), but for $\lfsr$ instead of $\lfdr$.
-Estimates of $\lfsr$ are consistently less conservative than $\lfdr$ when UA holds, and also less anti-conservative in bimodal scenario.}  \label{fig:lfsr}
-\end{subfigure}
-\end{center}
-\caption{Results of simulation studies (constant precision $s_j=1$).}
-\end{figure}
-
-For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
-data as follows:
-\begin{enumerate}
-\item Simulate $\pi_0  \sim U[0,1]$.
-\item For $j=1,\dots,J$, simulate $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)$.
-\item For $j=1,\dots,J$, simulate $\bhat_j | \beta_j \sim N(\beta_j,1)$.
-\end{enumerate}
-%Thus these simulations assume the same
-%precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).
-
-Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ashr ($y$ axis) with the true values ($x$ axis). 
-For \ashr we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
-and as a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
-as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
-The results show that \ashr provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
-in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ashr estimates can be slightly anti-conservative.
-We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
-are used. (In addition, the effects on $\lfsr$ estimates turn out to be relatively modest; see below).
-
-%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
-%
-%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
-%and so all effects are non-zero and no nulls hold. This scenario is rather different
-%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
-%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
-%``mostly non-null"
-%scenario, and we would agree! On the other hand, the distribution of $p$ values
-%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
-%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
-%Thus it seems desirable to have methods that perform well in both cases, and this is
-%an important feature of the methods we introduce here.
-%
-%TODO: put FDRs on graph? Simulate with 50\% null?
-
-\subsubsection*{The $\lfsr$ is more robust than \lfdr}
-
-The results above show that \ashr can improve on existing methods in producing smaller, more accurate, 
-estimates of $\pi_0$, which will lead to more accurate estimates of FDR.
-Nonetheless, in many scenarios \ashr continues
-to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
-This is because these scenarios include 
-an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
-estimation of $\pi_0$ impossible. Put another way, and as is well known, $\pi_0$ is not identifiable:
- the data can effectively provide an upper bound on plausible values of $\pi_0$,
-but not a lower bound (because the data cannot rule out that everything is non-null, but with miniscule effects).
-To obtain conservative behaviour we must estimate $\pi_0$ by this upper bound, which can
-be substantially larger than the true value.
-
-Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
-overestimation of $\pi_0$ is corresponding overestimation of FDR (and $\lfdr$, and $q$ values).
-To illustrate, Figure \ref{fig:lfdr} compares the estimated $\lfdr$ from ash.n with the true value (computed using
-Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios which involve many
-non-zero effects that are very near 0 (e.g.~the spiky scenario with $\pi_0$ small) where
-$\pi_0$ can be grossly overestimated.  
-(Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.)
-
-The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
- become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
- This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
-To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
-be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
-the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in $\lfsr$ than $\lfdr$.
-
-Note that, compared with previous debates regarding testing, 
-this section advances an additional reason for focussing on the sign of the effect, rather than just testing whether it is 0. 
-In previous debates authors have argued against testing whether an effect is 0
-because it is {\it implausible that effects are exactly 0}. Here we add that {\it even if one believes
-that some effects may be exactly zero}, it is still better to focus on the sign, because generally {\it the data are more informative about that question}
-and so inferences are more robust to, say, the inevitable mis-estimation of $\pi_0$.
-To provide some intuition, consider an observation with a $z$ score of 0. The $\lfdr$ of this observation can range from 0 (if $\pi_0=0$)
-to 1 (if $\pi_0=1$). But, assuming a symmetric $g$, the $\lfsr>0.5$ whatever the value of $\pi_0$, because the observation $z=0$ says
-nothing about the sign of the effect.
-Thus, are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
-whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.
-
-Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
-Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
-if we are prepared to assume that no effects are exactly zero, then removing the point mass 
-yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
-However, there is ``no free lunch":  if in fact some effects are exactly zero
-then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
-We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.
-
-\begin{figure}
-\begin{center}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s_nn.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be more accurate in this case.
-That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
-\end{subfigure}
-\end{center}
-\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
-\end{figure}
-
-\subsubsection*{Numerical Stability}
-
-The EM algorithm, which we use here to fit our model, is notorious for convergence to local optima. 
-However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
-caused by such behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
-in addition running it using our default initialization procedure (see Implementation details). We then compared the largest
-log-likelihood achieved across all 11 runs with the log-likelihood achieved by the default run.
-When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference ($< 0.02$), 
-and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference ($<0.02$),
-and 6\% of runs showed an appreciable log-likelihood difference ($>1$), with the largest difference being 5.0. However, perhaps suprisingly, 
-even for this largest difference results from the default run (e.g.
-the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood.
-See \url{http://github.com/stephens999/ash/blob/master/dsc-robust/summarize_dsc_robust.rmd} for further details.
-
-\subsubsection*{The UA helps provide reliable estimates of $g$}
-
-An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
-can estimate the {\it size} of each effect $\beta_j$.
-Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
-to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
-threshold $T$.
-%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
-%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
-%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
-%towards zero, with the strength of the shrinkage being estimated from the data.
-Further, because the posterior distribution is, by definition,
-conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
-been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
-In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
-\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
-how can confidence intervals be assigned to the true
-effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
-and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
-%but also notes the difficulties of implementing a stable general algorithm.
-
-The ability of the EB approach to provide valid ``post-selection" interval estimates is extremely attractive in principle.
-But its usefulness in practice
-depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
-which are notoriously difficult in general. Indeed, Efron emphasises
-the difficulties of implementing a stable general algorithm, noting in his rejoinder
-``the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying
-to be overly nonparametric ... but it is hard to imagine a
-generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
-Our key point here is that the UA greatly simplifies the deconvolution problem.
-While not meeting Efron's desire for an entirely general nonparametric approach, we believe 
-that the UA can handle many cases of practical interest.
-
-To illustrate this, Figure \ref{fig:egcdf} compares the
-estimated $g$ from \ashr  with that from \mixfdr which does not make the UA 
-(and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
-The greater reliability of estimates afforded
-by the UA is immediately apparent. In particular the estimated cdf from \mixfdr often has an almost-vertical
-segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. 
-The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
-While the UA is not the only way to achieve this, we find it an attractive, simple and effective approach.
-
-Interestingly, even in the ``bimodal" scenario $\ashr$ 
-is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
-Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
-Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
-to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
-(\url{http://github.com/stephens999/ash/blob/master/dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
-clearly visible in the observed $\bhat$.
-
-We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ashr tends to systematically
-overestimate the mass of $g$ near zero. On careful inspection, this is apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
-and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf}b),
-and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
-under-estimating $\pi_0$ (Figure \ref{fig:egcdf}c).
-
-\begin{figure}[h!] 
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=2in]{figures/egcdf.pdf}
-	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf_nopen.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
-Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
-\end{subfigure}
-\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.} \label{fig:egcdf}
-\end{figure}
-
-%making the UA provides an attractive and simple way to achieve estimating $g$  large
-%subset of cases of interest (those where the UA can reasonably be assumed to hold).
-
-%In practice, the utility of the posterior distributions noted above depends primarily on the accuracy
-%of the estimate of the distribution $g(\cdot)$.
-
- %quote from efron rejoinder , p46:
-%MorrisÕ Section 3 is especially pertinent. His formula
-%for p(?|z) is related to my discussion of the
-%BenjaminiÐYekutieli False Coverage Rate method in
-%Section 7, particularly (7.2)Ð(7.4). Originally I had
-%hoped to develop an empirical Bayes method for estimating
-%such models, but the effort foundered on practical
-%difficulties involving the perils of deconvolution.
-%Section 6 on the Òone-group modelÓ is the ugly duckling
-%of the current paper, but it bears on some important
-%points raised in the discussion. Figure 7 concerns
-%a fuzzy version of simultaneous hypothesis testing,
-%where, as in MorrisÕ hospital example, the usual singlepoint
-%null hypotheses seem unequal to the task. The
-%development from (6.6) onwards, particularly (6.12),
-%bears on the possibility of nonnormal null distributions,
-%and is about as far as I can go in answering Professors
-%Rice and SpiegelhalterÕs penultimate question.
-%With g(?) a normal distribution, model (6.1) returns
-%us to the realm of Stein estimation, the scene of my
-%happy collaborations with Carl Morris. I continue to
-%be surprised at how much harder bumpy, nonnormal
-%models like (7.1) are to deal with. JamesÐStein estimation
-%works fine with, say, N = 10 component problems,
-%while the RobbinsÕ form of empirical Bayes appropriate
-%to (7.1) seems to require hundreds or thousands.
-%The information calculations in Efron (2008)
-%reinforce this gloomy assessment. Maybe I am trying
-%to be overly nonparametric in constructing the empirical
-%Bayes Fdr estimates, but it is hard to imagine a
-%generally satisfactory parametric formulation for (6.1).
-%Or perhaps it is just that hypothesis testing is more demanding
-%than estimation.
-%My Note: (6.1) is the general model (mu \sim g()).
-%7.1 is the case where the effects are a mixture of zero and N(-3,1).
-
-\subsubsection*{Calibration of posterior intervals}
-
-To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over 100 simulations in each Scenario). 
-Specifically we examine the empirical coverage of nominal lower 95\% credible bounds for a) all observations; b) significant negative discoveries; c) significant positive discoveries.
-We examine only lower bounds because the results for upper bounds follow by
-symmetry (except for the one asymmetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
-in each case: for negative discoveries the lower bound is typically large and negative and limits how big (in absolute value) 
-the effect could be; for positive discoveries the lower bound 
-is positive, and limits how small (in absolute value) the effect could be. Intuitively, the lower bound for negative discoveries depends on the accuracy of $g$ in its tail,
-whereas for positive discoveries it is more dependent on the accuracy of $g$ in the center.
-
-The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
-as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
-
-% As noted above, the penalty term (\ref{eqn:penalty}) causes systematic over-estimation of the mass of $g$ near 0,  which one would expect to overshrink interval estimates towards zero. 
-%The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound is negative and overshrinking (moving it too far towards zero) makes the left tail too big; conversely, for positive discoveries, the lower confidence bound is positive, and
-%overshrinking (moving it too far towards zero) makes the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
-%and be prepared to put up with some overshrinkage to avoid it.
-
-%Coverage rates are given in:
-%\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
-
-%Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all.tex}
-	\caption{All observations.
-Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario.
-This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero.
-Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (Table \ref{tab:nopen}).
-Removing the penalty in the half-uniform case is not recommended (see Appendix).}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg.tex}
-\caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
-and the normal-based method in the flat-top scenario.
-These results likely reflect inaccurate estimates of the tails of $g$ due to a disconnect between the tail of $g$ and the component distributions in these cases.
-For example, the uniform methods sometimes substantially underestimate the length of the tail of $g$ in these long-tailed scenarios,
-causing over-shrinkage of the tail toward 0.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos.tex}
-\caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds} \label{tab:coverage}
-\end{table}
-
-%
-%\subsubsection*{Reverse asymptotics}
-%
-%Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
-%
-%Note: we'd like to make sure it behaves ok in the case where most observations are null.
-%Perhaps simulate with 999 null and 1 alternative?
-%Also maybe look at performance for small sample sizes - e.g. n=1?
-%
-%Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
-%can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
-%by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
-%
-
-\subsection*{Differing measurement precision across units}
-
- We turn now to the second important component of our work: allowing for varying
- measurement precision across units. The key to this is the use of a likelihood,
- (\ref{eqn:normlik}) or (\ref{eqn:tlik}), that explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
-
-  To illustrate, we conduct a simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
- ($s_j=10$).  In both cases,
-we assume that half the effects are null
-and the other half are normally distributed with standard deviation 1:
-\begin{equation}
-p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1).
-\end{equation}
-In this setting, the poor-precision measurements $(s_j=10)$ tell us very little, and any sane analysis should effectively ignore them. 
- However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor}). This is because the poor measurements 
- produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
- which, when combined with the good-precision measurements, dilute the overall signal (e.g. they reduce the density of $p$ values near 0).
- This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
-the estimated error rates ($q$-values, or $\lfdr$ values) for the good-precision observations increase when the low-precision observations are included in the analysis
-(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
-for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).
-
-\begin{figure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_hist.pdf} 
-    \label{fig:goodpoor-hist}
-    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations}
- \end{subfigure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_scatter.pdf}
-    \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) of a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
-\end{subfigure}
-\caption{Simulation illustrating how, for existing FDR methods, 
-poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
-$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
-a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
-In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
-\end{figure}
-
- \subsubsection*{Reordering of significance, and the ``$p$ value prior"}
-
-Another consequence of accounting for differences in measurement precision is that \ashr may re-order the significance of the observations compared
- with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
- We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measuments that have the same $p$ value.
- The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
- and so the $\lfsr$ for these poor-precision measurements is always high.
-
-The potential for Bayesian analyses to re-order the significance of observations, and specifically to downweight imprecise observations,
-was previously discussed in \cite{guan.stephens.08}.
-However, Wakefield \cite{wakefield:2009} showed that, under a certain prior assumption, the Bayesian analysis produces the same ranking of significance as $p$ values (or their $z$ scores).
-He named this prior the ``$p$-value prior" because it can be thought of as the
-implicit prior assumption that is being made if we rank the significance of observations by their $p$ value.
-Wakefield's $p$-value prior assumes that the less precise effect estimates correspond to larger true effects, and specifically that they scale proportional
-to the standard errors $s_j$. More specifically still, it assumes a normal prior for the non-zero $\beta_j$ with mean 0 and variance $K s^2_j$ for some constant $K$.
-Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. Specifically, if, instead of assuming that $\beta_j$ is independent
-of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
-then the $\lfsr$ computed by ash.n provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
-as is illustrated in Figure \ref{fig:lfsr_pval}, right panel. (This result does not hold when using the mixtures of uniforms prior, ash.u.)
-
-This $p$-value prior assumes that the $z$ scores $z_j=\beta_j/s_j$
-are identically distributed, independent of $s_j$. This is essentially the assumption made, implicitly or explicitly, by existing methods -- like \locfdr, \mixfdr and \qvalue --
-that model the $z_j$ or $p_j$ directly.  In contrast, we have assumed up to now that the $\beta_j$ are independent of $s_j$. 
-We can set both these assumptions within a more general framework, which allows that $\beta_j/s^\alpha$ is independent of $s_j$ for some $\alpha$ [Equation (\ref{eqn:beta-alpha}))].
-Setting $\alpha=0$ implies that $\beta_j$ is
-independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
-effects (in absolute value). This latter assumption may often be qualitatively plausible: 
-for example, in gene expression studies the standard error for gene $j$ depends partly on
-the variance of its expression among samples, and genes with a larger variance may tend to be less tightly regulated and
-so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
-On the other hand, there is no particular reason to expect that either $\alpha=1$ or $\alpha=0$ will be the optimal choice.
-Indeed, optimal choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$,
-which will be dataset-specific. 
-Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
-comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
-likelihood-based methods can be used to select $\alpha$. For example, following the logic of the EB approach it
-would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
-which has been implemented in our software by C. Dai.
-
-%\begin{result}
-%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
-%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
-%\end{result}
-
-%To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
-%Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
-%In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
-%Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
-%which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
-%(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
-%may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)
-
-%\begin{result}
-%Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and using a mixture of normals for the prior distribution 
-%(\ref{eqn:beta-alpha}) with $\alpha=1$, 
-%the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$.
-Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$)
-%is montonic with $|t_j|$. 
-%\end{result}
-
-%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
-%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.
-
-%Most previous authors of EB approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
-%have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
-%In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
-%This is analogous
-%to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
-%standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
-%that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.
-
-\begin{figure}
-\centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
-\caption{Figure illustrating affects of prior assumptions on re-ordering of significance.
-Left panel shows results under our ``default prior" which assumes that effects $\beta_j$ are identically distributed, independent of $s_j$.
-Right panel shows results under the ``$p$-value prior", which assumes that $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$.} \label{fig:lfsr_pval}
-\end{figure}
-
-%\subsubsection*{Adaptive Shrinkage}
-%
-%Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
-%when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
-%the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
-% $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
-% In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
-% Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
-% and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
-%
-%Our methods we build on this idea in two ways. First,
-%by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
-%Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
-%adapts to measurement precision: more precise measurements undergo
-% less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
-%than observations with
-%small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
-%To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
-%Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
-%
-%\begin{figure}[!ht]
-%\begin{center}
-%\includegraphics[width=400px]{../examples/ash.pdf} 
-%\end{center}
-%\caption{Figure illustrating adaptive shrinkage.
-Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0.
-The standard error of each observation, $s_j$, was simulated from Inverse Gamma(5,5), which resulted in $s_j$ varying from 0.34 to 6.7.
-The shrunken estimates ($y$ axis) are plotted against the observed estimates ($x$ axis), with color indicating the (square root of the) precision of each measurement (blue=lower-precision; red=higher-precision).
-The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
-%\end{figure}
-
-%DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
-%To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
-Because the prior distribution is unimodal about zero,
-%the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
-%amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
-%For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
-%than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
-%small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
-
-%previous work based on normal means:
-%http://link.springer.com/article/10.1007/BF02562623
-
-%\subsection{Null behaviour}
-
-%Results of simulations under the null.
-
-%\subsection{Comparison of different underlying component distributions}
-%
-%Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
-%than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
-%likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
-%For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
-%increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
-%expect $L_{hu}$ to be significantly larger than the others.
-%
-%Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
-%
-%\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
-%\begin{center}
-%\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
-%%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
-%\end{center}
-%\caption{}
-%\end{figure}
-
-%Notes from Adrian Raftery - 
-%Newcombe (1886)
-%Dean and Raftery
-%Uniform is super-efficient
-
-\section*{Discussion}
-
-We have presented an Empirical Bayes approach to large-scale multiple testing that emphasises two ideas.
-First, we emphasise the potential benefits of using two numbers ($\bhat$, and its standard error)
-rather than just one number (a $p$ value or $z$ score)  to summarize the information on each test.
-While requiring two numbers is slightly more onorous than requiring one, in many settings these
-numbers are easily available and if so we argue it makes sense to use them.
-Second, we note the potential benefits -- both statistical and computational -- of assuming that the effects come from a unimodal distribution, and provide
-flexible implementations for performing inference under this assumption. We also introduce the ``false sign rate" as
-an alternative measure of error to the FDR, and illustrate its improved robustness to errors in model fit, particularly mis-estimation
-of the proportion of null tests, $\pi_0$.
-
-Multiple testing is often referred to as a ``problem" or a ``burden". In our opinion, EB approaches turn this idea on its head,
-treating multiple testing as an {\it opportunity}: specifically, an opportunity to learn about the prior distributions, and other modelling assumptions,
-to improve inference and make informed decisions about significance (see also \cite{greenland1991empirical}). This view
-also emphasises that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
-Indeed, the FDR at a given fixed threshold does not depend on the number of tests: as the number of tests increases, both the true positives and false positives increase linearly,
-and the FDR remains the same.
-(If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example.
-The EB approach captures this dependence in an intuitive way:
-if there are lots of strong signals then we infer $\pi_0$ to be small, and the estimated FDR (or $\lfdr$, or $\lfsr$) at a given threshold may be low, even if a large number of tests were performed; and conversely
-if there are no strong signals then we infer $\pi_0$ to be large and the FDR at the same threshold may be high, even if relatively few tests were performed. 
-More generally, overall signal strength is reflected in the estimated $g$, which in turn influences the estimated FDR.
-
-Two important practical issues that we have not addressed here are correlations among tests,
-and the potential for deviations from the theoretical null distributions of test statistics. These two
-issues are connected: specifically, unmeasured confounding factors can cause both correlations among tests
-and deviations from the theoretical null \cite{efron2007correlation,leek:2007}. And although there are certainly other
-factors that could cause dependence among tests, unmeasured confounders are perhaps the most worrisome in practice
-because they can induce strong correlations among large numbers of tests and profoundly impact results.
-Approaches to deal with unmeasured confounders can be largely divided into two types: those
-that simply attempt to correct for the resulting inflation of test statistics \cite{devlin1999genomic,efron2004large}, and those
-that attempt to infer counfounders using clustering, principal components analysis, or factor models 
-\cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
-and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable,
-they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alterntively, our methods
-could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:lik}) using
-$\bhat_j | \shat_j \sim N(0,\lambda_1 \shat_j + \lambda_2)$, where $\lambda_1,\lambda_2$ are to be estimated. However,
-this immediately raises issues of identifiability and we do not pursue the idea further here.
-
-Another important practical issue is the challenge of small sample sizes. For example, in genomics applications researchers sometimes
-attempt to identify differences between two conditions based on only a handful of samples in each. In such settings the
-normal likelihood approximation (\ref{eqn:normlik}) will be inadequate. And, although the $t$ likelihood (\ref{eqn:tlik}) partially
-addresses this issue, it is also, it turns out, not entirely satisfactory.  The root of the problem is that, with small sample sizes, raw estimated standard
-errors $\shat_j$ can be horribly variable. In genomics it is routine to address this issue by applying EB methods \cite{smyth:2004} to ``moderate" (i.e.~shrink) variance estimates,
-before computing $p$ values from ``moderated" test statistics. We are currently investigating how
-our methods should incorporate such ``moderated" variance estimates to make it applicable to small sample settings.
-
-Our approach involves compromises between flexibility, generality, and simplicity on the one hand, and statistical efficiency and principle on the other.
-For example, in using an EB approach that uses a point estimate for $g$, rather than a fully Bayes approach that accounts for uncertainty in $g$,
-we have opted for simplicity over statistical principle. And in summarizing every test by
-two numbers and making a normal or $t$ approximation to the likelihood, we have aimed to produce generic methods that 
- can be applied whenever such summary data are available -- just as \qvalue can be applied to any set of $p$ values for example -- although
- possibly at the expense of statistical efficiency compared with developing multiple tailored approaches based on context-specific likelihoods. 
- Any attempt to produce generic methods will involve compromise between generality and efficiency.
- In genomics, many analyses -- not only FDR-based analyses -- involve first computing a series of $p$ values before
- subjecting them to some further downstream analysis. An important message here is 
- that working with two numbers ($\bhat_j,\shat_j$), rather than one ($p_j$ or $z_j$), 
- can yield substantial gains in functionality (e.g.~ estimating effect sizes, as well as testing; accounting for variations in
-measurement precision across units) while losing only a little in generality.  We hope that our work will
-encourate development of methods that exploit this idea in other contexts.
-
-%This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
-%from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
-%with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
-
-%
-%
-%Both \locfdr and \mixfdr allow for deviations from
-%the theoretical null distribution by allowing that, under the null, the $z$ scores are $N(0,\sigma^2)$ with $\sigma$ estimated from the data rather
-%than fixed to 1. Within our approach this idea could be incorporated by changing the likelihood - for example, to assume $\bhat_j \sim N(0,s_ 
-%
-%These are tricky issues that, as far as we are aware, can affect
-%all approaches to FDR analysis. An important focus of Efron's work, also implemented in \locfdr, 
-% is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
-%theoretical null distribution. 
-%
-%For example, we and others have worked on Bayesian approaches to genetic association testing, using subjective prior specification of
-%effect sizes based on external data (e.g.~\cite{stephens:2009}). However, the large numbers of tests brings the opportunity to learn about
-%priors, and here we aim to exploit this
-
-%Choice of mixture components. Use of likelihood to compare different mixture components?
-
-%  
-%Directly modeling the p values, or z scores, say via non-parametric methods, 
-%can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-%distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-%The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-%Before proceeding, we feel obliged to correct the common misconception
-% that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
-% not depend on the number of tests. 
-%  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
-% 
-% In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
-
-\section*{Detailed Methods} \label{sec:implement}
-
-\subsection*{Embellishments}
-
-\subsubsection*{More flexible unimodal distributions}
-
-Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
-Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
- Therefore, we have implemented a more general approach based on 
-\begin{equation} \label{eqn:g-general}
-g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) ,
-\end{equation}
-where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
-\begin{enumerate}[(i)]
-\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$, \qquad (``ash.n")
-\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,  \qquad (``ash.u")
-\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,  \qquad (``ash.hu")
-\end{enumerate}
-where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
-(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
-and $\pi$ is a $2K+1$ vector that sums to 1.)
-The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
-Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
-under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
-(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
-(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
-however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)
-
-Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
-using a large dense grid of $\sigma^2_k$ or $a_k$, (i)-(iii) can respectively approximate,
-with arbitrary accuracy,
-\begin{enumerate}[(i)]
-\item any scale mixture of normals, which includes as special cases
-the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
-\item any symmetric unimodal distribution about 0.
-\item any unimodal distribution about 0.
-\end{enumerate}
-The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
-In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.
-
-Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
-The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
-See Implementation for details of our software defaults.
-
-%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting.
-For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
-%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
-
-\subsubsection*{Dependence of effects on standard errors}
-
-Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
-This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
- \begin{equation} \label{eqn:beta-alpha}
- \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
- \end{equation}
-for any $\alpha$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
-assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
-it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
-the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.
-
-The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
-To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
- standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
- Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
-yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
-%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
-%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
-%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
-
-%Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
-%how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
-%it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
-%rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
-%and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
-%this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
-%do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
-
-\subsubsection*{Replace normal likelihood with $t$ likelihood}
-
-We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
-  \begin{equation} \label{eqn:tlik}
- \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
- \end{equation}
-where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
-and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
-The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
-This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms; see Implementation below.
-When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
-
-Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
-\begin{equation} \label{eqn:tdist}
-(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
-\end{equation}
-However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
-In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
-currently pursuing this improved approach (joint work with M.Lu) and results will be published elsewhere. 
-%We note here that
-%modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
-%where our simpler approach can produce unsatisfactory results.
-
-\subsubsection*{Non-zero mode}
-
-An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood, rather than fixed to 0.
-
-\subsection*{Implementation Details}
-
-\subsubsection*{Likelihood for $\pi$}
-
-We define the likelihood for $\pi$ to be the probability of the observed data $\bhat$ conditional on $\shat$: $L(\pi) := p(\bhat | \shat,\pi),$ which
-by our conditional indendence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$.
-[One might prefer to define the likelihood as $p(\bhat, \shat | \pi) = p(\bhat | \shat, \pi) p(\shat | \pi)$, in which case our definition comes down to assuming that the term $p(\shat | \pi)$ does not depend on $\pi$.]
-
-Using the prior $\beta_j \sim \sum_{k=0}^K \pi_k f_k(\beta_j)$ given by (\ref{eqn:g-general}), and the normal likelihood (\ref{eqn:normlik}), integrating over $\beta_j$ yields
-\begin{equation}
-p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
-\end{equation}
-where 
-\begin{equation}
-\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
-\end{equation}
-denotes the convolution of $f_k$ with a normal density.
-These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
-Specifically, 
-\begin{equation} \label{eqn:uconv}
-\tilde{f}_k(\bhat_j)  = 
-\begin{cases}
-N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
-\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
-\end{cases}
-\end{equation}
-where $\Psi$ denotes the cumulative distribution function (c.d.f.) of the standard normal distribution.
-If we replace the normal likelihood with the $t_\nu$ likelihood (\ref{eqn:tlik}) then the convolution 
-for $f_k$ uniform the convolution is still given by (\ref{eqn:uconv}) but with $\Psi$ the c.d.f. 
-of the $t_\nu$ distribution function. (The convolution for $f_k$ normal is tricky and we have not implemented it.)
-
-\subsubsection*{Penalty term on $\pi$}
-
-To make $\lfdr$ and $\lfsr$ estimates from our method ``conservative" we add a penalty 
-term $log(h(\pi;\lambda))$ to the log-likelihood $\log L(\pi)$ to encourage over-estimation of $\pi_0$:
-\begin{equation} \label{eqn:penalty}
-h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
-\end{equation}
-where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
-consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0sims}).
-
-Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
-we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
-
-\subsubsection*{Problems with removing the penalty term in the half-uniform case}
-
-It is straightforward to remove the penalty term by setting $\lambda_k=1$ in (\ref{eqn:penalty}).
-We note here an unanticipated problem we came across when using no penalty term in the half-uniform case 
-(i.e.~ $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$ in (\ref{eqn:g-general})): when the data
-are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
-sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
-for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
-and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
-
-\subsubsection*{EM algorithm}
-
-With this in place, the penalized log-likelihood for $\pi$ is given by:
-\begin{equation}
-\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
-\end{equation}
-where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This can be maximized using an EM algorithm \cite{dempster77}, whose
-one-step updates are:
-\begin{align} \label{eqn:w}
-w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
-n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
-\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}
-\end{align}
-Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
-and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
-probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$. (The $w_{kj}$ are sometimes referred to as the ``responsibilities".)
-Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
-We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple} to accelerate convergence of this EM algorithm.
-
-%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
-\subsubsection*{Initialization}
-
-By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
-Our rationale for initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
-but weak signal in the data cannot quickly draw the algorithm towards the null.
-
-In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
-$\pi_0=1; \pi_k=0 (k>0)$. If not then we replace the EM solution with the global null solution. This aims to guard against errors due to convergence to a local optimum when the
-data are consistent with the global null.
-
-\subsubsection*{Conditional distributions}
-
-Given $\hat\pi$, we compute the conditional distributions 
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
-\end{equation} 
-Each posterior is a mixture on $K+1$ components:
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
-\end{equation}
-where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
-and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
-$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
-All these posterior distributions are easily available.
-For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
-If $f_k$ is normal and $L$ is normal, then this is a normal distribution.
-
-\subsubsection*{Choice of grid for $\sigma_k, a_k$} \label{sec:grid}
-
-\def\sigmamax{\sigma_{\text{max}}}
-\def\sigmamin{\sigma_{\text{min}}}
-When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
-in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
-as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
-based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
-is negative we set $\sigmamax = 8\sigmamin$.)
-
-When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.
-
-Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
-a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.
-
-\appendix
-
-\begin{table}[!ht]
-\centering\begin{tabular}{c c } \toprule
-Scenario & Alternative distribution, $g_1$  \\ \midrule
-spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
-near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
-flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
- &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
-skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
-big-normal & $N(0,4^2)$ \\ 
-bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
-\end{tabular}
-\caption{Summary of simulation scenarios considered} \label{table:scenarios}
-\end{table}
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all_nopen.tex}
-	\caption{All observations}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg_nopen.tex}
-\caption{``Significant" negative discoveries.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos_nopen.tex}
-\caption{``Significant" positive discoveries.}
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{table:nopen}
-\end{table}
-%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
-%\begin{equation}
-%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
-%\end{equation}
-%which is close to the definition in \cite{storey.03}.
-
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgements}
-
-Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
-prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} , and includes contributions 
-from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.
-
-This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.
-
-%\section*{References}
-% The bibtex filename
-\bibliography{/Users/stephens/Dropbox/Documents/mainbib}
-
-\section*{Figure Legends}
-
-\clearpage
-
-\section*{Tables}
-
-\section*{Supporting Information Legends}
-
-Supplementary material can be found in {\bf Supplementary Information S1.}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
-\end{document}

### 2015-10-06 sectionMethods__subs.tex
New Article!

--- 2015-10-06 paper/main.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
-\documentclass[11pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-%\usepackage{natbib}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-\usepackage{booktabs}
-\usepackage{xspace}
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-\usepackage{sectsty}
-\subsectionfont{\normalsize}
-\subsubsectionfont{\mdseries\itshape}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\usepackage{color}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-% Use the PLoS provided bibtex style
-\bibliographystyle{/Users/stephens/Dropbox/Documents/stylefiles/plos2009}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
-\pagestyle{myheadings}
-%% ** EDIT HERE **
-\usepackage{enumerate}
-\usepackage{multirow} 
-\usepackage{url}
-\usepackage{xr} %for cross-referencing
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-\newtheorem{algorithm}{Algorithm}
-\newtheorem{proposition}{Proposition}
-\newtheorem{restateproposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{result}{Result}
-\newtheorem{note}{Note}
-\newtheorem{definition}{Definition}
-
-\def\lfdr{\textit{lfdr}}
-\def\lfsr{\textit{lfsr}}
-\def\x{\mbf{x}}
-\def\y{\mbf{y}}
-\def\e{\mbf{e}}
-\def\g{\mbf{g}}
-\def\p{p}
-\def\Yb{\hat{Y}_1^{\text{BAYES}}}
-\def\Xperp{X_{1 \perp 0}}
-\def\Pperp{P_{1 \perp 0}}
-\def\Pperpb{\Pperp^{\text{B}}}
-\def \hatd { \widehat{ d } }
-\def \dhatin {\hatd^{\mbox{\tiny IN}}}
-\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
-\def \mvn{\text{MVN}}
-\def\mn{\text{MN}}
-\def \iw{\text{W}^{-1}}
-\def \fp {fastPHASE\ }
-\def\bhat{\hat{\beta}}
-\def\shat{\hat{s}}
-\def\var{\rm Var}
-\def\mvr{\text{BMVR}}
-\def\bfuni{\text{BF}_\text{uni}}
-\def\one{\mbf{1}}
-\def\zero{\mbf{0}}
-\def\hall{H_{\text{all}}}
-\def\pall{\p_{\text{all}}}
-\def\bfall{\text{BF}_\text{all}}
-\def\bfav{\text{BF}_\text{av}}
-\def\gall{\gamma_\text{all}}
-\def\limbfall{\bfall^\rightarrow}
-\def\limbfav{\bfav^\rightarrow}
-\def\limbfgamma{\BF_\gamma^\rightarrow}
-\def\sxx{S_{xx}}
-\def\syy{S_{yy}}
-\def\vyy{V_{yy}}
-\def\vyx{V_{yx}}
-\def\vxx{{V_{xx}}}
-\def\rss{\text{\rm RSS}}
-\def\dss{\text{\rm $\Delta$SS}}
-\def\U{U}
-\def\D{D}
-\def\I{I}
-\def\luu{L_{UU}}
-\def\indep{\perp}
-\def\qvalue{{\tt qvalue}\xspace}
-\def\locfdr{{\tt locfdr}\xspace}
-\def\mixfdr{{\tt mixfdr}\xspace}
-\def\ashr{{\tt ashr}\xspace}
-
-\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
-                                % mode too
-\newcommand\height{{\it height}}
-\newcommand\weight{{\it weight}}
-\def\BF{\text{BF}}
-
-%% END MACROS SECTION
-%\externaldocument{plos_one_rev1_suppinfo}
-
-\begin{document}
-
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{False Discovery Rates: A New Deal}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Matthew Stephens$^{1*}$, 
-\\
-\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
-\\
-$\ast$ E-mail: Corresponding mstephens@uchicago.edu
-\end{flushleft}
-
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-%\section*{Author Summary}
-
-\section*{Introduction}
-
-Since its introduction in 1995 by Benjamini and Hochberg 
-\cite{benjamini1995controlling}, the ``False Discovery Rate" (FDR) has quickly established itself
-as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing
-in which the goal is to identify the non-zero ``effects" among a large number of imprecisely-measured effects.
-
-Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
-indeed, the notion that EB approaches could be helpful in handling multiple comparisons predates introduction of
-the FDR  (e.g.~\cite{greenland1991empirical}). More recently, EB approaches to the FDR
-have been extensively studied by several authors, 
-especially B.~Efron and co-authors 
-\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
-see also \cite{kendziorski2003parametric,newton2004detecting, 
-datta2005empirical,muralidharan2010empirical} for example.
-So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
-existing widely-used FDR pipelines) and can substantially affect
-inference. The first idea is to {\it assume that the distribution of effects is unimodal}. This yields a very simple, fast, and stable computer 
-implementation, as well as improving inference of FDR when the unimodal assumption is correct.  
-The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
-to summarize each measurement. This idea allows variations in measurement precision to be better accounted for,
- and avoids a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
-
-In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
-the idea that it may be preferable to focus on estimation rather than on testing.
-In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
-simply estimation with some positive prior probability that the effect is exactly zero.
-However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
-perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
-controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
-``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
-in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
-the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the lfsr, particularly its increased robustness to modelling assumptions.
-
-%terminology
-%``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
-%and correspondingly define the ``false signed discovery rate", or simply ``false sign rate" (FSR), in a natural way.
-
-%The methods we provide here can address both problems, and we provide results that highlight important benefits of focussing on estimation, and particularly on assessing the correctness of the {\it sign} of the estimated effects, rather than on testing for non-zero effects.
-
-Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
-(We address the reasons for this name, and connections with shrinkage analysis, in the Discussion.)
-
-%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.
-
-%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-%measurements are made on only a small number of
-%samples in each condition - sometimes as few as one
-%sample in each condition. Thus the error in estimates
-%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-%in estimates of $\beta_j$ still greater.
-%
-
-%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
-%Here we present new tools that bring two important ideas to bear on the problem.
-%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
-%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
-%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
-%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
-%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
-%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
-%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
-%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
-%a substantial increase in the number of discoveries at a given FDR).
-
-%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
-%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
-Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
-%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
-%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
-%generic methods that work directly with two numbers for each 
-%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
-Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
-%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
-
-\def\df{df}
-\def\FDR{\text{FDR}}
-\def\fdr{\text{\lfdr}}
-\def\FSR{\text{FSR}}
-\def\fsr{\text{lfsr}}
-
 \section*{Methods}
@@
 
-% Here we consider one setting in which FDR-based methods are widely used:
-%an experimenter has measured a large numbers of ``effects", each with an associated
-%measurement precision (standard error), 
-%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
-%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
-%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
-%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
-%For example, in genomics
-%a common goal is to identify ``differentially-expressed genes" -- that is, genes
-%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
-%mean expression level between the two conditions, often referred to as the ``log-fold change". 
 Here we describe the simplest version of our method, and briefly discuss embellishments we have also implemented.
@@
 which by Bayes theorem can be written as
+
 \begin{equation}
@@
 a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
-\begin{align} \label{eqn:beta}
+
+\begin{align}
+\label{eqn:beta}
 p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
@@
 For the likelihood $p(\bhat | \beta, \shat)$ we assume
-\begin{equation} \label{eqn:normlik}
+
+\begin{equation}
+\label{eqn:normlik}
 p(\bhat | \beta, \shat) = \prod_j N(\bhat_j; \beta_j, \shat_j^2).
@@
 However, this simple model also has several additional assumptions that can be relaxed. Specifically,
+
 \begin{enumerate}
@@
 
-%We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
-%$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
-%\begin{equation}
-%p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
-%\end{equation}
-%Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
-%\begin{equation} \label{eqn:normlik}
-%p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
-%
-%We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
-%as a mixture of null and alternative components:
-%\begin{equation} \label{eqn:betaj0}
-%g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
-%\end{equation}
-%where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
-%We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
-%\begin{equation} \label{eqn:beta-iid}
-%p(\beta | \shat) = \prod_j g(\beta_j).
-%\end{equation}
-%
-%Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
-%with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
-%to use a mixture of zero-mean normal distributions:
-%\begin{equation} \label{eqn:mixnorm}
-%g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
-%Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
-%represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
-% The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
-% more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below. 
-% 
-% Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
-% yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
-% \begin{equation} \label{eqn:g}
-% g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
-% \end{equation}
-% 
-% Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
-% \begin{align} \label{eqn:betaj}
-% p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
-% p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
-% \end{align}
-% with $g$ given by the mixture of normals $(\ref{eqn:g})$.
-
 \subsection*{Fitting the model}
@@
 Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
-\begin{equation} \label{eqn:marginallik}
+
+\begin{equation}
+\label{eqn:marginallik}
 p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
@@
 The usual EB approach to fitting this model would involve two simple steps:
+
 \begin{enumerate}
@@
 In other words it is the posterior probability that $\beta_j$ is actually zero:
-\begin{equation} \label{eqn:fdr}
+
+\begin{equation}
+\label{eqn:fdr}
 \fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
@@
 Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
+
 \begin{quote}
@@
 would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
-\begin{equation} \label{eqn:lfsr}
+
+\begin{equation}
+\label{eqn:lfsr}
 \lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
@@
 To illustrate, suppose that
+
 \begin{gather*}
@@
 $\Gamma \subset \{1,\dots,J\}$. For example,
-\begin{equation} \label{eqn:FDRhat}
+
+\begin{equation}
+\label{eqn:FDRhat}
 \widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
@@
 And
-\begin{equation} \label{eqn:qvalue}
+
+\begin{equation}
+\label{eqn:qvalue}
 q_j := \widehat \FDR(\{k: \fdr_k \leq \fdr_j\}) 
@@
 provides a measure of significance analogous to Storey's $q$ value \cite{storey.03}.
-
-%Because our methods involve explicitly modelling effects $\beta_j$, rather than 
-%$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
-
-%Indeed, even in settings where some
-%$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
-%analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
 
@@
  (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
-  %Here we highlight the key differences between our approach and these three existing approaches.
 
@@
 how it can improve performance by taking better account of variation in measurement precision among observations.
-%(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
-% Or more generally Section xx illustrates the potential practical effects
 
-%allows for different modelling assumptions to be compared through likelihood comparisons.
 Second, our unimodal assumption (UA) that the effects are unimodal about zero is quite different from assumptions made by $\qvalue, \locfdr$ or $\mixfdr$.
@@
 It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
-
-%Tukey "peri-null"
-
-%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
-%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}.
 
@@
 Indeed, one way of thinking about the UA is as a way to regularize $g$.
-
-%The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
-%encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-%$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-%$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-%the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-%observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-%corresondingly little effect on the estimated FDR of other observations. 
-%We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-%$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-%that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-%estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-%Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-\def\dt{\text{dt}}
-
-%
-%assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
-%$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
-% $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
-%  \begin{equation} \label{eqn:lik}
-% \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
-% \end{equation}
-% where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e.
-the density of $\mu+\sigma T_\nu$).
-In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
-%  \begin{equation} \label{eqn:normlik}
-% \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
-% \end{equation}
-%which, in light of (\ref{eqn:betahat}), is equivalent to
- % \begin{equation} \label{eqn:lik2}
-% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
- %\end{equation}
-
-%The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
-they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
-The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$.
-Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$.
-If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$.
-See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.
-
-%This likelihood approximation can be motivated by assuming that i) $\bhat$
-% $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
-% $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$. 
-%
-
-%Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.
-
-%While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
- %that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
-%there is an important difference: here,
-
-%The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
-For example, we can write $p(\bhat_j, \shat_j | \beta_j) = p(\bhat_j | \shat_j, \beta_j)p(\shat_j | \beta_j)$.
-Now, if we are willing to assume that $\shat_j$ alone contains no information about $\beta_j$, or equivalently that $p(\shat_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\bhat_j | \shat_j, \beta_j)$, and if $\bhat_j | \shat_j, \beta_j \sim N(\beta_j, \shat_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.
-
-\section*{Results}
-
-We compare results of \ashr with existing FDR-based methods implemented
-in the R packages \qvalue (v2.0 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
-In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
-to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.
-
-\subsection*{Effects of the Unimodal Assumption}
-
-Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
-standard error, $s_j=1$ and all methods are provided that information. That is,
-$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$.
-In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the only difference between our method and methods like $\locfdr$ and $\mixfdr$ are in how they estimate $g$.
-
-To briefly summarize the results in this section:
-\begin{enumerate}
-\item The UA can produce very different inferences compared with the ZA made by existing methods.
-\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
-\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
-\end{enumerate}
-
-\subsubsection*{The UA and ZA can produce different inferences}
-
-\begin{figure}
-\center \includegraphics[height=6in]{figures/decomp_ZA.pdf}
-\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
-The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
-In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
-The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
-(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
-\end{figure}
-
-To illustrate the different inferences from the UA and ZA we show results for a single
-dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
-but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
-\qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
-The results (Figure \ref{fig:ZA}) illustrate the clear difference between the existing methods and our method. The effects of the ZA made by \qvalue and \locfdr
-are visually clear, producing a ``hole"  in the alternative $z$ score distribution around 0. Although \mixfdr does not formally make the ZA, its decomposition
-exhibits a similar hole. In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
-effectively ``filling in" the hole.  (Of course the null distribution also has a peak at 0, and the local fdr 
- under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)
-
-Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA.
-Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
-This would increase the null component (dark blue) at the expense of
-the alternative component (light blue).
-Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0.
-The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears.
-In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
-See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
-demonstration.
-
- \subsubsection*{The UA can produce conservative estimates of $\pi_0$}
-
-The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
-Consequently \ashr will estimate smaller $\lfdr$s and FDRs than existing methods that make the ZA. 
-This is desirable, provided that these estimates remain conservative: that is,
-provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
-The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behaviour. To check its effectiveness
-we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
-for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
-They range from a ``spiky" distribution -- where many non-zero $\beta$ are
-too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
-flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
-making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
-which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.
-
-% How do these differences influence estimates of the FDR? To address this, note that  -- in this simple case where the 
-% $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
-% of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
-% strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
-% although the methods considered here differ from one another in many ways, having estimated $\pi_0$
-% they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
-% empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
-%
-
-%Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
-% and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
-%Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
-%At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
-%overestimates of $\pi_0$, which will yield overestimates of the FDR).
-
-\begin{figure} 
-\begin{center}
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=1.8in]{figures/scenarios_density.pdf} 
-	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/pi0_est.pdf} 
-\caption{Comparison of true and estimated values of $\pi_0$.
-When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ashr being least conservative, and hence most accurate.
-When the UA does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.} \label{fig:pi0sims}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfdr_est.png} 
-\caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
-Black line is $y=x$ and red line is $y=2x$.
-Estimates of $\lfdr$ are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est.png} 
-\caption{As in c), but for $\lfsr$ instead of $\lfdr$.
-Estimates of $\lfsr$ are consistently less conservative than $\lfdr$ when UA holds, and also less anti-conservative in bimodal scenario.}  \label{fig:lfsr}
-\end{subfigure}
-\end{center}
-\caption{Results of simulation studies (constant precision $s_j=1$).}
-\end{figure}
-
-For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
-data as follows:
-\begin{enumerate}
-\item Simulate $\pi_0  \sim U[0,1]$.
-\item For $j=1,\dots,J$, simulate $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)$.
-\item For $j=1,\dots,J$, simulate $\bhat_j | \beta_j \sim N(\beta_j,1)$.
-\end{enumerate}
-%Thus these simulations assume the same
-%precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).
-
-Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ashr ($y$ axis) with the true values ($x$ axis). 
-For \ashr we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
-and as a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
-as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
-The results show that \ashr provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
-in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ashr estimates can be slightly anti-conservative.
-We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
-are used. (In addition, the effects on $\lfsr$ estimates turn out to be relatively modest; see below).
-
-%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
-%
-%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
-%and so all effects are non-zero and no nulls hold. This scenario is rather different
-%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
-%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
-%``mostly non-null"
-%scenario, and we would agree! On the other hand, the distribution of $p$ values
-%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
-%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
-%Thus it seems desirable to have methods that perform well in both cases, and this is
-%an important feature of the methods we introduce here.
-%
-%TODO: put FDRs on graph? Simulate with 50\% null?
-
-\subsubsection*{The $\lfsr$ is more robust than \lfdr}
-
-The results above show that \ashr can improve on existing methods in producing smaller, more accurate, 
-estimates of $\pi_0$, which will lead to more accurate estimates of FDR.
-Nonetheless, in many scenarios \ashr continues
-to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
-This is because these scenarios include 
-an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
-estimation of $\pi_0$ impossible. Put another way, and as is well known, $\pi_0$ is not identifiable:
- the data can effectively provide an upper bound on plausible values of $\pi_0$,
-but not a lower bound (because the data cannot rule out that everything is non-null, but with miniscule effects).
-To obtain conservative behaviour we must estimate $\pi_0$ by this upper bound, which can
-be substantially larger than the true value.
-
-Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
-overestimation of $\pi_0$ is corresponding overestimation of FDR (and $\lfdr$, and $q$ values).
-To illustrate, Figure \ref{fig:lfdr} compares the estimated $\lfdr$ from ash.n with the true value (computed using
-Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios which involve many
-non-zero effects that are very near 0 (e.g.~the spiky scenario with $\pi_0$ small) where
-$\pi_0$ can be grossly overestimated.  
-(Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.)
-
-The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
- become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
- This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
-To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
-be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
-the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in $\lfsr$ than $\lfdr$.
-
-Note that, compared with previous debates regarding testing, 
-this section advances an additional reason for focussing on the sign of the effect, rather than just testing whether it is 0. 
-In previous debates authors have argued against testing whether an effect is 0
-because it is {\it implausible that effects are exactly 0}. Here we add that {\it even if one believes
-that some effects may be exactly zero}, it is still better to focus on the sign, because generally {\it the data are more informative about that question}
-and so inferences are more robust to, say, the inevitable mis-estimation of $\pi_0$.
-To provide some intuition, consider an observation with a $z$ score of 0. The $\lfdr$ of this observation can range from 0 (if $\pi_0=0$)
-to 1 (if $\pi_0=1$). But, assuming a symmetric $g$, the $\lfsr>0.5$ whatever the value of $\pi_0$, because the observation $z=0$ says
-nothing about the sign of the effect.
-Thus, are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
-whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.
-
-Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
-Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
-if we are prepared to assume that no effects are exactly zero, then removing the point mass 
-yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
-However, there is ``no free lunch":  if in fact some effects are exactly zero
-then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
-We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.
-
-\begin{figure}
-\begin{center}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s_nn.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be more accurate in this case.
-That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
-\end{subfigure}
-\end{center}
-\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
-\end{figure}
-
-\subsubsection*{Numerical Stability}
-
-The EM algorithm, which we use here to fit our model, is notorious for convergence to local optima. 
-However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
-caused by such behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
-in addition running it using our default initialization procedure (see Implementation details). We then compared the largest
-log-likelihood achieved across all 11 runs with the log-likelihood achieved by the default run.
-When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference ($< 0.02$), 
-and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference ($<0.02$),
-and 6\% of runs showed an appreciable log-likelihood difference ($>1$), with the largest difference being 5.0. However, perhaps suprisingly, 
-even for this largest difference results from the default run (e.g.
-the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood.
-See \url{http://github.com/stephens999/ash/blob/master/dsc-robust/summarize_dsc_robust.rmd} for further details.
-
-\subsubsection*{The UA helps provide reliable estimates of $g$}
-
-An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
-can estimate the {\it size} of each effect $\beta_j$.
-Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
-to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
-threshold $T$.
-%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
-%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
-%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
-%towards zero, with the strength of the shrinkage being estimated from the data.
-Further, because the posterior distribution is, by definition,
-conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
-been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
-In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
-\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
-how can confidence intervals be assigned to the true
-effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
-and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
-%but also notes the difficulties of implementing a stable general algorithm.
-
-The ability of the EB approach to provide valid ``post-selection" interval estimates is extremely attractive in principle.
-But its usefulness in practice
-depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
-which are notoriously difficult in general. Indeed, Efron emphasises
-the difficulties of implementing a stable general algorithm, noting in his rejoinder
-``the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying
-to be overly nonparametric ... but it is hard to imagine a
-generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
-Our key point here is that the UA greatly simplifies the deconvolution problem.
-While not meeting Efron's desire for an entirely general nonparametric approach, we believe 
-that the UA can handle many cases of practical interest.
-
-To illustrate this, Figure \ref{fig:egcdf} compares the
-estimated $g$ from \ashr  with that from \mixfdr which does not make the UA 
-(and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
-The greater reliability of estimates afforded
-by the UA is immediately apparent. In particular the estimated cdf from \mixfdr often has an almost-vertical
-segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. 
-The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
-While the UA is not the only way to achieve this, we find it an attractive, simple and effective approach.
-
-Interestingly, even in the ``bimodal" scenario $\ashr$ 
-is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
-Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
-Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
-to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
-(\url{http://github.com/stephens999/ash/blob/master/dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
-clearly visible in the observed $\bhat$.
-
-We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ashr tends to systematically
-overestimate the mass of $g$ near zero. On careful inspection, this is apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
-and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf}b),
-and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
-under-estimating $\pi_0$ (Figure \ref{fig:egcdf}c).
-
-\begin{figure}[h!] 
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=2in]{figures/egcdf.pdf}
-	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf_nopen.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
-Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
-\end{subfigure}
-\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.} \label{fig:egcdf}
-\end{figure}
-
-%making the UA provides an attractive and simple way to achieve estimating $g$  large
-%subset of cases of interest (those where the UA can reasonably be assumed to hold).
-
-%In practice, the utility of the posterior distributions noted above depends primarily on the accuracy
-%of the estimate of the distribution $g(\cdot)$.
-
- %quote from efron rejoinder , p46:
-%MorrisÕ Section 3 is especially pertinent. His formula
-%for p(?|z) is related to my discussion of the
-%BenjaminiÐYekutieli False Coverage Rate method in
-%Section 7, particularly (7.2)Ð(7.4). Originally I had
-%hoped to develop an empirical Bayes method for estimating
-%such models, but the effort foundered on practical
-%difficulties involving the perils of deconvolution.
-%Section 6 on the Òone-group modelÓ is the ugly duckling
-%of the current paper, but it bears on some important
-%points raised in the discussion. Figure 7 concerns
-%a fuzzy version of simultaneous hypothesis testing,
-%where, as in MorrisÕ hospital example, the usual singlepoint
-%null hypotheses seem unequal to the task. The
-%development from (6.6) onwards, particularly (6.12),
-%bears on the possibility of nonnormal null distributions,
-%and is about as far as I can go in answering Professors
-%Rice and SpiegelhalterÕs penultimate question.
-%With g(?) a normal distribution, model (6.1) returns
-%us to the realm of Stein estimation, the scene of my
-%happy collaborations with Carl Morris. I continue to
-%be surprised at how much harder bumpy, nonnormal
-%models like (7.1) are to deal with. JamesÐStein estimation
-%works fine with, say, N = 10 component problems,
-%while the RobbinsÕ form of empirical Bayes appropriate
-%to (7.1) seems to require hundreds or thousands.
-%The information calculations in Efron (2008)
-%reinforce this gloomy assessment. Maybe I am trying
-%to be overly nonparametric in constructing the empirical
-%Bayes Fdr estimates, but it is hard to imagine a
-%generally satisfactory parametric formulation for (6.1).
-%Or perhaps it is just that hypothesis testing is more demanding
-%than estimation.
-%My Note: (6.1) is the general model (mu \sim g()).
-%7.1 is the case where the effects are a mixture of zero and N(-3,1).
-
-\subsubsection*{Calibration of posterior intervals}
-
-To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over 100 simulations in each Scenario). 
-Specifically we examine the empirical coverage of nominal lower 95\% credible bounds for a) all observations; b) significant negative discoveries; c) significant positive discoveries.
-We examine only lower bounds because the results for upper bounds follow by
-symmetry (except for the one asymmetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
-in each case: for negative discoveries the lower bound is typically large and negative and limits how big (in absolute value) 
-the effect could be; for positive discoveries the lower bound 
-is positive, and limits how small (in absolute value) the effect could be. Intuitively, the lower bound for negative discoveries depends on the accuracy of $g$ in its tail,
-whereas for positive discoveries it is more dependent on the accuracy of $g$ in the center.
-
-The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
-as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
-
-% As noted above, the penalty term (\ref{eqn:penalty}) causes systematic over-estimation of the mass of $g$ near 0,  which one would expect to overshrink interval estimates towards zero. 
-%The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound is negative and overshrinking (moving it too far towards zero) makes the left tail too big; conversely, for positive discoveries, the lower confidence bound is positive, and
-%overshrinking (moving it too far towards zero) makes the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
-%and be prepared to put up with some overshrinkage to avoid it.
-
-%Coverage rates are given in:
-%\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
-
-%Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all.tex}
-	\caption{All observations.
-Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario.
-This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero.
-Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (Table \ref{tab:nopen}).
-Removing the penalty in the half-uniform case is not recommended (see Appendix).}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg.tex}
-\caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
-and the normal-based method in the flat-top scenario.
-These results likely reflect inaccurate estimates of the tails of $g$ due to a disconnect between the tail of $g$ and the component distributions in these cases.
-For example, the uniform methods sometimes substantially underestimate the length of the tail of $g$ in these long-tailed scenarios,
-causing over-shrinkage of the tail toward 0.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos.tex}
-\caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds} \label{tab:coverage}
-\end{table}
-
-%
-%\subsubsection*{Reverse asymptotics}
-%
-%Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
-%
-%Note: we'd like to make sure it behaves ok in the case where most observations are null.
-%Perhaps simulate with 999 null and 1 alternative?
-%Also maybe look at performance for small sample sizes - e.g. n=1?
-%
-%Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
-%can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
-%by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
-%
-
-\subsection*{Differing measurement precision across units}
-
- We turn now to the second important component of our work: allowing for varying
- measurement precision across units. The key to this is the use of a likelihood,
- (\ref{eqn:normlik}) or (\ref{eqn:tlik}), that explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
-
-  To illustrate, we conduct a simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
- ($s_j=10$).  In both cases,
-we assume that half the effects are null
-and the other half are normally distributed with standard deviation 1:
-\begin{equation}
-p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1).
-\end{equation}
-In this setting, the poor-precision measurements $(s_j=10)$ tell us very little, and any sane analysis should effectively ignore them. 
- However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor}). This is because the poor measurements 
- produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
- which, when combined with the good-precision measurements, dilute the overall signal (e.g. they reduce the density of $p$ values near 0).
- This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
-the estimated error rates ($q$-values, or $\lfdr$ values) for the good-precision observations increase when the low-precision observations are included in the analysis
-(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
-for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).
-
-\begin{figure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_hist.pdf} 
-    \label{fig:goodpoor-hist}
-    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations}
- \end{subfigure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_scatter.pdf}
-    \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) of a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
-\end{subfigure}
-\caption{Simulation illustrating how, for existing FDR methods, 
-poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
-$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
-a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
-In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
-\end{figure}
-
- \subsubsection*{Reordering of significance, and the ``$p$ value prior"}
-
-Another consequence of accounting for differences in measurement precision is that \ashr may re-order the significance of the observations compared
- with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
- We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measuments that have the same $p$ value.
- The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
- and so the $\lfsr$ for these poor-precision measurements is always high.
-
-The potential for Bayesian analyses to re-order the significance of observations, and specifically to downweight imprecise observations,
-was previously discussed in \cite{guan.stephens.08}.
-However, Wakefield \cite{wakefield:2009} showed that, under a certain prior assumption, the Bayesian analysis produces the same ranking of significance as $p$ values (or their $z$ scores).
-He named this prior the ``$p$-value prior" because it can be thought of as the
-implicit prior assumption that is being made if we rank the significance of observations by their $p$ value.
-Wakefield's $p$-value prior assumes that the less precise effect estimates correspond to larger true effects, and specifically that they scale proportional
-to the standard errors $s_j$. More specifically still, it assumes a normal prior for the non-zero $\beta_j$ with mean 0 and variance $K s^2_j$ for some constant $K$.
-Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. Specifically, if, instead of assuming that $\beta_j$ is independent
-of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
-then the $\lfsr$ computed by ash.n provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
-as is illustrated in Figure \ref{fig:lfsr_pval}, right panel. (This result does not hold when using the mixtures of uniforms prior, ash.u.)
-
-This $p$-value prior assumes that the $z$ scores $z_j=\beta_j/s_j$
-are identically distributed, independent of $s_j$. This is essentially the assumption made, implicitly or explicitly, by existing methods -- like \locfdr, \mixfdr and \qvalue --
-that model the $z_j$ or $p_j$ directly.  In contrast, we have assumed up to now that the $\beta_j$ are independent of $s_j$. 
-We can set both these assumptions within a more general framework, which allows that $\beta_j/s^\alpha$ is independent of $s_j$ for some $\alpha$ [Equation (\ref{eqn:beta-alpha}))].
-Setting $\alpha=0$ implies that $\beta_j$ is
-independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
-effects (in absolute value). This latter assumption may often be qualitatively plausible: 
-for example, in gene expression studies the standard error for gene $j$ depends partly on
-the variance of its expression among samples, and genes with a larger variance may tend to be less tightly regulated and
-so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
-On the other hand, there is no particular reason to expect that either $\alpha=1$ or $\alpha=0$ will be the optimal choice.
-Indeed, optimal choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$,
-which will be dataset-specific. 
-Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
-comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
-likelihood-based methods can be used to select $\alpha$. For example, following the logic of the EB approach it
-would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
-which has been implemented in our software by C. Dai.
-
-%\begin{result}
-%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
-%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
-%\end{result}
-
-%To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
-%Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
-%In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
-%Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
-%which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
-%(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
-%may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)
-
-%\begin{result}
-%Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and using a mixture of normals for the prior distribution 
-%(\ref{eqn:beta-alpha}) with $\alpha=1$, 
-%the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$.
-Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$)
-%is montonic with $|t_j|$. 
-%\end{result}
-
-%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
-%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.
-
-%Most previous authors of EB approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
-%have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
-%In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
-%This is analogous
-%to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
-%standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
-%that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.
-
-\begin{figure}
-\centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
-\caption{Figure illustrating affects of prior assumptions on re-ordering of significance.
-Left panel shows results under our ``default prior" which assumes that effects $\beta_j$ are identically distributed, independent of $s_j$.
-Right panel shows results under the ``$p$-value prior", which assumes that $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$.} \label{fig:lfsr_pval}
-\end{figure}
-
-%\subsubsection*{Adaptive Shrinkage}
-%
-%Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
-%when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
-%the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
-% $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
-% In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
-% Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
-% and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
-%
-%Our methods we build on this idea in two ways. First,
-%by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
-%Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
-%adapts to measurement precision: more precise measurements undergo
-% less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
-%than observations with
-%small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
-%To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
-%Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
-%
-%\begin{figure}[!ht]
-%\begin{center}
-%\includegraphics[width=400px]{../examples/ash.pdf} 
-%\end{center}
-%\caption{Figure illustrating adaptive shrinkage.
-Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0.
-The standard error of each observation, $s_j$, was simulated from Inverse Gamma(5,5), which resulted in $s_j$ varying from 0.34 to 6.7.
-The shrunken estimates ($y$ axis) are plotted against the observed estimates ($x$ axis), with color indicating the (square root of the) precision of each measurement (blue=lower-precision; red=higher-precision).
-The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
-%\end{figure}
-
-%DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
-%To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
-Because the prior distribution is unimodal about zero,
-%the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
-%amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
-%For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
-%than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
-%small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
-
-%previous work based on normal means:
-%http://link.springer.com/article/10.1007/BF02562623
-
-%\subsection{Null behaviour}
-
-%Results of simulations under the null.
-
-%\subsection{Comparison of different underlying component distributions}
-%
-%Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
-%than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
-%likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
-%For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
-%increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
-%expect $L_{hu}$ to be significantly larger than the others.
-%
-%Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
-%
-%\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
-%\begin{center}
-%\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
-%%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
-%\end{center}
-%\caption{}
-%\end{figure}
-
-%Notes from Adrian Raftery - 
-%Newcombe (1886)
-%Dean and Raftery
-%Uniform is super-efficient
-
-\section*{Discussion}
-
-We have presented an Empirical Bayes approach to large-scale multiple testing that emphasises two ideas.
-First, we emphasise the potential benefits of using two numbers ($\bhat$, and its standard error)
-rather than just one number (a $p$ value or $z$ score)  to summarize the information on each test.
-While requiring two numbers is slightly more onorous than requiring one, in many settings these
-numbers are easily available and if so we argue it makes sense to use them.
-Second, we note the potential benefits -- both statistical and computational -- of assuming that the effects come from a unimodal distribution, and provide
-flexible implementations for performing inference under this assumption. We also introduce the ``false sign rate" as
-an alternative measure of error to the FDR, and illustrate its improved robustness to errors in model fit, particularly mis-estimation
-of the proportion of null tests, $\pi_0$.
-
-Multiple testing is often referred to as a ``problem" or a ``burden". In our opinion, EB approaches turn this idea on its head,
-treating multiple testing as an {\it opportunity}: specifically, an opportunity to learn about the prior distributions, and other modelling assumptions,
-to improve inference and make informed decisions about significance (see also \cite{greenland1991empirical}). This view
-also emphasises that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
-Indeed, the FDR at a given fixed threshold does not depend on the number of tests: as the number of tests increases, both the true positives and false positives increase linearly,
-and the FDR remains the same.
-(If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example.
-The EB approach captures this dependence in an intuitive way:
-if there are lots of strong signals then we infer $\pi_0$ to be small, and the estimated FDR (or $\lfdr$, or $\lfsr$) at a given threshold may be low, even if a large number of tests were performed; and conversely
-if there are no strong signals then we infer $\pi_0$ to be large and the FDR at the same threshold may be high, even if relatively few tests were performed. 
-More generally, overall signal strength is reflected in the estimated $g$, which in turn influences the estimated FDR.
-
-Two important practical issues that we have not addressed here are correlations among tests,
-and the potential for deviations from the theoretical null distributions of test statistics. These two
-issues are connected: specifically, unmeasured confounding factors can cause both correlations among tests
-and deviations from the theoretical null \cite{efron2007correlation,leek:2007}. And although there are certainly other
-factors that could cause dependence among tests, unmeasured confounders are perhaps the most worrisome in practice
-because they can induce strong correlations among large numbers of tests and profoundly impact results.
-Approaches to deal with unmeasured confounders can be largely divided into two types: those
-that simply attempt to correct for the resulting inflation of test statistics \cite{devlin1999genomic,efron2004large}, and those
-that attempt to infer counfounders using clustering, principal components analysis, or factor models 
-\cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
-and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable,
-they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alterntively, our methods
-could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:lik}) using
-$\bhat_j | \shat_j \sim N(0,\lambda_1 \shat_j + \lambda_2)$, where $\lambda_1,\lambda_2$ are to be estimated. However,
-this immediately raises issues of identifiability and we do not pursue the idea further here.
-
-Another important practical issue is the challenge of small sample sizes. For example, in genomics applications researchers sometimes
-attempt to identify differences between two conditions based on only a handful of samples in each. In such settings the
-normal likelihood approximation (\ref{eqn:normlik}) will be inadequate. And, although the $t$ likelihood (\ref{eqn:tlik}) partially
-addresses this issue, it is also, it turns out, not entirely satisfactory.  The root of the problem is that, with small sample sizes, raw estimated standard
-errors $\shat_j$ can be horribly variable. In genomics it is routine to address this issue by applying EB methods \cite{smyth:2004} to ``moderate" (i.e.~shrink) variance estimates,
-before computing $p$ values from ``moderated" test statistics. We are currently investigating how
-our methods should incorporate such ``moderated" variance estimates to make it applicable to small sample settings.
-
-Our approach involves compromises between flexibility, generality, and simplicity on the one hand, and statistical efficiency and principle on the other.
-For example, in using an EB approach that uses a point estimate for $g$, rather than a fully Bayes approach that accounts for uncertainty in $g$,
-we have opted for simplicity over statistical principle. And in summarizing every test by
-two numbers and making a normal or $t$ approximation to the likelihood, we have aimed to produce generic methods that 
- can be applied whenever such summary data are available -- just as \qvalue can be applied to any set of $p$ values for example -- although
- possibly at the expense of statistical efficiency compared with developing multiple tailored approaches based on context-specific likelihoods. 
- Any attempt to produce generic methods will involve compromise between generality and efficiency.
- In genomics, many analyses -- not only FDR-based analyses -- involve first computing a series of $p$ values before
- subjecting them to some further downstream analysis. An important message here is 
- that working with two numbers ($\bhat_j,\shat_j$), rather than one ($p_j$ or $z_j$), 
- can yield substantial gains in functionality (e.g.~ estimating effect sizes, as well as testing; accounting for variations in
-measurement precision across units) while losing only a little in generality.  We hope that our work will
-encourate development of methods that exploit this idea in other contexts.
-
-%This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
-%from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
-%with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
-
-%
-%
-%Both \locfdr and \mixfdr allow for deviations from
-%the theoretical null distribution by allowing that, under the null, the $z$ scores are $N(0,\sigma^2)$ with $\sigma$ estimated from the data rather
-%than fixed to 1. Within our approach this idea could be incorporated by changing the likelihood - for example, to assume $\bhat_j \sim N(0,s_ 
-%
-%These are tricky issues that, as far as we are aware, can affect
-%all approaches to FDR analysis. An important focus of Efron's work, also implemented in \locfdr, 
-% is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
-%theoretical null distribution. 
-%
-%For example, we and others have worked on Bayesian approaches to genetic association testing, using subjective prior specification of
-%effect sizes based on external data (e.g.~\cite{stephens:2009}). However, the large numbers of tests brings the opportunity to learn about
-%priors, and here we aim to exploit this
-
-%Choice of mixture components. Use of likelihood to compare different mixture components?
-
-%  
-%Directly modeling the p values, or z scores, say via non-parametric methods, 
-%can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-%distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-%The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-%Before proceeding, we feel obliged to correct the common misconception
-% that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
-% not depend on the number of tests. 
-%  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
-% 
-% In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
-
-\section*{Detailed Methods} \label{sec:implement}
-
-\subsection*{Embellishments}
-
-\subsubsection*{More flexible unimodal distributions}
-
-Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
-Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
- Therefore, we have implemented a more general approach based on 
-\begin{equation} \label{eqn:g-general}
-g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) ,
-\end{equation}
-where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
-\begin{enumerate}[(i)]
-\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$, \qquad (``ash.n")
-\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,  \qquad (``ash.u")
-\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,  \qquad (``ash.hu")
-\end{enumerate}
-where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
-(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
-and $\pi$ is a $2K+1$ vector that sums to 1.)
-The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
-Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
-under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
-(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
-(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
-however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)
-
-Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
-using a large dense grid of $\sigma^2_k$ or $a_k$, (i)-(iii) can respectively approximate,
-with arbitrary accuracy,
-\begin{enumerate}[(i)]
-\item any scale mixture of normals, which includes as special cases
-the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
-\item any symmetric unimodal distribution about 0.
-\item any unimodal distribution about 0.
-\end{enumerate}
-The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
-In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.
-
-Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
-The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
-See Implementation for details of our software defaults.
-
-%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting.
-For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
-%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
-
-\subsubsection*{Dependence of effects on standard errors}
-
-Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
-This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
- \begin{equation} \label{eqn:beta-alpha}
- \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
- \end{equation}
-for any $\alpha$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
-assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
-it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
-the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.
-
-The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
-To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
- standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
- Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
-yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
-%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
-%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
-%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
-
-%Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
-%how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
-%it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
-%rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
-%and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
-%this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
-%do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
-
-\subsubsection*{Replace normal likelihood with $t$ likelihood}
-
-We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
-  \begin{equation} \label{eqn:tlik}
- \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
- \end{equation}
-where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
-and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
-The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
-This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms; see Implementation below.
-When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
-
-Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
-\begin{equation} \label{eqn:tdist}
-(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
-\end{equation}
-However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
-In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
-currently pursuing this improved approach (joint work with M.Lu) and results will be published elsewhere. 
-%We note here that
-%modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
-%where our simpler approach can produce unsatisfactory results.
-
-\subsubsection*{Non-zero mode}
-
-An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood, rather than fixed to 0.
-
-\subsection*{Implementation Details}
-
-\subsubsection*{Likelihood for $\pi$}
-
-We define the likelihood for $\pi$ to be the probability of the observed data $\bhat$ conditional on $\shat$: $L(\pi) := p(\bhat | \shat,\pi),$ which
-by our conditional indendence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$.
-[One might prefer to define the likelihood as $p(\bhat, \shat | \pi) = p(\bhat | \shat, \pi) p(\shat | \pi)$, in which case our definition comes down to assuming that the term $p(\shat | \pi)$ does not depend on $\pi$.]
-
-Using the prior $\beta_j \sim \sum_{k=0}^K \pi_k f_k(\beta_j)$ given by (\ref{eqn:g-general}), and the normal likelihood (\ref{eqn:normlik}), integrating over $\beta_j$ yields
-\begin{equation}
-p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
-\end{equation}
-where 
-\begin{equation}
-\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
-\end{equation}
-denotes the convolution of $f_k$ with a normal density.
-These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
-Specifically, 
-\begin{equation} \label{eqn:uconv}
-\tilde{f}_k(\bhat_j)  = 
-\begin{cases}
-N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
-\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
-\end{cases}
-\end{equation}
-where $\Psi$ denotes the cumulative distribution function (c.d.f.) of the standard normal distribution.
-If we replace the normal likelihood with the $t_\nu$ likelihood (\ref{eqn:tlik}) then the convolution 
-for $f_k$ uniform the convolution is still given by (\ref{eqn:uconv}) but with $\Psi$ the c.d.f. 
-of the $t_\nu$ distribution function. (The convolution for $f_k$ normal is tricky and we have not implemented it.)
-
-\subsubsection*{Penalty term on $\pi$}
-
-To make $\lfdr$ and $\lfsr$ estimates from our method ``conservative" we add a penalty 
-term $log(h(\pi;\lambda))$ to the log-likelihood $\log L(\pi)$ to encourage over-estimation of $\pi_0$:
-\begin{equation} \label{eqn:penalty}
-h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
-\end{equation}
-where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
-consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0sims}).
-
-Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
-we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
-
-\subsubsection*{Problems with removing the penalty term in the half-uniform case}
-
-It is straightforward to remove the penalty term by setting $\lambda_k=1$ in (\ref{eqn:penalty}).
-We note here an unanticipated problem we came across when using no penalty term in the half-uniform case 
-(i.e.~ $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$ in (\ref{eqn:g-general})): when the data
-are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
-sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
-for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
-and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
-
-\subsubsection*{EM algorithm}
-
-With this in place, the penalized log-likelihood for $\pi$ is given by:
-\begin{equation}
-\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
-\end{equation}
-where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This can be maximized using an EM algorithm \cite{dempster77}, whose
-one-step updates are:
-\begin{align} \label{eqn:w}
-w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
-n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
-\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}
-\end{align}
-Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
-and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
-probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$. (The $w_{kj}$ are sometimes referred to as the ``responsibilities".)
-Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
-We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple} to accelerate convergence of this EM algorithm.
-
-%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
-\subsubsection*{Initialization}
-
-By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
-Our rationale for initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
-but weak signal in the data cannot quickly draw the algorithm towards the null.
-
-In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
-$\pi_0=1; \pi_k=0 (k>0)$. If not then we replace the EM solution with the global null solution. This aims to guard against errors due to convergence to a local optimum when the
-data are consistent with the global null.
-
-\subsubsection*{Conditional distributions}
-
-Given $\hat\pi$, we compute the conditional distributions 
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
-\end{equation} 
-Each posterior is a mixture on $K+1$ components:
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
-\end{equation}
-where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
-and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
-$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
-All these posterior distributions are easily available.
-For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
-If $f_k$ is normal and $L$ is normal, then this is a normal distribution.
-
-\subsubsection*{Choice of grid for $\sigma_k, a_k$} \label{sec:grid}
-
-\def\sigmamax{\sigma_{\text{max}}}
-\def\sigmamin{\sigma_{\text{min}}}
-When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
-in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
-as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
-based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
-is negative we set $\sigmamax = 8\sigmamin$.)
-
-When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.
-
-Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
-a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.
-
-\appendix
-
-\begin{table}[!ht]
-\centering\begin{tabular}{c c } \toprule
-Scenario & Alternative distribution, $g_1$  \\ \midrule
-spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
-near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
-flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
- &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
-skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
-big-normal & $N(0,4^2)$ \\ 
-bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
-\end{tabular}
-\caption{Summary of simulation scenarios considered} \label{table:scenarios}
-\end{table}
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all_nopen.tex}
-	\caption{All observations}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg_nopen.tex}
-\caption{``Significant" negative discoveries.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos_nopen.tex}
-\caption{``Significant" positive discoveries.}
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{table:nopen}
-\end{table}
-%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
-%\begin{equation}
-%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
-%\end{equation}
-%which is close to the definition in \cite{storey.03}.
-
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgements}
-
-Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
-prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} , and includes contributions 
-from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.
-
-This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.
-
-%\section*{References}
-% The bibtex filename
-\bibliography{/Users/stephens/Dropbox/Documents/mainbib}
-
-\section*{Figure Legends}
-
-\clearpage
-
-\section*{Tables}
-
-\section*{Supporting Information Legends}
-
-Supplementary material can be found in {\bf Supplementary Information S1.}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
-\end{document}

### 2015-10-06 sectionResults__We_c.tex
New Article!

--- 2015-10-06 paper/main.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
-\documentclass[11pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-%\usepackage{natbib}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-\usepackage{booktabs}
-\usepackage{xspace}
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-\usepackage{sectsty}
-\subsectionfont{\normalsize}
-\subsubsectionfont{\mdseries\itshape}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\usepackage{color}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-% Use the PLoS provided bibtex style
-\bibliographystyle{/Users/stephens/Dropbox/Documents/stylefiles/plos2009}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
-\pagestyle{myheadings}
-%% ** EDIT HERE **
-\usepackage{enumerate}
-\usepackage{multirow} 
-\usepackage{url}
-\usepackage{xr} %for cross-referencing
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-\newtheorem{algorithm}{Algorithm}
-\newtheorem{proposition}{Proposition}
-\newtheorem{restateproposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{result}{Result}
-\newtheorem{note}{Note}
-\newtheorem{definition}{Definition}
-
-\def\lfdr{\textit{lfdr}}
-\def\lfsr{\textit{lfsr}}
-\def\x{\mbf{x}}
-\def\y{\mbf{y}}
-\def\e{\mbf{e}}
-\def\g{\mbf{g}}
-\def\p{p}
-\def\Yb{\hat{Y}_1^{\text{BAYES}}}
-\def\Xperp{X_{1 \perp 0}}
-\def\Pperp{P_{1 \perp 0}}
-\def\Pperpb{\Pperp^{\text{B}}}
-\def \hatd { \widehat{ d } }
-\def \dhatin {\hatd^{\mbox{\tiny IN}}}
-\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
-\def \mvn{\text{MVN}}
-\def\mn{\text{MN}}
-\def \iw{\text{W}^{-1}}
-\def \fp {fastPHASE\ }
-\def\bhat{\hat{\beta}}
-\def\shat{\hat{s}}
-\def\var{\rm Var}
-\def\mvr{\text{BMVR}}
-\def\bfuni{\text{BF}_\text{uni}}
-\def\one{\mbf{1}}
-\def\zero{\mbf{0}}
-\def\hall{H_{\text{all}}}
-\def\pall{\p_{\text{all}}}
-\def\bfall{\text{BF}_\text{all}}
-\def\bfav{\text{BF}_\text{av}}
-\def\gall{\gamma_\text{all}}
-\def\limbfall{\bfall^\rightarrow}
-\def\limbfav{\bfav^\rightarrow}
-\def\limbfgamma{\BF_\gamma^\rightarrow}
-\def\sxx{S_{xx}}
-\def\syy{S_{yy}}
-\def\vyy{V_{yy}}
-\def\vyx{V_{yx}}
-\def\vxx{{V_{xx}}}
-\def\rss{\text{\rm RSS}}
-\def\dss{\text{\rm $\Delta$SS}}
-\def\U{U}
-\def\D{D}
-\def\I{I}
-\def\luu{L_{UU}}
-\def\indep{\perp}
-\def\qvalue{{\tt qvalue}\xspace}
-\def\locfdr{{\tt locfdr}\xspace}
-\def\mixfdr{{\tt mixfdr}\xspace}
-\def\ashr{{\tt ashr}\xspace}
-
-\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
-                                % mode too
-\newcommand\height{{\it height}}
-\newcommand\weight{{\it weight}}
-\def\BF{\text{BF}}
-
-%% END MACROS SECTION
-%\externaldocument{plos_one_rev1_suppinfo}
-
-\begin{document}
-
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{False Discovery Rates: A New Deal}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Matthew Stephens$^{1*}$, 
-\\
-\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
-\\
-$\ast$ E-mail: Corresponding mstephens@uchicago.edu
-\end{flushleft}
-
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-%\section*{Author Summary}
-
-\section*{Introduction}
-
-Since its introduction in 1995 by Benjamini and Hochberg 
-\cite{benjamini1995controlling}, the ``False Discovery Rate" (FDR) has quickly established itself
-as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing
-in which the goal is to identify the non-zero ``effects" among a large number of imprecisely-measured effects.
-
-Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
-indeed, the notion that EB approaches could be helpful in handling multiple comparisons predates introduction of
-the FDR  (e.g.~\cite{greenland1991empirical}). More recently, EB approaches to the FDR
-have been extensively studied by several authors, 
-especially B.~Efron and co-authors 
-\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
-see also \cite{kendziorski2003parametric,newton2004detecting, 
-datta2005empirical,muralidharan2010empirical} for example.
-So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
-existing widely-used FDR pipelines) and can substantially affect
-inference. The first idea is to {\it assume that the distribution of effects is unimodal}. This yields a very simple, fast, and stable computer 
-implementation, as well as improving inference of FDR when the unimodal assumption is correct.  
-The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
-to summarize each measurement. This idea allows variations in measurement precision to be better accounted for,
- and avoids a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
-
-In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
-the idea that it may be preferable to focus on estimation rather than on testing.
-In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
-simply estimation with some positive prior probability that the effect is exactly zero.
-However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
-perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
-controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
-``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
-in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
-the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the lfsr, particularly its increased robustness to modelling assumptions.
-
-%terminology
-%``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
-%and correspondingly define the ``false signed discovery rate", or simply ``false sign rate" (FSR), in a natural way.
-
-%The methods we provide here can address both problems, and we provide results that highlight important benefits of focussing on estimation, and particularly on assessing the correctness of the {\it sign} of the estimated effects, rather than on testing for non-zero effects.
-
-Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
-(We address the reasons for this name, and connections with shrinkage analysis, in the Discussion.)
-
-%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.
-
-%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-%measurements are made on only a small number of
-%samples in each condition - sometimes as few as one
-%sample in each condition. Thus the error in estimates
-%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-%in estimates of $\beta_j$ still greater.
-%
-
-%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
-%Here we present new tools that bring two important ideas to bear on the problem.
-%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
-%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
-%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
-%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
-%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
-%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
-%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
-%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
-%a substantial increase in the number of discoveries at a given FDR).
-
-%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
-%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
-Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
-%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
-%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
-%generic methods that work directly with two numbers for each 
-%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
-Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
-%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
-
-\def\df{df}
-\def\FDR{\text{FDR}}
-\def\fdr{\text{\lfdr}}
-\def\FSR{\text{FSR}}
-\def\fsr{\text{lfsr}}
-
- \section*{Methods}
-
- \subsection*{Model Outine}
-
-% Here we consider one setting in which FDR-based methods are widely used:
-%an experimenter has measured a large numbers of ``effects", each with an associated
-%measurement precision (standard error), 
-%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
-%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
-%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
-%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
-%For example, in genomics
-%a common goal is to identify ``differentially-expressed genes" -- that is, genes
-%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
-%mean expression level between the two conditions, often referred to as the ``log-fold change". 
-Here we describe the simplest version of our method, and briefly discuss embellishments we have also implemented.
-
-Suppose that we are interested in the values of $J$ ``effects" $\beta=(\beta_1,\dots,\beta_J)$. For example, in a typical genomics application
-that aims to identify differentially expressed genes,  $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
-In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
-that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
-more general problem of estimating, and assessing uncertainty in, $\beta_j$.
-
-Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
-and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
-which by Bayes theorem can be written as 
-\begin{equation}
-p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
-\end{equation}
-
-For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
-This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
-A simple way to implement the UA is to assume that $g$ is
-a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
-\begin{align} \label{eqn:beta}
-p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
-g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
-\end{align}
-where $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyperparameters, which are non-negative and sum to one, and are to be estimated,
-while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
-represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
-(We encourate the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
-although of course in practice we use a finite grid -- see Implementation Details.)
-
-For the likelihood $p(\bhat | \beta, \shat)$ we assume  
-\begin{equation} \label{eqn:normlik}
-p(\bhat | \beta, \shat) = \prod_j N(\bhat_j; \beta_j, \shat_j^2).
-\end{equation}
-Here, in additional to some conditional independence assumptions, we are 
-effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-
-This simple model features both the key ideas we want to emphasise in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
-the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
-with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
-However, this simple model also has several additional assumptions that can be relaxed. Specifically,
-\begin{enumerate}
-\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) also implies that $g$ is symmetric about 0; more flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see (\ref{eqn:g-general})].
-\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
-We can relax this to allow for a relationship between these quantities [see (\ref{eqn:beta-alpha})].
-\item The likelihood (\ref{eqn:normlik}) effectively assumes that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-We can generalize this likelihood using a $t$ likelihood [see (\ref{eqn:tlik})].
-\end{enumerate}
-These embellishments are detailed in Embellishments and Implementation Details.
-Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
-(which are also made by most existing EB approaches to this problem). Correlations among tests certainly arise in practice, either due to genuine correlations
-in the system of study, or due to unmeasured confounders, and their potential to impact results of an FDR analysis is important to consider
-whatever analysis methods are used: see \cite{efron2007correlation,leek:2007} for relevant discussion.
-
-%We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
-%$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
-%\begin{equation}
-%p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
-%\end{equation}
-%Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
-%\begin{equation} \label{eqn:normlik}
-%p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
-%
-%We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
-%as a mixture of null and alternative components:
-%\begin{equation} \label{eqn:betaj0}
-%g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
-%\end{equation}
-%where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
-%We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
-%\begin{equation} \label{eqn:beta-iid}
-%p(\beta | \shat) = \prod_j g(\beta_j).
-%\end{equation}
-%
-%Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
-%with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
-%to use a mixture of zero-mean normal distributions:
-%\begin{equation} \label{eqn:mixnorm}
-%g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
-%Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
-%represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
-% The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
-% more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below. 
-% 
-% Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
-% yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
-% \begin{equation} \label{eqn:g}
-% g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
-% \end{equation}
-% 
-% Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
-% \begin{align} \label{eqn:betaj}
-% p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
-% p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
-% \end{align}
-% with $g$ given by the mixture of normals $(\ref{eqn:g})$.
-
-\subsection*{Fitting the model}
-
-In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
-each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
-Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
-\begin{equation} \label{eqn:marginallik}
-p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
-\end{equation}
-where we define $\sigma_0:=0$.
-
-The usual EB approach to fitting this model would involve two simple steps:
-\begin{enumerate}
-\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
-\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
-against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | | \bhat, \shat,\hat{\pi})$.
-\end{enumerate}
-Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(The simplicity of the EM algorithm in step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
-which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
-
-Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximise a penalized
-likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
-We introduce this penalty because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
-
-\subsection*{The local False Discovery Rate and local False Sign Rate}
-
-As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
-In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
-One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
-the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
-In other words it is the posterior probability that $\beta_j$ is actually zero:
-\begin{equation} \label{eqn:fdr}
-\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
-\end{equation}
-
-The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
-whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
-the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey1991philosophy}) argued that
-``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$.
-Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
-\begin{quote}
-...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
-\end{quote}
-Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
-focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.
-
-Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
-would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
-\begin{equation} \label{eqn:lfsr}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
-\end{equation}
-To illustrate, suppose that 
-\begin{gather*}
-p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
-p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
-p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
-\end{gather*} 
-Then 
-from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
-our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.
-
-As our notation suggests, $\lfsr_j$ is intended to be compared and contrasted with $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
-small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
-Of coure, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
-this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
-(this follows from the definition because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
-In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
-$\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.
-
-From these ``local" measures of significance, we can also compute average error rates over subsets of observations
-$\Gamma \subset \{1,\dots,J\}$. For example,   
-\begin{equation} \label{eqn:FDRhat}
-\widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
-\end{equation}
-estimates the FDR we would obtain if we were to declare all tests in $\Gamma$ significant.
-And 
-\begin{equation} \label{eqn:qvalue}
-q_j := \widehat \FDR(\{k: \fdr_k \leq \fdr_j\}) 
-\end{equation}
-provides a measure of significance analogous to Storey's $q$ value \cite{storey.03}.
-
-%Because our methods involve explicitly modelling effects $\beta_j$, rather than 
-%$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
-
-%Indeed, even in settings where some
-%$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
-%analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
-
-\subsection*{Related work}
-
-\subsubsection*{Previous approaches focussed on FDR}
-
-Among previous methods that explicitly consider the FDR and related quantities,
-our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
- (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
-  %Here we highlight the key differences between our approach and these three existing approaches.
-
-There are two key differences between our approach and all of these three existing methods. First, 
-whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
-we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
-for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
-Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 
-how it can improve performance by taking better account of variation in measurement precision among observations.
-%(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
-% Or more generally Section xx illustrates the potential practical effects
-
-%allows for different modelling assumptions to be compared through likelihood comparisons.
-Second, our unimodal assumption (UA) that the effects are unimodal about zero is quite different from assumptions made by $\qvalue, \locfdr$ or $\mixfdr$.
-Indeed, $\locfdr$ assumes that all $z$ scores near 0 are null (Efron calls this the Zero Assumption; ZA), which implies that 
-under the alternative hypothesis the distribution of $z$ scores has {\it no mass at 0}; this contrasts strikingly with the UA, which implies that this distribution has its peak at 0! 
-Similarly, $\qvalue$ assumes that all $p$ values near 1 are null,
-which is the same as the ZA because $p$ values near 1 correspond to $z$ scores near 0. And although \mixfdr does not formally make the ZA,
-we have found that in practice, with default settings, the results often approximately satisfy the ZA (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}).
-Thus, not only do these existing methods not make the UA, they actually make assumptions that are, in some sense, as different
-from the UA as they can be.
-
-Given that the UA and ZA are so different, it seems worth discussing why we generally favor the UA. 
-Although the UA will not apply to all situations, we believe that it will often be reasonable, especially in FDR-related contexts
-that have traditionally focussed on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis,  
-it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible. Further, it seems reasonable to expect that larger effects become
-decreasingly plausible, and so the distribution of the effects will be unimodal about 0.
-To paraphrase Tukey, ``All we know about the world teaches us that large effects are rare, whereas small effects abound."
-We emphasise that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
-It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
-
-%Tukey "peri-null"
-
-%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
-%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}.
-
-In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficents - 
-common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
-These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
-it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
-The UA assumption on regression coefficients is directly analagous to our UA here, and so its
-widespread use in the regression context supports its use here.
-
-Alternatively, we could motivate the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
-such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
-make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).
-
-Finally, the UA also has a considerable practical benefit: it yields simple algorithms that are both computationally
-and statistically stable. We illustrate these features in Results.
-
-\subsubsection*{Other work}
-
-There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
- Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
-indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
-This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
-Also, they provide no software implementation.
-More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
-Some pointers to recent papers in which other relevant
-citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}.
-Much of the literature focusses on the homoskedastic case (i.e.
-$\shat_j$ all equal) whereas we allow for heteroskedasticity.
-And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential.
-Several recent papers consider more flexible non-parametric assumptions on
-$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
-estimate (NPMLE) for $g$.
-These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive.
-However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
-may perform well for point estimation, it seems possible it will not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
-Indeed, one way of thinking about the UA is as a way to regularize $g$.
-
-%The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
-%encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-%$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-%$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-%the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-%observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-%corresondingly little effect on the estimated FDR of other observations. 
-%We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-%$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-%that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-%estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-%Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-\def\dt{\text{dt}}
-
-%
-%assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
-%$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
-% $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
-%  \begin{equation} \label{eqn:lik}
-% \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
-% \end{equation}
-% where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e.
-the density of $\mu+\sigma T_\nu$).
-In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
-%  \begin{equation} \label{eqn:normlik}
-% \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
-% \end{equation}
-%which, in light of (\ref{eqn:betahat}), is equivalent to
- % \begin{equation} \label{eqn:lik2}
-% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
- %\end{equation}
-
-%The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
-they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
-The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$.
-Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$.
-If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$.
-See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.
-
-%This likelihood approximation can be motivated by assuming that i) $\bhat$
-% $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
-% $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$. 
-%
-
-%Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.
-
-%While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
- %that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
-%there is an important difference: here,
-
-%The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
-For example, we can write $p(\bhat_j, \shat_j | \beta_j) = p(\bhat_j | \shat_j, \beta_j)p(\shat_j | \beta_j)$.
-Now, if we are willing to assume that $\shat_j$ alone contains no information about $\beta_j$, or equivalently that $p(\shat_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\bhat_j | \shat_j, \beta_j)$, and if $\bhat_j | \shat_j, \beta_j \sim N(\beta_j, \shat_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.
-
 \section*{Results}
@@
 To briefly summarize the results in this section:
+
 \begin{enumerate}
@@
 \subsubsection*{The UA and ZA can produce different inferences}
-
-\begin{figure}
-\center \includegraphics[height=6in]{figures/decomp_ZA.pdf}
-\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
-The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
-In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
-The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
-(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
-\end{figure}
-
-To illustrate the different inferences from the UA and ZA we show results for a single
-dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
-but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
-\qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
-The results (Figure \ref{fig:ZA}) illustrate the clear difference between the existing methods and our method. The effects of the ZA made by \qvalue and \locfdr
-are visually clear, producing a ``hole"  in the alternative $z$ score distribution around 0. Although \mixfdr does not formally make the ZA, its decomposition
-exhibits a similar hole. In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
-effectively ``filling in" the hole.  (Of course the null distribution also has a peak at 0, and the local fdr 
- under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)
-
-Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA.
-Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
-This would increase the null component (dark blue) at the expense of
-the alternative component (light blue).
-Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0.
-The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears.
-In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
-See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
-demonstration.
-
- \subsubsection*{The UA can produce conservative estimates of $\pi_0$}
-
-The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
-Consequently \ashr will estimate smaller $\lfdr$s and FDRs than existing methods that make the ZA. 
-This is desirable, provided that these estimates remain conservative: that is,
-provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
-The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behaviour. To check its effectiveness
-we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
-for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
-They range from a ``spiky" distribution -- where many non-zero $\beta$ are
-too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
-flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
-making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
-which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.
-
-% How do these differences influence estimates of the FDR? To address this, note that  -- in this simple case where the 
-% $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
-% of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
-% strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
-% although the methods considered here differ from one another in many ways, having estimated $\pi_0$
-% they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
-% empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
-%
-
-%Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
-% and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
-%Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
-%At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
-%overestimates of $\pi_0$, which will yield overestimates of the FDR).
-
-\begin{figure} 
-\begin{center}
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=1.8in]{figures/scenarios_density.pdf} 
-	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/pi0_est.pdf} 
-\caption{Comparison of true and estimated values of $\pi_0$.
-When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ashr being least conservative, and hence most accurate.
-When the UA does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.} \label{fig:pi0sims}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfdr_est.png} 
-\caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
-Black line is $y=x$ and red line is $y=2x$.
-Estimates of $\lfdr$ are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est.png} 
-\caption{As in c), but for $\lfsr$ instead of $\lfdr$.
-Estimates of $\lfsr$ are consistently less conservative than $\lfdr$ when UA holds, and also less anti-conservative in bimodal scenario.}  \label{fig:lfsr}
-\end{subfigure}
-\end{center}
-\caption{Results of simulation studies (constant precision $s_j=1$).}
-\end{figure}
-
-For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
-data as follows:
-\begin{enumerate}
-\item Simulate $\pi_0  \sim U[0,1]$.
-\item For $j=1,\dots,J$, simulate $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)$.
-\item For $j=1,\dots,J$, simulate $\bhat_j | \beta_j \sim N(\beta_j,1)$.
-\end{enumerate}
-%Thus these simulations assume the same
-%precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).
-
-Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ashr ($y$ axis) with the true values ($x$ axis). 
-For \ashr we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
-and as a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
-as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
-The results show that \ashr provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
-in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ashr estimates can be slightly anti-conservative.
-We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
-are used. (In addition, the effects on $\lfsr$ estimates turn out to be relatively modest; see below).
-
-%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
-%
-%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
-%and so all effects are non-zero and no nulls hold. This scenario is rather different
-%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
-%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
-%``mostly non-null"
-%scenario, and we would agree! On the other hand, the distribution of $p$ values
-%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
-%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
-%Thus it seems desirable to have methods that perform well in both cases, and this is
-%an important feature of the methods we introduce here.
-%
-%TODO: put FDRs on graph? Simulate with 50\% null?
-
-\subsubsection*{The $\lfsr$ is more robust than \lfdr}
-
-The results above show that \ashr can improve on existing methods in producing smaller, more accurate, 
-estimates of $\pi_0$, which will lead to more accurate estimates of FDR.
-Nonetheless, in many scenarios \ashr continues
-to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
-This is because these scenarios include 
-an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
-estimation of $\pi_0$ impossible. Put another way, and as is well known, $\pi_0$ is not identifiable:
- the data can effectively provide an upper bound on plausible values of $\pi_0$,
-but not a lower bound (because the data cannot rule out that everything is non-null, but with miniscule effects).
-To obtain conservative behaviour we must estimate $\pi_0$ by this upper bound, which can
-be substantially larger than the true value.
-
-Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
-overestimation of $\pi_0$ is corresponding overestimation of FDR (and $\lfdr$, and $q$ values).
-To illustrate, Figure \ref{fig:lfdr} compares the estimated $\lfdr$ from ash.n with the true value (computed using
-Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios which involve many
-non-zero effects that are very near 0 (e.g.~the spiky scenario with $\pi_0$ small) where
-$\pi_0$ can be grossly overestimated.  
-(Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.)
-
-The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
- become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
- This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
-To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
-be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
-the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in $\lfsr$ than $\lfdr$.
-
-Note that, compared with previous debates regarding testing, 
-this section advances an additional reason for focussing on the sign of the effect, rather than just testing whether it is 0. 
-In previous debates authors have argued against testing whether an effect is 0
-because it is {\it implausible that effects are exactly 0}. Here we add that {\it even if one believes
-that some effects may be exactly zero}, it is still better to focus on the sign, because generally {\it the data are more informative about that question}
-and so inferences are more robust to, say, the inevitable mis-estimation of $\pi_0$.
-To provide some intuition, consider an observation with a $z$ score of 0. The $\lfdr$ of this observation can range from 0 (if $\pi_0=0$)
-to 1 (if $\pi_0=1$). But, assuming a symmetric $g$, the $\lfsr>0.5$ whatever the value of $\pi_0$, because the observation $z=0$ says
-nothing about the sign of the effect.
-Thus, are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
-whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.
-
-Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
-Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
-if we are prepared to assume that no effects are exactly zero, then removing the point mass 
-yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
-However, there is ``no free lunch":  if in fact some effects are exactly zero
-then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
-We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.
-
-\begin{figure}
-\begin{center}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s_nn.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be more accurate in this case.
-That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
-\end{subfigure}
-\end{center}
-\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
-\end{figure}
-
-\subsubsection*{Numerical Stability}
-
-The EM algorithm, which we use here to fit our model, is notorious for convergence to local optima. 
-However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
-caused by such behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
-in addition running it using our default initialization procedure (see Implementation details). We then compared the largest
-log-likelihood achieved across all 11 runs with the log-likelihood achieved by the default run.
-When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference ($< 0.02$), 
-and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference ($<0.02$),
-and 6\% of runs showed an appreciable log-likelihood difference ($>1$), with the largest difference being 5.0. However, perhaps suprisingly, 
-even for this largest difference results from the default run (e.g.
-the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood.
-See \url{http://github.com/stephens999/ash/blob/master/dsc-robust/summarize_dsc_robust.rmd} for further details.
-
-\subsubsection*{The UA helps provide reliable estimates of $g$}
-
-An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
-can estimate the {\it size} of each effect $\beta_j$.
-Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
-to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
-threshold $T$.
-%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
-%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
-%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
-%towards zero, with the strength of the shrinkage being estimated from the data.
-Further, because the posterior distribution is, by definition,
-conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
-been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
-In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
-\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
-how can confidence intervals be assigned to the true
-effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
-and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
-%but also notes the difficulties of implementing a stable general algorithm.
-
-The ability of the EB approach to provide valid ``post-selection" interval estimates is extremely attractive in principle.
-But its usefulness in practice
-depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
-which are notoriously difficult in general. Indeed, Efron emphasises
-the difficulties of implementing a stable general algorithm, noting in his rejoinder
-``the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying
-to be overly nonparametric ... but it is hard to imagine a
-generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
-Our key point here is that the UA greatly simplifies the deconvolution problem.
-While not meeting Efron's desire for an entirely general nonparametric approach, we believe 
-that the UA can handle many cases of practical interest.
-
-To illustrate this, Figure \ref{fig:egcdf} compares the
-estimated $g$ from \ashr  with that from \mixfdr which does not make the UA 
-(and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
-The greater reliability of estimates afforded
-by the UA is immediately apparent. In particular the estimated cdf from \mixfdr often has an almost-vertical
-segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. 
-The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
-While the UA is not the only way to achieve this, we find it an attractive, simple and effective approach.
-
-Interestingly, even in the ``bimodal" scenario $\ashr$ 
-is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
-Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
-Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
-to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
-(\url{http://github.com/stephens999/ash/blob/master/dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
-clearly visible in the observed $\bhat$.
-
-We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ashr tends to systematically
-overestimate the mass of $g$ near zero. On careful inspection, this is apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
-and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf}b),
-and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
-under-estimating $\pi_0$ (Figure \ref{fig:egcdf}c).
-
-\begin{figure}[h!] 
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=2in]{figures/egcdf.pdf}
-	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf_nopen.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
-Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
-\end{subfigure}
-\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.} \label{fig:egcdf}
-\end{figure}
-
-%making the UA provides an attractive and simple way to achieve estimating $g$  large
-%subset of cases of interest (those where the UA can reasonably be assumed to hold).
-
-%In practice, the utility of the posterior distributions noted above depends primarily on the accuracy
-%of the estimate of the distribution $g(\cdot)$.
-
- %quote from efron rejoinder , p46:
-%MorrisÕ Section 3 is especially pertinent. His formula
-%for p(?|z) is related to my discussion of the
-%BenjaminiÐYekutieli False Coverage Rate method in
-%Section 7, particularly (7.2)Ð(7.4). Originally I had
-%hoped to develop an empirical Bayes method for estimating
-%such models, but the effort foundered on practical
-%difficulties involving the perils of deconvolution.
-%Section 6 on the Òone-group modelÓ is the ugly duckling
-%of the current paper, but it bears on some important
-%points raised in the discussion. Figure 7 concerns
-%a fuzzy version of simultaneous hypothesis testing,
-%where, as in MorrisÕ hospital example, the usual singlepoint
-%null hypotheses seem unequal to the task. The
-%development from (6.6) onwards, particularly (6.12),
-%bears on the possibility of nonnormal null distributions,
-%and is about as far as I can go in answering Professors
-%Rice and SpiegelhalterÕs penultimate question.
-%With g(?) a normal distribution, model (6.1) returns
-%us to the realm of Stein estimation, the scene of my
-%happy collaborations with Carl Morris. I continue to
-%be surprised at how much harder bumpy, nonnormal
-%models like (7.1) are to deal with. JamesÐStein estimation
-%works fine with, say, N = 10 component problems,
-%while the RobbinsÕ form of empirical Bayes appropriate
-%to (7.1) seems to require hundreds or thousands.
-%The information calculations in Efron (2008)
-%reinforce this gloomy assessment. Maybe I am trying
-%to be overly nonparametric in constructing the empirical
-%Bayes Fdr estimates, but it is hard to imagine a
-%generally satisfactory parametric formulation for (6.1).
-%Or perhaps it is just that hypothesis testing is more demanding
-%than estimation.
-%My Note: (6.1) is the general model (mu \sim g()).
-%7.1 is the case where the effects are a mixture of zero and N(-3,1).
-
-\subsubsection*{Calibration of posterior intervals}
-
-To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over 100 simulations in each Scenario). 
-Specifically we examine the empirical coverage of nominal lower 95\% credible bounds for a) all observations; b) significant negative discoveries; c) significant positive discoveries.
-We examine only lower bounds because the results for upper bounds follow by
-symmetry (except for the one asymmetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
-in each case: for negative discoveries the lower bound is typically large and negative and limits how big (in absolute value) 
-the effect could be; for positive discoveries the lower bound 
-is positive, and limits how small (in absolute value) the effect could be. Intuitively, the lower bound for negative discoveries depends on the accuracy of $g$ in its tail,
-whereas for positive discoveries it is more dependent on the accuracy of $g$ in the center.
-
-The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
-as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
-
-% As noted above, the penalty term (\ref{eqn:penalty}) causes systematic over-estimation of the mass of $g$ near 0,  which one would expect to overshrink interval estimates towards zero. 
-%The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound is negative and overshrinking (moving it too far towards zero) makes the left tail too big; conversely, for positive discoveries, the lower confidence bound is positive, and
-%overshrinking (moving it too far towards zero) makes the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
-%and be prepared to put up with some overshrinkage to avoid it.
-
-%Coverage rates are given in:
-%\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
-
-%Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all.tex}
-	\caption{All observations.
-Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario.
-This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero.
-Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (Table \ref{tab:nopen}).
-Removing the penalty in the half-uniform case is not recommended (see Appendix).}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg.tex}
-\caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
-and the normal-based method in the flat-top scenario.
-These results likely reflect inaccurate estimates of the tails of $g$ due to a disconnect between the tail of $g$ and the component distributions in these cases.
-For example, the uniform methods sometimes substantially underestimate the length of the tail of $g$ in these long-tailed scenarios,
-causing over-shrinkage of the tail toward 0.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos.tex}
-\caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds} \label{tab:coverage}
-\end{table}
-
-%
-%\subsubsection*{Reverse asymptotics}
-%
-%Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
-%
-%Note: we'd like to make sure it behaves ok in the case where most observations are null.
-%Perhaps simulate with 999 null and 1 alternative?
-%Also maybe look at performance for small sample sizes - e.g. n=1?
-%
-%Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
-%can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
-%by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
-%
-
-\subsection*{Differing measurement precision across units}
-
- We turn now to the second important component of our work: allowing for varying
- measurement precision across units. The key to this is the use of a likelihood,
- (\ref{eqn:normlik}) or (\ref{eqn:tlik}), that explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
-
-  To illustrate, we conduct a simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
- ($s_j=10$).  In both cases,
-we assume that half the effects are null
-and the other half are normally distributed with standard deviation 1:
-\begin{equation}
-p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1).
-\end{equation}
-In this setting, the poor-precision measurements $(s_j=10)$ tell us very little, and any sane analysis should effectively ignore them. 
- However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor}). This is because the poor measurements 
- produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
- which, when combined with the good-precision measurements, dilute the overall signal (e.g. they reduce the density of $p$ values near 0).
- This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
-the estimated error rates ($q$-values, or $\lfdr$ values) for the good-precision observations increase when the low-precision observations are included in the analysis
-(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
-for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).
-
-\begin{figure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_hist.pdf} 
-    \label{fig:goodpoor-hist}
-    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations}
- \end{subfigure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_scatter.pdf}
-    \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) of a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
-\end{subfigure}
-\caption{Simulation illustrating how, for existing FDR methods, 
-poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
-$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
-a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
-In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
-\end{figure}
-
- \subsubsection*{Reordering of significance, and the ``$p$ value prior"}
-
-Another consequence of accounting for differences in measurement precision is that \ashr may re-order the significance of the observations compared
- with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
- We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measuments that have the same $p$ value.
- The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
- and so the $\lfsr$ for these poor-precision measurements is always high.
-
-The potential for Bayesian analyses to re-order the significance of observations, and specifically to downweight imprecise observations,
-was previously discussed in \cite{guan.stephens.08}.
-However, Wakefield \cite{wakefield:2009} showed that, under a certain prior assumption, the Bayesian analysis produces the same ranking of significance as $p$ values (or their $z$ scores).
-He named this prior the ``$p$-value prior" because it can be thought of as the
-implicit prior assumption that is being made if we rank the significance of observations by their $p$ value.
-Wakefield's $p$-value prior assumes that the less precise effect estimates correspond to larger true effects, and specifically that they scale proportional
-to the standard errors $s_j$. More specifically still, it assumes a normal prior for the non-zero $\beta_j$ with mean 0 and variance $K s^2_j$ for some constant $K$.
-Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. Specifically, if, instead of assuming that $\beta_j$ is independent
-of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
-then the $\lfsr$ computed by ash.n provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
-as is illustrated in Figure \ref{fig:lfsr_pval}, right panel. (This result does not hold when using the mixtures of uniforms prior, ash.u.)
-
-This $p$-value prior assumes that the $z$ scores $z_j=\beta_j/s_j$
-are identically distributed, independent of $s_j$. This is essentially the assumption made, implicitly or explicitly, by existing methods -- like \locfdr, \mixfdr and \qvalue --
-that model the $z_j$ or $p_j$ directly.  In contrast, we have assumed up to now that the $\beta_j$ are independent of $s_j$. 
-We can set both these assumptions within a more general framework, which allows that $\beta_j/s^\alpha$ is independent of $s_j$ for some $\alpha$ [Equation (\ref{eqn:beta-alpha}))].
-Setting $\alpha=0$ implies that $\beta_j$ is
-independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
-effects (in absolute value). This latter assumption may often be qualitatively plausible: 
-for example, in gene expression studies the standard error for gene $j$ depends partly on
-the variance of its expression among samples, and genes with a larger variance may tend to be less tightly regulated and
-so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
-On the other hand, there is no particular reason to expect that either $\alpha=1$ or $\alpha=0$ will be the optimal choice.
-Indeed, optimal choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$,
-which will be dataset-specific. 
-Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
-comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
-likelihood-based methods can be used to select $\alpha$. For example, following the logic of the EB approach it
-would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
-which has been implemented in our software by C. Dai.
-
-%\begin{result}
-%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
-%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
-%\end{result}
-
-%To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
-%Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
-%In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
-%Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
-%which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
-%(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
-%may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)
-
-%\begin{result}
-%Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and using a mixture of normals for the prior distribution 
-%(\ref{eqn:beta-alpha}) with $\alpha=1$, 
-%the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$.
-Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$)
-%is montonic with $|t_j|$. 
-%\end{result}
-
-%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
-%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.
-
-%Most previous authors of EB approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
-%have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
-%In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
-%This is analogous
-%to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
-%standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
-%that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.
-
-\begin{figure}
-\centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
-\caption{Figure illustrating affects of prior assumptions on re-ordering of significance.
-Left panel shows results under our ``default prior" which assumes that effects $\beta_j$ are identically distributed, independent of $s_j$.
-Right panel shows results under the ``$p$-value prior", which assumes that $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$.} \label{fig:lfsr_pval}
-\end{figure}
-
-%\subsubsection*{Adaptive Shrinkage}
-%
-%Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
-%when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
-%the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
-% $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
-% In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
-% Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
-% and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
-%
-%Our methods we build on this idea in two ways. First,
-%by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
-%Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
-%adapts to measurement precision: more precise measurements undergo
-% less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
-%than observations with
-%small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
-%To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
-%Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
-%
-%\begin{figure}[!ht]
-%\begin{center}
-%\includegraphics[width=400px]{../examples/ash.pdf} 
-%\end{center}
-%\caption{Figure illustrating adaptive shrinkage.
-Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0.
-The standard error of each observation, $s_j$, was simulated from Inverse Gamma(5,5), which resulted in $s_j$ varying from 0.34 to 6.7.
-The shrunken estimates ($y$ axis) are plotted against the observed estimates ($x$ axis), with color indicating the (square root of the) precision of each measurement (blue=lower-precision; red=higher-precision).
-The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
-%\end{figure}
-
-%DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
-%To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
-Because the prior distribution is unimodal about zero,
-%the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
-%amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
-%For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
-%than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
-%small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
-
-%previous work based on normal means:
-%http://link.springer.com/article/10.1007/BF02562623
-
-%\subsection{Null behaviour}
-
-%Results of simulations under the null.
-
-%\subsection{Comparison of different underlying component distributions}
-%
-%Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
-%than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
-%likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
-%For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
-%increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
-%expect $L_{hu}$ to be significantly larger than the others.
-%
-%Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
-%
-%\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
-%\begin{center}
-%\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
-%%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
-%\end{center}
-%\caption{}
-%\end{figure}
-
-%Notes from Adrian Raftery - 
-%Newcombe (1886)
-%Dean and Raftery
-%Uniform is super-efficient
-
-\section*{Discussion}
-
-We have presented an Empirical Bayes approach to large-scale multiple testing that emphasises two ideas.
-First, we emphasise the potential benefits of using two numbers ($\bhat$, and its standard error)
-rather than just one number (a $p$ value or $z$ score)  to summarize the information on each test.
-While requiring two numbers is slightly more onorous than requiring one, in many settings these
-numbers are easily available and if so we argue it makes sense to use them.
-Second, we note the potential benefits -- both statistical and computational -- of assuming that the effects come from a unimodal distribution, and provide
-flexible implementations for performing inference under this assumption. We also introduce the ``false sign rate" as
-an alternative measure of error to the FDR, and illustrate its improved robustness to errors in model fit, particularly mis-estimation
-of the proportion of null tests, $\pi_0$.
-
-Multiple testing is often referred to as a ``problem" or a ``burden". In our opinion, EB approaches turn this idea on its head,
-treating multiple testing as an {\it opportunity}: specifically, an opportunity to learn about the prior distributions, and other modelling assumptions,
-to improve inference and make informed decisions about significance (see also \cite{greenland1991empirical}). This view
-also emphasises that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
-Indeed, the FDR at a given fixed threshold does not depend on the number of tests: as the number of tests increases, both the true positives and false positives increase linearly,
-and the FDR remains the same.
-(If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example.
-The EB approach captures this dependence in an intuitive way:
-if there are lots of strong signals then we infer $\pi_0$ to be small, and the estimated FDR (or $\lfdr$, or $\lfsr$) at a given threshold may be low, even if a large number of tests were performed; and conversely
-if there are no strong signals then we infer $\pi_0$ to be large and the FDR at the same threshold may be high, even if relatively few tests were performed. 
-More generally, overall signal strength is reflected in the estimated $g$, which in turn influences the estimated FDR.
-
-Two important practical issues that we have not addressed here are correlations among tests,
-and the potential for deviations from the theoretical null distributions of test statistics. These two
-issues are connected: specifically, unmeasured confounding factors can cause both correlations among tests
-and deviations from the theoretical null \cite{efron2007correlation,leek:2007}. And although there are certainly other
-factors that could cause dependence among tests, unmeasured confounders are perhaps the most worrisome in practice
-because they can induce strong correlations among large numbers of tests and profoundly impact results.
-Approaches to deal with unmeasured confounders can be largely divided into two types: those
-that simply attempt to correct for the resulting inflation of test statistics \cite{devlin1999genomic,efron2004large}, and those
-that attempt to infer counfounders using clustering, principal components analysis, or factor models 
-\cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
-and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable,
-they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alterntively, our methods
-could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:lik}) using
-$\bhat_j | \shat_j \sim N(0,\lambda_1 \shat_j + \lambda_2)$, where $\lambda_1,\lambda_2$ are to be estimated. However,
-this immediately raises issues of identifiability and we do not pursue the idea further here.
-
-Another important practical issue is the challenge of small sample sizes. For example, in genomics applications researchers sometimes
-attempt to identify differences between two conditions based on only a handful of samples in each. In such settings the
-normal likelihood approximation (\ref{eqn:normlik}) will be inadequate. And, although the $t$ likelihood (\ref{eqn:tlik}) partially
-addresses this issue, it is also, it turns out, not entirely satisfactory.  The root of the problem is that, with small sample sizes, raw estimated standard
-errors $\shat_j$ can be horribly variable. In genomics it is routine to address this issue by applying EB methods \cite{smyth:2004} to ``moderate" (i.e.~shrink) variance estimates,
-before computing $p$ values from ``moderated" test statistics. We are currently investigating how
-our methods should incorporate such ``moderated" variance estimates to make it applicable to small sample settings.
-
-Our approach involves compromises between flexibility, generality, and simplicity on the one hand, and statistical efficiency and principle on the other.
-For example, in using an EB approach that uses a point estimate for $g$, rather than a fully Bayes approach that accounts for uncertainty in $g$,
-we have opted for simplicity over statistical principle. And in summarizing every test by
-two numbers and making a normal or $t$ approximation to the likelihood, we have aimed to produce generic methods that 
- can be applied whenever such summary data are available -- just as \qvalue can be applied to any set of $p$ values for example -- although
- possibly at the expense of statistical efficiency compared with developing multiple tailored approaches based on context-specific likelihoods. 
- Any attempt to produce generic methods will involve compromise between generality and efficiency.
- In genomics, many analyses -- not only FDR-based analyses -- involve first computing a series of $p$ values before
- subjecting them to some further downstream analysis. An important message here is 
- that working with two numbers ($\bhat_j,\shat_j$), rather than one ($p_j$ or $z_j$), 
- can yield substantial gains in functionality (e.g.~ estimating effect sizes, as well as testing; accounting for variations in
-measurement precision across units) while losing only a little in generality.  We hope that our work will
-encourate development of methods that exploit this idea in other contexts.
-
-%This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
-%from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
-%with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
-
-%
-%
-%Both \locfdr and \mixfdr allow for deviations from
-%the theoretical null distribution by allowing that, under the null, the $z$ scores are $N(0,\sigma^2)$ with $\sigma$ estimated from the data rather
-%than fixed to 1. Within our approach this idea could be incorporated by changing the likelihood - for example, to assume $\bhat_j \sim N(0,s_ 
-%
-%These are tricky issues that, as far as we are aware, can affect
-%all approaches to FDR analysis. An important focus of Efron's work, also implemented in \locfdr, 
-% is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
-%theoretical null distribution. 
-%
-%For example, we and others have worked on Bayesian approaches to genetic association testing, using subjective prior specification of
-%effect sizes based on external data (e.g.~\cite{stephens:2009}). However, the large numbers of tests brings the opportunity to learn about
-%priors, and here we aim to exploit this
-
-%Choice of mixture components. Use of likelihood to compare different mixture components?
-
-%  
-%Directly modeling the p values, or z scores, say via non-parametric methods, 
-%can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-%distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-%The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-%Before proceeding, we feel obliged to correct the common misconception
-% that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
-% not depend on the number of tests. 
-%  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
-% 
-% In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
-
-\section*{Detailed Methods} \label{sec:implement}
-
-\subsection*{Embellishments}
-
-\subsubsection*{More flexible unimodal distributions}
-
-Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
-Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
- Therefore, we have implemented a more general approach based on 
-\begin{equation} \label{eqn:g-general}
-g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) ,
-\end{equation}
-where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
-\begin{enumerate}[(i)]
-\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$, \qquad (``ash.n")
-\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,  \qquad (``ash.u")
-\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,  \qquad (``ash.hu")
-\end{enumerate}
-where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
-(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
-and $\pi$ is a $2K+1$ vector that sums to 1.)
-The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
-Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
-under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
-(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
-(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
-however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)
-
-Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
-using a large dense grid of $\sigma^2_k$ or $a_k$, (i)-(iii) can respectively approximate,
-with arbitrary accuracy,
-\begin{enumerate}[(i)]
-\item any scale mixture of normals, which includes as special cases
-the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
-\item any symmetric unimodal distribution about 0.
-\item any unimodal distribution about 0.
-\end{enumerate}
-The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
-In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.
-
-Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
-The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
-See Implementation for details of our software defaults.
-
-%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting.
-For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
-%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
-
-\subsubsection*{Dependence of effects on standard errors}
-
-Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
-This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
- \begin{equation} \label{eqn:beta-alpha}
- \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
- \end{equation}
-for any $\alpha$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
-assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
-it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
-the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.
-
-The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
-To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
- standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
- Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
-yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
-%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
-%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
-%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
-
-%Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
-%how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
-%it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
-%rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
-%and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
-%this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
-%do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
-
-\subsubsection*{Replace normal likelihood with $t$ likelihood}
-
-We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
-  \begin{equation} \label{eqn:tlik}
- \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
- \end{equation}
-where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
-and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
-The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
-This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms; see Implementation below.
-When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
-
-Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
-\begin{equation} \label{eqn:tdist}
-(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
-\end{equation}
-However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
-In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
-currently pursuing this improved approach (joint work with M.Lu) and results will be published elsewhere. 
-%We note here that
-%modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
-%where our simpler approach can produce unsatisfactory results.
-
-\subsubsection*{Non-zero mode}
-
-An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood, rather than fixed to 0.
-
-\subsection*{Implementation Details}
-
-\subsubsection*{Likelihood for $\pi$}
-
-We define the likelihood for $\pi$ to be the probability of the observed data $\bhat$ conditional on $\shat$: $L(\pi) := p(\bhat | \shat,\pi),$ which
-by our conditional indendence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$.
-[One might prefer to define the likelihood as $p(\bhat, \shat | \pi) = p(\bhat | \shat, \pi) p(\shat | \pi)$, in which case our definition comes down to assuming that the term $p(\shat | \pi)$ does not depend on $\pi$.]
-
-Using the prior $\beta_j \sim \sum_{k=0}^K \pi_k f_k(\beta_j)$ given by (\ref{eqn:g-general}), and the normal likelihood (\ref{eqn:normlik}), integrating over $\beta_j$ yields
-\begin{equation}
-p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
-\end{equation}
-where 
-\begin{equation}
-\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
-\end{equation}
-denotes the convolution of $f_k$ with a normal density.
-These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
-Specifically, 
-\begin{equation} \label{eqn:uconv}
-\tilde{f}_k(\bhat_j)  = 
-\begin{cases}
-N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
-\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
-\end{cases}
-\end{equation}
-where $\Psi$ denotes the cumulative distribution function (c.d.f.) of the standard normal distribution.
-If we replace the normal likelihood with the $t_\nu$ likelihood (\ref{eqn:tlik}) then the convolution 
-for $f_k$ uniform the convolution is still given by (\ref{eqn:uconv}) but with $\Psi$ the c.d.f. 
-of the $t_\nu$ distribution function. (The convolution for $f_k$ normal is tricky and we have not implemented it.)
-
-\subsubsection*{Penalty term on $\pi$}
-
-To make $\lfdr$ and $\lfsr$ estimates from our method ``conservative" we add a penalty 
-term $log(h(\pi;\lambda))$ to the log-likelihood $\log L(\pi)$ to encourage over-estimation of $\pi_0$:
-\begin{equation} \label{eqn:penalty}
-h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
-\end{equation}
-where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
-consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0sims}).
-
-Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
-we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
-
-\subsubsection*{Problems with removing the penalty term in the half-uniform case}
-
-It is straightforward to remove the penalty term by setting $\lambda_k=1$ in (\ref{eqn:penalty}).
-We note here an unanticipated problem we came across when using no penalty term in the half-uniform case 
-(i.e.~ $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$ in (\ref{eqn:g-general})): when the data
-are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
-sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
-for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
-and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
-
-\subsubsection*{EM algorithm}
-
-With this in place, the penalized log-likelihood for $\pi$ is given by:
-\begin{equation}
-\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
-\end{equation}
-where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This can be maximized using an EM algorithm \cite{dempster77}, whose
-one-step updates are:
-\begin{align} \label{eqn:w}
-w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
-n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
-\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}
-\end{align}
-Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
-and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
-probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$. (The $w_{kj}$ are sometimes referred to as the ``responsibilities".)
-Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
-We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple} to accelerate convergence of this EM algorithm.
-
-%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
-\subsubsection*{Initialization}
-
-By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
-Our rationale for initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
-but weak signal in the data cannot quickly draw the algorithm towards the null.
-
-In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
-$\pi_0=1; \pi_k=0 (k>0)$. If not then we replace the EM solution with the global null solution. This aims to guard against errors due to convergence to a local optimum when the
-data are consistent with the global null.
-
-\subsubsection*{Conditional distributions}
-
-Given $\hat\pi$, we compute the conditional distributions 
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
-\end{equation} 
-Each posterior is a mixture on $K+1$ components:
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
-\end{equation}
-where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
-and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
-$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
-All these posterior distributions are easily available.
-For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
-If $f_k$ is normal and $L$ is normal, then this is a normal distribution.
-
-\subsubsection*{Choice of grid for $\sigma_k, a_k$} \label{sec:grid}
-
-\def\sigmamax{\sigma_{\text{max}}}
-\def\sigmamin{\sigma_{\text{min}}}
-When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
-in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
-as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
-based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
-is negative we set $\sigmamax = 8\sigmamin$.)
-
-When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.
-
-Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
-a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.
-
-\appendix
-
-\begin{table}[!ht]
-\centering\begin{tabular}{c c } \toprule
-Scenario & Alternative distribution, $g_1$  \\ \midrule
-spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
-near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
-flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
- &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
-skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
-big-normal & $N(0,4^2)$ \\ 
-bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
-\end{tabular}
-\caption{Summary of simulation scenarios considered} \label{table:scenarios}
-\end{table}
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all_nopen.tex}
-	\caption{All observations}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg_nopen.tex}
-\caption{``Significant" negative discoveries.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos_nopen.tex}
-\caption{``Significant" positive discoveries.}
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{table:nopen}
-\end{table}
-%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
-%\begin{equation}
-%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
-%\end{equation}
-%which is close to the definition in \cite{storey.03}.
-
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgements}
-
-Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
-prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} , and includes contributions 
-from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.
-
-This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.
-
-%\section*{References}
-% The bibtex filename
-\bibliography{/Users/stephens/Dropbox/Documents/mainbib}
-
-\section*{Figure Legends}
-
-\clearpage
-
-\section*{Tables}
-
-\section*{Supporting Information Legends}
-
-Supplementary material can be found in {\bf Supplementary Information S1.}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
-\end{document}

### 2015-10-06 subsectionDiffering_.tex
New Article!

--- 2015-10-06 paper/main.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
-\documentclass[11pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-%\usepackage{natbib}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-\usepackage{booktabs}
-\usepackage{xspace}
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-\usepackage{sectsty}
-\subsectionfont{\normalsize}
-\subsubsectionfont{\mdseries\itshape}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\usepackage{color}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-% Use the PLoS provided bibtex style
-\bibliographystyle{/Users/stephens/Dropbox/Documents/stylefiles/plos2009}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
-\pagestyle{myheadings}
-%% ** EDIT HERE **
-\usepackage{enumerate}
-\usepackage{multirow} 
-\usepackage{url}
-\usepackage{xr} %for cross-referencing
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-\newtheorem{algorithm}{Algorithm}
-\newtheorem{proposition}{Proposition}
-\newtheorem{restateproposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{result}{Result}
-\newtheorem{note}{Note}
-\newtheorem{definition}{Definition}
-
-\def\lfdr{\textit{lfdr}}
-\def\lfsr{\textit{lfsr}}
-\def\x{\mbf{x}}
-\def\y{\mbf{y}}
-\def\e{\mbf{e}}
-\def\g{\mbf{g}}
-\def\p{p}
-\def\Yb{\hat{Y}_1^{\text{BAYES}}}
-\def\Xperp{X_{1 \perp 0}}
-\def\Pperp{P_{1 \perp 0}}
-\def\Pperpb{\Pperp^{\text{B}}}
-\def \hatd { \widehat{ d } }
-\def \dhatin {\hatd^{\mbox{\tiny IN}}}
-\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
-\def \mvn{\text{MVN}}
-\def\mn{\text{MN}}
-\def \iw{\text{W}^{-1}}
-\def \fp {fastPHASE\ }
-\def\bhat{\hat{\beta}}
-\def\shat{\hat{s}}
-\def\var{\rm Var}
-\def\mvr{\text{BMVR}}
-\def\bfuni{\text{BF}_\text{uni}}
-\def\one{\mbf{1}}
-\def\zero{\mbf{0}}
-\def\hall{H_{\text{all}}}
-\def\pall{\p_{\text{all}}}
-\def\bfall{\text{BF}_\text{all}}
-\def\bfav{\text{BF}_\text{av}}
-\def\gall{\gamma_\text{all}}
-\def\limbfall{\bfall^\rightarrow}
-\def\limbfav{\bfav^\rightarrow}
-\def\limbfgamma{\BF_\gamma^\rightarrow}
-\def\sxx{S_{xx}}
-\def\syy{S_{yy}}
-\def\vyy{V_{yy}}
-\def\vyx{V_{yx}}
-\def\vxx{{V_{xx}}}
-\def\rss{\text{\rm RSS}}
-\def\dss{\text{\rm $\Delta$SS}}
-\def\U{U}
-\def\D{D}
-\def\I{I}
-\def\luu{L_{UU}}
-\def\indep{\perp}
-\def\qvalue{{\tt qvalue}\xspace}
-\def\locfdr{{\tt locfdr}\xspace}
-\def\mixfdr{{\tt mixfdr}\xspace}
-\def\ashr{{\tt ashr}\xspace}
-
-\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
-                                % mode too
-\newcommand\height{{\it height}}
-\newcommand\weight{{\it weight}}
-\def\BF{\text{BF}}
-
-%% END MACROS SECTION
-%\externaldocument{plos_one_rev1_suppinfo}
-
-\begin{document}
-
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{False Discovery Rates: A New Deal}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Matthew Stephens$^{1*}$, 
-\\
-\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
-\\
-$\ast$ E-mail: Corresponding mstephens@uchicago.edu
-\end{flushleft}
-
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-%\section*{Author Summary}
-
-\section*{Introduction}
-
-Since its introduction in 1995 by Benjamini and Hochberg 
-\cite{benjamini1995controlling}, the ``False Discovery Rate" (FDR) has quickly established itself
-as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing
-in which the goal is to identify the non-zero ``effects" among a large number of imprecisely-measured effects.
-
-Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
-indeed, the notion that EB approaches could be helpful in handling multiple comparisons predates introduction of
-the FDR  (e.g.~\cite{greenland1991empirical}). More recently, EB approaches to the FDR
-have been extensively studied by several authors, 
-especially B.~Efron and co-authors 
-\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
-see also \cite{kendziorski2003parametric,newton2004detecting, 
-datta2005empirical,muralidharan2010empirical} for example.
-So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
-existing widely-used FDR pipelines) and can substantially affect
-inference. The first idea is to {\it assume that the distribution of effects is unimodal}. This yields a very simple, fast, and stable computer 
-implementation, as well as improving inference of FDR when the unimodal assumption is correct.  
-The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
-to summarize each measurement. This idea allows variations in measurement precision to be better accounted for,
- and avoids a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
-
-In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
-the idea that it may be preferable to focus on estimation rather than on testing.
-In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
-simply estimation with some positive prior probability that the effect is exactly zero.
-However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
-perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
-controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
-``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
-in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
-the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the lfsr, particularly its increased robustness to modelling assumptions.
-
-%terminology
-%``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
-%and correspondingly define the ``false signed discovery rate", or simply ``false sign rate" (FSR), in a natural way.
-
-%The methods we provide here can address both problems, and we provide results that highlight important benefits of focussing on estimation, and particularly on assessing the correctness of the {\it sign} of the estimated effects, rather than on testing for non-zero effects.
-
-Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
-(We address the reasons for this name, and connections with shrinkage analysis, in the Discussion.)
-
-%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.
-
-%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-%measurements are made on only a small number of
-%samples in each condition - sometimes as few as one
-%sample in each condition. Thus the error in estimates
-%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-%in estimates of $\beta_j$ still greater.
-%
-
-%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
-%Here we present new tools that bring two important ideas to bear on the problem.
-%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
-%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
-%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
-%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
-%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
-%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
-%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
-%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
-%a substantial increase in the number of discoveries at a given FDR).
-
-%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
-%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
-Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
-%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
-%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
-%generic methods that work directly with two numbers for each 
-%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
-Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
-%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
-
-\def\df{df}
-\def\FDR{\text{FDR}}
-\def\fdr{\text{\lfdr}}
-\def\FSR{\text{FSR}}
-\def\fsr{\text{lfsr}}
-
- \section*{Methods}
-
- \subsection*{Model Outine}
-
-% Here we consider one setting in which FDR-based methods are widely used:
-%an experimenter has measured a large numbers of ``effects", each with an associated
-%measurement precision (standard error), 
-%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
-%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
-%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
-%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
-%For example, in genomics
-%a common goal is to identify ``differentially-expressed genes" -- that is, genes
-%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
-%mean expression level between the two conditions, often referred to as the ``log-fold change". 
-Here we describe the simplest version of our method, and briefly discuss embellishments we have also implemented.
-
-Suppose that we are interested in the values of $J$ ``effects" $\beta=(\beta_1,\dots,\beta_J)$. For example, in a typical genomics application
-that aims to identify differentially expressed genes,  $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
-In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
-that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
-more general problem of estimating, and assessing uncertainty in, $\beta_j$.
-
-Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
-and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
-which by Bayes theorem can be written as 
-\begin{equation}
-p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
-\end{equation}
-
-For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
-This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
-A simple way to implement the UA is to assume that $g$ is
-a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
-\begin{align} \label{eqn:beta}
-p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
-g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
-\end{align}
-where $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyperparameters, which are non-negative and sum to one, and are to be estimated,
-while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
-represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
-(We encourate the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
-although of course in practice we use a finite grid -- see Implementation Details.)
-
-For the likelihood $p(\bhat | \beta, \shat)$ we assume  
-\begin{equation} \label{eqn:normlik}
-p(\bhat | \beta, \shat) = \prod_j N(\bhat_j; \beta_j, \shat_j^2).
-\end{equation}
-Here, in additional to some conditional independence assumptions, we are 
-effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-
-This simple model features both the key ideas we want to emphasise in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
-the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
-with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
-However, this simple model also has several additional assumptions that can be relaxed. Specifically,
-\begin{enumerate}
-\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) also implies that $g$ is symmetric about 0; more flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see (\ref{eqn:g-general})].
-\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
-We can relax this to allow for a relationship between these quantities [see (\ref{eqn:beta-alpha})].
-\item The likelihood (\ref{eqn:normlik}) effectively assumes that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-We can generalize this likelihood using a $t$ likelihood [see (\ref{eqn:tlik})].
-\end{enumerate}
-These embellishments are detailed in Embellishments and Implementation Details.
-Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
-(which are also made by most existing EB approaches to this problem). Correlations among tests certainly arise in practice, either due to genuine correlations
-in the system of study, or due to unmeasured confounders, and their potential to impact results of an FDR analysis is important to consider
-whatever analysis methods are used: see \cite{efron2007correlation,leek:2007} for relevant discussion.
-
-%We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
-%$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
-%\begin{equation}
-%p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
-%\end{equation}
-%Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
-%\begin{equation} \label{eqn:normlik}
-%p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
-%
-%We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
-%as a mixture of null and alternative components:
-%\begin{equation} \label{eqn:betaj0}
-%g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
-%\end{equation}
-%where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
-%We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
-%\begin{equation} \label{eqn:beta-iid}
-%p(\beta | \shat) = \prod_j g(\beta_j).
-%\end{equation}
-%
-%Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
-%with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
-%to use a mixture of zero-mean normal distributions:
-%\begin{equation} \label{eqn:mixnorm}
-%g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
-%Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
-%represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
-% The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
-% more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below. 
-% 
-% Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
-% yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
-% \begin{equation} \label{eqn:g}
-% g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
-% \end{equation}
-% 
-% Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
-% \begin{align} \label{eqn:betaj}
-% p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
-% p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
-% \end{align}
-% with $g$ given by the mixture of normals $(\ref{eqn:g})$.
-
-\subsection*{Fitting the model}
-
-In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
-each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
-Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
-\begin{equation} \label{eqn:marginallik}
-p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
-\end{equation}
-where we define $\sigma_0:=0$.
-
-The usual EB approach to fitting this model would involve two simple steps:
-\begin{enumerate}
-\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
-\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
-against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | | \bhat, \shat,\hat{\pi})$.
-\end{enumerate}
-Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(The simplicity of the EM algorithm in step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
-which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
-
-Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximise a penalized
-likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
-We introduce this penalty because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
-
-\subsection*{The local False Discovery Rate and local False Sign Rate}
-
-As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
-In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
-One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
-the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
-In other words it is the posterior probability that $\beta_j$ is actually zero:
-\begin{equation} \label{eqn:fdr}
-\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
-\end{equation}
-
-The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
-whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
-the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey1991philosophy}) argued that
-``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$.
-Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
-\begin{quote}
-...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
-\end{quote}
-Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
-focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.
-
-Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
-would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
-\begin{equation} \label{eqn:lfsr}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
-\end{equation}
-To illustrate, suppose that 
-\begin{gather*}
-p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
-p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
-p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
-\end{gather*} 
-Then 
-from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
-our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.
-
-As our notation suggests, $\lfsr_j$ is intended to be compared and contrasted with $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
-small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
-Of coure, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
-this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
-(this follows from the definition because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
-In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
-$\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.
-
-From these ``local" measures of significance, we can also compute average error rates over subsets of observations
-$\Gamma \subset \{1,\dots,J\}$. For example,   
-\begin{equation} \label{eqn:FDRhat}
-\widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
-\end{equation}
-estimates the FDR we would obtain if we were to declare all tests in $\Gamma$ significant.
-And 
-\begin{equation} \label{eqn:qvalue}
-q_j := \widehat \FDR(\{k: \fdr_k \leq \fdr_j\}) 
-\end{equation}
-provides a measure of significance analogous to Storey's $q$ value \cite{storey.03}.
-
-%Because our methods involve explicitly modelling effects $\beta_j$, rather than 
-%$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
-
-%Indeed, even in settings where some
-%$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
-%analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
-
-\subsection*{Related work}
-
-\subsubsection*{Previous approaches focussed on FDR}
-
-Among previous methods that explicitly consider the FDR and related quantities,
-our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
- (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
-  %Here we highlight the key differences between our approach and these three existing approaches.
-
-There are two key differences between our approach and all of these three existing methods. First, 
-whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
-we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
-for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
-Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 
-how it can improve performance by taking better account of variation in measurement precision among observations.
-%(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
-% Or more generally Section xx illustrates the potential practical effects
-
-%allows for different modelling assumptions to be compared through likelihood comparisons.
-Second, our unimodal assumption (UA) that the effects are unimodal about zero is quite different from assumptions made by $\qvalue, \locfdr$ or $\mixfdr$.
-Indeed, $\locfdr$ assumes that all $z$ scores near 0 are null (Efron calls this the Zero Assumption; ZA), which implies that 
-under the alternative hypothesis the distribution of $z$ scores has {\it no mass at 0}; this contrasts strikingly with the UA, which implies that this distribution has its peak at 0! 
-Similarly, $\qvalue$ assumes that all $p$ values near 1 are null,
-which is the same as the ZA because $p$ values near 1 correspond to $z$ scores near 0. And although \mixfdr does not formally make the ZA,
-we have found that in practice, with default settings, the results often approximately satisfy the ZA (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}).
-Thus, not only do these existing methods not make the UA, they actually make assumptions that are, in some sense, as different
-from the UA as they can be.
-
-Given that the UA and ZA are so different, it seems worth discussing why we generally favor the UA. 
-Although the UA will not apply to all situations, we believe that it will often be reasonable, especially in FDR-related contexts
-that have traditionally focussed on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis,  
-it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible. Further, it seems reasonable to expect that larger effects become
-decreasingly plausible, and so the distribution of the effects will be unimodal about 0.
-To paraphrase Tukey, ``All we know about the world teaches us that large effects are rare, whereas small effects abound."
-We emphasise that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
-It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
-
-%Tukey "peri-null"
-
-%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
-%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}.
-
-In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficents - 
-common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
-These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
-it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
-The UA assumption on regression coefficients is directly analagous to our UA here, and so its
-widespread use in the regression context supports its use here.
-
-Alternatively, we could motivate the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
-such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
-make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).
-
-Finally, the UA also has a considerable practical benefit: it yields simple algorithms that are both computationally
-and statistically stable. We illustrate these features in Results.
-
-\subsubsection*{Other work}
-
-There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
- Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
-indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
-This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
-Also, they provide no software implementation.
-More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
-Some pointers to recent papers in which other relevant
-citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}.
-Much of the literature focusses on the homoskedastic case (i.e.
-$\shat_j$ all equal) whereas we allow for heteroskedasticity.
-And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential.
-Several recent papers consider more flexible non-parametric assumptions on
-$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
-estimate (NPMLE) for $g$.
-These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive.
-However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
-may perform well for point estimation, it seems possible it will not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
-Indeed, one way of thinking about the UA is as a way to regularize $g$.
-
-%The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
-%encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-%$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-%$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-%the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-%observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-%corresondingly little effect on the estimated FDR of other observations. 
-%We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-%$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-%that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-%estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-%Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-\def\dt{\text{dt}}
-
-%
-%assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
-%$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
-% $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
-%  \begin{equation} \label{eqn:lik}
-% \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
-% \end{equation}
-% where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e.
-the density of $\mu+\sigma T_\nu$).
-In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
-%  \begin{equation} \label{eqn:normlik}
-% \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
-% \end{equation}
-%which, in light of (\ref{eqn:betahat}), is equivalent to
- % \begin{equation} \label{eqn:lik2}
-% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
- %\end{equation}
-
-%The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
-they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
-The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$.
-Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$.
-If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$.
-See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.
-
-%This likelihood approximation can be motivated by assuming that i) $\bhat$
-% $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
-% $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$. 
-%
-
-%Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.
-
-%While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
- %that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
-%there is an important difference: here,
-
-%The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
-For example, we can write $p(\bhat_j, \shat_j | \beta_j) = p(\bhat_j | \shat_j, \beta_j)p(\shat_j | \beta_j)$.
-Now, if we are willing to assume that $\shat_j$ alone contains no information about $\beta_j$, or equivalently that $p(\shat_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\bhat_j | \shat_j, \beta_j)$, and if $\bhat_j | \shat_j, \beta_j \sim N(\beta_j, \shat_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.
-
-\section*{Results}
-
-We compare results of \ashr with existing FDR-based methods implemented
-in the R packages \qvalue (v2.0 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
-In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
-to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.
-
-\subsection*{Effects of the Unimodal Assumption}
-
-Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
-standard error, $s_j=1$ and all methods are provided that information. That is,
-$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$.
-In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the only difference between our method and methods like $\locfdr$ and $\mixfdr$ are in how they estimate $g$.
-
-To briefly summarize the results in this section:
-\begin{enumerate}
-\item The UA can produce very different inferences compared with the ZA made by existing methods.
-\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
-\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
-\end{enumerate}
-
-\subsubsection*{The UA and ZA can produce different inferences}
-
-\begin{figure}
-\center \includegraphics[height=6in]{figures/decomp_ZA.pdf}
-\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
-The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
-In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
-The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
-(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
-\end{figure}
-
-To illustrate the different inferences from the UA and ZA we show results for a single
-dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
-but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
-\qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
-The results (Figure \ref{fig:ZA}) illustrate the clear difference between the existing methods and our method. The effects of the ZA made by \qvalue and \locfdr
-are visually clear, producing a ``hole"  in the alternative $z$ score distribution around 0. Although \mixfdr does not formally make the ZA, its decomposition
-exhibits a similar hole. In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
-effectively ``filling in" the hole.  (Of course the null distribution also has a peak at 0, and the local fdr 
- under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)
-
-Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA.
-Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
-This would increase the null component (dark blue) at the expense of
-the alternative component (light blue).
-Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0.
-The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears.
-In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
-See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
-demonstration.
-
- \subsubsection*{The UA can produce conservative estimates of $\pi_0$}
-
-The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
-Consequently \ashr will estimate smaller $\lfdr$s and FDRs than existing methods that make the ZA. 
-This is desirable, provided that these estimates remain conservative: that is,
-provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
-The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behaviour. To check its effectiveness
-we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
-for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
-They range from a ``spiky" distribution -- where many non-zero $\beta$ are
-too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
-flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
-making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
-which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.
-
-% How do these differences influence estimates of the FDR? To address this, note that  -- in this simple case where the 
-% $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
-% of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
-% strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
-% although the methods considered here differ from one another in many ways, having estimated $\pi_0$
-% they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
-% empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
-%
-
-%Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
-% and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
-%Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
-%At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
-%overestimates of $\pi_0$, which will yield overestimates of the FDR).
-
-\begin{figure} 
-\begin{center}
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=1.8in]{figures/scenarios_density.pdf} 
-	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/pi0_est.pdf} 
-\caption{Comparison of true and estimated values of $\pi_0$.
-When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ashr being least conservative, and hence most accurate.
-When the UA does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.} \label{fig:pi0sims}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfdr_est.png} 
-\caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
-Black line is $y=x$ and red line is $y=2x$.
-Estimates of $\lfdr$ are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est.png} 
-\caption{As in c), but for $\lfsr$ instead of $\lfdr$.
-Estimates of $\lfsr$ are consistently less conservative than $\lfdr$ when UA holds, and also less anti-conservative in bimodal scenario.}  \label{fig:lfsr}
-\end{subfigure}
-\end{center}
-\caption{Results of simulation studies (constant precision $s_j=1$).}
-\end{figure}
-
-For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
-data as follows:
-\begin{enumerate}
-\item Simulate $\pi_0  \sim U[0,1]$.
-\item For $j=1,\dots,J$, simulate $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)$.
-\item For $j=1,\dots,J$, simulate $\bhat_j | \beta_j \sim N(\beta_j,1)$.
-\end{enumerate}
-%Thus these simulations assume the same
-%precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).
-
-Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ashr ($y$ axis) with the true values ($x$ axis). 
-For \ashr we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
-and as a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
-as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
-The results show that \ashr provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
-in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ashr estimates can be slightly anti-conservative.
-We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
-are used. (In addition, the effects on $\lfsr$ estimates turn out to be relatively modest; see below).
-
-%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
-%
-%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
-%and so all effects are non-zero and no nulls hold. This scenario is rather different
-%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
-%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
-%``mostly non-null"
-%scenario, and we would agree! On the other hand, the distribution of $p$ values
-%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
-%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
-%Thus it seems desirable to have methods that perform well in both cases, and this is
-%an important feature of the methods we introduce here.
-%
-%TODO: put FDRs on graph? Simulate with 50\% null?
-
-\subsubsection*{The $\lfsr$ is more robust than \lfdr}
-
-The results above show that \ashr can improve on existing methods in producing smaller, more accurate, 
-estimates of $\pi_0$, which will lead to more accurate estimates of FDR.
-Nonetheless, in many scenarios \ashr continues
-to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
-This is because these scenarios include 
-an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
-estimation of $\pi_0$ impossible. Put another way, and as is well known, $\pi_0$ is not identifiable:
- the data can effectively provide an upper bound on plausible values of $\pi_0$,
-but not a lower bound (because the data cannot rule out that everything is non-null, but with miniscule effects).
-To obtain conservative behaviour we must estimate $\pi_0$ by this upper bound, which can
-be substantially larger than the true value.
-
-Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
-overestimation of $\pi_0$ is corresponding overestimation of FDR (and $\lfdr$, and $q$ values).
-To illustrate, Figure \ref{fig:lfdr} compares the estimated $\lfdr$ from ash.n with the true value (computed using
-Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios which involve many
-non-zero effects that are very near 0 (e.g.~the spiky scenario with $\pi_0$ small) where
-$\pi_0$ can be grossly overestimated.  
-(Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.)
-
-The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
- become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
- This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
-To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
-be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
-the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in $\lfsr$ than $\lfdr$.
-
-Note that, compared with previous debates regarding testing, 
-this section advances an additional reason for focussing on the sign of the effect, rather than just testing whether it is 0. 
-In previous debates authors have argued against testing whether an effect is 0
-because it is {\it implausible that effects are exactly 0}. Here we add that {\it even if one believes
-that some effects may be exactly zero}, it is still better to focus on the sign, because generally {\it the data are more informative about that question}
-and so inferences are more robust to, say, the inevitable mis-estimation of $\pi_0$.
-To provide some intuition, consider an observation with a $z$ score of 0. The $\lfdr$ of this observation can range from 0 (if $\pi_0=0$)
-to 1 (if $\pi_0=1$). But, assuming a symmetric $g$, the $\lfsr>0.5$ whatever the value of $\pi_0$, because the observation $z=0$ says
-nothing about the sign of the effect.
-Thus, are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
-whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.
-
-Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
-Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
-if we are prepared to assume that no effects are exactly zero, then removing the point mass 
-yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
-However, there is ``no free lunch":  if in fact some effects are exactly zero
-then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
-We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.
-
-\begin{figure}
-\begin{center}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s_nn.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be more accurate in this case.
-That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
-\end{subfigure}
-\end{center}
-\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
-\end{figure}
-
-\subsubsection*{Numerical Stability}
-
-The EM algorithm, which we use here to fit our model, is notorious for convergence to local optima. 
-However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
-caused by such behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
-in addition running it using our default initialization procedure (see Implementation details). We then compared the largest
-log-likelihood achieved across all 11 runs with the log-likelihood achieved by the default run.
-When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference ($< 0.02$), 
-and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference ($<0.02$),
-and 6\% of runs showed an appreciable log-likelihood difference ($>1$), with the largest difference being 5.0. However, perhaps suprisingly, 
-even for this largest difference results from the default run (e.g.
-the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood.
-See \url{http://github.com/stephens999/ash/blob/master/dsc-robust/summarize_dsc_robust.rmd} for further details.
-
-\subsubsection*{The UA helps provide reliable estimates of $g$}
-
-An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
-can estimate the {\it size} of each effect $\beta_j$.
-Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
-to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
-threshold $T$.
-%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
-%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
-%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
-%towards zero, with the strength of the shrinkage being estimated from the data.
-Further, because the posterior distribution is, by definition,
-conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
-been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
-In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
-\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
-how can confidence intervals be assigned to the true
-effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
-and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
-%but also notes the difficulties of implementing a stable general algorithm.
-
-The ability of the EB approach to provide valid ``post-selection" interval estimates is extremely attractive in principle.
-But its usefulness in practice
-depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
-which are notoriously difficult in general. Indeed, Efron emphasises
-the difficulties of implementing a stable general algorithm, noting in his rejoinder
-``the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying
-to be overly nonparametric ... but it is hard to imagine a
-generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
-Our key point here is that the UA greatly simplifies the deconvolution problem.
-While not meeting Efron's desire for an entirely general nonparametric approach, we believe 
-that the UA can handle many cases of practical interest.
-
-To illustrate this, Figure \ref{fig:egcdf} compares the
-estimated $g$ from \ashr  with that from \mixfdr which does not make the UA 
-(and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
-The greater reliability of estimates afforded
-by the UA is immediately apparent. In particular the estimated cdf from \mixfdr often has an almost-vertical
-segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. 
-The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
-While the UA is not the only way to achieve this, we find it an attractive, simple and effective approach.
-
-Interestingly, even in the ``bimodal" scenario $\ashr$ 
-is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
-Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
-Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
-to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
-(\url{http://github.com/stephens999/ash/blob/master/dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
-clearly visible in the observed $\bhat$.
-
-We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ashr tends to systematically
-overestimate the mass of $g$ near zero. On careful inspection, this is apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
-and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf}b),
-and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
-under-estimating $\pi_0$ (Figure \ref{fig:egcdf}c).
-
-\begin{figure}[h!] 
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=2in]{figures/egcdf.pdf}
-	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf_nopen.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
-Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
-\end{subfigure}
-\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.} \label{fig:egcdf}
-\end{figure}
-
-%making the UA provides an attractive and simple way to achieve estimating $g$  large
-%subset of cases of interest (those where the UA can reasonably be assumed to hold).
-
-%In practice, the utility of the posterior distributions noted above depends primarily on the accuracy
-%of the estimate of the distribution $g(\cdot)$.
-
- %quote from efron rejoinder , p46:
-%MorrisÕ Section 3 is especially pertinent. His formula
-%for p(?|z) is related to my discussion of the
-%BenjaminiÐYekutieli False Coverage Rate method in
-%Section 7, particularly (7.2)Ð(7.4). Originally I had
-%hoped to develop an empirical Bayes method for estimating
-%such models, but the effort foundered on practical
-%difficulties involving the perils of deconvolution.
-%Section 6 on the Òone-group modelÓ is the ugly duckling
-%of the current paper, but it bears on some important
-%points raised in the discussion. Figure 7 concerns
-%a fuzzy version of simultaneous hypothesis testing,
-%where, as in MorrisÕ hospital example, the usual singlepoint
-%null hypotheses seem unequal to the task. The
-%development from (6.6) onwards, particularly (6.12),
-%bears on the possibility of nonnormal null distributions,
-%and is about as far as I can go in answering Professors
-%Rice and SpiegelhalterÕs penultimate question.
-%With g(?) a normal distribution, model (6.1) returns
-%us to the realm of Stein estimation, the scene of my
-%happy collaborations with Carl Morris. I continue to
-%be surprised at how much harder bumpy, nonnormal
-%models like (7.1) are to deal with. JamesÐStein estimation
-%works fine with, say, N = 10 component problems,
-%while the RobbinsÕ form of empirical Bayes appropriate
-%to (7.1) seems to require hundreds or thousands.
-%The information calculations in Efron (2008)
-%reinforce this gloomy assessment. Maybe I am trying
-%to be overly nonparametric in constructing the empirical
-%Bayes Fdr estimates, but it is hard to imagine a
-%generally satisfactory parametric formulation for (6.1).
-%Or perhaps it is just that hypothesis testing is more demanding
-%than estimation.
-%My Note: (6.1) is the general model (mu \sim g()).
-%7.1 is the case where the effects are a mixture of zero and N(-3,1).
-
-\subsubsection*{Calibration of posterior intervals}
-
-To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over 100 simulations in each Scenario). 
-Specifically we examine the empirical coverage of nominal lower 95\% credible bounds for a) all observations; b) significant negative discoveries; c) significant positive discoveries.
-We examine only lower bounds because the results for upper bounds follow by
-symmetry (except for the one asymmetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
-in each case: for negative discoveries the lower bound is typically large and negative and limits how big (in absolute value) 
-the effect could be; for positive discoveries the lower bound 
-is positive, and limits how small (in absolute value) the effect could be. Intuitively, the lower bound for negative discoveries depends on the accuracy of $g$ in its tail,
-whereas for positive discoveries it is more dependent on the accuracy of $g$ in the center.
-
-The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
-as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
-
-% As noted above, the penalty term (\ref{eqn:penalty}) causes systematic over-estimation of the mass of $g$ near 0,  which one would expect to overshrink interval estimates towards zero. 
-%The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound is negative and overshrinking (moving it too far towards zero) makes the left tail too big; conversely, for positive discoveries, the lower confidence bound is positive, and
-%overshrinking (moving it too far towards zero) makes the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
-%and be prepared to put up with some overshrinkage to avoid it.
-
-%Coverage rates are given in:
-%\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
-
-%Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all.tex}
-	\caption{All observations.
-Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario.
-This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero.
-Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (Table \ref{tab:nopen}).
-Removing the penalty in the half-uniform case is not recommended (see Appendix).}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg.tex}
-\caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
-and the normal-based method in the flat-top scenario.
-These results likely reflect inaccurate estimates of the tails of $g$ due to a disconnect between the tail of $g$ and the component distributions in these cases.
-For example, the uniform methods sometimes substantially underestimate the length of the tail of $g$ in these long-tailed scenarios,
-causing over-shrinkage of the tail toward 0.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos.tex}
-\caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds} \label{tab:coverage}
-\end{table}
-
-%
-%\subsubsection*{Reverse asymptotics}
-%
-%Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
-%
-%Note: we'd like to make sure it behaves ok in the case where most observations are null.
-%Perhaps simulate with 999 null and 1 alternative?
-%Also maybe look at performance for small sample sizes - e.g. n=1?
-%
-%Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
-%can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
-%by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
-%
-
 \subsection*{Differing measurement precision across units}
@@
 and the other half are normally distributed with standard deviation 1:
+
 \begin{equation}
@@
 for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).
-
-\begin{figure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_hist.pdf} 
-    \label{fig:goodpoor-hist}
-    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations}
- \end{subfigure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_scatter.pdf}
-    \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) of a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
-\end{subfigure}
-\caption{Simulation illustrating how, for existing FDR methods, 
-poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
-$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
-a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
-In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
-\end{figure}
-
- \subsubsection*{Reordering of significance, and the ``$p$ value prior"}
-
-Another consequence of accounting for differences in measurement precision is that \ashr may re-order the significance of the observations compared
- with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
- We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measuments that have the same $p$ value.
- The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
- and so the $\lfsr$ for these poor-precision measurements is always high.
-
-The potential for Bayesian analyses to re-order the significance of observations, and specifically to downweight imprecise observations,
-was previously discussed in \cite{guan.stephens.08}.
-However, Wakefield \cite{wakefield:2009} showed that, under a certain prior assumption, the Bayesian analysis produces the same ranking of significance as $p$ values (or their $z$ scores).
-He named this prior the ``$p$-value prior" because it can be thought of as the
-implicit prior assumption that is being made if we rank the significance of observations by their $p$ value.
-Wakefield's $p$-value prior assumes that the less precise effect estimates correspond to larger true effects, and specifically that they scale proportional
-to the standard errors $s_j$. More specifically still, it assumes a normal prior for the non-zero $\beta_j$ with mean 0 and variance $K s^2_j$ for some constant $K$.
-Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. Specifically, if, instead of assuming that $\beta_j$ is independent
-of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
-then the $\lfsr$ computed by ash.n provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
-as is illustrated in Figure \ref{fig:lfsr_pval}, right panel. (This result does not hold when using the mixtures of uniforms prior, ash.u.)
-
-This $p$-value prior assumes that the $z$ scores $z_j=\beta_j/s_j$
-are identically distributed, independent of $s_j$. This is essentially the assumption made, implicitly or explicitly, by existing methods -- like \locfdr, \mixfdr and \qvalue --
-that model the $z_j$ or $p_j$ directly.  In contrast, we have assumed up to now that the $\beta_j$ are independent of $s_j$. 
-We can set both these assumptions within a more general framework, which allows that $\beta_j/s^\alpha$ is independent of $s_j$ for some $\alpha$ [Equation (\ref{eqn:beta-alpha}))].
-Setting $\alpha=0$ implies that $\beta_j$ is
-independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
-effects (in absolute value). This latter assumption may often be qualitatively plausible: 
-for example, in gene expression studies the standard error for gene $j$ depends partly on
-the variance of its expression among samples, and genes with a larger variance may tend to be less tightly regulated and
-so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
-On the other hand, there is no particular reason to expect that either $\alpha=1$ or $\alpha=0$ will be the optimal choice.
-Indeed, optimal choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$,
-which will be dataset-specific. 
-Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
-comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
-likelihood-based methods can be used to select $\alpha$. For example, following the logic of the EB approach it
-would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
-which has been implemented in our software by C. Dai.
-
-%\begin{result}
-%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
-%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
-%\end{result}
-
-%To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
-%Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
-%In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
-%Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
-%which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
-%(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
-%may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)
-
-%\begin{result}
-%Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and using a mixture of normals for the prior distribution 
-%(\ref{eqn:beta-alpha}) with $\alpha=1$, 
-%the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$.
-Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$)
-%is montonic with $|t_j|$. 
-%\end{result}
-
-%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
-%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.
-
-%Most previous authors of EB approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
-%have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
-%In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
-%This is analogous
-%to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
-%standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
-%that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.
-
-\begin{figure}
-\centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
-\caption{Figure illustrating affects of prior assumptions on re-ordering of significance.
-Left panel shows results under our ``default prior" which assumes that effects $\beta_j$ are identically distributed, independent of $s_j$.
-Right panel shows results under the ``$p$-value prior", which assumes that $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$.} \label{fig:lfsr_pval}
-\end{figure}
-
-%\subsubsection*{Adaptive Shrinkage}
-%
-%Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
-%when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
-%the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
-% $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
-% In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
-% Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
-% and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
-%
-%Our methods we build on this idea in two ways. First,
-%by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
-%Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
-%adapts to measurement precision: more precise measurements undergo
-% less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
-%than observations with
-%small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
-%To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
-%Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
-%
-%\begin{figure}[!ht]
-%\begin{center}
-%\includegraphics[width=400px]{../examples/ash.pdf} 
-%\end{center}
-%\caption{Figure illustrating adaptive shrinkage.
-Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0.
-The standard error of each observation, $s_j$, was simulated from Inverse Gamma(5,5), which resulted in $s_j$ varying from 0.34 to 6.7.
-The shrunken estimates ($y$ axis) are plotted against the observed estimates ($x$ axis), with color indicating the (square root of the) precision of each measurement (blue=lower-precision; red=higher-precision).
-The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
-%\end{figure}
-
-%DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
-%To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
-Because the prior distribution is unimodal about zero,
-%the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
-%amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
-%For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
-%than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
-%small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
-
-%previous work based on normal means:
-%http://link.springer.com/article/10.1007/BF02562623
-
-%\subsection{Null behaviour}
-
-%Results of simulations under the null.
-
-%\subsection{Comparison of different underlying component distributions}
-%
-%Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
-%than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
-%likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
-%For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
-%increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
-%expect $L_{hu}$ to be significantly larger than the others.
-%
-%Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
-%
-%\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
-%\begin{center}
-%\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
-%%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
-%\end{center}
-%\caption{}
-%\end{figure}
-
-%Notes from Adrian Raftery - 
-%Newcombe (1886)
-%Dean and Raftery
-%Uniform is super-efficient
-
-\section*{Discussion}
-
-We have presented an Empirical Bayes approach to large-scale multiple testing that emphasises two ideas.
-First, we emphasise the potential benefits of using two numbers ($\bhat$, and its standard error)
-rather than just one number (a $p$ value or $z$ score)  to summarize the information on each test.
-While requiring two numbers is slightly more onorous than requiring one, in many settings these
-numbers are easily available and if so we argue it makes sense to use them.
-Second, we note the potential benefits -- both statistical and computational -- of assuming that the effects come from a unimodal distribution, and provide
-flexible implementations for performing inference under this assumption. We also introduce the ``false sign rate" as
-an alternative measure of error to the FDR, and illustrate its improved robustness to errors in model fit, particularly mis-estimation
-of the proportion of null tests, $\pi_0$.
-
-Multiple testing is often referred to as a ``problem" or a ``burden". In our opinion, EB approaches turn this idea on its head,
-treating multiple testing as an {\it opportunity}: specifically, an opportunity to learn about the prior distributions, and other modelling assumptions,
-to improve inference and make informed decisions about significance (see also \cite{greenland1991empirical}). This view
-also emphasises that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
-Indeed, the FDR at a given fixed threshold does not depend on the number of tests: as the number of tests increases, both the true positives and false positives increase linearly,
-and the FDR remains the same.
-(If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example.
-The EB approach captures this dependence in an intuitive way:
-if there are lots of strong signals then we infer $\pi_0$ to be small, and the estimated FDR (or $\lfdr$, or $\lfsr$) at a given threshold may be low, even if a large number of tests were performed; and conversely
-if there are no strong signals then we infer $\pi_0$ to be large and the FDR at the same threshold may be high, even if relatively few tests were performed. 
-More generally, overall signal strength is reflected in the estimated $g$, which in turn influences the estimated FDR.
-
-Two important practical issues that we have not addressed here are correlations among tests,
-and the potential for deviations from the theoretical null distributions of test statistics. These two
-issues are connected: specifically, unmeasured confounding factors can cause both correlations among tests
-and deviations from the theoretical null \cite{efron2007correlation,leek:2007}. And although there are certainly other
-factors that could cause dependence among tests, unmeasured confounders are perhaps the most worrisome in practice
-because they can induce strong correlations among large numbers of tests and profoundly impact results.
-Approaches to deal with unmeasured confounders can be largely divided into two types: those
-that simply attempt to correct for the resulting inflation of test statistics \cite{devlin1999genomic,efron2004large}, and those
-that attempt to infer counfounders using clustering, principal components analysis, or factor models 
-\cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
-and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable,
-they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alterntively, our methods
-could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:lik}) using
-$\bhat_j | \shat_j \sim N(0,\lambda_1 \shat_j + \lambda_2)$, where $\lambda_1,\lambda_2$ are to be estimated. However,
-this immediately raises issues of identifiability and we do not pursue the idea further here.
-
-Another important practical issue is the challenge of small sample sizes. For example, in genomics applications researchers sometimes
-attempt to identify differences between two conditions based on only a handful of samples in each. In such settings the
-normal likelihood approximation (\ref{eqn:normlik}) will be inadequate. And, although the $t$ likelihood (\ref{eqn:tlik}) partially
-addresses this issue, it is also, it turns out, not entirely satisfactory.  The root of the problem is that, with small sample sizes, raw estimated standard
-errors $\shat_j$ can be horribly variable. In genomics it is routine to address this issue by applying EB methods \cite{smyth:2004} to ``moderate" (i.e.~shrink) variance estimates,
-before computing $p$ values from ``moderated" test statistics. We are currently investigating how
-our methods should incorporate such ``moderated" variance estimates to make it applicable to small sample settings.
-
-Our approach involves compromises between flexibility, generality, and simplicity on the one hand, and statistical efficiency and principle on the other.
-For example, in using an EB approach that uses a point estimate for $g$, rather than a fully Bayes approach that accounts for uncertainty in $g$,
-we have opted for simplicity over statistical principle. And in summarizing every test by
-two numbers and making a normal or $t$ approximation to the likelihood, we have aimed to produce generic methods that 
- can be applied whenever such summary data are available -- just as \qvalue can be applied to any set of $p$ values for example -- although
- possibly at the expense of statistical efficiency compared with developing multiple tailored approaches based on context-specific likelihoods. 
- Any attempt to produce generic methods will involve compromise between generality and efficiency.
- In genomics, many analyses -- not only FDR-based analyses -- involve first computing a series of $p$ values before
- subjecting them to some further downstream analysis. An important message here is 
- that working with two numbers ($\bhat_j,\shat_j$), rather than one ($p_j$ or $z_j$), 
- can yield substantial gains in functionality (e.g.~ estimating effect sizes, as well as testing; accounting for variations in
-measurement precision across units) while losing only a little in generality.  We hope that our work will
-encourate development of methods that exploit this idea in other contexts.
-
-%This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
-%from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
-%with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
-
-%
-%
-%Both \locfdr and \mixfdr allow for deviations from
-%the theoretical null distribution by allowing that, under the null, the $z$ scores are $N(0,\sigma^2)$ with $\sigma$ estimated from the data rather
-%than fixed to 1. Within our approach this idea could be incorporated by changing the likelihood - for example, to assume $\bhat_j \sim N(0,s_ 
-%
-%These are tricky issues that, as far as we are aware, can affect
-%all approaches to FDR analysis. An important focus of Efron's work, also implemented in \locfdr, 
-% is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
-%theoretical null distribution. 
-%
-%For example, we and others have worked on Bayesian approaches to genetic association testing, using subjective prior specification of
-%effect sizes based on external data (e.g.~\cite{stephens:2009}). However, the large numbers of tests brings the opportunity to learn about
-%priors, and here we aim to exploit this
-
-%Choice of mixture components. Use of likelihood to compare different mixture components?
-
-%  
-%Directly modeling the p values, or z scores, say via non-parametric methods, 
-%can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-%distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-%The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-%Before proceeding, we feel obliged to correct the common misconception
-% that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
-% not depend on the number of tests. 
-%  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
-% 
-% In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
-
-\section*{Detailed Methods} \label{sec:implement}
-
-\subsection*{Embellishments}
-
-\subsubsection*{More flexible unimodal distributions}
-
-Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
-Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
- Therefore, we have implemented a more general approach based on 
-\begin{equation} \label{eqn:g-general}
-g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) ,
-\end{equation}
-where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
-\begin{enumerate}[(i)]
-\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$, \qquad (``ash.n")
-\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,  \qquad (``ash.u")
-\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,  \qquad (``ash.hu")
-\end{enumerate}
-where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
-(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
-and $\pi$ is a $2K+1$ vector that sums to 1.)
-The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
-Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
-under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
-(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
-(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
-however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)
-
-Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
-using a large dense grid of $\sigma^2_k$ or $a_k$, (i)-(iii) can respectively approximate,
-with arbitrary accuracy,
-\begin{enumerate}[(i)]
-\item any scale mixture of normals, which includes as special cases
-the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
-\item any symmetric unimodal distribution about 0.
-\item any unimodal distribution about 0.
-\end{enumerate}
-The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
-In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.
-
-Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
-The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
-See Implementation for details of our software defaults.
-
-%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting.
-For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
-%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
-
-\subsubsection*{Dependence of effects on standard errors}
-
-Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
-This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
- \begin{equation} \label{eqn:beta-alpha}
- \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
- \end{equation}
-for any $\alpha$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
-assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
-it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
-the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.
-
-The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
-To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
- standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
- Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
-yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
-%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
-%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
-%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
-
-%Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
-%how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
-%it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
-%rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
-%and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
-%this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
-%do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
-
-\subsubsection*{Replace normal likelihood with $t$ likelihood}
-
-We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
-  \begin{equation} \label{eqn:tlik}
- \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
- \end{equation}
-where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
-and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
-The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
-This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms; see Implementation below.
-When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
-
-Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
-\begin{equation} \label{eqn:tdist}
-(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
-\end{equation}
-However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
-In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
-currently pursuing this improved approach (joint work with M.Lu) and results will be published elsewhere. 
-%We note here that
-%modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
-%where our simpler approach can produce unsatisfactory results.
-
-\subsubsection*{Non-zero mode}
-
-An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood, rather than fixed to 0.
-
-\subsection*{Implementation Details}
-
-\subsubsection*{Likelihood for $\pi$}
-
-We define the likelihood for $\pi$ to be the probability of the observed data $\bhat$ conditional on $\shat$: $L(\pi) := p(\bhat | \shat,\pi),$ which
-by our conditional indendence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$.
-[One might prefer to define the likelihood as $p(\bhat, \shat | \pi) = p(\bhat | \shat, \pi) p(\shat | \pi)$, in which case our definition comes down to assuming that the term $p(\shat | \pi)$ does not depend on $\pi$.]
-
-Using the prior $\beta_j \sim \sum_{k=0}^K \pi_k f_k(\beta_j)$ given by (\ref{eqn:g-general}), and the normal likelihood (\ref{eqn:normlik}), integrating over $\beta_j$ yields
-\begin{equation}
-p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
-\end{equation}
-where 
-\begin{equation}
-\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
-\end{equation}
-denotes the convolution of $f_k$ with a normal density.
-These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
-Specifically, 
-\begin{equation} \label{eqn:uconv}
-\tilde{f}_k(\bhat_j)  = 
-\begin{cases}
-N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
-\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
-\end{cases}
-\end{equation}
-where $\Psi$ denotes the cumulative distribution function (c.d.f.) of the standard normal distribution.
-If we replace the normal likelihood with the $t_\nu$ likelihood (\ref{eqn:tlik}) then the convolution 
-for $f_k$ uniform the convolution is still given by (\ref{eqn:uconv}) but with $\Psi$ the c.d.f. 
-of the $t_\nu$ distribution function. (The convolution for $f_k$ normal is tricky and we have not implemented it.)
-
-\subsubsection*{Penalty term on $\pi$}
-
-To make $\lfdr$ and $\lfsr$ estimates from our method ``conservative" we add a penalty 
-term $log(h(\pi;\lambda))$ to the log-likelihood $\log L(\pi)$ to encourage over-estimation of $\pi_0$:
-\begin{equation} \label{eqn:penalty}
-h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
-\end{equation}
-where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
-consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0sims}).
-
-Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
-we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
-
-\subsubsection*{Problems with removing the penalty term in the half-uniform case}
-
-It is straightforward to remove the penalty term by setting $\lambda_k=1$ in (\ref{eqn:penalty}).
-We note here an unanticipated problem we came across when using no penalty term in the half-uniform case 
-(i.e.~ $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$ in (\ref{eqn:g-general})): when the data
-are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
-sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
-for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
-and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
-
-\subsubsection*{EM algorithm}
-
-With this in place, the penalized log-likelihood for $\pi$ is given by:
-\begin{equation}
-\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
-\end{equation}
-where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This can be maximized using an EM algorithm \cite{dempster77}, whose
-one-step updates are:
-\begin{align} \label{eqn:w}
-w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
-n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
-\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}
-\end{align}
-Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
-and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
-probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$. (The $w_{kj}$ are sometimes referred to as the ``responsibilities".)
-Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
-We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple} to accelerate convergence of this EM algorithm.
-
-%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
-\subsubsection*{Initialization}
-
-By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
-Our rationale for initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
-but weak signal in the data cannot quickly draw the algorithm towards the null.
-
-In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
-$\pi_0=1; \pi_k=0 (k>0)$. If not then we replace the EM solution with the global null solution. This aims to guard against errors due to convergence to a local optimum when the
-data are consistent with the global null.
-
-\subsubsection*{Conditional distributions}
-
-Given $\hat\pi$, we compute the conditional distributions 
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
-\end{equation} 
-Each posterior is a mixture on $K+1$ components:
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
-\end{equation}
-where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
-and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
-$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
-All these posterior distributions are easily available.
-For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
-If $f_k$ is normal and $L$ is normal, then this is a normal distribution.
-
-\subsubsection*{Choice of grid for $\sigma_k, a_k$} \label{sec:grid}
-
-\def\sigmamax{\sigma_{\text{max}}}
-\def\sigmamin{\sigma_{\text{min}}}
-When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
-in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
-as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
-based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
-is negative we set $\sigmamax = 8\sigmamin$.)
-
-When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.
-
-Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
-a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.
-
-\appendix
-
-\begin{table}[!ht]
-\centering\begin{tabular}{c c } \toprule
-Scenario & Alternative distribution, $g_1$  \\ \midrule
-spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
-near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
-flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
- &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
-skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
-big-normal & $N(0,4^2)$ \\ 
-bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
-\end{tabular}
-\caption{Summary of simulation scenarios considered} \label{table:scenarios}
-\end{table}
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all_nopen.tex}
-	\caption{All observations}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg_nopen.tex}
-\caption{``Significant" negative discoveries.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos_nopen.tex}
-\caption{``Significant" positive discoveries.}
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{table:nopen}
-\end{table}
-%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
-%\begin{equation}
-%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
-%\end{equation}
-%which is close to the definition in \cite{storey.03}.
-
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgements}
-
-Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
-prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} , and includes contributions 
-from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.
-
-This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.
-
-%\section*{References}
-% The bibtex filename
-\bibliography{/Users/stephens/Dropbox/Documents/mainbib}
-
-\section*{Figure Legends}
-
-\clearpage
-
-\section*{Tables}
-
-\section*{Supporting Information Legends}
-
-Supplementary material can be found in {\bf Supplementary Information S1.}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
-\end{document}

### 2015-10-06 subsubsectionCalibra.tex
New Article!

--- 2015-10-06 paper/main.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
-\documentclass[11pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-%\usepackage{natbib}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-\usepackage{booktabs}
-\usepackage{xspace}
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-\usepackage{sectsty}
-\subsectionfont{\normalsize}
-\subsubsectionfont{\mdseries\itshape}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\usepackage{color}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-% Use the PLoS provided bibtex style
-\bibliographystyle{/Users/stephens/Dropbox/Documents/stylefiles/plos2009}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
-\pagestyle{myheadings}
-%% ** EDIT HERE **
-\usepackage{enumerate}
-\usepackage{multirow} 
-\usepackage{url}
-\usepackage{xr} %for cross-referencing
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-\newtheorem{algorithm}{Algorithm}
-\newtheorem{proposition}{Proposition}
-\newtheorem{restateproposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{result}{Result}
-\newtheorem{note}{Note}
-\newtheorem{definition}{Definition}
-
-\def\lfdr{\textit{lfdr}}
-\def\lfsr{\textit{lfsr}}
-\def\x{\mbf{x}}
-\def\y{\mbf{y}}
-\def\e{\mbf{e}}
-\def\g{\mbf{g}}
-\def\p{p}
-\def\Yb{\hat{Y}_1^{\text{BAYES}}}
-\def\Xperp{X_{1 \perp 0}}
-\def\Pperp{P_{1 \perp 0}}
-\def\Pperpb{\Pperp^{\text{B}}}
-\def \hatd { \widehat{ d } }
-\def \dhatin {\hatd^{\mbox{\tiny IN}}}
-\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
-\def \mvn{\text{MVN}}
-\def\mn{\text{MN}}
-\def \iw{\text{W}^{-1}}
-\def \fp {fastPHASE\ }
-\def\bhat{\hat{\beta}}
-\def\shat{\hat{s}}
-\def\var{\rm Var}
-\def\mvr{\text{BMVR}}
-\def\bfuni{\text{BF}_\text{uni}}
-\def\one{\mbf{1}}
-\def\zero{\mbf{0}}
-\def\hall{H_{\text{all}}}
-\def\pall{\p_{\text{all}}}
-\def\bfall{\text{BF}_\text{all}}
-\def\bfav{\text{BF}_\text{av}}
-\def\gall{\gamma_\text{all}}
-\def\limbfall{\bfall^\rightarrow}
-\def\limbfav{\bfav^\rightarrow}
-\def\limbfgamma{\BF_\gamma^\rightarrow}
-\def\sxx{S_{xx}}
-\def\syy{S_{yy}}
-\def\vyy{V_{yy}}
-\def\vyx{V_{yx}}
-\def\vxx{{V_{xx}}}
-\def\rss{\text{\rm RSS}}
-\def\dss{\text{\rm $\Delta$SS}}
-\def\U{U}
-\def\D{D}
-\def\I{I}
-\def\luu{L_{UU}}
-\def\indep{\perp}
-\def\qvalue{{\tt qvalue}\xspace}
-\def\locfdr{{\tt locfdr}\xspace}
-\def\mixfdr{{\tt mixfdr}\xspace}
-\def\ashr{{\tt ashr}\xspace}
-
-\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
-                                % mode too
-\newcommand\height{{\it height}}
-\newcommand\weight{{\it weight}}
-\def\BF{\text{BF}}
-
-%% END MACROS SECTION
-%\externaldocument{plos_one_rev1_suppinfo}
-
-\begin{document}
-
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{False Discovery Rates: A New Deal}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Matthew Stephens$^{1*}$, 
-\\
-\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
-\\
-$\ast$ E-mail: Corresponding mstephens@uchicago.edu
-\end{flushleft}
-
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-%\section*{Author Summary}
-
-\section*{Introduction}
-
-Since its introduction in 1995 by Benjamini and Hochberg 
-\cite{benjamini1995controlling}, the ``False Discovery Rate" (FDR) has quickly established itself
-as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing
-in which the goal is to identify the non-zero ``effects" among a large number of imprecisely-measured effects.
-
-Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
-indeed, the notion that EB approaches could be helpful in handling multiple comparisons predates introduction of
-the FDR  (e.g.~\cite{greenland1991empirical}). More recently, EB approaches to the FDR
-have been extensively studied by several authors, 
-especially B.~Efron and co-authors 
-\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
-see also \cite{kendziorski2003parametric,newton2004detecting, 
-datta2005empirical,muralidharan2010empirical} for example.
-So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
-existing widely-used FDR pipelines) and can substantially affect
-inference. The first idea is to {\it assume that the distribution of effects is unimodal}. This yields a very simple, fast, and stable computer 
-implementation, as well as improving inference of FDR when the unimodal assumption is correct.  
-The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
-to summarize each measurement. This idea allows variations in measurement precision to be better accounted for,
- and avoids a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
-
-In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
-the idea that it may be preferable to focus on estimation rather than on testing.
-In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
-simply estimation with some positive prior probability that the effect is exactly zero.
-However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
-perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
-controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
-``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
-in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
-the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the lfsr, particularly its increased robustness to modelling assumptions.
-
-%terminology
-%``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
-%and correspondingly define the ``false signed discovery rate", or simply ``false sign rate" (FSR), in a natural way.
-
-%The methods we provide here can address both problems, and we provide results that highlight important benefits of focussing on estimation, and particularly on assessing the correctness of the {\it sign} of the estimated effects, rather than on testing for non-zero effects.
-
-Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
-(We address the reasons for this name, and connections with shrinkage analysis, in the Discussion.)
-
-%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.
-
-%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-%measurements are made on only a small number of
-%samples in each condition - sometimes as few as one
-%sample in each condition. Thus the error in estimates
-%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-%in estimates of $\beta_j$ still greater.
-%
-
-%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
-%Here we present new tools that bring two important ideas to bear on the problem.
-%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
-%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
-%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
-%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
-%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
-%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
-%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
-%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
-%a substantial increase in the number of discoveries at a given FDR).
-
-%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
-%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
-Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
-%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
-%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
-%generic methods that work directly with two numbers for each 
-%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
-Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
-%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
-
-\def\df{df}
-\def\FDR{\text{FDR}}
-\def\fdr{\text{\lfdr}}
-\def\FSR{\text{FSR}}
-\def\fsr{\text{lfsr}}
-
- \section*{Methods}
-
- \subsection*{Model Outine}
-
-% Here we consider one setting in which FDR-based methods are widely used:
-%an experimenter has measured a large numbers of ``effects", each with an associated
-%measurement precision (standard error), 
-%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
-%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
-%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
-%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
-%For example, in genomics
-%a common goal is to identify ``differentially-expressed genes" -- that is, genes
-%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
-%mean expression level between the two conditions, often referred to as the ``log-fold change". 
-Here we describe the simplest version of our method, and briefly discuss embellishments we have also implemented.
-
-Suppose that we are interested in the values of $J$ ``effects" $\beta=(\beta_1,\dots,\beta_J)$. For example, in a typical genomics application
-that aims to identify differentially expressed genes,  $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
-In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
-that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
-more general problem of estimating, and assessing uncertainty in, $\beta_j$.
-
-Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
-and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
-which by Bayes theorem can be written as 
-\begin{equation}
-p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
-\end{equation}
-
-For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
-This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
-A simple way to implement the UA is to assume that $g$ is
-a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
-\begin{align} \label{eqn:beta}
-p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
-g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
-\end{align}
-where $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyperparameters, which are non-negative and sum to one, and are to be estimated,
-while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
-represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
-(We encourate the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
-although of course in practice we use a finite grid -- see Implementation Details.)
-
-For the likelihood $p(\bhat | \beta, \shat)$ we assume  
-\begin{equation} \label{eqn:normlik}
-p(\bhat | \beta, \shat) = \prod_j N(\bhat_j; \beta_j, \shat_j^2).
-\end{equation}
-Here, in additional to some conditional independence assumptions, we are 
-effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-
-This simple model features both the key ideas we want to emphasise in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
-the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
-with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
-However, this simple model also has several additional assumptions that can be relaxed. Specifically,
-\begin{enumerate}
-\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) also implies that $g$ is symmetric about 0; more flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see (\ref{eqn:g-general})].
-\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
-We can relax this to allow for a relationship between these quantities [see (\ref{eqn:beta-alpha})].
-\item The likelihood (\ref{eqn:normlik}) effectively assumes that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-We can generalize this likelihood using a $t$ likelihood [see (\ref{eqn:tlik})].
-\end{enumerate}
-These embellishments are detailed in Embellishments and Implementation Details.
-Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
-(which are also made by most existing EB approaches to this problem). Correlations among tests certainly arise in practice, either due to genuine correlations
-in the system of study, or due to unmeasured confounders, and their potential to impact results of an FDR analysis is important to consider
-whatever analysis methods are used: see \cite{efron2007correlation,leek:2007} for relevant discussion.
-
-%We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
-%$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
-%\begin{equation}
-%p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
-%\end{equation}
-%Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
-%\begin{equation} \label{eqn:normlik}
-%p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
-%
-%We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
-%as a mixture of null and alternative components:
-%\begin{equation} \label{eqn:betaj0}
-%g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
-%\end{equation}
-%where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
-%We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
-%\begin{equation} \label{eqn:beta-iid}
-%p(\beta | \shat) = \prod_j g(\beta_j).
-%\end{equation}
-%
-%Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
-%with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
-%to use a mixture of zero-mean normal distributions:
-%\begin{equation} \label{eqn:mixnorm}
-%g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
-%Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
-%represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
-% The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
-% more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below. 
-% 
-% Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
-% yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
-% \begin{equation} \label{eqn:g}
-% g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
-% \end{equation}
-% 
-% Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
-% \begin{align} \label{eqn:betaj}
-% p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
-% p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
-% \end{align}
-% with $g$ given by the mixture of normals $(\ref{eqn:g})$.
-
-\subsection*{Fitting the model}
-
-In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
-each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
-Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
-\begin{equation} \label{eqn:marginallik}
-p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
-\end{equation}
-where we define $\sigma_0:=0$.
-
-The usual EB approach to fitting this model would involve two simple steps:
-\begin{enumerate}
-\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
-\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
-against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | | \bhat, \shat,\hat{\pi})$.
-\end{enumerate}
-Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(The simplicity of the EM algorithm in step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
-which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
-
-Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximise a penalized
-likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
-We introduce this penalty because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
-
-\subsection*{The local False Discovery Rate and local False Sign Rate}
-
-As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
-In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
-One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
-the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
-In other words it is the posterior probability that $\beta_j$ is actually zero:
-\begin{equation} \label{eqn:fdr}
-\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
-\end{equation}
-
-The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
-whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
-the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey1991philosophy}) argued that
-``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$.
-Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
-\begin{quote}
-...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
-\end{quote}
-Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
-focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.
-
-Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
-would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
-\begin{equation} \label{eqn:lfsr}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
-\end{equation}
-To illustrate, suppose that 
-\begin{gather*}
-p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
-p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
-p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
-\end{gather*} 
-Then 
-from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
-our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.
-
-As our notation suggests, $\lfsr_j$ is intended to be compared and contrasted with $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
-small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
-Of coure, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
-this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
-(this follows from the definition because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
-In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
-$\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.
-
-From these ``local" measures of significance, we can also compute average error rates over subsets of observations
-$\Gamma \subset \{1,\dots,J\}$. For example,   
-\begin{equation} \label{eqn:FDRhat}
-\widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
-\end{equation}
-estimates the FDR we would obtain if we were to declare all tests in $\Gamma$ significant.
-And 
-\begin{equation} \label{eqn:qvalue}
-q_j := \widehat \FDR(\{k: \fdr_k \leq \fdr_j\}) 
-\end{equation}
-provides a measure of significance analogous to Storey's $q$ value \cite{storey.03}.
-
-%Because our methods involve explicitly modelling effects $\beta_j$, rather than 
-%$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
-
-%Indeed, even in settings where some
-%$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
-%analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
-
-\subsection*{Related work}
-
-\subsubsection*{Previous approaches focussed on FDR}
-
-Among previous methods that explicitly consider the FDR and related quantities,
-our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
- (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
-  %Here we highlight the key differences between our approach and these three existing approaches.
-
-There are two key differences between our approach and all of these three existing methods. First, 
-whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
-we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
-for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
-Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 
-how it can improve performance by taking better account of variation in measurement precision among observations.
-%(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
-% Or more generally Section xx illustrates the potential practical effects
-
-%allows for different modelling assumptions to be compared through likelihood comparisons.
-Second, our unimodal assumption (UA) that the effects are unimodal about zero is quite different from assumptions made by $\qvalue, \locfdr$ or $\mixfdr$.
-Indeed, $\locfdr$ assumes that all $z$ scores near 0 are null (Efron calls this the Zero Assumption; ZA), which implies that 
-under the alternative hypothesis the distribution of $z$ scores has {\it no mass at 0}; this contrasts strikingly with the UA, which implies that this distribution has its peak at 0! 
-Similarly, $\qvalue$ assumes that all $p$ values near 1 are null,
-which is the same as the ZA because $p$ values near 1 correspond to $z$ scores near 0. And although \mixfdr does not formally make the ZA,
-we have found that in practice, with default settings, the results often approximately satisfy the ZA (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}).
-Thus, not only do these existing methods not make the UA, they actually make assumptions that are, in some sense, as different
-from the UA as they can be.
-
-Given that the UA and ZA are so different, it seems worth discussing why we generally favor the UA. 
-Although the UA will not apply to all situations, we believe that it will often be reasonable, especially in FDR-related contexts
-that have traditionally focussed on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis,  
-it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible. Further, it seems reasonable to expect that larger effects become
-decreasingly plausible, and so the distribution of the effects will be unimodal about 0.
-To paraphrase Tukey, ``All we know about the world teaches us that large effects are rare, whereas small effects abound."
-We emphasise that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
-It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
-
-%Tukey "peri-null"
-
-%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
-%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}.
-
-In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficents - 
-common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
-These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
-it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
-The UA assumption on regression coefficients is directly analagous to our UA here, and so its
-widespread use in the regression context supports its use here.
-
-Alternatively, we could motivate the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
-such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
-make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).
-
-Finally, the UA also has a considerable practical benefit: it yields simple algorithms that are both computationally
-and statistically stable. We illustrate these features in Results.
-
-\subsubsection*{Other work}
-
-There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
- Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
-indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
-This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
-Also, they provide no software implementation.
-More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
-Some pointers to recent papers in which other relevant
-citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}.
-Much of the literature focusses on the homoskedastic case (i.e.
-$\shat_j$ all equal) whereas we allow for heteroskedasticity.
-And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential.
-Several recent papers consider more flexible non-parametric assumptions on
-$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
-estimate (NPMLE) for $g$.
-These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive.
-However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
-may perform well for point estimation, it seems possible it will not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
-Indeed, one way of thinking about the UA is as a way to regularize $g$.
-
-%The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
-%encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-%$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-%$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-%the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-%observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-%corresondingly little effect on the estimated FDR of other observations. 
-%We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-%$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-%that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-%estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-%Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-\def\dt{\text{dt}}
-
-%
-%assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
-%$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
-% $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
-%  \begin{equation} \label{eqn:lik}
-% \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
-% \end{equation}
-% where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e.
-the density of $\mu+\sigma T_\nu$).
-In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
-%  \begin{equation} \label{eqn:normlik}
-% \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
-% \end{equation}
-%which, in light of (\ref{eqn:betahat}), is equivalent to
- % \begin{equation} \label{eqn:lik2}
-% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
- %\end{equation}
-
-%The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
-they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
-The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$.
-Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$.
-If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$.
-See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.
-
-%This likelihood approximation can be motivated by assuming that i) $\bhat$
-% $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
-% $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$. 
-%
-
-%Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.
-
-%While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
- %that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
-%there is an important difference: here,
-
-%The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
-For example, we can write $p(\bhat_j, \shat_j | \beta_j) = p(\bhat_j | \shat_j, \beta_j)p(\shat_j | \beta_j)$.
-Now, if we are willing to assume that $\shat_j$ alone contains no information about $\beta_j$, or equivalently that $p(\shat_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\bhat_j | \shat_j, \beta_j)$, and if $\bhat_j | \shat_j, \beta_j \sim N(\beta_j, \shat_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.
-
-\section*{Results}
-
-We compare results of \ashr with existing FDR-based methods implemented
-in the R packages \qvalue (v2.0 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
-In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
-to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.
-
-\subsection*{Effects of the Unimodal Assumption}
-
-Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
-standard error, $s_j=1$ and all methods are provided that information. That is,
-$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$.
-In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the only difference between our method and methods like $\locfdr$ and $\mixfdr$ are in how they estimate $g$.
-
-To briefly summarize the results in this section:
-\begin{enumerate}
-\item The UA can produce very different inferences compared with the ZA made by existing methods.
-\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
-\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
-\end{enumerate}
-
-\subsubsection*{The UA and ZA can produce different inferences}
-
-\begin{figure}
-\center \includegraphics[height=6in]{figures/decomp_ZA.pdf}
-\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
-The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
-In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
-The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
-(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
-\end{figure}
-
-To illustrate the different inferences from the UA and ZA we show results for a single
-dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
-but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
-\qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
-The results (Figure \ref{fig:ZA}) illustrate the clear difference between the existing methods and our method. The effects of the ZA made by \qvalue and \locfdr
-are visually clear, producing a ``hole"  in the alternative $z$ score distribution around 0. Although \mixfdr does not formally make the ZA, its decomposition
-exhibits a similar hole. In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
-effectively ``filling in" the hole.  (Of course the null distribution also has a peak at 0, and the local fdr 
- under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)
-
-Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA.
-Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
-This would increase the null component (dark blue) at the expense of
-the alternative component (light blue).
-Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0.
-The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears.
-In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
-See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
-demonstration.
-
- \subsubsection*{The UA can produce conservative estimates of $\pi_0$}
-
-The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
-Consequently \ashr will estimate smaller $\lfdr$s and FDRs than existing methods that make the ZA. 
-This is desirable, provided that these estimates remain conservative: that is,
-provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
-The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behaviour. To check its effectiveness
-we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
-for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
-They range from a ``spiky" distribution -- where many non-zero $\beta$ are
-too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
-flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
-making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
-which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.
-
-% How do these differences influence estimates of the FDR? To address this, note that  -- in this simple case where the 
-% $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
-% of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
-% strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
-% although the methods considered here differ from one another in many ways, having estimated $\pi_0$
-% they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
-% empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
-%
-
-%Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
-% and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
-%Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
-%At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
-%overestimates of $\pi_0$, which will yield overestimates of the FDR).
-
-\begin{figure} 
-\begin{center}
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=1.8in]{figures/scenarios_density.pdf} 
-	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/pi0_est.pdf} 
-\caption{Comparison of true and estimated values of $\pi_0$.
-When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ashr being least conservative, and hence most accurate.
-When the UA does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.} \label{fig:pi0sims}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfdr_est.png} 
-\caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
-Black line is $y=x$ and red line is $y=2x$.
-Estimates of $\lfdr$ are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est.png} 
-\caption{As in c), but for $\lfsr$ instead of $\lfdr$.
-Estimates of $\lfsr$ are consistently less conservative than $\lfdr$ when UA holds, and also less anti-conservative in bimodal scenario.}  \label{fig:lfsr}
-\end{subfigure}
-\end{center}
-\caption{Results of simulation studies (constant precision $s_j=1$).}
-\end{figure}
-
-For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
-data as follows:
-\begin{enumerate}
-\item Simulate $\pi_0  \sim U[0,1]$.
-\item For $j=1,\dots,J$, simulate $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)$.
-\item For $j=1,\dots,J$, simulate $\bhat_j | \beta_j \sim N(\beta_j,1)$.
-\end{enumerate}
-%Thus these simulations assume the same
-%precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).
-
-Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ashr ($y$ axis) with the true values ($x$ axis). 
-For \ashr we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
-and as a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
-as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
-The results show that \ashr provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
-in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ashr estimates can be slightly anti-conservative.
-We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
-are used. (In addition, the effects on $\lfsr$ estimates turn out to be relatively modest; see below).
-
-%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
-%
-%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
-%and so all effects are non-zero and no nulls hold. This scenario is rather different
-%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
-%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
-%``mostly non-null"
-%scenario, and we would agree! On the other hand, the distribution of $p$ values
-%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
-%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
-%Thus it seems desirable to have methods that perform well in both cases, and this is
-%an important feature of the methods we introduce here.
-%
-%TODO: put FDRs on graph? Simulate with 50\% null?
-
-\subsubsection*{The $\lfsr$ is more robust than \lfdr}
-
-The results above show that \ashr can improve on existing methods in producing smaller, more accurate, 
-estimates of $\pi_0$, which will lead to more accurate estimates of FDR.
-Nonetheless, in many scenarios \ashr continues
-to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
-This is because these scenarios include 
-an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
-estimation of $\pi_0$ impossible. Put another way, and as is well known, $\pi_0$ is not identifiable:
- the data can effectively provide an upper bound on plausible values of $\pi_0$,
-but not a lower bound (because the data cannot rule out that everything is non-null, but with miniscule effects).
-To obtain conservative behaviour we must estimate $\pi_0$ by this upper bound, which can
-be substantially larger than the true value.
-
-Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
-overestimation of $\pi_0$ is corresponding overestimation of FDR (and $\lfdr$, and $q$ values).
-To illustrate, Figure \ref{fig:lfdr} compares the estimated $\lfdr$ from ash.n with the true value (computed using
-Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios which involve many
-non-zero effects that are very near 0 (e.g.~the spiky scenario with $\pi_0$ small) where
-$\pi_0$ can be grossly overestimated.  
-(Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.)
-
-The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
- become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
- This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
-To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
-be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
-the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in $\lfsr$ than $\lfdr$.
-
-Note that, compared with previous debates regarding testing, 
-this section advances an additional reason for focussing on the sign of the effect, rather than just testing whether it is 0. 
-In previous debates authors have argued against testing whether an effect is 0
-because it is {\it implausible that effects are exactly 0}. Here we add that {\it even if one believes
-that some effects may be exactly zero}, it is still better to focus on the sign, because generally {\it the data are more informative about that question}
-and so inferences are more robust to, say, the inevitable mis-estimation of $\pi_0$.
-To provide some intuition, consider an observation with a $z$ score of 0. The $\lfdr$ of this observation can range from 0 (if $\pi_0=0$)
-to 1 (if $\pi_0=1$). But, assuming a symmetric $g$, the $\lfsr>0.5$ whatever the value of $\pi_0$, because the observation $z=0$ says
-nothing about the sign of the effect.
-Thus, are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
-whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.
-
-Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
-Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
-if we are prepared to assume that no effects are exactly zero, then removing the point mass 
-yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
-However, there is ``no free lunch":  if in fact some effects are exactly zero
-then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
-We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.
-
-\begin{figure}
-\begin{center}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s_nn.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be more accurate in this case.
-That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
-\end{subfigure}
-\end{center}
-\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
-\end{figure}
-
-\subsubsection*{Numerical Stability}
-
-The EM algorithm, which we use here to fit our model, is notorious for convergence to local optima. 
-However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
-caused by such behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
-in addition running it using our default initialization procedure (see Implementation details). We then compared the largest
-log-likelihood achieved across all 11 runs with the log-likelihood achieved by the default run.
-When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference ($< 0.02$), 
-and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference ($<0.02$),
-and 6\% of runs showed an appreciable log-likelihood difference ($>1$), with the largest difference being 5.0. However, perhaps suprisingly, 
-even for this largest difference results from the default run (e.g.
-the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood.
-See \url{http://github.com/stephens999/ash/blob/master/dsc-robust/summarize_dsc_robust.rmd} for further details.
-
-\subsubsection*{The UA helps provide reliable estimates of $g$}
-
-An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
-can estimate the {\it size} of each effect $\beta_j$.
-Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
-to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
-threshold $T$.
-%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
-%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
-%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
-%towards zero, with the strength of the shrinkage being estimated from the data.
-Further, because the posterior distribution is, by definition,
-conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
-been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
-In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
-\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
-how can confidence intervals be assigned to the true
-effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
-and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
-%but also notes the difficulties of implementing a stable general algorithm.
-
-The ability of the EB approach to provide valid ``post-selection" interval estimates is extremely attractive in principle.
-But its usefulness in practice
-depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
-which are notoriously difficult in general. Indeed, Efron emphasises
-the difficulties of implementing a stable general algorithm, noting in his rejoinder
-``the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying
-to be overly nonparametric ... but it is hard to imagine a
-generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
-Our key point here is that the UA greatly simplifies the deconvolution problem.
-While not meeting Efron's desire for an entirely general nonparametric approach, we believe 
-that the UA can handle many cases of practical interest.
-
-To illustrate this, Figure \ref{fig:egcdf} compares the
-estimated $g$ from \ashr  with that from \mixfdr which does not make the UA 
-(and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
-The greater reliability of estimates afforded
-by the UA is immediately apparent. In particular the estimated cdf from \mixfdr often has an almost-vertical
-segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. 
-The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
-While the UA is not the only way to achieve this, we find it an attractive, simple and effective approach.
-
-Interestingly, even in the ``bimodal" scenario $\ashr$ 
-is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
-Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
-Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
-to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
-(\url{http://github.com/stephens999/ash/blob/master/dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
-clearly visible in the observed $\bhat$.
-
-We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ashr tends to systematically
-overestimate the mass of $g$ near zero. On careful inspection, this is apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
-and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf}b),
-and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
-under-estimating $\pi_0$ (Figure \ref{fig:egcdf}c).
-
-\begin{figure}[h!] 
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=2in]{figures/egcdf.pdf}
-	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf_nopen.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
-Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
-\end{subfigure}
-\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.} \label{fig:egcdf}
-\end{figure}
-
-%making the UA provides an attractive and simple way to achieve estimating $g$  large
-%subset of cases of interest (those where the UA can reasonably be assumed to hold).
-
-%In practice, the utility of the posterior distributions noted above depends primarily on the accuracy
-%of the estimate of the distribution $g(\cdot)$.
-
- %quote from efron rejoinder , p46:
-%MorrisÕ Section 3 is especially pertinent. His formula
-%for p(?|z) is related to my discussion of the
-%BenjaminiÐYekutieli False Coverage Rate method in
-%Section 7, particularly (7.2)Ð(7.4). Originally I had
-%hoped to develop an empirical Bayes method for estimating
-%such models, but the effort foundered on practical
-%difficulties involving the perils of deconvolution.
-%Section 6 on the Òone-group modelÓ is the ugly duckling
-%of the current paper, but it bears on some important
-%points raised in the discussion. Figure 7 concerns
-%a fuzzy version of simultaneous hypothesis testing,
-%where, as in MorrisÕ hospital example, the usual singlepoint
-%null hypotheses seem unequal to the task. The
-%development from (6.6) onwards, particularly (6.12),
-%bears on the possibility of nonnormal null distributions,
-%and is about as far as I can go in answering Professors
-%Rice and SpiegelhalterÕs penultimate question.
-%With g(?) a normal distribution, model (6.1) returns
-%us to the realm of Stein estimation, the scene of my
-%happy collaborations with Carl Morris. I continue to
-%be surprised at how much harder bumpy, nonnormal
-%models like (7.1) are to deal with. JamesÐStein estimation
-%works fine with, say, N = 10 component problems,
-%while the RobbinsÕ form of empirical Bayes appropriate
-%to (7.1) seems to require hundreds or thousands.
-%The information calculations in Efron (2008)
-%reinforce this gloomy assessment. Maybe I am trying
-%to be overly nonparametric in constructing the empirical
-%Bayes Fdr estimates, but it is hard to imagine a
-%generally satisfactory parametric formulation for (6.1).
-%Or perhaps it is just that hypothesis testing is more demanding
-%than estimation.
-%My Note: (6.1) is the general model (mu \sim g()).
-%7.1 is the case where the effects are a mixture of zero and N(-3,1).
-
 \subsubsection*{Calibration of posterior intervals}
@@
 as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
-
-% As noted above, the penalty term (\ref{eqn:penalty}) causes systematic over-estimation of the mass of $g$ near 0,  which one would expect to overshrink interval estimates towards zero. 
-%The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound is negative and overshrinking (moving it too far towards zero) makes the left tail too big; conversely, for positive discoveries, the lower confidence bound is positive, and
-%overshrinking (moving it too far towards zero) makes the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
-%and be prepared to put up with some overshrinkage to avoid it.
-
-%Coverage rates are given in:
-%\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
-
-%Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all.tex}
-	\caption{All observations.
-Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario.
-This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero.
-Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (Table \ref{tab:nopen}).
-Removing the penalty in the half-uniform case is not recommended (see Appendix).}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg.tex}
-\caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
-and the normal-based method in the flat-top scenario.
-These results likely reflect inaccurate estimates of the tails of $g$ due to a disconnect between the tail of $g$ and the component distributions in these cases.
-For example, the uniform methods sometimes substantially underestimate the length of the tail of $g$ in these long-tailed scenarios,
-causing over-shrinkage of the tail toward 0.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos.tex}
-\caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds} \label{tab:coverage}
-\end{table}
-
-%
-%\subsubsection*{Reverse asymptotics}
-%
-%Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
-%
-%Note: we'd like to make sure it behaves ok in the case where most observations are null.
-%Perhaps simulate with 999 null and 1 alternative?
-%Also maybe look at performance for small sample sizes - e.g. n=1?
-%
-%Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
-%can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
-%by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
-%
-
-\subsection*{Differing measurement precision across units}
-
- We turn now to the second important component of our work: allowing for varying
- measurement precision across units. The key to this is the use of a likelihood,
- (\ref{eqn:normlik}) or (\ref{eqn:tlik}), that explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
-
-  To illustrate, we conduct a simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
- ($s_j=10$).  In both cases,
-we assume that half the effects are null
-and the other half are normally distributed with standard deviation 1:
-\begin{equation}
-p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1).
-\end{equation}
-In this setting, the poor-precision measurements $(s_j=10)$ tell us very little, and any sane analysis should effectively ignore them. 
- However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor}). This is because the poor measurements 
- produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
- which, when combined with the good-precision measurements, dilute the overall signal (e.g. they reduce the density of $p$ values near 0).
- This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
-the estimated error rates ($q$-values, or $\lfdr$ values) for the good-precision observations increase when the low-precision observations are included in the analysis
-(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
-for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).
-
-\begin{figure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_hist.pdf} 
-    \label{fig:goodpoor-hist}
-    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations}
- \end{subfigure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_scatter.pdf}
-    \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) of a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
-\end{subfigure}
-\caption{Simulation illustrating how, for existing FDR methods, 
-poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
-$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
-a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
-In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
-\end{figure}
-
- \subsubsection*{Reordering of significance, and the ``$p$ value prior"}
-
-Another consequence of accounting for differences in measurement precision is that \ashr may re-order the significance of the observations compared
- with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
- We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measuments that have the same $p$ value.
- The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
- and so the $\lfsr$ for these poor-precision measurements is always high.
-
-The potential for Bayesian analyses to re-order the significance of observations, and specifically to downweight imprecise observations,
-was previously discussed in \cite{guan.stephens.08}.
-However, Wakefield \cite{wakefield:2009} showed that, under a certain prior assumption, the Bayesian analysis produces the same ranking of significance as $p$ values (or their $z$ scores).
-He named this prior the ``$p$-value prior" because it can be thought of as the
-implicit prior assumption that is being made if we rank the significance of observations by their $p$ value.
-Wakefield's $p$-value prior assumes that the less precise effect estimates correspond to larger true effects, and specifically that they scale proportional
-to the standard errors $s_j$. More specifically still, it assumes a normal prior for the non-zero $\beta_j$ with mean 0 and variance $K s^2_j$ for some constant $K$.
-Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. Specifically, if, instead of assuming that $\beta_j$ is independent
-of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
-then the $\lfsr$ computed by ash.n provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
-as is illustrated in Figure \ref{fig:lfsr_pval}, right panel. (This result does not hold when using the mixtures of uniforms prior, ash.u.)
-
-This $p$-value prior assumes that the $z$ scores $z_j=\beta_j/s_j$
-are identically distributed, independent of $s_j$. This is essentially the assumption made, implicitly or explicitly, by existing methods -- like \locfdr, \mixfdr and \qvalue --
-that model the $z_j$ or $p_j$ directly.  In contrast, we have assumed up to now that the $\beta_j$ are independent of $s_j$. 
-We can set both these assumptions within a more general framework, which allows that $\beta_j/s^\alpha$ is independent of $s_j$ for some $\alpha$ [Equation (\ref{eqn:beta-alpha}))].
-Setting $\alpha=0$ implies that $\beta_j$ is
-independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
-effects (in absolute value). This latter assumption may often be qualitatively plausible: 
-for example, in gene expression studies the standard error for gene $j$ depends partly on
-the variance of its expression among samples, and genes with a larger variance may tend to be less tightly regulated and
-so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
-On the other hand, there is no particular reason to expect that either $\alpha=1$ or $\alpha=0$ will be the optimal choice.
-Indeed, optimal choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$,
-which will be dataset-specific. 
-Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
-comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
-likelihood-based methods can be used to select $\alpha$. For example, following the logic of the EB approach it
-would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
-which has been implemented in our software by C. Dai.
-
-%\begin{result}
-%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
-%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
-%\end{result}
-
-%To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
-%Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
-%In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
-%Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
-%which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
-%(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
-%may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)
-
-%\begin{result}
-%Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and using a mixture of normals for the prior distribution 
-%(\ref{eqn:beta-alpha}) with $\alpha=1$, 
-%the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$.
-Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$)
-%is montonic with $|t_j|$. 
-%\end{result}
-
-%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
-%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.
-
-%Most previous authors of EB approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
-%have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
-%In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
-%This is analogous
-%to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
-%standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
-%that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.
-
-\begin{figure}
-\centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
-\caption{Figure illustrating affects of prior assumptions on re-ordering of significance.
-Left panel shows results under our ``default prior" which assumes that effects $\beta_j$ are identically distributed, independent of $s_j$.
-Right panel shows results under the ``$p$-value prior", which assumes that $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$.} \label{fig:lfsr_pval}
-\end{figure}
-
-%\subsubsection*{Adaptive Shrinkage}
-%
-%Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
-%when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
-%the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
-% $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
-% In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
-% Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
-% and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
-%
-%Our methods we build on this idea in two ways. First,
-%by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
-%Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
-%adapts to measurement precision: more precise measurements undergo
-% less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
-%than observations with
-%small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
-%To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
-%Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
-%
-%\begin{figure}[!ht]
-%\begin{center}
-%\includegraphics[width=400px]{../examples/ash.pdf} 
-%\end{center}
-%\caption{Figure illustrating adaptive shrinkage.
-Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0.
-The standard error of each observation, $s_j$, was simulated from Inverse Gamma(5,5), which resulted in $s_j$ varying from 0.34 to 6.7.
-The shrunken estimates ($y$ axis) are plotted against the observed estimates ($x$ axis), with color indicating the (square root of the) precision of each measurement (blue=lower-precision; red=higher-precision).
-The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
-%\end{figure}
-
-%DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
-%To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
-Because the prior distribution is unimodal about zero,
-%the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
-%amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
-%For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
-%than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
-%small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
-
-%previous work based on normal means:
-%http://link.springer.com/article/10.1007/BF02562623
-
-%\subsection{Null behaviour}
-
-%Results of simulations under the null.
-
-%\subsection{Comparison of different underlying component distributions}
-%
-%Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
-%than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
-%likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
-%For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
-%increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
-%expect $L_{hu}$ to be significantly larger than the others.
-%
-%Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
-%
-%\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
-%\begin{center}
-%\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
-%%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
-%\end{center}
-%\caption{}
-%\end{figure}
-
-%Notes from Adrian Raftery - 
-%Newcombe (1886)
-%Dean and Raftery
-%Uniform is super-efficient
-
-\section*{Discussion}
-
-We have presented an Empirical Bayes approach to large-scale multiple testing that emphasises two ideas.
-First, we emphasise the potential benefits of using two numbers ($\bhat$, and its standard error)
-rather than just one number (a $p$ value or $z$ score)  to summarize the information on each test.
-While requiring two numbers is slightly more onorous than requiring one, in many settings these
-numbers are easily available and if so we argue it makes sense to use them.
-Second, we note the potential benefits -- both statistical and computational -- of assuming that the effects come from a unimodal distribution, and provide
-flexible implementations for performing inference under this assumption. We also introduce the ``false sign rate" as
-an alternative measure of error to the FDR, and illustrate its improved robustness to errors in model fit, particularly mis-estimation
-of the proportion of null tests, $\pi_0$.
-
-Multiple testing is often referred to as a ``problem" or a ``burden". In our opinion, EB approaches turn this idea on its head,
-treating multiple testing as an {\it opportunity}: specifically, an opportunity to learn about the prior distributions, and other modelling assumptions,
-to improve inference and make informed decisions about significance (see also \cite{greenland1991empirical}). This view
-also emphasises that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
-Indeed, the FDR at a given fixed threshold does not depend on the number of tests: as the number of tests increases, both the true positives and false positives increase linearly,
-and the FDR remains the same.
-(If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example.
-The EB approach captures this dependence in an intuitive way:
-if there are lots of strong signals then we infer $\pi_0$ to be small, and the estimated FDR (or $\lfdr$, or $\lfsr$) at a given threshold may be low, even if a large number of tests were performed; and conversely
-if there are no strong signals then we infer $\pi_0$ to be large and the FDR at the same threshold may be high, even if relatively few tests were performed. 
-More generally, overall signal strength is reflected in the estimated $g$, which in turn influences the estimated FDR.
-
-Two important practical issues that we have not addressed here are correlations among tests,
-and the potential for deviations from the theoretical null distributions of test statistics. These two
-issues are connected: specifically, unmeasured confounding factors can cause both correlations among tests
-and deviations from the theoretical null \cite{efron2007correlation,leek:2007}. And although there are certainly other
-factors that could cause dependence among tests, unmeasured confounders are perhaps the most worrisome in practice
-because they can induce strong correlations among large numbers of tests and profoundly impact results.
-Approaches to deal with unmeasured confounders can be largely divided into two types: those
-that simply attempt to correct for the resulting inflation of test statistics \cite{devlin1999genomic,efron2004large}, and those
-that attempt to infer counfounders using clustering, principal components analysis, or factor models 
-\cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
-and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable,
-they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alterntively, our methods
-could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:lik}) using
-$\bhat_j | \shat_j \sim N(0,\lambda_1 \shat_j + \lambda_2)$, where $\lambda_1,\lambda_2$ are to be estimated. However,
-this immediately raises issues of identifiability and we do not pursue the idea further here.
-
-Another important practical issue is the challenge of small sample sizes. For example, in genomics applications researchers sometimes
-attempt to identify differences between two conditions based on only a handful of samples in each. In such settings the
-normal likelihood approximation (\ref{eqn:normlik}) will be inadequate. And, although the $t$ likelihood (\ref{eqn:tlik}) partially
-addresses this issue, it is also, it turns out, not entirely satisfactory.  The root of the problem is that, with small sample sizes, raw estimated standard
-errors $\shat_j$ can be horribly variable. In genomics it is routine to address this issue by applying EB methods \cite{smyth:2004} to ``moderate" (i.e.~shrink) variance estimates,
-before computing $p$ values from ``moderated" test statistics. We are currently investigating how
-our methods should incorporate such ``moderated" variance estimates to make it applicable to small sample settings.
-
-Our approach involves compromises between flexibility, generality, and simplicity on the one hand, and statistical efficiency and principle on the other.
-For example, in using an EB approach that uses a point estimate for $g$, rather than a fully Bayes approach that accounts for uncertainty in $g$,
-we have opted for simplicity over statistical principle. And in summarizing every test by
-two numbers and making a normal or $t$ approximation to the likelihood, we have aimed to produce generic methods that 
- can be applied whenever such summary data are available -- just as \qvalue can be applied to any set of $p$ values for example -- although
- possibly at the expense of statistical efficiency compared with developing multiple tailored approaches based on context-specific likelihoods. 
- Any attempt to produce generic methods will involve compromise between generality and efficiency.
- In genomics, many analyses -- not only FDR-based analyses -- involve first computing a series of $p$ values before
- subjecting them to some further downstream analysis. An important message here is 
- that working with two numbers ($\bhat_j,\shat_j$), rather than one ($p_j$ or $z_j$), 
- can yield substantial gains in functionality (e.g.~ estimating effect sizes, as well as testing; accounting for variations in
-measurement precision across units) while losing only a little in generality.  We hope that our work will
-encourate development of methods that exploit this idea in other contexts.
-
-%This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
-%from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
-%with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
-
-%
-%
-%Both \locfdr and \mixfdr allow for deviations from
-%the theoretical null distribution by allowing that, under the null, the $z$ scores are $N(0,\sigma^2)$ with $\sigma$ estimated from the data rather
-%than fixed to 1. Within our approach this idea could be incorporated by changing the likelihood - for example, to assume $\bhat_j \sim N(0,s_ 
-%
-%These are tricky issues that, as far as we are aware, can affect
-%all approaches to FDR analysis. An important focus of Efron's work, also implemented in \locfdr, 
-% is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
-%theoretical null distribution. 
-%
-%For example, we and others have worked on Bayesian approaches to genetic association testing, using subjective prior specification of
-%effect sizes based on external data (e.g.~\cite{stephens:2009}). However, the large numbers of tests brings the opportunity to learn about
-%priors, and here we aim to exploit this
-
-%Choice of mixture components. Use of likelihood to compare different mixture components?
-
-%  
-%Directly modeling the p values, or z scores, say via non-parametric methods, 
-%can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-%distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-%The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-%Before proceeding, we feel obliged to correct the common misconception
-% that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
-% not depend on the number of tests. 
-%  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
-% 
-% In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
-
-\section*{Detailed Methods} \label{sec:implement}
-
-\subsection*{Embellishments}
-
-\subsubsection*{More flexible unimodal distributions}
-
-Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
-Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
- Therefore, we have implemented a more general approach based on 
-\begin{equation} \label{eqn:g-general}
-g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) ,
-\end{equation}
-where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
-\begin{enumerate}[(i)]
-\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$, \qquad (``ash.n")
-\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,  \qquad (``ash.u")
-\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,  \qquad (``ash.hu")
-\end{enumerate}
-where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
-(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
-and $\pi$ is a $2K+1$ vector that sums to 1.)
-The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
-Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
-under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
-(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
-(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
-however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)
-
-Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
-using a large dense grid of $\sigma^2_k$ or $a_k$, (i)-(iii) can respectively approximate,
-with arbitrary accuracy,
-\begin{enumerate}[(i)]
-\item any scale mixture of normals, which includes as special cases
-the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
-\item any symmetric unimodal distribution about 0.
-\item any unimodal distribution about 0.
-\end{enumerate}
-The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
-In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.
-
-Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
-The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
-See Implementation for details of our software defaults.
-
-%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting.
-For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
-%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
-
-\subsubsection*{Dependence of effects on standard errors}
-
-Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
-This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
- \begin{equation} \label{eqn:beta-alpha}
- \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
- \end{equation}
-for any $\alpha$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
-assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
-it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
-the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.
-
-The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
-To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
- standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
- Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
-yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
-%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
-%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
-%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
-
-%Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
-%how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
-%it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
-%rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
-%and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
-%this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
-%do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
-
-\subsubsection*{Replace normal likelihood with $t$ likelihood}
-
-We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
-  \begin{equation} \label{eqn:tlik}
- \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
- \end{equation}
-where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
-and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
-The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
-This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms; see Implementation below.
-When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
-
-Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
-\begin{equation} \label{eqn:tdist}
-(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
-\end{equation}
-However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
-In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
-currently pursuing this improved approach (joint work with M.Lu) and results will be published elsewhere. 
-%We note here that
-%modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
-%where our simpler approach can produce unsatisfactory results.
-
-\subsubsection*{Non-zero mode}
-
-An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood, rather than fixed to 0.
-
-\subsection*{Implementation Details}
-
-\subsubsection*{Likelihood for $\pi$}
-
-We define the likelihood for $\pi$ to be the probability of the observed data $\bhat$ conditional on $\shat$: $L(\pi) := p(\bhat | \shat,\pi),$ which
-by our conditional indendence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$.
-[One might prefer to define the likelihood as $p(\bhat, \shat | \pi) = p(\bhat | \shat, \pi) p(\shat | \pi)$, in which case our definition comes down to assuming that the term $p(\shat | \pi)$ does not depend on $\pi$.]
-
-Using the prior $\beta_j \sim \sum_{k=0}^K \pi_k f_k(\beta_j)$ given by (\ref{eqn:g-general}), and the normal likelihood (\ref{eqn:normlik}), integrating over $\beta_j$ yields
-\begin{equation}
-p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
-\end{equation}
-where 
-\begin{equation}
-\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
-\end{equation}
-denotes the convolution of $f_k$ with a normal density.
-These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
-Specifically, 
-\begin{equation} \label{eqn:uconv}
-\tilde{f}_k(\bhat_j)  = 
-\begin{cases}
-N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
-\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
-\end{cases}
-\end{equation}
-where $\Psi$ denotes the cumulative distribution function (c.d.f.) of the standard normal distribution.
-If we replace the normal likelihood with the $t_\nu$ likelihood (\ref{eqn:tlik}) then the convolution 
-for $f_k$ uniform the convolution is still given by (\ref{eqn:uconv}) but with $\Psi$ the c.d.f. 
-of the $t_\nu$ distribution function. (The convolution for $f_k$ normal is tricky and we have not implemented it.)
-
-\subsubsection*{Penalty term on $\pi$}
-
-To make $\lfdr$ and $\lfsr$ estimates from our method ``conservative" we add a penalty 
-term $log(h(\pi;\lambda))$ to the log-likelihood $\log L(\pi)$ to encourage over-estimation of $\pi_0$:
-\begin{equation} \label{eqn:penalty}
-h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
-\end{equation}
-where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
-consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0sims}).
-
-Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
-we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
-
-\subsubsection*{Problems with removing the penalty term in the half-uniform case}
-
-It is straightforward to remove the penalty term by setting $\lambda_k=1$ in (\ref{eqn:penalty}).
-We note here an unanticipated problem we came across when using no penalty term in the half-uniform case 
-(i.e.~ $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$ in (\ref{eqn:g-general})): when the data
-are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
-sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
-for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
-and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
-
-\subsubsection*{EM algorithm}
-
-With this in place, the penalized log-likelihood for $\pi$ is given by:
-\begin{equation}
-\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
-\end{equation}
-where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This can be maximized using an EM algorithm \cite{dempster77}, whose
-one-step updates are:
-\begin{align} \label{eqn:w}
-w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
-n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
-\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}
-\end{align}
-Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
-and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
-probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$. (The $w_{kj}$ are sometimes referred to as the ``responsibilities".)
-Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
-We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple} to accelerate convergence of this EM algorithm.
-
-%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
-\subsubsection*{Initialization}
-
-By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
-Our rationale for initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
-but weak signal in the data cannot quickly draw the algorithm towards the null.
-
-In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
-$\pi_0=1; \pi_k=0 (k>0)$. If not then we replace the EM solution with the global null solution. This aims to guard against errors due to convergence to a local optimum when the
-data are consistent with the global null.
-
-\subsubsection*{Conditional distributions}
-
-Given $\hat\pi$, we compute the conditional distributions 
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
-\end{equation} 
-Each posterior is a mixture on $K+1$ components:
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
-\end{equation}
-where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
-and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
-$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
-All these posterior distributions are easily available.
-For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
-If $f_k$ is normal and $L$ is normal, then this is a normal distribution.
-
-\subsubsection*{Choice of grid for $\sigma_k, a_k$} \label{sec:grid}
-
-\def\sigmamax{\sigma_{\text{max}}}
-\def\sigmamin{\sigma_{\text{min}}}
-When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
-in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
-as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
-based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
-is negative we set $\sigmamax = 8\sigmamin$.)
-
-When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.
-
-Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
-a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.
-
-\appendix
-
-\begin{table}[!ht]
-\centering\begin{tabular}{c c } \toprule
-Scenario & Alternative distribution, $g_1$  \\ \midrule
-spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
-near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
-flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
- &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
-skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
-big-normal & $N(0,4^2)$ \\ 
-bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
-\end{tabular}
-\caption{Summary of simulation scenarios considered} \label{table:scenarios}
-\end{table}
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all_nopen.tex}
-	\caption{All observations}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg_nopen.tex}
-\caption{``Significant" negative discoveries.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos_nopen.tex}
-\caption{``Significant" positive discoveries.}
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{table:nopen}
-\end{table}
-%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
-%\begin{equation}
-%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
-%\end{equation}
-%which is close to the definition in \cite{storey.03}.
-
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgements}
-
-Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
-prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} , and includes contributions 
-from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.
-
-This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.
-
-%\section*{References}
-% The bibtex filename
-\bibliography{/Users/stephens/Dropbox/Documents/mainbib}
-
-\section*{Figure Legends}
-
-\clearpage
-
-\section*{Tables}
-
-\section*{Supporting Information Legends}
-
-Supplementary material can be found in {\bf Supplementary Information S1.}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
-\end{document}

### 2015-10-06 subsubsectionReorder.tex
New Article!

--- 2015-10-06 paper/main.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
-\documentclass[11pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-%\usepackage{natbib}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-\usepackage{booktabs}
-\usepackage{xspace}
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-\usepackage{sectsty}
-\subsectionfont{\normalsize}
-\subsubsectionfont{\mdseries\itshape}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\usepackage{color}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-% Use the PLoS provided bibtex style
-\bibliographystyle{/Users/stephens/Dropbox/Documents/stylefiles/plos2009}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
-\pagestyle{myheadings}
-%% ** EDIT HERE **
-\usepackage{enumerate}
-\usepackage{multirow} 
-\usepackage{url}
-\usepackage{xr} %for cross-referencing
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-\newtheorem{algorithm}{Algorithm}
-\newtheorem{proposition}{Proposition}
-\newtheorem{restateproposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{result}{Result}
-\newtheorem{note}{Note}
-\newtheorem{definition}{Definition}
-
-\def\lfdr{\textit{lfdr}}
-\def\lfsr{\textit{lfsr}}
-\def\x{\mbf{x}}
-\def\y{\mbf{y}}
-\def\e{\mbf{e}}
-\def\g{\mbf{g}}
-\def\p{p}
-\def\Yb{\hat{Y}_1^{\text{BAYES}}}
-\def\Xperp{X_{1 \perp 0}}
-\def\Pperp{P_{1 \perp 0}}
-\def\Pperpb{\Pperp^{\text{B}}}
-\def \hatd { \widehat{ d } }
-\def \dhatin {\hatd^{\mbox{\tiny IN}}}
-\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
-\def \mvn{\text{MVN}}
-\def\mn{\text{MN}}
-\def \iw{\text{W}^{-1}}
-\def \fp {fastPHASE\ }
-\def\bhat{\hat{\beta}}
-\def\shat{\hat{s}}
-\def\var{\rm Var}
-\def\mvr{\text{BMVR}}
-\def\bfuni{\text{BF}_\text{uni}}
-\def\one{\mbf{1}}
-\def\zero{\mbf{0}}
-\def\hall{H_{\text{all}}}
-\def\pall{\p_{\text{all}}}
-\def\bfall{\text{BF}_\text{all}}
-\def\bfav{\text{BF}_\text{av}}
-\def\gall{\gamma_\text{all}}
-\def\limbfall{\bfall^\rightarrow}
-\def\limbfav{\bfav^\rightarrow}
-\def\limbfgamma{\BF_\gamma^\rightarrow}
-\def\sxx{S_{xx}}
-\def\syy{S_{yy}}
-\def\vyy{V_{yy}}
-\def\vyx{V_{yx}}
-\def\vxx{{V_{xx}}}
-\def\rss{\text{\rm RSS}}
-\def\dss{\text{\rm $\Delta$SS}}
-\def\U{U}
-\def\D{D}
-\def\I{I}
-\def\luu{L_{UU}}
-\def\indep{\perp}
-\def\qvalue{{\tt qvalue}\xspace}
-\def\locfdr{{\tt locfdr}\xspace}
-\def\mixfdr{{\tt mixfdr}\xspace}
-\def\ashr{{\tt ashr}\xspace}
-
-\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
-                                % mode too
-\newcommand\height{{\it height}}
-\newcommand\weight{{\it weight}}
-\def\BF{\text{BF}}
-
-%% END MACROS SECTION
-%\externaldocument{plos_one_rev1_suppinfo}
-
-\begin{document}
-
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{False Discovery Rates: A New Deal}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Matthew Stephens$^{1*}$, 
-\\
-\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
-\\
-$\ast$ E-mail: Corresponding mstephens@uchicago.edu
-\end{flushleft}
-
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-%\section*{Author Summary}
-
-\section*{Introduction}
-
-Since its introduction in 1995 by Benjamini and Hochberg 
-\cite{benjamini1995controlling}, the ``False Discovery Rate" (FDR) has quickly established itself
-as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing
-in which the goal is to identify the non-zero ``effects" among a large number of imprecisely-measured effects.
-
-Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
-indeed, the notion that EB approaches could be helpful in handling multiple comparisons predates introduction of
-the FDR  (e.g.~\cite{greenland1991empirical}). More recently, EB approaches to the FDR
-have been extensively studied by several authors, 
-especially B.~Efron and co-authors 
-\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
-see also \cite{kendziorski2003parametric,newton2004detecting, 
-datta2005empirical,muralidharan2010empirical} for example.
-So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
-existing widely-used FDR pipelines) and can substantially affect
-inference. The first idea is to {\it assume that the distribution of effects is unimodal}. This yields a very simple, fast, and stable computer 
-implementation, as well as improving inference of FDR when the unimodal assumption is correct.  
-The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
-to summarize each measurement. This idea allows variations in measurement precision to be better accounted for,
- and avoids a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
-
-In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
-the idea that it may be preferable to focus on estimation rather than on testing.
-In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
-simply estimation with some positive prior probability that the effect is exactly zero.
-However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
-perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
-controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
-``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
-in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
-the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the lfsr, particularly its increased robustness to modelling assumptions.
-
-%terminology
-%``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
-%and correspondingly define the ``false signed discovery rate", or simply ``false sign rate" (FSR), in a natural way.
-
-%The methods we provide here can address both problems, and we provide results that highlight important benefits of focussing on estimation, and particularly on assessing the correctness of the {\it sign} of the estimated effects, rather than on testing for non-zero effects.
-
-Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
-(We address the reasons for this name, and connections with shrinkage analysis, in the Discussion.)
-
-%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.
-
-%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-%measurements are made on only a small number of
-%samples in each condition - sometimes as few as one
-%sample in each condition. Thus the error in estimates
-%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-%in estimates of $\beta_j$ still greater.
-%
-
-%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
-%Here we present new tools that bring two important ideas to bear on the problem.
-%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
-%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
-%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
-%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
-%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
-%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
-%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
-%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
-%a substantial increase in the number of discoveries at a given FDR).
-
-%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
-%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
-Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
-%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
-%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
-%generic methods that work directly with two numbers for each 
-%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
-Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
-%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
-
-\def\df{df}
-\def\FDR{\text{FDR}}
-\def\fdr{\text{\lfdr}}
-\def\FSR{\text{FSR}}
-\def\fsr{\text{lfsr}}
-
- \section*{Methods}
-
- \subsection*{Model Outine}
-
-% Here we consider one setting in which FDR-based methods are widely used:
-%an experimenter has measured a large numbers of ``effects", each with an associated
-%measurement precision (standard error), 
-%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
-%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
-%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
-%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
-%For example, in genomics
-%a common goal is to identify ``differentially-expressed genes" -- that is, genes
-%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
-%mean expression level between the two conditions, often referred to as the ``log-fold change". 
-Here we describe the simplest version of our method, and briefly discuss embellishments we have also implemented.
-
-Suppose that we are interested in the values of $J$ ``effects" $\beta=(\beta_1,\dots,\beta_J)$. For example, in a typical genomics application
-that aims to identify differentially expressed genes,  $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
-In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
-that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
-more general problem of estimating, and assessing uncertainty in, $\beta_j$.
-
-Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
-and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
-which by Bayes theorem can be written as 
-\begin{equation}
-p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
-\end{equation}
-
-For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
-This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
-A simple way to implement the UA is to assume that $g$ is
-a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
-\begin{align} \label{eqn:beta}
-p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
-g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
-\end{align}
-where $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyperparameters, which are non-negative and sum to one, and are to be estimated,
-while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
-represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
-(We encourate the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
-although of course in practice we use a finite grid -- see Implementation Details.)
-
-For the likelihood $p(\bhat | \beta, \shat)$ we assume  
-\begin{equation} \label{eqn:normlik}
-p(\bhat | \beta, \shat) = \prod_j N(\bhat_j; \beta_j, \shat_j^2).
-\end{equation}
-Here, in additional to some conditional independence assumptions, we are 
-effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-
-This simple model features both the key ideas we want to emphasise in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
-the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
-with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
-However, this simple model also has several additional assumptions that can be relaxed. Specifically,
-\begin{enumerate}
-\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) also implies that $g$ is symmetric about 0; more flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see (\ref{eqn:g-general})].
-\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
-We can relax this to allow for a relationship between these quantities [see (\ref{eqn:beta-alpha})].
-\item The likelihood (\ref{eqn:normlik}) effectively assumes that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-We can generalize this likelihood using a $t$ likelihood [see (\ref{eqn:tlik})].
-\end{enumerate}
-These embellishments are detailed in Embellishments and Implementation Details.
-Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
-(which are also made by most existing EB approaches to this problem). Correlations among tests certainly arise in practice, either due to genuine correlations
-in the system of study, or due to unmeasured confounders, and their potential to impact results of an FDR analysis is important to consider
-whatever analysis methods are used: see \cite{efron2007correlation,leek:2007} for relevant discussion.
-
-%We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
-%$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
-%\begin{equation}
-%p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
-%\end{equation}
-%Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
-%\begin{equation} \label{eqn:normlik}
-%p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
-%
-%We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
-%as a mixture of null and alternative components:
-%\begin{equation} \label{eqn:betaj0}
-%g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
-%\end{equation}
-%where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
-%We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
-%\begin{equation} \label{eqn:beta-iid}
-%p(\beta | \shat) = \prod_j g(\beta_j).
-%\end{equation}
-%
-%Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
-%with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
-%to use a mixture of zero-mean normal distributions:
-%\begin{equation} \label{eqn:mixnorm}
-%g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
-%Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
-%represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
-% The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
-% more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below. 
-% 
-% Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
-% yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
-% \begin{equation} \label{eqn:g}
-% g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
-% \end{equation}
-% 
-% Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
-% \begin{align} \label{eqn:betaj}
-% p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
-% p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
-% \end{align}
-% with $g$ given by the mixture of normals $(\ref{eqn:g})$.
-
-\subsection*{Fitting the model}
-
-In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
-each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
-Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
-\begin{equation} \label{eqn:marginallik}
-p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
-\end{equation}
-where we define $\sigma_0:=0$.
-
-The usual EB approach to fitting this model would involve two simple steps:
-\begin{enumerate}
-\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
-\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
-against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | | \bhat, \shat,\hat{\pi})$.
-\end{enumerate}
-Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(The simplicity of the EM algorithm in step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
-which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
-
-Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximise a penalized
-likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
-We introduce this penalty because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
-
-\subsection*{The local False Discovery Rate and local False Sign Rate}
-
-As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
-In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
-One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
-the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
-In other words it is the posterior probability that $\beta_j$ is actually zero:
-\begin{equation} \label{eqn:fdr}
-\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
-\end{equation}
-
-The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
-whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
-the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey1991philosophy}) argued that
-``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$.
-Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
-\begin{quote}
-...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
-\end{quote}
-Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
-focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.
-
-Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
-would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
-\begin{equation} \label{eqn:lfsr}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
-\end{equation}
-To illustrate, suppose that 
-\begin{gather*}
-p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
-p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
-p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
-\end{gather*} 
-Then 
-from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
-our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.
-
-As our notation suggests, $\lfsr_j$ is intended to be compared and contrasted with $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
-small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
-Of coure, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
-this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
-(this follows from the definition because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
-In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
-$\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.
-
-From these ``local" measures of significance, we can also compute average error rates over subsets of observations
-$\Gamma \subset \{1,\dots,J\}$. For example,   
-\begin{equation} \label{eqn:FDRhat}
-\widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
-\end{equation}
-estimates the FDR we would obtain if we were to declare all tests in $\Gamma$ significant.
-And 
-\begin{equation} \label{eqn:qvalue}
-q_j := \widehat \FDR(\{k: \fdr_k \leq \fdr_j\}) 
-\end{equation}
-provides a measure of significance analogous to Storey's $q$ value \cite{storey.03}.
-
-%Because our methods involve explicitly modelling effects $\beta_j$, rather than 
-%$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
-
-%Indeed, even in settings where some
-%$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
-%analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
-
-\subsection*{Related work}
-
-\subsubsection*{Previous approaches focussed on FDR}
-
-Among previous methods that explicitly consider the FDR and related quantities,
-our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
- (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
-  %Here we highlight the key differences between our approach and these three existing approaches.
-
-There are two key differences between our approach and all of these three existing methods. First, 
-whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
-we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
-for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
-Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 
-how it can improve performance by taking better account of variation in measurement precision among observations.
-%(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
-% Or more generally Section xx illustrates the potential practical effects
-
-%allows for different modelling assumptions to be compared through likelihood comparisons.
-Second, our unimodal assumption (UA) that the effects are unimodal about zero is quite different from assumptions made by $\qvalue, \locfdr$ or $\mixfdr$.
-Indeed, $\locfdr$ assumes that all $z$ scores near 0 are null (Efron calls this the Zero Assumption; ZA), which implies that 
-under the alternative hypothesis the distribution of $z$ scores has {\it no mass at 0}; this contrasts strikingly with the UA, which implies that this distribution has its peak at 0! 
-Similarly, $\qvalue$ assumes that all $p$ values near 1 are null,
-which is the same as the ZA because $p$ values near 1 correspond to $z$ scores near 0. And although \mixfdr does not formally make the ZA,
-we have found that in practice, with default settings, the results often approximately satisfy the ZA (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}).
-Thus, not only do these existing methods not make the UA, they actually make assumptions that are, in some sense, as different
-from the UA as they can be.
-
-Given that the UA and ZA are so different, it seems worth discussing why we generally favor the UA. 
-Although the UA will not apply to all situations, we believe that it will often be reasonable, especially in FDR-related contexts
-that have traditionally focussed on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis,  
-it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible. Further, it seems reasonable to expect that larger effects become
-decreasingly plausible, and so the distribution of the effects will be unimodal about 0.
-To paraphrase Tukey, ``All we know about the world teaches us that large effects are rare, whereas small effects abound."
-We emphasise that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
-It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
-
-%Tukey "peri-null"
-
-%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
-%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}.
-
-In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficents - 
-common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
-These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
-it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
-The UA assumption on regression coefficients is directly analagous to our UA here, and so its
-widespread use in the regression context supports its use here.
-
-Alternatively, we could motivate the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
-such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
-make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).
-
-Finally, the UA also has a considerable practical benefit: it yields simple algorithms that are both computationally
-and statistically stable. We illustrate these features in Results.
-
-\subsubsection*{Other work}
-
-There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
- Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
-indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
-This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
-Also, they provide no software implementation.
-More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
-Some pointers to recent papers in which other relevant
-citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}.
-Much of the literature focusses on the homoskedastic case (i.e.
-$\shat_j$ all equal) whereas we allow for heteroskedasticity.
-And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential.
-Several recent papers consider more flexible non-parametric assumptions on
-$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
-estimate (NPMLE) for $g$.
-These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive.
-However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
-may perform well for point estimation, it seems possible it will not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
-Indeed, one way of thinking about the UA is as a way to regularize $g$.
-
-%The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
-%encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-%$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-%$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-%the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-%observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-%corresondingly little effect on the estimated FDR of other observations. 
-%We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-%$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-%that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-%estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-%Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-\def\dt{\text{dt}}
-
-%
-%assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
-%$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
-% $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
-%  \begin{equation} \label{eqn:lik}
-% \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
-% \end{equation}
-% where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e.
-the density of $\mu+\sigma T_\nu$).
-In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
-%  \begin{equation} \label{eqn:normlik}
-% \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
-% \end{equation}
-%which, in light of (\ref{eqn:betahat}), is equivalent to
- % \begin{equation} \label{eqn:lik2}
-% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
- %\end{equation}
-
-%The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
-they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
-The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$.
-Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$.
-If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$.
-See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.
-
-%This likelihood approximation can be motivated by assuming that i) $\bhat$
-% $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
-% $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$. 
-%
-
-%Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.
-
-%While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
- %that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
-%there is an important difference: here,
-
-%The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
-For example, we can write $p(\bhat_j, \shat_j | \beta_j) = p(\bhat_j | \shat_j, \beta_j)p(\shat_j | \beta_j)$.
-Now, if we are willing to assume that $\shat_j$ alone contains no information about $\beta_j$, or equivalently that $p(\shat_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\bhat_j | \shat_j, \beta_j)$, and if $\bhat_j | \shat_j, \beta_j \sim N(\beta_j, \shat_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.
-
-\section*{Results}
-
-We compare results of \ashr with existing FDR-based methods implemented
-in the R packages \qvalue (v2.0 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
-In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
-to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.
-
-\subsection*{Effects of the Unimodal Assumption}
-
-Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
-standard error, $s_j=1$ and all methods are provided that information. That is,
-$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$.
-In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the only difference between our method and methods like $\locfdr$ and $\mixfdr$ are in how they estimate $g$.
-
-To briefly summarize the results in this section:
-\begin{enumerate}
-\item The UA can produce very different inferences compared with the ZA made by existing methods.
-\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
-\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
-\end{enumerate}
-
-\subsubsection*{The UA and ZA can produce different inferences}
-
-\begin{figure}
-\center \includegraphics[height=6in]{figures/decomp_ZA.pdf}
-\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
-The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
-In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
-The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
-(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
-\end{figure}
-
-To illustrate the different inferences from the UA and ZA we show results for a single
-dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
-but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
-\qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
-The results (Figure \ref{fig:ZA}) illustrate the clear difference between the existing methods and our method. The effects of the ZA made by \qvalue and \locfdr
-are visually clear, producing a ``hole"  in the alternative $z$ score distribution around 0. Although \mixfdr does not formally make the ZA, its decomposition
-exhibits a similar hole. In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
-effectively ``filling in" the hole.  (Of course the null distribution also has a peak at 0, and the local fdr 
- under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)
-
-Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA.
-Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
-This would increase the null component (dark blue) at the expense of
-the alternative component (light blue).
-Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0.
-The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears.
-In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
-See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
-demonstration.
-
- \subsubsection*{The UA can produce conservative estimates of $\pi_0$}
-
-The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
-Consequently \ashr will estimate smaller $\lfdr$s and FDRs than existing methods that make the ZA. 
-This is desirable, provided that these estimates remain conservative: that is,
-provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
-The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behaviour. To check its effectiveness
-we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
-for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
-They range from a ``spiky" distribution -- where many non-zero $\beta$ are
-too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
-flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
-making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
-which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.
-
-% How do these differences influence estimates of the FDR? To address this, note that  -- in this simple case where the 
-% $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
-% of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
-% strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
-% although the methods considered here differ from one another in many ways, having estimated $\pi_0$
-% they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
-% empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
-%
-
-%Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
-% and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
-%Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
-%At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
-%overestimates of $\pi_0$, which will yield overestimates of the FDR).
-
-\begin{figure} 
-\begin{center}
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=1.8in]{figures/scenarios_density.pdf} 
-	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/pi0_est.pdf} 
-\caption{Comparison of true and estimated values of $\pi_0$.
-When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ashr being least conservative, and hence most accurate.
-When the UA does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.} \label{fig:pi0sims}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfdr_est.png} 
-\caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
-Black line is $y=x$ and red line is $y=2x$.
-Estimates of $\lfdr$ are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est.png} 
-\caption{As in c), but for $\lfsr$ instead of $\lfdr$.
-Estimates of $\lfsr$ are consistently less conservative than $\lfdr$ when UA holds, and also less anti-conservative in bimodal scenario.}  \label{fig:lfsr}
-\end{subfigure}
-\end{center}
-\caption{Results of simulation studies (constant precision $s_j=1$).}
-\end{figure}
-
-For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
-data as follows:
-\begin{enumerate}
-\item Simulate $\pi_0  \sim U[0,1]$.
-\item For $j=1,\dots,J$, simulate $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)$.
-\item For $j=1,\dots,J$, simulate $\bhat_j | \beta_j \sim N(\beta_j,1)$.
-\end{enumerate}
-%Thus these simulations assume the same
-%precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).
-
-Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ashr ($y$ axis) with the true values ($x$ axis). 
-For \ashr we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
-and as a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
-as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
-The results show that \ashr provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
-in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ashr estimates can be slightly anti-conservative.
-We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
-are used. (In addition, the effects on $\lfsr$ estimates turn out to be relatively modest; see below).
-
-%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
-%
-%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
-%and so all effects are non-zero and no nulls hold. This scenario is rather different
-%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
-%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
-%``mostly non-null"
-%scenario, and we would agree! On the other hand, the distribution of $p$ values
-%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
-%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
-%Thus it seems desirable to have methods that perform well in both cases, and this is
-%an important feature of the methods we introduce here.
-%
-%TODO: put FDRs on graph? Simulate with 50\% null?
-
-\subsubsection*{The $\lfsr$ is more robust than \lfdr}
-
-The results above show that \ashr can improve on existing methods in producing smaller, more accurate, 
-estimates of $\pi_0$, which will lead to more accurate estimates of FDR.
-Nonetheless, in many scenarios \ashr continues
-to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
-This is because these scenarios include 
-an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
-estimation of $\pi_0$ impossible. Put another way, and as is well known, $\pi_0$ is not identifiable:
- the data can effectively provide an upper bound on plausible values of $\pi_0$,
-but not a lower bound (because the data cannot rule out that everything is non-null, but with miniscule effects).
-To obtain conservative behaviour we must estimate $\pi_0$ by this upper bound, which can
-be substantially larger than the true value.
-
-Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
-overestimation of $\pi_0$ is corresponding overestimation of FDR (and $\lfdr$, and $q$ values).
-To illustrate, Figure \ref{fig:lfdr} compares the estimated $\lfdr$ from ash.n with the true value (computed using
-Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios which involve many
-non-zero effects that are very near 0 (e.g.~the spiky scenario with $\pi_0$ small) where
-$\pi_0$ can be grossly overestimated.  
-(Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.)
-
-The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
- become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
- This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
-To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
-be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
-the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in $\lfsr$ than $\lfdr$.
-
-Note that, compared with previous debates regarding testing, 
-this section advances an additional reason for focussing on the sign of the effect, rather than just testing whether it is 0. 
-In previous debates authors have argued against testing whether an effect is 0
-because it is {\it implausible that effects are exactly 0}. Here we add that {\it even if one believes
-that some effects may be exactly zero}, it is still better to focus on the sign, because generally {\it the data are more informative about that question}
-and so inferences are more robust to, say, the inevitable mis-estimation of $\pi_0$.
-To provide some intuition, consider an observation with a $z$ score of 0. The $\lfdr$ of this observation can range from 0 (if $\pi_0=0$)
-to 1 (if $\pi_0=1$). But, assuming a symmetric $g$, the $\lfsr>0.5$ whatever the value of $\pi_0$, because the observation $z=0$ says
-nothing about the sign of the effect.
-Thus, are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
-whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.
-
-Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
-Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
-if we are prepared to assume that no effects are exactly zero, then removing the point mass 
-yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
-However, there is ``no free lunch":  if in fact some effects are exactly zero
-then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
-We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.
-
-\begin{figure}
-\begin{center}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s_nn.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be more accurate in this case.
-That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
-\end{subfigure}
-\end{center}
-\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
-\end{figure}
-
-\subsubsection*{Numerical Stability}
-
-The EM algorithm, which we use here to fit our model, is notorious for convergence to local optima. 
-However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
-caused by such behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
-in addition running it using our default initialization procedure (see Implementation details). We then compared the largest
-log-likelihood achieved across all 11 runs with the log-likelihood achieved by the default run.
-When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference ($< 0.02$), 
-and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference ($<0.02$),
-and 6\% of runs showed an appreciable log-likelihood difference ($>1$), with the largest difference being 5.0. However, perhaps suprisingly, 
-even for this largest difference results from the default run (e.g.
-the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood.
-See \url{http://github.com/stephens999/ash/blob/master/dsc-robust/summarize_dsc_robust.rmd} for further details.
-
-\subsubsection*{The UA helps provide reliable estimates of $g$}
-
-An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
-can estimate the {\it size} of each effect $\beta_j$.
-Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
-to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
-threshold $T$.
-%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
-%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
-%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
-%towards zero, with the strength of the shrinkage being estimated from the data.
-Further, because the posterior distribution is, by definition,
-conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
-been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
-In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
-\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
-how can confidence intervals be assigned to the true
-effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
-and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
-%but also notes the difficulties of implementing a stable general algorithm.
-
-The ability of the EB approach to provide valid ``post-selection" interval estimates is extremely attractive in principle.
-But its usefulness in practice
-depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
-which are notoriously difficult in general. Indeed, Efron emphasises
-the difficulties of implementing a stable general algorithm, noting in his rejoinder
-``the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying
-to be overly nonparametric ... but it is hard to imagine a
-generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
-Our key point here is that the UA greatly simplifies the deconvolution problem.
-While not meeting Efron's desire for an entirely general nonparametric approach, we believe 
-that the UA can handle many cases of practical interest.
-
-To illustrate this, Figure \ref{fig:egcdf} compares the
-estimated $g$ from \ashr  with that from \mixfdr which does not make the UA 
-(and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
-The greater reliability of estimates afforded
-by the UA is immediately apparent. In particular the estimated cdf from \mixfdr often has an almost-vertical
-segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. 
-The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
-While the UA is not the only way to achieve this, we find it an attractive, simple and effective approach.
-
-Interestingly, even in the ``bimodal" scenario $\ashr$ 
-is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
-Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
-Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
-to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
-(\url{http://github.com/stephens999/ash/blob/master/dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
-clearly visible in the observed $\bhat$.
-
-We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ashr tends to systematically
-overestimate the mass of $g$ near zero. On careful inspection, this is apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
-and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf}b),
-and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
-under-estimating $\pi_0$ (Figure \ref{fig:egcdf}c).
-
-\begin{figure}[h!] 
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=2in]{figures/egcdf.pdf}
-	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf_nopen.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
-Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
-\end{subfigure}
-\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.} \label{fig:egcdf}
-\end{figure}
-
-%making the UA provides an attractive and simple way to achieve estimating $g$  large
-%subset of cases of interest (those where the UA can reasonably be assumed to hold).
-
-%In practice, the utility of the posterior distributions noted above depends primarily on the accuracy
-%of the estimate of the distribution $g(\cdot)$.
-
- %quote from efron rejoinder , p46:
-%MorrisÕ Section 3 is especially pertinent. His formula
-%for p(?|z) is related to my discussion of the
-%BenjaminiÐYekutieli False Coverage Rate method in
-%Section 7, particularly (7.2)Ð(7.4). Originally I had
-%hoped to develop an empirical Bayes method for estimating
-%such models, but the effort foundered on practical
-%difficulties involving the perils of deconvolution.
-%Section 6 on the Òone-group modelÓ is the ugly duckling
-%of the current paper, but it bears on some important
-%points raised in the discussion. Figure 7 concerns
-%a fuzzy version of simultaneous hypothesis testing,
-%where, as in MorrisÕ hospital example, the usual singlepoint
-%null hypotheses seem unequal to the task. The
-%development from (6.6) onwards, particularly (6.12),
-%bears on the possibility of nonnormal null distributions,
-%and is about as far as I can go in answering Professors
-%Rice and SpiegelhalterÕs penultimate question.
-%With g(?) a normal distribution, model (6.1) returns
-%us to the realm of Stein estimation, the scene of my
-%happy collaborations with Carl Morris. I continue to
-%be surprised at how much harder bumpy, nonnormal
-%models like (7.1) are to deal with. JamesÐStein estimation
-%works fine with, say, N = 10 component problems,
-%while the RobbinsÕ form of empirical Bayes appropriate
-%to (7.1) seems to require hundreds or thousands.
-%The information calculations in Efron (2008)
-%reinforce this gloomy assessment. Maybe I am trying
-%to be overly nonparametric in constructing the empirical
-%Bayes Fdr estimates, but it is hard to imagine a
-%generally satisfactory parametric formulation for (6.1).
-%Or perhaps it is just that hypothesis testing is more demanding
-%than estimation.
-%My Note: (6.1) is the general model (mu \sim g()).
-%7.1 is the case where the effects are a mixture of zero and N(-3,1).
-
-\subsubsection*{Calibration of posterior intervals}
-
-To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over 100 simulations in each Scenario). 
-Specifically we examine the empirical coverage of nominal lower 95\% credible bounds for a) all observations; b) significant negative discoveries; c) significant positive discoveries.
-We examine only lower bounds because the results for upper bounds follow by
-symmetry (except for the one asymmetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
-in each case: for negative discoveries the lower bound is typically large and negative and limits how big (in absolute value) 
-the effect could be; for positive discoveries the lower bound 
-is positive, and limits how small (in absolute value) the effect could be. Intuitively, the lower bound for negative discoveries depends on the accuracy of $g$ in its tail,
-whereas for positive discoveries it is more dependent on the accuracy of $g$ in the center.
-
-The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
-as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
-
-% As noted above, the penalty term (\ref{eqn:penalty}) causes systematic over-estimation of the mass of $g$ near 0,  which one would expect to overshrink interval estimates towards zero. 
-%The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound is negative and overshrinking (moving it too far towards zero) makes the left tail too big; conversely, for positive discoveries, the lower confidence bound is positive, and
-%overshrinking (moving it too far towards zero) makes the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
-%and be prepared to put up with some overshrinkage to avoid it.
-
-%Coverage rates are given in:
-%\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
-
-%Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all.tex}
-	\caption{All observations.
-Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario.
-This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero.
-Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (Table \ref{tab:nopen}).
-Removing the penalty in the half-uniform case is not recommended (see Appendix).}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg.tex}
-\caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
-and the normal-based method in the flat-top scenario.
-These results likely reflect inaccurate estimates of the tails of $g$ due to a disconnect between the tail of $g$ and the component distributions in these cases.
-For example, the uniform methods sometimes substantially underestimate the length of the tail of $g$ in these long-tailed scenarios,
-causing over-shrinkage of the tail toward 0.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos.tex}
-\caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds} \label{tab:coverage}
-\end{table}
-
-%
-%\subsubsection*{Reverse asymptotics}
-%
-%Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
-%
-%Note: we'd like to make sure it behaves ok in the case where most observations are null.
-%Perhaps simulate with 999 null and 1 alternative?
-%Also maybe look at performance for small sample sizes - e.g. n=1?
-%
-%Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
-%can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
-%by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
-%
-
-\subsection*{Differing measurement precision across units}
-
- We turn now to the second important component of our work: allowing for varying
- measurement precision across units. The key to this is the use of a likelihood,
- (\ref{eqn:normlik}) or (\ref{eqn:tlik}), that explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
-
-  To illustrate, we conduct a simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
- ($s_j=10$).  In both cases,
-we assume that half the effects are null
-and the other half are normally distributed with standard deviation 1:
-\begin{equation}
-p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1).
-\end{equation}
-In this setting, the poor-precision measurements $(s_j=10)$ tell us very little, and any sane analysis should effectively ignore them. 
- However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor}). This is because the poor measurements 
- produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
- which, when combined with the good-precision measurements, dilute the overall signal (e.g. they reduce the density of $p$ values near 0).
- This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
-the estimated error rates ($q$-values, or $\lfdr$ values) for the good-precision observations increase when the low-precision observations are included in the analysis
-(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
-for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).
-
-\begin{figure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_hist.pdf} 
-    \label{fig:goodpoor-hist}
-    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations}
- \end{subfigure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_scatter.pdf}
-    \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) of a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
-\end{subfigure}
-\caption{Simulation illustrating how, for existing FDR methods, 
-poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
-$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
-a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
-In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
-\end{figure}
-
 \subsubsection*{Reordering of significance, and the ``$p$ value prior"}
@@
 which has been implemented in our software by C. Dai.
-
-%\begin{result}
-%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
-%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
-%\end{result}
-
-%To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
-%Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
-%In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
-%Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
-%which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
-%(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
-%may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)
-
-%\begin{result}
-%Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and using a mixture of normals for the prior distribution 
-%(\ref{eqn:beta-alpha}) with $\alpha=1$, 
-%the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$.
-Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$)
-%is montonic with $|t_j|$. 
-%\end{result}
-
-%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
-%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.
-
-%Most previous authors of EB approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
-%have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
-%In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
-%This is analogous
-%to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
-%standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
-%that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.
-
-\begin{figure}
-\centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
-\caption{Figure illustrating affects of prior assumptions on re-ordering of significance.
-Left panel shows results under our ``default prior" which assumes that effects $\beta_j$ are identically distributed, independent of $s_j$.
-Right panel shows results under the ``$p$-value prior", which assumes that $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$.} \label{fig:lfsr_pval}
-\end{figure}
-
-%\subsubsection*{Adaptive Shrinkage}
-%
-%Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
-%when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
-%the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
-% $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
-% In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
-% Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
-% and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
-%
-%Our methods we build on this idea in two ways. First,
-%by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
-%Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
-%adapts to measurement precision: more precise measurements undergo
-% less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
-%than observations with
-%small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
-%To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
-%Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
-%
-%\begin{figure}[!ht]
-%\begin{center}
-%\includegraphics[width=400px]{../examples/ash.pdf} 
-%\end{center}
-%\caption{Figure illustrating adaptive shrinkage.
-Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0.
-The standard error of each observation, $s_j$, was simulated from Inverse Gamma(5,5), which resulted in $s_j$ varying from 0.34 to 6.7.
-The shrunken estimates ($y$ axis) are plotted against the observed estimates ($x$ axis), with color indicating the (square root of the) precision of each measurement (blue=lower-precision; red=higher-precision).
-The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
-%\end{figure}
-
-%DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
-%To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
-Because the prior distribution is unimodal about zero,
-%the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
-%amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
-%For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
-%than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
-%small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
-
-%previous work based on normal means:
-%http://link.springer.com/article/10.1007/BF02562623
-
-%\subsection{Null behaviour}
-
-%Results of simulations under the null.
-
-%\subsection{Comparison of different underlying component distributions}
-%
-%Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
-%than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
-%likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
-%For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
-%increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
-%expect $L_{hu}$ to be significantly larger than the others.
-%
-%Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
-%
-%\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
-%\begin{center}
-%\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
-%%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
-%\end{center}
-%\caption{}
-%\end{figure}
-
-%Notes from Adrian Raftery - 
-%Newcombe (1886)
-%Dean and Raftery
-%Uniform is super-efficient
-
-\section*{Discussion}
-
-We have presented an Empirical Bayes approach to large-scale multiple testing that emphasises two ideas.
-First, we emphasise the potential benefits of using two numbers ($\bhat$, and its standard error)
-rather than just one number (a $p$ value or $z$ score)  to summarize the information on each test.
-While requiring two numbers is slightly more onorous than requiring one, in many settings these
-numbers are easily available and if so we argue it makes sense to use them.
-Second, we note the potential benefits -- both statistical and computational -- of assuming that the effects come from a unimodal distribution, and provide
-flexible implementations for performing inference under this assumption. We also introduce the ``false sign rate" as
-an alternative measure of error to the FDR, and illustrate its improved robustness to errors in model fit, particularly mis-estimation
-of the proportion of null tests, $\pi_0$.
-
-Multiple testing is often referred to as a ``problem" or a ``burden". In our opinion, EB approaches turn this idea on its head,
-treating multiple testing as an {\it opportunity}: specifically, an opportunity to learn about the prior distributions, and other modelling assumptions,
-to improve inference and make informed decisions about significance (see also \cite{greenland1991empirical}). This view
-also emphasises that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
-Indeed, the FDR at a given fixed threshold does not depend on the number of tests: as the number of tests increases, both the true positives and false positives increase linearly,
-and the FDR remains the same.
-(If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example.
-The EB approach captures this dependence in an intuitive way:
-if there are lots of strong signals then we infer $\pi_0$ to be small, and the estimated FDR (or $\lfdr$, or $\lfsr$) at a given threshold may be low, even if a large number of tests were performed; and conversely
-if there are no strong signals then we infer $\pi_0$ to be large and the FDR at the same threshold may be high, even if relatively few tests were performed. 
-More generally, overall signal strength is reflected in the estimated $g$, which in turn influences the estimated FDR.
-
-Two important practical issues that we have not addressed here are correlations among tests,
-and the potential for deviations from the theoretical null distributions of test statistics. These two
-issues are connected: specifically, unmeasured confounding factors can cause both correlations among tests
-and deviations from the theoretical null \cite{efron2007correlation,leek:2007}. And although there are certainly other
-factors that could cause dependence among tests, unmeasured confounders are perhaps the most worrisome in practice
-because they can induce strong correlations among large numbers of tests and profoundly impact results.
-Approaches to deal with unmeasured confounders can be largely divided into two types: those
-that simply attempt to correct for the resulting inflation of test statistics \cite{devlin1999genomic,efron2004large}, and those
-that attempt to infer counfounders using clustering, principal components analysis, or factor models 
-\cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
-and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable,
-they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alterntively, our methods
-could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:lik}) using
-$\bhat_j | \shat_j \sim N(0,\lambda_1 \shat_j + \lambda_2)$, where $\lambda_1,\lambda_2$ are to be estimated. However,
-this immediately raises issues of identifiability and we do not pursue the idea further here.
-
-Another important practical issue is the challenge of small sample sizes. For example, in genomics applications researchers sometimes
-attempt to identify differences between two conditions based on only a handful of samples in each. In such settings the
-normal likelihood approximation (\ref{eqn:normlik}) will be inadequate. And, although the $t$ likelihood (\ref{eqn:tlik}) partially
-addresses this issue, it is also, it turns out, not entirely satisfactory.  The root of the problem is that, with small sample sizes, raw estimated standard
-errors $\shat_j$ can be horribly variable. In genomics it is routine to address this issue by applying EB methods \cite{smyth:2004} to ``moderate" (i.e.~shrink) variance estimates,
-before computing $p$ values from ``moderated" test statistics. We are currently investigating how
-our methods should incorporate such ``moderated" variance estimates to make it applicable to small sample settings.
-
-Our approach involves compromises between flexibility, generality, and simplicity on the one hand, and statistical efficiency and principle on the other.
-For example, in using an EB approach that uses a point estimate for $g$, rather than a fully Bayes approach that accounts for uncertainty in $g$,
-we have opted for simplicity over statistical principle. And in summarizing every test by
-two numbers and making a normal or $t$ approximation to the likelihood, we have aimed to produce generic methods that 
- can be applied whenever such summary data are available -- just as \qvalue can be applied to any set of $p$ values for example -- although
- possibly at the expense of statistical efficiency compared with developing multiple tailored approaches based on context-specific likelihoods. 
- Any attempt to produce generic methods will involve compromise between generality and efficiency.
- In genomics, many analyses -- not only FDR-based analyses -- involve first computing a series of $p$ values before
- subjecting them to some further downstream analysis. An important message here is 
- that working with two numbers ($\bhat_j,\shat_j$), rather than one ($p_j$ or $z_j$), 
- can yield substantial gains in functionality (e.g.~ estimating effect sizes, as well as testing; accounting for variations in
-measurement precision across units) while losing only a little in generality.  We hope that our work will
-encourate development of methods that exploit this idea in other contexts.
-
-%This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
-%from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
-%with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
-
-%
-%
-%Both \locfdr and \mixfdr allow for deviations from
-%the theoretical null distribution by allowing that, under the null, the $z$ scores are $N(0,\sigma^2)$ with $\sigma$ estimated from the data rather
-%than fixed to 1. Within our approach this idea could be incorporated by changing the likelihood - for example, to assume $\bhat_j \sim N(0,s_ 
-%
-%These are tricky issues that, as far as we are aware, can affect
-%all approaches to FDR analysis. An important focus of Efron's work, also implemented in \locfdr, 
-% is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
-%theoretical null distribution. 
-%
-%For example, we and others have worked on Bayesian approaches to genetic association testing, using subjective prior specification of
-%effect sizes based on external data (e.g.~\cite{stephens:2009}). However, the large numbers of tests brings the opportunity to learn about
-%priors, and here we aim to exploit this
-
-%Choice of mixture components. Use of likelihood to compare different mixture components?
-
-%  
-%Directly modeling the p values, or z scores, say via non-parametric methods, 
-%can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-%distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-%The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-%Before proceeding, we feel obliged to correct the common misconception
-% that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
-% not depend on the number of tests. 
-%  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
-% 
-% In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
-
-\section*{Detailed Methods} \label{sec:implement}
-
-\subsection*{Embellishments}
-
-\subsubsection*{More flexible unimodal distributions}
-
-Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
-Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
- Therefore, we have implemented a more general approach based on 
-\begin{equation} \label{eqn:g-general}
-g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) ,
-\end{equation}
-where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
-\begin{enumerate}[(i)]
-\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$, \qquad (``ash.n")
-\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,  \qquad (``ash.u")
-\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,  \qquad (``ash.hu")
-\end{enumerate}
-where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
-(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
-and $\pi$ is a $2K+1$ vector that sums to 1.)
-The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
-Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
-under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
-(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
-(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
-however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)
-
-Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
-using a large dense grid of $\sigma^2_k$ or $a_k$, (i)-(iii) can respectively approximate,
-with arbitrary accuracy,
-\begin{enumerate}[(i)]
-\item any scale mixture of normals, which includes as special cases
-the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
-\item any symmetric unimodal distribution about 0.
-\item any unimodal distribution about 0.
-\end{enumerate}
-The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
-In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.
-
-Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
-The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
-See Implementation for details of our software defaults.
-
-%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting.
-For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
-%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
-
-\subsubsection*{Dependence of effects on standard errors}
-
-Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
-This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
- \begin{equation} \label{eqn:beta-alpha}
- \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
- \end{equation}
-for any $\alpha$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
-assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
-it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
-the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.
-
-The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
-To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
- standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
- Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
-yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
-%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
-%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
-%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
-
-%Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
-%how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
-%it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
-%rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
-%and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
-%this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
-%do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
-
-\subsubsection*{Replace normal likelihood with $t$ likelihood}
-
-We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
-  \begin{equation} \label{eqn:tlik}
- \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
- \end{equation}
-where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
-and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
-The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
-This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms; see Implementation below.
-When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
-
-Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
-\begin{equation} \label{eqn:tdist}
-(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
-\end{equation}
-However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
-In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
-currently pursuing this improved approach (joint work with M.Lu) and results will be published elsewhere. 
-%We note here that
-%modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
-%where our simpler approach can produce unsatisfactory results.
-
-\subsubsection*{Non-zero mode}
-
-An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood, rather than fixed to 0.
-
-\subsection*{Implementation Details}
-
-\subsubsection*{Likelihood for $\pi$}
-
-We define the likelihood for $\pi$ to be the probability of the observed data $\bhat$ conditional on $\shat$: $L(\pi) := p(\bhat | \shat,\pi),$ which
-by our conditional indendence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$.
-[One might prefer to define the likelihood as $p(\bhat, \shat | \pi) = p(\bhat | \shat, \pi) p(\shat | \pi)$, in which case our definition comes down to assuming that the term $p(\shat | \pi)$ does not depend on $\pi$.]
-
-Using the prior $\beta_j \sim \sum_{k=0}^K \pi_k f_k(\beta_j)$ given by (\ref{eqn:g-general}), and the normal likelihood (\ref{eqn:normlik}), integrating over $\beta_j$ yields
-\begin{equation}
-p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
-\end{equation}
-where 
-\begin{equation}
-\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
-\end{equation}
-denotes the convolution of $f_k$ with a normal density.
-These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
-Specifically, 
-\begin{equation} \label{eqn:uconv}
-\tilde{f}_k(\bhat_j)  = 
-\begin{cases}
-N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
-\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
-\end{cases}
-\end{equation}
-where $\Psi$ denotes the cumulative distribution function (c.d.f.) of the standard normal distribution.
-If we replace the normal likelihood with the $t_\nu$ likelihood (\ref{eqn:tlik}) then the convolution 
-for $f_k$ uniform the convolution is still given by (\ref{eqn:uconv}) but with $\Psi$ the c.d.f. 
-of the $t_\nu$ distribution function. (The convolution for $f_k$ normal is tricky and we have not implemented it.)
-
-\subsubsection*{Penalty term on $\pi$}
-
-To make $\lfdr$ and $\lfsr$ estimates from our method ``conservative" we add a penalty 
-term $log(h(\pi;\lambda))$ to the log-likelihood $\log L(\pi)$ to encourage over-estimation of $\pi_0$:
-\begin{equation} \label{eqn:penalty}
-h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
-\end{equation}
-where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
-consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0sims}).
-
-Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
-we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
-
-\subsubsection*{Problems with removing the penalty term in the half-uniform case}
-
-It is straightforward to remove the penalty term by setting $\lambda_k=1$ in (\ref{eqn:penalty}).
-We note here an unanticipated problem we came across when using no penalty term in the half-uniform case 
-(i.e.~ $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$ in (\ref{eqn:g-general})): when the data
-are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
-sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
-for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
-and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
-
-\subsubsection*{EM algorithm}
-
-With this in place, the penalized log-likelihood for $\pi$ is given by:
-\begin{equation}
-\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
-\end{equation}
-where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This can be maximized using an EM algorithm \cite{dempster77}, whose
-one-step updates are:
-\begin{align} \label{eqn:w}
-w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
-n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
-\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}
-\end{align}
-Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
-and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
-probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$. (The $w_{kj}$ are sometimes referred to as the ``responsibilities".)
-Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
-We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple} to accelerate convergence of this EM algorithm.
-
-%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
-\subsubsection*{Initialization}
-
-By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
-Our rationale for initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
-but weak signal in the data cannot quickly draw the algorithm towards the null.
-
-In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
-$\pi_0=1; \pi_k=0 (k>0)$. If not then we replace the EM solution with the global null solution. This aims to guard against errors due to convergence to a local optimum when the
-data are consistent with the global null.
-
-\subsubsection*{Conditional distributions}
-
-Given $\hat\pi$, we compute the conditional distributions 
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
-\end{equation} 
-Each posterior is a mixture on $K+1$ components:
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
-\end{equation}
-where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
-and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
-$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
-All these posterior distributions are easily available.
-For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
-If $f_k$ is normal and $L$ is normal, then this is a normal distribution.
-
-\subsubsection*{Choice of grid for $\sigma_k, a_k$} \label{sec:grid}
-
-\def\sigmamax{\sigma_{\text{max}}}
-\def\sigmamin{\sigma_{\text{min}}}
-When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
-in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
-as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
-based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
-is negative we set $\sigmamax = 8\sigmamin$.)
-
-When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.
-
-Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
-a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.
-
-\appendix
-
-\begin{table}[!ht]
-\centering\begin{tabular}{c c } \toprule
-Scenario & Alternative distribution, $g_1$  \\ \midrule
-spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
-near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
-flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
- &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
-skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
-big-normal & $N(0,4^2)$ \\ 
-bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
-\end{tabular}
-\caption{Summary of simulation scenarios considered} \label{table:scenarios}
-\end{table}
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all_nopen.tex}
-	\caption{All observations}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg_nopen.tex}
-\caption{``Significant" negative discoveries.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos_nopen.tex}
-\caption{``Significant" positive discoveries.}
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{table:nopen}
-\end{table}
-%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
-%\begin{equation}
-%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
-%\end{equation}
-%which is close to the definition in \cite{storey.03}.
-
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgements}
-
-Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
-prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} , and includes contributions 
-from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.
-
-This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.
-
-%\section*{References}
-% The bibtex filename
-\bibliography{/Users/stephens/Dropbox/Documents/mainbib}
-
-\section*{Figure Legends}
-
-\clearpage
-
-\section*{Tables}
-
-\section*{Supporting Information Legends}
-
-Supplementary material can be found in {\bf Supplementary Information S1.}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
-\end{document}

### 2015-10-06 tab_coverage.tex
New Article!

--- 2015-10-06 paper/main.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
-\documentclass[11pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-%\usepackage{natbib}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-\usepackage{booktabs}
-\usepackage{xspace}
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-\usepackage{sectsty}
-\subsectionfont{\normalsize}
-\subsubsectionfont{\mdseries\itshape}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\usepackage{color}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-% Use the PLoS provided bibtex style
-\bibliographystyle{/Users/stephens/Dropbox/Documents/stylefiles/plos2009}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
-\pagestyle{myheadings}
-%% ** EDIT HERE **
-\usepackage{enumerate}
-\usepackage{multirow} 
-\usepackage{url}
-\usepackage{xr} %for cross-referencing
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-\newtheorem{algorithm}{Algorithm}
-\newtheorem{proposition}{Proposition}
-\newtheorem{restateproposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{result}{Result}
-\newtheorem{note}{Note}
-\newtheorem{definition}{Definition}
-
-\def\lfdr{\textit{lfdr}}
-\def\lfsr{\textit{lfsr}}
-\def\x{\mbf{x}}
-\def\y{\mbf{y}}
-\def\e{\mbf{e}}
-\def\g{\mbf{g}}
-\def\p{p}
-\def\Yb{\hat{Y}_1^{\text{BAYES}}}
-\def\Xperp{X_{1 \perp 0}}
-\def\Pperp{P_{1 \perp 0}}
-\def\Pperpb{\Pperp^{\text{B}}}
-\def \hatd { \widehat{ d } }
-\def \dhatin {\hatd^{\mbox{\tiny IN}}}
-\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
-\def \mvn{\text{MVN}}
-\def\mn{\text{MN}}
-\def \iw{\text{W}^{-1}}
-\def \fp {fastPHASE\ }
-\def\bhat{\hat{\beta}}
-\def\shat{\hat{s}}
-\def\var{\rm Var}
-\def\mvr{\text{BMVR}}
-\def\bfuni{\text{BF}_\text{uni}}
-\def\one{\mbf{1}}
-\def\zero{\mbf{0}}
-\def\hall{H_{\text{all}}}
-\def\pall{\p_{\text{all}}}
-\def\bfall{\text{BF}_\text{all}}
-\def\bfav{\text{BF}_\text{av}}
-\def\gall{\gamma_\text{all}}
-\def\limbfall{\bfall^\rightarrow}
-\def\limbfav{\bfav^\rightarrow}
-\def\limbfgamma{\BF_\gamma^\rightarrow}
-\def\sxx{S_{xx}}
-\def\syy{S_{yy}}
-\def\vyy{V_{yy}}
-\def\vyx{V_{yx}}
-\def\vxx{{V_{xx}}}
-\def\rss{\text{\rm RSS}}
-\def\dss{\text{\rm $\Delta$SS}}
-\def\U{U}
-\def\D{D}
-\def\I{I}
-\def\luu{L_{UU}}
-\def\indep{\perp}
-\def\qvalue{{\tt qvalue}\xspace}
-\def\locfdr{{\tt locfdr}\xspace}
-\def\mixfdr{{\tt mixfdr}\xspace}
-\def\ashr{{\tt ashr}\xspace}
-
-\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
-                                % mode too
-\newcommand\height{{\it height}}
-\newcommand\weight{{\it weight}}
-\def\BF{\text{BF}}
-
-%% END MACROS SECTION
-%\externaldocument{plos_one_rev1_suppinfo}
-
-\begin{document}
-
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{False Discovery Rates: A New Deal}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Matthew Stephens$^{1*}$, 
-\\
-\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
-\\
-$\ast$ E-mail: Corresponding mstephens@uchicago.edu
-\end{flushleft}
-
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-%\section*{Author Summary}
-
-\section*{Introduction}
-
-Since its introduction in 1995 by Benjamini and Hochberg 
-\cite{benjamini1995controlling}, the ``False Discovery Rate" (FDR) has quickly established itself
-as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing
-in which the goal is to identify the non-zero ``effects" among a large number of imprecisely-measured effects.
-
-Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
-indeed, the notion that EB approaches could be helpful in handling multiple comparisons predates introduction of
-the FDR  (e.g.~\cite{greenland1991empirical}). More recently, EB approaches to the FDR
-have been extensively studied by several authors, 
-especially B.~Efron and co-authors 
-\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
-see also \cite{kendziorski2003parametric,newton2004detecting, 
-datta2005empirical,muralidharan2010empirical} for example.
-So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
-existing widely-used FDR pipelines) and can substantially affect
-inference. The first idea is to {\it assume that the distribution of effects is unimodal}. This yields a very simple, fast, and stable computer 
-implementation, as well as improving inference of FDR when the unimodal assumption is correct.  
-The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
-to summarize each measurement. This idea allows variations in measurement precision to be better accounted for,
- and avoids a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
-
-In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
-the idea that it may be preferable to focus on estimation rather than on testing.
-In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
-simply estimation with some positive prior probability that the effect is exactly zero.
-However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
-perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
-controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
-``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
-in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
-the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the lfsr, particularly its increased robustness to modelling assumptions.
-
-%terminology
-%``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
-%and correspondingly define the ``false signed discovery rate", or simply ``false sign rate" (FSR), in a natural way.
-
-%The methods we provide here can address both problems, and we provide results that highlight important benefits of focussing on estimation, and particularly on assessing the correctness of the {\it sign} of the estimated effects, rather than on testing for non-zero effects.
-
-Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
-(We address the reasons for this name, and connections with shrinkage analysis, in the Discussion.)
-
-%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.
-
-%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
-%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
-%measurements are made on only a small number of
-%samples in each condition - sometimes as few as one
-%sample in each condition. Thus the error in estimates
-%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
-%in estimates of $\beta_j$ still greater.
-%
-
-%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
-%Here we present new tools that bring two important ideas to bear on the problem.
-%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
-%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
-%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
-%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
-%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
-%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
-%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
-%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
-%a substantial increase in the number of discoveries at a given FDR).
-
-%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
-%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
-Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
-%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
-%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
-%generic methods that work directly with two numbers for each 
-%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic).
-Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the
-%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.
-
-\def\df{df}
-\def\FDR{\text{FDR}}
-\def\fdr{\text{\lfdr}}
-\def\FSR{\text{FSR}}
-\def\fsr{\text{lfsr}}
-
- \section*{Methods}
-
- \subsection*{Model Outine}
-
-% Here we consider one setting in which FDR-based methods are widely used:
-%an experimenter has measured a large numbers of ``effects", each with an associated
-%measurement precision (standard error), 
-%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
-%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
-%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
-%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
-%For example, in genomics
-%a common goal is to identify ``differentially-expressed genes" -- that is, genes
-%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
-%mean expression level between the two conditions, often referred to as the ``log-fold change". 
-Here we describe the simplest version of our method, and briefly discuss embellishments we have also implemented.
-
-Suppose that we are interested in the values of $J$ ``effects" $\beta=(\beta_1,\dots,\beta_J)$. For example, in a typical genomics application
-that aims to identify differentially expressed genes,  $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
-In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
-that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
-more general problem of estimating, and assessing uncertainty in, $\beta_j$.
-
-Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
-and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
-which by Bayes theorem can be written as 
-\begin{equation}
-p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
-\end{equation}
-
-For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
-This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
-A simple way to implement the UA is to assume that $g$ is
-a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
-\begin{align} \label{eqn:beta}
-p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
-g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
-\end{align}
-where $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyperparameters, which are non-negative and sum to one, and are to be estimated,
-while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
-represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
-(We encourate the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
-although of course in practice we use a finite grid -- see Implementation Details.)
-
-For the likelihood $p(\bhat | \beta, \shat)$ we assume  
-\begin{equation} \label{eqn:normlik}
-p(\bhat | \beta, \shat) = \prod_j N(\bhat_j; \beta_j, \shat_j^2).
-\end{equation}
-Here, in additional to some conditional independence assumptions, we are 
-effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-
-This simple model features both the key ideas we want to emphasise in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
-the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
-with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
-However, this simple model also has several additional assumptions that can be relaxed. Specifically,
-\begin{enumerate}
-\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) also implies that $g$ is symmetric about 0; more flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see (\ref{eqn:g-general})].
-\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
-We can relax this to allow for a relationship between these quantities [see (\ref{eqn:beta-alpha})].
-\item The likelihood (\ref{eqn:normlik}) effectively assumes that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-We can generalize this likelihood using a $t$ likelihood [see (\ref{eqn:tlik})].
-\end{enumerate}
-These embellishments are detailed in Embellishments and Implementation Details.
-Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
-(which are also made by most existing EB approaches to this problem). Correlations among tests certainly arise in practice, either due to genuine correlations
-in the system of study, or due to unmeasured confounders, and their potential to impact results of an FDR analysis is important to consider
-whatever analysis methods are used: see \cite{efron2007correlation,leek:2007} for relevant discussion.
-
-%We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
-%$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
-%\begin{equation}
-%p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
-%\end{equation}
-%Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
-%\begin{equation} \label{eqn:normlik}
-%p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
-%
-%We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
-%as a mixture of null and alternative components:
-%\begin{equation} \label{eqn:betaj0}
-%g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
-%\end{equation}
-%where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
-%We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
-%\begin{equation} \label{eqn:beta-iid}
-%p(\beta | \shat) = \prod_j g(\beta_j).
-%\end{equation}
-%
-%Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
-%with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
-%to use a mixture of zero-mean normal distributions:
-%\begin{equation} \label{eqn:mixnorm}
-%g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
-%\end{equation}
-%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
-%Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
-%represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
-% The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
-% more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below. 
-% 
-% Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
-% yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
-% \begin{equation} \label{eqn:g}
-% g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
-% \end{equation}
-% 
-% Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
-% \begin{align} \label{eqn:betaj}
-% p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
-% p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
-% \end{align}
-% with $g$ given by the mixture of normals $(\ref{eqn:g})$.
-
-\subsection*{Fitting the model}
-
-In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
-each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
-Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
-\begin{equation} \label{eqn:marginallik}
-p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
-\end{equation}
-where we define $\sigma_0:=0$.
-
-The usual EB approach to fitting this model would involve two simple steps:
-\begin{enumerate}
-\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
-\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
-against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | | \bhat, \shat,\hat{\pi})$.
-\end{enumerate}
-Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(The simplicity of the EM algorithm in step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
-which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
-
-Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximise a penalized
-likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
-We introduce this penalty because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
-
-\subsection*{The local False Discovery Rate and local False Sign Rate}
-
-As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
-In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
-One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
-the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
-In other words it is the posterior probability that $\beta_j$ is actually zero:
-\begin{equation} \label{eqn:fdr}
-\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
-\end{equation}
-
-The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
-whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
-the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey1991philosophy}) argued that
-``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$.
-Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
-\begin{quote}
-...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
-\end{quote}
-Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
-focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.
-
-Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
-would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
-\begin{equation} \label{eqn:lfsr}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
-\end{equation}
-To illustrate, suppose that 
-\begin{gather*}
-p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
-p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
-p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
-\end{gather*} 
-Then 
-from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
-our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.
-
-As our notation suggests, $\lfsr_j$ is intended to be compared and contrasted with $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
-small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
-Of coure, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
-this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
-(this follows from the definition because both the events $\beta_j \geq 0$
-and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
-In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
-$\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.
-
-From these ``local" measures of significance, we can also compute average error rates over subsets of observations
-$\Gamma \subset \{1,\dots,J\}$. For example,   
-\begin{equation} \label{eqn:FDRhat}
-\widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
-\end{equation}
-estimates the FDR we would obtain if we were to declare all tests in $\Gamma$ significant.
-And 
-\begin{equation} \label{eqn:qvalue}
-q_j := \widehat \FDR(\{k: \fdr_k \leq \fdr_j\}) 
-\end{equation}
-provides a measure of significance analogous to Storey's $q$ value \cite{storey.03}.
-
-%Because our methods involve explicitly modelling effects $\beta_j$, rather than 
-%$p$ values or $z$ scores, they can provide estimates of the local false sign rates.
-
-%Indeed, even in settings where some
-%$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
-%analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.
-
-\subsection*{Related work}
-
-\subsubsection*{Previous approaches focussed on FDR}
-
-Among previous methods that explicitly consider the FDR and related quantities,
-our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
- (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
-  %Here we highlight the key differences between our approach and these three existing approaches.
-
-There are two key differences between our approach and all of these three existing methods. First, 
-whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
-we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
-for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
-Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 
-how it can improve performance by taking better account of variation in measurement precision among observations.
-%(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
-% Or more generally Section xx illustrates the potential practical effects
-
-%allows for different modelling assumptions to be compared through likelihood comparisons.
-Second, our unimodal assumption (UA) that the effects are unimodal about zero is quite different from assumptions made by $\qvalue, \locfdr$ or $\mixfdr$.
-Indeed, $\locfdr$ assumes that all $z$ scores near 0 are null (Efron calls this the Zero Assumption; ZA), which implies that 
-under the alternative hypothesis the distribution of $z$ scores has {\it no mass at 0}; this contrasts strikingly with the UA, which implies that this distribution has its peak at 0! 
-Similarly, $\qvalue$ assumes that all $p$ values near 1 are null,
-which is the same as the ZA because $p$ values near 1 correspond to $z$ scores near 0. And although \mixfdr does not formally make the ZA,
-we have found that in practice, with default settings, the results often approximately satisfy the ZA (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}).
-Thus, not only do these existing methods not make the UA, they actually make assumptions that are, in some sense, as different
-from the UA as they can be.
-
-Given that the UA and ZA are so different, it seems worth discussing why we generally favor the UA. 
-Although the UA will not apply to all situations, we believe that it will often be reasonable, especially in FDR-related contexts
-that have traditionally focussed on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis,  
-it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible. Further, it seems reasonable to expect that larger effects become
-decreasingly plausible, and so the distribution of the effects will be unimodal about 0.
-To paraphrase Tukey, ``All we know about the world teaches us that large effects are rare, whereas small effects abound."
-We emphasise that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
-It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
-
-%Tukey "peri-null"
-
-%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
-%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}.
-
-In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficents - 
-common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
-These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
-it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
-The UA assumption on regression coefficients is directly analagous to our UA here, and so its
-widespread use in the regression context supports its use here.
-
-Alternatively, we could motivate the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
-such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
-make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).
-
-Finally, the UA also has a considerable practical benefit: it yields simple algorithms that are both computationally
-and statistically stable. We illustrate these features in Results.
-
-\subsubsection*{Other work}
-
-There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
- Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
-indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
-This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
-Also, they provide no software implementation.
-More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
-Some pointers to recent papers in which other relevant
-citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}.
-Much of the literature focusses on the homoskedastic case (i.e.
-$\shat_j$ all equal) whereas we allow for heteroskedasticity.
-And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential.
-Several recent papers consider more flexible non-parametric assumptions on
-$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
-estimate (NPMLE) for $g$.
-These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive.
-However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
-may perform well for point estimation, it seems possible it will not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
-Indeed, one way of thinking about the UA is as a way to regularize $g$.
-
-%The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
-%encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
-%$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
-%$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
-%the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
-%observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
-%corresondingly little effect on the estimated FDR of other observations. 
-%We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
-%$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
-%that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
-%estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
-%Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.
-
-\def\dt{\text{dt}}
-
-%
-%assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
-%$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
-% $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
-%  \begin{equation} \label{eqn:lik}
-% \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
-% \end{equation}
-% where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e.
-the density of $\mu+\sigma T_\nu$).
-In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
-%  \begin{equation} \label{eqn:normlik}
-% \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
-% \end{equation}
-%which, in light of (\ref{eqn:betahat}), is equivalent to
- % \begin{equation} \label{eqn:lik2}
-% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
- %\end{equation}
-
-%The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e.
-they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$.
-The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$.
-Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$.
-If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$.
-See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.
-
-%This likelihood approximation can be motivated by assuming that i) $\bhat$
-% $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
-% $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$. 
-%
-
-%Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.
-
-%While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
- %that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
-%there is an important difference: here,
-
-%The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways.
-For example, we can write $p(\bhat_j, \shat_j | \beta_j) = p(\bhat_j | \shat_j, \beta_j)p(\shat_j | \beta_j)$.
-Now, if we are willing to assume that $\shat_j$ alone contains no information about $\beta_j$, or equivalently that $p(\shat_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\bhat_j | \shat_j, \beta_j)$, and if $\bhat_j | \shat_j, \beta_j \sim N(\beta_j, \shat_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.
-
-\section*{Results}
-
-We compare results of \ashr with existing FDR-based methods implemented
-in the R packages \qvalue (v2.0 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
-In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
-to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.
-
-\subsection*{Effects of the Unimodal Assumption}
-
-Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
-standard error, $s_j=1$ and all methods are provided that information. That is,
-$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$.
-In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the only difference between our method and methods like $\locfdr$ and $\mixfdr$ are in how they estimate $g$.
-
-To briefly summarize the results in this section:
-\begin{enumerate}
-\item The UA can produce very different inferences compared with the ZA made by existing methods.
-\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
-\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
-\end{enumerate}
-
-\subsubsection*{The UA and ZA can produce different inferences}
-
-\begin{figure}
-\center \includegraphics[height=6in]{figures/decomp_ZA.pdf}
-\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
-The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
-In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
-The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
-(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
-\end{figure}
-
-To illustrate the different inferences from the UA and ZA we show results for a single
-dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
-but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
-\qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
-The results (Figure \ref{fig:ZA}) illustrate the clear difference between the existing methods and our method. The effects of the ZA made by \qvalue and \locfdr
-are visually clear, producing a ``hole"  in the alternative $z$ score distribution around 0. Although \mixfdr does not formally make the ZA, its decomposition
-exhibits a similar hole. In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
-effectively ``filling in" the hole.  (Of course the null distribution also has a peak at 0, and the local fdr 
- under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)
-
-Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA.
-Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
-This would increase the null component (dark blue) at the expense of
-the alternative component (light blue).
-Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0.
-The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears.
-In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
-See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
-demonstration.
-
- \subsubsection*{The UA can produce conservative estimates of $\pi_0$}
-
-The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
-Consequently \ashr will estimate smaller $\lfdr$s and FDRs than existing methods that make the ZA. 
-This is desirable, provided that these estimates remain conservative: that is,
-provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
-The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behaviour. To check its effectiveness
-we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
-for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
-They range from a ``spiky" distribution -- where many non-zero $\beta$ are
-too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
-flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
-making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
-which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.
-
-% How do these differences influence estimates of the FDR? To address this, note that  -- in this simple case where the 
-% $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
-% of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
-% strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
-% although the methods considered here differ from one another in many ways, having estimated $\pi_0$
-% they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
-% empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
-%
-
-%Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
-% and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
-%Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
-%At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
-%overestimates of $\pi_0$, which will yield overestimates of the FDR).
-
-\begin{figure} 
-\begin{center}
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=1.8in]{figures/scenarios_density.pdf} 
-	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/pi0_est.pdf} 
-\caption{Comparison of true and estimated values of $\pi_0$.
-When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ashr being least conservative, and hence most accurate.
-When the UA does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.} \label{fig:pi0sims}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfdr_est.png} 
-\caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
-Black line is $y=x$ and red line is $y=2x$.
-Estimates of $\lfdr$ are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est.png} 
-\caption{As in c), but for $\lfsr$ instead of $\lfdr$.
-Estimates of $\lfsr$ are consistently less conservative than $\lfdr$ when UA holds, and also less anti-conservative in bimodal scenario.}  \label{fig:lfsr}
-\end{subfigure}
-\end{center}
-\caption{Results of simulation studies (constant precision $s_j=1$).}
-\end{figure}
-
-For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
-data as follows:
-\begin{enumerate}
-\item Simulate $\pi_0  \sim U[0,1]$.
-\item For $j=1,\dots,J$, simulate $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)$.
-\item For $j=1,\dots,J$, simulate $\bhat_j | \beta_j \sim N(\beta_j,1)$.
-\end{enumerate}
-%Thus these simulations assume the same
-%precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).
-
-Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ashr ($y$ axis) with the true values ($x$ axis). 
-For \ashr we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
-and as a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
-as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
-The results show that \ashr provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
-in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ashr estimates can be slightly anti-conservative.
-We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
-are used. (In addition, the effects on $\lfsr$ estimates turn out to be relatively modest; see below).
-
-%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
-%
-%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
-%and so all effects are non-zero and no nulls hold. This scenario is rather different
-%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
-%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
-%``mostly non-null"
-%scenario, and we would agree! On the other hand, the distribution of $p$ values
-%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
-%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
-%Thus it seems desirable to have methods that perform well in both cases, and this is
-%an important feature of the methods we introduce here.
-%
-%TODO: put FDRs on graph? Simulate with 50\% null?
-
-\subsubsection*{The $\lfsr$ is more robust than \lfdr}
-
-The results above show that \ashr can improve on existing methods in producing smaller, more accurate, 
-estimates of $\pi_0$, which will lead to more accurate estimates of FDR.
-Nonetheless, in many scenarios \ashr continues
-to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
-This is because these scenarios include 
-an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
-estimation of $\pi_0$ impossible. Put another way, and as is well known, $\pi_0$ is not identifiable:
- the data can effectively provide an upper bound on plausible values of $\pi_0$,
-but not a lower bound (because the data cannot rule out that everything is non-null, but with miniscule effects).
-To obtain conservative behaviour we must estimate $\pi_0$ by this upper bound, which can
-be substantially larger than the true value.
-
-Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
-overestimation of $\pi_0$ is corresponding overestimation of FDR (and $\lfdr$, and $q$ values).
-To illustrate, Figure \ref{fig:lfdr} compares the estimated $\lfdr$ from ash.n with the true value (computed using
-Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios which involve many
-non-zero effects that are very near 0 (e.g.~the spiky scenario with $\pi_0$ small) where
-$\pi_0$ can be grossly overestimated.  
-(Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.)
-
-The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
- become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
- This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
-To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
-be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
-the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in $\lfsr$ than $\lfdr$.
-
-Note that, compared with previous debates regarding testing, 
-this section advances an additional reason for focussing on the sign of the effect, rather than just testing whether it is 0. 
-In previous debates authors have argued against testing whether an effect is 0
-because it is {\it implausible that effects are exactly 0}. Here we add that {\it even if one believes
-that some effects may be exactly zero}, it is still better to focus on the sign, because generally {\it the data are more informative about that question}
-and so inferences are more robust to, say, the inevitable mis-estimation of $\pi_0$.
-To provide some intuition, consider an observation with a $z$ score of 0. The $\lfdr$ of this observation can range from 0 (if $\pi_0=0$)
-to 1 (if $\pi_0=1$). But, assuming a symmetric $g$, the $\lfsr>0.5$ whatever the value of $\pi_0$, because the observation $z=0$ says
-nothing about the sign of the effect.
-Thus, are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
-whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.
-
-Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
-Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
-if we are prepared to assume that no effects are exactly zero, then removing the point mass 
-yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
-However, there is ``no free lunch":  if in fact some effects are exactly zero
-then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
-We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.
-
-\begin{figure}
-\begin{center}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s_nn.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be more accurate in this case.
-That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
-\end{subfigure}
-\end{center}
-\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
-\end{figure}
-
-\subsubsection*{Numerical Stability}
-
-The EM algorithm, which we use here to fit our model, is notorious for convergence to local optima. 
-However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
-caused by such behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
-in addition running it using our default initialization procedure (see Implementation details). We then compared the largest
-log-likelihood achieved across all 11 runs with the log-likelihood achieved by the default run.
-When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference ($< 0.02$), 
-and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference ($<0.02$),
-and 6\% of runs showed an appreciable log-likelihood difference ($>1$), with the largest difference being 5.0. However, perhaps suprisingly, 
-even for this largest difference results from the default run (e.g.
-the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood.
-See \url{http://github.com/stephens999/ash/blob/master/dsc-robust/summarize_dsc_robust.rmd} for further details.
-
-\subsubsection*{The UA helps provide reliable estimates of $g$}
-
-An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
-can estimate the {\it size} of each effect $\beta_j$.
-Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
-to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
-threshold $T$.
-%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
-%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
-%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
-%towards zero, with the strength of the shrinkage being estimated from the data.
-Further, because the posterior distribution is, by definition,
-conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
-been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
-In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
-\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
-how can confidence intervals be assigned to the true
-effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
-and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
-%but also notes the difficulties of implementing a stable general algorithm.
-
-The ability of the EB approach to provide valid ``post-selection" interval estimates is extremely attractive in principle.
-But its usefulness in practice
-depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
-which are notoriously difficult in general. Indeed, Efron emphasises
-the difficulties of implementing a stable general algorithm, noting in his rejoinder
-``the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying
-to be overly nonparametric ... but it is hard to imagine a
-generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
-Our key point here is that the UA greatly simplifies the deconvolution problem.
-While not meeting Efron's desire for an entirely general nonparametric approach, we believe 
-that the UA can handle many cases of practical interest.
-
-To illustrate this, Figure \ref{fig:egcdf} compares the
-estimated $g$ from \ashr  with that from \mixfdr which does not make the UA 
-(and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
-The greater reliability of estimates afforded
-by the UA is immediately apparent. In particular the estimated cdf from \mixfdr often has an almost-vertical
-segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. 
-The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
-While the UA is not the only way to achieve this, we find it an attractive, simple and effective approach.
-
-Interestingly, even in the ``bimodal" scenario $\ashr$ 
-is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
-Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
-Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
-to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
-(\url{http://github.com/stephens999/ash/blob/master/dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
-clearly visible in the observed $\bhat$.
-
-We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ashr tends to systematically
-overestimate the mass of $g$ near zero. On careful inspection, this is apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
-and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf}b),
-and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
-under-estimating $\pi_0$ (Figure \ref{fig:egcdf}c).
-
-\begin{figure}[h!] 
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=2in]{figures/egcdf.pdf}
-	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{figures/mean_cdf_nopen.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
-Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
-\end{subfigure}
-\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.} \label{fig:egcdf}
-\end{figure}
-
-%making the UA provides an attractive and simple way to achieve estimating $g$  large
-%subset of cases of interest (those where the UA can reasonably be assumed to hold).
-
-%In practice, the utility of the posterior distributions noted above depends primarily on the accuracy
-%of the estimate of the distribution $g(\cdot)$.
-
- %quote from efron rejoinder , p46:
-%MorrisÕ Section 3 is especially pertinent. His formula
-%for p(?|z) is related to my discussion of the
-%BenjaminiÐYekutieli False Coverage Rate method in
-%Section 7, particularly (7.2)Ð(7.4). Originally I had
-%hoped to develop an empirical Bayes method for estimating
-%such models, but the effort foundered on practical
-%difficulties involving the perils of deconvolution.
-%Section 6 on the Òone-group modelÓ is the ugly duckling
-%of the current paper, but it bears on some important
-%points raised in the discussion. Figure 7 concerns
-%a fuzzy version of simultaneous hypothesis testing,
-%where, as in MorrisÕ hospital example, the usual singlepoint
-%null hypotheses seem unequal to the task. The
-%development from (6.6) onwards, particularly (6.12),
-%bears on the possibility of nonnormal null distributions,
-%and is about as far as I can go in answering Professors
-%Rice and SpiegelhalterÕs penultimate question.
-%With g(?) a normal distribution, model (6.1) returns
-%us to the realm of Stein estimation, the scene of my
-%happy collaborations with Carl Morris. I continue to
-%be surprised at how much harder bumpy, nonnormal
-%models like (7.1) are to deal with. JamesÐStein estimation
-%works fine with, say, N = 10 component problems,
-%while the RobbinsÕ form of empirical Bayes appropriate
-%to (7.1) seems to require hundreds or thousands.
-%The information calculations in Efron (2008)
-%reinforce this gloomy assessment. Maybe I am trying
-%to be overly nonparametric in constructing the empirical
-%Bayes Fdr estimates, but it is hard to imagine a
-%generally satisfactory parametric formulation for (6.1).
-%Or perhaps it is just that hypothesis testing is more demanding
-%than estimation.
-%My Note: (6.1) is the general model (mu \sim g()).
-%7.1 is the case where the effects are a mixture of zero and N(-3,1).
-
-\subsubsection*{Calibration of posterior intervals}
-
-To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over 100 simulations in each Scenario). 
-Specifically we examine the empirical coverage of nominal lower 95\% credible bounds for a) all observations; b) significant negative discoveries; c) significant positive discoveries.
-We examine only lower bounds because the results for upper bounds follow by
-symmetry (except for the one asymmetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
-in each case: for negative discoveries the lower bound is typically large and negative and limits how big (in absolute value) 
-the effect could be; for positive discoveries the lower bound 
-is positive, and limits how small (in absolute value) the effect could be. Intuitively, the lower bound for negative discoveries depends on the accuracy of $g$ in its tail,
-whereas for positive discoveries it is more dependent on the accuracy of $g$ in the center.
-
-The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
-as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
-
-% As noted above, the penalty term (\ref{eqn:penalty}) causes systematic over-estimation of the mass of $g$ near 0,  which one would expect to overshrink interval estimates towards zero. 
-%The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound is negative and overshrinking (moving it too far towards zero) makes the left tail too big; conversely, for positive discoveries, the lower confidence bound is positive, and
-%overshrinking (moving it too far towards zero) makes the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
-%and be prepared to put up with some overshrinkage to avoid it.
-
-%Coverage rates are given in:
-%\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
-
-%Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.
-
-\begin{table}[!ht]
+\begin{table}
+[!ht]
 \begin{subtable}{\textwidth}
@@
 \caption{Table of empirical coverage for nominal 95\% lower credible bounds} \label{tab:coverage}
+
 \end{table}
-
-%
-%\subsubsection*{Reverse asymptotics}
-%
-%Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
-%
-%Note: we'd like to make sure it behaves ok in the case where most observations are null.
-%Perhaps simulate with 999 null and 1 alternative?
-%Also maybe look at performance for small sample sizes - e.g. n=1?
-%
-%Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
-%can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
-%by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
-%
-
-\subsection*{Differing measurement precision across units}
-
- We turn now to the second important component of our work: allowing for varying
- measurement precision across units. The key to this is the use of a likelihood,
- (\ref{eqn:normlik}) or (\ref{eqn:tlik}), that explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.
-
-  To illustrate, we conduct a simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
- ($s_j=10$).  In both cases,
-we assume that half the effects are null
-and the other half are normally distributed with standard deviation 1:
-\begin{equation}
-p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1).
-\end{equation}
-In this setting, the poor-precision measurements $(s_j=10)$ tell us very little, and any sane analysis should effectively ignore them. 
- However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor}). This is because the poor measurements 
- produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
- which, when combined with the good-precision measurements, dilute the overall signal (e.g. they reduce the density of $p$ values near 0).
- This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
-the estimated error rates ($q$-values, or $\lfdr$ values) for the good-precision observations increase when the low-precision observations are included in the analysis
-(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
-for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).
-
-\begin{figure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_hist.pdf} 
-    \label{fig:goodpoor-hist}
-    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations}
- \end{subfigure}
-\begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_scatter.pdf}
-    \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) of a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
-\end{subfigure}
-\caption{Simulation illustrating how, for existing FDR methods, 
-poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
-$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
-a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
-In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
-\end{figure}
-
- \subsubsection*{Reordering of significance, and the ``$p$ value prior"}
-
-Another consequence of accounting for differences in measurement precision is that \ashr may re-order the significance of the observations compared
- with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
- We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measuments that have the same $p$ value.
- The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
- and so the $\lfsr$ for these poor-precision measurements is always high.
-
-The potential for Bayesian analyses to re-order the significance of observations, and specifically to downweight imprecise observations,
-was previously discussed in \cite{guan.stephens.08}.
-However, Wakefield \cite{wakefield:2009} showed that, under a certain prior assumption, the Bayesian analysis produces the same ranking of significance as $p$ values (or their $z$ scores).
-He named this prior the ``$p$-value prior" because it can be thought of as the
-implicit prior assumption that is being made if we rank the significance of observations by their $p$ value.
-Wakefield's $p$-value prior assumes that the less precise effect estimates correspond to larger true effects, and specifically that they scale proportional
-to the standard errors $s_j$. More specifically still, it assumes a normal prior for the non-zero $\beta_j$ with mean 0 and variance $K s^2_j$ for some constant $K$.
-Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. Specifically, if, instead of assuming that $\beta_j$ is independent
-of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
-then the $\lfsr$ computed by ash.n provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
-as is illustrated in Figure \ref{fig:lfsr_pval}, right panel. (This result does not hold when using the mixtures of uniforms prior, ash.u.)
-
-This $p$-value prior assumes that the $z$ scores $z_j=\beta_j/s_j$
-are identically distributed, independent of $s_j$. This is essentially the assumption made, implicitly or explicitly, by existing methods -- like \locfdr, \mixfdr and \qvalue --
-that model the $z_j$ or $p_j$ directly.  In contrast, we have assumed up to now that the $\beta_j$ are independent of $s_j$. 
-We can set both these assumptions within a more general framework, which allows that $\beta_j/s^\alpha$ is independent of $s_j$ for some $\alpha$ [Equation (\ref{eqn:beta-alpha}))].
-Setting $\alpha=0$ implies that $\beta_j$ is
-independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
-effects (in absolute value). This latter assumption may often be qualitatively plausible: 
-for example, in gene expression studies the standard error for gene $j$ depends partly on
-the variance of its expression among samples, and genes with a larger variance may tend to be less tightly regulated and
-so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
-On the other hand, there is no particular reason to expect that either $\alpha=1$ or $\alpha=0$ will be the optimal choice.
-Indeed, optimal choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$,
-which will be dataset-specific. 
-Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
-comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
-likelihood-based methods can be used to select $\alpha$. For example, following the logic of the EB approach it
-would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
-which has been implemented in our software by C. Dai.
-
-%\begin{result}
-%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
-%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
-%\end{result}
-
-%To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
-%Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
-%In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
-%Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
-%which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
-%(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
-%may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)
-
-%\begin{result}
-%Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and using a mixture of normals for the prior distribution 
-%(\ref{eqn:beta-alpha}) with $\alpha=1$, 
-%the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$.
-Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$)
-%is montonic with $|t_j|$. 
-%\end{result}
-
-%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
-%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.
-
-%Most previous authors of EB approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
-%have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
-%In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
-%This is analogous
-%to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
-%standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
-%that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.
-
-\begin{figure}
-\centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
-\caption{Figure illustrating affects of prior assumptions on re-ordering of significance.
-Left panel shows results under our ``default prior" which assumes that effects $\beta_j$ are identically distributed, independent of $s_j$.
-Right panel shows results under the ``$p$-value prior", which assumes that $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$.} \label{fig:lfsr_pval}
-\end{figure}
-
-%\subsubsection*{Adaptive Shrinkage}
-%
-%Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
-%when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
-%the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
-% $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
-% In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
-% Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
-% and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
-%
-%Our methods we build on this idea in two ways. First,
-%by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
-%Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
-%adapts to measurement precision: more precise measurements undergo
-% less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
-%than observations with
-%small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
-%To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
-%Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
-%
-%\begin{figure}[!ht]
-%\begin{center}
-%\includegraphics[width=400px]{../examples/ash.pdf} 
-%\end{center}
-%\caption{Figure illustrating adaptive shrinkage.
-Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0.
-The standard error of each observation, $s_j$, was simulated from Inverse Gamma(5,5), which resulted in $s_j$ varying from 0.34 to 6.7.
-The shrunken estimates ($y$ axis) are plotted against the observed estimates ($x$ axis), with color indicating the (square root of the) precision of each measurement (blue=lower-precision; red=higher-precision).
-The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
-%\end{figure}
-
-%DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
-%To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$.
-Because the prior distribution is unimodal about zero,
-%the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
-%amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
-%For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
-%than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
-%small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.
-
-%previous work based on normal means:
-%http://link.springer.com/article/10.1007/BF02562623
-
-%\subsection{Null behaviour}
-
-%Results of simulations under the null.
-
-%\subsection{Comparison of different underlying component distributions}
-%
-%Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
-%than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
-%likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
-%For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
-%increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
-%expect $L_{hu}$ to be significantly larger than the others.
-%
-%Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
-%
-%\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
-%\begin{center}
-%\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
-%%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
-%\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
-%\end{center}
-%\caption{}
-%\end{figure}
-
-%Notes from Adrian Raftery - 
-%Newcombe (1886)
-%Dean and Raftery
-%Uniform is super-efficient
-
-\section*{Discussion}
-
-We have presented an Empirical Bayes approach to large-scale multiple testing that emphasises two ideas.
-First, we emphasise the potential benefits of using two numbers ($\bhat$, and its standard error)
-rather than just one number (a $p$ value or $z$ score)  to summarize the information on each test.
-While requiring two numbers is slightly more onorous than requiring one, in many settings these
-numbers are easily available and if so we argue it makes sense to use them.
-Second, we note the potential benefits -- both statistical and computational -- of assuming that the effects come from a unimodal distribution, and provide
-flexible implementations for performing inference under this assumption. We also introduce the ``false sign rate" as
-an alternative measure of error to the FDR, and illustrate its improved robustness to errors in model fit, particularly mis-estimation
-of the proportion of null tests, $\pi_0$.
-
-Multiple testing is often referred to as a ``problem" or a ``burden". In our opinion, EB approaches turn this idea on its head,
-treating multiple testing as an {\it opportunity}: specifically, an opportunity to learn about the prior distributions, and other modelling assumptions,
-to improve inference and make informed decisions about significance (see also \cite{greenland1991empirical}). This view
-also emphasises that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
-Indeed, the FDR at a given fixed threshold does not depend on the number of tests: as the number of tests increases, both the true positives and false positives increase linearly,
-and the FDR remains the same.
-(If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example.
-The EB approach captures this dependence in an intuitive way:
-if there are lots of strong signals then we infer $\pi_0$ to be small, and the estimated FDR (or $\lfdr$, or $\lfsr$) at a given threshold may be low, even if a large number of tests were performed; and conversely
-if there are no strong signals then we infer $\pi_0$ to be large and the FDR at the same threshold may be high, even if relatively few tests were performed. 
-More generally, overall signal strength is reflected in the estimated $g$, which in turn influences the estimated FDR.
-
-Two important practical issues that we have not addressed here are correlations among tests,
-and the potential for deviations from the theoretical null distributions of test statistics. These two
-issues are connected: specifically, unmeasured confounding factors can cause both correlations among tests
-and deviations from the theoretical null \cite{efron2007correlation,leek:2007}. And although there are certainly other
-factors that could cause dependence among tests, unmeasured confounders are perhaps the most worrisome in practice
-because they can induce strong correlations among large numbers of tests and profoundly impact results.
-Approaches to deal with unmeasured confounders can be largely divided into two types: those
-that simply attempt to correct for the resulting inflation of test statistics \cite{devlin1999genomic,efron2004large}, and those
-that attempt to infer counfounders using clustering, principal components analysis, or factor models 
-\cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
-and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable,
-they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alterntively, our methods
-could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:lik}) using
-$\bhat_j | \shat_j \sim N(0,\lambda_1 \shat_j + \lambda_2)$, where $\lambda_1,\lambda_2$ are to be estimated. However,
-this immediately raises issues of identifiability and we do not pursue the idea further here.
-
-Another important practical issue is the challenge of small sample sizes. For example, in genomics applications researchers sometimes
-attempt to identify differences between two conditions based on only a handful of samples in each. In such settings the
-normal likelihood approximation (\ref{eqn:normlik}) will be inadequate. And, although the $t$ likelihood (\ref{eqn:tlik}) partially
-addresses this issue, it is also, it turns out, not entirely satisfactory.  The root of the problem is that, with small sample sizes, raw estimated standard
-errors $\shat_j$ can be horribly variable. In genomics it is routine to address this issue by applying EB methods \cite{smyth:2004} to ``moderate" (i.e.~shrink) variance estimates,
-before computing $p$ values from ``moderated" test statistics. We are currently investigating how
-our methods should incorporate such ``moderated" variance estimates to make it applicable to small sample settings.
-
-Our approach involves compromises between flexibility, generality, and simplicity on the one hand, and statistical efficiency and principle on the other.
-For example, in using an EB approach that uses a point estimate for $g$, rather than a fully Bayes approach that accounts for uncertainty in $g$,
-we have opted for simplicity over statistical principle. And in summarizing every test by
-two numbers and making a normal or $t$ approximation to the likelihood, we have aimed to produce generic methods that 
- can be applied whenever such summary data are available -- just as \qvalue can be applied to any set of $p$ values for example -- although
- possibly at the expense of statistical efficiency compared with developing multiple tailored approaches based on context-specific likelihoods. 
- Any attempt to produce generic methods will involve compromise between generality and efficiency.
- In genomics, many analyses -- not only FDR-based analyses -- involve first computing a series of $p$ values before
- subjecting them to some further downstream analysis. An important message here is 
- that working with two numbers ($\bhat_j,\shat_j$), rather than one ($p_j$ or $z_j$), 
- can yield substantial gains in functionality (e.g.~ estimating effect sizes, as well as testing; accounting for variations in
-measurement precision across units) while losing only a little in generality.  We hope that our work will
-encourate development of methods that exploit this idea in other contexts.
-
-%This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
-%from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
-%with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.
-
-%
-%
-%Both \locfdr and \mixfdr allow for deviations from
-%the theoretical null distribution by allowing that, under the null, the $z$ scores are $N(0,\sigma^2)$ with $\sigma$ estimated from the data rather
-%than fixed to 1. Within our approach this idea could be incorporated by changing the likelihood - for example, to assume $\bhat_j \sim N(0,s_ 
-%
-%These are tricky issues that, as far as we are aware, can affect
-%all approaches to FDR analysis. An important focus of Efron's work, also implemented in \locfdr, 
-% is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
-%theoretical null distribution. 
-%
-%For example, we and others have worked on Bayesian approaches to genetic association testing, using subjective prior specification of
-%effect sizes based on external data (e.g.~\cite{stephens:2009}). However, the large numbers of tests brings the opportunity to learn about
-%priors, and here we aim to exploit this
-
-%Choice of mixture components. Use of likelihood to compare different mixture components?
-
-%  
-%Directly modeling the p values, or z scores, say via non-parametric methods, 
-%can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
-%distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
-%The solution is to model beta hat as a convolution of some distribution g and an error component.
-
-%Before proceeding, we feel obliged to correct the common misconception
-% that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
-% not depend on the number of tests. 
-%  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
-% 
-% In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.
-
-\section*{Detailed Methods} \label{sec:implement}
-
-\subsection*{Embellishments}
-
-\subsubsection*{More flexible unimodal distributions}
-
-Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
-Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
- Therefore, we have implemented a more general approach based on 
-\begin{equation} \label{eqn:g-general}
-g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) ,
-\end{equation}
-where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
-\begin{enumerate}[(i)]
-\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$, \qquad (``ash.n")
-\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,  \qquad (``ash.u")
-\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,  \qquad (``ash.hu")
-\end{enumerate}
-where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
-(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
-and $\pi$ is a $2K+1$ vector that sums to 1.)
-The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
-Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
-under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
-(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
-(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
-however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)
-
-Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
-using a large dense grid of $\sigma^2_k$ or $a_k$, (i)-(iii) can respectively approximate,
-with arbitrary accuracy,
-\begin{enumerate}[(i)]
-\item any scale mixture of normals, which includes as special cases
-the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
-\item any symmetric unimodal distribution about 0.
-\item any unimodal distribution about 0.
-\end{enumerate}
-The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
-In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.
-
-Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
-The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
-See Implementation for details of our software defaults.
-
-%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting.
-For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint.
-(Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
-%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).
-
-\subsubsection*{Dependence of effects on standard errors}
-
-Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
-This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
- \begin{equation} \label{eqn:beta-alpha}
- \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
- \end{equation}
-for any $\alpha$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
-assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
-it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
-the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.
-
-The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
-To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
- standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
- Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
-yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
-%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
-%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
-%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.
-
-%Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
-%how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
-%it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
-%rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
-%and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
-%this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
-%do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.
-
-\subsubsection*{Replace normal likelihood with $t$ likelihood}
-
-We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
-  \begin{equation} \label{eqn:tlik}
- \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
- \end{equation}
-where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
-and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
-The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
-This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms; see Implementation below.
-When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.
-
-Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
-\begin{equation} \label{eqn:tdist}
-(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
-\end{equation}
-However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
-In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
-currently pursuing this improved approach (joint work with M.Lu) and results will be published elsewhere. 
-%We note here that
-%modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
-%where our simpler approach can produce unsatisfactory results.
-
-\subsubsection*{Non-zero mode}
-
-An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood, rather than fixed to 0.
-
-\subsection*{Implementation Details}
-
-\subsubsection*{Likelihood for $\pi$}
-
-We define the likelihood for $\pi$ to be the probability of the observed data $\bhat$ conditional on $\shat$: $L(\pi) := p(\bhat | \shat,\pi),$ which
-by our conditional indendence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$.
-[One might prefer to define the likelihood as $p(\bhat, \shat | \pi) = p(\bhat | \shat, \pi) p(\shat | \pi)$, in which case our definition comes down to assuming that the term $p(\shat | \pi)$ does not depend on $\pi$.]
-
-Using the prior $\beta_j \sim \sum_{k=0}^K \pi_k f_k(\beta_j)$ given by (\ref{eqn:g-general}), and the normal likelihood (\ref{eqn:normlik}), integrating over $\beta_j$ yields
-\begin{equation}
-p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
-\end{equation}
-where 
-\begin{equation}
-\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
-\end{equation}
-denotes the convolution of $f_k$ with a normal density.
-These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
-Specifically, 
-\begin{equation} \label{eqn:uconv}
-\tilde{f}_k(\bhat_j)  = 
-\begin{cases}
-N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
-\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
-\end{cases}
-\end{equation}
-where $\Psi$ denotes the cumulative distribution function (c.d.f.) of the standard normal distribution.
-If we replace the normal likelihood with the $t_\nu$ likelihood (\ref{eqn:tlik}) then the convolution 
-for $f_k$ uniform the convolution is still given by (\ref{eqn:uconv}) but with $\Psi$ the c.d.f. 
-of the $t_\nu$ distribution function. (The convolution for $f_k$ normal is tricky and we have not implemented it.)
-
-\subsubsection*{Penalty term on $\pi$}
-
-To make $\lfdr$ and $\lfsr$ estimates from our method ``conservative" we add a penalty 
-term $log(h(\pi;\lambda))$ to the log-likelihood $\log L(\pi)$ to encourage over-estimation of $\pi_0$:
-\begin{equation} \label{eqn:penalty}
-h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
-\end{equation}
-where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
-consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0sims}).
-
-Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
-we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.
-
-\subsubsection*{Problems with removing the penalty term in the half-uniform case}
-
-It is straightforward to remove the penalty term by setting $\lambda_k=1$ in (\ref{eqn:penalty}).
-We note here an unanticipated problem we came across when using no penalty term in the half-uniform case 
-(i.e.~ $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$ in (\ref{eqn:g-general})): when the data
-are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
-sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
-for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
-and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.
-
-\subsubsection*{EM algorithm}
-
-With this in place, the penalized log-likelihood for $\pi$ is given by:
-\begin{equation}
-\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
-\end{equation}
-where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This can be maximized using an EM algorithm \cite{dempster77}, whose
-one-step updates are:
-\begin{align} \label{eqn:w}
-w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
-n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
-\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}
-\end{align}
-Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
-and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
-probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$. (The $w_{kj}$ are sometimes referred to as the ``responsibilities".)
-Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
-We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple} to accelerate convergence of this EM algorithm.
-
-%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
-\subsubsection*{Initialization}
-
-By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
-Our rationale for initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
-but weak signal in the data cannot quickly draw the algorithm towards the null.
-
-In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
-$\pi_0=1; \pi_k=0 (k>0)$. If not then we replace the EM solution with the global null solution. This aims to guard against errors due to convergence to a local optimum when the
-data are consistent with the global null.
-
-\subsubsection*{Conditional distributions}
-
-Given $\hat\pi$, we compute the conditional distributions 
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
-\end{equation} 
-Each posterior is a mixture on $K+1$ components:
-\begin{equation}
-p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
-\end{equation}
-where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
-and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
-$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
-All these posterior distributions are easily available.
-For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
-If $f_k$ is normal and $L$ is normal, then this is a normal distribution.
-
-\subsubsection*{Choice of grid for $\sigma_k, a_k$} \label{sec:grid}
-
-\def\sigmamax{\sigma_{\text{max}}}
-\def\sigmamin{\sigma_{\text{min}}}
-When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
-in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
-as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
-based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
-is negative we set $\sigmamax = 8\sigmamin$.)
-
-When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.
-
-Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
-a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.
-
-\appendix
-
-\begin{table}[!ht]
-\centering\begin{tabular}{c c } \toprule
-Scenario & Alternative distribution, $g_1$  \\ \midrule
-spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
-near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
-flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
- &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
-skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
-big-normal & $N(0,4^2)$ \\ 
-bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
-\end{tabular}
-\caption{Summary of simulation scenarios considered} \label{table:scenarios}
-\end{table}
-
-\begin{table}[!ht]
-\begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all_nopen.tex}
-	\caption{All observations}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg_nopen.tex}
-\caption{``Significant" negative discoveries.}
-\end{subtable}
-
-\begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos_nopen.tex}
-\caption{``Significant" positive discoveries.}
-\end{subtable}
-
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{table:nopen}
-\end{table}
-%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
-%\begin{equation}
-%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
-%\end{equation}
-%which is close to the definition in \cite{storey.03}.
-
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgements}
-
-Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
-prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} , and includes contributions 
-from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.
-
-This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.
-
-%\section*{References}
-% The bibtex filename
-\bibliography{/Users/stephens/Dropbox/Documents/mainbib}
-
-\section*{Figure Legends}
-
-\clearpage
-
-\section*{Tables}
-
-\section*{Supporting Information Legends}
-
-Supplementary material can be found in {\bf Supplementary Information S1.}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
-\end{document}

### 2015-10-13 paper/main.tex
added abstract

@@
 
+We introduce a novel Empirical Bayes approach for performing large-scale hypothesis testing, including  
+estimating False Discovery Rates (FDRs), and estimating effect sizes. Compared with
+existing commonly-used approaches to FDR analysis, the method has two key differences. First, it
+assumes that the distribution of the actual (unobserved) effects being tested is unimodal, with a mode at 0.
+This ``unimodal assumption" (UA), which is common in other related contexts, but very different from assumptions usually made
+in FDR analyses, yields more accurate estimates of FDR than existing methods, provided the UA holds.
+Second, the method takes as it input two numbers for each test (an effect size estimate, and corresponding standard error),
+rather than the one number usually used ($p$ value, or $z$ score). Provided these two numbers are available,
+using them allows our method to better account for
+variations in measurement precision across tests. It also facilitates the estimation of actual 
+effect sizes, and our approach provides ``shrunken" interval estimates (credible regions) for each effect in addition to measures of ``significance".
+To provide a bridge between interval estimates and signicance measures we introduce the term ``local false sign rate"
+to refer to the probability of getting the sign of an effect wrong, and argue that it is a superior
+measure of significance than the local FDR because it is both more generally applicable, and
+can be more robustly estimated. Our methods are implemented in an R package {\tt ashr}
+available from \url{http://github.com/stephens999/ashr}.
+
 % Please keep the Author Summary between 150 and 200 words
@@
 in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
-the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the lfsr, particularly its increased robustness to modelling assumptions.
+the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the $\lfsr$, particularly its increased robustness to modelling assumptions.
 
@@
 Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
-(We address the reasons for this name, and connections with shrinkage analysis, in the Discussion.)
+(The names comes from the fact that EB methods in general, and -- due to the UA -- our EB method in particular, is a form of shrinkage estimation,
+and the shrinkage is adaptive to both the amount of signal in the data and the measurement precision of each observation. 
+This will be discussed further in Xing and Stephens (in preparation).)
 
@@
 \end{enumerate}
-These embellishments are detailed in Embellishments and Implementation Details.
+These embellishments are detailed in Detailed Methods.
 Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
@@
 We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.
-
-\begin{figure}
-\begin{center}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s_nn.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be more accurate in this case.
-That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{figures/lfsr_est_s.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
-Black line is $y=x$ and red line is $y=2x$.
-The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
-\end{subfigure}
-\end{center}
-\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
-\end{figure}
 
@@
 
+\begin{figure}
+\begin{center}
+\begin{subfigure}{\textwidth}
+\includegraphics[height=1.8in]{figures/lfsr_est_s_nn.png} 
+\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
+Black line is $y=x$ and red line is $y=2x$.
+The results illustrate how estimates of $\lfsr$ can be more accurate in this case.
+That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
+\end{subfigure}
+\begin{subfigure}{\textwidth}
+\includegraphics[height=1.8in]{figures/lfsr_est_s.png} 
+\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
+Black line is $y=x$ and red line is $y=2x$.
+The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
+\end{subfigure}
+\end{center}
+\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
+\end{figure}
+
 \begin{table}[!ht]
@@
 
-\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{table:nopen}
+\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{tab:nopen}
 \end{table}
@@
 
-Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the ggplot2 package \cite{ggplot2}, and text
+Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the {\tt ggplot2} package \cite{ggplot2}, and text
 prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} , and includes contributions 
+git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} and includes contributions 
 from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.

### 2015-10-13 examples/rosenblatt.rmd
added rosenblatt example

---
title: "mixture of zero and non-zero mean components"
output: html_document
---

Based on email exchange with Jonathan Rosenblatt (October 13, 2015) 
let's try to fit data to a model that is a 50-50 mixture of a distribution g0 centered at 0
and a distribution g1 centered at mu.

First I'll simulate from such a model:
```{r}
library(ashr)
set.seed(100)
ns = 1000 # number of samples
beta = c(rnorm(ns/2,0,1),rnorm(ns/2,3,1))
shat = 1 #just set noise to 1
betahat = rnorm(ns,beta,shat)
hist(betahat)
```

Now fit the model for various values of mu (I just used -1 to 5)
```{r}
mixsd = ashr:::autoselect.mixsd(betahat,shat,mult=sqrt(2)) #probably can improve choice of grid, but just to get something rolling
k= length(mixsd)

mu = seq(-1,5,length=7)
fit.ash=list()

for(i in 1:length(mu)){ #fit ash model for each mu
  g = normalmix(pi=rep(1/(2*k),2*k), mean = c(rep(0,k),rep(mu[i],k)), sd = c(mixsd,mixsd) )
  fit.ash[[i]] = ash(betahat,shat,g=g,prior="uniform")
}
```

Plot log-likelihood for different mu: it seems to be maximized near mu=3 which is promising.
```{r}
l=lapply(fit.ash,get_loglik)
plot(mu,l,xlab="mu",ylab="log-likelihood (maximized over g0,g1)")
```

Look at the estimated proportions of the zero vs 3 component
```{r}
x=seq(-1,5,length=100)
a = fit.ash[[which.max(l)]]
plot(get_density(a,x),type="l")
fit.ash[[which.max(l)]]
```

### 2015-10-14 talks/twentymin_seminar.Rmd
updated seminar for CSH

@@
 
-## How can we make more progress together?
-
-- Abstract our problems
-- Develop benchmarks
-- Build generic tools
-
 ## The power of generic tools
@@
 
-Goal: generic tools for shrinkage estimation
+Goal: generic tools for FDR and shrinkage estimation
 
@@
 
-- Analogously, one can assume that all Z scores near 0 are null. Efron refers to this as the ``Zero Assumption".
+- Analogously, one can assume that all Z scores near 0 are null. Efron refers to this as the "Zero Assumption".
 
@@
 ```{r, echo=FALSE, include=FALSE}
-require(fdrtool)
-  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue",plot=FALSE)
+#require(fdrtool)
+#  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue",plot=FALSE)
 require(locfdr)
@@
 
-  nullalthist(hh.zscore,hh.fdrtool$lfdr)  
+  nullalthist(hh.zscore,hh.q$lfdr)  
 ```
@@
 
-- In some cases the measurement precisions differ among units
+- Measurement precision can differ among units
 
-- Eg effect sizes of rare SNPs have larger standard error than those of common SNPs
+    - Eg effects of rare SNPs have larger standard error than common SNPs
 
-- Eg Expression levels of low-expressed genes have less precision than high-expressed genes
+    - Eg low-expressed genes less precise than high-expressed genes
 
-- If some effects are measured less precisely than others, those tests ``lack power"
-and dilute signal, increasing FDR
+- Low precision measurements add noise, dilute signal, increase FDR
 
@@
 
-- Following previous work (e.g. Newton, Efron, Muralidharan) we take an empirical Bayes approach to FDR.
+- Following previous work (e.g. Newton, Efron, Muralidharan) we take an empirical Bayes approach.
 
@@
 
-- Various semi-parametric approaches taken to estimating $f_1$. For example,
-Efron uses Poisson regression; Muralidharan uses mixture of normal distributions.
-
-- Once $f_1$ and $\pi_0$ estimated, FDR calculations are straightforward.
+- Once $f_1$ and $\pi_0$ estimated, FDR calculations straightforward.
 
@@
 
-(cf Bayes inference from summary statistics; V.Johnson, J.Wakefield)
-
 ## FDR - A New Deal
@@
 
-- Estimating $g$ comes down to estimating $\pi$. Joint estimation of $\pi_0,\pi$ easy by maximum likelihood (EM algorithm).
-
-- By allowing $K$ large, and $\sigma_k$ to span a dense grid of values,
-we get a flexible unimodal symmetric distribution.
-
-- Can approximate, arbitrarily closely, any scale mixture of normals.
-Includes almost all priors used for sparse regression problems (spike-and-slab, double exponential/Laplace/Bayesian Lasso, horseshoe).
-
-## FDR - A New Deal
-
-- Alternatively, a mixture of uniforms, with 0 as one end-point of the range,
-provides still more flexibility, and in particular allows for asymmetry.
-
-- If allow a very large number of uniforms this provides the non-parametric mle for $g$; cf Grenander 1953; Cordy + Thomas 1997.
+- Fix $K$ large, $\sigma_k$ fixed dense grid, estimate $\pi$.
 
@@
 
-- This approach actually provides a full posterior distribution for each $\beta_j$.
+- Approach provides not only FDR estimates, but posterior distribution for each $\beta_j$.
 
-- So easy to obtain point estimates and credible intervals.
-
-- Because $g(\beta)$ is unimodal, the point estimates (and CIs) will tend to be ``shrunk" towards the overall mean (0).
+- Because $g(\beta)$ is unimodal, the point estimates (and CIs) will tend to be "shrunk" towards the overall mean (0).
 
@@
 
-- So we call the approach ``Adaptive Shrinkage" (ASH).
+- So we call the approach "Adaptive Shrinkage" (ASH).
 
@@
 ```{r, echo=FALSE,fig.height=4,fig.cap=""}
-nullalthist(hh.zscore,hh.fdrtool$lfdr,main="qvalue")  
+nullalthist(hh.zscore,hh.q$lfdr,main="qvalue")  
 ```
@@
 
-## Some more extensive simulations
+## More extensive simulations
 
@@
 
-All methods are, generally, "conservative" (deliberately).
+- All methods are, generally, "conservative" (deliberately).
+
+- `ash` generally least conservative.
 
@@
 
-## A new problem: an embarrassment of riches
-
-- If the null is mostly false, this approach can  provide unsettling results
-
-- The FDR can be small for all observations, even those with $p \approx 1$!
-
-- In the illustrative example, the maximum $q$ value is `r round(max(hh.ash$qvalue),2)`
-
-## Perhaps we didn't really understand the question?
-
-- Problem arises only if we insist on asking question "is $\beta_j=0$?"
-
-- Given enough signal, we become convinced that very few of the $\beta_j=0$
-
-- But for some $\beta_j$ we still may have little information about actual value
-
-- Suggests a change of focus: ask for which $\beta_j$ are we confident about the sign (cf Gelman et al, 2012; Tukey 1961).
-
 ## Generic procedure: summary
@@
 $$\beta_j | s_j \sim g(), \text{unimodal}.$$
-
-- Output: estimate of $g$; "shrunken" posterior distribution for $\beta_j$.
-
-## Generic procedure: embellishments
-
-- Inputs: $(\hat{\beta}_1,s_1), \dots, (\hat{\beta}_p,s_p)$
-- Assumptions:
-
-$$\hat{\beta}_j | \beta_j, s_j \sim t(\beta_j,s_j)$$
-$$\beta_j | s_j \sim s_j^\alpha g(), \text{unimodal}.$$
 
@@
 
-- Incorporate shrinkage of variances and not just means. (e.g. ``moderated $t$ test", Smyth et al).
+- Incorporate shrinkage of variances and not just means. (e.g. "moderated $t$ test", Smyth et al).
 
@@
 
-- NHGRI, Moore Foundation for funding.
+- NHGRI, and the Gordon and Betty Moore Foundation for funding.
 

### 2015-10-24 paper/main.tex
minor update

@@
 in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
-as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
+as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=\min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
 based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$

### 2015-10-26 dsc-robust/run_dsc_robust.R
change location to which dsc-robust results are saved

@@ -33,7 +33,7 @@ add_score(dsc_robust,score)
 ######## Run the DSC #################
 
 res_robust=run_dsc(dsc_robust)
-save(dsc_robust,file="dsc_robust.RData")
+save(dsc_robust,file="dsc-robust-files/dsc_robust.RData")
 
 
 

### 2016-04-16 talks/one_hour_seminar.Rmd
update seminar

@@
 author: Matthew Stephens
-date: March 9, 2015
+date: April 15, 2016
 output: ioslides_presentation
@@
 - Great for sharing material across multiple people (eg student and advisor!)
-- An amateur example: [http://github.com/stephens999/ash](http://github.com/stephens999/ash)
+- For example: [https://github.com/jdblischak/singleCellSeq](https://github.com/jdblischak/singleCellSeq)
+- A template (in progress): [https://github.com/jhsiao999/ashlar](https://github.com/jhsiao999/ashlar)
 
@@
 
-## Commercial Break:
+<!-- ## Commercial Break:  -->
 
-A quick ad for a recently-started project I call "Dynamic Statistical Comparisons":
+<!-- A quick ad for a recently-started project I call "Dynamic Statistical Comparisons": -->
 
-- Overall Goal: Improve how the statistics and genomics (and other) communities make comparisons among methods.
+<!-- - Overall Goal: Improve how the statistics and genomics (and other) communities make comparisons among methods. -->
 
-- How? Centralize reproducible statistical comparisons in easily-extensible public repositories.
+<!-- - How? Centralize reproducible statistical comparisons in easily-extensible public repositories. -->
 
-## Dynamic Statistical Comparisons: advantages
+<!-- ## Dynamic Statistical Comparisons: advantages -->
 
-- Authors can run their own methods. 
-- Ease of adding new data can quickly uncover over-tuning to specific datasets.
-- Work is divided, and not duplicated.
-- Promotes collaboration through competition. 
-- Ensures results are reproducible, and extendable.
+<!-- - Authors can run their own methods.  -->
+<!-- - Ease of adding new data can quickly uncover over-tuning to specific datasets. -->
+<!-- - Work is divided, and not duplicated. -->
+<!-- - Promotes collaboration through competition.  -->
+<!-- - Ensures results are reproducible, and extendable. -->
 
-## More on git, github, knitr, reproducibility, DSCs
+## More on git, github, knitr, reproducibility
 
@@
 - Google "donohoe buckheit" for "Wavelab and reproducible research"
-- [http://stephens999.github.io/blog/](http://stephens999.github.io/blog/) for more on DSC
 
@@
 
-- Estimating $g$ comes down to estimating $\pi$. Joint estimation of $\pi_0,\pi$ easy by maximum likelihood (EM algorithm).
+- Estimating $g$ comes down to estimating $\pi$. Joint estimation of $\pi_0,\pi$ easy by maximum likelihood (convex optimization).
 
-- By allowing $K$ large, and $\sigma_k$ to span a dense grid of values,
-we get a flexible unimodal symmetric distribution.
+- Setting $K$ large, with dense grid of $\sigma_k$, yields flexible unimodal symmetric distribution.
 
@@
 
-- Because $g(\beta)$ is unimodal, the point estimates (and CIs) will tend to be ``shrunk" towards the overall mean (0).
+- Because $g(\beta)$ is unimodal, the point estimates (and CIs) will tend to be "shrunk" towards the overall mean (0).
 
@@
 
-## Simulated example: $\pi_0$ not identifiable.
+## More extensive simulations
 
-```{r,echo=FALSE,include=FALSE}
-load("../paper/Rcode/sim1.RData") #load simulation results
-source("../paper/Rcode/plot_pi0.R")
-```
+![Image](../paper/figures/scenarios_density.pdf)
 
-```{r, echo=FALSE,fig.height=4,fig.cap=""}
-plot_pi0(list(simres1a))
-```
+$$\beta_j \sim \pi_0 \delta_0(.) + (1-\pi_0) g(.)$$
 
-## Simulated Example: so fdr not identifiable
+Goals: Estimate $\pi_0$? Estimate $g$?
 
-```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-source("../paper/Rcode/plot_lfsr.R")
-  plot_lfsr(list(simres1a),0.1,ptype="lfdr")
-```
+## Estimation of $\pi_0$
 
-## Simulated Example: fsr much more identifiable
+![Image](../paper/figures/pi0_est.pdf)
 
-```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
-  plot_lfsr(list(simres1a),0.1,ptype="lfsr")
-```
+- All methods are, generally, "conservative" (deliberately).
+
+- `ash` generally least conservative.
+
+## Estimation of $g$
+
+- Estimating $g$ is a “deconvolution problem”.
+- Notoriously difficult in general. 
+- Efron emphasises the difficulties of implementing a stable general algorithm:
+
+<div class="red">
+“the effort foundered on practical difficulties involving the perils of deconvolution...
+Maybe I am trying to be overly nonparametric...
+but it is hard to imagine a generally satisfactory parametric formulation...”
+</div>
+
+## Unimodal assumption stabilizes estimation of $g$
+
+![Image](../paper/figures/egcdf.pdf)
 
@@
 
-## Other Applications
+## Applications/Generalizations
 
-- Widely applicable: requiring only an estimated
-effect size and standard error for each object.
+- The idea of inducing sparsity/shrinkage via unimodal distributions is very general!
 
-- E.g. Currently applying it to wavelet shrinkage applications.
+- smash: Smoothing via adaptive shrinkage (wavelet-based)
 
-## Next steps?
+- vash: shrinkage of variances and not just means. (e.g. ``moderated $t$ test", Smyth et al) [Mengyin Lu]
 
-- Incorporate shrinkage of variances and not just means. (e.g. ``moderated $t$ test", Smyth et al)
+- flash: shrinkage of factors and loadings in factor analysis [Wei Wang; David Gerard]
 
-- Allow $g(\cdot;\pi)$ to depend on covariates $X$.
+- mash: allow for matrix of observations/effects $\hat\beta_{ij}$ [Sarah Urbut, Gao Wang]
 
-- Allow for correlations in the measured $\hat\beta_j$.
+- cash: allow $g(\cdot;\pi)$ to depend on covariates $X$ [Scott Powers, Lei Sun]
 
@@
 
-- Including  Chaoxing Dai, Mengyin Lu, Ester Pantaleo, Scott Powers, Sen Tian, Wei Wang, Zhengrong Xing.
+- Including  Chaoxing Dai, Mengyin Lu, Ester Pantaleo, Sen Tian, Wei Wang, Zhengrong Xing, Sarah Urbut
 
